### 2011-11-14 ack.tex
restructured

So many researchers contribute to any dissertation that to name them all compromises the honor of those most involved. More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, for his commitment to my work and the principle of explanatory adequacy. My other internal committee members, Gene Buckley, Mark Liberman, and Rolf Noyer, and the faculty at IRCS, are responsible for creating an environment at Penn where this (admittedly contentious) dissertation could be written. Though his name appears nowhere else in this document, it is hard to imagine it could exist without the pioneering work of Bill Labov.

The subject of the first chapter was inspired by the mentoring of Jennifer Cole. Important contributions to this dissertation were made by Stephen Anderson, who graciously agreed to serve on my committee, Michael Wagner, who gave me a goal, and Morris Halle, who taught me to never apologize.

I would like to thank my fellow students at Penn, but especially my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody. Elsewhere, I have been shown a great deal of kindness by so many, but especially Morgan Sonderregger and Maryam Bakht. Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document. Among non-linguists, my good friend Carly Fisher stands out for her unflagging support.

### 2011-11-14 clusters.tex
restructured

As reviewed below, this was the view of many phonologists of the 1960s and 1970s. While several classic documents indicated problesm with this view, I will argue that the broader view of

phonological

At the same time, the hypothesis seems to have been ignored by work in the late 1980s and early 1990s, and that

and no compelling reasons have been offered to revise it. If anything, it appears that this hypothesis was forgotten or ignored during the 1980s and early '90s, and that despite the fact that it was championed by the founding Optimality Theory

 founding document forgotten or ignored during the 1980s and early '90s, and that despite the fact that it was championed by the founding document forgotten or ignored during the 1980s and early '90s, and that despite the fact that it was championed by the founding document forgotten or ignored during the 1980s and early '90s, and that despite the fact that it was championed by the founding document

 work was forgotten during the 1980s, briefly revived by the the earliset

Linguists of this era were not ignorant of the possibilities of obtaining experimental evidence bearing on MSCs. A classic study, and one that is a  demonstration of the fundamental correctness of the classic view

\footnote{\citeauthor{Zimmer1969}'s paper uses place features to distinguish vowels, so that front vowels are [$+$\textsc{palatal}], and so on. I have taken the liberty of using more familiar vowel features, such as [$\pm$\textsc{high}, $\pm$\textsc{back}].}

[$+$\textsc{back}] \emph{u, o, \i, a}
[$-$\textsc{back}] \emph{i, e, \"u, \"o}

questionnaire

Subjects were also permitted to check both words, indicating ``no preference''. 
\begin{tabular}{l r l r}
\toprule
front harmony &      & not \\
\midrule
temez     & $19$ & temaz   & $3$ & $1$ \\
tipez     & $21$ & tipaz   & $1$ & $1$ \\
ter\"uz   & $20$ & teruz   & $1$ & $2$ \\
t\"or\"uz & $19$ & t\"oriz & $1$ & $3$ \\
teriz     & $23$ & ter\i z & $0$ & $0$ \\
% tamuz   & $3$  & tam\I z & $16$ \\
% tafuz   & $3$  & taf\I z & $17$ \\
% tapuz   & $7$  & tap\I z & $9$ \\
\bottomrule
\end{tabular}

\citet[][313]{Zimmer1969}

\citep{Goodman1954}[][749]

\begin{equation}
\gamma = \frac{c - d}{c + d}
\end{equation}

$c = 102$, $d = 6$, $\gamma = 0.888$

\begin{tabular}{l r l r}
\toprule
[\emph{a}[labial]\emph{u}  &      & *\emph{a}[labial]\emph{\i] & 
\midrule
pamuz    & $17$ & pam\i z & $15$ \\
tafuz    & $21$ & taf\i z & $11$ \\
tapuz    & $17$ & tap\i z & $15$ \\
mavuz    & $16$ & mav\i z & $16$ \\
tabuz    & $16$ & tab\i z & $16$ \\
\bottomrule
\end{tabular}
\citet[][314]{Zimmer1969}

$c = 87$, $d = 73$, $\gamma = 0.088$

\section{The phenomenon}

% why syllable contact clusters
I adopt English syllable contact clusters in monomorphs. Two considerations guide this choice. First, English syllable contact clusters are the subject of many prior phonological and phonotactic analyses. Second, the data should be accessible to any reader.

% lexicon
The \emph{lexicon} has many senses: throughout, I use the term to refer to the set of underlying representations \citep[][269]{LANGUAGE}. I depart from \citeauthor{LANGUAGE}'s characteriztion of the lexicon in two ways. First, I assume that this set is merely a reflection of individual speakers' knowledge of their language, realized in the human; it has no further reality. Second, in using the term, I do not intend to refer to any knowledge of morphological irregularities, which also reside in the lexicon for \citeauthor{LANGUAGE} (ibid., p.~273).

\section{Phonotactic gaps and the null hypothesis}

Before studying

\section{History}

A major result of the American structuralists was the finding that the set of sound sequences found in morphs and across morpheme boundaries are different.\footnote{The discussion in this paragraph is based on \citealt[][267]{Anderson1985}.} \citet{Bloomfield1930} proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries. \citeauthor{Bloomfield1930} notes that [\c{c}] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [ku\lm x\textschwa n] `cake' $\sim$ \emph{Kuhchen} [ku\lm\c{c}\textschwa n] `cow-let'. \citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf. \emph{Kuh} `cow'). Under the assumption that [x] occurs in this position (before a back vowel) only when this preceding vowel belongs to the same word or morpheme, [\c{c}] and [x] can be cast as allophones of the same phoneme (likely /\c{c}/, the basic variant). A natural consequence of this view is that generalizations about sound sequences found within morphemes are necessarily different within and across morpheme boundaries, 
and thus a distributional analysis must take morphology into account, as argued forcefully by \citet{Pike1947b}. \citeauthor{Pike1947b} discusses limitations on possible vowel sequences in Mixteco monomorphs, but it is easy to find exceptions to these generalizations when a morpheme boundary intervenes in the short Mixteco text analyzed in \citealt{Pike1944}.

\ex Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}: \\ 
\begin{tabular}{r l l l} %\toprule
   & MSC & complex word \\ %\midrule
a. & *{C}a{C}e & k\'a-\textsuperscript{n}dee & `kept \ldots inside' \\
b. & *{C}\textipa{@}{C}e & n\`i-k\textipa{\`@b@-de} & `who entered'        \\
c. & *{C}e{C}i & te-n\'i-ke\textsuperscript{n}da & `was walking         \\
d. & *{C}i{C}e & te-n\`i-kee-t\`\textschwa & `and went away'      \\ %\toprule
%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
\end{tabular} \xe

\noindent
\citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, and this idea was further developed by \citet{Harris1955}, who proposes that this mechanism allows the linguist to identify morpheme boundary.

While the above work is widely known for introducing ``grammatical'' (which in the sense current at the time just meant ``supraphonemic'') components into the theory of phonology, albeit indirectly, similar (and arguably more sophisticated) intuitions about the interaction between phonotactic generalizations and morpheme boundaries were made by members of the Prague Circle at roughly the same time, a fact which has not been noted to my knowledge. \citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture. For instance, most Russian infinitives end with a final /t'/. The reflexive variant of these verbs is marked with a final /-sa/, which feeds a general process of regressive voicing assimilation, and this /\ldots t\textceltpal -s\ldots/ juncture triggers the loss of palatalization. However, palatalization of a stem-final labial or coronal---including /t'/, as in ?c ---is preserved in reflexive imperatives, which are formed by attaching the reflexive /-sa/ to the bare stem.\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which makes use of a highly abstract phonemic transcription. The broad phonetic transcriptions below are the result of consultations with N linguistically na\"ive native Russian speakers. Stress and palatalization are marked regardless of whether they are constrastive in that position.}

\ex Russian reflexives \citep[after][]{Jakobson1932}: \\
\begin{tabular}{l l l l} %\toprule
   &  infinitive & imperative \\ %\midrule
a. & sl\textceltpal av\textceltpal its\textschwa         & sl\'af\textceltpal s\textschwa            & `be glorious'    \\
   & upr\'amits\textschwa         & upr\'am\textceltpal s\textschwa          & `be stubborn'    \\
b. & kr\'as\textceltpal its\textschwa         & kr\'as\textceltpal s\textschwa            & `put on makeup'  \\
   & \v{z}\'arits\textschwa       & \v{z}\'ar\textceltpal s\textschwa         & `roast'          \\ 
c. & z\textschwa b\'yts\textschwa & z\textschwa b\'ut\textceltpal s\textschwa & `forget' \\ %\bottomrule
%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
\end{tabular} \xe

Unlike in \citeauthor{Bloomfield1930}'s analysis of German, the cluster of palatal coronal followed by /s/ has an intervening morpheme boundary in both infinitive and imperative. \citeauthor{Jakobson1932} proposes that the two /t'-sa/ junctures in c, in fact different, with the imperative not giving rise to the loess of palatalization. Obvious parallels of this more powerful notion can be found in Lexical Phonology or the model developed by \citet{Halle1987}, to name a few. A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.

Arguably, this more powerful system might be appropriate for \citeauthor{Bloomfield1930}'s data, if it is the case that the final \emph{-en} in \emph{Kuchen} is also a morpheme (perhaps related to \emph{K\"uche} `kitchen; cuisine').

%c. & l'etS'        & l'\'akka     & `lie down' &
%for the verb l'\'e\v{c} `to lie down'. From two present active forms (among other parts of the paradigm), 1sg. \emph{l'\'ago}, 3pl., \emph{l'\'agot}, one can detect an argument for analyzing the root as /g/-final. The imperative, \emph{l'\'akka},

### 2011-11-14 pre.tex
restructured

\section{``What is the most important fact?''}

Perhaps the most important fact for the study of speakers' internal knowledge of language is Wilhelm von Humboldt's observation that speakers make ``infinite use of finite means'' CITE \citep{Chomsky1966}.

An enduring demonstration of ``infinite use'' is the famous emph{wug-test} \citep{Berko1958}. Jean Berko-Gleason introduced English children aged ?-? to pictures of concrete objects (such as a small bird), labeled them with a nonce word (``that's a \emph{wug}''), and then asked them to finish the sentence

The children in the study produced the same response you may have: the plural is \emph{wugs} (where $\langle$s$\rangle$ writes a final [z]).

Of course, children have never heard the plural of these nonce words, but they extend /-z/ to these new words.

But what of the ``finite means'' from which grammars spring? On this, a good deal less has been said.
a study of spaekers' knowledge of natural
The most important fact for

This

### 2011-11-14 zipf_sim.py

@@ -1,0 +1,187 @@
+#!/usr/bin/env python
+""" Copyright (c) 2011 Kyle Gorman
+
+ Permission is hereby granted, free of charge, to any person obtaining a copy
+ of this software and associated documentation files (the "Software"), to deal
+ in the Software without restriction, including without limitation the rights
+ to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+ copies of the Software, and to permit persons to whom the Software is
+ furnished to do so, subject to the following conditions:
+
+ The above copyright notice and this permission notice shall be included in
+ all copies or substantial portions of the Software.
+
+ THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+ IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+ FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+ AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+ LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+ OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
+ THE SOFTWARE.
+
+ Zipfian distribution sampling simulations
+ Kyle Gorman <kgorman@ling.upenn.edu>
+ Usage is described below. """
+
+from sys import argv
+from bisect import bisect
+from random import random
+from math import log, fsum
+from collections import defaultdict
+
+
+def sparse_dist(N, alpha=1, beta=0):
+    return [(i + beta) ** -alpha for i in xrange(N, 0, -1)]
+
+
+def mean(x):
+    """ 
+    Return arithmetic mean of a list of continuous values 
+    """
+    return fsum(x) / len(x)
+
+
+def median(x, sorted=False):
+    """ 
+    Return median of an iterable of continuous values
+    """
+    my_x = x[:] # deep copy
+    if not sorted:
+        my_x.sort()
+    if len(x) % 2 == 0: # even 
+        l = len(my_x) / 2
+        return (my_x[l - 1] + my_x[l]) / 2.
+    else: # odd
+        return my_x[len(my_x) // 2]
+
+
+class ProbDist(object):
+
+    def __init__(self, item2prob=None):
+        """ 
[…]

### 2011-11-14 clusters.tex
daily commit

@@
-As reviewed below, this was the view of many phonologists of the 1960s and 1970s. While several classic documents indicated problesm with this view, I will argue that the broader view of
+The \emph{lexicon} has many senses: throughout, I use the term to refer to the set of underlying representations \citep[][269]{LANGUAGE}, but I depart from \citeauthor{LANGUAGE}'s understanding by further adopting the assumption that this set is merely a reflection of individual speakers' knowledge of their language, realized in the human; it has no further reality.
+
+\section{The null hypothesis}
+
+\subsection{Structuralist phonology}
+
+A major result of the American structuralists was the finding that generalizations which hold of the sound sequences found in underlying forms may not hold across morpheme boundaries.
+If this is correct, then there must be some sense in which phonology is sensitive to the presence of morpheme boundaries.
+In his history of phonology, \citet[][267]{Anderson1985} credits this innovation to \citet{Bloomfield1930}, who proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries.
+\citeauthor{Bloomfield1930} notes that [\c{c}] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [ku\lm x\textschwa n] `cake' $\sim$ \emph{Kuhchen} [ku\lm\c{c}\textschwa n] `cow-let'.
+\citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf.
+\emph{Kuh} `cow').
+Under the assumption that [x] occurs in this position (before a back vowel) only when this preceding vowel belongs to the same word or morpheme, [\c{c}] and [x] can be cast as allophones of the same phoneme (likely /\c{c}/, the basic variant).
+
+A natural consequence of this view is that generalizations about sound sequences found within morphemes are necessarily different within and across morpheme boundaries, and \citet{Pike1947b} argues that the distributional analysis of sound sequences must take morphology into account.
+\citeauthor{Pike1947b} discusses limitations on possible vowel sequences in Mixteco monomorphs, but none of these generalizations hold across morpheme boundaries, at least according to the phonological analysis of a short Mixteco text in one of \citeauthor{Pike1944}'s earlier papers.
+
+\ex Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}: \\ 
+\begin{tabular}{r l l l} %\toprule
+   & MSC & complex exception \\ %\midrule
+%a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
+%b. & *{C}\textipa{@}{C}e & [n\`i-k\`\textschwa b\textschwa-de] & `who entered'        \\
+%c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
+%d. & *{C}i{C}e & [te-n\`i-kee-t\`\textschwa] & `and went away'      \\ %\toprule
+a. & *{C}a{C}e & [k\'a\textsuperscript{n}dee] & `kept \ldots inside' \\
+b. & *{C}\textipa{@}{C}e & [n\`ik\`\textschwa b\textschwa de] & `who entered'        \\
+c. & *{C}e{C}i & [ten\'i ke\textsuperscript{n}da] & `was walking         \\
+d. & *{C}i{C}e & [ten\`i keet\`\textschwa] & `and went away'      \\ %\toprule
+%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
+%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
+\end{tabular} \xe
+
+\noindent
+\citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a morpheme discovery routine.
+
+While \citeauthor{Bloomfield1930} is known for introducing morphemic structure into phonology, a member of the Prague circle developed an arguably more sophisticated approach only two years later.
+\citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture.
+For instance, most Russian infinitives end with a final /t'/.
+The reflexive variant of these verbs is marked with a final /-sa/, which feeds a general process of regressive voicing assimilation, and this /\ldots t\textceltpal -s\ldots/ juncture triggers the loss of palatalization.
+However, palatalization of a stem-final labial or coronal---including /t'/, as in (\nextx c)---is preserved in reflexive imperatives, which are formed by attaching the reflexive /-sa/ to the bare stem.\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses a highly abstract phonemic transcription.
+The broad phonetic transcriptions below are the result of consultations with N linguistically na\"ive native Russian speakers.
+Palatalization is marked even where not constrastive.}
+
+\ex Russian reflexives \citep[after][]{Jakobson1932}: \\
+\begin{tabular}{l l l l} %\toprule
+   &  infinitive & imperative \\ %\midrule
+a. & [\textprimstress sl\textceltpal av\textceltpal its\textschwa]         & [\textprimstress slaf\textceltpal s\textschwa]            & `be glorious'    \\
+   & [u\textprimstress pramits\textschwa]         & [u\textprimstress pram\textceltpal s\textschwa]          & `be stubborn'    \\
+b. & [\textprimstress kras\textceltpal its\textschwa]         & [\textprimstress kras\textceltpal s\textschwa]           & `put on makeup'  \\
+   & [\textprimstress\textyogh arits\textschwa]       & [\textprimstress\textyogh ar\textceltpal s\textschwa]         & `roast'          \\ 
+c. & [\textprimstress z\textschwa byts\textschwa] & [\textprimstress z\textschwa but\textceltpal s\textschwa] & `forget' \\ %\bottomrule
+%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
+\end{tabular} \xe
+
+\noindent
+
+Both forms in (\lastx c) contain a /t\textceltpal-s/ juncture, though only one is faithful.
+\citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization.
+A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.
+
+Arguably, this more powerful system might be appropriate for \citeauthor{Bloomfield1930}'s data, if it is the case that the final \emph{-en} in \emph{Kuchen} is also a morpheme (perhaps related to \emph{K\"uche} `kitchen; cuisine').
+There are obvious parallels between \citeauthor{Jakobson1932}'s analysis and notions like readjustment in \emph{SPE}, or cyclic rules in lexical phonology or the \citet{Halle1987} model.
+
+Further, I promise that the only compelling reason to reconsider this view is no longer of force, and
+
+As reviewed below, this was the view of many phonologists of the 1960s and 1970s. While several classic documents indicated problems with this view, I will argue that the broader view of
+
+It is worthwhile to consider the appraisal offered by some key textbooks of the era.
+
+\citet{Dell1973}
+
+\citet{Anderson1974}
+
+\citet{Kenstowicz1979}
 
@@
 
-Linguists of this era were not ignorant of the possibilities of obtaining experimental evidence bearing on MSCs.
-A classic study, and one that is a  demonstration of the fundamental correctness of the classic view
+\subsection{Generative phonology}
 
-\footnote{\citeauthor{Zimmer1969}'s paper uses place features to distinguish vowels, so that front vowels are [$+$\textsc{palatal}], and so on.
-I have taken the liberty of using more familiar vowel features, such as [$\pm$\textsc{high}, $\pm$\textsc{back}].}
+\subsection{Early psycholinguistic evidence}
 
-[$+$\textsc{back}] \emph{u, o, \i, a}
-[$-$\textsc{back}] \emph{i, e, \"u, \"o}
+Linguists of this era were not ignorant of the possibilities of obtaining experimental evidence bearing on MSCs, and the experiments by \citet{Greenberg1964} and \citet{Scholes1966} will be a major topic of the following chapter.
+Here, I will discuss a study by \citet{Zimmer1969} concerning the synchronic status of three Turkish morpheme structure constraints and their relation to synchronic processes of vowel harmony, after touching on a few grammatical prerequisites.
 
-questionnaire
+Turksih vowel phonemes can be divided by height, roundness, and backness:\footnote{\citeauthor{Zimmer1969} uses place features to distinguish vowels, so that, for instance, \emph{i} is [$+$\textsc{palatal}, $-$\textsc{labial}].
+I have taken the liberty of using a more familiar feature specification, though nothing hinges on this choice.
+I have also structured my discussion of the Turkish facts to avoid the need to discuss the details of roundness harmony.
+See \citealt{Zimmer1969} for additional details.}
 
+\ex Feature specification for Turkish vowels: \par\nobreak
+\begin{tabular}{l c c c c}
+ & \multicolumn{2}{c}{[$-$\textsc{back}]} & \multicolumn{2}{c}{[$+$\textsc{back}]} \\
+& [$-$\textsc{round}] & [$+$\textsc{round}] & [$-$\textsc{round}] & [$+$\textsc{round}] \\
+\buf[$+$\textsc{high}] & \emph{i} [i]        & \emph{\"u} [y]      & \emph{\i} [\textturnm] & \emph{u} [u] \\
+\buf[$-$\textsc{high}] & \emph{e} [e]        & \emph{\"o} [\o]     & \emph{a} [\textschwa]  & \emph{o} [o] \\
+\end{tabular}
+\xe
+
+%into four [$+$\textsc{back}] vowels \emph{u}, \emph{\i}, \emph{o}, \emph{a}, denoting [u, \textturnm, o, a] and four [$-$\textsc{back}] vowels \emph{i}, \emph{\"u}, \emph{e}, \emph{\"o} denoting [i, y, e, \o].
+
+Native roots generally contain only [$+$\textsc{back}] vowels (e.g., \emph{g\"orev} `duty') or [$-$\textsc{back}] vowels (e.g., \emph{odun} `wood'), but loanwords need not conform to this generalization (e.g., \emph{sosis} `sausage').
+Despite this latter fact, there is some evidence that this constraint on roots is an induced one.
+
+\ex Turkish declensional suffix allomorphy (after \citealt[][315]{Zimmer1969}): \par\nobreak
+\begin{tabular}{l l l l l}
+   & nom.sg.      & acc.sg.         & gen.sg.              \\
+a. & \emph{it}    & \emph{iti}      & \emph{itin}      & `dog'  \\
+%a. & \emph{g\"ul} & \emph{g\"ul\"u} & \emph{g\"ul\"un} & `rose' \\
+c. & \emph{k\i z} & \emph{k\i z\i}  & \emph{k\i z\i n} & `girl' \\
+%c. & \emph{tuz}   & \emph{tuzu}     & \emph{tuzun}     & `salt' \\
+\end{tabular}
+\xe
+
+\noindent
+Similar facts are indicated by the rest of the declensional paradigms. It is possible to analyze the forms of the acc.sg. 
+(\emph{i}, \emph{\i})
+%(\emph{\"u}, \emph{u})
+%(\emph{i}, \emph{\"u}, \emph{\i}, \emph{u})
+and gen.sg. 
+(\emph{in}, \emph{\i n})
+%(\emph{\"un}, \emph{un})
+%(\emph{in}, \emph{\"un}, \emph{\i n}, and \emph{un}) 
+without suppletion, using the following phonological rule:
+
+\ex \buf[\textsc{back}] \textsc{Harmony}: %\par\nobreak
+V $\rightarrow$ [$\alpha$\textsc{back}] / V[$\alpha$\textsc{back}] C$_0$ \gap\gap \xe
+
+In other words, a vowel takes on the [\textsc{back}] specifications of the preceding vowel.
+The acc.sg.
+suffix is then [$+$\textsc{high}] (and for these cases, [$-$\textsc{round}]); there is no evidence bearing on whether it is one of /i, \i/ or simply underspecified for [back].
+The gen.sg.
+suffix is the same vowel followed by /n/.
+
+If (\lastx) is a correct synchronic generalization, it clearly can induce the aforementioned morpheme structure constraint requiring root vowels to agree on [\textsc{back}].
+However, (\lastx) as stated must contend with loanword exceptions.
+These exceptions might be used to argue that it is not a phonologically general rule, and should be restricted, for instance, to apply only in morphologically derived environments.
+I will return to this question shortly.
+
+Another morpheme structure constraint discussed by \citeauthor{Zimmer1969} is that a vowel immediately preceded by \emph{a} and a labial consonant (\emph{p}, \emph{b}, \emph{f}, \emph{v}) must be [$+$\textsc{round}], though once again there are a few roots that are exceptions to this generalization.
+What is crucial here is that this labial consonant constraint is not involved in any suffix alternation, and in fact, does not hold of complex words.
+For instance, the acc.sg.
+of \emph{av} `hunt' is not *\emph{avu}, but rather \emph{av\i}.
+This suggests that this constraint cannot be considered an induced restriction.
+
+\citeauthor{Zimmer1969} attempts to study whether speakers respond differently to induced restrictions and static restrictions in a wordlikeness task.
+The linking hypothesis for such a task and the grammatical system will be taken up in the following chapter: for the moment, I will simply assume, with \citeauthor{Zimmer1969}, that some reasonable linking hypothesis can be put forth.
+
+%In Zimmer's first experiment, native Turkish speakers were given a questionnaire where each line consisted of two nonce words written in the modern orthography.
+The subjects were asked to read the words allowed and to put a checkmark next to the word they preferred.
 Subjects were also permitted to check both words, indicating ``no preference''.
-\begin{tabular}{l r l r}
+Four of the pairs in the first study consisted of a pair of nonce words differing only whether or not they are consistent with the [\textsc{back}] \textsc{Harmony} generalization.
+These results are given in Table \ref{backharm}.
+%\begin{table}
+%\centering
+%\begin{tabular}{l r l r r}
+%\toprule
+%hi \\
+%\midrule
+%\emph{temez}   & $19$ & \emph{temaz} & $3$ & $1$ \\ %back
+%%\emph{tipez}   & $21$ & \emph{tipaz} & $1$ & $1$ \\ %back
+%\emph{ter\"uz} & $20$ & \emph{teruz} & $1$ & $2$ \\ %back
+%%t\"or\"uz & $19$ & t\"oriz & $1$ & $3$ \\ %round
+%\emph{teriz}   & $23$ & ter\i z & $0$ & $0$ \\ 
+%% tamuz   & $3$  & tam\I z & $16$ \\
+%% tafuz   & $3$  & taf\I z & $17$ \\
+%% tapuz   & $7$  & tap\I z & $9$ \\
+%\bottomrule
+%\end{tabular}
+%\caption{This is a test, \citep[][313]{Zimmer1969}}
+%\label{backharm}
+%\end{table}
+
+%\begin{tabular}{l r l r r}
+%\toprule
+%hi \\
+%\midrule
+%tamuz & $3$ & tam\i z & $16$ & $4$ \\
+%tafuz & $3$ & taf\i z & $17$ & $3$ \\
+%tapuz & $7$ & tap\i z & $9$  & $7$ \\
+%tavuz & $9$ & tav\i z & $4$  & $10$ \\
+%tabuz & $5$ & tab\i z & $12$ & $6$ \\
+%\bottomrule
+%\end{tabular}
+%\caption{\citet[][313]{Zimmer1969}}
+%\end{table}
+
+In Zimmer's Experiment II, native Turkish speakers heard pairs of nonce words spoken, by a native speaker.
+Fifteen minutes later, they heard a second presentation of each of the pairs, and were asked to select which of the two words in the pair sounded more like a Turkish word.
+Four of the pairs differ solely in whether they obey or disobey the [\textsc{back}] \textsc{Harmony} generalization.
+These results are given in Table \ref{backharm}.
+
+\begin{table}
+\centering
+\begin{tabular}{l r l r r}
 \toprule
-front harmony &      & not \\
+harmonic      & \#   & disharmonic & \#   \\
 \midrule
-temez     & $19$ & temaz   & $3$ & $1$ \\
-tipez     & $21$ & tipaz   & $1$ & $1$ \\
-ter\"uz   & $20$ & teruz   & $1$ & $2$ \\
-t\"or\"uz & $19$ & t\"oriz & $1$ & $3$ \\
-teriz     & $23$ & ter\i z & $0$ & $0$ \\
-% tamuz   & $3$  & tam\I z & $16$ \\
-% tafuz   & $3$  & taf\I z & $17$ \\
-% tapuz   & $7$  & tap\I z & $9$ \\
+\buf[pemez]   & $30$ & [pemaz]     & $2$  \\
+\buf[teriz]   & $28$ & [ter\i z]   & $3$  \\
+\buf[tokaz]   & $26$ & [tokez]     & $6$  \\
+\buf[tipez]   & $24$ & [tipaz]     & $8$  \\
+\buf[ter\"uz] & $19$ & [teruz]     & $13$ \\
 \bottomrule
 \end{tabular}
+\caption{this is at test \citet[][314]{Zimmer1969}}
+\label{backharm}
+\end{table}
 
-\citet[][313]{Zimmer1969}
-
-\citep{Goodman1954}[][749]
+In each case, the root obeying [\textsc{back}] \textsc{Harmony} is strongly preferred.
+One way to quantify the preference for the harmonic stems is as follows.
+An observation from this experiment is called experiment can be thought of being either \emph{concordant} with the hypothesis that speakers will prefer harmonic nonce words if the harmonic nonce word is selected as better.
+If the non-harmonic nonce word is selected, this observation is called \emph{discordant}.
+``No preference'' observations are ignored.
+\citet[][749]{Goodman1954} use the count of concordant pairs $c$ and discordant pairs $d$ to define a metric $\gamma$:
 
@@
 
-$c = 102$, $d = 6$, $\gamma = 0.888$
+This $\gamma$ ranges between $-1$, indicating complete disagreement, and $1$, indicating that all data is is consistent with the hypothesis.
+For the four nonce words in table one, $\gamma = 0.597$, which indicates that speakers do have a moderately preference for the harmonic nonce words.
+In the same experiment, subjects also encountered five pairs which differed on whether they obeyed the labial consonant generalization.
+The results are shown in Table \ref{labcons}.
 
+\begin{table}
+\centering
 \begin{tabular}{l r l r}
 \toprule
-[\emph{a}[labial]\emph{u}  &      & *\emph{a}[labial]\emph{\i] & 
+harmonic & \# & disharmonic & \# \\
 \midrule
-pamuz    & $17$ & pam\i z & $15$ \\
-tafuz    & $21$ & taf\i z & $11$ \\
-tapuz    & $17$ & tap\i z & $15$ \\
-mavuz    & $16$ & mav\i z & $16$ \\
-tabuz    & $16$ & tab\i z & $16$ \\
+\buf[tafuz] & $21$ & [taf\i z] & $11$ \\
+\buf[pamuz] & $17$ & [pam\i z] & $15$ \\
+\buf[tapuz] & $17$ & [tap\i z] & $15$ \\
+\buf[mavuz] & $16$ & [mav\i z] & $16$ \\
+\buf[tabuz] & $16$ & [tab\i z] & $16$ \\
 \bottomrule
 \end{tabular}
-\citet[][314]{Zimmer1969}
+\caption{this is at test \citet[][314]{Zimmer1969}}
+\label{labcons}
+\end{table}
 
-$c = 87$, $d = 73$, $\gamma = 0.088$
+Here is it clear that the preferences, while going in the expected direction, are far smaller ($\gamma = 0.088$).
+Further, only one of five pairs, \emph{tafuz} $\sim$ \emph{taf\i z}, matches the stronger preferences seen with backness harmony.
+It might be possible to though any failure to find significance here should surely be attributed to insufficient statistical power, not evidence for the null hypothesis.
 
-\section{The phenomenon}
+\subsection{Optimality Theory}
 
-% why syllable contact clusters
-I adopt English syllable contact clusters in monomorphs.
-Two considerations guide this choice.
-First, English syllable contact clusters are the subject of many prior phonological and phonotactic analyses.
-Second, the data should be accessible to any reader.
+\subsection{The present day}
 
-% lexicon
-The \emph{lexicon} has many senses: throughout, I use the term to refer to the set of underlying representations \citep[][269]{LANGUAGE}.
-I depart from \citeauthor{LANGUAGE}'s characteriztion of the lexicon in two ways.
-First, I assume that this set is merely a reflection of individual speakers' knowledge of their language, realized in the human; it has no further reality.
-Second, in using the term, I do not intend to refer to any knowledge of morphological irregularities, which also reside in the lexicon for \citeauthor{LANGUAGE} (ibid., p.~273).
+\section{English syllable contact}
 
-\section{Phonotactic gaps and the null hypothesis}
-
-Before studying
-
-\section{History}
-
-A major result of the American structuralists was the finding that the set of sound sequences found in morphs and across morpheme boundaries are different.\footnote{The discussion in this paragraph is based on \citealt[][267]{Anderson1985}.} \citet{Bloomfield1930} proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries.
-\citeauthor{Bloomfield1930} notes that [\c{c}] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [ku\lm x\textschwa n] `cake' $\sim$ \emph{Kuhchen} [ku\lm\c{c}\textschwa n] `cow-let'.
-\citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf.
-\emph{Kuh} `cow').
-Under the assumption that [x] occurs in this position (before a back vowel) only when this preceding vowel belongs to the same word or morpheme, [\c{c}] and [x] can be cast as allophones of the same phoneme (likely /\c{c}/, the basic variant).
-A natural consequence of this view is that generalizations about sound sequences found within morphemes are necessarily different within and across morpheme boundaries,
-and thus a distributional analysis must take morphology into account, as argued forcefully by \citet{Pike1947b}.
-\citeauthor{Pike1947b} discusses limitations on possible vowel sequences in Mixteco monomorphs, but it is easy to find exceptions to these generalizations when a morpheme boundary intervenes in the short Mixteco text analyzed in \citealt{Pike1944}.
-
-\ex Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}: \\ 
-\begin{tabular}{r l l l} %\toprule
-   & MSC & complex word \\ %\midrule
-a. & *{C}a{C}e & k\'a-\textsuperscript{n}dee & `kept \ldots inside' \\
-b. & *{C}\textipa{@}{C}e & n\`i-k\textipa{\`@b@-de} & `who entered'        \\
-c. & *{C}e{C}i & te-n\'i-ke\textsuperscript{n}da & `was walking         \\
-d. & *{C}i{C}e & te-n\`i-kee-t\`\textschwa & `and went away'      \\ %\toprule
-%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
-%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
-\end{tabular} \xe
-
-\noindent
-\citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, and this idea was further developed by \citet{Harris1955}, who proposes that this mechanism allows the linguist to identify morpheme boundary.
-
-While the above work is widely known for introducing ``grammatical'' (which in the sense current at the time just meant ``supraphonemic'') components into the theory of phonology, albeit indirectly, similar (and arguably more sophisticated) intuitions about the interaction between phonotactic generalizations and morpheme boundaries were made by members of the Prague Circle at roughly the same time, a fact which has not been noted to my knowledge.
-\citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture.
-For instance, most Russian infinitives end with a final /t'/.
-The reflexive variant of these verbs is marked with a final /-sa/, which feeds a general process of regressive voicing assimilation, and this /\ldots t\textceltpal -s\ldots/ juncture triggers the loss of palatalization.
-However, palatalization of a stem-final labial or coronal---including /t'/, as in ?c ---is preserved in reflexive imperatives, which are formed by attaching the reflexive /-sa/ to the bare stem.\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which makes use of a highly abstract phonemic transcription.
-The broad phonetic transcriptions below are the result of consultations with N linguistically na\"ive native Russian speakers.
-Stress and palatalization are marked regardless of whether they are constrastive in that position.}
-
-\ex Russian reflexives \citep[after][]{Jakobson1932}: \\
-\begin{tabular}{l l l l} %\toprule
-   &  infinitive & imperative \\ %\midrule
-a. & sl\textceltpal av\textceltpal its\textschwa         & sl\'af\textceltpal s\textschwa            & `be glorious'    \\
-   & upr\'amits\textschwa         & upr\'am\textceltpal s\textschwa          & `be stubborn'    \\
-b. & kr\'as\textceltpal its\textschwa         & kr\'as\textceltpal s\textschwa            & `put on makeup'  \\
-   & \v{z}\'arits\textschwa       & \v{z}\'ar\textceltpal s\textschwa         & `roast'          \\ 
-c. & z\textschwa b\'yts\textschwa & z\textschwa b\'ut\textceltpal s\textschwa & `forget' \\ %\bottomrule
-%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
-\end{tabular} \xe
-
-Unlike in \citeauthor{Bloomfield1930}'s analysis of German, the cluster of palatal coronal followed by /s/ has an intervening morpheme boundary in both infinitive and imperative.
-\citeauthor{Jakobson1932} proposes that the two /t'-sa/ junctures in c, in fact different, with the imperative not giving rise to the loess of palatalization.
-Obvious parallels of this more powerful notion can be found in Lexical Phonology or the model developed by \citet{Halle1987}, to name a few.
-A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.
-
-Arguably, this more powerful system might be appropriate for \citeauthor{Bloomfield1930}'s data, if it is the case that the final \emph{-en} in \emph{Kuchen} is also a morpheme (perhaps related to \emph{K\"uche} `kitchen; cuisine').
+The rest of this chapter concerns English syllable contact clusters in monomorphs.
+This phenomenon is chosen for two reasons: first, English syllable contact clusters are the subject of many prior phonological and phonotactic analyses, and second, the data is very accessible.
 

### 2011-12-01 abs.tex
went to XeLaTeX

This dissertation considers two types of ``gaps'' in the lexicon. In both cases, it is argued that current generative models can be modified minimally to account for the apparent overgeneration of gapped structures. The first part of the dissertation considers the underlying phonological structure of morphs, specifically medial consonant clusters in English. A number of prior studies have argued that the grammar must have a separate ``phonotactic'' module to account

at the level of prosodic word. In both cases, it is argued that the i

### 2011-12-01 ack.tex
daily commit
went to XeLaTeX

@@
 The subject of the first chapter was inspired by the mentoring of Jennifer Cole.
-Important contributions to this dissertation were made by Stephen Anderson, who graciously agreed to serve on my committee, Michael Wagner, who gave me a goal, and Morris Halle, who taught me to never apologize.
+Important contributions to this dissertation were made by Stephen Anderson, who graciously agreed to serve on my committee, Michael Wagner, who gave me a goal to work towards, and Morris Halle, who taught me to never apologize.
 
@@
 Among non-linguists, my good friend Carly Fisher stands out for her unflagging support.
+Among parents, mine stand out for having more faith in their son than he could ever merit.

### 2011-12-01 clusters.tex
went to XeLaTeX

@@
+ontogeny reveals ontology
+
+\chapter{Structural and accidental phonotactic gaps in the lexicon}
+
+Some (admittedly similar)
+
+Some evidence that this is active comes from a study by \citet[]{Wright1975}.
+\citeauthor{Wright1975} taught several members of a ``neighborhood gang'' a few words from an artificial language, and then asked them to repeat the words to each other in a game of ``telephone''.
+Crucially
+
+A particularly
+
+\ex Artificial language repetitions \citep[from][]{Wright1975}: \\
+\begin{tabular}{l l l l l l} %\toprule
+%a. & [go\textupsilon np] & $>$ & [gu\lm mp] \\
+%[g\textturnv np]  
+%b. & [\dh \textturnv\textipa{N}d] & $>$ & [\dh \textopeno\textipa{N}g\textschwa] \\ % $>$ [t\textopeno\textipa{N}g\textschwa] 
+%c. & [\textdyoghlig umg] & $>$ & [\textdyoghlig u\textipa{N}g\textschwa] \\
+\end{tabular} \xe
+
+\noindent
+While a number of other adaptations there 
+[np] becomes [mp],
+
+%[\textipa{N}d] becomes [nd], and [mg] becomes [\textipa{N}g].
+
+[ɳd]
+
+t
+If this task is viewed as something like loanword adaptation, then there are certainly good reasons to think that extragrammatical pressures \citep{Dupoux1999,Ussishkin2003,Peperkamp2008} could be responsible for what is observed.
+That said, it is certainly noteable that
+
+\footnote{In this latter view, it is interesting that the (\lastx c) final nasal-[g] cluster is preserved. In standard }
+
+\citep{Schleef2011}
+
+%[\textipa{N}g] is certainly
+
+though it can be found in the speech fo
+
+There is good phonological eivd
+
+In this latter view, it is interesting that the finIn this latter view, it is interesting that the finIn this latter view, it is interesting that the final [g] is preserved.
+
+As \citeauthor{Wright1975} notes, it is particularly interesting that
+
+It is particularly interesting that 
+%included a stimulus with [\textipa{N}] as a control, since
+
+There is a lengthy literature arguing for a phonotactic component to speakers' phonological knowledge.
+The fundamental argument pursued in this chapter is that many of these studies are misguided, and fail to take into account a powerful null hypothesis.
+
 The \emph{lexicon} has many senses: throughout, I use the term to refer to the set of underlying representations \citep[][269]{LANGUAGE}, but I depart from \citeauthor{LANGUAGE}'s understanding by further adopting the assumption that this set is merely a reflection of individual speakers' knowledge of their language, realized in the human; it has no further reality.
 
-\section{The null hypothesis}
+\section{Preliminaries}
+
+\subsection{Some definitions}
+
+\subsection{The null hypothesis}
+
+\subsubsection{History of the null}
 
@@
    &  infinitive & imperative \\ %\midrule
-a. & [\textprimstress sl\textceltpal av\textceltpal its\textschwa]         & [\textprimstress slaf\textceltpal s\textschwa]            & `be glorious'    \\
-   & [u\textprimstress pramits\textschwa]         & [u\textprimstress pram\textceltpal s\textschwa]          & `be stubborn'    \\
-b. & [\textprimstress kras\textceltpal its\textschwa]         & [\textprimstress kras\textceltpal s\textschwa]           & `put on makeup'  \\
-   & [\textprimstress\textyogh arits\textschwa]       & [\textprimstress\textyogh ar\textceltpal s\textschwa]         & `roast'          \\ 
-c. & [\textprimstress z\textschwa byts\textschwa] & [\textprimstress z\textschwa but\textceltpal s\textschwa] & `forget' \\ %\bottomrule
+a. & [slav\textceltpal its\textschwa]         & [slaf\textceltpal s\textschwa]            & `be glorious'    \\
+   & [pramits\textschwa]         & [upram\textceltpal s\textschwa]          & `be stubborn'    \\
+b. & [kras\textceltpal its\textschwa]         & [kras\textceltpal s\textschwa]           & `put on makeup'  \\
+   & [\textyogh arits\textschwa]       & [\textyogh ar\textceltpal s\textschwa]         & `roast'          \\ 
+c. & [z\textschwa byts\textschwa] & [z\textschwa but\textceltpal s\textschwa] & `forget' \\ %\bottomrule
 %   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
@@
 \midrule
-\buf[pemez]   & $30$ & [pemaz]     & $2$  \\
-\buf[teriz]   & $28$ & [ter\i z]   & $3$  \\
-\buf[tokaz]   & $26$ & [tokez]     & $6$  \\
-\buf[tipez]   & $24$ & [tipaz]     & $8$  \\
-\buf[ter\"uz] & $19$ & [teruz]     & $13$ \\
+\emph{pemez}  & $30$  & \emph{pemaz}     & $2$  \\
+\emph{teriz}   & $28$ & \emph{ter\i z}   & $3$  \\
+\emph{tokaz}   & $26$ & \emph{tokez}     & $6$  \\
+\emph{tipez}   & $24$ & \emph{tipaz}     & $8$  \\
+\emph{ter\"uz} & $19$ & \emph{teruz}     & $13$ \\
 \bottomrule
@@
 \end{table}
+% c = 127, d = 32, p = 5.233e-29
 
@@
 \midrule
-\buf[tafuz] & $21$ & [taf\i z] & $11$ \\
-\buf[pamuz] & $17$ & [pam\i z] & $15$ \\
-\buf[tapuz] & $17$ & [tap\i z] & $15$ \\
-\buf[mavuz] & $16$ & [mav\i z] & $16$ \\
-\buf[tabuz] & $16$ & [tab\i z] & $16$ \\
+\emph{tafuz} & $21$ & \emph{taf\i z} & $11$ \\
+\emph{pamuz} & $17$ & \emph{pam\i z} & $15$ \\
+\emph{tapuz} & $17$ & \emph{tap\i z} & $15$ \\
+\emph{mavuz} & $16$ & \emph{mav\i z} & $16$ \\
+\emph{tabuz} & $16$ & \emph{tab\i z} & $16$ \\
 \bottomrule
@@
 \end{table}
+% c = 87, d = 73, p = 0.101
 
 Here is it clear that the preferences, while going in the expected direction, are far smaller ($\gamma = 0.088$).
-Further, only one of five pairs, \emph{tafuz} $\sim$ \emph{taf\i z}, matches the stronger preferences seen with backness harmony.
+There are ties for two of the lexical items, and only one of the five pairs, \emph{tafuz} $\sim$ \emph{taf\i z}, matches the stronger preferences seen with backness harmony.
+If we remove this pair, $\gamma = 0.032$.
 It might be possible to though any failure to find significance here should surely be attributed to insufficient statistical power, not evidence for the null hypothesis.
@@
 The imperative, \emph{l'\'akka},
+
+Degemination
+
+\citet[][20]{Harris1994}

### 2011-12-01 lexinsertion.tex
went to XeLaTeX

\citep{Pater2000,Pater2005b,Pater2005a,Pater2006,Pater2009,Becker2009a}

``the smaller set of indexed morphemes could be built in'' \citep[][?]{Pater2009}

This fails in the case of the

\footnote{This observation is due to Charles Yang (personal communication).}

There are more seriosu failures:

if as standardly assumed 
(CITATION)
the initial state of the child's grammar ranks all markedness constraints above faithfulness constraints, then

it is hard to see how even regular inflection is applied to nonce words.

Consider the case of 
/bl\textsci k/

Consider separately a nonce word such 
\emph{vuwt} [v\textupsilon t]

/t-z/
a marked contrast in voicing, so clearly markedesss concerns favor [ts] is favored. As it happens, however, word-final [ts] is a marked structure in Latin,
where it is resolved by deleting the final consonant.

\ex \label{para} Latin plosive-final third declension nouns \citep[from][]{Gorman2011b}:

\begin{tabular}{ r l l l l}
       & \textsc{UR}  & {nom.sg.} & {gen.sg.}         & {gloss} \\ \cmidrule{2-5}
    a. & /stirp-/   & \emph{stirps}   & \emph{stirpis}  & `scion'  \\
       & /urb-/     & \emph{urbs}\footnote{}   & \emph{urbis}    & `city'   \\
       & /fak-/     & \emph{fax}   & \emph{facis}    & `torch' \\
%       & /re\lm g-/ & \emph{r\emac x} & \emph{r\emac gis}  & `king' \\
%    b. & /ko\lm t-/ & \emph{c\omac s}    & \emph{c\omac tis}  & `flint' \\
%       & /ped-/     & \emph{p\emac s}    & \emph{pedis}    & `foot'
\end{tabular} \xe

\footnotetext[\value{footnote}]{This is also written \emph{urps} in Classical texts, and it is generally accepted that the surface form was [urps] %\citep[21]{Lindsay1895} 
\citep[][137f.]{Neue1866}.}

### 2011-12-05 abs.tex
Went all Unicode, all the time.

@@
-This dissertation considers two types of ``gaps'' in the lexicon.
-In both cases, it is argued that current generative models can be modified minimally to account for the apparent overgeneration of gapped structures.
+This dissertation considers two types of gaps in the lexicon.
+In both cases, it is argued that current generative models can be modified minimally to account for the apparent overgeneration of missing structures.
+
 The first part of the dissertation considers the underlying phonological structure of morphs, specifically medial consonant clusters in English.
-A number of prior studies have argued that the grammar must have a separate ``phonotactic'' module to account
+Many prior studies have argued for morpheme structure constraints, generalizations that account for the underlying shape of morphs.
+It is shown, however, that many of the gaps in the inventory of medial consonant clusters are due to sampling from sparse data, cannot be reliably predicted by current computational models of phonotactic learning, and can thus be regarded as accidental gaps.
+In contrast, phonological alternations induce reliable structural gaps in the inventory of medial clusters.
+This suggests that language-specific phonotactics can be derived without a separate phonotactic module (\citealt{Halle1962}).
+%This model is validated with a meta-analysis of 5 English wordlikeness studies.
 
-at the level of prosodic word. In both cases, it is argued that the i
+The second part considers the inflectional patterns in a number of languages. There are many languages in which a certain class of words unexpectedly fails to 
+Hungarian and Kinande.
+
+The use of elsewhere rules to account for allomorphy fail to
+
+This model is validated using data on inflectional gaps in English, Spanish, Russian, Swedish, and Greek.
+
+% Keywords: Gaps, sparsity, Zipf's Law, acquisition, phonotactics, wordlikeness, syllable contact, inflectional gaps, defectivity, lexical exceptions
+% Languages: Turkish, English, Latin, Hungarian, Kinande, Spanish, Russian, Swedish, Greek

### 2011-12-05 ack.tex
Went all Unicode, all the time.

@@
 I would like to thank my fellow students at Penn, but especially my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody.
-Elsewhere, I have been shown a great deal of kindness by so many, but especially Morgan Sonderregger and Maryam Bakht.
+Elsewhere, I have been shown a great deal of kindness by so many, but especially Maryam Bakht.
 Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document.
-Among non-linguists, my good friend Carly Fisher stands out for her unflagging support.
-Among parents, mine stand out for having more faith in their son than he could ever merit.
+Katherine McEldoon provided a useful review of numerical cognition research.
+Carly Fisher gave me her unflagging support.

### 2011-12-05 clusters.tex
Went all Unicode, all the time.

@@
+\chapter{Structural and accidental phonotactic gaps in the lexicon}
+
+\citet[][111]{Vennemann1972}
+\citet[][94f.]{Shibatani1973}
+\citet[][84]{Hooper1973}
+\citet{Kaye1974}
+\citet{Clayton1976}
+
+Lang Acq
+
+\citet{White2008}
+
+%\citeauthor{Hooper1973} (ibid., p.~84) provides an example from the analysis of nasal place assimilation in Spanish given by \citet[][16]{Harris1969}
+
+Odawa provides
+
+peripherality
+\citet[][8f.]{LGB}
+\citep{Duanmu2008}
+\citet{Borowsky1989}
+
+\section{Evidence in phonotactics}
+
+\subsection{Evidence from acquisition}
+
 ontogeny reveals ontology
 
-\chapter{Structural and accidental phonotactic gaps in the lexicon}
+\subsection{Evidence from wordlikeness judgements}
+
+\subsection{Evidence from lexical statistics}
+
+\citet{Berkley1994a,Berkley1994b}
+
+\begin{quotation}
+\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
+\citep[][]{Fischer-Jorgensen1952}
+\end{quotation}
+
+Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters two years later.
+
+\begin{quotation}
+Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters.
+The material will never be complete.
+It will always contain accidental gaps \ldots~partly because some clusters by pure chance do not occur in the vocabulary.
+\citep[][30]{Vogt1954}
+\end{quotation}
+
+\begin{quotation}
+\ldots~in languages that allow relatively complex consonant clusters, the occurring clusters represent only a small percentage of all theoretically possible consonant clusters.
+\citep[][94]{Hooper1973}
+\end{quotation}
+
+Similar observations have also been made in recent years, for instance in the study of consonant co-occurrence restrictions in Arabic performed by \citeauthor{Frisch2004}.
+
+\begin{quotation}
+In describing and explaining the phonotatic patterns in any language, phonologists recognize a distinctino between accidental and systematic gaps in the lexcion.
+In practice, this distinction has often been applied in an intuitive and \emph{post-hoc} fashion.
+A gap is taken to be systematic if it belongs to a natural class of examples according to the current theory of the investigator.
+Otherwise, it is viewed as accidental.
+\citep[][180]{Frisch2004}
+\end{quotation}
+
+There are also a number of prior studies \citet{Kaeding1898,Weiss1961,Sigurd1968,Tambovtsev2007} which note that phonemes and phoneme sequences are sparsely distributed: that is, there are a few very frequent phonemes or sequences, and many infrequent phonemes or sequences.
 
@@
 %[g\textturnv np]  
-%b. & [\dh \textturnv\textipa{N}d] & $>$ & [\dh \textopeno\textipa{N}g\textschwa] \\ % $>$ [t\textopeno\textipa{N}g\textschwa] 
-%c. & [\textdyoghlig umg] & $>$ & [\textdyoghlig u\textipa{N}g\textschwa] \\
+%b. & [\dh \textturnv\textipa{N}d] & $>$ & [\dh \textopeno\textipa{N}gə] \\ % $>$ [t\textopeno\textipa{N}gə] 
+%c. & [\textdyoghlig umg] & $>$ & [\textdyoghlig u\textipa{N}gə] \\
 \end{tabular} \xe
@@
 t
-If this task is viewed as something like loanword adaptation, then there are certainly good reasons to think that extragrammatical pressures \citep{Dupoux1999,Ussishkin2003,Peperkamp2008} could be responsible for what is observed.
+If this task is viewed as something like loanword adaptation, then there are certainly good reasons to think that extragrammatical pressures \citep{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008} could be responsible for what is observed.
 That said, it is certainly noteable that
@@
 In his history of phonology, \citet[][267]{Anderson1985} credits this innovation to \citet{Bloomfield1930}, who proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries.
-\citeauthor{Bloomfield1930} notes that [\c{c}] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [ku\lm x\textschwa n] `cake' $\sim$ \emph{Kuhchen} [ku\lm\c{c}\textschwa n] `cow-let'.
+\citeauthor{Bloomfield1930} notes that [ç] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [ku\lm xən] `cake' $\sim$ \emph{Kuhchen} [ku\lm çən] `cow-let'.
 \citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf.
 \emph{Kuh} `cow').
-Under the assumption that [x] occurs in this position (before a back vowel) only when this preceding vowel belongs to the same word or morpheme, [\c{c}] and [x] can be cast as allophones of the same phoneme (likely /\c{c}/, the basic variant).
+Under the assumption that [x] occurs in this position (before a back vowel) only when this preceding vowel belongs to the same word or morpheme, [ç] and [x] can be cast as allophones of the same phoneme (likely /ç/, the basic variant).
 
@@
 %a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
-%b. & *{C}\textipa{@}{C}e & [n\`i-k\`\textschwa b\textschwa-de] & `who entered'        \\
+%b. & *{C}\textipa{@}{C}e & [n\`i-k\`ə bə-de] & `who entered'        \\
 %c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
-%d. & *{C}i{C}e & [te-n\`i-kee-t\`\textschwa] & `and went away'      \\ %\toprule
-a. & *{C}a{C}e & [k\'a\textsuperscript{n}dee] & `kept \ldots inside' \\
-b. & *{C}\textipa{@}{C}e & [n\`ik\`\textschwa b\textschwa de] & `who entered'        \\
-c. & *{C}e{C}i & [ten\'i ke\textsuperscript{n}da] & `was walking         \\
-d. & *{C}i{C}e & [ten\`i keet\`\textschwa] & `and went away'      \\ %\toprule
+%d. & *{C}i{C}e & [te-n\`i-kee-t\`ə] & `and went away'      \\ %\toprule
+a. & *{C}a{C}e & [ká\textsuperscript{n}dee] & `kept \ldots inside' \\
+b. & *{C}ə{C}e & [nìk\`əbəde] & `who entered'        \\
+c. & *{C}e{C}i & [teníke\textsuperscript{n}da] & `was walking         \\
+d. & *{C}i{C}e & [tenìkeet\`ə] & `and went away'      \\ %\toprule
 %e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
@@
 \citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture.
-For instance, most Russian infinitives end with a final /t'/.
-The reflexive variant of these verbs is marked with a final /-sa/, which feeds a general process of regressive voicing assimilation, and this /\ldots t\textceltpal -s\ldots/ juncture triggers the loss of palatalization.
+For instance, most Russian infinitives end with a final /t\pal/.
+The reflexive variant of these verbs is marked with a final /-sa/, which feeds a general process of regressive voicing assimilation, and this /\ldots t\pal-s\ldots/ juncture triggers the loss of palatalization.
 However, palatalization of a stem-final labial or coronal---including /t'/, as in (\nextx c)---is preserved in reflexive imperatives, which are formed by attaching the reflexive /-sa/ to the bare stem.\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses a highly abstract phonemic transcription.
@@
    &  infinitive & imperative \\ %\midrule
-a. & [slav\textceltpal its\textschwa]         & [slaf\textceltpal s\textschwa]            & `be glorious'    \\
-   & [pramits\textschwa]         & [upram\textceltpal s\textschwa]          & `be stubborn'    \\
-b. & [kras\textceltpal its\textschwa]         & [kras\textceltpal s\textschwa]           & `put on makeup'  \\
-   & [\textyogh arits\textschwa]       & [\textyogh ar\textceltpal s\textschwa]         & `roast'          \\ 
-c. & [z\textschwa byts\textschwa] & [z\textschwa but\textceltpal s\textschwa] & `forget' \\ %\bottomrule
+a. & [slav\pal itsə] & [slaf\pal sə] & `be glorious'    \\
+   & [pramitsə]              & [upram\pal sə]          & `be stubborn'    \\
+b. & [kras\pal itsə]         & [kras\pal sə]           & `put on makeup'  \\
+   & [ʒaritsə]       & [ʒar\pal sə]         & `roast'          \\ 
+c. & [zəbytsə] & [zəbut\pal sə] & `forget' \\ %\bottomrule
 %   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
@@
 
-Both forms in (\lastx c) contain a /t\textceltpal-s/ juncture, though only one is faithful.
+Both forms in (\lastx c) contain a /t\pal-s/ juncture, though only one is faithful.
 \citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization.
@@
 
-Arguably, this more powerful system might be appropriate for \citeauthor{Bloomfield1930}'s data, if it is the case that the final \emph{-en} in \emph{Kuchen} is also a morpheme (perhaps related to \emph{K\"uche} `kitchen; cuisine').
+Arguably, this more powerful system might be appropriate for \citeauthor{Bloomfield1930}'s data, if it is the case that the final \emph{-en} in \emph{Kuchen} is also a morpheme (perhaps related to \emph{Küche} `kitchen; cuisine').
 There are obvious parallels between \citeauthor{Jakobson1932}'s analysis and notions like readjustment in \emph{SPE}, or cyclic rules in lexical phonology or the \citet{Halle1987} model.
@@
 & [$-$\textsc{round}] & [$+$\textsc{round}] & [$-$\textsc{round}] & [$+$\textsc{round}] \\
-\buf[$+$\textsc{high}] & \emph{i} [i]        & \emph{\"u} [y]      & \emph{\i} [\textturnm] & \emph{u} [u] \\
-\buf[$-$\textsc{high}] & \emph{e} [e]        & \emph{\"o} [\o]     & \emph{a} [\textschwa]  & \emph{o} [o] \\
+\buf[$+$\textsc{high}] & \emph{i} [i]        & \emph{ü} [y]      & \emph{ı} [ɯ] & \emph{u} [u] \\
+\buf[$-$\textsc{high}] & \emph{e} [e]        & \emph{ö} [ø]     & \emph{a} [ə]  & \emph{o} [o] \\
 \end{tabular}
@@
 
-%into four [$+$\textsc{back}] vowels \emph{u}, \emph{\i}, \emph{o}, \emph{a}, denoting [u, \textturnm, o, a] and four [$-$\textsc{back}] vowels \emph{i}, \emph{\"u}, \emph{e}, \emph{\"o} denoting [i, y, e, \o].
+%into four [$+$\textsc{back}] vowels \emph{u}, \emph{ı}, \emph{o, \emph{a}, denoting [u, ɯ, o, a] and four [$-$\textsc{back}] vowels \emph{i}, \emph{ü}, \emph{e}, \emph{ö} denoting [i, y, e, ø].
 
-Native roots generally contain only [$+$\textsc{back}] vowels (e.g., \emph{g\"orev} `duty') or [$-$\textsc{back}] vowels (e.g., \emph{odun} `wood'), but loanwords need not conform to this generalization (e.g., \emph{sosis} `sausage').
+Native roots generally contain only [$+$\textsc{back}] vowels (e.g., \emph{görev} `duty') or [$-$\textsc{back}] vowels (e.g., \emph{odun} `wood'), but loanwords need not conform to this generalization (e.g., \emph{sosis} `sausage').
 Despite this latter fact, there is some evidence that this constraint on roots is an induced one.
@@
 a. & \emph{it}    & \emph{iti}      & \emph{itin}      & `dog'  \\
-%a. & \emph{g\"ul} & \emph{g\"ul\"u} & \emph{g\"ul\"un} & `rose' \\
-c. & \emph{k\i z} & \emph{k\i z\i}  & \emph{k\i z\i n} & `girl' \\
+%a. & \emph{gül} & \emph{gülü} & \emph{gülün} & `rose' \\
+c. & \emph{kız} & \emph{kızı}  & \emph{kızın} & `girl' \\
 %c. & \emph{tuz}   & \emph{tuzu}     & \emph{tuzun}     & `salt' \\
@@
 Similar facts are indicated by the rest of the declensional paradigms. It is possible to analyze the forms of the acc.sg. 
-(\emph{i}, \emph{\i})
-%(\emph{\"u}, \emph{u})
-%(\emph{i}, \emph{\"u}, \emph{\i}, \emph{u})
+(\emph{i}, \emph{ı})
+%(\emph{ü}, \emph{u})
+%(\emph{i}, \emph{ü}, \emph{ı}, \emph{u})
 and gen.sg. 
-(\emph{in}, \emph{\i n})
-%(\emph{\"un}, \emph{un})
-%(\emph{in}, \emph{\"un}, \emph{\i n}, and \emph{un}) 
+(\emph{in}, \emph{ı n})
+%(\emph{ün}, \emph{un})
+%(\emph{in}, \emph{ün}, \emph{ı n}, and \emph{un}) 
 without suppletion, using the following phonological rule:
@@
 The acc.sg.
-suffix is then [$+$\textsc{high}] (and for these cases, [$-$\textsc{round}]); there is no evidence bearing on whether it is one of /i, \i/ or simply underspecified for [back].
+suffix is then [$+$\textsc{high}] (and for these cases, [$-$\textsc{round}]); there is no evidence bearing on whether it is one of /i, ı/ or simply underspecified for [back].
 The gen.sg.
@@
 For instance, the acc.sg.
-of \emph{av} `hunt' is not *\emph{avu}, but rather \emph{av\i}.
+of \emph{av} `hunt' is not *\emph{avu}, but rather \emph{avı}.
 This suggests that this constraint cannot be considered an induced restriction.
@@
 %%\emph{tipez}   & $21$ & \emph{tipaz} & $1$ & $1$ \\ %back
-%\emph{ter\"uz} & $20$ & \emph{teruz} & $1$ & $2$ \\ %back
-%%t\"or\"uz & $19$ & t\"oriz & $1$ & $3$ \\ %round
-%\emph{teriz}   & $23$ & ter\i z & $0$ & $0$ \\ 
+%\emph{terüz} & $20$ & \emph{teruz} & $1$ & $2$ \\ %back
+%%törüz & $19$ & töriz & $1$ & $3$ \\ %round
+%\emph{teriz}   & $23$ & terı z & $0$ & $0$ \\ 
 %% tamuz   & $3$  & tam\I z & $16$ \\
@@
 %\midrule
-%tamuz & $3$ & tam\i z & $16$ & $4$ \\
-%tafuz & $3$ & taf\i z & $17$ & $3$ \\
-%tapuz & $7$ & tap\i z & $9$  & $7$ \\
-%tavuz & $9$ & tav\i z & $4$  & $10$ \\
-%tabuz & $5$ & tab\i z & $12$ & $6$ \\
+%tamuz & $3$ & tamı z & $16$ & $4$ \\
+%tafuz & $3$ & tafı z & $17$ & $3$ \\
+%tapuz & $7$ & tapı z & $9$  & $7$ \\
+%tavuz & $9$ & tavı z & $4$  & $10$ \\
+%tabuz & $5$ & tabı z & $12$ & $6$ \\
 %\bottomrule
@@
 \emph{pemez} & $30$ & \emph{pemaz} & $2$  \\
-\emph{teriz}   & $28$ & \emph{ter\i z}   & $3$  \\
+\emph{teriz} & $28$ & \emph{terız} & $3$  \\
+\emph{teriz} & $28$ & \emph{terız} & $3$  \\
 \emph{tokaz} & $26$ & \emph{tokez} & $6$  \\
 \emph{tipez} & $24$ & \emph{tipaz} & $8$  \\
-\emph{ter\"uz} & $19$ & \emph{teruz}     & $13$ \\
+\emph{terüz} & $19$ & \emph{teruz} & $13$ \\
 \bottomrule
@@
 \midrule
-\emph{tafuz} & $21$ & \emph{taf\i z} & $11$ \\
-\emph{pamuz} & $17$ & \emph{pam\i z} & $15$ \\
-\emph{tapuz} & $17$ & \emph{tap\i z} & $15$ \\
-\emph{mavuz} & $16$ & \emph{mav\i z} & $16$ \\
-\emph{tabuz} & $16$ & \emph{tab\i z} & $16$ \\
+\emph{tafuz} & $21$ & \emph{tafız} & $11$ \\
+\emph{pamuz} & $17$ & \emph{pamız} & $15$ \\
+\emph{tapuz} & $17$ & \emph{tapız} & $15$ \\
+\emph{mavuz} & $16$ & \emph{mavız} & $16$ \\
+\emph{tabuz} & $16$ & \emph{tabız} & $16$ \\
 \bottomrule
@@
 Here is it clear that the preferences, while going in the expected direction, are far smaller ($\gamma = 0.088$).
-There are ties for two of the lexical items, and only one of the five pairs, \emph{tafuz} $\sim$ \emph{taf\i z}, matches the stronger preferences seen with backness harmony.
+There are ties for two of the lexical items, and only one of the five pairs, \emph{tafuz} $\sim$ \emph{tafı z}, matches the stronger preferences seen with backness harmony.
 If we remove this pair, $\gamma = 0.032$.
@@
 \citet[][20]{Harris1994}
+
+soft-margin support vector machine \citep{Cortes1995}

### 2011-12-05 lexinsertion.tex
Went all Unicode, all the time.

@@
-\citep{Pater2000,Pater2005b,Pater2005a,Pater2006,Pater2009,Becker2009a}
+\section{The reality of inflectional gaps}
+
+Comparing the counts of the preterite and non-preterite forms of \emph{undergo} and \emph{forgo} in both corpora with the Fisher exact test demonstrates that the non-usage of a preterite of \emph{forgo} is unlikely to be accidental (BNC: $p = 3.4$\textsc{e}-31, COCA: $p = 1.7$\textsc{e}-186).
+This is even true of \emph{sightsee}, which occurs with low frequency as a verb (BNC: $p = 0.0227$).
+
+Surveys of inflectional gaps can be fonud in
+
+as surveyed in \citealt{Fanselow2002}, \citealt{Rice2009b}, and \citealt{Baerman2010b}.
+
+A characteristic of the traditional split between derivational morphology and inflectional morphology is
+
+As \citet{Marantz1997b} notes, there are many cases of productive category-changing morphology.
+In English the suffix \emph{-ness} productively forms abstract adjectives from nouns, and the suffix \emph{-ōs-} plays a similar productive role in Latin:
+
+\ex Classical Latin denominal adjectives:
+\begin{tabular}{l l l l l}
+   & nom.sg. noun  &            & nom.sg. masc. adjective & \\
+a. & \emph{macula} & `spot'     & \emph{maculōsus}       & `spotted'     \\
+b. & \emph{uentus} & `windy'    & \emph{uentōsus}        & `windy'       \\
+%c. & \emph{mo:ns}  & `mountain' & \emph{monto:sus}        & `mountainous' \\ 
+c. & \emph{lutum } & `mud'      & \emph{lutōsus}         & `muddy'       \\
+d. & \emph{ulcus}  & `ulcer'    & \emph{ulcerōsus}      & `uclerous'     \\
+\end{tabular} \xe
+
+%\ex English simple past suffixes:
+%\begin{tabular}{l l l l l l }
+%T[\textsc{Past}] & $\rightleftarrow$ & -\zr & \gap & \{$\surd$\textsc{hit}, $\surd$\textsc{sing}, \ldots\}
+%                 & $\rightleftarrow$ & -t   & \gap & \{$\surd$\textsc{leave}, $\surd$\textsc{bend}, \ldots\}
+%                 & $\rightleftarrow$ & -d   
+%\end{tabular} \xe
+
+Table \ref{fgo} provides some token counts from the 100 million word British National Corpus \citep{BNC,BYUBNC} and the 421 million word Corpus of Contemporary American English \citep{COCA}.
+
+\begin{table}
+\centering
+\begin{tabular}{l@{} l r r l@{} l r r} 
+\toprule
+    &            & BNC   & COCA   &   &             & BNC & COCA  \\ 
+\midrule
+    & undergo    & 613   & 2,757  &   & forgo       & 203 & 1,213 \\
+    & undergoes  & 123   & 594    &   & forgoes     & 4   & 89    \\
+    & undergoing & 581   & 2,464  &   & forgoing    & 24  & 232   \\
+    & underwent  & 550   & 2,440  & * & forwent     & 0   & 0     \\
+$*$ & undergoed  & 0     & 0      & * & forgoed     & 0   & 0     \\
+    & undergone  & 565   & 1,996  &   & forgone     & 64  & 190   \\ 
+\midrule
+    & foresee    & 271   & 1,020  &   & sightsee    & 4   & 55    \\
+    & foresees   & 58    & 300    &   & sightsees   & 0   & 0     \\
+    & foreseeing & 38    & 89     &   & sightseeing & 20  & 5     \\
+    & foresaw    & 132   & 431    & * & sightsaw    & 0   & 0     \\
+$*$ & foreseed   & 0     & 0      & * & sightseed   & 0   & 0     \\
+    & foreseen   & 275   & 568    & * & sightseen   & 0   & 0     \\ 
+\midrule
+    & ride       & 1,607 & 10,904 &   & stride      & 102 & 402   \\
+    & rides      & 230   & 2,495  &   & strides     & 54  & 682   \\
+    & riding     & 1,696 & 10,994 &   & striding    & 253 & 715   \\
+    & rode       & 1,079 & 5,997  &   & strode      & 641 & 1,752 \\
+$*$ & rided      & 1     & 0      & * & strided     & 0   & 2     \\
+    & ridden     & 456   & 1,303  & * & stridden    & 1   & 2     \\ 
+\bottomrule
+\end{tabular}
+\caption{Token counts the BNC and COCA for English verbs with inflectional gaps}
+\label{fgo}
+\end{table}
+
+\citep{Pater2000,Pater2005b,Pater2005a,Pater2006,Pater2009,Becker2009b}
 

### 2011-12-13 without.tex
Restructured chapter 1, and continued to move towards more Unicode.

\chapter{Life without morpheme structure constraints}

\section{Introduction}

\subsection{The claim}

\ex \textsc{Empirical principle}: // any constraint on the phonological contents of a morph operates on a domain either smaller or larger than the morph \xe

\subsection{On the desirability of a theory free of MSCs}

\subsubsection{Duplication and conspiracy}

noted first by stanly

\citet[][401]{Stanley1967}

Another major trade-off-, discussed by \citet[][?]{Harms1968}, is between rule and lexical economy; one can imagine models with both aspects.

\citet[][?]{Kisseberth1970}
and 
\citet[][?]{Kenstowicz1977}

conspiracy

\citet[][]{Shibatani1973}

\citet[][310f.]{Clayton1976} also claims that ther is no evidnece that MSCs can diachronically change without changes in the narrow phonology.

A key point, made by \citet[][301]{Clayton1976} with respect to MSCs, is that duplication amounts to claim that ``the two statements represent independent phenomena''.

% kisseberth & kenstowicz, stanley vs. hooper
\subsubsection{Computational inertness}

\begin{quotation}
Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computational inert and thus irrelevant to the input-output mappings that grammar is responsible for. \citep[][18-19]{PE}
\end{quotation}

\subsubsection{Falsifiability}          % wordlikeness

Also in Chapter 3, I will provide explicit quantitative evidence that highly statistically reliable generalizations identified by linguists may play no role in wordlikeness tasks. This does not eliminate the possibility that some constraints, either stated on the

but any study that uses statistics about the structure of underlying forms may be modeling a generalization that has no synchronic status at all.

outside the extension of a speaker's phonology \citep[][3]{Goldsmith1995}.

Generative phonologies are not constrained to be \emph{bijective},\footnote{Formally, a function $f(X) \rightarrow Y$ is bijective if only if $\forall y \in Y . \exists! x \in X . y = f(x)$. 
% SEE PMW
}

leaving the underlying form of a morph often underdetermined even when the URs of roots are known. For instance, the surface form \emph{frōns} in Latin could mean either `forehead', from /front-s/ (cf. gen.sg. \emph{frontis}) or `leaf', from /frond-s/ (cf. gen.sg. \emph{frondis}). From the surface form alone, there is no way to decide if an [ns] is from /nts/, /nds/, or perhaps /ns/ (though see \citealt{Gorman2011b}). Countless other examples of this type could easily be adduced.

%One might admit some indeterminacy in underlying form; for instance, giving \emph{frōns} a tentative UR /fronT-s/, where /T/ is an archiphoneme representing /t, d/. Even determining such degenerate URs from surface forms is itself not guaranteed to be efficiently computable \citep{Barton1987}, casting doubt on the plausibility of ``inverted phonology'' as the appropriate cognitive model \citep[see][530 for discussion]{Anderson1988}.

\subsection{Outline}

\section{A brief history of the MSC}

A brief note on the organization of this section: the subsections below correspond to chronologically ordered theoretical periods, but within each period, no attempt has been made to respect chronology.

\subsection{The structuralist MSC}

\subsubsection{Constraints on phoneme sequences}
% Vogt, Fischer-Jorgensen

\begin{quotation}
\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language. \citep[][]{Fischer-Jorgensen1952}
\end{quotation}

\noindent
Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters

\begin{quotation}
Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters. The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
\end{quotation}

\subsubsection{The morph in phonology}
% German, Russian
% Bloomfield, Jakobson

A major result of the American structuralists was the finding that generalizations which hold of the sound sequences found in underlying forms may not hold across morpheme boundaries. If this is correct, then there must be some sense in which phonology is sensitive to the presence of morpheme boundaries. In his history of phonology, \citet[][267]{Anderson1985} credits this innovation to \citet{Bloomfield1930}, who proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries. \citeauthor{Bloomfield1930} notes that [ç] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [ku\lm xən] `cake' $\sim$ \emph{Kuhchen} [ku\lm çən] `cow-let'. \citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf. \emph{Kuh} `cow'). Under the assumption that [x] occurs in this position (before a back vowel) only when this preceding vowel belongs to the same word or morpheme, [ç] and [x] can be cast as allophones of the same phoneme (likely /ç/, the basic variant).

While \citeauthor{Bloomfield1930} is known for introducing morphemic structure into phonology, a member of the Prague circle developed an arguably more sophisticated approach only two years later. \citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture. For instance, most Russian infinitives end with a final /t\pal/. The reflexive variant of these verbs is marked with a final /-sa/, which feeds a general process of regressive voicing assimilation, and this /\ldots t\pal-s\ldots/ juncture triggers the loss of palatalization. However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\nextx c)---is preserved in reflexive imperatives, which are formed by attaching the reflexive /-sa/ to the bare stem.\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses a highly abstract phonemic transcription. The broad phonetic transcriptions below are the result of consultations with N linguistically na\"ive native Russian speakers. Palatalization is marked even where not constrastive.}

\ex Russian reflexives \citep[after][]{Jakobson1932}: \\
\begin{tabular}{l l l l} %\toprule
   &  infinitive & imperative \\ %\midrule
a. & [slav\pal itsə] & [slaf\pal sə] & `be glorious'    \\
   & [upram\pal itsə]              & [upram\pal sə]          & `be stubborn'    \\
b. & [kras\pal itsə]         & [kras\pal sə]           & `put on makeup'  \\
   & [ʒar\pal itsə]       & [ʒar\pal sə]         & `roast'          \\
c. & [zəbytsə] & [zəbut\pal sə] & `forget' \\ %\bottomrule
%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
\end{tabular} \xe

\noindent
Both (\lastx c) forms contain a /t\pal-s/ juncture, though palatalization is only preserved in the imperative. \citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization. A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.

\subsubsection{The morpheme structure constraint}
% Mixteco
% Pike

A natural consequence of this view is that generalizations about sound sequences found within morphemes are necessarily different within and across morpheme boundaries, and \citet{Pike1947b} argues that the distributional analysis of sound sequences must take morphology into account. \citeauthor{Pike1947b} discusses limitations on possible vowel sequences in Mixteco monomorphs, but none of these generalizations hold across morpheme boundaries, at least according to the phonological analysis of a short Mixteco text in one of \citeauthor{Pike1944}'s earlier papers.

\ex Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}: \\
\begin{tabular}{r l l l} %\toprule
   & MSC & complex exception \\ %\midrule
%a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
%b. & *{C}\textipa{@}{C}e & [n\`i-k\`ə bə-de] & `who entered'        \\
%c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
%d. & *{C}i{C}e & [te-n\`i-kee-t\`ə] & `and went away'      \\ %\toprule
%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
a. & *{C}a{C}e & [ká\textsuperscript{n}dee] & `kept \ldots inside' \\
b. & *{C}ə{C}e & [nìk\`əbəde] & `who entered'        \\
c. & *{C}e{C}i & [teníke\textsuperscript{n}da] & `was walking         \\
d. & *{C}i{C}e & [tenìkeet\`ə] & `and went away'      \\ 
\end{tabular} \xe

\noindent
\citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a morpheme discovery routine.

\subsection{The early generativist MSC}

\subsubsection{Neutralization in phonology} 
% English (Can. raising, nasal allophones)
% Joos, Harris, Halle

\citet{SPR}
\citet[][143]{Joos1942} 
\citet[][69f., originally printed in 1951]{Harris1960}

Nasal allophone
\citet{Borowsky1986}

A minimal extension of \citeauthor{Harris1960}'s proposal gives rise to the rule ordering analysis given by \citet[][238]{Chomsky1957}.

\footnote{The discussion here is loosely adapted from \citealt{Idsardi2006}.}

\subsubsection{Lexical redundancy}

\citet[][101]{Chomsky1965}
[blɪk] vs. [bnɪk]

\citet[][101]{Chomsky1965}
\citet{Stanley1967}
\citet{SPE}

\begin{quotation}
If, on the other hand, we interpret the set of sequence structure rules as a statement of constraints on systematic phoneme sequences, then this set provides a characterization of the set of `possible' morphemes of Russian. This fact is clear if, as is quite natural, we regard a `possible'  morpheme as a form which may or may not occur in the lexicon, but whose phonological structure violates no sequence structure rule (and of course, no segment structure rule) of the language. \citep[][401]{Stanley1967}
\end{quotation}

\subsubsection{Surface constraints}

\citeauthor{Hale1965}

\begin{quotation}
To specify more exactly...which clusters can actually occur, and where, would require paraphrase of a subset of the morphophonemic rules, since the `possible clusters' in Papago, and their distributions, are automatically specified by a number of very general, independently motivated rules which impose a phonetic interpretation upon the morphophonemic representation. \citep[][297]{Hale1965}
\end{quotation}

\begin{quotation}
...we are forced to incorporate into every complete generative grammar a characterization of the distinction between admissible and inadmissible segment sequences. The fact ectively cuts the ground out from under the recent suggestion that generative grammars be supplemented with special phonological grammars, since the sole purpose of phonological grammars is to characterize the distinction between admissible and inadmissible segment sequences. \citep[][61-62]{Halle1962}
\end{quotation}

Cast in somewhat disparate theoretical terms,

\citet[][]{Postal1968} concludes, after a lengthy discussion of morpheme structure in Mohawk, that ``the restrictions on phoneme combination are completely a function of the grammar and morphophonemic rules'' (p.~212) and that ``an independent phonotactics \ldots~is in all cases useless and redundant in its entirety. It describes or accounts for not one fact which is not accounted for without it'' (p.~214).

German
\citet[][94f.]{Shibatani1973} 
\citet[][82f.]{Hooper1973}
\citet{Clayton1976}

\subsubsection{Stampean occultation}

English
\citet[][28f.]{Stampe1979}
\citet[][205f.]{Dell1973} \emph{restrictions induites}

Turkish
\citet{Zimmer1969}

Russian
\citet[][]{Anderson1974}

Odawa
\citet{Kaye1974}

Critiqued by 
\citet[][310]{Clayton1976}
\citet[][73f.]{Kiparsky1982b}

Desano
\citet{Kaye1971} and discussed at length by \citet{Leben1973}

Sanskrit
(which appears to be due to \citealt{MacEachern1999})
\citet{Gallagher2010a}

which \citet{Sag1974} traces back to the ancient Sanskrit grammarian Pāṇini. It is also, as \citeauthor{Sag1974} notes, the account given by \citet[][\S141f.]{Whitney1889} in his Sanskrit grammar. \citet[][59f.]{Hoenigswald1965} makes a similar claim.  This account is put forth by \citet{Sag1974,Sag1976} and \citet{Schindler1976}. \citet{Stemberger1980}, \citet{Borowsky1983}, and \citet{Kaye1985} present an autosegmental version of the aspirate spreading account; \citet{Janda1989} critique a number of the specifics of these accounts.

\citet[][109f.]{Zwicky1965} and \citet[][\S3.2]{Kiparsky1965} independently propose a solution with underlying diaspirates and Grassmannian dissimilation, one that is further developed by \citet{Anderson1970}, \citet{Phelps1973}, \citet{Phelps1975} and \citet{Hoard1975}.

Ofo
\citet{DeReuse1981}
\citet[][38-42]{MacEachern1999},

\subsection{MSCs beyond the segment}

\subsubsection{Suprasegmental MSCs}

Mende
\citet{Leben1973}

\subsubsection{Autosegmental MSCs}
% Greenberg, Goldsmith, McCarthy

Etung
\citet{Edmondson1966}
\citet[][132]{Goldsmith1976}

\citet{McCarthy1979}
\citet{McCarthy1981b}
\citet{McCarthy1986}

\citet{Greenberg1950},
\citet[][178]{Greenberg1950}, who observes that

``There are no Proto-Semitic roots with identical consonants in the first and second positions''

Tigrinya
\citet{Berhane1991}
\citet{Buckley1990a}
\citet{Buckley1997}
\citet{Buckley2000c}

\subsubsection{Prosodic licensing}

\citet{Ito1989}

English
\citet{McCarthy1979}
\citet[][63, 73]{Kiparsky1982b}
\citet[][19f.]{Wolf2009}

Yawelmani
%\citet{Newman1944}
\citet{Kuroda1967}
\citet{Kisseberth1969}
\citet{Kisseberth1970}

\subsection{The MSC in Optimality Theory}

\subsubsection{Richness of the base}

\subsubsection{Lexicon Optimization}

\subsubsection{Recent developments}
% lexical redunancy studies as evidence, gradience

\section{The MSC in acquisition}

\subsection{Phonotactic knowledge}

\subsubsection{Head turn preference studies}
\citet{Jusczyk1993b} % dutch

\citet{Jusczyk2002}

\subsubsection{Infant lexicons}

%an idea which is already endorsed by \citet{Shibatani1973}.

\citet{Werker1984}

\citet{Swingley2000,Swingley2002}
relatively detailed lexical entries, at least including contrastive features
by demonstrating that children recognize minor mispronunciations of familiar words

\citet{Jusczyk1995}, who find that 7.5-month-old infants can find words in sentential context (that is, running speech).

However, \citet[][563]{Johnson2001} caution that ``[e]ight-month-olds know very few words and are not likely to use lexical knowledge in segmenting speech.''

\citet{Bergelson2010}

\footnote{This should come as no surprise: \citet[][294]{Darwin1877} who writes that his seven-month-old child knew the name of his nurse.}

% My quick study
\citet{Jusczyk1994}

exs 1-2:
high prob [R IH S, G EH N, K AE Z], low prob [Y AW JH, SH AO CH, G UW SH]

ex 3:
high prob [CH AH N, T AA L S, K IY K, V EY T, M ER N], low prob [Y AW SH, SH AY B, G IY TH, TH EY JH, CH ER G]

\citet{CDI,Dale1996}
\citet{Stadthagen-Gonzalez2006}
\citet{Cortese2008}
\citet{Bird2001b}

\subsection{Phonological acquisition}

\subsubsection{Learning of alternations}

\citet{White2008}

\subsubsection{Learning of structural descriptions}

\citet{Hayes2008}

\begin{quotation}
The data presently available to us suggests that is is indeed the case that all the types of rule interaction sketched above---feeding, counterfeeding, bleeding, and counterbleeding---are required both for the application of separate rules and also \emph{for the multiple applications of a single rule}. \citep[][327, emphasis mine]{Kenstowicz1979} 
\end{quotation}

Lardil iteration
\citet{McCarthy2003a}

\citet{Anderson1969}
\citet{Howard1972}
\citet{Kenstowicz1973} 
For instance, reviewing work by \citet{Johnson1972} and \citet{Kaplan1994}

and is endorsed by \citet[][101]{Kiparsky1973c}

\citet{Gorman2011b}

For instance
\emph{targeted constraints}
\citet{Wilson2001}

\subsection{Word and morph segmentation}

%\citet{Anderson1988}

\subsubsection{In adults}
\citet{Hay2004a}

\footnote{Another concern, is that even the those forms that \citeauthor{Hay2004a} consider to be monomorphs may be morphologically complex. Some evidence for this is given in \S?.}

\subsubsection{In children}

\citet{Mattys2001a,Mattys2001b}

\citet{Onishi2002,Chambers2003}
% see gambell and yang

% Adriaans
\citet{Adriaans2010}

% word seg
\citet{Yang2004,Gambell2005,Lignos2010,Lignos2011}

\citet{Jusczyk1999c}
\citet{Johnson2001}

\citet{Gambell2005}:
if the infant extracts allophonic cues from the linguistic data, it seems...that she must have extracted a set of words to begin with. Hence, the assumptions and mechanisms required for the successful application of articulatory cues remain somewhat unclear.

%\section{Loanword adaptation}
% shibatani's rant on this
%satisfying
%\citep[e.g.,][]{Hyman1970,Danesi1985}
%those which do not \citep[e.g.,][]{Yip1993,Ito1994,Ito1995}
%with, perhaps, a relatively gradual cline of nativization \citep[e.g.,][]{Davidson1997}.

\section{Formal aspects}

\subsection{The generative architecture}

\ex \textsc{Representational Hypothesis}: UG does not allow the imposition of language-specific constraints on underlying representations \xe

\subsubsection{Underlying forms}
% rich base assumption?

\subsubsection{Complex word formation}

\subsubsection{Phonological rule application}

\subsection{Acquisition of underlying representations}

\ex \textsc{Acquisitional Hypothesis}: The LAD does not search for language-specific constraints on underlying representations \xe

\subsubsection{Stampean occultation}

insights of \citeauthor{Stampe1979}, \citeauthor{Dell1973}, and others

\subsubsection{The Alternation Condition and Lexicon Optimization}

\subsubsection{Richness of the Base}

%\citet{Prince2004}

\subsection{Falsifiability of the hypothesis}

\ex \textsc{Empirical Hypothesis}: // any constraint on the phonological contents of a morph operates on a domain either smaller or larger than the morph \xe

\subsubsection{Inventory gaps}

\subsubsection{Phonologically induced constraints}

\subsubsection{Accidental gaps}

practically a correlate of arbitrariness

\citet[][]{Fischer-Jorgensen1952} and \citet{Vogt1954}

\citet[94]{Hooper1973} claims that ``in languages that allow relatively complex consonant clusters, the occurring clusters represent only a small percentage of all theoretically possible consonant clusters.''

\begin{quotation}
In describing and explaining the phonotatic patterns in any language, phonologists recognize a distinction between accidental and systematic gaps in the lexcion. In practice, this distinction has often been applied in an intuitive and \emph{post-hoc} fashion. A gap is taken to be systematic if it belongs to a natural class of examples according to the current theory of the investigator. Otherwise, it is viewed as accidental. \citep[][180]{Frisch2004}
\end{quotation}

(and unfortunately, the approach taken by \citeauthor{Frisch2004}).

Surely it is the case that if there were no more phonological constraints, there would be no

### 2012-02-16 2.1.1.tex
Restructured into smaller files, wrote much of chapter 2.

% 2.1.1: Data sources

The English CELEX database \citep{CELEX},

\citet{Aronoff1976}

\citet{Harley2009}

\citet{Taft1975} ?
\citet{Taft1981} ?
\citet{Taft2004a} ?

\citet{Halle1962}

### 2012-02-16 2.1.2.tex
Restructured into smaller files, wrote much of chapter 2.

% 2.1.2: Syllabification

It is necessary to split medial consonant clusters into the codas and onsets that make them up. While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic. For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in RP \citep[][294]{AOE2}.}
This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language. Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which contrast \emph{chemistry} and \emph{ministry}.

An automated syllabification system was constructed and applied to medial clusters in the CELEX data. This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.  There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification. The technique used is a variation on the theme of onset maximization \citep[42f.]{Kahn1976}, which parses word-medial clusters so that the onset is as large as possible. Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty. However, in words like \emph{mi}[n.str]\emph{el}, the [nstr] cluster does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.

There is one case where unchecked maximization of medial onsets produces incorrect syllabifications. When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}. Onset maximization makes the right predictions for words like \emph{propaganda}, where the medial cluster cannot be an onset, but \citet[][55]{Harris1994} notes that it makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty. The approach adopted here is to first apply onset maximization, then to reassign the first consonant of a medial onset to the preceding coda if it is immediately preceded by a stressed lax vowel.

If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables. For this reason, it is assumed here that affricates are single segments for the purposes of syllabification. This assumption is motivated by the fact that affricates pattern with simple segments in many languages; for instance, the only word-initial consonant clusters in Classical Nahua are affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.

%In fact, [t.ʃ, d.ʒ] clusters are not found in simplex English words, despite the fact that that affricates occur as medial onsets in clusters (e.g., \emph{tru}[n.tʃ]\emph{eon}, \emph{so}[l.dʒ]\emph{er}).

### 2012-02-16 2.1.3.tex
Restructured into smaller files, wrote much of chapter 2.

% 2.1.3: Inventory considerations

The syllables derived by the above procedure must further be divided into onset, nucleus, and coda, and mapped to underlying representation. In most cases, this is trivial: the nucleus is a pure vowel and can be easily separated from the consonants, which surface faithfully. However, the status of several surface phones in the syllable and the relationship of surface segments to underlying forms is not always obvious. The treatment of those segments is described and motivated below.

\subsubsection{The velar nasal}
\label{velarnasal}

There is a long-standing debate concering whether English [ŋ] is a phoneme in its own right, demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}. The strongest piece of evidence for the pure allophonic analysis (and an attentant process of /g/-\textsc{Deletion}) is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. \citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is adopted here and formalized below in \S\ref{cnpasection}.
%\citealt[][62]{Halle1985a}),

\subsubsection{[j] onglides}

I assume that the front onglide is not part of the onset except in the case where the onset would ortherwise be null (e.g., \emph{jun}[j]\emph{or}). %\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}. However, when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm]. This defective distribution of vowels following [j] follows if the onglide in this context is the first component of a phonological diphthong, and thus part of the nucleus (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}). External support for this account comes from the fact that in words like \emph{spew}, [ju\lm] may act as a ``unit'' in language games \citep{Nevins2003,Idsardi2005}, to the exclusion of the rest of preceding consonants. Finally, \citet[][42]{Clements1983} note 
that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}, in which I assume the glide is nuclear.

%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [ju\lm] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant. 
%This is the only sign that [ju\lm] shows an affiliation for the onset.

\subsubsection{[w] onglides}

The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset. Whereas the front onglide shows only limited selectivity for preceding tautosylalbic consonants \citep{Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by any vowel.
%Whereas [Cju\lm] syllables attracts stress, [kwV] syllables do not \citep[][162f.]{Davis1995}.

\subsubsection{Post-vocalic \emph{r}}

Like this study, \citet{Pierrehumbert1994} uses a dictionary that transcribes RP, in which word-medial post-vocalic \emph{r} has been lost. There are independent reasons to believe that \emph{r}-full dialects of English assign post-vocalic \emph{r} to the nucleus and is not part of syllable contact clusters in such dialects. First, \citet[][255]{Harris1994} notes that the number of vowel contrasts is greatly reduced before \emph{r} when compared to other coronal consonants. For instance, the majority of North America has lost the historical contrast between \emph{Mary}, \emph{marry}, and \emph{merry}, and one holdout, Philadelphia, is losing the the contrast between \emph{merry} and \emph{Murray} \citep[14f.]{ANAE}. Another source of evidence for the nuclear status of post-vocalic \emph{r} comes from its failure to trigger two processes in English variable phonology. \citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.

\ex /t/-\textsc{Glottalization} in British English \citep[after][195, 258]{Harris1994}: \\
\begin{tabular}{l l l@{} l}
a. & fis[t]   & * & fis[ʔ]   \\
   & mis[t]er & * & mis[ʔ]er \\
b. & par[t]   &   & par[ʔ]   \\
   & car[t]on &   & car[ʔ]on \\
\end{tabular}
\xe

\noindent
Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\nextx a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.

\ex /t, d/-\textsc{Deletion} in American English: \\
\begin{tabular}{l l l@{} l}
a. & be[lt]  &   & be[l]  \\
%   & we[ld]  &   & we[l]  \\
%   & cha[nt] &   & cha[n] \\
   & me[nd]  &   & me[n]  \\
%b. & fl[ɜ˞]  & * & fl[ɜ˞]  \\
%b. & fl[ɝt]  & * & fl[ɝ]  \\
b. & sh[ɚt]  & * & sh[ɚ] \\
%   & w[ɚd]   & * & w[ɚ]  \\
   & c[ɚd]   & * & c[ɚ]  \\
\end{tabular}
\xe

%Finally, \citet[][251]{Fromkin1973} also presents evidence that post-vocalic \emph{r} may behave as if nuclear in speech errors.

### 2012-02-16 2.2.1.tex
Restructured into smaller files, wrote much of chapter 2.

% 2.2.1: Static and derived constraints

Having described the database of English syllable contact clsuters, I now consider constraints on sequences of URs as predictors of attested and unattested contact clusters. First, I describe the statistical technique used, then apply this technique to both phonologically derived constraints and a set of ``static'' phonotactic constraints proposed by \citet{Pierrehumbert1994}.

\subsubsection{Statistical technique}

Before the dawn of generative phonology, many linguists concerned themselves with documenting co-occurrence in various languages. \citet[][28]{Vogt1954} declares that a ``phonemic description of a language \ldots should also comprise a description of the phoneme combinations that occur or may be expected to occur in the language''. Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively. In the case of \citet{Greenberg1950}'s study, the primary generalization has no exceptions, but \citeauthor{Uhlenbeck1950}'s description of co-occurrence restrictions contains many exceptions. Many years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and suggests a statistical method to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.

\citeauthor{Mester1988}'s technique, which has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English), is based on the chi-square test.

In this study, I use the Fisher exact test, which is isomorphic to this chi-square test. The chi-square test has been favored historically simply because it can be easily computed by hand, but it is equally easy to perform either test on a modern computer. Compared to the Fisher test, which is appropriate for small samples or rare events, the chi-square test has a serious defect, since it depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples \citep[e.g.,][142]{Mester1988}.

The opposite problem, difficulties scaling to large samples, is an unfortunate defect of the Monte Carlo procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010}. This technique, which is the Monte Carlo isomorph of the one-tailed chi-square test, depends on the ability to generate random (i.e., unbiased) permutations of large lists. This is much more difficult than prior researchers have realized, as I explain below; less mathematically inclined readers may wish to skip ahead.

Any permutation of a list $L$ can be thought of as a list $S$ such that

$S_i$ 
the $i$th element in the permuted list is

the permuted position of the $i$th elemnt of $L$. Since

A permutation of a list of length $N$ can be thought of a vector $S$ such that $S_i$ is a counting number denoting the length of k

as a list of $N$ unique counting numbers which are less than or equal to $N$;

$N - 1$

The problem is that even for moderately-sized lists, the number of possible permutations quickly exceeds modern computational capabilities.

To explain why this is so requires a short discussion of

The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a ``period'' of approximately $2 ^ 48$; this means that it repeats the same 281-trillion number pattern indefinitiely.

. What this means has a period of approximately $2 ^ 48$. What this means has a period of approximately $2 ^ 48$. What this means has a period of approximately $2 ^ 48$. What this means

use the Perl scripting language, which generally provides a pseudo-random number generator

Both

A list of $N$ elements has $N!$ different permutations. T

When $N$ is large, t
When $N$ is large, current computational

this greatly exceeds the abilities of the pseudorandom number generators

This often a collosal quantity

When $N$ is large, this is a collosal quantita

with a ``period'' of approximately $2 ^ 48$, i.e., it will repeat itself after

I had a chance to write out my thoughts about the periodicity issue for Kessler's Monte Carlo method, and do a bit of research into Perl resources.

%If you're lucky, you have a version of Perl that uses the drand48 library function. The period is 2 ^ 48, around 281 trillion; after 281 trillion draws, it just regenerates the same 281 trillion numbers in the same order.

%A permutation of a list of N items can be thought of as a list of N - 1 random numbers; this is N - 1 because there's only one way to "permute" a one-element list (and in your method, only one outcome for the last element left).

%Given a period P, there are only P unique sequences of random numbers. Unfortunately, even for N = 18, N!, the number of possible permutations is far larger than the period of the Perl random number generator. That means there are many random permutations that can never be generated. When N gets significantly larger, virtually all random permutations can never be reached. In your CELEX compound study, for instance, N = 4758, and thus requires a period of approximately 2 ^ 51260 according to the Stirling approximation. But Perl can only reach 2 ^ 48 of them, a tiny fraction thereof.

%I only have browsed Kessler's book, but I saw that he spoke about how hard it was to calculate the Fisher exact test. With all due respect, I think he's just wrong. It's difficult to compute a p-value via Monte Carlo test that is provably unbiased (for instance, I'm unaware of any publically-available pseudo-random number generator with a period big enough for the CELEX compounds study), but the Fisher exact test, the one-tailed variant of which is isomorphic to the shuffle test, can be computed in the blink of an eye in any good stats environment.

 which in turn requires

Since totally-identical consonants occur quite frequently,

\ex Javanese root co-occurrence (\citealp[][264]{Uhlenbeck1950}, \citealp[][139]{Mester1988}): \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
                      & attested & unattested & saturation & $p$-value \\
\midrule
Same major place      & 21       & 15         & 0.583      & \multirow{2}{*}{6.10\e{-06}} \\
Different major place & 134      & 12         & 0.918 \\
\bottomrule
\end{tabular}
\xe

\noindent
It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing. This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance ($\chi^2 = 64.525$, two-tailed $p = 9.5$\e{-16}).

\citep[see also][]{Gorman2012}
contains the counts

%\ex $ \lim_{n \rightarrow \infty} n! = 2 ^ {(n \ln{n} - n) \cdot \log_2{e}} $ \xe

\citet{Pierrehumbert1994}

Applying the syllabification algorithm to the CELEX data produces 40 medial onset types and 22 medial coda types from 161 medial consonant clusters.

It must be stressed that under current assumptions, this is independent of whether predictable prosodic representations (particularly coda and onset identity) are part of the memorized underlying representation \citep[e.g.,][]{Vaux2003}.

It is an undeniable fact that URs need to be

The derivative nature of 
W

\citet{Ito1989a,Noske1993}

The nature of 
The prosodic characteristic of

The Fisher exact test is isomorphic to the chi-square test

and has been used by several researchers \citep{Zuraw2010}.

Unlike the chi-square test, which uses an approximation which is exact in the limit (i.e., only with a hypothetical infinite sample), the Fisher exact test is robust to small sample sizes.

The opposite problem, inability to scale to large samples, is a defect of the a statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence restrictions by \citet{Martin2007,Martin2011}. This procedure is a Monte Carlo approach to t

Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis. As

least some researchers (i.e., \citealt{Mester1988})

 \citep[e.g.,]{Brown2010} have argued for phonotactic constraints giving rise to overrepresentation, so I admit this possibility here and use two-tailed tests throughout.

%\footnote{For example, \citet{Martin2007} applies this technique to a list of 4,758 noun-noun compounds. The number of permutations of a list of this length can be estimated using Stirling's approximation.

%\ex $ \lim_{n \rightarrow \infty} n! = e ^ {n \ln{n} - n} = 2 ^ {\log_2{e} (n \ln{n} - n)} $ \xe

%\noindent
%$4,758!$ is approximately $2 ^ {51260}$. The author is unaware of any system that provides 51,260 bits of entropy; and some popular programming languages, such as ANSI C, provide as few as 15 bits.}

%%% START OF SERIOUSNESS

I will first consider three well-known English phonological alternations, showing that these give rise to constraints on the sequence structure of morphs. Simple autosegmental formalizations of these processes are provided, and the Fisher exact test is used to demonstrate their effect on the English lexicon.

\subsubsection{Obstruent Voice Assimilation}

English has a process of voice assimilation which is found in two inflectional suffixes. The suffix /-d/ forms regular pasts and denominal adjectives, and /-z/ forms regular noun plurals and genitives, and the 3sg. present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, \citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][211f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}

\ex Voice assimilation in the past tense and noun plural: \vspace{6pt} \\
\begin{tabular}{l l l l l l}
a. & /næp-d/     & \arrow & [næpt]     & (cf. \emph{nab}[d], \emph{haul}[d]) \\
b. & /læp-z/     & \arrow & [læps]     & (cf. \emph{lab}[z], \emph{clan}[z])  \\
\end{tabular}
\label{ovad}
\xe

\noindent
The voiced obstruent suffixes devoice when preceded by a voiceless obstruent.
This is formalized as a rule spreading the [\textsc{Voice}] specification rightward over adjacent obstruents.

\ex \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178): \\
\xymatrix@C=6pt@R=12pt{
\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{..}[dr] & \\
\txt{C}\ar@{-}[d]             & \txt{C}\ar@{-}[d] \\
\txt{[$+$\textsc{Obstruent}]} & \txt{[$+$\textsc{Obstruent}]}}
\label{ovar}
\xe

\noindent
This rules out the possibility of surface clusters of obstruents with differing  [\textsc{Voice}] specifications. While the data in (\ref{ovad}) make it clear \textsc{Obstruent Voice Assimilation} applies in morphologically derived environments, it is not obvious that it applies to obstruent clusters in non-derived environments (e.g., obstruent clusters in the URs of morphs) as well. \citet{Pierrehumbert1994} does not mention voice assimilation in her study of gaps in English syllable contact clusters. \citet[][74f.]{Hammond1999a} notes the existence of words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain obstruent voice clusters disagreeing in [\textsc{Voice}], and takes this as evidence that obstruent voicing is restricted to derived environments.

Despite these exceptions, I propose that \textsc{Obstruent Voice Assimilation} plays a major role in shaping the English lexicon. Only English syllable contact clusters which contain two adjacent obstruents speak to the status of \textsc{Obstruent Voice Assimilation}; the remaining 160 possible clusters are ignored.
A $2 \times 2$ contingency table is constructed by dividing the remaining possible clusters into those which agree on obstruent voicing and those which do not, 
and into those which are attested and those which are not.

\ex Lexical effects of \textsc{Obstruent Voice Assimilation}: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 80 & 370 & 0.178 & \multirow{2}{*}{1.1\e{-11}}\\
violating  &  6 & 264 & 0.022 \\
\bottomrule
\end{tabular}
\xe

\noindent 
18\% of the possible clusters containing adjacent obstruents with the same [\textsc{Voice}] specification are found, 
whereas only 2\%  of those which disagree in [\textsc{Voice}], 
such as in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, 
CELEX also contains \emph{jo}[d.p]\emph{urs}, \emph{pi}[ntʃ.b]\emph{eck}, \emph{sa}[k.b]\emph{ut}, and \emph{ja}[k.d]\emph{aw}. 
As the $p$-value shows, 
the rarity of these disagreeing clusters is unlikely to be due to chance.

I propose that this pattern should be attributed directly to \textsc{Obstruent Voice Assimilation} applying to non-derived environments, preventing learners from positing URs in which adjacent obstruents do not agree in voice. There is no duplication between constraints on sequence structure and surface forms if both are derived from the application of the phonological rule (\citealt[][401f.]{Stanley1967}, \citealt[][382]{SPE}). The handful of exceptions to this generalization can be marked with appropriate exception diacritics. Another possibility is suggested by recent psycholinguistic results. \citet{Mattys2001b} find that 9-month old infants are sensitive to the contrast between [vt], a cluster of obstruents disagreeing in [\textsc{Voice}] which is not found word-medially, and [ft], which conforms to the above generalization and is found medially (e.g. \emph{a}[f.t]\emph{er}). This ability to use phonotactic violations to segment continuous speech persists into adulthood \citep{McQueen1998}, suggesting that ``hetero-voiced'' obstruent clusters may be analyzed as semantically opaque compounds by native speakers (e.g., \emph{ja}[k\#d]\emph{aw}). 
%\footnote{An analogous account is the proposal that readjustment rules insert morphological boundaries for the purposes of assigning stress (\emph{SPE}:94).}

\subsubsection{Coda nasal place assimilation}
\label{cnpasection}

Above, it was assumed that the velar nasal is derived from underlying /n/. 
This analysis is not purely allophonic, as nasal place assimilation alternations target the final consonant of the \emph{im-}/\emph{in-} prefix. In (\lastx ab), the prefix-final consonant takes on the same major place of articulation as the following obstruent, and in (\lastx c), where the root is vowel-initial, the prefix-final consonant is realized as a default /n/.

\ex Nasal place assimilation in Latinate prefixes: \\
\begin{tabular}{l l l l}
%a. & possible & i[m.p]ossible & & balance & i[m.b]alance \\
%b. & tangible & i[n.t]angible & & decent  & i[n.d]ecent  \\
%c. & elegant  & i[n.]elegant  & & ability & i[n.]ability \\
a. & possible & i[m.p]ossible \\
   & balance  & i[m.b]alance  \\
b. & tangible & i[n.t]angible \\
   & decent   & i[n.d]ecent   \\
c. & elegant  & i[n.]elegant  \\
   & ability  & i[n.]ability  \\
\end{tabular} 
\xe

\noindent
The shape of the prefix before other obstruents is somewhat more complex. 
%Before /f/, CELEX transcribes the nasal as either labial or alveolar, the latter being very rare in simplex words. 
%There are also additional complexities regarding the shape of the prefix when attached to /k, g/-initial roots. 
\citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but this assimilation is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}. The prefix \emph{com-}/\emph{con-}, which patterns with \emph{im-}/\emph{in-} in showing assimilation (e.g., \emph{co}[m.b]\emph{ine}), participates in a similar contrast, between \emph{có}[ŋ.g]\emph{ress} and \emph{co}[n.g]\emph{réssional}. As \citeauthor{Borowsky1986} notes, stress does not block assimilation in simplex words (e.g., \emph{a}[ŋ.g]\emph{óra}), a fact which demonstrates the necessity of a partially morphological explanation for the blocking of assimliation in words like \emph{co}[n.g]\emph{réssional}. Regardless of these complexities, there is sufficient evidence to posit assimilation of the \textsc{Place} features of an obstruent to a preceding nasal consonant.

\ex \textsc{Coda Nasal Place Assimilation} (\emph{SPE}:85): \\
\label{CNPA}
\xymatrix@C=6pt{
                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{..}[dl] \\
\txt{C}\ar@{-}[d]      & \txt{C}\ar@{-}[d] \\
\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]} }
\xe

\citet[][175]{Pierrehumbert1994} observes that in simplex English words, ``nasal-stop sequences agree in labiality''; presumably, this generalization is limited to [\textsc{Labial}] because \citeauthor{Pierrehumbert1994} also adopts the assumption that the velar nasal is derived from /n/. However, \citeauthor{Pierrehumbert1994} does not propose any connection between the alternation and the generalization that nasal consonants in URs tend to agree with the following consonant. While clusters like \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey} are the norm, exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} are found.

In all, there are only 80 possible nasal-obstruent syllable contact clusters, the majority of which are attested. In addition, there are a small number of words which contain clusters like \emph{oi}[nt.m]\emph{ent}, in which a nasal-obstruent cluster are both parsed into a complex medial coda, but in this position, place agreement is exceptionless, and thus these clusters are ignored below.

\ex Lexical effects of \textsc{Coda Nasal Place Assimilation}: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 41 & 11 & 0.788 & \multirow{2}{*}{2.6\e{-05}}\\
violating  & 8  & 20 & 0.286 \\
\bottomrule
\end{tabular}
\xe

\noindent
Clusters which conform to this generalization are far more likely to be attested than those that violate it, and this is highly unlikely to be due to chance.

There is some evidence that even young infants show a preference for nonce words that conform to this generalization. \citet{Jusczyk2003}, \citet{Mattys1999}, and \citet{Jusczyk2002} play nonce words containing nasal-obstruent clusters to 4.5, 9, and 10-month-old infants (respectively) acquiring English, All three studies find that infants prefer to listen to stimuli which conform to the above generalization (e.g., \emph{u}[m.b]\emph{o}) over those which do not (e.g., \emph{u}[n.b]\emph{o}). \citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''. In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.

\ex Artificial language adaptations \citep[][394]{Wright1975}: \\
\begin{tabular}{l l l l l l}
a. & [gownp]    & > & [gump] \\
b. & [\v{j}umg] & > & [\v{j}úŋgə] \\
c. & [ðʌŋd]     & > & [tɔŋg] \\
\end{tabular} 
\xe

\noindent
If this game is akin to the process of loanword adaptation, it is possible that various extragrammatical pressures are at play \citep[e.g.,][]{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008}. Yet, the independent evidence for a phonological process of nasal place assimilation provides the most parsiminious account of the adaptations seen in \citeauthor{Wright1975}'s study.

%\citet{Hay2004a} demonstrates that [n.p, mθ] clusters, which are not found in English, are in fact rated better than many attested clusters.
%report that 10-month-old infants prefer to listen to nonsense words which conform to \textsc{Coda Nasal Place Assimilation} (e.g., \emph{u}[m.b]\emph{o}) compared to those which do not (e.g., \emph{u}[n.b]\emph{o}).

\subsubsection{Degemination}

A final well-known alternation in English is the simplification of geminates derived by certain suffixes. For instance, the suffix \emph{-ly} rarely forms a geminate when attaching to root-final /l/, whether forming denominal adjectives (e.g., \emph{sme}[l]\emph{y}) or deadjectival adverbs (e.g., \emph{norma}[l]\emph{y}). A small class of English verbs take a /-t/ suffix in the past; as this suffix does not trigger the \textsc{Epenthesis} found in the regular past (e.g., \citealt{Bakovic2005b}, \citealt{Fruehwald2011}), \textsc{Degemination} applies.

\ex Degemination in the /-t/ past (\citealp[][105]{Halle1985a}, \citealp[][492]{Myers1987}): \vspace{6pt} \\
\begin{tabular}{l l l l l}
a. & /bɛnd-t/ & \arrow & [bɛnt] & (cf. \emph{burn}[t]) \\
b. & /bɪld-t/ & \arrow & [bɪlt] & (cf. \emph{spil}[t]) \\
\end{tabular}
\xe

More complicated cases of degemination are found in English prefix morphology and described in \emph{SPE} (p.~148, 219f.) and \citealt[][]{Borowsky1986} (p. 102f.), and are not reviewed here.

%\ex \textsc{Coda Nasal Place Assimilation} feeds \textsc{Degemination} \citep[][116]{Borowsky1986}: \\
%\begin{tabular}{l l l l}
%a. & mature    & i[m]ature    & (cf. \emph{i}[m.b]\emph{alance})   \\
%b. & numerable & i[n]umerable & (cf. \emph{i}[n.d]\emph{ependent}) \\
%\end{tabular}
%\xe

The rule deletes the first of two adjacent consonants agreeing in \textsc{Place} and manner features, and is formalized using quantification over features \citep{Reiss2003b}.

\ex \textsc{Degemination} (\emph{SPE}:253): \\
\label{degem}
%\xymatrix@C=6pt{
%{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\
%\textsc{C}\ar@{-}[d]       & \longrightarrow & \emptyset & / & \gap\gap\gap & \textsc{C}\ar@{-}[d] \\
%*+[F(][F)]\txt{β \textsc{Continuant} \\ γ \textsc{Obstruent} \\ δ \textsc{Sonorant} \\ ε \textsc{Lateral} \\ ζ \textsc{Nasal}} & & & & & *+[F(][F)]\txt{β \textsc{Continuant} \\ γ \textsc{Obstruent} \\ δ \textsc{Sonorant} \\ ε \textsc{Lateral} \\ ζ \textsc{Nasal}} \\ \\
%}
\xymatrix@C=6pt{
%{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\
\textsc{C}_1       & \longrightarrow & \emptyset & / & \gap\gap\gap & \textsc{C}_2 }
\\ 
Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
\xe

\noindent
I assume that geminates crossing a syllable boundary are ``false geminates'', meaning that they are adjacent similar segments rather than a single feature matrix linked to two timing slots. Under the alternative account in which the absence of geminates is an inventory restriction (i.e., a \textsc{Segment Structure Condition}), an undesirable duplication of phonological derived surface constraints and constraints on URs would be unavoidable.

Neither identical consonants, nor those which differ only in voice (e.g., *[p.b]) occurs in simplex words in CELEX, and this absence is highly unlikely to be due to chance.

\ex Lexical effects of \textsc{Degemination}: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 157 & 596 & 0.208 & \multirow{2}{*}{7.2\e{-09}}\\
violating  &   0 &  87 & 0.000 \\
\bottomrule
\end{tabular}
\xe

%\footnote{Note that it is not the case that the English lexicon as envisioned by \emph{SPE} is free of underlying geminates; \citeauthor{SPE} use underlying geminates to attract stress to a preceding nucleus (p.~148--151). In \emph{SPE}, the underlying geminate-singleton contrast is absolutely neutralized, but has consequences for surface stress. In addition to the usual objections to absolute neutralization (i.e., that it is a learning conundrum mascarading as a phonological solution) that emerged in the wake of the publication of \emph{SPE}, the phonological particulars of this account have also come under attack \citep[e.g.,][]{Ross1972a}. It would not seem tenable to use underlying geminates to attract stress even if absolutely neutralized contrasts were admitted.} Unsurprisingly, this is unlikely to be due to chance.

\subsubsection{Dorsal/non-coronal clusters}

\citet{Pierrehumbert1994} surveys the English syllable contact cluster inventory, and discerns the effect of several ``static'' constraints, those which do not derive from neutralizing phonological rules.

\citet[][173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Since \citeauthor{Pierrehumbert1994}'s study is restricted to triconsonantal clusters, so words like \emph{a}[k.m]\emph{e}, \emph{ru}[g.b]\emph{y}, or \emph{pi}[g.m]\emph{ent} do not violate this narrower generalization. Further, \citeauthor{Pierrehumbert1994} adopts the traditional assumption \citep[e.g.,][66f.]{Borowsky1986} that coda [ŋ] is an allophone of /n/ before dorsal consonants, so words such as \emph{mo}[ŋ.gr]\emph{el} are not exceptions.

\citeauthor{Pierrehumbert1994} generalization is termed static simply because there is no evidence for any phonological process which changes clusters of dorsal obstruents followed by non-coronals into something else (perhaps a dorsal/coronal cluster). However, the statistical test, taking into account clusters of any length, does not indicate that this dispreference for dorsal/non-coronal clusters is so compelling that it could not be generated by chance, simply because possible dorsal/non-coronal clusters are about as likely to occur as are similar types of clusters such as those with a non-dorsal coda and a non-coronal onset (e.g., \emph{a}[t.m]\emph{osphere}) or a dorsal coda and coronal onset (e.g., \emph{ve}[k.t]\emph{or}). These clusters are marked as ``conforming'' to a constraint \textsc{*[+Dorsal][$-$Coronal]}, and dorsal/non-coronal consonants as ``violating'' it.

\ex Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 76 & 359 & 0.175 & \multirow{2}{*}{0.145} \\
violating  &  6 &  57 & 0.095 \\
\bottomrule
\end{tabular}
\xe

There is no exceptions to \citeauthor{Pierrehumbert1994}'s narrower generalization concerning triconsonantal clusters, but it is no more statistically convincing.

\ex Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]} on triconsonantal clusters: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 20 & 148 & 0.119 & \multirow{2}{*}{0.136} \\
violating  &  0 &  22 & 0.000 \\
\bottomrule
\end{tabular}
\xe

\noindent
The evidence for \citeauthor{Pierrehumbert1994}'s static constraint does not actually require that such a constraint exists at all, since it is quite possible that the apparent dispreference for dorsal/non-coronal clusters could result by accident.

\subsubsection{Coda coronal obstruents}

Another static generalization about triconsonantal syllable contact clusters proposed by \citet[][175]{Pierrehumbert1994} is that  ``clusters with a coronal obstruent in the coda do not occur.'' In an endnote (p. 186), \citet{Pierrehumbert1994} notes an exception in \emph{a}[nt.l]\emph{er}, and one can add the triconsonantal clusters in \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many biconsonantal clusters like \emph{a}[t.l\emph{as} or \emph{e}[s.t]\emph{er}. As before, applying the statistical test to clusters of any length reveals that coda coronal obstruent clusters are not significantly less likely to occur than clusters with non-coronal codas.

\ex Lexical effects of \textsc{*Coda Coronal Obstruent}: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming &  48 & 312 & 0.133 & \multirow{2}{*}{0.301} \\
violating  &  38 & 322 & 0.106 \\
\bottomrule
\end{tabular}
\xe

Restricting \citeauthor{Pierrehumbert1994}'s generalization to triconsonantal clusters does not strengthen the case for this static constraint.

\ex Lexical effects of \textsc{*Coda Coronal Obstruent} on triconsonantal clusters: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming &  4 & 104 & 0.037 & \multirow{2}{*}{0.419} \\
violating  &  2 & 124 & 0.016 \\
\bottomrule
\end{tabular}
\xe

%Note that [tʃ] has not been included since it so rarely occurs in clusters.

\subsubsection{ABA clusters}

Finally, \citet[][176]{Pierrehumbert1994} notes ``a lack of clusters with identical first and third elements''. For the determination of identity, I assume the same definition of geminate as used in the rule of \textsc{Degemination} above, namely consonants which share all the same \textsc{Place} and manner features, ignoring \textsc{Voice}. 
%and I also expand the definition of the constraint to also apply to the second and fourth consonant of quadraconsonantal clusters. 
\citeauthor{Pierrehumbert1994} notes one potential exception to this generalization, the cluster in words such as \emph{e}[ks.kl]\emph{ude}, which contains two /k/, does not arise here since such words have been marked morphologically complex. Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly rarer than other triconsonaontal and quadraconsonantal clusters, of which only 7.8\% occur.

\ex Lexical effects of \textsc{*ABA}: \vspace{6pt} \\
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 41 & 478 & 0.079 & \multirow{2}{*}{0.818} \\
violating  &  0 &  11 & 0.000 \\
\bottomrule
\end{tabular}
\xe

Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}.

%\subsubsection{Summary}
%
%The above results are summarized in Table \ref{constraints}.
%
%\begin{table}
%\centering
%\begin{tabular}{l r r r r r}
%\toprule
%                                       & \multicolumn{2}{c}{attested} & \multicolumn{2}{c}{unattested} & \multirow{2}{*}{$p$-value} \\
%                                       & conforming & violating & conforming & violating \\
%\midrule
%\textsc{O.V.A.} (\ref{ovar})        &  80 &   6 & 370 & 264 & 1.1\e{-11} \\
%\textsc{C.N.P.A} (\ref{CNPA})       &  41 &  25 &  12 &  42 & 1.7\e{-05} \\ 
%\textsc{Degemination} (\ref{degem}) & 161 & 632 &   0 &  87 & 1.4\e{-08} \\
%\midrule
%\textsc{*Dorsal-Labial} ()          &  68 &   4 & 246 &  40 & 0.069 \\
%\textsc{*C.C.O.} ()                 &  48 &  38 & 312 & 322 & 0.301 \\
%\textsc{*A,BA} ()                   & 161 &   0 & 708 &  11 & 0.231 \\
%\bottomrule
%\end{tabular}
%\caption{Counts and Fisher exact test $p$-values for the constraints discussed above.}
%\label{constraints}
%\end{table}

### 2012-02-16 2.2.2.tex
Restructured into smaller files, wrote much of chapter 2.

% 2.2.2: Computational models

Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}. There is now a growing literature on computational cognitive models of phonotactic learning. By applying these models to the syllable contact data, it is possible to exhaustively search for statistically motivated constraints, elimianting the possibility of human error.

To evaluate these statistical models, they are first trained on the set of attested clusters, then put to the task of classifying all possible clusters into those which are and are not attested. Four metrics are used to evaluate these computational models. Classification accuracy corresponds to the probability of a classification being either a true positive ($t.p.$), correctly predicting a cluster to occur, or a true negative ($t.n.$), correctly predicting a cluster to be absent. Incorrect outcomes are either false positives ($f.p.$), incorrectly predicting a cluster to occur, and false negatives ($f.n.$), incorrectly predicting a cluster to be absent.

%\ex Binary classification contingency table: \\
%\begin{tabular}{c | l l} \toprule
%                             & \multicolumn{2}{c}{prediction} \\
%\midrule
%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
%                             & false negative ($fn$) & true negative ($tn$) \\
%\end{tabular}
%\begin{tabular}{|l l|}
%\toprule
%true positive ($tp$)  & false positive ($fp$) \\
%false negative ($fn$) & true negative ($tn$) \\
%\bottomrule
%\end{tabular}
%\xe

\ex $\displaystyle \textrm{Accuracy} = \frac{t.p. + t.n.}{t.p. + t.n. + f.p. + f.n.}$ \xe

\noindent
A few other metrics can be used to assess classification performance. Precision corresponds to the probability that a cluster predicted to occur is observed in the data.

\ex $\displaystyle \textrm{Precision} = \frac{t.p.}{t.p. + f.p.}$ \xe

\noindent
Recall (also known as sensitivity) is the probability a observed cluster is predicted to occur.

\ex $\displaystyle \textrm{Recall} = \frac{t.p.}{t.p. + f.n.}$ \xe

It is possible to maximize recall at the expense of precision, by predicting more clusters to be attested, or to increase precision at the expense of recall by predicting fewer clusters to be attested. A common way to balance these concerns is the $F_1$ score, which is the harmonic mean of precision and recall.

\ex $\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$ \xe

\subsubsection{Null baseline}

The 840 possible clusters have an 18.7\% of saturdaion rate. As a consequence of this sparsity, a null model which classifies all clusters as unattested achieves 81.3\% accuracy without the need to posit any constraints either derived or static. Since this model does not have any positive classifications, either true or false, precision, recall, and $F_1$ are undefined.

\subsubsection{Alternations}

As was shown above, \textsc{Obstruent Voice Assimiation}, \textsc{Coda Nasal Place Assimilation}, and \textsc{Degemination} are reliable predictors of what clusters are and are not attested and have few lexical exceptions. I generate another baseline by predicting only those clusters which do not violate any of these generalizations to be attested. This provides a significant improvement to the null baseline (accuracy = 0.857). As precision (0.401) is smaller than recall (0.708), there are more false positives (i.e., unattested clusters that do not violate any of these three alternation-derived generalizations) than there are false negatives (i.e., attested exceptions to these derived generalizations).

\subsubsection{\citet{Pierrehumbert1994}}

\citet{Pierrehumbert1994} first proposes to measure the well-formedness of syllable contact clusters by the independent probabilities of codas and onsets, and \citet{Coleman1997} extend this model to whole words. \citet{Pierrehumbert1994} uses both medial and initial/final codas and onsets to compute these probabilities, but I obtain a better fit to the data by only using medial coda and onset frequencies. Medial coda and onset probabilities are multipled to obtain a well-formedness score. This produces a continuous value between 0 and 1. To derive a binary classification from these values, I use a soft-margin support vector machine \citep{Cortes1995} with a linear kernel. This technique derives a ``decision stump'' providing the best single split into attested and an unattested clusters. This is not put forth as a cognitively plausible model for relating this expected probability to attestation: it is simply the upper bound that such a model could reach.

Unfortunately, this score only provides a small improvement to the null baseline
 (accuracy = 0.839), and has a lower accuracy and $F_1$ (0.328) than the alternation baseline.

\subsubsection{\citet{Hayes2008a}}

\citeauthor{Hayes2008a} develop a sophisticated method of learning phonotactic distributions from positive data. They develop software for estimating probability distributions over sequences of (sets of) phonological features using the principle of maximum entropy and making few assumptions about the independence of features. As above, a support vector machine with a linear kernel is used to turn these probability values into binary predictions. The software provided by these authors exposes many ``metaparameters'', operational decisions to be made in conducting an experiment with this system. To avoid any potential bias induced by the choice of these metaparameters, I use the same settings as used by \citet{Hayes2008a} in their similar study of English onset consonants. This includes the ``accuracy schedule'' [0.001, 0.01, 0.1, 0.2, 0.3] and no constraints on the number of constraints learned, as well English consonantal features used by \citeauthor{Hayes2008a}. Since this model is inherently stochastic, producing slightly different outcomes on each run, I follow the practice of \citeauthor{Hayes2008a} (p.~396) and report the lowest-accuracy of ten runs, though in general there is not a great deal of variation between individual runs.

This model provides a small improvement in accuracy (0.877) and $F_1$ (0.703) compared to the alternation baseline, but it has lower recall (0.642) than this baseline. This indicates that the model produces many false negatives.
%; it assigns the lowest probability value to 122 unattested clusters.
This seems to further indicate that this model internalizes narrower phonotactic generalizations than those derived from the three alternation-derived constraints; for instance, it assigns the highest probability to unatteested *[m.kl], which is ruled out by \textsc{Coda Nasal Place Assimilation}. On the other hand, it also rules out clusters like those in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are consistent with all three alternation-derived constraints.

\citet{Berent2012} identify an important defect with the publically-available version of the \citeauthor{Hayes2008a} model used here: it cannot directly incorporate constraints like the alteration-derived constraint disfavoring geminate consonants, because it does not include variables over phonological features. However, the model appears to have experienced no problems internalizing the generalization that geminate clusters are unattested.

The results for the four models are summarized below.

\ex Summary of computational model accuracy: \vspace{6pt} \\
\begin{tabular}{l | r r r r}
\toprule
                          & accuracy & precision & recall & $F_1$ \\ 
\midrule
Null baseline             & 0.813    & n.a.      & n.a.   & n.a.  \\
Alternations              & 0.857    & 0.401     & 0.708  & 0.512 \\
\citet{Pierrehumbert1994} & 0.839    & 0.210     & 0.750  & 0.328 \\
\citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
\bottomrule
\end{tabular}
\xe

\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{I am grateful to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are more frequent than those like [t.j], with rising sonority. \citeauthor{McGowan2011} reports, however, that this accounts for only a tiny amount of the variance in frequency ($R^2 = 0.077$). Change in sonority provides no improvement to the null baseline, and so this model is not reported here.

### 2012-02-16 2.2.3.tex
Restructured into smaller files, wrote much of chapter 2.

% 2.2.3: Statistical properties of syllable contact

Above, we have seen that constraints on URs which are derive from phonological alternations restrict the contents of the lexicon. Applying these derived constraints to the lexicon increases saturation from 18.8\% to 28.8\% and incurs only a handful of exceptions. Yet, despite the fact that ``attestation'' is the minority pattern, attempts to identify static lexical constraints, whether by hand or computational model, lend little additional predictive power. There is no evidence that the English syllable contact inventory is subject to any static constraint at all. I am forced to conclude that most, if not all, of 71.2\% of possible clusters which are unattested are accidental gaps. Below, I show that this state of affairs follows directly from the sparse distribution of codas and onsets.

\subsubsection{The $Z_r$ transform}

In a sparse distribution, many types occur at the same low frequencies. \citet{Good1953} notes that this produces a type of quantization, an artificially long and flat right tail in the frequency spectrum. \citet[][29]{Church1991} propose a method to eliminate this quantization, the $Z_r$ transform, defined as follows.
Let $r, n$ be defined such that $n_i$ is the number of types which occur at frequency $r_i$ (i.e., $n$ contains frequencies of type frequencies). Further, let $r$ be sorted in increasing order. $Z$ is simply each element of $n$ scaled by the nearest points to the left and right.

\ex $\displaystyle Z_i = \frac{2 n_i}{r_{i + 1} - r_{i - 1}}$ \xe

\noindent
\citeauthor{Church1991} do not define this transform for the lowest and highest points (i.e., when $i = 1$ or $i = N$), and so I extend the definition by scaling these points using only one inward-facing point.

\ex $\displaystyle Z_1 = \frac{n_1}{r_2 - r_1}$ \xe
\ex $\displaystyle Z_N = \frac{n_N}{r_N - r_{N - 1}}$ \xe

%\indent
%The effect of this transform can be seen in Figure \ref{zr}.

%\begin{figure}
%\centering
%\includegraphics{zr.pdf}
%\caption{The SUBTLEX word frequency norms \citep{Brysbaert2009} show a long right tail in the log-frequency/log-rank space (left panel), and the $Z_r$ transform (right panel) smooths out this tail.}
%\label{zr}
%\end{figure}

\subsubsection{Coda, onset, and cluster sparsity}

A number of previous studies \citep[e.g.,][]{Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} have observed that phonemes and graphemes exhibit sparse type (i.e., lexical) and token frequency distributions. The lexical frequencies of the medial codas ond onsets which make up syllable contact clusters show a similar distribution. In Figure \ref{codaonset}, $Z_r$-transformed lexical frequencies of codas and onsets are plot against rank, in log-log spaces. The near-linear relationship that obtains indicates that the data can be modeled by a generalization of Zipf's law \citep{Zipf1949}.

\begin{figure}
\centering
\includegraphics{coda.pdf}
\includegraphics{onset.pdf}
\caption{Medial coda and medial onset lexical frequency exhibit the Zipfian log-log-linear relationship between rank and frequency.}
\label{codaonset}
\end{figure}

\ex $\displaystyle f(r~; C, \alpha) = \frac{C}{r^\alpha}$ \xe

\noindent
The parameters $C$, a constant sensitive to sample size, and $\alpha$, the slope of the rank/frequency relationship in log-log space, are easily computed using the method of least squares in a linear regression, where $\epsilon$ represents the error term.

\ex $\displaystyle \textrm{log}~Z_r \sim C + \alpha~\textrm{log}~r + \epsilon$ \xe

\noindent
Codas ($\alpha = -0.720$, $R^2 = 0.749$) and onsets ($\alpha = -1.056$, $R^2 = 0.828$) are well-fit by this distribution, though there are some outliers. Similarly, clusters, shown in Figure \ref{clus}, are also a good fit to the generalized Zipf's law ($\alpha = -0.588$, $R^2 = 0.924$).

\begin{figure}
\centering
\includegraphics{cluster.pdf}
\caption{Syllable contact clusters, themselves composed of Zipfian medial codas and medial onsets, also exhibit a Zipfian distribution.}
\label{clus}
\end{figure}

That clusters (and the codas and onsets that make them up conforms to Zipf's law is not itself a deep observation. Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also $n$-grams \citep{Baroni2009} and of randomly-generated texts \citep{Miller1957,Li1992}. The import of this statistical property for the study of cluster phonotactics is that it makes it difficult, on statistical grounds alone, to determine which unobserved events are excluded, and which are accidentally missing.

\citet{Good1953} puts forth a simple theory of just how frequent unobserved events might be. He proposes to estimate $\hat{p}_0$, the probability of all unseen events, as the ratio of events which occur just once $n_1$, to the number of observations, $\sum N$. This quantity is known as the Good-Turing estimate, since it was originally developed by Alan Turing for codebreaking during the second World War.

\ex $\displaystyle \hat{p}_0 = \frac{n_1}{\sum N}$ \xe

\noindent
In the CELEX data, $65$ clusters occur only once, and there are 873 clusters in all, so $\hat{p}_0 = 0.074$. One way to interpret this quantity is with respect to a ``replication''. If it were possible to generate a new corpus of syllable contact clusters, approximately 7\% of cluster tokens will be ones that were not attested in the previous sample, i.e., they accidentally failed to be sampled. This is a not-inconsiderable quantity.

Unfortunately, the Good-Turing estimate does not provide a way to determine which unattested clusters might be found in a replication. But I have already proposed that the null hypothesis of generative phonology, combined with the principle of Stampean occultation, provides a way out of this conundrum. Unattested clusters like *[m.kl] are structurally excluded (since such a cluster is occulted by \textsc{Coda Nasal Place Assimilation}), whereas *[s.l] is probably an accidental gap.

\subsubsection{Sampling simulation}

As a final investigation into this data, I consider the statistical properties of a randomly generated novel ``lexicon''. The simulated lexicon is generated by  generating a distribution of clusters, by repeatedly applying the following procedure.

\ex {Simulation procedure}: \\
    \begin{tabular}{l l}
    a. & Sample a medial coda according to the observed probabilities  \\
    b. & Sample a medial onset according to the observed probabilities \\
    c. & Apply the SPE rules to the resulting cluster                  \\
    \end{tabular}
\xe

\noindent
This is a simulation version of the model proposed by \citet{Pierrehumbert1994} and \citet{Coleman1997}, but also takes occulting phonology into account.

While this procedure can be repeated indefinitely (using the CELEX data, listed in Appendix \ref{A}), consideration of a single characteristic run of the simulation should suffice. The saturation rate of the simulated sample (18.2\%) is very close to the real sample (18.8\%). The slope of the log-rank/log-frequency relationship for the simulated data ($\alpha = -0.557$) is also very close to the observed slope for the CELEX data ($\alpha = -0.588$). And, as shown in Figure \ref{sim}, the two distributions are nearly indistinguishable.

\begin{figure}
\centering
\includegraphics{sim.pdf}
\caption{A comparison of the observed lexical frequencies of English syllable contact clusters to an ``English lexicon'' simulated by combining.}
\label{sim}
\end{figure}

It is important to note that the results of this simulation should not be construed as an endorsement of a revised version of the \citet{Pierrehumbert1994} and \citet{Coleman1997} model which also incorporates occultation. In fact, as was shown above, the independent frequencies of coda and onset are very poor predictors of what clusters will be attested even when occultation is taken into account, and the set of attested clusters only partially overlaps those produced by the simulation. I propose that this is because the lexicon is not only finite, but far smaller than the combinatoric possibilities provided by phoneme sequences, making accidental gaps unavoidable. That these gaps will be arbitary from the perspective of phonology is nothing more than a corrolary of the principle of \emph{l'arbitraire du signe}.

It is quite possible that phonology (and, perhaps, static phonotactics) acts as a filter on what URs are learned, but a cluster might be overrepresented or underrepresented for reasons that are quite external to phonology, e.g., language contact or historical change. Similarly, the factors that cause a /n/ to be a common word-medial coda in English may be independent of the propensity of that coda to combine with medial onset, assuming the resulting cluster is phonologically licit. The import of this simulation is somewhat more subtle; it simply shows free combination of codas and onsets according to their independent probabilities, constrained only by occultation, would produce the same sparse distribution of clusters that was used to argue for static phonotactics.

%\citet{Evert2004}
%\includegraphics{zipf.pdf}

### 2012-02-16 2.3.tex
Restructured into smaller files, wrote much of chapter 2.

% 2.3: Conclusion

The above is a stark reminder.

A reasonable objection to the arguments presented in this chapter is to view the results as little more than an indictment of the results of \citealt{Pierrehumbert1994}.

However, the fact that state-of-the-art models are not capable of providing large improvements to the predictive accuracy indicates

 do not have a statistically significant effect on the shape of the English lexicon, but that experiments might turn up evidence that speakers have internalized

shown to be aware of static constraints if they reached statistical significance.

 statistically reliable static constraints could be identified

The of

Regarding the historical developments,

\citet{Martin2007}

On the other hand, patterns created by sound change are not guaranteed to persist over time. 
One example of non-persistence is discussed by \citet{Iverson2005}.  
Around 1100 CE, Old English \emph{sk} became [ʃ]. 
This sound change introduced no alternations.
Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{V\lm sk\#} words when the
 change was actuated, and \emph{V\lm sh\#} continues to be rare in Modern English. 
What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.

A third pattern is that a historically inherited pattern

\citet[][140]{Frisch2004} suggest that the strong tendency for the first and second consonants of the Arabic root to be non-identic
al is the ``a diachronic result of a processing constraint that disfavors repetition.'' 
Unfortunately, there is no evidence that this pattern is diachronic other than in the sense that it appears to be inherited from the proto-language: there is simply no Proto-Semitic verb roots with identical first and second consonants \citep[][178]{Greenberg1950}. 
In other Semitic languages, the inherited patern has experienced considerable erosion.

\ex Tigrinya roots with identical first and second consonants \citep{Buckley1990a}: \\
\begin{tabular}{l l l l}
a. & lʌlʌw     & `scorch'                   & (< Ge'ez \emph{lʌwlʌw} `inflame')     \\
   & mʌmʌy     & `winnow'                   & (< Ge'ez \emph{mʌymʌy} `distinguish') \\
%   & mʌmʌt & `pick out loot' & (< 
b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
\end{tabular}
\xe

Similar exceptions are found in 
%Amharic (\citealp[][?]{Broselow1984}, \citealp[][?]{McCarthy1985}) and 
Hebrew \citep[][29]{Bat-El2005}.

The next two chapters return to the question of synchrony, addressing the relationship between statistical patterns in the lexicon and speakers' behaviors when presented with underrepresented sequences.

### 2012-02-16 2.tex
Restructured into smaller files, wrote much of chapter 2.

% 2: Accidental and structural gaps in English syllable contact

In the preceding chapter, it was proposed that

Reconsider the null hypothesis presented in the preceding chapter:

\ex The null hypothesis: \\
Universal grammar does not countenance constraints on the contents of URs
\xe

The idea

relationship to proposal

One primary way which this null hypothesis has been critiqued is by demonstrating that the lexicon of some language 
exhibits dispreferences or gaps corresponding to.

argue by trends in the lexicon

possibility of sparsity (C\&H, Fischer/Vogt and recent stats)

Of course, to obtain the desired results, we must guarantee that each sequence structure rule reflects a general systematic fact about the language, and not a fact which is due merely to the existence of accidental gaps in the lexicon. \citep[][401, fn.~8]{Stanley1967}

In the preceding chapter, it was proposed that

Lexicon

This appears to be exceptionlessly true of Arabic

That even gradient patterns

One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. 
Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a} but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Dmitrieva2008a,Dmitrieva2008b,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.

The absence of 
The claim here is that however stark the absence or underrepresentation of some form is, it

This hypothesis admits the possibility that there may be accidental gaps in the lexicon, a possibility which predates generative thinking:

\begin{quotation}
\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language. \citep[][16]{Fischer-Jorgensen1952}
\end{quotation}

\noindent
Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters

\begin{quotation}
Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters. The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
\end{quotation}

\citet{Chomsky1965}

This chaper demonstrates that apparently systematic gaps may arise even in the absence of phonotactic preferences, and that this possibility compromises attempts to show that such preferences shape the lexicon.

### 2012-02-16 abs.tex
Restructured into smaller files, wrote much of chapter 2.

--- 2011-12-01 abs.tex
@@
-This dissertation considers two types of ``gaps'' in the lexicon.
-In both cases, it is argued that current generative models can be modified minimally to account for the apparent overgeneration of gapped structures.
-The first part of the dissertation considers the underlying phonological structure of morphs, specifically medial consonant clusters in English.
-A number of prior studies have argued that the grammar must have a separate ``phonotactic'' module to account
+This dissertation uses the status of unattested words to provide evidence for the architecture of Universal Grammar (UG).
 
-at the level of prosodic word. In both cases, it is argued that the i
+The first part of the dissertation considers constraints on phonological underlying representations (URs). 
+It is proposed that UG countenances no language-specific constraints on URs. 
+Putative counterexamples are dismissed as either diachronic contingencies or as induced by synchronic constraints not coterminous with the morph.
+A corpus study of the inventory of medial consonant clusters in English simplex words reveals that a number of gaps in this system are accidental in nature, and do not require the structural explanations proposed in prior studies.
+A reanalysis of a study of Turkish wordlikeness judgements reveals that speakers are sensitive to constraints on underlying representations if and only if they are induced by phonological alternations.
+A systematic review of English wordlikeness studies reveals once the presence or absence of gross phonotactic violations is controlled for, current gradient phonotactic models are superfluous.
+
+The second part investigates inflectional gaps, complex words judged ineffable.
+It is proposed that UG countenances no language-specific constraints on phonological output representations which give rise to ineffability. 
+Putative counterexamples are dismissed.
+%This indicates that Universal Grammar does not incorporate any inviolable constraints on surface structure. 
+It is proven that a model of inviolable constraint learning fails to converge on the proper grammar. 
+Models which store complex words are shown to result in extensive undergeneration. 
+An analysis of inflectional gaps as reflexes of unproductive morphophonological generalizations is put forth and applied to cases from English, Spanish, Greek, and Russian.

### 2012-02-16 ack.tex
Restructured into smaller files, wrote much of chapter 2.

--- 2011-11-14 ack.tex
@@
-So many researchers contribute to any dissertation that to name them all compromises the honor of those most involved.
-More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, for his commitment to my work and the principle of explanatory adequacy.
-My other internal committee members, Gene Buckley, Mark Liberman, and Rolf Noyer, and the faculty at IRCS, are responsible for creating an environment at Penn where this (admittedly contentious) dissertation could be written.
-Though his name appears nowhere else in this document, it is hard to imagine it could exist without the pioneering work of Bill Labov.
+So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy. 
+More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, whose influence can be seen throughout. 
+Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are responsible for creating an environment at Penn where a student could write this admittedly contenious dissertation. 
+And though his name appears nowhere else on this document, it is hard to imagine this work could exist without Bill Labov.
 
-The subject of the first chapter was inspired by the mentoring of Jennifer Cole.
-Important contributions to this dissertation were made by Stephen Anderson, who graciously agreed to serve on my committee, Michael Wagner, who gave me a goal, and Morris Halle, who taught me to never apologize.
+My journey in linguistics begins with the mentoring of Jennifer Cole. 
+Steve Anderson graciously agreed to serve on my committee. 
+Morris Halle told me to never apologize.
 
 I would like to thank my fellow students at Penn, but especially my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody. 
-Elsewhere, I have been shown a great deal of kindness by so many, but especially Morgan Sonderregger and Maryam Bakht.
 Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document. 
-Among non-linguists, my good friend Carly Fisher stands out for her unflagging support.
+Maryam Bakht kept me centered.
+
+My gratitude also goes out to the administrative assistants at 3401 Walnut and 255 S.
+36th, Lauris Olson at the Van Pelt Library, Katherine McEldoon for a helpful review of numerical cognition research, and Carly Fisher for her unflagging support and friendship.

### 2012-02-16 zipf_sim.py

@@ -1,187 +1,0 @@
-#!/usr/bin/env python
-""" Copyright (c) 2011 Kyle Gorman
-
- Permission is hereby granted, free of charge, to any person obtaining a copy
- of this software and associated documentation files (the "Software"), to deal
- in the Software without restriction, including without limitation the rights
- to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
- copies of the Software, and to permit persons to whom the Software is
- furnished to do so, subject to the following conditions:
-
- The above copyright notice and this permission notice shall be included in
- all copies or substantial portions of the Software.
-
- THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
- IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
- FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
- AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
- LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
- OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
- THE SOFTWARE.
-
- Zipfian distribution sampling simulations
- Kyle Gorman <kgorman@ling.upenn.edu>
- Usage is described below. """
-
-from sys import argv
-from bisect import bisect
-from random import random
-from math import log, fsum
-from collections import defaultdict
-
-
-def sparse_dist(N, alpha=1, beta=0):
-    return [(i + beta) ** -alpha for i in xrange(N, 0, -1)]
-
-
-def mean(x):
-    """ 
-    Return arithmetic mean of a list of continuous values 
-    """
-    return fsum(x) / len(x)
-
-
-def median(x, sorted=False):
-    """ 
-    Return median of an iterable of continuous values
-    """
-    my_x = x[:] # deep copy
-    if not sorted:
-        my_x.sort()
-    if len(x) % 2 == 0: # even 
-        l = len(my_x) / 2
-        return (my_x[l - 1] + my_x[l]) / 2.
-    else: # odd
-        return my_x[len(my_x) // 2]
-
-
-class ProbDist(object):
-
-    def __init__(self, item2prob=None):
-        """ 
[…]

### 2012-02-16 2.tex
Spacing in appendices.

@@
 % 2: Accidental and structural gaps in English syllable contact
+
+%lexicon
+The \emph{lexicon} has many senses: throughout, I use the term to refer to the set of underlying representations \citep[][269]{LANGUAGE}, but I depart from \citeauthor{LANGUAGE}'s understanding by further adopting the assumption that this set is merely a reflection of individual speakers' knowledge of their language, realized in the human mind; it has no further reality.
 

### 2012-02-17 2.1.3.tex
Rearranged and some writing on 2-5.

@@
 
-There is a long-standing debate concering whether English [ŋ] is a phoneme in its own right, demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}.
+There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right, as demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}.
 The strongest piece of evidence for the pure allophonic analysis (and an attentant process of /g/-\textsc{Deletion}) is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
-\citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is adopted here and formalized below in \S\ref{cnpasection}.
+\citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is adopted here and formalized below in Section \ref{cnpasection}.
 %\citealt[][62]{Halle1985a}),
@@
 
-I assume that the front onglide is not part of the onset except in the case where the onset would ortherwise be null (e.g., \emph{jun}[j]\emph{or}).
+I assume that the [j] onglide in words such as \emph{ass}[j]\emph{ume} is present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J.
+\citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}).
+The primary motivation for this assumption is the observation that presence or absence of this glide is contrastive (e.g.
+\emph{coup} $\sim$ \emph{queue}, \emph{booty} $\sim$ \emph{beauty}).
+I further assume that the front glide is not part of the onset except in the case where the onset would ortherwise be null (e.g., \emph{jun}[j]\emph{or}).
 %\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
@@
 However, when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm].
-This defective distribution of vowels following [j] follows if the onglide in this context is the first component of a phonological diphthong, and thus part of the nucleus (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}).
-External support for this account comes from the fact that in words like \emph{spew}, [ju\lm] may act as a ``unit'' in language games \citep{Nevins2003,Idsardi2005}, to the exclusion of the rest of preceding consonants.
-Finally, \citet[][42]{Clements1983} note
-that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}, in which I assume the glide is nuclear.
+This defective distribution of vowels following [j] follows if the onglide in this context is the first component of a phonological diphthong, and thus part of the nucleus (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
+\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [ju\lm] also suggests the glide is nuclear.
+There is a great deal of external support for this position as well.
+The [ju\lm] in words like \emph{spew} may pattern together
+in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003}, to the exclusion of the rest of preceding tautosyllabic consonants.
+The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
+Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.
+
+\ex VBR's complex onsets \citep[79--80, his transcriptions]{Buchwald2005}: \\
+\begin{tabular}{l l l}
+a. & kəræb  & `crab'  \\
+   & bəlid  & `bleed' \\
+b. & kəwin  & `queen' \\
+   & kəwoʊt & `quote' \\
+c. & kut    & `cute'  \\
+   & musɪk  & `music  \\ 
+\end{tabular}
+\xe
+
+\noindent
+VBR breaks up complex onsets, including back onglide clusters in (\nextx b).
+However, in (\nextx c), which contain the front onglide in standard English, no epenthesis occurs: the glide is simply absent.
+This also suggests that the front onglide is part of the onset, not the nucleus.
 
@@
 The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset.
-Whereas the front onglide shows only limited selectivity for preceding tautosylalbic consonants \citep{Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
-Unlike the front glide, syllable-initial [kw] may be followed by any vowel.
+Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
+Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}.
+Unlike [ju\lm], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.
+
 %Whereas [Cju\lm] syllables attracts stress, [kwV] syllables do not \citep[][162f.]{Davis1995}.

### 2012-02-17 2.2.1.tex
Spacing in appendices.
Rearranged and some writing on 2-5.

@@
 Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively.
-In the case of \citet{Greenberg1950}'s study, the primary generalization has no exceptions, but \citeauthor{Uhlenbeck1950}'s description of co-occurrence restrictions contains many exceptions.
-Many years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and suggests a statistical method to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.
+\citeauthor{Uhlenbeck1950}, for instance, puts forth co-occurrence dispreferences, rather than exceptionless generalizations of the sort given by \citet{Greenberg1950}.
+Years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and proposes a statistical technique, based on the chi-square test, to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.
 
-\citeauthor{Mester1988}'s technique, which has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English), is based on the chi-square test.
+\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English).
+The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented.
+Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.
 
-In this study, I use the Fisher exact test, which is isomorphic to this chi-square test.
-The chi-square test has been favored historically simply because it can be easily computed by hand, but it is equally easy to perform either test on a modern computer.
-Compared to the Fisher test, which is appropriate for small samples or rare events, the chi-square test has a serious defect, since it depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples \citep[e.g.,][142]{Mester1988}.
+\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
+\chi^2$ \xe
 
-The opposite problem, difficulties scaling to large samples, is an unfortunate defect of the Monte Carlo procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010}.
-This technique, which is the Monte Carlo isomorph of the one-tailed chi-square test, depends on the ability to generate random (i.e., unbiased) permutations of large lists.
-This is much more difficult than prior researchers have realized, as I explain below; less mathematically inclined readers may wish to skip ahead.
+%\footnote{This is the one-tailed version of the test.}
 
-Any permutation of a list $L$ can be thought of as a list $S$ such that
+In this study, I use the \citet{Fisher1934} exact test, which is isomorphic to this chi-square test.
+Compared to the chi-square test, the Fisher exact test $p$-value is somewhat more difficult to compute, but both are automated by modern statistical packages and can be computed rapidly.
+The major advantage of the Fisher test over the chi-square test is that it is appropriate for small samples or rare events, whereas the chi-square test depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples \citep[see][]{Gorman2012a}.
+%\citep[e.g.,][142]{Mester1988}.
 
-$S_i$ 
-the $i$th element in the permuted list is
+On the other end of the spectrum, the Monte Carlo statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010} does not scale to large samples.
+The Monte Carlo procedure is isomorphic to the one-tailed Fisher Exact test, and involves comparing observed co-occurrence counts to those produced by randomly generated samples.
+This requires the researcher to generate random permutations, something which is particularly difficult for large samples, as I will explain; disinterested readers may wish to skip the following paragraph.
 
-the permuted position of the $i$th elemnt of $L$. Since
+Any permutation of a sequence $L$ can be represented as a sequence $S$ of the same length, where the value of $S_i$ corresponds to the permuted position of $L_i$.
+Generating random permutations requires an unbiased method to generate all $N!$ of these lists with equal probability.
+Unfortunately, this is near impossible for $N$s of moderate size with current computational resources.
+Any pseudo-random number generator can be characterized by its ``period'', the number of pseduo-random numbers it can emit before repeating itself; this is also the upper limit for the number of random permutations that can be generated.
+The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a period of approximately $2^{48}$, a number which turns out to be smaller than $17!$.
+For the large samples, sometimes in the thousands, used by these researchers, virtually all the possible permutations can never be generated by this method.
+As a consequence, the shuffling procedure is far from random and introducing the possibility that the statistical test is biased in unpredictable ways.
 
-A permutation of a list of length $N$ can be thought of a vector $S$ such that $S_i$ is a counting number denoting the length of k
+As an example of the Fisher exact test as it is applied here, consider some of the Javanese co-occurrence facts considered by \citet{Mester1988}.
+\citeauthor{Mester1988} is at pains to show that there is a gradient dispreference for first and second root consonants which share the same major place.
+However, perfectly identical first and second consonants are exempt.
+Since
 
-as a list of $N$ unique counting numbers which are less than or equal to $N$;
-
-$N - 1$
-
-The problem is that even for moderately-sized lists, the number of possible permutations quickly exceeds modern computational capabilities.
-
-To explain why this is so requires a short discussion of
-
-The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a ``period'' of approximately $2 ^ 48$; this means that it repeats the same 281-trillion number pattern indefinitiely.
-
-. What this means has a period of approximately $2 ^ 48$. What this means has a period of approximately $2 ^ 48$. What this means has a period of approximately $2 ^ 48$. What this means
-
-use the Perl scripting language, which generally provides a pseudo-random number generator
-
-Both
-
-A list of $N$ elements has $N!$ different permutations. T
-
-When $N$ is large, t
-When $N$ is large, current computational
-
-this greatly exceeds the abilities of the pseudorandom number generators
-
-This often a collosal quantity
-
-When $N$ is large, this is a collosal quantita
-
-with a ``period'' of approximately $2 ^ 48$, i.e., it will repeat itself after
-
-I had a chance to write out my thoughts about the periodicity issue for Kessler's Monte Carlo method, and do a bit of research into Perl resources.
-
-%If you're lucky, you have a version of Perl that uses the drand48 library function.
-The period is 2 ^ 48, around 281 trillion; after 281 trillion draws, it just regenerates the same 281 trillion numbers in the same order.
-
-%A permutation of a list of N items can be thought of as a list of N - 1 random numbers; this is N - 1 because there's only one way to "permute" a one-element list (and in your method, only one outcome for the last element left).
-
-%Given a period P, there are only P unique sequences of random numbers.
-Unfortunately, even for N = 18, N!, the number of possible permutations is far larger than the period of the Perl random number generator.
-That means there are many random permutations that can never be generated.
-When N gets significantly larger, virtually all random permutations can never be reached.
-In your CELEX compound study, for instance, N = 4758, and thus requires a period of approximately 2 ^ 51260 according to the Stirling approximation.
-But Perl can only reach 2 ^ 48 of them, a tiny fraction thereof.
-
-%I only have browsed Kessler's book, but I saw that he spoke about how hard it was to calculate the Fisher exact test.
-With all due respect, I think he's just wrong.
-It's difficult to compute a p-value via Monte Carlo test that is provably unbiased (for instance, I'm unaware of any publically-available pseudo-random number generator with a period big enough for the CELEX compounds study), but the Fisher exact test, the one-tailed variant of which is isomorphic to the shuffle test, can be computed in the blink of an eye in any good stats environment.
-
- which in turn requires
-
-Since totally-identical consonants occur quite frequently,
+% APPLY THIS TO JAVANESE
 
@@
 \midrule
-Same major place      & 21       & 15         & 0.583      & \multirow{2}{*}{6.10\e{-06}} \\
-Different major place & 134      & 12         & 0.918 \\
+conforming & 134      & 12         & 0.918      & \multirow{2}{*}{6.095\e{-06}} \\
+violating  & 21       & 15         & 0.583 \\
 \bottomrule
@@
 It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.
-This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance ($\chi^2 = 64.525$, two-tailed $p = 9.5$\e{-16}).
 
-\citep[see also][]{Gorman2012}
-contains the counts
+This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance
 
-%\ex $ \lim_{n \rightarrow \infty} n! = 2 ^ {(n \ln{n} - n) \cdot \log_2{e}} $ \xe
+That this is highly unlikely to be due to chance is apparent both from the chi-square ($\chi^2 = 25.572$, two-tailed $p = 4.3$\e{-07}) and Fisher exaact tests ($p =  6.095$\e{-06}).
+
+I use two-tailed statistical tests throughout. 
+Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis.
+As
 
@@
 
-Applying the syllabification algorithm to the CELEX data produces 40 medial onset types and 22 medial coda types from 161 medial consonant clusters.
-
 It must be stressed that under current assumptions, this is independent of whether predictable prosodic representations (particularly coda and onset identity) are part of the memorized underlying representation \citep[e.g.,][]{Vaux2003}.
-
 It is an undeniable fact that URs need to be 
-
 The derivative nature of 
-W
-
-\citet{Ito1989a,Noske1993}
-
-The nature of 
-The prosodic characteristic of
-
-The Fisher exact test is isomorphic to the chi-square test
-
-and has been used by several researchers \citep{Zuraw2010}.
-
-Unlike the chi-square test, which uses an approximation which is exact in the limit (i.e., only with a hypothetical infinite sample), the Fisher exact test is robust to small sample sizes.
-
-The opposite problem, inability to scale to large samples, is a defect of the a statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence restrictions by \citet{Martin2007,Martin2011}.
-This procedure is a Monte Carlo approach to t
-
-Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis.
-As
+\citet{Ito1989a,Noske1992}
 
@@
 The suffix /-d/ forms regular pasts and denominal adjectives, and /-z/ forms regular noun plurals and genitives, and the 3sg.
-present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, \citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][211f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
+present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S.
+\citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][211f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
 
@@
 
-\ex Artificial language adaptations \citep[][394]{Wright1975}: \\
+\ex Artificial language adaptations \citep[][394, his transcriptions]{Wright1975}: \\
 \begin{tabular}{l l l l l l}
 a. & [gownp] & > & [gump] \\
-b. & [\v{j}umg] & > & [\v{j}úŋgə] \\
+b. & [ǰumg]  & > & [ǰúŋgə] \\
 c. & [ðʌŋd]  & > & [tɔŋg] \\
@@
 Yet, the independent evidence for a phonological process of nasal place assimilation provides the most parsiminious account of the adaptations seen in \citeauthor{Wright1975}'s study.
+J.
+\citet{Myers1993} also shows that \textsc{Coda Nasal Place Assimilation} is fed by speech errors:
+
+\ex Speech errors and \textsc{Coda Nasal Place Assimilation} \citep[][228]{Myers1993}: \\
+\begin{tabular}{l l l}
+   & \emph{error}  & \emph{target} \\
+a. & primps        & prints        \\
+b. & rand orker    & rank order    \\
+%  & [sprɪg]time for [hɪ̃ntlɘr] & springtime for Hitler \\
+c. & po[ŋ]cutation & computation   \\
+\end{tabular}
+\xe
 
@@
 
-\ex Degemination in the /-t/ past (\citealp[][105]{Halle1985a}, \citealp[][492]{Myers1987}): \vspace{6pt} \\
+\ex Degemination in the /-t/ past (\citealp[][105]{Halle1985a}, S. \citealp[][492]{Myers1987}): \vspace{6pt} \\
 \begin{tabular}{l l l l l}

### 2012-02-17 2.2.3.tex
Rearranged and some writing on 2-5.

@@
 
-\ex $\displaystyle f(r~; C, \alpha) = \frac{C}{r^\alpha}$ \xe
+\ex $\displaystyle f(r; C, \alpha) = \frac{C}{r^\alpha}$ \xe
 
@@
 That clusters (and the codas and onsets that make them up conforms to Zipf's law is not itself a deep observation.
-Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also $n$-grams \citep{Baroni2009} and of randomly-generated texts \citep{Miller1957,Li1992}.
+Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also $n$-grams \citep{Baroni2009} and of non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} or randomly-generated texts \citep{Miller1957,Li1992}.
 The import of this statistical property for the study of cluster phonotactics is that it makes it difficult, on statistical grounds alone, to determine which unobserved events are excluded, and which are accidentally missing.

### 2012-02-17 3.1.1.tex
Rearranged and some writing on 2-5.

% 3.1.1: Alternation vs. phonotactic

\subsubsection{Roundness harmony}

\subsubsection{Labial attraction}

\citet{Zuraw2000}
\citet{Clements1982}
\citet{Inkelas1997}

### 2012-02-17 3.1.2.tex
Rearranged and some writing on 2-5.

% 3.1.2: Lexical statistics

This is essentially the hypothesis pursued in the previous chapter.

\citet{TELL}
\citet{Inkelas2001}

%We might, for instance, consider methods based not on statistical \citep{Yang2005,Gorman2011b}.

### 2012-02-17 3.1.tex
Rearranged and some writing on 2-5.

% 3.1: Wordlikeness predictions

\citet{Zimmer1969}

### 2012-02-17 3.2.2.tex
Rearranged and some writing on 2-5.

% 3.2.2: Wordlikeness results

\citet{Goodman1954}

\ex $\displaystyle \gamma = \frac{C - D}{C + D}$ \xe

\subsubsection{Roundness harmony}

($\gamma = 0.597$, $p = 5.2$\e{-29})

Roundness harmony wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
\midrule
pemez & 30                   & pemaz & 2 \\
teriz & 28                   & terız & 3 \\
tokaz & 26                   & tokez & 6 \\
tipez & 24                   & tipaz & 8 \\
terüz & 19                   & teruz & 13 \\
\bottomrule
\end{tabular}

\subsubsection{Labial harmony}

Labial attraction wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
\begin{tabular}{l r l r}
\toprule
\multicolumn{2}{l}{aPu} & \multicolumn{2}{l}{aPı} \\
\midrule
tafuz & 21              & tafız & 11 \\
pamuz & 17              & pamız & 15 \\
tapuz & 17              & tapız & 15 \\
mavuz & 16              & mavız & 16 \\
tabuz & 16              & tabız & 16 \\
\bottomrule
\end{tabular}

($\gamma = 0.087$, $p = 0.101$)

### 2012-02-17 3.3.1.tex
Rearranged and some writing on 2-5.

% 3.3.1: Naturalness

\citet{Becker2011}

This has an ad hoc nature to it; 
There are two senses in which this objection is ad hoc. 
First, \citeauthor{Becker2011} appeal to no particular theory of the naturalness of processes or static constraints which excludes \textsc{Labial Attraction}. 
Secondly, this appears to be a minority view: \textsc{Labial Attraction} was considered a true generalization by early specialists
\citep[e.g.,][]{Lees1966a}, and despite \citeauthor{Zimmer1969}'s suggestive psycholinguistic results, reviewed above, it also been treated as a plausible constraint by later theorists \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995a,Zuraw2000}.
Further, there is a real danger that if the label ``unnatural'' Labial Attraction} describes an impossible structural change or structural description, that one will fail to account for earlier forms of Turkish or sound changes therein.

### 2012-02-17 3.tex
Rearranged and some writing on 2-5.

% 3: Static and derived constraints on Turkish vowel harmony

In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon may occur

This brief chapter argues that even statistically valid phonotactic generalizations are ignroed by speakers. Two phonotactic generalizations over the Turksih lexicon are shown to be equally statistically reliable, but only the generalization which is derived by a phonological alternation influence wordlikeness judgements.

Research on gradient phonotactics has usually rejected this possibilty implicit
ly.

After reviewing co-occurrence restrictions in Gitksan roots, \citet{Brown2010} rejects this explicitly. He writes:

\begin{quotation}
\ldots the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
\end{quotation}

%Classical Arabic adjectives often have stative verbs in which the root is imposed onto the template CaCuCa:

%\ex Arabic verbs of coming into being: \\
%\begin{tabular}{r l l l}
%a. & kabura & `become big'       & (cf. \emph{kabiːr} `big') \\
%b. & saʁiir & `become small'     & (cf. \emph{saʁiːr} `small') \\
%c. & ħasuna & `become beautiful' & (cf. \emph{ħasan} `handsome') \\
%\end{tabular}
%\xe

### 2012-02-17 4.1.1.tex
Rearranged and some writing on 2-5.

% 4.1.1: Gradience in judgements

at least as far back as \emph{LSLT} \citep{LSLT}

%and has become a recent feature of the discussion
is that non-gradient models are totally incompatible with psycholinguistic results.

\subsubsection{Task model}

\citet[][16]{Schutze1996}

\citet{Schutze2005}
\citet{Schutze2011}

%\citet{Coleman1996}

\subsubsection{Falsifiability}

\citet{Armstrong1983}

% wordlikeness critique
\citet{Yang2008a}

a mix of ``declarative'' knowledge,
corresponding to rote facts and thus the lexicon,
and ``procedural'' knowledge, or knowledge of generative procedures 
(in the sense of J. \citealt{Anderson1993}).

% math cog
\citet{Logan1988}
\citet{Sfard1991}
\citet{Delazer2005}

### 2012-02-17 4.1.2.tex
Rearranged and some writing on 2-5.

% 4.1.2: The ontology of wordlikeness

\subsubsection{Gross phonotactic violation}

position in \citet{SPE}

\citet{Shibatani1972}
\citet{Hooper1973}

peripheral \citep[in the sense of][81]{LGB}

As an example of this applied ot phonotactics, 
\citet{Borowsky1989} assumes that there are peripheral morph-internal syllable contact clusters.

\subsubsection{Phonological statistics}

\citet{SPE}

This statement is made very strongly by \citet{Coleman1996}

\subsubsection{Lexical similarity}

\citet{LSLT}

\citet{Greenberg1964}
\citet{Ohala1986b}

A recent, very articulated model is presented by \citet{Bailey2001}.

% simple model
This is simply because more complex models fail to

% similarity vs. rules
% as a final note, \citet{Hahn1998}

### 2012-02-17 4.2.1.tex
Rearranged and some writing on 2-5.

% 4.2.1: Assessing gradience

\subsubsection{Response distributions}

This is shown in Figure \ref{density}.

\begin{figure}
\centering
\includegraphics{density.pdf}
\caption{Density of responses}
\label{density}
\end{figure}

\subsubsection{Prediction distributions}

\citet{Coltheart1977}

\citet{Hayes2008a}
\citet{Albright2009}

% stats thereof
%\citet{Schilling2002}
%\citet{Helguerro1904}
%\citet{Cohen1956}

% EM
%\citet{EM}

% -2 log like ratio test

### 2012-02-17 4.2.2.tex
Rearranged and some writing on 2-5.

% 4.2.2: Assessing models

\subsubsection{Basic correlations}
\subsubsection{Intramodal correlations}
\subsubsection{Residual correlations}

### 2012-02-17 4.tex
Rearranged and some writing on 2-5.

% 4: Gradient and categorical aspects of wordlikeness judgements

(though few if any phonologists regard $n$-gram segment probabilities as a core component of phonology)

\begin{quotation}
``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
\end{quotation}

### 2012-03-01 1.1.1.tex
Drastic restructuring and some writing on chapters 2-3.

% 1.1.1: Structuralism

Since, with the possible exception of Roman Jakobson (who might be labeled an early generativist), the mentalistic interpretation of structural linguistic analyses was anathema, one must take care to not assume that a description of MSCs is something that these linguists would posit as part of speakers' knowledge of language.

\subsubsection{Morpheme boundaries in phonology}

bloomfield

%While \citeauthor{Bloomfield1930} is known for introducing morphemic structure into phonology, a member of the Prague circle developed an arguably more sophisticated approach only two years later.

\citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture. The vast majority of Russian infinitives consist of the stem followed by a theme vowel and /t\pal/. The reflexive is marked wih a final /-sə/ which feeds a general process of regressive voicing assimilation. In the reflexive infinitive, /\ldots t\pal-s\ldots/ is realized [ts], with no palatalization. However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\nextx c)---is preserved in the reflexive imperative.

\ex Russian reflexives \citep[after][]{Jakobson1932}:\footnote{
I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses an abstract phonemic transcription. 
The broad transcription below corresponds to the speech of Lev Blumenfeld, who I thank for help with this data.
} \\
\begin{tabular}{l l l l} %\toprule
   &  infinitive      & imperative \\ %\midrule
a. & [slav\pal itsə]  & [slaf\pal s\pal ə]  & `be glorious'    \\
   & [upram\pal itsə] & [upram\pal s\pal ə] & `be stubborn'    \\
b. & [kras\pal itsə]  & [kras\pal s\pal ə]  & `put on makeup'  \\
   & [ʒar\pal itsə]   & [ʒar\pal s\pal ə]   & `roast'          \\
c. & [zəbytsə]        & [zəbut\pal s\pal ə] & `forget' \\ %\bottomrule
%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
\end{tabular} \xe

\noindent
The forms in (\lastx c) have different outcomes for their /t\pal-s/ clusters. \citeauthor{Jakobson1932} proposes that the preservation of palatalization is a special property of the imperative. %A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.

\subsubsection{The morph as a constraint domain}

A corrolary of this hypothesis is that the phoneme sequences found within morphs may be distinct from those which span multiple morphs.

This was apparent to structuralists

the phonologists 
This is apparent from 
a discussion of constraints on vowel sequences in Mixteco given by \citet{Pike1947b}.
It is apparent that \citeauthor{Pike1947b}'s constraints on vowel sequences are not intended to hold across morph boundaries,

according to the phonological analysis of a Mixteco text published a few years earlier \citep{Pike1944}.

\ex Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}: \\
\begin{tabular}{r l l l} %\toprule
   & MSC & complex exception \\ %\midrule
%a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
%b. & *{C}\textipa{@}{C}e & [n\`i-k\`ə bə-de] & `who entered'        \\
%c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
%d. & *{C}i{C}e & [te-n\`i-kee-t\`ə] & `and went away'      \\ %\toprule
%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
a. & *{C}a{C}e & [ká\textsuperscript{n}dee] & `kept \ldots inside' \\
b. & *{C}ə{C}e & [nìk\`əbəde] & `who entered'        \\
c. & *{C}e{C}i & [teníke\textsuperscript{n}da] & `was walking         \\
d. & *{C}i{C}e & [tenìkeet\`ə] & `and went away'      \\ 
\end{tabular} \xe

\noindent
\citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a morpheme discovery routine.

\subsubsection{Biuniqueness and neutralization}

joos
harris
chomsky

### 2012-03-01 1.1.2.tex
Drastic restructuring and some writing on chapters 2-3.

% 1.1.2: Early generativism

\subsubsection{\citealt{Stanley1967}}

inventory constraints
sequence constraints

\subsubsection{Surface constraints and occultation}

shibatani
stampe
dell

\emph{A}[mt] `office', \emph{Anbau} `cultivation'
\emph{ei}[n.g]\emph{reifen} `to intervene',

\subsubsection{The duplication problem}

inventory problems 
dell, chung, but also problems

conspiracies

### 2012-03-01 1.2.4.tex
Drastic restructuring and some writing on chapters 2-3.

% 1.2.4: Universal constraints?
% This may be cut.

(see \citealt[][323f.]{Hockett1947}, \citealt[][415f.]{Nida1948}, \citealt[][50f.]{Anderson1992}, \citealt[][36f.]{Stump2001a}). Irrespective of one's position on this debate, it does not seem correct to extend this hypothesis to lexical roots. For instance, there

imperfect subjunctive  & īrem     & īrēs     & īret     & īrēmus     & īrētis     & īrent \\
pluperfect subjunctive & issem    & issēs    & isset    & issēmus    & issētis    & issent \\
imperfect subjunctive  & venīrem  & venīrēs  & venīret  & venīrēmus  & venīrētis  & venīrent \\
pluperfect subjunctive & vēnissem & vēnissēs & vēnisset & vēnissēmus & vēnissētis & vēnissent \\

That this is not the

While it might be possible to analyze \emph{ueni:re} as ``athematic'' /weni:-re/ and \emph{i:re} as /i:-re/, the frequentive \emph{uentita:re} `come often, be wont to come', formed with the \emph{-tit-} frequentive suffix (see \citet[][\S263]{Allen1903}), which selects for the first conjugation (cf.~\emph{agere} `act, make' vs. \emph{actita:re} `act, make often/repeatedly') suggests /wen-/.

It is also interesting to note that the same pattern has emerged in the history of French for a set of verbs which do not share this property in Latin. and the future and conditional indicative forms of \emph{aller} `go' and \emph{cuire} `to cook' in French.\footnote{\emph{Cuire}, interestingly, is not descended from the same conjugation as Latin \emph{i:re}, but rather from Latin \emph{coquere}; thus this syncretism is not simply an etymological relic of Latin. The same holds for other French verbs that inflect in the same manner, such as \emph{conduire} `drive (a vehicle); behave' and \emph{d\'etruire} `destroy'.}

future indicative      & irai   & iras   & ira      & irons    & irez    & iront \\
conditional indicative & irais  & irais  & irait    & irions   & iriez   & iraient \\
future indicative      & cuirai & cuiras  & cuira   & cuirons  & cuirez  & cuiront \\
conditional indicative & curias & cuirais & cuirait & cuirions & cuiriez & cuiront \\

One could even conceive of a largely vacuous principle on morpheme structure which simply requires that some subset of morphs not be null.

%Anttila 2008
%Dmitrieva et al. 2008a,b
%Duanmu 2009
%Berkley 1994a,b, 2000
%Buckley 1997
%Coetzee 2008, Coetzee and Pater 2008
%Goad 2011
%Graff & Jaeger in press,
%Hammond 1999
%Colavin 2010,
%Frisch 1996, Frisch et al. 2004
%Hayes & Wilson 2008
%Kessler et al. 1997
%Martin 2007
%McGowan 2011
%Mester 1986
%Fudge 1969
%Padgett 1992
%Pierrehumbert 1993, 1994

### 2012-03-01 1.tex
Drastic restructuring and some writing on chapters 2-3.

\chapter{Morpheme structure constraints in generative grammar}
\label{msc}

The first part of this text concerns the empirical substance of the theory of \textsc{Morpheme Structure Constraints} (MSCs), that is, language-specific constraints on the shapes of morphs stored in lexical memory.
Evidence is presented to support the claim that such constraints are solely derived by other components of grammar, and do not have an ontology of their own.

The first section of this chapter outlines a brief history of MSCs.
The following section draws upon this material to present the above claim as a null hypothesis, and presents conditions for falsifying it.
The final section outlines the remainder of Part I (chapters 2--4), which present evidence in support of this hypothesis.

% it is not clear what follows from a phonotactic description
% pater on NC
% can't be right tho, see stanley, kisseberth, clayton, shibatani, hooper, etc.

\section{A brief history of MSCs} \input{1.1}
\section{The null hypothesis} \input{1.2}
\section{Outline of Part I} \input{1.3}

### 2012-03-01 2.1.3.tex
Drastic restructuring and some writing on chapters 2-3.

@@
 When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}.
-However, when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm].
+However, when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm], and speakers judge other following vowels to be anomalous in nonce words \citep{Moreland2009}.
 This defective distribution of vowels following [j] follows if the onglide in this context is the first component of a phonological diphthong, and thus part of the nucleus (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
@@
 There is a great deal of external support for this position as well.
-The [ju\lm] in words like \emph{spew} may pattern together
-in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003}, to the exclusion of the rest of preceding tautosyllabic consonants.
+The [ju\lm] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003}, to the exclusion of the rest of preceding tautosyllabic consonants.
 The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.

### 2012-03-01 2.2.1.tex
Drastic restructuring and some writing on chapters 2-3.

@@
 \subsubsection{Statistical technique}
+\label{stattech}
 
@@
 \ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
-\chi^2$ \xe
+χ^2$ \xe
 
@@
 present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S.
-\citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][211f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
+\citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][210f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
 
@@
 \ex \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178): \\
-\xymatrix@C=6pt@R=12pt{
+\xymatrix@R=12pt{
 \txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{..}[dr] & \\
@@
 \txt{[$+$\textsc{Obstruent}]} & \txt{[$+$\textsc{Obstruent}]}}
-\label{ovar}
-\xe
+\label{ovar} \xe
 
@@
 \bottomrule
-\end{tabular}
-\xe
+\end{tabular} \xe
 
@@
 \label{CNPA}
-\xymatrix@C=6pt{
+\xymatrix@R=12pt{
                           & \txt{\textsc{Place}}\ar@{-}[d]\ar@{..}[dl] \\
@@
 
-There is some evidence that even young infants show a preference for nonce words that conform to this generalization.
-\citet{Jusczyk2003}, \citet{Mattys1999}, and \citet{Jusczyk2002} play nonce words containing nasal-obstruent clusters to 4.5, 9, and 10-month-old infants (respectively) acquiring English, All three studies find that infants prefer to listen to stimuli which conform to the above generalization (e.g., \emph{u}[m.b]\emph{o}) over those which do not (e.g., \emph{u}[n.b]\emph{o}).
+There is some evidence that even young infants show a preference for nonce words that conform to this generalization. %\citet{Jusczyk2003}
+\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} play nonce words containing nasal-obstruent clusters to 4.5, 9, and 10-month-old infants (respectively) acquiring English, All three studies find that infants prefer to listen to stimuli which conform to the above generalization (e.g., \emph{u}[m.b]\emph{o}) over those which do not (e.g., \emph{u}[n.b]\emph{o}).
 \citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''.
@@
 %}
-\xymatrix@C=6pt{
+\xymatrix@R=12pt{
 %{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\

### 2012-03-01 2.2.3.tex
Drastic restructuring and some writing on chapters 2-3.

@@
 That clusters (and the codas and onsets that make them up conforms to Zipf's law is not itself a deep observation.
-Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also $n$-grams \citep{Baroni2009} and of non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} or randomly-generated texts \citep{Miller1957,Li1992}.
+Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also word \citep{Baroni2009} and phoneme \citep{Daland2011a} n-grams, and tokens in non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} or randomly-generated texts \citep{Miller1957,Li1992}.
 The import of this statistical property for the study of cluster phonotactics is that it makes it difficult, on statistical grounds alone, to determine which unobserved events are excluded, and which are accidentally missing.

### 2012-03-01 3.1.1.tex
Drastic restructuring and some writing on chapters 2-3.

@@
-% 3.1.1: Alternation vs. phonotactic
+% 3.1.1: Backness harmony
 
-\subsubsection{Roundness harmony}
+\subsubsection{Phonological description}
 
-\subsubsection{Labial attraction}
+One of the most salient aspects of Turkish phonology is the presence of vowel harmony.
+\citeauthor{Lees1966a} (\citeyear[][284]{Lees1966b}, \citeyear[][35]{Lees1966a}) proposes a number of phonological rules to describe the harmony processes that act on roots and regularly spread rightward to suffixes (Turkish has no prefixes); one such rule consists of rightward spreading of the feature [\textsc{Back}].
 
-\citet{Zuraw2000}
-\citet{Clements1982}
-\citet{Inkelas1997}
+\ex \textsc{Backness Harmony} \citep[e.g.,][229]{Clements1982}: \\
+\xymatrix@R=12pt{
+\txt{[α \textsc{Back]}}\ar@{-}[d]\ar@{..}[drr] &             \\
+\txt{V}                                        & \txt{C$_0$} & \txt{V}}
+\xe
+
+%\citet{Clements1982}
+%\citet{Inkelas1997}
+
+This rule accounts for much of the suffix allomorphy found in Turkish.
+The dat.sg.
+(and the vowel of the gen.sg.) is always a [$+$\textsc{High}] vowel, and in general, it agrees with the final stem vowel for backness and roundness.
+Similarly, the vowel of the nom.pl.
+is [$-$\textsc{High}, $-$\textsc{Round}], and takes its backness from the stem-final vowel.
+Similar patterns obtain for many other Turkish suffixes.
+
+\ex Turkish nominal suffix allomorphy: \vspace{6pt} \\
+\begin{tabular}{l l l l l l l}
+   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.} & \emph{nom.pl.} \\
+a. & ip             & ipi            & ipin           & ipler          & `rope' & (C\&S:216) \\
+   & el             & eli            & elin           & eller          & `hand'    \\
+   & kız            & kızı           & kızın          & kilzar         & `girl'    \\
+   & sap            & sapı           & sapın          & saplar         & `stalk'   \\
+   & yüz            & yüzü           & yüzün          & yüzler         & `face'    \\
+   & köy            & köyü           & köyün          & köyler         & `village' \\
+   & pul            & pulu           & pulun          & pullar         & `stamp'   \\
+   & son            & sonu           & sonun          & sonlar         & `end'     \\
+b. & neden          & nedeni         & nedenin        & nedenler       & `reason'  & (TELL) \\
+   & boğaz          & boğazı         & boğazın        & boğazlar       & `throat'  \\
+   & kiler          & kileri         & kilerin        & kilerler       & `pantry'  \\
+   & pelür          & pelürü         & pelürün        & pelürler       & `tissue paper' \\
+   & sapık          & sapığı         & sapığın        & sapıklar       & `pervert' \\
+\end{tabular} \xe
+
+\noindent
+If the vowel of the nom.pl.
+suffix is specified as [$-$\textsc{High}, $-$\textsc{Round}], then the \emph{a} $\sim$ \emph{e} pattern follows naturally from \textsc{Backness Harmony}.
+If the dat.sg.
+suffix is a vowel specified only as [$+$\textsc{High}], then \textsc{Backness Harmony}, along with the rightward spreading of \textsc{Round}, will produce the correct results.
+
+There are a few complications, however.
+First, while most polysyllabic roots conform to \textsc{Backness Harmony} (\lastx b), not all do.
+In this case, suffix vowels generally harmonize with the final root vowel (\nextx a).
+Finally, here is a small class of nouns which exhibit [$-$\textsc{Back}] suffix vowels despite the fact that their final root vowel is [$+$\textsc{Back}] (\nextx b).
+
+\ex Turkish nominal suffix allomorphy: \vspace{6pt} \\
+\begin{tabular}{l l l l l l l}
+   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.} & \emph{nom.pl.} \\
+a. & mezar          & mezarı         & mezarın        & mezarlar       & `grave' & (TELL) \\
+   & model          & modeli         & modelin        & modeller       & `model' \\
+   & silah          & silahı         & silahın        & silahlar       & `weapon'     \\
+   & memur          & memuru         & memurun        & memurlar       & `bureaucrat' \\
+   & sabun          & sabunu         & sabunun        & sabunlar       & `soap'       \\
+b. & saat           & saati          & saatin         & saatler        & `hour, clock' \\
+   & harf           & harfi          & harfin         & harfler        & `(alphabetic) letter' \\ %& \citep{Goksel2005}
+   & etol           & etolü          & etolün         & etoller        & `fur stole' \\
+\end{tabular} \xe
+
+While it is uncontroversial that disharmonic suffixes (\lastx b) constitute no more than sporadic exceptions to \textsc{Backness Harmony}, the status of root disharmony (\lastx a) has been the subject of much debate.
+
+1. \citet[][289]{Anderson1974} 
+2. \citet[][]{Kiparsky1981} makes a similar objection regarding Finnish vowel harmony.
+3. \citet{Clements1982}
+4. \citet{Inkelas1997}
+
+have identified (\lastx b) as sporadic exceptions to 
+A number of objections can be made for
+
+\subsubsection{Lexical statistics}
+
+\citet{TELL}
+%\citet{Inkelas2001}
+
+\ex Lexical effects of \textsc{Obstruent Voice Assimilation}: \vspace{6pt} \\
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 80 & 370 & 0.178 & \multirow{2}{*}{1.1\e{-11}}\\
+violating  &  6 & 264 & 0.022 \\
+\bottomrule
+\end{tabular} \xe
+
+\subsubsection{External evidence}
+
+The data reviewed thus far leaves open the possibility that speakers do not internalize the generalization that roots conform to \textsc{Backness Harmony}.
+One piece of evidence against this hypothesis comes from a Turkic language game \citep{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but it corresponds to a morphological process native to the related langauge Tuvan, and is quickly taught to even young Turkish speakers.
+The process consists of reduplication of the base with the base-initial vowel replaced by \emph{a} or \emph{u}, which conveys a sense of ``vagueness'' or ``jocularity''.
+Reduplication interacts with root harmony in both Tuvan and Turkish.
+When the base is harmonically [$-$\textsc{Back}], the non-initial vowels of the reduplicant do not surface faithfully: they surface as [$+$\textsc{Back}] (\nextx a).
+
+\ex Turkish reduplication game \citep[][231]{Harrison2001}: \\
+\begin{tabular}{l l l l}
+a. & kibrit & kibrit-kabrıt & `match'    \\
+   & bütün  & bütün-batın   & `whole'    \\
+b. & mali   & mali-muli     & `Mali'     \\
+   & butik  & butik-batik   & `boutique' \\
+\end{tabular} \xe
+
+\noindent
+Interestingly, reharmonization is absent in disharmonic bases (\lastx b).
+This is not an isolated fact about Turkic, as \citeauthor{Harrison2001} note similar outcomes in an unrelated Finnish language game \citep{Campbell1986}.
+The simplest explanation of this contrast is to treat \textsc{Backness Harmony} as a feature-filling process, as proposed by \citet{Clements1982} and \citet{Inkelas1997}.
+In harmonic roots, the non-initial vowels are underspecified for \textsc{Back} and are spelled out by \textsc{Backness Harmony}.
+In reduplication, this produces reharmonization.
+In disharmonic roots, the vowels are fully specified and do not meet the structural description of \textsc{Backness Harmony}.
+
+A number of studies have found that speakers of Finnish use disharmony as a clue to segment running speech.\footnote{Thanks to Charles Yang for pointing out the relevance of these studies to me.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which are harmonic for the feature \textsc{Back}.
+These stimuli are auditorily presented to subjects who are asked to press a button when the nonce trisyllable ends with a target nonce disyllable, or a real disyllabic word.
+Speakers are quicker to press the button when the prefix and disyllabic word disagree for \textsc{Back}.
+This is taken to indicate that speakers are attuned to the fact that disharmonic transistions are good predictors of word boundaries.
+On the other hand, if this also indicates that speakers recognize that harmonic transistions are more likely to be root-internal, then harmony is active not just where it is triggered by alternations (i.e., in Finnish affixes) but also in roots.
+
+\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect on word-spotting as it does in Finnish: speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when it preceded by the disharmonic pseudo-prefix \emph{gölü-} than when it is prefixed with harmonic nonce prefix \emph{golu-}.
+The effect of harmony does not obtain for speakers of French, which lacks vowel harmony.
+As in Finnish, the results imply speakers have internalized the predominance of root-internal harmony.
+
+It seems that root-internal harmony is not only present in adults, but in fact is learned very early.
+The pseudoword spotting experiment has been adapted for 9-month-old Turkish infants by \citet{Kampen2008}.
+Infants are familiarized with harmonic disyllabic pseudowords bearing a pseudo-prefix, which may be harmonic or disharmonic.
+At test time, the infants are played the disyllabic pseudowords in isolation using the head turn preference paradigm.
+Infants show a preference to listen to those pseudowords which were familiarized with a disharmonic pseudo-prefix over those which were familiarized with a harmonic pseudo-prefix.
+This preference is not observed in 9-month-old infants learning German, which lacks vowel harmony.
+Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants do not.
+
+%We might suggest methods not based only on statistical significance, but rather substantive computational properties of the linguistic objects in question \citep[e.g.,][]{Yang2005a}.

### 2012-03-01 3.1.2.tex
Drastic restructuring and some writing on chapters 2-3.

@@
-% 3.1.2: Lexical statistics
+% 3.1.2: Labial attraction
 
-This is essentially the hypothesis pursued in the previous chapter.
+I am using the dotted association line in a deliberately ambiguous fashion in this chapter.
+The autosegmental structures represent formalizations of the constraints, but I intend to remain ambiguous as to their ontology and do not intend them to be interpreted as derivational rules.
+These questions are taken up more carefully below.
 
+\citet{Clements1982}
+
+\citet{Zimmer1969}, \citet{Zuraw2000}, and \citet{Inkelas2001} all suggest
+
+\subsubsection{Phonological description}
+
+\citet[][36]{Lees1966a} describes \textsc{Labial Attraction} as a process by which ``a high, short harmonic vowel is rounded in the second syllable of a disyllabic word whose first vowel is /a/, and whose medial consonant cluster contains a labial /p, b, m, v/, and then it is de-harmonified''.
+This description is transalted into an autosegmental rule below.
+
+\ex \textsc{Labial Attraction} (after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}): \\
+\xymatrix@R=12pt{
+\txt{[$-$\textsc{Round}]}\ar@{-}[d] &                            & \txt{[$+$\textsc{Round}]}\ar@{..}[d] \\
+\txt{V}\ar@{-}[d]                    & \txt{C}\ar@{-}[d]          & \txt{V}\ar@{-}[d] \\
+\txt{[$-$\textsc{High}, $+$\textsc{Back}]}             & \txt{[$+$\textsc{Labial}]} & \txt{[$+$\textsc{High}, $+$\textsc{Back}]}}
+\xe
+
+\noindent
+A few examples of words conforming to this generalization are \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', and \emph{samur} `sable' \citep[][285]{Lees1966b}; \citeauthor{Lees1966b} also notes the existence of exceptions like \emph{tavır} `mode', but writes they are ``surprisingly rare'' (ibid., 286); \citet[][225]{Clements1982} lists a number of additional exceptions.
+
+\citeauthor{Clements1982} propose that there is an additional type of exception to \textsc{Labial Attraction}.
+
+\begin{quotation}
+Even more decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots~aCu~\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots We conclude that there is no systematic restriction on the set of consonants that may occur medially in rotos of the form /\ldots~aCu~\ldots/.
+\citep[][225]{Clements1982}
+\end{quotation}
+
+\noindent
+\citeauthor{Clements1982} are suggesting that \textsc{Labial Attraction} implies that \emph{aTu}, where \emph{T} represents a non-labial consonant, should be infrequent, but this fact is not inconsistent with the formulation of the constraint by \citet{Lees1966a,Lees1996b} and \citet{Zimmer1969}; \emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description.
+
+There is reason to believe that \textsc{Labial Attraction} is at best a static constraint: it never applies across morpheme boundaries.
+If, contrary to fact, there was a \textsc{Labial Attraction} alternation, then one would expect, for example, that the gen.sg.
+of \emph{sap} `stalk' would be *\emph{sapun} instead of the observed \emph{sapın}.
+Additional potential evidence for the static nature of \textsc{Labial Attraction} is provided by \citeauthor{Clements1982}, who note a class of bisyllabic words in which the second vowel, always [$+$\textsc{High}], alternates with zero.
+
+\ex High-vowel/zero alternations \citep[][243]{Clements1982}: \\
+\begin{tabular}{l l l l}
+   & nom.sg. & 3.poss. \\
+a. & fikir   & fikri & `idea' \\ 
+   & hüküm   & hükmü & `judgement' \\ 
+%   & filim   & filmi & `film' & \citep[][178]{Inkelas2001} \\
+%   & kojun   & kojun & `bosom' \\
+b. & vakit  & vakti  & `time' \\
+   & rahim  & rahmi  & `womb' \\
+\end{tabular}
+\xe
+
+\noindent 
+The vowel is present in the nom.sg. form and absent in the 3.poss. form. The examples in (\lastx b) shows that it need not harmonize with the preceding vowel
+\citeauthor{Inkelas2001} (ibid, 178) note that \textsc{Labial Attraction} does not affect this vowel (e.g., \emph{sabır}/\emph{sabrı} `patience'), citing this as evidence against \textsc{Labial Attraction}.
+However, it is not a foregone conclusion that \textsc{Labial Attraction} should apply to the ``unstable'' vowel.
+Since it is as a morpheme structure constraint, this vowel must be underlying present and deleted in the 3.poss.; \citeauthor{Clements1982} prefer the converse analysis, in which epenthesis applies in the nom.sg., though they note that either approach requires lexically arbitrary diacritics.
+I conclude that unstable vowels provide no evidence for or against \textsc{Labial Attraction}.
+
+\begin{quotation}
+Lee's rule of \textsc{Labial Attraction}\ldots is not a real generalization about the Turkish lexicon.
+It is not true synchronically, either of native or nonnative items; nor, according to the historical and dialectical literature, does \textsc{Labial Attraction} appear to have been true at any stage going back as far as Old Turkic.
+\citep[][196]{Inkelas2001}
+\end{quotation}
+
+\begin{quotation}
+Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
+\citep[][196, emphasis in original]{Inkelas2001}
+\end{quotation}
+
+\citet{Inkelas1997}
+\citet{Inkelas2001}
 \citet{TELL}
-\citet{Inkelas2001}
 
-%We might, for instance, consider methods based not on statistical \citep{Yang2005,Gorman2011b}.
+\subsubsection{Lexical statistics}
+
+%##TELL: 31,236
+%#Lexicon: 25,164
+
+\ex Lexical effects of \textsc{Labial Attraction} \citep[][186]{Inkelas2001}: \vspace{6pt} \\
+\begin{tabular}{l r r r | r r r}
+\toprule
+Corpus  & aPu & aPı & \% & aTu & aTı & $p$-value   \\ % & corpus size
+\midrule
+%TELL    & 378 & 248 & 446 & 1,140 & 2.83\e{-44} \\ % & 31,236 \\
+%TELL    & 77  & 30 & 72.6 & 115 & 356 & ?\e{-20} \\
+%Lexicon & 108 & 57 & 65.5 & 155 & 533 & 6.6\e{-25} \\
+\bottomrule
+\end{tabular} \xe
+
+BROAD labial attraction stats:
+
+TELL full:
+585 aPu (favored)
+230 aPI (disfavored)
+1,137 aTu
+1,461 aTI
+71.8\% labial attraction
+p = 3.4\e{-45}
+
+Lexicon:
+1,175 aPu (favored)
+446 aPI (disfavored)
+1,939 aTu
+2,878 aTI
+72.5\% labial attraction
+p = 3.1\e{-114}
+
+STUPID attraction stats:
+
+TELL full:
+267 aPu (favored)
+546 aPI (disfavored)
+346 aTu
+327 aTI
+67.1\% attraction
+p = 5.2\e{-13}
+
+Lexicon:
+357 aPu (favored)
+874 aPI (disfavored)
+488 aTu
+396 aTI
+71% attraction
+p = 9.7\e{-34}
+
+zimmer: 319-320 
+inkelas et al.: 196
+
+\subsubsection{External evidence}
+
+Beyond the experimental results of \citet{Zimmer1969} reviewed in the next section, there is no external evidence for or against \textsc{Labial Attraction}.
+
+%% HERE THERE BE DRAGONS
+%\ex Examples of \textsc{Labial Attraction} \citep[][285]{Lees1966b}: \\
+%\begin{tabular}{l l l}
+%a. & çapul &  `raid \\
+%b. & sabur & `patience' \\
+%c. & şaful & `wooden honey tub' \\
+%d. & avuç  & `palm of hand' \\
+%   & havuz & `hollow'       \\
+%e. & samur & sable'         \\
+%   & camus & `buffalo'      \\
+%\end{tabular}
+%\xe

### 2012-03-01 3.1.tex
Drastic restructuring and some writing on chapters 2-3.

@@
 \citet{Zimmer1969}
+
+I assume without discussion the following binary feature specifications for the vowel phonemes of Turkish.
+
+\ex Turkish vowel features: \vspace{6pt} \\
+\begin{tabular}{c | c c c c}
+                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
+                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
+\buf[$+$\textsc{High}] & i & ü [y] & ı [ɯ] & u \\
+\buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
+\end{tabular}
+\xe
+
+%\citep[][289]{Anderson1975}
+%\citep{Clements1982}
+%\citet{Inkelas1997}
+%\citet{Harrison2001}
+
+\subsection{Backness harmony}  \input{3.1.1}
+\subsection{Labial attraction} \input{3.1.2}

### 2012-03-01 3.2.2.tex
Drastic restructuring and some writing on chapters 2-3.

@@
 % 3.2.2: Wordlikeness results
+
+There is some disagreement in the literature about just what \citet{Zimmer1969} finds regarding \textsc{Labial Attraction}.
+
+\citet{Zimmer1969}
+(see also \citet[][4]{Zuraw2000})
+claims however, Zimmer's evidence suggests that the speakers did internalize a version of the labial attraction generalization.
+
+\citet{Inkelas2001} present a different variation on the original rule of \textsc{Labial Attraction}
+
+\begin{quotation}
+Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
+\citep[][196, emphasis in original]{Inkelas2001}
+\end{quotation}
+
+``These results support our view that there is no rule of Labial Attraction'' \citep[][225]{Clements1982}.
+
+since the opposing are all broader versions than the original formulation by \citeauthor{Lees1966a}.
+
+\citet{Ito1993} argue that the data shows that it holds of native words only
 

### 2012-03-01 3.tex
Drastic restructuring and some writing on chapters 2-3.

@@
-% 3: Static and derived constraints on Turkish vowel harmony
+\chapter{Static and derived constraints in Turkish}
+\label{turkish}
 
-In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon may occur
+One of the original motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
+\citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible.
+\emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy.
+In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities \citep[][283f.]{Anderson1974}.\footnote{This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
+Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
+This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.}
 
-This brief chapter argues that even statistically valid phonotactic generalizations are ignroed by speakers.
-Two phonotactic generalizations over the Turksih lexicon are shown to be equally statistically reliable, but only the generalization which is derived by a phonological alternation influence wordlikeness judgements.
+the other being the subject
 
-Research on gradient phonotactics has usually rejected this possibilty implicit
-ly.
+a principle which is one of the two motivations  motivated the introduction of morpheme structure constraints.
 
-After reviewing co-occurrence restrictions in Gitksan roots, \citet{Brown2010} rejects this explicitly. He writes:
+In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon does not always require static constraints on possible underlying representations, since such distributions arise by chance and static constraints defined over natural classes contribute little to predicting the contents of the lexicon.
+%This result does not directly bear on the question of whether static restrictions have any synchronic reality.
+
+Yet countless static constraints are statistically supported.
+Under the null hypothesis advanced in Chapter \ref{MSCs}, the lexical statistics of co-occurrence restrictions are irrelevant: the language acquisition device only attends to derived restrictions and ignores static restrictions regardless of their statistical reliability.
+On the other hand, many linguists assume that statistical significance provides sufficient conditions for positing synchronically active static constraints.
+For instance, \citet{Brown2010} writes:
 
@@
 
+\noindent
+A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions.
+Any explicit theory of static constraints will have to impose formal restrictions on possible co-occurrence restrictions.
+The set of possible constraints must respect results in formal learning theory, and may also be, for example, constrained by naturalness biases or formal markedness considerations.
+
+A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
+This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).
+
+Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
+In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.
+
+In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word.
+
+The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements.
+
+The data below, drawn from Turkish, shows that
+
+This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
+Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements.
+Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
+
 %Classical Arabic adjectives often have stative verbs in which the root is imposed onto the template CaCuCa:
 
-%\ex Arabic verbs of coming into being: \\
-%\begin{tabular}{r l l l}
-%a. & kabura & `become big'       & (cf. \emph{kabiːr} `big') \\
-%b. & saʁiir & `become small'     & (cf. \emph{saʁiːr} `small') \\
-%c. & ħasuna & `become beautiful' & (cf. \emph{ħasan} `handsome') \\
-%\end{tabular}
-%\xe
+%It is important to note that the nature of the \emph{blick}/\emph{bnick} contrast is no longer a mystery. 
+%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
+It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails.
+If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
+
+\section{Turkish vowel MSCs} \input{3.1}
+\section{Evaluation}         \input{3.2}
+\section{Conclusion}         \input{3.3}

### 2012-03-01 4.2.1.tex
Drastic restructuring and some writing on chapters 2-3.

@@
 
-\subsubsection{Prediction distributions}
+\ex EM test results: \vspace{6pt} \\
+\begin{tabular}{l | r r | r r r r r r | r r}
+\toprule
+     & \multicolumn{2}{c|}{one $\mathcal{N}$} & \multicolumn{6}{c|}{mixture of two $\mathcal{N}$s} & \multicolumn{2}{c}{log $\mathcal{L}$ test} \\
+     & $\mu$ & $\sigma$ & $\mu_1$ & $\sigma_1$ & mix$_1$ & $\mu_2$ & $\sigma_2$ & mix$_2$ & $\Lambda$ & $p$-value  \\
+\midrule
+G\&J & 0.378 & 0.295    & 0.150   & 0.090      & 0.436     & 0.554   & 0.278    & 0.564     & 5.190    & 0.158   \\
+S    & 0.596 & 0.361    & 0.236   & 0.232      & 0.437     & 0.876   & 0.105    & 0.563 & 48.965    & 1.3\e{-10} \\
+A\&H & 0.431 & 0.266    & 0.301   & 0.200      & 0.559     & 0.595   & 0.246    & 0.411 & 16.186    & 0.001      \\
+\bottomrule
+\end{tabular}
+\xe
+
+%\ex Residualized correlations
+%\begin{tabular}
+%\toprule
+%ASDF
+%\midrule
+%G\&J & 
+%S    & 
+%A\&H & 
+%\bottomrule
+%\xe
+
+%\subsubsection{Prediction distributions}
 
@@
 \citet{Hayes2008a}
-\citet{Albright2009}
+\citet{Albright2009a}
 
@@
 % EM
-%\citet{EM}
+\citet{EM}
+
+\footnote{In performing this evaluation, I have benefitted greatly from course notes by Mark Liberman and Stephen Isard.}
+%available at \url{http://www.ling.upenn.edu/courses/cogs501/K-meansHW.html}
 
 % -2 log like ratio test
+
+where the difference in degrees o freedom equals $D$
+
+%\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } 
+%\ex $\displaystyle \Lambda = -2 \textrm{ ln } L_0 + 2 \textrm{ ln } L_1}$ \xe
+% > χ^2_{d.f.=D}$ \xe

### 2012-03-01 4.3.tex
Drastic restructuring and some writing on chapters 2-3.

% 4.3: Conclusion

It is not the case that this evidence is inconsistent with the hypothesis that there are additional degrees of wordlikeness: it is just a demonstration that these data to not require it.

### 2012-03-01 5.tex
Drastic restructuring and some writing on chapters 2-3.

\chapter{Inflectional gaps in generative grammar}

\section{A brief history of inflectional gaps} \input{5.1}
\section{The null hypothesis}                  \input{5.2}
\section{Outline of Part II}                   \input{5.3}

### 2012-03-01 7.1.1.tex
Drastic restructuring and some writing on chapters 2-3.

\subsection{Proposal}

\citet{OT}
\citet{Orgun1999}

\subsubsection{Inviolable constraints in a rule-based phonology}

\citet{Vaux2008}
(28 ``violability'')

such as developed by Andrea \citet{Calabrese1995,Calabrese2005}.

citing the OCP as an example...it's not clear wtf he's talking about
\citet{Goldsmith1976}

\citet{Nevins2003} % shm-gaps
\citet{Rice2007} % is being produced but avoided?

\citet[][9f.]{Baerman2010b}
\citet[][437]{Heath2005}

\citet{OT}
\citet{Orgun1999}

\subsubsection{MParse}

\citet{OT}
\citet{Raffelsiefen1999}
\citet{Raffelsiefen2004}

\citet{Wolf2009}

\subsubsection{Control}

\citet{Orgun1999}
\citet{Orgun2009}

### 2012-03-01 7.1.2.tex
Drastic restructuring and some writing on chapters 2-3.

\subsection{Evaluation}

% 7.1.3.1: Turkish and Turkish'

\citet{Orgun1999}
\citet{Orgun2009}

\citet{Ito1989b}

\subsubsection{Kinande reduplication}

``hurriedness or repetitive action \citep[][85; henceforth, MH]{Mutaka1990}.

\citet{Orgun1999} provide an analysis of a similar gap in Tiene, but it appears that the judgements they report are inferred, as \citet{Ellington1977}, the source for the generalization they report, provides no relevant judgements or examples, and his discussion suggests the gap is morphological in nature.
Since I have been unable to find any other published material on Tiene reduplication, Kinande, which is in comparison well documented, is presented here in its place.
But I predict that any close study of Tiene reduplication will reveal similar lexical exceptions.

\footnote{Thanks to Toni Cook for bringing this example to my attention.}

The material making up the so-called ``pre-stem'' will not be a major object of concern here: it consists of the nominal augment vowel (\textsc{Aug}), a verbal prefix (\textsc{Pre}), and where appropriate, the object marker (\textsc{OM}). Valence-chagning suffixes, known as ``extensions'', may appear to the right (in the below example, the applicative \textsc{Appl} and the reciprocal \textsc{Recp}), and the final \textsc{FV}, which indicates aspect.

\ex Linear structure of the Kinande infinitive: \\
\begin{tabular}{l l l l l l l}
e- & -rí- & -bí & tum & -ir & -an & -a \\
\{\textsc{Aug} & \textsc{Pre} & \texsc{OM}\} & \{send & \textsc{Appl} & \textsc{FV}\} \\
\multicolumn{7}{l}{``to send them to each other''}
\xe

bisyllabic
\citet{Downing2000}

\citet{Mutaka2008}

\subsubsection{Swedish indefinite adjectives}

\citet{Lowenadler2010a}
\citet{Lowenadler2010b}

\citet[][ chapter 5]{Lofstedt2010}

\citet[][7]{Buchanan2007}

\citet{Linell1973}
\citet{Sunden1869}
\citet{Hylen1920}
\citet{Cederschiold1912}
\citet{Brunnmark1826}
\citet{Eliasson1973}
\citet{Sigurd1965}
\citet{Teleman1999}
\citet{Kiefer1970}
\citet{Hellberg1972}
\citet{Iverson1981}
\citet{Pettersson1990}

\subsubsection{Tagalog infixation}

\citet{OT}

\citet{Orgun1999}
\citet{Orgun2009}

\citet{Ross1995}

\citet{Baerman2010b}

### 2012-03-01 2.2.2.tex
More arranging.

@@
-% 2.2.2: Computational models
+\subsection{Computational models}
 
@@
 
-\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{I am grateful to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are more frequent than those like [t.j], with rising sonority.
-\citeauthor{McGowan2011} reports, however, that this accounts for only a tiny amount of the variance in frequency ($R^2 = 0.077$).
-Change in sonority provides no improvement to the null baseline, and so this model is not reported here.
+\footnote{\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.
+Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
+However, sonority of clusters accounts for only a tiny amount of the variance in frequency ($R^2 = 0.077$).
+Sonority distance provides no improvements to the null baseline in this task, and is not reported further.
+I am grateful to Maria Gouskova for bringing \citeauthor{McGowan2011}'s study to my attention.}

### 2012-03-01 3.1.2.tex
More arranging.

@@
-% 3.1.2: Labial attraction
+\subsection{Labial attraction}
 
@@
 \noindent
-\citeauthor{Clements1982} are suggesting that \textsc{Labial Attraction} implies that \emph{aTu}, where \emph{T} represents a non-labial consonant, should be infrequent, but this fact is not inconsistent with the formulation of the constraint by \citet{Lees1966a,Lees1996b} and \citet{Zimmer1969}; \emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description.
+\citeauthor{Clements1982} are suggesting that \textsc{Labial Attraction} implies that \emph{aTu}, where \emph{T} represents a non-labial consonant, should be infrequent, but this fact is not inconsistent with the formulation of the constraint by \citet{Lees1966a,Lees1966b} and \citet{Zimmer1969}; \emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description.
 

### 2012-03-01 3.3.1.tex
Drastic restructuring and some writing on chapters 2-3.
More arranging.

@@
 % 3.3.1: Naturalness
+\subsection{Naturalness}
 
@@
 \citep[e.g.,][]{Lees1966a}, and despite \citeauthor{Zimmer1969}'s suggestive psycholinguistic results, reviewed above, it also been treated as a plausible constraint by later theorists \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995a,Zuraw2000}.
-Further, there is a real danger that if the label ``unnatural'' Labial Attraction} describes an impossible structural change or structural description, that one will fail to account for earlier forms of Turkish or sound changes therein.
+Further, there is a real danger that if the label ``unnatural'' 
+%Labial Attraction} 
+describes an impossible structural change or structural description, that one will fail to account for earlier forms of Turkish or sound changes therein.
+
+%Classical Arabic adjectives often have stative verbs in which the root is imposed onto the template CaCuCa:
+
+%\ex Arabic verbs of coming into being: \\
+%\begin{tabular}{r l l l}
+%a. & kabura & `become big'       & (cf. \emph{kabiːr} `big') \\
+%b. & saʁiir & `become small'     & (cf. \emph{saʁiːr} `small') \\
+%c. & ħasuna & `become beautiful' & (cf. \emph{ħasan} `handsome') \\
+%\end{tabular}
+%\xe

### 2012-03-01 4.tex
Drastic restructuring and some writing on chapters 2-3.
More arranging.

@@
-% 4: Gradient and categorical aspects of wordlikeness judgements
+\chapter{Gradient and categorical aspects of wordlikeness judgements}
+\label{gradience}
 
@@
 
-\begin{quotation}
-``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
-\end{quotation}
+%\begin{quotation}
+%``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
+%\end{quotation}
+
+\input{4.1}
+\input{4.2}
+\input{4.3}

### 2012-03-04 3.1.1.tex
More arranging.
Writing on chapter 3

@@
-% 3.1.1: Backness harmony
+\subsection{Backness harmony}
+
+One of the most salient aspects of spoken Turkish is the presence of vowel harmony. To a first approximation, 
+the high vowels of a Turkish prosodic word are all either all round or all unround. Here we will focus on backness harmony
+
+vowels agree
+
+all the vowels in a Turkish prosodic words agree for backness, and in addition, high vowels agree in roundness.
 
@@
 
-One of the most salient aspects of Turkish phonology is the presence of vowel harmony.
 \citeauthor{Lees1966a} (\citeyear[][284]{Lees1966b}, \citeyear[][35]{Lees1966a}) proposes a number of phonological rules to describe the harmony processes that act on roots and regularly spread rightward to suffixes (Turkish has no prefixes); one such rule consists of rightward spreading of the feature [\textsc{Back}].
@@
 1. \citet[][289]{Anderson1974} 
-2. \citet[][]{Kiparsky1981} makes a similar objection regarding Finnish vowel harmony.
+%2. \citet[][]{Kiparsky1981} makes a similar objection regarding Finnish vowel harmony.
 3. \citet{Clements1982}

### 2012-03-04 3.1.tex
More arranging.
Writing on chapter 3

@@
-% 3.1: Wordlikeness predictions
+\section{Turkish vowel sequence structure constraints}
 
-\citet{Zimmer1969}
-
-I assume without discussion the following binary feature specifications for the vowel phonemes of Turkish.
+Below, I detail two sequence structure constraints on root vowels in Turkish proposed by \citet{Lees1966a,Lees1966b}, relating these generalizations to Turkish alternation phonology and various sources of external evidence.
+A standard binary feature specification for the Turkish vowel system, given below, is assumed throughout.
 
@@
 \buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
-\end{tabular}
-\xe
+\end{tabular} \xe
+
+\input{3.1.1}
+\input{3.1.2}
 
@@
 %\citet{Harrison2001}
-
-\subsection{Backness harmony}  \input{3.1.1}
-\subsection{Labial attraction} \input{3.1.2}

### 2012-03-04 3.2.2.tex
More arranging.
Writing on chapter 3

@@
-% 3.2.2: Wordlikeness results
+\subsection{Wordlikeness results}
 
@@
 
-\subsubsection{Roundness harmony}
+\subsubsection{Backness harmony}
 
@@
 
-Roundness harmony wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
+Backness harmony wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
 \begin{tabular}{l r l r}
@@
 tipez & 24                   & tipaz & 8  \\ 
-terüz & 19                   & teruz & 13 \\
+terüz & 19                   & teruz & 13 \\ % roundness violator
 \bottomrule

### 2012-03-04 3.tex
More arranging.
Writing on chapter 3

@@
 
-One of the original motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
+This is not a universal, but rather something English speakers must learn: some languages, such as Moroccan Arabic, do in fact permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet').
+
+%bni 'building'
+%bnin 'delicious'
+%bniqa 'closet' 
+%bnat 'girl' 
+%bnita 'girl (dim.)'
+
+a matter of universal
+
+In their illustration of
+
+One of the
+
+One of the most compelling findings
+
+The contrast between \emph{blick} and \emph{bnick} is what has been called a "projection problem" \citep{Baker1979}
+
+This is what we might call a projec
+In presenting the components of a 
+a term which
+
+by subphonemic details
+
+\citet{Davidson2005,Davidson2006a} shows that excresent schwa is acoustically and articulatorily distinct from "lexical schwa":
+\citet{Wilson2011} show taht speakers use this
+
+%English speakers' z^@k is more similar to \emph{scum} than \emph{succumb}
+
+/sC-/ onsets than /s@C-/.
+
+% Davidson 2005: z^@C is more like sC than s@C (ultrasound of articulations)
+% Davidson 2006a: lexical and excresant schwa are not the same (acoustics)
+
+% Davidson 2010: production
+% Wilson & Davidson 2011:
+
+\citet{Berent2007a}
+\citet{Berent2007b}
+\citet{Berent2008a}
+
+% Berent et al. 2007
+% 598: Highly marked inputs are repaired in perception to abide by the grammatical restrictions of the language
+% Exps 1-2: syllable count task (misperception due to linguistic experience; "disyllables were more likely to be perceived as monosyllables if their monosyllabic counterparts had a less marked cluster" in both Russian and English, not what OT predicts)
+% Exps 3-4: AB task
+% Exps 5-6: priming effect (non-epenthetic and epenethetic things represented the same)
+% Berent and Lennertz 2007: (640: If the high rate of monosyllabic errors to unmarked onsets is only due to phonetic failures to encode the input, then it is puzzling why the same onsets also yield the highest rate of accurate response.)
+% Berent et al. 2008a:
+\citet{Peperkamp2007}
+% Peperkamp 2007: during one stage of processing phonology is undone with help from the lexicon...it only takes licit inputs and is irrelevant (p. 633)
+\citet{Dupoux1999}
+% Dupoux et al. 1999: Japanese speakers still can't distinguish between ebzo and ebuzo in an ABX task using the SAME a or b.
+
+\citet{Daland2011}
+
+does not so much as explain the thing itself but rather suggests the
+
+I believe a number of the predictions of this model are wrong
+the original and still-compelling 
+findings that drove
+
+ motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
 \citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible.
 \emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy.
-In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities \citep[][283f.]{Anderson1974}.\footnote{This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
+In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities \citep[][283f.]{Anderson1974}.
+
+This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
 Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
-This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.}
+This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.
 
@@
 
-\section{Turkish vowel MSCs} \input{3.1}
-\section{Evaluation}         \input{3.2}
-\section{Conclusion}         \input{3.3}
+\input{3.1}
+\input{3.2}
+\input{3.3}

### 2012-04-18 1.1.2.tex
More arranging.
chapters 2 and 4 have been swapped

@@
-% 1.1.2: Early generativism
+\subsection{Early generativism}
 
@@
 
+\citeauthor{Anderson1974} sees the duplication of \textsc{Backness Harmony} as both sequence structure constraint and phonological rule as a foregone conclusion
+.The tendency of sequence structure constraints to duplicate phonological rules was noted by \citet[][401f.]{Stanley1967} and in \emph{SPE} (p.~382) in the discussion of morpheme structure constraints, and is labeled ``the duplication problem'' by \citet[][?]{Kenstowicz1977}.
+
 inventory problems 

### 2012-04-18 2.1.1.tex
More arranging.
chapters 2 and 4 have been swapped

@@
-% 2.1.1: Data sources
+\subsection{Gradience in judgements}
 
-The English CELEX database \citep{CELEX},
+at least as far back as \emph{LSLT} \citep{LSLT}
 
-\citet{Aronoff1976}
+%and has become a recent feature of the discussion
+is that non-gradient models are totally incompatible with psycholinguistic results.
 
-\citet{Harley2009}
+\subsubsection{Task model}
 
-\citet{Taft1975} ?
-\citet{Taft1981} ?
-\citet{Taft2004a} ?
+\citet[][16]{Schutze1996}
 
-\citet{Halle1962}
+\citet{Schutze2005}
+\citet{Schutze2011}
+
+\citet{Coleman1997}
+
+\subsubsection{Falsifiability}
+
+\citet{Armstrong1983}
+
+% wordlikeness critique
+\citet{Yang2008a}
+
+a mix of ``declarative'' knowledge,
+corresponding to rote facts and thus the lexicon,
+and ``procedural'' knowledge, or knowledge of generative procedures 
+(in the sense of J. \citealt{Anderson1993}).
+
+% math cog
+\citet{Logan1988}
+\citet{Sfard1991}
+\citet{Delazer2005}

### 2012-04-18 2.1.2.tex
More arranging.
chapters 2 and 4 have been swapped

@@
-% 2.1.2: Syllabification
+\subsection{The ontology of wordlikeness}
 
-It is necessary to split medial consonant clusters into the codas and onsets that make them up.
-While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic.
-For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in RP \citep[][294]{AOE2}.}
-This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language.
-Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which contrast \emph{chemistry} and \emph{ministry}.
+\subsubsection{Gross phonotactic violation}
 
-An automated syllabification system was constructed and applied to medial clusters in the CELEX data.
-This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.
-There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification.
-The technique used is a variation on the theme of onset maximization \citep[42f.]{Kahn1976}, which parses word-medial clusters so that the onset is as large as possible.
-Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty.
-However, in words like \emph{mi}[n.str]\emph{el}, the [nstr] cluster does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.
+position in \citet{SPE}
 
-There is one case where unchecked maximization of medial onsets produces incorrect syllabifications.
-When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
-Onset maximization makes the right predictions for words like \emph{propaganda}, where the medial cluster cannot be an onset, but \citet[][55]{Harris1994} notes that it makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
-The approach adopted here is to first apply onset maximization, then to reassign the first consonant of a medial onset to the preceding coda if it is immediately preceded by a stressed lax vowel.
+\citet{Shibatani1973}
+\citet{Hooper1973}
 
-If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables.
-For this reason, it is assumed here that affricates are single segments for the purposes of syllabification.
-This assumption is motivated by the fact that affricates pattern with simple segments in many languages; for instance, the only word-initial consonant clusters in Classical Nahua are affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.
+peripheral \citep[in the sense of][81]{LGB}
 
-%In fact, [t.ʃ, d.ʒ] clusters are not found in simplex English words, despite the fact that that affricates occur as medial onsets in clusters (e.g., \emph{tru}[n.tʃ]\emph{eon}, \emph{so}[l.dʒ]\emph{er}).
+As an example of this applied ot phonotactics, 
+\citet{Borowsky1989} assumes that there are peripheral morph-internal syllable contact clusters.
+
+\subsubsection{Phonological statistics}
+
+\citet{SPE}
+
+This statement is made very strongly by \citet{Coleman1997}
+
+\subsubsection{Lexical similarity}
+
+\citet{LSLT}
+
+\citet{Greenberg1964}
+\citet{Ohala1986b}
+
+A recent, very articulated model is presented by \citet{Bailey2001}.
+
+% simple model
+This is simply because more complex models fail to
+
+% similarity vs. rules
+% as a final note, \citet{Hahn1998}

### 2012-04-18 2.2.1.tex
More arranging.
chapters 2 and 4 have been swapped

--- 2012-02-17 2.2.1.tex
@@
-% 2.2.1: Static and derived constraints
+\subsection{Assessing gradience}
 
-Having described the database of English syllable contact clsuters, I now consider constraints on sequences of URs as predictors of attested and unattested contact clusters.
-First, I describe the statistical technique used, then apply this technique to both phonologically derived constraints and a set of ``static'' phonotactic constraints proposed by \citet{Pierrehumbert1994}.
+\subsubsection{Response distributions}
 
-\subsubsection{Statistical technique}
+This is shown in Figure \ref{density}.
 
-Before the dawn of generative phonology, many linguists concerned themselves with documenting co-occurrence in various languages.
-\citet[][28]{Vogt1954} declares that a ``phonemic description of a language \ldots should also comprise a description of the phoneme combinations that occur or may be expected to occur in the language''.
-Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively.
-\citeauthor{Uhlenbeck1950}, for instance, puts forth co-occurrence dispreferences, rather than exceptionless generalizations of the sort given by \citet{Greenberg1950}.
-Years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and proposes a statistical technique, based on the chi-square test, to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.
+\begin{figure}
+\centering
+\includegraphics{density.pdf}
+\caption{Density of responses}
+\label{density}
+\end{figure}
 
-\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English).
-The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented.
-Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.
+\begin{example}
+EM test results:
 
-\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
-\chi^2$ \xe
-
-%\footnote{This is the one-tailed version of the test.}
-
-In this study, I use the \citet{Fisher1934} exact test, which is isomorphic to this chi-square test.
-Compared to the chi-square test, the Fisher exact test $p$-value is somewhat more difficult to compute, but both are automated by modern statistical packages and can be computed rapidly.
-The major advantage of the Fisher test over the chi-square test is that it is appropriate for small samples or rare events, whereas the chi-square test depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples \citep[see][]{Gorman2012a}.
-%\citep[e.g.,][142]{Mester1988}.
-
-On the other end of the spectrum, the Monte Carlo statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010} does not scale to large samples.
-The Monte Carlo procedure is isomorphic to the one-tailed Fisher Exact test, and involves comparing observed co-occurrence counts to those produced by randomly generated samples.
-This requires the researcher to generate random permutations, something which is particularly difficult for large samples, as I will explain; disinterested readers may wish to skip the following paragraph.
-
-Any permutation of a sequence $L$ can be represented as a sequence $S$ of the same length, where the value of $S_i$ corresponds to the permuted position of $L_i$.
-Generating random permutations requires an unbiased method to generate all $N!$ of these lists with equal probability.
-Unfortunately, this is near impossible for $N$s of moderate size with current computational resources.
-Any pseudo-random number generator can be characterized by its ``period'', the number of pseduo-random numbers it can emit before repeating itself; this is also the upper limit for the number of random permutations that can be generated.
-The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a period of approximately $2^{48}$, a number which turns out to be smaller than $17!$.
-For the large samples, sometimes in the thousands, used by these researchers, virtually all the possible permutations can never be generated by this method.
-As a consequence, the shuffling procedure is far from random and introducing the possibility that the statistical test is biased in unpredictable ways.
-
-As an example of the Fisher exact test as it is applied here, consider some of the Javanese co-occurrence facts considered by \citet{Mester1988}.
-\citeauthor{Mester1988} is at pains to show that there is a gradient dispreference for first and second root consonants which share the same major place.
-However, perfectly identical first and second consonants are exempt.
-Since
-
-% APPLY THIS TO JAVANESE
-
-\ex Javanese root co-occurrence (\citealp[][264]{Uhlenbeck1950}, \citealp[][139]{Mester1988}): \vspace{6pt} \\
-\begin{tabular}{l r r r r}
+\vspace{0.5\baselineskip}
+\begin{tabular}{l | r r | r r r r r r | r r}
 \toprule
-           & attested & unattested & saturation & $p$-value \\
+     & \multicolumn{2}{c|}{one $\mathcal{N}$} & \multicolumn{6}{c|}{mixture of two $\mathcal{N}$s} & \multicolumn{2}{c}{log $\mathcal{L}$ test} \\
+     & $\mu$ & $\sigma$ & $\mu_1$ & $\sigma_1$ & mix$_1$ & $\mu_2$ & $\sigma_2$ & mix$_2$ & $\Lambda$ & $p$-value  \\
 \midrule
-conforming & 134      & 12         & 0.918      & \multirow{2}{*}{6.095\e{-06}} \\
-violating  & 21       & 15         & 0.583 \\
+G\&J & 0.378 & 0.295    & 0.150   & 0.090      & 0.436     & 0.554   & 0.278    & 0.564     & 5.190    & 0.158   \\
+S    & 0.596 & 0.361    & 0.236   & 0.232      & 0.437     & 0.876   & 0.105    & 0.563 & 48.965    & 1.3\e{-10} \\
+A\&H & 0.431 & 0.266    & 0.301   & 0.200      & 0.559     & 0.595   & 0.246    & 0.411 & 16.186    & 0.001      \\
 \bottomrule
 \end{tabular}
-\xe
+\end{example}
 
-\noindent
-It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.
-
-This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance
-
-That this is highly unlikely to be due to chance is apparent both from the chi-square ($\chi^2 = 25.572$, two-tailed $p = 4.3$\e{-07}) and Fisher exaact tests ($p =  6.095$\e{-06}).
-
-I use two-tailed statistical tests throughout. 
-Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis.
-As
-
-\citet{Pierrehumbert1994}
-
-It must be stressed that under current assumptions, this is independent of whether predictable prosodic representations (particularly coda and onset identity) are part of the memorized underlying representation \citep[e.g.,][]{Vaux2003}.
-It is an undeniable fact that URs need to be 
-The derivative nature of 
-\citet{Ito1989a,Noske1992}
-
-least some researchers (i.e., \citealt{Mester1988})
-
- \citep[e.g.,]{Brown2010} have argued for phonotactic constraints giving rise to overrepresentation, so I admit this possibility here and use two-tailed tests throughout.
-
-%\footnote{For example, \citet{Martin2007} applies this technique to a list of 4,758 noun-noun compounds.
-The number of permutations of a list of this length can be estimated using Stirling's approximation.
-
-%\ex $ \lim_{n \rightarrow \infty} n! = e ^ {n \ln{n} - n} = 2 ^ {\log_2{e} (n \ln{n} - n)} $ \xe
-
-%\noindent
-%$4,758!$ is approximately $2 ^ {51260}$. The author is unaware of any system that provides 51,260 bits of entropy; and some popular programming languages, such as ANSI C, provide as few as 15 bits.}
-
-%%% START OF SERIOUSNESS
-
-I will first consider three well-known English phonological alternations, showing that these give rise to constraints on the sequence structure of morphs.
-Simple autosegmental formalizations of these processes are provided, and the Fisher exact test is used to demonstrate their effect on the English lexicon.
-
-\subsubsection{Obstruent Voice Assimilation}
-
-English has a process of voice assimilation which is found in two inflectional suffixes.
-The suffix /-d/ forms regular pasts and denominal adjectives, and /-z/ forms regular noun plurals and genitives, and the 3sg.
-present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S.
-\citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][211f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
-
-\ex Voice assimilation in the past tense and noun plural: \vspace{6pt} \\
-\begin{tabular}{l l l l l l}
-a. & /næp-d/     & \arrow & [næpt]     & (cf. \emph{nab}[d], \emph{haul}[d]) \\
-b. & /læp-z/     & \arrow & [læps]     & (cf. \emph{lab}[z], \emph{clan}[z])  \\
-\end{tabular}
-\label{ovad}
-\xe
-
-\noindent
-The voiced obstruent suffixes devoice when preceded by a voiceless obstruent.
-This is formalized as a rule spreading the [\textsc{Voice}] specification rightward over adjacent obstruents.
-
-\ex \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178): \\
-\xymatrix@C=6pt@R=12pt{
-\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{..}[dr] & \\
-\txt{C}\ar@{-}[d]             & \txt{C}\ar@{-}[d] \\
-\txt{[$+$\textsc{Obstruent}]} & \txt{[$+$\textsc{Obstruent}]}}
-\label{ovar}
-\xe
-
-\noindent
-This rules out the possibility of surface clusters of obstruents with differing  [\textsc{Voice}] specifications.
-While the data in (\ref{ovad}) make it clear \textsc{Obstruent Voice Assimilation} applies in morphologically derived environments, it is not obvious that it applies to obstruent clusters in non-derived environments (e.g., obstruent clusters in the URs of morphs) as well.
-\citet{Pierrehumbert1994} does not mention voice assimilation in her study of gaps in English syllable contact clusters.
-\citet[][74f.]{Hammond1999a} notes the existence of words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain obstruent voice clusters disagreeing in [\textsc{Voice}], and takes this as evidence that obstruent voicing is restricted to derived environments.
-
-Despite these exceptions, I propose that \textsc{Obstruent Voice Assimilation} plays a major role in shaping the English lexicon.
-Only English syllable contact clusters which contain two adjacent obstruents speak to the status of \textsc{Obstruent Voice Assimilation}; the remaining 160 possible clusters are ignored.
-A $2 \times 2$ contingency table is constructed by dividing the remaining possible clusters into those which agree on obstruent voicing and those which do not, 
-and into those which are attested and those which are not.
-
-\ex Lexical effects of \textsc{Obstruent Voice Assimilation}: \vspace{6pt} \\
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming & 80 & 370 & 0.178 & \multirow{2}{*}{1.1\e{-11}}\\
-violating  &  6 & 264 & 0.022 \\
-\bottomrule
-\end{tabular}
-\xe
-
-\noindent 
-18\% of the possible clusters containing adjacent obstruents with the same [\textsc{Voice}] specification are found, 
-whereas only 2\%  of those which disagree in [\textsc{Voice}], 
-such as in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, 
-CELEX also contains \emph{jo}[d.p]\emph{urs}, \emph{pi}[ntʃ.b]\emph{eck}, \emph{sa}[k.b]\emph{ut}, and \emph{ja}[k.d]\emph{aw}. 
-As the $p$-value shows, 
-the rarity of these disagreeing clusters is unlikely to be due to chance.
-
-I propose that this pattern should be attributed directly to \textsc{Obstruent Voice Assimilation} applying to non-derived environments, preventing learners from positing URs in which adjacent obstruents do not agree in voice.
-There is no duplication between constraints on sequence structure and surface forms if both are derived from the application of the phonological rule (\citealt[][401f.]{Stanley1967}, \citealt[][382]{SPE}).
-The handful of exceptions to this generalization can be marked with appropriate exception diacritics.
-Another possibility is suggested by recent psycholinguistic results.
-\citet{Mattys2001b} find that 9-month old infants are sensitive to the contrast between [vt], a cluster of obstruents disagreeing in [\textsc{Voice}] which is not found word-medially, and [ft], which conforms to the above generalization and is found medially (e.g.
-\emph{a}[f.t]\emph{er}).
-This ability to use phonotactic violations to segment continuous speech persists into adulthood \citep{McQueen1998}, suggesting that ``hetero-voiced'' obstruent clusters may be analyzed as semantically opaque compounds by native speakers (e.g., \emph{ja}[k\#d]\emph{aw}).
-%\footnote{An analogous account is the proposal that readjustment rules insert morphological boundaries for the purposes of assigning stress (\emph{SPE}:94).}
-
-\subsubsection{Coda nasal place assimilation}
-\label{cnpasection}
-
-Above, it was assumed that the velar nasal is derived from underlying /n/. 
-This analysis is not purely allophonic, as nasal place assimilation alternations target the final consonant of the \emph{im-}/\emph{in-} prefix.
-In (\lastx ab), the prefix-final consonant takes on the same major place of articulation as the following obstruent, and in (\lastx c), where the root is vowel-initial, the prefix-final consonant is realized as a default /n/.
-
-\ex Nasal place assimilation in Latinate prefixes: \\
-\begin{tabular}{l l l l}
-%a. & possible & i[m.p]ossible & & balance & i[m.b]alance \\
-%b. & tangible & i[n.t]angible & & decent  & i[n.d]ecent  \\
-%c. & elegant  & i[n.]elegant  & & ability & i[n.]ability \\
-a. & possible & i[m.p]ossible \\
-   & balance  & i[m.b]alance  \\
-b. & tangible & i[n.t]angible \\
-   & decent   & i[n.d]ecent   \\
-c. & elegant  & i[n.]elegant  \\
-   & ability  & i[n.]ability  \\
-\end{tabular} 
-\xe
-
-\noindent
-The shape of the prefix before other obstruents is somewhat more complex. 
-%Before /f/, CELEX transcribes the nasal as either labial or alveolar, the latter being very rare in simplex words. 
-%There are also additional complexities regarding the shape of the prefix when attached to /k, g/-initial roots. 
-\citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but this assimilation is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
-The prefix \emph{com-}/\emph{con-}, which patterns with \emph{im-}/\emph{in-} in showing assimilation (e.g., \emph{co}[m.b]\emph{ine}), participates in a similar contrast, between \emph{có}[ŋ.g]\emph{ress} and \emph{co}[n.g]\emph{réssional}.
-As \citeauthor{Borowsky1986} notes, stress does not block assimilation in simplex words (e.g., \emph{a}[ŋ.g]\emph{óra}), a fact which demonstrates the necessity of a partially morphological explanation for the blocking of assimliation in words like \emph{co}[n.g]\emph{réssional}.
-Regardless of these complexities, there is sufficient evidence to posit assimilation of the \textsc{Place} features of an obstruent to a preceding nasal consonant.
-
-\ex \textsc{Coda Nasal Place Assimilation} (\emph{SPE}:85): \\
-\label{CNPA}
-\xymatrix@C=6pt{
-                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{..}[dl] \\
-\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d] \\
-\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]} }
-\xe
-
-\citet[][175]{Pierrehumbert1994} observes that in simplex English words, ``nasal-stop sequences agree in labiality''; presumably, this generalization is limited to [\textsc{Labial}] because \citeauthor{Pierrehumbert1994} also adopts the assumption that the velar nasal is derived from /n/.
-However, \citeauthor{Pierrehumbert1994} does not propose any connection between the alternation and the generalization that nasal consonants in URs tend to agree with the following consonant.
-While clusters like \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey} are the norm, exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} are found.
-
-In all, there are only 80 possible nasal-obstruent syllable contact clusters, the majority of which are attested.
-In addition, there are a small number of words which contain clusters like \emph{oi}[nt.m]\emph{ent}, in which a nasal-obstruent cluster are both parsed into a complex medial coda, but in this position, place agreement is exceptionless, and thus these clusters are ignored below.
-
-\ex Lexical effects of \textsc{Coda Nasal Place Assimilation}: \vspace{6pt} \\
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming & 41 & 11 & 0.788 & \multirow{2}{*}{2.6\e{-05}}\\
-violating  & 8  & 20 & 0.286 \\
-\bottomrule
-\end{tabular}
-\xe
-
-\noindent
-Clusters which conform to this generalization are far more likely to be attested than those that violate it, and this is highly unlikely to be due to chance.
-
-There is some evidence that even young infants show a preference for nonce words that conform to this generalization.
-\citet{Jusczyk2003}, \citet{Mattys1999}, and \citet{Jusczyk2002} play nonce words containing nasal-obstruent clusters to 4.5, 9, and 10-month-old infants (respectively) acquiring English, All three studies find that infants prefer to listen to stimuli which conform to the above generalization (e.g., \emph{u}[m.b]\emph{o}) over those which do not (e.g., \emph{u}[n.b]\emph{o}).
-\citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''.
-In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.
-
-\ex Artificial language adaptations \citep[][394, his transcriptions]{Wright1975}: \\
-\begin{tabular}{l l l l l l}
-a. & [gownp] & > & [gump] \\
-b. & [ǰumg]  & > & [ǰúŋgə] \\
-c. & [ðʌŋd]  & > & [tɔŋg] \\
-\end{tabular} 
-\xe
-
-\noindent
-If this game is akin to the process of loanword adaptation, it is possible that various extragrammatical pressures are at play \citep[e.g.,][]{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008}.
-Yet, the independent evidence for a phonological process of nasal place assimilation provides the most parsiminious account of the adaptations seen in \citeauthor{Wright1975}'s study.
-J.
-\citet{Myers1993} also shows that \textsc{Coda Nasal Place Assimilation} is fed by speech errors:
-
-\ex Speech errors and \textsc{Coda Nasal Place Assimilation} \citep[][228]{Myers1993}: \\
-\begin{tabular}{l l l}
-   & \emph{error}  & \emph{target} \\
-a. & primps        & prints        \\
-b. & rand orker    & rank order    \\
-%  & [sprɪg]time for [hɪ̃ntlɘr] & springtime for Hitler \\
-c. & po[ŋ]cutation & computation   \\
-\end{tabular}
-\xe
-
-%\citet{Hay2004a} demonstrates that [n.p, mθ] clusters, which are not found in English, are in fact rated better than many attested clusters.
-%report that 10-month-old infants prefer to listen to nonsense words which conform to \textsc{Coda Nasal Place Assimilation} (e.g., \emph{u}[m.b]\emph{o}) compared to those which do not (e.g., \emph{u}[n.b]\emph{o}).
-
-\subsubsection{Degemination}
-
-A final well-known alternation in English is the simplification of geminates derived by certain suffixes.
-For instance, the suffix \emph{-ly} rarely forms a geminate when attaching to root-final /l/, whether forming denominal adjectives (e.g., \emph{sme}[l]\emph{y}) or deadjectival adverbs (e.g., \emph{norma}[l]\emph{y}).
-A small class of English verbs take a /-t/ suffix in the past; as this suffix does not trigger the \textsc{Epenthesis} found in the regular past (e.g., \citealt{Bakovic2005b}, \citealt{Fruehwald2011}), \textsc{Degemination} applies.
-
-\ex Degemination in the /-t/ past (\citealp[][105]{Halle1985a}, S. \citealp[][492]{Myers1987}): \vspace{6pt} \\
-\begin{tabular}{l l l l l}
-a. & /bɛnd-t/ & \arrow & [bɛnt] & (cf. \emph{burn}[t]) \\
-b. & /bɪld-t/ & \arrow & [bɪlt] & (cf. \emph{spil}[t]) \\
-\end{tabular}
-\xe
-
-More complicated cases of degemination are found in English prefix morphology and described in \emph{SPE} (p.~148, 219f.) and \citealt[][]{Borowsky1986} (p. 102f.), and are not reviewed here.
-
-%\ex \textsc{Coda Nasal Place Assimilation} feeds \textsc{Degemination} \citep[][116]{Borowsky1986}: \\
-%\begin{tabular}{l l l l}
-%a. & mature    & i[m]ature    & (cf. \emph{i}[m.b]\emph{alance})   \\
-%b. & numerable & i[n]umerable & (cf. \emph{i}[n.d]\emph{ependent}) \\
-%\end{tabular}
+%\ex Residualized correlations
+%\begin{tabular}
+%\toprule
+%ASDF
+%\midrule
+%G\&J & 
+%S    & 
+%A\&H & 
+%\bottomrule
 %\xe
 
-The rule deletes the first of two adjacent consonants agreeing in \textsc{Place} and manner features, and is formalized using quantification over features \citep{Reiss2003b}.
+%\subsubsection{Prediction distributions}
 
-\ex \textsc{Degemination} (\emph{SPE}:253): \\
-\label{degem}
-%\xymatrix@C=6pt{
-%{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\
-%\textsc{C}\ar@{-}[d]       & \longrightarrow & \emptyset & / & \gap\gap\gap & \textsc{C}\ar@{-}[d] \\
-%*+[F(][F)]\txt{β \textsc{Continuant} \\ γ \textsc{Obstruent} \\ δ \textsc{Sonorant} \\ ε \textsc{Lateral} \\ ζ \textsc{Nasal}} & & & & & *+[F(][F)]\txt{β \textsc{Continuant} \\ γ \textsc{Obstruent} \\ δ \textsc{Sonorant} \\ ε \textsc{Lateral} \\ ζ \textsc{Nasal}} \\ \\
-%}
-\xymatrix@C=6pt{
-%{\textsc{α Place}}\ar@{-}[d] &                 &           &   &              & \textsc{α Place}\ar@{-}[d] \\
-\textsc{C}_1       & \longrightarrow & \emptyset & / & \gap\gap\gap & \textsc{C}_2 }
-\\ 
-Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
-\xe
+\citet{Coltheart1977}
 
-\noindent
-I assume that geminates crossing a syllable boundary are ``false geminates'', meaning that they are adjacent similar segments rather than a single feature matrix linked to two timing slots.
-Under the alternative account in which the absence of geminates is an inventory restriction (i.e., a \textsc{Segment Structure Condition}), an undesirable duplication of phonological derived surface constraints and constraints on URs would be unavoidable.
+\citet{Hayes2008a}
+\citet{Albright2009a}
 
-Neither identical consonants, nor those which differ only in voice (e.g., *[p.b]) occurs in simplex words in CELEX, and this absence is highly unlikely to be due to chance.
+% stats thereof
+%\citet{Schilling2002}
+%\citet{Helguerro1904}
+%\citet{Cohen1956}
 
-\ex Lexical effects of \textsc{Degemination}: \vspace{6pt} \\
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming & 157 & 596 & 0.208 & \multirow{2}{*}{7.2\e{-09}}\\
-violating  &   0 &  87 & 0.000 \\
-\bottomrule
-\end{tabular}
-\xe
+% EM
+\citet{EM}
 
-%\footnote{Note that it is not the case that the English lexicon as envisioned by \emph{SPE} is free of underlying geminates; \citeauthor{SPE} use underlying geminates to attract stress to a preceding nucleus (p.~148--151).
-In \emph{SPE}, the underlying geminate-singleton contrast is absolutely neutralized, but has consequences for surface stress.
-In addition to the usual objections to absolute neutralization (i.e., that it is a learning conundrum mascarading as a phonological solution) that emerged in the wake of the publication of \emph{SPE}, the phonological particulars of this account have also come under attack \citep[e.g.,][]{Ross1972a}.
-It would not seem tenable to use underlying geminates to attract stress even if absolutely neutralized contrasts were admitted.} Unsurprisingly, this is unlikely to be due to chance.
+\footnote{In performing this evaluation, I have benefitted greatly from course notes by Mark Liberman and Stephen Isard.}
+%available at \url{http://www.ling.upenn.edu/courses/cogs501/K-meansHW.html}
 
-\subsubsection{Dorsal/non-coronal clusters}
+% -2 log like ratio test
 
-\citet{Pierrehumbert1994} surveys the English syllable contact cluster inventory, and discerns the effect of several ``static'' constraints, those which do not derive from neutralizing phonological rules.
+where the difference in degrees o freedom equals $D$
 
-\citet[][173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Since \citeauthor{Pierrehumbert1994}'s study is restricted to triconsonantal clusters, so words like \emph{a}[k.m]\emph{e}, \emph{ru}[g.b]\emph{y}, or \emph{pi}[g.m]\emph{ent} do not violate this narrower generalization.
-Further, \citeauthor{Pierrehumbert1994} adopts the traditional assumption \citep[e.g.,][66f.]{Borowsky1986} that coda [ŋ] is an allophone of /n/ before dorsal consonants, so words such as \emph{mo}[ŋ.gr]\emph{el} are not exceptions.
-
-\citeauthor{Pierrehumbert1994} generalization is termed static simply because there is no evidence for any phonological process which changes clusters of dorsal obstruents followed by non-coronals into something else (perhaps a dorsal/coronal cluster).
-However, the statistical test, taking into account clusters of any length, does not indicate that this dispreference for dorsal/non-coronal clusters is so compelling that it could not be generated by chance, simply because possible dorsal/non-coronal clusters are about as likely to occur as are similar types of clusters such as those with a non-dorsal coda and a non-coronal onset (e.g., \emph{a}[t.m]\emph{osphere}) or a dorsal coda and coronal onset (e.g., \emph{ve}[k.t]\emph{or}).
-These clusters are marked as ``conforming'' to a constraint \textsc{*[+Dorsal][$-$Coronal]}, and dorsal/non-coronal consonants as ``violating'' it.
-
-\ex Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}: \vspace{6pt} \\
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming & 76 & 359 & 0.175 & \multirow{2}{*}{0.145} \\
-violating  &  6 &  57 & 0.095 \\
-\bottomrule
-\end{tabular}
-\xe
-
-There is no exceptions to \citeauthor{Pierrehumbert1994}'s narrower generalization concerning triconsonantal clusters, but it is no more statistically convincing.
-
-\ex Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]} on triconsonantal clusters: \vspace{6pt} \\
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming & 20 & 148 & 0.119 & \multirow{2}{*}{0.136} \\
-violating  &  0 &  22 & 0.000 \\
-\bottomrule
-\end{tabular}
-\xe
-
-\noindent
-The evidence for \citeauthor{Pierrehumbert1994}'s static constraint does not actually require that such a constraint exists at all, since it is quite possible that the apparent dispreference for dorsal/non-coronal clusters could result by accident.
-
-\subsubsection{Coda coronal obstruents}
-
-Another static generalization about triconsonantal syllable contact clusters proposed by \citet[][175]{Pierrehumbert1994} is that  ``clusters with a coronal obstruent in the coda do not occur.'' In an endnote (p.
-186), \citet{Pierrehumbert1994} notes an exception in \emph{a}[nt.l]\emph{er}, and one can add the triconsonantal clusters in \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many biconsonantal clusters like \emph{a}[t.l\emph{as} or \emph{e}[s.t]\emph{er}.
-As before, applying the statistical test to clusters of any length reveals that coda coronal obstruent clusters are not significantly less likely to occur than clusters with non-coronal codas.
-
-\ex Lexical effects of \textsc{*Coda Coronal Obstruent}: \vspace{6pt} \\
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming &  48 & 312 & 0.133 & \multirow{2}{*}{0.301} \\
-violating  &  38 & 322 & 0.106 \\
-\bottomrule
-\end{tabular}
-\xe
-
-Restricting \citeauthor{Pierrehumbert1994}'s generalization to triconsonantal clusters does not strengthen the case for this static constraint.
-
-\ex Lexical effects of \textsc{*Coda Coronal Obstruent} on triconsonantal clusters: \vspace{6pt} \\
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming &  4 & 104 & 0.037 & \multirow{2}{*}{0.419} \\
-violating  &  2 & 124 & 0.016 \\
-\bottomrule
-\end{tabular}
-\xe
-
-%Note that [tʃ] has not been included since it so rarely occurs in clusters.
-
-\subsubsection{ABA clusters}
-
-Finally, \citet[][176]{Pierrehumbert1994} notes ``a lack of clusters with identical first and third elements''.
-For the determination of identity, I assume the same definition of geminate as used in the rule of \textsc{Degemination} above, namely consonants which share all the same \textsc{Place} and manner features, ignoring \textsc{Voice}.
-%and I also expand the definition of the constraint to also apply to the second and fourth consonant of quadraconsonantal clusters. 
-\citeauthor{Pierrehumbert1994} notes one potential exception to this generalization, the cluster in words such as \emph{e}[ks.kl]\emph{ude}, which contains two /k/, does not arise here since such words have been marked morphologically complex.
-Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly rarer than other triconsonaontal and quadraconsonantal clusters, of which only 7.8\% occur.
-
-\ex Lexical effects of \textsc{*ABA}: \vspace{6pt} \\
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming & 41 & 478 & 0.079 & \multirow{2}{*}{0.818} \\
-violating  &  0 &  11 & 0.000 \\
-\bottomrule
-\end{tabular}
-\xe
-
-Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}.
-
-%\subsubsection{Summary}
-%
-%The above results are summarized in Table \ref{constraints}.
-%
-%\begin{table}
-%\centering
-%\begin{tabular}{l r r r r r}
-%\toprule
-%                                       & \multicolumn{2}{c}{attested} & \multicolumn{2}{c}{unattested} & \multirow{2}{*}{$p$-value} \\
-%                                       & conforming & violating & conforming & violating \\
-%\midrule
-%\textsc{O.V.A.} (\ref{ovar})        &  80 &   6 & 370 & 264 & 1.1\e{-11} \\
-%\textsc{C.N.P.A} (\ref{CNPA})       &  41 &  25 &  12 &  42 & 1.7\e{-05} \\ 
-%\textsc{Degemination} (\ref{degem}) & 161 & 632 &   0 &  87 & 1.4\e{-08} \\
-%\midrule
-%\textsc{*Dorsal-Labial} ()          &  68 &   4 & 246 &  40 & 0.069 \\
-%\textsc{*C.C.O.} ()                 &  48 &  38 & 312 & 322 & 0.301 \\
-%\textsc{*A,BA} ()                   & 161 &   0 & 708 &  11 & 0.231 \\
-%\bottomrule
-%\end{tabular}
-%\caption{Counts and Fisher exact test $p$-values for the constraints discussed above.}
-%\label{constraints}
-%\end{table}
+%\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } 
+%\ex $\displaystyle \Lambda = -2 \textrm{ ln } L_0 + 2 \textrm{ ln } L_1}$ \xe
+% > χ^2_{d.f.=D}$ \xe

### 2012-04-18 2.2.2.tex
chapters 2 and 4 have been swapped

--- 2012-02-16 2.2.2.tex
@@
-% 2.2.2: Computational models
+\subsection{Assessing models}
 
-Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}.
-There is now a growing literature on computational cognitive models of phonotactic learning.
-By applying these models to the syllable contact data, it is possible to exhaustively search for statistically motivated constraints, elimianting the possibility of human error.
-
-To evaluate these statistical models, they are first trained on the set of attested clusters, then put to the task of classifying all possible clusters into those which are and are not attested.
-Four metrics are used to evaluate these computational models.
-Classification accuracy corresponds to the probability of a classification being either a true positive ($t.p.$), correctly predicting a cluster to occur, or a true negative ($t.n.$), correctly predicting a cluster to be absent.
-Incorrect outcomes are either false positives ($f.p.$), incorrectly predicting a cluster to occur, and false negatives ($f.n.$), incorrectly predicting a cluster to be absent.
-
-%\ex Binary classification contingency table: \\
-%\begin{tabular}{c | l l} \toprule
-%                             & \multicolumn{2}{c}{prediction} \\
-%\midrule
-%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
-%                             & false negative ($fn$) & true negative ($tn$) \\
-%\end{tabular}
-%\begin{tabular}{|l l|}
-%\toprule
-%true positive ($tp$)  & false positive ($fp$) \\
-%false negative ($fn$) & true negative ($tn$) \\
-%\bottomrule
-%\end{tabular}
-%\xe
-
-\ex $\displaystyle \textrm{Accuracy} = \frac{t.p. + t.n.}{t.p. + t.n. + f.p. + f.n.}$ \xe
-
-\noindent
-A few other metrics can be used to assess classification performance. Precision corresponds to the probability that a cluster predicted to occur is observed in the data.
-
-\ex $\displaystyle \textrm{Precision} = \frac{t.p.}{t.p. + f.p.}$ \xe
-
-\noindent
-Recall (also known as sensitivity) is the probability a observed cluster is predicted to occur.
-
-\ex $\displaystyle \textrm{Recall} = \frac{t.p.}{t.p. + f.n.}$ \xe
-
-It is possible to maximize recall at the expense of precision, by predicting more clusters to be attested, or to increase precision at the expense of recall by predicting fewer clusters to be attested.
-A common way to balance these concerns is the $F_1$ score, which is the harmonic mean of precision and recall.
-
-\ex $\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$ \xe
-
-\subsubsection{Null baseline}
-
-The 840 possible clusters have an 18.7\% of saturdaion rate.
-As a consequence of this sparsity, a null model which classifies all clusters as unattested achieves 81.3\% accuracy without the need to posit any constraints either derived or static.
-Since this model does not have any positive classifications, either true or false, precision, recall, and $F_1$ are undefined.
-
-\subsubsection{Alternations}
-
-As was shown above, \textsc{Obstruent Voice Assimiation}, \textsc{Coda Nasal Place Assimilation}, and \textsc{Degemination} are reliable predictors of what clusters are and are not attested and have few lexical exceptions.
-I generate another baseline by predicting only those clusters which do not violate any of these generalizations to be attested.
-This provides a significant improvement to the null baseline (accuracy = 0.857).
-As precision (0.401) is smaller than recall (0.708), there are more false positives (i.e., unattested clusters that do not violate any of these three alternation-derived generalizations) than there are false negatives (i.e., attested exceptions to these derived generalizations).
-
-\subsubsection{\citet{Pierrehumbert1994}}
-
-\citet{Pierrehumbert1994} first proposes to measure the well-formedness of syllable contact clusters by the independent probabilities of codas and onsets, and \citet{Coleman1997} extend this model to whole words.
-\citet{Pierrehumbert1994} uses both medial and initial/final codas and onsets to compute these probabilities, but I obtain a better fit to the data by only using medial coda and onset frequencies.
-Medial coda and onset probabilities are multipled to obtain a well-formedness score.
-This produces a continuous value between 0 and 1.
-To derive a binary classification from these values, I use a soft-margin support vector machine \citep{Cortes1995} with a linear kernel.
-This technique derives a ``decision stump'' providing the best single split into attested and an unattested clusters.
-This is not put forth as a cognitively plausible model for relating this expected probability to attestation: it is simply the upper bound that such a model could reach.
-
-Unfortunately, this score only provides a small improvement to the null baseline
- (accuracy = 0.839), and has a lower accuracy and $F_1$ (0.328) than the alternation baseline.
-
-\subsubsection{\citet{Hayes2008a}}
-
-\citeauthor{Hayes2008a} develop a sophisticated method of learning phonotactic distributions from positive data.
-They develop software for estimating probability distributions over sequences of (sets of) phonological features using the principle of maximum entropy and making few assumptions about the independence of features.
-As above, a support vector machine with a linear kernel is used to turn these probability values into binary predictions.
-The software provided by these authors exposes many ``metaparameters'', operational decisions to be made in conducting an experiment with this system.
-To avoid any potential bias induced by the choice of these metaparameters, I use the same settings as used by \citet{Hayes2008a} in their similar study of English onset consonants.
-This includes the ``accuracy schedule'' [0.001, 0.01, 0.1, 0.2, 0.3] and no constraints on the number of constraints learned, as well English consonantal features used by \citeauthor{Hayes2008a}.
-Since this model is inherently stochastic, producing slightly different outcomes on each run, I follow the practice of \citeauthor{Hayes2008a} (p.~396) and report the lowest-accuracy of ten runs, though in general there is not a great deal of variation between individual runs.
-
-This model provides a small improvement in accuracy (0.877) and $F_1$ (0.703) compared to the alternation baseline, but it has lower recall (0.642) than this baseline.
-This indicates that the model produces many false negatives.
-%; it assigns the lowest probability value to 122 unattested clusters.
-This seems to further indicate that this model internalizes narrower phonotactic generalizations than those derived from the three alternation-derived constraints; for instance, it assigns the highest probability to unatteested *[m.kl], which is ruled out by \textsc{Coda Nasal Place Assimilation}.
-On the other hand, it also rules out clusters like those in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are consistent with all three alternation-derived constraints.
-
-\citet{Berent2012} identify an important defect with the publically-available version of the \citeauthor{Hayes2008a} model used here: it cannot directly incorporate constraints like the alteration-derived constraint disfavoring geminate consonants, because it does not include variables over phonological features.
-However, the model appears to have experienced no problems internalizing the generalization that geminate clusters are unattested.
-
-The results for the four models are summarized below.
-
-\ex Summary of computational model accuracy: \vspace{6pt} \\
-\begin{tabular}{l | r r r r}
-\toprule
-                          & accuracy & precision & recall & $F_1$ \\ 
-\midrule
-Null baseline             & 0.813    & n.a.      & n.a.   & n.a.  \\
-Alternations              & 0.857    & 0.401     & 0.708  & 0.512 \\
-\citet{Pierrehumbert1994} & 0.839    & 0.210     & 0.750  & 0.328 \\
-\citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
-\bottomrule
-\end{tabular}
-\xe
-
-\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{I am grateful to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are more frequent than those like [t.j], with rising sonority.
-\citeauthor{McGowan2011} reports, however, that this accounts for only a tiny amount of the variance in frequency ($R^2 = 0.077$).
-Change in sonority provides no improvement to the null baseline, and so this model is not reported here.
+\subsubsection{Basic correlations}
+\subsubsection{Intramodal correlations}
+\subsubsection{Residual correlations}

### 2012-04-18 2.3.tex
Rearranged and some writing on 2-5.
Drastic restructuring and some writing on chapters 2-3.
More arranging.
chapters 2 and 4 have been swapped

@@
-% 2.3: Conclusion
+\section{Conclusions}
 
-The above is a stark reminder.
-
-A reasonable objection to the arguments presented in this chapter is to view the results as little more than an indictment of the results of \citealt{Pierrehumbert1994}.
-
-However, the fact that state-of-the-art models are not capable of providing large improvements to the predictive accuracy indicates
-
- do not have a statistically significant effect on the shape of the English lexicon, but that experiments might turn up evidence that speakers have internalized
-
-shown to be aware of static constraints if they reached statistical significance.
-
- statistically reliable static constraints could be identified
-
-The of
-
-Regarding the historical developments,
-
-\citet{Martin2007}
-
-On the other hand, patterns created by sound change are not guaranteed to persist over time. 
-One example of non-persistence is discussed by \citet{Iverson2005}.  
-Around 1100 CE, Old English \emph{sk} became [ʃ]. 
-This sound change introduced no alternations.
-Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{V\lm sk\#} words when the
- change was actuated, and \emph{V\lm sh\#} continues to be rare in Modern English. 
-What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.
-
-A third pattern is that a historically inherited pattern
-
-\citet[][140]{Frisch2004} suggest that the strong tendency for the first and second consonants of the Arabic root to be non-identic
-al is the ``a diachronic result of a processing constraint that disfavors repetition.'' 
-Unfortunately, there is no evidence that this pattern is diachronic other than in the sense that it appears to be inherited from the proto-language: there is simply no Proto-Semitic verb roots with identical first and second consonants \citep[][178]{Greenberg1950}.
-In other Semitic languages, the inherited patern has experienced considerable erosion.
-
-\ex Tigrinya roots with identical first and second consonants \citep{Buckley1990a}: \\
-\begin{tabular}{l l l l}
-a. & lʌlʌw     & `scorch'                   & (< Ge'ez \emph{lʌwlʌw} `inflame')     \\
-   & mʌmʌy     & `winnow'                   & (< Ge'ez \emph{mʌymʌy} `distinguish') \\
-%   & mʌmʌt & `pick out loot' & (< 
-b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
-   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
-c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
-   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
-\end{tabular}
-\xe
-
-Similar exceptions are found in 
-%Amharic (\citealp[][?]{Broselow1984}, \citealp[][?]{McCarthy1985}) and 
-Hebrew \citep[][29]{Bat-El2005}.
-
-The next two chapters return to the question of synchrony, addressing the relationship between statistical patterns in the lexicon and speakers' behaviors when presented with underrepresented sequences.
+It is not the case that this evidence is inconsistent with the hypothesis that there are additional degrees of wordlikeness: it is just a demonstration that these data to not require it.

### 2012-04-18 2.tex
Drastic restructuring and some writing on chapters 2-3.
More arranging.
chapters 2 and 4 have been swapped

--- 2012-02-16 2.tex
@@
-% 2: Accidental and structural gaps in English syllable contact
+\chapter{Gradient and categorical aspects of wordlikeness judgements}
+\label{gradience}
 
-In the preceding chapter, it was proposed that
+(though few if any phonologists regard $n$-gram segment probabilities as a core component of phonology)
 
-Reconsider the null hypothesis presented in the preceding chapter:
+%\begin{quotation}
+%``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
+%\end{quotation}
 
-\ex The null hypothesis: \\
-Universal grammar does not countenance constraints on the contents of URs
-\xe
-
-The idea
-
-relationship to proposal
-
-One primary way which this null hypothesis has been critiqued is by demonstrating that the lexicon of some language 
-exhibits dispreferences or gaps corresponding to.
-
-argue by trends in the lexicon
-
-possibility of sparsity (C\&H, Fischer/Vogt and recent stats)
-
-Of course, to obtain the desired results, we must guarantee that each sequence structure rule reflects a general systematic fact about the language, and not a fact which is due merely to the existence of accidental gaps in the lexicon.
-\citep[][401, fn.~8]{Stanley1967}
-
-In the preceding chapter, it was proposed that
-
-Lexicon
-
-This appears to be exceptionlessly true of Arabic
-
-That even gradient patterns
-
-One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. 
-Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
-This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a} but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Dmitrieva2008a,Dmitrieva2008b,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.
-
-The absence of 
-The claim here is that however stark the absence or underrepresentation of some form is, it
-
-This hypothesis admits the possibility that there may be accidental gaps in the lexicon, a possibility which predates generative thinking:
-
-\begin{quotation}
-\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
-\citep[][16]{Fischer-Jorgensen1952}
-\end{quotation}
-
-\noindent
-Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters
-
-\begin{quotation}
-Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters.
-The material will never be complete.
-It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary.
-\citep[][30]{Vogt1954}
-\end{quotation}
-
-\citet{Chomsky1965}
-
-This chaper demonstrates that apparently systematic gaps may arise even in the absence of phonotactic preferences, and that this possibility compromises attempts to show that such preferences shape the lexicon.
+\input{2.1}
+\input{2.2}
+\input{2.3}

### 2012-04-18 3.1.1.tex
chapters 2 and 4 have been swapped

--- 2012-02-17 3.1.1.tex
@@
-% 3.1.1: Alternation vs. phonotactic
+\subsection{Backness harmony}
 
-\subsubsection{Roundness harmony}
+One of the most salient properties of Turkish is its harmony system, which \citeauthor{Lees1966b} (\citeyear[][35]{Lees1966b}, \citeyear[][284]{Lees1966a}) derives by means of feature spreading rules.
+The least constrained of these processes is the rule of backness harmony described in this section.
 
-\subsubsection{Labial attraction}
+\subsubsection{Phonological description}
 
-\citet{Zuraw2000}
-\citet{Clements1982}
-\citet{Inkelas1997}
+Backness harmony spreads the value of \textsc{Back} rightward over any intervening consonants onto the next vowel.
+
+%\citep[after][229]{Clements1982}: \\
+\begin{example}
+\textsc{Backness Harmony}:
+
+\xymatrix@R=24pt@C=24pt{
+\txt{V}                                        & \txt{C$_0$} & \txt{V} \\
+\txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{..}[urr] \\
+}
+\end{example}
+
+Along with the rule of \textsc{Roundness Harmony} discussed in the next section, this rule accounts for a great deal of the suffix allomorphy found in Turkish.
+The nominative plural (nom.pl.) is formed by adding the suffixes \emph{-ler} or \emph{-lar} to the nominative singular (nom.sg.).
+
+\begin{example}
+The Turkish nominative:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l}
+   & \emph{nom.sg.} & \emph{nom.pl.} \\ 
+a. & ip             & ipler          & `rope' & \citep[][216]{Clements1982} \\
+   & el             & eller          & `hand'    \\
+   & köy            & köyler         & `village' \\
+   & yüz            & yüzler         & `face'    \\
+   & kız            & kizlar         & `girl'    \\
+   & sap            & saplar         & `stalk'   \\
+   & son            & sonlar         & `end'     \\
+   & pul            & pullar         & `stamp'   \\
+b. & neden          & nedenler       & `reason'  & (TELL) \\
+   & boğaz          & boğazlar       & `throat'  \\
+   & kiler          & kilerler       & `pantry'  \\
+   & pelür          & pelürler       & `tissue paper' \\
+   & sapık          & sapıklar       & `pervert' \\
+\end{tabular}
+\end{example}
+
+\noindent
+\textsc{Backness Harmony} correctly predicts the choice of suffix.
+There are a few complications, however.
+First, while most polysyllabic roots conform to \textsc{Backness Harmony} (\lastx b), not all do.
+In this case, suffix vowels generally harmonize with the final root vowel (\nextx a).
+Yet there is also a small class of nouns which exhibit [$-$\textsc{Back}] suffix vowels despite the fact that their final root vowel is [$+$\textsc{Back}] (\nextx b).
+
+\begin{example}
+Exceptional Turkish nominatives:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l}
+   & \emph{nom.sg.} & \emph{nom.pl.} \\
+a. & mezar          & mezarlar       & `grave' & \citep{TELL} \\
+   & model          & modeller       & `model' \\
+   & silah          & silahlar       & `weapon'     \\
+   & memur          & memurlar       & `bureaucrat' \\
+   & sabun          & sabunlar       & `soap'       \\
+b. & saat           & saatler        & `hour, clock' \\
+   & harf           & harfler        & `(alphabetic) letter' \\ %& \citep{Goksel2005}
+   & etol           & etoller        & `fur stole' \\
+\end{tabular}
+\end{example}
+
+While it is uncontroversial that disharmonic suffixes (\lastx b) are no more than sporadic exceptions to \textsc{Backness Harmony}, the status of root disharmony (\lastx a) has been the subject of much debate.
+\citet[][212, 289]{Anderson1974} views the fact that disharmonic roots still trigger suffix harmony as evidence that suffix harmony is distinct from root harmony, the latter being a sequence structure constraint.
+
+%Once additional source of evidence on root (dis)harmony is inconclusive.
+There is a small class of bisyllabic words in which the second vowel, always [$+$\textsc{High}, $-$\textsc{Back}], alternates with zero.
+
+%\ex High-vowel/zero alternations \citep[][243]{Clements1982}: \\
+%\begin{tabular}{l l l l}
+%   & nom.sg. & gen.sg. \\
+%a. & fikir   & fikri  & `idea' \\
+%   & hüküm   & hükmün & `judgement' \\
+%%  & filim   & filmi & `film' & \citep[][178]{Inkelas2001} \\
+%b. & vakit   & vaktin & `time' \\
+%   & rahim   & rahmin & `womb' \\
+%\end{tabular} \xe
+%
+%\noindent
+%It is possible that \textsc{Backness Harmony} might produce a fluctuating \emph{ı} after root \emph{a}, but this does not obtain (\lastx b).
+However, this might simply indicate that the fluctuating vowel is epenthetic and that harmony applies before epenthesis (see \citealt{Clements1982} for both sides of this argument), making it less than a counterexample.
+
+Underspecification provides an alternative under which the exceptionality of disharmonic roots is only apparent.
+While it is clear that the first vowel of Turkish roots are contrastively specified for backness (e.g., \emph{kül} `ash' vs.
+\emph{kul} `servant', \emph{kepek} `bran' vs.
+\emph{kapak} `lid'), there are also some minimal pairs with regards to backness (dis)harmony (e.g., \emph{deve} `camel' vs.
+\emph{deva} `medicine', \emph{sene} `year' vs.
+\emph{sena} `praise').
+This supports the possibility, first suggested by \citet{Clements1982}, that harmonic roots may be underspecified and disharmonic roots fully specified, exemplified below.
+
+%There is some further evidence that individual vowels may differ in specification for this feature even within individual roots or affixes.
+For instance, the present continuous suffix has harmony-determined allomorphs \emph{-iyor}, \emph{-üyor}, \emph{-ıyor}, \emph{-uyor}, but the \emph{o} of the suffix is invariant.
+A similar situation might obtain in Turkish roots.
+
+\begin{example}
+Underlying specification of \textsc{Back} for (dis)harmonic roots:
+
+\xymatrix@R=24pt@C=24pt{
+\txt{a.} & \txt{s} & \txt{V} & \txt{n} & \txt{V}  & \txt{\emph{sene} `year'} \\
+&        & \txt{[$-$\textsc{Back}]}\ar@{-}[u]     \\
+\txt{b.} & \txt{s} & \txt{V} & \txt{n} & \txt{V} & \txt{\emph{sena} `praise'} \\
+         &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u] & & \txt{[$+$\textsc{Back}]}\ar@{-}[u]
+}
+\end{example}
+
+\noindent
+Harmonizing suffix vowels will also be underspecified for \textsc{Back}.
+Of course, \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being the use of a a \textsc{Structure Preservation} condition \citep{Kiparsky1985}.
+Unfortunately, this detail reintroduces the duplication alluded to above.
+Any condition on the rule of \textsc{Backness Harmony} which prevents overwriting entails that it has no control over the distribution of harmonic and disharmonic roots, and thus fails to account for predominance of harmonic roots.
+
+However, the theory of exceptionality presented in \emph{SPE} (p.~374f.) provides a direct account of suffix harmony in disharmonic roots without duplication.
+In \emph{SPE}, the specification of the target (i.e., the segment to be changed) of every rule $R$ must be [$+$rule $R$] by convention.
+A root or affix which fails to undergo $R$ despite otherwise matching the structural description is simply said to be marked [$-$rule $R$], and thus failing to meet the full structural description.
+If disharmonic roots are [$-$\textsc{Backness Harmony}], then suffix vowels are correctly predicted to undergo \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the environment, which is not required to be [$+$rule $R$].
+
+\subsubsection{Psycholinguistic evidence}
+
+The evidence from alternations leaves open the question of whether speakers internalize a generalization regarding the tendency of roots to conform to \textsc{Backness Harmony}.
+One piece of evidence that speaks in favor of root-internal harmony comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but it corresponds to a morphological process native to the related language Tuvan, in which it conveys a sense of ``vagueness or jocularity'', and \citeauthor{Harrison2001} report that it can be quickly taught to even young Turkish speakers.
+The game consists of reduplication of the base and replacement of the first vowel in the reduplicant with \emph{a} or \emph{u}.
+
+In both Tuvan and Turkish, the unchanged reduplicant vowel is also affected.
+Reduplication interacts with root harmony in both Tuvan and Turkish.
+When the base is harmonically [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] results in what \citeauthor{Harrison2001} call ``reharmonization'' (\nextx a).
+
+\begin{example}
+\label{redupgame}
+Turkish reduplication game \citep[][231]{Harrison2001}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l}
+a. & kibrit & kibrit-kabrıt & `match'    \\
+   & bütün  & bütün-batın   & `whole'    \\
+b. & mali   & mali-muli     & `Mali'     \\
+   & butik  & butik-batik   & `boutique' \\
+\end{tabular}
+\end{example}
+
+\noindent
+Under the underspecification hypothesis, non-initial harmonic vowels lack an underlying \textsc{Back} feature, so it comes as no surprise that changing the \textsc{Back} specification of the root-initial vowel results in reharmonization.
+And the full specification of disharmonic roots correctly predicts that they will be exempt from reharmonization (\lastx b), which is also borne out in Tuvan and in an unrelated Finnish language game \citep{Campbell1986}.
+
+A number of studies have found that speakers of Finnish use disharmony in speech processing experiments.\footnote{Thanks to Charles Yang for pointing out the relevance of these studies to me.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which are harmonic for the feature \textsc{Back}.
+These stimuli are auditorily presented to subjects who are asked to press a button when the nonce trisyllable ends with a target nonce disyllable, or a real disyllabic word.
+Speakers are quicker to press the button when the prefix and disyllabic word disagree for \textsc{Back}.
+These results suggest that speakers are attuned to the fact that disharmonic transisitions are good predictors of word boundaries.
+If speakers have also internalized the converse generalization, that harmonic transistions are more likely to be root-internal, then there is additional evidence that harmony is active not just in Finnish affix alternations but also in roots.
+
+\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect on word-spotting as it does in Finnish: speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by the pseudo-prefix \emph{gölü-}, a disharmonic transistion, than when it is preceded by the pseudo-prefix \emph{golu-}, a harmonic transition.
+\citeauthor{Kabak2010} find that effect of harmony does not obtain for speakers of French, a language which lacks vowel harmony.
+As in Finnish, the results imply speakers have internalized the predominance of root-internal harmony.
+
+It seems that the root-internal harmony bias is in fact learned by Turkish speakers very early.
+The pseudoword spotting experiment has been adapted for 9-month-old Turkish infants by \citet{Kampen2008}.
+Infants are familiarized with harmonic disyllabic pseudowords bearing a pseudo-prefix, which may be harmonic or disharmonic.
+At test time, the infants are played the disyllabic pseudowords in isolation using the head turn preference paradigm.
+Infants show a preference to listen to those pseudowords which were familiarized with a disharmonic pseudo-prefix over those which were familiarized with a harmonic pseudo-prefix.
+This preference is not observed in 9-month-old infants learning German, which also lacks vowel harmony.
+Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants show no such preference.

### 2012-04-18 3.1.2.tex
chapters 2 and 4 have been swapped

--- 2012-02-17 3.1.2.tex
@@
-% 3.1.2: Lexical statistics
+\subsection{Roundness Harmony}
 
-This is essentially the hypothesis pursued in the previous chapter.
+The rule of roundness harmony and the data that motivates it overlaps with the preceding evidence for \textsc{Backness Harmony}.
 
-\citet{TELL}
-\citet{Inkelas2001}
+\subsubsection{Phonological description}
 
-%We might, for instance, consider methods based not on statistical \citep{Yang2005,Gorman2011b}.
+The environment for roundness harmony differs from \textsc{Backness Harmony} in that it requiers the target to be [$+$\textsc{High}].
+
+\begin{example}
+\textsc{Roundness Harmony}:
+
+\xymatrix@R=24pt@C=24pt{
+\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{..}[drr] &             & \txt{[$+$\textsc{High}]} \\
+\txt{V}                                         & \txt{C$_0$} & \txt{V}\ar@{-}[u] \\
+}
+\end{example}
+
+This rule, in concert with \textsc{Backness Harmony}, accounts for the forms of the dative singular (dat.sg.) and genitive singular (gen.sg.), among other suffixes.
+
+\begin{example}
+Turkish nominal suffix allomorphy:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l l l}
+   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
+a. & ip             & ipi            & ipin           & `rope' & (CS:216) \\
+   & el             & eli            & elin           & `hand'    \\
+   & kız            & kızı           & kızın          & `girl'    \\
+   & sap            & sapı           & sapın          & `stalk'   \\
+   & yüz            & yüzü           & yüzün          & `face'    \\
+   & köy            & köyü           & köyün          & `village' \\
+   & pul            & pulu           & pulun          & `stamp'   \\
+   & son            & sonu           & sonun          & `end'     \\
+b. & boğaz          & boğazı         & boğazın        & `throat'  & (TELL) \\
+   & pelür          & pelürü         & pelürün        & `tissue paper' \\
+   & döviz          & dövizi         & dövizin        & `currency' \\
+   & yamuk          & yamuğu         & yamuğun        & `trapezoid' \\
+   & ümit           & ümiti          & ümitin         & `hope'     \\
+\end{tabular}
+\end{example}
+
+\noindent
+Much as was the case for \textsc{Backness Harmony}, there are disharmonic roots which participate in suffix harmony.
+Once again, underspecification of harmonic roots and full specification of disharmonic roots allows for the use of the \emph{SPE} exception convention.
+
+\subsubsection{Psycholinguistic evidence}
+
+The only psycholinguistic evidence for \textsc{Roundness Harmony} comes from \citeauthor{Harrison2001}'s language game (\ref{redupgame}).
+As shown above, the second vowel in the reduplicated form of \emph{bütün} `whole' is \emph{bütün-batın}.
+The second vowel in the base, \emph{ü}, is reharmonized as [$+$\textsc{Back}, $-$\textsc{Round}], indicating that \textsc{Roundness Harmony} also participates in reharmonization.

### 2012-04-18 3.1.3.tex
chapters 2 and 4 have been swapped

\subsection{Labial attraction}

The final rule

\subsubsection{Phonological description}

\citet[][36]{Lees1966a} describes \textsc{Labial Attraction} as a process by which ``a high, short harmonic vowel is rounded in the second syllable of a disyllabic word whose first vowel is /a/, and whose medial consonant cluster contains a labial /p, b, m, v/, and then it is de-harmonified''. This description is transalted into an autosegmental rule below.

\begin{example}
\textsc{Labial Attraction} (after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}):

\xymatrix@R=24pt@C=8pt{
\txt{[$-$\textsc{Round}]} &                                         & \txt{[$-$\textsc{High}]} & \txt{[$+$\textsc{Labial}]}    & \txt{[$+$\textsc{High}]}\ar@{-}[dr] &         & \txt{[$+$\textsc{Round}]}\ar@{..}[dl] \\
                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C$_0$ C C$_0$}\ar@{-}[u] &                                      & \txt{V} & \\
                         & \txt{[$+$\textsc{Back}]}\ar@{-}[urrrr] 
}
\end{example}

The formalization of this rule is naturally complex, and perhaps obscures the fact that \textsc{Labial Attraction} generates exceptions to \textsc{Roundness Harmony}; it produces \emph{aCu} sequences (e.g., \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', and \emph{samur} `sable' \citep[][285]{Lees1966b}) instead of the otherwise expected \emph{aCı}. \citeauthor{Lees1966b} notes the existence of exceptions (e.g., \emph{tavır} `mode') but writes that they are they are ``surprisingly rare'' (ibid., 286); \citet[][225]{Clements1982} note additional exceptions.

There is reason to believe that \textsc{Labial Attraction} is at best a sequence structure constraint as it never applies in derived environments. If, contrary to fact, there was a \textsc{Labial Attraction} alternation, then one would expect, for example, that the gen.sg. of \emph{sap} `stalk' would be *\emph{sapun} instead of the observed \emph{sapın}.

\citet{Lees1966a}
\citet{Zimmer1969}

\begin{quote}
\ldots decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots~aCu~\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots We conclude that there is no systematic restriction on the set of consonants that may occur medially in rotos of the form /\ldots~aCu~\ldots/. \citep[][225]{Clements1982}
\end{quote}

\noindent
\citeauthor{Clements1982} appear to be suggesting that \textsc{Labial Attraction} implies that \emph{aTu}, where \emph{T} represents a non-labial consonant, should be infrequent, but this fact is not inconsistent with the formulation of the constraint by \citet{Lees1966a,Lees1966b} and \citet{Zimmer1969}; \emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description. This is less than conclusive, since the fluctuating vowel alternation also fails to undergo harmony, and might just indicate that the flucutating vowel is epenthesized relatively late in the derivation; the frequency of \emph{aTu} provides only indirect information about \textsc{Labial Attraction} in that it provides a baseline estimate for just how frequent this disharmonic sequence of vowels is.

\citet{Inkelas1997}
\citet{Inkelas2001}

\begin{quote}
Lee's rule of \textsc{Labial Attraction}\ldots is not a real generalization about the Turkish lexicon. It is not true synchronically, either of native or nonnative items; nor, according to the historical and dialectical literature, does \textsc{Labial Attraction} appear to have been true at any stage going back as far as Old Turkic. \citep[][196]{Inkelas2001}
\end{quote}

zimmer: 319-320 
inkelas et al.: 196

\subsubsection{Psycholinguistic evidence}

To the author's knowledge, there are no psycholinguistic studies investigating \textsc{Labial Attraction} beyond the experimental results of \citet{Zimmer1969} reviewed in the next section.

### 2012-04-18 3.1.tex
chapters 2 and 4 have been swapped

--- 2012-02-17 3.1.tex
@@
-% 3.1: Wordlikeness predictions
+\section{Turkish vowel sequence structure constraints}
 
-\citet{Zimmer1969}
+\citet{Lees1966a,Lees1966b} proposes three sequence structure constraints on Turkish vowels, and these constraints have been discussed extensively in the subsequent literature.
+Below, these constraints are formalized and related to Turkish alternation phonology and various sources of external evidence.
+A traditional binary feature specification for the Turkish vowel system, given below, is assumed without discussion.
+
+\begin{example}
+Turkish vowel features:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{c | c c c c}
+                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
+                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
+\buf[$+$\textsc{High}] & i & ü [y] & ı [ɯ] & u \\
+\buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
+\end{tabular}
+\end{example}
+
+\input{3.1.1}
+\input{3.1.2}
+\input{3.1.3}

### 2012-04-18 3.2.2.tex
chapters 2 and 4 have been swapped

@@
 
-\begin{quotation}
+\begin{quote}
 Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
 \citep[][196, emphasis in original]{Inkelas2001}
-\end{quotation}
+\end{quote}
 
@@
 
-\ex $\displaystyle \gamma = \frac{C - D}{C + D}$ \xe
+\begin{example}
+$\displaystyle \gamma = \frac{C - D}{C + D}$
+\end{example}
 
@@
 
-Backness harmony wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
+\begin{example}
+Backness harmony wordlikeness forced choices \citep[314]{Zimmer1969}:
+
+\vspace{0.5\baselineskip}
 \begin{tabular}{l r l r}
@@
 pemez & 30                   & pemaz & 2  \\
-teriz & 28                   & terız & 3  \\
-tokaz & 26                   & tokez & 6  \\
 tipez & 24                   & tipaz & 8  \\ 
 terüz & 19                   & teruz & 13 \\ % roundness violator
+teriz & 28                   & terız & 3  \\
+tokaz & 26                   & tokez & 6  \\ % roundness violator
 \bottomrule
 \end{tabular}
+\end{example}
+
+\subsubsection{Roundness harmony}
+
+\begin{example}
+Roundness harmony forced choice wordlikeness judgements:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r l r}
+\toprule
+\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
+\midrule
+pörüz & 32 & pöriz & 0  \\
+tatız & 20 & tatuz & 12 \\
+tüpüz & 31 & tüpiz & 1  \\
+takız & 22 & takuz & 10 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+%tamaz & 26 & tamoz & 6  \\
+%putoz & 25 & putaz & 7  \\
 
@@
 
-Labial attraction wordlikeness forced choices \citep[314]{Zimmer1969}: \vspace{6pt} \\ 
+\begin{example}
+Labial attraction wordlikeness forced choices \citep[314]{Zimmer1969}:
+
+\vspace{0.5\baselineskip}
 \begin{tabular}{l r l r}
@@
 \midrule
+pamuz & 17              & pamız & 15 \\
 tafuz & 21              & tafız & 11 \\
-pamuz & 17              & pamız & 15 \\
 tapuz & 17              & tapız & 15 \\
@@
 \end{tabular}
+\end{example}
 

### 2012-04-18 3.tex
chapters 2 and 4 have been swapped

@@
 
-This is not a universal, but rather something English speakers must learn: some languages, such as Moroccan Arabic, do in fact permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet').
+glossematics
 
-%bni 'building'
-%bnin 'delicious'
-%bniqa 'closet' 
-%bnat 'girl' 
-%bnita 'girl (dim.)'
+\citet{Chomsky1965} propose
 
-a matter of universal
-
-In their illustration of
-
-One of the
-
-One of the most compelling findings
+In retrospect, it seems somewhat odd that \citeauthor{Chomsky1965} would coh
+It might 
+In defining the
 
 The contrast between \emph{blick} and \emph{bnick} is what has been called a "projection problem" \citep{Baker1979}
+
+This is not a universal, but rather something English speakers must learn, as some languages, such as Moroccan Arabic, do in fact permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet');
+
+\citet{Lees1966b,Lees1966a}
 
@@
 % Dupoux et al. 1999: Japanese speakers still can't distinguish between ebzo and ebuzo in an ABX task using the SAME a or b.
-
-\citet{Daland2011}
 

### 2012-04-18 4.1.1.tex
chapters 2 and 4 have been swapped

@@
-% 4.1.1: Gradience in judgements
+\subsection{Data sources}
 
-at least as far back as \emph{LSLT} \citep{LSLT}
+The English CELEX database \citep{CELEX},
 
-%and has become a recent feature of the discussion
-is that non-gradient models are totally incompatible with psycholinguistic results.
+\citet{Aronoff1976}
 
-\subsubsection{Task model}
+\citet{Harley2009}
 
-\citet[][16]{Schutze1996}
+\citet{Taft1975} ?
+\citet{Taft1981} ?
+\citet{Taft2004a} ?
 
-\citet{Schutze2005}
-\citet{Schutze2011}
+\citet{Halle1962}
 
-%\citet{Coleman1996}
+% 2.1.4: k
 
-\subsubsection{Falsifiability}
+\subsubsection{Corpus counts}
 
-\citet{Armstrong1983}
-
-% wordlikeness critique
-\citet{Yang2008a}
-
-a mix of ``declarative'' knowledge,
-corresponding to rote facts and thus the lexicon,
-and ``procedural'' knowledge, or knowledge of generative procedures 
-(in the sense of J. \citealt{Anderson1993}).
-
-% math cog
-\citet{Logan1988}
-\citet{Sfard1991}
-\citet{Delazer2005}
+Under the phonological analyses described in the next two sections, the 6,876 simplex words in CELEX contain 21 unique medial codas and 40 unique medial onsets.
+Of these 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, for a 18.8\% saturation rate.

### 2012-04-18 4.1.2.tex
More arranging.
chapters 2 and 4 have been swapped

@@
-% 4.1.2: The ontology of wordlikeness
+\subsection{Syllabification}
 
-\subsubsection{Gross phonotactic violation}
+It is necessary to split medial consonant clusters into the codas and onsets that make them up.
+While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic.
+For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language.
+Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which contrast \emph{chemistry} and \emph{ministry}.
 
-position in \citet{SPE}
+An automated syllabification system was constructed and applied to medial clusters in the CELEX data.
+This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.
+There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification.
+The technique used is a variation on the theme of onset maximization \citep[42f.]{Kahn1976}, which parses word-medial clusters so that the onset is as large as possible.
+Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty.
+In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.
 
-\citet{Shibatani1972}
-\citet{Hooper1973}
+There is one case where unchecked maximization of medial onsets produces incorrect syllabifications.
+When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
+Onset maximization makes the right predictions for words like \emph{propaganda}, where the medial cluster cannot be an onset, but \citet[][55]{Harris1994} notes that it makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
+The approach adopted here is to first apply onset maximization, then to reassign the first consonant of a medial onset to the preceding coda if it is immediately preceded by a stressed lax vowel.
 
-peripheral \citep[in the sense of][81]{LGB}
+If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables.
+For this reason, it is assumed here that affricates are single segments for the purposes of syllabification.
+This assumption is motivated by the fact that affricates pattern with simple segments in many languages; for instance, the complex onsets in Classical Nahua are affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.
 
-As an example of this applied ot phonotactics, 
-\citet{Borowsky1989} assumes that there are peripheral morph-internal syllable contact clusters.
-
-\subsubsection{Phonological statistics}
-
-\citet{SPE}
-
-This statement is made very strongly by \citet{Coleman1996}
-
-\subsubsection{Lexical similarity}
-
-\citet{LSLT}
-
-\citet{Greenberg1964}
-\citet{Ohala1986b}
-
-A recent, very articulated model is presented by \citet{Bailey2001}.
-
-% simple model
-This is simply because more complex models fail to
-
-% similarity vs. rules
-% as a final note, \citet{Hahn1998}
+%In fact, [t.ʃ, d.ʒ] clusters are not found in simplex English words, despite the fact that that affricates occur as medial onsets in clusters (e.g., \emph{tru}[n.tʃ]\emph{eon}, \emph{so}[l.dʒ]\emph{er}).

### 2012-04-18 4.1.3.tex
chapters 2 and 4 have been swapped

--- 2012-02-17 2.1.3.tex
@@
-% 2.1.3: Inventory considerations
+\subsection{Inventory considerations}
 
-The syllables derived by the above procedure must further be divided into onset, nucleus, and coda, and mapped to underlying representation.
-In most cases, this is trivial: the nucleus is a pure vowel and can be easily separated from the consonants, which surface faithfully.
-However, the status of several surface phones in the syllable and the relationship of surface segments to underlying forms is not always obvious.
-The treatment of those segments is described and motivated below.
+To identify the components of syllable contact clusters, the syllables identified by the above procedure must be further separated into onset, nucleus, and coda, and mapped to underlying representation.
+This is trivial in most cases: the nucleus is easily separated from onset and coda and surface consonants are faithful to underlying representation.
+The treatment of the few difficult cases is described below.
+Ambiguous segments are assigned to onset, nucleus, or coda, and to underlying form, using a distributional heuristic.
 
@@
 There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right, as demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}.
-The strongest piece of evidence for the pure allophonic analysis (and an attentant process of /g/-\textsc{Deletion}) is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
-\citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is adopted here and formalized below in Section \ref{cnpasection}.
+The strongest piece of evidence for treating the velar nasal as a pure allophone is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
+\citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here.
+English nasal place allophony is formalized in Section \ref{cnpasection} below.
 %\citealt[][62]{Halle1985a}),
@@
 
-I assume that the [j] onglide in words such as \emph{ass}[j]\emph{ume} is present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J.
-\citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}).
-The primary motivation for this assumption is the observation that presence or absence of this glide is contrastive (e.g.
-\emph{coup} $\sim$ \emph{queue}, \emph{booty} $\sim$ \emph{beauty}).
-I further assume that the front glide is not part of the onset except in the case where the onset would ortherwise be null (e.g., \emph{jun}[j]\emph{or}).
+The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J.
+\citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}).
+The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).
 %\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
-When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}.
-However, when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm].
-This defective distribution of vowels following [j] follows if the onglide in this context is the first component of a phonological diphthong, and thus part of the nucleus (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
+There is extensive motivation for this principle.
+When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm].
+In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [u\lm] to be anomalous \citep{Moreland2009}.
+This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
 \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [ju\lm] also suggests the glide is nuclear.
-There is a great deal of external support for this position as well.
-The [ju\lm] in words like \emph{spew} may pattern together
-in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003}, to the exclusion of the rest of preceding tautosyllabic consonants.
+There is also strong external support for this principle.
+The [ju\lm] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants.
 The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
@@
 
-\ex VBR's complex onsets \citep[79--80, his transcriptions]{Buchwald2005}: \\
+\begin{example}
+VBR's complex onsets \citep[79--80, his transcriptions]{Buchwald2005}:
+
+\vspace{0.5\baselineskip}
 \begin{tabular}{l l l}
@@
 \end{tabular}
-\xe
+\end{example}
 
 \noindent
-VBR breaks up complex onsets, including back onglide clusters in (\nextx b).
-However, in (\nextx c), which contain the front onglide in standard English, no epenthesis occurs: the glide is simply absent.
-This also suggests that the front onglide is part of the onset, not the nucleus.
+VBR breaks up complex onsets with epenthesis, including those that consist of a consonant and a back onglide cluster (\lastx b).
+However, no epenthesis occurs between a consonant and a front onglide; rather, the glide is absent (\lastx c).
+The failure of the front onglide to pattern with other consonant clusters suggests once again that the glide is part of the nucleus.
 
@@
 Unlike [ju\lm], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.
+And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.
 
@@
 
-Like this study, \citet{Pierrehumbert1994} uses a dictionary that transcribes RP, in which word-medial post-vocalic \emph{r} has been lost.
-There are independent reasons to believe that \emph{r}-full dialects of English assign post-vocalic \emph{r} to the nucleus and is not part of syllable contact clusters in such dialects.
-First, \citet[][255]{Harris1994} notes that the number of vowel contrasts is greatly reduced before \emph{r} when compared to other coronal consonants.
-For instance, the majority of North America has lost the historical contrast between \emph{Mary}, \emph{marry}, and \emph{merry}, and one holdout, Philadelphia, is losing the the contrast between \emph{merry} and \emph{Murray} \citep[14f.]{ANAE}.
-Another source of evidence for the nuclear status of post-vocalic \emph{r} comes from its failure to trigger two processes in English variable phonology.
+Both CELEX and the dictionary used by \citet{Pierrehumbert1994} in her study transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost.
+In \emph{r}-full dialects, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters.
+Many vowel contrasts are suspended before \emph{r} \citep[][255]{Harris1994}; for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}.
+Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which post-vocalic \emph{r} patterns differently than internal coda consonants.
 \citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.
 
-\ex /t/-\textsc{Glottalization} in British English \citep[after][195, 258]{Harris1994}: \\
+\begin{example}
+/t/-\textsc{Glottalization} in British English \citep[after][195, 258]{Harris1994}:
+
+\vspace{0.5\baselineskip}
 \begin{tabular}{l l l@{} l}
@@
 \end{tabular}
-\xe
+\end{example}
 
@@
 
-\ex /t, d/-\textsc{Deletion} in American English: \\
+\begin{example}
+/t, d/-\textsc{Deletion} in American English:
+
+\vspace{0.5\baselineskip}
 \begin{tabular}{l l l@{} l}

### 2012-04-18 4.2.1.tex
More arranging.
chapters 2 and 4 have been swapped

--- 2012-02-17 4.2.1.tex
@@
-% 4.2.1: Assessing gradience
+\subsection{Static and derived constraints}
 
-\subsubsection{Response distributions}
+With this database of English syllable contact clusters it is now possible to evaluate constraints on sequences of underlying phonemes.
 
-This is shown in Figure \ref{density}.
+Here, the metric used is simply
 
-\begin{figure}
-\centering
-\includegraphics{density.pdf}
-\caption{Density of responses}
-\label{density}
-\end{figure}
+The evaluation
 
-\subsubsection{Prediction distributions}
+The foremost evidence for
 
-\citet{Coltheart1977}
+The evaluation used here is simply
 
-\citet{Hayes2008a}
-\citet{Albright2009}
+Having described the database of English syllable contact clusters,
 
-% stats thereof
-%\citet{Schilling2002}
-%\citet{Helguerro1904}
-%\citet{Cohen1956}
+I now consider constraints on sequences of URs as predictors of attested and unattested contact clusters.
+First, I describe the statistical technique used, then apply this technique to both phonologically derived constraints and a set of ``static'' phonotactic constraints proposed by \citet{Pierrehumbert1994}.
 
-% EM
-%\citet{EM}
+\subsubsection{Statistical technique}
+\label{stattech}
 
-% -2 log like ratio test
+Before the dawn of generative phonology, many linguists concerned themselves with documenting co-occurrence in various languages.
+\citet[][28]{Vogt1954} declares that a ``phonemic description of a language \ldots should also comprise a description of the phoneme combinations that occur or may be expected to occur in the language''.
+Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively.
+\citeauthor{Uhlenbeck1950}, for instance, puts forth co-occurrence dispreferences, rather than exceptionless generalizations of the sort given by \citet{Greenberg1950}.
+Years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and proposes a statistical technique, based on the chi-square test, to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.
+
+\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English).
+The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented.
+Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.
+
+\begin{example}
+$\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
+χ^2$
+\end{example}
+
+%\footnote{This is the one-tailed version of the test.}
+
+In this study, I use the \citet{Fisher1934} exact test, which is isomorphic to this chi-square test.
+Compared to the chi-square test, the Fisher exact test $p$-value is somewhat more difficult to compute, but both are automated by modern statistical packages and can be computed rapidly.
+The major advantage of the Fisher test over the chi-square test is that it is appropriate for small samples or rare events, whereas the chi-square test depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples \citep[see][]{Gorman2012a}.
+%\citep[e.g.,][142]{Mester1988}.
+
+On the other end of the spectrum, the Monte Carlo statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010} does not scale to large samples.
+The Monte Carlo procedure is isomorphic to the one-tailed Fisher Exact test, and involves comparing observed co-occurrence counts to those produced by randomly generated samples.
+This requires the researcher to generate random permutations, something which is particularly difficult for large samples, as I will explain; disinterested readers may wish to skip the following paragraph.
+
+Any permutation of a sequence $L$ can be represented as a sequence $S$ of the same length, where the value of $S_i$ corresponds to the permuted position of $L_i$.
+Generating random permutations requires an unbiased method to generate all $N!$ of these lists with equal probability.
+Unfortunately, this is near impossible for $N$s of moderate size with current computational resources.
+Any pseudo-random number generator can be characterized by its ``period'', the number of pseduo-random numbers it can emit before repeating itself; this is also the upper limit for the number of random permutations that can be generated.
+The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a period of approximately $2^{48}$, a number which turns out to be smaller than $17!$.
+For the large samples, sometimes in the thousands, used by these researchers, virtually all the possible permutations can never be generated by this method.
+As a consequence, the shuffling procedure is far from random and introducing the possibility that the statistical test is biased in unpredictable ways.
+
+As an example of the Fisher exact test as it is applied here, consider some of the Javanese co-occurrence facts considered by \citet{Mester1988}.
+\citeauthor{Mester1988} is at pains to show that there is a gradient dispreference for first and second root consonants which share the same major place.
+However, perfectly identical first and second consonants are exempt.
+Since
+
+% APPLY THIS TO JAVANESE
+
+\begin{example}
+Javanese root co-occurrence (\citealp[][264]{Uhlenbeck1950}, \citealp[][139]{Mester1988}):
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 134      & 12         & 0.918      & \multirow{2}{*}{6.095\e{-06}} \\
+violating  & 21       & 15         & 0.583 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent
+It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.
+
+This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance
+
+That this is highly unlikely to be due to chance is apparent both from the chi-square ($\chi^2 = 25.572$, two-tailed $p = 4.3$\e{-07}) and Fisher exaact tests ($p =  6.095$\e{-06}).
+
+I use two-tailed statistical tests throughout. 
+Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis.
+As
+
+\citet{Pierrehumbert1994}
+
+It must be stressed that under current assumptions, this is independent of whether predictable prosodic representations (particularly coda and onset identity) are part of the memorized underlying representation \citep[e.g.,][]{Vaux2003}.
+It is an undeniable fact that URs need to be 
+The derivative nature of 
+\citet{Ito1989a,Noske1992}
+
+least some researchers (i.e., \citealt{Mester1988})
+
+ \citep[e.g.,]{Brown2010} have argued for phonotactic constraints giving rise to overrepresentation, so I admit this possibility here and use two-tailed tests throughout.
+
+%\footnote{For example, \citet{Martin2007} applies this technique to a list of 4,758 noun-noun compounds.
+The number of permutations of a list of this length can be estimated using Stirling's approximation.
+
+%\ex $ \lim_{n \rightarrow \infty} n! = e ^ {n \ln{n} - n} = 2 ^ {\log_2{e} (n \ln{n} - n)} $ \xe
+
+%\noindent
+%$4,758!$ is approximately $2 ^ {51260}$. The author is unaware of any system that provides 51,260 bits of entropy; and some popular programming languages, such as ANSI C, provide as few as 15 bits.}
+
+%%% START OF SERIOUSNESS
+
+I will first consider three well-known English phonological alternations, showing that these give rise to constraints on the sequence structure of morphs.
+Simple autosegmental formalizations of these processes are provided, and the Fisher exact test is used to demonstrate their effect on the English lexicon.
+
+\subsubsection{Obstruent Voice Assimilation}
+
+English has a process of voice assimilation which is found in two inflectional suffixes.
+The suffix /-d/ forms regular pasts and denominal adjectives, and /-z/ forms regular noun plurals and genitives, and the 3sg.
+present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S.
+\citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][210f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
+
+\begin{example}
+Voice assimilation in the past tense and noun plural:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l l}
+a. & /næp-d/     & \goesto & [næpt]     & (cf. \emph{nab}[d], \emph{haul}[d]) \\
+b. & /læp-z/     & \goesto & [læps]     & (cf. \emph{lab}[z], \emph{clan}[z]) \\
+\end{tabular}
+\end{example}
+
+\noindent
+The voiced obstruent suffixes devoice when preceded by a voiceless obstruent.
+This is formalized as a rule spreading the [\textsc{Voice}] specification rightward over adjacent obstruents.
+
+\begin{example}
+\textsc{Obstruent Voice Assimilation} (\emph{SPE}:178):
+
+\xymatrix@R=24pt@C=24pt{
+\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{..}[dr] & \\
+\txt{C}\ar@{-}[d]             & \txt{C}\ar@{-}[d] \\
+\txt{[$+$\textsc{Obstruent}]} & \txt{[$+$\textsc{Obstruent}]}\\
+}
+\end{example}
+
+\noindent
+This rules out the possibility of surface clusters of obstruents with differing  [\textsc{Voice}] specifications.
+While the data make it clear \textsc{Obstruent Voice Assimilation} applies in morphologically derived environments, it is not obvious that it applies to obstruent clusters in non-derived environments (e.g., obstruent clusters in the URs of morphs) as well.
+\citet{Pierrehumbert1994} does not mention voice assimilation in her study of gaps in English syllable contact clusters.
+\citet[][74f.]{Hammond1999a} notes the existence of words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain obstruent voice clusters disagreeing in [\textsc{Voice}], and takes this as evidence that obstruent voicing is restricted to derived environments.
+
+Despite these exceptions, I propose that \textsc{Obstruent Voice Assimilation} plays a major role in shaping the English lexicon.
+Only English syllable contact clusters which contain two adjacent obstruents speak to the status of \textsc{Obstruent Voice Assimilation}; the remaining 160 possible clusters are ignored.
+A $2 \times 2$ contingency table is constructed by dividing the remaining possible clusters into those which agree on obstruent voicing and those which do not, and into those which are attested and those which are not.
+
+\begin{example}
+Lexical effects of \textsc{Obstruent Voice Assimilation}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 80 & 370 & 0.178 & \multirow{2}{*}{1.1\e{-11}}\\
+violating  &  6 & 264 & 0.022 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent 
+18\% of the possible clusters containing adjacent obstruents with the same [\textsc{Voice}] specification are found, whereas only 2\%  of those which disagree in [\textsc{Voice}], such as in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, CELEX also contains \emph{jo}[d.p]\emph{urs}, \emph{pi}[ntʃ.b]\emph{eck}, \emph{sa}[k.b]\emph{ut}, and \emph{ja}[k.d]\emph{aw}.
+As the $p$-value shows, the rarity of these disagreeing clusters is unlikely to be due to chance.
+
+I propose that this pattern should be attributed directly to \textsc{Obstruent Voice Assimilation} applying to non-derived environments, preventing learners from positing URs in which adjacent obstruents do not agree in voice.
+There is no duplication between constraints on sequence structure and surface forms if both are derived from the application of the phonological rule (\citealt[][401f.]{Stanley1967}, \citealt[][382]{SPE}).
+The handful of exceptions to this generalization can be marked with appropriate exception diacritics.
+Another possibility is suggested by recent psycholinguistic results.
+\citet{Mattys2001b} find that 9-month old infants are sensitive to the contrast between [vt], a cluster of obstruents disagreeing in [\textsc{Voice}] which is not found word-medially, and [ft], which conforms to the above generalization and is found medially (e.g.
+\emph{a}[f.t]\emph{er}).
+This ability to use phonotactic violations to segment continuous speech persists into adulthood \citep{McQueen1998}, suggesting that ``hetero-voiced'' obstruent clusters may be analyzed as semantically opaque compounds by native speakers (e.g., \emph{ja}[k\#d]\emph{aw}).
+%\footnote{An analogous account is the proposal that readjustment rules insert morphological boundaries for the purposes of assigning stress (\emph{SPE}:94).}
+
+\subsubsection{Coda nasal place assimilation}
+\label{cnpasection}
+
+Above, it was assumed that the velar nasal is derived from underlying /n/. 
+This analysis is not purely allophonic, as nasal place assimilation alternations target the final consonant of the \emph{im-}/\emph{in-} prefix.
+In (\lastx ab), the prefix-final consonant takes on the same major place of articulation as the following obstruent, and in (\lastx c), where the root is vowel-initial, the prefix-final consonant is realized as a default /n/.
+
+\begin{example}
+Nasal place assimilation in Latinate prefixes:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l}
+%a. & possible & i[m.p]ossible & & balance & i[m.b]alance \\
+%b. & tangible & i[n.t]angible & & decent  & i[n.d]ecent  \\
+%c. & elegant  & i[n.]elegant  & & ability & i[n.]ability \\
+a. & possible & i[m.p]ossible \\
+   & balance  & i[m.b]alance  \\
+b. & tangible & i[n.t]angible \\
+   & decent   & i[n.d]ecent   \\
+c. & elegant  & i[n.]elegant  \\
+   & ability  & i[n.]ability  \\
+\end{tabular} 
+\end{example}
+
+\noindent
+The shape of the prefix before other obstruents is somewhat more complex. 
+%Before /f/, CELEX transcribes the nasal as either labial or alveolar, the latter being very rare in simplex words. 
+%There are also additional complexities regarding the shape of the prefix when attached to /k, g/-initial roots. 
+\citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but this assimilation is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
+The prefix \emph{com-}/\emph{con-}, which patterns with \emph{im-}/\emph{in-} in showing assimilation (e.g., \emph{co}[m.b]\emph{ine}), participates in a similar contrast, between \emph{có}[ŋ.g]\emph{ress} and \emph{co}[n.g]\emph{réssional}.
+As \citeauthor{Borowsky1986} notes, stress does not block assimilation in simplex words (e.g., \emph{a}[ŋ.g]\emph{óra}), a fact which demonstrates the necessity of a partially morphological explanation for the blocking of assimliation in words like \emph{co}[n.g]\emph{réssional}.
+Regardless of these complexities, there is sufficient evidence to posit assimilation of the \textsc{Place} features of an obstruent to a preceding nasal consonant.
+
+\begin{example}
+\textsc{Coda Nasal Place Assimilation} (\emph{SPE}:85):
+
+\xymatrix@R=24pt@C=24pt{
+                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{..}[dl] \\
+\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d] \\
+\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]} \\
+}
+\end{example}
+
+\citet[][175]{Pierrehumbert1994} observes that in simplex English words, ``nasal-stop sequences agree in labiality''; presumably, this generalization is limited to [\textsc{Labial}] because \citeauthor{Pierrehumbert1994} also adopts the assumption that the velar nasal is derived from /n/.
+However, \citeauthor{Pierrehumbert1994} does not propose any connection between the alternation and the generalization that nasal consonants in URs tend to agree with the following consonant.
+While clusters like \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey} are the norm, exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} are found.
+
+In all, there are only 80 possible nasal-obstruent syllable contact clusters, the majority of which are attested.
+In addition, there are a small number of words which contain clusters like \emph{oi}[nt.m]\emph{ent}, in which a nasal-obstruent cluster are both parsed into a complex medial coda, but in this position, place agreement is exceptionless, and thus these clusters are ignored below.
+
+\begin{example}
+Lexical effects of \textsc{Coda Nasal Place Assimilation}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 41 & 11 & 0.788 & \multirow{2}{*}{2.6\e{-05}}\\
+violating  & 8  & 20 & 0.286 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent
+Clusters which conform to this generalization are far more likely to be attested than those that violate it, and this is highly unlikely to be due to chance.
+
+There is some evidence that even young infants show a preference for nonce words that conform to this generalization. %\citet{Jusczyk2003}
+\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} play nonce words containing nasal-obstruent clusters to 4.5, 9, and 10-month-old infants (respectively) acquiring English, All three studies find that infants prefer to listen to stimuli which conform to the above generalization (e.g., \emph{u}[m.b]\emph{o}) over those which do not (e.g., \emph{u}[n.b]\emph{o}).
+\citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''.
+In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.
+
+\begin{example}
+Artificial language adaptations \citep[][394, his transcriptions]{Wright1975}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l l}
+a. & [gownp] & > & [gump] \\
+b. & [ǰumg]  & > & [ǰúŋgə] \\
+c. & [ðʌŋd]  & > & [tɔŋg] \\
+\end{tabular}
+\end{example}
+
+\noindent
+If this game is akin to the process of loanword adaptation, it is possible that various extragrammatical pressures are at play \citep[e.g.,][]{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008}.
+Yet, the independent evidence for a phonological process of nasal place assimilation provides the most parsiminious account of the adaptations seen in \citeauthor{Wright1975}'s study.
+J.
+\citet{Myers1993} also shows that \textsc{Coda Nasal Place Assimilation} is fed by speech errors:
+
+\begin{example} 
+Speech errors and \textsc{Coda Nasal Place Assimilation} \citep[][228]{Myers1993}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l}
+   & \emph{error}  & \emph{target} \\
+a. & primps        & prints        \\
+b. & rand orker    & rank order    \\
+%  & [sprɪg]time for [hɪ̃ntlɘr] & springtime for Hitler \\
+c. & po[ŋ]cutation & computation   \\
+\end{tabular} 
+\end{example}
+
+%\citet{Hay2004a} demonstrates that [n.p, mθ] clusters, which are not found in English, are in fact rated better than many attested clusters.
+%report that 10-month-old infants prefer to listen to nonsense words which conform to \textsc{Coda Nasal Place Assimilation} (e.g., \emph{u}[m.b]\emph{o}) compared to those which do not (e.g., \emph{u}[n.b]\emph{o}).
+
+\subsubsection{Degemination}
+
+A final well-known alternation in English is the simplification of geminates derived by certain suffixes.
+For instance, the suffix \emph{-ly} rarely forms a geminate when attaching to root-final /l/, whether forming denominal adjectives (e.g., \emph{sme}[l]\emph{y}) or deadjectival adverbs (e.g., \emph{norma}[l]\emph{y}).
+A small class of English verbs take a /-t/ suffix in the past; as this suffix does not trigger the \textsc{Epenthesis} found in the regular past (e.g., \citealt{Bakovic2005b}, \citealt{Fruehwald2011}), \textsc{Degemination} applies.
+
+\begin{example}
+Degemination in the /-t/ past (\citealp[][105]{Halle1985a}, S. \citealp[][492]{Myers1987}):
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l}
+a. & /bɛnd-t/ & \goesto & [bɛnt] & (cf. \emph{burn}[t]) \\
+b. & /bɪld-t/ & \goesto & [bɪlt] & (cf. \emph{spil}[t]) \\
+\end{tabular}
+\end{example}
+
+More complicated cases of degemination are found in English prefix morphology and described in \emph{SPE} (p.~148, 219f.) and \citealt[][]{Borowsky1986} (p. 102f.), and are not reviewed here.
+
+%\ex \textsc{Coda Nasal Place Assimilation} feeds \textsc{Degemination} \citep[][116]{Borowsky1986}: \\
+%\begin{tabular}{l l l l}
+%a. & mature    & i[m]ature    & (cf. \emph{i}[m.b]\emph{alance})   \\
+%b. & numerable & i[n]umerable & (cf. \emph{i}[n.d]\emph{ependent}) \\
+%\end{tabular}
+%\xe
+
+The rule deletes the first of two adjacent consonants agreeing in \textsc{Place} and manner features, and is formalized using quantification over features \citep{Reiss2003b}.
+% cf. \citep{Berent2012}
+
+\begin{example}
+\textsc{Degemination} (\emph{SPE}:253):
+
+\xymatrix@R=24pt@C=24pt{
+\textsc{C}_1       & \longrightarrow & \emptyset & / & A~\gap\gap~A & \textsc{C}_2 \\
+}
+Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
+\end{example}
+
+\noindent
+I assume that geminates crossing a syllable boundary are ``false geminates'', meaning that they are adjacent similar segments rather than a single feature matrix linked to two timing slots.
+Under the alternative account in which the absence of geminates is an inventory restriction (i.e., a \textsc{Segment Structure Condition}), an undesirable duplication of phonological derived surface constraints and constraints on URs would be unavoidable.
+
+Neither identical consonants, nor those which differ only in voice (e.g., *[p.b]) occurs in simplex words in CELEX, and this absence is highly unlikely to be due to chance.
+
+\begin{example} Lexical effects of \textsc{Degemination}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 157 & 596 & 0.208 & \multirow{2}{*}{7.2\e{-09}}\\
+violating  &   0 &  87 & 0.000 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+%\footnote{Note that it is not the case that the English lexicon as envisioned by \emph{SPE} is free of underlying geminates; \citeauthor{SPE} use underlying geminates to attract stress to a preceding nucleus (p.~148--151).
+In \emph{SPE}, the underlying geminate-singleton contrast is absolutely neutralized, but has consequences for surface stress.
+In addition to the usual objections to absolute neutralization (i.e., that it is a learning conundrum mascarading as a phonological solution) that emerged in the wake of the publication of \emph{SPE}, the phonological particulars of this account have also come under attack \citep[e.g.,][]{Ross1972a}.
+It would not seem tenable to use underlying geminates to attract stress even if absolutely neutralized contrasts were admitted.} Unsurprisingly, this is unlikely to be due to chance.
+
+\subsubsection{Dorsal/non-coronal clusters}
+
+\citet{Pierrehumbert1994} surveys the English syllable contact cluster inventory, and discerns the effect of several ``static'' constraints, those which do not derive from neutralizing phonological rules.
+
+\citet[][173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Since \citeauthor{Pierrehumbert1994}'s study is restricted to triconsonantal clusters, so words like \emph{a}[k.m]\emph{e}, \emph{ru}[g.b]\emph{y}, or \emph{pi}[g.m]\emph{ent} do not violate this narrower generalization.
+Further, \citeauthor{Pierrehumbert1994} adopts the traditional assumption \citep[e.g.,][66f.]{Borowsky1986} that coda [ŋ] is an allophone of /n/ before dorsal consonants, so words such as \emph{mo}[ŋ.gr]\emph{el} are not exceptions.
+
+\citeauthor{Pierrehumbert1994} generalization is termed static simply because there is no evidence for any phonological process which changes clusters of dorsal obstruents followed by non-coronals into something else (perhaps a dorsal/coronal cluster).
+However, the statistical test, taking into account clusters of any length, does not indicate that this dispreference for dorsal/non-coronal clusters is so compelling that it could not be generated by chance, simply because possible dorsal/non-coronal clusters are about as likely to occur as are similar types of clusters such as those with a non-dorsal coda and a non-coronal onset (e.g., \emph{a}[t.m]\emph{osphere}) or a dorsal coda and coronal onset (e.g., \emph{ve}[k.t]\emph{or}).
+These clusters are marked as ``conforming'' to a constraint \textsc{*[+Dorsal][$-$Coronal]}, and dorsal/non-coronal consonants as ``violating'' it.
+
+\begin{example}
+Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 76 & 359 & 0.175 & \multirow{2}{*}{0.145} \\
+violating  &  6 &  57 & 0.095 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+There is no exceptions to \citeauthor{Pierrehumbert1994}'s narrower generalization concerning triconsonantal clusters, but it is no more statistically convincing.
+
+\begin{example}
+Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]} on triconsonantal clusters:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 20 & 148 & 0.119 & \multirow{2}{*}{0.136} \\
+violating  &  0 &  22 & 0.000 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent
+The evidence for \citeauthor{Pierrehumbert1994}'s static constraint does not actually require that such a constraint exists at all, since it is quite possible that the apparent dispreference for dorsal/non-coronal clusters could result by accident.
+
+\subsubsection{Coda coronal obstruents}
+
+Another static generalization about triconsonantal syllable contact clusters proposed by \citet[][175]{Pierrehumbert1994} is that  ``clusters with a coronal obstruent in the coda do not occur.'' In an endnote (p.
+186), \citet{Pierrehumbert1994} notes an exception in \emph{a}[nt.l]\emph{er}, and one can add the triconsonantal clusters in \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many biconsonantal clusters like \emph{a}[t.l\emph{as} or \emph{e}[s.t]\emph{er}.
+As before, applying the statistical test to clusters of any length reveals that coda coronal obstruent clusters are not significantly less likely to occur than clusters with non-coronal codas.
+
+\begin{example}
+Lexical effects of \textsc{*Coda Coronal Obstruent}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming &  48 & 312 & 0.133 & \multirow{2}{*}{0.301} \\
+violating  &  38 & 322 & 0.106 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+Restricting \citeauthor{Pierrehumbert1994}'s generalization to triconsonantal clusters does not strengthen the case for this static constraint.
+
+\begin{example}
+Lexical effects of \textsc{*Coda Coronal Obstruent} on triconsonantal clusters:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming &  4 & 104 & 0.037 & \multirow{2}{*}{0.419} \\
+violating  &  2 & 124 & 0.016 \\
+\bottomrule
+\end{tabular} 
+\end{example}
+
+%Note that [tʃ] has not been included since it so rarely occurs in clusters.
+
+\subsubsection{ABA clusters}
+
+Finally, \citet[][176]{Pierrehumbert1994} notes ``a lack of clusters with identical first and third elements''.
+For the determination of identity, I assume the same definition of geminate as used in the rule of \textsc{Degemination} above, namely consonants which share all the same \textsc{Place} and manner features, ignoring \textsc{Voice}.
+%and I also expand the definition of the constraint to also apply to the second and fourth consonant of quadraconsonantal clusters. 
+\citeauthor{Pierrehumbert1994} notes one potential exception to this generalization, the cluster in words such as \emph{e}[ks.kl]\emph{ude}, which contains two /k/, does not arise here since such words have been marked morphologically complex.
+Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly rarer than other triconsonaontal and quadraconsonantal clusters, of which only 7.8\% occur.
+
+\begin{example}
+Lexical effects of \textsc{*ABA}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 41 & 478 & 0.079 & \multirow{2}{*}{0.818} \\
+violating  &  0 &  11 & 0.000 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}.
+
+%\subsubsection{Summary}
+%
+%The above results are summarized in Table \ref{constraints}.
+%
+%\begin{table}
+%\centering
+%\begin{tabular}{l r r r r r}
+%\toprule
+%                                       & \multicolumn{2}{c}{attested} & \multicolumn{2}{c}{unattested} & \multirow{2}{*}{$p$-value} \\
+%                                       & conforming & violating & conforming & violating \\
+%\midrule
+%\textsc{O.V.A.} (\ref{ovar})        &  80 &   6 & 370 & 264 & 1.1\e{-11} \\
+%\textsc{C.N.P.A} (\ref{CNPA})       &  41 &  25 &  12 &  42 & 1.7\e{-05} \\ 
+%\textsc{Degemination} (\ref{degem}) & 161 & 632 &   0 &  87 & 1.4\e{-08} \\
+%\midrule
+%\textsc{*Dorsal-Labial} ()          &  68 &   4 & 246 &  40 & 0.069 \\
+%\textsc{*C.C.O.} ()                 &  48 &  38 & 312 & 322 & 0.301 \\
+%\textsc{*A,BA} ()                   & 161 &   0 & 708 &  11 & 0.231 \\
+%\bottomrule
+%\end{tabular}
+%\caption{Counts and Fisher exact test $p$-values for the constraints discussed above.}
+%\label{constraints}
+%\end{table}

### 2012-04-18 4.2.2.tex
chapters 2 and 4 have been swapped

@@
-% 4.2.2: Assessing models
+\subsection{Computational models}
 
-\subsubsection{Basic correlations}
-\subsubsection{Intramodal correlations}
-\subsubsection{Residual correlations}
+Applying these derived constraints to the lexicon increases saturation from 18.8\% to 28.8\% and incurs only a handful of exceptions.
+Yet, despite the fact that ``attestation'' is the minority pattern, attempts to identify static lexical constraints, whether by hand or computational model, lend little additional predictive power.
+There is no evidence that the English syllable contact inventory is subject to any static constraint at all.
+I am forced to conclude that most, if not all, of 71.2\% of possible clusters which are unattested are accidental gaps.
+Below, I show that this state of affairs follows directly from the sparse distribution of codas and onsets.
+
+The three alternations targeting English syllable contact clusters have been shown to have the expected robust effects on the English lexicon.
+
+Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}.
+There is now a growing literature on computational cognitive models of phonotactic learning.
+By applying these models to the syllable contact data, it is possible to exhaustively search for statistically motivated constraints, elimianting the possibility of human error.
+
+To evaluate these statistical models, they are first trained on the set of attested clusters, then put to the task of classifying all possible clusters into those which are and are not attested.
+Four metrics are used to evaluate these computational models.
+Classification accuracy corresponds to the probability of a classification being either a true positive ($t.p.$), correctly predicting a cluster to occur, or a true negative ($t.n.$), correctly predicting a cluster to be absent.
+Incorrect outcomes are either false positives ($f.p.$), incorrectly predicting a cluster to occur, and false negatives ($f.n.$), incorrectly predicting a cluster to be absent.
+
+%\ex Binary classification contingency table: \\
+%\begin{tabular}{c | l l} \toprule
+%                             & \multicolumn{2}{c}{prediction} \\
+%\midrule
+%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
+%                             & false negative ($fn$) & true negative ($tn$) \\
+%\end{tabular}
+%\begin{tabular}{|l l|}
+%\toprule
+%true positive ($tp$)  & false positive ($fp$) \\
+%false negative ($fn$) & true negative ($tn$) \\
+%\bottomrule
+%\end{tabular}
+%\xe
+
+\begin{example}
+$\displaystyle \textrm{Accuracy} = \frac{t.p. + t.n.}{t.p. + t.n. + f.p. + f.n.}$
+\end{example}
+
+\noindent
+A few other metrics can be used to assess classification performance. Precision corresponds to the probability that a cluster predicted to occur is observed in the data. 
+\begin{example}
+$\displaystyle \textrm{Precision} = \frac{t.p.}{t.p. + f.p.}$ 
+\end{example}
+
+\noindent
+Recall (also known as sensitivity) is the probability a observed cluster is predicted to occur.
+
+\begin{example}
+$\displaystyle \textrm{Recall} = \frac{t.p.}{t.p. + f.n.}$
+\end{example}
+
+It is possible to maximize recall at the expense of precision, by predicting more clusters to be attested, or to increase precision at the expense of recall by predicting fewer clusters to be attested.
+A common way to balance these concerns is the $F_1$ score, which is the harmonic mean of precision and recall.
+
+\begin{example}
+$\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$ 
+\end{example}
+
+\subsubsection{Null baseline}
+
+The 840 possible clusters have an 18.7\% of saturdaion rate.
+As a consequence of this sparsity, a null model which classifies all clusters as unattested achieves 81.3\% accuracy without the need to posit any constraints either derived or static.
+Since this model does not have any positive classifications, either true or false, precision, recall, and $F_1$ are undefined.
+
+\subsubsection{Alternations}
+
+As was shown above, \textsc{Obstruent Voice Assimiation}, \textsc{Coda Nasal Place Assimilation}, and \textsc{Degemination} are reliable predictors of what clusters are and are not attested and have few lexical exceptions.
+I generate another baseline by predicting only those clusters which do not violate any of these generalizations to be attested.
+This provides a significant improvement to the null baseline (accuracy = 0.857).
+As precision (0.401) is smaller than recall (0.708), there are more false positives (i.e., unattested clusters that do not violate any of these three alternation-derived generalizations) than there are false negatives (i.e., attested exceptions to these derived generalizations).
+
+\subsubsection{\citet{Pierrehumbert1994}}
+
+\citet{Pierrehumbert1994} first proposes to measure the well-formedness of syllable contact clusters by the independent probabilities of codas and onsets, and \citet{Coleman1997} extend this model to whole words.
+\citet{Pierrehumbert1994} uses both medial and initial/final codas and onsets to compute these probabilities, but I obtain a better fit to the data by only using medial coda and onset frequencies.
+Medial coda and onset probabilities are multipled to obtain a well-formedness score.
+This produces a continuous value between 0 and 1.
+To derive a binary classification from these values, I use a soft-margin support vector machine \citep{Cortes1995} with a linear kernel.
+This technique derives a ``decision stump'' providing the best single split into attested and an unattested clusters.
+This is not put forth as a cognitively plausible model for relating this expected probability to attestation: it is simply the upper bound that such a model could reach.
+
+Unfortunately, this score only provides a small improvement to the null baseline
+ (accuracy = 0.839), and has a lower accuracy and $F_1$ (0.328) than the alternation baseline.
+
+\subsubsection{\citet{Hayes2008a}}
+
+\citeauthor{Hayes2008a} develop a sophisticated method of learning phonotactic distributions from positive data.
+They develop software for estimating probability distributions over sequences of (sets of) phonological features using the principle of maximum entropy and making few assumptions about the independence of features.
+As above, a support vector machine with a linear kernel is used to turn these probability values into binary predictions.
+The software provided by these authors exposes many ``metaparameters'', operational decisions to be made in conducting an experiment with this system.
+To avoid any potential bias induced by the choice of these metaparameters, I use the same settings as used by \citet{Hayes2008a} in their similar study of English onset consonants.
+This includes the ``accuracy schedule'' [0.001, 0.01, 0.1, 0.2, 0.3] and no constraints on the number of constraints learned, as well English consonantal features used by \citeauthor{Hayes2008a}.
+Since this model is inherently stochastic, producing slightly different outcomes on each run, I follow the practice of \citeauthor{Hayes2008a} (p.~396) and report the lowest-accuracy of ten runs, though in general there is not a great deal of variation between individual runs.
+
+This model provides a small improvement in accuracy (0.877) and $F_1$ (0.703) compared to the alternation baseline, but it has lower recall (0.642) than this baseline, indicating that the \citeauthor{Hayes2008a} model introduces many false negatives.
+%; it assigns the lowest probability value to 122 unattested clusters.
+More specifically, it suggests that this model prefers narrower phonotactic generalizations than those derived from alternations.
+For instance, unattested *[m.kl] is assigned the highest probability, whereas it is ruled out by \textsc{Coda Nasal Place Assimilation}.
+On the other hand, the \citeauthor{Hayes2008a} model assigns a low score to clusters like those found in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.
+
+\citet{Berent2012} identify an important defect with the \citeauthor{Hayes2008a} model used here: it does not make use of variables over features and thus fails to unify the constraint against geminate clusters.
+While the model may miss this unifying generalization, it correctly predicts all clusters containing geminates to be unattested.
+
+The results for the four models are summarized below.
+
+\begin{example}
+Summary of computational model accuracy:
+
+\vspace{0.5\baselineskip} 
+\begin{tabular}{l | r r r r}
+\toprule
+                          & accuracy & precision & recall & $F_1$ \\ 
+\midrule
+Null baseline             & 0.813    & n.a.      & n.a.   & n.a.  \\
+Alternations              & 0.857    & 0.401     & 0.708  & 0.512 \\
+\citet{Pierrehumbert1994} & 0.839    & 0.210     & 0.750  & 0.328 \\
+\citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+One additional model was evaluated for this study.
+\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
+However, \citeauthor{McGowan2011} finds that sonority of clusters accounts for only a tiny amount of the variance in cluster frequency ($R^2 = 0.077$).
+In this study, inclusion of sonority distance provided no improvement over the null baseline, and therefore was not considered further.

### 2012-04-18 4.2.3.tex
chapters 2 and 4 have been swapped

--- 2012-02-16 2.2.3.tex
@@
-% 2.2.3: Statistical properties of syllable contact
+\subsection{Statistical properties of syllable contact}
+
+Constraints on
 
 Above, we have seen that constraints on URs which are derive from phonological alternations restrict the contents of the lexicon.
+
 Applying these derived constraints to the lexicon increases saturation from 18.8\% to 28.8\% and incurs only a handful of exceptions.
@@
 
-\ex $\displaystyle Z_i = \frac{2 n_i}{r_{i + 1} - r_{i - 1}}$ \xe
+\begin{example}
+$\displaystyle Z_i = \frac{2 n_i}{r_{i + 1} - r_{i - 1}}$
+\end{example}
 
@@
 
-\ex $\displaystyle Z_1 = \frac{n_1}{r_2 - r_1}$ \xe
-\ex $\displaystyle Z_N = \frac{n_N}{r_N - r_{N - 1}}$ \xe
+\begin{example}
+$\displaystyle Z_1 = \frac{n_1}{r_2 - r_1}$
+\end{example}
+
+\begin{example}
+$\displaystyle Z_N = \frac{n_N}{r_N - r_{N - 1}}$
+\end{example}
 
@@
 
-\ex $\displaystyle f(r~; C, \alpha) = \frac{C}{r^\alpha}$ \xe
+\begin{example}
+$\displaystyle f(r; C, \alpha) = \frac{C}{r^\alpha}$ 
+\end{example}
 
@@
 
-\ex $\displaystyle \textrm{log}~Z_r \sim C + \alpha~\textrm{log}~r + \epsilon$ \xe
+\begin{example}
+$\displaystyle \textrm{log}~Z_r \sim C + \alpha~\textrm{log}~r + \epsilon$ 
+\end{example}
 
@@
 That clusters (and the codas and onsets that make them up conforms to Zipf's law is not itself a deep observation.
-Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also $n$-grams \citep{Baroni2009} and of randomly-generated texts \citep{Miller1957,Li1992}.
+Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also word \citep{Teahan1998,Ha2002,Baroni2009} and phoneme \citep{Daland2011a} n-grams, and tokens in non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} or randomly-generated texts \citep{Miller1957,Li1992}.
 The import of this statistical property for the study of cluster phonotactics is that it makes it difficult, on statistical grounds alone, to determine which unobserved events are excluded, and which are accidentally missing.
@@
 
-\ex $\displaystyle \hat{p}_0 = \frac{n_1}{\sum N}$ \xe
+\begin{example}
+$\displaystyle \hat{p}_0 = \frac{n_1}{\sum N}$ 
+\end{example}
 
@@
 
-\ex {Simulation procedure}: \\
+\begin{example}
+Simulation procedure: \\
     \begin{tabular}{l l}
@@
     \end{tabular}
-\xe
+\end{example}
 

### 2012-04-18 4.3.tex
chapters 2 and 4 have been swapped

@@
-% 4.3: Conclusion
+\section{Conclusions}
 
-It is not the case that this evidence is inconsistent with the hypothesis that there are additional degrees of wordlikeness: it is just a demonstration that these data to not require it.
+\citet{Borowsky1989} on peripherality.
+
+productive \citet{Duanmu2008}
+
+The above is a stark reminder.
+
+A reasonable objection to the arguments presented in this chapter is to view the results as little more than an indictment of the results of \citealt{Pierrehumbert1994}.
+
+However, the fact that state-of-the-art models are not capable of providing large improvements to the predictive accuracy indicates
+
+ do not have a statistically significant effect on the shape of the English lexicon, but that experiments might turn up evidence that speakers have internalized
+
+shown to be aware of static constraints if they reached statistical significance.
+
+ statistically reliable static constraints could be identified
+
+The of
+
+Regarding the historical developments,
+
+\citet{Martin2007}
+
+On the other hand, patterns created by sound change are not guaranteed to persist over time. 
+One example of non-persistence is discussed by \citet{Iverson2005}.  
+Around 1100 CE, Old English \emph{sk} became [ʃ]. 
+This sound change introduced no alternations.
+Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{V\lm sk\#} words when the
+ change was actuated, and \emph{V\lm sh\#} continues to be rare in Modern English. 
+What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.
+
+A third pattern is that a historically inherited pattern
+
+\citet[][140]{Frisch2004} suggest that the strong tendency for the first and second consonants of the Arabic root to be non-identic
+al is the ``a diachronic result of a processing constraint that disfavors repetition.'' 
+Unfortunately, there is no evidence that this pattern is diachronic other than in the sense that it appears to be inherited from the proto-language: there is simply no Proto-Semitic verb roots with identical first and second consonants \citep[][178]{Greenberg1950}.
+In other Semitic languages, the inherited patern has experienced considerable erosion.
+
+\begin{example}
+Tigrinya roots with identical first and second consonants \citep{Buckley1990a}: \\
+\begin{tabular}{l l l l}
+a. & lʌlʌw     & `scorch'                   & (< Ge'ez \emph{lʌwlʌw} `inflame')     \\
+   & mʌmʌy     & `winnow'                   & (< Ge'ez \emph{mʌymʌy} `distinguish') \\
+%   & mʌmʌt & `pick out loot' & (< 
+b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
+   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
+c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
+   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
+\end{tabular}
+\end{example}
+
+Similar exceptions are found in 
+%Amharic (\citealp[][?]{Broselow1984}, \citealp[][?]{McCarthy1985}) and 
+Hebrew \citep[][29]{Bat-El2005}.
+
+The next two chapters return to the question of synchrony, addressing the relationship between statistical patterns in the lexicon and speakers' behaviors when presented with underrepresented sequences.
+
+accidental gaps
+\citet[][419f.]{Hayes2008a}

### 2012-04-18 4.tex
chapters 2 and 4 have been swapped

--- 2012-02-17 4.tex
@@
-% 4: Gradient and categorical aspects of wordlikeness judgements
+\chapter{Accidental and structural gaps in English syllable contact}
+\label{clusters}
 
-(though few if any phonologists regard $n$-gram segment probabilities as a core component of phonology)
+The term \emph{lexicon} has many senses: throughout, I use it to refer simply to the set of underlying representations.
+%\citep[][269]{LANGUAGE}.
+
+In the preceding chapter, it was proposed that
+
+Reconsider the null hypothesis presented in the preceding chapter:
+
+\begin{example}
+The null hypothesis: \\
+Universal grammar does not countenance constraints on the contents of URs
+\end{example}
+
+The idea
+
+relationship to proposal
+
+One primary way which this null hypothesis has been critiqued is by demonstrating that the lexicon of some language 
+exhibits dispreferences or gaps corresponding to.
+
+argue by trends in the lexicon
+
+possibility of sparsity (C\&H, Fischer/Vogt and recent stats)
+
+Of course, to obtain the desired results, we must guarantee that each sequence structure rule reflects a general systematic fact about the language, and not a fact which is due merely to the existence of accidental gaps in the lexicon.
+\citep[][401, fn.~8]{Stanley1967}
+
+In the preceding chapter, it was proposed that
+
+Lexicon
+
+This appears to be exceptionlessly true of Arabic
+
+That even gradient patterns
+
+One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. 
+Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
+This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a} but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Dmitrieva2008a,Dmitrieva2008b,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.
+
+The absence of 
+The claim here is that however stark the absence or underrepresentation of some form is, it
+
+This hypothesis admits the possibility that there may be accidental gaps in the lexicon, a possibility which predates generative thinking:
 
 \begin{quotation}
-``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
+\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
+\citep[][16]{Fischer-Jorgensen1952}
 \end{quotation}
+
+\noindent
+Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters
+
+\begin{quotation}
+Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters.
+The material will never be complete.
+It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary.
+\citep[][30]{Vogt1954}
+\end{quotation}
+
+\citet{Chomsky1965}
+
+This chaper demonstrates that apparently systematic gaps may arise even in the absence of phonotactic preferences, and that this possibility compromises attempts to show that such preferences shape the lexicon.
+
+\input{4.1}
+\input{4.2}
+\input{4.3}

### 2012-04-18 7.1.2.tex
More arranging.
chapters 2 and 4 have been swapped

@@
 
-\footnote{Thanks to Toni Cook for bringing this example to my attention.}
+\footnote{Thanks to Toni Cook for bringing this study to my attention.}
 
@@
 e- & -rí- & -bí & tum & -ir & -an & -a \\
-\{\textsc{Aug} & \textsc{Pre} & \texsc{OM}\} & \{send & \textsc{Appl} & \textsc{FV}\} \\
+\{\textsc{Aug} & \textsc{Pre} & \textsc{OM}\} & $\surd$\textsc{send} & \textsc{Appl} & \textsc{FV}\} \\
 \multicolumn{7}{l}{``to send them to each other''}
+\end{tabular}
 \xe

### 2012-04-18 pre.tex
Went all Unicode, all the time.
Restructured into smaller files, wrote much of chapter 2.
Drastic restructuring and some writing on chapters 2-3.
chapters 2 and 4 have been swapped

@@
-\section{``What is the most important fact?''}
-
 Perhaps the most important fact for the study of speakers' internal knowledge of language is Wilhelm von Humboldt's observation that speakers make ``infinite use of finite means'' CITE \citep{Chomsky1966}.
@@
 
-This
+In this dissertation, I critique two such epicyclic proposals, and present solutions using independently-motivated grammatical machinery to account for the data.
+
+The first half of this dissertation began with a suggestion from Charles Yang that the sparse nature of the lexicon might itself account for at least some phonotactic gaps.
+The success of the resulting study, in chapter four, led me to question the received wisdom that speakers' knowledge of phonotactic generalizations is gradient and potentially independent of the phonological system.
+
+These results support a null hypothesis in which phonotactic knowledge is derivative of the language acquisition device and phonological processes.
+While I present this theory as a falsifiable scientific hypothesis, an alternative would treat this proposal as a \emph{negative heuristic}, a
+
+guides us away from ``phonotactic grammar'' as an explanatory device.
+
+even those who are not inclined to accept this proposal should note how much of the coverage traditionally attributed to
+
+can be accounted for by the nu
+
+The second half of the dissertation develops a theory of inflectional gaps.
+I was pleased to hear (from Greville Corbett) that Stephen Anderson regarded inflectional gaps as ``the biggest scandal in morphology'', and even more so when Prof.
+Anderson agreed to serve as the external member of my committee.
+
+In developing this theory, I have focused primarily on how inflectional gaps could be learned, and thus my approach is does not directly bear on competing morphological theories.
+At the same time, this theory is consistent with a proposal by \citet{Pesetsky1977} that inflectional gaps provide novel evidence for the ordering of various components of morphology.
+As an example, consider a few Russian verbs exhibiting inflectional gaps in the 1sg non-past.
+
+\ex Russian 1sg. imperfect gaps \citep[after][]{Pesetsky1977}: \\
+\begin{tabular}{l l l@{}l@{}l@{}l@{}l@{} l}
+   & infinitive &   & 1sg. non-past \\
+a. & pobedit{j} & * & pobeʒu  & / & * & pobeʒdu  & `conquer'   \\
+   & ubedit{j}  & * & ubeʒu   & / & * & ubeʒdu   & `persuade'  \\
+b. & tʃudit{j}  & * & tʃuʒu   & / & * & tʃuʒdu & `act oddly'   \\
+   & otʃudit{j} & * & natʃuʒu & / & * & natʃuʒdu & `alienate'  \\
+\end{tabular}
+
+\noindent
+As \citealt{Pesetsky1977} notes, the ill-formedness of the 1sg. non-past forms is shared by verbs related by prefixation. The opposite pattern can be seen in English.
+
+\ex English preterite gaps: \\
+\begin{tabular}{l l l@{}l@{}l@{}l@{}l@{}}
+   & present  &   & preterite &   &   &           \\
+a. & foresee  &   & foresaw   &   &   &           \\
+   & sightsee & * & sightsaw  & / & * & sightseed \\
+b. & undergo  &   & underwent &   &   &           \\
+   & forgo    & * & forwent   & / & * & forgoed   \\
+\end{tabular}
+
+Unlike the pattern observed in Russian, verbs that appear to share the same root do not show the same patterns of defectivity.
+If we simply label the process of verbal inflection as the source of inflectional gaps, we can understand this contrast as follows: in Russian, inflection precedes and feeds prefixation; in English, the formation of complex verb stems and feeds the inflectional system.
+
+While at this time I do not have any insights derived from these facts, there exists a potential to use inflectional gaps for evidence of ordered operations in morphology.
+These inferences are available to any account that recognizes inflectional gaps as something about the inflectional system, including the one presented here, but not if inflectional gaps are due to phonological properties of the derived objects.
+
+The two parts of this dissertation were conceived and executed separately, but some uniting principles can be identified.
+Both parts argue against inviolable, language-specific constraints in favor of a derivational explanation for the facts at hand.
+Both parts derive the phenomena at hand using novel acquisitional strategies (\emph{occultation} and \emph{tolerance}, respectively) and independently motivated components of the grammatical architecture (alternation phonology and the elsewhere principle).
+And the accounts presented in both parts are inspired by concerns for the sparsity of the primary linguistic data.

### 2012-04-26 3.1.1.tex
Albright MS data and some writing on chapter 3

@@
 One of the most salient properties of Turkish is its harmony system, which \citeauthor{Lees1966b} (\citeyear[][35]{Lees1966b}, \citeyear[][284]{Lees1966a}) derives by means of feature spreading rules.
-The least constrained of these processes is the rule of backness harmony described in this section.
+Backness harmony is the the most general of these processes.
 
@@
 
-Backness harmony spreads the value of \textsc{Back} rightward over any intervening consonants onto the next vowel.
+The rule of \textsc{Backness Harmony} spreads a vowel's \textsc{Back} specification rightward over any intervening consonants onto the next vowel.
 
@@
 \txt{V}                                        & \txt{C$_0$} & \txt{V} \\
-\txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{..}[urr] \\
+\txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
 }
@@
 
-Along with the rule of \textsc{Roundness Harmony} discussed in the next section, this rule accounts for a great deal of the suffix allomorphy found in Turkish.
-The nominative plural (nom.pl.) is formed by adding the suffixes \emph{-ler} or \emph{-lar} to the nominative singular (nom.sg.).
+This rule accounts for a great deal of the suffix allomorphy found in Turkish.
+For instance, the nominative plural (nom.pl.) suffix is \emph{-ler} when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}], as predicted by Backness Harmony.
 
 \begin{example}
+\label{turknom}
 The Turkish nominative:
@@
 
-\noindent
-\textsc{Backness Harmony} correctly predicts the choice of suffix.
-There are a few complications, however.
-First, while most polysyllabic roots conform to \textsc{Backness Harmony} (\lastx b), not all do.
-In this case, suffix vowels generally harmonize with the final root vowel (\nextx a).
-Yet there is also a small class of nouns which exhibit [$-$\textsc{Back}] suffix vowels despite the fact that their final root vowel is [$+$\textsc{Back}] (\nextx b).
+There are however a few complications.
+First, not all polysyllabic roots conform to \textsc{Backness Harmony}.
+In this case, as can be seen from (\ref{turkexcept}a), suffix vowels generally harmonize with the final root vowel.
+There is also a small class of nouns, shown in (\ref{turkexcept}b), which exhibit [$-$\textsc{Back}] suffix vowels despite the fact that their final root vowel is [$+$\textsc{Back}].
 
 \begin{example}
+\label{turkexcept}
 Exceptional Turkish nominatives:
@@
 
-While it is uncontroversial that disharmonic suffixes (\lastx b) are no more than sporadic exceptions to \textsc{Backness Harmony}, the status of root disharmony (\lastx a) has been the subject of much debate.
-\citet[][212, 289]{Anderson1974} views the fact that disharmonic roots still trigger suffix harmony as evidence that suffix harmony is distinct from root harmony, the latter being a sequence structure constraint.
+It is uncontroversial that the disharmonic suffixes of (\ref{turkexcept}b) are no more than very sporadic exceptions to \textsc{Backness Harmony}, root disharmony has ben the subject of much debate.
+As disharmonic roots still trigger suffix harmony, \citet[][212, 289]{Anderson1974} proposes to distinguish suffix harmony (an alternation) from a sequence structure constraint governing root harmony.
+
+However, as \citet[][?]{Zonneveld1978} notes, the theory of exceptionality proposed in \emph{SPE} (p.~374f.) provides a direct account of suffix harmony in disharmonic roots.
+\citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule $R$ must be marked [$+$rule $R$] by convention.
+A root or affix which fails to undergo $R$ despite otherwise matching the structural description is simply said to be marked [$-$rule $R$].
+Under this account, exceptionality boils down to failing to match an extended structural description.
+If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the environment, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.
 
@@
 
-Underspecification provides an alternative under which the exceptionality of disharmonic roots is only apparent.
-While it is clear that the first vowel of Turkish roots are contrastively specified for backness (e.g., \emph{kül} `ash' vs.
+An alternative account, first proposed by \citet{Clements1982}, uses underspecificatio to derive the non-exceptionality of harmonic roots.
+The initial-syllable vowel of a Turkish root is contrastively specified for backness (e.g., \emph{kül} `ash' vs.
 \emph{kul} `servant', \emph{kepek} `bran' vs.
-\emph{kapak} `lid'), there are also some minimal pairs with regards to backness (dis)harmony (e.g., \emph{deve} `camel' vs.
-\emph{deva} `medicine', \emph{sene} `year' vs.
-\emph{sena} `praise').
-This supports the possibility, first suggested by \citet{Clements1982}, that harmonic roots may be underspecified and disharmonic roots fully specified, exemplified below.
+\emph{kapak} `lid'); harmonic roots are ones in which non-initial syllable vowels cannot contrast for this feature, and \citeauthor{Clements1982} proposes to leave them underspecified.
+Only in disharmonic roots does \textsc{Back} need to be specified for non-initial syllable vowels.
+%This is exemplified below in (\ref{spec}).
+%(e.g., \emph{deve} `camel' vs. \emph{deva} `medicine', \emph{sene} `year' vs. \emph{sena} `praise').
 
@@
 \begin{example}
-Underlying specification of \textsc{Back} for (dis)harmonic roots:
+\label{spec}
+\textsc{Back} specification for (dis)harmonic roots:
 
+\vspace{0.5\baselineskip}
 \xymatrix@R=24pt@C=24pt{
-\txt{a.} & \txt{s} & \txt{V} & \txt{n} & \txt{V}  & \txt{\emph{sene} `year'} \\
-&        & \txt{[$-$\textsc{Back}]}\ar@{-}[u]     \\
-\txt{b.} & \txt{s} & \txt{V} & \txt{n} & \txt{V} & \txt{\emph{sena} `praise'} \\
-         &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u] & & \txt{[$+$\textsc{Back}]}\ar@{-}[u]
+\txt{a.} & \txt{harmonic root:~~~~} & \txt{C} & \txt{V} & \txt{C} & \txt{V} \\
+&   &    & \txt{[$-$\textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
+\txt{b.} & \txt{disharmonic root:} & \txt{C} & \txt{V} & \txt{C} & \txt{V} \\
+    &    &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u] & & \txt{[$+$\textsc{Back}]}\ar@{-}[u]
 }
@@
 
-However, the theory of exceptionality presented in \emph{SPE} (p.~374f.) provides a direct account of suffix harmony in disharmonic roots without duplication.
-In \emph{SPE}, the specification of the target (i.e., the segment to be changed) of every rule $R$ must be [$+$rule $R$] by convention.
-A root or affix which fails to undergo $R$ despite otherwise matching the structural description is simply said to be marked [$-$rule $R$], and thus failing to meet the full structural description.
-If disharmonic roots are [$-$\textsc{Backness Harmony}], then suffix vowels are correctly predicted to undergo \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the environment, which is not required to be [$+$rule $R$].
+\citet{Clements1982} and \citet{Kaun1999}
+
+\footnote{Thanks to Kie Zuraw for bringing this data to my attention.}
+
+the epenthetic vowel is always [$+$\textsc{High}], and when the first consonant of the cluster is non-dorsal, it shows backness harmony
+
+\begin{example}
+Adaptation of initial foreign clusters \citep[][247]{Clements1982}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l l}
+a. & spiker  & \alt{} & sipiker  & `announcer' \\
+   & fren    & \alt{} & firen    & `break'     \\
+   & trablus & \alt{} & tırablus & `Tripoli'   \\
+   & brom    & \alt{} & burom    & `bromide'   \\
+   & prusya  & \alt{} & purusya  & `Prussia'   \\
+b. & kral    & \alt{} & kıral    & `king'      \\
+   & grup    & \alt{} & gurup    & `group'     \\
+c. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
+   & kredi   & \alt{} & kıredi   & `credit'    \\
+\end{tabular}
+\end{example}
 
@@
 Reduplication interacts with root harmony in both Tuvan and Turkish.
-When the base is harmonically [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] results in what \citeauthor{Harrison2001} call ``reharmonization'' (\nextx a).
+When the base is harmonically [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] results in what \citeauthor{Harrison2001} call ``reharmonization'' (\ref{redupgame}a).
 
@@
 Under the underspecification hypothesis, non-initial harmonic vowels lack an underlying \textsc{Back} feature, so it comes as no surprise that changing the \textsc{Back} specification of the root-initial vowel results in reharmonization.
-And the full specification of disharmonic roots correctly predicts that they will be exempt from reharmonization (\lastx b), which is also borne out in Tuvan and in an unrelated Finnish language game \citep{Campbell1986}.
+And the full specification of disharmonic roots correctly predicts that they will be exempt from reharmonization (\ref{redupgame}b), which is also borne out in Tuvan and in an unrelated Finnish language game \citep{Campbell1986}.
 

### 2012-04-26 3.1.tex
Albright MS data and some writing on chapter 3

@@
 
-\citet{Lees1966a,Lees1966b} proposes three sequence structure constraints on Turkish vowels, and these constraints have been discussed extensively in the subsequent literature.
-Below, these constraints are formalized and related to Turkish alternation phonology and various sources of external evidence.
-A traditional binary feature specification for the Turkish vowel system, given below, is assumed without discussion.
+\citet{Lees1966a,Lees1966b} proposes three constraints on Turkish vowel sequences 
+which have been the attention of many subsequent studies on vowel harmony and on sequence structure constraints.
+In this section, these constraints are formalized, and where possible, related to phonological alternations in Turkish and to various sources of external evidence.
+The following feature specification for the Turkish vowels is assumed throughout.
 
@@
                        & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
-\buf[$+$\textsc{High}] & i & ü [y] & ı [ɯ] & u \\
-\buf[$-$\textsc{High}] & e & ö [ø] & a [ɑ] & o \\
+\buf[$+$\textsc{High}] & {i} & {ü} [y] & {ı} [ɯ] & {u} \\
+\buf[$-$\textsc{High}] & {e} & {ö} [ø] & {a} [ɑ] & {o} \\
 \end{tabular}

### 2012-04-26 3.tex
Albright MS data and some writing on chapter 3

@@
-\chapter{Static and derived constraints in Turkish}
+\chapter{Static and derived constraints}
 \label{turkish}
@@
 
-The contrast between \emph{blick} and \emph{bnick} is what has been called a "projection problem" \citep{Baker1979}
+The contrast between \emph{blick} and \emph{bnick} is what has been called a ``projection problem'' \citep{Baker1979}
 
-This is not a universal, but rather something English speakers must learn, as some languages, such as Moroccan Arabic, do in fact permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet');
+This is not a universal, but rather something English speakers must learn: Moroccan Arabic, for instance, does permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet').
 
@@
 % Dupoux et al. 1999: Japanese speakers still can't distinguish between ebzo and ebuzo in an ABX task using the SAME a or b.
+%Elliot Moreton (p.c.)
 

### 2012-04-26 4.1.2.tex
Albright MS data and some writing on chapter 3

@@
 
-It is necessary to split medial consonant clusters into the codas and onsets that make them up.
+For this task, it is necessary to separate medial consonant clusters into codas and onsets.
 While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic.
 For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language.
-Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which contrast \emph{chemistry} and \emph{ministry}.
+Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which explain the \emph{chemistry}/\emph{ministry} syllabification contrast.
 
@@
 When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
-Onset maximization makes the right predictions for words like \emph{propaganda}, where the medial cluster cannot be an onset, but \citet[][55]{Harris1994} notes that it makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
+\citet[][55]{Harris1994} notes that onset maximization makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
 The approach adopted here is to first apply onset maximization, then to reassign the first consonant of a medial onset to the preceding coda if it is immediately preceded by a stressed lax vowel.

### 2012-04-26 4.1.3.tex
Albright MS data and some writing on chapter 3

@@
 \begin{example}
+\label{VBR}
 VBR's complex onsets \citep[79--80, his transcriptions]{Buchwald2005}:
@@
 \noindent
-VBR breaks up complex onsets with epenthesis, including those that consist of a consonant and a back onglide cluster (\lastx b).
-However, no epenthesis occurs between a consonant and a front onglide; rather, the glide is absent (\lastx c).
+VBR breaks up complex onsets with epenthesis, including those that consist of a consonant and a back onglide cluster (\ref{VBR}b).
+However, no epenthesis occurs between a consonant and a front onglide; rather, the glide is absent (\ref{VBR}c).
 The failure of the front onglide to pattern with other consonant clusters suggests once again that the glide is part of the nucleus.
@@
 \noindent
-Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\nextx a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.
+Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.
 
 \begin{example}
+\label{td}
 /t, d/-\textsc{Deletion} in American English:

### 2012-04-26 4.2.1.tex
Albright MS data and some writing on chapter 3

@@
 This analysis is not purely allophonic, as nasal place assimilation alternations target the final consonant of the \emph{im-}/\emph{in-} prefix.
-In (\lastx ab), the prefix-final consonant takes on the same major place of articulation as the following obstruent, and in (\lastx c), where the root is vowel-initial, the prefix-final consonant is realized as a default /n/.
+In (\ref{npa}ab), the prefix-final consonant takes on the same major place of articulation as the following obstruent, and in (\ref{npa}c), where the root is vowel-initial, the prefix-final consonant is realized as a default /n/.
 
 \begin{example}
+\label{npa}
 Nasal place assimilation in Latinate prefixes:

### 2012-05-03 2.2.1.tex
another day at the office

@@
 
-%\subsubsection{Prediction distributions}
+\subsubsection{Prediction distributions}
 
-\citet{Coltheart1977}
+\begin{table}
+\centering
 
-\citet{Hayes2008a}
-\citet{Albright2009a}
+\begin{tabular}{l l l l l}
+\toprule
+Spearman $\rho$             & MaxEnt         & density        & bigram         & attested onsets/rimes \\
+\midrule
+\citeauthor{Greenberg1964a} & 0.765          & 0.648          & 0.863          & \textbf{0.845}        \\
+\citeauthor{Scholes1966}    & 0.762          & 0.718          & \textbf{0.827} & 0.791                 \\
+\citeauthor{Albright2003a}  & 0.429          & \textbf{0.742} & 0.708          & 0.725                 \\
+\citeauthor{Albright2007}   & \textbf{0.926} & 0.619          & 0.523          & 0.793                 \\
+\bottomrule
+Kendall $\tau_b$            & MaxEnt         & density & bigram & attested onsets/rimes \
+\midrule
+\citeauthor{Greenberg1964a} & 0.585          & 0.462   & 0.692  & \textbf{0.716}      \\
+\citeauthor{Scholes1966}    & 0.597          & 0.565   & 0.652  & \textbf{0.664}      \\
+\citeauthor{Albright2003a}  & 0.343          & 0.556   & 0.506  & \textbf{0.599}      \\
+\citeauthor{Albright2007}   & \textbf{0.788} & 0.462   & 0.371  & 0.659               \\
+\bottomrule
+Goodman-Kruskal $\gamma$    & MaxEnt & density & bigram & attested onsets/rimes
+\midrule
+\citeauthor{Greenberg1964a} & 0.684  & 0.462   & 0.692  & \textbf{1.000} \\
+\citeauthor{Scholes1966}    & 0.634  & 0.614   & 0.667  & \textbf{0.995} \\
+\citeauthor{Albright2003a}  & 0.656  & 0.575   & 0.509  & \textbf{0.953} \\
+\citeauthor{Albright2007}   & 0.871  & 0.498   & 0.373  & \textbf{0.977} \\
+\bottomrule
+\end{tabular}
+
+\caption{Here is my caption.  All values are significant at $\alpha = 0.05$}
+\end{table}
+
+\begin{table}
+\centering
+
+\begin{tabular}{l r r r}
+\toprule
+Spearman $\rho$             & MaxEnt              & density             & bigram              \\
+\midrule
+\citeauthor{Greenberg1964a} & 0.185               & 0.154               & \textbf{0.319}      \\
+\citeauthor{Scholes1966}    & 0.122               & 0.174               & \ul{\textbf{0.292}} \\
+\citeauthor{Albright2003a}  & 0.011               & \ul{\textbf{0.246}} & 0.203               \\
+\citeauthor{Albright2007}   & \ul{\textbf{0.440}} & \ul{0.340}          & 0.016               \\ 
+\bottomrule
+Kendall $\tau_b$            & MaxEnt              & density             & bigram \\
+\midrule
+\citeauthor{Greenberg1964a} & 0.135               & 0.077               & \textbf{0.205}      \\
+\citeauthor{Scholes1966}    & 0.098               & 0.124               & \ul{\textbf{0.206}} \\
+\citeauthor{Albright2003a}  & 0.005               & \ul{\textbf{0.157}} & 0.126               \\
+\citeauthor{Albright2007}   & \ul{\textbf{0.340}} & 0.215               & $-$0.017            \\
+\bottomrule
+Goodman-Kruskal $\gamma$    & MaxEnt              & density             & bigram \\
+\midrule
+\citeauthor{Greenberg1964a} & \ul{0.158}          & 0.077               & \ul{\textbf{0.205}} \\
+\citeauthor{Scholes1966}    & 0.105               & 0.137               & \ul{\textbf{0.214}} \\
+\citeauthor{Albright2003a}  & 0.010               & \ul{\textbf{0.162}} & \ul{0.127}          \\
+\citeauthor{Albright2007}   & \ul{\textbf{0.376}} & 0.232               & 0.017               \\
+\bottomrule
+\end{tabular}
+
+\caption{Here is my caption; underlined values are significant at $\alpha = 0.05$}
+\end{table}
+
+%\citet{Coltheart1977}
+%
+%\citet{Hayes2008a}
+%\citet{Albright2009a}
 

### 2012-05-03 3.1.1.tex
another day at the office

@@
 
-There are however a few complications.
+There are a few complications, however.
 First, not all polysyllabic roots conform to \textsc{Backness Harmony}.
@@
 As disharmonic roots still trigger suffix harmony, \citet[][212, 289]{Anderson1974} proposes to distinguish suffix harmony (an alternation) from a sequence structure constraint governing root harmony.
+This point is echoed by \citet{Iverson1978}.
 
-However, as \citet[][?]{Zonneveld1978} notes, the theory of exceptionality proposed in \emph{SPE} (p.~374f.) provides a direct account of suffix harmony in disharmonic roots.
+However, as \citet[][197f.]{Zonneveld1978} notes, the theory of exceptionality proposed in \emph{SPE} (p.~374f.) provides a direct account of suffix harmony in disharmonic roots.
 \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule $R$ must be marked [$+$rule $R$] by convention.

### 2012-05-07 3.1.1.tex
more writing

@@
 
-One of the most salient properties of Turkish is its harmony system, which \citeauthor{Lees1966b} (\citeyear[][35]{Lees1966b}, \citeyear[][284]{Lees1966a}) derives by means of feature spreading rules.
+\citeauthor{Lees1966b} (\citeyear[][35]{Lees1966b}, \citeyear[][284]{Lees1966a}) models the Turkish vowel harmony systems by means of three feature spreading rules.
 Backness harmony is the the most general of these processes.
@@
 
-The rule of \textsc{Backness Harmony} spreads a vowel's \textsc{Back} specification rightward over any intervening consonants onto the next vowel.
+\textsc{Backness Harmony} spreads the backness specification of a vowel rightward ignoring any intervening consonants.
 
@@
 \xymatrix@R=24pt@C=24pt{
-\txt{V}                                        & \txt{C$_0$} & \txt{V} \\
+\txt{V}                                        & \txt{C$_0$~~~~~} & \txt{V} & \txt{(condition: rightward application)} \\
 \txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
@@
 
-This rule accounts for a great deal of the suffix allomorphy found in Turkish.
-For instance, the nominative plural (nom.pl.) suffix is \emph{-ler} when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}], as predicted by Backness Harmony.
+This rule accounts for the tendency of polysyllablic roots to either contain all [$+$\textsc{Back}] or all [$-$\textsc{Back}];
+%, as can be seen in the nominative singulars in (\ref{turknom}b); 
+this tendency is quantified in \S\ref{lexstats} below.
+\textsc{Backness Harmony} also accounts for the allomorphs of several inflectional suffixes.
+For instance, the nominative plural (nom.pl.) suffix is \emph{-ler}, when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].
 
@@
 
-\vspace{0.5\baselineskip}
+%\vspace{0.5\baselineskip}
 \begin{tabular}{l l l l l}
@@
    & pul            & pullar         & `stamp'   \\
-b. & neden          & nedenler       & `reason'  & (TELL) \\
-   & boğaz          & boğazlar       & `throat'  \\
-   & kiler          & kilerler       & `pantry'  \\
-   & pelür          & pelürler       & `tissue paper' \\
-   & sapık          & sapıklar       & `pervert' \\
+b. & neden          & nedenler       & `reason'  & (TELL) \\ % front
+   & kiler          & kilerler       & `pantry'           \\ % front
+   & pelür          & pelürler       & `tissue paper'     \\ % back
+   & boğaz          & boğazlar       & `throat'           \\ % back
+   & sapık          & sapıklar       & `pervert'          \\ % back
 \end{tabular}
@@
 First, not all polysyllabic roots conform to \textsc{Backness Harmony}.
-In this case, as can be seen from (\ref{turkexcept}a), suffix vowels generally harmonize with the final root vowel.
-There is also a small class of nouns, shown in (\ref{turkexcept}b), which exhibit [$-$\textsc{Back}] suffix vowels despite the fact that their final root vowel is [$+$\textsc{Back}].
+In this case, as shown in (\ref{turkexcept}a), suffix vowels generally harmonize with the final root vowel.
+There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} despite the fact that their final root vowel is [$+$\textsc{Back}].
 
@@
 
-It is uncontroversial that the disharmonic suffixes of (\ref{turkexcept}b) are no more than very sporadic exceptions to \textsc{Backness Harmony}, root disharmony has ben the subject of much debate.
-As disharmonic roots still trigger suffix harmony, \citet[][212, 289]{Anderson1974} proposes to distinguish suffix harmony (an alternation) from a sequence structure constraint governing root harmony.
-This point is echoed by \citet{Iverson1978}.
+While it is uncontroversial that the disharmonic suffixes of (\ref{turkexcept}b) are no more than very sporadic exceptions to \textsc{Backness Harmony}, root disharmony has ben the subject of much debate.
+As disharmonic roots still trigger suffix harmony, \citet[][212, 289]{Anderson1974} and \citet{Iverson1978} propose to separate suffix harmony, an alternation, from a sequence structure constraint governing root harmony.
 
-However, as \citet[][197f.]{Zonneveld1978} notes, the theory of exceptionality proposed in \emph{SPE} (p.~374f.) provides a direct account of suffix harmony in disharmonic roots.
-\citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule $R$ must be marked [$+$rule $R$] by convention.
-A root or affix which fails to undergo $R$ despite otherwise matching the structural description is simply said to be marked [$-$rule $R$].
-Under this account, exceptionality boils down to failing to match an extended structural description.
-If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the environment, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.
+The disadvantage of this account is that it introduces a sort of ``duplication'' (in the sense of \citealt{Kenstowicz1977}) of sequence structure and phonological generalizations, differning only in their patterns of exceptionality.
+%\citep{Kisseberth1970b} (see also \citealp[][401]{Stanley1967} and \emph{SPE}:382)
+However, \citet[][197f.]{Zonneveld1978} observes that the theory of exceptionality proposed in \emph{SPE}
+% (p.~374f.)
+can account for suffix harmony in disharmonic roots.
+\citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
+A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}].
+In other words, a form is never truly an ``exception'': it simply fails to match an extended structural description including the rule feature.
+If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.\footnote{The definition of ``target of a rule'' is not obvious in autosegmental rules, but any useful formalization of this notion should, for example, identify a harmonic suffix vowel as the target of \textsc{Backness Harmony}.} With this in place, \textsc{Backness Harmony} is sufficient to account for both root and suffix harmony.
 
@@
 
-An alternative account, first proposed by \citet{Clements1982}, uses underspecificatio to derive the non-exceptionality of harmonic roots.
-The initial-syllable vowel of a Turkish root is contrastively specified for backness (e.g., \emph{kül} `ash' vs.
+Before moving on, it is necessary to dispense with an alternative analysis originally proposed by \citet{Clements1982}.
+Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
 \emph{kul} `servant', \emph{kepek} `bran' vs.
-\emph{kapak} `lid'); harmonic roots are ones in which non-initial syllable vowels cannot contrast for this feature, and \citeauthor{Clements1982} proposes to leave them underspecified.
-Only in disharmonic roots does \textsc{Back} need to be specified for non-initial syllable vowels.
+\emph{kapak} `lid'), whereas harmonic roots are those in which the backness of any remaining vowels is predictable.
+\citeauthor{Clements1982} propose that these vowels are underspecified for backness, whereas the non-initial vowels of disharmonic roots are fully specified.
+This is schematized below.
+
 %This is exemplified below in (\ref{spec}).
 %(e.g., \emph{deve} `camel' vs. \emph{deva} `medicine', \emph{sene} `year' vs. \emph{sena} `praise'). 
-
 %There is some further evidence that individual vowels may differ in specification for this feature even within individual roots or affixes.
@@
 \label{spec}
-\textsc{Back} specification for (dis)harmonic roots:
+Underspecification in harmonic roots (after \citealp{Clements1982}):
 
-\vspace{0.5\baselineskip}
+%\vspace{0.5\baselineskip}
 \xymatrix@R=24pt@C=24pt{
@@
 
-\noindent
-Harmonizing suffix vowels will also be underspecified for \textsc{Back}.
-Of course, \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being the use of a a \textsc{Structure Preservation} condition \citep{Kiparsky1985}.
-Unfortunately, this detail reintroduces the duplication alluded to above.
-Any condition on the rule of \textsc{Backness Harmony} which prevents overwriting entails that it has no control over the distribution of harmonic and disharmonic roots, and thus fails to account for predominance of harmonic roots.
+Harmonizing suffix vowels will also be underspecified for backness.
 
-\citet{Clements1982} and \citet{Kaun1999}
+Two additional details are needed to complete this analysis.
+First, there must be some kind of constraint which prevents speakers from positing redundant backness specifications for non-initial vowels in harmonic roots, a constraint on identical adjacent specifications in underlying representations, the lexical ``obligatory contour principle'' \citep[][OCP]{Leben1973}, on surface tonal representations \citep{Goldsmith1976}, or both (\citealp{Leben1978}, \citealp{McCarthy1986}; see \citealp{Odden1986,Odden1988} for criticism).
+Second, \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being the use of a structure preservation condition \citep{Kiparsky1985}.
+However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are formally marked in Turkish.\footnote{On the other hand, if one interprets ``harmony'', that is, a single backness specification per root, as the learner's default (as \citealp{Odden1986} proposes for the tonal obligatory contour principle), this account is a mere notational variant of the rule exceptionality account.}
+%\footnote{Kie Zuraw (p.c.) suggests that the markedness of disharmonic roots might be attributed to the presence of more structure, namely, multiple specifications for backness.}
+I reject the underspecification analysis on these grounds. Rather, I assume a unitary phonological rule of \textsc{Backness Harmony} and assign [$-$\textsc{Backness Harmony}] to disharmonic roots.
 
-\footnote{Thanks to Kie Zuraw for bringing this data to my attention.}
+\subsubsection{External evidence}
+\label{backharmexternal}
 
-the epenthetic vowel is always [$+$\textsc{High}], and when the first consonant of the cluster is non-dorsal, it shows backness harmony
+While suffix harmony can be inferred from alternations, no evidence has yet been presented to show that speakers internalize in any way tendency for roots to exhibit backness harmony.
+
+%\begin{quote}
+%\ldots{}these generalizations\ldots{}have no observable consequences in the course of the normal use of the language. \citep[][320]{Zimmer1969}
+%\end{quote}
+
+Indeed, \citeauthor{PE} adopt the null hypothesis that speakers do not internalize any generalizations about possible or likely sequences in underlying representations.
+
+\begin{quote}
+Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for.
+\citep[][18]{PE}
+\end{quote}
+
+Several independent pieces of external evidence argue that this is not the case for root harmony in Turkish. 
+The adaptation of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, is one such case.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} While some speakers are said to be capable of pronouncing these non-native clusters, in normal speech the cluster is resolved by epenthesis of a [$+$\textsc{High}] vowel.
+In most cases, the epenthetic vowel matches the following root vowel for backness, suggesting that adapt foreign words in a way that conforms to the root harmony generalization.
+
+%the epenthetic vowel is always [$+$\textsc{High}],
+%and in general, it conforms to \textsc{Backness Harmony}.
+%For reasons that are 
+%and when the first consonant of the cluster is non-dorsal, it shows backness harmony
 
@@
 
-\vspace{0.5\baselineskip}
+%\vspace{0.5\baselineskip}
 \begin{tabular}{l l l l l l}
@@
    & fren    & \alt{} & firen    & `break'     \\
-   & trablus & \alt{} & tırablus & `Tripoli'   \\
-   & brom    & \alt{} & burom    & `bromide'   \\
+b. & brom    & \alt{} & burom    & `bromide'   \\
    & prusya  & \alt{} & purusya  & `Prussia'   \\
-b. & kral    & \alt{} & kıral    & `king'      \\
-   & grup    & \alt{} & gurup    & `group'     \\
-c. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
-   & kredi   & \alt{} & kıredi   & `credit'    \\
+%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
+%   & kredi   & \alt{} & kıredi   & `credit'    \\
+%   & trablus & \alt{} & tırablus & `Tripoli'   \\
+%b. & kral    & \alt{} & kıral    & `king'      \\
+%   & grup    & \alt{} & gurup    & `group'     \\
 \end{tabular}
@@
 
-\subsubsection{Psycholinguistic evidence}
-
-The evidence from alternations leaves open the question of whether speakers internalize a generalization regarding the tendency of roots to conform to \textsc{Backness Harmony}.
-One piece of evidence that speaks in favor of root-internal harmony comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but it corresponds to a morphological process native to the related language Tuvan, in which it conveys a sense of ``vagueness or jocularity'', and \citeauthor{Harrison2001} report that it can be quickly taught to even young Turkish speakers.
-The game consists of reduplication of the base and replacement of the first vowel in the reduplicant with \emph{a} or \emph{u}.
-
-In both Tuvan and Turkish, the unchanged reduplicant vowel is also affected.
-Reduplication interacts with root harmony in both Tuvan and Turkish.
-When the base is harmonically [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] results in what \citeauthor{Harrison2001} call ``reharmonization'' (\ref{redupgame}a).
+Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but is identical to a morphological process in the related language Tuvan, in which it conveys a sense of ``vagueness or jocularity'', and \citeauthor{Harrison2001} report that it can be quickly taught to Turkish speakers.
+The game consists of reduplication of the base and replacement of the first vowel in the reduplicant with the [$+$\textsc{Back}] vowels \emph{a} or \emph{u}.
+Both in the native Tuvan process and in the Turkish game, the second vowel of the reduplicant (shown in braces below) is affected by this process: when the base is harmonically [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] results in what \citeauthor{Harrison2001} call ``reharmonization'', shown in  (\ref{redupgame}a).
 
@@
 \begin{tabular}{l l l l}
-a. & kibrit & kibrit-kabrıt & `match'    \\
-   & bütün  & bütün-batın   & `whole'    \\
-b. & mali   & mali-muli     & `Mali'     \\
-   & butik  & butik-batik   & `boutique' \\
+a. & kibrit & kibrit-\{kabrıt\} & `match'    \\
+   & bütün  & bütün-\{batın\}   & `whole'    \\
+b. & mali   & mali-\{muli\}     & `Mali'     \\
+   & butik  & butik-\{batik\}   & `boutique' \\
 \end{tabular}
@@
 
-\noindent
-Under the underspecification hypothesis, non-initial harmonic vowels lack an underlying \textsc{Back} feature, so it comes as no surprise that changing the \textsc{Back} specification of the root-initial vowel results in reharmonization.
-And the full specification of disharmonic roots correctly predicts that they will be exempt from reharmonization (\ref{redupgame}b), which is also borne out in Tuvan and in an unrelated Finnish language game \citep{Campbell1986}.
+\noindent \citeauthor{Harrison2001} analyze this in terms of the \citeauthor{Clements1982} underspecification analysis, but it is equally consistent with full specification.
+Reharmonization is simply the spread of the backness specification of the \emph{a} or \emph{u} in the reduplicant.
+On the other hand, (\ref{redupgame}b) shows that this does not obtain for disharmonic roots; presumably, the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication as well.
+Similar results obtain both in Tuvan and in an unrelated Finnish language game \citep{Campbell1986}.
 
-A number of studies have found that speakers of Finnish use disharmony in speech processing experiments.\footnote{Thanks to Charles Yang for pointing out the relevance of these studies to me.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which are harmonic for the feature \textsc{Back}.
+A number of studies on Finnish suggest that speakers of that language, which has a vowel harmony system similar to that of Turkish, make use of (dis)harmony to processing running speech.\footnote{Thanks to Charles Yang for bringing these studies to my attention.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which conform to backness harmony.
 These stimuli are auditorily presented to subjects who are asked to press a button when the nonce trisyllable ends with a target nonce disyllable, or a real disyllabic word.
-Speakers are quicker to press the button when the prefix and disyllabic word disagree for \textsc{Back}.
+Speakers are quicker to press the button when the prefix and disyllabic word have different backness specifications.
 These results suggest that speakers are attuned to the fact that disharmonic transisitions are good predictors of word boundaries.
@@
 
-\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect on word-spotting as it does in Finnish: speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by the pseudo-prefix \emph{gölü-}, a disharmonic transistion, than when it is preceded by the pseudo-prefix \emph{golu-}, a harmonic transition.
+\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect in word-spotting tasks as it does in Finnish: speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by the pseudo-prefix \emph{gölü-}, a disharmonic transistion, than when it is preceded by the pseudo-prefix \emph{golu-}, a harmonic transition.
 \citeauthor{Kabak2010} find that effect of harmony does not obtain for speakers of French, a language which lacks vowel harmony.

### 2012-05-07 3.1.2.tex
more writing

@@
 
-The rule of roundness harmony and the data that motivates it overlaps with the preceding evidence for \textsc{Backness Harmony}.
+Roundness harmony is a closely related process with a somewhat narrower scope.
 
@@
 
-The environment for roundness harmony differs from \textsc{Backness Harmony} in that it requiers the target to be [$+$\textsc{High}].
+Roundness harmony is quite similar to \textsc{Backness Harmony}, with the additional stipulation that the target must be [$+$\textsc{High}].
 
@@
 \xymatrix@R=24pt@C=24pt{
-\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{..}[drr] &             & \txt{[$+$\textsc{High}]} \\
+\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{--}[drr] &             & \txt{[$+$\textsc{High}]} \\
 \txt{V}                                         & \txt{C$_0$} & \txt{V}\ar@{-}[u] \\
@@
 
-\vspace{0.5\baselineskip}
+%\vspace{0.5\baselineskip}
 \begin{tabular}{l l l l l l l}
    & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
-a. & ip             & ipi            & ipin           & `rope' & (CS:216) \\
+a. & ip             & ipi            & ipin           & `rope' & \citep[][216]{Clements1982} \\
    & el             & eli            & elin           & `hand'    \\
@@
 
-\noindent
-Much as was the case for \textsc{Backness Harmony}, there are disharmonic roots which participate in suffix harmony.
-Once again, underspecification of harmonic roots and full specification of disharmonic roots allows for the use of the \emph{SPE} exception convention.
+\noindent As was the case for \textsc{Backness Harmony}, disharmonic roots are found, and the final vowel of disharmonic roots triggesr suffix harmony.
 
-\subsubsection{Psycholinguistic evidence}
+\subsubsection{External evidence}
 
-The only psycholinguistic evidence for \textsc{Roundness Harmony} comes from \citeauthor{Harrison2001}'s language game (\ref{redupgame}).
-As shown above, the second vowel in the reduplicated form of \emph{bütün} `whole' is \emph{bütün-batın}.
-The second vowel in the base, \emph{ü}, is reharmonized as [$+$\textsc{Back}, $-$\textsc{Round}], indicating that \textsc{Roundness Harmony} also participates in reharmonization.
+Though few studies have directly sought external evidence for speaker's awareness of root \textsc{Roundness Harmony}, two of the findings reviewed in \S\ref{backharmexternal} above speak in support of this claim.
+First, the epenthetic vowel in non-native word-initial onsets
+like \emph{trablus} \alt{} \emph{tırablus} `Tripoli' conform also to \textsc{Roundness Harmony}.
+%\citealp[][247]{Clements1982}). 
+Second, \textsc{Roundness Harmony} participates in reharmonization in \citeauthor{Harrison2001}'s language game: the reduplicated \emph{bütün-batın} exhibits replacement of the second \emph{ü}, a [$+$\textsc{Round}] vowel, with [$-$\textsc{Round}] \emph{ı}.

### 2012-05-07 3.1.3.tex
more writing

@@
 
-The final rule
+%\begin{quote}
+%\ldots{}these generalizations\ldots{}have no observable consequences in the course of the normal use of the language. \citep[][320]{Zimmer1969}
+%\end{quote}
 
@@
 \xymatrix@R=24pt@C=8pt{
-\txt{[$-$\textsc{Round}]} &                                         & \txt{[$-$\textsc{High}]} & \txt{[$+$\textsc{Labial}]}    & \txt{[$+$\textsc{High}]}\ar@{-}[dr] &         & \txt{[$+$\textsc{Round}]}\ar@{..}[dl] \\
+\txt{[$-$\textsc{Round}]} &                                         & \txt{[$-$\textsc{High}]} & \txt{[$+$\textsc{Labial}]}    & \txt{[$+$\textsc{High}]}\ar@{-}[dr] &         & \txt{[$+$\textsc{Round}]}\ar@{--}[dl] \\
                          & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C$_0$ C C$_0$}\ar@{-}[u] &                                      & \txt{V} & \\

### 2012-05-07 3.1.tex
more writing

@@
 
-\citet{Lees1966a,Lees1966b} proposes three constraints on Turkish vowel sequences 
-which have been the attention of many subsequent studies on vowel harmony and on sequence structure constraints.
-In this section, these constraints are formalized, and where possible, related to phonological alternations in Turkish and to various sources of external evidence.
+\citet{Lees1966a,Lees1966b} proposes three constraints on Turkish vowel sequences which have been the focus of many subsequent studies.
+In this section, these constraints are formalized, and where possible, related to phonological alternations in Turkish and to evidence that speakers have internalized these restrictions.
 The following feature specification for the Turkish vowels is assumed throughout.

### 2012-05-07 3.2.2.tex
another day at the office
more writing

@@
 
-There is some disagreement in the literature about just what \citet{Zimmer1969} finds regarding \textsc{Labial Attraction}.
+There is some disagreement in the literature about just what \citeauthor{Zimmer1969}'s study reveals about \textsc{Labial Attraction}.
 
-\citet{Zimmer1969}
-(see also \citet[][4]{Zuraw2000})
-claims however, Zimmer's evidence suggests that the speakers did internalize a version of the labial attraction generalization.
+zimmer
+clements
+\citet{Ito1993} argue that the data shows that it holds of native words only
+\citet{Inkelas2001} refute this point
+
+\citet{Zuraw2000}:4
 
@@
 
-since the opposing are all broader versions than the original formulation by \citeauthor{Lees1966a}.
-
-\citet{Ito1993} argue that the data shows that it holds of native words only
+%since the opposing are all broader versions than the original formulation by \citeauthor{Lees1966a}.
 
@@
 \subsubsection{Backness harmony}
-
-($\gamma = 0.597$, $p = 5.2$\e{-29})
 
@@
 
+$\gamma = 0.597$, $p = 5.2$\e{-29}
+
 \subsubsection{Roundness harmony}
@@
 \end{example}
+
+$\gamma = 0.641$, $p = 7.7$\e{-27}
 
@@
 
-($\gamma = 0.087$, $p = 0.101$)
+$\gamma = 0.087$, $p = 0.101$

### 2012-05-07 3.tex
more writing

@@
-\chapter{Static and derived constraints}
 \label{turkish}
 
-glossematics
+%In this chapter, I present evidence that speakers 
+%glossematics
+%\citet{Chomsky1965} propose
 
-\citet{Chomsky1965} propose
+%In retrospect, it seems somewhat odd that \citeauthor{Chomsky1965} would coh
+%The contrast between \emph{blick} and \emph{bnick} is what has been called a ``projection problem'' \citep{Baker1979}
+%This is not a universal, but rather something English speakers must learn: Moroccan Arabic, for instance, does permit \emph{bn} onsets (e.g., \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet').
+%\citet{Lees1966b,Lees1966a}
 
-In retrospect, it seems somewhat odd that \citeauthor{Chomsky1965} would coh
-It might 
-In defining the
+%\citet{Davidson2005,Davidson2006a} shows that this ``excrescent'' schwa is distinct both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.
+%\citet{Wilson2011} show that speakers use this
 
-The contrast between \emph{blick} and \emph{bnick} is what has been called a ``projection problem'' \citep{Baker1979}
-
-This is not a universal, but rather something English speakers must learn: Moroccan Arabic, for instance, does permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet').
-
-\citet{Lees1966b,Lees1966a}
-
-This is what we might call a projec
-In presenting the components of a 
-a term which
-
-by subphonemic details
-
-\citet{Davidson2005,Davidson2006a} shows that excresent schwa is acoustically and articulatorily distinct from "lexical schwa":
-\citet{Wilson2011} show taht speakers use this
-
-%English speakers' z^@k is more similar to \emph{scum} than \emph{succumb}
-
-/sC-/ onsets than /s@C-/.
-
-% Davidson 2005: z^@C is more like sC than s@C (ultrasound of articulations)
-% Davidson 2006a: lexical and excresant schwa are not the same (acoustics)
-
-% Davidson 2010: production
-% Wilson & Davidson 2011:
-
-\citet{Berent2007a}
-\citet{Berent2007b}
-\citet{Berent2008a}
-
-% Berent et al. 2007
+%\citet{Berent2007a}
+%\citet{Berent2007b}
+%\citet{Berent2008a}
 % 598: Highly marked inputs are repaired in perception to abide by the grammatical restrictions of the language
@@
 % Berent et al. 2008a:
-\citet{Peperkamp2007}
+%\citet{Peperkamp2007}
 % Peperkamp 2007: during one stage of processing phonology is undone with help from the lexicon...it only takes licit inputs and is irrelevant (p. 633)
-\citet{Dupoux1999}
+%\citet{Dupoux1999}
 % Dupoux et al. 1999: Japanese speakers still can't distinguish between ebzo and ebuzo in an ABX task using the SAME a or b.
@@
 
-does not so much as explain the thing itself but rather suggests the
-
-I believe a number of the predictions of this model are wrong
-the original and still-compelling 
-findings that drove
-
- motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
+% motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
 \citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible.
@@
 
-This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
+%This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
 Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
@@
 
-the other being the subject
+%a principle which is one of the two motivations  motivated the introduction of morpheme structure constraints.
 
-a principle which is one of the two motivations  motivated the introduction of morpheme structure constraints.
-
-In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon does not always require static constraints on possible underlying representations, since such distributions arise by chance and static constraints defined over natural classes contribute little to predicting the contents of the lexicon.
+%In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon does not always require static constraints on possible underlying representations, since such distributions arise by chance and static constraints defined over natural classes contribute little to predicting the contents of the lexicon.
 %This result does not directly bear on the question of whether static restrictions have any synchronic reality.
 
-Yet countless static constraints are statistically supported.
+%Yet countless static constraints are statistically supported.
 Under the null hypothesis advanced in Chapter \ref{MSCs}, the lexical statistics of co-occurrence restrictions are irrelevant: the language acquisition device only attends to derived restrictions and ignores static restrictions regardless of their statistical reliability.
@@
 
-\begin{quotation}
-\ldots the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
-\end{quotation}
+%\begin{quotation}
+%\ldots the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
+%\end{quotation}
 
-\noindent
-A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions.
+%\noindent A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions.
 Any explicit theory of static constraints will have to impose formal restrictions on possible co-occurrence restrictions.
-The set of possible constraints must respect results in formal learning theory, and may also be, for example, constrained by naturalness biases or formal markedness considerations.
+The set of possible constraints must respect results in formal learning theory, and may also be constrained by markedness considerations, for example.
 
-A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
+%A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
 This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).
 
-Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
+%Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
 In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.
 
-In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word.
+%In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word. 
+%The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements. 
+%The data below, drawn from Turkish, shows that
 
-The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements.
-
-The data below, drawn from Turkish, shows that
-
-This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
+%This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
 Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements.

### 2012-05-07 4.1.3.tex
more writing

@@
 There is extensive motivation for this principle.
-When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm].
-In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [u\lm] to be anomalous \citep{Moreland2009}.
+When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː].
+In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}.
 This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
-\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [ju\lm] also suggests the glide is nuclear.
+\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear.
 There is also strong external support for this principle.
-The [ju\lm] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants.
-The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
+The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants.
+The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor homponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
 Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.
@@
 
-%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [ju\lm] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant.
-%This is the only sign that [ju\lm] shows an affiliation for the onset.
+%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [juː] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant.
+%This is the only sign that [juː] shows an affiliation for the onset.
 
@@
 Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}.
-Unlike [ju\lm], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.
+Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.
 And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.
 
-%Whereas [Cju\lm] syllables attracts stress, [kwV] syllables do not \citep[][162f.]{Davis1995}.
+%Whereas [Cjuː] syllables attracts stress, [kwV] syllables do not \citep[][162f.]{Davis1995}.
 

### 2012-05-07 4.2.3.tex
more writing

@@
 
-While this procedure can be repeated indefinitely (using the CELEX data, listed in Appendix \ref{A}), consideration of a single characteristic run of the simulation should suffice.
+While this procedure can be repeated indefinitely 
+%(using the CELEX data, listed in Appendix \ref{A}), 
+consideration of a single characteristic run of the simulation should suffice.
 The saturation rate of the simulated sample (18.2\%) is very close to the real sample (18.8\%).

### 2012-05-07 ack.tex
more writing

@@
 So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy. 
-More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, whose influence can be seen throughout. 
-Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are responsible for creating an environment at Penn where a student could write this admittedly contenious dissertation. 
+More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, 
+who did some much to make this work, and me as a linguist,
+into what it is now.
+Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are to be thanked for creating an environment at Penn where a student could write this admittedly contenious dissertation. 
 And though his name appears nowhere else on this document, it is hard to imagine this work could exist without Bill Labov.
@@
 Steve Anderson graciously agreed to serve on my committee.
-Morris Halle told me to never apologize.
+Morris Halle taught me to speak clearly and never apologize.
 
-I would like to thank my fellow students at Penn, but especially my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody. 
-Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document. 
-Maryam Bakht kept me centered.
+While so many of my fellow students at Penn have a hand in this document, 
+I would like to especially thank my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody.
+
+%Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document.
+%Maryam Bakht kept me centered.
 

### 2012-05-13 1.1.1.tex
More arranging.
chapters 2 and 4 have been swapped
include at the chapter level

@@
-% 1.1.1: Structuralism
+\subsection{Structuralism}
 
@@
 In the reflexive infinitive, /\ldots t\pal-s\ldots/ is realized [ts], with no palatalization.
-However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\nextx c)---is preserved in the reflexive imperative.
+However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\ref{russian}c)---is preserved in the reflexive imperative.
 
-\ex Russian reflexives \citep[after][]{Jakobson1932}:\footnote{
-I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses an abstract phonemic transcription. 
-The broad transcription below corresponds to the speech of Lev Blumenfeld, who I thank for help with this data.
-} \\
+\begin{example}
+\label{russian}
+Russian reflexives \citep[after][]{Jakobson1932}:\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses an abstract phonemic transcription.
+Thanks to Lev Blumenfeld (p.c.) for help with the transcription of this data.}
+
 \begin{tabular}{l l l l} %\toprule
@@
 %   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
-\end{tabular} \xe
+\end{tabular}
+\end{example}
 
-\noindent
-The forms in (\lastx c) have different outcomes for their /t\pal-s/ clusters.
+\noindent The forms in (\ref{russian}c) have different outcomes for their /t\pal-s/ clusters.
 \citeauthor{Jakobson1932} proposes that the preservation of palatalization is a special property of the imperative.
@@
 
-\ex Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}: \\
+\begin{example}
+Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}:
+
 \begin{tabular}{r l l l} %\toprule
@@
 d. & *{C}i{C}e & [tenìkeet\`ə] & `and went away'      \\ 
-\end{tabular} \xe
+\end{tabular}
+\end{example}
 
-\noindent
-\citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a morpheme discovery routine.
+\noindent \citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a morpheme discovery routine.
 

### 2012-05-13 2.2.1.tex
include at the chapter level

--- 2012-04-18 2.2.1.tex
@@
-\subsection{Assessing gradience}
+\subsection{Data sources}
 
-\subsubsection{Response distributions}
+While these are by no means the only wordlikeness s studies in Englihs, these studies all have the following characteristics: 
+the item-averaged ratings are reported in the document,
+the stimuli are presented auditorily,
+and at least some stimuli are not phonotactically valid in English.
 
-This is shown in Figure \ref{density}.
+(some caveats about subjects
 
-\begin{figure}
-\centering
-\includegraphics{density.pdf}
-\caption{Density of responses}
-\label{density}
-\end{figure}
+\citep{Shademan2007}
 
-\begin{example}
-EM test results:
+cost of prior averaging
+\citep{Baayen2004}
 
-\vspace{0.5\baselineskip}
-\begin{tabular}{l | r r | r r r r r r | r r}
-\toprule
-     & \multicolumn{2}{c|}{one $\mathcal{N}$} & \multicolumn{6}{c|}{mixture of two $\mathcal{N}$s} & \multicolumn{2}{c}{log $\mathcal{L}$ test} \\
-     & $\mu$ & $\sigma$ & $\mu_1$ & $\sigma_1$ & mix$_1$ & $\mu_2$ & $\sigma_2$ & mix$_2$ & $\Lambda$ & $p$-value  \\
-\midrule
-G\&J & 0.378 & 0.295    & 0.150   & 0.090      & 0.436     & 0.554   & 0.278    & 0.564     & 5.190    & 0.158   \\
-S    & 0.596 & 0.361    & 0.236   & 0.232      & 0.437     & 0.876   & 0.105    & 0.563 & 48.965    & 1.3\e{-10} \\
-A\&H & 0.431 & 0.266    & 0.301   & 0.200      & 0.559     & 0.595   & 0.246    & 0.411 & 16.186    & 0.001      \\
-\bottomrule
-\end{tabular}
-\end{example}
+\subsubsection{\citealt{Greenberg1964}}
 
-%\ex Residualized correlations
-%\begin{tabular}
-%\toprule
-%ASDF
-%\midrule
-%G\&J & 
-%S    & 
-%A\&H & 
-%\bottomrule
-%\xe
+\citet{Greenberg1964} investigate English wordlikeness using the technique of free magnitude estimation, a mechanism which has become increasingly popular in syntax \citep[e.g.,][]{Bard1996,Sprouse2011}.
+At the beginning of each run of the experiment, a subject heard a recording of the word \emph{stick}.
+In each subsequent trial, the subject heard a new item and was asked to assign a value for ``How far would you say that is from English?'', with \emph{stick} representing ``1''.
+They were explicitly instructed that a word which was ``twice as far from English'' as \emph{stick} should be given ``a number that is twice as large'' (op.
+cit., 160).
+The data used here are from their ``experiment B'', in which 17 undergraduates were presented 13 items.
+This is the only study used here in which actual English words (\emph{stick}, \emph{grass}, \emph{spell}, and \emph{truck}) were intentionalyl included among the stimuli.
 
-%\subsubsection{Prediction distributions}
+\subsubsection{\citealt{Scholes1966}}
 
-\citet{Coltheart1977}
+\citet{Scholes1966} conducts a number of English wordlikeness judgement with middle school children.
+The data used here come from his ``experiment 5'', in which 66 items were presented to 33 seventh-grade students.
+For each stimulus, the subjects produced forced choice ``yes''/``no'' answers to the question of whether the item ``is likely to be usable as a word of English''.
+\citet{Hayes2008a} and \citet{Albright2009a} analyse this data, which is binary at the trial level, as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus.
 
-\citet{Hayes2008a}
-\citet{Albright2009a}
+%mrupation problem
 
-% stats thereof
-%\citet{Schilling2002}
-%\citet{Helguerro1904}
-%\citet{Cohen1956}
+\subsubsection{\citealt{Albright2003b}}
 
-% EM
-\citet{EM}
+\citet{Albright2003b} gather wordlikeness data to provide norms for the \emph{wug}-task that is the focus of their study.
+87 items were presented to 20 undergraduate subjects, who rated each word on a seven-point Likert scale; though \citeauthor{Albright2003b} do not provide the precise instructions the subjects were given, they do report that ``1'', the lowest point on the scale, was labeled
+``completely bizarre, impossible as an English word'', and that the highest point, ``7'', was named ``complete normal, would make a fine English word''.
 
-\footnote{In performing this evaluation, I have benefitted greatly from course notes by Mark Liberman and Stephen Isard.}
-%available at \url{http://www.ling.upenn.edu/courses/cogs501/K-meansHW.html}
+\subsubsection{\citet{Albright2007}}
 
-% -2 log like ratio test
-
-where the difference in degrees o freedom equals $D$
-
-%\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } 
-%\ex $\displaystyle \Lambda = -2 \textrm{ ln } L_0 + 2 \textrm{ ln } L_1}$ \xe
-% > χ^2_{d.f.=D}$ \xe
+Wordlikeness is the focus of an unpublished study by \citet{Albright2007}.
+\citeauthor{Albright2007} presetns an unknown number of subjects with 40 nonce words in addition to 170 fillers, which are rated on the same seven-point Likert scale used by \citet{Albright2003b}.

### 2012-05-13 2.2.2.tex
include at the chapter level

@@
-\subsection{Assessing models}
+\subsection{Response bimodality}
 
-\subsubsection{Basic correlations}
-\subsubsection{Intramodal correlations}
-\subsubsection{Residual correlations}
+\subsubsection{Prediction distributions}
+
+%\citet{Coltheart1977}
+%\citet{Hayes2008a}
+%\citet{Albright2009a}
+
+% stats thereof
+%\citet{Schilling2002}
+%\citet{Helguerro1904}
+%\citet{Cohen1956}
+
+% EM
+\citet{EM}
+
+\footnote{In performing this evaluation, I have benefitted greatly from course notes by Mark Liberman and Stephen Isard.}
+%available at \url{http://www.ling.upenn.edu/courses/cogs501/K-meansHW.html}
+
+% -2 log like ratio test
+
+where the difference in degrees o freedom equals $D$
+
+%\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } 
+%\ex $\displaystyle \Lambda = -2 \textrm{ ln } L_0 + 2 \textrm{ ln } L_1}$ \xe
+% > χ^2_{d.f.=D}$ \xe
+
+\subsubsection{Response distributions}
+
+This is shown in Figure \ref{density}.
+
+\begin{figure}
+\centering
+\includegraphics{density.pdf}
+\caption{Density of responses}
+\label{density}
+\end{figure}
+
+\begin{example}
+EM test results:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l | r r | r r r r r r | r r}
+\toprule
+     & \multicolumn{2}{c|}{one $\mathcal{N}$} & \multicolumn{6}{c|}{mixture of two $\mathcal{N}$s} & \multicolumn{2}{c}{log $\mathcal{L}$ test} \\
+     & $\mu$ & $\sigma$ & $\mu_1$ & $\sigma_1$ & mix$_1$ & $\mu_2$ & $\sigma_2$ & mix$_2$ & $\Lambda$ & $p$-value  \\
+\midrule
+G\&J & 0.378 & 0.295    & 0.150   & 0.090      & 0.436     & 0.554   & 0.278    & 0.564     & 5.190    & 0.158   \\
+S    & 0.596 & 0.361    & 0.236   & 0.232      & 0.437     & 0.876   & 0.105    & 0.563 & 48.965    & 1.3\e{-10} \\
+A\&H & 0.431 & 0.266    & 0.301   & 0.200      & 0.559     & 0.595   & 0.246    & 0.411 & 16.186    & 0.001      \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+%\ex Residualized correlations
+%\begin{tabular}
+%\toprule
+%ASDF
+%\midrule
+%G\&J & 
+%S    & 
+%A\&H & 
+%\bottomrule
+%\xe

### 2012-05-13 2.2.3.tex
More arranging.
include at the chapter level

--- 2012-02-16 2.2.3.tex
@@
-% 2.2.3: Statistical properties of syllable contact
+\subsection{Model comparison}
 
-Above, we have seen that constraints on URs which are derive from phonological alternations restrict the contents of the lexicon.
-Applying these derived constraints to the lexicon increases saturation from 18.8\% to 28.8\% and incurs only a handful of exceptions.
-Yet, despite the fact that ``attestation'' is the minority pattern, attempts to identify static lexical constraints, whether by hand or computational model, lend little additional predictive power.
-There is no evidence that the English syllable contact inventory is subject to any static constraint at all.
-I am forced to conclude that most, if not all, of 71.2\% of possible clusters which are unattested are accidental gaps.
-Below, I show that this state of affairs follows directly from the sparse distribution of codas and onsets.
+\subsubsection{Correlation with judgements}
 
-\subsubsection{The $Z_r$ transform}
+\begin{table}
+\centering
+\begin{tabular}{l l l l l}
+\toprule
+Pearson $R$                 & MaxEnt         & density        & bigram         & attested onsets/rimes \\
+\citealt{Greenberg1964a} & 0.589          & 0.537          & \textbf{0.892} & 0.824                 \\
+\citealt{Scholes1966}    & 0.744          & 0.716          & 0.824          & 0.893                 \\
+\citealt{Albright2003a}  & 0.437          & 0.633          & 0.775          & 0.766                 \\
+\citealt{Albright2007}   & 0.848          & 0.700          & 0.597          & 0.858                 \\
+\bottomrule
+Spearman $\rho$             & MaxEnt         & density        & bigram         & attested onsets/rimes \\
+\midrule
+\citealt{Greenberg1964a} & 0.765          & 0.648          & 0.863          & \textbf{0.845}        \\
+\citealt{Scholes1966}    & 0.762          & 0.718          & \textbf{0.827} & 0.791                 \\
+\citealt{Albright2003a}  & 0.429          & \textbf{0.742} & 0.708          & 0.725                 \\
+\citealt{Albright2007}   & \textbf{0.926} & 0.619          & 0.523          & 0.793                 \\
+\bottomrule
+Kendall $\tau_b$            & MaxEnt         & density & bigram & attested onsets/rimes                \\
+\midrule
+\citealt{Greenberg1964a} & 0.585          & 0.462          & 0.692          & \textbf{0.716}        \\
+\citealt{Scholes1966}    & 0.597          & 0.565          & 0.652          & \textbf{0.664}        \\
+\citealt{Albright2003a}  & 0.343          & 0.556          & 0.506          & \textbf{0.599}        \\
+\citealt{Albright2007}   & \textbf{0.788} & 0.462          & 0.371          & 0.659                 \\
+\bottomrule
+Goodman-Kruskal $\gamma$    & MaxEnt & density & bigram & attested onsets/rimes
+\midrule
+\citealt{Greenberg1964a} & 0.684          & 0.462          & 0.692          & \textbf{1.000}        \\
+\citealt{Scholes1966}    & 0.634          & 0.614          & 0.667          & \textbf{0.995}        \\
+\citealt{Albright2003a}  & 0.656          & 0.575          & 0.509          & \textbf{0.953}        \\
+\citealt{Albright2007}   & 0.871          & 0.498          & 0.373          & \textbf{0.977}        \\
+\bottomrule
+\end{tabular}
+\caption{Here is my caption.  All values are significant at $\alpha = 0.05$}
+\end{table}
 
-In a sparse distribution, many types occur at the same low frequencies.
-\citet{Good1953} notes that this produces a type of quantization, an artificially long and flat right tail in the frequency spectrum.
-\citet[][29]{Church1991} propose a method to eliminate this quantization, the $Z_r$ transform, defined as follows.
-Let $r, n$ be defined such that $n_i$ is the number of types which occur at frequency $r_i$ (i.e., $n$ contains frequencies of type frequencies).
-Further, let $r$ be sorted in increasing order.
-$Z$ is simply each element of $n$ scaled by the nearest points to the left and right.
+\begin{table}
+\centering
+\begin{tabular}{l r r r}
+\toprule
+Pearson $R$                 & MaxEnt              & density             & bigram              \\
+\citealt{Greenberg1964a} & 0.027               & 0.136               & \textbf{0.299}      \\
+\citealt{Scholes1966}    & 0.075               & $-0.064$            & \textbf{0.129}      \\
+\citealt{Albright2003a}  & $-0.020$            & \textbf{0.255}      & 0.112               \\
+\citealt{Albright2007}   & \ul{\textbf{0.352}} & 0.134               & 0.025               \\
+\bottomrule
+Spearman $\rho$             & MaxEnt              & density             & bigram              \\
+\midrule
+\citealt{Greenberg1964a} & 0.185               & 0.154               & \textbf{0.319}      \\
+\citealt{Scholes1966}    & 0.122               & 0.174               & \ul{\textbf{0.292}} \\
+\citealt{Albright2003a}  & 0.011               & \ul{\textbf{0.246}} & 0.203               \\
+\citealt{Albright2007}   & \ul{\textbf{0.440}} & \ul{0.340}          & 0.016               \\ 
+\bottomrule
+Kendall $\tau_b$            & MaxEnt              & density             & bigram \\
+\midrule
+\citealt{Greenberg1964a} & 0.135               & 0.077               & \textbf{0.205}      \\
+\citealt{Scholes1966}    & 0.098               & 0.124               & \ul{\textbf{0.206}} \\
+\citealt{Albright2003a}  & 0.005               & \ul{\textbf{0.157}} & 0.126               \\
+\citealt{Albright2007}   & \ul{\textbf{0.340}} & 0.215               & $-$0.017            \\
+\bottomrule
+Goodman-Kruskal $\gamma$    & MaxEnt              & density             & bigram \\
+\midrule
+\citealt{Greenberg1964a} & \ul{0.158}          & 0.077               & \ul{\textbf{0.205}} \\
+\citealt{Scholes1966}    & 0.105               & 0.137               & \ul{\textbf{0.214}} \\
+\citealt{Albright2003a}  & 0.010               & \ul{\textbf{0.162}} & \ul{0.127}          \\
+\citealt{Albright2007}   & \ul{\textbf{0.376}} & 0.232               & 0.017               \\
+\bottomrule
+\end{tabular}
 
-\ex $\displaystyle Z_i = \frac{2 n_i}{r_{i + 1} - r_{i - 1}}$ \xe
+\caption{Here is my caption; underlined values are significant at $\alpha = 0.05$}
+\end{table}
 
-\noindent
-\citeauthor{Church1991} do not define this transform for the lowest and highest points (i.e., when $i = 1$ or $i = N$), and so I extend the definition by scaling these points using only one inward-facing point.
+%\citet{Coltheart1977}
+%\citet{Hayes2008a}
+%\citet{Albright2009a}
 
-\ex $\displaystyle Z_1 = \frac{n_1}{r_2 - r_1}$ \xe
-\ex $\displaystyle Z_N = \frac{n_N}{r_N - r_{N - 1}}$ \xe
+% stats thereof
+%\citet{Schilling2002}
+%\citet{Helguerro1904}
+%\citet{Cohen1956}
 
-%\indent
-%The effect of this transform can be seen in Figure \ref{zr}.
+% EM
+\citet{EM}
 
-%\begin{figure}
-%\centering
-%\includegraphics{zr.pdf}
-%\caption{The SUBTLEX word frequency norms \citep{Brysbaert2009} show a long right tail in the log-frequency/log-rank space (left panel), and the $Z_r$ transform (right panel) smooths out this tail.}
-%\label{zr}
-%\end{figure}
+\footnote{In performing this evaluation, I have benefitted greatly from course notes by Mark Liberman and Stephen Isard.}
+%available at \url{http://www.ling.upenn.edu/courses/cogs501/K-meansHW.html}
 
-\subsubsection{Coda, onset, and cluster sparsity}
+% -2 log like ratio test
 
-A number of previous studies \citep[e.g.,][]{Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} have observed that phonemes and graphemes exhibit sparse type (i.e., lexical) and token frequency distributions.
-The lexical frequencies of the medial codas ond onsets which make up syllable contact clusters show a similar distribution.
-In Figure \ref{codaonset}, $Z_r$-transformed lexical frequencies of codas and onsets are plot against rank, in log-log spaces.
-The near-linear relationship that obtains indicates that the data can be modeled by a generalization of Zipf's law \citep{Zipf1949}.
+where the difference in degrees o freedom equals $D$
+
+%\ex $\displaystyle \textrm{Reject } H_0 \textrm{ iff } 
+%\ex $\displaystyle \Lambda = -2 \textrm{ ln } L_0 + 2 \textrm{ ln } L_1}$ \xe
+% > χ^2_{d.f.=D}$ \xe
+
+\subsubsection{Response distributions}
+
+This is shown in Figure \ref{density}.
 
@@
 \centering
-\includegraphics{coda.pdf}
-\includegraphics{onset.pdf}
-\caption{Medial coda and medial onset lexical frequency exhibit the Zipfian log-log-linear relationship between rank and frequency.}
-\label{codaonset}
+\includegraphics{density.pdf}
+\caption{Density of responses}
+\label{density}
 \end{figure}
 
-\ex $\displaystyle f(r~; C, \alpha) = \frac{C}{r^\alpha}$ \xe
+\begin{example}
+EM test results:
 
-\noindent
-The parameters $C$, a constant sensitive to sample size, and $\alpha$, the slope of the rank/frequency relationship in log-log space, are easily computed using the method of least squares in a linear regression, where $\epsilon$ represents the error term.
+\vspace{0.5\baselineskip}
+\begin{tabular}{l | r r | r r r r r r | r r}
+\toprule
+     & \multicolumn{2}{c|}{one $\mathcal{N}$} & \multicolumn{6}{c|}{mixture of two $\mathcal{N}$s} & \multicolumn{2}{c}{log $\mathcal{L}$ test} \\
+     & $\mu$ & $\sigma$ & $\mu_1$ & $\sigma_1$ & mix$_1$ & $\mu_2$ & $\sigma_2$ & mix$_2$ & $\Lambda$ & $p$-value  \\
+\midrule
+G\&J & 0.378 & 0.295    & 0.150   & 0.090      & 0.436     & 0.554   & 0.278    & 0.564     & 5.190    & 0.158   \\
+S    & 0.596 & 0.361    & 0.236   & 0.232      & 0.437     & 0.876   & 0.105    & 0.563 & 48.965    & 1.3\e{-10} \\
+A\&H & 0.431 & 0.266    & 0.301   & 0.200      & 0.559     & 0.595   & 0.246    & 0.411 & 16.186    & 0.001      \\
+\bottomrule
+\end{tabular}
+\end{example}
 
-\ex $\displaystyle \textrm{log}~Z_r \sim C + \alpha~\textrm{log}~r + \epsilon$ \xe
+%\ex Residualized correlations
+%\begin{tabular}
+%\toprule
+%ASDF
+%\midrule
+%G\&J & 
+%S    & 
+%A\&H & 
+%\bottomrule
+%\xe
 
-\noindent
-Codas ($\alpha = -0.720$, $R^2 = 0.749$) and onsets ($\alpha = -1.056$, $R^2 = 0.828$) are well-fit by this distribution, though there are some outliers.
-Similarly, clusters, shown in Figure \ref{clus}, are also a good fit to the generalized Zipf's law ($\alpha = -0.588$, $R^2 = 0.924$).
+\subsubsection{Residual correlations}
 
-\begin{figure}
-\centering
-\includegraphics{cluster.pdf}
-\caption{Syllable contact clusters, themselves composed of Zipfian medial codas and medial onsets, also exhibit a Zipfian distribution.}
-\label{clus}
-\end{figure}
+\subsection{Evaluating ?}
 
-That clusters (and the codas and onsets that make them up conforms to Zipf's law is not itself a deep observation.
-Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also $n$-grams \citep{Baroni2009} and of randomly-generated texts \citep{Miller1957,Li1992}.
-The import of this statistical property for the study of cluster phonotactics is that it makes it difficult, on statistical grounds alone, to determine which unobserved events are excluded, and which are accidentally missing.
+\begin{tabular}
+\toprule
+                         & accuracy & precision & recall & F$_1$ & MCC   \\
+\midrule
+\citealt{Greenberg1964a} & 0.846    & 1.000     & 0.800  & 0.889 & 0.693 \\
+\citealt{Scholes1966}    & 0.952    & 0.947     & 0.900  & 0.923 & 0.889 \\
+\citealt{Albright2003a}  & 0.920    & 0.938     & 0.953  & 0.946 & 0.797 \\
+\citealt{Alrirght2007}   & 0.925    & 0.923     & 0.857  & 0.889 & 0.834 \\
+\bottomrule
+\end{tabular}
 
-\citet{Good1953} puts forth a simple theory of just how frequent unobserved events might be.
-He proposes to estimate $\hat{p}_0$, the probability of all unseen events, as the ratio of events which occur just once $n_1$, to the number of observations, $\sum N$.
-This quantity is known as the Good-Turing estimate, since it was originally developed by Alan Turing for codebreaking during the second World War.
-
-\ex $\displaystyle \hat{p}_0 = \frac{n_1}{\sum N}$ \xe
-
-\noindent
-In the CELEX data, $65$ clusters occur only once, and there are 873 clusters in all, so $\hat{p}_0 = 0.074$.
-One way to interpret this quantity is with respect to a ``replication''.
-If it were possible to generate a new corpus of syllable contact clusters, approximately 7\% of cluster tokens will be ones that were not attested in the previous sample, i.e., they accidentally failed to be sampled.
-This is a not-inconsiderable quantity.
-
-Unfortunately, the Good-Turing estimate does not provide a way to determine which unattested clusters might be found in a replication.
-But I have already proposed that the null hypothesis of generative phonology, combined with the principle of Stampean occultation, provides a way out of this conundrum.
-Unattested clusters like *[m.kl] are structurally excluded (since such a cluster is occulted by \textsc{Coda Nasal Place Assimilation}), whereas *[s.l] is probably an accidental gap.
-
-\subsubsection{Sampling simulation}
-
-As a final investigation into this data, I consider the statistical properties of a randomly generated novel ``lexicon''.
-The simulated lexicon is generated by  generating a distribution of clusters, by repeatedly applying the following procedure.
-
-\ex {Simulation procedure}: \\
-    \begin{tabular}{l l}
-    a. & Sample a medial coda according to the observed probabilities  \\
-    b. & Sample a medial onset according to the observed probabilities \\
-    c. & Apply the SPE rules to the resulting cluster                  \\
-    \end{tabular}
-\xe
-
-\noindent
-This is a simulation version of the model proposed by \citet{Pierrehumbert1994} and \citet{Coleman1997}, but also takes occulting phonology into account.
-
-While this procedure can be repeated indefinitely (using the CELEX data, listed in Appendix \ref{A}), consideration of a single characteristic run of the simulation should suffice.
-The saturation rate of the simulated sample (18.2\%) is very close to the real sample (18.8\%).
-The slope of the log-rank/log-frequency relationship for the simulated data ($\alpha = -0.557$) is also very close to the observed slope for the CELEX data ($\alpha = -0.588$).
-And, as shown in Figure \ref{sim}, the two distributions are nearly indistinguishable.
-
-\begin{figure}
-\centering
-\includegraphics{sim.pdf}
-\caption{A comparison of the observed lexical frequencies of English syllable contact clusters to an ``English lexicon'' simulated by combining.}
-\label{sim}
-\end{figure}
-
-It is important to note that the results of this simulation should not be construed as an endorsement of a revised version of the \citet{Pierrehumbert1994} and \citet{Coleman1997} model which also incorporates occultation.
-In fact, as was shown above, the independent frequencies of coda and onset are very poor predictors of what clusters will be attested even when occultation is taken into account, and the set of attested clusters only partially overlaps those produced by the simulation.
-I propose that this is because the lexicon is not only finite, but far smaller than the combinatoric possibilities provided by phoneme sequences, making accidental gaps unavoidable.
-That these gaps will be arbitary from the perspective of phonology is nothing more than a corrolary of the principle of \emph{l'arbitraire du signe}.
-
-It is quite possible that phonology (and, perhaps, static phonotactics) acts as a filter on what URs are learned, but a cluster might be overrepresented or underrepresented for reasons that are quite external to phonology, e.g., language contact or historical change.
-Similarly, the factors that cause a /n/ to be a common word-medial coda in English may be independent of the propensity of that coda to combine with medial onset, assuming the resulting cluster is phonologically licit.
-The import of this simulation is somewhat more subtle; it simply shows free combination of codas and onsets according to their independent probabilities, constrained only by occultation, would produce the same sparse distribution of clusters that was used to argue for static phonotactics.
-
-%\citet{Evert2004}
-%\includegraphics{zipf.pdf}
+One could imagine a case in which unattested things might be actually less marked than certain attested, but low frequency clusters.

### 2012-05-13 2.3.tex
include at the chapter level

@@
 It is not the case that this evidence is inconsistent with the hypothesis that there are additional degrees of wordlikeness: it is just a demonstration that these data to not require it.
+
+\subsection{Future directions}
+
+Imagine a run of a wordlikeness experiment in which a subject rates the nonce word \emph{dresp} higher than the nonce word \emph{srest}. This observation should indicates the need for many more.
+
+Did the subject faithfully perceive the unattested [sr] onset of \emph{srest}, or was it perceived as, e.g., [ʃr]?
+Does the subject have intuitions about how to ``correct'' \emph{srest} as to make it more English-like?\footnote{This appears to be a prediction of constraint-based theories of phonotactic grammaticality.
+Thanks to Colin Wilson (p.c.) for some discussion on this point.} Does the subject rate the relative wordlikeness of \emph{dresp} and \emph{srest} the same a day, a week, a month, a year later?
+%Are relative judgements ``commutative'' in the sense of \citep{Sprouse2011}, so that if another word is \emph
+Do other subjects agree with this relative judgement?
+And, if not, does this contrast correlate with other subject-specific properties such as age, level of education, exposure to foreign languages, and so on?

### 2012-05-13 2.tex
more writing
include at the chapter level

@@
 
+%This has
+%I am unaware of any study which empiritically
+
+\begin{quote}
+A defect of current grammatical acounts of phonotactics is that they render simple up-or-down- decisions concerning well-formedness and cannot account for gradient judgements.
+But when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values.
+\citep[][371]{Shademan2006}
+
+In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale\ldots{}we consider the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models.
+\citep[][382]{Hayes2008a}
+
+%Gradient intuition can be found even among forms that satisfy the categorical phonotactics of the language, but contain rare sequences\ldots{}
+
+When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing.
+In the usual case, novel items fall along a gradient cline of acceptability.
+\citep[][9]{Albright2009a}
+\end{quote}
+
 (though few if any phonologists regard $n$-gram segment probabilities as a core component of phonology)
 
-%\begin{quotation}
+%\begin{quote}
 %``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
-%\end{quotation}
+%\end{quote}
 

### 2012-05-13 3.1.3.tex
include at the chapter level

@@
 \begin{example}
-\textsc{Labial Attraction} (after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}):
+\textsc{Labial Attraction}: %(after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}):
 
@@
 \txt{[$-$\textsc{Round}]} &                                         & \txt{[$-$\textsc{High}]} & \txt{[$+$\textsc{Labial}]}    & \txt{[$+$\textsc{High}]}\ar@{-}[dr] &         & \txt{[$+$\textsc{Round}]}\ar@{--}[dl] \\
+%                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C}\ar@{-}[u] &                                      & \txt{V} & \\
                          & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C$_0$ C C$_0$}\ar@{-}[u] &                                      & \txt{V} & \\
@@
 
-The formalization of this rule is naturally complex, and perhaps obscures the fact that \textsc{Labial Attraction} generates exceptions to \textsc{Roundness Harmony}; it produces \emph{aCu} sequences (e.g., \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', and \emph{samur} `sable' \citep[][285]{Lees1966b}) instead of the otherwise expected \emph{aCı}.
-\citeauthor{Lees1966b} notes the existence of exceptions (e.g., \emph{tavır} `mode') but writes that they are they are ``surprisingly rare'' (ibid., 286); \citet[][225]{Clements1982} note additional exceptions.
+The formalization of this rule is naturally complex, and perhaps obscures the fact that \textsc{Labial Attraction} generates exceptions to \textsc{Roundness Harmony}, producing \emph{aCu} sequences (e.g., \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', \emph{samur} `sable'; \citealp[285]{Lees1966b}) rather than the expected \emph{aCı}.
+\citet[286]{Lees1966b} and \citet[311]{Zimmer1969} note the existence of exceptions (e.g., \emph{tavır} `mode') but both agree that they are surprisingly rare.
+
+\citet[225]{Clements1982} and \citet{Inkelas2001}
 

### 2012-05-13 3.2.2.tex
include at the chapter level

@@
 clements
-\citet{Ito1993} argue that the data shows that it holds of native words only
+\citet{Ito1993} 
+%and \citet{NiChiosain1993} 
+argue that the data shows that it holds of native words only
+
 \citet{Inkelas2001} refute this point
@@
 \end{quote}
+
+\noindent \citeauthor{Inkelas2001} persistently refer to ``statistical support'', but in fact their study contains no statistics (that is, quantities calculated from a set of data) beyond counts and percentages.
 

### 2012-05-13 3.tex
include at the chapter level

@@
+\chapter{Static and derived phonotactic constraints in Turkish} 
 \label{turkish}
 
-%In this chapter, I present evidence that speakers 
-%glossematics
-%\citet{Chomsky1965} propose
+%This chapter attempts to derive which phonotactic judgemetns are internalized by
 
-%In retrospect, it seems somewhat odd that \citeauthor{Chomsky1965} would coh
-%The contrast between \emph{blick} and \emph{bnick} is what has been called a ``projection problem'' \citep{Baker1979}
-%This is not a universal, but rather something English speakers must learn: Moroccan Arabic, for instance, does permit \emph{bn} onsets (e.g., \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet').
+\citet{Chomsky1965} argue that speakers internalize not only generalizations governing phonological alternations in their language, but also about the shapes of possible words in their language.
+It is certainly possible to imagine that speakers would not bother at all.
+\citet[][320]{Zimmer1969} claims these generalizations ``have no observable consequences in the course of the normal use of the language'' and a similar sentiment is echoed more recently by \citeauthor{PE}:
+
+\begin{quote}
+Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for.
+\citep[][18]{PE}
+\end{quote}
+
+This is a principled null hypothesis, but one which can be quickly rejected.
+\citeauthor{Chomsky1965} point to robust contrasts between equally unattested items like *\emph{blick}, which is recognized by speakers a possible word of English, and *\emph{bnick}, which is not.\footnote{Any readers who are skeptical that \citeauthor{Chomsky1965}'s intuitions generalize will find support for their claim in the wordlikeness judgement data reported in Appendix \ref{albright}.}
+%As a quick validation of this claim, consider the results reported by \citet{Albright2007} and discussed in the previous chapter.
+English speakers asked to rate non-words on a scale from 1-7, where 7 indicates ``most like English'', favor non-words [blʌs, blɑd] (average scores 4.67 and 5.13) over [bnʌs, bnɑd] (average scores 2.06, 2.00).
+This preference must be regarded as a learned behavior, since many languages permit  \emph{bn} onsets (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet').
+It is also possible to show speakers access this knowledge not just in wordlikeness judgement tasks but also in naturalistic and quotidien tasks like segmenting spontaneous speech into words, as \citet{McQueen1998} and \citet{Mattys2001b} argue for adult and infant subjects, respectively.
+
+\citeauthor{Chomsky1965} were not the first to take an interest in characterizing the lexicon, which was also a major concern for the scholars of the Copenhagen school.
+In two sophisticated studies dealing with possible and impossible onsets in Georgian, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} recognize the inherent difficulty of distinguishing between structural and accidental phonotactic gaps in the lexicon on the basis of statistical criteria alone.
+What distinguishes the early generativist approach to phonotactics is the mentalistic concern with what generalizations speakers have internalized.
+By grounding the study of phonotactics in the minds of speakers, it is possible to give a precise definition to the previously vague notions; an accidental phonotactic gap is simply one which speakers do not attend to or internalize.
+Yet few studies of phonotactics are explicitly concerned with what speakers have internalized, despite the increasing sophistication of the phonological representations used.
+In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique, a chi-square test, in an attempt to validate a generalization about Arabic root consonant co-occurrence faces originally due to \citet{Greenberg1950}.
+While this particular generalization was ultimately also evaluated using a wordlikeness task \citep{Frisch2004}, there is now an extensive literature in which phonotactic knowledge is inferred directly from lexical statistics, with no psycholinguistic component of any kind \citep[e.g.,][]{Mester1988,Padgett1992,Padgett1995,Berkley1994b,Berkley1994a,Berkley2000,Buckley1997,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Pozdniakov2007,Anttila2008,Coetzee2008a,Dmitrieva2008a,Dmitrieva2008b,Graff2011}.
+The (usually tacit) assumption of most of this work is made explicit by \citet{Brown2010} in his study of consonant cluster co-occurrence restrictions in Gitksan: any statistically significant pattern in the lexicon is one that is internalized by speakers.
+
+\begin{quote}
+\ldots{}the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
+\end{quote}
+
+In the loanword adaptation literature, it has recently been recognized that not all statistically reliable generalizations about the native lexicon take part in nativization of borrowings \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}.
+In this chapter, data from Turkish is used to show that a widely reported, statistically reliable phonotactic generalization (\textsc{Labial Attraction}) has no effect on wordlikeness judgements.
+This undermines the widespread use of statistic criteria to determine which of the vast number of the overlapping, competing phonotactic generalizations are internalized by speakers and which are ignored.
+Comparison of \textsc{Labial Attraction} to related constraints in Turkish suggests a novel necessary condition for identifying psychologically real phonotactic constraints: namely, speakers attend to phonotactic generalizations only when they are derived from phonological processes (including syllabification).
+
+There are a number of precedents for
+
+This hypothesis was anticipated by \cite[][283]{Anderson1974}.
+
+\begin{quote}
+It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
+\citep[][283]{Anderson1974}
+\end{quote}
+
+This hypothesis is not ahdered to by early research on sequence structure constraints \citep[e.g.,][]{SPR,Chomsky1965,SPE,Stanley1967}.
+
+Work in this vein sought to use sequence structure constraints to compress the lexicon as much as possible
+
+This work has a second interest, namely in ``compressing'' the lexicon as much as possible,
+
+It is important to know that this wor It is important to know that this wor It is important to know that this wor It is important to know that this work
+
+% motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
+\citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible.
+\emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy.
+In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities
+
+This condition (henceforth, the \textsc{derived} condition) 
+On a historical note, this hypothesis is more constrained than what was proposed
+such constraints must be derived from phonological processes in the language in question
+synchronically motivated phonological processes such as alternation or syllabification.
+
+%It is important to note that the nature of the \emph{blick}/\emph{bnick} contrast is no longer a mystery. 
+%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
+It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails.
+If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
+
+%\footnote{I note a potential problem with this account: \citet{Davidson2005,Davidson2006a} argue that this ``excrescent'' schwa is distinct from the both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.}
+
+%This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
+Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
+This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.
+
 %\citet{Lees1966b,Lees1966a}
-
-%\citet{Davidson2005,Davidson2006a} shows that this ``excrescent'' schwa is distinct both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.
-%\citet{Wilson2011} show that speakers use this
 
@@
 %Elliot Moreton (p.c.)
-
-% motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
-\citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible.
-\emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy.
-In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities \citep[][283f.]{Anderson1974}.
-
-%This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
-Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
-This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.
-
-%a principle which is one of the two motivations  motivated the introduction of morpheme structure constraints.
-
-%In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon does not always require static constraints on possible underlying representations, since such distributions arise by chance and static constraints defined over natural classes contribute little to predicting the contents of the lexicon.
-%This result does not directly bear on the question of whether static restrictions have any synchronic reality.
-
-%Yet countless static constraints are statistically supported.
-Under the null hypothesis advanced in Chapter \ref{MSCs}, the lexical statistics of co-occurrence restrictions are irrelevant: the language acquisition device only attends to derived restrictions and ignores static restrictions regardless of their statistical reliability.
-On the other hand, many linguists assume that statistical significance provides sufficient conditions for positing synchronically active static constraints.
-For instance, \citet{Brown2010} writes:
-
-%\begin{quotation}
-%\ldots the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
-%\end{quotation}
 
@@
 
-%It is important to note that the nature of the \emph{blick}/\emph{bnick} contrast is no longer a mystery. 
-%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
-It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails.
-If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
-
-\input{3.1}
-\input{3.2}
-\input{3.3}
+%\input{3.1}
+%\input{3.2}
+%\input{3.3}

### 2012-05-13 4.1.2.tex
include at the chapter level

@@
 If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables.
-For this reason, it is assumed here that affricates are single segments for the purposes of syllabification.
-This assumption is motivated by the fact that affricates pattern with simple segments in many languages; for instance, the complex onsets in Classical Nahua are affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.
+For this reason, an additional constraint is assumed which prevents the two components of the affricates from being split by syllabification.
+This is motivated by the tendency of affricates to pattern with single segments in many languages.
+For instance, the only complex onsets in Classical Nahua are the affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.
 

### 2012-05-13 4.1.3.tex
include at the chapter level

@@
 
-To identify the components of syllable contact clusters, the syllables identified by the above procedure must be further separated into onset, nucleus, and coda, and mapped to underlying representation.
-This is trivial in most cases: the nucleus is easily separated from onset and coda and surface consonants are faithful to underlying representation.
-The treatment of the few difficult cases is described below.
-Ambiguous segments are assigned to onset, nucleus, or coda, and to underlying form, using a distributional heuristic.
+To identify the inventory of syllable contact clusters, it is also necessary to parse syllables into onset, nucleus, and coda.
+In most cases this is trivial, but a few distributional heuristics for difficult cases are described below.
 
@@
 There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right, as demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}.
-The strongest piece of evidence for treating the velar nasal as a pure allophone is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
+The strongest piece of evidence for treating the velar nasal as a pure allophone is the general absence of ?
+nset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
 \citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here.
@@
 The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants.
-The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor homponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
+The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor} [h]\emph{omponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
 Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.
 
-\begin{example}
-\label{VBR}
-VBR's complex onsets \citep[79--80, his transcriptions]{Buchwald2005}:
+\begin{example}[VBR's complex onsets (\citealp{Buchwald2005}:79--80, his transcriptions)]
+%\protect\citep[79--80, his transcriptions]{Buchwald2005}]
 
-\vspace{0.5\baselineskip}
 \begin{tabular}{l l l}
@@
 \end{tabular}
+\label{VBR}
 \end{example}

### 2012-05-13 4.3.tex
include at the chapter level

@@
 
-\begin{example}
-Tigrinya roots with identical first and second consonants \citep{Buckley1990a}: \\
-\begin{tabular}{l l l l}
-a. & lʌlʌw     & `scorch'                   & (< Ge'ez \emph{lʌwlʌw} `inflame')     \\
-   & mʌmʌy     & `winnow'                   & (< Ge'ez \emph{mʌymʌy} `distinguish') \\
+%\begin{example}
+%Tigrinya roots with identical first and second consonants \citep{Buckley1990a}:
+%\begin{tabular}{l l l l}
+%a. & lʌlʌw     & `scorch'                   & (< Ge'ez \emph{lʌwlʌw} `inflame')     \\
+%   & mʌmʌy     & `winnow'                   & (< Ge'ez \emph{mʌymʌy} `distinguish') \\
 %   & mʌmʌt & `pick out loot' & (< 
-b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
-   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
-c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
-   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
-\end{tabular}
-\end{example}
+%b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
+%   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
+%c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
+%   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
+%\end{tabular}
+%\end{example}
 

### 2012-05-13 4.tex
more writing
include at the chapter level

@@
-\chapter{Accidental and structural gaps in English syllable contact}
+%\chapter{Accidental and structural gaps in English syllable contact\protect\footnotemark{*}\protect\footnotetext{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics.}}
+\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
 \label{clusters}
@@
 
-\begin{example}
-The null hypothesis: \\
+\begin{example}[null hypothesis]
 Universal grammar does not countenance constraints on the contents of URs
@@
 
-\begin{quotation}
-\ldots the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
+\begin{quote}
+\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
 \citep[][16]{Fischer-Jorgensen1952}
-\end{quotation}
+\end{quote}
 
@@
 
-\begin{quotation}
+\begin{quote}
 Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters.
@@
 \citep[][30]{Vogt1954}
-\end{quotation}
+\end{quote}
 

### 2012-05-24 1.tex
More arranging.
include at the chapter level
back to chapter files, lots of writing

@@
-\chapter{Morpheme structure constraints in generative grammar}
+\chapter{Sequence structure constraints in generative grammar}
 \label{msc}
@@
 
-\section{A brief history of MSCs} \input{1.1}
-\section{The null hypothesis} \input{1.2}
-\section{Outline of Part I} \input{1.3}
+\section{A brief history of MSCs}
+
+\subsection{Structuralism}
+
+Since, with the possible exception of Roman Jakobson (who might be labeled an early generativist), the mentalistic interpretation of structural linguistic analyses was anathema, one must take care to not assume that a description of MSCs is something that these linguists would posit as part of speakers' knowledge of language.
+
+\subsubsection{Morpheme boundaries in phonology}
+
+bloomfield
+
+%While \citeauthor{Bloomfield1930} is known for introducing morphemic structure into phonology, a member of the Prague circle developed an arguably more sophisticated approach only two years later.
+
+\citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture.
+The vast majority of Russian infinitives consist of the stem followed by a theme vowel and /t\pal/.
+The reflexive is marked wih a final /-sə/ which feeds a general process of regressive voicing assimilation.
+In the reflexive infinitive, /\ldots t\pal-s\ldots/ is realized [ts], with no palatalization.
+However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\ref{russian}c)---is preserved in the reflexive imperative.
+
+\begin{example}
+\label{russian}
+Russian reflexives \citep[after][]{Jakobson1932}:\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses an abstract phonemic transcription.
+Thanks to Lev Blumenfeld (p.c.) for help with the transcription of this data.}
+
+\begin{tabular}{l l l l} %\toprule
+   &  infinitive      & imperative \\ %\midrule
+a. & [slav\pal itsə]  & [slaf\pal s\pal ə]  & `be glorious'    \\
+   & [upram\pal itsə] & [upram\pal s\pal ə] & `be stubborn'    \\
+b. & [kras\pal itsə]  & [kras\pal s\pal ə]  & `put on makeup'  \\
+   & [ʒar\pal itsə]   & [ʒar\pal s\pal ə]   & `roast'          \\
+c. & [zəbytsə]        & [zəbut\pal s\pal ə] & `forget' \\ %\bottomrule
+%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
+\end{tabular}
+\end{example}
+
+\noindent The forms in (\ref{russian}c) have different outcomes for their /t\pal-s/ clusters.
+\citeauthor{Jakobson1932} proposes that the preservation of palatalization is a special property of the imperative.
+%A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.
+
+\subsubsection{The morph as a constraint domain}
+
+A corrolary of this hypothesis is that the phoneme sequences found within morphs may be distinct from those which span multiple morphs.
+
+This was apparent to structuralists
+
+the phonologists 
+This is apparent from 
+a discussion of constraints on vowel sequences in Mixteco given by \citet{Pike1947b}.
+It is apparent that \citeauthor{Pike1947b}'s constraints on vowel sequences are not intended to hold across morph boundaries,
+
+according to the phonological analysis of a Mixteco text published a few years earlier \citep{Pike1944}.
+
+\begin{example}
+Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}:
+
+\begin{tabular}{r l l l} %\toprule
+   & MSC & complex exception \\ %\midrule
+%a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
+%b. & *{C}\textipa{@}{C}e & [n\`i-k\`ə bə-de] & `who entered'        \\
+%c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
+%d. & *{C}i{C}e & [te-n\`i-kee-t\`ə] & `and went away'      \\ %\toprule
+%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
+%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
+a. & *{C}a{C}e & [ká\textsuperscript{n}dee] & `kept \ldots inside' \\
+b. & *{C}ə{C}e & [nìk\`əbəde] & `who entered'        \\
+c. & *{C}e{C}i & [teníke\textsuperscript{n}da] & `was walking         \\
+d. & *{C}i{C}e & [tenìkeet\`ə] & `and went away'      \\ 
+\end{tabular}
+\end{example}
+
+\noindent \citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a morpheme discovery routine.
+
+\subsubsection{Biuniqueness and neutralization}
+
+joos
+harris
+chomsky
+
+\subsection{Early generativism}
+
+\subsubsection{\citealt{Stanley1967}}
+
+inventory constraints
+sequence constraints
+
+\subsubsection{Surface constraints and occultation}
+
+shibatani
+stampe
+dell
+
+\emph{A}[mt] `office', \emph{Anbau} `cultivation'
+\emph{ei}[n.g]\emph{reifen} `to intervene',
+
+\subsubsection{The duplication problem}
+
+\citeauthor{Anderson1974} sees the duplication of \textsc{Backness Harmony} as both sequence structure constraint and phonological rule as a foregone conclusion
+.The tendency of sequence structure constraints to duplicate phonological rules was noted by \citet[][401f.]{Stanley1967} and in \emph{SPE} (p.~382) in the discussion of morpheme structure constraints, and is labeled ``the duplication problem'' by \citet[][?]{Kenstowicz1977}.
+
+inventory problems 
+dell, chung, but also problems
+
+conspiracies
+
+\subsection{Autosegmentalism}
+
+also ``supersegmentalism''
+
+The proposals during this period are very diverse.
+
+\subsubsection{Syllable structure}
+
+hooper
+noske
+
+\subsubsection{Morph features}
+
+Leben
+Kaye
+
+\subsubsection{The Obligatory Contour Principle}
+
+\subsection{Classic Optimality Theory}
+
+%(page numbers here are taken from the published version)
+
+\subsubsection{Richness of the Base}
+
+some quotes from OT doc
+\citet{PE}
+\citet{Bye2001}
+
+\subsubsection{Freedom of Analysis}
+
+\citet{Smolensky1996}
+\citet{PE}
+
+\subsubsection{Undominated constraints}
+
+% complaints about maceahern, gallagher
+% bit on loanword adaptation
+
+\subsection{Recent developments}
+
+\subsubsection{Language acquisition}
+% jusczyk forever
+% white et al.
+% early word learning
+% syllables
+
+\subsubsection{Psycholinguistic tasks in adults}
+
+% wordlikeness
+% spotting
+% error correction, etc.
+
+\subsubsection{Non-native word adaptation}
+
+% perceptual factors
+
+\section{The null hypothesis}
+
+\subsection{The empirical proposal}
+
+\subsection{The acquisition principle}
+
+\subsection{Falsifiability}
+
+%\subsection{Universal constraints?}    \input{1.2.4}
+
+\section{Outline of the dissertation}
+
+%Chapter 2
+%Chapter 3
+%Chapter 4

### 2012-05-24 2.tex
back to chapter files, lots of writing

@@
 
-%This has
-%I am unaware of any study which empiritically
+\citet{Chomsky1965} observe that speakers internalize not only generalizations governing phonological alternations in their language, but also about the shapes of possible words in their language.
+\citeauthor{Chomsky1965} illustrate this latter type of knowledge with two nonce words *\emph{blick} [blɪk] and *\emph{bnick} [bnɪk]; whereas \emph{blick} is a possible word of English, \emph{bnick} can be immediately recognized not only as unattested but also impossible.
+
+What is peculiar about
+
+% actually a mystery
+% follows from syllabification
+
+The problems of this account have
+
+Richer representations in
+
+the increasingly rich representations in
+
+A number of innovations
+
+The requirement that
+
+Citing the presence of epenthesis in loanword adaptation, \citet{Hooper1973}
+
+%\citet{Hooper1973} is an early proponent of introducing syllable representations into generative phonology; citing loanword adaptation process, she writes they ``show that syllable-formation is productive, and that there is very possibly a rule that inserts V's into syllables in the correct position.'' \citep[][101-102]{Hooper1973}.
+%\citet{Kiparsky1982b} also recognizes that the requirement that outputs be syllabifiable may derive generalizations about the shape of possible morphs.
+% presumably by Stampean occultation %(for which Kiparsky cites \citealt{McCarthy1979b})
+
+%\citet{Ito1989a}
+%\citet{Noske1992}
+%s unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
+
+Others have suggested that cluster simplification might result: \citet[][19f.]{Wolf2009}, for instance, proposes that /bnɪk/ would be realized [nɪk].
+
+\citep[e.g.,][]{Pertz1975,Albright2007}
+
+% fails gradience
 
@@
 
-In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale\ldots{}we consider the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models.
+In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
 \citep[][382]{Hayes2008a}
-
-%Gradient intuition can be found even among forms that satisfy the categorical phonotactics of the language, but contain rare sequences\ldots{}
 
@@
 
-(though few if any phonologists regard $n$-gram segment probabilities as a core component of phonology)
+%\ldots{}we consider the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models. \citep[][382]{Hayes2008a}
+%If adhered to literally, however, the gradience hypothesis becomes an axiom.
 
-%\begin{quote}
-%``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
-%\end{quote}
+\section{The gradience hypothesis}
 
-\input{2.1}
-\input{2.2}
-\input{2.3}
+The gradience hypothesis, as defined below, has both weak and strong forms.
+
+% weak version
+% strong version
+\subsection{A brief history}
+\subsubsection{In syntax}
+% LSLT
+% Aspects
+% CH
+% LGB
+\subsubsection{In phonology}
+% SPR
+% GH
+% CH
+% SPE
+% Borowsky
+\subsubsection{The last two decades}
+% OT/HG
+% Coleman 1997
+\subsection{Ontology}
+% intro text
+\subsubsection{Representational likelihood}
+% C&P, positional probs, bigrams
+\subsubsection{Gradient wellformedness}
+% hayes & wilson, anttilla, coetzee & pater
+\subsubsection{Lexical similarity}
+% greenberg & jenkins, ohala & ohala, bailey & hahn
+\subsection{Falsifiability}
+% considers how one might falsify
+\subsubsection{Against an axiomatic interpretation}
+% not an axiom: hayes & wilson
+\subsubsection{The weak hypothesis}
+% \citet{Armstrong1983}
+% schutze quote
+% stuff about mathcog
+\subsubsection{The strong hypothesis}
+% comparison of gradient models
+% incremental contribution of gradient models
+\section{Evaluation}
+
+The strong gradience hypothesis can only be evaluated by comparing individual gradient models to actual wordlikeness data.
+The remainder of this chapter describes such a study using state-of-the-art phonotactic models and data from English.
+
+\subsection{Data sources}
+
+The evaluation makes exclusive use of previously published wordlikeness data sets.
+To be included in this evaluation, a study must have the three following characteristics.
+First, the stimuli must consist only of monomorphs presented auditorily.
+Secondly, a significant portion of the stimuli must contain gross phonotactic violations (e.g., *\emph{bnick}), and finally, the ratings, averaged across subjects, must be publicly available.
+The study by \citet{Bailey2001} which is ineligible because very few stimuli contain phonotactic violations, and the studies by \citet{Shademan2007} is not considered here because the data is not available to the public.
+
+\subsubsection{\citealt{Greenberg1964}}
+
+\citet{Greenberg1964} investigated wordlikeness using the technique of free magnitude estimation, a mechanism which has become increasingly popular among syntactians \citep[e.g.,][]{Bard1996}.
+At the beginning of the experiment, the subject heard a recording of the word \emph{stick}.
+In subsequent trials, the subjects heard a new (possibly nonce) word and is asked to report ``how far would you say that is from English?'', with \emph{stick} at ``1''; subjects are told that a word that is ``twice as far from English'' as \emph{stick} should be scored ``2''.
+The data used here are from \citeauthor{Greenberg1964}'s Experiment B, in which 17 undergraduates were presented 17 stimuli in all.
+In addition to \emph{stick}, several other English words were included in the stimuli; these four items were excluded from further analyses, leaving 13 stimuli.
+As is standard practice in psychophysics \citep[e.g,]{Butler1987}, these magnitude ratings were log-transformed before analysis.
+
+\subsubsection{\citealt{Scholes1966}}
+
+\citet{Scholes1966} conducted a number of English wordlikeness judgement with middle school children.
+The data used here come from his ``experiment 5'', in which 63 monosyllabic items were presented to 33 seventh-grade students.
+For each stimulus, the subjects produced forced choice ``yes''/``no'' answers to the question of whether the item ``is likely to be usable as a word of English''.
+\citet{Hayes2008a} and \citet{Albright2009a} analyse this data as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus.
+For instance, 22 of the 33 students answered ``yes'' for \emph{shlerk} [ʃlɚk], so it is assigned a score of $0.666$.
+These stimuli are all ostensibly non-words, but include \emph{clung} [klʌŋ], the preterite and past participle of the verb \emph{cling}, and \emph{brung} [brʌŋ], a dialectical past participle for \emph{bring}.
+These two words were excluded and the remaining 61 stimuli were submitted to analysis.
+
+\subsubsection{\citealt{Albright2003b}}
+
+\citet{Albright2003b} gathered wordlikeness judgements to serve as norms for a \emph{wug}-test.
+87 items were presented to 20 undergraduate subjects, who rated each word on a seven-point Likert scale.
+The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``complete normal, would make a fine English word''.
+
+\subsection{Model comparison}
+
+These models are compared against wordlikeness ratings using non-parametric rank correlation statistics.
+While the use of the Pearson correlation is standard in the phonotactics literature, rank correlation statistics make none of the troublesome assumptions associated with this former method.
+The Spearman $\rho$ is most widely known rank correlation statistic, but the Kendall $\tau_b$ and Goodman-Kruskal $\gamma$ have much simpler interpretations, relating to the probabilities of {concordant} and {discordant} pairs \citep{Noether1981}.
+Consider the case here, in which model scores are compared with wordlikeness ratings.
+If a model predicts that *\emph{dresp} [?] is more highly rated than *\emph{srest}, this pair is {concordant} if speakers also rate *\emph{dresp} more English-like than *\emph{srest} and it is discordant if *\emph{srest} is rated more English-like.
+Both $\tau_b$ and $\gamma$ are based on ratios of the number of concordant and discordant pairs, differing only in their treatment of ``ties'' (in which, e.g., *\emph{dresp} and *\emph{srest} are scored the same, or rated the same); $\tau_b$ includes a correction for ties, whereas $\gamma$ discards tied pairs.
+Much like the familiar Pearson correlation, $\rho$, $\tau_b$ and $\gamma$ are all in the range [-1, 1].
+The correlations of the four models, discussed below, are given in Table \ref{cor}.
+
+\begin{table} \label{cor}
+\centering
+\begin{tabular}{l l l l l}
+\toprule
+Spearman $\rho$          & binary           & MaxEnt  & bigram           & density          \\
+\midrule
+\citealt{Greenberg1964}  & {0.845}          & {0.765} & {\textbf{0.863}} & {0.648}          \\
+\citealt{Scholes1966}    & {0.791}          & {0.762} & {\textbf{0.827}} & {0.827}          \\
+\citealt{Albright2003b}  & {0.725}          & {0.429} & {0.708}          & {\textbf{0.742}} \\
+\midrule
+Kendall $\tau_b$         & binary           & MaxEnt  & bigram           & density          \\
+\midrule
+\citealt{Greenberg1964}  & {\textbf{0.716}} & {0.585} & {0.692}          & {0.462}          \\
+\citealt{Scholes1966}    & {\textbf{0.664}} & {0.597} & {0.652}          & {0.565}          \\
+\citealt{Albright2003b}  & {\textbf{0.599}} & {0.343} & {0.506}          & {0.556}          \\
+\midrule
+Goodman-Kruskal $\gamma$ & bigram           & MaxEnt  & bigram           & density          \\
+\midrule
+\citealt{Greenberg1964}  & {\textbf{1.000}} & {0.684} & {0.692}          & {0.462}          \\
+\citealt{Scholes1966}    & {\textbf{0.995}} & {0.634} & {0.667}          & {0.614}          \\
+\citealt{Albright2003b}  & {\textbf{0.953}} & {0.656} & {0.509}          & {0.575}          \\
+\bottomrule
+\end{tabular}
+\caption{Rank correlations between wordlikeness ratings and phonotactic models surprisingly reveal that the binary baseline meeds or exceeds the coverage of three state-of-the-art phonotactic models.
+All correlations are significant at $p = 0.05$.}
+\end{table}
+
+\subsubsection{Binary baseline}
+
+The binary baseline here corrresponds to the null hypothesis that there are no gradient effects in wordlikeness judgements.
+To create such a baseline, it is necessary to distinguish between nonce words which contain a gross phonotactic violation and those which do not.
+Since all the observed phonotactic violations in the data can either be localized to the onset or rime, this problem reduces to the problem of determining whether a nonce word consists of a valid onset and rime.
+Rather than making this partition in a way based on any theoretical insights, the binary baseline is further ``cursed'' with the inability to abstract beyond seen data: any word which contains an unattested onset or rime is marked invalid.
+Despite the simplistic nature of this model, it outperforms the gradient models described below in most contexts.
+Further comparison of this model and possible improvements are sketched in \S\ref{limitations}.
+
+\subsubsection{Maximum Entropy phonotactics}
+
+\citeauthor{Hayes2008a} (\citeyear{Hayes2008a}; henceforth H\&W) develop a sophisticated model of phonotactic grammaticality which estimates a probability distribution over phoneme sequences by weighing constraints according to the principle of maximum entropy.
+H\&W find that the predictions of their model are closely correlated with the \citet{Scholes1966} wordlikeness ratings.
+A direct replication of this evaluation was attempted using the software, the model parameters, and training data used by H\&W.
+Since the training of model is inherently stochastic, producing slightly different outcomes on each run, the lowest scoring of ten runs is reported (H\&W:396) though in general there is not a great deal of variation between individual runs.
+One limitation of this model is that it is not feasible to score whole words, as the number of constraints which must be inspected grows expontentially as their scope increases.
+Following H\&W and of \citet{Albright2009a}, who also applies the maximum entropy model to the \citet{Albright2003b} norms, the model is trained and scored only on stimulus onsets.
+However, as a consequence, the maximum entropy model performs particularly poorly on this data set, in which many stimuli contain phonotactic violations in the rime.
+
+\subsubsection{Segment bigram probability}
+
+The bigram probability of a sequence $ijk$ is simply the product of three probabilities: the probability of an sequence-initial $i$, the probability that $j$ follows $i$, and the probability that $k$ follows $j$.
+
+\begin{unlabeledexample}
+$\displaystyle \hat{p}(ijk) = p(i|\textrm{start}) \cdot p(j|i) \cdot p(k|j)$
+\end{unlabeledexample}
+
+\noindent Bigram models are widely used in natural language processing, and \citet{Albright2009a} considers their relevance to modeling wordlikeness judgements.
+While the focus of \citeauthor{Albright2009a}'s study is on developing a model which uses bigrams over phonological features rather than segments themselves, \citeauthor{Albright2009a}'s evaluation, which includes both the \citeauthor{Scholes1966} and \citeauthor{Albright2003b} data sets, finds an advantage for segmental bigrams.
+
+\citeauthor{Albright2009a} estimates bigram probabilities using familiar maximum likelihood method over types in the lexicon.
+The variant of segmental bigrams used here computes probabilities with a simple type of smoothing in which the count of every bigram (including those never observed) is incremented by one, a technique known as Laplacian, or ``add one'' smoothing.
+This has the desirable effect that no nonce word is ever assigned a zero probability, and produces in a small increase in the correlation between the \citeauthor{Albright2003b} wordlikeness norms compared with the maximum likelihood estimate (Table \ref{albrightimproved}).
+This model consistenly outperformed the positional probability probability models of \citet{Vitevitch2004} and \citet{Vaden2009}.
+The bigram model consistently performs well in all the evaluations, and has the highest Spearman correlation with the \citeauthor{Greenberg1964} and \citeauthor{Scholes1966}, and is frequently second place model to the binary baseline elsewhere.
+
+\begin{table} \label{albrightimproved}
+\centering
+\begin{tabular}{l r r r}
+\toprule
+                         & maximum likelihood & Laplacian smoothing \\
+\midrule
+Spearman $\rho$          & 0.660              & \textbf{0.708} \\
+Kendall $\tau_b$         & 0.467              & \textbf{0.506} \\
+Goodman-Kruskal $\gamma$ & 0.473              & \textbf{0.509} \\
+\bottomrule
+\end{tabular}
+\caption{Laplacian smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
+All correlations are significant at $p = 0.05$.}
+\end{table}
+
+\subsubsection{Lexical neighbors}
+
+The hypothesis that wordlikeness judgements reflect similarity to known words has a long history and has been promoted by linguists as diverse as \citet{Greenberg1964}, \citet{Ohala1986b}, and \citet{Schutze2005}.
+There are now an extensive number of methods for computing similarity in the psycholinguistic literature.
+The simplest of these methods, neighborhood density \citep{Coltheart1977}, consistenly outperformed recent proposals by \citet{Bailey2001}, \citet{Vaden2009}, and \citet{Suarez2011}.
+\citeauthor{Coltheart1977} define the neighborhood density as the number of real monomorphemic words which can be changed into the target nonce word by a single insertion, deletion, or substition.
+For instance, the neighbors of \emph{blick} include \emph{blink} (insertion), \emph{lick} (deletion) and \emph{black} (substitution).
+While many studies \citep[e.g.,][]{Bailey2001} report robust lexical similarity effects, it may be that these effects are diminished in the presence of gross phonotactic violations, since dnesity performs more poorly than the competing models.
+
+\subsection{Discussion}
+\label{2dis}
+
+\subsubsection{The strong gradience hypothesis}
+
+The evaluation provides overwhelming evidence that the gradient phonotactic models are closely correlated with the binary baseline, but no gradient model is reliably superior to the baseline.
+This is not sufficient, however, to rule out the strong gradience hypothesis, which predicts that even once gross phonotactic violatoins are controlled for, gradient models contribute additional information about speakers' judgements.
+
+As an exploratory analysis, the densities of ratings from the three studies, linearly transformed to the interval [0, 1] and split according to the binary baseline, are shown in Figure \ref{density}.
+In all three studies, it is possible to discern a relatively sharp separation between valid and invalid clusters.
+
+\begin{figure} \centering \label{density}
+\includegraphics{density.pdf}
+\caption{Average ratings of individual nonce words, linearly transformed to the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
+\end{figure}
+
+Figure \ref{density} also reveals intermediate ratings are used.
+The weak gradience hypothesis requires a demonstration that these intermediate ratings are correlated once the categorical effects are taken into account.
+The residual contribution of the three gradient models is quantified in the following manner.
+Instead of calculating rank correlations directly on the model scores as in Table \ref{cor}, the model scores are mapped to ranks with the additional constraint that all ``valid'' stimlui are ranked higher than all ``invalid'' stimuli, and used to compute new correlation statistics.
+To estimate the improvement in correlation generating by augmenting the binary model with gradient predictions, the binary baseline correlation statistics from this number; these differences are given in \ref{controlled}.
+In most cases, including the gradient models on top of the binary baseline results in a reduction of performance in stark contrast to the predictions of the strong gradience hypothesis.
+
+\begin{table} \label{controlled}
+\centering
+\begin{tabular}{l r r r}
+\toprule
+Spearman $\rho$          & MaxEnt            & bigram            & density  \\
+\midrule
+\citealt{Greenberg1964}  & $-0.060$          &  \textbf{0.038} & $-0.017$ \\
+\citealt{Scholes1966}    & $-0.029$          &  \textbf{0.047} & $-0.035$ \\
+\citealt{Albright2003b}  & $-0.008$          & $-0.015$          & \textbf{0.018} \\
+\midrule
+Kendall $\tau_b$         & MaxEnt            & bigram            & density  \\
+\midrule
+\citealt{Greenberg1964}  & $-0.114$          & \textbf{$-$0.007} & $-0.084$ \\
+\citealt{Scholes1966}    & $-0.067$          & \textbf{0.003}  & $-0.061$ \\
+\citealt{Albright2003b}  & \textbf{$-$0.038} & $-0.092$          & $-0.049$ \\
+\midrule
+Goodman-Kruskal $\gamma$ & MaxEnt            & bigram            & density  \\
+\midrule
+\citealt{Greenberg1964}  & $-0.268$          & \textbf{$-$0.260} & $-0.337$ \\
+\citealt{Scholes1966}    & $-0.361$          & \textbf{$-$0.313} & $-0.345$ \\
+\citealt{Albright2003b}  & \textbf{$-$0.137} & $-0.443$          & $-0.386$ \\
+\bottomrule
+\end{tabular}
+\caption{The change in rank correlation generated by augmenting the purely binary model with gradient predictions is small and in many cases it is negative.}
+\end{table}
+
+\subsubsection{Limitations of the baseline}
+\label{limitations}
+
+The most serious limitation of this study is that the binary model, while surprisingly performant, is not intended as a satisfactory model of wordlikeness judgements, merely as a baseline.
+The results suggest the possibility of advancing a serious categorical model of English phonotactics.
+The binary model used here is deficient insofar as it uses attestedness of onsets and rimes as a sole criterion, but there may be unattested rimes which are wellformed, demonstrating the need to abstract beyond mere absense of evidence in distinguishing between possible and impossible clusters.
+For instance, the rime [ɛsp] is unattested, but *\emph{dresp} [drɛsp] receives high wordlikeness ratings \citep{Albright2009a}.
+This problem is more acute in languages with more permissive syllable structures than those of English.
+Discussing the inventory of consonant clusters in Georgian, which admit clusters of as many as five consonants, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory.
+A similar result is reached in Chapter 4, which considers the inventory of syllable contact clusters in English.
+% FIXME
+
+A further extension to categorical models would introduce one or more intermediate, peripheral levels to the current dichotomy.
+While *\emph{zhlick} [ʒlɪk] contains an unattested onset, it seems quite likely that speakers might recognize it as more English-like than *\emph{bnick}.
+A satisfactory model of this type is nothing more than a categorical formalization of phonotactic possibility and peripherality.
+
+\section{Conclusion}
+
+This chapter finds that virtually all of the apparent coverage of state-of-the-art phonotactic models is simply a reflection of their overall ability to distinguish between possible and impossible.
+These state-of-the-art phonotactic models fail to predict intermediate wordlikeness ratings with accuracy.
+
+The results here demonstrate the importance of including simple categorical baselines, which turn the axiom of gradience into a falsifiable alternative hypothesis.
+Any model which cannot outperform this trivial baseline is no better than the list of attested onsets and rimes which make it up.
+Given that the state of the art is inferior to such a list,
+
+one can only conclude that
+there does not yet exist a model which
+
+ there exists no model which can as of yet project for no model can yet project from
+
+, and fails the task of projecting from the observed primary data to speakers' intuitions about possible words in their language.
+
+%\subsection{Future directions}
+%\subsubsection{Interspeaker variation}
+%The studies analyzed above all use ratings averaged across subjects.
+While aggregation of this type is quite common in psycholinguistics, it is undesirable as a statistical practice: it drastically increases the chance of a Type II error, the error of failing to reject a null hypothesis when the null hypothesis is false \citep[][8f.]{Baayen2004}.
+Further, little is known about intraspeaker variation in wordlikeness judgements and even less about intraspeaker variation, but \citet{Shademan2007} does find suggestive differences between older and younger adults.
+It would be extremely interesting to correlate differences in speakers with external factors like age, level of education, and exposure to foreign languages.
+%\subsubsection{Mixed modalities}
+%While the use of auditory stimulus presentation seems preferable to rating orthographic words, the possibiltiy of misperception introduces a potential confound.
+\citet{Scholes1966} mentions that a few of his stimuli were frequently misperceived, and a number of studies \citep{Massaro1983,Halle1998b,Pitt1998,Dupoux1999,Berent2007a,Davidson2007} uncover positive correlations between phonotactic acceptability and accurate perception.
+If the phonotactically invalid cluster in *\emph{srest} is perceived as a phonotactically valid cluster with similar acoustic features, such as [ʃr], it will likely receive an inflated wordlikeness rating.
+In many psycholinguistic tasks, the behavioral measure of interest comes with a validation of correct perception: for instance, in shadowing tasks, reaction times are the primary behavioral measure but shadowing errors are also informative behavioral measures \citep[e.g.,][]{Marslen-Wilson1978}.
+Wordlikeness studies could combine transcription or repetition tasks with judgement tasks to confirm correct perception and perhaps to enhance the online nature of the task.
+%Several recent studies have investigated the role of phonotactics using an auditory lexical decision tasks \citep[e.g.,][]{Berent2001a,Berent2004,Shatzman2007a}.
+A recent study by \citet{Coetzee2008b} is particularly relevant to the larger questions of this dissertation, though the results are ultimately inconclusive.
+\citeauthor{Coetzee2008b} claims that words of the form [spVp, skVk] (e.g., *\emph{spep} [spɛp], *\emph{skeek} [skiːk]) are impossible in English, whereas [stVt] words (e.g., *\emph{stoit} [stɔɪt]) are wellformed.
+In addition to a traditional wordlikeness task, \citet{Coetzee2008b} includes stimuli of this type in a lexical decision task, under the hypothesis that the illformed nonce words, those of the shape [spVp, skVk], will be more quickly rejected than valid [stVt] stimuli, a prediction which is borne out.
+However, there is no need to provide a phonotactic interpretation of these effects.
+It is known that auditory tasks reaction times are significantly slower for non-words with dense cohorts, that is, nonce words like *\emph{hup} [hʌp] which share initial segments with many real words (e.g., \emph{hull}, \emph{hut}, etc.).
+These effects are found across modalities \citep[e.g.,][]{Marslen-Wilson1978,Marslen-Wilson1984} and appear to be particularly attentuated for non-word onsets \citep{Cole1980,Vitevitch2002}.
+Since [st] onsets are far more frequent than [sp] or [sk] onsets \citep[][395]{Hayes2008a}, this alone could explain the variation in reaction times.
+Despite this, nonce word lexical decision with controls for lexical cohort effects shows potential as a substitute for wordlikeness judgements in phonotactics research.
+%\subsubsection{Data availability}
+%There are still very few publicly available wordlikeness databases, and several authors who were contacted for this project declined to share data.
+Not only does this make large-scale model comparison difficult, it also means that most wordlikeness studies cannot be replicated except at the conceptual level.
+Lexical decision researchers now have databases of response times for American \citep{ELP} and British English \citep{BLP}, Dutch \citep{DLP}, French \citep{FLP}, and Malay \citep{MLP}.
+Given the simplicity of wordlikeness judgements (a sort of linguistic \emph{Drosophilia}), the time for a publicly available ``English Wordlikeness Project'' has come.
+%\subsubsection{Non-parametric statistics}
+%The surprising results of this evaluation can be traced directly to two methodological improvements which should be more widely adopted.
+First, model comparison studies must replace parametric statistical models, such as Pearson correlation, with non-parametric equivalents.
+Given the tendency of ratings to cluster at endpoints, Pearson correlations are inflated \citep[][23, fn.
+12]{Albright2009a} and the assumptions of the test are flagrantly violated.
+A shift towards non-parametric comparison appears to be well underway: \citet{Hayes2008a} also report model correlations using the Spearman $\rho$ and \citet{Albright2009a} focuses on a variant of the Kendall correlation called $tau_c$.
+%On the other hand, both \citeauthor{Hayes2008a} and \citeauthor{Albright2009a} ultimately use parametric statistics for model comparison.
+%\begin{table}
+%\centering
+%\begin{tabular}
+%\toprule
+%                        & accuracy & precision & recall & F$_1$ & MCC   \\
+%\midrule
+%\citealt{Greenberg1964} & 0.846    & 1.000     & 0.800  & 0.889 & 0.693 \\
+%\citealt{Scholes1966}   & 0.952    & 0.947     & 0.900  & 0.923 & 0.889 \\
+%\citealt{Albright2003b} & 0.920    & 0.938     & 0.953  & 0.946 & 0.797 \\
+%\bottomrule
+%\end{tabular}
+%\end{table}
+% EM \citet{EM} 
+%\begin{table}
+%\centering
+%\begin{tabular}{l | r r}
+%\toprule
+%                        & $\chi^2_{3}$ & $p$-value  \\
+%\midrule
+%\citealt{Greenberg1964} &  5.190       & 0.158      \\
+%\citealt{Scholes1966}   & 48.964       & 1.3\e{-15} \\
+%\citealt{Albright2003b} & 16.186       & 0.001      \\
+%\bottomrule
+%\end{tabular}
+%\caption{This is my caption}
+%\end{table}
+%\footnote{This is conceptually similar to the technique of residualization discussed by \citet{Gorman2010c}.}
+%\begin{unlabeledexample}
+%$\displaystyle P = s^{1/7.4}$
+%\end{unlabeledexample}
+%\noindent Even then, \citet[][23, fn. 12]{Albright2009a} 
+%An interesting feature of this data set is that includes a diverse set of onsets ranging from frequent to impossible, but only a few common rimes; \citet{Albright2009a} speculates this might have made the onsets more salient to the subjects than tehy would be otherwise.
+%The Pearson correlation depends on an interval assumption, which holds that the difference between, for example, ``1.
+completely bi disagree'' and ``2.
+agree'' is in some sense identical to that between ``3.
+neither agree nor disagree'' and ``4.
+agree''.
+\citet{Stevens1946} already observes that there is no a priori basis for this assumption, and similar objections can be leveled for other rating methods.
+Secondly,
+%the Pearson correlation assumes a linear relationship between model scores and ratings; when this assumption is violated, researchers must search for a satisfactory transformation to produce linearity.

### 2012-05-24 3.tex
back to chapter files, lots of writing

@@
 
-%\input{3.1}
-%\input{3.2}
-%\input{3.3}
+\section{Turkish vowel sequence structure constraints}
+
+%\citet{Lees1966a,Lees1966b} 
+\citet{Lees1966b,Lees1966a} 
+proposes three constraints on Turkish vowel sequences which have been the focus of many subsequent studies.
+In this section, these constraints are formalized, and where possible, related to phonological alternations in Turkish and to evidence that speakers have internalized these restrictions.
+The following feature specification for the Turkish vowels is assumed throughout.
+
+\begin{example}
+Turkish vowel features:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{c | c c c c}
+                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
+                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
+\buf[$+$\textsc{High}] & {i} & {ü} [y] & {ı} [ɯ] & {u} \\
+\buf[$-$\textsc{High}] & {e} & {ö} [ø] & {a} [ɑ] & {o} \\
+\end{tabular}
+\end{example}
+
+\subsection{Backness harmony}
+
+\citeauthor{Lees1966b} (\citeyear[][35]{Lees1966b}, \citeyear[][284]{Lees1966a}) models the Turkish vowel harmony systems by means of three feature spreading rules.
+Backness harmony is the the most general of these processes.
+
+\subsubsection{Phonological description}
+
+\textsc{Backness Harmony} spreads the backness specification of a vowel rightward ignoring any intervening consonants.
+
+%\citep[after][229]{Clements1982}: \\
+\begin{example}
+\textsc{Backness Harmony}:
+
+\xymatrix@R=24pt@C=24pt{
+\txt{V}                                        & \txt{C$_0$~~~~~} & \txt{V} & \txt{(condition: rightward application)} \\
+\txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
+}
+\end{example}
+
+This rule accounts for the tendency of polysyllablic roots to either contain all [$+$\textsc{Back}] or all [$-$\textsc{Back}];
+%, as can be seen in the nominative singulars in (\ref{turknom}b); 
+this tendency is quantified in \S\ref{lexstats} below.
+\textsc{Backness Harmony} also accounts for the allomorphs of several inflectional suffixes.
+For instance, the nominative plural (nom.pl.) suffix is \emph{-ler}, when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].
+
+\begin{example}
+\label{turknom}
+The Turkish nominative:
+
+%\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l}
+   & \emph{nom.sg.} & \emph{nom.pl.} \\
+a. & ip             & ipler          & `rope' & \citep[][216]{Clements1982} \\
+   & el             & eller          & `hand'    \\
+   & köy            & köyler         & `village' \\
+   & yüz            & yüzler         & `face'    \\
+   & kız            & kizlar         & `girl'    \\
+   & sap            & saplar         & `stalk'   \\
+   & son            & sonlar         & `end'     \\
+   & pul            & pullar         & `stamp'   \\
+b. & neden          & nedenler       & `reason'  & (TELL) \\ % front
+   & kiler          & kilerler       & `pantry'           \\ % front
+   & pelür          & pelürler       & `tissue paper'     \\ % back
+   & boğaz          & boğazlar       & `throat'           \\ % back
+   & sapık          & sapıklar       & `pervert'          \\ % back
+\end{tabular}
+\end{example}
+
+There are a few complications, however.
+First, not all polysyllabic roots conform to \textsc{Backness Harmony}.
+In this case, as shown in (\ref{turkexcept}a), suffix vowels generally harmonize with the final root vowel.
+There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} despite the fact that their final root vowel is [$+$\textsc{Back}].
+
+\begin{example}
+\label{turkexcept}
+Exceptional Turkish nominatives:
+
+%\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l}
+   & \emph{nom.sg.} & \emph{nom.pl.} \\
+a. & mezar          & mezarlar       & `grave' & \citep{TELL} \\
+   & model          & modeller       & `model' \\
+   & silah          & silahlar       & `weapon'     \\
+   & memur          & memurlar       & `bureaucrat' \\
+   & sabun          & sabunlar       & `soap'       \\
+b. & saat           & saatler        & `hour, clock' \\
+   & harf           & harfler        & `(alphabetic) letter' \\ %& \citep{Goksel2005}
+   & etol           & etoller        & `fur stole' \\
+\end{tabular}
+\end{example}
+
+While it is uncontroversial that the disharmonic suffixes of (\ref{turkexcept}b) are no more than very sporadic exceptions to \textsc{Backness Harmony}, root disharmony has ben the subject of much debate.
+As disharmonic roots still trigger suffix harmony, \citet[][212, 289]{Anderson1974} and \citet{Iverson1978} propose to separate suffix harmony, an alternation, from a sequence structure constraint governing root harmony.
+
+The disadvantage of this account is that it introduces a sort of ``duplication'' (in the sense of \citealt{Kenstowicz1977}) of sequence structure and phonological generalizations, differning only in their patterns of exceptionality.
+%\citep{Kisseberth1970b} (see also \citealp[][401]{Stanley1967} and \emph{SPE}:382)
+However, \citet[][197f.]{Zonneveld1978} observes that the theory of exceptionality proposed in \emph{SPE}
+% (p.~374f.)
+can account for suffix harmony in disharmonic roots.
+\citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
+A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}].
+In other words, a form is never truly an ``exception'': it simply fails to match an extended structural description including the rule feature.
+If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.\footnote{The definition of ``target of a rule'' is not obvious in autosegmental rules, but any useful formalization of this notion should, for example, identify a harmonic suffix vowel as the target of \textsc{Backness Harmony}.} With this in place, \textsc{Backness Harmony} is sufficient to account for both root and suffix harmony.
+
+%Once additional source of evidence on root (dis)harmony is inconclusive.
+There is a small class of bisyllabic words in which the second vowel, always [$+$\textsc{High}, $-$\textsc{Back}], alternates with zero.
+
+%\ex High-vowel/zero alternations \citep[][243]{Clements1982}: \\
+%\begin{tabular}{l l l l}
+%   & nom.sg. & gen.sg. \\
+%a. & fikir   & fikri  & `idea' \\
+%   & hüküm   & hükmün & `judgement' \\
+%%  & filim   & filmi & `film' & \citep[][178]{Inkelas2001} \\
+%b. & vakit   & vaktin & `time' \\
+%   & rahim   & rahmin & `womb' \\
+%\end{tabular} \xe
+%
+%\noindent
+%It is possible that \textsc{Backness Harmony} might produce a fluctuating \emph{ı} after root \emph{a}, but this does not obtain (\lastx b).
+However, this might simply indicate that the fluctuating vowel is epenthetic and that harmony applies before epenthesis (see \citealt{Clements1982} for both sides of this argument), making it less than a counterexample.
+
+Before moving on, it is necessary to dispense with an alternative analysis originally proposed by \citet{Clements1982} and further developed by \citet{Inkelas1997}.
+Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
+\emph{kul} `servant', \emph{kepek} `bran' vs.
+\emph{kapak} `lid'), whereas harmonic roots are those in which the backness of any remaining vowels is predictable.
+\citeauthor{Clements1982} propose that these vowels are underspecified for backness, whereas the non-initial vowels of disharmonic roots are fully specified.
+This is schematized below.
+
+%This is exemplified below in (\ref{spec}).
+%(e.g., \emph{deve} `camel' vs. \emph{deva} `medicine', \emph{sene} `year' vs. \emph{sena} `praise'). 
+%There is some further evidence that individual vowels may differ in specification for this feature even within individual roots or affixes.
+For instance, the present continuous suffix has harmony-determined allomorphs \emph{-iyor}, \emph{-üyor}, \emph{-ıyor}, \emph{-uyor}, but the \emph{o} of the suffix is invariant.
+A similar situation might obtain in Turkish roots.
+
+\begin{example}
+\label{spec}
+Underspecification in harmonic roots (after \citealp{Clements1982}):
+
+%\vspace{0.5\baselineskip}
+\xymatrix@R=24pt@C=24pt{
+\txt{a.} & \txt{harmonic root:~~~~} & \txt{C} & \txt{V} & \txt{C} & \txt{V} \\
+&   &    & \txt{[$-$\textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
+\txt{b.} & \txt{disharmonic root:} & \txt{C} & \txt{V} & \txt{C} & \txt{V} \\
+    &    &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u] & & \txt{[$+$\textsc{Back}]}\ar@{-}[u]
+}
+\end{example}
+
+Harmonizing suffix vowels will also be underspecified for backness.
+
+Two additional details are needed to complete this analysis.
+First, there must be some kind of constraint which prevents speakers from positing redundant backness specifications for non-initial vowels in harmonic roots, a constraint on identical adjacent specifications in underlying representations, the lexical ``obligatory contour principle'' \citep[][OCP]{Leben1973}, on surface tonal representations \citep{Goldsmith1976}, or both (\citealp{Leben1978}, \citealp{McCarthy1986}; see \citealp{Odden1986,Odden1988} for criticism).
+Second, \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being the use of a structure preservation condition \citep{Kiparsky1985}.
+However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are formally marked in Turkish.\footnote{On the other hand, if one interprets ``harmony'', that is, a single backness specification per root, as the learner's default (as \citealp{Odden1986} proposes for the tonal obligatory contour principle), this account is a mere notational variant of the rule exceptionality account.}
+%\footnote{Kie Zuraw (p.c.) suggests that the markedness of disharmonic roots might be attributed to the presence of more structure, namely, multiple specifications for backness.}
+I reject the underspecification analysis on these grounds. Rather, I assume a unitary phonological rule of \textsc{Backness Harmony} and assign [$-$\textsc{Backness Harmony}] to disharmonic roots.
+
+\subsubsection{External evidence}
+\label{backharmexternal}
+
+While suffix harmony can be inferred from alternations, no evidence has yet been presented to show that speakers internalize in any way tendency for roots to exhibit backness harmony.
+
+%\begin{quote}
+%\ldots{}these generalizations\ldots{}have no observable consequences in the course of the normal use of the language. \citep[][320]{Zimmer1969}
+%\end{quote}
+
+Indeed, \citeauthor{PE} adopt the null hypothesis that speakers do not internalize any generalizations about possible or likely sequences in underlying representations.
+
+\begin{quote}
+Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for.
+\citep[][18]{PE}
+\end{quote}
+
+Several independent pieces of external evidence argue that this is not the case for root harmony in Turkish. 
+The adaptation of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, is one such case.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} While some speakers are said to be capable of pronouncing these non-native clusters, in normal speech the cluster is resolved by epenthesis of a [$+$\textsc{High}] vowel.
+In most cases, the epenthetic vowel matches the following root vowel for backness, suggesting that adapt foreign words in a way that conforms to the root harmony generalization.
+
+%the epenthetic vowel is always [$+$\textsc{High}],
+%and in general, it conforms to \textsc{Backness Harmony}.
+%For reasons that are 
+%and when the first consonant of the cluster is non-dorsal, it shows backness harmony
+
+\begin{example}
+Adaptation of initial foreign clusters \citep[][247]{Clements1982}:
+
+%\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l l}
+a. & spiker  & \alt{} & sipiker  & `announcer' \\
+   & fren    & \alt{} & firen    & `break'     \\
+b. & brom    & \alt{} & burom    & `bromide'   \\
+   & prusya  & \alt{} & purusya  & `Prussia'   \\
+%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
+%   & kredi   & \alt{} & kıredi   & `credit'    \\
+%   & trablus & \alt{} & tırablus & `Tripoli'   \\
+%b. & kral    & \alt{} & kıral    & `king'      \\
+%   & grup    & \alt{} & gurup    & `group'     \\
+\end{tabular}
+\end{example}
+
+Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but is identical to a morphological process in the related language Tuvan, in which it conveys a sense of ``vagueness or jocularity'', and \citeauthor{Harrison2001} report that it can be quickly taught to Turkish speakers.
+The game consists of reduplication of the base and replacement of the first vowel in the reduplicant with the [$+$\textsc{Back}] vowels \emph{a} or \emph{u}.
+Both in the native Tuvan process and in the Turkish game, the second vowel of the reduplicant (shown in braces below) is affected by this process: when the base is harmonically [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] results in what \citeauthor{Harrison2001} call ``reharmonization'', shown in  (\ref{redupgame}a).
+
+\begin{example}
+\label{redupgame}
+Turkish reduplication game \citep[][231]{Harrison2001}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l}
+a. & kibrit & kibrit-\{kabrıt\} & `match'    \\
+   & bütün  & bütün-\{batın\}   & `whole'    \\
+b. & mali   & mali-\{muli\}     & `Mali'     \\
+   & butik  & butik-\{batik\}   & `boutique' \\
+\end{tabular}
+\end{example}
+
+\noindent \citeauthor{Harrison2001} analyze this in terms of the \citeauthor{Clements1982} underspecification analysis, but it is equally consistent with full specification.
+Reharmonization is simply the spread of the backness specification of the \emph{a} or \emph{u} in the reduplicant.
+On the other hand, (\ref{redupgame}b) shows that this does not obtain for disharmonic roots; presumably, the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication as well.
+Similar results obtain both in Tuvan and in an unrelated Finnish language game \citep{Campbell1986}.
+
+A number of studies on Finnish suggest that speakers of that language, which has a vowel harmony system similar to that of Turkish, make use of (dis)harmony to processing running speech.\footnote{Thanks to Charles Yang for bringing these studies to my attention.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which conform to backness harmony.
+These stimuli are auditorily presented to subjects who are asked to press a button when the nonce trisyllable ends with a target nonce disyllable, or a real disyllabic word.
+Speakers are quicker to press the button when the prefix and disyllabic word have different backness specifications.
+These results suggest that speakers are attuned to the fact that disharmonic transisitions are good predictors of word boundaries.
+If speakers have also internalized the converse generalization, that harmonic transistions are more likely to be root-internal, then there is additional evidence that harmony is active not just in Finnish affix alternations but also in roots.
+
+\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect in word-spotting tasks as it does in Finnish: speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by the pseudo-prefix \emph{gölü-}, a disharmonic transistion, than when it is preceded by the pseudo-prefix \emph{golu-}, a harmonic transition.
+\citeauthor{Kabak2010} find that effect of harmony does not obtain for speakers of French, a language which lacks vowel harmony.
+As in Finnish, the results imply speakers have internalized the predominance of root-internal harmony.
+
+It seems that the root-internal harmony bias is in fact learned by Turkish speakers very early.
+The pseudoword spotting experiment has been adapted for 9-month-old Turkish infants by \citet{Kampen2008}.
+Infants are familiarized with harmonic disyllabic pseudowords bearing a pseudo-prefix, which may be harmonic or disharmonic.
+At test time, the infants are played the disyllabic pseudowords in isolation using the head turn preference paradigm.
+Infants show a preference to listen to those pseudowords which were familiarized with a disharmonic pseudo-prefix over those which were familiarized with a harmonic pseudo-prefix.
+This preference is not observed in 9-month-old infants learning German, which also lacks vowel harmony.
+Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants show no such preference.
+
+\subsection{Roundness Harmony}
+
+Roundness harmony is a closely related process with a somewhat narrower scope.
+
+\subsubsection{Phonological description}
+
+Roundness harmony is quite similar to \textsc{Backness Harmony}, with the additional stipulation that the target must be [$+$\textsc{High}].
+
+\begin{example}
+\textsc{Roundness Harmony}:
+
+\xymatrix@R=24pt@C=24pt{
+\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{--}[drr] &             & \txt{[$+$\textsc{High}]} \\
+\txt{V}                                         & \txt{C$_0$} & \txt{V}\ar@{-}[u] \\
+}
+\end{example}
+
+This rule, in concert with \textsc{Backness Harmony}, accounts for the forms of the dative singular (dat.sg.) and genitive singular (gen.sg.), among other suffixes.
+
+\begin{example}
+Turkish nominal suffix allomorphy:
+
+%\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l l l}
+   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
+a. & ip             & ipi            & ipin           & `rope' & \citep[][216]{Clements1982} \\
+   & el             & eli            & elin           & `hand'    \\
+   & kız            & kızı           & kızın          & `girl'    \\
+   & sap            & sapı           & sapın          & `stalk'   \\
+   & yüz            & yüzü           & yüzün          & `face'    \\
+   & köy            & köyü           & köyün          & `village' \\
+   & pul            & pulu           & pulun          & `stamp'   \\
+   & son            & sonu           & sonun          & `end'     \\
+b. & boğaz          & boğazı         & boğazın        & `throat'  & (TELL) \\
+   & pelür          & pelürü         & pelürün        & `tissue paper' \\
+   & döviz          & dövizi         & dövizin        & `currency' \\
+   & yamuk          & yamuğu         & yamuğun        & `trapezoid' \\
+   & ümit           & ümiti          & ümitin         & `hope'     \\
+\end{tabular}
+\end{example}
+
+\noindent As was the case for \textsc{Backness Harmony}, disharmonic roots are found, and the final vowel of disharmonic roots triggesr suffix harmony.
+
+\subsubsection{External evidence}
+
+Though few studies have directly sought external evidence for speaker's awareness of root \textsc{Roundness Harmony}, two of the findings reviewed in \S\ref{backharmexternal} above speak in support of this claim.
+First, the epenthetic vowel in non-native word-initial onsets
+like \emph{trablus} \alt{} \emph{tırablus} `Tripoli' conform also to \textsc{Roundness Harmony}.
+%\citealp[][247]{Clements1982}). 
+Second, \textsc{Roundness Harmony} participates in reharmonization in \citeauthor{Harrison2001}'s language game: the reduplicated \emph{bütün-batın} exhibits replacement of the second \emph{ü}, a [$+$\textsc{Round}] vowel, with [$-$\textsc{Round}] \emph{ı}.
+
+\subsection{Labial attraction}
+
+%\begin{quote}
+%\ldots{}these generalizations\ldots{}have no observable consequences in the course of the normal use of the language. \citep[][320]{Zimmer1969}
+%\end{quote}
+
+\subsubsection{Phonological description}
+
+\citet[][36]{Lees1966a} describes \textsc{Labial Attraction} as a process by which ``a high, short harmonic vowel is rounded in the second syllable of a disyllabic word whose first vowel is /a/, and whose medial consonant cluster contains a labial /p, b, m, v/, and then it is de-harmonified''.
+This description is transalted into an autosegmental rule below.
+
+\begin{example}
+\textsc{Labial Attraction}: %(after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}):
+
+\xymatrix@R=24pt@C=8pt{
+\txt{[$-$\textsc{Round}]} &                                         & \txt{[$-$\textsc{High}]} & \txt{[$+$\textsc{Labial}]}    & \txt{[$+$\textsc{High}]}\ar@{-}[dr] &         & \txt{[$+$\textsc{Round}]}\ar@{--}[dl] \\
+%                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C}\ar@{-}[u] &                                      & \txt{V} & \\
+                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C$_0$ C C$_0$}\ar@{-}[u] &                                      & \txt{V} & \\
+                         & \txt{[$+$\textsc{Back}]}\ar@{-}[urrrr]
+}
+\end{example}
+
+The formalization of this rule is naturally complex, and perhaps obscures the fact that \textsc{Labial Attraction} generates exceptions to \textsc{Roundness Harmony}, producing \emph{aCu} sequences (e.g., \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', \emph{samur} `sable'; \citealp[285]{Lees1966b}) rather than the expected \emph{aCı}.
+\citet[286]{Lees1966b} and \citet[311]{Zimmer1969} note the existence of exceptions (e.g., \emph{tavır} `mode') but both agree that they are surprisingly rare.
+
+\citet[225]{Clements1982} and \citet{Inkelas2001}
+
+There is reason to believe that \textsc{Labial Attraction} is at best a sequence structure constraint as it never applies in derived environments.
+If, contrary to fact, there was a \textsc{Labial Attraction} alternation, then one would expect, for example, that the gen.sg.
+of \emph{sap} `stalk' would be *\emph{sapun} instead of the observed \emph{sapın}.
+
+\citet{Lees1966a}
+\citet{Zimmer1969}
+
+\begin{quote}
+\ldots decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots~aCu~\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots We conclude that there is no systematic restriction on the set of consonants that may occur medially in rotos of the form /\ldots~aCu~\ldots/.
+\citep[][225]{Clements1982}
+\end{quote}
+
+\noindent
+\citeauthor{Clements1982} appear to be suggesting that \textsc{Labial Attraction} implies that \emph{aTu}, where \emph{T} represents a non-labial consonant, should be infrequent, but this fact is not inconsistent with the formulation of the constraint by \citet{Lees1966a,Lees1966b} and \citet{Zimmer1969}; \emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description.
+This is less than conclusive, since the fluctuating vowel alternation also fails to undergo harmony, and might just indicate that the flucutating vowel is epenthesized relatively late in the derivation; the frequency of \emph{aTu} provides only indirect information about \textsc{Labial Attraction} in that it provides a baseline estimate for just how frequent this disharmonic sequence of vowels is.
+
+\citet{Inkelas1997}
+\citet{Inkelas2001}
+
+\begin{quote}
+Lee's rule of \textsc{Labial Attraction}\ldots is not a real generalization about the Turkish lexicon.
+It is not true synchronically, either of native or nonnative items; nor, according to the historical and dialectical literature, does \textsc{Labial Attraction} appear to have been true at any stage going back as far as Old Turkic.
+\citep[][196]{Inkelas2001}
+\end{quote}
+
+zimmer: 319-320 
+inkelas et al.: 196
+
+\subsubsection{Psycholinguistic evidence}
+
+To the author's knowledge, there are no psycholinguistic studies investigating \textsc{Labial Attraction} beyond the experimental results of \citet{Zimmer1969} reviewed in the next section.
+
+\section{Evaluation}
+
+The results reveal a dissociation between the lexical statistics and wordlikeness judgements, a dissociation that finds a natural explanation from the principle of ...
+
+There is some disagreement in the literature about just what \citet{Zimmer1969} finds regarding \textsc{Backness Harmony} and \textsc{Labial Attraction}.
+
+\subsection{Lexical statistics}
+\label{lexstats}
+
+This choice of statistical test is justified in Section \ref{stattech}.
+
+\subsubsection{Backness harmony}
+
+\subsubsection{Roundness harmony}
+
+%\ex Lexical effects of \textsc{Roundness Harmony} \citep{TELL}: \vspace{6pt} \\
+%\begin{tabular}{l r r r r r}
+%\toprule              & 
+%Corpus                & $p$-value \\
+%\midrule
+%Full TELL             &  
+%Elicited TELL         & 
+%TELL with etymologies & 
+%\end{tabular}
+%\xe
+
+%\citet{Harrison2004} 73\% 
+%of lexical types are harmonic (both back and round)
+
+\subsubsection{Labial attraction}
+
+\begin{quote}
+Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
+\citep[][196, emphasis in original]{Inkelas2001}
+\end{quote}
+
+%\ex Lexical effects of \textsc{Labial Attraction} \citep[][186]{Inkelas2001}: \vspace{6pt} \\
+%\begin{tabular}{l r r r r r}
+%\toprule
+%Corpus                & aPu & aPı & aTu & aTı   & $p$-value   \\ % & corpus size
+%\midrule
+%Full TELL             & 378 & 248 & 446 & 1,140 & 2.83\e{-44} \\ % & 31,236 \\
+%Elicited TELL         & 152 & 265 & 101 & 1,839 & 9.84\e{-60} \\ % & 16,541 \\
+%TELL with etymologies & 128 & 109 &  79 &   470 & 6.56\e{-32} \\ 
+%\bottomrule
+%\end{tabular}
+%\xe
+
+%Etymological subset of TELL
+%Native    & Foreign   \\
+%aBu & aBI & aBu & aBI \\
+%12  & 11  & 84  & 28  \\
+%p = 0.0417
+
+\subsection{Wordlikeness results}
+
+There is some disagreement in the literature about just what \citeauthor{Zimmer1969}'s study reveals about \textsc{Labial Attraction}.
+
+zimmer
+clements
+\citet{Ito1993} 
+%and \citet{NiChiosain1993} 
+argue that the data shows that it holds of native words only
+
+\citet{Inkelas2001} refute this point
+
+\citet{Zuraw2000}:4
+
+\citet{Inkelas2001} present a different variation on the original rule of \textsc{Labial Attraction}
+
+\begin{quote}
+Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
+\citep[][196, emphasis in original]{Inkelas2001}
+\end{quote}
+
+\noindent \citeauthor{Inkelas2001} persistently refer to ``statistical support'', but in fact their study contains no statistics (that is, quantities calculated from a set of data) beyond counts and percentages.
+
+``These results support our view that there is no rule of Labial Attraction'' \citep[][225]{Clements1982}.
+
+%since the opposing are all broader versions than the original formulation by \citeauthor{Lees1966a}.
+
+\citet{Goodman1954}
+
+\begin{example}
+$\displaystyle \gamma = \frac{C - D}{C + D}$
+\end{example}
+
+\subsubsection{Backness harmony}
+
+\begin{example}
+Backness harmony wordlikeness forced choices \citep[314]{Zimmer1969}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r l r}
+\toprule
+\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
+\midrule
+pemez & 30                   & pemaz & 2  \\
+tipez & 24                   & tipaz & 8  \\ 
+terüz & 19                   & teruz & 13 \\ % roundness violator
+teriz & 28                   & terız & 3  \\
+tokaz & 26                   & tokez & 6  \\ % roundness violator
+\bottomrule
+\end{tabular}
+\end{example}
+
+$\gamma = 0.597$, $p = 5.2$\e{-29}
+
+\subsubsection{Roundness harmony}
+
+\begin{example}
+Roundness harmony forced choice wordlikeness judgements:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r l r}
+\toprule
+\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
+\midrule
+pörüz & 32 & pöriz & 0  \\
+tatız & 20 & tatuz & 12 \\
+tüpüz & 31 & tüpiz & 1  \\
+takız & 22 & takuz & 10 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+$\gamma = 0.641$, $p = 7.7$\e{-27}
+
+%tamaz & 26 & tamoz & 6  \\
+%putoz & 25 & putaz & 7  \\
+
+\subsubsection{Labial harmony}
+
+\begin{example}
+Labial attraction wordlikeness forced choices \citep[314]{Zimmer1969}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r l r}
+\toprule
+\multicolumn{2}{l}{aPu} & \multicolumn{2}{l}{aPı} \\
+\midrule
+pamuz & 17              & pamız & 15 \\
+tafuz & 21              & tafız & 11 \\
+tapuz & 17              & tapız & 15 \\
+mavuz & 16              & mavız & 16 \\
+tabuz & 16              & tabız & 16 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+$\gamma = 0.087$, $p = 0.101$
+
+\section{Conclusions}
+
+\citet{Becker2011}
+
+It is 
+%further, even long-distance co-occurrence restrictions frequently are ``conjunctive'': for instance, \citet{Gallagher2009} report that in Chol, ejectives must be totally identical, that is to say, 
+%\citet{Bakovic2005b}
+
+This has an ad hoc nature to it; 
+There are two senses in which this objection is ad hoc. 
+First, \citeauthor{Becker2011} appeal to no particular theory of the naturalness of processes or static constraints which excludes \textsc{Labial Attraction}. 
+Secondly, this appears to be a minority view: \textsc{Labial Attraction} was considered a true generalization by early specialists
+\citep[e.g.,][]{Lees1966a}, and despite \citeauthor{Zimmer1969}'s suggestive psycholinguistic results, reviewed above, it also been treated as a plausible constraint by later theorists \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995a,Zuraw2000}.
+Further, there is a real danger that if the label ``unnatural'' 
+%Labial Attraction} 
+describes an impossible structural change or structural description, that one will fail to account for earlier forms of Turkish or sound changes therein.
+
+%Classical Arabic adjectives often have stative verbs in which the root is imposed onto the template CaCuCa:
+Classical Arabic adjectives often have stative verbs in which the root melody is rewritten with \emph{a\ldots{}u}.\footnote{Thanks to Uri Horesh (p.c.) for assistance with this data.}
+
+\begin{example}
+Arabic verbs of coming into being: \\
+\begin{tabular}{l l l l}
+kabiːr  & `big'        & kabura  & `become big'       \\
+ħasan   & `handsome'   & ħasuna  & `become beautiful' \\
+dʒadiːb & `barren'     & dʒaduba & `become dry'       \\
+\end{tabular}
+%b. & saʁiir & }`become small'     & (cf. \emph{saʁiːr} `small') \\
+\end{example}
+
+% FORMER 3.3.2: Diachronic factors
+
+% etymological issues
+\citet{Inkelas2001}
+
+\begin{example}
+\textsc{Labial harmony} and etymology in TELL \citep[][187]{Inkelas2001}:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r}
+        & aPu & aPı & $p$-value \\
+Native  & 12  & 11  & \multirow{2}{*}{0.042} \\
+Foreign & 84  & 28  \\
+\end{tabular}
+\end{example}
+
+\citet{NiChiosain1993} and \citet{Ito1995b}
+
+``sonority projection''
+Ther
+
+\citet{Daland2011b} finds that this can be learned easily from positive data but a number of different psych
+
+%It is important to note that the nature of the \emph{blick}/\emph{bnick} contrast is no longer a mystery. 
+%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
+It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails.
+If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
+
+%\footnote{I note a potential problem with this account: \citet{Davidson2005,Davidson2006a} argue that this ``excrescent'' schwa is distinct from the both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.}
+
+%\noindent A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions.
+Any explicit theory of static constraints will have to impose formal restrictions on possible co-occurrence restrictions.
+The set of possible constraints must respect results in formal learning theory, and may also be constrained by markedness considerations, for example.
+
+%A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
+This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).
+
+%Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
+In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.
+
+%In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word. 
+%The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements. 
+%The data below, drawn from Turkish, shows that
+
+%This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
+Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements.
+Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
+
+%Classical Arabic adjectives often have stative verbs in which the root is imposed onto the template CaCuCa:

### 2012-05-24 4.tex
back to chapter files, lots of writing

@@
-%\chapter{Accidental and structural gaps in English syllable contact\protect\footnotemark{*}\protect\footnotetext{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics.}}
 \chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
@@
 
-\input{4.1}
-\input{4.2}
-\input{4.3}
+\section{Domain}
+
+% WHY I'm DOING THIS STUDY
+% describes makeup of the corpus and how p
+
+\subsection{Data sources}
+
+The English CELEX database \citep{CELEX},
+
+\citet{Aronoff1976}
+
+\citet{Harley2009}
+
+\citet{Taft1975} ?
+\citet{Taft1981} ?
+\citet{Taft2004a} ?
+
+\citet{Halle1962}
+
+% 2.1.4: k
+
+\subsubsection{Corpus counts}
+
+Under the phonological analyses described in the next two sections, the 6,876 simplex words in CELEX contain 21 unique medial codas and 40 unique medial onsets.
+Of these 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, for a 18.8\% saturation rate.
+
+\subsection{Syllabification}
+
+For this task, it is necessary to separate medial consonant clusters into codas and onsets.
+While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic.
+For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language.
+Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which explain the \emph{chemistry}/\emph{ministry} syllabification contrast.
+
+An automated syllabification system was constructed and applied to medial clusters in the CELEX data.
+This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.
+There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification.
+The technique used is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948}, \citealt[42f.]{Kahn1976}, \citealt[][?]{Selkirk1983}) which parses word-medial clusters so that the onset is as large as possible.
+Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty.
+In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.
+
+There is one case where unchecked maximization of medial onsets produces incorrect syllabifications.
+When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
+\citet[][55]{Harris1994} notes that onset maximization makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
+The approach adopted here is to first apply onset maximization, then to reassign the first consonant of a medial onset to the preceding coda if it is immediately preceded by a stressed lax vowel.\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}
+
+If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables.
+For this reason, an additional constraint is assumed which prevents the two components of the affricates from being split by syllabification.
+This is motivated by the tendency of affricates to pattern with single segments in many languages.
+For instance, the only complex onsets in Classical Nahua are the affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.
+
+%In fact, [t.ʃ, d.ʒ] clusters are not found in simplex English words, despite the fact that that affricates occur as medial onsets in clusters (e.g., \emph{tru}[n.tʃ]\emph{eon}, \emph{so}[l.dʒ]\emph{er}).
+
+\subsection{Inventory considerations}
+
+To identify the inventory of syllable contact clusters, it is also necessary to parse syllables into onset, nucleus, and coda.
+In most cases this is trivial, but a few distributional heuristics for difficult cases are described below.
+
+\subsubsection{The velar nasal}
+\label{velarnasal}
+
+There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right, as demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}.
+The strongest piece of evidence for treating the velar nasal as a pure allophone is the general absence of ?
+nset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
+\citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here.
+English nasal place allophony is formalized in Section \ref{cnpasection} below.
+%\citealt[][62]{Halle1985a}),
+
+\subsubsection{[j] onglides}
+
+The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J.
+\citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}).
+The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).
+%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
+There is extensive motivation for this principle.
+When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː].
+In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}.
+This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
+\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear.
+There is also strong external support for this principle.
+The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants.
+The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor} [h]\emph{omponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
+Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.
+
+\begin{example}[VBR's complex onsets (\citealp{Buchwald2005}:79--80, his transcriptions)] \label{VBR}
+\begin{tabular}{l l l}
+a. & kəræb  & `crab'  \\
+   & bəlid  & `bleed' \\
+b. & kəwin  & `queen' \\
+   & kəwoʊt & `quote' \\
+c. & kut    & `cute'  \\
+   & musɪk  & `music  \\ 
+\end{tabular}
+\end{example}
+
+\noindent
+VBR breaks up complex onsets with epenthesis, including those that consist of a consonant and a back onglide cluster (\ref{VBR}b).
+However, no epenthesis occurs between a consonant and a front onglide; rather, the glide is absent (\ref{VBR}c).
+The failure of the front onglide to pattern with other consonant clusters suggests once again that the glide is part of the nucleus.
+
+%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [juː] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant.
+%This is the only sign that [juː] shows an affiliation for the onset.
+
+\subsubsection{[w] onglides}
+
+The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset.
+Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
+Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}.
+Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.
+And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.
+
+%Whereas [Cjuː] syllables attracts stress, [kwV] syllables do not \citep[][162f.]{Davis1995}.
+
+\subsubsection{Post-vocalic \emph{r}}
+
+Both CELEX and the dictionary used by \citet{Pierrehumbert1994} in her study transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost.
+In \emph{r}-full dialects, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters.
+Many vowel contrasts are suspended before \emph{r} \citep[][255]{Harris1994}; for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}.
+Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which post-vocalic \emph{r} patterns differently than internal coda consonants.
+\citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.
+
+\begin{example}[/t/-\textsc{Glottalization} in British English (after \citealp{Harris1994}:195, 258)]
+\begin{tabular}{l l l@{} l}
+a. & fis[t]   & * & fis[ʔ]   \\
+   & mis[t]er & * & mis[ʔ]er \\
+b. & par[t]   &   & par[ʔ]   \\
+   & car[t]on &   & car[ʔ]on \\
+\end{tabular}
+\end{example}
+
+\noindent Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.
+
+\begin{example}[/t, d/-\textsc{Deletion}in American English] \label{td}
+\begin{tabular}{l l l@{} l}
+a. & be[lt]  &   & be[l]  \\
+   & me[nd]  &   & me[n]  \\
+b. & sh[ɚt]  & * & sh[ɚ] \\
+   & c[ɚd]   & * & c[ɚ]  \\
+\end{tabular}
+\end{example}
+
+%Finally, \citet[][251]{Fromkin1973} also presents evidence that post-vocalic \emph{r} may behave as if nuclear in speech errors.
+
+\section{Evaluation}
+
+% what is being evaluated
+% summary
+
+\subsection{Static and derived constraints}
+
+With this database of English syllable contact clusters it is now possible to evaluate constraints on sequences of underlying phonemes.
+
+Here, the metric used is simply
+
+The evaluation
+
+The foremost evidence for
+
+The evaluation used here is simply
+
+Having described the database of English syllable contact clusters,
+
+I now consider constraints on sequences of URs as predictors of attested and unattested contact clusters.
+First, I describe the statistical technique used, then apply this technique to both phonologically derived constraints and a set of ``static'' phonotactic constraints proposed by \citet{Pierrehumbert1994}.
+
+\subsubsection{Statistical technique}
+\label{stattech}
+
+Before the dawn of generative phonology, many linguists concerned themselves with documenting co-occurrence in various languages.
+\citet[][28]{Vogt1954} declares that a ``phonemic description of a language \ldots should also comprise a description of the phoneme combinations that occur or may be expected to occur in the language''.
+Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively.
+\citeauthor{Uhlenbeck1950}, for instance, puts forth co-occurrence dispreferences, rather than exceptionless generalizations of the sort given by \citet{Greenberg1950}.
+
+\citet{McCarthy1988}
+
+Years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and proposes a statistical technique, based on the chi-square test, to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.
+
+\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English).
+The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented.
+Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.
+
+\begin{example}[null hypothesis testing]
+$\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
+χ^2$
+\end{example}
+
+%\footnote{This is the one-tailed version of the test.}
+
+In this study, I use the \citet{Fisher1934} exact test, which is isomorphic to this chi-square test.
+Compared to the chi-square test, the Fisher exact test $p$-value is somewhat more difficult to compute, but both are automated by modern statistical packages and can be computed rapidly.
+The major advantage of the Fisher test over the chi-square test is that it is appropriate for small samples or rare events, whereas the chi-square test depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples \citep[see][]{Gorman2012a}.
+%\citep[e.g.,][142]{Mester1988}.
+
+On the other end of the spectrum, the Monte Carlo statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010} does not scale to large samples.
+The Monte Carlo procedure is isomorphic to the one-tailed Fisher Exact test, and involves comparing observed co-occurrence counts to those produced by randomly generated samples.
+This requires the researcher to generate random permutations, something which is particularly difficult for large samples, as I will explain; disinterested readers may wish to skip the following paragraph.
+
+Any permutation of a sequence $L$ can be represented as a sequence $S$ of the same length, where the value of $S_i$ corresponds to the permuted position of $L_i$.
+Generating random permutations requires an unbiased method to generate all $N!$ of these lists with equal probability.
+Unfortunately, this is near impossible for $N$s of moderate size with current computational resources.
+Any pseudo-random number generator can be characterized by its ``period'', the number of pseduo-random numbers it can emit before repeating itself; this is also the upper limit for the number of random permutations that can be generated.
+The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a period of approximately $2^{48}$, a number which turns out to be smaller than $17!$.
+For the large samples, sometimes in the thousands, used by these researchers, virtually all the possible permutations can never be generated by this method.
+As a consequence, the shuffling procedure is far from random and introducing the possibility that the statistical test is biased in unpredictable ways.
+
+As an example of the Fisher exact test as it is applied here, consider some of the Javanese co-occurrence facts considered by \citet{Mester1988}.
+\citeauthor{Mester1988} is at pains to show that there is a gradient dispreference for first and second root consonants which share the same major place.
+However, perfectly identical first and second consonants are exempt.
+Since
+
+% APPLY THIS TO JAVANESE
+
+\begin{example}[Javanese root co-occurrence] %(\citealp[][264]{Uhlenbeck1950}, \citealp[][139]{Mester1988})]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 134      & 12         & 0.918      & \multirow{2}{*}{6.095\e{-06}} \\
+violating  & 21       & 15         & 0.583 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent
+It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.
+
+This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance
+
+That this is highly unlikely to be due to chance is apparent both from the chi-square ($\chi^2 = 25.572$, two-tailed $p = 4.3$\e{-07}) and Fisher exaact tests ($p =  6.095$\e{-06}).
+
+I use two-tailed statistical tests throughout. 
+Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis.
+As
+
+\citet{Pierrehumbert1994}
+
+It must be stressed that under current assumptions, this is independent of whether predictable prosodic representations (particularly coda and onset identity) are part of the memorized underlying representation \citep[e.g.,][]{Vaux2003}.
+It is an undeniable fact that URs need to be 
+The derivative nature of 
+\citet{Ito1989a,Noske1992}
+
+least some researchers (i.e., \citealt{Mester1988})
+
+ \citep[e.g.,]{Brown2010} have argued for phonotactic constraints giving rise to overrepresentation, so I admit this possibility here and use two-tailed tests throughout.
+
+%\footnote{For example, \citet{Martin2007} applies this technique to a list of 4,758 noun-noun compounds.
+The number of permutations of a list of this length can be estimated using Stirling's approximation.
+
+%\ex $ \lim_{n \rightarrow \infty} n! = e ^ {n \ln{n} - n} = 2 ^ {\log_2{e} (n \ln{n} - n)} $ \xe
+
+%\noindent
+%$4,758!$ is approximately $2 ^ {51260}$. The author is unaware of any system that provides 51,260 bits of entropy; and some popular programming languages, such as ANSI C, provide as few as 15 bits.}
+
+%%% START OF SERIOUSNESS
+
+I will first consider three well-known English phonological alternations, showing that these give rise to constraints on the sequence structure of morphs.
+Simple autosegmental formalizations of these processes are provided, and the Fisher exact test is used to demonstrate their effect on the English lexicon.
+
+\subsubsection{Obstruent Voice Assimilation}
+
+English has a process of voice assimilation which is found in two inflectional suffixes.
+The suffix /-d/ forms regular pasts and denominal adjectives, and /-z/ forms regular noun plurals and genitives, and the 3sg.
+present active indicative verbs.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S.
+\citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}, and alternative views are presented by \citet[][210f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
+
+\begin{example}
+Voice assimilation in the past tense and noun plural:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l l l}
+a. & nap & nap[t] \\
+   & nab & nab[d] \\
+b. & lap & lap[s] \\
+   & lab & lab[z]  \\
+\end{tabular}
+\end{example}
+
+\noindent
+The voiced obstruent suffixes devoice when preceded by a voiceless obstruent.
+This is formalized as a rule spreading the [\textsc{Voice}] specification rightward over adjacent obstruents.
+
+\begin{example}
+\textsc{Obstruent Voice Assimilation} (\emph{SPE}:178):
+
+\xymatrix@R=24pt@C=24pt{
+\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{..}[dr] & \\
+\txt{C}\ar@{-}[d]             & \txt{C}\ar@{-}[d] \\
+\txt{[$+$\textsc{Obstruent}]} & \txt{[$+$\textsc{Obstruent}]}\\
+}
+\end{example}
+
+\noindent
+This rules out the possibility of surface clusters of obstruents with differing  [\textsc{Voice}] specifications.
+While the data make it clear \textsc{Obstruent Voice Assimilation} applies in morphologically derived environments, it is not obvious that it applies to obstruent clusters in non-derived environments (e.g., obstruent clusters in the URs of morphs) as well.
+\citet{Pierrehumbert1994} does not mention voice assimilation in her study of gaps in English syllable contact clusters.
+\citet[][74f.]{Hammond1999a} notes the existence of words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain obstruent voice clusters disagreeing in [\textsc{Voice}], and takes this as evidence that obstruent voicing is restricted to derived environments.
+
+Despite these exceptions, I propose that \textsc{Obstruent Voice Assimilation} plays a major role in shaping the English lexicon.
+Only English syllable contact clusters which contain two adjacent obstruents speak to the status of \textsc{Obstruent Voice Assimilation}; the remaining 160 possible clusters are ignored.
+A $2 \times 2$ contingency table is constructed by dividing the remaining possible clusters into those which agree on obstruent voicing and those which do not, and into those which are attested and those which are not.
+
+\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}]
+\vspace{0.5\baselineskip}
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 80 & 370 & 0.178 & \multirow{2}{*}{1.1\e{-11}}\\
+violating  &  6 & 264 & 0.022 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent 
+18\% of the possible clusters containing adjacent obstruents with the same [\textsc{Voice}] specification are found, whereas only 2\%  of those which disagree in [\textsc{Voice}], such as in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, CELEX also contains \emph{jo}[d.p]\emph{urs}, \emph{pi}[ntʃ.b]\emph{eck}, \emph{sa}[k.b]\emph{ut}, and \emph{ja}[k.d]\emph{aw}.
+As the $p$-value shows, the rarity of these disagreeing clusters is unlikely to be due to chance.
+
+I propose that this pattern should be attributed directly to \textsc{Obstruent Voice Assimilation} applying to non-derived environments, preventing learners from positing URs in which adjacent obstruents do not agree in voice.
+There is no duplication between constraints on sequence structure and surface forms if both are derived from the application of the phonological rule (\citealt[][401f.]{Stanley1967}, \citealt[][382]{SPE}).
+The handful of exceptions to this generalization can be marked with appropriate exception diacritics.
+Another possibility is suggested by recent psycholinguistic results.
+\citet{Mattys2001b} find that 9-month old infants are sensitive to the contrast between [vt], a cluster of obstruents disagreeing in [\textsc{Voice}] which is not found word-medially, and [ft], which conforms to the above generalization and is found medially (e.g.
+\emph{a}[f.t]\emph{er}).
+This ability to use phonotactic violations to segment continuous speech persists into adulthood \citep{McQueen1998}, suggesting that ``hetero-voiced'' obstruent clusters may be analyzed as semantically opaque compounds by native speakers (e.g., \emph{ja}[k\#d]\emph{aw}).
+%\footnote{An analogous account is the proposal that readjustment rules insert morphological boundaries for the purposes of assigning stress (\emph{SPE}:94).}
+
+\subsubsection{Coda nasal place assimilation}
+\label{cnpasection}
+
+Above, it was assumed that the velar nasal is derived from underlying /n/. 
+This analysis is not purely allophonic, as nasal place assimilation alternations target the final consonant of the \emph{im-}/\emph{in-} prefix.
+In (\ref{npa}ab), the prefix-final consonant takes on the same major place of articulation as the following obstruent, and in (\ref{npa}c), where the root is vowel-initial, the prefix-final consonant is realized as a default /n/.
+
+\begin{example}
+\label{npa}
+Nasal place assimilation in Latinate prefixes:
+
+\vspace{0.5\baselineskip}
+\begin{tabular}{l l l l}
+%a. & possible & i[m.p]ossible & & balance & i[m.b]alance \\
+%b. & tangible & i[n.t]angible & & decent  & i[n.d]ecent  \\
+%c. & elegant  & i[n.]elegant  & & ability & i[n.]ability \\
+a. & possible & i[m.p]ossible \\
+   & balance  & i[m.b]alance  \\
+b. & tangible & i[n.t]angible \\
+   & decent   & i[n.d]ecent   \\
+c. & elegant  & i[n]elegant  \\
+   & ability  & i[n]ability  \\
+\end{tabular} 
+\end{example}
+
+\noindent
+The shape of the prefix before other obstruents is somewhat more complex. 
+%Before /f/, CELEX transcribes the nasal as either labial or alveolar, the latter being very rare in simplex words. 
+%There are also additional complexities regarding the shape of the prefix when attached to /k, g/-initial roots. 
+\citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but this assimilation is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
+The prefix \emph{com-}/\emph{con-}, which patterns with \emph{im-}/\emph{in-} in showing assimilation (e.g., \emph{co}[m.b]\emph{ine}), participates in a similar contrast, between \emph{có}[ŋ.g]\emph{ress} and \emph{co}[n.g]\emph{réssional}.
+As \citeauthor{Borowsky1986} notes, stress does not block assimilation in simplex words (e.g., \emph{a}[ŋ.g]\emph{óra}), a fact which demonstrates the necessity of a partially morphological explanation for the blocking of assimliation in words like \emph{co}[n.g]\emph{réssional}.
+Regardless of these complexities, there is sufficient evidence to posit assimilation of the \textsc{Place} features of an obstruent to a preceding nasal consonant.
+
+\begin{example}
+\textsc{Coda Nasal Place Assimilation} (\emph{SPE}:85):
+
+\xymatrix@R=24pt@C=24pt{
+                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{..}[dl] \\
+\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d] \\
+\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]} \\
+}
+\end{example}
+
+\citet[][175]{Pierrehumbert1994} observes that in simplex English words, ``nasal-stop sequences agree in labiality''; presumably, this generalization is limited to [\textsc{Labial}] because \citeauthor{Pierrehumbert1994} also adopts the assumption that the velar nasal is derived from /n/.
+However, \citeauthor{Pierrehumbert1994} does not propose any connection between the alternation and the generalization that nasal consonants in URs tend to agree with the following consonant.
+While clusters like \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey} are the norm, exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} are found.
+
+In all, there are only 80 possible nasal-obstruent syllable contact clusters, the majority of which are attested.
+In addition, there are a small number of words which contain clusters like \emph{oi}[nt.m]\emph{ent}, in which a nasal-obstruent cluster are both parsed into a complex medial coda, but in this position, place agreement is exceptionless, and thus these clusters are ignored below.
+
+\begin{example}[Lexical effects of \textsc{Coda Nasal Place Assimilation}]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 41 & 11 & 0.788 & \multirow{2}{*}{2.6\e{-05}}\\
+violating  & 8  & 20 & 0.286 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent Clusters which conform to this generalization are far more likely to be attested than those that violate it, and this is highly unlikely to be due to chance.
+
+There is some evidence that even young infants show a preference for nonce words that conform to this generalization. %\citet{Jusczyk2003}
+\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} play nonce words containing nasal-obstruent clusters to 4.5, 9, and 10-month-old infants (respectively) acquiring English, All three studies find that infants prefer to listen to stimuli which conform to the above generalization (e.g., \emph{u}[m.b]\emph{o}) over those which do not (e.g., \emph{u}[n.b]\emph{o}).
+\citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''.
+In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.
+
+\begin{example}[Artificial language adaptations (\citealt{Wright1975}:394, his transcriptions)]
+\begin{tabular}{l l l l l l}
+a. & [gownp] & > & [gump] \\
+b. & [ǰumg]  & > & [ǰúŋgə] \\
+c. & [ðʌŋd]  & > & [tɔŋg] \\
+\end{tabular}
+\end{example}
+
+\noindent If this game is akin to the process of loanword adaptation, it is possible that various extragrammatical pressures are at play \citep[e.g.,][]{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008}.
+Yet, the independent evidence for a phonological process of nasal place assimilation provides the most parsiminious account of the adaptations seen in \citeauthor{Wright1975}'s study.
+J.
+\citet{Myers1993} also shows that \textsc{Coda Nasal Place Assimilation} is fed by speech errors:
+
+\begin{example}[Speech errors and \textsc{Coda Nasal Place Assimilation} (\citealp{Myers1993}:228)]
+\begin{tabular}{l l l}
+   & \emph{error}  & \emph{target} \\
+a. & primps        & prints        \\
+b. & rand orker    & rank order    \\
+%  & [sprɪg]time for [hɪ̃ntlɘr] & springtime for Hitler \\
+c. & po[ŋ]cutation & computation   \\
+\end{tabular} 
+\end{example}
+
+%\citet{Hay2004a} demonstrates that [n.p, mθ] clusters, which are not found in English, are in fact rated better than many attested clusters.
+%report that 10-month-old infants prefer to listen to nonsense words which conform to \textsc{Coda Nasal Place Assimilation} (e.g., \emph{u}[m.b]\emph{o}) compared to those which do not (e.g., \emph{u}[n.b]\emph{o}).
+
+\subsubsection{Degemination}
+
+A final well-known alternation in English is the simplification of geminates derived by certain suffixes.
+For instance, the suffix \emph{-ly} rarely forms a geminate when attaching to root-final /l/, whether forming denominal adjectives (e.g., \emph{sme}[l]\emph{y}) or deadjectival adverbs (e.g., \emph{norma}[l]\emph{y}).
+A small class of English verbs take a /-t/ suffix in the past; as this suffix does not trigger the \textsc{Epenthesis} found in the regular past (e.g., \citealt{Bakovic2005b}, \citealt{Fruehwald2011}), \textsc{Degemination} applies.
+
+\begin{example}[Degemination in /-t/ pasts (\citealp{Halle1985a}:105, \citealp{Myers1987}:492)]
+\begin{tabular}{l l l l l}
+a. & burn    & burn[t] \\
+   & spill   & spil[t] \\
+b. & ben[d]  & ben[t]  \\
+   & buil[d] & buil[t] \\
+\end{tabular}
+\end{example}
+
+More complicated cases of degemination are found in English prefix morphology and described in \emph{SPE} (p.~148, 219f.) and \citealt[][]{Borowsky1986} (p. 102f.), and are not reviewed here.
+
+%\ex \textsc{Coda Nasal Place Assimilation} feeds \textsc{Degemination} \citep[][116]{Borowsky1986}: \\
+%\begin{tabular}{l l l l}
+%a. & mature    & i[m]ature    & (cf. \emph{i}[m.b]\emph{alance})   \\
+%b. & numerable & i[n]umerable & (cf. \emph{i}[n.d]\emph{ependent}) \\
+%\end{tabular}
+%\xe
+
+The rule deletes the first of two adjacent consonants agreeing in \textsc{Place} and manner features, and is formalized using quantification over features \citep{Reiss2003b}.
+% cf. \citep{Berent2012}
+
+\begin{example}[\textsc{Degemination} (\emph{SPE}:253)]
+\xymatrix@R=24pt@C=24pt{
+\textsc{C}_1       & \longrightarrow & \emptyset & / & A~\gap\gap~A & \textsc{C}_2 \\
+}
+Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
+\end{example}
+
+\noindent I assume that geminates crossing a syllable boundary are ``false geminates'', meaning that they are adjacent similar segments rather than a single feature matrix linked to two timing slots.
+Under the alternative account in which the absence of geminates is an inventory restriction (i.e., a \textsc{Segment Structure Condition}), an undesirable duplication of phonological derived surface constraints and constraints on URs would be unavoidable.
+
+Neither identical consonants, nor those which differ only in voice (e.g., *[p.b]) occurs in simplex words in CELEX, and this absence is highly unlikely to be due to chance.
+
+\begin{example}[Lexical effects of \textsc{Degemination}] 
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value                   \\
+\midrule
+conforming & 157      & 596        & 0.208      & \multirow{2}{*}{7.2\e{-09}} \\
+violating  & 0        &  87        & 0.000                                    \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+%\footnote{Note that it is not the case that the English lexicon as envisioned by \emph{SPE} is free of underlying geminates; \citeauthor{SPE} use underlying geminates to attract stress to a preceding nucleus (p.~148--151).
+In \emph{SPE}, the underlying geminate-singleton contrast is absolutely neutralized, but has consequences for surface stress.
+In addition to the usual objections to absolute neutralization (i.e., that it is a learning conundrum mascarading as a phonological solution) that emerged in the wake of the publication of \emph{SPE}, the phonological particulars of this account have also come under attack \citep[e.g.,][]{Ross1972a}.
+It would not seem tenable to use underlying geminates to attract stress even if absolutely neutralized contrasts were admitted.} Unsurprisingly, this is unlikely to be due to chance.
+
+\subsubsection{Dorsal/non-coronal clusters}
+
+\citet{Pierrehumbert1994} surveys the English syllable contact cluster inventory, and discerns the effect of several ``static'' constraints, those which do not derive from neutralizing phonological rules.
+
+\citet[][173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Since \citeauthor{Pierrehumbert1994}'s study is restricted to triconsonantal clusters, so words like \emph{a}[k.m]\emph{e}, \emph{ru}[g.b]\emph{y}, or \emph{pi}[g.m]\emph{ent} do not violate this narrower generalization.
+Further, \citeauthor{Pierrehumbert1994} adopts the traditional assumption \citep[e.g.,][66f.]{Borowsky1986} that coda [ŋ] is an allophone of /n/ before dorsal consonants, so words such as \emph{mo}[ŋ.gr]\emph{el} are not exceptions.
+
+\citeauthor{Pierrehumbert1994} generalization is termed static simply because there is no evidence for any phonological process which changes clusters of dorsal obstruents followed by non-coronals into something else (perhaps a dorsal/coronal cluster).
+However, the statistical test, taking into account clusters of any length, does not indicate that this dispreference for dorsal/non-coronal clusters is so compelling that it could not be generated by chance, simply because possible dorsal/non-coronal clusters are about as likely to occur as are similar types of clusters such as those with a non-dorsal coda and a non-coronal onset (e.g., \emph{a}[t.m]\emph{osphere}) or a dorsal coda and coronal onset (e.g., \emph{ve}[k.t]\emph{or}).
+These clusters are marked as ``conforming'' to a constraint \textsc{*[+Dorsal][$-$Coronal]}, and dorsal/non-coronal consonants as ``violating'' it.
+
+\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 76 & 359 & 0.175 & \multirow{2}{*}{0.145} \\
+violating  &  6 &  57 & 0.095 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+There is no exceptions to \citeauthor{Pierrehumbert1994}'s narrower generalization concerning triconsonantal clusters, but it is no more statistically convincing.
+
+\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}, triconsonantal clusters only]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value              \\
+\midrule
+conforming & 20       & 148        & 0.119      & \multirow{2}{*}{0.136} \\
+violating  &  0       &  22        & 0.000                               \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent The evidence for \citeauthor{Pierrehumbert1994}'s static constraint does not actually require that such a constraint exists at all, since it is quite possible that the apparent dispreference for dorsal/non-coronal clusters could result by accident.
+
+\subsubsection{Coda coronal obstruents}
+
+Another static generalization about triconsonantal syllable contact clusters proposed by \citet[][175]{Pierrehumbert1994} is that  ``clusters with a coronal obstruent in the coda do not occur.'' In an endnote (p.
+186), \citet{Pierrehumbert1994} notes an exception in \emph{a}[nt.l]\emph{er}, and one can add the triconsonantal clusters in \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many biconsonantal clusters like \emph{a}[t.l\emph{as} or \emph{e}[s.t]\emph{er}.
+As before, applying the statistical test to clusters of any length reveals that coda coronal obstruent clusters are not significantly less likely to occur than clusters with non-coronal codas.
+
+\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value              \\
+\midrule
+conforming & 48       & 312        & 0.133      & \multirow{2}{*}{0.301} \\
+violating  & 38       & 322        & 0.106                               \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+Restricting \citeauthor{Pierrehumbert1994}'s generalization to triconsonantal clusters does not strengthen the case for this static constraint.
+
+% there isn't really any theoretical argument for triconsonantal clusters only
+
+\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}, triconsonantal clusters only]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 4        & 104        & 0.037      & \multirow{2}{*}{0.419} \\
+violating  & 2        & 124        & 0.016                               \\
+\bottomrule
+\end{tabular} 
+\end{example}
+
+%Note that [tʃ] has not been included since it so rarely occurs in clusters.
+
+\subsubsection{ABA clusters}
+
+Finally, \citet[][176]{Pierrehumbert1994} notes ``a lack of clusters with identical first and third elements''.
+For the determination of identity, I assume the same definition of geminate as used in the rule of \textsc{Degemination} above, namely consonants which share all the same \textsc{Place} and manner features, ignoring \textsc{Voice}.
+%and I also expand the definition of the constraint to also apply to the second and fourth consonant of quadraconsonantal clusters. 
+\citeauthor{Pierrehumbert1994} notes one potential exception to this generalization, the cluster in words such as \emph{e}[ks.kl]\emph{ude}, which contains two /k/, does not arise here since such words have been marked morphologically complex.
+Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly rarer than other triconsonaontal and quadraconsonantal clusters, of which only 7.8\% occur.
+
+\begin{example}[Lexical effects of \textsc{*aba}]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 41       & 478        & 0.079      & \multirow{2}{*}{0.818} \\
+violating  &  0       &  11        & 0.000                               \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}.
+
+%\subsubsection{Summary}
+%
+%The above results are summarized in Table \ref{constraints}.
+%
+%\begin{table}
+%\centering
+%\begin{tabular}{l r r r r r}
+%\toprule
+%                                       & \multicolumn{2}{c}{attested} & \multicolumn{2}{c}{unattested} & \multirow{2}{*}{$p$-value} \\
+%                                       & conforming & violating & conforming & violating \\
+%\midrule
+%\textsc{O.V.A.} (\ref{ovar})        &  80 &   6 & 370 & 264 & 1.1\e{-11} \\
+%\textsc{C.N.P.A} (\ref{CNPA})       &  41 &  25 &  12 &  42 & 1.7\e{-05} \\ 
+%\textsc{Degemination} (\ref{degem}) & 161 & 632 &   0 &  87 & 1.4\e{-08} \\
+%\midrule
+%\textsc{*Dorsal-Labial} ()          &  68 &   4 & 246 &  40 & 0.069 \\
+%\textsc{*C.C.O.} ()                 &  48 &  38 & 312 & 322 & 0.301 \\
+%\textsc{*A,BA} ()                   & 161 &   0 & 708 &  11 & 0.231 \\
+%\bottomrule
+%\end{tabular}
+%\caption{Counts and Fisher exact test $p$-values for the constraints discussed above.}
+%\label{constraints}
+%\end{table}
+
+\subsection{Computational models}
+
+Applying these derived constraints to the lexicon increases saturation from 18.8\% to 28.8\% and incurs only a handful of exceptions.
+Yet, despite the fact that ``attestation'' is the minority pattern, attempts to identify static lexical constraints, whether by hand or computational model, lend little additional predictive power.
+There is no evidence that the English syllable contact inventory is subject to any static constraint at all.
+I am forced to conclude that most, if not all, of 71.2\% of possible clusters which are unattested are accidental gaps.
+Below, I show that this state of affairs follows directly from the sparse distribution of codas and onsets.
+
+The three alternations targeting English syllable contact clusters have been shown to have the expected robust effects on the English lexicon.
+
+Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}.
+There is now a growing literature on computational cognitive models of phonotactic learning.
+By applying these models to the syllable contact data, it is possible to exhaustively search for statistically motivated constraints, elimianting the possibility of human error.
+
+To evaluate these statistical models, they are first trained on the set of attested clusters, then put to the task of classifying all possible clusters into those which are and are not attested.
+Four metrics are used to evaluate these computational models.
+Classification accuracy corresponds to the probability of a classification being either a true positive ($t.p.$), correctly predicting a cluster to occur, or a true negative ($t.n.$), correctly predicting a cluster to be absent.
+Incorrect outcomes are either false positives ($f.p.$), incorrectly predicting a cluster to occur, and false negatives ($f.n.$), incorrectly predicting a cluster to be absent.
+
+%Binary classification contingency table: \\
+%\begin{tabular}{c | l l} \toprule
+%                             & \multicolumn{2}{c}{prediction} \\
+%\midrule
+%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
+%                             & false negative ($fn$) & true negative ($tn$) \\
+%\end{tabular}
+%\begin{tabular}{|l l|}
+%\toprule
+%true positive ($tp$)  & false positive ($fp$) \\
+%false negative ($fn$) & true negative ($tn$) \\
+%\bottomrule
+%\end{tabular}
+
+\begin{unlabeledexample}
+$\displaystyle \textrm{Accuracy} = \frac{t.p. + t.n.}{t.p. + t.n. + f.p. + f.n.}$
+\end{unlabeledexample}
+
+\noindent A few other metrics can be used to assess classification performance. Precision corresponds to the probability that a cluster predicted to occur is observed in the data.
+
+\begin{unlabeledexample}
+$\displaystyle \textrm{Precision} = \frac{t.p.}{t.p. + f.p.}$ 
+\end{unlabeledexample}
+
+\noindent Recall (also known as sensitivity) is the probability a observed cluster is predicted to occur.
+
+\begin{unlabeledexample}
+$\displaystyle \textrm{Recall} = \frac{t.p.}{t.p. + f.n.}$
+\end{unlabeledexample}
+
+It is possible to maximize recall at the expense of precision, by predicting more clusters to be attested, or to increase precision at the expense of recall by predicting fewer clusters to be attested.
+A common way to balance these concerns is the $F_1$ score, which is the harmonic mean of precision and recall.
+
+\begin{unlabeledexample}
+$\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$ 
+\end{unlabeledexample}
+
+\subsubsection{Null baseline}
+
+The 840 possible clusters have an 18.7\% of saturdaion rate.
+As a consequence of this sparsity, a null model which classifies all clusters as unattested achieves 81.3\% accuracy without the need to posit any constraints either derived or static.
+Since this model does not have any positive classifications, either true or false, precision, recall, and $F_1$ are undefined.
+
+\subsubsection{Alternations}
+
+As was shown above, \textsc{Obstruent Voice Assimiation}, \textsc{Coda Nasal Place Assimilation}, and \textsc{Degemination} are reliable predictors of what clusters are and are not attested and have few lexical exceptions.
+I generate another baseline by predicting only those clusters which do not violate any of these generalizations to be attested.
+This provides a significant improvement to the null baseline (accuracy = 0.857).
+As precision (0.401) is smaller than recall (0.708), there are more false positives (i.e., unattested clusters that do not violate any of these three alternation-derived generalizations) than there are false negatives (i.e., attested exceptions to these derived generalizations).
+
+\subsubsection{\citet{Pierrehumbert1994}}
+
+\citet{Pierrehumbert1994} first proposes to measure the well-formedness of syllable contact clusters by the independent probabilities of codas and onsets, and \citet{Coleman1997} extend this model to whole words.
+\citet{Pierrehumbert1994} uses both medial and initial/final codas and onsets to compute these probabilities, but I obtain a better fit to the data by only using medial coda and onset frequencies.
+Medial coda and onset probabilities are multipled to obtain a well-formedness score.
+This produces a continuous value between 0 and 1.
+To derive a binary classification from these values, I use a soft-margin support vector machine \citep{Cortes1995} with a linear kernel.
+This technique derives a ``decision stump'' providing the best single split into attested and an unattested clusters.
+This is not put forth as a cognitively plausible model for relating this expected probability to attestation: it is simply the upper bound that such a model could reach.
+
+Unfortunately, this score only provides a small improvement to the null baseline
+ (accuracy = 0.839), and has a lower accuracy and $F_1$ (0.328) than the alternation baseline.
+
+\subsubsection{\citet{Hayes2008a}}
+
+\citeauthor{Hayes2008a} develop a sophisticated method of learning phonotactic distributions from positive data.
+They develop software for estimating probability distributions over sequences of (sets of) phonological features using the principle of maximum entropy and making few assumptions about the independence of features.
+As above, a support vector machine with a linear kernel is used to turn these probability values into binary predictions.
+The software provided by these authors exposes many ``metaparameters'', operational decisions to be made in conducting an experiment with this system.
+To avoid any potential bias induced by the choice of these metaparameters, I use the same settings as used by \citet{Hayes2008a} in their similar study of English onset consonants.
+This includes the ``accuracy schedule'' [0.001, 0.01, 0.1, 0.2, 0.3] and no constraints on the number of constraints learned, as well English consonantal features used by \citeauthor{Hayes2008a}.
+Since this model is inherently stochastic, producing slightly different outcomes on each run, I follow the practice of \citeauthor{Hayes2008a} (p.~396) and report the lowest-accuracy of ten runs, though in general there is not a great deal of variation between individual runs.
+
+This model provides a small improvement in accuracy (0.877) and $F_1$ (0.703) compared to the alternation baseline, but it has lower recall (0.642) than this baseline, indicating that the \citeauthor{Hayes2008a} model introduces many false negatives.
+%; it assigns the lowest probability value to 122 unattested clusters.
+More specifically, it suggests that this model prefers narrower phonotactic generalizations than those derived from alternations.
+For instance, unattested *[m.kl] is assigned the highest probability, whereas it is ruled out by \textsc{Coda Nasal Place Assimilation}.
+On the other hand, the \citeauthor{Hayes2008a} model assigns a low score to clusters like those found in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.
+
+\citet{Berent2012} identify an important defect with the \citeauthor{Hayes2008a} model used here: it does not make use of variables over features and thus fails to unify the constraint against geminate clusters.
+While the model may miss this unifying generalization, it correctly predicts all clusters containing geminates to be unattested.
+
+The results for the four models are summarized below.
+
+\begin{example}[Summary of computational models]
+\begin{tabular}{l | r r r r}
+\toprule
+                          & accuracy & precision & recall & $F_1$ \\ 
+\midrule
+Null baseline             & 0.813    & n.a.      & n.a.   & n.a.  \\
+Alternations              & 0.857    & 0.401     & 0.708  & 0.512 \\
+\citet{Pierrehumbert1994} & 0.839    & 0.210     & 0.750  & 0.328 \\
+\citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+One additional model was evaluated for this study.
+\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
+However, \citeauthor{McGowan2011} finds that sonority of clusters accounts for only a tiny amount of the variance in cluster frequency ($R^2 = 0.077$).
+In this study, inclusion of sonority distance provided no improvement over the null baseline, and therefore was not considered further.
+
+\subsection{Computational models}
+
+Applying these derived constraints to the lexicon increases saturation from 18.8\% to 28.8\% and incurs only a handful of exceptions.
+Yet, despite the fact that ``attestation'' is the minority pattern, attempts to identify static lexical constraints, whether by hand or computational model, lend little additional predictive power.
+There is no evidence that the English syllable contact inventory is subject to any static constraint at all.
+I am forced to conclude that most, if not all, of 71.2\% of possible clusters which are unattested are accidental gaps.
+Below, I show that this state of affairs follows directly from the sparse distribution of codas and onsets.
+
+The three alternations targeting English syllable contact clusters have been shown to have the expected robust effects on the English lexicon.
+
+Whereas the three alternations targeting English syllable contact clusters have robust lexical reflexes as measured by the Fisher exact test, such tests reveal no evidence for the static dispreferences identified by \citeauthor{Pierrehumbert1994}.
+There is now a growing literature on computational cognitive models of phonotactic learning.
+By applying these models to the syllable contact data, it is possible to exhaustively search for statistically motivated constraints, elimianting the possibility of human error.
+
+To evaluate these statistical models, they are first trained on the set of attested clusters, then put to the task of classifying all possible clusters into those which are and are not attested.
+Four metrics are used to evaluate these computational models.
+Classification accuracy corresponds to the probability of a classification being either a true positive ($t.p.$), correctly predicting a cluster to occur, or a true negative ($t.n.$), correctly predicting a cluster to be absent.
+Incorrect outcomes are either false positives ($f.p.$), incorrectly predicting a cluster to occur, and false negatives ($f.n.$), incorrectly predicting a cluster to be absent.
+
+%\ex Binary classification contingency table: \\
+%\begin{tabular}{c | l l} \toprule
+%                             & \multicolumn{2}{c}{prediction} \\
+%\midrule
+%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
+%                             & false negative ($fn$) & true negative ($tn$) \\
+%\end{tabular}
+%\begin{tabular}{|l l|}
+%\toprule
+%true positive ($tp$)  & false positive ($fp$) \\
+%false negative ($fn$) & true negative ($tn$) \\
+%\bottomrule
+%\end{tabular}
+%\xe
+
+It is possible to maximize recall at the expense of precision, by predicting more clusters to be attested, or to increase precision at the expense of recall by predicting fewer clusters to be attested.
+A common way to balance these concerns is the $F_1$ score, which is the harmonic mean of precision and recall.
+
+\subsubsection{Null baseline}
+
+The 840 possible clusters have an 18.7\% of saturdaion rate.
+As a consequence of this sparsity, a null model which classifies all clusters as unattested achieves 81.3\% accuracy without the need to posit any constraints either derived or static.
+Since this model does not have any positive classifications, either true or false, precision, recall, and $F_1$ are undefined.
+
+\subsubsection{Alternations}
+
+As was shown above, \textsc{Obstruent Voice Assimiation}, \textsc{Coda Nasal Place Assimilation}, and \textsc{Degemination} are reliable predictors of what clusters are and are not attested and have few lexical exceptions.
+I generate another baseline by predicting only those clusters which do not violate any of these generalizations to be attested.
+This provides a significant improvement to the null baseline (accuracy = 0.857).
+As precision (0.401) is smaller than recall (0.708), there are more false positives (i.e., unattested clusters that do not violate any of these three alternation-derived generalizations) than there are false negatives (i.e., attested exceptions to these derived generalizations).
+
+\subsubsection{\citet{Pierrehumbert1994}}
+
+\citet{Pierrehumbert1994} first proposes to measure the well-formedness of syllable contact clusters by the independent probabilities of codas and onsets, and \citet{Coleman1997} extend this model to whole words.
+\citet{Pierrehumbert1994} uses both medial and initial/final codas and onsets to compute these probabilities, but I obtain a better fit to the data by only using medial coda and onset frequencies.
+Medial coda and onset probabilities are multipled to obtain a well-formedness score.
+This produces a continuous value between 0 and 1.
+To derive a binary classification from these values, I use a soft-margin support vector machine \citep{Cortes1995} with a linear kernel.
+This technique derives a ``decision stump'' providing the best single split into attested and an unattested clusters.
+This is not put forth as a cognitively plausible model for relating this expected probability to attestation: it is simply the upper bound that such a model could reach.
+
+Unfortunately, this score only provides a small improvement to the null baseline
+ (accuracy = 0.839), and has a lower accuracy and $F_1$ (0.328) than the alternation baseline.
+
+\subsubsection{\citet{Hayes2008a}}
+
+\citeauthor{Hayes2008a} develop a sophisticated method of learning phonotactic distributions from positive data.
+They develop software for estimating probability distributions over sequences of (sets of) phonological features using the principle of maximum entropy and making few assumptions about the independence of features.
+As above, a support vector machine with a linear kernel is used to turn these probability values into binary predictions.
+The software provided by these authors exposes many ``metaparameters'', operational decisions to be made in conducting an experiment with this system.
+To avoid any potential bias induced by the choice of these metaparameters, I use the same settings as used by \citet{Hayes2008a} in their similar study of English onset consonants.
+This includes the ``accuracy schedule'' [0.001, 0.01, 0.1, 0.2, 0.3] and no constraints on the number of constraints learned, as well English consonantal features used by \citeauthor{Hayes2008a}.
+Since this model is inherently stochastic, producing slightly different outcomes on each run, I follow the practice of \citeauthor{Hayes2008a} (p.~396) and report the lowest-accuracy of ten runs, though in general there is not a great deal of variation between individual runs.
+
+This model provides a small improvement in accuracy (0.877) and $F_1$ (0.703) compared to the alternation baseline, but it has lower recall (0.642) than this baseline, indicating that the \citeauthor{Hayes2008a} model introduces many false negatives.
+%; it assigns the lowest probability value to 122 unattested clusters.
+More specifically, it suggests that this model prefers narrower phonotactic generalizations than those derived from alternations.
+For instance, unattested *[m.kl] is assigned the highest probability, whereas it is ruled out by \textsc{Coda Nasal Place Assimilation}.
+On the other hand, the \citeauthor{Hayes2008a} model assigns a low score to clusters like those found in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.
+
+\citet{Berent2012} identify an important defect with the \citeauthor{Hayes2008a} model used here: it does not make use of variables over features and thus fails to unify the constraint against geminate clusters.
+While the model may miss this unifying generalization, it correctly predicts all clusters containing geminates to be unattested.
+
+The results for the four models are summarized below.
+
+\begin{example}
+Summary of computational model accuracy:
+
+\vspace{0.5\baselineskip} 
+\begin{tabular}{l | r r r r}
+\toprule
+                          & accuracy & precision & recall & $F_1$ \\ 
+\midrule
+Null baseline             & 0.813    & n.a.      & n.a.   & n.a.  \\
+Alternations              & 0.857    & 0.401     & 0.708  & 0.512 \\
+\citet{Pierrehumbert1994} & 0.839    & 0.210     & 0.750  & 0.328 \\
+\citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+One additional model was evaluated for this study.
+\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
+However, \citeauthor{McGowan2011} finds that sonority of clusters accounts for only a tiny amount of the variance in cluster frequency ($R^2 = 0.077$).
+In this study, inclusion of sonority distance provided no improvement over the null baseline, and therefore was not considered further.
+
+\subsection{Statistical properties of syllable contact}
+
+Constraints on
+
+Above, we have seen that constraints on URs which are derive from phonological alternations restrict the contents of the lexicon.
+
+Applying these derived constraints to the lexicon increases saturation from 18.8\% to 28.8\% and incurs only a handful of exceptions.
+Yet, despite the fact that ``attestation'' is the minority pattern, attempts to identify static lexical constraints, whether by hand or computational model, lend little additional predictive power.
+There is no evidence that the English syllable contact inventory is subject to any static constraint at all.
+I am forced to conclude that most, if not all, of 71.2\% of possible clusters which are unattested are accidental gaps.
+Below, I show that this state of affairs follows directly from the sparse distribution of codas and onsets.
+
+\subsubsection{The $Z_r$ transform}
+
+In a sparse distribution, many types occur at the same low frequencies.
+\citet{Good1953} notes that this produces a type of quantization, an artificially long and flat right tail in the frequency spectrum.
+\citet[][29]{Church1991} propose a method to eliminate this quantization, the $Z_r$ transform, defined as follows.
+Let $r, n$ be defined such that $n_i$ is the number of types which occur at frequency $r_i$ (i.e., $n$ contains frequencies of type frequencies).
+Further, let $r$ be sorted in increasing order.
+$Z$ is simply each element of $n$ scaled by the nearest points to the left and right.
+
+\begin{unlabeledexample}
+$\displaystyle Z_i = \frac{2 n_i}{r_{i + 1} - r_{i - 1}}$
+\end{unlabeledexample}
+
+\noindent \citeauthor{Church1991} do not define this transform for the lowest and highest points (i.e., when $i = 1$ or $i = N$), and so I extend the definition by scaling these points using only one inward-facing point.
+
+\begin{unlabeledexample}
+$\displaystyle Z_1 = \frac{n_1}{r_2 - r_1}$
+\end{unlabeledexample}
+
+\begin{unlabeledexample}
+$\displaystyle Z_N = \frac{n_N}{r_N - r_{N - 1}}$
+\end{unlabeledexample}
+
+%\indent
+%The effect of this transform can be seen in Figure \ref{zr}.
+
+%\begin{figure}
+%\centering
+%\includegraphics{zr.pdf}
+%\caption{The SUBTLEX word frequency norms \citep{Brysbaert2009} show a long right tail in the log-frequency/log-rank space (left panel), and the $Z_r$ transform (right panel) smooths out this tail.}
+%\label{zr}
+%\end{figure}
+
+\subsubsection{Coda, onset, and cluster sparsity}
+
+A number of previous studies \citep[e.g.,][]{Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} have observed that phonemes and graphemes exhibit sparse type (i.e., lexical) and token frequency distributions.
+The lexical frequencies of the medial codas ond onsets which make up syllable contact clusters show a similar distribution.
+In Figure \ref{codaonset}, $Z_r$-transformed lexical frequencies of codas and onsets are plot against rank, in log-log spaces.
+The near-linear relationship that obtains indicates that the data can be modeled by a generalization of Zipf's law \citep{Zipf1949}.
+
+%\begin{figure}
+%\centering
+%\includegraphics{coda.pdf} 
+%\includegraphics{onset.pdf}
+%\caption{Medial coda and medial onset lexical frequency exhibit the Zipfian log-log-linear relationship between rank and frequency.}
+%\label{codaonset}
+%\end{figure}
+
+\begin{unlabeledexample}
+$\displaystyle f(r; C, \alpha) = \frac{C}{r^\alpha}$ 
+\end{unlabeledexample}
+
+\noindent The parameters $C$, a constant sensitive to sample size, and $\alpha$, the slope of the rank/frequency relationship in log-log space, are easily computed using the method of least squares in a linear regression, where $\epsilon$ represents the error term.
+
+\begin{example}[example]
+$\displaystyle \textrm{log}~Z_r \sim C + \alpha~\textrm{log}~r + \epsilon$ 
+\end{example}
+
+\noindent Codas ($\alpha = -0.720$, $R^2 = 0.749$) and onsets ($\alpha = -1.056$, $R^2 = 0.828$) are well-fit by this distribution, though there are some outliers.
+Similarly, clusters, shown in Figure \ref{clus}, are also a good fit to the generalized Zipf's law ($\alpha = -0.588$, $R^2 = 0.924$).
+
+\begin{figure}
+\centering
+\includegraphics{cluster.pdf}
+\caption{Syllable contact clusters, themselves composed of Zipfian medial codas and medial onsets, also exhibit a Zipfian distribution.}
+\label{clus}
+\end{figure}
+
+That clusters (and the codas and onsets that make them up conforms to Zipf's law is not itself a deep observation.
+Zipfian distributions are characteristic of both syntactic rules \citep{Yang2009}, but also word \citep{Teahan1998,Ha2002,Baroni2009} and phoneme \citep{Daland2011a} n-grams, and tokens in non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} or randomly-generated texts \citep{Miller1957,Li1992}.
+The import of this statistical property for the study of cluster phonotactics is that it makes it difficult, on statistical grounds alone, to determine which unobserved events are excluded, and which are accidentally missing.
+
+\citet{Good1953} puts forth a simple theory of just how frequent unobserved events might be.
+He proposes to estimate $\hat{p}_0$, the probability of all unseen events, as the ratio of events which occur just once $n_1$, to the number of observations, $\sum N$.
+This quantity is known as the Good-Turing estimate, since it was originally developed by Alan Turing for codebreaking during the second World War.
+
+\begin{unlabeledexample}
+$\displaystyle \hat{p}_0 = \frac{n_1}{\sum N}$ 
+\end{unlabeledexample}
+
+\noindent In the CELEX data, $65$ clusters occur only once, and there are 873 clusters in all, so $\hat{p}_0 = 0.074$.
+One way to interpret this quantity is with respect to a ``replication''.
+If it were possible to generate a new corpus of syllable contact clusters, approximately 7\% of cluster tokens will be ones that were not attested in the previous sample, i.e., they accidentally failed to be sampled.
+This is a not-inconsiderable quantity.
+
+Unfortunately, the Good-Turing estimate does not provide a way to determine which unattested clusters might be found in a replication.
+But I have already proposed that the null hypothesis of generative phonology, combined with the principle of Stampean occultation, provides a way out of this conundrum.
+Unattested clusters like *[m.kl] are structurally excluded (since such a cluster is occulted by \textsc{Coda Nasal Place Assimilation}), whereas *[s.l] is probably an accidental gap.
+
+\subsubsection{Sampling simulation}
+
+As a final investigation into this data, I consider the statistical properties of a randomly generated novel ``lexicon''.
+The simulated lexicon is generated by  generating a distribution of clusters, by repeatedly applying the following procedure.
+
+\begin{example}[Simulation procedure]
+\begin{tabular}{l l}
+a. & Sample a medial coda according to the observed probabilities  \\
+b. & Sample a medial onset according to the observed probabilities \\
+c. & Apply the SPE rules to the resulting cluster                  \\
+\end{tabular}
+\end{example}
+
+\noindent This is a simulation version of the model proposed by \citet{Pierrehumbert1994} and \citet{Coleman1997}, but also takes occulting phonology into account.
+
+While this procedure can be repeated indefinitely 
+%(using the CELEX data, listed in Appendix \ref{A}), 
+consideration of a single characteristic run of the simulation should suffice.
+The saturation rate of the simulated sample (18.2\%) is very close to the real sample (18.8\%).
+The slope of the log-rank/log-frequency relationship for the simulated data ($\alpha = -0.557$) is also very close to the observed slope for the CELEX data ($\alpha = -0.588$).
+And, as shown in Figure \ref{sim}, the two distributions are nearly indistinguishable.
+
+\begin{figure}
+\centering
+\includegraphics{sim.pdf}
+\caption{A comparison of the observed lexical frequencies of English syllable contact clusters to an ``English lexicon'' simulated by combining.}
+\label{sim}
+\end{figure}
+
+It is important to note that the results of this simulation should not be construed as an endorsement of a revised version of the \citet{Pierrehumbert1994} and \citet{Coleman1997} model which also incorporates occultation.
+In fact, as was shown above, the independent frequencies of coda and onset are very poor predictors of what clusters will be attested even when occultation is taken into account, and the set of attested clusters only partially overlaps those produced by the simulation.
+I propose that this is because the lexicon is not only finite, but far smaller than the combinatoric possibilities provided by phoneme sequences, making accidental gaps unavoidable.
+That these gaps will be arbitary from the perspective of phonology is nothing more than a corrolary of the principle of \emph{l'arbitraire du signe}.
+
+It is quite possible that phonology (and, perhaps, static phonotactics) acts as a filter on what URs are learned, but a cluster might be overrepresented or underrepresented for reasons that are quite external to phonology, e.g., language contact or historical change.
+Similarly, the factors that cause a /n/ to be a common word-medial coda in English may be independent of the propensity of that coda to combine with medial onset, assuming the resulting cluster is phonologically licit.
+The import of this simulation is somewhat more subtle; it simply shows free combination of codas and onsets according to their independent probabilities, constrained only by occultation, would produce the same sparse distribution of clusters that was used to argue for static phonotactics.
+
+%\citet{Evert2004}
+%\includegraphics{zipf.pdf}
+
+\section{Conclusions}
+
+\citet{Borowsky1989} on peripherality.
+
+productive \citet{Duanmu2008}
+
+The above is a stark reminder.
+
+A reasonable objection to the arguments presented in this chapter is to view the results as little more than an indictment of the results of \citealt{Pierrehumbert1994}.
+
+However, the fact that state-of-the-art models are not capable of providing large improvements to the predictive accuracy indicates
+
+ do not have a statistically significant effect on the shape of the English lexicon, but that experiments might turn up evidence that speakers have internalized
+
+shown to be aware of static constraints if they reached statistical significance.
+
+ statistically reliable static constraints could be identified
+
+The of
+
+Regarding the historical developments,
+
+\citet{Martin2007}
+
+On the other hand, patterns created by sound change are not guaranteed to persist over time.
+One example of non-persistence is discussed by \citet{Iverson2005}.
+Around 1100 CE, Old English \emph{sk} became [ʃ].
+This sound change introduced no alternations.
+Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{V\lm sk\#} words when the change was actuated, and \emph{V\lm sh\#} continues to be rare in Modern English.
+What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.
+
+\citet[][140]{Frisch2004} suggest that the strong tendency for the first and second consonants of the Arabic root to be non-identical is the ``a diachronic result of a processing constraint that disfavors repetition.'' Unfortunately, there is no evidence that this pattern is diachronic other than in the sense that it appears to be inherited from the proto-language: there is simply no Proto-Semitic verb roots with identical first and second consonants \citep[][178]{Greenberg1950}.
+In other Semitic languages, the inherited patern has experienced considerable erosion.
+
+\begin{example}[Tigrinya roots with identical first and second consonants \citep{Buckley1990a}]
+\begin{tabular}{l l l l}
+a. & lʌlʌw     & `scorch'                   & (\textless{} Ge'ez \emph{lʌwlʌw} `inflame')     \\
+   & mʌmʌy     & `winnow'                   & (\textless{} Ge'ez \emph{mʌymʌy} `distinguish') \\
+   & mʌmʌt     & `pick out loot'            \\
+b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
+   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
+c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
+   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
+\end{tabular}
+\end{example}
+
+Similar exceptions are found in 
+%Amharic (\citealp[][?]{Broselow1984}, \citealp[][?]{McCarthy1985}) and 
+Hebrew \citep[][29]{Bat-El2005}.
+
+The next two chapters return to the question of synchrony, addressing the relationship between statistical patterns in the lexicon and speakers' behaviors when presented with underrepresented sequences.
+
+accidental gaps
+\citet[][419f.]{Hayes2008a}

### 2012-06-03 1.tex
Chapter 2 draft done.

@@
-\chapter{Sequence structure constraints in generative grammar}
+%\chapter{Sequence structure constraints in generative grammar}
 \label{msc}
@@
 In the reflexive infinitive, /\ldots t\pal-s\ldots/ is realized [ts], with no palatalization.
-However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\ref{russian}c)---is preserved in the reflexive imperative.
-
-\begin{example}
-\label{russian}
-Russian reflexives \citep[after][]{Jakobson1932}:\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses an abstract phonemic transcription.
+However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\ref{russian}c)---is preserved in the reflexive imperative.\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses an abstract phonemic transcription.
 Thanks to Lev Blumenfeld (p.c.) for help with the transcription of this data.}
 
+\begin{example}[Russian reflexives (after \citealt[][]{Jakobson1932}] \label{russian}
 \begin{tabular}{l l l l} %\toprule

### 2012-06-03 2.tex
Chapter 2 draft done.

--- 2012-04-18 2.tex
@@
 
-(though few if any phonologists regard $n$-gram segment probabilities as a core component of phonology)
+\citet{Halle1962} and \citet{Chomsky1965} observe that speakers internalize generalizations about possible and impossible words in their language.
+\citeauthor{Chomsky1965} illustrate this knowledge with two nonce words \emph{blick} [blɪk] and \emph{bnick} [bnɪk]; whereas \emph{blick} is unattested, but possible word of English,
+speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word of the language.
 
-%\begin{quotation}
-%``...the greatest defect of classical philosophy of the mind...seems to me to be its unquestioned assumption that the properties and content of the mind are accessible to introspection.'' \citep[][22]{Chomsky1968}
-%\end{quotation}
+The classic account of this contrast is as follows.
+Segments must be assigned to prosodic structures like the syllable, or undergo phonological repair (e.g., \citealt{Hooper1973}:10f., \citealt{Ito1989a}, \citealt{Noske1992}).
+In English, [bl] is a permissible onset, but unlike some languages (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet'), [bn] is not.
+\citet[][19f.]{Wolf2009} suggest, for instance, that an underlying /bnɪk/ would be realized as [nɪk]; it is also possible to imagine resolution by prothesis---[əbnɪk]---or anapytxis---[bənɪk], and indeed \citet{Davidson2006b} finds that English speakers produce these repairs when attempting to mimic foreign pronunciations of obstruent-nasal onsets.
 
-\input{2.1}
-\input{2.2}
-\input{2.3}
+Many linguists view this account as overly simpliest, for, as \citeauthor{Shademan2006} argues, it draws a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report wordlikeness intuitions.
+
+\begin{quote}
+A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
+But when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values.
+\citep[][371]{Shademan2006}
+\end{quote}
+
+This chapter focuses on the hypothesis that speakers' knowledge of wordlikeness is gradient.
+\S\ref{gradhyp} places this debate in a historical perspective, and argues that the naïve but widely accepted form of this hypothesis fails the test of falsifiability.
+In \S\ref{2evaluation}, a revision of this hypothesis is subjected to an quantitative evaluation and is found wanting.
+
+[Explain that there's not going to be any issue of, e.g., heterovoiced obstruents, that stuff is all in chap. 4]
+
+\section{The gradience hypothesis} \label{gradhyp}
+
+%In this section, the gradience hypothesis is reviewed from historical and scientific perspectives.
+
+\subsection{A brief history of the gradience hypothesis} \label{history}
+
+\citet{Chomsky1965} were not the first to consider contrasts between possible and impossible words.
+Their primary contribution to this question, is, essentially, mentalism: they recognize that naïve speakers effortlessly acquire language-specific wordlikeness generalizations and can report them with ease.
+But the nature of these contrasts were also considered by structuralists.
+In an early study, \citet{Fischer-Jorgensen1952} proposes that possible and impossible words represent endpoints on a continuum, and that no non-arbitrary line can be drawn between them.
+While it would anachronistic to assign this a mentalistic interpretation, this appears to be the earliest proposal that wordlikeness contrasts are in some sense gradient.
+
+\subsubsection{Early generative grammar}
+
+Gradience is also discussed in the earliest work in generative syntax.
+\citet[][132]{LSLT} writes that ``there is little doubt that speakers can fairly consistently order new utterances, never previously heard, with respect to their degree of `belongingness' to the language''.
+\citet{ASPECTS} proposes to link different degrees of ungrammaticality to different kinds of syntactic violations.
+Unlike \citeauthor{Fischer-Jorgensen1952}, there is for \citeauthor{LSLT} a hard-and-fast boundary between the grammatical and ungrammatical, only the latter showing any gradience.
+For \citeauthor{LSLT}, grammatical utterances are all equally grammatical, but ungrammatical utterances may be deviant in their their own ways \citep[][61]{Schutze1996}.
+
+However, most of the early literature on wordlikeness concerns apparently categorical contrasts the possible and impossible.
+\citet[][31]{Vogt1954} recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
+Generalizing somewhat from \citeauthor{Vogt1954}'s discussion, allophony may account for the absence of certain phone sequences, but it does not provide a suitable explanation for the absence of initial [bn] in English, nor does it make correct predictions about the surface realization of an underlying initial /bn/.
+\citeauthor{Vogt1954} concludes that additional grammatical machinery will be needed to account for possible and impossible words.
+\citet{Halle1962} and \citet{Stanley1967} propose to derive wordlikeness effects from morphophonemic redundancy rules acting on underlying representations.
+The discussion of \emph{bnick} by \citet[][101]{Chomsky1965} is a good example of this type of analysis.
+They correctly observe that before a word-initial stop, a English consonant must be liquid.
+A side effect of a redundancy rule which specifies a consonant in this position as [$+$\textsc{Liquid}] precludes the derivation of \emph{bnick}.\footnote{This is not the only redundancy that might be used to rule out \emph{bnick}; for instance, the only word-initial obstruent that may follow a nasal in English is /s/.}
+
+\subsubsection{\emph{The sound pattern of English}}
+
+In \emph{The sound pattern of English} (SPE), \citet{SPE} extend this model of wordlikeness to account for
+
+[Explain the SPE model]
+
+\subsubsection{Autosegmental phonology and beyond}
+
+The \emph{SPE} model was not widely adopted, and research on wordlikeness turned to other principles.
+Syllabification plays no role in \emph{SPE} theory, though the syllable is found in many earlier studies (see \citealp{Goldsmith2011b}).
+\citet{Hooper1973} and \citet{Kahn1976} argue that wordlikeness generalizations argue for the need for syllabification as part of the phonological computation.
+This classic syllabification-based model is further enriched by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as envisioned by][]{Pike1947a}, and the theory of prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
+
+\subsection{A naïve gradience axiom}
+
+Recent critiques of this syllabification-based model focus on the existence of intermediate wordlikeness ratings (see also \citealt{Coleman1997}, \citealt{Anttila2008}).
+
+\begin{quote}
+In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
+\citep[][382]{Hayes2008a}
+
+When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing.
+In the usual case, novel items fall along a gradient cline of acceptability.
+\citep[][9]{Albright2009a}
+\end{quote}
+
+The observation itself cannot be questioned, but \citeauthor{Hayes2008a} and \citeauthor{Albright2009a} do not explicitly state why this data is relevant to the construction of models of wordlikeness.
+The following seems the most likely: \citeauthor{Hayes2008a} and \citeauthor{Albright2009a} believe this proves that wordlikeness, as an internal state, is gradient simply because subjects make use of intermediate degrees of wordlikeness in judgement tasks.
+This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, and an influential view in the cognitive sciences (see \citealt{Fodor1981a} for a critique).
+
+\begin{unlabeledexample}
+\textsc{Naïve Gradience Proposition}: If graded judgements of a concept or mental state use intermediate ratings, that concept or mental state is gradient
+\end{unlabeledexample}
+
+This proposition is an indicative conditional.
+While the antecedent of this condition, the presence or absence of intermediate ratings in a graded judgement task, is readily observable, it has no bearing on the truth or falsity of the conditional proposition.
+However, whereas a false consequent (e.g., an all-or-nothing concept) would potentially falsify the proposition, the consequent, a mental state, cannot be directly observed.
+If this is the case, then both the proposition and the consequent (the proposition that wordlikeness is itself gradient) fail the test of falsification, placing it beyond the reach of scientific inquiry.
+
+\citet{Armstrong1983} cut the Gordian knot by asserting the possibility of observing a false consequent; they find the existence of all-or-nothing concepts like ``odd number'' or (more controversially) ``female'' to be apparent.
+
+\begin{quote}
+Are there definitional concepts?
+Of course.
+For example.
+consider the concept \emph{odd number}.
+This seems to have a clear definition, a precise description\ldots{}No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd.
+No odd number seems odder than any other odd number.
+\citep[274]{Armstrong1983}
+\end{quote}
+
+\citeauthor{Armstrong1983} note that if subjects produce intermediate ratings for these categories, then the naïve proposition is false,
+%(i.e., $p \wedge \neg q \vdash \neg (p rightarrow q)$
+and this is what \citeauthor{Armstrong1983} find in their experiments.
+They ask subjects to rate, on a seven-point scale, the extent to which, e.g., certain odd counting numbers represent the concept ``odd number'', and find that speakers do make use of intermediate values.
+For instance, \emph{7} is rated 1.4 out of 7 on average, whereas the average ratings of \emph{447} and \emph{501} are 3.5 and 3.7, respectively.
+
+In summary, the proposition that intermediate ratings indicate mental gradience is either untestable, or it is false, and is thus rejected.
+
+\subsection{A falsifiable gradience hypothesis}
+
+\citeauthor{Armstrong1983} do not view their results as evidence that subjects have an internal model of odd numbers which is at odds with the formal, all-or-nothing definition, and in fact, several additional experiments show that subjects deploy the extension of the formal definition in other tasks.
+Reviewing this study, \citet[215]{Schutze2011} writes that the experiments show that judgements are ``sensitive to factors other than our underlying competence''.
+In the case of wordlikeness, a case can be made that one such factor is similarity to existing words (e.g., \citealt[][151, fn.
+27]{LSLT}, \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Bailey2001}).
+However, the mere fact that subjects use intermediate ratings does not show that they do so with any systematicity, or that, e.g., the contrast between \emph{447} and \emph{7} with respect could be given any satisfying explanation.
+\citeauthor{Schutze2011} further suggests that gradience responses is a property of tasks, not concepts.
+
+\begin{quote}
+\ldots{}when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different.
+\citep[][215]{Schutze2011}
+\end{quote}
+
+This introduces a troubling possibility, that subjects ``oblige'' experimenters by introducing random noise to their categorical judgements when presented with a Likert task.
+Were this the case, the observed intermediate judgements would scarcely be worthy of modeling.
+The alternative considered here is that there exists some model in which the gradience in ratings is systematic.
+
+\begin{unlabeledexample}
+\textsc{Gradience Hypothesis}: If graded judgements of a concept or mental state use intermediate ratings, there is some model which can predict this behavior
+\end{unlabeledexample}
+
+\noindent There are now many computational models of wordlikeness which proport to do just this.
+An unfortunate defect of prior evaluations of these models, however, is that they adopt the naïve gradience hypothesis.
+\citet[382]{Hayes2008a} are just some of the authors who consider the ability to model gradient intuitions to be so important that they do not include any categorical models in their evaluation.
+As a result, the literature contains no serious attempt to evaluate the gradience hypothesis.
+
+\section{Evaluation} \label{2evaluation}
+
+The remainder of this chapter is devoted to evaluting the gradience hypothesis, using state-of-the-art computational wordlikeness models and data from English.
+
+\subsection{Data sources}
+
+The evaluation makes exclusive use of previously published English wordlikeness data.
+For inclusion in the evaluation, a study must conform to all three of the following conditions.
+First, the stimuli must consist only of monosyllabic words presented auditorily.
+Secondly, a significant portion of the stimuli must contain gross phonotactic violations (e.g., \emph{bnick}).
+Finally, the ratings, averaged across subjects, must be publicly available.
+The large-scale studies by \citet{Bailey2001} and \citet{Shademan2006,Shademan2007} are ineligible because the former lack stimuli with gross phonotactic violations and the latter data are not available to the public in any form.
+
+\subsubsection{\citealt{Greenberg1964}}
+
+\citet{Greenberg1964} investigated wordlikeness using the technique of free magnitude estimation, a mechanism which has become increasingly popular among syntactians \citep[e.g.,][]{Bard1996}.
+At the beginning of the experiment, the subject heard a recording of the word \emph{stick}.
+In subsequent trials, the subjects heard a nonce word and were asked to report ``how far would you say that is from English?'', with \emph{stick} at ``1''; subjects are told that a word that is ``twice as far from English'' as \emph{stick} should be scored ``2''.
+The data used here are from \citeauthor{Greenberg1964}'s Experiment B, in which 17 undergraduates were presented 17 stimuli in all.
+In addition to \emph{stick}, the stimuli include three other English words; these four items were excluded from further analyses, leaving 13 stimuli.
+As is standard practice in psychophysics \citep[e.g.,][]{Butler1987}, these magnitude ratings were log-transformed before analysis.
+
+\subsubsection{\citealt{Scholes1966}}
+
+\citet{Scholes1966} conducted a number of English wordlikeness judgement with middle school children.
+The data used here come from his ``experiment 5'', in which 63 monosyllabic items were presented to 33 seventh-grade students.
+For each stimulus, the subjects produced forced choice ``yes''/``no'' answers to the question of whether the item ``is likely to be usable as a word of English''.
+\citet{Hayes2008a} and \citet{Albright2009a} analyse this data as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus (see also \citealt{Pierrehumbert1994}, \citealt{Coleman1997}, \citealt{Frisch2000}).
+For instance, 22 of the 33 students answered ``yes'' for \emph{shlerk} [ʃlɚk], so it is assigned a score of $0.666$.\footnote{This procedure conflates intraspeaker variation (which may be considerable; see \citealt{Shademan2007}) and speaker-internal gradience, and its adoption here should not be construed as an endorsement.
+It is provided solely for comparison with prior studies using the \citeauthor{Scholes1966} data.} These stimuli are all ostensibly non-words, but include \emph{clung} [klʌŋ], the preterite and past participle of the verb \emph{cling}, and \emph{brung} [brʌŋ], a dialectical past participle for \emph{bring}.
+These two words were excluded and the remaining 61 stimuli were submitted to analysis.
+
+\subsubsection{\citealt{Albright2003b}}
+
+\citet{Albright2003b} gathered wordlikeness judgements to serve as norms for a \emph{wug}-test.
+87 items were presented to 20 undergraduate subjects, who rated each word on a seven-point Likert scale.
+The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``complete normal, would make a fine English word''.
+
+\subsection{Model comparison}
+
+The four models used here consist of a binary baseline and three computationally implemented gradient models.
+These three models are chosen because prior studies have shown they are correlated with wordlikeness judgements.
+Non-parametric rank correlation statistics are used evaluate the correlation between model predictions and wordlikeness ratings.
+Rank correlation are used rather than the parametric Pearson correlation statistic that is used by \citet{Hayes2008a}, for instance, because they make none of the potentially troublesome assumptions of the latter method (see \citealt{Albright2009a}:23, fn.
+12).
+The Spearman $\rho$ is most widely known rank correlation statistic, but it is difficult to give a natural interpretation to this quantity.
+On the other hand, the Kendall $\tau_b$ and Goodman-Kruskal $\gamma$ can be interpreted as fractions of the number of \emph{concordant} and \emph{discordant} pairs \citep{Noether1981}.
+Consider the case here, in which model scores are compared with wordlikeness ratings.
+If a model rates \emph{dresp} [dɹɛsp] more wordlike than *\emph{srest} [sɹɛst], this pair is concordant if speakers also rate \emph{dresp} more English-like than \emph{srest}, and it is discordant if *\emph{srest} is rated more English-like.
+The $\tau_b$ and $\gamma$ differ only in their treatment of ``ties'' (i.e., if e.g., *\emph{dresp} and *\emph{srest} are scored the same, or rated the same); $\tau_b$ includes a correction for ties, whereas $\gamma$ ignores tied pairs.
+Much like the familiar Pearson correlation, $\rho$, $\tau_b$ and $\gamma$ are all in the range [-1, 1].
+Correlations of the four models are given in Table \ref{cor}.
+
+\begin{table} \label{cor}
+\centering
+\begin{tabular}{l l l l l}
+\toprule
+Spearman $\rho$          & baseline         & maxent  & bigram           & density          \\
+\midrule
+\citealt{Greenberg1964}  & {0.845}          & {0.765} & {\textbf{0.863}} & {0.648}          \\
+\citealt{Scholes1966}    & {0.791}          & {0.762} & {\textbf{0.827}} & {0.827}          \\
+\citealt{Albright2003b}  & {0.725}          & {0.429} & {0.708}          & {\textbf{0.742}} \\
+\midrule
+Kendall $\tau_b$         & baseline         & maxent  & bigram           & density          \\
+\midrule
+\citealt{Greenberg1964}  & {\textbf{0.716}} & {0.585} & {0.692}          & {0.462}          \\
+\citealt{Scholes1966}    & {\textbf{0.664}} & {0.597} & {0.652}          & {0.565}          \\
+\citealt{Albright2003b}  & {\textbf{0.599}} & {0.343} & {0.506}          & {0.556}          \\
+\midrule
+Goodman-Kruskal $\gamma$ & baseline         & maxent  & bigram           & density          \\
+\midrule
+\citealt{Greenberg1964}  & {\textbf{1.000}} & {0.684} & {0.692}          & {0.462}          \\
+\citealt{Scholes1966}    & {\textbf{0.995}} & {0.634} & {0.667}          & {0.614}          \\
+\citealt{Albright2003b}  & {\textbf{0.953}} & {0.656} & {0.509}          & {0.575}          \\
+\bottomrule
+\end{tabular}
+\caption{Rank correlations between wordlikeness ratings and phonotactic models surprisingly reveal that the binary baseline meeds or exceeds the coverage of three state-of-the-art phonotactic models.
+All correlations are significant at $p = 0.05$.}
+\end{table}
+
+\subsubsection{Binary baseline}
+
+The binary baseline used here is a crude implementation of the null hypothesis that there are no gradient effects in wordlikeness judgements.
+To create such a baseline, it is necessary to distinguish between nonce words which contain a gross phonotactic violation and those which do not.
+As all stimuli here are monosyllables, this task can be further simplified by separately considering the two major subcomponents of the syllable, the onset and rime.
+Speakers are particularly adept as separating onsets from rimes (\citealt{Treiman1986,Treiman1995}, \citealt{Fowler1993}), and a large portion of phonotactic violations can be localized to one or the other unit (e.g., \citealt{Fudge1969}, \citealt{Treiman2000}).
+
+The baseline considers a nonce syllable to be well-formed if it consists of both an attested onset and an attested rime; the free combination of these two components is the only mechanism by which this model can generalize beyond attested words.
+It is surely the case that this is an ultimately insufficient model: \citet{Albright2009a} observes, for instance, that [ɛsp] is a well-formed rime, even though it is found in no word of English.
+This problem is more acute in languages with more permissive syllable structures than those of English, and thus a sparser lexicon with regard to phonotactic constraints.
+For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.
+A similar result can be found in Chapter 4, which considers the inventory of syllable contact clusters in English.
+% FIXME
+Despite these defects, the baseline outperforms the gradient models in most contexts.
+
+In Figure \ref{density}, the densities of ratings from the three studies, linearly transformed to the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{density}.
+In all three studies, it is possible to discern a relatively sharp separation between valid and invalid clusters, but also the presence of intermediate values.
+
+\begin{figure} \centering \label{density}
+\includegraphics{density.pdf}
+\caption{Average ratings of individual nonce words, linearly transformed to the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
+\end{figure}
+
+\subsubsection{Maximum entropy phonotactics}
+
+\citeauthor{Hayes2008a} (\citeyear{Hayes2008a}; henceforth H\&W) develop a sophisticated model of phonotactic grammaticality which estimates a probability distribution over phoneme sequences by weighing constraints according to the principle of maximum entropy.
+H\&W find that the predictions of their model are closely correlated with the \citet{Scholes1966} wordlikeness ratings.
+A direct replication of their predictions was attempted by using the software, model parameters, and training data as described in that study.
+Since the training of the maximum entropy model is inherently stochastic, producing slightly different outcomes on each run, the lowest scoring of ten runs is reported (H\&W:396), though in general there is not a great deal of variation between individual runs.
+One limitation of this model is that it is not feasible to score whole words, as the number of constraints which must be inspected grows expontentially as their scope increases.
+Following H\&W and of \citet{Albright2009a}, who also applies the maximum entropy model to the \citet{Albright2003b} norms, the model is trained and scored only on stimulus onsets.
+However, as a consequence, the maximum entropy model performs particularly poorly on this data set, as many stimuli contain phonotactic violations in other positions.
+
+\subsubsection{Segment bigram probability}
+
+The bigram probability of a sequence $ijk$ is the product of the probability of an sequence-initial $i$, the probability that $j$ follows $i$, and the probability that $k$ follows $j$, and the product of sequence-final \emph{k}.
+
+\begin{unlabeledexample}
+$\displaystyle \hat{p}(ijk) = p(i|\textrm{start}) \cdot p(j|i) \cdot p(k|j) \cdot p(\textrm{stop}|k)$
+\end{unlabeledexample}
+
+\noindent Bigram models are widely used in natural language processing, and \citet{Albright2009a} considers their relevance to modeling wordlikeness judgements.
+While the focus of \citeauthor{Albright2009a}'s study is on developing a model which uses bigrams over phonological features rather than segments themselves, \citeauthor{Albright2009a}'s evaluation, which includes both the \citeauthor{Scholes1966} and \citeauthor{Albright2003b} data sets, finds an advantage for segmental bigrams.
+
+\citeauthor{Albright2009a} estimates bigram probabilities using the method of maximum likelihood over types in the lexicon.
+The variant of segmental bigrams used here computes probabilities with a simple type of smoothing in which the count of all possible bigrams (including those never observed) are incremented by one.
+This technique is known as Laplacian, or ``add one'' smoothing.
+This has the desirable effect that no nonce word is ever assigned a zero probability, and produces in a small increase in the correlation between the \citeauthor{Albright2003b} wordlikeness norms compared with the maximum likelihood estimate (Table \ref{albrightimproved}).
+For all three data sets, this model also consistenly outperforms ``positional'' probability models implemented by \citet{Vitevitch2004} and \citet{Vaden2009}, and thus these are not considered further.
+The bigram model consistently performs well in all the evaluations, and has the highest Spearman correlation with the \citeauthor{Greenberg1964} and \citeauthor{Scholes1966}, and is frequently second place model to the binary baseline elsewhere.
+
+\begin{table} \label{albrightimproved}
+\centering
+\begin{tabular}{l r r r}
+\toprule
+                         & maximum likelihood & Laplacian smoothing \\
+\midrule
+Spearman $\rho$          & 0.660              & \textbf{0.708} \\
+Kendall $\tau_b$         & 0.467              & \textbf{0.506} \\
+Goodman-Kruskal $\gamma$ & 0.473              & \textbf{0.509} \\
+\bottomrule
+\end{tabular}
+\caption{Laplacian smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
+All correlations are significant at $p = 0.05$.}
+\end{table}
+
+\subsubsection{Neighborhood density}
+
+There are now many methods for computing similarity between nonce words and existing words, long thought to be reflected in wordlikeness judgements. 
+For this study, a number of such methods were evaluated, including the Generalized Neighborhood Model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a number of variations on neighborhood density \citep{Coltheart1977} provided by \citet{Vaden2009}.
+The best performance was obtained with the simplest version of neighborhood density, which is defined as the number of real monomorphemic words which can be changed into the target nonce word by a single insertion, deletion, or substitution of a phone.
+For instance, the neighbors of \emph{blick} include \emph{blink} (insertion), \emph{lick} (deletion) and \emph{black} (substitution).
+While many studies \citep[e.g.,][]{Bailey2001} report robust lexical similarity effects, it may be that the relatively weak performance of neighborhood density is the result of the presence of gross phonotactic violations.
+
+\subsection{Discussion} \label{2discussion}
+
+\subsubsection{The gradience hypothesis}
+
+The primary result is that no gradient model reliably exceeds the accuracy of the binary baseline.
+Despite this, there are relatively strong correlations between the binary baseline and these gradient models (see Table \ref{bcor}).
+However, the fact that the gradient models are generally outperformed by the binary baseline suggest that they do not reliably predict intermediate ratings.
+To quantify this, the following method was used to estimate the residual contribution of the three gradient models once gross phonotactic violations are taken into account.
+Instead of calculating rank correlations directly on the model scores as in Table \ref{cor}, the model scores are mapped to ranks with the additional constraint that all ``valid'' stimlui be ranked above all ``invalid'' stimuli.
+The resulting ranks are used to compute new correlation statistics.
+Finally, the binary baseline correlation is subtracted from this number, so that the resulting value is the amount of improvement derived from augmenting the binary model with gradience.
+These difference numbers are shown in Table \ref{controlled}.
+In most cases, including the gradient models on top of the binary baseline produces a worse correlation than is obtained with the binary baseline alone.
+The intepretation of this is direct, at least in the case of $\tau_b$ and $\gamma$.
+Each gradient model draws contrasts within the sets of phonotactically valid and invalid clusters, respectively.
+For instance, the bigram model favors \emph{troog} [tɹuːg] over \emph{swach} [swætʃ], though neither contains any gross phonotactic violation.
+Among these contrasts, however, the majority are not reflected in relative wordlikeness judgements.
+For instance, \emph{troog} is rated less English-like than \emph{swach}.
+This is stark evidence against the gradience hypothesis.
+
+\begin{table} \label{bcor}
+\centering
+\begin{tabular}{l r r r}
+\toprule
+Kendall $\tau_b$          & maxent         & bigram         & density  \\
+\midrule
+\citealt{Greenberg1964}   & 0.670          & \textbf{0.680} & 0.501 \\
+\citealt{Scholes1966}     & \textbf{0.685} & 0.632          & 0.639 \\
+\citealt{Albright2003b}   & 0.542          & 0.603          & \textbf{0.623} \\
+\bottomrule
+\end{tabular}
+\caption{The binary baseline is strongly correlated with the three gradient model scores; all correlations are significant at $p = 0.05$.}
+\end{table}
+
+\begin{table} \label{controlled}
+\centering
+\begin{tabular}{l r r r}
+\toprule
+$\Delta$ Spearman $\rho$          & maxent            & bigram            & density  \\
+\midrule
+\citealt{Greenberg1964}  & $-0.060$          &  \textbf{0.038} & $-0.017$ \\
+\citealt{Scholes1966}    & $-0.029$          &  \textbf{0.047} & $-0.035$ \\
+\citealt{Albright2003b}  & $-0.008$          & $-0.015$          & \textbf{0.018} \\
+\midrule
+$\Delta$ Kendall $\tau_b$         & maxent            & bigram            & density  \\
+\midrule
+\citealt{Greenberg1964}  & $-0.114$          & \textbf{$-$0.007} & $-0.084$ \\
+\citealt{Scholes1966}    & $-0.067$          & \textbf{0.003}  & $-0.061$ \\
+\citealt{Albright2003b}  & \textbf{$-$0.038} & $-0.092$          & $-0.049$ \\
+\midrule
+$\Delta$ Goodman-Kruskal $\gamma$ & maxent            & bigram            & density  \\
+\midrule
+\citealt{Greenberg1964}  & $-0.268$          & \textbf{$-$0.260} & $-0.337$ \\
+\citealt{Scholes1966}    & $-0.361$          & \textbf{$-$0.313} & $-0.345$ \\
+\citealt{Albright2003b}  & \textbf{$-$0.137} & $-0.443$          & $-0.386$ \\
+\bottomrule
+\end{tabular}
+\caption{The change in rank correlation generated by augmenting the purely binary model with gradient predictions is small and in most cases it is negative.}
+\end{table}
+
+\subsubsection{Possible extensions to the binary baseline}
+
+The strong performance of the binary baseline should not be taken as evidence either that wordlikenesss judgements are binary, or that the binary baseline is a plausible model.
+The most serious limitation of this evaluation is the primitive nature of the binary baseline.
+The inability to generalization within onsets and rimes is a serious flaw, as is the assumption of independence of onset and rime.
+A cognitively plausible version of this model must entertain phonotactic generalizations that are larger than these units.
+
+A possible further extension to the binary baseline would be the introduction of additional levels of wellformedness.
+While the evaluation has shown that current gradient models do not reliably identify intermediate wellformedness, it does seem possible to identify at least three levels of grammaticality: for instance, one might share the intuition that \emph{zhlick} [ʒlɪk] is more similar to English than \emph{bnick}, though both have unattested onsets.
+
+There are precedents for labeling certain attested words as phonotactically ``peripheral'' (see, e.g., the appendices in \citealt{Myers1987} and \citealt{Borowsky1989}); such words are regarded as lexical exceptions to language-general principles of syllabification.
+If this extends to nonce words, then an intermediate level of grammaticality could be assigned to ``possible'' but formally marked words.
+Another likely source of additional levels of grammaticality is the cumulative effect of multiple phonotactic violations.
+While, as \citet{Coleman1997} note, classic model predict taht a nonce word is as ill-formed as its worst deviation from syllable structure,
+it is possible to imagine that multiple phonotactic violations would result in greater degrees of ill-formedness.
+Many competing models make this prediction (e.g., \citealt{Legendre1990}, \citealt{Levelt2000}, \citealt{Goldwater2003}, \citealt{Jager2007}, \citealt{Albright2008}, \citealt{Pater2009b}) but despite this, there is currently no data bearing on whether cumulative effects are found in wordlikeness tasks.
+
+\section{Conclusion}
+
+This chapter has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
+The surprising result is that virtually all of the apparent coverage of state-of-the-art gradient phonotactic models is simply a reflection of their ability to distinguish between the possible and the totally impossible; beyond this, they are unreliable.
+A trivial baseline, endowed with few abilities to project beyond the observed data, generally outperforms the state of the art.
+It follows that the projections made by the state-of-the-art gradient models are not like those made by speakers.
+
+%\subsection{Future directions}
+%\subsubsection{Interspeaker variation}
+%The studies analyzed above all use ratings averaged across subjects.
+While aggregation of this type is quite common in psycholinguistics, it is undesirable as a statistical practice: it drastically increases the chance of a Type II error, the error of failing to reject a null hypothesis when the null hypothesis is false \citep[][8f.]{Baayen2004}.
+Further, little is known about intraspeaker variation in wordlikeness judgements and even less about intraspeaker variation, but \citet{Shademan2007} does find suggestive differences between older and younger adults.
+It would be extremely interesting to correlate differences in speakers with external factors like age, level of education, and exposure to foreign languages.
+%\subsubsection{Mixed modalities}
+%While the use of auditory stimulus presentation seems preferable to rating orthographic words, the possibiltiy of misperception introduces a potential confound.
+\citet{Scholes1966} mentions that a few of his stimuli were frequently misperceived, and a number of studies \citep{Massaro1983,Halle1998b,Pitt1998,Dupoux1999,Berent2007a,Davidson2007} uncover positive correlations between phonotactic acceptability and accurate perception.
+If the phonotactically invalid cluster in *\emph{srest} is perceived as a phonotactically valid cluster with similar acoustic features, such as [ʃr], it will likely receive an inflated wordlikeness rating.
+In many psycholinguistic tasks, the behavioral measure of interest comes with a validation of correct perception: for instance, in shadowing tasks, reaction times are the primary behavioral measure but shadowing errors are also informative behavioral measures \citep[e.g.,][]{Marslen-Wilson1978}.
+Wordlikeness studies could combine transcription or repetition tasks with judgement tasks to confirm correct perception and perhaps to enhance the online nature of the task.
+%Several recent studies have investigated the role of phonotactics using an auditory lexical decision tasks \citep[e.g.,][]{Berent2001a,Berent2004,Shatzman2007a}.
+A recent study by \citet{Coetzee2008b} is particularly relevant to the larger questions of this dissertation, though the results are ultimately inconclusive.
+\citeauthor{Coetzee2008b} claims that words of the form [spVp, skVk] (e.g., *\emph{spep} [spɛp], *\emph{skeek} [skiːk]) are impossible in English, whereas [stVt] words (e.g., *\emph{stoit} [stɔɪt]) are wellformed.
+In addition to a traditional wordlikeness task, \citet{Coetzee2008b} includes stimuli of this type in a lexical decision task, under the hypothesis that the illformed nonce words, those of the shape [spVp, skVk], will be more quickly rejected than valid [stVt] stimuli, a prediction which is borne out.
+However, there is no need to provide a phonotactic interpretation of these effects.
+It is known that auditory tasks reaction times are significantly slower for non-words with dense cohorts, that is, nonce words like *\emph{hup} [hʌp] which share initial segments with many real words (e.g., \emph{hull}, \emph{hut}, etc.).
+These effects are found across modalities \citep[e.g.,][]{Marslen-Wilson1978,Marslen-Wilson1984} and appear to be particularly attentuated for non-word onsets \citep{Cole1980,Vitevitch2002}.
+Since [st] onsets are far more frequent than [sp] or [sk] onsets \citep[][395]{Hayes2008a}, this alone could explain the variation in reaction times.
+Despite this, nonce word lexical decision with controls for lexical cohort effects shows potential as a substitute for wordlikeness judgements in phonotactics research.
+%\subsubsection{Data availability}
+%There are still very few publicly available wordlikeness databases, and several authors who were contacted for this project declined to share data.
+Not only does this make large-scale model comparison difficult, it also means that most wordlikeness studies cannot be replicated except at the conceptual level.
+Lexical decision researchers now have databases of response times for American \citep{ELP} and British English \citep{BLP}, Dutch \citep{DLP}, French \citep{FLP}, and Malay \citep{MLP}.
+Given the simplicity of wordlikeness judgements (a sort of linguistic \emph{Drosophilia}), the time for a publicly available ``English Wordlikeness Project'' has come.
+%\subsubsection{Non-parametric statistics}
+%The surprising results of this evaluation can be traced directly to two methodological improvements which should be more widely adopted.
+First, model comparison studies must replace parametric statistical models, such as Pearson correlation, with non-parametric equivalents.
+Given the tendency of ratings to cluster at endpoints, Pearson correlations are inflated \citep[][23, fn.
+12]{Albright2009a} and the assumptions of the test are flagrantly violated.
+A shift towards non-parametric comparison appears to be well underway: \citet{Hayes2008a} also report model correlations using the Spearman $\rho$ and \citet{Albright2009a} focuses on a variant of the Kendall correlation called $tau_c$.
+%On the other hand, both \citeauthor{Hayes2008a} and \citeauthor{Albright2009a} ultimately use parametric statistics for model comparison.
+%\begin{table}
+%\centering
+%\begin{tabular}
+%\toprule
+%                        & accuracy & precision & recall & F$_1$ & MCC   \\
+%\midrule
+%\citealt{Greenberg1964} & 0.846    & 1.000     & 0.800  & 0.889 & 0.693 \\
+%\citealt{Scholes1966}   & 0.952    & 0.947     & 0.900  & 0.923 & 0.889 \\
+%\citealt{Albright2003b} & 0.920    & 0.938     & 0.953  & 0.946 & 0.797 \\
+%\bottomrule
+%\end{tabular}
+%\end{table}
+% EM \citet{EM} 
+%\begin{table}
+%\centering
+%\begin{tabular}{l | r r}
+%\toprule
+%                        & $\chi^2_{3}$ & $p$-value  \\
+%\midrule
+%\citealt{Greenberg1964} &  5.190       & 0.158      \\
+%\citealt{Scholes1966}   & 48.964       & 1.3\e{-15} \\
+%\citealt{Albright2003b} & 16.186       & 0.001      \\
+%\bottomrule
+%\end{tabular}
+%\caption{This is my caption}
+%\end{table}
+%\footnote{This is conceptually similar to the technique of residualization discussed by \citet{Gorman2010c}.}
+%\begin{unlabeledexample}
+%$\displaystyle P = s^{1/7.4}$
+%\end{unlabeledexample}
+%\noindent Even then, \citet[][23, fn. 12]{Albright2009a} 
+%An interesting feature of this data set is that includes a diverse set of onsets ranging from frequent to impossible, but only a few common rimes; \citet{Albright2009a} speculates this might have made the onsets more salient to the subjects than tehy would be otherwise.
+%The Pearson correlation depends on an interval assumption, which holds that the difference between, for example, ``1.
+completely bi disagree'' and ``2.
+agree'' is in some sense identical to that between ``3.
+neither agree nor disagree'' and ``4.
+agree''.
+\citet{Stevens1946} already observes that there is no a priori basis for this assumption, and similar objections can be leveled for other rating methods.
+Secondly,
+%the Pearson correlation assumes a linear relationship between model scores and ratings; when this assumption is violated, researchers must search for a satisfactory transformation to produce linearity.
+%As \citet{Halle1962} argues, this eliminates the need for additional grammatical machinery.
+%\begin{quote}
+%\ldots{}we are forced to incorporate into every complete generative grammar a characterization of the distinction between admissible and inadmissible segment sequences.
+The fact effectively cuts the ground out from under the recent suggestion that generative grammars be supplemented with special phonological grammars\ldots{} \citep[][61]{Halle1962}
+%\end{quote}

### 2012-06-03 3.tex
Chapter 2 draft done.

@@
-\chapter{Static and derived phonotactic constraints in Turkish} 
+%\chapter{Static and derived phonotactic constraints in Turkish} 
 \label{turkish}
-
-%This chapter attempts to derive which phonotactic judgemetns are internalized by
 
@@
 
-\begin{example}
-Backness harmony wordlikeness forced choices \citep[314]{Zimmer1969}:
-
-\vspace{0.5\baselineskip}
+\begin{example}[Backness harmony wordlikeness forced choices (\citealp{Zimmer1967}:314)]
 \begin{tabular}{l r l r}
@@
 
-\begin{example}
-Roundness harmony forced choice wordlikeness judgements:
-
-\vspace{0.5\baselineskip}
+\begin{example}[Wordlikeness comparisons, roundness harmony]
 \begin{tabular}{l r l r}
@@
 
-\begin{example}
-Labial attraction wordlikeness forced choices \citep[314]{Zimmer1969}:
-
-\vspace{0.5\baselineskip}
+\begin{example}[Wordlikeness comparisons]
 \begin{tabular}{l r l r}
@@
 
-\begin{example}
-Arabic verbs of coming into being: \\
+\begin{example}[Arabic stative verbs]
 \begin{tabular}{l l l l}
@@
 
-\begin{example}
-\textsc{Labial harmony} and etymology in TELL \citep[][187]{Inkelas2001}:
-
-\vspace{0.5\baselineskip}
+\begin{example}[\textsc{Labial harmony} and etymology in TELL \citealp{Inkelas2001}:187]
 \begin{tabular}{l r r r}
@@
 Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
-
-%Classical Arabic adjectives often have stative verbs in which the root is imposed onto the template CaCuCa:

### 2012-06-03 4.tex
Chapter 2 draft done.

--- 2012-05-13 4.tex
@@
-%\chapter{Accidental and structural gaps in English syllable contact\protect\footnotemark{*}\protect\footnotetext{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics.}}
-\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
+%\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
 \label{clusters}
@@
 
-\input{4.1}
-\input{4.2}
-\input{4.3}
+\section{Domain}
+
+% WHY I'm DOING THIS STUDY
+% describes makeup of the corpus and how p
+
+\subsection{Data sources}
+
+The English CELEX database \citep{CELEX},
+
+\citet{Aronoff1976}
+
+\citet{Harley2009}
+
+\citet{Taft1975} ?
+\citet{Taft1981} ?
+\citet{Taft2004a} ?
+
+\citet{Halle1962}
+
+% 2.1.4: k
+
+Under the parses and phonologization described immediately below, the 6,876 simplex CELEX entries contain 21 medial coda types and 40 unique medial onsets.
+Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, a saturation rate of 18.8\%.
+
+\subsection{Syllabification}
+
+For this study, it is necessary to separate medial consonant clusters into codas and onsets.
+While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic.
+For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language.
+Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which explain the \emph{chemistry}/\emph{ministry} syllabification contrast.
+
+An automated syllabification system was constructed and applied to medial clusters in the CELEX data.
+This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.
+There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification.
+The technique used is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948},
+\citealt[75]{Pulgram1970}, \citealt[42f.]{Kahn1976}, \citealt[][358f.]{Selkirk1982b}) which parses word-medial clusters so that the onset is as large as possible.
+Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty.
+In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.
+
+There is one case where unchecked maximization of medial onsets produces incorrect syllabifications.
+When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
+\citet[][55]{Harris1994} notes that onset maximization makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
+The approach adopted here is essentially that of \citet{Pulgram1970}: the first consonant of a medial cluster is assigned to the coda of a preceding coda if is immediately preceded by a stressed lax vowel, which do not occur syllable-finally in English.\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}
+
+If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables.
+For this reason, an additional constraint is assumed which prevents the two components of the affricates from being split by syllabification.
+This is motivated by the tendency of affricates to pattern with single segments in many languages.
+For instance, the only complex onsets in Classical Nahua are the affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.
+
+\subsection{Inventory consdierations and subsyllabic parsing}
+
+To identify the inventory of syllable contact clusters, it is also necessary to parse syllables into onset, nucleus, and coda.
+In most cases this is trivial, but in a few cases, it is necessary to use distributional heuristics to determine the underlying inventory and subsyllabic affiliations.
+
+\subsubsection{The velar nasal} \label{velarnasal}
+
+There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right \citep[e.g.,][]{Sapir1925}, an analysis which is a consequence of principles like the Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ (e.g., \emph{SPE}:85, \citealt{Borowsky1986}:65f.).
+The strongest piece of evidence for the latter position, in which the velar nasal is a pure allophone, is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
+\citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here.
+English nasal place allophony is formalized in Section \ref{cnpasection} below.
+
+\subsubsection{[j] onglides}
+
+The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J.
+\citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}).
+The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).
+
+There is considerable evidence for this assumption.
+When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː].
+In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}.
+This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
+\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear.
+There is also strong external support for this principle.
+The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants.
+The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor} [h]\emph{omponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
+
+\subsubsection{[w] onglides}
+
+The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset.
+Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
+Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}.
+Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.
+And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.
+
+\subsubsection{Post-vocalic \emph{r}}
+
+Both CELEX and the dictionary used by \citet{Pierrehumbert1994} in her study transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost.
+In \emph{r}-full dialects, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters.
+Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}); for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}.
+Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which post-vocalic \emph{r} patterns differently than internal coda consonants.
+\citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.
+
+\begin{example}[/t/-\textsc{Glottalization} in British English (after \citealp{Harris1994}:195, 258)]
+\begin{tabular}{l l l@{} l}
+a. & par[t]   &   & par[ʔ]   \\
+   & car[t]on &   & car[ʔ]on \\
+b. & fis[t]   & * & fis[ʔ]   \\
+   & mis[t]er & * & mis[ʔ]er \\
+\end{tabular}
+\end{example}
+
+\noindent Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.
+
+\begin{example}[/t, d/-\textsc{Deletion} in American English] \label{td}
+\begin{tabular}{l l l@{} l}
+a. & be[lt]  &   & be[l] \\
+   & me[nd]  &   & me[n] \\
+b. & sh[ɚt]  & * & sh[ɚ] \\
+   & c[ɚd]   & * & c[ɚ]  \\
+\end{tabular}
+\end{example}
+
+\section{Evaluation}
+
+\subsection{Static and derived constraints}
+
+With this database of English syllable contact clusters it is now possible to evaluate constraints on sequences of underlying phonemes.
+
+Here, the metric used is simply
+
+The evaluation
+
+The foremost evidence for
+
+The evaluation used here is simply
+
+Having described the database of English syllable contact clusters,
+
+I now consider constraints on sequences of URs as predictors of attested and unattested contact clusters.
+First, I describe the statistical technique used, then apply this technique to both phonologically derived constraints and a set of ``static'' phonotactic constraints proposed by \citet{Pierrehumbert1994}.
+
+Before the dawn of generative phonology, many linguists concerned themselves with documenting co-occurrence in various languages.
+\citet[][28]{Vogt1954} declares that a ``phonemic description of a language \ldots should also comprise a description of the phoneme combinations that occur or may be expected to occur in the language''.
+Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively.
+\citeauthor{Uhlenbeck1950}, for instance, puts forth co-occurrence dispreferences, rather than exceptionless generalizations of the sort given by \citet{Greenberg1950}.
+
+\citet{McCarthy1988}
+
+Years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and proposes a statistical technique, based on the chi-square test, to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.
+
+\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English).
+The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented.
+Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.
+
+\begin{example}[null hypothesis testing]
+$\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
+χ^2$
+\end{example}
+
+%\footnote{This is the one-tailed version of the test.}
+
+In this study, I use the \citet{Fisher1934} exact test, which is isomorphic to this chi-square test.
+Compared to the chi-square test, the Fisher exact test $p$-value is somewhat more difficult to compute, but both are automated by modern statistical packages and can be computed rapidly.
+The major advantage of the Fisher test over the chi-square test is that it is appropriate for small samples or rare events, whereas the chi-square test depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples.
+
+On the other end of the spectrum, the Monte Carlo statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010} does not scale to large samples.
+The Monte Carlo procedure is isomorphic to the one-tailed Fisher Exact test, and involves comparing observed co-occurrence counts to those produced by randomly generated samples.
+This requires the researcher to generate random permutations, something which is particularly difficult for large samples, as I will explain; disinterested readers may wish to skip the following paragraph.
+
+Any permutation of a sequence $L$ can be represented as a sequence $S$ of the same length, where the value of $S_i$ corresponds to the permuted position of $L_i$.
+Generating random permutations requires an unbiased method to generate all $N!$ of these lists with equal probability.
+Unfortunately, this is near impossible for $N$s of moderate size with current computational resources.
+Any pseudo-random number generator can be characterized by its ``period'', the number of pseduo-random numbers it can emit before repeating itself; this is also the upper limit for the number of random permutations that can be generated.
+The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a period of approximately $2^{48}$, a number which turns out to be smaller than $17!$.
+For the large samples, sometimes in the thousands, used by these researchers, virtually all the possible permutations can never be generated by this method.
+As a consequence, the shuffling procedure is far from random and introducing the possibility that the statistical test is biased in unpredictable ways.
+
+As an example of the Fisher exact test as it is applied here, consider some of the Javanese co-occurrence facts considered by \citet{Mester1988}.
+\citeauthor{Mester1988} is at pains to show that there is a gradient dispreference for first and second root consonants which share the same major place.
+However, perfectly identical first and second consonants are exempt.
+Since
+
+% APPLY THIS TO JAVANESE
+
+\begin{example}[Javanese root co-occurrence] %(\citealp[][264]{Uhlenbeck1950}, \citealp[][139]{Mester1988})]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 134      & 12         & 0.918      & \multirow{2}{*}{6.095\e{-06}} \\
+violating  & 21       & 15         & 0.583 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent
+It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.
+
+This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance
+
+That this is highly unlikely to be due to chance is apparent both from the chi-square ($\chi^2 = 25.572$, two-tailed $p = 4.3$\e{-07}) and Fisher exaact tests ($p =  6.095$\e{-06}).
+
+I use two-tailed statistical tests throughout. 
+Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis.
+As
+
+\citet{Pierrehumbert1994}
+
+It must be stressed that under current assumptions, this is independent of whether predictable prosodic representations (particularly coda and onset identity) are part of the memorized underlying representation \citep[e.g.,][]{Vaux2003}.
+It is an undeniable fact that URs need to be 
+The derivative nature of 
+\citet{Ito1989a,Noske1992}
+
+least some researchers (i.e., \citealt{Mester1988})
+
+ \citep[e.g.,]{Brown2010} have argued for phonotactic constraints giving rise to overrepresentation, so I admit this possibility here and use two-tailed tests throughout.
+
+%\footnote{For example, \citet{Martin2007} applies this technique to a list of 4,758 noun-noun compounds.
+The number of permutations of a list of this length can be estimated using Stirling's approximation.
+
+%\ex $ \lim_{n \rightarrow \infty} n! = e ^ {n \ln{n} - n} = 2 ^ {\log_2{e} (n \ln{n} - n)} $ \xe
+
+%\noindent
+%$4,758!$ is approximately $2 ^ {51260}$. The author is unaware of any system that provides 51,260 bits of entropy; and some popular programming languages, such as ANSI C, provide as few as 15 bits.}
+
+%%% START OF SERIOUSNESS
+
+\subsection{Derived constraints}
+
+\emph{SPE} describes three phonological alternations which target syllable contact clusters. The corresponding constraints on simplex syllable contact clusters are considered below.
+
+\subsubsection{Obstruent voice assimilation}
+
+Voice assimilation in English is exemplified by the allomorphs of /-d/ suffix, forming regular pasts and denominal adjectives, and the /-z/ suffix, forming regular noun plurals, genitives, and the 3sg.
+present active indicative.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S.
+\citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}; alternative URs are argued for by \citet[][210f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
+
+\begin{example}[Inflectional affix voice assimilation]
+\begin{tabular}{l l l l l l}
+a. & nap & nap[t] \\
+   & nab & nab[d] \\
+b. & lap & lap[s] \\
+   & lab & lab[z] \\
+\end{tabular}
+\end{example}
+
+\noindent The devoicing of the two voiced obstruent suffixes can be formalized as a rule spreading the \textsc{Voice} specification of an obstruent rightward.
+
+\begin{example}[\textsc{Obstruent Voice Assimilation} (after \emph{SPE}:178)]
+\xymatrix@R=24pt@C=24pt{
+\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{--}[dr] &                               \\
+\txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
+\txt{[$+$\textsc{Obstruent}]}                  & \txt{[$+$\textsc{Obstruent}]} \\
+}
+\end{example}
+
+\citet{Pierrehumbert1994} does not mention voice assimilation in her study of English syllable contact clusters.
+\citet[][74f.]{Hammond1999a} notes the existence of words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain obstruent voice clusters disagreeing in [\textsc{Voice}].
+This might be taken as evidence that \textsc{Obstruent Voice Assimilation} does not apply to simplex words, and indeed, the \textsc{Revised Alternation Condition} (RAC) proposed by \citeauthor{Kiparsky1973a} (\citeyear{Kiparsky1973a}:163, \citeyear{Kiparsky1982a}:152) blocks the application of obligatory neutralization processes like \textsc{Obstruent Voice Assimilation} in root-internal (i.e., non-derived) environments.\footnote{The ``obligatory'' condition on the RAC has the result that \textsc{Obstruent Voice Assimilation} is always consistent with the RAC whether or not it applies in non-derived environments.
+If the rule applies in non-derived environments, then it has lexical exceptions (e.g., \emph{a}[s.b]\emph{estos}) and is not obligatory and thus free from the RAC.
+On the other hand, if the rule is subject to the RAC, it only applies in derived environments, in which it is obligatory.} Despite this, root-internal clusters of obstruents which disagree in voicing (e.g., [s.b, b.s]) are much rarer than those which contain adjacent voiced or voiceless obstruents (e.g., [s.p, b.z]), suggesting that \textsc{Obstruent Voice Assimilation} applies in non-derived environments, preventing learners from positing URs in which adjacent obstruents do not agree in voice.
+
+\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value                   \\
+\midrule
+conforming & 80       & 370        & 0.178      & \multirow{2}{*}{1.1\e{-11}} \\
+violating  &  6       & 264        & 0.022                                    \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\noindent 18\% of the possible clusters containing adjacent obstruents with the same [\textsc{Voice}] specification occur in the corpus, whereas only 2\%  of those which disagree in [\textsc{Voice}].
+The small $p$-value indicates that the rarity of the disagreeing clusters is unlikely to be due to chance.
+
+The failure of a small number of roots to conform to this generalization can be derived by marking such roots [$-$\textsc{Obstruent Voice Assimilation}].
+Another possibility is that disagreeing obstruent clusters might be analyzed as semantically opaque compounds by native speakers (e.g., \emph{ja}[k\#d]\emph{aw}).
+There are precedents for using morphological boundaries solely to preserve phonological generalizations.
+In \emph{SPE} (p.~94), for instance, morphological boundaries are used to simplify stress assignment, and \citet[][546]{Rice2009d} analyses many words in the Athapaskan language Slave as compounds simply because they contain consonant clusters that rarely occur otherwise.
+Recent experiments find that 9-month-old infants are sensitive to the contrast between [vt], a disagreeing cluster which does not occur word-medially, and [ft], found in medial position \citep{Mattys2001b}, and this ability seems to persist into adulthood \citep{McQueen1998}.
+
+\subsubsection{Nasal place assimilation} \label{npasection}
+
+\citet[][175]{Pierrehumbert1994} observes that ``nasal-stop sequences agree in labiality'' in simplex English words; \citeauthor{Pierrehumbert1994} presumably makes no statement about other place features as she also assumes the velar nasal is derived from underlying /n/.
+While \citeauthor{Pierrehumbert1994} makes no connection between this sequence structure constraint and the any phonological alternation, the nasal place assimilation that derives [ŋ] is also responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix.
+In (\ref{npa}ab), the prefix-final consonant takes on the same major place of articulation as the following obstruent.
+The vowel-initial bases in (\ref{npa}c) demonstrate that the prefix is underlyingly /ɪn-/.
+
+\begin{example}[\emph{im-}/\emph{in-} allomorphy] \label{npa}
+\begin{tabular}{l l l l}
+a. & possible  & i[m.p]ossible \\
+   & balance   & i[m.b]alance  \\
+b. & tangible  & i[n.t]angible \\
+   & decent    & i[n.d]ecent   \\
+c. & elegant   & i[n]elegant   \\
+   & ability   & i[n]ability   \\
+d. & mature    & i[m]ature     \\
+   & numerable & i[n]umerable  \\
+\end{tabular} 
+\end{example}
+
+The shape of the prefix before the velar stops /k, g/ is somewhat more complex.
+\citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but that assimilation of \emph{im-}/\emph{in-} is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
+However, the stress condition does not hold in simplex words (e.g., \emph{a}[ŋ.g]\emph{óra}), and thus can be safely ignored here.
+
+\begin{example}[\textsc{Nasal Place Assimilation} (after \emph{SPE}:85)]
+\xymatrix@R=24pt@C=24pt{
+                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
+\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
+\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]}              \\
+}
+\end{example}
+
+Exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} do occur, they are much rarer than homorganic nasal-obstruent clusters such as \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey}, a preference which is unlikely to be due to chance.\footnote{There are also a small class of simplex words which contain clusters like \emph{oi}[nt.m]\emph{ent}, in which a nasal-obstruent cluster is found within a complex coda, but place assimilation is without exception in this context, and thus these clusters are ignored here.}
+
+\begin{example}[Lexical effects of \textsc{Nasal Place Assimilation}]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value                   \\
+\midrule
+conforming & 41       & 11         & 0.788      & \multirow{2}{*}{2.6\e{-05}} \\
+violating  & 8        & 20         & 0.286                                    \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+There is some evidence that even young infants acquiring English are aware of this generalization.
+\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} find that 4.5-, 9-, and 10-month old infants (respectively) prefer to listen to nonce words which contain the homorganic nasal-obstruent clusters (e.g., \emph{u}[m.b]\emph{o}) over those which contain heterorganic clusters (e.g., \emph{u}[n.b]\emph{o}).
+Further, adult speech errors may derive obstruent clusters which then undergo \textsc{Nasal Place Assimilation}.
+
+\begin{example}[Speech errors feeding \textsc{Nasal Place Assimilation} (\citealp{Myers1993}:228)]
+\begin{tabular}{l l l}
+a. & primps        & (prints)      \\
+b. & rand orker    & (rank order)  \\
+c. & po[ŋ]cutation & (computation) \\
+\end{tabular} 
+\end{example}
+
+\subsubsection{Degemination}
+
+A final alternation targeting English syllable contact clusters consists of the simplification of geminates derived by certain affixes.
+For instance, the deadjectival adverbial suffix \emph{-ly} rarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{norma}[l]\emph{y}, cf.
+\emph{calm}[l]\emph{y}).
+Degemination is also found in /d/-final roots selecting the irregular /-t/ past tense suffix.
+
+\begin{example}[Degemination in /-t/ pasts (\citealp{Myers1987}:492)]
+\begin{tabular}{l l l l l}
+a. & burn    & burn[t] \\
+   & spill   & spil[t] \\
+b. & ben[d]  & ben[t]  \\
+   & buil[d] & buil[t] \\
+\end{tabular}
+\end{example}
+
+More complicated cases of degemination are also found in English prefix morphology (\emph{SPE}:148, \citealt{Borowsky1986}:102f.); one can be found in (\ref{npa}d) above, in which \textsc{Nasal Place Assimilation}.
+To formalize this process of degemination, it is assumed that geminates are adjacent similar segments.
+The alternative, that geminates are single feature matrices linked to multiple timing slots, places this constraint in the inventory and makes it difficult to link degemination and the absence of geminates in underlying representations.
+Geminates are identified using quantification over features \citep{Reiss2003b}.
+
+\begin{example}[\textsc{Degemination} (\emph{SPE}:253)]
+\xymatrix@R=24pt@C=24pt{
+\txt{C}_1       & \rightarrow & \emptyset & / & \gap\gap & \txt{C}_2 \\
+}
+
+\noindent Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
+\end{example}
+
+Neither identical consonants, nor those which differ only in voice (e.g., *[p.b]) occurs in simplex words in CELEX, and this absence is highly unlikely to be due to chance.\footnote{It is not the case that the English lexicon envisioned in \emph{SPE} is free of underlying geminates; \citeauthor{SPE} use underlying (but never surfacing) geminates to produce exceptions to the stress rules (see also \citealt{Burzio1994}:52f., \citealt{Halle1998c}:551, fn.
+9).
+In addition to the usual objections to absolute neutralization (i.e., that it is a learning conundrum mascarading as a phonological solution), the phonological particulars of this account are critiqued by \citet{Ross1972a}.}
+
+\begin{example}[Lexical effects of \textsc{Degemination}] 
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value                   \\
+\midrule
+conforming & 157      & 596        & 0.208      & \multirow{2}{*}{7.2\e{-09}} \\
+violating  & 0        &  87        & 0.000                                    \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\subsection{Static constraints}
+
+\citet{Pierrehumbert1994} surveys the English syllable contact cluster inventory, and discerns the effect of several static constraints.
+These constraints are termed ``static'' because there is no evidence for any phonological process, either allophonic or neutralizing, which targets the clusters the constraints disfavor.
+
+\subsubsection{Dorsal/non-coronal clusters}
+
+\citet[][173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Since \citeauthor{Pierrehumbert1994}'s study is restricted to triconsonantal clusters, so words like \emph{a}[k.m]\emph{e}, \emph{ru}[g.b]\emph{y}, or \emph{pi}[g.m]\emph{ent} do not violate this narrower generalization.
+Further, \citeauthor{Pierrehumbert1994} adopts the traditional assumption \citep[e.g.,][66f.]{Borowsky1986} that coda [ŋ] is an allophone of /n/ before dorsal consonants, so words such as \emph{mo}[ŋ.gr]\emph{el} are not exceptions.
+While there are no exceptions to the generalization about triconsonantal clusters, the statistical evidence does not rule out the possibility the absence of such clusters is due to chance.
+Clusters of any length also fail to support a broader version of this generalization.
+As dorsal/non-coronal clusters are not significantly less likely to occur than similar types of ``conforming'' clusters such as those with a non-dorsal coda and a non-coronal onset (e.g., \emph{a}[t.m]\emph{osphere}) or a dorsal coda and coronal onset (e.g., \emph{ve}[k.t]\emph{or}), there is no evidence for any dispreference for dorsal/non-coronal clusters.
+
+\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}, triconsonantal clusters only]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value              \\
+\midrule
+conforming & 20       & 148        & 0.119      & \multirow{2}{*}{0.136} \\
+violating  &  0       &  22        & 0.000                               \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}, all size clusters]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 76 & 359 & 0.175 & \multirow{2}{*}{0.145} \\
+violating  &  6 &  57 & 0.095 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\subsubsection{Coda coronal obstruents}
+
+\citet[][175]{Pierrehumbert1994} proposes that ``clusters with a coronal obstruent in the coda do not occur'', though \citet{Pierrehumbert1994} notes an exception in \emph{a}[nt.l]\emph{er}, and one can add the triconsonantal clusters in \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many biconsonantal clusters like \emph{a}[t.l\emph{as} or \emph{e}[s.t]\emph{er}.
+Triconsonantal coda coronal obstruent clusters are not significantly less likely to occur than ``conforming'' clusters with non-coronal or non-obstruents.
+The same is true if the arbitrary restriction to triconsonantal clusters is lifted.
+
+\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}, triconsonantal clusters only]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value              \\
+\midrule
+conforming & 20       & 148        & 0.119      & \multirow{2}{*}{0.136} \\
+violating  & 0        & 22         & 0.000                               \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}, all size clusters]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value              \\
+\midrule
+conforming & 48       & 312        & 0.133      & \multirow{2}{*}{0.301} \\
+violating  & 38       & 322        & 0.106                               \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\subsubsection{ABA clusters}
+
+Finally, \citet[][176]{Pierrehumbert1994} observes ``a lack of clusters with identical first and third elements''.
+This is formalized by treating as ``identical'' any consonants which share all the same place and manner features and ignoring voicing, as was done for the definition of \textsc{Degemination} above.
+\citeauthor{Pierrehumbert1994} notes one potential exception to this generalization, the cluster in words such as \emph{e}[ks.kl]\emph{ude}, which contains two /k/, does not arise here since such words are marked as morphologically complex in CELEX.
+Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonaontal and quadraconsonantal clusters.
+
+\begin{example}[Lexical effects of \textsc{*aba}]
+\begin{tabular}{l r r r r}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 41       & 478        & 0.079      & \multirow{2}{*}{0.818} \\
+violating  &  0       &  11        & 0.000                               \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\subsection{Computational models}
+
+The three constraints derived from phonological alternations have been shown to have robust effects on the English lexicon.
+In constrast, the static constraints proposed by \citet{Pierrehumbert1994} do not have a statistically reliable effect on the cluster inventory.
+Obviously this does not indicate, however, that there are no static constraints on the cluster inventory.
+By applying computational models of phonotactics to the syllable contact data, however, it is possible to exhaustively search for any such static constraints.
+
+\subsubsection{Evaluation of computational models}
+
+The task of the computational models is the same as above: they are used to predict which of the possible clusters are attested and which are not.
+Models are scored using four metrics.
+Accuracy is the probability of that a possible cluster is either a true positive ($tp$), correctly predicted to occur, or a true negative ($tn$), correctly predicted to be unattested.
+Incorrect outcomes are either false positives ($fp$), incorrectly predicting a cluster to occur, and false negatives ($fn$), incorrectly predicting a cluster to be unattested.
+
+\begin{unlabeledexample}
+$\displaystyle \textrm{Accuracy} = \frac{tp + tn}{tp + tn + fp + fn}$
+\end{unlabeledexample}
+
+Precision is the probability that a cluster predicted to occur is attested.
+
+\begin{unlabeledexample}
+$\displaystyle \textrm{Precision} = \frac{tp}{tp + fp}$ 
+\end{unlabeledexample}
+
+Recall (or sensitivity) is the probability a observed cluster is predicted to occur.
+
+\begin{unlabeledexample}
+$\displaystyle \textrm{Recall} = \frac{tp}{tp + fn}$
+\end{unlabeledexample}
+
+It is possible to maximize recall at the expense of precision, by predicting more clusters to be attested, or to increase precision at the expense of recall by predicting fewer clusters to be attested.
+$F_1$, the harmonic mean of these two measures, quantifies this tradeoff.
+
+\begin{unlabeledexample}
+$\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$
+\end{unlabeledexample}
+
+The results for the four models are summarized in Table \ref{cmresults}.
+
+\begin{table} \label{cmresults}
+\centering
+\begin{tabular}{l | r r r r}
+\toprule
+                          & accuracy & precision & recall & $F_1$ \\ 
+\midrule
+Null baseline             & 0.813    & n.a.      & n.a.   & n.a.  \\
+Derived constraints       & 0.857    & 0.401     & 0.708  & 0.512 \\
+\citet{Pierrehumbert1994} & 0.839    & 0.210     & 0.750  & 0.328 \\
+\citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
+\bottomrule
+\end{tabular}
+\caption{The \citeauthor{Hayes2008a} model makes the most accurate predictions about possible and impossible syllable contact clusters, though it is only slightly better than the derived constraint baseline.}
+\end{table}
+
+\subsubsection{Null baseline}
+
+18.8\% of the 840 possible clusters are attested.
+As a consequence of this sparsity, a null baseline which classifies all clusters as unattested will achieve 81.2\% accuracy.
+Precision, recall, and $F_1$ are undefined for this baseline.
+
+One model was evaluated for this study, but performed no better than this baseline.
+\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
+However, \citeauthor{McGowan2011} finds that sonority of clusters accounts for only a tiny amount of the variance in cluster frequency ($R^2 = 0.077$).
+As sonority distance provided no improvement over the null baseline, it is not reported above.
+
+\subsubsection{Derived constraints}
+
+\textsc{Obstruent Voice Assimiation}, \textsc{Coda Nasal Place Assimilation}, and \textsc{Degemination} are reliable predictors of what clusters are and are not attested and admit few exceptions.
+A slightly more complex baseline can be produced by predicting any cluster which does not violate these generalizations to be attested.
+This results in a significant improvement to the null baseline (accuracy = 0.857).
+As might be expected given the sparse nature of the syllable contact inventory, there are more unattested clusters which do not violate any of these alternation-derived constraints than exceptions to these generalizations, reflected in the fact that recall (0.708) is somewhat higher than precision (0.401).
+
+\subsubsection{\citet{Pierrehumbert1994}}
+
+\citet{Pierrehumbert1994} proposes a baseline in which the well-formedness of syllable contact clusters is proportional to the product of the probabilities of their codas and onsets.
+\citeauthor{Coleman1997} summarize the results as follows.
+
+\begin{quote}
+\citet{Pierrehumbert1994} showed that the independent probabilities of the coda and the following onset was the single biggest factor in predicting which complex clusters exist.
+\citep[50]{Coleman1997}
+\end{quote}
+
+\noindent \citet{Pierrehumbert1994} uses both medial and initial/final codas and onsets to compute these probabilities, but a better fit is obtained by using only word-medial coda and onset frequencies.
+\footnote{The use of medial frequencies is likely to cross-linguistically superior, as there are many languages, like Finnish, in which medial codas are not possible final codas \citep[][306]{Fischer-Jorgensen1952}.} The resulting probabilities are transformed into a classification of ``attested'' or ``unattested'' using a soft-margin support vector machine \citep{Cortes1995} with a linear kernel.
+This technique finds the optimal point for separating attested and unattested clusters.
+This is not a cognitively plausible learning model: it simply provides an upper bound on the performance of the \citeauthor{Pierrehumbert1994} model.
+By definition, this model imposes no constraints on sequences crossing the syllable boundary, and unfortunately, this only provides a small improvement to the null baseline (accuracy = 0.839), and has a lower accuracy and $F_1$ (0.328) than the derived constraint baseline.
+
+%\citep{Haugen1956}. Thanks to Stephen Anderson for this observation
+
+\subsubsection{\citet{Hayes2008a}}
+
+\citet{Hayes2008a} (henceforth, H\&W) develop a sophisticated method of learning phonotactic distributions from positive data.
+They estimating probability distributions over sequences of sets of phonological features using the principle of maximum entropy.
+As above, a support vector machine with a linear kernel is used to turn these probability values into binary predictions.
+A direct replication of the experiments by H\&W are attempted, by using their software, feature specfication, and model settings.
+Since the training of this model is inherently stochastic, producing slightly different outcomes on each run, the results reported here are thoes of the lowest-accuracy run out of ten.
+This practice is used by H\&W (p.~396), though is not a great deal of variation between individual runs.
+
+This model provides a small improvement in accuracy (0.877) and $F_1$ (0.703) compared to the alternation baseline, but it has lower recall (0.642) than this baseline, indicating the presence of false negatives.
+More specifically, the H\&W model learns narrower phonotactic generalizations than those derived from alternations.
+For instance, unattested *[m.kl] is assigned the highest probability, whereas it is ruled out by \textsc{Coda Nasal Place Assimilation}.
+On the other hand, the \citeauthor{Hayes2008a} model assigns a low score to clusters like those found in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.\footnote{\citet{Berent2012} discuss a potential defect of the H\&W model: it does not use of variables over features and thus fails to unify constraints against geminate clusters.
+This does not appear to be relevant to this evaluation, as the H\&W model correctly predicts the total absence of geminate clusters.}
+
+\subsection{Discussion}
+
+Though the H\&W model has the overall best performance, it is not greatly superior to the derived constraint baseline.
+This suggests that most, if not all, of the 71.2\% of possible clusters which are unattested are not ill-formed, but are simply missing due to the sparsity of the data: that is, they are accidental gaps.
+Below, the role of sparsity in the syllable contact inventory is quantified in a number of ways.
+
+\subsubsection{The Zipfian distribution}
+
+A number of previous studies \citep[e.g.,][]{Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phonemes and graphemes exhibit sparse type (i.e., lexical) and token frequency distributions.
+The lexical frequencies of the medial codas ond onsets which make up syllable contact clusters show a similar distribution.
+
+In a sparse distribution, many types occur at the same low frequencies.
+\citet{Good1953} notes that this produces a type of quantization, an artificially long and flat right tail in the frequency spectrum.
+\citet[][29]{Church1991} propose a method to eliminate this quantization, the $Z_r$ transform, which is defined as follows.
+The vectors $r, n$ are such that $n_1$ is the number of types which occur at frequency $r_i$ (that is, $n$ is a vector of frequencies of type frequencies), and $r$ is sorted in increasing order.
+$Z$ is simply each element of $n$ scaled by the nearest points to the left and right.
+
+\begin{unlabeledexample}
+$\displaystyle Z_i = \frac{2 f(rn_i}{r_{i + 1} - r_{i - 1}}$
+\end{unlabeledexample}
+
+\noindent \citeauthor{Church1991} do not define this transform for the lowest and highest points (i.e., when $i = 1$ or $i = N$), but the definition can be extended by scaling the edge points according to next innermost point.
+
+\begin{unlabeledexample}
+$\displaystyle Z_1 = \frac{n_1}{r_2 - r_1}$
+\end{unlabeledexample}
+
+\begin{unlabeledexample}
+$\displaystyle Z_N = \frac{n_N}{r_N - r_{N - 1}}$
+\end{unlabeledexample}
+
+In Figure \ref{clus}, $Z_r$-transformed lexical frequencies of codas and onsets are plotted against rank, in log-log spaces.
+The near-linear relationship that obtains ($R^2 = 0.924$) indicates that the data conform to a generalization of Zipf's (\citeyear{Zipf1949}) law.
+
+\begin{unlabeledexample}
+$\displaystyle Z_r(C, \alpha) = \frac{C}{r^\alpha}$
+\end{unlabeledexample}
+
+\noindent The parameters $C$, a constant sensitive to sample size, and $\alpha$, the slope of the log-rank/log-frequency line, can be computed using the method of least squares, with $\epsilon$ representing the error terms.
+
+\begin{unlabeledexample}
+$\displaystyle \textrm{log}~Z_r \sim C + \alpha~\textrm{log}~r + \epsilon$ 
+\end{unlabeledexample}
+
+\begin{figure}
+\centering
+\includegraphics{cluster.pdf}
+\caption{The log-log linear relationship between frequency and rank demonstrates that syllable contact clusters are Zipf distributed.}
+\label{clus}
+\end{figure}
+
+It is important to note that there is nothing inherently meaningful about the fact that clusters conform to a Zipfian distribution.
+These distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2009} to word \citep{Teahan1998,Ha2002,Baroni2009} and phoneme \citep{Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
+The importance of this statistical property is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps due only to the sparsity of the data.
+
+\subsubsection{The Good-Turing estimate}
+
+\citet{Good1953} proposes an estimate of $\hat{p}_0$, the probability of all unseen events, originally developed in collaboration with Alan Turing.
+This probability is simply $n_1$, the number of events which occur just once, divided by the number of observations.
+
+\begin{unlabeledexample}
+$\displaystyle \hat{p}_0 = \frac{n_1}{\displaystyle\sum N}$ 
+\end{unlabeledexample}
+
+\noindent In CELEX, $65$ clusters occur only once, and there are 873 cluster tokens, so $\hat{p}_0 = 0.074$.
+The interpretation of this quantity is as follows: were possible to generate a new corpus of syllable contact clusters, approximately 7\% of the tokens would be clusters not found in the previous sample.
+Unfortunately, the Good-Turing estimate does not provide a way to determine which unattested clusters might be found in a replication.
+
+\subsubsection{Sampling simulation}
+
+As a final consideration of this data, consider the statistical properties of a randomly generated ``lexicon'' as they compare to the observed distribution of clusters.
+A simulated lexicon can be generated by repeatedly applying the following procedure, which combines the derived constraint baseline with the model proposed by \citet{Pierrehumbert1994}.
+
+\begin{example}[Simulation procedure]
+\begin{tabular}{l l}
+a. & Sample a medial coda according to the observed probabilities  \\
+b. & Sample a medial onset according to the observed probabilities \\
+c. & Apply all matching \emph{SPE} rules to the resulting clusters \\
+\end{tabular}
+\end{example}
+
+While this procedure can be repeated indefinitely, it is worthwhile to consider the results of a single characteristic ``lexicon'' produced in this fashion.
+The slope of the log-rank/log-frequency relationship for the simulated data ($\alpha = -0.524$) is also close to the observed slope for the CELEX data ($\alpha = -0.588$).
+And, as shown in Figure \ref{sim}, the two distributions are nearly indistinguishable.
+
+\begin{figure}
+\centering
+\includegraphics{sim.pdf}
+\caption{The observed cluster frequencies are closely matched by a simulated cluster inventory generated by random sampling.}
+\label{sim}
+\end{figure}
+
+This result should not be construed as an endorsement of the \citet{Pierrehumbert1994} model.
+Rather, it shows that even if the only constraints on syllable contact clusters are that they must consist of valid medial codas and onsets which further conform to the three derived constraints, there still will be many gaps in the cluster inventory, gaps which are quite arbitrary from the perspective of the phonology.
+
+\section{Conclusion}
+
+\citet{Borowsky1989} on peripherality.
+productive \citet{Duanmu2008}
+
+The next two chapters return to the question of synchrony, addressing the relationship between statistical patterns in the lexicon and speakers' behaviors when presented with underrepresented sequences.
+
+accidental gaps
+\citet[][419f.]{Hayes2008a}
+
+%as in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} CELEX also contains \emph{jo}[d.p]\emph{urs}, \emph{pi}[ntʃ.b]\emph{eck}, \emph{sa}[k.b]\emph{ut}, and \emph{ja}[k.d]\emph{aw}. 
+%There is no duplication between constraints on sequence structure and surface forms if both are derived from the application of the phonological rule (\citealt[][401f.]{Stanley1967}, \citealt[][382]{SPE}).
+%\citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''.
+In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.
+%\begin{example}[Artificial language adaptations (\citealt{Wright1975}:394, his transcriptions)]
+%\begin{tabular}{l l l l}
+%a. & [gownp] & \textgreater{} & [gump]  \\
+%b. & [ǰumg]  & \textgreater{} & [ǰúŋgə] \\
+%c. & [ðʌŋd]  & \textgreater{} & [tɔŋg]  \\
+%\end{tabular}
+%\end{example}
+%\noindent If this game is akin to the process of loanword adaptation, it is possible that various extragrammatical pressures are at play \citep[e.g.,][]{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008}.
+Yet, the independent evidence for a phonological process of nasal place assimilation provides the most parsiminious account of the adaptations seen in \citeauthor{Wright1975}'s study.
+%\citet{Hay2004a} demonstrates that [n.p, mθ] clusters, which are not found in English, are in fact rated better than many attested clusters.
+%The same is true of the prefix \emph{com-}/\emph{con-}, as shown by the presence of assimilation in \emph{có}[ŋ.g]\emph{ress} but not in \emph{co}[n.g]\emph{réssional}. 
+%(e.g., \citealt{Bakovic2005b}, \citealt{Fruehwald2011}),
+
+%\ex \textsc{Coda Nasal Place Assimilation} feeds \textsc{Degemination} \citep[][116]{Borowsky1986}: \\
+%\begin{tabular}{l l l l}
+%a. & mature    & i[m]ature    & (cf. \emph{i}[m.b]\emph{alance})   \\
+%b. & numerable & i[n]umerable & (cf. \emph{i}[n.d]\emph{ependent}) \\
+%\end{tabular}
+%\xe
+%(\citealp{Halle1985a}:105, 
+%; presumably, this generalization is limited to this one place feature because  \citeauthor{Pierrehumbert1994} also adopts the assumption that the velar nasal is derived from /n/. 
+%Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets. 
+%
+%\begin{example}[VBR's complex onsets (\citealp{Buchwald2005}:79--80, his transcriptions)] \label{VBR}
+%\begin{tabular}{l l l}
+%a. & kəræb  & `crab'  \\
+%   & bəlid  & `bleed' \\
+%b. & kəwin  & `queen' \\
+%   & kəwoʊt & `quote' \\
+%c. & kut    & `cute'  \\
+%   & musɪk  & `music  \\ 
+%\end{tabular}
+%\end{example}
+%
+%\noindent VBR breaks up complex onsets with epenthesis, including those that consist of a consonant and a back onglide cluster (\ref{VBR}b).
+However, no epenthesis occurs between a consonant and a front onglide; rather, the glide is absent (\ref{VBR}c).
+The failure of the front onglide to pattern with other consonant clusters suggests once again that the glide is part of the nucleus.
+%j
+%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [juː] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant.This is the only sign that [juː] shows an affiliation for the onset.
+%In fact, [t.ʃ, d.ʒ] clusters are not found in simplex English words, despite the fact that that affricates occur as medial onsets in clusters (e.g., \emph{tru}[n.tʃ]\emph{eon}, \emph{so}[l.dʒ]\emph{er}).
+%Finally, \citet[][251]{Fromkin1973} also presents evidence that post-vocalic \emph{r} may behave as if nuclear in speech errors. 
+%\citep[e.g.,][142]{Mester1988}.
+%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or a part of a contrast in any known language \citep{Rubach2002}.}
+% Note that [tʃ] has not been included since it so rarely occurs in clusters.
+%\subsubsection{Summary}
+%
+%The above results are summarized in Table \ref{constraints}.
+%
+%\begin{table}
+%\centering
+%\begin{tabular}{l r r r r r}
+%\toprule
+%                                       & \multicolumn{2}{c}{attested} & \multicolumn{2}{c}{unattested} & \multirow{2}{*}{$p$-value} \\
+%                                       & conforming & violating & conforming & violating \\
+%\midrule
+%\textsc{O.V.A.} (\ref{ovar})        &  80 &   6 & 370 & 264 & 1.1\e{-11} \\
+%\textsc{C.N.P.A} (\ref{CNPA})       &  41 &  25 &  12 &  42 & 1.7\e{-05} \\ 
+%\textsc{Degemination} (\ref{degem}) & 161 & 632 &   0 &  87 & 1.4\e{-08} \\
+%\midrule
+%\textsc{*Dorsal-Labial} ()          &  68 &   4 & 246 &  40 & 0.069 \\
+%\textsc{*C.C.O.} ()                 &  48 &  38 & 312 & 322 & 0.301 \\
+%\textsc{*A,BA} ()                   & 161 &   0 & 708 &  11 & 0.231 \\
+%\bottomrule
+%\end{tabular}
+%\caption{Counts and Fisher exact test $p$-values for the constraints discussed above.}
+%\label{constraints}
+%\end{table}
+%Whereas 18.8.\% of the ``possible'' clusters are attested, 28.8\% of those which conform to these three constraints are found. 
+%Binary classification contingency table: \\
+%\begin{tabular}{c | l l} \toprule
+%                             & \multicolumn{2}{c}{prediction} \\
+%\midrule
+%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
+%                             & false negative ($fn$) & true negative ($tn$) \\
+%\end{tabular}
+%\begin{tabular}{|l l|}
+%\toprule
+%true positive ($tp$)  & false positive ($fp$) \\
+%false negative ($fn$) & true negative ($tn$) \\
+%\bottomrule
+%\end{tabular}
+%\begin{figure}
+%\centering
+%\includegraphics{coda.pdf} 
+%\includegraphics{onset.pdf}
+%\caption{Medial coda and medial onset lexical frequency exhibit the Zipfian log-log-linear relationship between rank and frequency.}
+%\label{codaonset}
+%\end{figure}
+%Sparsity alone is not the only reason for cluster underrepresentation which is external to the phonology.
+For instance, a cluster could be underrepresented due to language change.
+\citet{Iverson2005} discuss a case of this type: In English, long vowels before word-final [ʃ] in English are generally rare (though note ``affective'' \emph{swoosh}, \emph{sheesh}, etc.).
+This is ultimately due to the fact that native [ʃ] is descended from Old English [sk], and long vowels are not fonud not occur before complex codas.
+\citeauthor{Iverson2005} conclude that this gap is accidental because it does not inhibit borrowing and coinage.
+%On the other hand, patterns created by sound change are not guaranteed to persist over time.
+One example of non-persistence is discussed by \citet{Iverson2005}.
+Around 1100 CE, Old English \emph{sk} became [ʃ].
+This sound change introduced no alternations.
+Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{V\lm sk\#} words when the change was actuated, and \emph{V\lm sh\#} continues to be rare in Modern English.
+What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.
+%\citet[][140]{Frisch2004} suggest that the strong tendency for the first and second consonants of the Arabic root to be non-identical is the ``a diachronic result of a processing constraint that disfavors repetition.'' Unfortunately, there is no evidence that this pattern is diachronic other than in the sense that it appears to be inherited from the proto-language: there is simply no Proto-Semitic verb roots with identical first and second consonants \citep[][178]{Greenberg1950}.
+In other Semitic languages, the inherited patern has experienced considerable erosion.
+%
+%\begin{example}[Tigrinya roots with identical first and second consonants \citep{Buckley1990a}]
+%\begin{tabular}{l l l l}
+%a. & lʌlʌw     & `scorch'                   & (\textless{} Ge'ez \emph{lʌwlʌw} `inflame')     \\
+%   & mʌmʌy     & `winnow'                   & (\textless{} Ge'ez \emph{mʌymʌy} `distinguish') \\
+%   & mʌmʌt     & `pick out loot'            \\
+%b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
+%   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
+%c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
+%   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
+%\end{tabular}
+%\end{example}
+
+%Similar exceptions are found in 
+%Amharic (\citealp[][?]{Broselow1984}, \citealp[][?]{McCarthy1985}) and 
+%Hebrew \citep[][29]{Bat-El2005}.

### 2012-06-03 pre.tex
back to chapter files, lots of writing
Chapter 2 draft done.

@@
 
-\ex Russian 1sg. imperfect gaps \citep[after][]{Pesetsky1977}: \\
+\begin{example}[Russian 1sg. imperfect gaps (after \citealp{Pesetsky1977})]
 \begin{tabular}{l l l@{}l@{}l@{}l@{}l@{} l}
@@
 b. & tʃudit{j}  & * & tʃuʒu   & / & * & tʃuʒdu & `act oddly'   \\
-   & otʃudit{j} & * & natʃuʒu & / & * & natʃuʒdu & `alienate'  \\
+   & naʃudit{j} & * & natʃuʒu & / & * & natʃuʒdu & `alienate'  \\
 \end{tabular}
+\end{example}
 
-\noindent
-As \citealt{Pesetsky1977} notes, the ill-formedness of the 1sg. non-past forms is shared by verbs related by prefixation. The opposite pattern can be seen in English.
+\noindent As \citeauthor{Pesetsky1977} notes, the ill-formedness of the 1sg. non-past forms is shared by verbs related by prefixation. The opposite pattern can be seen in English.
 
-\ex English preterite gaps: \\
+\begin{example}[English preterite gaps]
 \begin{tabular}{l l l@{}l@{}l@{}l@{}l@{}}
-   & present  &   & preterite &   &   &           \\
+   & present  &   & preterite                     \\
 a. & foresee  &   & foresaw   &   &   &           \\

### 2012-06-30 2.tex
Feverish writing on chapters 2-4

@@
-\chapter{Gradient and categorical aspects of wordlikeness judgements}
-\label{gradience}
+\chapter{Gradient and categorical aspects of wordlikeness judgements} \label{gradience}
 
 \citet{Halle1962} and \citet{Chomsky1965} observe that speakers internalize generalizations about possible and impossible words in their language.
-\citeauthor{Chomsky1965} illustrate this knowledge with two nonce words \emph{blick} [blɪk] and \emph{bnick} [bnɪk]; whereas \emph{blick} is unattested, but possible word of English,
-speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word of the language.
+\citeauthor{Chomsky1965} illustrate this knowledge with two nonce words \emph{blick} [blɪk] and \emph{bnick} [bnɪk].
+Whereas \emph{blick} is unattested, but in some sense a ``possible'' word of English, speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word of the language.
 
-The classic account of this contrast is as follows.
-Segments must be assigned to prosodic structures like the syllable, or undergo phonological repair (e.g., \citealt{Hooper1973}:10f., \citealt{Ito1989a}, \citealt{Noske1992}).
+A naïve account of this contrast is based on the assumption that segments or timing units must either be parsed into prosodic structures like the syllable, or be subject to phonological repair.
+Early precedents for this can be found in dissertations by \citet[10f.]{Hooper1973} and \citet[57f.]{Kahn1976}, and it is a fundamental component of many later frameworks (e.g., \citealt{Ito1989a}, \citealt{Noske1992}, \citealt{OT}).
 In English, [bl] is a permissible onset, but unlike some languages (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet'), [bn] is not.
-\citet[][19f.]{Wolf2009} suggest, for instance, that an underlying /bnɪk/ would be realized as [nɪk]; it is also possible to imagine resolution by prothesis---[əbnɪk]---or anapytxis---[bənɪk], and indeed \citet{Davidson2006b} finds that English speakers produce these repairs when attempting to mimic foreign pronunciations of obstruent-nasal onsets.
+\citet[][19f.]{Wolf2009} suggest, for instance, that an underlying /bnɪk/ would be realized as [nɪk]; it is also possible to imagine resolution by prothesis---[əbnɪk]---or anapytxis---[bənɪk], and indeed \citet{Davidson2006b} finds that English speakers use all three of these repairs when asked to mimic foreign pronunciations of obstruent-nasal onsets like /bn/.
 
-Many linguists view this account as overly simpliest, for, as \citeauthor{Shademan2006} argues, it draws a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report wordlikeness intuitions.
+Many linguists who have concerned themselves with ``wordlikeness'' now regard the above as overly simplistic, for, as \citeauthor{Shademan2006} argues, such accounts draw a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report ``wordlikeness'' intuitions.
 
@@
 
-This chapter focuses on the hypothesis that speakers' knowledge of wordlikeness is gradient.
+This chapter focuses on the hypothesis, implicit in \citeauthor{Shademan2006}'s statement, that there is an inherent gradient or continuum between possible and impossible words, against an null hypothesis that the observed patterns of judgements are little more than an artifact of the wordlikeness task itself.
 \S\ref{gradhyp} places this debate in a historical perspective, and argues that the naïve but widely accepted form of this hypothesis fails the test of falsifiability.
-In \S\ref{2evaluation}, a revision of this hypothesis is subjected to an quantitative evaluation and is found wanting.
+In \S\ref{2evaluation}, a quantitative evaluation of a revised version of this hypothesis reveals a poor match between the gradience in speakers' judgements and contemporary computational models of wordlikeness.
+Taken together, the findings argue against the gradience hypothesis.
 
-[Explain that there's not going to be any issue of, e.g., heterovoiced obstruents, that stuff is all in chap. 4]
+% What \citet[32]{Cohn2006} calls ``gradient well-formedness''
 
@@
 
-%In this section, the gradience hypothesis is reviewed from historical and scientific perspectives.
-
 \subsection{A brief history of the gradience hypothesis} \label{history}
 
-\citet{Chomsky1965} were not the first to consider contrasts between possible and impossible words.
-Their primary contribution to this question, is, essentially, mentalism: they recognize that naïve speakers effortlessly acquire language-specific wordlikeness generalizations and can report them with ease.
-But the nature of these contrasts were also considered by structuralists.
+\citeauthor{Chomsky1965} were not the first to consider the notion of possible and impossible words.
+Their primary contribution is that their mentalist perspective: they recognize that naïve speakers effortlessly acquire language-specific generalizations about possible and impossible words and can report them without any explicit training.
+
+\subsubsection{The Copenhagen school}
+
+Wordlikeness received some attention by structuralists of the Copenhagen school.
 In an early study, \citet{Fischer-Jorgensen1952} proposes that possible and impossible words represent endpoints on a continuum, and that no non-arbitrary line can be drawn between them.
@@
 
+However, not all early literature is concerned with gradience.
+\citet[][31]{Vogt1954}, for instance, recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
+\citeauthor{Vogt1954} observes that allophony may account for the absence of certain phone sequences, but it does not provide a suitable explanation for the absence of initial [bn] in English, nor does it make correct predictions about the surface realization of an underlying initial /bn/.
+\citeauthor{Vogt1954} concludes that additional grammatical machinery will be needed to account for possible and impossible words.
+
 \subsubsection{Early generative grammar}
 
-Gradience is also discussed in the earliest work in generative syntax.
-\citet[][132]{LSLT} writes that ``there is little doubt that speakers can fairly consistently order new utterances, never previously heard, with respect to their degree of `belongingness' to the language''.
-\citet{ASPECTS} proposes to link different degrees of ungrammaticality to different kinds of syntactic violations.
-Unlike \citeauthor{Fischer-Jorgensen1952}, there is for \citeauthor{LSLT} a hard-and-fast boundary between the grammatical and ungrammatical, only the latter showing any gradience.
-For \citeauthor{LSLT}, grammatical utterances are all equally grammatical, but ungrammatical utterances may be deviant in their their own ways \citep[][61]{Schutze1996}.
+Gradient grammaticality is assumed in the earliest work on generative syntax, \emph{The logical structure of linguistic theory}, in which \citet[][132]{LSLT} writes that ``there is little doubt that speakers can fairly consistently order new utterances, never previously heard, with respect to their degree of `belongingness' to the language''.
+Later, \citet{ASPECTS} proposes that different syntactic violations give rise to different degrees of ungrammaticality.
+Unlike \citeauthor{Fischer-Jorgensen1952}, however, for \citeauthor{LSLT} this continuum does incorporate a hard-and-fast boundary between the grammatical and ungrammatical: there are no constrasts between grammatical utterances, merely between different degrees of ungrammaticality.
+For \citeauthor{LSLT}, all grammatical utterances are alike with respect to their grammaticality, but there are many degrees of ungrammaticality \citep[][61]{Schutze1996}.
 
-However, most of the early literature on wordlikeness concerns apparently categorical contrasts the possible and impossible.
-\citet[][31]{Vogt1954} recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
-Generalizing somewhat from \citeauthor{Vogt1954}'s discussion, allophony may account for the absence of certain phone sequences, but it does not provide a suitable explanation for the absence of initial [bn] in English, nor does it make correct predictions about the surface realization of an underlying initial /bn/.
-\citeauthor{Vogt1954} concludes that additional grammatical machinery will be needed to account for possible and impossible words.
-\citet{Halle1962} and \citet{Stanley1967} propose to derive wordlikeness effects from morphophonemic redundancy rules acting on underlying representations.
+\citet{Halle1962} and \citet{Stanley1967} concern themselves only with categorical wordlikeness contrasts, however.
+Both authors derive wordlikeness effects from morphophonemic redundancy rules acting on underlying representations.
 The discussion of \emph{bnick} by \citet[][101]{Chomsky1965} is a good example of this type of analysis.
-They correctly observe that before a word-initial stop, a English consonant must be liquid.
-A side effect of a redundancy rule which specifies a consonant in this position as [$+$\textsc{Liquid}] precludes the derivation of \emph{bnick}.\footnote{This is not the only redundancy that might be used to rule out \emph{bnick}; for instance, the only word-initial obstruent that may follow a nasal in English is /s/.}
+\citeauthor{Chomsky1965} observe that a consonant before a word-initial stop can only be a liquid in English.
+A corresponding redundancy rule which specifies a consonant in this position as [$+$\textsc{Liquid}] precludes the derivation of \emph{bnick}, since there is no way to realize the second consonant as a nasal.\footnote{This is not the only redundancy that could rule out \emph{bnick}.
+For instance, /s/ is the only word-initial obstruent that may precede a nasal consonant in English.}
 
@@
 
-In \emph{The sound pattern of English} (SPE), \citet{SPE} extend this model of wordlikeness to account for
+In \emph{The sound pattern of English} (SPE), \citet{SPE} argue that a complete solution to wordlikeness must countenance gradience.
+In addition to the contrast of \emph{blick} and \emph{bnick}, they introduce a third nonce word \emph{bnzk}, which they take to be even less well-formed than ``impossible'' \emph{bnick}.
 
-[Explain the SPE model]
+\begin{quote}
+Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
+\end{quote}
+
+In some languages (e.g., Imdlawn Tashlhiyt Berber [tm.zħ] `she jested', [tzmt] `it (f.) is stifling'; \citealt{Dell1985}), \emph{bnzk} is syllabifiable, so its ill-formedness is once again a property of English, not a universal.
+
+\citeauthor{SPE} propose to derive gradience from the complexity (using a simple feature-counting metric) of the redundancy rules needed to map a nonce word to an attested word, and they show that this derives the increasing cline of wordlikeness running from ``consonant soup'' \emph{bnzk} to ill-formed \emph{bnick}, possible \emph{blick}, and finally lexical \emph{brick}.
+
+Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no attention in the literature.
+At the risk of explaining what might be no more than an an accidental gap in the historical record, the novel aspects of the \emph{SPE} model---gradience derived from similarity to existing lexical entries---seem to have been overshadowed by compelling arguments against \citeauthor{SPE}'s assumption (see also \citealt{Halle1962}, \citealt{Stanley1967}) that wordlikeness contrasts derive from properties of underlying forms.
+\citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
+An example from German is shown in (\ref{fd}) below.
+
+\begin{example}[German final devoicing] \label{fd}
+\begin{tabular}{l l l l}
+   & nom.sg. & nom.pl.    \\
+a. & [piːp]    & [piːpə]  & `cheep(s)'      \\
+   & [diːp]    & [diːbə]  & `thief/thieves' \\
+b. & [ɡʀaːt]   & [ɡʀaːtə] & `ridge(s)'      \\
+   & [ɡʀaːt]   & [ɡʀaːdə] & `degree(s)'     \\
+\end{tabular}
+\end{example}
+
+\noindent The plural appears to preserve a contrast in final obstruent voicing which is absent in the singular:\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing (e.g., \citealt{Fourakis1984}) or not \citep[e.g.,][]{Port1985}, it is irrelevant to the discussion at hand.} word-final voiced stops are never found in German, and \citet[95]{Shibatani1973} reports that ``it is easy to show that a native speaker of German rejects those forms \emph{on the ground that they end in voiced obstruents}'' (emphasis in original).
+However, given that root-final voicing is not predictable (e.g., /ɡʀaːt \alt{} ɡʀaːd/), the process of \textsc{Final Devoicing} cannot be any kind of lexical redundancy and is mysterious under the \emph{SPE} account.
+The alternative proposed by \citeauthor{Shibatani1973} and by \citet{Clayton1976}, is that to say that nonce words like [ɡʀaːd] are ill-formed not beacuse of any underlying property, but because they fail to undergo an otherwise-exceptionless phonological process, \textsc{Final Devoicing}, or equivalently, the surface-true generalization it derives.
+
+\citet{Sommerstein1974} further notes that the computation of hypothetical underlying forms from nonce surface forms which is implied by the \emph{SPE} theory is non-trivial in the presence of phonological opacity \citep[see][528f.]{Anderson1988a}, and the rejection of the biuniqueness principle means that there is not always a unique solution, as is illustrated by the two underlying forms corresponding to [ɡʀaːt] in (\ref{fd}) above.
 
@@
 
-The \emph{SPE} model was not widely adopted, and research on wordlikeness turned to other principles.
-Syllabification plays no role in \emph{SPE} theory, though the syllable is found in many earlier studies (see \citealp{Goldsmith2011b}).
-\citet{Hooper1973} and \citet{Kahn1976} argue that wordlikeness generalizations argue for the need for syllabification as part of the phonological computation.
-This classic syllabification-based model is further enriched by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as envisioned by][]{Pike1947a}, and the theory of prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
+The arguments of \citeauthor{Shibatani1973} and others led theorists to focus their attention on properties of surface representations as determinants of wordlikeness.
+Though syllabification plays no role in \emph{SPE}, the syllable is discussed in many earlier studies, and it received particular attention in the 1970s.
+\citet{Hooper1973} and \citet{Kahn1976} argue that the syllable is useful for defining wordlikeness generalizations.\footnote{\citet{Steriade1999} and \citet{Blevins2003}, however, argue that a number of phonotactic generalizations previously stated in syllabic terms can be reanalyzed without making reference to syllables.} \citeauthor{Hooper1973} argues, for instance, that [bn], impossible as an English onset, is unobjectionable as a syllable contact cluster in nonce words like \emph{stabnik} (or in names like \emph{Abner}), and that this demonstrates the superiority of syllable-based wordlikeness generalizations.
+This already signals further trouble for alternative accounts which focus on underlying forms.
+Syllabification may span morphs, is generally predictable, and is universally non-contrastive, and as a consequence, few posit in to be present in underlying representations \citep[though see, e.g.,][]{Vaux2003}.
+\citeauthor{Hooper1973} also points to loanword adaptations which produce native syllable structure \citep[e.g.,][]{Carlisle1991} as evidence that syllabification is part of the phonological computation.
+
+Further enrichments to the theory are provided by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as envisioned by][]{Pike1947a}, and theories like prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
+
+While none of these authors discuss gradience, the syllable plays a role in the definition of a gradient measure, ``positional probability'', though to correlate closely with human judgements of wordlikeness.
+The positional probability of a nonce monosyllabic word is derived from the combined probabilities at which segments occur in onset, nucleus, and coda positions in the lexicon of a given language.
+
+Using the head-term preference paradigm, \citet{Jusczyk1993b} and \citet{Friederici1993} find that typically-developing children as young as 9 months of age distinguish between nonce words which are and are not phonotactically valid in their target language.
+\citet{Jusczyk1994} report that 9-month-old children acquiring English also show preferences for nonce words with high positional probability over those with low positional probability.
+Faciliatory effects of positional probability (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements (\citealt{Vitevitch1999a}, \citealt{Luce2001}, \citealt{Lipinski2005}, \citealt{Vitevitch2005}), and lexical decision \citep{Pylkkanen2002a}.
+
+%\citet{Mattys1999,Mattys2001}
+%\citet{McQueen1998}
+
+The most interesting result of these studies comes from the first experiment of \citealt{Vitevitch1997}, which finds that speakers' wordlikeness ratings of bisyllabic words are correlated with the positional probabilities of the constituent syllables.
+Unfortunately, neither \citeauthor{Vitevitch1997} nor any of the other aforementioned psycholinguists make any effort to eliminate the possibility that the low positional probability stimuli are ``impossible'' words of English.
+In fact, in Chapter \ref{clusters}, it is argued that many of the stimuli used by \citet{Vitevitch1997}, including even the high probability stimuli, contain illicit word-medial consonant clusters.
+While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness to a post-hoc test, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and the wordlikeness judgements.
+\citeauthor{Vitevitch1997} ultimately conclude that it cannot explain all the variation in wordlikeness.
+
+The aforementioned studies all conclude that the gradient measure of positional probability correlates with behavioral results.
+As the flaws of the \citet{Vitevitch1997} study demonstrate, the aforementioned studies do little to tease apart the gradient and categorical aspects of phonotactics.
+More generally, they do little to distinguish between positional probability and closely correlated measures like bigram probability (see \S\ref{bigram} below) or neighborhood density, since these studies carefully select stimuli which either have high or low values for all of positional probability, bigram probablility, and neighborhood density.
+This is particularly troublesome given that positional probability was created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production (e.g., \citealt{Luce1998}, \citealt{Luce2000}) and perception (e.g., \citealt{Marslen-Wilson1984,Marslen-Wilson1987}, \citealt{McClelland1986}, \citealt{Norris1994}, \citealt{Norris2000}).
 
@@
 The observation itself cannot be questioned, but \citeauthor{Hayes2008a} and \citeauthor{Albright2009a} do not explicitly state why this data is relevant to the construction of models of wordlikeness.
-The following seems the most likely: \citeauthor{Hayes2008a} and \citeauthor{Albright2009a} believe this proves that wordlikeness, as an internal state, is gradient simply because subjects make use of intermediate degrees of wordlikeness in judgement tasks.
-This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, and an influential view in the cognitive sciences (see \citealt{Fodor1981a} for a critique).
+Presumably, these authors believe that these patterns of judgements demonstrate that wordlikeness, as an internal state, is gradient simply because subjects make use of intermediate degrees of wordlikeness in judgement tasks.
+This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, and in general, an influential view in the cognitive sciences (see \citealt{Fodor1981a} for a critique).
 
@@
 Of course.
-For example.
-consider the concept \emph{odd number}.
+For example, consider the concept \emph{odd number}.
 This seems to have a clear definition, a precise description\ldots{}No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd.
@@
 
-\citeauthor{Armstrong1983} note that if subjects produce intermediate ratings for these categories, then the naïve proposition is false,
-%(i.e., $p \wedge \neg q \vdash \neg (p rightarrow q)$
-and this is what \citeauthor{Armstrong1983} find in their experiments.
-They ask subjects to rate, on a seven-point scale, the extent to which, e.g., certain odd counting numbers represent the concept ``odd number'', and find that speakers do make use of intermediate values.
-For instance, \emph{7} is rated 1.4 out of 7 on average, whereas the average ratings of \emph{447} and \emph{501} are 3.5 and 3.7, respectively.
+\citeauthor{Armstrong1983} note that if subjects produce intermediate ratings for these categories, then the naïve proposition is false, and this is what \citeauthor{Armstrong1983} find in their experiments.
+They ask subjects to rate, on a seven-point scale, the extent to which, e.g., certain odd counting numbers represent the concept ``odd number'', and find that speakers do make use of intermediate values, as can be seen in Figure \ref{agg}.
+In summary, the proposition that intermediate ratings indicate mental gradience is either untestable, or it is false, and is thus rejected.
 
-In summary, the proposition that intermediate ratings indicate mental gradience is either untestable, or it is false, and is thus rejected.
+\begin{figure} \centering
+\includegraphics{agg.pdf}
+\caption{Subjects asked to rate the degree to which certain even and odd numbers represent the categories ``even'' and ``odd'', respectively, freely use intermediate ratings.}
+\label{agg} 
+\end{figure}
 
@@
 In the case of wordlikeness, a case can be made that one such factor is similarity to existing words (e.g., \citealt[][151, fn.
-27]{LSLT}, \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Bailey2001}).
+27]{LSLT}, \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Bailey2001}, \citealt{Schutze2005}).
 However, the mere fact that subjects use intermediate ratings does not show that they do so with any systematicity, or that, e.g., the contrast between \emph{447} and \emph{7} with respect could be given any satisfying explanation.
@@
 \begin{quote}
-\ldots{}when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different.
+Putting it another way, when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different.
 \citep[][215]{Schutze2011}
@@
 
-The remainder of this chapter is devoted to evaluting the gradience hypothesis, using state-of-the-art computational wordlikeness models and data from English.
+The remainder of this chapter is devoted to evaluating the gradience hypothesis, using state-of-the-art computational wordlikeness models and data from English.
 
@@
 The large-scale studies by \citet{Bailey2001} and \citet{Shademan2006,Shademan2007} are ineligible because the former lack stimuli with gross phonotactic violations and the latter data are not available to the public in any form.
+All data are reproduced in Appendix \ref{appendixA}
 
@@
 
-\citet{Greenberg1964} investigated wordlikeness using the technique of free magnitude estimation, a mechanism which has become increasingly popular among syntactians \citep[e.g.,][]{Bard1996}.
+\citet{Greenberg1964} investigated wordlikeness using the technique of free magnitude estimation, a mechanism which has become increasingly popular among syntacticians \citep[e.g.,][]{Bard1996}.
 At the beginning of the experiment, the subject heard a recording of the word \emph{stick}.
@@
 For each stimulus, the subjects produced forced choice ``yes''/``no'' answers to the question of whether the item ``is likely to be usable as a word of English''.
-\citet{Hayes2008a} and \citet{Albright2009a} analyse this data as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus (see also \citealt{Pierrehumbert1994}, \citealt{Coleman1997}, \citealt{Frisch2000}).
-For instance, 22 of the 33 students answered ``yes'' for \emph{shlerk} [ʃlɚk], so it is assigned a score of $0.666$.\footnote{This procedure conflates intraspeaker variation (which may be considerable; see \citealt{Shademan2007}) and speaker-internal gradience, and its adoption here should not be construed as an endorsement.
-It is provided solely for comparison with prior studies using the \citeauthor{Scholes1966} data.} These stimuli are all ostensibly non-words, but include \emph{clung} [klʌŋ], the preterite and past participle of the verb \emph{cling}, and \emph{brung} [brʌŋ], a dialectical past participle for \emph{bring}.
+\citet{Hayes2008a} and \citet{Albright2009a} analyze this data as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus (see also \citealt{Pierrehumbert1994}, \citealt{Coleman1997}, \citealt{Frisch2000}).
+For instance, 22 of the 33 students answered ``yes'' for \emph{shlerk} [ʃlɚk], so it is assigned a score of $0.666$.\footnote{This procedure conflates intraspeaker variation, which in some cases is considerable \citep{Shademan2007} with speaker-internal gradience, and its adoption here should not be construed as an endorsement.
+Rather, it is provided solely for comparison with prior studies using the \citeauthor{Scholes1966} data.} These stimuli are all ostensibly non-words, but include \emph{clung} [klʌŋ], the preterite and past participle of the verb \emph{cling}, and \emph{brung} [brʌŋ], a dialectical past participle for \emph{bring}.
 These two words were excluded and the remaining 61 stimuli were submitted to analysis.
@@
 Consider the case here, in which model scores are compared with wordlikeness ratings.
-If a model rates \emph{dresp} [dɹɛsp] more wordlike than *\emph{srest} [sɹɛst], this pair is concordant if speakers also rate \emph{dresp} more English-like than \emph{srest}, and it is discordant if *\emph{srest} is rated more English-like.
+If a model rates \emph{dresp} [dɹɛsp] more wordlike than *\emph{srest} [sɹɛst] and if speakers further rate \emph{dresp} more English-like than \emph{srest}, than \emph{dresp}/\emph{srest} are a concordant pair.
+If, however, a model and speakers disagree on the relative ranking of \emph{dresp} and \emph{srest}, the pair is discordant.
 The $\tau_b$ and $\gamma$ differ only in their treatment of ``ties'' (i.e., if e.g., *\emph{dresp} and *\emph{srest} are scored the same, or rated the same); $\tau_b$ includes a correction for ties, whereas $\gamma$ ignores tied pairs.
-Much like the familiar Pearson correlation, $\rho$, $\tau_b$ and $\gamma$ are all in the range [-1, 1].
+Much like the familiar Pearson correlation, $\rho$, $\tau_b$ and $\gamma$ are all in the range [$-1$, $1$].
 Correlations of the four models are given in Table \ref{cor}.
 
-\begin{table} \label{cor}
-\centering
+\begin{table} \centering
 \begin{tabular}{l l l l l}
@@
 All correlations are significant at $p = 0.05$.}
+\label{cor}
 \end{table}
@@
 This problem is more acute in languages with more permissive syllable structures than those of English, and thus a sparser lexicon with regard to phonotactic constraints.
-For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.
-A similar result can be found in Chapter 4, which considers the inventory of syllable contact clusters in English.
-% FIXME
-Despite these defects, the baseline outperforms the gradient models in most contexts.
+For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.\footnote{Chapter \ref{clusters} reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
 
-In Figure \ref{density}, the densities of ratings from the three studies, linearly transformed to the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{density}.
+The rating densities from the three studies, linearly transformed to fit within the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{dsn}.
 In all three studies, it is possible to discern a relatively sharp separation between valid and invalid clusters, but also the presence of intermediate values.
 
-\begin{figure} \centering \label{density}
+\begin{figure} \centering
 \includegraphics{density.pdf}
 \caption{Average ratings of individual nonce words, linearly transformed to the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
+\label{dsn}
 \end{figure}
@@
 
-\citeauthor{Hayes2008a} (\citeyear{Hayes2008a}; henceforth H\&W) develop a sophisticated model of phonotactic grammaticality which estimates a probability distribution over phoneme sequences by weighing constraints according to the principle of maximum entropy.
+\citeauthor{Hayes2008a} (\citeyear{Hayes2008a}; henceforth H\&W) develop a sophisticated model of phonotactic grammaticality which estimates a probability distribution over phoneme sequences by weighing constraints according to the principle of maximum entropy, following \citet{Goldwater2003} and \citet{Jager2007}.
 H\&W find that the predictions of their model are closely correlated with the \citet{Scholes1966} wordlikeness ratings.
@@
 Since the training of the maximum entropy model is inherently stochastic, producing slightly different outcomes on each run, the lowest scoring of ten runs is reported (H\&W:396), though in general there is not a great deal of variation between individual runs.
-One limitation of this model is that it is not feasible to score whole words, as the number of constraints which must be inspected grows expontentially as their scope increases.
+One limitation of this model is that it is not feasible to score whole words, as the number of constraints which must be inspected grows exponentially as the size of possible constraints increases.
 Following H\&W and of \citet{Albright2009a}, who also applies the maximum entropy model to the \citet{Albright2003b} norms, the model is trained and scored only on stimulus onsets.
@@
 
-\subsubsection{Segment bigram probability}
+\subsubsection{Segment bigram probability} \label{bigram}
 
@@
 The variant of segmental bigrams used here computes probabilities with a simple type of smoothing in which the count of all possible bigrams (including those never observed) are incremented by one.
-This technique is known as Laplacian, or ``add one'' smoothing.
+This technique is known as Laplace, or ``add one'' smoothing.
 This has the desirable effect that no nonce word is ever assigned a zero probability, and produces in a small increase in the correlation between the \citeauthor{Albright2003b} wordlikeness norms compared with the maximum likelihood estimate (Table \ref{albrightimproved}).
-For all three data sets, this model also consistenly outperforms ``positional'' probability models implemented by \citet{Vitevitch2004} and \citet{Vaden2009}, and thus these are not considered further.
+For all three data sets, this model also consistenly outperforms positional probability models as defined by \citet{Vitevitch2004} and \citet{Vaden2009}.
+Given that these probabilities are highly correlatd with bigram probabilities \citep[][54]{Vitevitch1997}, they are not considered further.
 The bigram model consistently performs well in all the evaluations, and has the highest Spearman correlation with the \citeauthor{Greenberg1964} and \citeauthor{Scholes1966}, and is frequently second place model to the binary baseline elsewhere.
 
-\begin{table} \label{albrightimproved}
-\centering
+\begin{table} \centering
 \begin{tabular}{l r r r}
 \toprule
-                         & maximum likelihood & Laplacian smoothing \\
+                         & maximum likelihood & Laplace smoothing \\
 \midrule
@@
 \end{tabular}
-\caption{Laplacian smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
+\caption{Laplace smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
 All correlations are significant at $p = 0.05$.}
+\label{albrightimproved}
 \end{table}
 
-\subsubsection{Neighborhood density}
+\subsubsection{Neighborhood density} \label{density}
 
@@
 For this study, a number of such methods were evaluated, including the Generalized Neighborhood Model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a number of variations on neighborhood density \citep{Coltheart1977} provided by \citet{Vaden2009}.
-The best performance was obtained with the simplest version of neighborhood density, which is defined as the number of real monomorphemic words which can be changed into the target nonce word by a single insertion, deletion, or substitution of a phone.
-For instance, the neighbors of \emph{blick} include \emph{blink} (insertion), \emph{lick} (deletion) and \emph{black} (substitution).
+The best performance was obtained with the simplest version of neighborhood density, which is defined as the number of real monomorphemic words which can be changed into the target nonce word by a single insertion, deletion, or substitution of a phone.\footnote{\citet{Greenberg1964} use a variant in which only substitutions are counted.} For instance, the neighbors of \emph{blick} include \emph{blink} (insertion), \emph{lick} (deletion) and \emph{black} (substitution).
 While many studies \citep[e.g.,][]{Bailey2001} report robust lexical similarity effects, it may be that the relatively weak performance of neighborhood density is the result of the presence of gross phonotactic violations.
 
-\subsection{Discussion} \label{2discussion}
-
-\subsubsection{The gradience hypothesis}
+\subsection{Modeling residual gradience}
 
@@
 To quantify this, the following method was used to estimate the residual contribution of the three gradient models once gross phonotactic violations are taken into account.
-Instead of calculating rank correlations directly on the model scores as in Table \ref{cor}, the model scores are mapped to ranks with the additional constraint that all ``valid'' stimlui be ranked above all ``invalid'' stimuli.
+Instead of calculating rank correlations directly on the model scores as in Table \ref{cor}, the model scores are mapped to ranks with the additional constraint that all ``valid'' stimuli be ranked above all ``invalid'' stimuli.
 The resulting ranks are used to compute new correlation statistics.
@@
 In most cases, including the gradient models on top of the binary baseline produces a worse correlation than is obtained with the binary baseline alone.
-The intepretation of this is direct, at least in the case of $\tau_b$ and $\gamma$.
-Each gradient model draws contrasts within the sets of phonotactically valid and invalid clusters, respectively.
+
+For $\tau_b$ and $\gamma$, the interpretation of this result is clear.
+The gradient models assign rankings to the sets of phonotactically valid and invalid clusters, respectively.
 For instance, the bigram model favors \emph{troog} [tɹuːg] over \emph{swach} [swætʃ], though neither contains any gross phonotactic violation.
-Among these contrasts, however, the majority are not reflected in relative wordlikeness judgements.
-For instance, \emph{troog} is rated less English-like than \emph{swach}.
+Similarly, the bigram model favors \emph{choop} [tʃuːp] over \emph{zhrick} [ʒɹɪk], even though neither word is a possible word of English.
+However, the majority of such predicted contrasts are not reflected in speakers' judgements; for instance, \emph{troog} is rated less English-like than \emph{swach} \citep{Greenberg1964}, contrary to the model predictions.
 This is stark evidence against the gradience hypothesis.
 
-\begin{table} \label{bcor}
-\centering
+\begin{table} \centering
 \begin{tabular}{l r r r}
@@
 \caption{The binary baseline is strongly correlated with the three gradient model scores; all correlations are significant at $p = 0.05$.}
+\label{bcor}
 \end{table}
 
-\begin{table} \label{controlled}
-\centering
+\begin{table} \centering
 \begin{tabular}{l r r r}
@@
 \caption{The change in rank correlation generated by augmenting the purely binary model with gradient predictions is small and in most cases it is negative.}
+\label{controlled}
 \end{table}
 
-\subsubsection{Possible extensions to the binary baseline}
+\section{Conclusions}
+
+This chapter
+
+has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
+
+The surprising result is that virtually all of the apparent coverage of state-of-the-art gradient phonotactic models is simply a reflection of their ability to distinguish between the possible and the totally impossible; beyond this, they are unreliable.
+A trivial baseline, endowed with few abilities to project beyond the observed data, generally outperforms the state of the art.
+It follows that the projections made by the state-of-the-art gradient models are not like those made by speakers.
+
+%MORE SUMMARY
+
+\subsection{Possible extensions to the binary baseline}
 
@@
 Another likely source of additional levels of grammaticality is the cumulative effect of multiple phonotactic violations.
-While, as \citet{Coleman1997} note, classic model predict taht a nonce word is as ill-formed as its worst deviation from syllable structure,
-it is possible to imagine that multiple phonotactic violations would result in greater degrees of ill-formedness.
-Many competing models make this prediction (e.g., \citealt{Legendre1990}, \citealt{Levelt2000}, \citealt{Goldwater2003}, \citealt{Jager2007}, \citealt{Albright2008}, \citealt{Pater2009b}) but despite this, there is currently no data bearing on whether cumulative effects are found in wordlikeness tasks.
-
-\section{Conclusion}
-
-This chapter has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
-The surprising result is that virtually all of the apparent coverage of state-of-the-art gradient phonotactic models is simply a reflection of their ability to distinguish between the possible and the totally impossible; beyond this, they are unreliable.
-A trivial baseline, endowed with few abilities to project beyond the observed data, generally outperforms the state of the art.
-It follows that the projections made by the state-of-the-art gradient models are not like those made by speakers.
+While, as \citet{Coleman1997} note, classic model predict that a nonce word is as ill-formed as its worst deviation from syllable structure, it is possible to imagine that multiple phonotactic violations would result in greater degrees of ill-formedness.
+The bigram and maxent models make this prediction, as do many others (e.g., \citealt{Legendre1990}, \citealt{Levelt2000}, \citealt{Albright2008}, \citealt{Anttila2008}, \citealt{Pater2009b}) but despite this, there is still no data demonstrating cumulative effects in wordlikeness tasks.
 
@@
 %the Pearson correlation assumes a linear relationship between model scores and ratings; when this assumption is violated, researchers must search for a satisfactory transformation to produce linearity.
-%As \citet{Halle1962} argues, this eliminates the need for additional grammatical machinery.
-%\begin{quote}
-%\ldots{}we are forced to incorporate into every complete generative grammar a characterization of the distinction between admissible and inadmissible segment sequences.
-The fact effectively cuts the ground out from under the recent suggestion that generative grammars be supplemented with special phonological grammars\ldots{} \citep[][61]{Halle1962}
 %\end{quote}
+%The similarity between sequence structure constraints and neutralizing phonological rules was noted by \citet[][401f.]{Stanley1967}.
+Later discussions tend to view it as a problem (\citealt{Kisseberth1970b}, \citealt{Kenstowicz1977}).
+The relationship between surface and underlying constraints is taken up in the next two chapters.
+%\citealt{Barton1986}, \citealt{Barton1987}, \citealt{Anderson1988a}, \citealt{Ristad1994}
+%\citet{Vitevitch1999b} % summary paper
+%\citealt{Kahlen-Halstenbach1990}, \citealt{Inozuka1991})

### 2012-06-30 3.tex
Feverish writing on chapters 2-4

@@
-%\chapter{Static and derived phonotactic constraints in Turkish} 
-\label{turkish}
+\chapter{Static and derived phonotactic preferences in Turkish}
 
@@
 This is a principled null hypothesis, but one which can be quickly rejected.
-\citeauthor{Chomsky1965} point to robust contrasts between equally unattested items like *\emph{blick}, which is recognized by speakers a possible word of English, and *\emph{bnick}, which is not.\footnote{Any readers who are skeptical that \citeauthor{Chomsky1965}'s intuitions generalize will find support for their claim in the wordlikeness judgement data reported in Appendix \ref{albright}.}
-%As a quick validation of this claim, consider the results reported by \citet{Albright2007} and discussed in the previous chapter.
-English speakers asked to rate non-words on a scale from 1-7, where 7 indicates ``most like English'', favor non-words [blʌs, blɑd] (average scores 4.67 and 5.13) over [bnʌs, bnɑd] (average scores 2.06, 2.00).
+\citeauthor{Chomsky1965} point to robust contrasts between equally unattested items like *\emph{blick}, which is recognized by speakers a possible word of English, and *\emph{bnick}, which is not.
 This preference must be regarded as a learned behavior, since many languages permit  \emph{bn} onsets (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet').
@@
 
-\begin{example}
-Turkish vowel features:
-
-\vspace{0.5\baselineskip}
+\begin{example}[Turkish vowel features]
 \begin{tabular}{c | c c c c}
@@
                        & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
-\buf[$+$\textsc{High}] & {i} & {ü} [y] & {ı} [ɯ] & {u} \\
-\buf[$-$\textsc{High}] & {e} & {ö} [ø] & {a} [ɑ] & {o} \\
+\buf[$+$\textsc{High}] & \emph{i} & \emph{ü} [y] & \emph{ı} [ɯ] & \emph{u} \\
+\buf[$-$\textsc{High}] & \emph{e} & \emph{ö} [ø] & \emph{a} [ɑ] & \emph{o} \\
 \end{tabular}
@@
 %\citep[after][229]{Clements1982}: \\
-\begin{example}
-\textsc{Backness Harmony}:
-
+\begin{example}[\textsc{Backness Harmony}]
 \xymatrix@R=24pt@C=24pt{
@@
 
-\begin{example}
-\label{turknom}
-The Turkish nominative:
-
-%\vspace{0.5\baselineskip}
+\begin{example}[The Turkish nominative] \label{turknom}
 \begin{tabular}{l l l l l}
-   & \emph{nom.sg.} & \emph{nom.pl.} \\
-a. & ip             & ipler          & `rope' & \citep[][216]{Clements1982} \\
-   & el             & eller          & `hand'    \\
-   & köy            & köyler         & `village' \\
-   & yüz            & yüzler         & `face'    \\
-   & kız            & kizlar         & `girl'    \\
-   & sap            & saplar         & `stalk'   \\
-   & son            & sonlar         & `end'     \\
-   & pul            & pullar         & `stamp'   \\
-b. & neden          & nedenler       & `reason'  & (TELL) \\ % front
-   & kiler          & kilerler       & `pantry'           \\ % front
-   & pelür          & pelürler       & `tissue paper'     \\ % back
-   & boğaz          & boğazlar       & `throat'           \\ % back
-   & sapık          & sapıklar       & `pervert'          \\ % back
+   & nom.sg.      & nom.pl. \\
+a. & \emph{ip}    & \emph{ipler}    & `rope'        & \citep[][216]{Clements1982} \\
+   & \emph{el}    & \emph{eller}    & `hand'    \\
+   & \emph{köy}   & \emph{köyler}   & `village' \\
+   & \emph{yüz}   & \emph{yüzler}   & `face'    \\
+   & \emph{kız}   & \emph{kızlar}   & `girl'    \\
+   & \emph{sap}   & \emph{saplar}   & `stalk'   \\
+   & \emph{son}   & \emph{sonlar}   & `end'     \\
+   & \emph{pul}   & \emph{pullar}   & `stamp'   \\
+b. & \emph{neden} & \emph{nedenler} & `reason'       & (TELL) \\ % front
+   & \emph{kiler} & \emph{kilerler} & `pantry'       \\ % front
+   & \emph{pelür} & \emph{pelürler} & `tissue paper' \\ % back
+   & \emph{boğaz} & \emph{boğazlar} & `throat'       \\ % back
+   & \emph{sapık} & \emph{sapıklar} & `pervert'      \\ % back
 \end{tabular}
@@
 
-\begin{example}
-\label{turkexcept}
-Exceptional Turkish nominatives:
-
-%\vspace{0.5\baselineskip}
+\begin{example}[Exceptional Turkish nominatives] \label{turkexcept}
 \begin{tabular}{l l l l l}
-   & \emph{nom.sg.} & \emph{nom.pl.} \\
-a. & mezar          & mezarlar       & `grave' & \citep{TELL} \\
-   & model          & modeller       & `model' \\
-   & silah          & silahlar       & `weapon'     \\
-   & memur          & memurlar       & `bureaucrat' \\
-   & sabun          & sabunlar       & `soap'       \\
-b. & saat           & saatler        & `hour, clock' \\
-   & harf           & harfler        & `(alphabetic) letter' \\ %& \citep{Goksel2005}
-   & etol           & etoller        & `fur stole' \\
+   & {nom.sg.}    & {nom.pl.} \\
+a. & \emph{mezar} & \emph{mezarlar} & `grave' & \citep{TELL} \\
+   & \emph{model} & \emph{modeller} & `model' \\
+   & \emph{silah} & \emph{silahlar} & `weapon'     \\
+   & \emph{memur} & \emph{memurlar} & `bureaucrat' \\
+   & \emph{sabun} & \emph{sabunlar} & `soap'       \\
+b. & \emph{saat}  & \emph{saatler}  & `hour, clock' \\
+   & \emph{harf}  & \emph{harfler}  & `(alphabetic) letter' \\ %& \citep{Goksel2005}
+   & \emph{etol}  & \emph{etoller}  & `fur stole' \\
 \end{tabular}
@@
 
-\begin{example}
-\label{spec}
-Underspecification in harmonic roots (after \citealp{Clements1982}):
-
-%\vspace{0.5\baselineskip}
+\begin{example}[Underspecification in harmonic roots (after \citealp{Clements1982})] \label{spec}
 \xymatrix@R=24pt@C=24pt{
-\txt{a.} & \txt{harmonic root:~~~~} & \txt{C} & \txt{V} & \txt{C} & \txt{V} \\
+\txt{a.} & \txt{harmonic root:~~~~} & \txt{C} & \txt{V} & \txt{C} & \txt{V} & \txt{\ldots} \\
 &   &    & \txt{[$-$\textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
-\txt{b.} & \txt{disharmonic root:} & \txt{C} & \txt{V} & \txt{C} & \txt{V} \\
+\txt{b.} & \txt{disharmonic root:} & \txt{C} & \txt{V} & \txt{C} & \txt{V} & \txt{\ldots} \\
     &    &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u] & & \txt{[$+$\textsc{Back}]}\ar@{-}[u]
@@
 The adaptation of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, is one such case.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} While some speakers are said to be capable of pronouncing these non-native clusters, in normal speech the cluster is resolved by epenthesis of a [$+$\textsc{High}] vowel.
-In most cases, the epenthetic vowel matches the following root vowel for backness, suggesting that adapt foreign words in a way that conforms to the root harmony generalization.
+In most cases, the epenthetic vowel matches the following root vowel for backness, suggesting that adapt foreign words in a way that conforms to the root harmony generalization.\footnote{At the same time, this shows that adaptation is not simply applying rules, like \textsc{Backness Harmony}, to foreign forms.
+Rather, foreign forms are restructured as non-exceptional Turkish lexical items.}
 
@@
 
-\begin{example}
-Adaptation of initial foreign clusters \citep[][247]{Clements1982}:
-
-%\vspace{0.5\baselineskip}
+\begin{example}[Adaptation of initial foreign clusters (\citealp{Clements1982}:247)] 
 \begin{tabular}{l l l l l l}
-a. & spiker  & \alt{} & sipiker  & `announcer' \\
-   & fren    & \alt{} & firen    & `break'     \\
-b. & brom    & \alt{} & burom    & `bromide'   \\
-   & prusya  & \alt{} & purusya  & `Prussia'   \\
+a. & \emph{spiker}  & \alt{} & \emph{sipiker}  & `announcer' \\
+   & \emph{fren}    & \alt{} & \emph{firen}    & `break'     \\
+b. & \emph{brom}    & \alt{} & \emph{burom}    & `bromide'   \\
+   & \emph{prusya}  & \alt{} & \emph{purusya}  & `Prussia'   \\
 %b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
@@
 
-\begin{example}
-\label{redupgame}
-Turkish reduplication game \citep[][231]{Harrison2001}:
-
-\vspace{0.5\baselineskip}
+\begin{example}[Turkish reduplication game (\citealp{Harrison2001}:231)] \label{redupgame}
 \begin{tabular}{l l l l}
-a. & kibrit & kibrit-\{kabrıt\} & `match'    \\
-   & bütün  & bütün-\{batın\}   & `whole'    \\
-b. & mali   & mali-\{muli\}     & `Mali'     \\
-   & butik  & butik-\{batik\}   & `boutique' \\
+a. & \emph{kibrit} & \emph{kibrit}-\{\emph{kabrıt}\} & `match'    \\
+   & \emph{bütün}  & \emph{bütün}-\{\emph{batın}\}   & `whole'    \\
+b. & \emph{mali}   & \emph{mali}-\{\emph{muli}\}     & `Mali'     \\
+   & \emph{butik}  & \emph{butik}-\{\emph{batik}\}   & `boutique' \\
 \end{tabular}
@@
 
-A number of studies on Finnish suggest that speakers of that language, which has a vowel harmony system similar to that of Turkish, make use of (dis)harmony to processing running speech.\footnote{Thanks to Charles Yang for bringing these studies to my attention.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which conform to backness harmony.
+\label{harmonyandsegmentation} A number of studies on Finnish suggest that speakers of that language, which has a vowel harmony system similar to that of Turkish, make use of (dis)harmony to processing running speech.\footnote{Thanks to Charles Yang for bringing these studies to my attention.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which conform to backness harmony.
 These stimuli are auditorily presented to subjects who are asked to press a button when the nonce trisyllable ends with a target nonce disyllable, or a real disyllabic word.
@@
 
-\begin{example}
-\textsc{Roundness Harmony}:
-
+\begin{example}[\textsc{Roundness Harmony}]
 \xymatrix@R=24pt@C=24pt{
@@
 
-\begin{example}
-Turkish nominal suffix allomorphy:
-
-%\vspace{0.5\baselineskip}
+\begin{example}[Turkish nominal suffix allomorphy]
 \begin{tabular}{l l l l l l l}
-   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
-a. & ip             & ipi            & ipin           & `rope' & \citep[][216]{Clements1982} \\
-   & el             & eli            & elin           & `hand'    \\
-   & kız            & kızı           & kızın          & `girl'    \\
-   & sap            & sapı           & sapın          & `stalk'   \\
-   & yüz            & yüzü           & yüzün          & `face'    \\
-   & köy            & köyü           & köyün          & `village' \\
-   & pul            & pulu           & pulun          & `stamp'   \\
-   & son            & sonu           & sonun          & `end'     \\
-b. & boğaz          & boğazı         & boğazın        & `throat'  & (TELL) \\
-   & pelür          & pelürü         & pelürün        & `tissue paper' \\
-   & döviz          & dövizi         & dövizin        & `currency' \\
-   & yamuk          & yamuğu         & yamuğun        & `trapezoid' \\
-   & ümit           & ümiti          & ümitin         & `hope'     \\
+   & {nom.sg.}    & {dat.sg.}     & {gen.sg.}  \\
+a. & \emph{ip}    & \emph{ipi}    & \emph{ipin}    & `rope' & \citep[][216]{Clements1982} \\
+   & \emph{el}    & \emph{eli}    & \emph{elin}    & `hand'    \\
+   & \emph{kız}   & \emph{kızı}   & \emph{kızın}   & `girl'    \\
+   & \emph{sap}   & \emph{sapı}   & \emph{sapın}   & `stalk'   \\
+   & \emph{yüz}   & \emph{yüzü}   & \emph{yüzün}   & `face'    \\
+   & \emph{köy}   & \emph{köyü}   & \emph{köyün}   & `village' \\
+   & \emph{pul}   & \emph{pulu}   & \emph{pulun}   & `stamp'   \\
+   & \emph{son}   & \emph{sonu}   & \emph{sonun}   & `end'     \\
+b. & \emph{boğaz} & \emph{boğazı} & \emph{boğazın} & `throat'  & (TELL) \\
+   & \emph{pelür} & \emph{pelürü} & \emph{pelürün} & `tissue paper' \\
+   & \emph{döviz} & \emph{dövizi} & \emph{dövizin} & `currency' \\
+   & \emph{yamuk} & \emph{yamuğu} & \emph{yamuğun} & `trapezoid' \\
+   & \emph{ümit}  & \emph{ümiti}  & \emph{ümitin}  & `hope'     \\
 \end{tabular}
@@
 
-\begin{example}
-\textsc{Labial Attraction}: %(after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}):
-
+\begin{example}[\textsc{Labial Attraction}]
+%(after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}): 
 \xymatrix@R=24pt@C=8pt{
@@
 \begin{quote}
-\ldots decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots~aCu~\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots We conclude that there is no systematic restriction on the set of consonants that may occur medially in rotos of the form /\ldots~aCu~\ldots/.
+\ldots{}decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots~aCu~\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots We conclude that there is no systematic restriction on the set of consonants that may occur medially in roots of the form /\ldots~aCu~\ldots/.
 \citep[][225]{Clements1982}
@@
 
-\subsection{Lexical statistics}
-\label{lexstats}
+\subsection{Lexical statistics} \label{lexstats}
 
@@
 
-\begin{example}
-$\displaystyle \gamma = \frac{C - D}{C + D}$
-\end{example}
-
 \subsubsection{Backness harmony}
 
-\begin{example}[Backness harmony wordlikeness forced choices (\citealp{Zimmer1967}:314)]
+\begin{example}[Backness harmony wordlikeness forced choices (\citealp{Zimmer1969}:314)]
 \begin{tabular}{l r l r}
@@
 \midrule
-pemez & 30                   & pemaz & 2  \\
-tipez & 24                   & tipaz & 8  \\ 
-terüz & 19                   & teruz & 13 \\ % roundness violator
-teriz & 28                   & terız & 3  \\
-tokaz & 26                   & tokez & 6  \\ % roundness violator
+\emph{pemez} & 30            & \emph{pemaz} & 2  \\
+\emph{tipez} & 24            & \emph{tipaz} & 8  \\ 
+\emph{terüz} & 19            & \emph{teruz} & 13 \\ % roundness violator
+\emph{teriz} & 28            & \emph{terız} & 3  \\
+\emph{tokaz} & 26            & \emph{tokez} & 6  \\ % roundness violator
 \bottomrule
@@
 \midrule
-pörüz & 32 & pöriz & 0  \\
-tatız & 20 & tatuz & 12 \\
-tüpüz & 31 & tüpiz & 1  \\
-takız & 22 & takuz & 10 \\
+\emph{pörüz} & 32            & \emph{pöriz} & 0  \\
+\emph{tatız} & 20            & \emph{tatuz} & 12 \\
+\emph{tüpüz} & 31            & \emph{tüpiz} & 1  \\
+\emph{takız} & 22            & \emph{takuz} & 10 \\
 \bottomrule
@@
 \toprule
-\multicolumn{2}{l}{aPu} & \multicolumn{2}{l}{aPı} \\
+\multicolumn{2}{l}{\emph{a}P\emph{u}} & \multicolumn{2}{l}{\emph{a}P\emph{ı}} \\
 \midrule
-pamuz & 17              & pamız & 15 \\
-tafuz & 21              & tafız & 11 \\
-tapuz & 17              & tapız & 15 \\
-mavuz & 16              & mavız & 16 \\
-tabuz & 16              & tabız & 16 \\
+\emph{pamuz} & 17       & \emph{pamız} & 15 \\
+\emph{tafuz} & 21       & \emph{tafız} & 11 \\
+\emph{tapuz} & 17       & \emph{tapız} & 15 \\
+\emph{mavuz} & 16       & \emph{mavız} & 16 \\
+\emph{tabuz} & 16       & \emph{tabız} & 16 \\
 \bottomrule
@@
 \begin{tabular}{l l l l}
-kabiːr  & `big'        & kabura  & `become big'       \\
-ħasan   & `handsome'   & ħasuna  & `become beautiful' \\
-dʒadiːb & `barren'     & dʒaduba & `become dry'       \\
+\hspace{0pt}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
+\hspace{0pt}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
+\hspace{0pt}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
 \end{tabular}
@@
 \begin{tabular}{l r r r}
-        & aPu & aPı & $p$-value \\
+        & \emph{a}P\emph{u} & \emph{a}P\emph{ı} & $p$-value \\
 Native  & 12                & 11                & \multirow{2}{*}{0.042} \\
@@
 If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
-
 %\footnote{I note a potential problem with this account: \citet{Davidson2005,Davidson2006a} argue that this ``excrescent'' schwa is distinct from the both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.}
-
 %\noindent A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions.
@@
 The set of possible constraints must respect results in formal learning theory, and may also be constrained by markedness considerations, for example.
-
 %A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
 This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).
-
 %Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
 In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.
-
 %In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word. 
@@
 %The data below, drawn from Turkish, shows that 
-
 %This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
@@
 Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
+%\begin{unlabeledexample}
+%$\displaystyle \gamma = \frac{C - D}{C + D}$
+%\end{unlabeledexample}
+%\footnote{Any readers who are skeptical that \citeauthor{Chomsky1965}'s intuitions generalize will find support for their claim in the wordlikeness judgement data reported in Appendix \ref{albright}.}
+%As a quick validation of this claim, consider the results reported by \citet{Albright2007} and discussed in the previous chapter.
+English speakers asked to rate non-words on a scale from 1-7, where 7 indicates ``most like English'', favor non-words [blʌs, blɑd] (average scores 4.67 and 5.13) over [bnʌs, bnɑd] (average scores 2.06, 2.00).

### 2012-06-30 4.tex
Feverish writing on chapters 2-4

@@
-%\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
-\label{clusters}
+%\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter was presented at the 30th West Coast Conference on Formal Linguistics and will appear in the proceedings.}} \label{clusters}
+\chapter{Structural and accidental gaps} \label{clusters}
+
+This chapter
+
+It has often been noted that these regularities, while perhaps statistical in nature,
+
+\begin{quote}
+These regularities, although not the products of alternations, are described and defined by elements of phonological theories: segments\ldots{}and distinctive features. \citep[13]{Berkley2000}
+\end{quote}
 
@@
 %\citep[][269]{LANGUAGE}.
-
-In the preceding chapter, it was proposed that
-
-Reconsider the null hypothesis presented in the preceding chapter:
 
@@
 Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
-This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a} but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Dmitrieva2008a,Dmitrieva2008b,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.
+This technique has been applied to consonants in Semitic, especially Arabic 
+(\citealt{McCarthy1988,McCarthy1994}, \citealt{Pierrehumbert1993}, \citealt{Frisch1996,Frisch2004}, \citealt{Coetzee2008a}) but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English (\citealt{Berkley1994b,Berkley1994a}, \citealt{Pierrehumbert1994}, \citealt{Coetzee2008b}), Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.
 
@@
 
-\section{Domain}
+\section{The static phonotactics hypothesis and the lexicon}
 
@@
 
+The focus of this study are the constraints on syllable contact clusters (i.e., word-medial consonant clusters spanning syllable boundaries), also known as \emph{interludes}, in English monomorphs.
+
+\citep{Haugen1956}
+\footnote{Thanks to Stephen Anderson for bringing this study to my attention.}
+
+\begin{quote}
+\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
+\end{quote}
+
+\noindent Before considering any ``additional provisos'', it is important to note that neither this study nor the study of \citet{Pierrehumbert1994} are not so much concerned so much with what clusters are ``possible'', but rather which are attested in the English lexicon.
+
+One such proviso 
+The neutralizing effect of phonological processes are one such proviso. Consider
+
+\subsection{The focus on monomorphs}
+
+This requires a bit of history.
+
+A major result of American structuralism is that generalizations which hold of the sound sequences found in underlying forms may not hold across morph boundaries.
+In his history of phonology, \citet[][267]{Anderson1985} indirectly credits this insight to \citet{Bloomfield1930}, who proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries.
+\citeauthor{Bloomfield1930} notes that [ç] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [kuːxən] `cake' $\sim$ \emph{Kuhchen} [kuːçən] `cow-let'.
+\citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf.
+\emph{Kuh} `cow').
+Under the standard analysis that [x] is the allophone of /ç/ before back vowels,
+all that needs to be added is a restriction that the conditioning vowel belongs to the same word (or perhaps morph).
+
+A more sophisticated approach to problems of this type, making no reference to \citeauthor{Bloomfield1930}, was published only two years later.
+\citet{Jakobson1932}, then a member of the Prague circle,
+
+attributes a number of phonological properties of the Russian imperative to ``juncture''.
+
+Nearly all Russian infinitives are marked with a final theme vowel and /-t\pal{}/; in reflexives, this is followed by a suffix realized as [s\pal{}ə], which triggers a general process of regressive voicing assimilation.
+In the reflexive infinitive, the palatalization of infinitive /-t\pal/ is lost, but in the reflexive imperative, palatalization of a stem-final consonant is preserved, even when the root-final consonant is /t\pal/, as in (\ref{rusreflex}c).\footnote{Thanks to Lev Blumenfeld for assistance with these transcriptions.}
+
+\begin{example}[Russian reflexives] \label{rusreflex}
+\begin{tabular}{l l l l}
+   &  infinitive       & imperative \\
+a. & [slav\pal{}itsə]  & [slaf\pal{}s\pal{}ə]  & `be glorious'   \\
+   & [upram\pal{}itsə] & [upram\pal{}s\pal{}ə] & `be stubborn'   \\
+b. & [kras\pal{}itsə]  & [kras\pal{}s\pal{}ə]  & `put on makeup' \\
+   & [ʒar\pal{}itsə]   & [ʒar\pal{}{}s\pal{}ə] & `roast'         \\
+c. & [zəbytsə]         & [zəbut\pal{}s\pal{}ə] & `forget'        \\
+\end{tabular}
+\end{example}
+
+\noindent \citeauthor{Jakobson1932} proposes that the preservation of palatalization in /\ldots{}t\pal{}-s\ldots{}/ is a special property of the imperative.
+Both (\ref{rusreflex}c) forms contain a /t\pal-s/ juncture, though palatalization is only preserved in the imperative.
+\citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization.
+A few years later, \citet{Trnka1936} makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.
+
+This is apparent from a discussion of constraints on vowel sequences in Mixteco (a dialect cluster also known as Mixtec) given by \citet{Pike1947b}.
+It is apparent that \citeauthor{Pike1947b}'s constraints on vowel sequences are not intended to hold across morph boundaries, according to the phonological analysis of a Mixteco text published a few years earlier \citep{Pike1944}.
+A few examples are given below; the transcriptions are \citeauthor{Pike1944,Pike1947b}'s.
+
+\begin{example}[Mixteco MSCs and complex words \citep{Pike1944,Pike1947b}]
+\begin{tabular}{l l@{} l l l}
+a. & * & {C}a{C}e & [ká\textsuperscript{n}dee]    & `kept inside'   \\
+b. & * & {C}ə{C}e & [nìk\`əbəde]                  & `who entered'   \\
+c. & * & {C}e{C}i & [teníke\textsuperscript{n}da] & `was walking    \\
+d. & * & {C}i{C}e & [tenìkeet\`ə]                 & `and went away' \\
+\end{tabular}
+\end{example}
+
+\noindent \citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a discovery procedure for morphs (see also \citet{Hafer1974}).
+
+and lead children \citep{Mattys2001b,Jusczyk2002,Davidson2004} and adults \citep{McQueen1998,Hay2004a} to posit morph boundaries (see also \S\ref{harmonyandsegmentation}).
+
+\section{Evaluation}
+
+% 
+%
+
 \subsection{Data sources}
+
+a categorization of morphemic complexity 
+and broad transcriptions
 
@@
 
-\citet{Aronoff1976}
+\subsubsection{Identifying monomorphs}
 
-\citet{Harley2009}
+% Pierrehumbert informal judgements * \citet{Pierrehumbert1994} and \citet{Hammond1999a} technique is not replicable, so we turn to CELEX
+% not replicable, either the experience nor the data itself
+% Some of Pierrehumbert's and some of Hammond's are ridiculous
+% Despite this, we preserve some Pierrehumbertian generalizations
 
-\citet{Taft1975} ?
-\citet{Taft1981} ?
-\citet{Taft2004a} ?
+Beyond this, there is a potential for bias or circularity of argumentation.
+As mentioned above (\S\ref{pike}), a sequence which is rare in monomorphs is sometimes said to mark a juncture even in the absense of other evidence for segmentation.
+But it is certainly incorrect to use this principle to argue for a juncture if the argument is only being made to preserve this very phonotactic generalization, as \citet[186]{Pierrehumbert1994} does with the word \emph{antler}.
 
-\citet{Halle1962}
+% Compounds, regulars, and more productive derivatives are pretty clear 
+% Latin is less clear: \citet{Bolinger1961} on stress, and \emph{SPE} and \citet[11--13]{Aronoff1976} on forms of prefixes and bound stems
+%There has been a good deal of attention paid to Latinate prefixes
+%\citet{Bolinger1961} suggests this for English stress.
+%\citet{SPE}
 
-% 2.1.4: k
+\begin{example}[Bound stem-specific allomorphy]
+\begin{tabular}{l l l}
+a. & \emph{adhere}   & \emph{adhesion}   \\
+   & \emph{cohere}   & \emph{cohesion}   \\
+b. & \emph{conceive} & \emph{conception} \\
+   & \emph{perceive} & \emph{perception} \\
+\end{tabular}
+\end{example}
 
-Under the parses and phonologization described immediately below, the 6,876 simplex CELEX entries contain 21 medial coda types and 40 unique medial onsets.
-Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, a saturation rate of 18.8\%.
+Note that this is common behavior in other languages in which the independent nature of bound stems is less controversial, such as Russian \citep{Pesetsky1977}.
 
-\subsection{Syllabification}
+% Harley2009 \citet{Gropen1989} \citet[56]{Coppock2008} observes that there is every reason to believe that the ditransitive is productive in the sense that it can be applied to new words; consider recent conversion \emph{text} (e.g., \emph{text me your number}).
 
-For this study, it is necessary to separate medial consonant clusters into codas and onsets.
-While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic.
-For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language.
-Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which explain the \emph{chemistry}/\emph{ministry} syllabification contrast.
+\begin{example}[Latinate verbs and small clauses]
+\begin{tabular}{l l@{} l l l@{}}
+a. & ~ & {show it off}           & * & {exhibit it off}           \\
+b. & ~ & {drink himself stupid}  & * & {imbibe himself stupid}    \\
+c. & ~ & {show him the painting} & * & {exhibit him the painting} \\
+\end{tabular}
+\end{example}
 
-An automated syllabification system was constructed and applied to medial clusters in the CELEX data.
-This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.
-There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification.
-The technique used is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948},
-\citealt[75]{Pulgram1970}, \citealt[42f.]{Kahn1976}, \citealt[][358f.]{Selkirk1982b}) which parses word-medial clusters so that the onset is as large as possible.
-Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty.
-In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.
+% While this is precisely the abstractness that  ???, it is reasonably well supported by psycholinguistic results in lexical decision research.
+
+% description of taft model \citet{Taft1975} propose an influential model of morphological processing.
+The first step in the \citeauthor{Taft1975} model is to strip affixes from possible roots.
+Root and affix are then looked up in lexical memory.
+At the final stage, they are recombined.
+% It has been suggested since then that full decomposition is only one of two ways to recognize complex words, the other being a whole word storage route running in cascade \citep{Caramazza1988} or in parallel \citep{Baayen1997b}.
+Whatever evidence is amassed for this (see \citealp{Lignos2012} for a critical review) but few if any lexical decision researchers would deny the possibility of \citep{Butterworth1983}.
+
+% first experiment \citet{Manelis1977} who suggest poor stimulus controls \citeauthor{Taft1975}
+
+% that *\emph{devive} in the Latinate prefix \emph{de-} applies to a bound stem \emph{vive} (e.g., \emph{revive}) is slower to be recognized than *\emph{debel}, since \emph{bel} is not recognized as abound stem in words like \emph{rebel}.
+
+% pseudoprefixed 
+% \citet{Taft1981} 
+% \citet{Rubin1979}
+%\emph{regatta}
+%\emph{retrieve}
+
+%\citet{Taft2004b}
+%\emph{virus} is primed by \emph{viral}
+
+\subsubsection{Lexical statistics}
+
+The English portion of CELEX contains 6,876 simplex words.
+Under the syllabification procedure described immediately below, this yields contain 21 medial coda types and 40 unique medial onsets.
+Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, an 18.8\% saturation rate.
+
+Cluster frequencies are reproduced in Appendix \ref{appendixB}.
+
+\subsubsection{Syllabification}
+
+CELEX provides transcriptions of individual wordforms, include syllabification.
+However, the authors do not describe the syllabification procedure, and the syllable boundaries appear to be placed in an unsystematic fashion.
+For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][II.294]{AOE}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast does not conform to native speaker intuition, and it is typologically dubious: there is no convincing evidence for contrastive syllabification in any language.
+Apparent counterexamples appear to be interactions of syllabification with vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which might predict a different syllabification for \emph{chemistry} and \emph{ministry}.
+Many other instances of contrasts in syllabification can be found in the CELEX data.
+
+For this reason, an automated syllabification procedure was developed to separate medial clusters in the CELEX  monomorphs into coda and onset.
+Since this procedure is applied only to medial clusters, it is possible to remain agnostic on several contentious issues concerning syllabification of English; for instance, there is no need to address the status of so-called ``ambisyllabic'' consonants, since this does not occur with medial clusters, or morphological effects on syllabification, since only monomorphs are used.
+
+The automated syllabifier is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948}, \citealt[75]{Pulgram1970}, \citealt[42f.]{Kahn1976}, \citealt[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset.
+Just what is a ``possible'' medial onset is defined by the onsets that occur in initial position.
+While this equation may not be appropriate for all languages \citep[306]{Fischer-Jorgensen1952}, it is for English.
+Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), so onset maximization assigns the entire medial cluster to the onset, leaving the medial coda empty.
+In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} is not found word-initially; the maximal onset here is [str], and the residual [n] is assigned to the coda.
 
 There is one case where unchecked maximization of medial onsets produces incorrect syllabifications.
-When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
-\citet[][55]{Harris1994} notes that onset maximization makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
-The approach adopted here is essentially that of \citet{Pulgram1970}: the first consonant of a medial cluster is assigned to the coda of a preceding coda if is immediately preceded by a stressed lax vowel, which do not occur syllable-finally in English.\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}
+When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
+\citet[][55]{Harris1994} note that onset maximization produces an incorrect syllabification when the medial cluster is a valid onset: in \emph{whisper}, \emph{vestige}, and \emph{musket}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
+The approach adopted here is is to assign the first consonant of a medial cluster to the coda of a preceding syllable if it follows a stressed lax vowel \citep[cf.][48]{Pulgram1970}.%\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}
 
-If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables.
-For this reason, an additional constraint is assumed which prevents the two components of the affricates from being split by syllabification.
-This is motivated by the tendency of affricates to pattern with single segments in many languages.
-For instance, the only complex onsets in Classical Nahua are the affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.
+\paragraph{The status of affricates} At least as far back as the 1939 first edition of the \emph{Grundz{ü}ge der Phonologie} \citep{Trubetzkoy1958}, affricates have been analyzed as single plosive segments distinguished from stops by the presence of stridency \citep[24]{Jakobson1961} or a delayed release (\emph{SPE}:321f).
+If however, the English affricates [tʃ, dʒ] are treated as stop-fricative sequences (e.g., \citealt{Hualde1988}, \citealt{Lombardi1990}), this procedure will incorrectly assign the two components of the medial affricate to separate syllables before stressed lax vowels (e.g., \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}) unless there is an ad-hoc constraint that treats these sequences as indivisible \citep[e.g.,][]{Wells1990}.
+Here, the traditional treatment of affricates as single segments is assumed.
+This assumption can be motivated by the tendency of affriates to pattern with individual segments in phonotactic generalizations.
+For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but otherwise does not permit other onset clusters \citep[9]{Launey2011}, and \citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.
 
-\subsection{Inventory consdierations and subsyllabic parsing}
+\subsubsection{Subsyllabic parsing}
 
-To identify the inventory of syllable contact clusters, it is also necessary to parse syllables into onset, nucleus, and coda.
-In most cases this is trivial, but in a few cases, it is necessary to use distributional heuristics to determine the underlying inventory and subsyllabic affiliations.
+It is not trivial in all cases to separate interludes from the nuclei on either side. Immediately below, distributional heuristics are used to develop a subsyllabic parsing procedure for English.
 
-\subsubsection{The velar nasal} \label{velarnasal}
+\paragraph{The velar nasal} There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right \citep[e.g.,][]{Sapir1925}, an analysis which is a consequence of principles like the Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ when followed by underlying /k, ɡ/ (e.g., \emph{SPE}:85, \citealt{Borowsky1986}:65f.), sometimes followed by the deletion of /ɡ/.
+The strongest piece of evidence for the latter position, in which the velar nasal is a pure allophone, is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
+Following \citet{Pierrehumbert1994}, the pure allophonic analysis is adopted here, and English nasal place allophony is formalized in Section \ref{npasection} below.
 
-There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right \citep[e.g.,][]{Sapir1925}, an analysis which is a consequence of principles like the Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ (e.g., \emph{SPE}:85, \citealt{Borowsky1986}:65f.).
-The strongest piece of evidence for the latter position, in which the velar nasal is a pure allophone, is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
-\citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here.
-English nasal place allophony is formalized in Section \ref{cnpasection} below.
-
-\subsubsection{[j] onglides}
-
-The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J.
-\citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}).
+\paragraph{[j] onglides} The [j] onglide in words such as \emph{val}[j]\emph{ue} are assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}).
 The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).
-
 There is considerable evidence for this assumption.
 When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː].
-In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}.
+%In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}. 
 This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
-\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear.
+\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters, though they may be followed by [ju] in words like [mj]\emph{use} or [vj]\emph{iew}, also suggesting the nuclear affiliation of the glide.
 There is also strong external support for this principle.
-The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants.
-The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor} [h]\emph{omponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
+The [ju] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005}, \emph{shm}-reduplication \citep{Nevins2003}, and speech errors, such as [kjuː]\emph{mor} [h]\emph{omponent} for intended [hju]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
 
-\subsubsection{[w] onglides}
-
-The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset.
+\paragraph{[w] onglides} The selective properties of the back onglide [w] contrast sharply with those of the front onglide.
 Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
@@
 Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.
-And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.
+Together, this indicates taht the back onglide is assigned to the onset.
 
-\subsubsection{Post-vocalic \emph{r}}
+\paragraph{Post-vocalic \emph{r}} Both CELEX and the dictionary used by \citet{Pierrehumbert1994} transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost.
+Even in \emph{r}-full dialects, though, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters.
+Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}); for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}.
+Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which it patterns differently than internal coda consonants.
+\citet[258]{Harris1994} reports that post-vocalic \emph{r} is the only consonant which does not block a variable process of /t/-\textsc{Glottalization} found in many British dialects.
 
-Both CELEX and the dictionary used by \citet{Pierrehumbert1994} in her study transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost.
-In \emph{r}-full dialects, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters.
-Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}); for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}.
-Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which post-vocalic \emph{r} patterns differently than internal coda consonants.
-\citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.
-
-\begin{example}[/t/-\textsc{Glottalization} in British English (after \citealp{Harris1994}:195, 258)]
-\begin{tabular}{l l l@{} l}
-a. & par[t]   &   & par[ʔ]   \\
-   & car[t]on &   & car[ʔ]on \\
-b. & fis[t]   & * & fis[ʔ]   \\
-   & mis[t]er & * & mis[ʔ]er \\
+\begin{example}[/t/-\textsc{Glottalization} in British English]
+\begin{tabular}{l l l l@{} l}
+a. & \emph{des}[ɚt]         & \alt{} &   & \emph{des}[ɚʔ]         \\
+   & \emph{c}[ɚt]\emph{ain} & \alt{} &   & \emph{c}[ɚʔ]\emph{ain} \\
+b. & \emph{fi}[st]          & \alt{} & * & \emph{fi}[sʔ]          \\
+   & \emph{mi}[st]\emph{er} & \alt{} & * & \emph{mi}[sʔ]\emph{er} \\
 \end{tabular}
@@
 
-\noindent Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.
+Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.
 
 \begin{example}[/t, d/-\textsc{Deletion} in American English] \label{td}
-\begin{tabular}{l l l@{} l}
-a. & be[lt]  &   & be[l] \\
-   & me[nd]  &   & me[n] \\
-b. & sh[ɚt]  & * & sh[ɚ] \\
-   & c[ɚd]   & * & c[ɚ]  \\
+\begin{tabular}{l l l l@{} l}
+a. & \emph{be}[lt] & \alt{} &   & \emph{be}[l] \\
+   & \emph{me}[nd] & \alt{} &   & \emph{me}[n] \\
+b. & \emph{sh}[ɚt] & \alt{} & * & \emph{sh}[ɚ] \\
+   & \emph{c}[ɚd]  & \alt{} & * & \emph{c}[ɚ]  \\
 \end{tabular}
 \end{example}
-
-\section{Evaluation}
 
@@
 
-\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English).
+\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000} on English).
 The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented.
@@
 
-\noindent
-It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.
+\noindent It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.
 
@@
 
-\subsection{Derived constraints}
-
 \emph{SPE} describes three phonological alternations which target syllable contact clusters. The corresponding constraints on simplex syllable contact clusters are considered below.
@@
 \begin{tabular}{l l l l l l}
-a. & nap & nap[t] \\
-   & nab & nab[d] \\
-b. & lap & lap[s] \\
-   & lab & lab[z] \\
+a. & \emph{nap} & \emph{nap}[t] \\
+   & \emph{nab} & \emph{nab}[d] \\
+b. & \emph{lap} & \emph{lap}[s] \\
+   & \emph{lab} & \emph{lab}[z] \\
 \end{tabular}
@@
 In \emph{SPE} (p.~94), for instance, morphological boundaries are used to simplify stress assignment, and \citet[][546]{Rice2009d} analyses many words in the Athapaskan language Slave as compounds simply because they contain consonant clusters that rarely occur otherwise.
-Recent experiments find that 9-month-old infants are sensitive to the contrast between [vt], a disagreeing cluster which does not occur word-medially, and [ft], found in medial position \citep{Mattys2001b}, and this ability seems to persist into adulthood \citep{McQueen1998}.
+Recent experiments find that 9-month-old infants are sensitive to the contrast between [vt], a disagreeing cluster which does not occur word-medially, and [ft], found in medial position \citep{Mattys2001b}.
+And this ability seems to persist into adulthood (\citealt{Brown1956}, \citealt{Norris1997}, \citealt{McQueen1998}).
 
@@
 \begin{tabular}{l l l l}
-a. & possible  & i[m.p]ossible \\
-   & balance   & i[m.b]alance  \\
-b. & tangible  & i[n.t]angible \\
-   & decent    & i[n.d]ecent   \\
-c. & elegant   & i[n]elegant   \\
-   & ability   & i[n]ability   \\
-d. & mature    & i[m]ature     \\
-   & numerable & i[n]umerable  \\
+a. & \emph{possible}  & \emph{i}[m.p]\emph{ossible} \\
+   & \emph{balance}   & \emph{i}[m.b]\emph{alance}  \\
+b. & \emph{tangible}  & \emph{i}[n.t]\emph{angible} \\
+   & \emph{decent}    & \emph{i}[n.d]\emph{ecent}   \\
+c. & \emph{elegant}   & \emph{i}[n]\emph{elegant}   \\
+   & \emph{ability}   & \emph{i}[n]\emph{ability}   \\
+d. & \emph{mature}    & \emph{i}[m]\emph{ature}     \\
+   & \emph{numerable} & \emph{i}[n]\emph{umerable}  \\
 \end{tabular}
@@
 \citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but that assimilation of \emph{im-}/\emph{in-} is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
-However, the stress condition does not hold in simplex words (e.g., \emph{a}[ŋ.g]\emph{óra}), and thus can be safely ignored here.
+However, the stress condition does not hold in simplex words (e.g., \emph{a}[ŋ.ɡ]\emph{óra}), and thus can be safely ignored here.
 
@@
 \begin{tabular}{l l l}
-a. & primps        & (prints)      \\
-b. & rand orker    & (rank order)  \\
-c. & po[ŋ]cutation & (computation) \\
+a. & \emph{pri}[mp]\emph{s}      & (\emph{prints})      \\
+b. & \emph{ra}[nd] \emph{orker}  & (\emph{rank order})  \\
+c. & \emph{po}[ŋk]\emph{utation} & (\emph{computation}) \\
 \end{tabular} 
@@
 \begin{tabular}{l l l l l}
-a. & burn    & burn[t] \\
-   & spill   & spil[t] \\
-b. & ben[d]  & ben[t]  \\
-   & buil[d] & buil[t] \\
+a. & \emph{burn}    & \emph{burn}[t] \\
+   & \emph{spill}   & \emph{spil}[t] \\
+b. & \emph{ben}[d]  & \emph{ben}[t]  \\
+   & \emph{buil}[d] & \emph{buil}[t] \\
 \end{tabular}
@@
 
-\subsection{Static constraints}
+\subsubsection{Local summary}
+
+All three \emph{SPE} rules affecting syllable consonant clusters are reliable predictors of attested and unattested medial clusters.
+
+\subsubsection{Static constraints}
 
@@
 
-\citet[][173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Since \citeauthor{Pierrehumbert1994}'s study is restricted to triconsonantal clusters, so words like \emph{a}[k.m]\emph{e}, \emph{ru}[g.b]\emph{y}, or \emph{pi}[g.m]\emph{ent} do not violate this narrower generalization.
-Further, \citeauthor{Pierrehumbert1994} adopts the traditional assumption \citep[e.g.,][66f.]{Borowsky1986} that coda [ŋ] is an allophone of /n/ before dorsal consonants, so words such as \emph{mo}[ŋ.gr]\emph{el} are not exceptions.
+\citet[][173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Since \citeauthor{Pierrehumbert1994}'s study is restricted to triconsonantal clusters, so words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} do not violate this narrower generalization.
+Further, \citeauthor{Pierrehumbert1994} adopts the traditional assumption \citep[e.g.,][66f.]{Borowsky1986} that coda [ŋ] is an allophone of /n/ before dorsal consonants, so words such as \emph{mo}[ŋ.ɡr]\emph{el} are not exceptions.
 While there are no exceptions to the generalization about triconsonantal clusters, the statistical evidence does not rule out the possibility the absence of such clusters is due to chance.
@@
 
+\subsubsection{Local summary: static constraints}
+
+The inventory of attested clusters shows no statistically reliable traces of the three static constraints proposed by \citet{Pierrehumbert1994}.
+
 \subsection{Computational models}
 
-The three constraints derived from phonological alternations have been shown to have robust effects on the English lexicon.
-In constrast, the static constraints proposed by \citet{Pierrehumbert1994} do not have a statistically reliable effect on the cluster inventory.
+The absence of a statistical trace of these three static constraints does not indicate that there is no evidence for such static constraints.
+
+static constraint or system of constraints can be detected from can be detected from can be detected from can be detected from
+
+ static constraint
+
+The failure of these static constraints to correspond
+
 Obviously this does not indicate, however, that there are no static constraints on the cluster inventory.
@@
 
-\subsubsection{Evaluation of computational models}
+\subsubsection{Performance metrics}
 
@@
 One model was evaluated for this study, but performed no better than this baseline.
-\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
-However, \citeauthor{McGowan2011} finds that sonority of clusters accounts for only a tiny amount of the variance in cluster frequency ($R^2 = 0.077$).
+\citet{McGowan2009} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2009} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
+However, \citeauthor{McGowan2009} finds that sonority of clusters accounts for only a tiny amount of the variance in cluster frequency ($R^2 = 0.077$).
 As sonority distance provided no improvement over the null baseline, it is not reported above.
@@
 
-\subsection{Discussion}
+\section{Conclusions}
 
@@
 
-\subsubsection{The Zipfian distribution}
+\subsection{The Zipfian distribution}
 
-A number of previous studies \citep[e.g.,][]{Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phonemes and graphemes exhibit sparse type (i.e., lexical) and token frequency distributions.
+A number of previous studies \citep[e.g.,][]{Weiss1961,Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phonemes and graphemes exhibit sparse type (i.e., lexical) and token frequency distributions.
 The lexical frequencies of the medial codas ond onsets which make up syllable contact clusters show a similar distribution.
@@
 It is important to note that there is nothing inherently meaningful about the fact that clusters conform to a Zipfian distribution.
-These distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2009} to word \citep{Teahan1998,Ha2002,Baroni2009} and phoneme \citep{Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
+These distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2009} to word \citep{Teahan1998,Baroni2009} and phoneme \citep{Belevitch1956,Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Mandelbrot1954,Miller1957,Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
 The importance of this statistical property is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps due only to the sparsity of the data.
 
-\subsubsection{The Good-Turing estimate}
+\subsection{The Good-Turing estimate}
 
@@
 
-\subsubsection{Sampling simulation}
+\subsection{Sampling simulation}
+
+A final
 
 As a final consideration of this data, consider the statistical properties of a randomly generated ``lexicon'' as they compare to the observed distribution of clusters.
+
+The following procedure was developed to 
 A simulated lexicon can be generated by repeatedly applying the following procedure, which combines the derived constraint baseline with the model proposed by \citet{Pierrehumbert1994}.
@@
 
-\section{Conclusion}
-
 \citet{Borowsky1989} on peripherality.
 productive \citet{Duanmu2008}
-
 The next two chapters return to the question of synchrony, addressing the relationship between statistical patterns in the lexicon and speakers' behaviors when presented with underrepresented sequences.
-
 accidental gaps
@@
 %\begin{tabular}{l l l l}
-%a. & [gownp] & \textgreater{} & [gump]  \\
-%b. & [ǰumg]  & \textgreater{} & [ǰúŋgə] \\
-%c. & [ðʌŋd]  & \textgreater{} & [tɔŋg]  \\
+%a. & [ɡownp] & \textgreater{} & [ɡump]  \\
+%b. & [ǰumɡ]  & \textgreater{} & [ǰúŋɡuə] \\
+%c. & [ðʌŋd]  & \textgreater{} & [tɔŋɡ]  \\
 %\end{tabular}
@@
 %\citet{Hay2004a} demonstrates that [n.p, mθ] clusters, which are not found in English, are in fact rated better than many attested clusters.
-%The same is true of the prefix \emph{com-}/\emph{con-}, as shown by the presence of assimilation in \emph{có}[ŋ.g]\emph{ress} but not in \emph{co}[n.g]\emph{réssional}. 
+%The same is true of the prefix \emph{com-}/\emph{con-}, as shown by the presence of assimilation in \emph{có}[ŋ.ɡ]\emph{ress} but not in \emph{co}[n.g]\emph{réssional}. 
 %(e.g., \citealt{Bakovic2005b}, \citealt{Fruehwald2011}),
@@
 This sound change introduced no alternations.
-Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{V\lm sk\#} words when the change was actuated, and \emph{V\lm sh\#} continues to be rare in Modern English.
+Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{Vːsk\#} words when the change was actuated, and \emph{Vːsh\#} continues to be rare in Modern English.
 What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.
@@
 %\end{example}
-
 %Similar exceptions are found in 
@@
 %Hebrew \citep[][29]{Bat-El2005}.
+%And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.

### 2012-06-30 ack.tex
Feverish writing on chapters 2-4

@@
 So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy.
-More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, 
-who did some much to make this work, and me as a linguist,
-into what it is now.
-Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are to be thanked for creating an environment at Penn where a student could write this admittedly contenious dissertation. 
-And though his name appears nowhere else on this document, it is hard to imagine this work could exist without Bill Labov.
+More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, who has done so much for me as a linguist.
+Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are to be thanked for creating tan environment receptive to this admittedly contentious thesis.
+And though his name appears nowhere else in this document, it is hard to imagine this dissertation would exist without Bill Labov.
 
-My journey in linguistics begins with the mentoring of Jennifer Cole.
+While so many of my fellow students at Penn have a hand in this document, I would like to especially thank Toni Cook and Laurel MacKenzie for all their support over the years.
+I also must thank the administrative assistants at 3401 Walnut and 255 S.
+36th, Lauris Olson at the Van Pelt Library, Katherine McEldoon for a helpful review of numerical cognition research, and Carly Fisher for ten years of her unflagging friendship.
+
+I have many debts beyond Penn.
+My journey in linguistics began with the careful mentoring of Jennifer Cole and Richard Sproat.
 Steve Anderson graciously agreed to serve on my committee.
-Morris Halle taught me to speak clearly and never apologize.
-
-While so many of my fellow students at Penn have a hand in this document, 
-I would like to especially thank my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody.
+Morris Halle taught me to speak clearly, and to never apologize.
 
@@
 %Maryam Bakht kept me centered.
-
-My gratitude also goes out to the administrative assistants at 3401 Walnut and 255 S.
-36th, Lauris Olson at the Van Pelt Library, Katherine McEldoon for a helpful review of numerical cognition research, and Carly Fisher for her unflagging support and friendship.

### 2012-06-30 pre.tex
Feverish writing on chapters 2-4

--- 2011-11-14 pre.tex
@@
 
-This
+In this dissertation, I critique two such epicyclic proposals, and present solutions using independently-motivated grammatical machinery to account for the data.
+
+The first half of this dissertation began with a suggestion from Charles Yang that the sparse nature of the lexicon might itself account for at least some phonotactic gaps.
+The success of the resulting study, in chapter four, led me to question the received wisdom that speakers' knowledge of phonotactic generalizations is gradient and potentially independent of the phonological system.
+
+These results support a null hypothesis in which phonotactic knowledge is derivative of the language acquisition device and phonological processes.
+While I present this theory as a falsifiable scientific hypothesis, an alternative would treat this proposal as a \emph{negative heuristic}, a
+
+guides us away from ``phonotactic grammar'' as an explanatory device.
+
+even those who are not inclined to accept this proposal should note how much of the coverage traditionally attributed to

### 2012-07-10 1.tex
Feverish writing on chapters 2-4
Writing on chapters 2, 4

--- 2012-03-01 1.tex
@@
-\chapter{Morpheme structure constraints in generative grammar}
-\label{msc}
+\chapter{Sequence structure constraints in generative grammar} \label{msc}
 
@@
 
-\section{A brief history of MSCs} \input{1.1}
-\section{The null hypothesis} \input{1.2}
-\section{Outline of Part I} \input{1.3}
+\section{A brief history of MSCs}
+
+A more expressive approach to problems of this type, making no reference to \citeauthor{Bloomfield1930}'s study, was published only two years later.
+\citet{Jakobson1932}, then a member of the Prague circle, attributes a number of phonological properties of the Russian imperative to particular types of juncture, a precursor of morphological levels \citep[e.g.,][]{Siegel1974} in Lexical Phonology.
+Nearly all Russian infinitives are marked with a final theme vowel and /-t\pal{}/.
+In reflexives, this is followed by a suffix realized as [s\pal{}ə], which triggers a general process of regressive voicing assimilation.
+In the reflexive infinitive, the palatalization of infinitive /-t\pal/ is lost, but palatalization of a stem-final consonant is preserved in the reflexive imperative, even when the root-final consonant is /t\pal/, as in (\ref{rusreflex}c).\footnote{Thanks to Lev Blumenfeld for discussion of this data.}
+
+\begin{example}[Russian reflexives] \label{rusreflex}
+\begin{tabular}{l l l l}   & \emph{infinitive} & \emph{imperative} \\
+a. & [slav\pal{}itsə]  & [slaf\pal{}s\pal{}ə]  & `be glorious'   \\
+   & [upram\pal{}itsə] & [upram\pal{}s\pal{}ə] & `be stubborn'   \\
+b. & [kras\pal{}itsə]  & [kras\pal{}s\pal{}ə]  & `put on makeup' \\
+   & [ʒar\pal{}itsə]   & [ʒar\pal{}{}s\pal{}ə] & `roast'         \\
+c. & [zəbytsə]         & [zəbut\pal{}s\pal{}ə] & `forget'        \\
+\end{tabular}
+\end{example}
+
+\noindent \citeauthor{Jakobson1932} proposes that the preservation of palatalization in /\ldots{}t\pal{}-s\ldots{}/ is a special property of the imperative.
+Both (\ref{rusreflex}c) forms contain a /t\pal-s/ juncture, though palatalization is only preserved in the imperative.
+\citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization.
+%A few years later, \citet{Trnka1936} makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic gener
+
+nstraints.
+
+\subsection{Static phonotactics and the lexicon}
+
+If juncture may give rise to sound sequences not found in what are now called \emph{non-derived environments}, it follows that the morph represents a privileged domain for phonotactic generalizations.
+\citet{SPR,Halle1962} and \citet{Chomsky1965,SPE} posit a system of a system of \emph{morpheme structure constraints}---more precisely, \emph{sequence structure constraints} \citep{Stanley1967}---used to fill out redundant specifications in underlying forms, and to account for speakers' knowledge of possible words.
+
+This too can be traced to late structuralist thinkers, for example Kenneth Pike.
+\citet{Pike1947b} discusses a number of constraints on vowel sequences in Mixteco.
+It is apparent that \citeauthor{Pike1947b} does not expect these constraints to hold across junctures when comparing these constraints to his analysis of a Mixteco text published a few years earlier \citep{Pike1944}.\footnote{The junctures posited by \citet{Pike1944} are indicated with a dash, but this is not intended to imply that juncture is necessarily a phonetic event \citep[cf.][55f.]{Scheer2011}.}
+
+\begin{example}[Mixteco MSCs and complex words \citep{Pike1944,Pike1947b}]\begin{tabular}{l l@{} l l l}
+a. & * & {C}a{C}e & [ká-\textsuperscript{n}dee]     & `kept inside'   \\
+b. & * & {C}ə{C}e & [nì-k\`əbə]                     & `and entered'   \\
+c. & * & {C}e{C}i & [te-ní-ke\textsuperscript{n}da] & `was walking    \\
+d. & * & {C}i{C}e & [te-nì-kee]                     & `and went away' \\
+\end{tabular}
+\end{example}
+
+\noindent \citet[][166]{Pike1947b} writes that junctures are ``marked'' by the presence of sequences which cannot be found morph-internal, an idea further developed by \citet{Harris1955} as a discovery procedure for morphs (see also \citet{H
+
+Much later, \citet{McCarthy1988} and \citet{Mester1988} assume that these lexical constraints do not have to be exceptionless to be psychologically real; any statistically reliable pattern is potential evidence for the organization of feature systems, structural descriptions, and the like.
+Though these two studies have been widely imitated and in dozens of languages (a near-exhaustive list can be found in \citealp[?]{Brown2010} ), \citeauthor{McCarthy1988} and \citeauthor{Mester1988} never present a linking hypothesis for the connection between static lexical patterns and phonology.
+What follows is a sketch of such an argument.
+First, lexical patterns are delimited in the units of phonology \citep[9]{Berkley2000}: phonemes, syllables, and the like.
+The acquisition heuristic known as \emph{Stampean occultation} \citep[54]{OT} imposes phonological alternations root-internally, so that statistically reliable lexical patterns can plausibly be traced to the phonology of the recent past.
+For instance, virtually all obstruent-obstruent sequences in the English lexicon consist either of two voiced, or two unvoiced, consonants (see \S\ref{ova} below).
+This corresponds to the fact that English has---and has had for at least a millenium---a process of \textsc{Obstruent Voice Assimilation}.
+Unfortunately, few of these later studies distinguish between phonotactics derived by alternations and those which are have no analogue in the synchronic phonology and thus are ``static''.
+This chapter is a first attempt to fill this lacuna.
+
+\subsubsection{Biuniqueness and neutralization}
+
+joos
+harris
+chomsky
+
+\subsection{Early generativism}
+
+\subsubsection{\citealt{Stanley1967}}
+
+inventory constraints
+sequence constraints
+
+\subsubsection{Surface constraints and occultation}
+
+shibatani
+stampe
+dell
+
+\emph{A}[mt] `office', \emph{Anbau} `cultivation'
+\emph{ei}[n.g]\emph{reifen} `to intervene',
+
+\subsubsection{The duplication problem}
+
+\citeauthor{Anderson1974} sees the duplication of \textsc{Backness Harmony} as both sequence structure constraint and phonological rule as a foregone conclusion
+.The tendency of sequence structure constraints to duplicate phonological rules was noted by \citet[][401f.]{Stanley1967} and in \emph{SPE} (p.~382) in the discussion of morpheme structure constraints, and is labeled ``the duplication problem'' by \citet[][?]{Kenstowicz1977}.
+
+inventory problems 
+dell, chung, but also problems
+
+conspiracies
+
+\subsection{Autosegmentalism}
+
+also ``supersegmentalism''
+
+The proposals during this period are very diverse.
+
+\subsubsection{Syllable structure}
+
+hooper
+noske
+
+\subsubsection{Morph features}
+
+Leben
+Kaye
+
+\subsubsection{The Obligatory Contour Principle}
+
+\subsection{Classic Optimality Theory}
+
+%(page numbers here are taken from the published version)
+
+\subsubsection{Richness of the Base}
+
+some quotes from OT doc
+\citet{PE}
+\citet{Bye2001}
+
+\subsubsection{Freedom of Analysis}
+
+\citet{Smolensky1996}
+\citet{PE}
+
+\subsubsection{Undominated constraints}
+
+% complaints about maceahern, gallagher
+% bit on loanword adaptation
+
+\subsection{Recent developments}
+
+\subsubsection{Language acquisition}
+% jusczyk forever
+% white et al.
+% early word learning
+% syllables
+
+\subsubsection{Psycholinguistic tasks in adults}
+
+% wordlikeness
+% spotting
+% error correction, etc.
+
+\subsubsection{Non-native word adaptation}
+
+% perceptual factors
+
+\section{The null hypothesis}
+
+\subsection{The empirical proposal}
+
+\subsection{The acquisition principle}
+
+\subsection{Falsifiability}
+
+%\subsection{Universal constraints?}    \input{1.2.4}
+
+\section{Outline of the dissertation}
+
+%Chapter 2
+%Chapter 3
+%Chapter 4

### 2012-07-10 2.tex
Writing on chapters 2, 4

@@
 A naïve account of this contrast is based on the assumption that segments or timing units must either be parsed into prosodic structures like the syllable, or be subject to phonological repair.
-Early precedents for this can be found in dissertations by \citet[10f.]{Hooper1973} and \citet[57f.]{Kahn1976}, and it is a fundamental component of many later frameworks (e.g., \citealt{Ito1989a}, \citealt{Noske1992}, \citealt{OT}).
+Early precedents for this can be found in dissertations by \citet[10f.]{Hooper1973} and \citet[57f.]{Kahn1976}, and it is a fundamental component of many later frameworks \citep[e.g.,][]{Ito1989a,Noske1992,OT}.
+
 In English, [bl] is a permissible onset, but unlike some languages (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet'), [bn] is not.
@@
 This chapter focuses on the hypothesis, implicit in \citeauthor{Shademan2006}'s statement, that there is an inherent gradient or continuum between possible and impossible words, against an null hypothesis that the observed patterns of judgements are little more than an artifact of the wordlikeness task itself.
-\S\ref{gradhyp} places this debate in a historical perspective, and argues that the naïve but widely accepted form of this hypothesis fails the test of falsifiability.
+\S\ref{2background} places this debate in a historical perspective, and argues that the naïve but widely accepted form of this hypothesis fails the test of falsifiability.
 In \S\ref{2evaluation}, a quantitative evaluation of a revised version of this hypothesis reveals a poor match between the gradience in speakers' judgements and contemporary computational models of wordlikeness.
@@
 
-% What \citet[32]{Cohn2006} calls ``gradient well-formedness''
-
-\section{The gradience hypothesis} \label{gradhyp}
+\section{Background} \label{2background}
 
@@
 
-\subsubsection{\emph{The sound pattern of English}}
+\subsubsection{\emph{The sound pattern of English}} \label{2spe}
 
-In \emph{The sound pattern of English} (SPE), \citet{SPE} argue that a complete solution to wordlikeness must countenance gradience.
+In \emph{The sound pattern of English} (Henceforth \emph{SPE}), \citet{SPE} argue that a complete solution to wordlikeness must countenance gradience.
 In addition to the contrast of \emph{blick} and \emph{bnick}, they introduce a third nonce word \emph{bnzk}, which they take to be even less well-formed than ``impossible'' \emph{bnick}.
@@
 Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no attention in the literature.
-At the risk of explaining what might be no more than an an accidental gap in the historical record, the novel aspects of the \emph{SPE} model---gradience derived from similarity to existing lexical entries---seem to have been overshadowed by compelling arguments against \citeauthor{SPE}'s assumption (see also \citealt{Halle1962}, \citealt{Stanley1967}) that wordlikeness contrasts derive from properties of underlying forms.
+At the risk of explaining what might be no more than an an accidental gap in the historical record, the novel aspects of the \emph{SPE} model---gradience derived from similarity to existing lexical entries---seem to have been overshadowed by compelling arguments against \citeauthor{SPE}'s assumption \citep[see also][]{Halle1962,Stanley1967} that wordlikeness contrasts derive from properties of underlying forms.
 \citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
@@
 
-\noindent The plural appears to preserve a contrast in final obstruent voicing which is absent in the singular:\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing (e.g., \citealt{Fourakis1984}) or not \citep[e.g.,][]{Port1985}, it is irrelevant to the discussion at hand.} word-final voiced stops are never found in German, and \citet[95]{Shibatani1973} reports that ``it is easy to show that a native speaker of German rejects those forms \emph{on the ground that they end in voiced obstruents}'' (emphasis in original).
+\noindent The plural appears to preserve a contrast in final obstruent voicing which is absent in the singular:\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984}) or not \citep[e.g.,][]{Port1985}, it is irrelevant to the discussion at hand.} word-final voiced stops are never found in German, and \citet[95]{Shibatani1973} reports that ``it is easy to show that a native speaker of German rejects those forms \emph{on the ground that they end in voiced obstruents}'' (emphasis in original).
 However, given that root-final voicing is not predictable (e.g., /ɡʀaːt \alt{} ɡʀaːd/), the process of \textsc{Final Devoicing} cannot be any kind of lexical redundancy and is mysterious under the \emph{SPE} account.
@@
 The arguments of \citeauthor{Shibatani1973} and others led theorists to focus their attention on properties of surface representations as determinants of wordlikeness.
-Though syllabification plays no role in \emph{SPE}, the syllable is discussed in many earlier studies, and it received particular attention in the 1970s.
+Though syllabification plays no role in \emph{SPE}, it is crucial to many earlier studies \citep[see][]{Goldsmith2011b}, and it received particular attention in the 1970s.
 \citet{Hooper1973} and \citet{Kahn1976} argue that the syllable is useful for defining wordlikeness generalizations.\footnote{\citet{Steriade1999} and \citet{Blevins2003}, however, argue that a number of phonotactic generalizations previously stated in syllabic terms can be reanalyzed without making reference to syllables.} \citeauthor{Hooper1973} argues, for instance, that [bn], impossible as an English onset, is unobjectionable as a syllable contact cluster in nonce words like \emph{stabnik} (or in names like \emph{Abner}), and that this demonstrates the superiority of syllable-based wordlikeness generalizations.
@@
 \citeauthor{Hooper1973} also points to loanword adaptations which produce native syllable structure \citep[e.g.,][]{Carlisle1991} as evidence that syllabification is part of the phonological computation.
-
 Further enrichments to the theory are provided by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as envisioned by][]{Pike1947a}, and theories like prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
@@
 \citet{Jusczyk1994} report that 9-month-old children acquiring English also show preferences for nonce words with high positional probability over those with low positional probability.
-Faciliatory effects of positional probability (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements (\citealt{Vitevitch1999a}, \citealt{Luce2001}, \citealt{Lipinski2005}, \citealt{Vitevitch2005}), and lexical decision \citep{Pylkkanen2002a}.
+Faciliatory effects of positional probability (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements \citep{Vitevitch1999a,Luce2001,Lipinski2005,Vitevitch2005}, and lexical decision \citep{Pylkkanen2002a}.
 
-%\citet{Mattys1999,Mattys2001}
-%\citet{McQueen1998}
-
-The most interesting result of these studies comes from the first experiment of \citealt{Vitevitch1997}, which finds that speakers' wordlikeness ratings of bisyllabic words are correlated with the positional probabilities of the constituent syllables.
-Unfortunately, neither \citeauthor{Vitevitch1997} nor any of the other aforementioned psycholinguists make any effort to eliminate the possibility that the low positional probability stimuli are ``impossible'' words of English.
-In fact, in Chapter \ref{clusters}, it is argued that many of the stimuli used by \citet{Vitevitch1997}, including even the high probability stimuli, contain illicit word-medial consonant clusters.
-While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness to a post-hoc test, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and the wordlikeness judgements.
+Most relevant to the question at hand, \citet{Vitevitch1997} and \citet{Frisch2000} find that speakers' wordlikeness ratings of multisyllabic words are correlated wtih the positional probailities of the constituent syllables.
+Unfortunately, none of these researchers make any effort to eliminate the possibility that the low positional probability stimuli are ``impossible'' words of English.
+In fact,
+Chapter \ref{clusters} argues
+%the author has argued elsewhere \citep{Gorman2012c}
+that many of the stimuli used by \citet{Vitevitch1997} and \citet{Frisch2000} contain illicit word-medial consonant clusters.
+While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and the wordlikeness judgements.
 \citeauthor{Vitevitch1997} ultimately conclude that it cannot explain all the variation in wordlikeness.
@@
 More generally, they do little to distinguish between positional probability and closely correlated measures like bigram probability (see \S\ref{bigram} below) or neighborhood density, since these studies carefully select stimuli which either have high or low values for all of positional probability, bigram probablility, and neighborhood density.
-This is particularly troublesome given that positional probability was created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production (e.g., \citealt{Luce1998}, \citealt{Luce2000}) and perception (e.g., \citealt{Marslen-Wilson1984,Marslen-Wilson1987}, \citealt{McClelland1986}, \citealt{Norris1994}, \citealt{Norris2000}).
+This is particularly troublesome given that positional probability was created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production \citep[e.g.,][]{Luce1998,Luce2000} and perception \citep{Marslen-Wilson1984,Marslen-Wilson1987,McClelland1986,Norris1994,Norris2000}.
 
-\subsection{A naïve gradience axiom}
+\subsection{A naïve gradience hypothesis}
 
-Recent critiques of this syllabification-based model focus on the existence of intermediate wordlikeness ratings (see also \citealt{Coleman1997}, \citealt{Anttila2008}).
+Recent critiques of this syllabification-based model focus on the existence of intermediate wordlikeness ratings \citep[see also][]{Coleman1997,Anttila2008}.
 
@@
 
-The observation itself cannot be questioned, but \citeauthor{Hayes2008a} and \citeauthor{Albright2009a} do not explicitly state why this data is relevant to the construction of models of wordlikeness.
+As discussed in \S\ref{2spe} above, this is also the view of \citeauthor{SPE} in \emph{SPE}.
+The observation itself cannot be questioned, but \citeauthor{SPE}, \citeauthor{Hayes2008a}, and \citeauthor{Albright2009a} do not explicitly state why this data is relevant to the construction of models of wordlikeness.
 Presumably, these authors believe that these patterns of judgements demonstrate that wordlikeness, as an internal state, is gradient simply because subjects make use of intermediate degrees of wordlikeness in judgement tasks.
-This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, and in general, an influential view in the cognitive sciences (see \citealt{Fodor1981a} for a critique).
+This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, an influential view in the cognitive sciences in general (see \citealt{Fodor1981a} for a critique).
 
 \begin{unlabeledexample}
-\textsc{Naïve Gradience Proposition}: If graded judgements of a concept or mental state use intermediate ratings, that concept or mental state is gradient
+\textsc{Naïve Gradience Hypothesis}: If graded judgements of a concept or mental state use intermediate ratings, that concept or mental state is gradient
 \end{unlabeledexample}
 
-This proposition is an indicative conditional.
+This proposition has the form of an indicative conditional.
 While the antecedent of this condition, the presence or absence of intermediate ratings in a graded judgement task, is readily observable, it has no bearing on the truth or falsity of the conditional proposition.
@@
 In the case of wordlikeness, a case can be made that one such factor is similarity to existing words (e.g., \citealt[][151, fn.
-27]{LSLT}, \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Bailey2001}, \citealt{Schutze2005}).
+27]{LSLT}, \citealt{Greenberg1964}, \citealt{Vitz1973}, \citealt{Ohala1986b}, \citealt{Sendlmeier1987}, \citealt{Bailey2001}, \citealt{Schutze2005}).
 However, the mere fact that subjects use intermediate ratings does not show that they do so with any systematicity, or that, e.g., the contrast between \emph{447} and \emph{7} with respect could be given any satisfying explanation.
@@
 
-The remainder of this chapter is devoted to evaluating the gradience hypothesis, using state-of-the-art computational wordlikeness models and data from English.
+The remainder of this chapter is devoted to testing the revised hypothesis.
+This is accomplished by an evaluation of computational models of wordlikeness in which categorical and gradient models are placed on an equal footing.
 
@@
 The large-scale studies by \citet{Bailey2001} and \citet{Shademan2006,Shademan2007} are ineligible because the former lack stimuli with gross phonotactic violations and the latter data are not available to the public in any form.
-All data are reproduced in Appendix \ref{appendixA}
+All data are reproduced in Appendix \ref{appendixA}.
 
@@
 For each stimulus, the subjects produced forced choice ``yes''/``no'' answers to the question of whether the item ``is likely to be usable as a word of English''.
-\citet{Hayes2008a} and \citet{Albright2009a} analyze this data as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus (see also \citealt{Pierrehumbert1994}, \citealt{Coleman1997}, \citealt{Frisch2000}).
+\citet{Hayes2008a} and \citet{Albright2009a} analyze this data as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus \citep[see also][]{Pierrehumbert1994,Coleman1997,Frisch2000}.
 For instance, 22 of the 33 students answered ``yes'' for \emph{shlerk} [ʃlɚk], so it is assigned a score of $0.666$.\footnote{This procedure conflates intraspeaker variation, which in some cases is considerable \citep{Shademan2007} with speaker-internal gradience, and its adoption here should not be construed as an endorsement.
@@
 Non-parametric rank correlation statistics are used evaluate the correlation between model predictions and wordlikeness ratings.
-Rank correlation are used rather than the parametric Pearson correlation statistic that is used by \citet{Hayes2008a}, for instance, because they make none of the potentially troublesome assumptions of the latter method (see \citealt{Albright2009a}:23, fn.
-12).
+Rank correlation are used rather than the parametric Pearson correlation statistic that is used by \citet{Hayes2008a}, for instance, because they make none of the potentially troublesome assumptions of the latter method \citep[see][23, fn.
+12]{Albright2009a}.
+
 The Spearman $\rho$ is most widely known rank correlation statistic, but it is difficult to give a natural interpretation to this quantity.
@@
 \citealt{Greenberg1964}  & {0.845}          & {0.765} & {\textbf{0.863}} & {0.648}          \\
-\citealt{Scholes1966}    & {0.791}          & {0.762} & {\textbf{0.827}} & {0.827}          \\
+\citealt{Scholes1966}    & {0.791}          & {0.762} & {\textbf{0.827}} & {\textbf{0.827}} \\
 \citealt{Albright2003b}  & {0.725}          & {0.429} & {0.708}          & {\textbf{0.742}} \\
@@
 As all stimuli here are monosyllables, this task can be further simplified by separately considering the two major subcomponents of the syllable, the onset and rime.
-Speakers are particularly adept as separating onsets from rimes (\citealt{Treiman1986,Treiman1995}, \citealt{Fowler1993}), and a large portion of phonotactic violations can be localized to one or the other unit (e.g., \citealt{Fudge1969}, \citealt{Treiman2000}).
+Speakers are particularly adept as separating onsets from rimes \citep{Treiman1986,Treiman1995,Fowler1993}, and a large portion of phonotactic generalizations and violations can be localized to either onset or rime \citep[e.g.,][]{Fudge1969,Treiman1995,Kessler1997,Treiman2000}.
 
@@
 This problem is more acute in languages with more permissive syllable structures than those of English, and thus a sparser lexicon with regard to phonotactic constraints.
-For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.\footnote{Chapter \ref{clusters} reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
+For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.\footnote{
+%Chapter \ref{clusters} 
+\citet{Gorman2012c}
+reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
 
@@
 The gradient models assign rankings to the sets of phonotactically valid and invalid clusters, respectively.
-For instance, the bigram model favors \emph{troog} [tɹuːg] over \emph{swach} [swætʃ], though neither contains any gross phonotactic violation.
-Similarly, the bigram model favors \emph{choop} [tʃuːp] over \emph{zhrick} [ʒɹɪk], even though neither word is a possible word of English.
+For instance, the bigram model favors \emph{troog} [tɹuːɡ] over \emph{swach} [swætʃ], though neither contains any gross phonotactic violation.
+Similarly, the bigram model favors \emph{chwoop} [tʃwuːp] over \emph{zhrick} [ʒɹɪk], even though neither word is a possible word of English.
 However, the majority of such predicted contrasts are not reflected in speakers' judgements; for instance, \emph{troog} is rated less English-like than \emph{swach} \citep{Greenberg1964}, contrary to the model predictions.
@@
 
-This chapter
+\subsection{The gradience hypothesis}
 
-has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
-
+This chapter has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
 The surprising result is that virtually all of the apparent coverage of state-of-the-art gradient phonotactic models is simply a reflection of their ability to distinguish between the possible and the totally impossible; beyond this, they are unreliable.
@@
 It follows that the projections made by the state-of-the-art gradient models are not like those made by speakers.
+What remains to be seen is whether any model can be put forth which accurately predicts these intermediate ratings, and whether such a model deploys linguistic representations.
 
-%MORE SUMMARY
+These result provide support for recent findings that speakers asked to perform gradient syntactic judgements produce responses closely corresponding to a widely recognized categorical grammatical/ungrammatical distinction \citep{Sprouse2007}.
 
-\subsection{Possible extensions to the binary baseline}
+\subsection{Extensions to the binary baseline}
 
@@
 While, as \citet{Coleman1997} note, classic model predict that a nonce word is as ill-formed as its worst deviation from syllable structure, it is possible to imagine that multiple phonotactic violations would result in greater degrees of ill-formedness.
-The bigram and maxent models make this prediction, as do many others (e.g., \citealt{Legendre1990}, \citealt{Levelt2000}, \citealt{Albright2008}, \citealt{Anttila2008}, \citealt{Pater2009b}) but despite this, there is still no data demonstrating cumulative effects in wordlikeness tasks.
+The bigram and maxent models make this prediction, as do many others \citep[e.g.,][]{Legendre1990,Levelt2000,Albright2008,Anttila2008,Pater2009b} but despite this, there is still little data demonstrating cumulative effects in wordlikeness tasks.
 
-%\subsection{Future directions}
-%\subsubsection{Interspeaker variation}
-%The studies analyzed above all use ratings averaged across subjects.
-While aggregation of this type is quite common in psycholinguistics, it is undesirable as a statistical practice: it drastically increases the chance of a Type II error, the error of failing to reject a null hypothesis when the null hypothesis is false \citep[][8f.]{Baayen2004}.
-Further, little is known about intraspeaker variation in wordlikeness judgements and even less about intraspeaker variation, but \citet{Shademan2007} does find suggestive differences between older and younger adults.
-It would be extremely interesting to correlate differences in speakers with external factors like age, level of education, and exposure to foreign languages.
-%\subsubsection{Mixed modalities}
-%While the use of auditory stimulus presentation seems preferable to rating orthographic words, the possibiltiy of misperception introduces a potential confound.
-\citet{Scholes1966} mentions that a few of his stimuli were frequently misperceived, and a number of studies \citep{Massaro1983,Halle1998b,Pitt1998,Dupoux1999,Berent2007a,Davidson2007} uncover positive correlations between phonotactic acceptability and accurate perception.
-If the phonotactically invalid cluster in *\emph{srest} is perceived as a phonotactically valid cluster with similar acoustic features, such as [ʃr], it will likely receive an inflated wordlikeness rating.
-In many psycholinguistic tasks, the behavioral measure of interest comes with a validation of correct perception: for instance, in shadowing tasks, reaction times are the primary behavioral measure but shadowing errors are also informative behavioral measures \citep[e.g.,][]{Marslen-Wilson1978}.
-Wordlikeness studies could combine transcription or repetition tasks with judgement tasks to confirm correct perception and perhaps to enhance the online nature of the task.
-%Several recent studies have investigated the role of phonotactics using an auditory lexical decision tasks \citep[e.g.,][]{Berent2001a,Berent2004,Shatzman2007a}.
-A recent study by \citet{Coetzee2008b} is particularly relevant to the larger questions of this dissertation, though the results are ultimately inconclusive.
-\citeauthor{Coetzee2008b} claims that words of the form [spVp, skVk] (e.g., *\emph{spep} [spɛp], *\emph{skeek} [skiːk]) are impossible in English, whereas [stVt] words (e.g., *\emph{stoit} [stɔɪt]) are wellformed.
-In addition to a traditional wordlikeness task, \citet{Coetzee2008b} includes stimuli of this type in a lexical decision task, under the hypothesis that the illformed nonce words, those of the shape [spVp, skVk], will be more quickly rejected than valid [stVt] stimuli, a prediction which is borne out.
-However, there is no need to provide a phonotactic interpretation of these effects.
-It is known that auditory tasks reaction times are significantly slower for non-words with dense cohorts, that is, nonce words like *\emph{hup} [hʌp] which share initial segments with many real words (e.g., \emph{hull}, \emph{hut}, etc.).
-These effects are found across modalities \citep[e.g.,][]{Marslen-Wilson1978,Marslen-Wilson1984} and appear to be particularly attentuated for non-word onsets \citep{Cole1980,Vitevitch2002}.
-Since [st] onsets are far more frequent than [sp] or [sk] onsets \citep[][395]{Hayes2008a}, this alone could explain the variation in reaction times.
-Despite this, nonce word lexical decision with controls for lexical cohort effects shows potential as a substitute for wordlikeness judgements in phonotactics research.
-%\subsubsection{Data availability}
-%There are still very few publicly available wordlikeness databases, and several authors who were contacted for this project declined to share data.
-Not only does this make large-scale model comparison difficult, it also means that most wordlikeness studies cannot be replicated except at the conceptual level.
-Lexical decision researchers now have databases of response times for American \citep{ELP} and British English \citep{BLP}, Dutch \citep{DLP}, French \citep{FLP}, and Malay \citep{MLP}.
-Given the simplicity of wordlikeness judgements (a sort of linguistic \emph{Drosophilia}), the time for a publicly available ``English Wordlikeness Project'' has come.
-%\subsubsection{Non-parametric statistics}
-%The surprising results of this evaluation can be traced directly to two methodological improvements which should be more widely adopted.
-First, model comparison studies must replace parametric statistical models, such as Pearson correlation, with non-parametric equivalents.
-Given the tendency of ratings to cluster at endpoints, Pearson correlations are inflated \citep[][23, fn.
-12]{Albright2009a} and the assumptions of the test are flagrantly violated.
-A shift towards non-parametric comparison appears to be well underway: \citet{Hayes2008a} also report model correlations using the Spearman $\rho$ and \citet{Albright2009a} focuses on a variant of the Kendall correlation called $tau_c$.
-%On the other hand, both \citeauthor{Hayes2008a} and \citeauthor{Albright2009a} ultimately use parametric statistics for model comparison.
-%\begin{table}
-%\centering
-%\begin{tabular}
-%\toprule
-%                        & accuracy & precision & recall & F$_1$ & MCC   \\
-%\midrule
-%\citealt{Greenberg1964} & 0.846    & 1.000     & 0.800  & 0.889 & 0.693 \\
-%\citealt{Scholes1966}   & 0.952    & 0.947     & 0.900  & 0.923 & 0.889 \\
-%\citealt{Albright2003b} & 0.920    & 0.938     & 0.953  & 0.946 & 0.797 \\
-%\bottomrule
-%\end{tabular}
-%\end{table}
-% EM \citet{EM} 
-%\begin{table}
-%\centering
-%\begin{tabular}{l | r r}
-%\toprule
-%                        & $\chi^2_{3}$ & $p$-value  \\
-%\midrule
-%\citealt{Greenberg1964} &  5.190       & 0.158      \\
-%\citealt{Scholes1966}   & 48.964       & 1.3\e{-15} \\
-%\citealt{Albright2003b} & 16.186       & 0.001      \\
-%\bottomrule
-%\end{tabular}
-%\caption{This is my caption}
-%\end{table}
-%\footnote{This is conceptually similar to the technique of residualization discussed by \citet{Gorman2010c}.}
-%\begin{unlabeledexample}
-%$\displaystyle P = s^{1/7.4}$
-%\end{unlabeledexample}
-%\noindent Even then, \citet[][23, fn. 12]{Albright2009a} 
-%An interesting feature of this data set is that includes a diverse set of onsets ranging from frequent to impossible, but only a few common rimes; \citet{Albright2009a} speculates this might have made the onsets more salient to the subjects than tehy would be otherwise.
-%The Pearson correlation depends on an interval assumption, which holds that the difference between, for example, ``1.
-completely bi disagree'' and ``2.
-agree'' is in some sense identical to that between ``3.
-neither agree nor disagree'' and ``4.
-agree''.
-\citet{Stevens1946} already observes that there is no a priori basis for this assumption, and similar objections can be leveled for other rating methods.
-Secondly,
-%the Pearson correlation assumes a linear relationship between model scores and ratings; when this assumption is violated, researchers must search for a satisfactory transformation to produce linearity.
-%\end{quote}
-%The similarity between sequence structure constraints and neutralizing phonological rules was noted by \citet[][401f.]{Stanley1967}.
-Later discussions tend to view it as a problem (\citealt{Kisseberth1970b}, \citealt{Kenstowicz1977}).
-The relationship between surface and underlying constraints is taken up in the next two chapters.
-%\citealt{Barton1986}, \citealt{Barton1987}, \citealt{Anderson1988a}, \citealt{Ristad1994}
-%\citet{Vitevitch1999b} % summary paper
-%\citealt{Kahlen-Halstenbach1990}, \citealt{Inozuka1991})
+\subsection{Language acquisition}
+
+The findings here also have interesting rammifications for language acquisition.
+\citet{Saffran1996} famously argues that infants use statistical dependencies between segments to extract words from continuous speech.
+Among those who ascribe to the Continuity Hypothesis \citep[e.g.,][]{Macnamara1982,Pinker1984,Crain1991,Carey1995,deVilliers2001,Legate2007}, which holds that children acquiring language deploy the same types of linguistic representations as adults, evidence for gradient phonotactic knowledge in infants is evidence for this knowledge in adults, and vis versa.
+The results here then stand to weaken the hypothesis of \citeauthor{Saffran1996}, as well as the hypothesis that infants acquire a system of phonotactics (gradient or otherwise) before learning of alternations and lexical entries that is a direct prediction of algorithms for learning Optimality Theory grammars \citep{Tesar2004,Hayes2004b}.
+
+%\footnote{The hypothesis that infants acquire a system of phonotactics (gradient or otherwise) before learning of alternations and lexical entries has some traction among formal linguists, for \citet{Hayes2004b} and \citet{Tesar2004} note that is a direct prediction of algorithms for learning Optimality Theory grammars.
+However, there is extensive evidence that infants who are younger than 9 months, the age at which \citet{Jusczyk1994} argue that the system probabilistic phonotactics comes online, have been demonstrated have already begun extracting words \citep[e.g.,][]{Darwin1877,Mandel1995,Jusczyk1995,Tincoff1999,Jusczyk1999b,Brent2001,Bortfeld2005,Seidl2006,Shukla2011,Bergelson2012} and alternations \citep{White2008} from multiword utterances they encounter.} Each of the tenets of this argument has been subject to serious critique, however.
+
+First, it appears that children treat syllables holistically as early as 4 days after birth \citep{Bijeljac-Babic1993} and rapidly learn native-language syllable-based phonotactics \citep{Onishi2002,Chambers2003}, suggesting that any co-occurrence statistics should be delimited on units larger than the segments proposed by \citeauthor{Saffran1996}.
+Secondly, a good deal of the literature has focused on segmentation according to near exceptionless constraints that hold between adjacent syllables.
+For instance, infants as young as 4.5 months seem to be aware that English nasal codas agree in place with following obstruents \citep{Mattys2001b,Jusczyk2002,Davidson2004}.
+It might be the case that syllable co-occurrence statistics might be little more than a reflection of infants' learning of categorical ``lexical viability'' constraints \citep{Johnson2003} of the sort also seen in adult speech processing \citep[e.g.,][]{Norris1997}.
+Further, the effect of these statistical dependencies appear to be filtered by expectations about word stress \citep{Johnson2001,Mattys2001a,Shukla2007}, word length \citep{Lew-Williams2012}, and allophonic generalizations \citep{Hohne1994,Jusczyk1999c}.
+Finally, and perhaps most importantly, \citet{Gambell2005} find that a word segmentation strategy based only on statistical dependencies performs very poorly on realistic data.
+The emerging view of the role on statistics seems to be that infants attend to transitional probabilities or an analogue thereof only when they are faced with artificial stimuli with few of the properties of natural language \citep[e.g.,][]{Shukla2007,Lew-Williams2011}.
+The argument from Continuity is weak, in either direction.

### 2012-07-10 4.tex
Writing on chapters 2, 4

--- 2012-06-03 4.tex
@@
-%\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
-\label{clusters}
+\chapter{Structural and accidental gaps} \label{clusters}
 
-The term \emph{lexicon} has many senses: throughout, I use it to refer simply to the set of underlying representations.
-%\citep[][269]{LANGUAGE}.
-
-In the preceding chapter, it was proposed that
-
-Reconsider the null hypothesis presented in the preceding chapter:
-
-\begin{example}[null hypothesis]
-Universal grammar does not countenance constraints on the contents of URs
-\end{example}
-
-The idea
-
-relationship to proposal
-
-One primary way which this null hypothesis has been critiqued is by demonstrating that the lexicon of some language 
-exhibits dispreferences or gaps corresponding to.
-
-argue by trends in the lexicon
-
-possibility of sparsity (C\&H, Fischer/Vogt and recent stats)
-
-Of course, to obtain the desired results, we must guarantee that each sequence structure rule reflects a general systematic fact about the language, and not a fact which is due merely to the existence of accidental gaps in the lexicon.
-\citep[][401, fn.~8]{Stanley1967}
-
-In the preceding chapter, it was proposed that
-
-Lexicon
-
-This appears to be exceptionlessly true of Arabic
-
-That even gradient patterns
-
-One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. 
-Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
-This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a} but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Dmitrieva2008a,Dmitrieva2008b,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.
-
-The absence of 
-The claim here is that however stark the absence or underrepresentation of some form is, it
-
-This hypothesis admits the possibility that there may be accidental gaps in the lexicon, a possibility which predates generative thinking:
+Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1979}, \citealp[410f.]{Stanley1967}).
+Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the \emph{communis opinio}
+(including many of the authors cited above) holds that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
 
 \begin{quote}
-\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
-\citep[][16]{Fischer-Jorgensen1952}
+It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
+\citep[283]{Anderson1974}
 \end{quote}
 
-\noindent
-Hans \citeauthor{Vogt1954} makes a similar observation in his study of Georgian clusters
+English word-medial consonant clusters have been used, by \citet{Pierrehumbert1994} and others, to argue for the necessity of an independent phonotactics module and illustrate proposals for its architecture.
+This chapter argues, however, that the only structural constraints on the concatenation of codas and onsets to form word-medial clusters are derivative of well-documented phonological alternations in English.
+
+\section{Background}
+
+English admits a variety of word-medial consonant clusters spanning syllable boundaries, known as \emph{syllable contact clusters} or \emph{interludes}.
+These clusters can be as short as two consonants (e.g., \emph{a}[n.t]\emph{ics}; singleton intervocalic consonants are ignored here) or as long as four (e.g., \emph{mi}[n.str]\emph{el}).
+\citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in any given language \citep[cf.][]{Haugen1956}.
 
 \begin{quote}
-Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters.
-The material will never be complete.
-It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary.
-\citep[][30]{Vogt1954}
+\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
 \end{quote}
 
-\citet{Chomsky1965}
+This chapter is concerned with the nature of these ``provisos'' as reflected by which of the ``possible'' clusters, so defined, are attested.
+While many phonotactic generalizations can be thought of as inventory restrictions imposed by prosodic constituents, additional filters on the set of possible clusters are essentially combinatoric in nature and thus distinct, there is no prosodic constituent consisting of the coda of one consonant-final syllable and the onset of a following consonant-initial syllable \citep[though see][]{Steriade1999}.
 
-This chaper demonstrates that apparently systematic gaps may arise even in the absence of phonotactic preferences, and that this possibility compromises attempts to show that such preferences shape the lexicon.
+\section{Evaluation}
 
-\section{Domain}
+\citeauthor{Pierrehumbert1994} argues that the inventory of clusters in English shows the effect of a small number of static phonotactic constraints.
+The remainder of this chapter provides an exhaustive quantitative evaluation of this hypothesis.
+The most important constraints on medial clusters are those derived from phonological rules which apply across syllable boundaries; there is no remaining role for static constraints.
+The sparse nature of the lexicon plays a secondary role, giving rise to apparent phonotactic gaps even in the absence of static phonotactic constraints.
 
-% WHY I'm DOING THIS STUDY
-% describes makeup of the corpus and how p
+\subsection{Corpus construction}
 
-\subsection{Data sources}
+A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database.
+The resulting list of clusters is reproduced in Appendix \ref{appendixB} and in electronic form at the author's website.
 
-The English CELEX database \citep{CELEX},
+\subsubsection{Simplex words}
 
-\citet{Aronoff1976}
+An analysis of lexical constraints requires a corpus of lexical representations.
+So as to remain agnostic about particulars of morphological theory, the goal here is to identify words which bear no overt inflection, and which do not admit any plausible further decomposition, henceforth ``simplex words''.
+For the structuralists, a procedure for parsing words into morphs, just the thing for locating simplex words, held the status of ``philosopher's stone'', a sort of near-mythical objective of the endeavor, but actual attempts to develop such a procedure \citep[e.g.,][]{Harris1955,Nida1948} are now recognized as anything but theory-neutral.
+Yet, there is no alternative to adopting some computational procedure or human-coded database as the ``gold standard'' for a study of this type.
+\citet{Pierrehumbert1994} analyzes a list of entries in the Collins English dictionary by hand, using her intuitions to filter out ``complex'' words as well as those which are not ``reasonably familiar''.
+Unfortunately, the resulting list was neither published nor circulated, and the possibility of replicating \citeauthor{Pierrehumbert1994}'s' sensations of morphological complexity is remote.
+Following prior studies of syllable contact clusters by \citet[ chapter 3]{Hammond1999a} and \citet[ chapter 8]{Duanmu2009}, a list of simplex words was derived from the ``lemmas'' (that is, uninflected forms) in the English portion of CELEX \citep{CELEX}, a database of morphological, phonological, and syntactic annotations based on the COBUILD corpus, a resource that was also used to construct the Collins English dictionary employed by \citeauthor{Pierrehumbert1994}.
 
-\citet{Harley2009}
+Two further filters were applied to this list.
+First, any lemma coded as an unassimilated loanword is excluded.
+Secondly, all lemmas that have a ``morphological status'' other than ``monomorpheme'' are excluded.
+This is crucial because CELEX makes no guarantee that lemmas are simplex; indeed, many lemmas are products of derivation (e.g., \emph{abusive}, from \emph{abuse}).
+These stringent criteria have a profound effect on the makeup of the data.
+For instance, nearly all exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) cited by \citet[74]{Hammond1999a} are excluded as unassimilated loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as morphologically complex (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}).
 
-\citet{Taft1975} ?
-\citet{Taft1981} ?
-\citet{Taft2004a} ?
+``Neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}, have an ambiguous and long-debated morphological status.
+\citeauthor{Pierrehumbert1994} attempts to distinguish between words like \emph{complete}, \emph{extreme}, \emph{inspect}, \emph{obtuse}, which she assumes to be simplex, and \emph{excrete} and \emph{transparent}, which she considers complex.
+However, CELEX codes these words as complex and they are thus excluded.
+This is consistent with converging evidence that speakers decompose these Latinate forms.
+The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.
+In a similar vein, \citet[11--13]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.
 
-\citet{Halle1962}
+%\footnote{Similar behaviors are found in Russian verbs which share roots but which take different prefixes or the reflexive \citep{Pesetsky1977}.}
 
-% 2.1.4: k
-
-Under the parses and phonologization described immediately below, the 6,876 simplex CELEX entries contain 21 medial coda types and 40 unique medial onsets.
-Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, a saturation rate of 18.8\%.
-
-\subsection{Syllabification}
-
-For this study, it is necessary to separate medial consonant clusters into codas and onsets.
-While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic.
-For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language.
-Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which explain the \emph{chemistry}/\emph{ministry} syllabification contrast.
-
-An automated syllabification system was constructed and applied to medial clusters in the CELEX data.
-This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.
-There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification.
-The technique used is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948},
-\citealt[75]{Pulgram1970}, \citealt[42f.]{Kahn1976}, \citealt[][358f.]{Selkirk1982b}) which parses word-medial clusters so that the onset is as large as possible.
-Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty.
-In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.
-
-There is one case where unchecked maximization of medial onsets produces incorrect syllabifications.
-When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
-\citet[][55]{Harris1994} notes that onset maximization makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty.
-The approach adopted here is essentially that of \citet{Pulgram1970}: the first consonant of a medial cluster is assigned to the coda of a preceding coda if is immediately preceded by a stressed lax vowel, which do not occur syllable-finally in English.\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}
-
-If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables.
-For this reason, an additional constraint is assumed which prevents the two components of the affricates from being split by syllabification.
-This is motivated by the tendency of affricates to pattern with single segments in many languages.
-For instance, the only complex onsets in Classical Nahua are the affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.
-
-\subsection{Inventory consdierations and subsyllabic parsing}
-
-To identify the inventory of syllable contact clusters, it is also necessary to parse syllables into onset, nucleus, and coda.
-In most cases this is trivial, but in a few cases, it is necessary to use distributional heuristics to determine the underlying inventory and subsyllabic affiliations.
-
-\subsubsection{The velar nasal} \label{velarnasal}
-
-There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right \citep[e.g.,][]{Sapir1925}, an analysis which is a consequence of principles like the Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ (e.g., \emph{SPE}:85, \citealt{Borowsky1986}:65f.).
-The strongest piece of evidence for the latter position, in which the velar nasal is a pure allophone, is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone.
-\citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here.
-English nasal place allophony is formalized in Section \ref{cnpasection} below.
-
-\subsubsection{[j] onglides}
-
-The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J.
-\citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}).
-The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).
-
-There is considerable evidence for this assumption.
-When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː].
-In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}.
-This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}).
-\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear.
-There is also strong external support for this principle.
-The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants.
-The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor} [h]\emph{omponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.
-
-\subsubsection{[w] onglides}
-
-The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset.
-Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
-Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}.
-Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.
-And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.
-
-\subsubsection{Post-vocalic \emph{r}}
-
-Both CELEX and the dictionary used by \citet{Pierrehumbert1994} in her study transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost.
-In \emph{r}-full dialects, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters.
-Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}); for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}.
-Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which post-vocalic \emph{r} patterns differently than internal coda consonants.
-\citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.
-
-\begin{example}[/t/-\textsc{Glottalization} in British English (after \citealp{Harris1994}:195, 258)]
-\begin{tabular}{l l l@{} l}
-a. & par[t]   &   & par[ʔ]   \\
-   & car[t]on &   & car[ʔ]on \\
-b. & fis[t]   & * & fis[ʔ]   \\
-   & mis[t]er & * & mis[ʔ]er \\
+\begin{example}[Bound stem-specific allomorphy]
+\begin{tabular}{l l l l l l}
+a. & {adhere}   & {adhesion}   \\
+   & {cohere}   & {cohesion}   \\
+b. & {conceive} & {conception} \\
+   & {perceive} & {perception} \\
 \end{tabular}
@@
 
-\noindent Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.
+There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement.
+\citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.
+\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
 
-\begin{example}[/t, d/-\textsc{Deletion} in American English] \label{td}
-\begin{tabular}{l l l@{} l}
-a. & be[lt]  &   & be[l] \\
-   & me[nd]  &   & me[n] \\
-b. & sh[ɚt]  & * & sh[ɚ] \\
-   & c[ɚd]   & * & c[ɚ]  \\
+\begin{example}[Latinate verbs and small clauses] \label{harley}
+\begin{tabular}{ll@{}ll@{}l}
+a. & * & {exhibit him the painting} & ~ & {show him the painting} \\
+b. & * & {imbibe himself stupid}    & ~ & {drink himself stupid}  \\
+c. & * & {exhibit it off}           & ~ & {show it off}           \\
 \end{tabular}
@@
 
-\section{Evaluation}
+Decomposition is also supported by lexical decision results.
+Among nonce words, those that can be exhaustively decomposed into prefix and bound stem, like *\emph{de}-\emph{juvenate}, are processed slower that contain a licit morph but cannot be exhaustively decomposed, like *\emph{de-pertoire}, and these in turn are processed slower than those which admit no such decomposition \citep{Taft1975}.\footnote{The same contrasts are present in many other domains where decomposition into root and affix(es) is less controversial, for instance, inflected verbs in Italian \citep{Caramazza1988}.} \citet{Emmorey1989} reports facilitative auditory priming between pairs of words which share a bound stem, like \emph{permit}-\emph{submit} \citep[though see][]{Marslen-Wilson1994}.\footnote{It may be the case that full decomposition is only one of the two methods for the processing of complex words, the other being whole word looking, either running in serial \citep{Caramazza1988} or in parallel \citep{Baayen1997b}.
+Whatever evidence that is amassed in support of a whole word lookup procedure, however, does not undermine the considerable evidence for decomposition.}
 
-\subsection{Static and derived constraints}
+\subsubsection{Syllabification and phonologization}
 
-With this database of English syllable contact clusters it is now possible to evaluate constraints on sequences of underlying phonemes.
+CELEX includes broad, syllabified Received Pronunciation transcriptions of individual wordforms.
+However, a casual inspection of the data reveals the unsystematic nature of these syllabifications.
+Any two words sharing a nucleus-medial cluster-nucleus sequence should be given the same parse, since syllabification is universally non-contrastive.
+Yet many putative syllabification contrasts can be found in CELEX; for instance, compare the parses given to the clusters in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][II.294]{AOE}.}
 
-Here, the metric used is simply
+A systematic syllabification requires an automated procedure for delimiting medial clusters and parsing them into coda and onset.
+Since this procedure is applied only to medial clusters, this study remains agnostic on several contentious issues concerning syllabification of English; for instance, there is no need to address the status of so-called ``ambisyllabic'' consonants, since this does not occur with medial clusters, or morphological effects on syllabification, since no complex words are included.
+The technique used here is a variation on the theme of onset maximization (e.g., \citealp[42f.]{Kahn1976}, \citealp{Kurylowicz1948}, \citealp[75]{Pulgram1970}, \citealp[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset.
+Initial onsets are used to define what is a ``possible'' medial onset \citep[though cf.][36]{Fischer-Jorgensen1952}.
+Medial clusters in words like \emph{neu}[.tɹ]\emph{on} or \emph{bi}[.stɹ]\emph{o} are found in word-initial position (e.g., [tɹ]\emph{ansit}, [stɹ]\emph{ike}), so onset maximization assigns the entire medial cluster to the onset, leaving the medial coda empty.
+In contrast, the cluster in \emph{mi}[n.stɹ]\emph{el} is not found word-initially; the maximal onset here is [stɹ], and the residual [n] is assigned to the coda.
 
-The evaluation
+\paragraph{Stressed lax vowels} When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
+\citet[][55]{Harris1994} notes that onset maximization produces an incorrect result when, in addition, the medial cluster is a valid onset: in \emph{whisper}, \emph{vestige}, and \emph{musket}, onset maximization incorrectly assigns [sp, st, sk] to the medial onset, leaving the lax vowel unchecked.
+Consequently, the first consonant of a medial cluster is assigned to the coda of a preceding syllable before a stressed lax vowel \citep[cf.][48]{Pulgram1970}, a minimal modification of the onset maximization parse.
 
-The foremost evidence for
+\paragraph{Affricates} If the affricates [tʃ, dʒ] are treated as stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990}, the previous principle will incorrectly assign the two segments of the affricate to separate syllables before stressed lax vowels (e.g., *\emph{ra}[t.ʃ]\emph{et} or \emph{a}[d.ʒ]\emph{ile}) unless some ad hoc constraint \citep[e.g.,][]{Wells1990} is in place.
+A more traditional analysis of affricates, in which they are simple segments distinguished from pure stops by stridency \citep[24]{Jakobson1961} or delayed release (\emph{SPE}:321f.) is assumed here, following \citeauthor{Pierrehumbert1994}.
+This assumption can be motivated by the tendency of affricates to pattern with individual segments in phonotactic generalizations.
+For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but bans all other onset clusters \citep[9]{Launey2011}.
+%\citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.
 
-The evaluation used here is simply
+\paragraph{The velar nasal} According to some \citep[e.g.,][]{Sapir1925}, [ŋ] in English is a phoneme in its own right, but others (e.g., \emph{SPE}:85, \citealp{Borowsky1986}:65f.) hold that it is simply the allophone of /n/ before underlying /k, ɡ/ (with later deletion of /ɡ/ in some contexts).
+The total absence of onset [ŋ], where it cannot be followed by a dorsal consonant which conditions the velar allophone, is strong evidence that [ŋ] is a pure allophone of /n/, the analysis assumed here.
 
-Having described the database of English syllable contact clusters,
+\paragraph{[j] onglides} While some have argued that the front onglide in words such as \emph{val}[j]\emph{ue} is inserted by rule (\emph{SPE}:196, \citealp[][89]{Halle1985a}, \citealp[][217]{McMahon1990}), the very presence or absence of the glide is contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}) indicating that it is present in underlying representation (\citealp{Anderson1988b}, \citealp[278]{Borowsky1986}).
+The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).
+There is considerable evidence for this assumption.
+When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u], suggesting that the glide is nuclear (\citealp{Davis1995}, \citealp[][61f.]{Harris1994}, \citealp[][232]{Hayes1980}).
+\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters, though they may be followed by [ju] in words like [mj]\emph{use} or [vj]\emph{iew}, also suggesting the nuclear affiliation of the glide.
+There is also external support for this analysis: the [ju] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005}, \emph{shm}-reduplication \citep{Nevins2003}, and speech errors (e.g., [kju]\emph{mor} [h]\emph{omponent}, intended [hju]\emph{mor component}; \citealp[130]{Shattuck-Hufnagel1986}).
 
-I now consider constraints on sequences of URs as predictors of attested and unattested contact clusters.
-First, I describe the statistical technique used, then apply this technique to both phonologically derived constraints and a set of ``static'' phonotactic constraints proposed by \citet{Pierrehumbert1994}.
+\paragraph{[w] onglides} The phonotactic properties of the back onglide [w] are the opposite of the front onglide.
+Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
+Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[161]{Davis1995}.
+Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[166]{Davis1995}.
+Taken together, these facts indicate that the back onglide is assigned to the onset.
 
-Before the dawn of generative phonology, many linguists concerned themselves with documenting co-occurrence in various languages.
-\citet[][28]{Vogt1954} declares that a ``phonemic description of a language \ldots should also comprise a description of the phoneme combinations that occur or may be expected to occur in the language''.
-Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively.
-\citeauthor{Uhlenbeck1950}, for instance, puts forth co-occurrence dispreferences, rather than exceptionless generalizations of the sort given by \citet{Greenberg1950}.
+\paragraph{Post-vocalic \emph{r}} Word-medial post-vocalic \emph{r} has been lost in Received Pronunciation, the standard accent of the CELEX transcriptions.
+Even in \emph{r}-ful dialects, though, there is reason to believe that post-vocalic \emph{r} is fact nuclear, and thus irrelevant to syllable contact clusters.
+Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}): compare American English \emph{fern}, \emph{fir}, \emph{fur} to \emph{pet}, \emph{pit}, \emph{putt}.
+Post-vocalic \emph{r} patterns with vowels and not other consonants for several phonological processes: it is the only consonant which does not block variable glottalization of following /t/ in British dialects \citep[258]{Harris1994}.
 
-\citet{McCarthy1988}
-
-Years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and proposes a statistical technique, based on the chi-square test, to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon.
-
-\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English).
-The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented.
-Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.
-
-\begin{example}[null hypothesis testing]
-$\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} >
-χ^2$
-\end{example}
-
-%\footnote{This is the one-tailed version of the test.}
-
-In this study, I use the \citet{Fisher1934} exact test, which is isomorphic to this chi-square test.
-Compared to the chi-square test, the Fisher exact test $p$-value is somewhat more difficult to compute, but both are automated by modern statistical packages and can be computed rapidly.
-The major advantage of the Fisher test over the chi-square test is that it is appropriate for small samples or rare events, whereas the chi-square test depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples.
-
-On the other end of the spectrum, the Monte Carlo statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010} does not scale to large samples.
-The Monte Carlo procedure is isomorphic to the one-tailed Fisher Exact test, and involves comparing observed co-occurrence counts to those produced by randomly generated samples.
-This requires the researcher to generate random permutations, something which is particularly difficult for large samples, as I will explain; disinterested readers may wish to skip the following paragraph.
-
-Any permutation of a sequence $L$ can be represented as a sequence $S$ of the same length, where the value of $S_i$ corresponds to the permuted position of $L_i$.
-Generating random permutations requires an unbiased method to generate all $N!$ of these lists with equal probability.
-Unfortunately, this is near impossible for $N$s of moderate size with current computational resources.
-Any pseudo-random number generator can be characterized by its ``period'', the number of pseduo-random numbers it can emit before repeating itself; this is also the upper limit for the number of random permutations that can be generated.
-The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a period of approximately $2^{48}$, a number which turns out to be smaller than $17!$.
-For the large samples, sometimes in the thousands, used by these researchers, virtually all the possible permutations can never be generated by this method.
-As a consequence, the shuffling procedure is far from random and introducing the possibility that the statistical test is biased in unpredictable ways.
-
-As an example of the Fisher exact test as it is applied here, consider some of the Javanese co-occurrence facts considered by \citet{Mester1988}.
-\citeauthor{Mester1988} is at pains to show that there is a gradient dispreference for first and second root consonants which share the same major place.
-However, perfectly identical first and second consonants are exempt.
-Since
-
-% APPLY THIS TO JAVANESE
-
-\begin{example}[Javanese root co-occurrence] %(\citealp[][264]{Uhlenbeck1950}, \citealp[][139]{Mester1988})]
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming & 134      & 12         & 0.918      & \multirow{2}{*}{6.095\e{-06}} \\
-violating  & 21       & 15         & 0.583 \\
-\bottomrule
+\begin{example}[/t/-\textsc{Glottalization} in British English]
+\begin{tabular}{l l l l@{} l l l}
+a. & {des}[ɚt]    & \alt{} &   & {des}[ɚʔ]    \\
+   & {c}[ɚt]{ain} & \alt{} &   & {c}[ɚʔ]{ain} \\
+b. & {fi}[st]     & \alt{} & * & {fi}[sʔ]     \\
+   & {mi}[st]{er} & \alt{} & * & {mi}[sʔ]{er} \\
 \end{tabular}
@@
 
-\noindent
-It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing.
+\noindent Post-vocalic \emph{r} is the only consonant which does not trigger variable deletion of a following /t, d/ in American dialects \citep[8]{Guy1980}.
 
-This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance
+\begin{example}[/t, d/-\textsc{Deletion} in American English]
+\begin{tabular}{l l l l@{} l l l}
+a. & {be}[lt] & \alt{} &   & {be}[l] \\
+   & {me}[nd] & \alt{} &   & {me}[n] \\
+b. & {sh}[ɚt] & \alt{} & * & {sh}[ɚ] \\
+   & {c}[ɚd]  & \alt{} & * & {c}[ɚ]  \\
+\end{tabular}
+\end{example}
 
-That this is highly unlikely to be due to chance is apparent both from the chi-square ($\chi^2 = 25.572$, two-tailed $p = 4.3$\e{-07}) and Fisher exaact tests ($p =  6.095$\e{-06}).
+\subsubsection{Lexical statistics}
 
-I use two-tailed statistical tests throughout. 
-Under a one-tailed test, the alternative hypothesis is underrepresentation, whereas a two-tailed test also allows the possibility that a class of clusters would be overrepresented, i.e., more frequent than predicted by the null hypothesis.
-As
+CELEX contains 6,876 simplex words of English, comprising 21 unique medial codas and 40 unique medial onsets.
+Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested.
 
-\citet{Pierrehumbert1994}
+\subsection{Evaluating derived constraints}
 
-It must be stressed that under current assumptions, this is independent of whether predictable prosodic representations (particularly coda and onset identity) are part of the memorized underlying representation \citep[e.g.,][]{Vaux2003}.
-It is an undeniable fact that URs need to be 
-The derivative nature of 
-\citet{Ito1989a,Noske1992}
+\emph{SPE} describes three phonological alternations which target syllable contact clusters.
+The corresponding constraints on simplex syllable contact clusters are considered below.
+The first such derived constraint, obstruent voice assimilation, is presented as a tutorial introduction.
+The second set of constraints consists of three static phonotactic constraints proposed by \citet{Pierrehumbert1994}.
 
-least some researchers (i.e., \citealt{Mester1988})
+\subsubsection{Obstruent voice assimilation} \label{ova}
 
- \citep[e.g.,]{Brown2010} have argued for phonotactic constraints giving rise to overrepresentation, so I admit this possibility here and use two-tailed tests throughout.
-
-%\footnote{For example, \citet{Martin2007} applies this technique to a list of 4,758 noun-noun compounds.
-The number of permutations of a list of this length can be estimated using Stirling's approximation.
-
-%\ex $ \lim_{n \rightarrow \infty} n! = e ^ {n \ln{n} - n} = 2 ^ {\log_2{e} (n \ln{n} - n)} $ \xe
-
-%\noindent
-%$4,758!$ is approximately $2 ^ {51260}$. The author is unaware of any system that provides 51,260 bits of entropy; and some popular programming languages, such as ANSI C, provide as few as 15 bits.}
-
-%%% START OF SERIOUSNESS
-
-\subsection{Derived constraints}
-
-\emph{SPE} describes three phonological alternations which target syllable contact clusters. The corresponding constraints on simplex syllable contact clusters are considered below.
-
-\subsubsection{Obstruent voice assimilation}
-
-Voice assimilation in English is exemplified by the allomorphs of /-d/ suffix, forming regular pasts and denominal adjectives, and the /-z/ suffix, forming regular noun plurals, genitives, and the 3sg.
-present active indicative.\footnote{The URs /-d, -z/ are argued for by \citet[][282]{Hockett1958}, \citet[][210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, S.
-\citet[][]{Anderson1973a}, \citet[][102]{Pinker1988}, and \citet[][284f.]{Bakovic2005b}; alternative URs are argued for by \citet[][210f.]{LANGUAGE}, \citet[][426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet[]{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[][135]{Borowsky1986}.}
+The [t \alt{} d] allomorphs of the regular past (e.g.
+\emph{nap}[t] $\sim$ \emph{nab}[d]), and the [s \alt{} z] allomorphs of the regular noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), Saxon genitive, and third person singular active indicative verb agreement, exemplify \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178).\footnote{Underlying /-d, -z/ are argued for by \citet[282]{Hockett1958}, \citet[210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, \citet{Anderson1973a}, \citet[102]{Pinker1988}, and \citet[284f.]{Bakovic2005b}; alternatives are suggested by \citet[210f.]{LANGUAGE}, \citet[426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[135]{Borowsky1986}.}
 
@@
 \begin{tabular}{l l l l l l}
-a. & nap & nap[t] \\
-   & nab & nab[d] \\
-b. & lap & lap[s] \\
-   & lab & lab[z] \\
+a. & {nap} & {nap}[t] & & {nab} & {nab}[d] \\
+b. & {lap} & {lap}[s] & & {lab} & {lab}[z] \\
 \end{tabular}
@@
 
-\noindent The devoicing of the two voiced obstruent suffixes can be formalized as a rule spreading the \textsc{Voice} specification of an obstruent rightward.
+\noindent The devoicing of the two voiced obstruent suffixes can be formalized as a 
+rule spreading the voicing specification of an obstruent rightward.
 
-\begin{example}[\textsc{Obstruent Voice Assimilation} (after \emph{SPE}:178)]
+\begin{example}[\textsc{Obstruent Voice Assimilation}]
 \xymatrix@R=24pt@C=24pt{
@@
 
-\citet{Pierrehumbert1994} does not mention voice assimilation in her study of English syllable contact clusters.
-\citet[][74f.]{Hammond1999a} notes the existence of words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain obstruent voice clusters disagreeing in [\textsc{Voice}].
-This might be taken as evidence that \textsc{Obstruent Voice Assimilation} does not apply to simplex words, and indeed, the \textsc{Revised Alternation Condition} (RAC) proposed by \citeauthor{Kiparsky1973a} (\citeyear{Kiparsky1973a}:163, \citeyear{Kiparsky1982a}:152) blocks the application of obligatory neutralization processes like \textsc{Obstruent Voice Assimilation} in root-internal (i.e., non-derived) environments.\footnote{The ``obligatory'' condition on the RAC has the result that \textsc{Obstruent Voice Assimilation} is always consistent with the RAC whether or not it applies in non-derived environments.
+It is quite apparent that root-internal hetero-voiced obstruent clusters like [s.b] or [z.t] are also rare compared to clusters which are uniformly voiced or uniformly voiceless, as predicted the assimilation rule.
+Yet, \citet{Pierrehumbert1994} does not mention voice assimilation in her study of English syllable contact clusters.
+\citet[][74f.]{Hammond1999a} cites words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain hetero-voiced obstruent clusters, as evidence that the process does not apply root-internally, though few of his examples are in fact simplex according to CELEX.\footnote{The \textsc{Revised Alternation Condition} (RAC) proposed by \citeauthor{Kiparsky1973a} (\citeyear{Kiparsky1973a}:163, \citeyear{Kiparsky1982a}:152) blocks the application of obligatory neutralization processes like \textsc{Obstruent Voice Assimilation} in root-internal (i.e., non-derived) environments.
+Simply because there are exceptions in non-derived environments, \textsc{Obstruent Voice Assimilation} is always consistent with the RAC whether or not it actually applies in non-derived environments.
+This is due to the ``obligatory'' condition of the RAC.
 If the rule applies in non-derived environments, then it has lexical exceptions (e.g., \emph{a}[s.b]\emph{estos}) and is not obligatory and thus free from the RAC.
-On the other hand, if the rule is subject to the RAC, it only applies in derived environments, in which it is obligatory.} Despite this, root-internal clusters of obstruents which disagree in voicing (e.g., [s.b, b.s]) are much rarer than those which contain adjacent voiced or voiceless obstruents (e.g., [s.p, b.z]), suggesting that \textsc{Obstruent Voice Assimilation} applies in non-derived environments, preventing learners from positing URs in which adjacent obstruents do not agree in voice.
+On the other hand, if the rule is subject to the RAC, it only applies in derived environments, in which it is obligatory.}
 
-\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}]
+The question is: does \textsc{Obstruent Voice Assimilation} contribute a meaningful characterization the English lexicon?
+More specifically, are hetero-voiced clusters underrepresented in a way that is unlikely under the hypothesis of free combination?
+To answer this question, the 720 clusters in which both the final coda consonant and initial onset consonant are obstruent are sorted into four bins, according to whether they are attested or not, and whether or not the conform to the derived constraint---both obstruents are either voiced or voiceless, like in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}, respectively---or violate it, like the aforementioned \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}.
+
+This table, shown in (\ref{ovatable}), is then submitted to the Fisher Exact Test. This statistical test is so named because it is isomorphic to the chi-square test 
+for contingency tables, but computes an exact $p$-value, whereas the chi-square test $p$-value depends on an approximation which is inappropriate for small samples.
+The small $p$-value produced by the Fisher test indicates that the rarity of the disagreeing clusters is unlikely to be due to chance.
+The simplest interpretation is that \textsc{Obstruent Voice Assimilation} applies in non-derived environments, preventing learners from positing underlying representations in which adjacent obstruents do not agree in voice.
+
+\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}] \label{ovatable}
 \begin{tabular}{l r r r r}
 \toprule
-           & attested & unattested & saturation & $p$-value                   \\
+           & attested & unattested & \% saturation & $p$-value                   \\
 \midrule
-conforming & 80       & 370        & 0.178      & \multirow{2}{*}{1.1\e{-11}} \\
-violating  &  6       & 264        & 0.022                                    \\
+conforming & 80       & 370        & 17.8          & \multirow{2}{*}{1.1\e{-11}} \\
+violating  &  6       & 264        & 2.2           \\
 \bottomrule
@@
 
-\noindent 18\% of the possible clusters containing adjacent obstruents with the same [\textsc{Voice}] specification occur in the corpus, whereas only 2\%  of those which disagree in [\textsc{Voice}].
-The small $p$-value indicates that the rarity of the disagreeing clusters is unlikely to be due to chance.
+There are two possible analyses of the small number of roots that violate the derived generalization.
+Some hetero-voiced obstruent clusters, like the one in \emph{ja}[k.d]\emph{aw}, for instance, might be analyzed as compounds by native speakers, and this analysis might itself be driven by the rarity of morph-internal adjacent hetero-voiced obstruents (cf.
+\citealp[546]{Rice2009d} on compounds in Slave).
+In fact, \citet{Mattys2001b} find that 9-month-old infants are sensitive to the contrast between [f.t] and exceptional (and unattested) [v.t], treating the latter as an indication of a word boundary and this heuristic is also available to adults \citep{Brown1956,McQueen1998}.
+Less explanatorily, exceptional roots can be marked [$-$\textsc{Obstruent Voice Assimilation}].
 
-The failure of a small number of roots to conform to this generalization can be derived by marking such roots [$-$\textsc{Obstruent Voice Assimilation}].
-Another possibility is that disagreeing obstruent clusters might be analyzed as semantically opaque compounds by native speakers (e.g., \emph{ja}[k\#d]\emph{aw}).
-There are precedents for using morphological boundaries solely to preserve phonological generalizations.
-In \emph{SPE} (p.~94), for instance, morphological boundaries are used to simplify stress assignment, and \citet[][546]{Rice2009d} analyses many words in the Athapaskan language Slave as compounds simply because they contain consonant clusters that rarely occur otherwise.
-Recent experiments find that 9-month-old infants are sensitive to the contrast between [vt], a disagreeing cluster which does not occur word-medially, and [ft], found in medial position \citep{Mattys2001b}, and this ability seems to persist into adulthood \citep{McQueen1998}.
+\subsubsection{Nasal place assimilation} \label{npa}
 
-\subsubsection{Nasal place assimilation} \label{npasection}
+In addition to [n $\sim$ ŋ] allophony, \textsc{Nasal Place Assimilation} (\emph{SPE}:85) is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant.
 
-\citet[][175]{Pierrehumbert1994} observes that ``nasal-stop sequences agree in labiality'' in simplex English words; \citeauthor{Pierrehumbert1994} presumably makes no statement about other place features as she also assumes the velar nasal is derived from underlying /n/.
-While \citeauthor{Pierrehumbert1994} makes no connection between this sequence structure constraint and the any phonological alternation, the nasal place assimilation that derives [ŋ] is also responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix.
-In (\ref{npa}ab), the prefix-final consonant takes on the same major place of articulation as the following obstruent.
-The vowel-initial bases in (\ref{npa}c) demonstrate that the prefix is underlyingly /ɪn-/.
+\begin{example}[\emph{im-}/\emph{in-} allomorphy] \label{nparule}
+\begin{tabular}{l l l l l l l}
+a. & {polite}   & {i}[m.p]{olite}   & & {balance} & {i}[m.b]{alance} \\
+b. & {tangible} & {i}[n.t]{angible} & & {decent}  & {i}[n.d]{ecent}  \\
+\end{tabular}
+\end{example}
 
-\begin{example}[\emph{im-}/\emph{in-} allomorphy] \label{npa}
-\begin{tabular}{l l l l}
-a. & possible  & i[m.p]ossible \\
-   & balance   & i[m.b]alance  \\
-b. & tangible  & i[n.t]angible \\
-   & decent    & i[n.d]ecent   \\
-c. & elegant   & i[n]elegant   \\
-   & ability   & i[n]ability   \\
-d. & mature    & i[m]ature     \\
-   & numerable & i[n]umerable  \\
-\end{tabular} 
+\begin{example}[\textsc{Nasal Place Assimilation}]
+\xymatrix@R=24pt@C=24pt{
+                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
+\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
+\txt{[$+$\textsc{Nasal}]} & \txt{[$-$\textsc{Sonorant}]}              \\
+}
 \end{example}
@@
 \citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but that assimilation of \emph{im-}/\emph{in-} is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
-However, the stress condition does not hold in simplex words (e.g., \emph{a}[ŋ.g]\emph{óra}), and thus can be safely ignored here.
+However, the stress condition does not hold for simplex words (e.g., \emph{a}[ŋ.ɡ]\emph{óra}), and is ignored here.
 
-\begin{example}[\textsc{Nasal Place Assimilation} (after \emph{SPE}:85)]
-\xymatrix@R=24pt@C=24pt{
-                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
-\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
-\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]}              \\
-}
-\end{example}
+There is evidence that even young infants acquiring English are aware of this generalization.
+\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} find that infants at 4.5 months, 9 moths, and 10 months of age, respectively, prefer to listen to nonce words which contain the homorganic nasal-obstruent clusters (e.g., \emph{u}[m.b]\emph{o}) over those which contain heterorganic clusters (e.g., \emph{u}[n.b]\emph{o}).
+Nasal-obstruent clusters arising in speech errors also undergo \textsc{Nasal Place Assimilation} (e.g.
+\emph{ra}[nd] \emph{orker}, intended \emph{ra}[ŋk] \emph{order}; \citealt[228]{Myers1993}).
 
-Exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} do occur, they are much rarer than homorganic nasal-obstruent clusters such as \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey}, a preference which is unlikely to be due to chance.\footnote{There are also a small class of simplex words which contain clusters like \emph{oi}[nt.m]\emph{ent}, in which a nasal-obstruent cluster is found within a complex coda, but place assimilation is without exception in this context, and thus these clusters are ignored here.}
+\citet[175]{Pierrehumbert1994} observes that ``nasal-stop sequences agree in labiality'', but makes no reference to \textsc{Nasal Place Assimilation}.
+Exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} do occur, but they are much rarer than homorganic nasal-obstruent clusters such as \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey}.
 
@@
 \toprule
-           & attested & unattested & saturation & $p$-value                   \\
+           & attested & unattested & \% saturation & $p$-value                   \\
 \midrule
-conforming & 41       & 11         & 0.788      & \multirow{2}{*}{2.6\e{-05}} \\
-violating  & 8        & 20         & 0.286                                    \\
+conforming & 41       & 11         & 78.8          & \multirow{2}{*}{2.6\e{-05}} \\
+violating  & 8        & 20         & 28.6                                    \\
 \bottomrule
@@
 
-There is some evidence that even young infants acquiring English are aware of this generalization.
-\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} find that 4.5-, 9-, and 10-month old infants (respectively) prefer to listen to nonce words which contain the homorganic nasal-obstruent clusters (e.g., \emph{u}[m.b]\emph{o}) over those which contain heterorganic clusters (e.g., \emph{u}[n.b]\emph{o}).
-Further, adult speech errors may derive obstruent clusters which then undergo \textsc{Nasal Place Assimilation}.
+\subsubsection{Degemination} \label{deg}
 
-\begin{example}[Speech errors feeding \textsc{Nasal Place Assimilation} (\citealp{Myers1993}:228)]
-\begin{tabular}{l l l}
-a. & primps        & (prints)      \\
-b. & rand orker    & (rank order)  \\
-c. & po[ŋ]cutation & (computation) \\
-\end{tabular} 
-\end{example}
-
-\subsubsection{Degemination}
-
-A final alternation targeting English syllable contact clusters consists of the simplification of geminates derived by certain affixes.
+A final alternation targeting English syllable contact clusters is the simplification of geminates derived by certain affixes.
 For instance, the deadjectival adverbial suffix \emph{-ly} rarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{norma}[l]\emph{y}, cf.
 \emph{calm}[l]\emph{y}).
-Degemination is also found in /d/-final roots selecting the irregular /-t/ past tense suffix.
-
-\begin{example}[Degemination in /-t/ pasts (\citealp{Myers1987}:492)]
-\begin{tabular}{l l l l l}
-a. & burn    & burn[t] \\
-   & spill   & spil[t] \\
-b. & ben[d]  & ben[t]  \\
-   & buil[d] & buil[t] \\
-\end{tabular}
-\end{example}
-
-More complicated cases of degemination are also found in English prefix morphology (\emph{SPE}:148, \citealt{Borowsky1986}:102f.); one can be found in (\ref{npa}d) above, in which \textsc{Nasal Place Assimilation}.
-To formalize this process of degemination, it is assumed that geminates are adjacent similar segments.
-The alternative, that geminates are single feature matrices linked to multiple timing slots, places this constraint in the inventory and makes it difficult to link degemination and the absence of geminates in underlying representations.
-Geminates are identified using quantification over features \citep{Reiss2003b}.
-
-\begin{example}[\textsc{Degemination} (\emph{SPE}:253)]
-\xymatrix@R=24pt@C=24pt{
-\txt{C}_1       & \rightarrow & \emptyset & / & \gap\gap & \txt{C}_2 \\
-}
-
-\noindent Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
-\end{example}
-
-Neither identical consonants, nor those which differ only in voice (e.g., *[p.b]) occurs in simplex words in CELEX, and this absence is highly unlikely to be due to chance.\footnote{It is not the case that the English lexicon envisioned in \emph{SPE} is free of underlying geminates; \citeauthor{SPE} use underlying (but never surfacing) geminates to produce exceptions to the stress rules (see also \citealt{Burzio1994}:52f., \citealt{Halle1998c}:551, fn.
-9).
-In addition to the usual objections to absolute neutralization (i.e., that it is a learning conundrum mascarading as a phonological solution), the phonological particulars of this account are critiqued by \citet{Ross1972a}.}
+Degemination is also found in /d/-final roots selecting the irregular /-t/ past tense suffixes (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).
+More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}).
+Identical adjacent consonants do not occur in the corpus, nor do those which differ only in voice, such as *[p.b], which would be made identical by the operation of \textsc{Obstruent Voice Assimilation}.
+This absence is unlikely to be due to chance.
 
@@
 \toprule
-           & attested & unattested & saturation & $p$-value                   \\
+           & attested & unattested & \% saturation & $p$-value                   \\
 \midrule
-conforming & 157      & 596        & 0.208      & \multirow{2}{*}{7.2\e{-09}} \\
-violating  & 0        &  87        & 0.000                                    \\
+conforming & 157      & 596        & 20.8      & \multirow{2}{*}{7.2\e{-09}} \\
+violating  & 0        &  87        & 0.00                                    \\
 \bottomrule
@@
 
-\subsection{Static constraints}
+\subsubsection{Local summary}
 
-\citet{Pierrehumbert1994} surveys the English syllable contact cluster inventory, and discerns the effect of several static constraints.
-These constraints are termed ``static'' because there is no evidence for any phonological process, either allophonic or neutralizing, which targets the clusters the constraints disfavor.
+The three \emph{SPE} rules affecting syllable consonant clusters are reliably reflected in the lexicon.
+Possible clusters that are surface-exceptions to these three rules are much more likely to be unattested than those that conform to them.
 
-\subsubsection{Dorsal/non-coronal clusters}
+\subsection{Evaluating static constraints}
 
-\citet[][173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Since \citeauthor{Pierrehumbert1994}'s study is restricted to triconsonantal clusters, so words like \emph{a}[k.m]\emph{e}, \emph{ru}[g.b]\emph{y}, or \emph{pi}[g.m]\emph{ent} do not violate this narrower generalization.
-Further, \citeauthor{Pierrehumbert1994} adopts the traditional assumption \citep[e.g.,][66f.]{Borowsky1986} that coda [ŋ] is an allophone of /n/ before dorsal consonants, so words such as \emph{mo}[ŋ.gr]\emph{el} are not exceptions.
-While there are no exceptions to the generalization about triconsonantal clusters, the statistical evidence does not rule out the possibility the absence of such clusters is due to chance.
-Clusters of any length also fail to support a broader version of this generalization.
-As dorsal/non-coronal clusters are not significantly less likely to occur than similar types of ``conforming'' clusters such as those with a non-dorsal coda and a non-coronal onset (e.g., \emph{a}[t.m]\emph{osphere}) or a dorsal coda and coronal onset (e.g., \emph{ve}[k.t]\emph{or}), there is no evidence for any dispreference for dorsal/non-coronal clusters.
+The second set of constraints, proposed by \citet{Pierrehumbert1994}, are ``static'' in that they have no analogue in English morphophonemics.
+Clusters of all length are considered here, not only the triconsonantal clusters considered by \citeauthor{Pierrehumbert1994}, as this restriction appears to be arbitrary.
 
-\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}, triconsonantal clusters only]
+\subsubsection{Dorsal/non-coronal consonants}
+
+\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Two-consonant clusters like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} are found, however.
+Despite this, dorsal/non-coronal clusters are no less likely to occur than non-dorsal codas with non-coronal onsets (e.g.
+\emph{oi}[nt.m]\emph{ent}) or dorsal codas with coronal onsets (e.g., \emph{ve}[k.t]\emph{or}).
+
+\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}]
 \begin{tabular}{l r r r r}
 \toprule
-           & attested & unattested & saturation & $p$-value              \\
+           & attested & unattested & \% saturation & $p$-value \\
 \midrule
-conforming & 20       & 148        & 0.119      & \multirow{2}{*}{0.136} \\
-violating  &  0       &  22        & 0.000                               \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}, all size clusters]
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value \\
-\midrule
-conforming & 76 & 359 & 0.175 & \multirow{2}{*}{0.145} \\
-violating  &  6 &  57 & 0.095 \\
+conforming & 76 & 359 & 17.5 & \multirow{2}{*}{0.145} \\
+violating  &  6 &  57 & 9.5 \\
 \bottomrule
@@
 
-\citet[][175]{Pierrehumbert1994} proposes that ``clusters with a coronal obstruent in the coda do not occur'', though \citet{Pierrehumbert1994} notes an exception in \emph{a}[nt.l]\emph{er}, and one can add the triconsonantal clusters in \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many biconsonantal clusters like \emph{a}[t.l\emph{as} or \emph{e}[s.t]\emph{er}.
-Triconsonantal coda coronal obstruent clusters are not significantly less likely to occur than ``conforming'' clusters with non-coronal or non-obstruents.
-The same is true if the arbitrary restriction to triconsonantal clusters is lifted.
+\citet[175]{Pierrehumbert1994} proposes that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many clusters such as \emph{a}[t.l]\emph{as} or \emph{e}[s.t]\emph{er}.
+Once again, coda coronal obstruent clusters are no less likely to occur than non-coronal obstruent clusters like \emph{re}[p.t]\emph{ile} or coda coronal sonorant clusters such as \emph{co}[n.d]\emph{or}.
 
-\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}, triconsonantal clusters only]
+\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}]
 \begin{tabular}{l r r r r}
 \toprule
-           & attested & unattested & saturation & $p$-value              \\
+           & attested & unattested & \% saturation & $p$-value              \\
 \midrule
-conforming & 20       & 148        & 0.119      & \multirow{2}{*}{0.136} \\
-violating  & 0        & 22         & 0.000                               \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}, all size clusters]
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & saturation & $p$-value              \\
-\midrule
-conforming & 48       & 312        & 0.133      & \multirow{2}{*}{0.301} \\
-violating  & 38       & 322        & 0.106                               \\
+conforming & 48       & 312        & 13.3      & \multirow{2}{*}{0.301} \\
+violating  & 38       & 322        & 10.6                               \\
 \bottomrule
@@
 Finally, \citet[][176]{Pierrehumbert1994} observes ``a lack of clusters with identical first and third elements''.
-This is formalized by treating as ``identical'' any consonants which share all the same place and manner features and ignoring voicing, as was done for the definition of \textsc{Degemination} above.
-\citeauthor{Pierrehumbert1994} notes one potential exception to this generalization, the cluster in words such as \emph{e}[ks.kl]\emph{ude}, which contains two /k/, does not arise here since such words are marked as morphologically complex in CELEX.
-Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonaontal and quadraconsonantal clusters.
+Here, this is operationalized in the same fashion as \textsc{Degemination}, by definining ``identity'' as consonants which share the same place and manner features, but which do not necessarily agree on voicing.
+Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters.
 
-\begin{example}[Lexical effects of \textsc{*aba}]
+\begin{example}[Lexical effects of \textsc{*ABA}]
 \begin{tabular}{l r r r r}
 \toprule
-           & attested & unattested & saturation & $p$-value \\
+           & attested & unattested & \% saturation & $p$-value \\
 \midrule
-conforming & 41       & 478        & 0.079      & \multirow{2}{*}{0.818} \\
-violating  &  0       &  11        & 0.000                               \\
+conforming & 41       & 478        & 7.9      & \multirow{2}{*}{0.818} \\
+violating  &  0       &  11        & 0.0                               \\
 \bottomrule
@@
 
-\subsection{Computational models}
+\subsubsection{Local summary}
 
-The three constraints derived from phonological alternations have been shown to have robust effects on the English lexicon.
-In constrast, the static constraints proposed by \citet{Pierrehumbert1994} do not have a statistically reliable effect on the cluster inventory.
-Obviously this does not indicate, however, that there are no static constraints on the cluster inventory.
-By applying computational models of phonotactics to the syllable contact data, however, it is possible to exhaustively search for any such static constraints.
+There is no statistically reliable evidence for the three static constraints proposed by \citeauthor{Pierrehumbert1994}.
+However, it is premature to conclude that no static constraint exists.
+It is possible that an explicit model of ``constraint discovery'', such as those proposed by \citet{Pierrehumbert1994} and \citet{Hayes2008a}, could identify reliable static constraints, and predict which of the possible clusters are attested and unattested beyond the derived constraints.
 
-\subsubsection{Evaluation of computational models}
+\subsection{Evaluating computational models}
 
-The task of the computational models is the same as above: they are used to predict which of the possible clusters are attested and which are not.
-Models are scored using four metrics.
-Accuracy is the probability of that a possible cluster is either a true positive ($tp$), correctly predicted to occur, or a true negative ($tn$), correctly predicted to be unattested.
-Incorrect outcomes are either false positives ($fp$), incorrectly predicting a cluster to occur, and false negatives ($fn$), incorrectly predicting a cluster to be unattested.
+The computational models of \citeauthor{Pierrehumbert1994} and \citeauthor{Hayes2008a}, as well as two baselines, are scored for their ability to predict which of the possible medial clusters are attested and which are not.
+For any single cluster, there are four possible outcomes.
+
+\begin{example}[Outcomes in cluster classification task]
+\begin{tabular}{c l | l l}
+                   & & \multicolumn{2}{c}{\textbf{actual outcome}}            \\
+                   & & \emph{attested}   & \emph{unattested}            \\
+\midrule
+\textbf{predicted} & \emph{attested}   & true positive ($tp$)  & false positive ($fp$) \\
+\textbf{outcome}   & \emph{unattested} & false negative ($fn$) & true negative ($tn$)  \\
+\end{tabular}
+\end{example}
+
+\noindent Accuracy is the probability of a correct classification, whether a true positive or negative.
 
@@
 
-Precision is the probability that a cluster predicted to occur is attested.
+\noindent Precision is the probability that a cluster predicted to occur is attested.
 
@@
 
-Recall (or sensitivity) is the probability a observed cluster is predicted to occur.
+\noindent Recall (or sensitivity) is the probability an attested cluster is predicted to occur.
 
@@
 
-It is possible to maximize recall at the expense of precision, by predicting more clusters to be attested, or to increase precision at the expense of recall by predicting fewer clusters to be attested.
-$F_1$, the harmonic mean of these two measures, quantifies this tradeoff.
+\noindent It is possible to maximize recall at the expense of precision, by predicting attestation for a greater number of clusters, or to increase precision at the expense of recall by predicting non-attestation for a greater number of clusters.
+$F_1$, the harmonic mean of these two measures, quantifies this trade-off.
 
@@
 
-The results for the four models are summarized in Table \ref{cmresults}.
+A soft-margin support vector machine \citep{Cortes1995} with a linear kernel is used to convert from the per-cluster scores to attestation predictions, by finding a single cutoff which optimally separates attested and unattested clusters according to their model scores.
+This is in no way intended to be a cognitively plausible model for learning phonotactics; it simply represents an upper bound on the performance of these models.
+The results for all four models are summarized in Table \ref{cmresults}.
 
-\begin{table} \label{cmresults}
-\centering
+\begin{table} \centering
 \begin{tabular}{l | r r r r}
@@
 Derived constraints       & 0.857    & 0.401     & 0.708  & 0.512 \\
-\citet{Pierrehumbert1994} & 0.839    & 0.210     & 0.750  & 0.328 \\
+Expected frequency        & 0.839    & 0.210     & 0.750  & 0.328 \\
 \citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
@@
 \caption{The \citeauthor{Hayes2008a} model makes the most accurate predictions about possible and impossible syllable contact clusters, though it is only slightly better than the derived constraint baseline.}
+\label{cmresults}
 \end{table}
@@
 18.8\% of the 840 possible clusters are attested.
-As a consequence of this sparsity, a null baseline which classifies all clusters as unattested will achieve 81.2\% accuracy.
-Precision, recall, and $F_1$ are undefined for this baseline.
-
-One model was evaluated for this study, but performed no better than this baseline.
-\citet{McGowan2011} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2011} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
-However, \citeauthor{McGowan2011} finds that sonority of clusters accounts for only a tiny amount of the variance in cluster frequency ($R^2 = 0.077$).
-As sonority distance provided no improvement over the null baseline, it is not reported above.
+Consequently, a null baseline which predicts no clusters to be attested trivially achieves 81.2\% accuracy.
+The other metrics are undefined for this baseline.
 
@@
 
-\textsc{Obstruent Voice Assimiation}, \textsc{Coda Nasal Place Assimilation}, and \textsc{Degemination} are reliable predictors of what clusters are and are not attested and admit few exceptions.
-A slightly more complex baseline can be produced by predicting any cluster which does not violate these generalizations to be attested.
-This results in a significant improvement to the null baseline (accuracy = 0.857).
-As might be expected given the sparse nature of the syllable contact inventory, there are more unattested clusters which do not violate any of these alternation-derived constraints than exceptions to these generalizations, reflected in the fact that recall (0.708) is somewhat higher than precision (0.401).
+The three derived constraints define a structured baseline, in which attestation is the predicted outcome for all clusters which not violate the \emph{SPE}-derived generalizations, and non-attestation otherwise.
+This produces a significant improvement to the null baseline (accuracy = 0.857; sign test $p = 2.16$\e{-05}).
+The high recall (0.708) but low precision (0.401) of this model indicates that there are more false positives (i.e., unattested clusters which do not violate any constraint) than false negatives (i.e., exceptions to the three rules).
+It is not clear that low precision is a flaw, since some of these false positives might be accidental gaps---that is, clusters that are well-formed, but unattested.
 
-\subsubsection{\citet{Pierrehumbert1994}}
+\subsubsection{Expected frequency}
 
-\citet{Pierrehumbert1994} proposes a baseline in which the well-formedness of syllable contact clusters is proportional to the product of the probabilities of their codas and onsets.
-\citeauthor{Coleman1997} summarize the results as follows.
+\citet{Pierrehumbert1994} proposes a simple model in which the wellformedness of syllable contact clusters is proportional to the joint probability, estimated from word-final coda and word-initial onset frequencies and assuming that coda and onset are statistically independent.
+There are two problems with this procedure, however.
+First, there exist languages, such as Finnish \citep[36]{Fischer-Jorgensen1952}, in which there are medial codas not found in final position; this alone suggests that ``outer'' frequencies are not the best model of what is possible medially.
+More to the point, a pilot study determined that using medial coda and onset frequencies produced a better fit, and this variant is adopted here.
+%\citeauthor{Pierrehumbert1994}'s procedure would assign clusters of this type the lowest score. 
+While \citeauthor{Pierrehumbert1994} reports that this is the best predictor of which complex clusters occur, it imposes no constraints on sequences spanning the syllable boundary, and consequently has lower accuracy (0.839) and $F_1$ (0.328) than the derived constraint baseline, though it is a small but significant improvement over the null baseline (sign test $p = 1.23$\e{-04}).
+
+\subsubsection{\citealt{Hayes2008a}}
+
+\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints.
+Since the model has many experimenter-determined parameters, a direct replication of the experiments reported by \citeauthor{Hayes2008a} is attempted; their software, feature specification, and model settings are all used.
+Since the training of this model is inherently stochastic, producing slightly different outcomes on each run, the results reported here from the lowest accuracy of ten independent runs, a practice also used by \citeauthor{Hayes2008a}.
+The model achieves the highest accuracy (0.877) and $F_1$ (0.703), but lower recall (0.642) than the alternation baseline, a significant improvement over the null baseline (sign test $p = 2.9$\e{-05}), but not significantly different from the performance of the alternation baseline (sign test $p = 0.070$).
+
+The generalizations extracted by the \citeauthor{Hayes2008a} model are in general quite similar to those derived from the \emph{SPE} alternations, but there are some notable differences.
+Unattested (and ill-formed) *[m.kl] is assigned the highest score, though it is ruled out by \textsc{Coda Nasal Place Assimilation}.
+On the other hand, very low scores are assigned to the clusters in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.
+
+\subsubsection{\citealt{McGowan2009}}
+
+\citet{McGowan2009} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2009} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
+However, \citeauthor{McGowan2009} finds that sonority of clusters accounts for a very small of the variance in cluster frequency ($R^2 = 0.077$).
+As a predictor of cluster attestation, sonority distance provided no improvement over the null baseline; consequently, it is not included in Table \ref{cmresults}.
+
+\subsubsection{Local summary}
+
+Compared to phonological alternations, current computational models of phonotactic learning are not significantly better predictors of which syllable contact clusters a
+
+\section{Conclusions}
+
+None of the foregoing results have revealed any evidence for structural constraints on the English syllable contact cluster, whether ``discovered'' by linguist or computer, except those derived from phonological alternations.
+What then is to be said about the countless clusters which do not violate any phonologically derived constraint, but yet are unattested, like [b.z] or [z.n]?
+Since these gaps appear to be quite arbitrary from a phonological perspective, all that can be said at this juncture is that these clusters must be regarded \emph{accidental gaps}: they are well-formed clusters, but simply missing from the English lexicon due to the sparsity thereof, a possibility that has long been recognized.
 
 \begin{quote}
-\citet{Pierrehumbert1994} showed that the independent probabilities of the coda and the following onset was the single biggest factor in predicting which complex clusters exist.
-\citep[50]{Coleman1997}
+\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
+\citep[][16]{Fischer-Jorgensen1952}
+
+The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
 \end{quote}
 
-\noindent \citet{Pierrehumbert1994} uses both medial and initial/final codas and onsets to compute these probabilities, but a better fit is obtained by using only word-medial coda and onset frequencies.
-\footnote{The use of medial frequencies is likely to cross-linguistically superior, as there are many languages, like Finnish, in which medial codas are not possible final codas \citep[][306]{Fischer-Jorgensen1952}.} The resulting probabilities are transformed into a classification of ``attested'' or ``unattested'' using a soft-margin support vector machine \citep{Cortes1995} with a linear kernel.
-This technique finds the optimal point for separating attested and unattested clusters.
-This is not a cognitively plausible learning model: it simply provides an upper bound on the performance of the \citeauthor{Pierrehumbert1994} model.
-By definition, this model imposes no constraints on sequences crossing the syllable boundary, and unfortunately, this only provides a small improvement to the null baseline (accuracy = 0.839), and has a lower accuracy and $F_1$ (0.328) than the derived constraint baseline.
+\noindent Below, attempts are made to quantify the extent of accidental gaps in this domain below.
 
-%\citep{Haugen1956}. Thanks to Stephen Anderson for this observation
+\subsection{Zipfian distribution}
 
-\subsubsection{\citet{Hayes2008a}}
+A number of previous studies \citep[e.g.,][]{Weiss1961,Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phoneme or letter frequencies exhibit sparse distributions, in which a few types account for a large amount of the probability mass, and the remaining is divided up among a large number of rare types.
+The same is true for the type frequencies of the medial codas and onsets which make up syllable contact clusters, as well as cluster frequencies themselves.
+In Figure \ref{clus}, cluster frequencies are plotted against rank in the log-log space, and a near-linear relationship obtains ($R^2 = 0.924$), indicating that the data conform to a generalization of Zipf's (\citeyear{Zipf1949}) Law.
+See Appendix \ref{zr} for estimation details.
 
-\citet{Hayes2008a} (henceforth, H\&W) develop a sophisticated method of learning phonotactic distributions from positive data.
-They estimating probability distributions over sequences of sets of phonological features using the principle of maximum entropy.
-As above, a support vector machine with a linear kernel is used to turn these probability values into binary predictions.
-A direct replication of the experiments by H\&W are attempted, by using their software, feature specfication, and model settings.
-Since the training of this model is inherently stochastic, producing slightly different outcomes on each run, the results reported here are thoes of the lowest-accuracy run out of ten.
-This practice is used by H\&W (p.~396), though is not a great deal of variation between individual runs.
-
-This model provides a small improvement in accuracy (0.877) and $F_1$ (0.703) compared to the alternation baseline, but it has lower recall (0.642) than this baseline, indicating the presence of false negatives.
-More specifically, the H\&W model learns narrower phonotactic generalizations than those derived from alternations.
-For instance, unattested *[m.kl] is assigned the highest probability, whereas it is ruled out by \textsc{Coda Nasal Place Assimilation}.
-On the other hand, the \citeauthor{Hayes2008a} model assigns a low score to clusters like those found in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.\footnote{\citet{Berent2012} discuss a potential defect of the H\&W model: it does not use of variables over features and thus fails to unify constraints against geminate clusters.
-This does not appear to be relevant to this evaluation, as the H\&W model correctly predicts the total absence of geminate clusters.}
-
-\subsection{Discussion}
-
-Though the H\&W model has the overall best performance, it is not greatly superior to the derived constraint baseline.
-This suggests that most, if not all, of the 71.2\% of possible clusters which are unattested are not ill-formed, but are simply missing due to the sparsity of the data: that is, they are accidental gaps.
-Below, the role of sparsity in the syllable contact inventory is quantified in a number of ways.
-
-\subsubsection{The Zipfian distribution}
-
-A number of previous studies \citep[e.g.,][]{Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phonemes and graphemes exhibit sparse type (i.e., lexical) and token frequency distributions.
-The lexical frequencies of the medial codas ond onsets which make up syllable contact clusters show a similar distribution.
-
-In a sparse distribution, many types occur at the same low frequencies.
-\citet{Good1953} notes that this produces a type of quantization, an artificially long and flat right tail in the frequency spectrum.
-\citet[][29]{Church1991} propose a method to eliminate this quantization, the $Z_r$ transform, which is defined as follows.
-The vectors $r, n$ are such that $n_1$ is the number of types which occur at frequency $r_i$ (that is, $n$ is a vector of frequencies of type frequencies), and $r$ is sorted in increasing order.
-$Z$ is simply each element of $n$ scaled by the nearest points to the left and right.
-
-\begin{unlabeledexample}
-$\displaystyle Z_i = \frac{2 f(rn_i}{r_{i + 1} - r_{i - 1}}$
-\end{unlabeledexample}
-
-\noindent \citeauthor{Church1991} do not define this transform for the lowest and highest points (i.e., when $i = 1$ or $i = N$), but the definition can be extended by scaling the edge points according to next innermost point.
-
-\begin{unlabeledexample}
-$\displaystyle Z_1 = \frac{n_1}{r_2 - r_1}$
-\end{unlabeledexample}
-
-\begin{unlabeledexample}
-$\displaystyle Z_N = \frac{n_N}{r_N - r_{N - 1}}$
-\end{unlabeledexample}
-
-In Figure \ref{clus}, $Z_r$-transformed lexical frequencies of codas and onsets are plotted against rank, in log-log spaces.
-The near-linear relationship that obtains ($R^2 = 0.924$) indicates that the data conform to a generalization of Zipf's (\citeyear{Zipf1949}) law.
-
-\begin{unlabeledexample}
-$\displaystyle Z_r(C, \alpha) = \frac{C}{r^\alpha}$
-\end{unlabeledexample}
-
-\noindent The parameters $C$, a constant sensitive to sample size, and $\alpha$, the slope of the log-rank/log-frequency line, can be computed using the method of least squares, with $\epsilon$ representing the error terms.
-
-\begin{unlabeledexample}
-$\displaystyle \textrm{log}~Z_r \sim C + \alpha~\textrm{log}~r + \epsilon$ 
-\end{unlabeledexample}
-
-\begin{figure}
-\centering
+\begin{figure} \centering
 \includegraphics{cluster.pdf}
-\caption{The log-log linear relationship between frequency and rank demonstrates that syllable contact clusters are Zipf distributed.}
+\caption{Syllable contact clusters exhibit a log-log linear relationship between frequency and rank that is consistent with Zipf's Law.}
 \label{clus}
@@
 
-It is important to note that there is nothing inherently meaningful about the fact that clusters conform to a Zipfian distribution.
-These distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2009} to word \citep{Teahan1998,Ha2002,Baroni2009} and phoneme \citep{Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
-The importance of this statistical property is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps due only to the sparsity of the data.
+Sparse distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2009} to word \citep{Teahan1998,Baroni2009} and phoneme \citep{Belevitch1956,Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Mandelbrot1954,Miller1957,Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
+The importance of sparsity in this context is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps.
 
-\subsubsection{The Good-Turing estimate}
+\subsection{Good-Turing estimate}
 
-\citet{Good1953} proposes an estimate of $\hat{p}_0$, the probability of all unseen events, originally developed in collaboration with Alan Turing.
-This probability is simply $n_1$, the number of events which occur just once, divided by the number of observations.
+\citet{Good1953} proposes an estimate for the probability of all accidental gaps, $\hat{p}_0$, as the number of events that occur just once, $n_1$ divided by the number of observations.
 
@@
 
-\noindent In CELEX, $65$ clusters occur only once, and there are 873 cluster tokens, so $\hat{p}_0 = 0.074$.
-The interpretation of this quantity is as follows: were possible to generate a new corpus of syllable contact clusters, approximately 7\% of the tokens would be clusters not found in the previous sample.
-Unfortunately, the Good-Turing estimate does not provide a way to determine which unattested clusters might be found in a replication.
+\noindent In the CELEX data, $65$ clusters occur only once, and there are 873 cluster tokens, and thus $\hat{p}_0 = 0.074$.
+There is thus a non-trivial probability that many of the ``false positive'' clusters are in fact accidental gaps.
 
-\subsubsection{Sampling simulation}
+\subsection{Sampling simulation}
 
-As a final consideration of this data, consider the statistical properties of a randomly generated ``lexicon'' as they compare to the observed distribution of clusters.
-A simulated lexicon can be generated by repeatedly applying the following procedure, which combines the derived constraint baseline with the model proposed by \citet{Pierrehumbert1994}.
+It is finally possible to ask what the lexicon might look like if it was stochastically generated from independent coda and onset frequencies, \emph{a la} \citeauthor{Pierrehumbert1994}, but also constrained by phonological neutralizations.
+The following method was repeated to generate a simulated lexicon.
 
@@
 b. & Sample a medial onset according to the observed probabilities \\
-c. & Apply all matching \emph{SPE} rules to the resulting clusters \\
+c. & Apply all matching \emph{SPE} rules to cluster formed by their concatenation \\
 \end{tabular}
@@
 
-While this procedure can be repeated indefinitely, it is worthwhile to consider the results of a single characteristic ``lexicon'' produced in this fashion.
-The slope of the log-rank/log-frequency relationship for the simulated data ($\alpha = -0.524$) is also close to the observed slope for the CELEX data ($\alpha = -0.588$).
-And, as shown in Figure \ref{sim}, the two distributions are nearly indistinguishable.
+The cluster frequencies for a single characteristic simulation are plotted in Figure \ref{sim}, along with observed frequencies; the two distributions are nearly indistinguishable.
+The sparse cluster inventory, previously taken as evidence for static constraints on syllable contact, is approximately what one would expect even if the only constraints on possible clusters are phonologically derived.
 
@@
 \includegraphics{sim.pdf}
-\caption{The observed cluster frequencies are closely matched by a simulated cluster inventory generated by random sampling.}
+\caption{The observed cluster frequencies are closely matched by a simulated cluster inventory generated by random sampling, assuming independence and applying phonological neutralizations.
+The $Z_r$ transform \citep[][29]{Church1991} has been used to eliminate quantization in lower frequency bands.}
 \label{sim}
 \end{figure}
-
-This result should not be construed as an endorsement of the \citet{Pierrehumbert1994} model.
-Rather, it shows that even if the only constraints on syllable contact clusters are that they must consist of valid medial codas and onsets which further conform to the three derived constraints, there still will be many gaps in the cluster inventory, gaps which are quite arbitrary from the perspective of the phonology.
-
-\section{Conclusion}
-
-\citet{Borowsky1989} on peripherality.
-productive \citet{Duanmu2008}
-
-The next two chapters return to the question of synchrony, addressing the relationship between statistical patterns in the lexicon and speakers' behaviors when presented with underrepresented sequences.
-
-accidental gaps
-\citet[][419f.]{Hayes2008a}
-
-%as in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} CELEX also contains \emph{jo}[d.p]\emph{urs}, \emph{pi}[ntʃ.b]\emph{eck}, \emph{sa}[k.b]\emph{ut}, and \emph{ja}[k.d]\emph{aw}. 
-%There is no duplication between constraints on sequence structure and surface forms if both are derived from the application of the phonological rule (\citealt[][401f.]{Stanley1967}, \citealt[][382]{SPE}).
-%\citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''.
-In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.
-%\begin{example}[Artificial language adaptations (\citealt{Wright1975}:394, his transcriptions)]
-%\begin{tabular}{l l l l}
-%a. & [gownp] & \textgreater{} & [gump]  \\
-%b. & [ǰumg]  & \textgreater{} & [ǰúŋgə] \\
-%c. & [ðʌŋd]  & \textgreater{} & [tɔŋg]  \\
-%\end{tabular}
-%\end{example}
-%\noindent If this game is akin to the process of loanword adaptation, it is possible that various extragrammatical pressures are at play \citep[e.g.,][]{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008}.
-Yet, the independent evidence for a phonological process of nasal place assimilation provides the most parsiminious account of the adaptations seen in \citeauthor{Wright1975}'s study.
-%\citet{Hay2004a} demonstrates that [n.p, mθ] clusters, which are not found in English, are in fact rated better than many attested clusters.
-%The same is true of the prefix \emph{com-}/\emph{con-}, as shown by the presence of assimilation in \emph{có}[ŋ.g]\emph{ress} but not in \emph{co}[n.g]\emph{réssional}. 
-%(e.g., \citealt{Bakovic2005b}, \citealt{Fruehwald2011}),
-
-%\ex \textsc{Coda Nasal Place Assimilation} feeds \textsc{Degemination} \citep[][116]{Borowsky1986}: \\
-%\begin{tabular}{l l l l}
-%a. & mature    & i[m]ature    & (cf. \emph{i}[m.b]\emph{alance})   \\
-%b. & numerable & i[n]umerable & (cf. \emph{i}[n.d]\emph{ependent}) \\
-%\end{tabular}
-%\xe
-%(\citealp{Halle1985a}:105, 
-%; presumably, this generalization is limited to this one place feature because  \citeauthor{Pierrehumbert1994} also adopts the assumption that the velar nasal is derived from /n/. 
-%Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets. 
-%
-%\begin{example}[VBR's complex onsets (\citealp{Buchwald2005}:79--80, his transcriptions)] \label{VBR}
-%\begin{tabular}{l l l}
-%a. & kəræb  & `crab'  \\
-%   & bəlid  & `bleed' \\
-%b. & kəwin  & `queen' \\
-%   & kəwoʊt & `quote' \\
-%c. & kut    & `cute'  \\
-%   & musɪk  & `music  \\ 
-%\end{tabular}
-%\end{example}
-%
-%\noindent VBR breaks up complex onsets with epenthesis, including those that consist of a consonant and a back onglide cluster (\ref{VBR}b).
-However, no epenthesis occurs between a consonant and a front onglide; rather, the glide is absent (\ref{VBR}c).
-The failure of the front onglide to pattern with other consonant clusters suggests once again that the glide is part of the nucleus.
-%j
-%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [juː] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant.This is the only sign that [juː] shows an affiliation for the onset.
-%In fact, [t.ʃ, d.ʒ] clusters are not found in simplex English words, despite the fact that that affricates occur as medial onsets in clusters (e.g., \emph{tru}[n.tʃ]\emph{eon}, \emph{so}[l.dʒ]\emph{er}).
-%Finally, \citet[][251]{Fromkin1973} also presents evidence that post-vocalic \emph{r} may behave as if nuclear in speech errors. 
-%\citep[e.g.,][142]{Mester1988}.
-%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or a part of a contrast in any known language \citep{Rubach2002}.}
-% Note that [tʃ] has not been included since it so rarely occurs in clusters.
-%\subsubsection{Summary}
-%
-%The above results are summarized in Table \ref{constraints}.
-%
-%\begin{table}
-%\centering
-%\begin{tabular}{l r r r r r}
-%\toprule
-%                                       & \multicolumn{2}{c}{attested} & \multicolumn{2}{c}{unattested} & \multirow{2}{*}{$p$-value} \\
-%                                       & conforming & violating & conforming & violating \\
-%\midrule
-%\textsc{O.V.A.} (\ref{ovar})        &  80 &   6 & 370 & 264 & 1.1\e{-11} \\
-%\textsc{C.N.P.A} (\ref{CNPA})       &  41 &  25 &  12 &  42 & 1.7\e{-05} \\ 
-%\textsc{Degemination} (\ref{degem}) & 161 & 632 &   0 &  87 & 1.4\e{-08} \\
-%\midrule
-%\textsc{*Dorsal-Labial} ()          &  68 &   4 & 246 &  40 & 0.069 \\
-%\textsc{*C.C.O.} ()                 &  48 &  38 & 312 & 322 & 0.301 \\
-%\textsc{*A,BA} ()                   & 161 &   0 & 708 &  11 & 0.231 \\
-%\bottomrule
-%\end{tabular}
-%\caption{Counts and Fisher exact test $p$-values for the constraints discussed above.}
-%\label{constraints}
-%\end{table}
-%Whereas 18.8.\% of the ``possible'' clusters are attested, 28.8\% of those which conform to these three constraints are found. 
-%Binary classification contingency table: \\
-%\begin{tabular}{c | l l} \toprule
-%                             & \multicolumn{2}{c}{prediction} \\
-%\midrule
-%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
-%                             & false negative ($fn$) & true negative ($tn$) \\
-%\end{tabular}
-%\begin{tabular}{|l l|}
-%\toprule
-%true positive ($tp$)  & false positive ($fp$) \\
-%false negative ($fn$) & true negative ($tn$) \\
-%\bottomrule
-%\end{tabular}
-%\begin{figure}
-%\centering
-%\includegraphics{coda.pdf} 
-%\includegraphics{onset.pdf}
-%\caption{Medial coda and medial onset lexical frequency exhibit the Zipfian log-log-linear relationship between rank and frequency.}
-%\label{codaonset}
-%\end{figure}
-%Sparsity alone is not the only reason for cluster underrepresentation which is external to the phonology.
-For instance, a cluster could be underrepresented due to language change.
-\citet{Iverson2005} discuss a case of this type: In English, long vowels before word-final [ʃ] in English are generally rare (though note ``affective'' \emph{swoosh}, \emph{sheesh}, etc.).
-This is ultimately due to the fact that native [ʃ] is descended from Old English [sk], and long vowels are not fonud not occur before complex codas.
-\citeauthor{Iverson2005} conclude that this gap is accidental because it does not inhibit borrowing and coinage.
-%On the other hand, patterns created by sound change are not guaranteed to persist over time.
-One example of non-persistence is discussed by \citet{Iverson2005}.
-Around 1100 CE, Old English \emph{sk} became [ʃ].
-This sound change introduced no alternations.
-Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{V\lm sk\#} words when the change was actuated, and \emph{V\lm sh\#} continues to be rare in Modern English.
-What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.
-%\citet[][140]{Frisch2004} suggest that the strong tendency for the first and second consonants of the Arabic root to be non-identical is the ``a diachronic result of a processing constraint that disfavors repetition.'' Unfortunately, there is no evidence that this pattern is diachronic other than in the sense that it appears to be inherited from the proto-language: there is simply no Proto-Semitic verb roots with identical first and second consonants \citep[][178]{Greenberg1950}.
-In other Semitic languages, the inherited patern has experienced considerable erosion.
-%
-%\begin{example}[Tigrinya roots with identical first and second consonants \citep{Buckley1990a}]
-%\begin{tabular}{l l l l}
-%a. & lʌlʌw     & `scorch'                   & (\textless{} Ge'ez \emph{lʌwlʌw} `inflame')     \\
-%   & mʌmʌy     & `winnow'                   & (\textless{} Ge'ez \emph{mʌymʌy} `distinguish') \\
-%   & mʌmʌt     & `pick out loot'            \\
-%b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
-%   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
-%c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
-%   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
-%\end{tabular}
-%\end{example}
-
-%Similar exceptions are found in 
-%Amharic (\citealp[][?]{Broselow1984}, \citealp[][?]{McCarthy1985}) and 
-%Hebrew \citep[][29]{Bat-El2005}.

### 2012-10-03 1.tex
Edits to 2 and 4, and writing in 3

--- 2012-05-24 1.tex
@@
-\chapter{Sequence structure constraints in generative grammar}
-\label{msc}
+\chapter{A generative theory of phonotactics and phonotactic learning}
 
-The first part of this text concerns the empirical substance of the theory of \textsc{Morpheme Structure Constraints} (MSCs), that is, language-specific constraints on the shapes of morphs stored in lexical memory.
-Evidence is presented to support the claim that such constraints are solely derived by other components of grammar, and do not have an ontology of their own.
+\section{\emph{Táxis} and generativity}
 
-The first section of this chapter outlines a brief history of MSCs.
-The following section draws upon this material to present the above claim as a null hypothesis, and presents conditions for falsifying it.
-The final section outlines the remainder of Part I (chapters 2--4), which present evidence in support of this hypothesis.
+%\emph{τάξις}
+(\emph{táxis} `order, arrangement; the prefix \emph{syn-} means `together')
 
-% it is not clear what follows from a phonotactic description
-% pater on NC
-% can't be right tho, see stanley, kisseberth, clayton, shibatani, hooper, etc.
+This etymology gives little clue to just what syntacticians do, however.
+This is not merely a consequence of the ``generative revolution'', as the predecessors of transformational grammars associated with Zellig Harris and Noam Chomsky were also more than theories of arrangement, but also instantiate the notion of derivation \citep[e.g.,][]{Post1943}.
 
-\section{A brief history of MSCs}
+\subsection{?}
 
-\subsection{Structuralism}
+\subsection{Units of description}
+\subsubsection{Segment inventory}
+\subsubsection{Surface segment sequences}
+\subsubsection{Underlying segment sequences}
 
-Since, with the possible exception of Roman Jakobson (who might be labeled an early generativist), the mentalistic interpretation of structural linguistic analyses was anathema, one must take care to not assume that a description of MSCs is something that these linguists would posit as part of speakers' knowledge of language.
+\subsection{The morph}
 
-\subsubsection{Morpheme boundaries in phonology}
+The notion of blocking in non-derived environments (\citealp[163]{Kiparsky1973a}, \citeyear[152]{Kiparsky1982a}, \citealp{Mascaro1976}) is well-known, but \citet{Hall2006} also amasses evidence for phonological processes restricted to non-derived environments.
+One of the most famous examples is the distribution of \emph{ich-laut} [ç] and \emph{ach-laut} [x] in German \citep{Bloomfield1930}.
+As shown in (\label{ichach}), [x] appears only before back vowels; in all other contexts, the only dorsal fricative is [ç].
 
-bloomfield
-
-%While \citeauthor{Bloomfield1930} is known for introducing morphemic structure into phonology, a member of the Prague circle developed an arguably more sophisticated approach only two years later.
-
-\citet{Jakobson1932} attributes a number of phonological properites of the Russian imperative to morpheme juncture.
-The vast majority of Russian infinitives consist of the stem followed by a theme vowel and /t\pal/.
-The reflexive is marked wih a final /-sə/ which feeds a general process of regressive voicing assimilation.
-In the reflexive infinitive, /\ldots t\pal-s\ldots/ is realized [ts], with no palatalization.
-However, palatalization of a stem-final labial or coronal---including /t\pal/, as in (\ref{russian}c)---is preserved in the reflexive imperative.
-
-\begin{example}
-\label{russian}
-Russian reflexives \citep[after][]{Jakobson1932}:\footnote{I have taken a number of liberties with \citeauthor{Jakobson1932}'s presentation of the data, which uses an abstract phonemic transcription.
-Thanks to Lev Blumenfeld (p.c.) for help with the transcription of this data.}
-
-\begin{tabular}{l l l l} %\toprule
-   &  infinitive      & imperative \\ %\midrule
-a. & [slav\pal itsə]  & [slaf\pal s\pal ə]  & `be glorious'    \\
-   & [upram\pal itsə] & [upram\pal s\pal ə] & `be stubborn'    \\
-b. & [kras\pal itsə]  & [kras\pal s\pal ə]  & `put on makeup'  \\
-   & [ʒar\pal itsə]   & [ʒar\pal s\pal ə]   & `roast'          \\
-c. & [zəbytsə]        & [zəbut\pal s\pal ə] & `forget' \\ %\bottomrule
-%   & ab\'utsa      & ab\'ujsa     & `put on shoes'   \\
+\begin{example}[German \emph{ich}- and \emph{ach-laut}]
+\begin{tabular}{l l l}
+a. & \emph{B}[uːx]          & `book'           \\
+   & \emph{T}[ɔx]\emph{ter} & `daughter'       \\
+   & \emph{N}[ax]\emph{t}   & `night'          \\
+b. & \emph{S}[iç]\emph{t}   & `view'           \\
+   & \emph{Sp}[eç]\emph{t}  & `woodpecker'     \\
+   & \emph{Ger}[yç]\emph{t} & `rumor'          \\
+   & \emph{Kn}[øç]\emph{el} & `ankle, knuckle' \\
+   & \emph{Fl}[ɛç]\emph{e}  & `surface'        \\
 \end{tabular}
@@
 
-\noindent The forms in (\ref{russian}c) have different outcomes for their /t\pal-s/ clusters.
-\citeauthor{Jakobson1932} proposes that the preservation of palatalization is a special property of the imperative.
-%A few years later, \citet{Trnka1936}, another member of the Prague Circle, makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.
+%a. & \emph{B}[uːx]          & \emph{B}[yːç]\emph{er} & `book' \\
+%   & \emph{L}[ɔx]           & \emph{L}[øç]\emph{er}  & `hole' \\
+%%   &\emph{T}[ɔx]\emph{ter} & \emph{T}[øç]\emph{ter}
+%   & \emph{N}[ax]\emph{t}   & \emph{N}[ɛç]\emph{te} & `night' \\
 
-\subsubsection{The morph as a constraint domain}
+\emph{K}[uːx]\emph{en}
+\emph{K}[uːç]\emph{en}
 
-A corrolary of this hypothesis is that the phoneme sequences found within morphs may be distinct from those which span multiple morphs.
+%\emph{Mas}[oːx]-\emph{Mas}
+\footnote{Examples like \emph{Mas}[oːç]\emph{ist} `masochist', \emph{Eun}[uːç]\emph{ismus} `eunichism', first noted by \citet{Merchant1994}, suggest that this should also be restricted to assimilation within the same foot \citep[226f.]{Jensen2000}.}
 
-This was apparent to structuralists
+%\emph{Mas}[oːx]-\emph{Mas}[oːç]\emph{ist}
 
-the phonologists 
-This is apparent from 
-a discussion of constraints on vowel sequences in Mixteco given by \citet{Pike1947b}.
-It is apparent that \citeauthor{Pike1947b}'s constraints on vowel sequences are not intended to hold across morph boundaries,
+The concept of non-derived environment blocking
 
-according to the phonological analysis of a Mixteco text published a few years earlier \citep{Pike1944}.
+\section{The model}
 
-\begin{example}
-Mixteco MSCs \citep{Pike1947b} and complex words \citep{Pike1944}:
+\subsection{Prosodic parsing}
+\subsubsection{Syllabification}
+\subsubsection{Autosegmental tiers}
 
-\begin{tabular}{r l l l} %\toprule
-   & MSC & complex exception \\ %\midrule
-%a. & *{C}a{C}e & [k\'a-\textsuperscript{n}dee] & `kept \ldots inside' \\
-%b. & *{C}\textipa{@}{C}e & [n\`i-k\`ə bə-de] & `who entered'        \\
-%c. & *{C}e{C}i & [te-n\'i-ke\textsuperscript{n}da] & `was walking         \\
-%d. & *{C}i{C}e & [te-n\`i-kee-t\`ə] & `and went away'      \\ %\toprule
-%e. & *{C}e{C}o & b\'e\textglotstopvari e-\v{z}\'o & `our house'          \\
-%f. & *{C}eo & ke-o-d\'e & `we eat him'         \\
-a. & *{C}a{C}e & [ká\textsuperscript{n}dee] & `kept \ldots inside' \\
-b. & *{C}ə{C}e & [nìk\`əbəde] & `who entered'        \\
-c. & *{C}e{C}i & [teníke\textsuperscript{n}da] & `was walking         \\
-d. & *{C}i{C}e & [tenìkeet\`ə] & `and went away'      \\ 
-\end{tabular}
-\end{example}
+\subsection{Rule application}
+\subsubsection{Violation detection}
+\subsubsection{Rule exceptionality}
 
-\noindent \citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a morpheme discovery routine.
+\section{Consequences of the model}
 
-\subsubsection{Biuniqueness and neutralization}
+\subsection{Gradience}
+\subsubsection{Probabilistic well-formedness}
+\subsubsection{Cumulativity}
+\subsubsection{External factors}
 
-joos
-harris
-chomsky
+\subsection{Static phonotactics}
+\subsubsection{Statistical insensitivity}
+\subsubsection{Saussurean arbitrariness}
 
-\subsection{Early generativism}
+\subsection{Order of acquisition}
+\subsubsection{Lexical acquisition}
+\subsubsection{Prosodic parsing}
+\subsubsection{Rule application}
 
-\subsubsection{\citealt{Stanley1967}}
-
-inventory constraints
-sequence constraints
-
-\subsubsection{Surface constraints and occultation}
-
-shibatani
-stampe
-dell
-
-\emph{A}[mt] `office', \emph{Anbau} `cultivation'
-\emph{ei}[n.g]\emph{reifen} `to intervene',
-
-\subsubsection{The duplication problem}
-
-\citeauthor{Anderson1974} sees the duplication of \textsc{Backness Harmony} as both sequence structure constraint and phonological rule as a foregone conclusion
-.The tendency of sequence structure constraints to duplicate phonological rules was noted by \citet[][401f.]{Stanley1967} and in \emph{SPE} (p.~382) in the discussion of morpheme structure constraints, and is labeled ``the duplication problem'' by \citet[][?]{Kenstowicz1977}.
-
-inventory problems 
-dell, chung, but also problems
-
-conspiracies
-
-\subsection{Autosegmentalism}
-
-also ``supersegmentalism''
-
-The proposals during this period are very diverse.
-
-\subsubsection{Syllable structure}
-
-hooper
-noske
-
-\subsubsection{Morph features}
-
-Leben
-Kaye
-
-\subsubsection{The Obligatory Contour Principle}
-
-\subsection{Classic Optimality Theory}
-
-%(page numbers here are taken from the published version)
-
-\subsubsection{Richness of the Base}
-
-some quotes from OT doc
-\citet{PE}
-\citet{Bye2001}
-
-\subsubsection{Freedom of Analysis}
-
-\citet{Smolensky1996}
-\citet{PE}
-
-\subsubsection{Undominated constraints}
-
-% complaints about maceahern, gallagher
-% bit on loanword adaptation
-
-\subsection{Recent developments}
-
-\subsubsection{Language acquisition}
-% jusczyk forever
-% white et al.
-% early word learning
-% syllables
-
-\subsubsection{Psycholinguistic tasks in adults}
-
-% wordlikeness
-% spotting
-% error correction, etc.
-
-\subsubsection{Non-native word adaptation}
-
-% perceptual factors
-
-\section{The null hypothesis}
-
-\subsection{The empirical proposal}
-
-\subsection{The acquisition principle}
-
-\subsection{Falsifiability}
-
-%\subsection{Universal constraints?}    \input{1.2.4}
-
-\section{Outline of the dissertation}
-
-%Chapter 2
-%Chapter 3
-%Chapter 4
+%\subsubsection{Exceptionality}

### 2012-10-03 2.tex
Edits to 2 and 4, and writing in 3

@@
 
-\citet{Halle1962} and \citet{Chomsky1965} observe that speakers internalize generalizations about possible and impossible words in their language.
-\citeauthor{Chomsky1965} illustrate this knowledge with two nonce words \emph{blick} [blɪk] and \emph{bnick} [bnɪk].
-Whereas \emph{blick} is unattested, but in some sense a ``possible'' word of English, speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word of the language.
+\citet{Halle1962} and \citet{Chomsky1965} use two nonce words, \emph{blick} [blɪk] and \emph{bnick} [bnɪk], to illustrate that speakers internalize generalizations about possible and impossible words in their language.
+While \emph{blick} is unattested, but in some sense a ``possible'' word of English, \citeauthor{Chomsky1965} observe that speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word, despite the fact that it is well-formed in other languages (e.g., Moroccan Arabic \emph{bniqa} `closet').
 
 A naïve account of this contrast is based on the assumption that segments or timing units must either be parsed into prosodic structures like the syllable, or be subject to phonological repair.
-Early precedents for this can be found in dissertations by \citet[10f.]{Hooper1973} and \citet[57f.]{Kahn1976}, and it is a fundamental component of many later frameworks \citep[e.g.,][]{Ito1989a,Noske1992,OT}.
+Early precedents for this assumption can be found in work by \citet[10f.]{Hooper1973} and \citet[57f.]{Kahn1976}, and it is fundamental to many later frameworks \citep[e.g.,][]{Ito1989a,Noske1992,OT}.
 
@@
 
-Many linguists who have concerned themselves with ``wordlikeness'' now regard the above as overly simplistic, for, as \citeauthor{Shademan2006} argues, such accounts draw a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report ``wordlikeness'' intuitions.
+However, many linguists now regard the above as overly simplistic, for, as \citeauthor{Shademan2006} argues, such accounts draw a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report ``wordlikeness'' intuitions.
 
@@
 But when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values.
-\citep[][371]{Shademan2006}
+\citep[371]{Shademan2006}
 \end{quote}
@@
 This chapter focuses on the hypothesis, implicit in \citeauthor{Shademan2006}'s statement, that there is an inherent gradient or continuum between possible and impossible words, against an null hypothesis that the observed patterns of judgements are little more than an artifact of the wordlikeness task itself.
-\S\ref{2background} places this debate in a historical perspective, and argues that the naïve but widely accepted form of this hypothesis fails the test of falsifiability.
+\S\ref{2background} places this debate in a historical perspective, and argues that the naïve but widely accepted form of this hypothesis cannot be falsified.
 In \S\ref{2evaluation}, a quantitative evaluation of a revised version of this hypothesis reveals a poor match between the gradience in speakers' judgements and contemporary computational models of wordlikeness.
-Taken together, the findings argue against the gradience hypothesis.
+Taken together, the findings undermine the gradience hypothesis.
 
@@
 However, not all early literature is concerned with gradience.
-\citet[][31]{Vogt1954}, for instance, recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
+\citet[31]{Vogt1954}, for instance, recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
 \citeauthor{Vogt1954} observes that allophony may account for the absence of certain phone sequences, but it does not provide a suitable explanation for the absence of initial [bn] in English, nor does it make correct predictions about the surface realization of an underlying initial /bn/.
@@
 
-Gradient grammaticality is assumed in the earliest work on generative syntax, \emph{The logical structure of linguistic theory}, in which \citet[][132]{LSLT} writes that ``there is little doubt that speakers can fairly consistently order new utterances, never previously heard, with respect to their degree of `belongingness' to the language''.
+Gradient grammaticality is assumed in the earliest work on generative syntax, \emph{The logical structure of linguistic theory}, in which \citet[132]{LSLT} writes that ``there is little doubt that speakers can fairly consistently order new utterances, never previously heard, with respect to their degree of `belongingness' to the language''.
 Later, \citet{ASPECTS} proposes that different syntactic violations give rise to different degrees of ungrammaticality.
 Unlike \citeauthor{Fischer-Jorgensen1952}, however, for \citeauthor{LSLT} this continuum does incorporate a hard-and-fast boundary between the grammatical and ungrammatical: there are no constrasts between grammatical utterances, merely between different degrees of ungrammaticality.
-For \citeauthor{LSLT}, all grammatical utterances are alike with respect to their grammaticality, but there are many degrees of ungrammaticality \citep[][61]{Schutze1996}.
+For \citeauthor{LSLT}, all grammatical utterances are alike with respect to their grammaticality, but there are many degrees of ungrammaticality \citep[61]{Schutze1996}.
 
@@
 Both authors derive wordlikeness effects from morphophonemic redundancy rules acting on underlying representations.
-The discussion of \emph{bnick} by \citet[][101]{Chomsky1965} is a good example of this type of analysis.
-\citeauthor{Chomsky1965} observe that a consonant before a word-initial stop can only be a liquid in English.
+The discussion of \emph{bnick} by \citet[101]{Chomsky1965} is a good example: \citeauthor{Chomsky1965} observe that a consonant before a word-initial stop can only be a liquid in English.
 A corresponding redundancy rule which specifies a consonant in this position as [$+$\textsc{Liquid}] precludes the derivation of \emph{bnick}, since there is no way to realize the second consonant as a nasal.\footnote{This is not the only redundancy that could rule out \emph{bnick}.
-For instance, /s/ is the only word-initial obstruent that may precede a nasal consonant in English.}
+For instance, /s/ is the only word-initial consonant that may precede a nasal consonant in English.}
 
@@
 In addition to the contrast of \emph{blick} and \emph{bnick}, they introduce a third nonce word \emph{bnzk}, which they take to be even less well-formed than ``impossible'' \emph{bnick}.
+They write:
 
@@
 
-In some languages (e.g., Imdlawn Tashlhiyt Berber [tm.zħ] `she jested', [tzmt] `it (f.) is stifling'; \citealt{Dell1985}), \emph{bnzk} is syllabifiable, so its ill-formedness is once again a property of English, not a universal.
+%In some languages (e.g., Imdlawn Tashlhiyt Berber [tm.zħ] `she jested', [tzmt] `it (f.) is stifling'; \citealt{Dell1985}), \emph{bnzk} is syllabifiable, so its ill-formedness is once again a property of English, not a universal.
 
@@
 Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no attention in the literature.
-At the risk of explaining what might be no more than an an accidental gap in the historical record, the novel aspects of the \emph{SPE} model---gradience derived from similarity to existing lexical entries---seem to have been overshadowed by compelling arguments against \citeauthor{SPE}'s assumption \citep[see also][]{Halle1962,Stanley1967} that wordlikeness contrasts derive from properties of underlying forms.
+At the risk of explaining what might be no more than an an accidental gap in the historical record, the novel aspects of the \emph{SPE} model---gradience derived from similarity to existing lexical entries---seem to have been overshadowed by compelling arguments against \citeauthor{SPE}'s assumption \citep[see also][]{Halle1962,Stanley1967} that wordlikeness contrasts derive solely from properties of underlying forms.
 \citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
@@
 
-\noindent The plural appears to preserve a contrast in final obstruent voicing which is absent in the singular:\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984}) or not \citep[e.g.,][]{Port1985}, it is irrelevant to the discussion at hand.} word-final voiced stops are never found in German, and \citet[95]{Shibatani1973} reports that ``it is easy to show that a native speaker of German rejects those forms \emph{on the ground that they end in voiced obstruents}'' (emphasis in original).
+\noindent The plural appears to preserve a contrast in final obstruent voicing which is absent in the singular:\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984} or not \citep[e.g.,][]{Port1985}, it is irrelevant to the discussion at hand.} word-final voiced stops are never found in German, and \citet[95]{Shibatani1973} reports that ``it is easy to show that a native speaker of German rejects those forms \emph{on the ground that they end in voiced obstruents}'' (emphasis in original).
 However, given that root-final voicing is not predictable (e.g., /ɡʀaːt \alt{} ɡʀaːd/), the process of \textsc{Final Devoicing} cannot be any kind of lexical redundancy and is mysterious under the \emph{SPE} account.
-The alternative proposed by \citeauthor{Shibatani1973} and by \citet{Clayton1976}, is that to say that nonce words like [ɡʀaːd] are ill-formed not beacuse of any underlying property, but because they fail to undergo an otherwise-exceptionless phonological process, \textsc{Final Devoicing}, or equivalently, the surface-true generalization it derives.
+The alternative proposed by \citeauthor{Shibatani1973} and by \citet{Clayton1976}, is that to say that nonce words like [ɡʀaːd] are ill-formed not because of any underlying property, but since they fail to undergo an otherwise-exceptionless phonological process, \textsc{Final Devoicing}, or equivalently, the surface-true generalization it derives.
 
@@
 The arguments of \citeauthor{Shibatani1973} and others led theorists to focus their attention on properties of surface representations as determinants of wordlikeness.
-Though syllabification plays no role in \emph{SPE}, it is crucial to many earlier studies \citep[see][]{Goldsmith2011b}, and it received particular attention in the 1970s.
+Though syllabification plays no role in \emph{SPE}, it is crucial to many earlier studies \citep[for a review, see][]{Goldsmith2011b}, and it received particular attention in the 1970s.
 \citet{Hooper1973} and \citet{Kahn1976} argue that the syllable is useful for defining wordlikeness generalizations.\footnote{\citet{Steriade1999} and \citet{Blevins2003}, however, argue that a number of phonotactic generalizations previously stated in syllabic terms can be reanalyzed without making reference to syllables.} \citeauthor{Hooper1973} argues, for instance, that [bn], impossible as an English onset, is unobjectionable as a syllable contact cluster in nonce words like \emph{stabnik} (or in names like \emph{Abner}), and that this demonstrates the superiority of syllable-based wordlikeness generalizations.
@@
 \citeauthor{Hooper1973} also points to loanword adaptations which produce native syllable structure \citep[e.g.,][]{Carlisle1991} as evidence that syllabification is part of the phonological computation.
-Further enrichments to the theory are provided by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as envisioned by][]{Pike1947a}, and theories like prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
+Further enrichments to the theory are provided by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as first envisioned by][]{Pike1947a}, and theories like prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
 
@@
 
-Most relevant to the question at hand, \citet{Vitevitch1997} and \citet{Frisch2000} find that speakers' wordlikeness ratings of multisyllabic words are correlated wtih the positional probailities of the constituent syllables.
+Most relevant to the question at hand, \citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated wtih the positional probailities of the constituent syllables.
 Unfortunately, none of these researchers make any effort to eliminate the possibility that the low positional probability stimuli are ``impossible'' words of English.
@@
 %the author has argued elsewhere \citep{Gorman2012c}
-that many of the stimuli used by \citet{Vitevitch1997} and \citet{Frisch2000} contain illicit word-medial consonant clusters.
-While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and the wordlikeness judgements.
-\citeauthor{Vitevitch1997} ultimately conclude that it cannot explain all the variation in wordlikeness.
+that many of the stimuli used by \citeauthor{Frisch2000} and \citeauthor{Vitevitch1997} contain illicit word-medial consonant clusters.
+While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and wordlikeness judgements, but \citeauthor{Vitevitch1997} ultimately conclude this cannot explain all the variation in wordlikeness.
 
@@
 More generally, they do little to distinguish between positional probability and closely correlated measures like bigram probability (see \S\ref{bigram} below) or neighborhood density, since these studies carefully select stimuli which either have high or low values for all of positional probability, bigram probablility, and neighborhood density.
-This is particularly troublesome given that positional probability was created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production \citep[e.g.,][]{Luce1998,Luce2000} and perception \citep{Marslen-Wilson1984,Marslen-Wilson1987,McClelland1986,Norris1994,Norris2000}.
+This is particularly troublesome given that no justification has ever been given for the positional probability measure in the first place; it appears to have been created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production \citep[e.g.,][]{Luce1998,Luce2000} and perception \citep{Marslen-Wilson1984,Marslen-Wilson1987,McClelland1986,Norris1994,Norris2000}.
 
@@
 However, whereas a false consequent (e.g., an all-or-nothing concept) would potentially falsify the proposition, the consequent, a mental state, cannot be directly observed.
-If this is the case, then both the proposition and the consequent (the proposition that wordlikeness is itself gradient) fail the test of falsification, placing it beyond the reach of scientific inquiry.
+If this is the case, then both the proposition and the consequent (the proposition that wordlikeness is itself gradient) fail the test of falsification, placing it beyond the scope of scientific inquiry.
 
@@
 \citeauthor{Armstrong1983} note that if subjects produce intermediate ratings for these categories, then the naïve proposition is false, and this is what \citeauthor{Armstrong1983} find in their experiments.
-They ask subjects to rate, on a seven-point scale, the extent to which, e.g., certain odd counting numbers represent the concept ``odd number'', and find that speakers do make use of intermediate values, as can be seen in Figure \ref{agg}.
+They ask subjects to rate, on a seven-point scale, the extent to which, e.g., certain odd counting numbers represent the concept ``odd number'', and find that speakers do make use of intermediate values, as shown in Figure \ref{agg}.
 In summary, the proposition that intermediate ratings indicate mental gradience is either untestable, or it is false, and is thus rejected.
@@
 Reviewing this study, \citet[215]{Schutze2011} writes that the experiments show that judgements are ``sensitive to factors other than our underlying competence''.
-In the case of wordlikeness, a case can be made that one such factor is similarity to existing words (e.g., \citealt[][151, fn.
-27]{LSLT}, \citealt{Greenberg1964}, \citealt{Vitz1973}, \citealt{Ohala1986b}, \citealt{Sendlmeier1987}, \citealt{Bailey2001}, \citealt{Schutze2005}).
+In the case of wordlikeness, a case can be made that one such factor is similarity to existing words (e.g., \citealt{Bailey2001}, \citealt[][151, fn.
+27]{LSLT}, \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Schutze2005}, \citealt{Sendlmeier1987}, \citealt{Vitz1973}).
 However, the mere fact that subjects use intermediate ratings does not show that they do so with any systematicity, or that, e.g., the contrast between \emph{447} and \emph{7} with respect could be given any satisfying explanation.
-\citeauthor{Schutze2011} further suggests that gradience responses is a property of tasks, not concepts.
+\citeauthor{Schutze2011} further suggests that gradient responses are the product of the task itself:
 
@@
 Putting it another way, when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different.
-\citep[][215]{Schutze2011}
+\citep[215]{Schutze2011}
 \end{quote}
@@
 This introduces a troubling possibility, that subjects ``oblige'' experimenters by introducing random noise to their categorical judgements when presented with a Likert task.
-Were this the case, the observed intermediate judgements would scarcely be worthy of modeling.
+Were this the case, the observed intermediate judgements are unlikely to yield information about speakers' knowledge of possible and impossible words.
 The alternative considered here is that there exists some model in which the gradience in ratings is systematic.
@@
 An unfortunate defect of prior evaluations of these models, however, is that they adopt the naïve gradience hypothesis.
-\citet[382]{Hayes2008a} are just some of the authors who consider the ability to model gradient intuitions to be so important that they do not include any categorical models in their evaluation.
+\citet[382]{Hayes2008a} state that consider the ability to model gradient intuitions to be so important that they do not include any categorical models in their evaluation, as do many other authors.
 As a result, the literature contains no serious attempt to evaluate the gradience hypothesis.
@@
 In addition to \emph{stick}, the stimuli include three other English words; these four items were excluded from further analyses, leaving 13 stimuli.
-As is standard practice in psychophysics \citep[e.g.,][]{Butler1987}, these magnitude ratings were log-transformed before analysis.
+As is standard practice in psychophysics \citep[e.g.,][]{Butler1987}, magnitudes were log-transformed prior to analysis.
 
@@
 
+\subsubsection{Binary baseline}
+
+The binary baseline used here is a crude implementation of the null hypothesis that there are no gradient effects in wordlikeness judgements.
+To create such a baseline, it is necessary to distinguish between nonce words which contain a gross phonotactic violation and those which do not.
+As all stimuli here are monosyllables, this task can be further simplified by separately considering the two major subcomponents of the syllable, the onset and rime.
+Speakers are particularly adept as separating onsets from rimes \citep{Treiman1986,Treiman1995,Fowler1993}, and a large portion of phonotactic generalizations and violations can be localized to either onset or rime \citep[e.g.,][]{Borowsky1989,Fudge1969,Treiman1995,Kessler1997,Treiman2000}.
+
+The baseline considers a nonce syllable to be well-formed if it consists of both an attested onset and an attested rime; the free combination of these two components is the only mechanism by which this model can generalize beyond attested words.
+It is surely the case that this is an ultimately insufficient model: \citet{Albright2009a} observes, for instance, that [ɛsp] is a well-formed rime, even though it is found in no word of English.
+This problem is more acute in languages with more permissive syllable structures than those of English.
+For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.\footnote{
+Chapter \ref{clusters} 
+%\citet{Gorman2012c}
+reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
+
+The rating densities from the three studies, linearly transformed to fit within the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{dsn}.
+In all three studies, it is possible to discern a relatively sharp separation between valid and invalid clusters, but also the presence of intermediate values.
+
+\begin{figure} \centering
+\includegraphics{density.pdf}
+\caption{Average ratings of individual nonce words, linearly transformed to fit the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
+\label{dsn}
+\end{figure}
+
 \begin{table} \centering
@@
 
-\subsubsection{Binary baseline}
-
-The binary baseline used here is a crude implementation of the null hypothesis that there are no gradient effects in wordlikeness judgements.
-To create such a baseline, it is necessary to distinguish between nonce words which contain a gross phonotactic violation and those which do not.
-As all stimuli here are monosyllables, this task can be further simplified by separately considering the two major subcomponents of the syllable, the onset and rime.
-Speakers are particularly adept as separating onsets from rimes \citep{Treiman1986,Treiman1995,Fowler1993}, and a large portion of phonotactic generalizations and violations can be localized to either onset or rime \citep[e.g.,][]{Fudge1969,Treiman1995,Kessler1997,Treiman2000}.
-
-The baseline considers a nonce syllable to be well-formed if it consists of both an attested onset and an attested rime; the free combination of these two components is the only mechanism by which this model can generalize beyond attested words.
-It is surely the case that this is an ultimately insufficient model: \citet{Albright2009a} observes, for instance, that [ɛsp] is a well-formed rime, even though it is found in no word of English.
-This problem is more acute in languages with more permissive syllable structures than those of English, and thus a sparser lexicon with regard to phonotactic constraints.
-For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.\footnote{
-%Chapter \ref{clusters} 
-\citet{Gorman2012c}
-reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
-
-The rating densities from the three studies, linearly transformed to fit within the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{dsn}.
-In all three studies, it is possible to discern a relatively sharp separation between valid and invalid clusters, but also the presence of intermediate values.
-
-\begin{figure} \centering
-\includegraphics{density.pdf}
-\caption{Average ratings of individual nonce words, linearly transformed to the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
-\label{dsn}
-\end{figure}
-
 \subsubsection{Maximum entropy phonotactics}
@@
 \citeauthor{Hayes2008a} (\citeyear{Hayes2008a}; henceforth H\&W) develop a sophisticated model of phonotactic grammaticality which estimates a probability distribution over phoneme sequences by weighing constraints according to the principle of maximum entropy, following \citet{Goldwater2003} and \citet{Jager2007}.
-H\&W find that the predictions of their model are closely correlated with the \citet{Scholes1966} wordlikeness ratings.
+H\&W report that the predictions of their model are closely correlated with the \citet{Scholes1966} wordlikeness ratings.
 A direct replication of their predictions was attempted by using the software, model parameters, and training data as described in that study.
 Since the training of the maximum entropy model is inherently stochastic, producing slightly different outcomes on each run, the lowest scoring of ten runs is reported (H\&W:396), though in general there is not a great deal of variation between individual runs.
-One limitation of this model is that it is not feasible to score whole words, as the number of constraints which must be inspected grows exponentially as the size of possible constraints increases.
+One limitation of this model is that it is not feasible to score whole words, as the number of constraints which must be inspected grows exponentially as the span of possible constraints increases.
 Following H\&W and of \citet{Albright2009a}, who also applies the maximum entropy model to the \citet{Albright2003b} norms, the model is trained and scored only on stimulus onsets.
-However, as a consequence, the maximum entropy model performs particularly poorly on this data set, as many stimuli contain phonotactic violations in other positions.
+However, as a consequence, the maximum entropy model performs particularly poorly on this data set, as many stimuli contain phonotactic violations in rime positions.
 
@@
 
-\noindent Bigram models are widely used in natural language processing, and \citet{Albright2009a} considers their relevance to modeling wordlikeness judgements.
+\noindent \citet{Albright2009a} employs bigram probability to model wordlikeness judgements.
 While the focus of \citeauthor{Albright2009a}'s study is on developing a model which uses bigrams over phonological features rather than segments themselves, \citeauthor{Albright2009a}'s evaluation, which includes both the \citeauthor{Scholes1966} and \citeauthor{Albright2003b} data sets, finds an advantage for segmental bigrams.
+\citeauthor{Albright2009a} does not provide an implementation of the featural bigram model, nor does his study describe it in sufficient detail to allow for a new implementation, but segmental bigram scores for the \citeauthor{Albright2003a} data are reported in the appendix.
+As reported by \citeauthor{Albright2009a}, segmental bigrams outperform featural bigrams (see Table \ref{albrightimproved}).
 
@@
 This technique is known as Laplace, or ``add one'' smoothing.
-This has the desirable effect that no nonce word is ever assigned a zero probability, and produces in a small increase in the correlation between the \citeauthor{Albright2003b} wordlikeness norms compared with the maximum likelihood estimate (Table \ref{albrightimproved}).
-For all three data sets, this model also consistenly outperforms positional probability models as defined by \citet{Vitevitch2004} and \citet{Vaden2009}.
-Given that these probabilities are highly correlatd with bigram probabilities \citep[][54]{Vitevitch1997}, they are not considered further.
-The bigram model consistently performs well in all the evaluations, and has the highest Spearman correlation with the \citeauthor{Greenberg1964} and \citeauthor{Scholes1966}, and is frequently second place model to the binary baseline elsewhere.
+This has the desirable effect that no nonce word is ever assigned a zero probability, and produces a small increase in the correlation between the \citeauthor{Albright2003b} wordlikeness norms compared with the maximum likelihood estimate (Table \ref{albrightimproved}).
+For all three data sets, this model also consistenly outperforms positional probability models defined by \citet{Vitevitch2004} and \citet{Vaden2009}.
+Given this and the fact that these model scores are highly correlatd with bigram probability \citep[][54]{Vitevitch1997}, they are not considered further.
+The bigram model consistently performs well in all the evaluations, and has the highest Spearman correlation with the \citeauthor{Greenberg1964} and \citeauthor{Scholes1966} data, and is frequently second place model to the binary baseline elsewhere.
 
 \begin{table} \centering
-\begin{tabular}{l r r r}
+\begin{tabular}{l r r r r}
 \toprule
-                         & maximum likelihood & Laplace smoothing \\
+                         & \multicolumn{1}{c}{featural bigrams} & \multicolumn{2}{c}{segmental bigrams}  \\
+                         & maximum likelihood                   & maximum likelihood & Laplace smoothing \\
 \midrule
-Spearman $\rho$          & 0.660              & \textbf{0.708} \\
-Kendall $\tau_b$         & 0.467              & \textbf{0.506} \\
-Goodman-Kruskal $\gamma$ & 0.473              & \textbf{0.509} \\
+%Pearson $\rho
+Spearman $\rho$          & 0.638                                & 0.660              & \textbf{0.708}    \\
+Kendall $\tau_b$         & 0.457                                & 0.467              & \textbf{0.506}    \\
+Goodman-Kruskal $\gamma$ & 0.462                                & 0.473              & \textbf{0.509}    \\
 \bottomrule
 \end{tabular}
-\caption{Laplace smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
+\caption{Segmental bigrams outperform ``featural bigrams'', and Laplace smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
 All correlations are significant at $p = 0.05$.}
@@
 For instance, the bigram model favors \emph{troog} [tɹuːɡ] over \emph{swach} [swætʃ], though neither contains any gross phonotactic violation.
-Similarly, the bigram model favors \emph{chwoop} [tʃwuːp] over \emph{zhrick} [ʒɹɪk], even though neither word is a possible word of English.
+Similarly, the bigram model favors \emph{chwoop} [tʃwuːp] over \emph{zhrick} [ʒɹɪk], even though both contain ill-formed onsets.
 However, the majority of such predicted contrasts are not reflected in speakers' judgements; for instance, \emph{troog} is rated less English-like than \emph{swach} \citep{Greenberg1964}, contrary to the model predictions.
-This is stark evidence against the gradience hypothesis.
+This shows quite starkly that these models fail to reliably predict intermediate ratings.
 
@@
 A trivial baseline, endowed with few abilities to project beyond the observed data, generally outperforms the state of the art.
-It follows that the projections made by the state-of-the-art gradient models are not like those made by speakers.
-What remains to be seen is whether any model can be put forth which accurately predicts these intermediate ratings, and whether such a model deploys linguistic representations.
+The projections made by the state-of-the-art gradient models are not like those made by speakers.
+It remains to be seen is whether any model can be put forth which accurately predicts these intermediate ratings.
 
@@
 The inability to generalization within onsets and rimes is a serious flaw, as is the assumption of independence of onset and rime.
-A cognitively plausible version of this model must entertain phonotactic generalizations that are larger than these units.
+Regarding the rime, \citet{Borowsky1989} proposes a theory of possible rimes in English, which does make the correct prediction regarding the unattested but well-formed [ɛsp].
+On the other hand, a cognitively plausible version of this model might need to entertain phonotactic generalizations that are larger than these units, since syllable-sized phonotactic generalizations have been proposed for English \citep[e.g.,][]{Berkley1994a,Berkley1994b,Coetzee2008b,Fudge1969}.
 
 A possible further extension to the binary baseline would be the introduction of additional levels of wellformedness.
-While the evaluation has shown that current gradient models do not reliably identify intermediate wellformedness, it does seem possible to identify at least three levels of grammaticality: for instance, one might share the intuition that \emph{zhlick} [ʒlɪk] is more similar to English than \emph{bnick}, though both have unattested onsets.
-
+While the evaluation has shown that current gradient models do not reliably identify intermediate wellformedness, it does seem possible to identify at least three levels of grammaticality: for instance, one might encode the intuition that \emph{zhlick} [ʒlɪk] is more English-like than \emph{bnick}, though both have unattested onsets.
 There are precedents for labeling certain attested words as phonotactically ``peripheral'' (see, e.g., the appendices in \citealt{Myers1987} and \citealt{Borowsky1989}); such words are regarded as lexical exceptions to language-general principles of syllabification.
@@
 Another likely source of additional levels of grammaticality is the cumulative effect of multiple phonotactic violations.
-While, as \citet{Coleman1997} note, classic model predict that a nonce word is as ill-formed as its worst deviation from syllable structure, it is possible to imagine that multiple phonotactic violations would result in greater degrees of ill-formedness.
+While, as \citet{Coleman1997} note, classical Optimality Theory predicts that a nonce word is as ill-formed as its worst deviation from syllable structure, it is possible to imagine that multiple phonotactic violations would result in greater degrees of ill-formedness.
 The bigram and maxent models make this prediction, as do many others \citep[e.g.,][]{Legendre1990,Levelt2000,Albright2008,Anttila2008,Pater2009b} but despite this, there is still little data demonstrating cumulative effects in wordlikeness tasks.
@@
 
-First, it appears that children treat syllables holistically as early as 4 days after birth \citep{Bijeljac-Babic1993} and rapidly learn native-language syllable-based phonotactics \citep{Onishi2002,Chambers2003}, suggesting that any co-occurrence statistics should be delimited on units larger than the segments proposed by \citeauthor{Saffran1996}.
+First, it appears that the youngest treat syllables holistically \citep{Bertoncini1981,Jusczyk1987} and rapidly learn native-language syllable-based phonotactics \citep{Onishi2002,Chambers2003}, suggesting that any co-occurrence statistics should be delimited on units larger than the segments proposed by \citeauthor{Saffran1996}.
 Secondly, a good deal of the literature has focused on segmentation according to near exceptionless constraints that hold between adjacent syllables.

### 2012-10-03 3.tex
Writing on chapters 2, 4
Edits to 2 and 4, and writing in 3

@@
 
-\citet{Chomsky1965} argue that speakers internalize not only generalizations governing phonological alternations in their language, but also about the shapes of possible words in their language.
-It is certainly possible to imagine that speakers would not bother at all.
+While \citet{Chomsky1965} argue that speakers internalize facts about the shapes of possible words in their language, it is certainly possible to imagine that speakers would not bother at all.
 \citet[][320]{Zimmer1969} claims these generalizations ``have no observable consequences in the course of the normal use of the language'' and a similar sentiment is echoed more recently by \citeauthor{PE}:
@@
 
-This is a principled null hypothesis, but one which can be quickly rejected.
-\citeauthor{Chomsky1965} point to robust contrasts between equally unattested items like *\emph{blick}, which is recognized by speakers a possible word of English, and *\emph{bnick}, which is not.
-This preference must be regarded as a learned behavior, since many languages permit  \emph{bn} onsets (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet').
-It is also possible to show speakers access this knowledge not just in wordlikeness judgement tasks but also in naturalistic and quotidien tasks like segmenting spontaneous speech into words, as \citet{McQueen1998} and \citet{Mattys2001b} argue for adult and infant subjects, respectively.
+This represents a principled null hypothesis, but it can be quickly dismissed. 
+This might represent
+
+This would be a principled null hypothesis, but it can be quickly rejected in light of the indisputable contrast between non-words like \emph{blick} and \emph{bnick} noted by \citeauthor{Chomsky1965}.
+It is also possible to show speakers access this knowledge not just in metalinguist tasks like wordlikeness judgements, but also in such naturalistic and quotidien tasks as segmenting spontaneous speech into words \citep{Mattys2001b,McQueen1998}.
 
@@
 In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique, a chi-square test, in an attempt to validate a generalization about Arabic root consonant co-occurrence faces originally due to \citet{Greenberg1950}.
-While this particular generalization was ultimately also evaluated using a wordlikeness task \citep{Frisch2004}, there is now an extensive literature in which phonotactic knowledge is inferred directly from lexical statistics, with no psycholinguistic component of any kind \citep[e.g.,][]{Mester1988,Padgett1992,Padgett1995,Berkley1994b,Berkley1994a,Berkley2000,Buckley1997,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Pozdniakov2007,Anttila2008,Coetzee2008a,Dmitrieva2008a,Dmitrieva2008b,Graff2011}.
-The (usually tacit) assumption of most of this work is made explicit by \citet{Brown2010} in his study of consonant cluster co-occurrence restrictions in Gitksan: any statistically significant pattern in the lexicon is one that is internalized by speakers.
+While this particular generalization was ultimately also evaluated using a wordlikeness task \citep{Frisch2004}, there is now an extensive literature in which phonotactic knowledge is inferred directly from lexical statistics, with no psycholinguistic component of any kind
+\citep[e.g.,][]{Anttila2008,
+Berkley1994b,Berkley1994a,Berkley2000,
+Brown2010,
+Buckley1997,
+Coetzee2008a,
+Dmitrieva2008a,Dmitrieva2008b,
+Elmedlaoui1995,
+Graff2011,
+Padgett1992,Padgett1995,
+MacEachern1999,
+Kinney2005,
+Kawahara2006,
+Martin2007,Martin2011,
+Mester1988,
+Miller-Ockhuizen2003,
+Pozdniakov2007,
+Yip1989}.
+%find amanda miller-ockhuizen on ju|'hoansi
+The (usually tacit) assumption of most of this work is made explicit by \citet{Brown2010} in a study of consonant cluster co-occurrence restrictions in Gitksan: he claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers.
 
@@
 \begin{example}[Turkish vowel features]
-\begin{tabular}{c | c c c c}
+\begin{tabular}{c c c c c}
                        & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
-                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ \midrule
-\buf[$+$\textsc{High}] & \emph{i} & \emph{ü} [y] & \emph{ı} [ɯ] & \emph{u} \\
-\buf[$-$\textsc{High}] & \emph{e} & \emph{ö} [ø] & \emph{a} [ɑ] & \emph{o} \\
+                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ %\midrule 
+\buf[$+$\textsc{High}] & {i} & {ü} [y] & {ı} [ɯ] & {u} \\
+\buf[$-$\textsc{High}] & {e} & {ö} [ø] & {a} [ɑ] & {o} \\
 \end{tabular}
@@
 
+\begin{example}[\textsc{Backness Harmony}]
+$\begin{bmatrix} +\textsc{Vocalic} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~\big /~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~\textrm{C}_0~\gap{}$
+\end{example}
+
 %\citep[after][229]{Clements1982}: \\
-\begin{example}[\textsc{Backness Harmony}]
-\xymatrix@R=24pt@C=24pt{
-\txt{V}                                        & \txt{C$_0$~~~~~} & \txt{V} & \txt{(condition: rightward application)} \\
-\txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
-}
-\end{example}
+%\xymatrix@R=24pt@C=24pt{
+%\txt{V}                                        & \txt{C$_0$~~~~~} & \txt{V} & \txt{(condition: rightward application)} \\
+%\txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
+%}
 
@@
 \begin{example}[The Turkish nominative] \label{turknom}
-\begin{tabular}{l l l l l}
-   & nom.sg.      & nom.pl. \\
-a. & \emph{ip}    & \emph{ipler}    & `rope'        & \citep[][216]{Clements1982} \\
-   & \emph{el}    & \emph{eller}    & `hand'    \\
-   & \emph{köy}   & \emph{köyler}   & `village' \\
-   & \emph{yüz}   & \emph{yüzler}   & `face'    \\
-   & \emph{kız}   & \emph{kızlar}   & `girl'    \\
-   & \emph{sap}   & \emph{saplar}   & `stalk'   \\
-   & \emph{son}   & \emph{sonlar}   & `end'     \\
-   & \emph{pul}   & \emph{pullar}   & `stamp'   \\
-b. & \emph{neden} & \emph{nedenler} & `reason'       & (TELL) \\ % front
-   & \emph{kiler} & \emph{kilerler} & `pantry'       \\ % front
-   & \emph{pelür} & \emph{pelürler} & `tissue paper' \\ % back
-   & \emph{boğaz} & \emph{boğazlar} & `throat'       \\ % back
-   & \emph{sapık} & \emph{sapıklar} & `pervert'      \\ % back
+\begin{tabular}{l l l l@{ }l}
+   & \emph{nom.sg.} & \emph{nom.pl.} \\
+a. & {ip}           & {ipler}    & `rope'        & \citep[][216]{Clements1982} \\
+   & {el}           & {eller}    & `hand'    \\
+   & {köy}          & {köyler}   & `village' \\
+   & {yüz}          & {yüzler}   & `face'    \\
+   & {kız}          & {kızlar}   & `girl'    \\
+   & {sap}          & {saplar}   & `stalk'   \\
+   & {son}          & {sonlar}   & `end'     \\
+   & {pul}          & {pullar}   & `stamp'   \\
+b. & {neden}        & {nedenler} & `reason'       & (TELL) \\ % front
+   & {kiler}        & {kilerler} & `pantry'       \\ % front
+   & {pelür}        & {pelürler} & `tissue paper' \\ % back
+   & {boğaz}        & {boğazlar} & `throat'       \\ % back
+   & {sapık}        & {sapıklar} & `pervert'      \\ % back
 \end{tabular}
@@
 \begin{example}[Exceptional Turkish nominatives] \label{turkexcept}
-\begin{tabular}{l l l l l}
-   & {nom.sg.}    & {nom.pl.} \\
-a. & \emph{mezar} & \emph{mezarlar} & `grave' & \citep{TELL} \\
-   & \emph{model} & \emph{modeller} & `model' \\
-   & \emph{silah} & \emph{silahlar} & `weapon'     \\
-   & \emph{memur} & \emph{memurlar} & `bureaucrat' \\
-   & \emph{sabun} & \emph{sabunlar} & `soap'       \\
-b. & \emph{saat}  & \emph{saatler}  & `hour, clock' \\
-   & \emph{harf}  & \emph{harfler}  & `(alphabetic) letter' \\ %& \citep{Goksel2005}
-   & \emph{etol}  & \emph{etoller}  & `fur stole' \\
+\begin{tabular}{l l l l@{ }l}
+   & \emph{nom.sg.} & \emph{nom.pl.} \\
+a. & {mezar}        & {mezarlar} & `grave' & \citep{TELL} \\
+   & {model}        & {modeller} & `model' \\
+   & {silah}        & {silahlar} & `weapon'     \\
+   & {memur}        & {memurlar} & `bureaucrat' \\
+   & {sabun}        & {sabunlar} & `soap'       \\
+b. & {saat}         & {saatler}  & `hour, clock' \\
+   & {harf}         & {harfler}  & `(alphabetic) letter' \\ %& \citep{Goksel2005}
+   & {etol}         & {etoller}  & `fur stole' \\
 \end{tabular}
@@
 
-\begin{example}[Underspecification in harmonic roots (after \citealp{Clements1982})] \label{spec}
+\begin{example}[Autosegmental underspecification in harmonic roots (after \citealp{Clements1982})] \label{spec}
 \xymatrix@R=24pt@C=24pt{
@@
 \begin{tabular}{l l l l l l}
-a. & \emph{spiker}  & \alt{} & \emph{sipiker}  & `announcer' \\
-   & \emph{fren}    & \alt{} & \emph{firen}    & `break'     \\
-b. & \emph{brom}    & \alt{} & \emph{burom}    & `bromide'   \\
-   & \emph{prusya}  & \alt{} & \emph{purusya}  & `Prussia'   \\
+a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
+   & {fren}    & \alt{} & {firen}    & `break'     \\
+b. & {brom}    & \alt{} & {burom}    & `bromide'   \\
+   & {prusya}  & \alt{} & {purusya}  & `Prussia'   \\
 %b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
@@
 \begin{tabular}{l l l l}
-a. & \emph{kibrit} & \emph{kibrit}-\{\emph{kabrıt}\} & `match'    \\
-   & \emph{bütün}  & \emph{bütün}-\{\emph{batın}\}   & `whole'    \\
-b. & \emph{mali}   & \emph{mali}-\{\emph{muli}\}     & `Mali'     \\
-   & \emph{butik}  & \emph{butik}-\{\emph{batik}\}   & `boutique' \\
+a. & {kibrit} & {kibrit}-\{{kabrıt}\} & `match'    \\
+   & {bütün}  & {bütün}-\{{batın}\}   & `whole'    \\
+b. & {mali}   & {mali}-\{{muli}\}     & `Mali'     \\
+   & {butik}  & {butik}-\{{batik}\}   & `boutique' \\
 \end{tabular}
@@
 
-\begin{example}[\textsc{Roundness Harmony}]
-\xymatrix@R=24pt@C=24pt{
-\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{--}[drr] &             & \txt{[$+$\textsc{High}]} \\
-\txt{V}                                         & \txt{C$_0$} & \txt{V}\ar@{-}[u] \\
-}
-\end{example}
+%\begin{example}[\textsc{Roundness Harmony}]
+%\xymatrix@R=24pt@C=24pt{
+%\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{--}[drr] &             & \txt{[$+$\textsc{High}]} \\
+%\txt{V}                                         & \txt{C$_0$} & \txt{V}\ar@{-}[u] \\
+%}
+%\end{example}
 
@@
 \begin{example}[Turkish nominal suffix allomorphy]
-\begin{tabular}{l l l l l l l}
-   & {nom.sg.}    & {dat.sg.}     & {gen.sg.}  \\
-a. & \emph{ip}    & \emph{ipi}    & \emph{ipin}    & `rope' & \citep[][216]{Clements1982} \\
-   & \emph{el}    & \emph{eli}    & \emph{elin}    & `hand'    \\
-   & \emph{kız}   & \emph{kızı}   & \emph{kızın}   & `girl'    \\
-   & \emph{sap}   & \emph{sapı}   & \emph{sapın}   & `stalk'   \\
-   & \emph{yüz}   & \emph{yüzü}   & \emph{yüzün}   & `face'    \\
-   & \emph{köy}   & \emph{köyü}   & \emph{köyün}   & `village' \\
-   & \emph{pul}   & \emph{pulu}   & \emph{pulun}   & `stamp'   \\
-   & \emph{son}   & \emph{sonu}   & \emph{sonun}   & `end'     \\
-b. & \emph{boğaz} & \emph{boğazı} & \emph{boğazın} & `throat'  & (TELL) \\
-   & \emph{pelür} & \emph{pelürü} & \emph{pelürün} & `tissue paper' \\
-   & \emph{döviz} & \emph{dövizi} & \emph{dövizin} & `currency' \\
-   & \emph{yamuk} & \emph{yamuğu} & \emph{yamuğun} & `trapezoid' \\
-   & \emph{ümit}  & \emph{ümiti}  & \emph{ümitin}  & `hope'     \\
+\begin{tabular}{l l l l l l@{ }l}
+   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
+a. & {ip}           & {ipi}          & {ipin}         & `rope' & \citep[][216]{Clements1982} \\
+   & {el}           & {eli}          & {elin}         & `hand'    \\
+   & {kız}          & {kızı}         & {kızın}        & `girl'    \\
+   & {sap}          & {sapı}         & {sapın}        & `stalk'   \\
+   & {yüz}          & {yüzü}         & {yüzün}        & `face'    \\
+   & {köy}          & {köyü}         & {köyün}        & `village' \\
+   & {pul}          & {pulu}         & {pulun}        & `stamp'   \\
+   & {son}          & {sonu}         & {sonun}        & `end'     \\
+b. & {boğaz}        & {boğazı}       & {boğazın}      & `throat'  & (TELL) \\
+   & {pelür}        & {pelürü}       & {pelürün}      & `tissue paper' \\
+   & {döviz}        & {dövizi}       & {dövizin}      & `currency' \\
+   & {yamuk}        & {yamuğu}       & {yamuğun}      & `trapezoid' \\
+   & {ümit}         & {ümiti}        & {ümitin}       & `hope'     \\
 \end{tabular}
@@
 
-\begin{example}[\textsc{Labial Attraction}]
-%(after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}): 
-\xymatrix@R=24pt@C=8pt{
-\txt{[$-$\textsc{Round}]} &                                         & \txt{[$-$\textsc{High}]} & \txt{[$+$\textsc{Labial}]}    & \txt{[$+$\textsc{High}]}\ar@{-}[dr] &         & \txt{[$+$\textsc{Round}]}\ar@{--}[dl] \\
-%                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C}\ar@{-}[u] &                                      & \txt{V} & \\
-                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C$_0$ C C$_0$}\ar@{-}[u] &                                      & \txt{V} & \\
-                         & \txt{[$+$\textsc{Back}]}\ar@{-}[urrrr]
-}
-\end{example}
+%\begin{example}[\textsc{Labial Attraction}]
+%%(after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}): 
+%\xymatrix@R=24pt@C=8pt{
+%\txt{[$-$\textsc{Round}]} &                                         & \txt{[$-$\textsc{High}]} & \txt{[$+$\textsc{Labial}]}    & \txt{[$+$\textsc{High}]}\ar@{-}[dr] &         & \txt{[$+$\textsc{Round}]}\ar@{--}[dl] \\
+%%                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C}\ar@{-}[u] &                                      & \txt{V} & \\
+%                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C$_0$ C C$_0$}\ar@{-}[u] &                                      & \txt{V} & \\
+%                         & \txt{[$+$\textsc{Back}]}\ar@{-}[urrrr]
+%}
+%\end{example}
 
@@
 \midrule
-\emph{pemez} & 30            & \emph{pemaz} & 2  \\
-\emph{tipez} & 24            & \emph{tipaz} & 8  \\ 
-\emph{terüz} & 19            & \emph{teruz} & 13 \\ % roundness violator
-\emph{teriz} & 28            & \emph{terız} & 3  \\
-\emph{tokaz} & 26            & \emph{tokez} & 6  \\ % roundness violator
+{pemez} & 30            & {pemaz} & 2  \\
+{tipez} & 24            & {tipaz} & 8  \\ 
+{terüz} & 19            & {teruz} & 13 \\ % roundness violator
+{teriz} & 28            & {terız} & 3  \\
+{tokaz} & 26            & {tokez} & 6  \\ % roundness violator
 \bottomrule
@@
 \midrule
-\emph{pörüz} & 32            & \emph{pöriz} & 0  \\
-\emph{tatız} & 20            & \emph{tatuz} & 12 \\
-\emph{tüpüz} & 31            & \emph{tüpiz} & 1  \\
-\emph{takız} & 22            & \emph{takuz} & 10 \\
+{pörüz} & 32            & {pöriz} & 0  \\
+{tatız} & 20            & {tatuz} & 12 \\
+{tüpüz} & 31            & {tüpiz} & 1  \\
+{takız} & 22            & {takuz} & 10 \\
 \bottomrule
@@
 \toprule
-\multicolumn{2}{l}{\emph{a}P\emph{u}} & \multicolumn{2}{l}{\emph{a}P\emph{ı}} \\
+\multicolumn{2}{l}{{a}P{u}} & \multicolumn{2}{l}{{a}P{ı}} \\
 \midrule
-\emph{pamuz} & 17       & \emph{pamız} & 15 \\
-\emph{tafuz} & 21       & \emph{tafız} & 11 \\
-\emph{tapuz} & 17       & \emph{tapız} & 15 \\
-\emph{mavuz} & 16       & \emph{mavız} & 16 \\
-\emph{tabuz} & 16       & \emph{tabız} & 16 \\
+{pamuz} & 17       & {pamız} & 15 \\
+{tafuz} & 21       & {tafız} & 11 \\
+{tapuz} & 17       & {tapız} & 15 \\
+{mavuz} & 16       & {mavız} & 16 \\
+{tabuz} & 16       & {tabız} & 16 \\
 \bottomrule
@@
 \begin{tabular}{l l l l}
-\hspace{0pt}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
-\hspace{0pt}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
-\hspace{0pt}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
+\buf{}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
+\buf{}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
+\buf{}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
 \end{tabular}
@@
 \begin{tabular}{l r r r}
-        & \emph{a}P\emph{u} & \emph{a}P\emph{ı} & $p$-value \\
-Native  & 12                & 11                & \multirow{2}{*}{0.042} \\
-Foreign & 84                & 28                \\
+        & {a}P{u} & {a}P{ı} & $p$-value \\
+native  & 12                & 11                & \multirow{2}{*}{0.042} \\
+foreign & 84                & 28                \\
 \end{tabular}

### 2012-10-03 4.tex
Edits to 2 and 4, and writing in 3

@@
 Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1979}, \citealp[410f.]{Stanley1967}).
-Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the \emph{communis opinio}
-(including many of the authors cited above) holds that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
+Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the \emph{communis opinio} (including many of the authors cited above) holds that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
 
@@
 
-\paragraph{Affricates} If the affricates [tʃ, dʒ] are treated as stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990}, the previous principle will incorrectly assign the two segments of the affricate to separate syllables before stressed lax vowels (e.g., *\emph{ra}[t.ʃ]\emph{et} or \emph{a}[d.ʒ]\emph{ile}) unless some ad hoc constraint \citep[e.g.,][]{Wells1990} is in place.
+\paragraph{Affricates} If the ffricates [tʃ, dʒ] are treated as stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990}, the previous principle will incorrectly assign the two segments of the affricate to separate syllables before stressed lax vowels (e.g., *\emph{ra}[t.ʃ]\emph{et} or \emph{a}[d.ʒ]\emph{ile}) unless some ad hoc constraint \citep[e.g.,][]{Wells1990} is in place.
 A more traditional analysis of affricates, in which they are simple segments distinguished from pure stops by stridency \citep[24]{Jakobson1961} or delayed release (\emph{SPE}:321f.) is assumed here, following \citeauthor{Pierrehumbert1994}.
 This assumption can be motivated by the tendency of affricates to pattern with individual segments in phonotactic generalizations.
-For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but bans all other onset clusters \citep[9]{Launey2011}.
+For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but bans true onset clusters \citep[9]{Launey2011}.
 %\citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.
@@
 
-\paragraph{[j] onglides} While some have argued that the front onglide in words such as \emph{val}[j]\emph{ue} is inserted by rule (\emph{SPE}:196, \citealp[][89]{Halle1985a}, \citealp[][217]{McMahon1990}), the very presence or absence of the glide is contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}) indicating that it is present in underlying representation (\citealp{Anderson1988b}, \citealp[278]{Borowsky1986}).
-The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).
+\paragraph{[j] onglides} While some have argued that the front onglide in words such as \emph{val}[j]\emph{ue} is inserted by rule (\emph{SPE}:196, \citealp[][89]{Halle1985a}, \citealp[][217]{McMahon1990}), the very presence or absence of the glide is contrastive (e.g., \emph{coot} \alt{} \emph{cute}, \emph{booty} \alt{} \emph{beauty}) indicating that it is present in underlying representation (\citealp{Anderson1988b}, \citealp[278]{Borowsky1986}).
+The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{ju}[n.j]\emph{or}).
 There is considerable evidence for this assumption.
@@
 
-\paragraph{Post-vocalic \emph{r}} Word-medial post-vocalic \emph{r} has been lost in Received Pronunciation, the standard accent of the CELEX transcriptions.
-Even in \emph{r}-ful dialects, though, there is reason to believe that post-vocalic \emph{r} is fact nuclear, and thus irrelevant to syllable contact clusters.
+\paragraph{Syllabic \emph{r}} Word-medial syllabic \emph{r} has been lost in Received Pronunciation, the accent used for the CELEX transcriptions.
+Even in \emph{r}-ful dialects, though, there is reason to believe that syllabic \emph{r} is fact nuclear, and thus irrelevant to syllable contact clusters.
 Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}): compare American English \emph{fern}, \emph{fir}, \emph{fur} to \emph{pet}, \emph{pit}, \emph{putt}.
-Post-vocalic \emph{r} patterns with vowels and not other consonants for several phonological processes: it is the only consonant which does not block variable glottalization of following /t/ in British dialects \citep[258]{Harris1994}.
+Syllabic \emph{r} patterns with vowels and not other consonants for several phonological processes: it is the only consonant which does not block variable glottalization of following /t/ in British dialects \citep[258]{Harris1994}.
 
@@
 
-\noindent Post-vocalic \emph{r} is the only consonant which does not trigger variable deletion of a following /t, d/ in American dialects \citep[8]{Guy1980}.
+\noindent Syllabic \emph{r} is the only consonant which does not trigger variable deletion of a following /t, d/ in American dialects \citep[8]{Guy1980}.
 
@@
 \txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
-\txt{[$+$\textsc{Obstruent}]}                  & \txt{[$+$\textsc{Obstruent}]} \\
+\txt{[$-$\textsc{Sonorant}]}                  & \txt{[$-$\textsc{Sonorant}]} \\
 }
@@
 A final alternation targeting English syllable contact clusters is the simplification of geminates derived by certain affixes.
-For instance, the deadjectival adverbial suffix \emph{-ly} rarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{norma}[l]\emph{y}, cf.
-\emph{calm}[l]\emph{y}).
+For instance, the deadjectival adverbial suffix \emph{-ly} rarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{fu}[l]\emph{y}, cf.
+\emph{free}[l]\emph{y}).
 Degemination is also found in /d/-final roots selecting the irregular /-t/ past tense suffixes (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).
-More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}).
-Identical adjacent consonants do not occur in the corpus, nor do those which differ only in voice, such as *[p.b], which would be made identical by the operation of \textsc{Obstruent Voice Assimilation}.
-This absence is unlikely to be due to chance.
+More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (see \emph{SPE}:148 and \citealt[102]{Borowsky1986}).
+Identical adjacent consonants do not occur in the corpus, nor do obstruent clusters which differ only in voice (e.g., *[p.b]), which would be made identical by the operation of \textsc{Obstruent Voice Assimilation}.
+As shown by the results of the Fisher test, this absence is unlikely to be due to chance.
 

### 2012-10-10 1.tex
Writing on front matter, chapters 1-4

@@
-\chapter{A generative theory of phonotactics and phonotactic learning}
+\chapter{A generative theory of phonotactic knowledge and acquisition}
+\label{intro}
+
+Syntax can hardly be described as a theory of word ``arrangement''; it would be inconceivable for syntacticians to engage in a lengthy debate about whether some word sequence is permitted in some language.
+Yet it is possible for two reasonable phonologists to dispute whether [k.p] is a permissible medial consonant cluster in English.
+While lexical entries are thought to consist of segments, and segments to consist of feature specifications, the phonological form of a word like \emph{spectre} is not ``generated'', in any relevant sense, by the concatenation of segments /s/, /p/, and so on, or by the concatenation of syllables \{spɛk\}$_{\sigma}$ and \{tɚ\}$_{\sigma}$.
+It would be possible to posit ``lexical concatenation rules'', on analogy with the phrase structure rules of syntax, but it is unclear how the strings produced would be linked with meanings, or how such a system would distinguish between, e.g., \emph{brick} and meaningless \emph{blick}.
+
+The more appealing alternative is that the phonological form of \emph{spectre}, whatever it is, is an atom of unit of lexical memory.
+It remains to account for any restrictions on the contents of underlying forms.
+The theory of \emph{phonotactics} (which is related to \emph{syntax} through the Greek root \emph{táxis} `order') is concerned with these restrictions and how they are related to the mappings between underlying and surface forms for which the phonological component is responsible.
+
+%\footnote{It is tempting, however, to describe binding principles as ``syntactics''.}
+
+I believe that the boundaries of this knowledge
+
+The boundaries of this knowledge
+and what might be acquired.
+These are not unrelated issues:
 
 \section{\emph{Táxis} and generativity}
+%\emph{τάξις}
 
-%\emph{τάξις}
-(\emph{táxis} `order, arrangement; the prefix \emph{syn-} means `together')
+(\emph{táxis} `order, arrangement'; the prefix \emph{syn-} means `together')
+
+This etymology is almost completley uninformative about the modern study of syntax. 
+The notion of derivation, most associated with the transformational grammars of Zellig Harris and Noam Chomsky
 
 This etymology gives little clue to just what syntacticians do, however.
-This is not merely a consequence of the ``generative revolution'', as the predecessors of transformational grammars associated with Zellig Harris and Noam Chomsky were also more than theories of arrangement, but also instantiate the notion of derivation \citep[e.g.,][]{Post1943}.
+This is not merely a consequence of the ``generative revolution'', as the predecessors of transformational grammars associated with Zellig Harris and Noam Chomsky were also more than theories of arrangement, but also instantiate the notion of derivation.
 
-\subsection{?}
+\subsection{Units of phonotactic description}
 
-\subsection{Units of description}
-\subsubsection{Segment inventory}
-\subsubsection{Surface segment sequences}
-\subsubsection{Underlying segment sequences}
+\subsubsection{Segments}
 
-\subsection{The morph}
+\subsubsection{Underlying representations}
+
+\subsubsection{Surface sequences}
+
+% discussion of Ernestus & Baayen 2003, Becker et al. 2011
 
@@
 One of the most famous examples is the distribution of \emph{ich-laut} [ç] and \emph{ach-laut} [x] in German \citep{Bloomfield1930}.
-As shown in (\label{ichach}), [x] appears only before back vowels; in all other contexts, the only dorsal fricative is [ç].
+The dorsal fricative is [x] before back vowels and[ç] in other contexts.
 
@@
 \begin{tabular}{l l l}
-a. & \emph{B}[uːx]          & `book'           \\
-   & \emph{T}[ɔx]\emph{ter} & `daughter'       \\
-   & \emph{N}[ax]\emph{t}   & `night'          \\
-b. & \emph{S}[iç]\emph{t}   & `view'           \\
-   & \emph{Sp}[eç]\emph{t}  & `woodpecker'     \\
-   & \emph{Ger}[yç]\emph{t} & `rumor'          \\
-   & \emph{Kn}[øç]\emph{el} & `ankle, knuckle' \\
-   & \emph{Fl}[ɛç]\emph{e}  & `surface'        \\
+a. & [buːx]   & `book'           \\
+   & [tɔxtər] & `daughter'       \\
+   & [naxt]   & `night'          \\
+b. & [siçt]   & `view'           \\
+   & [ʃpeçt]  & `woodpecker'     \\
+   & [ɡərʏçt] & `rumor'          \\
+   & [knøçəl] & `ankle, knuckle' \\
+   & [flɛçə]  & `surface'        \\
 \end{tabular}
@@
 
-%a. & \emph{B}[uːx]          & \emph{B}[yːç]\emph{er} & `book' \\
-%   & \emph{L}[ɔx]           & \emph{L}[øç]\emph{er}  & `hole' \\
-%%   &\emph{T}[ɔx]\emph{ter} & \emph{T}[øç]\emph{ter}
-%   & \emph{N}[ax]\emph{t}   & \emph{N}[ɛç]\emph{te} & `night' \\
+\emph{Umlaut}, the fronting (and raising) of back vowels in certain morphological contexts, produces the front variant of the dorsal fricative; e.g., [lɔx]-[løçər] `hole-holes'.
+%\emph{B}[uːx]-\emph{B}[yːç]\emph{er} `book-books',
 
@@
 \footnote{Examples like \emph{Mas}[oːç]\emph{ist} `masochist', \emph{Eun}[uːç]\emph{ismus} `eunichism', first noted by \citet{Merchant1994}, suggest that this should also be restricted to assimilation within the same foot \citep[226f.]{Jensen2000}.}
-
 %\emph{Mas}[oːx]-\emph{Mas}[oːç]\emph{ist}
-
-The concept of non-derived environment blocking
 
@@
 
-\subsection{Prosodic parsing}
-\subsubsection{Syllabification}
-\subsubsection{Autosegmental tiers}
+\subsection{Syllabification}
 
-\subsection{Rule application}
-\subsubsection{Violation detection}
-\subsubsection{Rule exceptionality}
+\subsection{Rule application and occultation}
 
-\section{Consequences of the model}
+\subsection{Detecting violations}
+
+\subsection{Exceptionality}
+% autosegments
+
+\section{Predictions of the model}
 
@@
 \subsubsection{Probabilistic well-formedness}
-\subsubsection{Cumulativity}
 \subsubsection{External factors}
@@
 \subsection{Static phonotactics}
-\subsubsection{Statistical insensitivity}
+\subsubsection{Insensitivity to statistics}
 \subsubsection{Saussurean arbitrariness}
@@
 \subsection{Order of acquisition}
+
+A point of departure is a review of phonotactic acquisition by \citet{Hayes2004b}.
+
+\emph{pure phonotactic learner}
+%...
+However, \citeauthor{Hayes2004b} ignores the considerable evidence that infants younger than nine months of age---the \emph{terminus ante quem} for the onset of phonotactic knowledge---already have relatively rich and detailed lexical knowledge
+
+A word of caution is in order.
+An experiment which finds typically-developing infants of a certain age insensitive to an adult-like contrast using a certain task is by no means sufficient evidence to conclude that infants of this age lack this ability.
+A negative result may simply indicate that the task lacks ecological validity or requires domain-general cognitive resources beyond those of the infants.
+For instance, \citet{Werker2002} report that 14-month-old infants are insensitive to small phonological differences between novel words, but \citet{Fennell2006} shows that they are sensitive to the same contrasts when the novel words are situated in a more naturalistic naming scenario and cognitive demands are minimized.
+Other negative findings might be attributed to failure to control for properties of the stimulus, or to an experimental design which lacks sufficient statistical power.
+
+\citet{Jusczyk1993b} native vs. non-native
+\citet{Jusczyk1994} no preference for high prob
+\citet{Friederici1993}
+
+Finally, \citet{Mattys1999} find that 9-month-old infants treat hetero-organic nasal-stop clusters in nonce words as indications of word juncture, though this effect is small compared to the effect of other cues like primary stress: they do not test younger infants.
+
+\subsubsection{Prosodic parsing}
+
+Newborn infants are already able to distinguish between monosyllabic and bisyllabic words, but these infants are thought to recognize syllables ``holistically'' rather than as sequences at least until 4 months \citep[e.g.,]{Bertoncini1981,Eimas1999,Jusczyk1987}.
+The earliest evidence for segmental representations comes not from phonotactic preferences, but from infants' ability to dicriminate between familiar words like \emph{cup} and mispronunciations like \emph{*tup} at 7.5 months of age \citep{Jusczyk1995}.
+
+\subsubsection{Phonological processes}
+
+There are very few investigating infants' knowledge of phonological alternations.
+\citet{White2008} expose 8.5-month-old and 10-month-old infants to an artificial language in which the voicing of fricatives is contrastive, but voiced and voiceless variants of stops are in a complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively.
+After familiarization, infants at both ages prefer to listen to stimuli which preserve the complementary distribution over those which disrupt it (e.g., \emph{na poli}, \emph{rot boli}).
+As \citeauthor{White2008} note, this suggests that infants have grouped \emph{pevi} and \emph{bevi} together and extracted the stop voicing alternation from the artifical language.
+
 \subsubsection{Lexical acquisition}
-\subsubsection{Prosodic parsing}
-\subsubsection{Rule application}
 
-%\subsubsection{Exceptionality}
+As first observed by \citet{Darwin1877}, the first words infants recognize are names, their own and those of caretakers.
+Recent studies suggest that infants learn these names as early as 4 months of age \citep{Bortfeld2005,Mandel1995,Tincoff1999}.
+Infants as young as 6 months of age have learned the meaning of familiar words \citep{Bergelson2012}.
+By 8 months of age, infants are able to locate familiar and novel words in utterances \citep{Jusczyk1997,Seidl2006}.
+\section{Outline of the dissertation}
+
+The remainder of the dissertation consists of three case studies which provide support for the novel and contentious predictions of the minimal phonotactic model sketched above
+
+Chapter \ref{gradience} considers evidence from wordlikeness rating tasks.
+It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
+A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' gradient wordlikeness judgements as well as state-of-the-art gradient wellformedness models.
+The gradient models do not contribute any additional coverage once categorical wellformedness is accounted for.
+
+Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns.
+It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.
+%Some arguments against Becker et al.
+
+Chapter \ref{clusters} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English.
+Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound.
+On the other hand, phonological alternations impose robust restrictions on the cluster inventory.
+The remaining gaps in the cluster inventory are due to the sparse nature of the lexicon, not static phonotactic restrictions.
+
+%% NOTES HERE
+%a. & \emph{B}[uːx]          & \emph{B}[yːç]\emph{er} & `book(s)' \\
+%   & \emph{L}[ɔx]           & \emph{L}[øç]\emph{er}  & `hole(s)' \\
+%   &\emph{T}[ɔx]\emph{ter} & \emph{T}[øç]\emph{ter}
+%   & \emph{N}[ax]\emph{t}   & \emph{N}[ɛç]\emph{te} & `night' \\
+%
+% While the bulk of evidence for detailed phonological entries is found in infants over a year of age \citep[e.g.,]{Werker2002,Fennell2003,Fennell2006,Stager1997,Swingley2000,Zamuner2006},
+%9-month-old infants already have learned the predominant stress pattern of their language \citep{Jusczyk1993a}.

### 2012-10-10 2.tex
Writing on front matter, chapters 1-4

@@
-\chapter{Gradient and categorical aspects of wordlikeness judgements} \label{gradience}
+\chapter{Gradient and categorical aspects of wordlikeness judgements} 
+\label{gradience}
+
+%\citet{Saporta1958}
+%\citet{Saporta1957}
 
@@
 In English, [bl] is a permissible onset, but unlike some languages (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet'), [bn] is not.
-\citet[][19f.]{Wolf2009} suggest, for instance, that an underlying /bnɪk/ would be realized as [nɪk]; it is also possible to imagine resolution by prothesis---[əbnɪk]---or anapytxis---[bənɪk], and indeed \citet{Davidson2006b} finds that English speakers use all three of these repairs when asked to mimic foreign pronunciations of obstruent-nasal onsets like /bn/.
-
-However, many linguists now regard the above as overly simplistic, for, as \citeauthor{Shademan2006} argues, such accounts draw a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report ``wordlikeness'' intuitions.
+\citet[][19f.]{Wolf2009} suggest, for instance, that an underlying /bnɪk/ would be realized as [nɪk]; it is also possible to imagine resolution by prothesis---[əbnɪk]---or anapytxis---[bənɪk], and indeed \citet{Davidson2006b} finds that English speakers use all three of these repairs when asked to mimic foreign pronunciations of obstruent-nasal onsets like /bn/.\footnote{However, \citet{Davidson2005,Davidson2006a} observes an acoustic and articulatory distinction between ``exscrescent'' echwa found in non-native cluster production and ``lexical'' schwa: the anaptyctic [z${^\textrm{ə}}$k] produced by English speakers attempting to mimic word-initial [zk] is more similar to \emph{scum} than to \emph{succumb}.
+This may undermine any link between non-native realizations and epenthesis found elsewhere in English (e.g., allomorphs of the regular past tense and noun plural).} However, many linguists now regard the above as overly simplistic, for, as \citeauthor{Shademan2006} argues, such accounts draw a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report ``wordlikeness'' intuitions.
 
@@
 
-\section{Background} \label{2background}
+\section{Background}
+\label{2background}
 
-\subsection{A brief history of the gradience hypothesis} \label{history}
+\subsection{A brief history of the gradience hypothesis} 
+\label{history}
 
@@
 
-Gradient grammaticality is assumed in the earliest work on generative syntax, \emph{The logical structure of linguistic theory}, in which \citet[132]{LSLT} writes that ``there is little doubt that speakers can fairly consistently order new utterances, never previously heard, with respect to their degree of `belongingness' to the language''.
-Later, \citet{ASPECTS} proposes that different syntactic violations give rise to different degrees of ungrammaticality.
+Gradient grammaticality was a tenet of the earliest work on generative syntax.
+For instance, \citet[132]{LSLT} writes that ``there is little doubt that speakers can fairly consistently order new utterances, never previously heard, with respect to their degree of `belongingness' to the language''.
+\citet{ASPECTS} proposes that different syntactic violations give rise to different degrees of ungrammaticality.
 Unlike \citeauthor{Fischer-Jorgensen1952}, however, for \citeauthor{LSLT} this continuum does incorporate a hard-and-fast boundary between the grammatical and ungrammatical: there are no constrasts between grammatical utterances, merely between different degrees of ungrammaticality.
@@
 
-\citet{Halle1962} and \citet{Stanley1967} concern themselves only with categorical wordlikeness contrasts, however.
+\citet{SPR,Halle1962} and \citet{Stanley1967} concern themselves only with categorical wordlikeness contrasts, however.
 Both authors derive wordlikeness effects from morphophonemic redundancy rules acting on underlying representations.
@@
 
-\subsubsection{\emph{The sound pattern of English}} \label{2spe}
+\subsubsection{\emph{The sound pattern of English}} 
+\label{2spe}
 
@@
 
-\begin{example}[German final devoicing] \label{fd}
+\begin{example}[German final devoicing] 
+\label{fd}
 \begin{tabular}{l l l l}
@@
 
-Recent critiques of this syllabification-based model focus on the existence of intermediate wordlikeness ratings \citep[see also][]{Coleman1997,Anttila2008}.
+Recent critiques of this syllabification-based model focus on the existence of intermediate wordlikeness ratings \citep[see also][]{Coleman1997,Anttila2008a}.
 
@@
 
-\begin{figure} \centering
+\begin{figure} 
+\centering
 \includegraphics{agg.pdf}
@@
 
-\section{Evaluation} \label{2evaluation}
+\section{Evaluation} 
+\label{2evaluation}
 
@@
 \citet{Albright2003b} gathered wordlikeness judgements to serve as norms for a \emph{wug}-test.
-87 items were presented to 20 undergraduate subjects, who rated each word on a seven-point Likert scale.
+87 items were presented to 24 undergraduate subjects, who rated each word on a seven-point Likert scale.
 The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``complete normal, would make a fine English word''.
+
+\begin{table}
+\label{counts}
+\centering
+\begin{tabular}{l r r r}
+\toprule
+        & subjects & items & observations \\
+\midrule
+Scholes & 33       & 66    & 2,079        \\
+A\&H    & 24       & 88    & 2,112        \\
+\bottomrule
+\end{tabular}
+\caption{Idk.}
+\end{table}
 
@@
 
-\begin{figure} \centering
+\begin{figure} 
+\centering
 \includegraphics{density.pdf}
@@
 
-\begin{table} \centering
+\begin{table} 
+\centering
 \begin{tabular}{l l l l l}
@@
 
-\subsubsection{Segment bigram probability} \label{bigram}
+\subsubsection{Segment bigram probability} 
+\label{bigram}
 
@@
 
-\begin{table} \centering
+\begin{table} 
+\centering
 \begin{tabular}{l r r r r}
@@
 
-\subsubsection{Neighborhood density} \label{density}
+\subsubsection{Neighborhood density} 
+\label{density}
 
@@
 
-\begin{table} \centering
+\begin{table} 
+\centering
 \begin{tabular}{l r r r}
@@
 While, as \citet{Coleman1997} note, classical Optimality Theory predicts that a nonce word is as ill-formed as its worst deviation from syllable structure, it is possible to imagine that multiple phonotactic violations would result in greater degrees of ill-formedness.
-The bigram and maxent models make this prediction, as do many others \citep[e.g.,][]{Legendre1990,Levelt2000,Albright2008,Anttila2008,Pater2009b} but despite this, there is still little data demonstrating cumulative effects in wordlikeness tasks.
+The bigram and maxent models make this prediction, as do many others \citep[e.g.,][]{Legendre1990,Levelt2000,Albright2008,Anttila2008a,Pater2009b} but despite this, there is still little data demonstrating cumulative effects in wordlikeness tasks.
 
@@
 
-The findings here also have interesting rammifications for language acquisition.
-\citet{Saffran1996} famously argues that infants use statistical dependencies between segments to extract words from continuous speech.
-Among those who ascribe to the Continuity Hypothesis \citep[e.g.,][]{Macnamara1982,Pinker1984,Crain1991,Carey1995,deVilliers2001,Legate2007}, which holds that children acquiring language deploy the same types of linguistic representations as adults, evidence for gradient phonotactic knowledge in infants is evidence for this knowledge in adults, and vis versa.
-The results here then stand to weaken the hypothesis of \citeauthor{Saffran1996}, as well as the hypothesis that infants acquire a system of phonotactics (gradient or otherwise) before learning of alternations and lexical entries that is a direct prediction of algorithms for learning Optimality Theory grammars \citep{Tesar2004,Hayes2004b}.
+The weak empirical status of gradient phonotactic knowledge as reflected in adults has rammifications for language acquisition under the hypothesis that that infants acquiring language deploy the same representations as adults \citep[e.g.,][]{Macnamara1982,Pinker1984,Crain1991,Carey1995,deVilliers2001,Legate2007}.
+Gradient wordlikeness judgements in adults would provide support for claims that infants recognize statistical(inherently gradient) dependencies between segments \citep{Jusczyk1994} and use these to segment words \citep{Saffran1996}.
+An emerging consensus suggests, however, that infants attend to transitional probabilities primarily in the absence of grammatical cues \citep{Gambell2005,Hohne1994,Johnson2001,Jusczyk1999c,Lignos2012,Mattys2001a,Shukla2007,Lew-Williams2012}.
+The vacuous nature of current evidence for gradient phonotactic knowledge in adults further weakens any hypothesis that would link statistical learning in infants to adults' behaviors.
 
-%\footnote{The hypothesis that infants acquire a system of phonotactics (gradient or otherwise) before learning of alternations and lexical entries has some traction among formal linguists, for \citet{Hayes2004b} and \citet{Tesar2004} note that is a direct prediction of algorithms for learning Optimality Theory grammars.
-However, there is extensive evidence that infants who are younger than 9 months, the age at which \citet{Jusczyk1994} argue that the system probabilistic phonotactics comes online, have been demonstrated have already begun extracting words \citep[e.g.,][]{Darwin1877,Mandel1995,Jusczyk1995,Tincoff1999,Jusczyk1999b,Brent2001,Bortfeld2005,Seidl2006,Shukla2011,Bergelson2012} and alternations \citep{White2008} from multiword utterances they encounter.} Each of the tenets of this argument has been subject to serious critique, however.
-
-First, it appears that the youngest treat syllables holistically \citep{Bertoncini1981,Jusczyk1987} and rapidly learn native-language syllable-based phonotactics \citep{Onishi2002,Chambers2003}, suggesting that any co-occurrence statistics should be delimited on units larger than the segments proposed by \citeauthor{Saffran1996}.
-Secondly, a good deal of the literature has focused on segmentation according to near exceptionless constraints that hold between adjacent syllables.
-For instance, infants as young as 4.5 months seem to be aware that English nasal codas agree in place with following obstruents \citep{Mattys2001b,Jusczyk2002,Davidson2004}.
-It might be the case that syllable co-occurrence statistics might be little more than a reflection of infants' learning of categorical ``lexical viability'' constraints \citep{Johnson2003} of the sort also seen in adult speech processing \citep[e.g.,][]{Norris1997}.
-Further, the effect of these statistical dependencies appear to be filtered by expectations about word stress \citep{Johnson2001,Mattys2001a,Shukla2007}, word length \citep{Lew-Williams2012}, and allophonic generalizations \citep{Hohne1994,Jusczyk1999c}.
-Finally, and perhaps most importantly, \citet{Gambell2005} find that a word segmentation strategy based only on statistical dependencies performs very poorly on realistic data.
-The emerging view of the role on statistics seems to be that infants attend to transitional probabilities or an analogue thereof only when they are faced with artificial stimuli with few of the properties of natural language \citep[e.g.,][]{Shukla2007,Lew-Williams2011}.
-The argument from Continuity is weak, in either direction.
+%For instance, infants as young as 4.5 months seem to be aware that English nasal codas agree in place with following obstruents \citep{Mattys2001b,Jusczyk2002,Davidson2004}. 
+%It might be the case that syllable co-occurrence statistics might be little more than a reflection of infants' learning of categorical ``lexical viability'' constraints \citep{Johnson2003} of the sort also seen in adult speech processing \citep[e.g.,][]{Norris1997}.

### 2012-10-10 3.tex
Writing on front matter, chapters 1-4

@@
 \chapter{Static and derived phonotactic preferences in Turkish}
+\label{turkish}
 
-While \citet{Chomsky1965} argue that speakers internalize facts about the shapes of possible words in their language, it is certainly possible to imagine that speakers would not bother at all.
-\citet[][320]{Zimmer1969} claims these generalizations ``have no observable consequences in the course of the normal use of the language'' and a similar sentiment is echoed more recently by \citeauthor{PE}:
+\citet{Chomsky1965} argue that speakers internalize language-specific generalizations about possible and impossible words, but it is certainly possible to imagine otherwise.
+\citet[][320]{Zimmer1969} writes that phonotactic generalizations ``have no observable consequences in the course of the normal use of the language'', and \citeauthor{PE} echo this sentiment more recently:
 
@@
 
-This represents a principled null hypothesis, but it can be quickly dismissed. 
-This might represent
+This represents a principled null hypothesis, but can be quickly dismissed in light of speaker's ready judgements of possible and impossible words observed by \citeauthor{Chomsky1965}.
+As shown in Chapter \ref{gradience}, wordlikeness tasks implicate prosodic wellformedness and cannot be reduced to purely statistical generalizations.
+Metalinguistic judgements are not the only task in which phonotactic knowledge is on display; speakers are also thought to make use \emph{possible word constraint} in more quotidien tasks like recognizing words in running speech \citep[e.g.,][]{Brown1956,McQueen1998b,Norris1997}.
 
-This would be a principled null hypothesis, but it can be quickly rejected in light of the indisputable contrast between non-words like \emph{blick} and \emph{bnick} noted by \citeauthor{Chomsky1965}.
-It is also possible to show speakers access this knowledge not just in metalinguist tasks like wordlikeness judgements, but also in such naturalistic and quotidien tasks as segmenting spontaneous speech into words \citep{Mattys2001b,McQueen1998}.
+struct/act
+copenhagen
+pure statistical critique
 
-\citeauthor{Chomsky1965} were not the first to take an interest in characterizing the lexicon, which was also a major concern for the scholars of the Copenhagen school.
-In two sophisticated studies dealing with possible and impossible onsets in Georgian, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} recognize the inherent difficulty of distinguishing between structural and accidental phonotactic gaps in the lexicon on the basis of statistical criteria alone.
-What distinguishes the early generativist approach to phonotactics is the mentalistic concern with what generalizations speakers have internalized.
-By grounding the study of phonotactics in the minds of speakers, it is possible to give a precise definition to the previously vague notions; an accidental phonotactic gap is simply one which speakers do not attend to or internalize.
-Yet few studies of phonotactics are explicitly concerned with what speakers have internalized, despite the increasing sophistication of the phonological representations used.
-In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique, a chi-square test, in an attempt to validate a generalization about Arabic root consonant co-occurrence faces originally due to \citet{Greenberg1950}.
-While this particular generalization was ultimately also evaluated using a wordlikeness task \citep{Frisch2004}, there is now an extensive literature in which phonotactic knowledge is inferred directly from lexical statistics, with no psycholinguistic component of any kind
-\citep[e.g.,][]{Anttila2008,
-Berkley1994b,Berkley1994a,Berkley2000,
-Brown2010,
-Buckley1997,
-Coetzee2008a,
-Dmitrieva2008a,Dmitrieva2008b,
-Elmedlaoui1995,
-Graff2011,
-Padgett1992,Padgett1995,
-MacEachern1999,
-Kinney2005,
-Kawahara2006,
-Martin2007,Martin2011,
-Mester1988,
-Miller-Ockhuizen2003,
-Pozdniakov2007,
-Yip1989}.
-%find amanda miller-ockhuizen on ju|'hoansi
-The (usually tacit) assumption of most of this work is made explicit by \citet{Brown2010} in a study of consonant cluster co-occurrence restrictions in Gitksan: he claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers.
+\citet{Fischer-Jorgensen1952} and \citet{Vogt1954}
+note the inherent difficulty of distinguishing between structural and accidental phonotactic gaps.
+
+Even if the criteria differentiating structural and accidental generalizations were purely statistical, an accidental generalization is one which lacks an antecedent cause, and it falls to the linguist to delimit a theory of these causes.
+
+\citeauthor{Chomsky1965} were not the first linguists to take an interest in possible wordhood.
+
+The early generativists identify these causes in mentalism, grounding the study of accidental gaps in what patterns are internalized, and ultimately how they are learned and revealed in psycholinguistic tasks.
+Consequently, for the generativist an accidental phonotactic generalization can only be one which speakers do not learn, do not attend to, and thus a generalization which is without mental antecedent.
+
+Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, expresses limited interest in what generalizations speakers internalize.
+In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique, a chi-square test, in an attempt to validate a generalization about Arabic root consonant co-occurrence facts, observations originally due to \citet{Greenberg1950}.
+The assumption appears to be that statistically valid static co-occurence restrictions will be reflected by wordlikeness judgements.
+While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies in the \citeauthor{McCarthy1988}ian model; rather, the phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Dmitrieva2008a,Dmitrieva2008b,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}.
+In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} makes the tacit assumption of this approach explicit, claiming that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers.
 
 \begin{quote}
-\ldots{}the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
+\ldots{} the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
 \end{quote}
 
-In the loanword adaptation literature, it has recently been recognized that not all statistically reliable generalizations about the native lexicon take part in nativization of borrowings \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}.
-In this chapter, data from Turkish is used to show that a widely reported, statistically reliable phonotactic generalization (\textsc{Labial Attraction}) has no effect on wordlikeness judgements.
-This undermines the widespread use of statistic criteria to determine which of the vast number of the overlapping, competing phonotactic generalizations are internalized by speakers and which are ignored.
-Comparison of \textsc{Labial Attraction} to related constraints in Turkish suggests a novel necessary condition for identifying psychologically real phonotactic constraints: namely, speakers attend to phonotactic generalizations only when they are derived from phonological processes (including syllabification).
+There are several arguments against this inference.
+First, linguists since Saussure have noted diachronic causes for statistical generalizations which lack any synchronic reality.
+For instance, Saussure \citeyear[202f.]{CLG} recognizes the a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}, and that this makes intervocalic \emph{s} rare in Classical Latin.
+However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.} Secondly, wordlikeness generalizations are constrained by negative results in learnability (e.g., \citealp{Gold1967}; see \citealt{Yang2012} for a recent review), and may also be restricted by markedness constraints \citep[e.g.,][]{Becker2011,Hayes2009,Moreton2002}.
+Finally, there are a number of results which suggest patterns obvious to the linguist go unlearned \citep{Becker2011,Hayes2006,Zimmer1969}.
+This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations.
+Consequently, grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may not be appropriate models of speakers' phonotactic knowledge.
 
-There are a number of precedents for
+This chapter is a comprehensive investigation of three well-known phonotactic generalizations in Turkish, as well as the results of a wordlikeness task designed to investigate the synchronic reality of these generalizations \citep{Zimmer1969}.
+This study merits reconsideration because there is widespread disagreement about the analysis of \citeauthor{Zimmer1969}, and prior discussions fail to relate the behavioral data to lexical statistics and competing formalizations of the constraints involved.
+One of these constraints, \textsc{Labial Attraction}, is shown to be highly statistically reliable but to be inactive in the wordlikeness task.
+This undermines the widespread use of statistical criteria to determine which of the vast number of phonotactic generalizations are internalized by speakers and which are not.
+The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is phonologically inactive a well.
 
-This hypothesis was anticipated by \cite[][283]{Anderson1974}.
+\section{Turkish vowel sequence structure constraints}
+
+\citet{Lees1966b,Lees1966a} proposes three constraints on Turkish vowel sequences, constraints which are the focus of many subsequent studies.
+In this setion, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence suggesting that speakers internalize restrictions.
+The following feature specification for the eight vowels of Turkish vowels is assumed throughout.
+
+\begin{example}[Turkish vowel features]
+\begin{tabular}{c c c c c}
+                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
+                       & [$-$\textsc{Rnd}] & [$+$\textsc{Rnd}] & [$-$\textsc{Rnd}] & [$+$\textsc{Rnd}] \\ 
+\cmidrule{2-5}
+\buf[$+$\textsc{High}] & {i} & {ü} [y] & {ı} [ɯ] & {u} \\
+\buf[$-$\textsc{High}] & {e} & {ö} [ø] & {a} [ɑ] & {o} \\
+\end{tabular}
+\end{example}
+
+Two non-conventional notations are used in this chapter.
+First, directional application conditions are used when necessary to model the directionality of harmony, rather than ad hoc contextual restrictions (e.g., that the trigger be aligned with the left of a prosodic word).
+Directional application provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}) and simplifies the analysis of harmony in Turkish \citep[209f.]{Anderson1974}.
+Further, \citet{Johnson1972} and \citet{Kaplan1994} prove that rules specified for directional application are formally learnable.
+Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used.
+Following \citet{McCawley1973}, a structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses the two potential counterexamples against this restrict, both adduced in \emph{SPE} (352-353).}
+
+\subsection{Backness harmony}
+
+\citeauthor{Lees1966b} (\citeyear[35]{Lees1966b}, \citeyear[284]{Lees1966a}) models the Turkish vowel harmony system with three rules.
+The most general of these rules spreads the specification [\textsc{Back}] rightward.
+
+\begin{example}[\textsc{Backness Harmony} (condition: rightward application)]
+$\begin{bmatrix} -\textsc{Cons} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~\big /~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~\textrm{C}_0~\gap$
+\end{example}
+
+\noindent
+A vowel becomes [$+$\textsc{Back}] after a [$+$\textsc{Back}] vowel, and [$-$\textsc{Back}] after a [$-$\textsc{Back}] vowel, ignoring any intervening consonants.
+The application of this rule proceeds from left to right; no vowel may be skipped.
+
+If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllablic roots to contain only [$+$\textsc{Back}] or [$-$\textsc{Back}] vowels, a tendency which will be quantified below.
+\textsc{Backness Harmony} also triggers alternations in inflectional suffix vowels.
+For instance, the nominative plural (nom.pl.) suffix is \emph{-ler}, when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].
+
+\begin{example}[The Turkish nominative] \label{turknom}
+\begin{tabular}{l l l l@{ }l}
+   & \emph{nom.sg.} & \emph{nom.pl.} \\
+a. & {ip}           & {ipler}    & `rope'         & \citep[][216]{Clements1982} \\
+%   & {el}           & {eller}    & `hand'         \\
+   & {köy}          & {köyler}   & `village'      \\
+   & {yüz}          & {yüzler}   & `face'         \\
+   & {kız}          & {kızlar}   & `girl'         \\
+%   & {sap}          & {saplar}   & `stalk'        \\
+%   & {son}          & {sonlar}   & `end'          \\
+   & {pul}          & {pullar}   & `stamp'        \\
+b. & {neden}        & {nedenler} & `reason'       & \citep{TELL} \\
+   & {kiler}        & {kilerler} & `pantry'       \\ % front
+   & {pelür}        & {pelürler} & `tissue paper' \\ % back
+   & {boğaz}        & {boğazlar} & `throat'       \\ % back
+   & {sapık}        & {sapıklar} & `pervert'      \\ % back
+\end{tabular}
+\end{example}
+
+A few complications arise, however.
+First, as shown in (\ref{turkexcept}a), not all polysyllabic roots conform to \textsc{Backness Harmony}.
+In this case, suffix vowels generally exhibit harmony with the final root vowel.
+
+\begin{example}[Exceptional Turkish nominatives] \label{turkexcept}
+\begin{tabular}{l l l l@{ }l}
+   & \emph{nom.sg.} & \emph{nom.pl.} \\
+a. & {mezar}        & {mezarlar} & `grave' & \citep{TELL} \\
+   & {model}        & {modeller} & `model' \\
+   & {silah}        & {silahlar} & `weapon'     \\
+   & {memur}        & {memurlar} & `bureaucrat' \\
+   & {sabun}        & {sabunlar} & `soap'       \\
+b. & {saat}         & {saatler}  & `hour, clock' & \citep{Goksel2005} \\
+   & {harf}         & {harfler}  & `(alphabetic) letter' \\ 
+   & {etol}         & {etoller}  & `fur stole' \\
+\end{tabular}
+\end{example}
+
+\noindent
+There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} despite the fact that their final root vowel is [$+$\textsc{Back}].
+While it is uncontroversial that the disharmonic suffixes of (\ref{turkexcept}b) are no more than very sporadic exceptions to \textsc{Backness Harmony}, root disharmony (\ref{turkexcept}a) has been the subject of much debate.
+As disharmonic roots still trigger suffix harmony, \citet[212, 289]{Anderson1974} and \citet{Iverson1978} distinguish suffix harmony alternations from the sequence structure constraint governing root harmony.
+
+The disadvantage of this account is that it introduces ``duplication'' (in the sense of \citealt{Kisseberth1970b} and \citealt{Kenstowicz1977}) of sequence structure and phonological generalizations, differning only in their patterns of exceptionality.
+However, \citet[][197f.]{Zonneveld1978} shows that the theory of exceptionality proposed in \emph{SPE} can account for suffix harmony in disharmonic roots.\footnote{An isomorphic analysis is available within less restrictive theories of phonological exceptionality \citep[e.g.,][]{Kisseberth1970,Pater2009}.} \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
+A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}].
+In other words, no form is never truly an ``exception''; rather, some forms fail to match the extended structural description of $R$ which requires that the target be [$+R$].
+If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.
+Thus \textsc{Backness Harmony} is able to account for both root and suffix harmony.
+
+It is necessary to dispense with an alternative analysis proposed by \citet{Clements1982} and \citet{Inkelas1997}.
+Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
+\emph{kul} `servant', \emph{kepek} `bran' vs.
+\emph{kapak} `lid'), whereas harmonic roots are those in which the backness of any remaining vowels is predictable.
+\citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots are fully specified.
+This is schematized below.
+
+\begin{example}[Autosegmental underspecification in harmonic roots (after \citealp{Clements1982})] \label{spec}
+\xymatrix@R=24pt@C=24pt{
+\txt{a.} & \txt{harmonic root:} & \txt{C} & \txt{V} & \txt{C} & \txt{V} & \txt{C} \\
+         &                      &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
+\txt{b.} & \txt{disharmonic root:} & \txt{C} & \txt{V} & \txt{C} & \txt{V} & \txt{C} \\
+         &                      &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u] & & \txt{[$+$\textsc{Back}]}\ar@{-}[u]
+}
+\end{example}
+
+One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being a Structure Preservation condition \citep{Kiparsky1985}.
+However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a perceptual default.
+A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone.
+However, this interpretation makes the underspecification analysis a mere notational variant of the rule exceptionality account, in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.
+
+While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show internalize the tendency for roots to conform to backness harmony.
+If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it.
+Several other ``external'' facts demonstrate that this is not the case.
+
+The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} While some speakers are said to be capable of pronouncing these non-native clusters, in fast speech the cluster is reportedly split by anaptyxis.
+In most cases, this vowel matches the following root vowel for backness.
+
+\begin{example}[Variable non-native cluster adaptation (\citealp{Clements1982}:247)] 
+\begin{tabular}{l l l l l l}
+a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
+   & {fren}    & \alt{} & {firen}    & `break'     \\
+b. & {trablus} & \alt{} & {tırablus} & `Tripoli'   \\
+   & {kral}    & \alt{} & {kıral}    & `king'      \\
+c. & {brom}    & \alt{} & {burom}    & `bromide'   \\
+   & {prusya}  & \alt{} & {purusya}  & `Prussia'   \\
+%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
+%   & kredi   & \alt{} & kıredi   & `credit'    \\
+\end{tabular}
+\end{example}
+
+\noindent
+It is worth noting that this cannot be merely the consequence of application of the native \textsc{Backness Harmony} rule, since the leftmost vowel of a prosodic word is not subject to harmonization.
+Rather, the adaptation of non-native onset clusters proceeds in such a fashion so that the lexical items in question are rarely [$-$\textsc{Backness Harmony}].
+
+Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but it is found in the grammar of the Turkic language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; \citeauthor{Harrison2001} report that it can be quickly taught to Turkish speakers.
+The game consists of reduplication of the base, with the first vowel of the reduplicant replaced with the [$+$\textsc{Back}] vowels \emph{a} or \emph{u}.
+Both in Tuvan and in the Turkish game, the second vowel of the reduplicant (shown in braces below) may also be effected by this process; when the base vowels are both [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] vowel results in what \citeauthor{Harrison2001} call ``reharmonization'', shown in  (\ref{redupgame}a).
+On the other hand, as shown in (\ref{redupgame}b), disharmonic roots do not reharmonize.\footnote{A similar pattern is found in Tuvan, and in an unrelated Finnish language game \citep{Campbell1986}.}
+
+\begin{example}[Turkish reduplication game (\citealp{Harrison2001}:231)] \label{redupgame}
+\begin{tabular}{l l l l}
+a. & {kibrit} & {kibrit}-\{{kabrıt}\} & `match'    \\
+   & {bütün}  & {bütün}-\{{batın}\}   & `whole'    \\
+b. & {mali}   & {mali}-\{{muli}\}     & `Mali'     \\
+   & {butik}  & {butik}-\{{batik}\}   & `boutique' \\
+\end{tabular}
+\end{example}
+
+\noindent 
+\citeauthor{Harrison2001} present an analysis of this game couched in the vowel underspecification proposed by \citeauthor{Clements1982}, but the data is equally consistent with a full specification analysis.
+Reharmonization is the result of \textsc{Backness Harmony} in the reduplicant.
+On the other hand, the lack of reharmonization in the disharmonic roots indicates that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.
+
+A number of studies have investigated the role of harmony in word-spotting tasks, though to reveal how speakers identify word boundaries in running speech.
+Many of these studies have been carried out in Finnish, which has a vowel harmony system similar to Turkish.
+\citet{Suomi1997} and \citet{Vroomen1998} task speakers with identifying harmonic disyllables in an auditory stream.
+Finnish speakers are better able to complete this task when the preceding syllable has a different backness specification than the disyllabic target.
+\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect on word-spotting tasks as it does in Finnish: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}), but this effect does not obtain for speakers of French, a language which lacks vowel harmony.
+These results imply that speakers of Turkish have internalized the tendency of harmonic sequences to be root-internal and of disharmonic transitions to mark word juncture.
+
+The word-spotting experiment has been adapted for infants by \citep{Kampen2008}.
+In this study, 9-month-old infants are familiarized with an auditory stream containing harmonic disyllabic nonce words.
+At test time, the infants listen to the disyllabic nonce words in isolation using the head turn preference paradigm.
+Infants acquiring Turkish listen longer to nonce words preceded by a disharmonic juncture during familiarization (e.g., \emph{lo-NETIS}), whereas infants acquiring German, a language which lacks vowel harmony, do not exhibit this preference.
+Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants show no such preference.
+
+\subsection{Roundness harmony}
+
+\subsubsection{Phonological description}
+
+This rule is quite similar to \textsc{Backness Harmony}, but imposes an additional restriction, that targets be [$+$\textsc{High}].
+
+\begin{example}[\textsc{Roundness Harmony} (condition: rightward application)]
+$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~/~\begin{bmatrix}~=\textsc{Back}~\end{bmatrix}~\textrm{C}_0~\gap$
+\end{example}
+
+\noindent
+A [$+$\textsc{High}] vowel becomes [$+$\textsc{Rnd}] after a [$+$\textsc{Rnd}] vowel, and [$-$\textsc{Rnd}] after a [$-$\textsc{Rnd}] vowel, ignoring any intervening consonants, and applying from left to right.
+
+If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only [$+$\textsc{Rnd}] or [$-$\textsc{Rnd}] if the non-initial vowels are [$+$\textsc{High}].
+In concert with \textsc{Backness Harmony}, \textsc{Roundness Harmony} also triggers alternations which account for the shape of the dative singular (dat.sg.) and genitive singular (gen.sg.) among other suffixes.
+As was the case for \textsc{Backness Harmony}, disharmonic roots are found, and the final vowel of disharmonic roots triggers suffix harmony.
+
+\begin{example}[Turkish nominal suffix allomorphy]
+\begin{tabular}{l l l l l l@{ }l}
+   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
+a. & {ip}           & {ipi}          & {ipin}         & `rope' & \citep[][216]{Clements1982} \\
+%  & {el}           & {eli}          & {elin}         & `hand'    \\
+   & {kız}          & {kızı}         & {kızın}        & `girl'    \\
+   & {sap}          & {sapı}         & {sapın}        & `stalk'   \\
+%  & {yüz}          & {yüzü}         & {yüzün}        & `face'    \\
+   & {köy}          & {köyü}         & {köyün}        & `village' \\
+%  & {pul}          & {pulu}         & {pulun}        & `stamp'   \\
+   & {son}          & {sonu}         & {sonun}        & `end'     \\
+b. & {boğaz}        & {boğazı}       & {boğazın}      & `throat'  & \citep{TELL} \\
+   & {pelür}        & {pelürü}       & {pelürün}      & `tissue paper' \\
+   & {döviz}        & {dövizi}       & {dövizin}      & `currency' \\
+   & {yamuk}        & {yamuğu}       & {yamuğun}      & `trapezoid' \\
+   & {ümit}         & {ümiti}        & {ümitin}       & `hope'     \\
+\end{tabular}
+\end{example}
+
+Few studies have directly investigated whether speakers are aware of the tendency for roots to conform to \textsc{Roundness Harmony}.
+However, two pieces of the external evidence on \textsc{Backness Harmony} bear on this question.
+First, the epenthetic vowel in non-native word-initial onsets like \emph{trablus} \alt{} \emph{tırablus} `Tripoli' conform both to \textsc{Backness Harmony} and to \textsc{Roundness Harmony}.
+Secondly, \textsc{Roundness Harmony} participates in reharmonization in the language game described by \citeauthor{Harrison2001}: the second \emph{ü} in \emph{bütün} `whole' reharmonizes to the [$-$\textsc{Rnd}] vowel \emph{ı} in reduplicated \emph{bütün-batın}.
+
+\subsection{Labial attraction}
+
+\citet{Lees1966b} describes \textsc{Labial Attraction} as a phonological process by which ``a high, short harmonic vowel is rounded in the second syllable of a disyllabic word whose first vowel is /a/, and whose medial consonant cluster contains a labial /p, b, m, v/, and then it is de-harmonified'' (36).
+
+\begin{example}[\textsc{Labial Attraction}]
+$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} -\textsc{Cons} \\ -\textsc{Rnd} \\ +\textsc{Back} \\ -\textsc{High} \end{bmatrix}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
+\end{example}
+
+The formalization of this rule is naturally complex, and perhaps obscures the fact that \textsc{Labial Attraction} generates exceptions to \textsc{Roundness Harmony}, producing \emph{a}P\emph{u} sequences (where P represents a labial consonant; e.g., \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', \emph{samur} `sable'; \citealp[285]{Lees1966a}) rather than the expected \emph{a}P\emph{ı}.
+However, \textsc{Labial Attraction} does not apply in derived environments: for example, the gen.sg.
+of \emph{sap} `stalk is \emph{sapın} rather than *\emph{sapun}.
+
+\citet[286]{Lees1966a} and \citet[311]{Zimmer1969} note the existence of root-internal exceptions to this rule (e.g., \emph{tavır} `mode') but agree that they are surprisingly rare; this is disputed in turn by  \citet{Clements1982} and \citet{Inkelas2001}.
+A more exhaustive review of the lexical statistics will be provided below.
+
+\section{Evaluation}
+
+by comparing lexical statistics with the results of wordlikeness task performed by \citet{Zimmer1969}.
+
+This resolves the debate concerning the nature of \textsc{Labial Atttraction}: it is shown to be a statistically robust generalization over the Turkish lexicon (\citealt{Lees1966a}, \citealt{Zimmer1969}, \emph{pace} \citealt{Clements1982}, \citealt{Inkelas2001}); however, no variant of \textsc{Labial Attraction} is reflected in \citeauthor{Zimmer1969}'s wordlikeness study.
+This dissociation between lexical generalizations and wordlikeness results provides further evidence against the common practice of inferring phonotactic knowledge via purely statistical criteria.
+
+\subsection{Lexical statistics}
+
+%\citet{Becker2011,Hayes2009}
+
+\subsubsection{Backness harmony}
+\subsubsection{Roundness harmony}
+\subsubsection{Labial attraction}
 
 \begin{quote}
-It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
-\citep[][283]{Anderson1974}
+\ldots{}decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots{}aCu\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots{}We conclude that there is no systematic restriction on the set of consonants that may occur medially in roots of the form /\ldots{}aCu\ldots/.
+\citep[225]{Clements1982}
 \end{quote}
 
-This hypothesis is not ahdered to by early research on sequence structure constraints \citep[e.g.,][]{SPR,Chomsky1965,SPE,Stanley1967}.
+but this fact is not inconsistent with the formulation of the constraint:
+\emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description.
 
-Work in this vein sought to use sequence structure constraints to compress the lexicon as much as possible
+\noindent
+\citeauthor{Clements1982} are suggesting that \textsc{Labial Attraction} implies that \emph{a}\ldots{}\emph{u} sequences separated by non-labial consonants should be infrequent.
+This is not immediately clear, since this is beyond the scope of \citeauthor{Lees1966a}'s description of \textsc{Labial Attraction}, but the frequency of these \emph{a}T\emph{u} sequences does provide an estimate of how often ???.
 
+\begin{quote}
+Lee's rule of \textsc{Labial Attraction}\ldots is not a real generalization about the Turkish lexicon.
+It is not true synchronically, either of native or nonnative items; nor, according to the historical and dialectical literature, does \textsc{Labial Attraction} appear to have been true at any stage going back as far as Old Turkic [\ldots] Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
+\citep[][196, emphasis in original]{Inkelas2001}
+\end{quote}
+
+\begin{example}[Lexical effects of \textsc{Labial Attraction}]
+\begin{tabular}{l r r r r r}
+\toprule
+Corpus                & aPu & aPı & aTu & aTı   & $p$-value   \\ % & corpus size
+\midrule
+Full TELL             & 378 & 248 & 446 & 1,140 & 2.83\e{-44} \\ % & 31,236 \\
+Elicited TELL         & 152 & 265 & 101 & 1,839 & 9.84\e{-60} \\ % & 16,541 \\
+TELL with etymologies & 128 & 109 &  79 &   470 & 6.56\e{-32} \\ 
+\bottomrule
+\end{tabular}
+\end{example}
+
+%\citet{Harrison2004} 73\% of lexical types are harmonic (both back and round)
+
+%Etymological subset of TELL
+%Native    & Foreign   \\
+%aBu & aBI & aBu & aBI \\
+%12  & 11  & 84  & 28  \\
+%p = 0.0417
+
+\subsection{Wordlikeness ratings}
+
+%\citet{Gigerenzer1990} argue that ratings tasks should be replaced with paired comparison to minimize error.
+
+%\citet{Goodman1954}
+%The results reveal a dissociation between the lexical statistics and wordlikeness judgements, a dissociation that finds a natural explanation from the principle of ...
+%There is some disagreement in the literature about just what \citeauthor{Zimmer1969}'s study reveals about \textsc{Labial Attraction}.
+
+%\citet{Ito1993} and \citet{NiChiosain1993} argue that the data shows that it holds of native words only, but \citet{Inkelas2001} refute this point
+
+%\citet{Zuraw2000}:4
+%\citet{Inkelas2001} present a different variation on the original rule of \textsc{Labial Attraction}
+
+%\begin{quote}
+%Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
+\citep[][196, emphasis in original]{Inkelas2001}
+%\end{quote}
+
+%\noindent 
+%\citeauthor{Inkelas2001} persistently refer to ``statistical support'', but in fact their study contains no statistics beyond counts and percentages.
+%since the opposing are all broader versions than the original formulation by \citeauthor{Lees1966a}.
+
+\subsubsection{Backness harmony}
+
+\begin{example}[Wordlikeness comparisons, backness harmony]
+\begin{tabular}{lrlr|lrlr}
+\toprule
+\multicolumn{4}{c|}{Experiment 1} & \multicolumn{4}{c}{Experiment 2} \\
+\multicolumn{2}{c}{\textsc{harmonic}} & \multicolumn{2}{c|}{\textsc{disharmonic}} & \multicolumn{2}{c}{\textsc{harmonic}} & \multicolumn{2}{c}{\textsc{disharmonic}} \\
+\midrule
+{temez} & 19            & {temaz} & 3 & {pemez} & 30            & {pemaz} & 2 \\
+{teriz} & 23            & {terız} & 0 & {teriz} & 28            & {terız} & 3 \\
+{tokaz} & 21            & {tokez} & 1 & {tokaz} & 26            & {tokez} & 6 \\
+{tipez} & 21            & {tipaz} & 1 & {tipez} & 24            & {tipaz} & 8 \\
+{terüz} & 20            & {teruz} & 1 & {terüz} & 19            & {teruz} & 13 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+% harmonic responses are favored approximately 6 to 1.
+%$\gamma = 0.717$, $p = 7.5$\e{-69}
+
+Note that both of the pair \emph{terüz}-\emph{teruz} violate \textsc{Roundness Harmony}
+
+\subsubsection{Roundness harmony}
+
+\begin{example}[Wordlikeness comparisons, roundness harmony]
+\begin{tabular}{lrlr|lrlr}
+\toprule
+\multicolumn{4}{c|}{Experiment 1} & \multicolumn{4}{c}{Experiment 2} \\
+\multicolumn{2}{c}{\textsc{harmonic}} & \multicolumn{2}{c|}{\textsc{disharmonic}} & \multicolumn{2}{c}{\textsc{harmonic}} & \multicolumn{2}{c}{\textsc{disharmonic}} \\
+\midrule
+{törüz} & 19 & {töriz} & 1 & {pörüz} & 32 & {pöriz} & 0  \\
+{tüpüz} & 22 & {tüpiz} & 0 & {tüpüz} & 31 & {tüpiz} & 1  \\
+{takız} & 15 & {takuz} & 3 & {takız} & 22 & {takuz} & 10 \\
+{tatız} & 12 & {tatuz} & 6 & {tatız} & 20 & {tatuz} & 12 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+% more than 5 to 1
+% gamma = 0.680 p = 1.1e-47
+
+\subsubsection{Labial attraction}
+
+\begin{example}[Wordlikeness comparisons, labial attraction]
+\begin{tabular}{lrlr|lrlr}
+\toprule
+\multicolumn{4}{c|}{Experiment 1} & \multicolumn{4}{c}{Experiment 2} \\
+\multicolumn{2}{c}{aPu} & \multicolumn{2}{c|}{\textsc{disharmonic}} & \multicolumn{2}{c}{\textsc{harmonic}} & \multicolumn{2}{c}{\textsc{disharmonic}} \\
+\midrule
+{tamuz} & 3 & {tamız} & 16 & {pamuz} & 15 & {pamız} & 17 \\
+{tafuz} & 3 & {tafız} & 17 & {tafuz} & 21 & {tafız} & 11 \\
+{tavuz} & 9 & {tavız} & 4  & {mavuz} & 16 & {mavız} & 16 \\
+{tapuz} & 7 & {tapız} & 9  & {tapuz} & 17 & {tapız} & 15 \\
+{tabuz} & 5 & {tabız} & 12 & {tabuz} & 16 & {tabız} & 16 \\
+\bottomrule
+\end{tabular}
+\end{example}
+
+\subsection{Discussion}
+
+etymological issues
+
+%\citet{Inkelas2001}
+%\citet{NiChiosain1993} and \citet{Ito1995b} 
+%\begin{example}[\textsc{Labial harmony} and etymology in TELL \citealp{Inkelas2001}:187]
+%\begin{tabular}{l r r r r}
+%\toprule
+%        & {a}P{u} & {a}P{ı} & \% aPu & $p$-value \\
+%\midrule
+%native  & 12      & 11      & 52.2\% & \multirow{2}{*}{0.042} \\
+%foreign & 84      & 28      & 75.0\% \\
+%\bottomrule
+%\end{tabular}
+%\end{example}
+
+lack of theory of naturalness
+
+%\citet{Becker2011}
+%This is difficult to evaluate insofar as \citeauthor{Becker2011} neither propose nor refer to any theory of naturalness which might exclude \textsc{Labial Attraction}. 
+%There are several other problems with this claim.
+Both early specialists \citep[e.g.,][]{Lees1966a} and later theorists working within a functionally-oriented framework \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995a} have considered it plausible.
+Further, \citet{Hayes2009} claim that Hungarian speakers internalize similarly ``unnatural'' generalizations about exceptions to vowel harmony.
+%Finally, there is at least some reason to admit similar structural descriptions in other languages.
+For example, some Classical Arabic adjectives have stative verbs in which the root melody is overwritten with \emph{a\ldots{}u}.
+If the stative verbs are derived from the adjectives, as has recently been argued for Semitic \citep[e.g.,]{Ussishkin1999,Ussishkin2000,Tucker2010}, than there is little reason to believe that this sequence is somehow marked.
+%\begin{example}[Arabic stative verbs]
+%\begin{tabular}{l l l l}
+%\buf{}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
+%\buf{}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
+%\buf{}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
+%\end{tabular}
+%\end{example}
+%\footnote{Thanks to Uri Horesh (p.c.) for assistance with this data.}
+
+%\noindent
+%Further, labial consonants frequently cause rounding of following vowels \citep[e.g.,][]{Vaux1993}. The relevant constraints, conjoined, are a close approximation of \textsc{Labial Attraction}.
+
+\section{Conclusions}
+
+%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
+It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails.
+If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
+%A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
+This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).
+%Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
+In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.
+%In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word. 
+%The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements. 
+%The data below, drawn from Turkish, shows that 
+%This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
+Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements.
+Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
+%\begin{unlabeledexample}
+%$\displaystyle \gamma = \frac{C - D}{C + D}$
+%\end{unlabeledexample}
+%\footnote{Any readers who are skeptical that \citeauthor{Chomsky1965}'s intuitions generalize will find support for their claim in the wordlikeness judgement data reported in Appendix \ref{albright}.}
+%As a quick validation of this claim, consider the results reported by \citet{Albright2007} and discussed in the previous chapter.
+English speakers asked to rate non-words on a scale from 1-7, where 7 indicates ``most like English'', favor non-words [blʌs, blɑd] (average scores 4.67 and 5.13) over [bnʌs, bnɑd] (average scores 2.06, 2.00).
+%This hypothesis is not ahdered to by early research on sequence structure constraints \citep[e.g.,][]{SPR,Chomsky1965,SPE,Stanley1967}.
+Work in this vein sought to use sequence structure constraints to compress the lexicon as much as possible.
 This work has a second interest, namely in ``compressing'' the lexicon as much as possible,
-
-It is important to know that this wor It is important to know that this wor It is important to know that this wor It is important to know that this work
-
-% motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
+%motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
 \citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible.
@@
 In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities
-
-This condition (henceforth, the \textsc{derived} condition) 
-On a historical note, this hypothesis is more constrained than what was proposed
-such constraints must be derived from phonological processes in the language in question
-synchronically motivated phonological processes such as alternation or syllabification.
-
-%It is important to note that the nature of the \emph{blick}/\emph{bnick} contrast is no longer a mystery. 
 %Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
@@
 If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
-
-%\footnote{I note a potential problem with this account: \citet{Davidson2005,Davidson2006a} argue that this ``excrescent'' schwa is distinct from the both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.}
-
 %This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
-Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
+%Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
 This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.
-
-%\citet{Lees1966b,Lees1966a}
 
@@
 %Elliot Moreton (p.c.)
-
-%\noindent A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions.
-Any explicit theory of static constraints will have to impose formal restrictions on possible co-occurrence restrictions.
-The set of possible constraints must respect results in formal learning theory, and may also be constrained by markedness considerations, for example.
-
-%A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
-This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).
-
-%Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
-In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.
-
-%In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word. 
-%The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements. 
-%The data below, drawn from Turkish, shows that
-
-%This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
-Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements.
-Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
-
-%Classical Arabic adjectives often have stative verbs in which the root is imposed onto the template CaCuCa:
-
-\section{Turkish vowel sequence structure constraints}
-
-%\citet{Lees1966a,Lees1966b} 
-\citet{Lees1966b,Lees1966a} 
-proposes three constraints on Turkish vowel sequences which have been the focus of many subsequent studies.
-In this section, these constraints are formalized, and where possible, related to phonological alternations in Turkish and to evidence that speakers have internalized these restrictions.
-The following feature specification for the Turkish vowels is assumed throughout.
-
-\begin{example}[Turkish vowel features]
-\begin{tabular}{c c c c c}
-                       & \multicolumn{2}{c}{[$-$\textsc{Back}]} & \multicolumn{2}{c}{[$+$\textsc{Back}]} \\
-                       & [$-$\textsc{Round}] & [$+$\textsc{Round}] & [$-$\textsc{Round}] & [$+$\textsc{Round}] \\ %\midrule 
-\buf[$+$\textsc{High}] & {i} & {ü} [y] & {ı} [ɯ] & {u} \\
-\buf[$-$\textsc{High}] & {e} & {ö} [ø] & {a} [ɑ] & {o} \\
-\end{tabular}
-\end{example}
-
-\subsection{Backness harmony}
-
-\citeauthor{Lees1966b} (\citeyear[][35]{Lees1966b}, \citeyear[][284]{Lees1966a}) models the Turkish vowel harmony systems by means of three feature spreading rules.
-Backness harmony is the the most general of these processes.
-
-\subsubsection{Phonological description}
-
-\textsc{Backness Harmony} spreads the backness specification of a vowel rightward ignoring any intervening consonants.
-
-\begin{example}[\textsc{Backness Harmony}]
-$\begin{bmatrix} +\textsc{Vocalic} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~\big /~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~\textrm{C}_0~\gap{}$
-\end{example}
-
-%\citep[after][229]{Clements1982}: \\
-%\xymatrix@R=24pt@C=24pt{
-%\txt{V}                                        & \txt{C$_0$~~~~~} & \txt{V} & \txt{(condition: rightward application)} \\
-%\txt{[α \textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
-%}
-
-This rule accounts for the tendency of polysyllablic roots to either contain all [$+$\textsc{Back}] or all [$-$\textsc{Back}];
-%, as can be seen in the nominative singulars in (\ref{turknom}b); 
-this tendency is quantified in \S\ref{lexstats} below.
-\textsc{Backness Harmony} also accounts for the allomorphs of several inflectional suffixes.
-For instance, the nominative plural (nom.pl.) suffix is \emph{-ler}, when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].
-
-\begin{example}[The Turkish nominative] \label{turknom}
-\begin{tabular}{l l l l@{ }l}
-   & \emph{nom.sg.} & \emph{nom.pl.} \\
-a. & {ip}           & {ipler}    & `rope'        & \citep[][216]{Clements1982} \\
-   & {el}           & {eller}    & `hand'    \\
-   & {köy}          & {köyler}   & `village' \\
-   & {yüz}          & {yüzler}   & `face'    \\
-   & {kız}          & {kızlar}   & `girl'    \\
-   & {sap}          & {saplar}   & `stalk'   \\
-   & {son}          & {sonlar}   & `end'     \\
-   & {pul}          & {pullar}   & `stamp'   \\
-b. & {neden}        & {nedenler} & `reason'       & (TELL) \\ % front
-   & {kiler}        & {kilerler} & `pantry'       \\ % front
-   & {pelür}        & {pelürler} & `tissue paper' \\ % back
-   & {boğaz}        & {boğazlar} & `throat'       \\ % back
-   & {sapık}        & {sapıklar} & `pervert'      \\ % back
-\end{tabular}
-\end{example}
-
-There are a few complications, however.
-First, not all polysyllabic roots conform to \textsc{Backness Harmony}.
-In this case, as shown in (\ref{turkexcept}a), suffix vowels generally harmonize with the final root vowel.
-There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} despite the fact that their final root vowel is [$+$\textsc{Back}].
-
-\begin{example}[Exceptional Turkish nominatives] \label{turkexcept}
-\begin{tabular}{l l l l@{ }l}
-   & \emph{nom.sg.} & \emph{nom.pl.} \\
-a. & {mezar}        & {mezarlar} & `grave' & \citep{TELL} \\
-   & {model}        & {modeller} & `model' \\
-   & {silah}        & {silahlar} & `weapon'     \\
-   & {memur}        & {memurlar} & `bureaucrat' \\
-   & {sabun}        & {sabunlar} & `soap'       \\
-b. & {saat}         & {saatler}  & `hour, clock' \\
-   & {harf}         & {harfler}  & `(alphabetic) letter' \\ %& \citep{Goksel2005}
-   & {etol}         & {etoller}  & `fur stole' \\
-\end{tabular}
-\end{example}
-
-While it is uncontroversial that the disharmonic suffixes of (\ref{turkexcept}b) are no more than very sporadic exceptions to \textsc{Backness Harmony}, root disharmony has ben the subject of much debate.
-As disharmonic roots still trigger suffix harmony, \citet[][212, 289]{Anderson1974} and \citet{Iverson1978} propose to separate suffix harmony, an alternation, from a sequence structure constraint governing root harmony.
-
-The disadvantage of this account is that it introduces a sort of ``duplication'' (in the sense of \citealt{Kenstowicz1977}) of sequence structure and phonological generalizations, differning only in their patterns of exceptionality.
-%\citep{Kisseberth1970b} (see also \citealp[][401]{Stanley1967} and \emph{SPE}:382)
-However, \citet[][197f.]{Zonneveld1978} observes that the theory of exceptionality proposed in \emph{SPE}
-% (p.~374f.)
-can account for suffix harmony in disharmonic roots.
-\citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
-A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}].
-In other words, a form is never truly an ``exception'': it simply fails to match an extended structural description including the rule feature.
-If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.\footnote{The definition of ``target of a rule'' is not obvious in autosegmental rules, but any useful formalization of this notion should, for example, identify a harmonic suffix vowel as the target of \textsc{Backness Harmony}.} With this in place, \textsc{Backness Harmony} is sufficient to account for both root and suffix harmony.
-
 %Once additional source of evidence on root (dis)harmony is inconclusive.
@@
 However, this might simply indicate that the fluctuating vowel is epenthetic and that harmony applies before epenthesis (see \citealt{Clements1982} for both sides of this argument), making it less than a counterexample.
+%duplication problem: \citep{Kisseberth1970b} (see also \citealp[][401]{Stanley1967} and \emph{SPE}:382)
+%(e.g., \emph{deve} `camel' vs. \emph{deva} `medicine', \emph{sene} `year' vs. \emph{sena} `praise').
 
-Before moving on, it is necessary to dispense with an alternative analysis originally proposed by \citet{Clements1982} and further developed by \citet{Inkelas1997}.
-Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
-\emph{kul} `servant', \emph{kepek} `bran' vs.
-\emph{kapak} `lid'), whereas harmonic roots are those in which the backness of any remaining vowels is predictable.
-\citeauthor{Clements1982} propose that these vowels are underspecified for backness, whereas the non-initial vowels of disharmonic roots are fully specified.
-This is schematized below.
+%This is not to imply uncritical acceptance of \citeauthor{Becker2011}'s account.
+\citeauthor{Becker2011} place this generalization ``in the grammar'' to account for their claim that ``naturalness'', which they equate with Universal Grammar, constrains the types of generalizations speakers extract in this task.
+% 
+%It is difficult to evaluate their claim in the absense of a theory of phonotactic naturalness, something which has consisteny eluded formalization for decades.
+Even the informal definition used by \citeauthor{Becker2011} is flawed.
+\citeauthor{Becker2011} label generalizations unnatural if they lack phonetic precursors, or if they are not unattested---no description of the methodology used to determine attestedness is given.
+These two varieties of evidence are not fully independent, however: extragrammatical \emph{channel bias} effects are thought to be a first-order predictor of attestation \citep{Blevins2004,Moreton2008}.
+In light of extensive evidence for generalizations which are phonetically unnatural \citep[e.g.,]{Anderson1981,Bach1972,Buckley2000a}, even as revealed in nonce word studies \citep[e.g.,][]{Hayes2009}, the fact that the generalizations extracted in the \citet{Becker2011} study are phonetically natural is either accident or conundrum.
+It is certainly not something that can be explained by the tacit theory of possible gneeralizations the authors adhere to.
 
-%This is exemplified below in (\ref{spec}).
-%(e.g., \emph{deve} `camel' vs. \emph{deva} `medicine', \emph{sene} `year' vs. \emph{sena} `praise'). 
-%There is some further evidence that individual vowels may differ in specification for this feature even within individual roots or affixes.
-For instance, the present continuous suffix has harmony-determined allomorphs \emph{-iyor}, \emph{-üyor}, \emph{-ıyor}, \emph{-uyor}, but the \emph{o} of the suffix is invariant.
-A similar situation might obtain in Turkish roots.
-
-\begin{example}[Autosegmental underspecification in harmonic roots (after \citealp{Clements1982})] \label{spec}
-\xymatrix@R=24pt@C=24pt{
-\txt{a.} & \txt{harmonic root:~~~~} & \txt{C} & \txt{V} & \txt{C} & \txt{V} & \txt{\ldots} \\
-&   &    & \txt{[$-$\textsc{Back}]}\ar@{-}[u]\ar@{--}[urr] \\
-\txt{b.} & \txt{disharmonic root:} & \txt{C} & \txt{V} & \txt{C} & \txt{V} & \txt{\ldots} \\
-    &    &         & \txt{[$-$\textsc{Back}]}\ar@{-}[u] & & \txt{[$+$\textsc{Back}]}\ar@{-}[u]
-}
-\end{example}
-
-Harmonizing suffix vowels will also be underspecified for backness.
-
-Two additional details are needed to complete this analysis.
-First, there must be some kind of constraint which prevents speakers from positing redundant backness specifications for non-initial vowels in harmonic roots, a constraint on identical adjacent specifications in underlying representations, the lexical ``obligatory contour principle'' \citep[][OCP]{Leben1973}, on surface tonal representations \citep{Goldsmith1976}, or both (\citealp{Leben1978}, \citealp{McCarthy1986}; see \citealp{Odden1986,Odden1988} for criticism).
-Second, \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being the use of a structure preservation condition \citep{Kiparsky1985}.
-However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are formally marked in Turkish.\footnote{On the other hand, if one interprets ``harmony'', that is, a single backness specification per root, as the learner's default (as \citealp{Odden1986} proposes for the tonal obligatory contour principle), this account is a mere notational variant of the rule exceptionality account.}
-%\footnote{Kie Zuraw (p.c.) suggests that the markedness of disharmonic roots might be attributed to the presence of more structure, namely, multiple specifications for backness.}
-I reject the underspecification analysis on these grounds. Rather, I assume a unitary phonological rule of \textsc{Backness Harmony} and assign [$-$\textsc{Backness Harmony}] to disharmonic roots.
-
-\subsubsection{External evidence}
-\label{backharmexternal}
-
-While suffix harmony can be inferred from alternations, no evidence has yet been presented to show that speakers internalize in any way tendency for roots to exhibit backness harmony.
-
-%\begin{quote}
-%\ldots{}these generalizations\ldots{}have no observable consequences in the course of the normal use of the language. \citep[][320]{Zimmer1969}
-%\end{quote}
-
-Indeed, \citeauthor{PE} adopt the null hypothesis that speakers do not internalize any generalizations about possible or likely sequences in underlying representations.
-
-\begin{quote}
-Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for.
-\citep[][18]{PE}
-\end{quote}
-
-Several independent pieces of external evidence argue that this is not the case for root harmony in Turkish. 
-The adaptation of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, is one such case.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} While some speakers are said to be capable of pronouncing these non-native clusters, in normal speech the cluster is resolved by epenthesis of a [$+$\textsc{High}] vowel.
-In most cases, the epenthetic vowel matches the following root vowel for backness, suggesting that adapt foreign words in a way that conforms to the root harmony generalization.\footnote{At the same time, this shows that adaptation is not simply applying rules, like \textsc{Backness Harmony}, to foreign forms.
-Rather, foreign forms are restructured as non-exceptional Turkish lexical items.}
-
-%the epenthetic vowel is always [$+$\textsc{High}],
-%and in general, it conforms to \textsc{Backness Harmony}.
-%For reasons that are 
-%and when the first consonant of the cluster is non-dorsal, it shows backness harmony
-
-\begin{example}[Adaptation of initial foreign clusters (\citealp{Clements1982}:247)] 
-\begin{tabular}{l l l l l l}
-a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
-   & {fren}    & \alt{} & {firen}    & `break'     \\
-b. & {brom}    & \alt{} & {burom}    & `bromide'   \\
-   & {prusya}  & \alt{} & {purusya}  & `Prussia'   \\
-%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
-%   & kredi   & \alt{} & kıredi   & `credit'    \\
-%   & trablus & \alt{} & tırablus & `Tripoli'   \\
-%b. & kral    & \alt{} & kıral    & `king'      \\
-%   & grup    & \alt{} & gurup    & `group'     \\
-\end{tabular}
-\end{example}
-
-Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but is identical to a morphological process in the related language Tuvan, in which it conveys a sense of ``vagueness or jocularity'', and \citeauthor{Harrison2001} report that it can be quickly taught to Turkish speakers.
-The game consists of reduplication of the base and replacement of the first vowel in the reduplicant with the [$+$\textsc{Back}] vowels \emph{a} or \emph{u}.
-Both in the native Tuvan process and in the Turkish game, the second vowel of the reduplicant (shown in braces below) is affected by this process: when the base is harmonically [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] results in what \citeauthor{Harrison2001} call ``reharmonization'', shown in  (\ref{redupgame}a).
-
-\begin{example}[Turkish reduplication game (\citealp{Harrison2001}:231)] \label{redupgame}
-\begin{tabular}{l l l l}
-a. & {kibrit} & {kibrit}-\{{kabrıt}\} & `match'    \\
-   & {bütün}  & {bütün}-\{{batın}\}   & `whole'    \\
-b. & {mali}   & {mali}-\{{muli}\}     & `Mali'     \\
-   & {butik}  & {butik}-\{{batik}\}   & `boutique' \\
-\end{tabular}
-\end{example}
-
-\noindent \citeauthor{Harrison2001} analyze this in terms of the \citeauthor{Clements1982} underspecification analysis, but it is equally consistent with full specification.
-Reharmonization is simply the spread of the backness specification of the \emph{a} or \emph{u} in the reduplicant.
-On the other hand, (\ref{redupgame}b) shows that this does not obtain for disharmonic roots; presumably, the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication as well.
-Similar results obtain both in Tuvan and in an unrelated Finnish language game \citep{Campbell1986}.
-
-\label{harmonyandsegmentation} A number of studies on Finnish suggest that speakers of that language, which has a vowel harmony system similar to that of Turkish, make use of (dis)harmony to processing running speech.\footnote{Thanks to Charles Yang for bringing these studies to my attention.} \citet{Suomi1997} and \citet{Vroomen1998} generate nonce trisyllabic words by adding a monosyllabic pseudo-prefix to real and nonce disyllabic words, all of which conform to backness harmony.
-These stimuli are auditorily presented to subjects who are asked to press a button when the nonce trisyllable ends with a target nonce disyllable, or a real disyllabic word.
-Speakers are quicker to press the button when the prefix and disyllabic word have different backness specifications.
-These results suggest that speakers are attuned to the fact that disharmonic transisitions are good predictors of word boundaries.
-If speakers have also internalized the converse generalization, that harmonic transistions are more likely to be root-internal, then there is additional evidence that harmony is active not just in Finnish affix alternations but also in roots.
-
-\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect in word-spotting tasks as it does in Finnish: speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by the pseudo-prefix \emph{gölü-}, a disharmonic transistion, than when it is preceded by the pseudo-prefix \emph{golu-}, a harmonic transition.
-\citeauthor{Kabak2010} find that effect of harmony does not obtain for speakers of French, a language which lacks vowel harmony.
-As in Finnish, the results imply speakers have internalized the predominance of root-internal harmony.
-
-It seems that the root-internal harmony bias is in fact learned by Turkish speakers very early.
-The pseudoword spotting experiment has been adapted for 9-month-old Turkish infants by \citet{Kampen2008}.
-Infants are familiarized with harmonic disyllabic pseudowords bearing a pseudo-prefix, which may be harmonic or disharmonic.
-At test time, the infants are played the disyllabic pseudowords in isolation using the head turn preference paradigm.
-Infants show a preference to listen to those pseudowords which were familiarized with a disharmonic pseudo-prefix over those which were familiarized with a harmonic pseudo-prefix.
-This preference is not observed in 9-month-old infants learning German, which also lacks vowel harmony.
-Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants show no such preference.
-
-\subsection{Roundness Harmony}
-
-Roundness harmony is a closely related process with a somewhat narrower scope.
-
-\subsubsection{Phonological description}
-
-Roundness harmony is quite similar to \textsc{Backness Harmony}, with the additional stipulation that the target must be [$+$\textsc{High}].
-
-%\begin{example}[\textsc{Roundness Harmony}]
-%\xymatrix@R=24pt@C=24pt{
-%\txt{[α \textsc{Round}]}\ar@{-}[d]\ar@{--}[drr] &             & \txt{[$+$\textsc{High}]} \\
-%\txt{V}                                         & \txt{C$_0$} & \txt{V}\ar@{-}[u] \\
-%}
-%\end{example}
-
-This rule, in concert with \textsc{Backness Harmony}, accounts for the forms of the dative singular (dat.sg.) and genitive singular (gen.sg.), among other suffixes.
-
-\begin{example}[Turkish nominal suffix allomorphy]
-\begin{tabular}{l l l l l l@{ }l}
-   & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
-a. & {ip}           & {ipi}          & {ipin}         & `rope' & \citep[][216]{Clements1982} \\
-   & {el}           & {eli}          & {elin}         & `hand'    \\
-   & {kız}          & {kızı}         & {kızın}        & `girl'    \\
-   & {sap}          & {sapı}         & {sapın}        & `stalk'   \\
-   & {yüz}          & {yüzü}         & {yüzün}        & `face'    \\
-   & {köy}          & {köyü}         & {köyün}        & `village' \\
-   & {pul}          & {pulu}         & {pulun}        & `stamp'   \\
-   & {son}          & {sonu}         & {sonun}        & `end'     \\
-b. & {boğaz}        & {boğazı}       & {boğazın}      & `throat'  & (TELL) \\
-   & {pelür}        & {pelürü}       & {pelürün}      & `tissue paper' \\
-   & {döviz}        & {dövizi}       & {dövizin}      & `currency' \\
-   & {yamuk}        & {yamuğu}       & {yamuğun}      & `trapezoid' \\
-   & {ümit}         & {ümiti}        & {ümitin}       & `hope'     \\
-\end{tabular}
-\end{example}
-
-\noindent As was the case for \textsc{Backness Harmony}, disharmonic roots are found, and the final vowel of disharmonic roots triggesr suffix harmony.
-
-\subsubsection{External evidence}
-
-Though few studies have directly sought external evidence for speaker's awareness of root \textsc{Roundness Harmony}, two of the findings reviewed in \S\ref{backharmexternal} above speak in support of this claim.
-First, the epenthetic vowel in non-native word-initial onsets
-like \emph{trablus} \alt{} \emph{tırablus} `Tripoli' conform also to \textsc{Roundness Harmony}.
-%\citealp[][247]{Clements1982}). 
-Second, \textsc{Roundness Harmony} participates in reharmonization in \citeauthor{Harrison2001}'s language game: the reduplicated \emph{bütün-batın} exhibits replacement of the second \emph{ü}, a [$+$\textsc{Round}] vowel, with [$-$\textsc{Round}] \emph{ı}.
-
-\subsection{Labial attraction}
-
-%\begin{quote}
-%\ldots{}these generalizations\ldots{}have no observable consequences in the course of the normal use of the language. \citep[][320]{Zimmer1969}
-%\end{quote}
-
-\subsubsection{Phonological description}
-
-\citet[][36]{Lees1966a} describes \textsc{Labial Attraction} as a process by which ``a high, short harmonic vowel is rounded in the second syllable of a disyllabic word whose first vowel is /a/, and whose medial consonant cluster contains a labial /p, b, m, v/, and then it is de-harmonified''.
-This description is transalted into an autosegmental rule below.
-
-%\begin{example}[\textsc{Labial Attraction}]
-%%(after \citealt[][286]{Lees1966b}, \citealt[][171]{Inkelas2001}): 
-%\xymatrix@R=24pt@C=8pt{
-%\txt{[$-$\textsc{Round}]} &                                         & \txt{[$-$\textsc{High}]} & \txt{[$+$\textsc{Labial}]}    & \txt{[$+$\textsc{High}]}\ar@{-}[dr] &         & \txt{[$+$\textsc{Round}]}\ar@{--}[dl] \\
-%%                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C}\ar@{-}[u] &                                      & \txt{V} & \\
-%                         & \txt{V}\ar@{-}[ul]\ar@{-}[ur]\ar@{-}[d] &                           & \txt{C$_0$ C C$_0$}\ar@{-}[u] &                                      & \txt{V} & \\
-%                         & \txt{[$+$\textsc{Back}]}\ar@{-}[urrrr]
-%}
-%\end{example}
-
-The formalization of this rule is naturally complex, and perhaps obscures the fact that \textsc{Labial Attraction} generates exceptions to \textsc{Roundness Harmony}, producing \emph{aCu} sequences (e.g., \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', \emph{samur} `sable'; \citealp[285]{Lees1966b}) rather than the expected \emph{aCı}.
-\citet[286]{Lees1966b} and \citet[311]{Zimmer1969} note the existence of exceptions (e.g., \emph{tavır} `mode') but both agree that they are surprisingly rare.
-
-\citet[225]{Clements1982} and \citet{Inkelas2001}
-
-There is reason to believe that \textsc{Labial Attraction} is at best a sequence structure constraint as it never applies in derived environments.
-If, contrary to fact, there was a \textsc{Labial Attraction} alternation, then one would expect, for example, that the gen.sg.
-of \emph{sap} `stalk' would be *\emph{sapun} instead of the observed \emph{sapın}.
-
-\citet{Lees1966a}
-\citet{Zimmer1969}
-
-\begin{quote}
-\ldots{}decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots~aCu~\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots We conclude that there is no systematic restriction on the set of consonants that may occur medially in roots of the form /\ldots~aCu~\ldots/.
-\citep[][225]{Clements1982}
-\end{quote}
-
-\noindent
-\citeauthor{Clements1982} appear to be suggesting that \textsc{Labial Attraction} implies that \emph{aTu}, where \emph{T} represents a non-labial consonant, should be infrequent, but this fact is not inconsistent with the formulation of the constraint by \citet{Lees1966a,Lees1966b} and \citet{Zimmer1969}; \emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description.
-This is less than conclusive, since the fluctuating vowel alternation also fails to undergo harmony, and might just indicate that the flucutating vowel is epenthesized relatively late in the derivation; the frequency of \emph{aTu} provides only indirect information about \textsc{Labial Attraction} in that it provides a baseline estimate for just how frequent this disharmonic sequence of vowels is.
-
-\citet{Inkelas1997}
-\citet{Inkelas2001}
-
-\begin{quote}
-Lee's rule of \textsc{Labial Attraction}\ldots is not a real generalization about the Turkish lexicon.
-It is not true synchronically, either of native or nonnative items; nor, according to the historical and dialectical literature, does \textsc{Labial Attraction} appear to have been true at any stage going back as far as Old Turkic.
-\citep[][196]{Inkelas2001}
-\end{quote}
-
-zimmer: 319-320 
-inkelas et al.: 196
-
-\subsubsection{Psycholinguistic evidence}
-
-To the author's knowledge, there are no psycholinguistic studies investigating \textsc{Labial Attraction} beyond the experimental results of \citet{Zimmer1969} reviewed in the next section.
-
-\section{Evaluation}
-
-The results reveal a dissociation between the lexical statistics and wordlikeness judgements, a dissociation that finds a natural explanation from the principle of ...
-
-There is some disagreement in the literature about just what \citet{Zimmer1969} finds regarding \textsc{Backness Harmony} and \textsc{Labial Attraction}.
-
-\subsection{Lexical statistics} \label{lexstats}
-
-This choice of statistical test is justified in Section \ref{stattech}.
-
-\subsubsection{Backness harmony}
-
-\subsubsection{Roundness harmony}
-
-%\ex Lexical effects of \textsc{Roundness Harmony} \citep{TELL}: \vspace{6pt} \\
-%\begin{tabular}{l r r r r r}
-%\toprule              & 
-%Corpus                & $p$-value \\
-%\midrule
-%Full TELL             &  
-%Elicited TELL         & 
-%TELL with etymologies & 
-%\end{tabular}
-%\xe
-
-%\citet{Harrison2004} 73\% 
-%of lexical types are harmonic (both back and round)
-
-\subsubsection{Labial attraction}
-
-\begin{quote}
-Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
-\citep[][196, emphasis in original]{Inkelas2001}
-\end{quote}
-
-%\ex Lexical effects of \textsc{Labial Attraction} \citep[][186]{Inkelas2001}: \vspace{6pt} \\
-%\begin{tabular}{l r r r r r}
-%\toprule
-%Corpus                & aPu & aPı & aTu & aTı   & $p$-value   \\ % & corpus size
-%\midrule
-%Full TELL             & 378 & 248 & 446 & 1,140 & 2.83\e{-44} \\ % & 31,236 \\
-%Elicited TELL         & 152 & 265 & 101 & 1,839 & 9.84\e{-60} \\ % & 16,541 \\
-%TELL with etymologies & 128 & 109 &  79 &   470 & 6.56\e{-32} \\ 
-%\bottomrule
-%\end{tabular}
-%\xe
-
-%Etymological subset of TELL
-%Native    & Foreign   \\
-%aBu & aBI & aBu & aBI \\
-%12  & 11  & 84  & 28  \\
-%p = 0.0417
-
-\subsection{Wordlikeness results}
-
-There is some disagreement in the literature about just what \citeauthor{Zimmer1969}'s study reveals about \textsc{Labial Attraction}.
-
-zimmer
-clements
-\citet{Ito1993} 
-%and \citet{NiChiosain1993} 
-argue that the data shows that it holds of native words only
-
-\citet{Inkelas2001} refute this point
-
-\citet{Zuraw2000}:4
-
-\citet{Inkelas2001} present a different variation on the original rule of \textsc{Labial Attraction}
-
-\begin{quote}
-Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
-\citep[][196, emphasis in original]{Inkelas2001}
-\end{quote}
-
-\noindent \citeauthor{Inkelas2001} persistently refer to ``statistical support'', but in fact their study contains no statistics (that is, quantities calculated from a set of data) beyond counts and percentages.
-
-``These results support our view that there is no rule of Labial Attraction'' \citep[][225]{Clements1982}.
-
-%since the opposing are all broader versions than the original formulation by \citeauthor{Lees1966a}.
-
-\citet{Goodman1954}
-
-\subsubsection{Backness harmony}
-
-\begin{example}[Backness harmony wordlikeness forced choices (\citealp{Zimmer1969}:314)]
-\begin{tabular}{l r l r}
-\toprule
-\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
-\midrule
-{pemez} & 30            & {pemaz} & 2  \\
-{tipez} & 24            & {tipaz} & 8  \\ 
-{terüz} & 19            & {teruz} & 13 \\ % roundness violator
-{teriz} & 28            & {terız} & 3  \\
-{tokaz} & 26            & {tokez} & 6  \\ % roundness violator
-\bottomrule
-\end{tabular}
-\end{example}
-
-$\gamma = 0.597$, $p = 5.2$\e{-29}
-
-\subsubsection{Roundness harmony}
-
-\begin{example}[Wordlikeness comparisons, roundness harmony]
-\begin{tabular}{l r l r}
-\toprule
-\multicolumn{2}{l}{harmonic} & \multicolumn{2}{l}{disharmonic} \\
-\midrule
-{pörüz} & 32            & {pöriz} & 0  \\
-{tatız} & 20            & {tatuz} & 12 \\
-{tüpüz} & 31            & {tüpiz} & 1  \\
-{takız} & 22            & {takuz} & 10 \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-$\gamma = 0.641$, $p = 7.7$\e{-27}
-
-%tamaz & 26 & tamoz & 6  \\
-%putoz & 25 & putaz & 7  \\
-
-\subsubsection{Labial harmony}
-
-\begin{example}[Wordlikeness comparisons]
-\begin{tabular}{l r l r}
-\toprule
-\multicolumn{2}{l}{{a}P{u}} & \multicolumn{2}{l}{{a}P{ı}} \\
-\midrule
-{pamuz} & 17       & {pamız} & 15 \\
-{tafuz} & 21       & {tafız} & 11 \\
-{tapuz} & 17       & {tapız} & 15 \\
-{mavuz} & 16       & {mavız} & 16 \\
-{tabuz} & 16       & {tabız} & 16 \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-$\gamma = 0.087$, $p = 0.101$
-
-\section{Conclusions}
-
-\citet{Becker2011}
-
-It is 
-%further, even long-distance co-occurrence restrictions frequently are ``conjunctive'': for instance, \citet{Gallagher2009} report that in Chol, ejectives must be totally identical, that is to say, 
-%\citet{Bakovic2005b}
-
-This has an ad hoc nature to it; 
-There are two senses in which this objection is ad hoc. 
-First, \citeauthor{Becker2011} appeal to no particular theory of the naturalness of processes or static constraints which excludes \textsc{Labial Attraction}. 
-Secondly, this appears to be a minority view: \textsc{Labial Attraction} was considered a true generalization by early specialists
-\citep[e.g.,][]{Lees1966a}, and despite \citeauthor{Zimmer1969}'s suggestive psycholinguistic results, reviewed above, it also been treated as a plausible constraint by later theorists \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995a,Zuraw2000}.
-Further, there is a real danger that if the label ``unnatural'' 
-%Labial Attraction} 
-describes an impossible structural change or structural description, that one will fail to account for earlier forms of Turkish or sound changes therein.
-
-%Classical Arabic adjectives often have stative verbs in which the root is imposed onto the template CaCuCa:
-Classical Arabic adjectives often have stative verbs in which the root melody is rewritten with \emph{a\ldots{}u}.\footnote{Thanks to Uri Horesh (p.c.) for assistance with this data.}
-
-\begin{example}[Arabic stative verbs]
-\begin{tabular}{l l l l}
-\buf{}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
-\buf{}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
-\buf{}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
-\end{tabular}
-%b. & saʁiir & }`become small'     & (cf. \emph{saʁiːr} `small') \\
-\end{example}
-
-% FORMER 3.3.2: Diachronic factors
-
-% etymological issues
-\citet{Inkelas2001}
-
-\begin{example}[\textsc{Labial harmony} and etymology in TELL \citealp{Inkelas2001}:187]
-\begin{tabular}{l r r r}
-        & {a}P{u} & {a}P{ı} & $p$-value \\
-native  & 12                & 11                & \multirow{2}{*}{0.042} \\
-foreign & 84                & 28                \\
-\end{tabular}
-\end{example}
-
-\citet{NiChiosain1993} and \citet{Ito1995b}
-
-``sonority projection''
-Ther
-
-\citet{Daland2011b} finds that this can be learned easily from positive data but a number of different psych
-
-%It is important to note that the nature of the \emph{blick}/\emph{bnick} contrast is no longer a mystery. 
-%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
-It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails.
-If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
-%\footnote{I note a potential problem with this account: \citet{Davidson2005,Davidson2006a} argue that this ``excrescent'' schwa is distinct from the both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.}
-%\noindent A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions.
-Any explicit theory of static constraints will have to impose formal restrictions on possible co-occurrence restrictions.
-The set of possible constraints must respect results in formal learning theory, and may also be constrained by markedness considerations, for example.
-%A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
-This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).
-%Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
-In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.
-%In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word. 
-%The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements. 
-%The data below, drawn from Turkish, shows that 
-%This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
-Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements.
-Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
-%\begin{unlabeledexample}
-%$\displaystyle \gamma = \frac{C - D}{C + D}$
-%\end{unlabeledexample}
-%\footnote{Any readers who are skeptical that \citeauthor{Chomsky1965}'s intuitions generalize will find support for their claim in the wordlikeness judgement data reported in Appendix \ref{albright}.}
-%As a quick validation of this claim, consider the results reported by \citet{Albright2007} and discussed in the previous chapter.
-English speakers asked to rate non-words on a scale from 1-7, where 7 indicates ``most like English'', favor non-words [blʌs, blɑd] (average scores 4.67 and 5.13) over [bnʌs, bnɑd] (average scores 2.06, 2.00).
+%Whereas the classic analysis of this phenomenon, by \citet{Inkelas1997}, does not make use of lexical exceptionality, \citeauthor{Becker2011}'s analysis of final aspirated stops and affricates as the underlying form requires that _every_ lexical entry with a final stop or affricate be treated as exceptional according to some phonological constraint.
+This projection of a lexical contrast into the grammar has another potential flaw: the constraints used to produce the voiced variants in, e.g., the dative, are not surface true.
+Consider the constraint *VpV used to induce alternations like \emph{kap}-\emph{kabı} `coat'.
+Ignoring the case of non-alternating root-final \emph{p} (e.g., \emph{ip}-\emph{ipi} `rope'), this is still not a surface-true generalization: \emph{p} occurs freely in intervocalic position: \emph{ahtapot} `octopus', \emph{köpük} `bubble', \emph{öpücük} `kiss'.
+More serious is the presence of root-internal intervocalic \emph{p} in roots which simultaneously exhibit the root-final \emph{p}-\emph{b} alternation: \emph{supap}-\emph{supabı} `valve', \emph{hipermetrop}-\emph{hipermetrobu} `far-sightedness'.
+One prediction shared by virtually all theories of lexical exceptionality is that a lexical entry cannot simultaneously be a target for, and exception to, a phonological generalization, but this precisely what is needed to account for the behavior of \emph{supap}.
+% \citet{inkelas1997}
+%küp-kübü `cube'
+%kasap-kasabı `butcher'
+%Further, the empirical status of the core instances of NDEB have recently been quite stridently disputed \citep{InkelasInPress}.
+There is some reason to suspect that NDEB is a symptom of interface restrictions on phonological application, rather than a diagnosis itself.

### 2012-10-10 4.tex
Writing on front matter, chapters 1-4

@@
-\chapter{Structural and accidental gaps} \label{clusters}
+\chapter{Structural and accidental gaps} 
+\label{clusters}
 
-Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1979}, \citealp[410f.]{Stanley1967}).
+Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}).
 Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the \emph{communis opinio} (including many of the authors cited above) holds that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
@@
 
+%\citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''
+%\citet[?]{Lombardi1999} says that English lacks devoicing in Level II.
+
 \paragraph{Stressed lax vowels} When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
@@
 \begin{example}[\textsc{Obstruent Voice Assimilation}]
-\xymatrix@R=24pt@C=24pt{
-\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{--}[dr] &                               \\
-\txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
-\txt{[$-$\textsc{Sonorant}]}                  & \txt{[$-$\textsc{Sonorant}]} \\
-}
+%\xymatrix@R=24pt@C=24pt{
+%\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{--}[dr] &                               \\
+%\txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
+%\txt{[$-$\textsc{Sonorant}]}                  & \txt{[$-$\textsc{Sonorant}]} \\
+%}
+$\begin{bmatrix} -\textsc{Son} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Voi} \end{bmatrix}~/~\gap~\begin{bmatrix} =\textsc{Voi} \\ -\textsc{Son} \end{bmatrix}$
 \end{example}
@@
 \citealp[546]{Rice2009d} on compounds in Slave).
-In fact, \citet{Mattys2001b} find that 9-month-old infants are sensitive to the contrast between [f.t] and exceptional (and unattested) [v.t], treating the latter as an indication of a word boundary and this heuristic is also available to adults \citep{Brown1956,McQueen1998}.
+In fact, \citet{Mattys2001b} find that 9-month-old infants are sensitive to the contrast between [f.t] and exceptional (and unattested) [v.t], treating the latter as an indication of a word boundary and this heuristic is also available to adults \citep{Brown1956,McQueen1998b}.
 Less explanatorily, exceptional roots can be marked [$-$\textsc{Obstruent Voice Assimilation}].
 
-\subsubsection{Nasal place assimilation} \label{npa}
+\subsubsection{Nasal place assimilation} 
+\label{npa}
 
@@
 
+% mention Jensen here
+
 \begin{example}[\textsc{Nasal Place Assimilation}]
-\xymatrix@R=24pt@C=24pt{
-                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
-\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
-\txt{[$+$\textsc{Nasal}]} & \txt{[$-$\textsc{Sonorant}]}              \\
-}
+%\xymatrix@R=24pt@C=24pt{
+%                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
+%\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
+%\txt{[$+$\textsc{Nasal}]} & \txt{[$-$\textsc{Sonorant}]}              \\
+%}
+$\begin{bmatrix} $+$\textsc{Nas} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \end{bmatrix}~/~\gap{}~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ -\textsc{Son} \end{bmatrix}$
 \end{example}
@@
 As shown by the results of the Fisher test, this absence is unlikely to be due to chance.
+
+\begin{example}[\textsc{Degemination}]
+$\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ =\textsc{Nas} \\ =\textsc{Cont} \end{bmatrix}~\goesto~\zero~/~\gap~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ =\textsc{Nas} \\ =\textsc{Cont} \end{bmatrix}$
+\end{example}
+
+% consider what coda-final consonants there are
 
@@
 
-Sparse distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2009} to word \citep{Teahan1998,Baroni2009} and phoneme \citep{Belevitch1956,Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Mandelbrot1954,Miller1957,Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
+Sparse distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2011b} to word \citep{Baroni2009} and phoneme \citep{Belevitch1956,Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Mandelbrot1954,Miller1957,Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
 The importance of sparsity in this context is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps.

### 2012-10-10 abs.tex
more writing
Writing on front matter, chapters 1-4

--- 2011-12-01 abs.tex
@@
-This dissertation considers two types of ``gaps'' in the lexicon.
-In both cases, it is argued that current generative models can be modified minimally to account for the apparent overgeneration of gapped structures.
-The first part of the dissertation considers the underlying phonological structure of morphs, specifically medial consonant clusters in English.
-A number of prior studies have argued that the grammar must have a separate ``phonotactic'' module to account
+This dissertation argues for that phonotactic knowledge is primarily derived not from static generalizations but rather from phonological alternations and prosodic parsing.
+Chapter \ref{intro} outlines the proposal, historical precursors, and predictions.
+%especially with regards to order of acquisition.
 
-at the level of prosodic word. In both cases, it is argued that the i
+Chapter \ref{gradience} considers evidence from wordlikeness rating tasks.
+It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
+A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements at least as acccurately as state-of-the-art gradient wellformedness models.
+Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.
+
+Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns.
+It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.
+
+Chapter \ref{clusters} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English.
+Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound, whereas phonological alternations impose robust restrictions on the cluster inventory.
+The remaining gaps in the cluster inventory are attributed to the sparse nature of the lexicon, not static phonotactic restrictions.

### 2012-10-10 ack.tex
Writing on chapters 2, 4
Writing on front matter, chapters 1-4

--- 2012-02-16 ack.tex
@@
 So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy.
-More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, whose influence can be seen throughout. 
-Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are responsible for creating an environment at Penn where a student could write this admittedly contenious dissertation. 
-And though his name appears nowhere else on this document, it is hard to imagine this work could exist without Bill Labov.
+More than anyone else, I owe a debt of gratitude to Charles Yang, who has done (and put up with) so much for me.
+I am grateful to my committee members, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment for quantitative and theoretical approaches to meet.
+And though his name does not appear again in this document, so much here has been inspired by Bill Labov.
 
-My journey in linguistics begins with the mentoring of Jennifer Cole. 
-Steve Anderson graciously agreed to serve on my committee. 
-Morris Halle told me to never apologize.
+The faculty, staff, and students of the Institute For Research in Cognitive Science have provided me with not merely an office, but a home, complete with dated early-90s decor.
+Thanks to my officemates, Chandan Narayan, John Trueswell, Laurel Sweeney and the rest of the staff there, as well as Amy Forsyth back at Williams Hall.
 
-I would like to thank my fellow students at Penn, but especially my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody. 
-Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document. 
-Maryam Bakht kept me centered.
+So many of my fellow students at Penn have had a hand in this document, none so much as Toni Cook and Laurel MaKenzie, who have supported me so much over the past few years.
+Without Joe Fruehwald, Marielle Lerner, Constantine Lignos, Brittany McLaughlin, and Hilary Prichard, it wouldn't have been nearly as much fun.
+Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar.
 
-My gratitude also goes out to the administrative assistants at 3401 Walnut and 255 S.
-36th, Lauris Olson at the Van Pelt Library, Katherine McEldoon for a helpful review of numerical cognition research, and Carly Fisher for her unflagging support and friendship.
+The ideas I first encountered under the tutelage of Jennifer Cole and Richard Sproat at the University of Illinois continue to guide my research.
+Insofar as I have acknowledged historical precursors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee.
+Morris Halle taught me to speak clearly and to never apologize.

### 2012-10-10 pre.tex
Writing on chapters 2, 4
Writing on front matter, chapters 1-4

@@
-Perhaps the most important fact for the study of speakers' internal knowledge of language is Wilhelm von Humboldt's observation that speakers make ``infinite use of finite means'' CITE \citep{Chomsky1966}.
+In a justly-famous study, \citet{Pierrehumbert1994} notes the general absense of velar-labial consonant clusters in English (though note the existence of a few exceptions like \emph{pigment} and \emph{rugby}).
+When I first encountered this study, as an undergraduate, I simply could not integrate this arbitrary gap with a view of the language faculty as the inexhaustable fount of productivity responsible for such wonders as [wʌgz] and \emph{Colorless green ideas sleep furiously}.
+This shocking observation was filed away for further meditation, in the hope that one day I might grok.
 
-An enduring demonstration of ``infinite use'' is the famous emph{wug-test} \citep{Berko1958}.
-Jean Berko-Gleason introduced English children aged ?-?
-to pictures of concrete objects (such as a small bird), labeled them with a nonce word (``that's a \emph{wug}''), and then asked them to finish the sentence
+The atoms of linguistic theory tend to be sparsely distributed, with ``few giants'' (/t/, \emph{the}) and ``many dwarves'' (/ʒ/, \emph{zymurgy}) in any sample.
+Sparsity is exacerbated for complex objects: however common \emph{the} is, any two word sequence containing it will be far less frequent.
+This dissertation began in earnest with a suggestion from Charles Yang that perhaps sparsity---the missing and rare data which have long plagued statistical approaches to language learning and processing---might be the only reason, e.g., English has no words containing a [k.p] cluster.
 
-The children in the study produced the same response you may have: the plural is \emph{wugs} (where $\langle$s$\rangle$ writes a final [z]).
+Linguists-in-training are inculcated, early and often, with an respect for productivity---the ``infinite use of finite means'' \citep[8]{ASPECTS}---over all else.
+Psycholinguists have been more free to pursue less combinatorial, less granular approaches to language.
+Many developmentalists have argued that the impoverished nature of young children's speech reflects whole phrases stored in memory, not phrase structure rules.
+It is therefore informative that the relatively sparse distribution of phrases in child speech is a practical \emph{necessity} of generativity and sparsity \citep{Yang2011b}.
+Natural language engineers, unburdened by the need to construct models which conform to the cognitive constraints imposed by human minds, have endowed statistical parsers with phrasal storage mechanisms ``bilexical'' rules \citep[e.g.,][]{Collins1999}, but these offer little no additional coverage \citep{Bikel2004}.
+I suspect this is not because stored phrases are inherently uninformative, but rather because only a tin small number of two-word sequences can be reasonably expected to occur with sufficient frequency in any corpus.
 
-Of course, children have never heard the plural of these nonce words, but they extend /-z/ to these new words.
+Similarly, theorists of lexical processing postulate that complex words are stored in memory to account for the facilitatory effect of whole word frequency on reaction times to complex words \citep[e.g.,][]{Alegre1999,Baayen1997a,Sereno1997}.
+But this effect of word frequency obtains regardless of whether or not whole word storage is postulated \citep{Taft2004b,LignosInPress}.
+Furthermore, in languages with rich agglutinative morphology, the role played by a whole word storage mechanism, should one be evidenced, would necessarily be quite marginal \citep{Hankamer1992,Chan2008}.
+According to one estimate, children acquiring American English hear 165,000 word tokens a week \citep{Swingley2007}.
+At that rate, it would take more than two months for a child to encounter the 1.5 milllion inflectional variants of a single verb in Archi \citep[467]{Kibrik1998},
+%, cited in \citealt[122]{Bauer2001}), 
+even if adult speech consisted of nothing but orderly recitation of the inflectional forms of this one verb, one after the other, without repetition.
 
-But what of the ``finite means'' from which grammars spring? On this, a good deal less has been said.
-a study of spaekers' knowledge of natural
-The most important fact for
+By
+some
 
-In this dissertation, I critique two such epicyclic proposals, and present solutions using independently-motivated grammatical machinery to account for the data.
+I can only hope that this humble theory has even a fraction of the influence on phonotactic thinking that
 
-The first half of this dissertation began with a suggestion from Charles Yang that the sparse nature of the lexicon might itself account for at least some phonotactic gaps.
-The success of the resulting study, in chapter four, led me to question the received wisdom that speakers' knowledge of phonotactic generalizations is gradient and potentially independent of the phonological system.
+There are two ways to read this thesis.
+First, it can be regarded as evidence for a particular, falsifiable theory of phonotactic learning and knowledge.
+If, for instance, some future study finds that English speakers rate nonce words containing the unattested [k.p] cluster to be less well-formed than similar words containing [k.t], a cluster found in dozens of English words, then this theory will need to be seriously revised.
+Suggestive evidence comes an auditory wordlikeness task performed by \citet[782]{Munson2001} which includes this contrast.
+Speakers rated [bɪk.pɘt] and [fæk.pɘt] slightly more English-like than [bɪk.tɘp] and [fæk.tɘp], respectively, and drew no contrast between the pair [nɪk.pɘt]-[nɪk.tɘp].
 
-These results support a null hypothesis in which phonotactic knowledge is derivative of the language acquisition device and phonological processes.
-While I present this theory as a falsifiable scientific hypothesis, an alternative would treat this proposal as a \emph{negative heuristic}, a
+Other claims are not essential to the theory here, but are rather principled null hypotheses, and evidence to the contrary would result only small modifications to the theory.
+For instance, it is not claimed that wordlikeness judgements are inherently binary, but rather that the evidence currently available is insufficient to justify the considerable additional complexities of gradient wellformedness models.
 
-guides us away from ``phonotactic grammar'' as an explanatory device.
+Under another reading, the claims here represent a \emph{negative heuristic} guiding the phonologist away from static phonotactics as an explanatory device.
+Then this work is a claim about something grammars do not contain, not a theory of what they do, much like \emph{Richness of the Base} (RotB), an important intellectual precursor to this work (see \citealt[6]{Bye2001} and \citealt[73]{PE} for further discussion).
+\emph{Lexicon optimization}, however, is an example of a theory which produces grammars which conform to RotB.
 
-even those who are not inclined to accept this proposal should note how much of the coverage traditionally attributed to
+A strict application of this negative heuristic produces some interesting predictions.
+If speakers are shown to have internalize a phonotactic generalization, then there must be a phonological process to blame.
+One clear case where processes, and not merely restrictions on underlying representations, are implicated is in Quecha, where ejective co-occurrence restrictions are reflected in speakers' difficulties in producing nonce words containing multiple ejectives \citep{GallagherInPress}.
+
+This negative heuristic also allows for the possibility of using psycholinguistic evidence to decide between competing phonological analyses.
+For instance, there is a long-standing debate concerning the representation of Sanskrit aspiration alternations like those represented by \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
+One analysis, going back to the ancient grammarian Pāṇini \citep{Sag1974} and to \citet[\S141f.]{Whitney1889}, posits an underlying representation /budh-/ with aspiration shifting leftward in certain contexts: modern proponents of this analysis include \citet{Borowsky1983}, \citet[][59f.]{Hoenigswald1965}, \citet{Kaye1985}, \citet{Sag1974,Sag1976} \citet{Schindler1976}, \citet{Stemberger1980}, and \citet[][\S141f.]{Whitney1889}.
+An alternative analysis posits an underlying representation /bhudh-/ and an apsirate dissimilation process (\citealt{Anderson1970}, \citealt{Hoard1975}, \citealt[\S3.2]{Kiparsky1965}, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt[109f.]{Zwicky1965}).
+It is not my intention to adjudicate between these competing analyses, or to evaluate the claim that both accounts are deeply flawed \citep{Janda1989}, but simply to note that they make different predictions under the theory proposed here.
+While the former, aspirate shift analysis denies the existence of diaspirate roots, this fact lacks synchronic motivation, so the co-occurrence restriction may go unlearned.
+On the other hand, under the aspirate dissimilation account, underlying diaspirate roots exist but speakers internalize synchronic processes responsible for their absence on the surface.
+
+%\citet[38f.]{MacEachern1999} discusses a similar restriction on co-occurign aspirate roots in Ofo, but also notes, following \citet{DeReuse1981}, that this generalization holds both within roots and also within complex words.
+For instance, the aspiration of ['osk\asp{}a] `crane' disappears when combined with [a'f\asp{}ã] `white': [oskaf\asp{}ã] `white or American egret'.
+Under the negative heuristic, the root co-occurrence restriction and this alternation are both the result of a phonologically general process of aspirate dissimilation.

### 2012-10-11 1.tex
Completion of draft of chapter 3

@@
 
-\section{\emph{Táxis} and generativity}
-%\emph{τάξις}
+One question is: why bother?
 
-(\emph{táxis} `order, arrangement'; the prefix \emph{syn-} means `together')
+It is certainly possible to imagine otherwise.
+\citet[][320]{Zimmer1969} writes that phonotactic generalizations ``have no observable consequences in the course of the normal use of the language'', and \citeauthor{PE} echo this sentiment more recently:
 
-This etymology is almost completley uninformative about the modern study of syntax. 
-The notion of derivation, most associated with the transformational grammars of Zellig Harris and Noam Chomsky
+\begin{quote}
+Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for.
+\citep[][18]{PE}
+\end{quote}
 
-This etymology gives little clue to just what syntacticians do, however.
-This is not merely a consequence of the ``generative revolution'', as the predecessors of transformational grammars associated with Zellig Harris and Noam Chomsky were also more than theories of arrangement, but also instantiate the notion of derivation.
+This represents a principled null hypothesis, but can be quickly dismissed in light of speaker's ready judgements of possible and impossible words observed by \citeauthor{Chomsky1965}.
+Metalinguistic judgements of these sorts, reviewed in Chapter \ref{wordlikeness}, are not the only task which exemplify phonotactic knowledge; infants and adults are also thought to make use \emph{possible word constraint} in more quotidien tasks like recognizing words in running speech \citep[e.g.,][]{Brown1956,Mattys1999,McQueen1998b,Norris1997}.
 
@@
 
+This is precisely the case for
+
 \subsection{Exceptionality}
@@
 
-There are very few investigating infants' knowledge of phonological alternations.
-\citet{White2008} expose 8.5-month-old and 10-month-old infants to an artificial language in which the voicing of fricatives is contrastive, but voiced and voiceless variants of stops are in a complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively.
-After familiarization, infants at both ages prefer to listen to stimuli which preserve the complementary distribution over those which disrupt it (e.g., \emph{na poli}, \emph{rot boli}).
-As \citeauthor{White2008} note, this suggests that infants have grouped \emph{pevi} and \emph{bevi} together and extracted the stop voicing alternation from the artifical language.
+Very few studies have investigated young infants' knowledge of phonological alternations.
+One exception is a fascinating study by \citet{White2008}.
+Simplifying somewhat, the experimenters expose 8.5-month-old and 10-month-old infants to an artificial language in which the voicing of fricatives is contrastive, but voiced and voiceless variants of plosives are in a complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively.
+After familiarization, infants at both ages prefer to listen to stimuli which preserve the complementary distribution over those which disrupt it (e.g., \emph{na-poli}, \emph{rot-boli}).
+As \citeauthor{White2008} note, this suggests that infants have both extracted the plosive voicing alternation and grouped \emph{pevi} and \emph{bevi} together.
 
@@
 By 8 months of age, infants are able to locate familiar and novel words in utterances \citep{Jusczyk1997,Seidl2006}.
+
 \section{Outline of the dissertation}
 
-The remainder of the dissertation consists of three case studies which provide support for the novel and contentious predictions of the minimal phonotactic model sketched above
+The remainder of the dissertation consists of three case studies which provide support for the novel and contentious predictions of the minimal phonotactic model sketched above.
 
@@
 It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
-A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' gradient wordlikeness judgements as well as state-of-the-art gradient wellformedness models.
-The gradient models do not contribute any additional coverage once categorical wellformedness is accounted for.
+A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements at least as acccurately as state-of-the-art gradient wellformedness models.
+Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.
 
@@
 It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.
-%Some arguments against Becker et al.
 
 Chapter \ref{clusters} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English.
-Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound.
-On the other hand, phonological alternations impose robust restrictions on the cluster inventory.
-The remaining gaps in the cluster inventory are due to the sparse nature of the lexicon, not static phonotactic restrictions.
+Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound, whereas phonological alternations impose robust restrictions on the cluster inventory.
+The remaining gaps in the cluster inventory are attributed to the sparse nature of the lexicon, not static phonotactic restrictions.
 
 %% NOTES HERE
+%
 %a. & \emph{B}[uːx]          & \emph{B}[yːç]\emph{er} & `book(s)' \\
@@
 %9-month-old infants already have learned the predominant stress pattern of their language \citep{Jusczyk1993a}.
+%\emph{τάξις}

### 2012-10-11 2.tex
Completion of draft of chapter 3

@@
 
+%\citet{Fischer-Jorgensen1952}
 %\citet{Saporta1958}
@@
 
-Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no attention in the literature.
-At the risk of explaining what might be no more than an an accidental gap in the historical record, the novel aspects of the \emph{SPE} model---gradience derived from similarity to existing lexical entries---seem to have been overshadowed by compelling arguments against \citeauthor{SPE}'s assumption \citep[see also][]{Halle1962,Stanley1967} that wordlikeness contrasts derive solely from properties of underlying forms.
+Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature.
+At the risk of explaining what might be no more than an an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms.
 \citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
@@
 For example, consider the concept \emph{odd number}.
-This seems to have a clear definition, a precise description\ldots{}No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd.
+This seems to have a clear definition, a precise description.
+[\ldots{}] No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd.
 No odd number seems odder than any other odd number.

### 2012-10-11 3.tex
Completion of draft of chapter 3

@@
 
-\citet{Chomsky1965} argue that speakers internalize language-specific generalizations about possible and impossible words, but it is certainly possible to imagine otherwise.
-\citet[][320]{Zimmer1969} writes that phonotactic generalizations ``have no observable consequences in the course of the normal use of the language'', and \citeauthor{PE} echo this sentiment more recently:
+There is a long tradition of distinguishing between structural phonotactic gaps and those which are simply accidental \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.
+In purely statistical terms, an accidental gap is one which is likely to arise without any antecedent cause, i.e., by chance.
+The contribution of early generative discussions of this contrast \citep[e.g.,][]{Halle1962,Chomsky1965,SPE} is to ground the definitions of \emph{structural} and \emph{accidental} gaps in mentalism: an accidental gap is one which which lacks a mental antecedent because it goes unlearned by speakers.
+
+Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, takes less interest in the question of what generalizations speakers internalize.
+In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique to validate a generalization about Arabic root consonant co-occurrence facts originally due to \citet{Greenberg1950}.
+The assumption appears to be that statistically valid static co-occurrence restrictions will be reflected by wordlikeness judgements.
+While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}.
+In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:
 
 \begin{quote}
-Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for.
-\citep[][18]{PE}
-\end{quote}
-
-This represents a principled null hypothesis, but can be quickly dismissed in light of speaker's ready judgements of possible and impossible words observed by \citeauthor{Chomsky1965}.
-As shown in Chapter \ref{gradience}, wordlikeness tasks implicate prosodic wellformedness and cannot be reduced to purely statistical generalizations.
-Metalinguistic judgements are not the only task in which phonotactic knowledge is on display; speakers are also thought to make use \emph{possible word constraint} in more quotidien tasks like recognizing words in running speech \citep[e.g.,][]{Brown1956,McQueen1998b,Norris1997}.
-
-struct/act
-copenhagen
-pure statistical critique
-
-\citet{Fischer-Jorgensen1952} and \citet{Vogt1954}
-note the inherent difficulty of distinguishing between structural and accidental phonotactic gaps.
-
-Even if the criteria differentiating structural and accidental generalizations were purely statistical, an accidental generalization is one which lacks an antecedent cause, and it falls to the linguist to delimit a theory of these causes.
-
-\citeauthor{Chomsky1965} were not the first linguists to take an interest in possible wordhood.
-
-The early generativists identify these causes in mentalism, grounding the study of accidental gaps in what patterns are internalized, and ultimately how they are learned and revealed in psycholinguistic tasks.
-Consequently, for the generativist an accidental phonotactic generalization can only be one which speakers do not learn, do not attend to, and thus a generalization which is without mental antecedent.
-
-Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, expresses limited interest in what generalizations speakers internalize.
-In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique, a chi-square test, in an attempt to validate a generalization about Arabic root consonant co-occurrence facts, observations originally due to \citet{Greenberg1950}.
-The assumption appears to be that statistically valid static co-occurence restrictions will be reflected by wordlikeness judgements.
-While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies in the \citeauthor{McCarthy1988}ian model; rather, the phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Dmitrieva2008a,Dmitrieva2008b,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}.
-In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} makes the tacit assumption of this approach explicit, claiming that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers.
-
-\begin{quote}
-\ldots{} the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
+\ldots{} the patterns outlined above are statistically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
 \end{quote}
@@
 There are several arguments against this inference.
-First, linguists since Saussure have noted diachronic causes for statistical generalizations which lack any synchronic reality.
-For instance, Saussure \citeyear[202f.]{CLG} recognizes the a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}, and that this makes intervocalic \emph{s} rare in Classical Latin.
-However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.} Secondly, wordlikeness generalizations are constrained by negative results in learnability (e.g., \citealp{Gold1967}; see \citealt{Yang2012} for a recent review), and may also be restricted by markedness constraints \citep[e.g.,][]{Becker2011,Hayes2009,Moreton2002}.
-Finally, there are a number of results which suggest patterns obvious to the linguist go unlearned \citep{Becker2011,Hayes2006,Zimmer1969}.
+First, linguists have long noted statistical generalizations which have a diachronic cause but which lack any synchronic reality.
+One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}), who recognizes the a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}, and that Classical Latin intervocalic \emph{s} is rare as a result.
+However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.} Similarly, [ʃ] in Modern English derives from Old English [sk] (e.g., \emph{fisc} 'fish'), and since long vowels before complex codas are not found in Old English, [ʃ] is rarely preceded by long vowels in word-final syllables to this day.
+\citet{Iverson2005} label this constraint ``phonologically accidental'', since a millennium of speakers have produced coinages (e.g., \emph{posh}, \emph{swoosh}) and borrowings (e.g., \emph{douche}, \emph{capiche}) in utter disregard of this generalization.
+Secondly, wordlikeness generalizations are absolutely constrained by negative results in learnability (see \citealt{Yang2012} for a recent review).
+Finally, there are a number of results which suggest patterns which were obvious to the linguist go unlearned \citep[e.g.,][]{Becker2011,Hayes2006,Hayes2009,HayesInPress}.
 This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations.
-Consequently, grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may not be appropriate models of speakers' phonotactic knowledge.
+The danger is that grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may not match speakers' knowledge.
 
-This chapter is a comprehensive investigation of three well-known phonotactic generalizations in Turkish, as well as the results of a wordlikeness task designed to investigate the synchronic reality of these generalizations \citep{Zimmer1969}.
-This study merits reconsideration because there is widespread disagreement about the analysis of \citeauthor{Zimmer1969}, and prior discussions fail to relate the behavioral data to lexical statistics and competing formalizations of the constraints involved.
-One of these constraints, \textsc{Labial Attraction}, is shown to be highly statistically reliable but to be inactive in the wordlikeness task.
-This undermines the widespread use of statistical criteria to determine which of the vast number of phonotactic generalizations are internalized by speakers and which are not.
-The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is phonologically inactive a well.
+This chapter is a comprehensive investigation of three well-known phonotactic generalizations in Turkish using lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}.
+\citeauthor{Zimmer1969}'s experiments merit reconsideration because there is widespread disagreement concerning the results and prior discussions fail to relate the behavioral data to lexical statistics and to competing formalizations of the generalizations involved.
+
+One of these generalizations, \textsc{Labial Attraction}, is shown to be highly statistically reliable but to be inactive in the wordlikeness task.
+This further undermines the widespread use of statistical criteria to determine which of the vast number of statistically reliable phonotactic generalizations are internalized by speakers and which are not.
+The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is phonologically inactive, corresponding to no phonological process in the language.
 
@@
 \citet{Lees1966b,Lees1966a} proposes three constraints on Turkish vowel sequences, constraints which are the focus of many subsequent studies.
-In this setion, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence suggesting that speakers internalize restrictions.
+In this section, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence bearing on speakers' knowlege of the restrictions.
 The following feature specification for the eight vowels of Turkish vowels is assumed throughout.
@@
 
-Two non-conventional notations are used in this chapter.
-First, directional application conditions are used when necessary to model the directionality of harmony, rather than ad hoc contextual restrictions (e.g., that the trigger be aligned with the left of a prosodic word).
-Directional application provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}) and simplifies the analysis of harmony in Turkish \citep[209f.]{Anderson1974}.
-Further, \citet{Johnson1972} and \citet{Kaplan1994} prove that rules specified for directional application are formally learnable.
+Two less familiar notations are used in this chapter.
+First, directional application conditions are used to derive the left-to-right, iterative properties of.
+Directional application explains provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}).
+In fact, harmony in Turkish is one such case, insofar as simultaneous application makes the wrong predictions.
+Further, \citet{Johnson1972} and \citet{Kaplan1994} prove that rules specified for directional application are no more complex in the terms of formal learnability than that do not.
 Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used.
-Following \citet{McCawley1973}, a structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses the two potential counterexamples against this restrict, both adduced in \emph{SPE} (352-353).}
+Following \citet{McCawley1973}, a structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses well-known counterexamples against this restriction adduced in \emph{SPE} (352-353).}
 
@@
 
-If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllablic roots to contain only [$+$\textsc{Back}] or [$-$\textsc{Back}] vowels, a tendency which will be quantified below.
+If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only [$+$\textsc{Back}] or [$-$\textsc{Back}] vowels, a tendency which will be quantified below.
 \textsc{Backness Harmony} also triggers alternations in inflectional suffix vowels.
@@
 
-\begin{example}[The Turkish nominative] \label{turknom}
-\begin{tabular}{l l l l@{ }l}
+\begin{example}[The Turkish nominative]
+\label{turknom}
+\begin{tabular}{lllll}
    & \emph{nom.sg.} & \emph{nom.pl.} \\
@@
 b. & {neden}        & {nedenler} & `reason'       & \citep{TELL} \\
-   & {kiler}        & {kilerler} & `pantry'       \\ % front
-   & {pelür}        & {pelürler} & `tissue paper' \\ % back
-   & {boğaz}        & {boğazlar} & `throat'       \\ % back
-   & {sapık}        & {sapıklar} & `pervert'      \\ % back
+   & {kiler}        & {kilerler} & `pantry'       \\
+   & {pelür}        & {pelürler} & `tissue paper' \\
+   & {boğaz}        & {boğazlar} & `throat'       \\
+   & {sapık}        & {sapıklar} & `pervert'      \\
 \end{tabular}
@@
 
-\begin{example}[Exceptional Turkish nominatives] \label{turkexcept}
-\begin{tabular}{l l l l@{ }l}
-   & \emph{nom.sg.} & \emph{nom.pl.} \\
+\begin{example}[Exceptional Turkish nominatives] 
+\label{turkexcept}
+\begin{tabular}{lllll}
+   & \emph{nom.sg.} & \emph{nom.pl.}             &                    \\
 a. & {mezar}        & {mezarlar} & `grave'       & \citep{TELL}       \\
@@
    & {sabun}        & {sabunlar} & `soap'                             \\
-b. & {saat}         & {saatler}  & `hour, clock' & \citep{Goksel2005} \\
-   & {harf}         & {harfler}  & `(alphabetic) letter' \\ 
-   & {etol}         & {etoller}  & `fur stole' \\
+b. & {etol}         & {etoller}  & `fur stole'   & \citep{Goksel2005} \\
+   & {saat}         & {saatler}  & `hour, clock' 	                  \\
+%  & {harf}         & {harfler}  & `(alphabetic) letter' \\ 
 \end{tabular}
@@
 There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} despite the fact that their final root vowel is [$+$\textsc{Back}].
-While it is uncontroversial that the disharmonic suffixes of (\ref{turkexcept}b) are no more than very sporadic exceptions to \textsc{Backness Harmony}, root disharmony (\ref{turkexcept}a) has been the subject of much debate.
-As disharmonic roots still trigger suffix harmony, \citet[212, 289]{Anderson1974} and \citet{Iverson1978} distinguish suffix harmony alternations from the sequence structure constraint governing root harmony.
+While suffix disharmony appears to very sporadic, the co-occurrence of suffix harmony and root disharmony has been the subject of much debate.
+As disharmonic roots still trigger suffix harmony,
+%\citet[212, 289]{Anderson1974} and
+\citet{Iverson1978} distinguish suffix harmony alternations from the sequence structure constraint governing root harmony.\footnote{\citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process.
+\citet[171f.]{Howard1972} correctly observes that this duplication is unnecessary.}
 
-The disadvantage of this account is that it introduces ``duplication'' (in the sense of \citealt{Kisseberth1970b} and \citealt{Kenstowicz1977}) of sequence structure and phonological generalizations, differning only in their patterns of exceptionality.
-However, \citet[][197f.]{Zonneveld1978} shows that the theory of exceptionality proposed in \emph{SPE} can account for suffix harmony in disharmonic roots.\footnote{An isomorphic analysis is available within less restrictive theories of phonological exceptionality \citep[e.g.,][]{Kisseberth1970,Pater2009}.} \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
+The disadvantage of this account is that it introduces ``duplication'' (in the sense of \citealt{Kisseberth1970b}) of sequence structure and phonological generalizations, differing only in their pattern of exceptionality.
+\citet[][197f.]{Zonneveld1978} shows that the theory of exceptionality proposed in \emph{SPE} can in fact account for suffix harmony in disharmonic roots.\footnote{An isomorphic analysis is available within less restrictive theories of phonological exceptionality \citep[e.g.,][]{Kisseberth1970,Pater2009}.} \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
 A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}].
@@
 If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.
-Thus \textsc{Backness Harmony} is able to account for both root and suffix harmony.
+Thus \textsc{Backness Harmony} is able to account for both root and suffix harmony, and duplication is unnecessary.
 
-It is necessary to dispense with an alternative analysis proposed by \citet{Clements1982} and \citet{Inkelas1997}.
+It is necessary to dispense with an alternative proposed by \citet{Clements1982} and \citet{Inkelas1997} which reintroduces this duplication.
 Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
@@
 
-\begin{example}[Autosegmental underspecification in harmonic roots (after \citealp{Clements1982})] \label{spec}
+\begin{example}[Autosegmental underspecification in harmonic roots (after \citealp{Clements1982})] 
+\label{spec}
 \xymatrix@R=24pt@C=24pt{
@@
 One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being a Structure Preservation condition \citep{Kiparsky1985}.
-However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a perceptual default.
+However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default.
 A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone.
-However, this interpretation makes the underspecification analysis a mere notational variant of the rule exceptionality account, in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.
+However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.
 
@@
 If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it.
-Several other ``external'' facts demonstrate that this is not the case.
+Several other ``external'' facts suggest  that this is not the case, with the caveat that uncritical acceptance of hypotheses linking loanword adaptation and language games to the grammar is no way being implied.
 
@@
 \begin{example}[Variable non-native cluster adaptation (\citealp{Clements1982}:247)] 
-\begin{tabular}{l l l l l l}
+\begin{tabular}{lllll}
 a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
@@
 
-Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} This game is not indigenous to Turkish, but it is found in the grammar of the Turkic language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; \citeauthor{Harrison2001} report that it can be quickly taught to Turkish speakers.
-The game consists of reduplication of the base, with the first vowel of the reduplicant replaced with the [$+$\textsc{Back}] vowels \emph{a} or \emph{u}.
-Both in Tuvan and in the Turkish game, the second vowel of the reduplicant (shown in braces below) may also be effected by this process; when the base vowels are both [$-$\textsc{Back}], the insertion of a [$+$\textsc{Back}] vowel results in what \citeauthor{Harrison2001} call ``reharmonization'', shown in  (\ref{redupgame}a).
-On the other hand, as shown in (\ref{redupgame}b), disharmonic roots do not reharmonize.\footnote{A similar pattern is found in Tuvan, and in an unrelated Finnish language game \citep{Campbell1986}.}
+Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} 
+The game is native to the related language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; it is not indigenous to Turkish, but can be taught quickly to children or adults.
+In this game, the base is reduplicated and the first vowel of the reduplicant replaced with a [$+$\textsc{Back}] vowel.
+In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the the inserted [$+$\textsc{Back}] vowel.
+The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}
 
-\begin{example}[Turkish reduplication game (\citealp{Harrison2001}:231)] \label{redupgame}
-\begin{tabular}{l l l l}
+\begin{example}[Turkish reduplication game (\citealp{Harrison2001}:231)] 
+\label{redupgame}
+\begin{tabular}{llll}
 a. & {kibrit} & {kibrit}-\{{kabrıt}\} & `match'    \\
@@
 
-A number of studies have investigated the role of harmony in word-spotting tasks, though to reveal how speakers identify word boundaries in running speech.
-Many of these studies have been carried out in Finnish, which has a vowel harmony system similar to Turkish.
-\citet{Suomi1997} and \citet{Vroomen1998} task speakers with identifying harmonic disyllables in an auditory stream.
-Finnish speakers are better able to complete this task when the preceding syllable has a different backness specification than the disyllabic target.
-\citet{Kabak2010} report that Turkish \textsc{Backness Harmony} has the same effect on word-spotting tasks as it does in Finnish: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}), but this effect does not obtain for speakers of French, a language which lacks vowel harmony.
-These results imply that speakers of Turkish have internalized the tendency of harmonic sequences to be root-internal and of disharmonic transitions to mark word juncture.
+A number of studies have investigated the role of harmony in word-spotting tasks, though to mimic auditory word recognition and segmentation in natural settings.
+Many of these studies have been carried out in Finnish, which has a vowel harmony system similar to that of Turkish.
+\citet{Suomi1997} and \citet{Vroomen1998} task Finnish speakers with identifying harmonic disyllables in an auditory stream.
+When the syllable preceding the disyllabic target has a different backness specification than the target, this faciliates speakers' recognition abilities.
+Presumably, disharmony facilitates the recognition of word boundaries.
+\citet{Kabak2010} find that Turkish \textsc{Backness Harmony} has a a similar effect: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}).
+\citet{Kabak2010} find that this effect does not obtain for speakers of French, a language which lacks vowel harmony.
+This implies that Turkish speakers have internalized the tendency of harmonic sequences to be root-internal and of disharmonic transitions to cross word boundaries.
 
-The word-spotting experiment has been adapted for infants by \citep{Kampen2008}.
+The Turkish word-spotting experiment has been adapted for infants by \citet{Kampen2008}.
 In this study, 9-month-old infants are familiarized with an auditory stream containing harmonic disyllabic nonce words.
@@
 Infants acquiring Turkish listen longer to nonce words preceded by a disharmonic juncture during familiarization (e.g., \emph{lo-NETIS}), whereas infants acquiring German, a language which lacks vowel harmony, do not exhibit this preference.
-Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic pseudowords such as \emph{paroz} over disharmonic pseudowords like \emph{nelok}, but German 6-month-old infants show no such preference.
+Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic nonce words such as \emph{paroz} over disharmonic nonce words like \emph{nelok}, but German 6-month-old infants show no such preference.
 
 \subsection{Roundness harmony}
-
-\subsubsection{Phonological description}
 
@@
 \begin{example}[\textsc{Roundness Harmony} (condition: rightward application)]
-$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~/~\begin{bmatrix}~=\textsc{Back}~\end{bmatrix}~\textrm{C}_0~\gap$
+$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix}~=\textsc{Rnd}~\end{bmatrix}~\textrm{C}_0~\gap$
 \end{example}
@@
 \begin{example}[Turkish nominal suffix allomorphy]
-\begin{tabular}{l l l l l l@{ }l}
+\begin{tabular}{lllllll}
    & \emph{nom.sg.} & \emph{dat.sg.} & \emph{gen.sg.}  \\
@@
 b. & {boğaz}        & {boğazı}       & {boğazın}      & `throat'  & \citep{TELL} \\
-   & {pelür}        & {pelürü}       & {pelürün}      & `tissue paper' \\
+   & {pelür}        & {pelürü}       & {pelürün}      & `tissue' \\
    & {döviz}        & {dövizi}       & {dövizin}      & `currency' \\
@@
 However, two pieces of the external evidence on \textsc{Backness Harmony} bear on this question.
-First, the epenthetic vowel in non-native word-initial onsets like \emph{trablus} \alt{} \emph{tırablus} `Tripoli' conform both to \textsc{Backness Harmony} and to \textsc{Roundness Harmony}.
+First, the epenthetic vowel in non-native word-initial onsets like \emph{trablus} \alt{} \emph{tırablus} conforms to \textsc{Roundness Harmony} in addition to \textsc{Backness Harmony}.
 Secondly, \textsc{Roundness Harmony} participates in reharmonization in the language game described by \citeauthor{Harrison2001}: the second \emph{ü} in \emph{bütün} `whole' reharmonizes to the [$-$\textsc{Rnd}] vowel \emph{ı} in reduplicated \emph{bütün-batın}.
@@
 \begin{example}[\textsc{Labial Attraction}]
-$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} -\textsc{Cons} \\ -\textsc{Rnd} \\ +\textsc{Back} \\ -\textsc{High} \end{bmatrix}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
+%$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} -\textsc{Cons} \\ -\textsc{Rnd} \\ +\textsc{Back} \\ -\textsc{High} \end{bmatrix}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
+$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\textrm{ɑ}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
 \end{example}
 
-The formalization of this rule is naturally complex, and perhaps obscures the fact that \textsc{Labial Attraction} generates exceptions to \textsc{Roundness Harmony}, producing \emph{a}P\emph{u} sequences (where P represents a labial consonant; e.g., \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', \emph{samur} `sable'; \citealp[285]{Lees1966a}) rather than the expected \emph{a}P\emph{ı}.
-However, \textsc{Labial Attraction} does not apply in derived environments: for example, the gen.sg.
+This rule is significantly more complex than the harmony rules, and this may obscure the fact that it produces exceptions to \textsc{Roundness Harmony}, producing \emph{a}C$_0$\emph{u} (e.g., \emph{çapul} `raid', \emph{sabur} `patient', \emph{şaful} `wooden honey tub', \emph{avuç} `palm of hand', \emph{samur} `sable'; \citealp[285]{Lees1966a}) rather than the expected \emph{a}C$_0$\emph{ı}.
+However, \textsc{Labial Attraction} does not apply in derived environments: the gen.sg.
 of \emph{sap} `stalk is \emph{sapın} rather than *\emph{sapun}.
-
-\citet[286]{Lees1966a} and \citet[311]{Zimmer1969} note the existence of root-internal exceptions to this rule (e.g., \emph{tavır} `mode') but agree that they are surprisingly rare; this is disputed in turn by  \citet{Clements1982} and \citet{Inkelas2001}.
-A more exhaustive review of the lexical statistics will be provided below.
+\citet[286]{Lees1966a} and \citet[311]{Zimmer1969} note the existence of root-internal exceptions to this rule (e.g., \emph{tavır} `mode') but agree that they are surprisingly rare.
 
@@
 
-by comparing lexical statistics with the results of wordlikeness task performed by \citet{Zimmer1969}.
+Others dispute that \textsc{Labial Attraction} is a reliable generalization about Turkish roots.
+\citet{Clements1982} point to additional exceptions to the rule, and \citet{Inkelas2001} provide lexical counts from the Turkish Electronic Living Lexicon \citep[TELL;][]{TELL}, a database of words known to two native Turkish speakers.
+However, no prior study reports any inferential statistics.
 
-This resolves the debate concerning the nature of \textsc{Labial Atttraction}: it is shown to be a statistically robust generalization over the Turkish lexicon (\citealt{Lees1966a}, \citealt{Zimmer1969}, \emph{pace} \citealt{Clements1982}, \citealt{Inkelas2001}); however, no variant of \textsc{Labial Attraction} is reflected in \citeauthor{Zimmer1969}'s wordlikeness study.
-This dissociation between lexical generalizations and wordlikeness results provides further evidence against the common practice of inferring phonotactic knowledge via purely statistical criteria.
+\citet[311]{Zimmer1969} administers two binary paired wordlikeness task designed to evaluate native speakers' knowledge of \textsc{Backness Harmony}, \textsc{Roundness Harmony}, and \textsc{Labial Attraction} in roots.
+Speakers are presented with two nonce words and simply have to indicate the nonce word that is more Turkish-like.\footnote{Compared to the unpaired ratings tasks  commonly used in wordlikeness research, paired rating tasks have considerably more statistical power \citep{Gigerenzer2004}.
+When items under comparison differ only in whether or not they conform to the phonotactic generalizations, it is less likely that the any contrast between the members of the paris are caused by some omitted variable.
+Also, there is no unnecessary deception regarding the purpose of the experiment, which is known to reduce spurious response variability.
+Consequently, paired tasks are particularly sensitive to small phonotactic contrasts and are ideal for collecting wordlikeness judgements.} \citeauthor{Zimmer1969} concludes that the former two rules are reflected in wordlikeness judgements, whereas \textsc{Labial Attraction} is not.
+At the same time, \citeauthor{Zimmer1969} also suggests that speakers do internalize a variant of the rule that does not require a labial consonant in the trigger, a proposal that is endorsed by \citet{Clements1982} and \citet[4]{Zuraw2000}.
+\citet[196]{Inkelas2001}, argue instead that the trigger must contain a labial consonant but does not require a preceding \emph{a} vowel.
+Once again, no inferential statistics over the wordlikeness results are reported.
+
+Below, both lexical statistics and \citeauthor{Zimmer1969}'s wordlikeness results are analyzed statistically, so as to resolve the debate concerning the synchronic reality of \textsc{Labial Attraction} and the connection between speakers' wordlikeness judgements and statistical patterns in the lexicon.
+\textsc{Labial Attraction} is shown to be a statistically robust generalization over the Turkish lexicon, but no variant of \textsc{Labial Attraction} is reflected in \citeauthor{Zimmer1969}'s wordlikeness study.
+In contrast, the two harmony processes have robust effects both on the lexicon and on wordlikeness.
+This dissociation between lexical generalizations and wordlikeness results provides further evidence against the common practice of inferring phonotactic knowledge directly from lexical statistics.
 
@@
 
-%\citet{Becker2011,Hayes2009}
+Counts were computed by regular expression matching on a 9,601-root subset of the TELL database, which were elicited from a native speaker, and which show no surface variation between or within speakers in any inflected form.
+
+To test for associations between the process (more specifically, the constraint that it imposes on roots) and type frequency in this database, each root was sorted into a $2 \times 2$ contingency table; the contents of each cell are specific to the process in question.
+The counts in this table are not expected to add up to 9,601, since many roots neither could exemplify nor violate the process in question; for instance, monosyllabic words are irrelevant to root harmony.
+The Fisher exact test is used to compute a $p$-value representing the probability of the observed data arising under the null hypothesis that there is no association between the process and type frequency; if this $p$-value is small, the null hypothesis is rejected and there is evidence that the process is reflected in the lexicon.
 
 \subsubsection{Backness harmony}
+
+\textsc{Backness Harmony} is exemplified in the lexicon insofar as there is an association between the backness of vowels in all adjacent syllables.
+Any disagreement on the backness specification of vowels in adjacent syllables are counted solely as exceptions, even if other vowel transitions in the root are harmonic.
+This follows from the hypothesis that phonological exceptionality is a property of whole underlying representations, not individual segments.
+To construct the contingency table, roots are binned according to the backness specification of the first nucleus, and that of following nuclei.
+For example, the first two syllables of \emph{adalet} `justice' are harmonic, but it is coded as disharmonic because there is a \emph{a\ldots{}e} transition later in the word.
+The resulting counts are shown in Table \ref{bhs}.
+61\% of roots conform to \textsc{Backness Harmony}, and the interaction between the backness of the first and of the subsequent vowels
+%predicted by root-internal \textsc{Backness Harmony} 
+is significant.
+
+\begin{table}
+\centering
+\begin{tabular}{lrrr}
+\toprule
+                             & [$+$\textsc{Back}]$_1$ & [$-$\textsc{Back}]$_1$ & $p$-value                     \\
+\midrule
+\buf{}[$+$\textsc{Back}]$_{2\ldots{}n}$ & 3,089                     & 1,704              & \multirow{2}{*}{$1.19$\e{-89}} \\
+\buf{}[$-$\textsc{Back}]$_{2\ldots{}n}$ & 1,698                     & 2,250                                               \\
+\bottomrule
+\end{tabular}
+\caption{TELL roots sorted according to \textsc{Backness Harmony}}
+\label{bhs}
+\end{table}
+
 \subsubsection{Roundness harmony}
+
+\textsc{Roundness Harmony} consists of an association between the roundness of a vowel and the roundness of high vowels in the next syllable.
+Any root for which a vowel does not agree in roundness with a high vowel in the following syllable (e.g., \emph{ümit}) is considered to be an exception.
+Roots are binned according to the roundness of non-initial high vowels, and according to the roundness of the preceding vowel.
+%Roots lacking non-initial high vowels do not bear on the status of root \textsc{Roundness Harmony}. 
+The resulting counts are shown in Table \ref{rhs}.
+83\% of the roots conform to \textsc{Roundness Harmony}, and the interaction between the roundness of the $i$th vowel and the roundness of the $(i +1)$th high vowel is significant.
+
+\begin{table}
+\centering
+\begin{tabular}{lrrr}
+\toprule
+                                              & [$+$\textsc{Rnd}]$_i$ & [$-$\textsc{Rnd}]$_i$ & $p$-value                      \\
+\midrule
+\buf{}[$+$\textsc{High}, $+$\textsc{Rnd}]$_{i+1}$ & 613                   &   261                 & \multirow{2}{*}{$1.02$\e{-36}} \\
+\buf{}[$+$\textsc{High}, $-$\textsc{Rnd}]$_{i+1}$ & 581                   & 2,841                                                  \\
+\bottomrule
+\end{tabular}
+\caption{TELL roots sorted according to \textsc{Roundness Harmony}}
+\label{rhs}
+\end{table}
+
+\noindent
+The counts in bottom row of Table \ref{rhs} contain a number of roots which are apparent exceptions to \textsc{Roundness Harmony} but conform to \textsc{Labial Attraction} (bottom left), and which conform to \textsc{Roundness Harmony} at the expense of \textsc{Labial Attraction} (bottom right).
+Excluding these types of roots would have the effect of slightly increasing the overall rate of \textsc{Roundness Harmony}, since the former is more common.
+
 \subsubsection{Labial attraction}
 
+\citet{Clements1982} object to \textsc{Labial Attraction}:
+
 \begin{quote}
-\ldots{}decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots{}aCu\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots{}We conclude that there is no systematic restriction on the set of consonants that may occur medially in roots of the form /\ldots{}aCu\ldots/.
+Even decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots{}aCu\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots{}We conclude that there is no systematic restriction on the set of consonants that may occur medially in roots of the form /\ldots{}aCu\ldots/.
 \citep[225]{Clements1982}
@@
 
-but this fact is not inconsistent with the formulation of the constraint:
-\emph{aTu} clusters do not meet \textsc{Labial Attraction}'s structural description.
+\noindent 
+This claim yields itself naturally to the Fisher exact test.
+Let P denote a sequence of one or more consonants, one of which is labial, and let T denote a sequence of one or more consonants none of which is labial.
+The null hypothesis is that \emph{a}P\emph{u} sequences, with P representing a cluster containing a labial consonant, which conform to \textsc{Labial Attraction}, are no more likely than would be expected from other \emph{a}T\emph{u} sequences violating \textsc{Roundness Harmony}.
+The resulting counts are shown in Table \ref{las}.
+Whereas the sequence \emph{a}P\emph{u} is more than twice as likely as \emph{a}P\emph{ı}, the sequence \emph{a}T\emph{u} is four times less likely than \emph{a}T\emph{ı}.
+This is a significant interaction predicted by \textsc{Labial Harmony}, but contradicting a statistical interpretation of the claim by \citeauthor{Clements1982}.
+
+\begin{table}
+\centering
+\begin{tabular}{lrrr}
+\toprule
+       & a\ldots{}u & a\ldots{}ı & $p$-value                      \\
+\midrule
+aP\ldots{} & 124    & 57     & \multirow{2}{*}{$1.02$\e{-36}} \\
+aT\ldots{} & 136    & 590    &                                \\
+\bottomrule
+\end{tabular}
+\caption{TELL roots sorted according to \textsc{Labial Attraction}}
+\label{las}
+\end{table}
+
+Alternative formulations of \textsc{Labial Harmony} have been alluded to above. \citeauthor{Inkelas2001}, for instance, propose to eliminate the requirement of a preceding \emph{a} in the trigger:
+
+\begin{quote}
+Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
+\citep[][196]{Inkelas2001}
+\end{quote}
 
 \noindent
-\citeauthor{Clements1982} are suggesting that \textsc{Labial Attraction} implies that \emph{a}\ldots{}\emph{u} sequences separated by non-labial consonants should be infrequent.
-This is not immediately clear, since this is beyond the scope of \citeauthor{Lees1966a}'s description of \textsc{Labial Attraction}, but the frequency of these \emph{a}T\emph{u} sequences does provide an estimate of how often ???.
+Counts from TELL for this formulation of the process are shown in Table \ref{lasi}.
+This triples the number of roots which exemplify \textsc{Labial Attraction} (top left cell) but also slightly increases the number of exceptions.
+The interaction remains significant.
 
-\begin{quote}
-Lee's rule of \textsc{Labial Attraction}\ldots is not a real generalization about the Turkish lexicon.
-It is not true synchronically, either of native or nonnative items; nor, according to the historical and dialectical literature, does \textsc{Labial Attraction} appear to have been true at any stage going back as far as Old Turkic [\ldots] Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
-\citep[][196, emphasis in original]{Inkelas2001}
-\end{quote}
-
-\begin{example}[Lexical effects of \textsc{Labial Attraction}]
-\begin{tabular}{l r r r r r}
+\begin{table}
+\centering
+\begin{tabular}{lrrr}
 \toprule
-Corpus                & aPu & aPı & aTu & aTı   & $p$-value   \\ % & corpus size
+       & \ldots{}u  & \ldots{}ı & $p$-value                      \\
 \midrule
-Full TELL             & 378 & 248 & 446 & 1,140 & 2.83\e{-44} \\ % & 31,236 \\
-Elicited TELL         & 152 & 265 & 101 & 1,839 & 9.84\e{-60} \\ % & 16,541 \\
-TELL with etymologies & 128 & 109 &  79 &   470 & 6.56\e{-32} \\ 
+P\ldots{}  & 371    & 71        & \multirow{2}{*}{$6.98$\e{-49}} \\
+T\ldots{}  & 811    & 922       &                                \\
 \bottomrule
 \end{tabular}
-\end{example}
+\caption{TELL roots sorted according to the reformulation of \textsc{Labial Attraction} proposed by \citet{Inkelas2001}}
+\label{lasi}
+\end{table}
 
-%\citet{Harrison2004} 73\% of lexical types are harmonic (both back and round)
-
-%Etymological subset of TELL
-%Native    & Foreign   \\
-%aBu & aBI & aBu & aBI \\
-%12  & 11  & 84  & 28  \\
-%p = 0.0417
+Reviewing the results of wordlikeness experiments, \citet{Zimmer1969} proposes a third variant of \textsc{Labial Attraction} which is insensitive to intervening consonant place.
+This has no support in the TELL database, however: \emph{a\ldots{}u} sequences are less, not more, common than \emph{a\ldots{}ı}.
+This variant is not considered further.
 
@@
 
-%\citet{Gigerenzer1990} argue that ratings tasks should be replaced with paired comparison to minimize error.
+\citet{Zimmer1969} administers two variants of the paired nonce word rating task.
+The first used 23 native adult speakers who were permitted to select either nonce word as more like Turkish, or to indicate `no preference'.
+For the purposes below, `no preference' results are discarded.
+The second experiment used 32 native adults, none of whom appeared in the preceding study, and used a forced binary choice with some minor changes to the stimulus set.
 
-%\citet{Goodman1954}
-%The results reveal a dissociation between the lexical statistics and wordlikeness judgements, a dissociation that finds a natural explanation from the principle of ...
-%There is some disagreement in the literature about just what \citeauthor{Zimmer1969}'s study reveals about \textsc{Labial Attraction}.
+Each response is coded as \emph{concordant} if the the nonce word conforming the the process is preferred, and \emph{discordant} if the disharmonic word is selected.
+To test for an association between the constraints, a non-parametric statistic, the Goodman-Kruskal (\citeyear{Goodman1954}) $\gamma$ is computed from to the count of concordant ($c$) and discordant ($d$) pairs:
 
-%\citet{Ito1993} and \citet{NiChiosain1993} argue that the data shows that it holds of native words only, but \citet{Inkelas2001} refute this point
+\begin{unlabeledexample}
+$\displaystyle \gamma = \frac{c - d}{c + d}$
+\end{unlabeledexample}
 
-%\citet{Zuraw2000}:4
-%\citet{Inkelas2001} present a different variation on the original rule of \textsc{Labial Attraction}
-
-%\begin{quote}
-%Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
-\citep[][196, emphasis in original]{Inkelas2001}
-%\end{quote}
-
-%\noindent 
-%\citeauthor{Inkelas2001} persistently refer to ``statistical support'', but in fact their study contains no statistics beyond counts and percentages.
-%since the opposing are all broader versions than the original formulation by \citeauthor{Lees1966a}.
+\noindent
+The $\gamma$ statistic ranges between -1, in the case that all paired choices are discordant, and 1, if all paired choices are concordant.\footnote{It is possible to perform statistical tests aggregating over items, but for this small number of items, such tests have very low (in some cases null) statistical power.}
 
@@
 
-\begin{example}[Wordlikeness comparisons, backness harmony]
+Both of the experiments performed by \citet{Zimmer1969} include 5 pairs which differ on in whether or not the nonce words conform to, or violate, \textsc{Backness Harmony}.
+As can be seen from Table \ref{bhw}, harmonic pairs are preferred approximately 6-to-1, and aggregating over speakers, no disharmonic member of a pair is favored.
+Speakers have a highly reliable preference for nonce words which exhibit \textsc{Backness Harmony} ($\gamma = 0.694$, $p = 1.7$\e{-59}).
+It is interesting to note that the disharmonic nonce word which has the highest rating is found in the pair \emph{terüz}-\emph{teruz}, both of which violate \textsc{Roundness Harmony}.
+While this is little more than an anecdote, this may be  indicative of a link between the two processes, and their exceptions, in the minds of native speakers.
+
+\begin{table}
+\centering
 \begin{tabular}{lrlr|lrlr}
@@
 \end{tabular}
-\end{example}
-
+\caption{Effects of \textsc{Backness Harmony} on wordlikeness \citep[from][]{Zimmer1969}}
+\label{bhw}
+\end{table}
 % harmonic responses are favored approximately 6 to 1.
@@
 
-Note that both of the pair \emph{terüz}-\emph{teruz} violate \textsc{Roundness Harmony}
-
 \subsubsection{Roundness harmony}
 
-\begin{example}[Wordlikeness comparisons, roundness harmony]
+Both experiments include 5 pairs which are differ in the presence of absence of \textsc{Roundness Harmony}.
+As shown in Table \ref{rhw}, there is an approximately 5-to-1 preference for harmonic nonce words, and as was the case above, no disharmonic member of any pair is overall preferred.
+Turkish speakers have a reliable preference for nonce words to conform to \textsc{Roundness Harmony} ($\gamma = 0.680$, $p = 1.1$\e{-47}).
+
+\begin{table}
+\center
 \begin{tabular}{lrlr|lrlr}
@@
 \end{tabular}
-\end{example}
-
-% more than 5 to 1
-% gamma = 0.680 p = 1.1e-47
+\caption{Effects of \textsc{Roundness Harmony} on wordlikeness \citep[from][]{Zimmer1969}}
+\label{rhw}
+\end{table}
 
@@
 
-\begin{example}[Wordlikeness comparisons, labial attraction]
+Both experiments include five pairs which either conform to \textsc{Labial Attraction} and violate \textsc{Roundness Harmony}, or vis versa; the preferences are shown in Table \ref{law}.
+There is a small preference against \textsc{Labial Harmony}, though this is non-significant ($\gamma = -0.043$, $p = 0.305$).
+This shows that in general speakers do not have preferences as predicted by \textsc{Labial Harmony}.
+This result also extends for the variants of \textsc{Labial Attraction} proposed by \citet{Zimmer1969} and \citet{Inkelas2001}, since these variants have structural descriptions targeting a superset of the original formulation by \citet{Lees1966a}.
+It is interesting to note that speakers do not have the clear opposite preference, as is predicted by \textsc{Roundness Harmony}.
+A possible explanation for this is that \textsc{Roundness Harmony} is suspended in the context defined by \textsc{Labial Attraction} or some variant thereof.
+
+\begin{table}
+\centering
 \begin{tabular}{lrlr|lrlr}
@@
 \end{tabular}
-\end{example}
+\caption{Effects of \textsc{Labial Attraction} on wordlikeness \citep[from][]{Zimmer1969}}
+\label{law}
+\end{table}
 
@@
 
-etymological issues
+It has been shown that while \textsc{Labial Attraction} is a highly reliable generalization about Turkish root, it is not reflected in wordlikeness judgements.
+In contrast, harmony processes have a similar statistical profile, but have unambigious traces on wordlikeness.
+The most plausible explanation for this is that  \textsc{Labial Attraction} does not trigger alternations; indeed, it is counter-exemplified by the effects of \textsc{Roundness Harmony} in suffixes.
+As is noted by \citet[412f.]{Inkelas1997}, the lexicon of Turkish will, under the assumptions here, remain as it is whether or not \textsc{Labial Attraction} has a synchronic reality.
+A principle of simplicity suggests that it does not exist at all.
 
-%\citet{Inkelas2001}
-%\citet{NiChiosain1993} and \citet{Ito1995b} 
-%\begin{example}[\textsc{Labial harmony} and etymology in TELL \citealp{Inkelas2001}:187]
-%\begin{tabular}{l r r r r}
-%\toprule
-%        & {a}P{u} & {a}P{ı} & \% aPu & $p$-value \\
-%\midrule
-%native  & 12      & 11      & 52.2\% & \multirow{2}{*}{0.042} \\
-%foreign & 84      & 28      & 75.0\% \\
-%\bottomrule
-%\end{tabular}
-%\end{example}
+\citet{NiChiosain1993} and \citet{Ito1995b} argue that \textsc{Labial Attraction} is true only of native vocabulary.
+This would be a potential confound for the experimental results, since it is not implausible that the subjects' in \citeauthor{Zimmer1969}'s study treated nonce words as if they were loanwords.\footnote{Indeed, many wordlikeness studies include instructions to the participants to treat the stimuli much as if they were loanwords \citep[e.g.,][]{Hay2004a}.
+This may influence the choice of task model used by participants.} However, this was not the opinion of \citet[266]{Lees1966a}, and is conclusively refuted by  \citet{Inkelas2001}.
+They observe that in the subset of the TELL database with etymological coding, a full 75\% of foreign vocabulary conform to \textsc{Labial Attraction}, which is surprisingly a smaller percentage than that of the foreign vocabulary (52\%), a significant trend (Fisher exact test, $p = 0.042$).
+One plausible explanation is that many of the languages in contact with Turkish lack the \emph{ı} phoneme needed to violate \textsc{Labial Attraction}.
 
-lack of theory of naturalness
+\citet{Becker2011} attribute the inactivity of \textsc{Labial Attraction} in wordlikeness to indicate that the generalization is phonologically unnatural, and therefore goes unlearned.
+This claim is difficult to evaluate insofar as phonotactic naturalness has continually eluded formalization and \citeauthor{Becker2011} neither propose nor refer to any theory thereof.
+\citeauthor{Becker2011} use the lack of phonetic precursors as a diagnostic of unnaturalness; however, there is extensive evidence that ``unnatural'' generalizations are in fact learned \citep[e.g.,][]{Anderson1981,Bach1972,Blevins2004,Buckley2000a,Hayes2009,Pierrehumbert2006c}.
+Furthermore, other functionally-oriented theorists have considered \textsc{Labial Attraction} sufficiently natural \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995b}.
 
-%\citet{Becker2011}
-%This is difficult to evaluate insofar as \citeauthor{Becker2011} neither propose nor refer to any theory of naturalness which might exclude \textsc{Labial Attraction}. 
-%There are several other problems with this claim.
-Both early specialists \citep[e.g.,][]{Lees1966a} and later theorists working within a functionally-oriented framework \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995a} have considered it plausible.
-Further, \citet{Hayes2009} claim that Hungarian speakers internalize similarly ``unnatural'' generalizations about exceptions to vowel harmony.
-%Finally, there is at least some reason to admit similar structural descriptions in other languages.
-For example, some Classical Arabic adjectives have stative verbs in which the root melody is overwritten with \emph{a\ldots{}u}.
-If the stative verbs are derived from the adjectives, as has recently been argued for Semitic \citep[e.g.,]{Ussishkin1999,Ussishkin2000,Tucker2010}, than there is little reason to believe that this sequence is somehow marked.
-%\begin{example}[Arabic stative verbs]
-%\begin{tabular}{l l l l}
-%\buf{}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
-%\buf{}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
-%\buf{}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
-%\end{tabular}
-%\end{example}
-%\footnote{Thanks to Uri Horesh (p.c.) for assistance with this data.}
+It is not even obvious that it is desirable to exclude \textsc{Labial Attraction} from the grammar.
+\citet[394, fn.
+2]{Inkelas1997} suggest that \textsc{Labial Attraction} may have even induced alternations at one point in the history of Turkish.
+Beyond Turkish, the rounding of a high back vowel after a labial consonant is both widely attested \citep[e.g.,][]{Vaux1993}.
+It is also phonetically natural, as both labial consonants and high round vowels are distinguished by a very low first and second formant.
+Regarding the non-local condition holding between the \emph{a} portion of the trigger and the \emph{u} target, Classical Arabic has many adjectives which form corresponding inchoative verbs by overwriting the vocalic melody with /a\ldots{}u/ (the final /-a/ is an inflectional suffix).
 
-%\noindent
-%Further, labial consonants frequently cause rounding of following vowels \citep[e.g.,][]{Vaux1993}. The relevant constraints, conjoined, are a close approximation of \textsc{Labial Attraction}.
+\begin{example}[Arabic derived inchoatives]
+\begin{tabular}{l l l l}
+\buf{}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
+\buf{}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
+\buf{}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
+\end{tabular}
+\end{example}
+
+\noindent
+In an Optimality Theory framework, for instance, constraints responsible for rounding of high back vowels before labials and for /a\ldots{}u/ overwriting can be conjoined to approximate the structural description of \textsc{Labial Attraction}.
 
 \section{Conclusions}
+
+It has been shown that statistical reliability is not a sufficient condition for a sequence structure generalization to be reflected in psycholinguistic tasks.
+However, in the case of constraints on Turkish vowels, it is a necessary condition that the constraint in question be derived from an alternation.
+Even alternations that have large numbers of lexical exceptions have have robust effects on a constrained wordlikeness task.
+Purely statistical approaches to phonotactic knowledge fail to draw this important distinguish.
 
@@
 Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
-%\begin{unlabeledexample}
-%$\displaystyle \gamma = \frac{C - D}{C + D}$
-%\end{unlabeledexample}
-%\footnote{Any readers who are skeptical that \citeauthor{Chomsky1965}'s intuitions generalize will find support for their claim in the wordlikeness judgement data reported in Appendix \ref{albright}.}
 %As a quick validation of this claim, consider the results reported by \citet{Albright2007} and discussed in the previous chapter.
@@
 If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
-%This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations.
 %Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
 This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.
-
 %\citet{Berent2007a}
@@
 % Dupoux et al. 1999: Japanese speakers still can't distinguish between ebzo and ebuzo in an ABX task using the SAME a or b.
-%Elliot Moreton (p.c.)
 %Once additional source of evidence on root (dis)harmony is inconclusive.
 There is a small class of bisyllabic words in which the second vowel, always [$+$\textsc{High}, $-$\textsc{Back}], alternates with zero.
-
 %\ex High-vowel/zero alternations \citep[][243]{Clements1982}: \\
@@
 %\end{tabular} \xe
-%
-%\noindent
 %It is possible that \textsc{Backness Harmony} might produce a fluctuating \emph{ı} after root \emph{a}, but this does not obtain (\lastx b).
@@
 %(e.g., \emph{deve} `camel' vs. \emph{deva} `medicine', \emph{sene} `year' vs. \emph{sena} `praise').
-
 %This is not to imply uncritical acceptance of \citeauthor{Becker2011}'s account.
 \citeauthor{Becker2011} place this generalization ``in the grammar'' to account for their claim that ``naturalness'', which they equate with Universal Grammar, constrains the types of generalizations speakers extract in this task.
-% 
 %It is difficult to evaluate their claim in the absense of a theory of phonotactic naturalness, something which has consisteny eluded formalization for decades.
@@
 It is certainly not something that can be explained by the tacit theory of possible gneeralizations the authors adhere to.
-
 %Whereas the classic analysis of this phenomenon, by \citet{Inkelas1997}, does not make use of lexical exceptionality, \citeauthor{Becker2011}'s analysis of final aspirated stops and affricates as the underlying form requires that _every_ lexical entry with a final stop or affricate be treated as exceptional according to some phonological constraint.
@@
 There is some reason to suspect that NDEB is a symptom of interface restrictions on phonological application, rather than a diagnosis itself.
+%The loanword adaptation literature also contains many reports that statistically reliable phonotactic generalizations about the native lexicon are absent in processes of nativization \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}.
+It is not immediately clear that this is evidence that the generalizations in question are external to the synchronic grammar, however.
+%\begin{example}[\textsc{Backness Harmony} lexical statistics]
+%\begin{example}[Wordlikeness comparisons, backness harmony]
+%\begin{example}[Wordlikeness comparisons, roundness harmony]
+%\begin{example}[\textsc{Labial harmony} and etymology in TELL (counts from \citealp{Inkelas2001}:187)]
+%\begin{example}[Wordlikeness comparisons, labial attraction]
+%\begin{table}
+%\centering
+%\begin{tabular}{lrrrr}
+%\toprule
+%        & {a}P{u} & {a}P{ı} & \% aPu & $p$-value \\
+%\midrule
+%native  & 12      & 11      & 52     & \multirow{2}{*}{0.042} \\
+%foreign & 84      & 28      & 75                              \\
+%\bottomrule
+%\end{tabular}
+%\caption{?}
+%\label{lae}
+%\end{table}
+%The loanword adaptation literature also contains many reports that statistically reliable phonotactic generalizations about the native lexicon are absent in processes of nativization \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}.
+It is not immediately clear that this is evidence that the generalizations in question are external to the synchronic grammar, however.

### 2012-10-11 4.tex
Completion of draft of chapter 3

@@
 
-%\footnote{Similar behaviors are found in Russian verbs which share roots but which take different prefixes or the reflexive \citep{Pesetsky1977}.}
-
 \begin{example}[Bound stem-specific allomorphy]
@@
 
-\begin{example}[Latinate verbs and small clauses] \label{harley}
+\begin{example}[Latinate verbs and small clauses] 
+\label{harley}
 \begin{tabular}{ll@{}ll@{}l}
@@
 
-\subsubsection{Obstruent voice assimilation} \label{ova}
+\subsubsection{Obstruent voice assimilation} 
+\label{ova}
 
@@
 \midrule
-conforming & 80       & 370        & 17.8          & \multirow{2}{*}{1.1\e{-11}} \\
-violating  &  6       & 264        & 2.2           \\
+conforming & 80       & 370        & 18          & \multirow{2}{*}{1.1\e{-11}} \\
+violating  &  6       & 264        & 2           \\
 \bottomrule
@@
 
-\begin{example}[\emph{im-}/\emph{in-} allomorphy] \label{nparule}
+\begin{example}[\emph{im-}/\emph{in-} allomorphy] 
+\label{nparule}
 \begin{tabular}{l l l l l l l}
@@
 \midrule
-conforming & 41       & 11         & 78.8          & \multirow{2}{*}{2.6\e{-05}} \\
-violating  & 8        & 20         & 28.6                                    \\
+conforming & 41       & 11         & 79          & \multirow{2}{*}{2.6\e{-05}} \\
+violating  & 8        & 20         & 29                                    \\
 \bottomrule
@@
 
-\subsubsection{Degemination} \label{deg}
+\subsubsection{Degemination} 
+\label{deg}
 
@@
 \midrule
-conforming & 157      & 596        & 20.8      & \multirow{2}{*}{7.2\e{-09}} \\
-violating  & 0        &  87        & 0.00                                    \\
+conforming & 157      & 596        & 21      & \multirow{2}{*}{7.2\e{-09}} \\
+violating  & 0        &  87        & 0                                    \\
 \bottomrule
@@
 \midrule
-conforming & 76 & 359 & 17.5 & \multirow{2}{*}{0.145} \\
-violating  &  6 &  57 & 9.5 \\
+conforming & 76 & 359 & 18 & \multirow{2}{*}{0.145} \\
+violating  &  6 &  57 & 10 \\
 \bottomrule
@@
 \midrule
-conforming & 48       & 312        & 13.3      & \multirow{2}{*}{0.301} \\
-violating  & 38       & 322        & 10.6                               \\
+conforming & 48       & 312        & 13            & \multirow{2}{*}{0.301} \\
+violating  & 38       & 322        & 11            \\
 \bottomrule
@@
 \midrule
-conforming & 41       & 478        & 7.9      & \multirow{2}{*}{0.818} \\
-violating  &  0       &  11        & 0.0                               \\
+conforming & 41       & 478        & 8             & \multirow{2}{*}{0.818} \\
+violating  &  0       &  11        & 0             \\
 \bottomrule
@@
 
-The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
+The material will never be complete. It will always contain accidental gaps\ldots{}partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
 \end{quote}

### 2012-10-11 abs.tex
Completion of draft of chapter 3

@@
 It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
-A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements at least as acccurately as state-of-the-art gradient wellformedness models.
+A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements at least as accurately as state-of-the-art gradient wellformedness models.
 Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.

### 2012-10-11 pre.tex
Completion of draft of chapter 3

@@
 But this effect of word frequency obtains regardless of whether or not whole word storage is postulated \citep{Taft2004b,LignosInPress}.
-Furthermore, in languages with rich agglutinative morphology, the role played by a whole word storage mechanism, should one be evidenced, would necessarily be quite marginal \citep{Hankamer1992,Chan2008}.
+Furthermore, in languages with rich agglutinative morphology, the role played by a whole word storage mechanism, should one be evidenced, would necessarily be quite marginal \citep{Anderson1988a,Hankamer1992,Chan2008}.
 According to one estimate, children acquiring American English hear 165,000 word tokens a week \citep{Swingley2007}.
@@
 If speakers are shown to have internalize a phonotactic generalization, then there must be a phonological process to blame.
-One clear case where processes, and not merely restrictions on underlying representations, are implicated is in Quecha, where ejective co-occurrence restrictions are reflected in speakers' difficulties in producing nonce words containing multiple ejectives \citep{GallagherInPress}.
+One clear case where processes, and not merely restrictions on underlying representations,
+is where co-occurrence restrictions cause difficulties in producing nonce words violating those restrictions
+\citep{Rose2007,GallagherInPress}
+
+are implicated is in Quecha, where ejective co-occurrence restrictions are reflected in speakers' difficulties in producing nonce words containing multiple ejectives \citep{GallagherInPress}.
 
@@
 For instance, there is a long-standing debate concerning the representation of Sanskrit aspiration alternations like those represented by \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
-One analysis, going back to the ancient grammarian Pāṇini \citep{Sag1974} and to \citet[\S141f.]{Whitney1889}, posits an underlying representation /budh-/ with aspiration shifting leftward in certain contexts: modern proponents of this analysis include \citet{Borowsky1983}, \citet[][59f.]{Hoenigswald1965}, \citet{Kaye1985}, \citet{Sag1974,Sag1976} \citet{Schindler1976}, \citet{Stemberger1980}, and \citet[][\S141f.]{Whitney1889}.
-An alternative analysis posits an underlying representation /bhudh-/ and an apsirate dissimilation process (\citealt{Anderson1970}, \citealt{Hoard1975}, \citealt[\S3.2]{Kiparsky1965}, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt[109f.]{Zwicky1965}).
+One analysis, going back to the ancient grammarian Pāṇini \citep{Sag1974} and to \citet[\S141f.]{Whitney1889}, posits an underlying representation /bud\asp{}-/ with aspiration shifting leftward in certain contexts: modern proponents of this analysis include \citet{Borowsky1983}, \citet[][59f.]{Hoenigswald1965}, \citet{Kaye1985}, \citet{Sag1974,Sag1976} \citet{Schindler1976}, \citet{Stemberger1980}, and \citet[][\S141f.]{Whitney1889}.
+An alternative analysis posits an underlying representation /b\asp{}ud\asp{}-/ and an apsirate dissimilation process (\citealt{Anderson1970}, \citealt{Hoard1975}, \citealt[\S3.2]{Kiparsky1965}, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt[109f.]{Zwicky1965}).
 It is not my intention to adjudicate between these competing analyses, or to evaluate the claim that both accounts are deeply flawed \citep{Janda1989}, but simply to note that they make different predictions under the theory proposed here.
-While the former, aspirate shift analysis denies the existence of diaspirate roots, this fact lacks synchronic motivation, so the co-occurrence restriction may go unlearned.
+While the former, aspirate shift analysis denies the existence of diaspirate roots, this fact lacks synchronic motivation, so the co-occurrence restriction go unlearned.
 On the other hand, under the aspirate dissimilation account, underlying diaspirate roots exist but speakers internalize synchronic processes responsible for their absence on the surface.

### 2012-10-14 3.tex
Checking on glosses

@@
    & {kiler}        & {kilerler} & `pantry'       \\
-   & {pelür}        & {pelürler} & `tissue paper' \\
+   & {pelür}        & {pelürler} & `onionskin'    \\
    & {boğaz}        & {boğazlar} & `throat'       \\
@@
    & {silah}        & {silahlar} & `weapon'                           \\
-   & {memur}        & {memurlar} & `bureaucrat'                       \\
+   & {memur}        & {memurlar} & `official'                         \\
    & {sabun}        & {sabunlar} & `soap'                             \\
@@
 \noindent
-There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} despite the fact that their final root vowel is [$+$\textsc{Back}].
+There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} though their final root vowel is [$+$\textsc{Back}].
 While suffix disharmony appears to very sporadic, the co-occurrence of suffix harmony and root disharmony has been the subject of much debate.
@@
 a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
-   & {fren}    & \alt{} & {firen}    & `break'     \\
+   & {fren}    & \alt{} & {firen}    & `brake'     \\
 b. & {trablus} & \alt{} & {tırablus} & `Tripoli'   \\
@@
 b. & {boğaz}        & {boğazı}       & {boğazın}      & `throat'  & \citep{TELL} \\
-   & {pelür}        & {pelürü}       & {pelürün}      & `tissue' \\
+   & {pelür}        & {pelürü}       & {pelürün}      & `onionskin' \\
    & {döviz}        & {dövizi}       & {dövizin}      & `currency'  \\
@@
 \citet{NiChiosain1993} and \citet{Ito1995b} argue that \textsc{Labial Attraction} is true only of native vocabulary.
-This would be a potential confound for the experimental results, since it is not implausible that the subjects' in \citeauthor{Zimmer1969}'s study treated nonce words as if they were loanwords.\footnote{Indeed, many wordlikeness studies include instructions to the participants to treat the stimuli much as if they were loanwords \citep[e.g.,][]{Hay2004a}.
-This may influence the choice of task model used by participants.} However, this was not the opinion of \citet[266]{Lees1966a}, and is conclusively refuted by  \citet{Inkelas2001}.
+This would be a potential confound for the experimental results, since it is not implausible that the subjects' in \citeauthor{Zimmer1969}'s study treated nonce words as if they were loanwords.
+%\footnote{Indeed, many wordlikeness studies include instructions to the participants to treat the stimuli much as if they were loanwords \citep[e.g.,][]{Hay2004a}.
+This may influence the choice of task model used by participants.}
+This was not the opinion of \citet[266]{Lees1966a}, and it is conclusively refuted by  \citet{Inkelas2001}.
 They observe that in the subset of the TELL database with etymological coding, a full 75\% of foreign vocabulary conform to \textsc{Labial Attraction}, which is surprisingly a smaller percentage than that of the foreign vocabulary (52\%), a significant trend (Fisher exact test, $p = 0.042$).
@@
 These two varieties of evidence are not fully independent, however: extragrammatical \emph{channel bias} effects are thought to be a first-order predictor of attestation \citep{Blevins2004,Moreton2008}.
-In light of extensive evidence for generalizations which are phonetically unnatural \citep[e.g.,]{Anderson1981,Bach1972,Buckley2000a}, even as revealed in nonce word studies \citep[e.g.,][]{Hayes2009}, the fact that the generalizations extracted in the \citet{Becker2011} study are phonetically natural is either accident or conundrum.
+In light of extensive evidence for generalizations which are phonetically unnatural \citep[e.g.,]{Anderson1981,Bach1972,Buckley2000a}, even as revealed in nonce word studies \citep[e.g.,][]{Hayes2009}, the generalizations extracted in the \citet{Becker2011} study are phonetically natural is either accident or conundrum.
 It is certainly not something that can be explained by the tacit theory of possible gneeralizations the authors adhere to.

### 2012-10-22 3.tex
edits on 3

@@
 There are several arguments against this inference.
-First, linguists have long noted statistical generalizations which have a diachronic cause but which lack any synchronic reality.
+First, there are many cases of statistical generalizations with a known diachronic cause but lacking any synchronic reality.
 One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}), who recognizes the a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}, and that Classical Latin intervocalic \emph{s} is rare as a result.
@@
 This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations.
-The danger is that grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may not match speakers' knowledge.
+The danger is that grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may overstate the phonotactic generalizations internalized by speakers.
 
 This chapter is a comprehensive investigation of three well-known phonotactic generalizations in Turkish using lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}.
-\citeauthor{Zimmer1969}'s experiments merit reconsideration because there is widespread disagreement concerning the results and prior discussions fail to relate the behavioral data to lexical statistics and to competing formalizations of the generalizations involved.
+\citeauthor{Zimmer1969}'s experiments merit reconsideration because there is widespread disagreement concerning the results and prior discussions do not relate behavioral data to lexical statistics and to competing formalizations of the generalizations involved.
 
@@
 Two less familiar notations are used in this chapter.
-First, directional application conditions are used to derive the left-to-right, iterative properties of.
+First, directional application conditions are used to derive the left-to-right, iterative properties of the harmony rules.
 Directional application explains provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}).
@@
 Further, \citet{Johnson1972} and \citet{Kaplan1994} prove that rules specified for directional application are no more complex in the terms of formal learnability than that do not.
-Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used.
-Following \citet{McCawley1973}, a structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses well-known counterexamples against this restriction adduced in \emph{SPE} (352-353).}
+Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used \citep{McCawley1973}.
+A structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses well-known counterexamples against this restriction adduced in \emph{SPE} (352-353).}
 
@@
 
-It is necessary to dispense with an alternative proposed by \citet{Clements1982} and \citet{Inkelas1997} which reintroduces this duplication.
+It is important to dispense with an alternative proposed by \citet{Clements1982} and \citet{Inkelas1997} which reintroduces this duplication.
 Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
@@
 
-While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show internalize the tendency for roots to conform to backness harmony.
+While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show that Turkish speakers internalize the tendency for roots to conform to backness harmony.
 If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it.
@@
 \citet{Suomi1997} and \citet{Vroomen1998} task Finnish speakers with identifying harmonic disyllables in an auditory stream.
-When the syllable preceding the disyllabic target has a different backness specification than the target, this faciliates speakers' recognition abilities.
+When the syllable preceding the disyllabic target has a different backness specification than the target, recognition of the target is facilitated.
 Presumably, disharmony facilitates the recognition of word boundaries.
@@
 \noindent 
-This claim yields itself naturally to the Fisher exact test.
+This claim can be evaluated using the Fisher exact test.
 Let P denote a sequence of one or more consonants, one of which is labial, and let T denote a sequence of one or more consonants none of which is labial.
@@
 
-Both of the experiments performed by \citet{Zimmer1969} include 5 pairs which differ on in whether or not the nonce words conform to, or violate, \textsc{Backness Harmony}.
+Both of the \citet{Zimmer1969} experiments include 5 pairs which differ on in whether or not the nonce words conform to, or violate, \textsc{Backness Harmony}.
 As can be seen from Table \ref{bhw}, harmonic pairs are preferred approximately 6-to-1, and aggregating over speakers, no disharmonic member of a pair is favored.
@@
 This was not the opinion of \citet[266]{Lees1966a}, and it is conclusively refuted by  \citet{Inkelas2001}.
-They observe that in the subset of the TELL database with etymological coding, a full 75\% of foreign vocabulary conform to \textsc{Labial Attraction}, which is surprisingly a smaller percentage than that of the foreign vocabulary (52\%), a significant trend (Fisher exact test, $p = 0.042$).
+They observe that in the subset of the TELL database with etymological coding, a full 75\% of foreign vocabulary conform to \textsc{Labial Attraction}.
+Among native lexical items, only 52\% do, a significant trend but in an unexpected direction (Fisher exact test, $p = 0.042$).
 One plausible explanation is that many of the languages in contact with Turkish lack the \emph{ı} phoneme needed to violate \textsc{Labial Attraction}.
@@
 However, in the case of constraints on Turkish vowels, it is a necessary condition that the constraint in question be derived from an alternation.
-Even alternations that have large numbers of lexical exceptions have have robust effects on a constrained wordlikeness task.
-Purely statistical approaches to phonotactic knowledge fail to draw this important distinguish.
+Even alternations that have large numbers of lexical exceptions have robust effects on a constrained wordlikeness task.
+Purely statistical approaches to phonotactic knowledge fail to draw this important distinction.
 

### 2012-10-23 2.tex
Writing on 1-4, especially edits on 3

@@
 \citet{Halle1962} and \citet{Chomsky1965} use two nonce words, \emph{blick} [blɪk] and \emph{bnick} [bnɪk], to illustrate that speakers internalize generalizations about possible and impossible words in their language.
-While \emph{blick} is unattested, but in some sense a ``possible'' word of English, \citeauthor{Chomsky1965} observe that speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word, despite the fact that it is well-formed in other languages (e.g., Moroccan Arabic \emph{bniqa} `closet').
+While \emph{blick} is unattested, but in some sense a ``possible'' word of English, \citeauthor{Chomsky1965} observe that speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word, though it is well-formed in other languages (e.g., Moroccan Arabic \emph{bniqa} `closet').
 
@@
 87 items were presented to 24 undergraduate subjects, who rated each word on a seven-point Likert scale.
-The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``complete normal, would make a fine English word''.
+The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``completely normal, would make a fine English word''.
 
@@
 This has the desirable effect that no nonce word is ever assigned a zero probability, and produces a small increase in the correlation between the \citeauthor{Albright2003b} wordlikeness norms compared with the maximum likelihood estimate (Table \ref{albrightimproved}).
-For all three data sets, this model also consistenly outperforms positional probability models defined by \citet{Vitevitch2004} and \citet{Vaden2009}.
-Given this and the fact that these model scores are highly correlatd with bigram probability \citep[][54]{Vitevitch1997}, they are not considered further.
+For all three data sets, this model also consistenly outperforms positional probability models defined by \citet{Vitevitch2004} and \citet{Vaden2009}; given that these model scores are highly correlatd with bigram probability \citep[][54]{Vitevitch1997}, they are not considered further.
 The bigram model consistently performs well in all the evaluations, and has the highest Spearman correlation with the \citeauthor{Greenberg1964} and \citeauthor{Scholes1966} data, and is frequently second place model to the binary baseline elsewhere.
@@
 Despite this, there are relatively strong correlations between the binary baseline and these gradient models (see Table \ref{bcor}).
-However, the fact that the gradient models are generally outperformed by the binary baseline suggest that they do not reliably predict intermediate ratings.
+From the strong performance of the categorical model one can infer that the gradient models do not reliably predict intermediate ratings, or contrasts in ratings between words which are grouped together.
 To quantify this, the following method was used to estimate the residual contribution of the three gradient models once gross phonotactic violations are taken into account.

### 2012-10-23 4.tex
Writing on 1-4, especially edits on 3

@@
 Here, this is operationalized in the same fashion as \textsc{Degemination}, by definining ``identity'' as consonants which share the same place and manner features, but which do not necessarily agree on voicing.
-Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters.
+Though the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters.
 

### 2012-10-23 pre.tex
Writing on 1-4, especially edits on 3

@@
 
-Similarly, theorists of lexical processing postulate that complex words are stored in memory to account for the facilitatory effect of whole word frequency on reaction times to complex words \citep[e.g.,][]{Alegre1999,Baayen1997a,Sereno1997}.
+Similarly, theorists of lexical processing postulate that complex words are stored in memory to account for the facilitatory effect of whole word frequency on reaction times to complex words \citep[e.g.,][]{Alegre1999,Baayen1997b,Sereno1997}.
 But this effect of word frequency obtains regardless of whether or not whole word storage is postulated \citep{Taft2004b,LignosInPress}.

### 2012-10-25 3.tex
chapter 3 draft is done.

@@
 In purely statistical terms, an accidental gap is one which is likely to arise without any antecedent cause, i.e., by chance.
-The contribution of early generative discussions of this contrast \citep[e.g.,][]{Halle1962,Chomsky1965,SPE} is to ground the definitions of \emph{structural} and \emph{accidental} gaps in mentalism: an accidental gap is one which which lacks a mental antecedent because it goes unlearned by speakers.
+The contribution of early generative discussions of this contrast \citep[e.g.,][]{Halle1962,Chomsky1965,SPE} is to ground the definitions of \emph{structural} and \emph{accidental} gaps not in statistical expectations, but rather in mentalism.
+For \citeauthor{SPE}, an accidental gap is one which is not learned, one which lacks a mental antecedent.
 
@@
 In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique to validate a generalization about Arabic root consonant co-occurrence facts originally due to \citet{Greenberg1950}.
-The assumption appears to be that statistically valid static co-occurrence restrictions will be reflected by wordlikeness judgements.
+The assumption appears to be that static co-occurrence restrictions which have a statistically significant trace in the lexicon will also be reflected in psycholinguistic tasks such as wordlikeness judgements.
 While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}.
@@
 There are several arguments against this inference.
-First, there are many cases of statistical generalizations with a known diachronic cause but lacking any synchronic reality.
-One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}), who recognizes the a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}, and that Classical Latin intervocalic \emph{s} is rare as a result.
-However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.} Similarly, [ʃ] in Modern English derives from Old English [sk] (e.g., \emph{fisc} 'fish'), and since long vowels before complex codas are not found in Old English, [ʃ] is rarely preceded by long vowels in word-final syllables to this day.
-\citet{Iverson2005} label this constraint ``phonologically accidental'', since a millennium of speakers have produced coinages (e.g., \emph{posh}, \emph{swoosh}) and borrowings (e.g., \emph{douche}, \emph{capiche}) in utter disregard of this generalization.
-Secondly, wordlikeness generalizations are absolutely constrained by negative results in learnability (see \citealt{Yang2012} for a recent review).
-Finally, there are a number of results which suggest patterns which were obvious to the linguist go unlearned \citep[e.g.,][]{Becker2011,Hayes2006,Hayes2009,HayesInPress}.
+First, negative results in formal learnability theory (see \citealt{Yang2012} for a recent review) show that there are whole classes of rules and constraints which cannot be learned under standard assumptions about acquisition (e.g., that the data is finite, can appear in any order, and that speakers do not have to request grammatical judgements from adults to achieve adult-like competence).
+
+Secondly, there are a number of results which statistically reliable patterns are not learned by speakers \citep[e.g.,][]{Becker2011,Hayes2006,Hayes2009,HayesInPress}.
 This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations.
@@
 
-This chapter is a comprehensive investigation of three well-known phonotactic generalizations in Turkish using lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}.
-\citeauthor{Zimmer1969}'s experiments merit reconsideration because there is widespread disagreement concerning the results and prior discussions do not relate behavioral data to lexical statistics and to competing formalizations of the generalizations involved.
+Another type of non-learning occurs when generalizations are statistically reliable but demonstrably incorrect.
+\citet{Legate2012}, for instance, note that the overwhelming majority of English words have initial primary stress.
+While this suggests a quantity-insensitive stress system, English stress is in fact quantity-sensitive (see \citealt{Halle1998c} and references therein).
 
-One of these generalizations, \textsc{Labial Attraction}, is shown to be highly statistically reliable but to be inactive in the wordlikeness task.
+Finally, there are many cases of statistical generalizations with a known diachronic cause but lacking any synchronic reality.
+One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}) in a discussion of rhotacism, a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}.
+As a result, Classical Latin intervocalic \emph{s} is rare.
+However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.}
+
+Similarly, [ʃ] in Modern English derives from Old English [sk] (e.g., \emph{fisc} `fish'), and since long vowels before complex codas are not found in Old English, [ʃ] is rarely preceded by long vowels in word-final syllables to this day according to \citet{Iverson2005}.
+The type frequencies of short and long vowels before final [ʃ] and a similar segment, [s] are shown in Table \ref{ssh}.
+As can be seen, long vowels before final [s] are twice as common before final [ʃ]; according to the appropriate statistical test (the Fisher exact test, described below), this is significant.
+Yet \citet{Iverson2005} label this constraint ``phonologically accidental'' since a millennium of speakers have produced coinages (e.g., \emph{posh}, \emph{swoosh}) and  borrowings (e.g., \emph{douche}, \emph{capiche}) that disregard this generalization.\footnote{In the loanword adaptation literature, there are many reports that statistically reliable generalizations over the native lexicon are ignored in nativization processes \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}.
+On the other hand, \citet{Frisch2001} suggest that novel borrowings in Egyptian Arabic rarely violate co-occurrence restrictions in that language; however, the generalization in question was exceptionless from Proto-Semitic \citep{Ehret1989} through Classical Arabic \citep{Greenberg1950}, so any violations of this generalization are unexpected.}
+
+\begin{table}
+\centering
+\begin{tabular}{l r r r r}
+\toprule
+          & \{ɪ, ɛ, æ, ʌ, ʊ\}\gap{}\# & \{i, e, ɑ, ɔ, u\}\gap{}\# & \% long & $p$-value \\
+\midrule
+\gap{}ʃ\# & 78                & 9                 & 8\%      & \multirow{2}{*}{0.026} \\
+\gap{}s\# & 410               & 107               & 16\%     & \\
+\bottomrule
+\end{tabular}
+\caption{Type frequencies of short and long vowels before word-final [ʃ] and [s] in a subset of the CMU pronunciation dictionary containing all words with token frequency greater than 1 per million words in the SUBTLEX-US frequency norms; the 9 words containing ending in a long vowel followed by ʃ are \emph{douche}, \emph{leash}, \emph{gosh}, \emph{josh}, \emph{posh}, \emph{squash}, \emph{unleash}, \emph{wash}, and \emph{woosh}.}
+\label{ssh}
+\end{table}
+
+This chapter is a comprehensive investigation of three phonotactic generalizations in Turkish, comparing lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}.
+Both the lexical statistics and \citeauthor{Zimmer1969}'s experimental results merit reconsideration, because prior discussions do not relate the lexical statistics to competing formalizations of the generalizations involved, and there is widespread disagreement concerning the experimental findings.
+
+One of the three generalizations, \textsc{Labial Attraction}, is shown to   state a statistically reliable generalization about the Turkish lexicon, but has no effect on speakers' wordlikeness judgements.
 This further undermines the widespread use of statistical criteria to determine which of the vast number of statistically reliable phonotactic generalizations are internalized by speakers and which are not.
-The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is phonologically inactive, corresponding to no phonological process in the language.
+The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is static, corresponding to no phonological process in the language.
 
@@
 
-\citet{Lees1966b,Lees1966a} proposes three constraints on Turkish vowel sequences, constraints which are the focus of many subsequent studies.
+\citet{Lees1966b,Lees1966a} proposes three constraints on Turkish vowel sequences; these constraints are the focus of many subsequent studies.
 In this section, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence bearing on speakers' knowlege of the restrictions.
-The following feature specification for the eight vowels of Turkish vowels is assumed throughout.
+The following feature specification for the eight vowels of Turkish is assumed throughout.
 
@@
 Two less familiar notations are used in this chapter.
-First, directional application conditions are used to derive the left-to-right, iterative properties of the harmony rules.
-Directional application explains provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}).
-In fact, harmony in Turkish is one such case, insofar as simultaneous application makes the wrong predictions.
-Further, \citet{Johnson1972} and \citet{Kaplan1994} prove that rules specified for directional application are no more complex in the terms of formal learnability than that do not.
+First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules.
+Directional application provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}).
+Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
+
 Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used \citep{McCawley1973}.
-A structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses well-known counterexamples against this restriction adduced in \emph{SPE} (352-353).}
+A structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].
+This is more restrictive than Greek-letter notation, in that it prevents the value of one feature being applied to another.
+\citet{Odden2012} argues that the two counterexamples against this restriction adduced in \emph{SPE} (352-353) are not probative.
 
@@
 \textsc{Backness Harmony} also triggers alternations in inflectional suffix vowels.
-For instance, the nominative plural (nom.pl.) suffix is \emph{-ler}, when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].
+For instance, the nominative plural (nom.pl.) suffix is \emph{-ler} when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].
 
@@
 a. & {ip}           & {ipler}    & `rope'         & \citep[][216]{Clements1982} \\
-%  & {el}           & {eller}    & `hand'         \\
    & {köy}          & {köyler}   & `village'      \\
@@
    & {kız}          & {kızlar}   & `girl'         \\
-%  & {sap}          & {saplar}   & `stalk'        \\
-%  & {son}          & {sonlar}   & `end'          \\
    & {pul}          & {pullar}   & `stamp'        \\
@@
 In this case, suffix vowels generally exhibit harmony with the final root vowel.
+There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler}, although their final root vowel is [$+$\textsc{Back}].
+Interestingly, they may themselves be harmonic.
 
@@
 \begin{tabular}{lllll}
-   & \emph{nom.sg.} & \emph{nom.pl.}             &                    \\
+   & \emph{nom.sg.} & \emph{nom.pl.}&                    \\
 a. & {mezar}        & {mezarlar}    & `grave'       & \citep{TELL}       \\
@@
    & {saat}         & {saatler}     & `hour, clock' 	              \\
-%  & {harf}         & {harfler}  & `(alphabetic) letter' \\ 
+   & {kahabat}      & {kahabatler}  & `fault'       \\
 \end{tabular}
@@
 
-\noindent
-There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} though their final root vowel is [$+$\textsc{Back}].
-While suffix disharmony appears to very sporadic, the co-occurrence of suffix harmony and root disharmony has been the subject of much debate.
-As disharmonic roots still trigger suffix harmony,
-%\citet[212, 289]{Anderson1974} and
-\citet{Iverson1978} distinguish suffix harmony alternations from the sequence structure constraint governing root harmony.\footnote{\citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process.
+\citet[212]{Anderson1974} and \citet{Iverson1978} argue that suffix harmony in disharmonic roots found in (\ref{turkexcept}a) requires the rule governing suffix harmony alternations to be distinguished from a sequence structure constraint governing root harmony.
+Since the rule and sequence structure constraint are otherwise identical, this constitutes an undesirable ``duplication'' in the sense of \citealt{Kisseberth1970b}.
+However, the theory of exceptionality proposed in \emph{SPE} can account for these facts without duplication \citep[197f.]{Zonneveld1978}.\footnote{\citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process.
 \citet[171f.]{Howard1972} correctly observes that this duplication is unnecessary.}
 
-The disadvantage of this account is that it introduces ``duplication'' (in the sense of \citealt{Kisseberth1970b}) of sequence structure and phonological generalizations, differing only in their pattern of exceptionality.
-\citet[][197f.]{Zonneveld1978} shows that the theory of exceptionality proposed in \emph{SPE} can in fact account for suffix harmony in disharmonic roots.\footnote{An isomorphic analysis is available within less restrictive theories of phonological exceptionality \citep[e.g.,][]{Kisseberth1970,Pater2009}.} \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
+\citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
 A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}].
-In other words, no form is never truly an ``exception''; rather, some forms fail to match the extended structural description of $R$ which requires that the target be [$+R$].
+In other words, no form is never truly an ``exception''; rather, some underlying representations have non-default features which do not match the extended structural description of $R$ which requires that the target be [$+R$].
 If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.
-Thus \textsc{Backness Harmony} is able to account for both root and suffix harmony, and duplication is unnecessary.
 
-It is important to dispense with an alternative proposed by \citet{Clements1982} and \citet{Inkelas1997} which reintroduces this duplication.
+Under this account, root and suffix harmony are derived from \textsc{Backness Harmony}, but they do not have the same ontology: suffix harmony is direct result of phonological rule application, whereas any ``effects'' of the  generalization concerning harmonic roots arises from a dispreference for lexical exceptionality.
+
+\citeauthor{Anderson1974} also notes that roots which fail to undergo suffix harmony, like (\ref{turkexcept}b), may themselves be harmonic. He takes to be evidence for the necessity of duplication:
+
+\begin{quote}
+\ldots{}there are words which are exceptions to harmony across boundaries (e.g., \emph{kabahat} `fault', \emph{kabahattı} `his fault') but which are perfectly regular internally.
+Since the morpheme structure condition and the phonological rule in this case have distinct classes of exceptions, it is clear that they cannot be identified.
+\citep[289]{Anderson1974}
+\end{quote}
+
+\noindent
+Under the assumptions so far, there is no reason to think that \emph{kabahat} is [$+$\textsc{Backness Harmony}], however: even roots with all back vowels need not undergo this rule.
+Whatever the ultimate analysis of those few words which fail to show suffix harmony, nothing depends on whether or not they are harmonic.
+
+%\begin{example}[Turkish progressives]
+%\begin{tabular}{l l l l l}
+%   & \emph{1sg. present}  & \emph{1sg. progressive} & \\
+%a. & gelirim              & geliyorum               & `come'  \\
+%b. & görürüm              & görüyorum               & `see'   \\
+%c. & atarım               & atıyorum                & `throw' \\
+%d. & bulurum              & buluyorum               & `find'  \\
+%\end{tabular}
+%\end{example}
+
+It is necessary to dispense with an alternative analysis proposed by \citet{Clements1982} and \citet{Inkelas1997}.
+This analysis has desirable properties, but a subtlety of it reintroduces the duplication of rule and sequence structure constraint.
 Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
 \emph{kul} `servant', \emph{kepek} `bran' vs.
-\emph{kapak} `lid'), whereas harmonic roots are those in which the backness of any remaining vowels is predictable.
-\citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots are fully specified.
+\emph{kapak} `lid'), whereas backness of vowels in non-initial syllables is predictable in harmonic roots; there are no prefixes in Turkish.
+\citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots and of certain exceptional suffixes are fully specified.
 This is schematized below.
@@
 
-One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being a Structure Preservation condition \citep{Kiparsky1985}.
+One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, either by non-derived environment blocking \citep{Mascaro1976} or structure preservation condition \citep{Kiparsky1985}.
 However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default.
@@
 If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it.
-Several other ``external'' facts suggest  that this is not the case, with the caveat that uncritical acceptance of hypotheses linking loanword adaptation and language games to the grammar is no way being implied.
+Several other ``external'' facts suggest that this is not the case.
+The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.
 
-The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} While some speakers are said to be capable of pronouncing these non-native clusters, in fast speech the cluster is reportedly split by anaptyxis.
+The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} Some speakers are said to be capable of pronouncing these non-native clusters, but in fast speech the cluster is split by anaptyxis.
 In most cases, this vowel matches the following root vowel for backness.
@@
    & {prusya}  & \alt{} & {purusya}  & `Prussia'   \\
-%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
-%   & kredi   & \alt{} & kıredi   & `credit'    \\
 \end{tabular}
+\label{spiker}
 \end{example}
@@
 \noindent
-It is worth noting that this cannot be merely the consequence of application of the native \textsc{Backness Harmony} rule, since the leftmost vowel of a prosodic word is not subject to harmonization.
-Rather, the adaptation of non-native onset clusters proceeds in such a fashion so that the lexical items in question are rarely [$-$\textsc{Backness Harmony}].
+It is unclear whether the cluster-splitting vowel is deleted in the non-native variant or epenthesized in the fast speech variant.
+Under either analysis, there is no ready explanation for the tendency of the cluster-splitting vowel to have the backness features of following consonants; if anything, one might expect it to determine the backness features of following consonants.
+All that can be said with certainty is that the adaptation of non-native onset clusters appears to proceed in such a fashion so that the lexical items in question are [$-$\textsc{Backness Harmony}].
 
@@
 In this game, the base is reduplicated and the first vowel of the reduplicant replaced with a [$+$\textsc{Back}] vowel.
-In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the the inserted [$+$\textsc{Back}] vowel.
+In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the inserted [$+$\textsc{Back}] vowel.
 The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}
@@
 \noindent 
-\citeauthor{Harrison2001} present an analysis of this game couched in the vowel underspecification proposed by \citeauthor{Clements1982}, but the data is equally consistent with a full specification analysis.
-Reharmonization is the result of \textsc{Backness Harmony} in the reduplicant.
-On the other hand, the lack of reharmonization in the disharmonic roots indicates that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.
+In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant.
+On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{The strongest argument that reduplication preserves diacritic features of the root would be a case where lexical exceptionality surfaces only in reduplicants.
+Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, provides such a case.
+Verb reduplication requires a bisyllabic reduplicant.
+This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots actually have three copies of the root, and since reduplicated forms of many trisyllable verbs are ineffable.
+However, a few trisyllabic verbs exceptionally show full reduplication.
+Exceptionality appears to be a lexical property (since it is not predictable), but it surfaces only in the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.
 
@@
 Presumably, disharmony facilitates the recognition of word boundaries.
-\citet{Kabak2010} find that Turkish \textsc{Backness Harmony} has a a similar effect: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}).
+\citet{Kabak2010} find that Turkish \textsc{Backness Harmony} has a similar effect: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}).
 \citet{Kabak2010} find that this effect does not obtain for speakers of French, a language which lacks vowel harmony.
@@
 
-The Turkish word-spotting experiment has been adapted for infants by \citet{Kampen2008}.
+The Turkish word-spotting experiment is adapted for infants by \citet{Kampen2008}.
 In this study, 9-month-old infants are familiarized with an auditory stream containing harmonic disyllabic nonce words.
@@
 Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic nonce words such as \emph{paroz} over disharmonic nonce words like \emph{nelok}, but German 6-month-old infants show no such preference.
+
+It is not clear, however, that the segmentation effects are evidence for grammatical generalizations; for instance, while the effect is defined in terms of harmony, it operates over a domain much larger than the prosodic word over which harmony is computed.
 
@@
 a. & {ip}           & {ipi}          & {ipin}         & `rope' & \citep[][216]{Clements1982} \\
-%  & {el}           & {eli}          & {elin}         & `hand'    \\
    & {kız}          & {kızı}         & {kızın}        & `girl'    \\
    & {sap}          & {sapı}         & {sapın}        & `stalk'  \\
-%  & {yüz}          & {yüzü}         & {yüzün}        & `face'    \\
    & {köy}          & {köyü}         & {köyün}        & `village' \\
-%  & {pul}          & {pulu}         & {pulun}        & `stamp'   \\
    & {son}          & {sonu}         & {sonun}        & `end'     \\
@@
 However, two pieces of the external evidence on \textsc{Backness Harmony} bear on this question.
-First, the epenthetic vowel in non-native word-initial onsets like \emph{trablus} \alt{} \emph{tırablus} conforms to \textsc{Roundness Harmony} in addition to \textsc{Backness Harmony}.
+First, the cluster-splitting vowel found in non-native onset clusters
+, discussed above, tends to agree in roundness with following high vowels (e.g., \emph{prusya}-\emph{purusya} `Prussia').
 Secondly, \textsc{Roundness Harmony} participates in reharmonization in the language game described by \citeauthor{Harrison2001}: the second \emph{ü} in \emph{bütün} `whole' reharmonizes to the [$-$\textsc{Rnd}] vowel \emph{ı} in reduplicated \emph{bütün-batın}.
@@
 \begin{example}[\textsc{Labial Attraction}]
-%$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} -\textsc{Cons} \\ -\textsc{Rnd} \\ +\textsc{Back} \\ -\textsc{High} \end{bmatrix}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
 $\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\textrm{ɑ}~\textrm{C}_0~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\textrm{C}_0~\gap{}$
@@
 However, \textsc{Labial Attraction} does not apply in derived environments: the gen.sg.
-of \emph{sap} `stalk is \emph{sapın} rather than *\emph{sapun}.
-\citet[286]{Lees1966a} and \citet[311]{Zimmer1969} note the existence of root-internal exceptions to this rule (e.g., \emph{tavır} `mode') but agree that they are surprisingly rare.
+of \emph{sap} `stalk is \emph{sapın} rather than *\emph{sapun} that would be predicted if \textsc{Labial Attraction} triggered alternations.
+\citet[286]{Lees1966a} and \citet[311]{Zimmer1969} note the existence of root-internal exceptions to this rule (e.g., \emph{tavır} `mode') but they agree that they are surprisingly rare.
 
@@
 
-\citet[311]{Zimmer1969} administers two binary paired wordlikeness task designed to evaluate native speakers' knowledge of \textsc{Backness Harmony}, \textsc{Roundness Harmony}, and \textsc{Labial Attraction} in roots.
+\citet[311]{Zimmer1969} administers two paired wordlikeness tasks designed to evaluate native speakers' knowledge of \textsc{Backness Harmony}, \textsc{Roundness Harmony}, and \textsc{Labial Attraction} in roots.
 Speakers are presented with two nonce words and simply have to indicate the nonce word that is more Turkish-like.\footnote{Compared to the unpaired ratings tasks  commonly used in wordlikeness research, paired rating tasks have considerably more statistical power \citep{Gigerenzer2004}.
-When items under comparison differ only in whether or not they conform to the phonotactic generalizations, it is less likely that the any contrast between the members of the paris are caused by some omitted variable.
+When items under comparison differ only in whether or not they conform to the phonotactic generalizations, it is less likely that any contrast between the members of the pairs are caused by some omitted variable.
 Also, there is no unnecessary deception regarding the purpose of the experiment, which is known to reduce spurious response variability.
 Consequently, paired tasks are particularly sensitive to small phonotactic contrasts and are ideal for collecting wordlikeness judgements.} \citeauthor{Zimmer1969} concludes that the former two rules are reflected in wordlikeness judgements, whereas \textsc{Labial Attraction} is not.
-At the same time, \citeauthor{Zimmer1969} also suggests that speakers do internalize a variant of the rule that does not require a labial consonant in the trigger, a proposal that is endorsed by \citet{Clements1982} and \citet[4]{Zuraw2000}.
-\citet[196]{Inkelas2001}, argue instead that the trigger must contain a labial consonant but does not require a preceding \emph{a} vowel.
-Once again, no inferential statistics over the wordlikeness results are reported.
 
-Below, both lexical statistics and \citeauthor{Zimmer1969}'s wordlikeness results are analyzed statistically, so as to resolve the debate concerning the synchronic reality of \textsc{Labial Attraction} and the connection between speakers' wordlikeness judgements and statistical patterns in the lexicon.
-\textsc{Labial Attraction} is shown to be a statistically robust generalization over the Turkish lexicon, but no variant of \textsc{Labial Attraction} is reflected in \citeauthor{Zimmer1969}'s wordlikeness study.
+Below, both lexical statistics and \citeauthor{Zimmer1969}'s wordlikeness results are analyzed statistically; \textsc{Labial Attraction} is shown to be a statistically robust generalization over the Turkish lexicon, but no variant of \textsc{Labial Attraction} is reflected in \citeauthor{Zimmer1969}'s wordlikeness study.
 In contrast, the two harmony processes have robust effects both on the lexicon and on wordlikeness.
@@
 
-Counts were computed by regular expression matching on a 9,601-root subset of the TELL database, which were elicited from a native speaker, and which show no surface variation between or within speakers in any inflected form.
+Counts were computed by regular expression matching on a 9,601-root subset of the TELL database which consists of words which show no surface variation between the two informants in any inflected form.
 
@@
 The counts in this table are not expected to add up to 9,601, since many roots neither could exemplify nor violate the process in question; for instance, monosyllabic words are irrelevant to root harmony.
-The Fisher exact test is used to compute a $p$-value representing the probability of the observed data arising under the null hypothesis that there is no association between the process and type frequency; if this $p$-value is small, the null hypothesis is rejected and there is evidence that the process is reflected in the lexicon.
+The Fisher exact test is used to compute a $p$-value representing the probability of the observed data arising under the null hypothesis that there is no association between the process and type frequency.
 
@@
 Any disagreement on the backness specification of vowels in adjacent syllables are counted solely as exceptions, even if other vowel transitions in the root are harmonic.
-This follows from the hypothesis that phonological exceptionality is a property of whole underlying representations, not individual segments.
+%This follows from \emph{SPE} theory of lexical exceptionality (see also \citealt{Gouskova2012}), in which phonological exceptionality is a property of underlying representations, not individual segments.
 To construct the contingency table, roots are binned according to the backness specification of the first nucleus, and that of following nuclei.
@@
 
-\textsc{Roundness Harmony} consists of an association between the roundness of a vowel and the roundness of high vowels in the next syllable.
+\textsc{Roundness Harmony} predicts correlation between the roundness of a vowel and the roundness of high vowels in the next syllable.
 Any root for which a vowel does not agree in roundness with a high vowel in the following syllable (e.g., \emph{ümit}) is considered to be an exception.
@@
 \noindent
-The counts in bottom row of Table \ref{rhs} contain a number of roots which are apparent exceptions to \textsc{Roundness Harmony} but conform to \textsc{Labial Attraction} (bottom left), and which conform to \textsc{Roundness Harmony} at the expense of \textsc{Labial Attraction} (bottom right).
+The counts in the bottom row of Table \ref{rhs} contain a number of roots which are apparent exceptions to \textsc{Roundness Harmony} but conform to \textsc{Labial Attraction} (bottom left), and which conform to \textsc{Roundness Harmony} at the expense of \textsc{Labial Attraction} (bottom right).
 Excluding these types of roots would have the effect of slightly increasing the overall rate of \textsc{Roundness Harmony}, since the former is more common.
@@
 
-\citet{Clements1982} object to \textsc{Labial Attraction}:
+Reviewing the results of wordlikeness experiments, \citet{Zimmer1969} proposes a third variant of \textsc{Labial Attraction} which is insensitive to intervening consonant place.
+This is formalized below.
+
+\begin{example}[\textsc{Labial Attraction} (revised by \citealt{Zimmer1969})]
+$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\textrm{ɑ}~\textrm{C}_0~\gap{}$
+\end{example}
+
+\noindent
+\citet{Clements1982} also  to \textsc{Labial Attraction}, arguing that it is not ``systematic''.
 
 \begin{quote}
-Even decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots{}aCu\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots{}We conclude that there is no systematic restriction on the set of consonants that may occur medially in roots of the form /\ldots{}aCu\ldots/.
+Even more decisive evidence against a rule of Labial Attraction is the existence of a further, much larger set of roots containing /\ldots{}aCu\ldots/ sequences in which the intervening consonant or consonant cluster does not contain a labial\ldots{}We conclude that there is no systematic restriction on the set of consonants that may occur medially in roots of the form /\ldots{}aCu\ldots/.
 \citep[225]{Clements1982}
@@
 Let P denote a sequence of one or more consonants, one of which is labial, and let T denote a sequence of one or more consonants none of which is labial.
-The null hypothesis is that \emph{a}P\emph{u} sequences, with P representing a cluster containing a labial consonant, which conform to \textsc{Labial Attraction}, are no more likely than would be expected from other \emph{a}T\emph{u} sequences violating \textsc{Roundness Harmony}.
+The null hypothesis is that \emph{a}P\emph{u} sequences, which conform to \textsc{Labial Attraction}, are no more likely than would be expected from other \emph{a}T\emph{u} sequences violating \textsc{Roundness Harmony}.
 The resulting counts are shown in Table \ref{las}.
-Whereas the sequence \emph{a}P\emph{u} is more than twice as likely as \emph{a}P\emph{ı}, the sequence \emph{a}T\emph{u} is four times less likely than \emph{a}T\emph{ı}.
-This is a significant interaction predicted by \textsc{Labial Harmony}, but contradicting a statistical interpretation of the claim by \citeauthor{Clements1982}.
+Whereas the sequence \emph{a}P\emph{u} is more than twice as likely as \emph{a}P\emph{ı}, the sequence \emph{a}T\emph{u} is 5 times less likely than \emph{a}T\emph{ı}.
+This interaction is significant, as predicted by \textsc{Labial Attraction}, but contradicting a statistical interpretation of the \citeauthor{Clements1982}'s claim.
+In fact, \emph{a\ldots{}u} sequences are less, not more, common than \emph{a\ldots{}ı}.
 
@@
 
-Alternative formulations of \textsc{Labial Harmony} have been alluded to above. \citeauthor{Inkelas2001}, for instance, propose to eliminate the requirement of a preceding \emph{a} in the trigger:
+\noindent
+\citet[196]{Inkelas2001} propose instead that the trigger must contain a labial consonant but does not require a preceding \emph{a} vowel.
 
@@
 \noindent
+This is formalized below.
+
+\begin{example}[\textsc{Labial Attraction} (revised by \citealt{Inkelas2001})]
+$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\gap{}$
+\end{example}
+
 Counts from TELL for this formulation of the process are shown in Table \ref{lasi}.
@@
 
-Reviewing the results of wordlikeness experiments, \citet{Zimmer1969} proposes a third variant of \textsc{Labial Attraction} which is insensitive to intervening consonant place.
-This has no support in the TELL database, however: \emph{a\ldots{}u} sequences are less, not more, common than \emph{a\ldots{}ı}.
-This variant is not considered further.
-
 \subsection{Wordlikeness ratings}
@@
 
-Each response is coded as \emph{concordant} if the the nonce word conforming the the process is preferred, and \emph{discordant} if the disharmonic word is selected.
-To test for an association between the constraints, a non-parametric statistic, the Goodman-Kruskal (\citeyear{Goodman1954}) $\gamma$ is computed from to the count of concordant ($c$) and discordant ($d$) pairs:
+Each response is coded as \emph{concordant} if the nonce word conforming to the process is preferred, and \emph{discordant} if the disharmonic word is selected.
+To test for an association between the constraints, a non-parametric statistic, the Goodman-Kruskal (\citeyear{Goodman1954}) $\gamma$ is computed from the count of concordant ($c$) and discordant ($d$) pairs:
 
@@
 \noindent
-The $\gamma$ statistic ranges between -1, in the case that all paired choices are discordant, and 1, if all paired choices are concordant.\footnote{It is possible to perform statistical tests aggregating over items, but for this small number of items, such tests have very low (in some cases null) statistical power.}
+The $\gamma$ statistic ranges between -1, in the case that all paired choices are discordant, and 1, if all paired choices are concordant.\footnote{It also is possible to perform statistical tests aggregating over items, but for this small number of items, such tests have very low statistical power.}
 
@@
 
-Both of the \citet{Zimmer1969} experiments include 5 pairs which differ on in whether or not the nonce words conform to, or violate, \textsc{Backness Harmony}.
+Both of the \citet{Zimmer1969} experiments include 5 pairs which differ in whether or not the nonce words conform to, or violate, \textsc{Backness Harmony}.
 As can be seen from Table \ref{bhw}, harmonic pairs are preferred approximately 6-to-1, and aggregating over speakers, no disharmonic member of a pair is favored.
@@
 
-Both experiments include 5 pairs which are differ in the presence of absence of \textsc{Roundness Harmony}.
+Both experiments include 5 pairs which differ in the presence or absence of \textsc{Roundness Harmony}.
 As shown in Table \ref{rhw}, there is an approximately 5-to-1 preference for harmonic nonce words, and as was the case above, no disharmonic member of any pair is overall preferred.
@@
 
-Both experiments include five pairs which either conform to \textsc{Labial Attraction} and violate \textsc{Roundness Harmony}, or vis versa; the preferences are shown in Table \ref{law}.
-There is a small preference against \textsc{Labial Harmony}, though this is non-significant ($\gamma = -0.043$, $p = 0.305$).
-This shows that in general speakers do not have preferences as predicted by \textsc{Labial Harmony}.
+Both experiments include 5 pairs which either conform to \textsc{Labial Attraction} and violate \textsc{Roundness Harmony}, or vice versa; the preferences are shown in Table \ref{law}.
+There is a small preference against \textsc{Labial Attraction}, though this is non-significant ($\gamma = -0.043$, $p = 0.305$).
+Speakers do not have the preferences predicted by \textsc{Labial Attraction}.
 This result also extends for the variants of \textsc{Labial Attraction} proposed by \citet{Zimmer1969} and \citet{Inkelas2001}, since these variants have structural descriptions targeting a superset of the original formulation by \citet{Lees1966a}.
-It is interesting to note that speakers do not have the clear opposite preference, as is predicted by \textsc{Roundness Harmony}.
-A possible explanation for this is that \textsc{Roundness Harmony} is suspended in the context defined by \textsc{Labial Attraction} or some variant thereof.
+It is interesting to note, however, that speakers do not have the clear opposite preference, as is predicted by \textsc{Roundness Harmony}.
 
@@
 
-It has been shown that while \textsc{Labial Attraction} is a highly reliable generalization about Turkish root, it is not reflected in wordlikeness judgements.
-In contrast, harmony processes have a similar statistical profile, but have unambigious traces on wordlikeness.
+It has been shown that while \textsc{Labial Attraction} is a highly reliable generalization about Turkish roots, it is not reflected in wordlikeness judgements.
+In contrast, harmony processes have a similar statistical profile, but have large effects on wordlikeness.
 The most plausible explanation for this is that  \textsc{Labial Attraction} does not trigger alternations; indeed, it is counter-exemplified by the effects of \textsc{Roundness Harmony} in suffixes.
@@
 \citet{NiChiosain1993} and \citet{Ito1995b} argue that \textsc{Labial Attraction} is true only of native vocabulary.
-This would be a potential confound for the experimental results, since it is not implausible that the subjects' in \citeauthor{Zimmer1969}'s study treated nonce words as if they were loanwords.
+This would be a potential confound for the experimental results, since it is not implausible that the subject's in \citeauthor{Zimmer1969}'s study treated nonce words as if they were loanwords.
 %\footnote{Indeed, many wordlikeness studies include instructions to the participants to treat the stimuli much as if they were loanwords \citep[e.g.,][]{Hay2004a}.
 This may influence the choice of task model used by participants.}
-This was not the opinion of \citet[266]{Lees1966a}, and it is conclusively refuted by  \citet{Inkelas2001}.
+This was not the opinion of \citet[266]{Lees1966a}, though, and it is conclusively refuted by  \citet{Inkelas2001}.
 They observe that in the subset of the TELL database with etymological coding, a full 75\% of foreign vocabulary conform to \textsc{Labial Attraction}.
 Among native lexical items, only 52\% do, a significant trend but in an unexpected direction (Fisher exact test, $p = 0.042$).
-One plausible explanation is that many of the languages in contact with Turkish lack the \emph{ı} phoneme needed to violate \textsc{Labial Attraction}.
+One possible explanation is that many of the languages in contact with Turkish lack the \emph{ı} phoneme, and therefore cannot contribute exceptions to \textsc{Labial Attraction}.
 
-\citet{Becker2011} attribute the inactivity of \textsc{Labial Attraction} in wordlikeness to indicate that the generalization is phonologically unnatural, and therefore goes unlearned.
+\citet{Becker2011} attribute the inactivity of \textsc{Labial Attraction} in wordlikeness to a claim that it is phonologically unnatural, and therefore goes unlearned.
 This claim is difficult to evaluate insofar as phonotactic naturalness has continually eluded formalization and \citeauthor{Becker2011} neither propose nor refer to any theory thereof.
@@
 2]{Inkelas1997} suggest that \textsc{Labial Attraction} may have even induced alternations at one point in the history of Turkish.
-Beyond Turkish, the rounding of a high back vowel after a labial consonant is both widely attested \citep[e.g.,][]{Vaux1993}.
+Beyond Turkish, the rounding of a high back vowel after a labial consonant, as in the alternative formulation of \textsc{Labial Attraction} proposed by \citet{Inkelas2001}, is widely attested \citep[e.g.,][]{Vaux1993}.
 It is also phonetically natural, as both labial consonants and high round vowels are distinguished by a very low first and second formant.
-Regarding the non-local condition holding between the \emph{a} portion of the trigger and the \emph{u} target, Classical Arabic has many adjectives which form corresponding inchoative verbs by overwriting the vocalic melody with /a\ldots{}u/ (the final /-a/ is an inflectional suffix).
+As shown in Table \ref{lasi}, it is even a statistically reliable generalization about the Turkish lexicon.
+However, it is not reflected in alternations or in wordlikeness judgements.
 
-\begin{example}[Arabic derived inchoatives]
-\begin{tabular}{l l l l}
-\buf{}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
-\buf{}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
-\buf{}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
-\end{tabular}
-\end{example}
+%Regarding the non-local condition holding between the \emph{a} portion of the trigger and the \emph{u} target, Classical Arabic has many adjectives which form corresponding inchoative verbs by overwriting the vocalic melody with /a\ldots{}u/ (the final /-a/ is an inflectional suffix).
 
-\noindent
-In an Optimality Theory framework, for instance, constraints responsible for rounding of high back vowels before labials and for /a\ldots{}u/ overwriting can be conjoined to approximate the structural description of \textsc{Labial Attraction}.
+%\begin{example}[Arabic derived inchoatives]
+%\begin{tabular}{l l l l}
+%\buf{}[kabiːr]  & `big'      & [kabura]  & `become big'       \\
+%\buf{}[ħasan]   & `handsome' & [ħasuna]  & `become beautiful' \\
+%\buf{}[dʒadiːb] & `barren'   & [dʒaduba] & `become dry'       \\
+%\end{tabular}
+%\end{example}
+
+%\noindent
+%In an Optimality Theory framework, for instance, constraints responsible for rounding of high back vowels before labials and for /a\ldots{}u/ overwriting can be conjoined to approximate the structural description of \textsc{Labial Attraction}.
 
@@
 Ignoring the case of non-alternating root-final \emph{p} (e.g., \emph{ip}-\emph{ipi} `rope'), this is still not a surface-true generalization: \emph{p} occurs freely in intervocalic position: \emph{ahtapot} `octopus', \emph{köpük} `bubble', \emph{öpücük} `kiss'.
-More serious is the presence of root-internal intervocalic \emph{p} in roots which simultaneously exhibit the root-final \emph{p}-\emph{b} alternation: \emph{supap}-\emph{supabı} `valve', \emph{hipermetrop}-\emph{hipermetrobu} `far-sightedness'.
+More serious is the presence of root-internal intervocalic \emph{p} in ooots which simultaneously exhibit the root-final \emph{p}-\emph{b} alternation: \emph{supap}-\emph{supabı} `valve', \emph{hipermetrop}-\emph{hipermetrobu} `far-sightedness'.
 One prediction shared by virtually all theories of lexical exceptionality is that a lexical entry cannot simultaneously be a target for, and exception to, a phonological generalization, but this precisely what is needed to account for the behavior of \emph{supap}.
@@
 There is some reason to suspect that NDEB is a symptom of interface restrictions on phonological application, rather than a diagnosis itself.
-%The loanword adaptation literature also contains many reports that statistically reliable phonotactic generalizations about the native lexicon are absent in processes of nativization \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}.
-It is not immediately clear that this is evidence that the generalizations in question are external to the synchronic grammar, however.
 %\begin{example}[\textsc{Backness Harmony} lexical statistics]
@@
 It is not immediately clear that this is evidence that the generalizations in question are external to the synchronic grammar, however.
+
+%  & {harf}         & {harfler}  & `(alphabetic) letter' \\ 
+%  & {el}           & {eller}    & `hand'         \\
+%  & {sap}          & {saplar}   & `stalk'        \\
+%  & {son}          & {sonlar}   & `end'          \\
+%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
+%   & kredi   & \alt{} & kıredi   & `credit'    \\
+%  & {el}           & {eli}          & {elin}         & `hand'    \\
+%  & {yüz}          & {yüzü}         & {yüzün}        & `face'    \\
+%  & {pul}          & {pulu}         & {pulun}        & `stamp'   \\

### 2012-10-27 2.tex
Improvements to the appendices, primarily.

@@
 
-%\citet{Fischer-Jorgensen1952}
-%\citet{Saporta1958}
-%\citet{Saporta1957}
+\citet{Halle1962} and \citet{Chomsky1965} use the nonce words \emph{blick} [blɪk] and \emph{bnick} [bnɪk] to illustrate the contrast between possible words, accidental gaps in the lexicon, and impossible words, structurally excluded words, respectively.
+In \emph{SPE}, \citealt{SPE} introduce a third nonce word, \emph{bnzk}, so to illustrate that speakers have intuitions about multiple degrees of wellformedness.
+Both \emph{bnick} and \emph{bnzk} are possible words in other languages: stop-nasal onsets are found in Moroccan Arabic (e.g., \emph{bniqa} `closet') and stop-nasal-fricative-stop words in Imdlawn Tashlhiyt Berber \citep{Dell1984}.
+While neither are permitted in English, there is some real sense in which \emph{bnzk} is even less English-like than \emph{bnick}.
+\citeauthor{SPE} write:
 
-\citet{Halle1962} and \citet{Chomsky1965} use two nonce words, \emph{blick} [blɪk] and \emph{bnick} [bnɪk], to illustrate that speakers internalize generalizations about possible and impossible words in their language.
-While \emph{blick} is unattested, but in some sense a ``possible'' word of English, \citeauthor{Chomsky1965} observe that speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word, though it is well-formed in other languages (e.g., Moroccan Arabic \emph{bniqa} `closet').
+\begin{quote}
+Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
+\end{quote}
+
+In recent years there has been increased interest in the ``problem of admissibility'', and particularly in the multiple degrees of wellformedness inferred from wordlikeness tasks:
+
+\begin{quote}
+When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing.
+In the usual case, novel items fall along a gradient cline of acceptability.
+\citep[][9]{Albright2009a}
+
+In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
+\citep[][382]{Hayes2008a}
+
+\ldots{}when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
+\end{quote}
+
+%A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements. 
+% shademan
+
+There has been increased interest in the gradient aspects of wordlikeness judgements in recent years.
 
@@
 This may undermine any link between non-native realizations and epenthesis found elsewhere in English (e.g., allomorphs of the regular past tense and noun plural).} However, many linguists now regard the above as overly simplistic, for, as \citeauthor{Shademan2006} argues, such accounts draw a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report ``wordlikeness'' intuitions.
-
-\begin{quote}
-A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
-But when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values.
-\citep[371]{Shademan2006}
-\end{quote}
 
@@
 They write:
-
-\begin{quote}
-Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
-\end{quote}
 
@@
 \midrule
-Scholes & 33       & 66    & 2,079        \\
-A\&H    & 24       & 88    & 2,112        \\
+
+\begin{table}
+\centering
+\begin{tabular}{l r r r}
+\toprule
+                        & subjects & items & trials \\
+\midrule
+\citealt{Albright2003a} & 24       & 87    & 2,088  \\
+\citealt{Albright2007}  & 68       & 40    & 2,720  \\
+\citealt{HayesInPress}  & 29       & 80    & 2,320  \\
+\citealt{Scholes1966}   & 33       & 63    & 2,178  \\
+\midrule
+\textsc{Total}          & 154      & 270   & 9,306  \\
+\bottomrule
+\end{tabular}
+\caption{Subject and item counts}
+\label{counts}
+\end{table}
+
 \bottomrule
@@
 %It might be the case that syllable co-occurrence statistics might be little more than a reflection of infants' learning of categorical ``lexical viability'' constraints \citep{Johnson2003} of the sort also seen in adult speech processing \citep[e.g.,][]{Norris1997}.
+
+% other ratings, but por qué?:
+% 
+% Hayes2000
+% 
+% Albright2003a
+% Albright2003b
+% Prasada1993 
+% 
+% Bard1996

### 2012-10-27 3.tex
Improvements to the appendices, primarily.

@@
 In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant.
-On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{The strongest argument that reduplication preserves diacritic features of the root would be a case where lexical exceptionality surfaces only in reduplicants.
+On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{There is reason to believe that reduplicants bear lexical diacritics.
 Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, provides such a case.
@@
 However, a few trisyllabic verbs exceptionally show full reduplication.
-Exceptionality appears to be a lexical property (since it is not predictable), but it surfaces only in the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.
+Exceptionality appears to be a lexical property (since it is specific to certain lexical entries), but it surfaces only in the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.
 
@@
 This claim is difficult to evaluate insofar as phonotactic naturalness has continually eluded formalization and \citeauthor{Becker2011} neither propose nor refer to any theory thereof.
-\citeauthor{Becker2011} use the lack of phonetic precursors as a diagnostic of unnaturalness; however, there is extensive evidence that ``unnatural'' generalizations are in fact learned \citep[e.g.,][]{Anderson1981,Bach1972,Blevins2004,Buckley2000a,Hayes2009,Pierrehumbert2006c}.
+\citeauthor{Becker2011} use the lack of phonetic precursors as a diagnostic of unnaturalness; however, there is extensive formal and experimental evidence for the learning of ``unnatural'' generalizations \citep[e.g.,][]{Anderson1981,Bach1972,Blevins2003,Buckley2000a,Hayes2009,Pierrehumbert2006c,Seidl2005}.
 Furthermore, other functionally-oriented theorists have considered \textsc{Labial Attraction} sufficiently natural \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995b}.
 
-It is not even obvious that it is desirable to exclude \textsc{Labial Attraction} from the grammar.
+It is not even obvious that it is desirable to exclude \textsc{Labial Attraction} as a possible rule.
 \citet[394, fn.

### 2012-10-27 4.tex
Improvements to the appendices, primarily.

@@
 
-Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}).
-Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the \emph{communis opinio} (including many of the authors cited above) holds that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
+Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints both on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}).
+
+In the literature there are both isolated claims that 
+There are both isolated claims that
+
+Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}),
+
+\citet{Kisseberth1970b}
+
+however, the \emph{communis opinio} still is that that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
+
+%FIXME
+% \citet[255]{Anderson1974}, discussing this fact, writes that \citeauthor{Haugen} ``gives an argument that shows the necessity of including syllables as units in underlying representations''.
+
+%FIXME
+(for dissenting views, see \citealt{Blevins2003} and \citealt{Steriade1999})
+
+Linguists have long recognized underlying representations as privileged domains for stating phonotactic generalizations (e.g., \citealt{Bloomfield1930,Jakobson1932,Pike1947b}).
+
+More recently, a similar role for the syllable has been recognized 
+(e.g., \citet{Haugen1956}, \citet{Hooper1973}, \citealt{Kahn1976})
+
+\begin{quote}
+It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
+\citep[283]{Anderson1974}
+\end{quote}
 
@@
 A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database.
-The resulting list of clusters is reproduced in Appendix \ref{appendixB} and in electronic form at the author's website.
+The resulting list of clusters is reproduced in Appendix \ref{appendixC} and in electronic form at the author's website.
 
@@
 \midrule
-conforming & 80       & 370        & 18          & \multirow{2}{*}{1.1\e{-11}} \\
-violating  &  6       & 264        & 2           \\
+conforming & 80       & 370        & 18\%          & \multirow{2}{*}{1.1\e{-11}} \\
+violating  &  6       & 264        & 2\%           \\
 \bottomrule
@@
 \begin{table} \centering
-\begin{tabular}{l | r r r r}
+\begin{tabular}{l | r r r r | r}
 \toprule
@@
 \midrule
-Null baseline        & 0.813    & n.a.      & n.a.   & n.a.  \\
-Derived constraints  & 0.857    & 0.401     & 0.708  & 0.512 \\
-Expected frequency   & 0.839    & 0.210     & 0.750  & 0.328 \\
-\citet{Hayes2008a}   & 0.877    & 0.777     & 0.642  & 0.703 \\
+Null baseline        &     .808 &      .808 &  1.000 &  .894 \\
+\citet{Hayes2008a}   &     .837 &      .948 &   .852 &  .898 \\ %& .002       \\
+Expected frequency   &     .835 &      .845 &   .983 &  .909 \\ %& 1.1\e{-05} \\
+Derived constraints  &     .840 &      .922 &   .885 &  .903 \\
+DC and EF            &     .864 &      .871 &   .989 &  .926 \\ %& 1.9\e{-13} \\
 \bottomrule
 \end{tabular}
-\caption{The \citeauthor{Hayes2008a} model makes the most accurate predictions about possible and impossible syllable contact clusters, though it is only slightly better than the derived constraint baseline.}
+\caption{FIXME}
 \label{cmresults}
@@
 
-18.8\% of the 840 possible clusters are attested.
-Consequently, a null baseline which predicts no clusters to be attested trivially achieves 81.2\% accuracy.
-The other metrics are undefined for this baseline.
+19\% of the 880 possible clusters are attested.
+Consequently, a null baseline which predicts that no clusters are possible will be 81\% accurate, trivially.
+Because this baseline FIXME RECALL FIXME PRECISION
 
@@
 The three derived constraints define a structured baseline, in which attestation is the predicted outcome for all clusters which not violate the \emph{SPE}-derived generalizations, and non-attestation otherwise.
-This produces a significant improvement to the null baseline (accuracy = 0.857; sign test $p = 2.16$\e{-05}).
-The high recall (0.708) but low precision (0.401) of this model indicates that there are more false positives (i.e., unattested clusters which do not violate any constraint) than false negatives (i.e., exceptions to the three rules).
-It is not clear that low precision is a flaw, since some of these false positives might be accidental gaps---that is, clusters that are well-formed, but unattested.
+This results in a significant improvement to the null baseline (sign test $p = 0.006$).
+This model has poor recall, however, because many well-formed clusters are unattested.
+This is not a flaw, but rather the signature of accidental gaps.
 

### 2012-10-27 pre.tex
Improvements to the appendices, primarily.

@@
 One analysis, going back to the ancient grammarian Pāṇini \citep{Sag1974} and to \citet[\S141f.]{Whitney1889}, posits an underlying representation /bud\asp{}-/ with aspiration shifting leftward in certain contexts: modern proponents of this analysis include \citet{Borowsky1983}, \citet[][59f.]{Hoenigswald1965}, \citet{Kaye1985}, \citet{Sag1974,Sag1976} \citet{Schindler1976}, \citet{Stemberger1980}, and \citet[][\S141f.]{Whitney1889}.
-An alternative analysis posits an underlying representation /b\asp{}ud\asp{}-/ and an apsirate dissimilation process (\citealt{Anderson1970}, \citealt{Hoard1975}, \citealt[\S3.2]{Kiparsky1965}, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt[109f.]{Zwicky1965}).
+An alternative analysis posits an underlying representation /b\asp{}ud\asp{}-/ and an aspirate dissimilation process (\citealt{Anderson1970}, \citeyear[46f.]{Anderson1974}, \citealt{Hoard1975}, \citealt[\S3.2]{Kiparsky1965}, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt[109f.]{Zwicky1965}).
 It is not my intention to adjudicate between these competing analyses, or to evaluate the claim that both accounts are deeply flawed \citep{Janda1989}, but simply to note that they make different predictions under the theory proposed here.

### 2012-10-29 1.tex
Writing on 1-4, especially edits on 3
Writing on 2-4 and appendices

@@
 %\emph{τάξις}
+%Whether phonological computations or representations themselves are graded \citep[e.g.,][]{Lakoff1973} is besides the point, as metalinguistic judgements are behaviors, not mental states; they can no more be compared than can ``fear'' and ``flight response''.
+
+\citet{Silverman2000}
+\citet{Kiparsky1995}
+\citet[352f.]{Hale2003a}
+%$\begin{bmatrix} +\textsc{Dorsal} \\ +\textsc{Continuant \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~/~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~\gap$
+\citeauthor{SPE} propose to derive gradience from the complexity (using a simple feature-counting metric) of the redundancy rules needed to map a nonce word to an attested word, and they show that this derives the increasing cline of wordlikeness running from ``consonant soup'' \emph{bnzk} to ill-formed \emph{bnick}, possible \emph{blick}, and finally lexical \emph{brick}.
+
+Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature.
+At the risk of explaining what might be no more than an an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms.
+\citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
+An example from German is shown in (\ref{fd}) below.
+
+\begin{example}[German final devoicing] 
+\label{fd}
+\begin{tabular}{l l l l}
+   & nom.sg. & nom.pl.    \\
+a. & [piːp]    & [piːpə]  & `cheep(s)'      \\
+   & [diːp]    & [diːbə]  & `thief/thieves' \\
+b. & [ɡʀaːt]   & [ɡʀaːtə] & `ridge(s)'      \\
+   & [ɡʀaːt]   & [ɡʀaːdə] & `degree(s)'     \\
+\end{tabular}
+\end{example}
+
+\noindent The plural appears to preserve a contrast in final obstruent voicing which is absent in the singular:\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984} or not \citep[e.g.,][]{Port1985}, it is irrelevant to the discussion at hand.} word-final voiced stops are never found in German, and \citet[95]{Shibatani1973} reports that ``it is easy to show that a native speaker of German rejects those forms \emph{on the ground that they end in voiced obstruents}'' (emphasis in original).
+However, given that root-final voicing is not predictable (e.g., /ɡʀaːt \alt{} ɡʀaːd/), the process of \textsc{Final Devoicing} cannot be any kind of lexical redundancy and is mysterious under the \emph{SPE} account.
+The alternative proposed by \citeauthor{Shibatani1973} and by \citet{Clayton1976}, is that to say that nonce words like [ɡʀaːd] are ill-formed not because of any underlying property, but since they fail to undergo an otherwise-exceptionless phonological process, \textsc{Final Devoicing}, or equivalently, the surface-true generalization it derives.
+
+\citet{Sommerstein1974} further notes that the computation of hypothetical underlying forms from nonce surface forms which is implied by the \emph{SPE} theory is non-trivial in the presence of phonological opacity \citep[see][528f.]{Anderson1988a}, and the rejection of the biuniqueness principle means that there is not always a unique solution, as is illustrated by the two underlying forms corresponding to [ɡʀaːt] in (\ref{fd}) above.
+
+\subsubsection{Autosegmental phonology and beyond}
+
+The arguments of \citeauthor{Shibatani1973} and others led theorists to focus their attention on properties of surface representations as determinants of wordlikeness.
+Though syllabification plays no role in \emph{SPE}, it is crucial to many earlier studies \citep[for a review, see][]{Goldsmith2011b}, and it received particular attention in the 1970s.
+\citet{Hooper1973} and \citet{Kahn1976} argue that the syllable is useful for defining wordlikeness generalizations.\footnote{\citet{Steriade1999} and \citet{Blevins2003}, however, argue that a number of phonotactic generalizations previously stated in syllabic terms can be reanalyzed without making reference to syllables.} \citeauthor{Hooper1973} argues, for instance, that [bn], impossible as an English onset, is unobjectionable as a syllable contact cluster in nonce words like \emph{stabnik} (or in names like \emph{Abner}), and that this demonstrates the superiority of syllable-based wordlikeness generalizations.
+This already signals further trouble for alternative accounts which focus on underlying forms.
+Syllabification may span morphs, is generally predictable, and is universally non-contrastive, and as a consequence, few posit in to be present in underlying representations \citep[though see, e.g.,][]{Vaux2003}.
+\citeauthor{Hooper1973} also points to loanword adaptations which produce native syllable structure \citep[e.g.,][]{Carlisle1991} as evidence that syllabification is part of the phonological computation.
+Further enrichments to the theory are provided by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as first envisioned by][]{Pike1947a}, and theories like prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
+
+While none of these authors discuss gradience, the syllable plays a role in the definition of a gradient measure, ``positional probability'', though to correlate closely with human judgements of wordlikeness.
+The positional probability of a nonce monosyllabic word is derived from the combined probabilities at which segments occur in onset, nucleus, and coda positions in the lexicon of a given language.

### 2012-10-29 2.tex
Writing on 2-4 and appendices

--- 2012-10-11 2.tex
@@
-\chapter{Gradient and categorical aspects of wordlikeness judgements} 
+\chapter{Categorical and gradient aspects of wordlikeness judgements} 
 \label{gradience}
 
-%\citet{Fischer-Jorgensen1952}
-%\citet{Saporta1958}
-%\citet{Saporta1957}
-
-\citet{Halle1962} and \citet{Chomsky1965} use two nonce words, \emph{blick} [blɪk] and \emph{bnick} [bnɪk], to illustrate that speakers internalize generalizations about possible and impossible words in their language.
-While \emph{blick} is unattested, but in some sense a ``possible'' word of English, \citeauthor{Chomsky1965} observe that speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word, despite the fact that it is well-formed in other languages (e.g., Moroccan Arabic \emph{bniqa} `closet').
-
-A naïve account of this contrast is based on the assumption that segments or timing units must either be parsed into prosodic structures like the syllable, or be subject to phonological repair.
-Early precedents for this assumption can be found in work by \citet[10f.]{Hooper1973} and \citet[57f.]{Kahn1976}, and it is fundamental to many later frameworks \citep[e.g.,][]{Ito1989a,Noske1992,OT}.
-
-In English, [bl] is a permissible onset, but unlike some languages (e.g., Moroccan Arabic \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet'), [bn] is not.
-\citet[][19f.]{Wolf2009} suggest, for instance, that an underlying /bnɪk/ would be realized as [nɪk]; it is also possible to imagine resolution by prothesis---[əbnɪk]---or anapytxis---[bənɪk], and indeed \citet{Davidson2006b} finds that English speakers use all three of these repairs when asked to mimic foreign pronunciations of obstruent-nasal onsets like /bn/.\footnote{However, \citet{Davidson2005,Davidson2006a} observes an acoustic and articulatory distinction between ``exscrescent'' echwa found in non-native cluster production and ``lexical'' schwa: the anaptyctic [z${^\textrm{ə}}$k] produced by English speakers attempting to mimic word-initial [zk] is more similar to \emph{scum} than to \emph{succumb}.
-This may undermine any link between non-native realizations and epenthesis found elsewhere in English (e.g., allomorphs of the regular past tense and noun plural).} However, many linguists now regard the above as overly simplistic, for, as \citeauthor{Shademan2006} argues, such accounts draw a categorical boundary between the possible and impossible which fails to match the granularity at which speakers may report ``wordlikeness'' intuitions.
-
-\begin{quote}
-A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
-But when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values.
-\citep[371]{Shademan2006}
-\end{quote}
-
-This chapter focuses on the hypothesis, implicit in \citeauthor{Shademan2006}'s statement, that there is an inherent gradient or continuum between possible and impossible words, against an null hypothesis that the observed patterns of judgements are little more than an artifact of the wordlikeness task itself.
-\S\ref{2background} places this debate in a historical perspective, and argues that the naïve but widely accepted form of this hypothesis cannot be falsified.
-In \S\ref{2evaluation}, a quantitative evaluation of a revised version of this hypothesis reveals a poor match between the gradience in speakers' judgements and contemporary computational models of wordlikeness.
-Taken together, the findings undermine the gradience hypothesis.
-
-\section{Background}
-\label{2background}
-
-\subsection{A brief history of the gradience hypothesis} 
-\label{history}
-
-\citeauthor{Chomsky1965} were not the first to consider the notion of possible and impossible words.
-Their primary contribution is that their mentalist perspective: they recognize that naïve speakers effortlessly acquire language-specific generalizations about possible and impossible words and can report them without any explicit training.
-
-\subsubsection{The Copenhagen school}
-
-Wordlikeness received some attention by structuralists of the Copenhagen school.
-In an early study, \citet{Fischer-Jorgensen1952} proposes that possible and impossible words represent endpoints on a continuum, and that no non-arbitrary line can be drawn between them.
-While it would anachronistic to assign this a mentalistic interpretation, this appears to be the earliest proposal that wordlikeness contrasts are in some sense gradient.
-
-However, not all early literature is concerned with gradience.
-\citet[31]{Vogt1954}, for instance, recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
-\citeauthor{Vogt1954} observes that allophony may account for the absence of certain phone sequences, but it does not provide a suitable explanation for the absence of initial [bn] in English, nor does it make correct predictions about the surface realization of an underlying initial /bn/.
-\citeauthor{Vogt1954} concludes that additional grammatical machinery will be needed to account for possible and impossible words.
-
-\subsubsection{Early generative grammar}
-
-Gradient grammaticality was a tenet of the earliest work on generative syntax.
-For instance, \citet[132]{LSLT} writes that ``there is little doubt that speakers can fairly consistently order new utterances, never previously heard, with respect to their degree of `belongingness' to the language''.
-\citet{ASPECTS} proposes that different syntactic violations give rise to different degrees of ungrammaticality.
-Unlike \citeauthor{Fischer-Jorgensen1952}, however, for \citeauthor{LSLT} this continuum does incorporate a hard-and-fast boundary between the grammatical and ungrammatical: there are no constrasts between grammatical utterances, merely between different degrees of ungrammaticality.
-For \citeauthor{LSLT}, all grammatical utterances are alike with respect to their grammaticality, but there are many degrees of ungrammaticality \citep[61]{Schutze1996}.
-
-\citet{SPR,Halle1962} and \citet{Stanley1967} concern themselves only with categorical wordlikeness contrasts, however.
-Both authors derive wordlikeness effects from morphophonemic redundancy rules acting on underlying representations.
-The discussion of \emph{bnick} by \citet[101]{Chomsky1965} is a good example: \citeauthor{Chomsky1965} observe that a consonant before a word-initial stop can only be a liquid in English.
-A corresponding redundancy rule which specifies a consonant in this position as [$+$\textsc{Liquid}] precludes the derivation of \emph{bnick}, since there is no way to realize the second consonant as a nasal.\footnote{This is not the only redundancy that could rule out \emph{bnick}.
-For instance, /s/ is the only word-initial consonant that may precede a nasal consonant in English.}
-
-\subsubsection{\emph{The sound pattern of English}} 
-\label{2spe}
-
-In \emph{The sound pattern of English} (Henceforth \emph{SPE}), \citet{SPE} argue that a complete solution to wordlikeness must countenance gradience.
-In addition to the contrast of \emph{blick} and \emph{bnick}, they introduce a third nonce word \emph{bnzk}, which they take to be even less well-formed than ``impossible'' \emph{bnick}.
-They write:
+Two nonce words, \emph{blick} [blɪk] and \emph{bnick} [bnɪk], underlie the claim that speakers can rapidly distinguish between possible and impossible words \citet{Halle1962}.
+In \emph{SPE}, \citet{SPE} use a third word, \emph{bnzk}, to argue that nonce words fall on a cline of well-formedness.
+Neither \emph{bnick} nor \emph{bnzk} are possible words of English, they are possible words in other languages: stop-nasal onsets are found in Moroccan Arabic (e.g., \emph{bniqa} `closet') and stop-nasal-fricative-stop words in Imdlawn Tashlhiyt Berber \citep{Dell1985}.
+However, there is some sense in which \emph{bnzk} is even less English-like than \emph{bnick}.
+Of this, \citeauthor{SPE} write:
 
@@
 
-%In some languages (e.g., Imdlawn Tashlhiyt Berber [tm.zħ] `she jested', [tzmt] `it (f.) is stifling'; \citealt{Dell1985}), \emph{bnzk} is syllabifiable, so its ill-formedness is once again a property of English, not a universal.
+\noindent
+The position that the well-formedness of nonce words is consistent the view of syntactic grammaticality taken in \emph{LSLT} and \emph{Aspects} \citep{LSLT,ASPECTS}, where it is claimed that different syntactic vioaltions result in different degrees of ungrammaticality.
 
-\citeauthor{SPE} propose to derive gradience from the complexity (using a simple feature-counting metric) of the redundancy rules needed to map a nonce word to an attested word, and they show that this derives the increasing cline of wordlikeness running from ``consonant soup'' \emph{bnzk} to ill-formed \emph{bnick}, possible \emph{blick}, and finally lexical \emph{brick}.
-
-Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature.
-At the risk of explaining what might be no more than an an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms.
-\citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
-An example from German is shown in (\ref{fd}) below.
-
-\begin{example}[German final devoicing] 
-\label{fd}
-\begin{tabular}{l l l l}
-   & nom.sg. & nom.pl.    \\
-a. & [piːp]    & [piːpə]  & `cheep(s)'      \\
-   & [diːp]    & [diːbə]  & `thief/thieves' \\
-b. & [ɡʀaːt]   & [ɡʀaːtə] & `ridge(s)'      \\
-   & [ɡʀaːt]   & [ɡʀaːdə] & `degree(s)'     \\
-\end{tabular}
-\end{example}
-
-\noindent The plural appears to preserve a contrast in final obstruent voicing which is absent in the singular:\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984} or not \citep[e.g.,][]{Port1985}, it is irrelevant to the discussion at hand.} word-final voiced stops are never found in German, and \citet[95]{Shibatani1973} reports that ``it is easy to show that a native speaker of German rejects those forms \emph{on the ground that they end in voiced obstruents}'' (emphasis in original).
-However, given that root-final voicing is not predictable (e.g., /ɡʀaːt \alt{} ɡʀaːd/), the process of \textsc{Final Devoicing} cannot be any kind of lexical redundancy and is mysterious under the \emph{SPE} account.
-The alternative proposed by \citeauthor{Shibatani1973} and by \citet{Clayton1976}, is that to say that nonce words like [ɡʀaːd] are ill-formed not because of any underlying property, but since they fail to undergo an otherwise-exceptionless phonological process, \textsc{Final Devoicing}, or equivalently, the surface-true generalization it derives.
-
-\citet{Sommerstein1974} further notes that the computation of hypothetical underlying forms from nonce surface forms which is implied by the \emph{SPE} theory is non-trivial in the presence of phonological opacity \citep[see][528f.]{Anderson1988a}, and the rejection of the biuniqueness principle means that there is not always a unique solution, as is illustrated by the two underlying forms corresponding to [ɡʀaːt] in (\ref{fd}) above.
-
-\subsubsection{Autosegmental phonology and beyond}
-
-The arguments of \citeauthor{Shibatani1973} and others led theorists to focus their attention on properties of surface representations as determinants of wordlikeness.
-Though syllabification plays no role in \emph{SPE}, it is crucial to many earlier studies \citep[for a review, see][]{Goldsmith2011b}, and it received particular attention in the 1970s.
-\citet{Hooper1973} and \citet{Kahn1976} argue that the syllable is useful for defining wordlikeness generalizations.\footnote{\citet{Steriade1999} and \citet{Blevins2003}, however, argue that a number of phonotactic generalizations previously stated in syllabic terms can be reanalyzed without making reference to syllables.} \citeauthor{Hooper1973} argues, for instance, that [bn], impossible as an English onset, is unobjectionable as a syllable contact cluster in nonce words like \emph{stabnik} (or in names like \emph{Abner}), and that this demonstrates the superiority of syllable-based wordlikeness generalizations.
-This already signals further trouble for alternative accounts which focus on underlying forms.
-Syllabification may span morphs, is generally predictable, and is universally non-contrastive, and as a consequence, few posit in to be present in underlying representations \citep[though see, e.g.,][]{Vaux2003}.
-\citeauthor{Hooper1973} also points to loanword adaptations which produce native syllable structure \citep[e.g.,][]{Carlisle1991} as evidence that syllabification is part of the phonological computation.
-Further enrichments to the theory are provided by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as first envisioned by][]{Pike1947a}, and theories like prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
-
-While none of these authors discuss gradience, the syllable plays a role in the definition of a gradient measure, ``positional probability'', though to correlate closely with human judgements of wordlikeness.
-The positional probability of a nonce monosyllabic word is derived from the combined probabilities at which segments occur in onset, nucleus, and coda positions in the lexicon of a given language.
-
-Using the head-term preference paradigm, \citet{Jusczyk1993b} and \citet{Friederici1993} find that typically-developing children as young as 9 months of age distinguish between nonce words which are and are not phonotactically valid in their target language.
-\citet{Jusczyk1994} report that 9-month-old children acquiring English also show preferences for nonce words with high positional probability over those with low positional probability.
-Faciliatory effects of positional probability (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements \citep{Vitevitch1999a,Luce2001,Lipinski2005,Vitevitch2005}, and lexical decision \citep{Pylkkanen2002a}.
-
-Most relevant to the question at hand, \citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated wtih the positional probailities of the constituent syllables.
-Unfortunately, none of these researchers make any effort to eliminate the possibility that the low positional probability stimuli are ``impossible'' words of English.
-In fact,
-Chapter \ref{clusters} argues
-%the author has argued elsewhere \citep{Gorman2012c}
-that many of the stimuli used by \citeauthor{Frisch2000} and \citeauthor{Vitevitch1997} contain illicit word-medial consonant clusters.
-While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and wordlikeness judgements, but \citeauthor{Vitevitch1997} ultimately conclude this cannot explain all the variation in wordlikeness.
-
-The aforementioned studies all conclude that the gradient measure of positional probability correlates with behavioral results.
-As the flaws of the \citet{Vitevitch1997} study demonstrate, the aforementioned studies do little to tease apart the gradient and categorical aspects of phonotactics.
-More generally, they do little to distinguish between positional probability and closely correlated measures like bigram probability (see \S\ref{bigram} below) or neighborhood density, since these studies carefully select stimuli which either have high or low values for all of positional probability, bigram probablility, and neighborhood density.
-This is particularly troublesome given that no justification has ever been given for the positional probability measure in the first place; it appears to have been created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production \citep[e.g.,][]{Luce1998,Luce2000} and perception \citep{Marslen-Wilson1984,Marslen-Wilson1987,McClelland1986,Norris1994,Norris2000}.
-
-\subsection{A naïve gradience hypothesis}
-
-Recent critiques of this syllabification-based model focus on the existence of intermediate wordlikeness ratings \citep[see also][]{Coleman1997,Anttila2008a}.
+Most recent discussions of the ``problem of admissibility'' focus on this cline of wellformedness as it is evidenced in wordlikeness tasks.
 
 \begin{quote}
+When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing.
+In the usual case, novel items fall along a gradient cline of acceptability.
+\citep[][9]{Albright2009a}
+
 In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
@@
 
-When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing.
-In the usual case, novel items fall along a gradient cline of acceptability.
-\citep[][9]{Albright2009a}
+\ldots{}when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
 \end{quote}
 
-As discussed in \S\ref{2spe} above, this is also the view of \citeauthor{SPE} in \emph{SPE}.
-The observation itself cannot be questioned, but \citeauthor{SPE}, \citeauthor{Hayes2008a}, and \citeauthor{Albright2009a} do not explicitly state why this data is relevant to the construction of models of wordlikeness.
-Presumably, these authors believe that these patterns of judgements demonstrate that wordlikeness, as an internal state, is gradient simply because subjects make use of intermediate degrees of wordlikeness in judgement tasks.
-This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, an influential view in the cognitive sciences in general (see \citealt{Fodor1981a} for a critique).
+At first blush, this would seem to argue against a naïve account of wellformedness judgements in which ill-formedness results when prosodic parsing fails (e.g., \citealt{Ito1989a}, \citealt{Noske1992}, \citealt{OT}).
+Consider the possibility that a nonce word is ill-formed if it cannot be syllabified without modification.
+There is a long history for the proposal that impossible words are those whose surface form cannot be parsed into prosodic structures like syllables without further phonological modification (e.g., \citealt[10f.]{Hooper1973}, \citealt[57f.]{Kahn1976}).
+There is no doubt that prosodic context is involved in determining where segments may occur.
+For instance, [ŋ] is a possible coda in English, but not a possible onset.
+Similarly, the [bn] of \emph{bnick} is judged to be admissible (though unattested) in medial position in English (e.g., \emph{sta}[b.n]\emph{ick}; \citealt[97]{Hooper1973}) but not as an onset.
+English is unable to sylalbify initial onset /bn/ without further modification, so one might expect underlying /bnɪk/ to be subject to deletion \citep[19f.]{Wolf2009}, prothesis, or anaptyxis.
+Indeed, \citet{Davidson2006b} finds that English speakers use all three of these repairs when asked to mimic foreign pronunciations of obstruent-nasal onsets impossible in English.\footnote{However, \citet{Davidson2005,Davidson2006a} finds acoustic and articulatory distinctions between the ``exscrescent'' schwa used to resolve non-native onset clusters, and the ``lexical'' schwa found in real words.
+For example, the anaptyctic [z${^\textrm{ə}}$k] produced by English speakers attempting to mimic word-initial [zk] more closely resembles the onset of \emph{scum} than the start of \emph{succumb}.
+This may undermine any link between non-native realizations and epenthesis of schwa in native words (e.g., in the syllabic allomorphs of the regular past tense and noun plural).}
 
-\begin{unlabeledexample}
-\textsc{Naïve Gradience Hypothesis}: If graded judgements of a concept or mental state use intermediate ratings, that concept or mental state is gradient
-\end{unlabeledexample}
-
-This proposition has the form of an indicative conditional.
-While the antecedent of this condition, the presence or absence of intermediate ratings in a graded judgement task, is readily observable, it has no bearing on the truth or falsity of the conditional proposition.
-However, whereas a false consequent (e.g., an all-or-nothing concept) would potentially falsify the proposition, the consequent, a mental state, cannot be directly observed.
-If this is the case, then both the proposition and the consequent (the proposition that wordlikeness is itself gradient) fail the test of falsification, placing it beyond the scope of scientific inquiry.
-
-\citet{Armstrong1983} cut the Gordian knot by asserting the possibility of observing a false consequent; they find the existence of all-or-nothing concepts like ``odd number'' or (more controversially) ``female'' to be apparent.
+However, prosodic well-formedness distinctions do not provide a ready explanation for the ``numerous other distinctions'' posited in \emph{SPE}.
+This is thought to rule out simple prosodic account of wordlikeness judgements.
 
 \begin{quote}
-Are there definitional concepts?
+A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements. \citep[371]{Shademan2006}
+\end{quote}
+
+Implicit in this claim is the assumption that the graded judgements obtained from wordlikeness tasks which permit gradient responses require a gradient model of well-formedness.
+However, it is argued here this is not pertinent to the question of whether wordlikeness is gradient or categorical, because graded judgements are a property of all graded tasks, even when applied to judgements of discrete categories.
+
+\citet[382]{Hayes2008a} ``consider the ability to model gradient intuitions to be an important criterion for evaluation phonotactic models''.
+However, they fail to ask consider whether categorical models provide a reasonable fit to gradient judgements.
+Rather than directly evaluating the categorical model of English phonotactics proposed by \citet{Clements1983}, \citeauthor{Hayes2008a} convert these constraints, many of which are exceptionless, into probabilitistic constraints.
+This suggests that \citeauthor{Hayes2008a} equate ``the ability to model gradient intuitions'' with the ability to make continuous predictions about well-formedness.
+But if gradient judgements are nothing more than a task effect, a model need not make gradient predictions to fit gradient judgements.
+
+Despite many claims that categorical models are empirically insufficient, there is not a single quantitative investigation of categorical wordlikeness models, and no study which compares categorical and gradient wordlikeness models on an equal footing.
+This chapter represents a first attempt to fill that lacuna.
+
+OUTLINE HERE
+
+\section{What wordlikeness might not be}
+
+\subsection{Gradience in wordlikeness judgements}
+
+\subsection{Gradience as a task effect}
+
+\section{Evaluation} 
+\label{2evaluation}
+
+\subsection{Method}
+
+\subsubsection{Materials and subjects}
+
+\subsubsection{Computational models}
+
+\subsection{Results}
+
+\subsection{Discussion}
+
+\section{Conclusions}
+
+%\citet{Armstrong1983} cut the Gordian knot by asserting the possibility of observing a false consequent; they find the existence of all-or-nothing concepts like ``odd number'' or (more controversially) ``female'' to be apparent.
+%
+%\begin{quote}
+%Are there definitional concepts?
 Of course.
@@
 \citep[274]{Armstrong1983}
-\end{quote}
-
-\citeauthor{Armstrong1983} note that if subjects produce intermediate ratings for these categories, then the naïve proposition is false, and this is what \citeauthor{Armstrong1983} find in their experiments.
+%\end{quote}
+%
+%This proposition has the form of an indicative conditional.
+While the antecedent of this condition, the presence or absence of intermediate ratings in a graded judgement task, is readily observable, it has no bearing on the truth or falsity of the conditional proposition.
+However, whereas a false consequent (e.g., an all-or-nothing concept) would potentially falsify the proposition, the consequent, a mental state, cannot be directly observed.
+If this is the case, then both the proposition and the consequent (the proposition that wordlikeness is itself gradient) fail the test of falsification, placing it beyond the scope of scientific inquiry.
+%
+%\subsection{Gradience as a liability}
+%
+%One might reasonably ask whether 
+%representations 
+%
+%
+%\citeauthor{Armstrong1983} note that if subjects produce intermediate ratings for these categories, then the naïve proposition is false, and this is what \citeauthor{Armstrong1983} find in their experiments.
 They ask subjects to rate, on a seven-point scale, the extent to which, e.g., certain odd counting numbers represent the concept ``odd number'', and find that speakers do make use of intermediate values, as shown in Figure \ref{agg}.
 In summary, the proposition that intermediate ratings indicate mental gradience is either untestable, or it is false, and is thus rejected.
-
-\begin{figure} 
-\centering
-\includegraphics{agg.pdf}
-\caption{Subjects asked to rate the degree to which certain even and odd numbers represent the categories ``even'' and ``odd'', respectively, freely use intermediate ratings.}
-\label{agg} 
-\end{figure}
-
-\subsection{A falsifiable gradience hypothesis}
-
-\citeauthor{Armstrong1983} do not view their results as evidence that subjects have an internal model of odd numbers which is at odds with the formal, all-or-nothing definition, and in fact, several additional experiments show that subjects deploy the extension of the formal definition in other tasks.
+%
+%\begin{figure} 
+%\centering
+%\includegraphics{agg.pdf}
+%\caption{Subjects asked to rate the degree to which certain even and odd numbers represent the categories ``even'' and ``odd'', respectively, freely use intermediate ratings.}
+%\label{agg} 
+%\end{figure}
+%
+%\citeauthor{Armstrong1983} do not view their results as evidence that subjects have an internal model of odd numbers which is at odds with the formal, all-or-nothing definition, and in fact, several additional experiments show that subjects deploy the extension of the formal definition in other tasks.
 Reviewing this study, \citet[215]{Schutze2011} writes that the experiments show that judgements are ``sensitive to factors other than our underlying competence''.
@@
 \citeauthor{Schutze2011} further suggests that gradient responses are the product of the task itself:
-
-\begin{quote}
-Putting it another way, when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different.
+%
+%\begin{quote}
+%Putting it another way, when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different.
 \citep[215]{Schutze2011}
-\end{quote}
-
-This introduces a troubling possibility, that subjects ``oblige'' experimenters by introducing random noise to their categorical judgements when presented with a Likert task.
+%\end{quote}
+%
+%This introduces a troubling possibility, that subjects ``oblige'' experimenters by introducing random noise to their categorical judgements when presented with a Likert task.
 Were this the case, the observed intermediate judgements are unlikely to yield information about speakers' knowledge of possible and impossible words.
 The alternative considered here is that there exists some model in which the gradience in ratings is systematic.
-
-\begin{unlabeledexample}
-\textsc{Gradience Hypothesis}: If graded judgements of a concept or mental state use intermediate ratings, there is some model which can predict this behavior
-\end{unlabeledexample}
-
-\noindent There are now many computational models of wordlikeness which proport to do just this.
-An unfortunate defect of prior evaluations of these models, however, is that they adopt the naïve gradience hypothesis.
-\citet[382]{Hayes2008a} state that consider the ability to model gradient intuitions to be so important that they do not include any categorical models in their evaluation, as do many other authors.
-As a result, the literature contains no serious attempt to evaluate the gradience hypothesis.
-
-\section{Evaluation} 
-\label{2evaluation}
-
-The remainder of this chapter is devoted to testing the revised hypothesis.
-This is accomplished by an evaluation of computational models of wordlikeness in which categorical and gradient models are placed on an equal footing.
-
-\subsection{Data sources}
-
-The evaluation makes exclusive use of previously published English wordlikeness data.
+%
+%The evaluation makes exclusive use of previously published English wordlikeness data.
 For inclusion in the evaluation, a study must conform to all three of the following conditions.
@@
 All data are reproduced in Appendix \ref{appendixA}.
-
-\subsubsection{\citealt{Greenberg1964}}
-
-\citet{Greenberg1964} investigated wordlikeness using the technique of free magnitude estimation, a mechanism which has become increasingly popular among syntacticians \citep[e.g.,][]{Bard1996}.
-At the beginning of the experiment, the subject heard a recording of the word \emph{stick}.
-In subsequent trials, the subjects heard a nonce word and were asked to report ``how far would you say that is from English?'', with \emph{stick} at ``1''; subjects are told that a word that is ``twice as far from English'' as \emph{stick} should be scored ``2''.
-The data used here are from \citeauthor{Greenberg1964}'s Experiment B, in which 17 undergraduates were presented 17 stimuli in all.
-In addition to \emph{stick}, the stimuli include three other English words; these four items were excluded from further analyses, leaving 13 stimuli.
-As is standard practice in psychophysics \citep[e.g.,][]{Butler1987}, magnitudes were log-transformed prior to analysis.
-
-\subsubsection{\citealt{Scholes1966}}
-
-\citet{Scholes1966} conducted a number of English wordlikeness judgement with middle school children.
+%
+%\subsubsection{\citealt{Scholes1966}}
+%
+%\citet{Scholes1966} conducted a number of English wordlikeness judgement with middle school children.
 The data used here come from his ``experiment 5'', in which 63 monosyllabic items were presented to 33 seventh-grade students.
@@
 These two words were excluded and the remaining 61 stimuli were submitted to analysis.
-
-\subsubsection{\citealt{Albright2003b}}
-
-\citet{Albright2003b} gathered wordlikeness judgements to serve as norms for a \emph{wug}-test.
+%
+%\subsubsection{\citealt{Albright2003b}}
+%
+%\citet{Albright2003b} gathered wordlikeness judgements to serve as norms for a \emph{wug}-test.
 87 items were presented to 24 undergraduate subjects, who rated each word on a seven-point Likert scale.
-The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``complete normal, would make a fine English word''.
-
-\begin{table}
-\label{counts}
-\centering
-\begin{tabular}{l r r r}
-\toprule
-        & subjects & items & observations \\
-\midrule
-Scholes & 33       & 66    & 2,079        \\
-A\&H    & 24       & 88    & 2,112        \\
-\bottomrule
-\end{tabular}
-\caption{Idk.}
-\end{table}
-
-\subsection{Model comparison}
-
-The four models used here consist of a binary baseline and three computationally implemented gradient models.
+The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``completely normal, would make a fine English word''.
+%
+%\begin{table}
+%\centering
+%\begin{tabular}{l r r r}
+%\toprule
+%                        & subjects & items & trials \\
+%\midrule
+%\citealt{Albright2003a} & 24       & 87    & 2,088  \\
+%\citealt{Albright2007}  & 68       & 40    & 2,720  \\
+%\citealt{HayesInPress}  & 29       & 80    & 2,320  \\
+%\citealt{Scholes1966}   & 33       & 63    & 2,178  \\
+%\midrule
+%\textsc{Total}          & 154      & 270   & 9,306  \\
+%\bottomrule
+%\end{tabular}
+%\caption{Subject and item counts}
+%\label{counts}
+%\end{table}
+%
+%\subsection{Model comparison}
+%
+%The four models used here consist of a binary baseline and three computationally implemented gradient models.
 These three models are chosen because prior studies have shown they are correlated with wordlikeness judgements.
@@
 12]{Albright2009a}.
-
-The Spearman $\rho$ is most widely known rank correlation statistic, but it is difficult to give a natural interpretation to this quantity.
+%
+%The Spearman $\rho$ is most widely known rank correlation statistic, but it is difficult to give a natural interpretation to this quantity.
 On the other hand, the Kendall $\tau_b$ and Goodman-Kruskal $\gamma$ can be interpreted as fractions of the number of \emph{concordant} and \emph{discordant} pairs \citep{Noether1981}.
@@
 Correlations of the four models are given in Table \ref{cor}.
-
-\subsubsection{Binary baseline}
-
-The binary baseline used here is a crude implementation of the null hypothesis that there are no gradient effects in wordlikeness judgements.
+%
+%\subsubsection{Binary baseline}
+%
+%The binary baseline used here is a crude implementation of the null hypothesis that there are no gradient effects in wordlikeness judgements.
 To create such a baseline, it is necessary to distinguish between nonce words which contain a gross phonotactic violation and those which do not.
@@
 Speakers are particularly adept as separating onsets from rimes \citep{Treiman1986,Treiman1995,Fowler1993}, and a large portion of phonotactic generalizations and violations can be localized to either onset or rime \citep[e.g.,][]{Borowsky1989,Fudge1969,Treiman1995,Kessler1997,Treiman2000}.
-
-The baseline considers a nonce syllable to be well-formed if it consists of both an attested onset and an attested rime; the free combination of these two components is the only mechanism by which this model can generalize beyond attested words.
+%
+%The baseline considers a nonce syllable to be well-formed if it consists of both an attested onset and an attested rime; the free combination of these two components is the only mechanism by which this model can generalize beyond attested words.
 It is surely the case that this is an ultimately insufficient model: \citet{Albright2009a} observes, for instance, that [ɛsp] is a well-formed rime, even though it is found in no word of English.
@@
 For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.\footnote{
-Chapter \ref{clusters} 
-%\citet{Gorman2012c}
-reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
-
-The rating densities from the three studies, linearly transformed to fit within the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{dsn}.
+%Chapter \ref{clusters} 
+%%\citet{Gorman2012c}
+%reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
+%
+%The rating densities from the three studies, linearly transformed to fit within the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{dsn}.
 In all three studies, it is possible to discern a relatively sharp separation between valid and invalid clusters, but also the presence of intermediate values.
-
-\begin{figure} 
-\centering
-\includegraphics{density.pdf}
-\caption{Average ratings of individual nonce words, linearly transformed to fit the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
-\label{dsn}
-\end{figure}
-
-\begin{table} 
-\centering
-\begin{tabular}{l l l l l}
-\toprule
-Spearman $\rho$          & baseline         & maxent  & bigram           & density          \\
-\midrule
-\citealt{Greenberg1964}  & {0.845}          & {0.765} & {\textbf{0.863}} & {0.648}          \\
-\citealt{Scholes1966}    & {0.791}          & {0.762} & {\textbf{0.827}} & {\textbf{0.827}} \\
-\citealt{Albright2003b}  & {0.725}          & {0.429} & {0.708}          & {\textbf{0.742}} \\
-\midrule
-Kendall $\tau_b$         & baseline         & maxent  & bigram           & density          \\
-\midrule
-\citealt{Greenberg1964}  & {\textbf{0.716}} & {0.585} & {0.692}          & {0.462}          \\
-\citealt{Scholes1966}    & {\textbf{0.664}} & {0.597} & {0.652}          & {0.565}          \\
-\citealt{Albright2003b}  & {\textbf{0.599}} & {0.343} & {0.506}          & {0.556}          \\
-\midrule
-Goodman-Kruskal $\gamma$ & baseline         & maxent  & bigram           & density          \\
-\midrule
-\citealt{Greenberg1964}  & {\textbf{1.000}} & {0.684} & {0.692}          & {0.462}          \\
-\citealt{Scholes1966}    & {\textbf{0.995}} & {0.634} & {0.667}          & {0.614}          \\
-\citealt{Albright2003b}  & {\textbf{0.953}} & {0.656} & {0.509}          & {0.575}          \\
-\bottomrule
-\end{tabular}
-\caption{Rank correlations between wordlikeness ratings and phonotactic models surprisingly reveal that the binary baseline meeds or exceeds the coverage of three state-of-the-art phonotactic models.
+%
+%\begin{figure} 
+%\centering
+%\includegraphics{density.pdf}
+%\caption{Average ratings of individual nonce words, linearly transformed to fit the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
+%\label{dsn}
+%\end{figure}
+%
+%\begin{table} 
+%\centering
+%\begin{tabular}{l l l l l}
+%\toprule
+%Spearman $\rho$          & baseline         & maxent  & bigram           & density          \\
+%\midrule
+%\citealt{Greenberg1964}  & {0.845}          & {0.765} & {\textbf{0.863}} & {0.648}          \\
+%\citealt{Scholes1966}    & {0.791}          & {0.762} & {\textbf{0.827}} & {\textbf{0.827}} \\
+%\citealt{Albright2003b}  & {0.725}          & {0.429} & {0.708}          & {\textbf{0.742}} \\
+%\midrule
+%Kendall $\tau_b$         & baseline         & maxent  & bigram           & density          \\
+%\midrule
+%\citealt{Greenberg1964}  & {\textbf{0.716}} & {0.585} & {0.692}          & {0.462}          \\
+%\citealt{Scholes1966}    & {\textbf{0.664}} & {0.597} & {0.652}          & {0.565}          \\
+%\citealt{Albright2003b}  & {\textbf{0.599}} & {0.343} & {0.506}          & {0.556}          \\
+%\midrule
+%Goodman-Kruskal $\gamma$ & baseline         & maxent  & bigram           & density          \\
+%\midrule
+%\citealt{Greenberg1964}  & {\textbf{1.000}} & {0.684} & {0.692}          & {0.462}          \\
+%\citealt{Scholes1966}    & {\textbf{0.995}} & {0.634} & {0.667}          & {0.614}          \\
+%\citealt{Albright2003b}  & {\textbf{0.953}} & {0.656} & {0.509}          & {0.575}          \\
+%\bottomrule
+%\end{tabular}
+%\caption{Rank correlations between wordlikeness ratings and phonotactic models surprisingly reveal that the binary baseline meeds or exceeds the coverage of three state-of-the-art phonotactic models.
 All correlations are significant at $p = 0.05$.}
-\label{cor}
-\end{table}
-
-\subsubsection{Maximum entropy phonotactics}
-
-\citeauthor{Hayes2008a} (\citeyear{Hayes2008a}; henceforth H\&W) develop a sophisticated model of phonotactic grammaticality which estimates a probability distribution over phoneme sequences by weighing constraints according to the principle of maximum entropy, following \citet{Goldwater2003} and \citet{Jager2007}.
+%\label{cor}
+%\end{table}
+%
+%
+%\subsubsection{Maximum entropy phonotactics}
+%
+%\citeauthor{Hayes2008a} (\citeyear{Hayes2008a}; henceforth H\&W) develop a sophisticated model of phonotactic grammaticality which estimates a probability distribution over phoneme sequences by weighing constraints according to the principle of maximum entropy, following \citet{Goldwater2003} and \citet{Jager2007}.
 H\&W report that the predictions of their model are closely correlated with the \citet{Scholes1966} wordlikeness ratings.
@@
 However, as a consequence, the maximum entropy model performs particularly poorly on this data set, as many stimuli contain phonotactic violations in rime positions.
-
-\subsubsection{Segment bigram probability} 
-\label{bigram}
-
-The bigram probability of a sequence $ijk$ is the product of the probability of an sequence-initial $i$, the probability that $j$ follows $i$, and the probability that $k$ follows $j$, and the product of sequence-final \emph{k}.
-
-\begin{unlabeledexample}
-$\displaystyle \hat{p}(ijk) = p(i|\textrm{start}) \cdot p(j|i) \cdot p(k|j) \cdot p(\textrm{stop}|k)$
-\end{unlabeledexample}
-
-\noindent \citet{Albright2009a} employs bigram probability to model wordlikeness judgements.
+%
+%\subsubsection{Segment bigram probability} 
+%\label{bigram}
+%
+%The bigram probability of a sequence $ijk$ is the product of the probability of an sequence-initial $i$, the probability that $j$ follows $i$, and the probability that $k$ follows $j$, and the product of sequence-final \emph{k}.
+%
+%\begin{unlabeledexample}
+%$\displaystyle \hat{p}(ijk) = p(i|\textrm{start}) \cdot p(j|i) \cdot p(k|j) \cdot p(\textrm{stop}|k)$
+%\end{unlabeledexample}
+%
+%\noindent \citet{Albright2009a} employs bigram probability to model wordlikeness judgements.
 While the focus of \citeauthor{Albright2009a}'s study is on developing a model which uses bigrams over phonological features rather than segments themselves, \citeauthor{Albright2009a}'s evaluation, which includes both the \citeauthor{Scholes1966} and \citeauthor{Albright2003b} data sets, finds an advantage for segmental bigrams.
@@
 As reported by \citeauthor{Albright2009a}, segmental bigrams outperform featural bigrams (see Table \ref{albrightimproved}).
-
-\citeauthor{Albright2009a} estimates bigram probabilities using the method of maximum likelihood over types in the lexicon.
+%
+%\citeauthor{Albright2009a} estimates bigram probabilities using the method of maximum likelihood over types in the lexicon.
 The variant of segmental bigrams used here computes probabilities with a simple type of smoothing in which the count of all possible bigrams (including those never observed) are incremented by one.
@@
 This has the desirable effect that no nonce word is ever assigned a zero probability, and produces a small increase in the correlation between the \citeauthor{Albright2003b} wordlikeness norms compared with the maximum likelihood estimate (Table \ref{albrightimproved}).
-For all three data sets, this model also consistenly outperforms positional probability models defined by \citet{Vitevitch2004} and \citet{Vaden2009}.
-Given this and the fact that these model scores are highly correlatd with bigram probability \citep[][54]{Vitevitch1997}, they are not considered further.
+For all three data sets, this model also consistenly outperforms positional probability models defined by \citet{Vitevitch2004} and \citet{Vaden2009}; given that these model scores are highly correlatd with bigram probability \citep[][54]{Vitevitch1997}, they are not considered further.
 The bigram model consistently performs well in all the evaluations, and has the highest Spearman correlation with the \citeauthor{Greenberg1964} and \citeauthor{Scholes1966} data, and is frequently second place model to the binary baseline elsewhere.
-
-\begin{table} 
-\centering
-\begin{tabular}{l r r r r}
-\toprule
-                         & \multicolumn{1}{c}{featural bigrams} & \multicolumn{2}{c}{segmental bigrams}  \\
-                         & maximum likelihood                   & maximum likelihood & Laplace smoothing \\
-\midrule
-%Pearson $\rho
-Spearman $\rho$          & 0.638                                & 0.660              & \textbf{0.708}    \\
-Kendall $\tau_b$         & 0.457                                & 0.467              & \textbf{0.506}    \\
-Goodman-Kruskal $\gamma$ & 0.462                                & 0.473              & \textbf{0.509}    \\
-\bottomrule
-\end{tabular}
-\caption{Segmental bigrams outperform ``featural bigrams'', and Laplace smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
+%
+%\begin{table} 
+%\centering
+%\begin{tabular}{l r r r r}
+%\toprule
+%                         & \multicolumn{1}{c}{featural bigrams} & \multicolumn{2}{c}{segmental bigrams}  \\
+%                         & maximum likelihood                   & maximum likelihood & Laplace smoothing \\
+%\midrule
+%%Pearson $\rho
+%Spearman $\rho$          & 0.638                                & 0.660              & \textbf{0.708}    \\
+%Kendall $\tau_b$         & 0.457                                & 0.467              & \textbf{0.506}    \\
+%Goodman-Kruskal $\gamma$ & 0.462                                & 0.473              & \textbf{0.509}    \\
+%\bottomrule
+%\end{tabular}
+%\caption{Segmental bigrams outperform ``featural bigrams'', and Laplace smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
 All correlations are significant at $p = 0.05$.}
-\label{albrightimproved}
-\end{table}
-
-\subsubsection{Neighborhood density} 
-\label{density}
-
-There are now many methods for computing similarity between nonce words and existing words, long thought to be reflected in wordlikeness judgements. 
-For this study, a number of such methods were evaluated, including the Generalized Neighborhood Model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a number of variations on neighborhood density \citep{Coltheart1977} provided by \citet{Vaden2009}.
+%\label{albrightimproved}
+%\end{table}
+%
+%\subsubsection{Neighborhood density} 
+%\label{density}
+%
+%There are now many methods for computing similarity between nonce words and existing words, long thought to be reflected in wordlikeness judgements. 
+%For this study, a number of such methods were evaluated, including the Generalized Neighborhood Model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a number of variations on neighborhood density \citep{Coltheart1977} provided by \citet{Vaden2009}.
 The best performance was obtained with the simplest version of neighborhood density, which is defined as the number of real monomorphemic words which can be changed into the target nonce word by a single insertion, deletion, or substitution of a phone.\footnote{\citet{Greenberg1964} use a variant in which only substitutions are counted.} For instance, the neighbors of \emph{blick} include \emph{blink} (insertion), \emph{lick} (deletion) and \emph{black} (substitution).
 While many studies \citep[e.g.,][]{Bailey2001} report robust lexical similarity effects, it may be that the relatively weak performance of neighborhood density is the result of the presence of gross phonotactic violations.
-
-\subsection{Modeling residual gradience}
-
-The primary result is that no gradient model reliably exceeds the accuracy of the binary baseline.
+%
+%\subsection{Modeling residual gradience}
+%
+%The primary result is that no gradient model reliably exceeds the accuracy of the binary baseline.
 Despite this, there are relatively strong correlations between the binary baseline and these gradient models (see Table \ref{bcor}).
-However, the fact that the gradient models are generally outperformed by the binary baseline suggest that they do not reliably predict intermediate ratings.
+From the strong performance of the categorical model one can infer that the gradient models do not reliably predict intermediate ratings, or contrasts in ratings between words which are grouped together.
 To quantify this, the following method was used to estimate the residual contribution of the three gradient models once gross phonotactic violations are taken into account.
@@
 In most cases, including the gradient models on top of the binary baseline produces a worse correlation than is obtained with the binary baseline alone.
-
-For $\tau_b$ and $\gamma$, the interpretation of this result is clear.
+%
+%For $\tau_b$ and $\gamma$, the interpretation of this result is clear.
 The gradient models assign rankings to the sets of phonotactically valid and invalid clusters, respectively.
@@
 This shows quite starkly that these models fail to reliably predict intermediate ratings.
-
-\begin{table} 
-\centering
-\begin{tabular}{l r r r}
-\toprule
-Kendall $\tau_b$          & maxent         & bigram         & density  \\
-\midrule
-\citealt{Greenberg1964}   & 0.670          & \textbf{0.680} & 0.501 \\
-\citealt{Scholes1966}     & \textbf{0.685} & 0.632          & 0.639 \\
-\citealt{Albright2003b}   & 0.542          & 0.603          & \textbf{0.623} \\
-\bottomrule
-\end{tabular}
-\caption{The binary baseline is strongly correlated with the three gradient model scores; all correlations are significant at $p = 0.05$.}
-\label{bcor}
-\end{table}
-
-\begin{table} 
-\centering
-\begin{tabular}{l r r r}
-\toprule
-$\Delta$ Spearman $\rho$          & maxent            & bigram            & density  \\
-\midrule
-\citealt{Greenberg1964}  & $-0.060$          &  \textbf{0.038} & $-0.017$ \\
-\citealt{Scholes1966}    & $-0.029$          &  \textbf{0.047} & $-0.035$ \\
-\citealt{Albright2003b}  & $-0.008$          & $-0.015$          & \textbf{0.018} \\
-\midrule
-$\Delta$ Kendall $\tau_b$         & maxent            & bigram            & density  \\
-\midrule
-\citealt{Greenberg1964}  & $-0.114$          & \textbf{$-$0.007} & $-0.084$ \\
-\citealt{Scholes1966}    & $-0.067$          & \textbf{0.003}  & $-0.061$ \\
-\citealt{Albright2003b}  & \textbf{$-$0.038} & $-0.092$          & $-0.049$ \\
-\midrule
-$\Delta$ Goodman-Kruskal $\gamma$ & maxent            & bigram            & density  \\
-\midrule
-\citealt{Greenberg1964}  & $-0.268$          & \textbf{$-$0.260} & $-0.337$ \\
-\citealt{Scholes1966}    & $-0.361$          & \textbf{$-$0.313} & $-0.345$ \\
-\citealt{Albright2003b}  & \textbf{$-$0.137} & $-0.443$          & $-0.386$ \\
-\bottomrule
-\end{tabular}
-\caption{The change in rank correlation generated by augmenting the purely binary model with gradient predictions is small and in most cases it is negative.}
-\label{controlled}
-\end{table}
-
-\section{Conclusions}
-
-\subsection{The gradience hypothesis}
-
-This chapter has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
+%
+%\begin{table} 
+%\centering
+%\begin{tabular}{l r r r}
+%\toprule
+%Kendall $\tau_b$          & maxent         & bigram         & density  \\
+%\midrule
+%\citealt{Greenberg1964}   & 0.670          & \textbf{0.680} & 0.501 \\
+%\citealt{Scholes1966}     & \textbf{0.685} & 0.632          & 0.639 \\
+%\citealt{Albright2003b}   & 0.542          & 0.603          & \textbf{0.623} \\
+%\bottomrule
+%\end{tabular}
+%\caption{The binary baseline is strongly correlated with the three gradient model scores; all correlations are significant at $p = 0.05$.}
+%\label{bcor}
+%\end{table}
+%
+%\begin{table} 
+%\centering
+%\begin{tabular}{l r r r}
+%\toprule
+%$\Delta$ Spearman $\rho$          & maxent            & bigram            & density  \\
+%\midrule
+%\citealt{Greenberg1964}  & $-0.060$          &  \textbf{0.038} & $-0.017$ \\
+%\citealt{Scholes1966}    & $-0.029$          &  \textbf{0.047} & $-0.035$ \\
+%\citealt{Albright2003b}  & $-0.008$          & $-0.015$          & \textbf{0.018} \\
+%\midrule
+%$\Delta$ Kendall $\tau_b$         & maxent            & bigram            & density  \\
+%\midrule
+%\citealt{Greenberg1964}  & $-0.114$          & \textbf{$-$0.007} & $-0.084$ \\
+%\citealt{Scholes1966}    & $-0.067$          & \textbf{0.003}  & $-0.061$ \\
+%\citealt{Albright2003b}  & \textbf{$-$0.038} & $-0.092$          & $-0.049$ \\
+%\midrule
+%$\Delta$ Goodman-Kruskal $\gamma$ & maxent            & bigram            & density  \\
+%\midrule
+%\citealt{Greenberg1964}  & $-0.268$          & \textbf{$-$0.260} & $-0.337$ \\
+%\citealt{Scholes1966}    & $-0.361$          & \textbf{$-$0.313} & $-0.345$ \\
+%\citealt{Albright2003b}  & \textbf{$-$0.137} & $-0.443$          & $-0.386$ \\
+%\bottomrule
+%\end{tabular}
+%\caption{The change in rank correlation generated by augmenting the purely binary model with gradient predictions is small and in most cases it is negative.}
+%\label{controlled}
+%\end{table}
+%
+%\subsection{The gradience hypothesis}
+%
+%This chapter has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
 The surprising result is that virtually all of the apparent coverage of state-of-the-art gradient phonotactic models is simply a reflection of their ability to distinguish between the possible and the totally impossible; beyond this, they are unreliable.
@@
 It remains to be seen is whether any model can be put forth which accurately predicts these intermediate ratings.
-
-These result provide support for recent findings that speakers asked to perform gradient syntactic judgements produce responses closely corresponding to a widely recognized categorical grammatical/ungrammatical distinction \citep{Sprouse2007}.
-
-\subsection{Extensions to the binary baseline}
-
-The strong performance of the binary baseline should not be taken as evidence either that wordlikenesss judgements are binary, or that the binary baseline is a plausible model.
+%
+%These result provide support for recent findings that speakers asked to perform gradient syntactic judgements produce responses closely corresponding to a widely recognized categorical grammatical/ungrammatical distinction \citep{Sprouse2007}.
+%
+%\subsection{Extensions to the binary baseline}
+%
+%The strong performance of the binary baseline should not be taken as evidence either that wordlikenesss judgements are binary, or that the binary baseline is a plausible model.
 The most serious limitation of this evaluation is the primitive nature of the binary baseline.
@@
 On the other hand, a cognitively plausible version of this model might need to entertain phonotactic generalizations that are larger than these units, since syllable-sized phonotactic generalizations have been proposed for English \citep[e.g.,][]{Berkley1994a,Berkley1994b,Coetzee2008b,Fudge1969}.
-
-A possible further extension to the binary baseline would be the introduction of additional levels of wellformedness.
+%
+%A possible further extension to the binary baseline would be the introduction of additional levels of wellformedness.
 While the evaluation has shown that current gradient models do not reliably identify intermediate wellformedness, it does seem possible to identify at least three levels of grammaticality: for instance, one might encode the intuition that \emph{zhlick} [ʒlɪk] is more English-like than \emph{bnick}, though both have unattested onsets.
@@
 The bigram and maxent models make this prediction, as do many others \citep[e.g.,][]{Legendre1990,Levelt2000,Albright2008,Anttila2008a,Pater2009b} but despite this, there is still little data demonstrating cumulative effects in wordlikeness tasks.
-
-\subsection{Language acquisition}
-
-The weak empirical status of gradient phonotactic knowledge as reflected in adults has rammifications for language acquisition under the hypothesis that that infants acquiring language deploy the same representations as adults \citep[e.g.,][]{Macnamara1982,Pinker1984,Crain1991,Carey1995,deVilliers2001,Legate2007}.
+%
+%\subsection{Language acquisition}
+%
+%The weak empirical status of gradient phonotactic knowledge as reflected in adults has rammifications for language acquisition under the hypothesis that that infants acquiring language deploy the same representations as adults \citep[e.g.,][]{Macnamara1982,Pinker1984,Crain1991,Carey1995,deVilliers2001,Legate2007}.
 Gradient wordlikeness judgements in adults would provide support for claims that infants recognize statistical(inherently gradient) dependencies between segments \citep{Jusczyk1994} and use these to segment words \citep{Saffran1996}.
-An emerging consensus suggests, however, that infants attend to transitional probabilities primarily in the absence of grammatical cues \citep{Gambell2005,Hohne1994,Johnson2001,Jusczyk1999c,Lignos2012,Mattys2001a,Shukla2007,Lew-Williams2012}.
+An emerging consensus suggests, however, that infants attend to transitional probabilities primarily in the absence of grammatical cues \citep{Gambell2005,Hohne1994,Johnson2001,Jusczyk1999c,Lignos2012b,Mattys2001a,Shukla2007,Lew-Williams2012}.
 The vacuous nature of current evidence for gradient phonotactic knowledge in adults further weakens any hypothesis that would link statistical learning in infants to adults' behaviors.
-
-%For instance, infants as young as 4.5 months seem to be aware that English nasal codas agree in place with following obstruents \citep{Mattys2001b,Jusczyk2002,Davidson2004}. 
-%It might be the case that syllable co-occurrence statistics might be little more than a reflection of infants' learning of categorical ``lexical viability'' constraints \citep{Johnson2003} of the sort also seen in adult speech processing \citep[e.g.,][]{Norris1997}.
+%
+%\section{Conclusion}
+%
+%%For instance, infants as young as 4.5 months seem to be aware that English nasal codas agree in place with following obstruents \citep{Mattys2001b,Jusczyk2002,Davidson2004}. 
+%%It might be the case that syllable co-occurrence statistics might be little more than a reflection of infants' learning of categorical ``lexical viability'' constraints \citep{Johnson2003} of the sort also seen in adult speech processing \citep[e.g.,][]{Norris1997}.
+%
+%% other ratings, but por qué?:
+%% 
+%% Hayes2000
+%% 
+%% Albright2003a
+%% Albright2003b
+%% Prasada1993 
+%% 
+%% Bard1996
+%% 
+%% Koo2009
+%% 
+%% Treiman2000
+%% 
+%% Warker2006
+%% 
+%% Massaro1983
+%% 
+%% Rusaw2009
+%do not explicitly state why this data is relevant to the construction of models of wordlikeness.
+Presumably, these authors believe that these patterns of judgements demonstrate that wordlikeness, as an internal state, is gradient simply because subjects make use of intermediate degrees of wordlikeness in judgement tasks.
+This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, an influential view in the cognitive sciences in general (see \citealt{Fodor1981a} for a critique).
+%
+%\citeauthor{Chomsky1965} were not the first to consider the notion of possible and impossible words.
+Their primary contribution is that their mentalist perspective: they recognize that naïve speakers effortlessly acquire language-specific generalizations about possible and impossible words and can report them without any explicit training.
+%
+%However, not all early literature is concerned with gradience.
+\citet[31]{Vogt1954}, for instance, recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
+\citeauthor{Vogt1954} observes that allophony may account for the absence of certain phone sequences, but it does not provide a suitable explanation for the absence of initial [bn] in English, nor does it make correct predictions about the surface realization of an underlying initial /bn/.
+\citeauthor{Vogt1954} concludes that additional grammatical machinery will be needed to account for possible and impossible words.
+%
+%Most relevant to the question at hand, \citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated wtih the positional probailities of the constituent syllables.
+Unfortunately, none of these researchers make any effort to eliminate the possibility that the low positional probability stimuli are ``impossible'' words of English.
+In fact,
+%Chapter \ref{clusters} argues
+%%the author has argued elsewhere \citep{Gorman2012c}
+%that many of the stimuli used by \citeauthor{Frisch2000} and \citeauthor{Vitevitch1997} contain illicit word-medial consonant clusters.
+While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and wordlikeness judgements, but \citeauthor{Vitevitch1997} ultimately conclude this cannot explain all the variation in wordlikeness.
+%
+%Using the head-term preference paradigm, \citet{Jusczyk1993b} and \citet{Friederici1993} find that typically-developing children as young as 9 months of age distinguish between nonce words which are and are not phonotactically valid in their target language.
+\citet{Jusczyk1994} report that 9-month-old children acquiring English also show preferences for nonce words with high positional probability over those with low positional probability.
+Faciliatory effects of positional probability (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements \citep{Vitevitch1999a,Luce2001,Lipinski2005,Vitevitch2005}, and lexical decision \citep{Pylkkanen2002a}.
+%
+%The aforementioned studies all conclude that the gradient measure of positional probability correlates with behavioral results.
+As the flaws of the \citet{Vitevitch1997} study demonstrate, the aforementioned studies do little to tease apart the gradient and categorical aspects of phonotactics.
+More generally, they do little to distinguish between positional probability and closely correlated measures like bigram probability (see \S\ref{bigram} below) or neighborhood density, since these studies carefully select stimuli which either have high or low values for all of positional probability, bigram probablility, and neighborhood density.
+This is particularly troublesome given that no justification has ever been given for the positional probability measure in the first place; it appears to have been created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production \citep[e.g.,][]{Luce1998,Luce2000} and perception \citep{Marslen-Wilson1984,Marslen-Wilson1987,McClelland1986,Norris1994,Norris2000}.
+%\subsubsection{\citealt{Greenberg1964}}
+%
+%\citet{Greenberg1964} investigated wordlikeness using the technique of free magnitude estimation, a mechanism which has become increasingly popular among syntacticians \citep[e.g.,][]{Bard1996}.
+At the beginning of the experiment, the subject heard a recording of the word \emph{stick}.
+In subsequent trials, the subjects heard a nonce word and were asked to report ``how far would you say that is from English?'', with \emph{stick} at ``1''; subjects are told that a word that is ``twice as far from English'' as \emph{stick} should be scored ``2''.
+The data used here are from \citeauthor{Greenberg1964}'s Experiment B, in which 17 undergraduates were presented 17 stimuli in all.
+In addition to \emph{stick}, the stimuli include three other English words; these four items were excluded from further analyses, leaving 13 stimuli.
+As is standard practice in psychophysics \citep[e.g.,][]{Butler1987}, magnitudes were log-transformed prior to analysis.
+%

### 2012-10-29 4.tex
Writing on 2-4 and appendices

--- 2012-10-23 4.tex
@@
 \chapter{Structural and accidental gaps} 
-\label{clusters}
+\label{gaps}
 
-Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}).
-Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the \emph{communis opinio} (including many of the authors cited above) holds that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
+Early work in generative phonology notes that neutralizing, surface-true phonological alternations impose constraints both on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kisseberth1970b}, \citealp{Kenstowicz1977}: chap.~3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}).
 
-\begin{quote}
-It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
-\citep[283]{Anderson1974}
-\end{quote}
+Two domains are essential to defining
 
-English word-medial consonant clusters have been used, by \citet{Pierrehumbert1994} and others, to argue for the necessity of an independent phonotactics module and illustrate proposals for its architecture.
-This chapter argues, however, that the only structural constraints on the concatenation of codas and onsets to form word-medial clusters are derivative of well-documented phonological alternations in English.
+Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}),
 
-\section{Background}
+\section{English syllable contact clusters}
 
-English admits a variety of word-medial consonant clusters spanning syllable boundaries, known as \emph{syllable contact clusters} or \emph{interludes}.
-These clusters can be as short as two consonants (e.g., \emph{a}[n.t]\emph{ics}; singleton intervocalic consonants are ignored here) or as long as four (e.g., \emph{mi}[n.str]\emph{el}).
-\citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in any given language \citep[cf.][]{Haugen1956}.
+English admits a variety of word-medial consonant clusters spanning syllable boundaries; these are known as \emph{syllable contact clusters} or \emph{interludes}.
+\citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in some language, echoing a similar proposal by \citet{Haugen1956}.
 
@@
 
-This chapter is concerned with the nature of these ``provisos'' as reflected by which of the ``possible'' clusters, so defined, are attested.
-While many phonotactic generalizations can be thought of as inventory restrictions imposed by prosodic constituents, additional filters on the set of possible clusters are essentially combinatoric in nature and thus distinct, there is no prosodic constituent consisting of the coda of one consonant-final syllable and the onset of a following consonant-initial syllable \citep[though see][]{Steriade1999}.
+Many phonotactic generalization can be thought of as restrictions on the inventory of different prosodic constituents defining what is, e.g., a ``well-formed onset''.
+Since few linguists posit the interlude as a prosodic constituent \citep[though cf.][]{Steriade1999}, further constraints on syllable contact clusters, however, are combinatoric.
 
 \section{Evaluation}
+\label{4evaluation}
 
-\citeauthor{Pierrehumbert1994} argues that the inventory of clusters in English shows the effect of a small number of static phonotactic constraints.
-The remainder of this chapter provides an exhaustive quantitative evaluation of this hypothesis.
-The most important constraints on medial clusters are those derived from phonological rules which apply across syllable boundaries; there is no remaining role for static constraints.
-The sparse nature of the lexicon plays a secondary role, giving rise to apparent phonotactic gaps even in the absence of static phonotactic constraints.
+\subsection{Method}
 
-\subsection{Corpus construction}
+\subsubsection{Materials and procedure}
 
-A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database.
-The resulting list of clusters is reproduced in Appendix \ref{appendixB} and in electronic form at the author's website.
+Following \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, 
+a wordlist was produced using the English portion of the CELEX database \citep{CELEX}.
+Any words which were coded as borrowings were excluded, as were all words not coded as ``monomorphemic''.
+The use of this stringent criterion eliminates many of the exceptions noted by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies.
+For instance, many of the exceptions to \textsc{Obstruent Voice Assimilation} (see \S?
+below) noted by \citet[74]{Hammond1999a} are coded either as loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as complex words (e.g., \emph{jurisdiction}, \emph{madcap} \emph{tadpole}, \emph{scapegoat}, \emph{magpie}).
 
-\subsubsection{Simplex words}
+Unlike prior studies, ``neo-classical compounds'', words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from this wordlist.
+This is done in response to converging evidence that English speakers regard such words as complex.
+This assumption has long been made to simplify phonological \citep{SPE}, allomorphic \citep[11f.]{Aronoff1976}, and syntactic \citep{Harley2009} generalizations.
+Results from lexical decision experiments also suggest that neoclassical compounds are complex.
+\citet{Taft1986} and \citet{Taft1975,Taft1976} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure such as *\emph{refant}).
+Bound stems also show frequency effects which are independent of their whole word frequency \citep{Taft1979,Taft2006}.
+Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect suggestive of a shared morphological identity.
 
-An analysis of lexical constraints requires a corpus of lexical representations.
-So as to remain agnostic about particulars of morphological theory, the goal here is to identify words which bear no overt inflection, and which do not admit any plausible further decomposition, henceforth ``simplex words''.
-For the structuralists, a procedure for parsing words into morphs, just the thing for locating simplex words, held the status of ``philosopher's stone'', a sort of near-mythical objective of the endeavor, but actual attempts to develop such a procedure \citep[e.g.,][]{Harris1955,Nida1948} are now recognized as anything but theory-neutral.
-Yet, there is no alternative to adopting some computational procedure or human-coded database as the ``gold standard'' for a study of this type.
-\citet{Pierrehumbert1994} analyzes a list of entries in the Collins English dictionary by hand, using her intuitions to filter out ``complex'' words as well as those which are not ``reasonably familiar''.
-Unfortunately, the resulting list was neither published nor circulated, and the possibility of replicating \citeauthor{Pierrehumbert1994}'s' sensations of morphological complexity is remote.
-Following prior studies of syllable contact clusters by \citet[ chapter 3]{Hammond1999a} and \citet[ chapter 8]{Duanmu2009}, a list of simplex words was derived from the ``lemmas'' (that is, uninflected forms) in the English portion of CELEX \citep{CELEX}, a database of morphological, phonological, and syntactic annotations based on the COBUILD corpus, a resource that was also used to construct the Collins English dictionary employed by \citeauthor{Pierrehumbert1994}.
+This filtering results in a list of 6,619 simplex words.
+The CELEX transcriptions of these words were then syllabified and phonologized using a technique described in Appendix \ref{syllabification}.
+This produces 23 unique medial codas and 40 unique medial onsets.
+Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination, 174 (19\%) are attested.
+The full set of clusters and their frequencies are reproduced in Appendix \ref{clusters}.
 
-Two further filters were applied to this list.
-First, any lemma coded as an unassimilated loanword is excluded.
-Secondly, all lemmas that have a ``morphological status'' other than ``monomorpheme'' are excluded.
-This is crucial because CELEX makes no guarantee that lemmas are simplex; indeed, many lemmas are products of derivation (e.g., \emph{abusive}, from \emph{abuse}).
-These stringent criteria have a profound effect on the makeup of the data.
-For instance, nearly all exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) cited by \citet[74]{Hammond1999a} are excluded as unassimilated loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as morphologically complex (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}).
+\subsection{Results}
 
-``Neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}, have an ambiguous and long-debated morphological status.
-\citeauthor{Pierrehumbert1994} attempts to distinguish between words like \emph{complete}, \emph{extreme}, \emph{inspect}, \emph{obtuse}, which she assumes to be simplex, and \emph{excrete} and \emph{transparent}, which she considers complex.
-However, CELEX codes these words as complex and they are thus excluded.
-This is consistent with converging evidence that speakers decompose these Latinate forms.
-The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.
-In a similar vein, \citet[11--13]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.
+\subsubsection{Static constraints}
 
-\begin{example}[Bound stem-specific allomorphy]
-\begin{tabular}{l l l l l l}
-a. & {adhere}   & {adhesion}   \\
-   & {cohere}   & {cohesion}   \\
-b. & {conceive} & {conception} \\
-   & {perceive} & {perception} \\
-\end{tabular}
+In summary, there is no statistical support for any of the three static constraints proposed by \citeauthor{Pierrehumbert1994}. However, it is premature to conclude that no static constraint exists.
+
+%It is possible that an explicit model of ``constraint discovery'', could identify reliable static constraints,
+
+\subsubsection{Derived constraints}
+
+The allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and of the regular noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]) illustrate the well-known process of \textsc{Obstruent Voice Assimilation} (e.g., \emph{SPE}:178) in which the voicing specification of the first of two adjacent obstruents perserveres.
+
+\begin{example}[\textsc{Obstruent Voice Assimilation}]
+$\begin{bmatrix} -\alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\goesto{}~\begin{bmatrix} \alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\big /~\begin{bmatrix} \alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\gap{}$
 \end{example}
 
-There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement.
-\citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.
-\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
+Though it is not mentioned by \citet{Pierrehumbert1994} in her study of medial clusters, the vast majority of medial obstruent clusters are either uniformly voiced or voiceless, as in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}.
+``Disagreeing'' clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rare
 
-\begin{example}[Latinate verbs and small clauses] 
-\label{harley}
-\begin{tabular}{ll@{}ll@{}l}
-a. & * & {exhibit him the painting} & ~ & {show him the painting} \\
-b. & * & {imbibe himself stupid}    & ~ & {drink himself stupid}  \\
-c. & * & {exhibit it off}           & ~ & {show it off}           \\
-\end{tabular}
-\end{example}
+\subsubsection{Computational models}
 
-Decomposition is also supported by lexical decision results.
-Among nonce words, those that can be exhaustively decomposed into prefix and bound stem, like *\emph{de}-\emph{juvenate}, are processed slower that contain a licit morph but cannot be exhaustively decomposed, like *\emph{de-pertoire}, and these in turn are processed slower than those which admit no such decomposition \citep{Taft1975}.\footnote{The same contrasts are present in many other domains where decomposition into root and affix(es) is less controversial, for instance, inflected verbs in Italian \citep{Caramazza1988}.} \citet{Emmorey1989} reports facilitative auditory priming between pairs of words which share a bound stem, like \emph{permit}-\emph{submit} \citep[though see][]{Marslen-Wilson1994}.\footnote{It may be the case that full decomposition is only one of the two methods for the processing of complex words, the other being whole word looking, either running in serial \citep{Caramazza1988} or in parallel \citep{Baayen1997b}.
-Whatever evidence that is amassed in support of a whole word lookup procedure, however, does not undermine the considerable evidence for decomposition.}
+A soft-margin support vector machine \citep{Cortes1995} with a linear kerne
+l is used to convert from the per-cluster scores to attestation predictions
+, by finding a single cutoff which optimally separates attested and unattes
+ted clusters according to their model scores. This is in no way intended to
+ be a cognitively plausible model for learning phonotactics; it simply repr
+esents an upper bound on the performance of these models. The results for a
+ll four models are summarized in Table \ref{cmresults}.
 
-\subsubsection{Syllabification and phonologization}
-
-CELEX includes broad, syllabified Received Pronunciation transcriptions of individual wordforms.
-However, a casual inspection of the data reveals the unsystematic nature of these syllabifications.
-Any two words sharing a nucleus-medial cluster-nucleus sequence should be given the same parse, since syllabification is universally non-contrastive.
-Yet many putative syllabification contrasts can be found in CELEX; for instance, compare the parses given to the clusters in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][II.294]{AOE}.}
-
-A systematic syllabification requires an automated procedure for delimiting medial clusters and parsing them into coda and onset.
-Since this procedure is applied only to medial clusters, this study remains agnostic on several contentious issues concerning syllabification of English; for instance, there is no need to address the status of so-called ``ambisyllabic'' consonants, since this does not occur with medial clusters, or morphological effects on syllabification, since no complex words are included.
-The technique used here is a variation on the theme of onset maximization (e.g., \citealp[42f.]{Kahn1976}, \citealp{Kurylowicz1948}, \citealp[75]{Pulgram1970}, \citealp[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset.
-Initial onsets are used to define what is a ``possible'' medial onset \citep[though cf.][36]{Fischer-Jorgensen1952}.
-Medial clusters in words like \emph{neu}[.tɹ]\emph{on} or \emph{bi}[.stɹ]\emph{o} are found in word-initial position (e.g., [tɹ]\emph{ansit}, [stɹ]\emph{ike}), so onset maximization assigns the entire medial cluster to the onset, leaving the medial coda empty.
-In contrast, the cluster in \emph{mi}[n.stɹ]\emph{el} is not found word-initially; the maximal onset here is [stɹ], and the residual [n] is assigned to the coda.
-
-%\citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''
-%\citet[?]{Lombardi1999} says that English lacks devoicing in Level II.
-
-\paragraph{Stressed lax vowels} When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
-\citet[][55]{Harris1994} notes that onset maximization produces an incorrect result when, in addition, the medial cluster is a valid onset: in \emph{whisper}, \emph{vestige}, and \emph{musket}, onset maximization incorrectly assigns [sp, st, sk] to the medial onset, leaving the lax vowel unchecked.
-Consequently, the first consonant of a medial cluster is assigned to the coda of a preceding syllable before a stressed lax vowel \citep[cf.][48]{Pulgram1970}, a minimal modification of the onset maximization parse.
-
-\paragraph{Affricates} If the ffricates [tʃ, dʒ] are treated as stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990}, the previous principle will incorrectly assign the two segments of the affricate to separate syllables before stressed lax vowels (e.g., *\emph{ra}[t.ʃ]\emph{et} or \emph{a}[d.ʒ]\emph{ile}) unless some ad hoc constraint \citep[e.g.,][]{Wells1990} is in place.
-A more traditional analysis of affricates, in which they are simple segments distinguished from pure stops by stridency \citep[24]{Jakobson1961} or delayed release (\emph{SPE}:321f.) is assumed here, following \citeauthor{Pierrehumbert1994}.
-This assumption can be motivated by the tendency of affricates to pattern with individual segments in phonotactic generalizations.
-For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but bans true onset clusters \citep[9]{Launey2011}.
-%\citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.
-
-\paragraph{The velar nasal} According to some \citep[e.g.,][]{Sapir1925}, [ŋ] in English is a phoneme in its own right, but others (e.g., \emph{SPE}:85, \citealp{Borowsky1986}:65f.) hold that it is simply the allophone of /n/ before underlying /k, ɡ/ (with later deletion of /ɡ/ in some contexts).
-The total absence of onset [ŋ], where it cannot be followed by a dorsal consonant which conditions the velar allophone, is strong evidence that [ŋ] is a pure allophone of /n/, the analysis assumed here.
-
-\paragraph{[j] onglides} While some have argued that the front onglide in words such as \emph{val}[j]\emph{ue} is inserted by rule (\emph{SPE}:196, \citealp[][89]{Halle1985a}, \citealp[][217]{McMahon1990}), the very presence or absence of the glide is contrastive (e.g., \emph{coot} \alt{} \emph{cute}, \emph{booty} \alt{} \emph{beauty}) indicating that it is present in underlying representation (\citealp{Anderson1988b}, \citealp[278]{Borowsky1986}).
-The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{ju}[n.j]\emph{or}).
-There is considerable evidence for this assumption.
-When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u], suggesting that the glide is nuclear (\citealp{Davis1995}, \citealp[][61f.]{Harris1994}, \citealp[][232]{Hayes1980}).
-\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters, though they may be followed by [ju] in words like [mj]\emph{use} or [vj]\emph{iew}, also suggesting the nuclear affiliation of the glide.
-There is also external support for this analysis: the [ju] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005}, \emph{shm}-reduplication \citep{Nevins2003}, and speech errors (e.g., [kju]\emph{mor} [h]\emph{omponent}, intended [hju]\emph{mor component}; \citealp[130]{Shattuck-Hufnagel1986}).
-
-\paragraph{[w] onglides} The phonotactic properties of the back onglide [w] are the opposite of the front onglide.
-Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
-Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[161]{Davis1995}.
-Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[166]{Davis1995}.
-Taken together, these facts indicate that the back onglide is assigned to the onset.
-
-\paragraph{Syllabic \emph{r}} Word-medial syllabic \emph{r} has been lost in Received Pronunciation, the accent used for the CELEX transcriptions.
-Even in \emph{r}-ful dialects, though, there is reason to believe that syllabic \emph{r} is fact nuclear, and thus irrelevant to syllable contact clusters.
-Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}): compare American English \emph{fern}, \emph{fir}, \emph{fur} to \emph{pet}, \emph{pit}, \emph{putt}.
-Syllabic \emph{r} patterns with vowels and not other consonants for several phonological processes: it is the only consonant which does not block variable glottalization of following /t/ in British dialects \citep[258]{Harris1994}.
-
-\begin{example}[/t/-\textsc{Glottalization} in British English]
-\begin{tabular}{l l l l@{} l l l}
-a. & {des}[ɚt]    & \alt{} &   & {des}[ɚʔ]    \\
-   & {c}[ɚt]{ain} & \alt{} &   & {c}[ɚʔ]{ain} \\
-b. & {fi}[st]     & \alt{} & * & {fi}[sʔ]     \\
-   & {mi}[st]{er} & \alt{} & * & {mi}[sʔ]{er} \\
-\end{tabular}
-\end{example}
-
-\noindent Syllabic \emph{r} is the only consonant which does not trigger variable deletion of a following /t, d/ in American dialects \citep[8]{Guy1980}.
-
-\begin{example}[/t, d/-\textsc{Deletion} in American English]
-\begin{tabular}{l l l l@{} l l l}
-a. & {be}[lt] & \alt{} &   & {be}[l] \\
-   & {me}[nd] & \alt{} &   & {me}[n] \\
-b. & {sh}[ɚt] & \alt{} & * & {sh}[ɚ] \\
-   & {c}[ɚd]  & \alt{} & * & {c}[ɚ]  \\
-\end{tabular}
-\end{example}
-
-\subsubsection{Lexical statistics}
-
-CELEX contains 6,876 simplex words of English, comprising 21 unique medial codas and 40 unique medial onsets.
-Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested.
-
-\subsection{Evaluating derived constraints}
-
-\emph{SPE} describes three phonological alternations which target syllable contact clusters.
-The corresponding constraints on simplex syllable contact clusters are considered below.
-The first such derived constraint, obstruent voice assimilation, is presented as a tutorial introduction.
-The second set of constraints consists of three static phonotactic constraints proposed by \citet{Pierrehumbert1994}.
-
-\subsubsection{Obstruent voice assimilation} 
-\label{ova}
-
-The [t \alt{} d] allomorphs of the regular past (e.g.
-\emph{nap}[t] $\sim$ \emph{nab}[d]), and the [s \alt{} z] allomorphs of the regular noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), Saxon genitive, and third person singular active indicative verb agreement, exemplify \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178).\footnote{Underlying /-d, -z/ are argued for by \citet[282]{Hockett1958}, \citet[210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, \citet{Anderson1973a}, \citet[102]{Pinker1988}, and \citet[284f.]{Bakovic2005b}; alternatives are suggested by \citet[210f.]{LANGUAGE}, \citet[426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[135]{Borowsky1986}.}
-
-\begin{example}[Inflectional affix voice assimilation]
-\begin{tabular}{l l l l l l}
-a. & {nap} & {nap}[t] & & {nab} & {nab}[d] \\
-b. & {lap} & {lap}[s] & & {lab} & {lab}[z] \\
-\end{tabular}
-\end{example}
-
-\noindent The devoicing of the two voiced obstruent suffixes can be formalized as a 
-rule spreading the voicing specification of an obstruent rightward.
-
-\begin{example}[\textsc{Obstruent Voice Assimilation}]
-%\xymatrix@R=24pt@C=24pt{
-%\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{--}[dr] &                               \\
-%\txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
-%\txt{[$-$\textsc{Sonorant}]}                  & \txt{[$-$\textsc{Sonorant}]} \\
-%}
-$\begin{bmatrix} -\textsc{Son} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Voi} \end{bmatrix}~/~\gap~\begin{bmatrix} =\textsc{Voi} \\ -\textsc{Son} \end{bmatrix}$
-\end{example}
-
-It is quite apparent that root-internal hetero-voiced obstruent clusters like [s.b] or [z.t] are also rare compared to clusters which are uniformly voiced or uniformly voiceless, as predicted the assimilation rule.
-Yet, \citet{Pierrehumbert1994} does not mention voice assimilation in her study of English syllable contact clusters.
-\citet[][74f.]{Hammond1999a} cites words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain hetero-voiced obstruent clusters, as evidence that the process does not apply root-internally, though few of his examples are in fact simplex according to CELEX.\footnote{The \textsc{Revised Alternation Condition} (RAC) proposed by \citeauthor{Kiparsky1973a} (\citeyear{Kiparsky1973a}:163, \citeyear{Kiparsky1982a}:152) blocks the application of obligatory neutralization processes like \textsc{Obstruent Voice Assimilation} in root-internal (i.e., non-derived) environments.
-Simply because there are exceptions in non-derived environments, \textsc{Obstruent Voice Assimilation} is always consistent with the RAC whether or not it actually applies in non-derived environments.
-This is due to the ``obligatory'' condition of the RAC.
-If the rule applies in non-derived environments, then it has lexical exceptions (e.g., \emph{a}[s.b]\emph{estos}) and is not obligatory and thus free from the RAC.
-On the other hand, if the rule is subject to the RAC, it only applies in derived environments, in which it is obligatory.}
-
-The question is: does \textsc{Obstruent Voice Assimilation} contribute a meaningful characterization the English lexicon?
-More specifically, are hetero-voiced clusters underrepresented in a way that is unlikely under the hypothesis of free combination?
-To answer this question, the 720 clusters in which both the final coda consonant and initial onset consonant are obstruent are sorted into four bins, according to whether they are attested or not, and whether or not the conform to the derived constraint---both obstruents are either voiced or voiceless, like in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}, respectively---or violate it, like the aforementioned \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}.
-
-This table, shown in (\ref{ovatable}), is then submitted to the Fisher Exact Test. This statistical test is so named because it is isomorphic to the chi-square test 
-for contingency tables, but computes an exact $p$-value, whereas the chi-square test $p$-value depends on an approximation which is inappropriate for small samples.
-The small $p$-value produced by the Fisher test indicates that the rarity of the disagreeing clusters is unlikely to be due to chance.
-The simplest interpretation is that \textsc{Obstruent Voice Assimilation} applies in non-derived environments, preventing learners from positing underlying representations in which adjacent obstruents do not agree in voice.
-
-\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}] \label{ovatable}
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & \% saturation & $p$-value                   \\
-\midrule
-conforming & 80       & 370        & 18          & \multirow{2}{*}{1.1\e{-11}} \\
-violating  &  6       & 264        & 2           \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-There are two possible analyses of the small number of roots that violate the derived generalization.
-Some hetero-voiced obstruent clusters, like the one in \emph{ja}[k.d]\emph{aw}, for instance, might be analyzed as compounds by native speakers, and this analysis might itself be driven by the rarity of morph-internal adjacent hetero-voiced obstruents (cf.
-\citealp[546]{Rice2009d} on compounds in Slave).
-In fact, \citet{Mattys2001b} find that 9-month-old infants are sensitive to the contrast between [f.t] and exceptional (and unattested) [v.t], treating the latter as an indication of a word boundary and this heuristic is also available to adults \citep{Brown1956,McQueen1998b}.
-Less explanatorily, exceptional roots can be marked [$-$\textsc{Obstruent Voice Assimilation}].
-
-\subsubsection{Nasal place assimilation} 
-\label{npa}
-
-In addition to [n $\sim$ ŋ] allophony, \textsc{Nasal Place Assimilation} (\emph{SPE}:85) is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant.
-
-\begin{example}[\emph{im-}/\emph{in-} allomorphy] 
-\label{nparule}
-\begin{tabular}{l l l l l l l}
-a. & {polite}   & {i}[m.p]{olite}   & & {balance} & {i}[m.b]{alance} \\
-b. & {tangible} & {i}[n.t]{angible} & & {decent}  & {i}[n.d]{ecent}  \\
-\end{tabular}
-\end{example}
-
-% mention Jensen here
-
-\begin{example}[\textsc{Nasal Place Assimilation}]
-%\xymatrix@R=24pt@C=24pt{
-%                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
-%\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
-%\txt{[$+$\textsc{Nasal}]} & \txt{[$-$\textsc{Sonorant}]}              \\
-%}
-$\begin{bmatrix} $+$\textsc{Nas} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \end{bmatrix}~/~\gap{}~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ -\textsc{Son} \end{bmatrix}$
-\end{example}
-
-The shape of the prefix before the velar stops /k, g/ is somewhat more complex.
-\citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but that assimilation of \emph{im-}/\emph{in-} is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
-However, the stress condition does not hold for simplex words (e.g., \emph{a}[ŋ.ɡ]\emph{óra}), and is ignored here.
-
-There is evidence that even young infants acquiring English are aware of this generalization.
-\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} find that infants at 4.5 months, 9 moths, and 10 months of age, respectively, prefer to listen to nonce words which contain the homorganic nasal-obstruent clusters (e.g., \emph{u}[m.b]\emph{o}) over those which contain heterorganic clusters (e.g., \emph{u}[n.b]\emph{o}).
-Nasal-obstruent clusters arising in speech errors also undergo \textsc{Nasal Place Assimilation} (e.g.
-\emph{ra}[nd] \emph{orker}, intended \emph{ra}[ŋk] \emph{order}; \citealt[228]{Myers1993}).
-
-\citet[175]{Pierrehumbert1994} observes that ``nasal-stop sequences agree in labiality'', but makes no reference to \textsc{Nasal Place Assimilation}.
-Exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} do occur, but they are much rarer than homorganic nasal-obstruent clusters such as \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey}.
-
-\begin{example}[Lexical effects of \textsc{Nasal Place Assimilation}]
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & \% saturation & $p$-value                   \\
-\midrule
-conforming & 41       & 11         & 79          & \multirow{2}{*}{2.6\e{-05}} \\
-violating  & 8        & 20         & 29                                    \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-\subsubsection{Degemination} 
-\label{deg}
-
-A final alternation targeting English syllable contact clusters is the simplification of geminates derived by certain affixes.
-For instance, the deadjectival adverbial suffix \emph{-ly} rarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{fu}[l]\emph{y}, cf.
-\emph{free}[l]\emph{y}).
-Degemination is also found in /d/-final roots selecting the irregular /-t/ past tense suffixes (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).
-More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (see \emph{SPE}:148 and \citealt[102]{Borowsky1986}).
-Identical adjacent consonants do not occur in the corpus, nor do obstruent clusters which differ only in voice (e.g., *[p.b]), which would be made identical by the operation of \textsc{Obstruent Voice Assimilation}.
-As shown by the results of the Fisher test, this absence is unlikely to be due to chance.
-
-\begin{example}[\textsc{Degemination}]
-$\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ =\textsc{Nas} \\ =\textsc{Cont} \end{bmatrix}~\goesto~\zero~/~\gap~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ =\textsc{Nas} \\ =\textsc{Cont} \end{bmatrix}$
-\end{example}
-
-% consider what coda-final consonants there are
-
-\begin{example}[Lexical effects of \textsc{Degemination}]
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & \% saturation & $p$-value                   \\
-\midrule
-conforming & 157      & 596        & 21      & \multirow{2}{*}{7.2\e{-09}} \\
-violating  & 0        &  87        & 0                                    \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-\subsubsection{Local summary}
-
-The three \emph{SPE} rules affecting syllable consonant clusters are reliably reflected in the lexicon.
-Possible clusters that are surface-exceptions to these three rules are much more likely to be unattested than those that conform to them.
-
-\subsection{Evaluating static constraints}
-
-The second set of constraints, proposed by \citet{Pierrehumbert1994}, are ``static'' in that they have no analogue in English morphophonemics.
-Clusters of all length are considered here, not only the triconsonantal clusters considered by \citeauthor{Pierrehumbert1994}, as this restriction appears to be arbitrary.
-
-\subsubsection{Dorsal/non-coronal consonants}
-
-\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Two-consonant clusters like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} are found, however.
-Despite this, dorsal/non-coronal clusters are no less likely to occur than non-dorsal codas with non-coronal onsets (e.g.
-\emph{oi}[nt.m]\emph{ent}) or dorsal codas with coronal onsets (e.g., \emph{ve}[k.t]\emph{or}).
-
-\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}]
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & \% saturation & $p$-value \\
-\midrule
-conforming & 76 & 359 & 18 & \multirow{2}{*}{0.145} \\
-violating  &  6 &  57 & 10 \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-\subsubsection{Coda coronal obstruents}
-
-\citet[175]{Pierrehumbert1994} proposes that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many clusters such as \emph{a}[t.l]\emph{as} or \emph{e}[s.t]\emph{er}.
-Once again, coda coronal obstruent clusters are no less likely to occur than non-coronal obstruent clusters like \emph{re}[p.t]\emph{ile} or coda coronal sonorant clusters such as \emph{co}[n.d]\emph{or}.
-
-\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}]
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & \% saturation & $p$-value \\
-\midrule
-conforming & 48       & 312        & 13            & \multirow{2}{*}{0.301} \\
-violating  & 38       & 322        & 11            \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-\subsubsection{ABA clusters}
-
-Finally, \citet[][176]{Pierrehumbert1994} observes ``a lack of clusters with identical first and third elements''.
-Here, this is operationalized in the same fashion as \textsc{Degemination}, by definining ``identity'' as consonants which share the same place and manner features, but which do not necessarily agree on voicing.
-Though the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters.
-
-\begin{example}[Lexical effects of \textsc{*ABA}]
-\begin{tabular}{l r r r r}
-\toprule
-           & attested & unattested & \% saturation & $p$-value \\
-\midrule
-conforming & 41       & 478        & 8             & \multirow{2}{*}{0.818} \\
-violating  &  0       &  11        & 0             \\
-\bottomrule
-\end{tabular}
-\end{example}
-
-\subsubsection{Local summary}
-
-There is no statistically reliable evidence for the three static constraints proposed by \citeauthor{Pierrehumbert1994}.
-However, it is premature to conclude that no static constraint exists.
-It is possible that an explicit model of ``constraint discovery'', such as those proposed by \citet{Pierrehumbert1994} and \citet{Hayes2008a}, could identify reliable static constraints, and predict which of the possible clusters are attested and unattested beyond the derived constraints.
-
-\subsection{Evaluating computational models}
-
-The computational models of \citeauthor{Pierrehumbert1994} and \citeauthor{Hayes2008a}, as well as two baselines, are scored for their ability to predict which of the possible medial clusters are attested and which are not.
-For any single cluster, there are four possible outcomes.
-
-\begin{example}[Outcomes in cluster classification task]
-\begin{tabular}{c l | l l}
-                   & & \multicolumn{2}{c}{\textbf{actual outcome}}            \\
-                   & & \emph{attested} & \emph{unattested}            \\
-\midrule
-\textbf{predicted} & \emph{attested}   & true positive ($tp$)  & false positive ($fp$) \\
-\textbf{outcome}   & \emph{unattested} & false negative ($fn$) & true negative ($tn$)  \\
-\end{tabular}
-\end{example}
-
-\noindent Accuracy is the probability of a correct classification, whether a true positive or negative.
-
-\begin{unlabeledexample}
-$\displaystyle \textrm{Accuracy} = \frac{tp + tn}{tp + tn + fp + fn}$
-\end{unlabeledexample}
-
-\noindent Precision is the probability that a cluster predicted to occur is attested.
-
-\begin{unlabeledexample}
-$\displaystyle \textrm{Precision} = \frac{tp}{tp + fp}$
-\end{unlabeledexample}
-
-\noindent Recall (or sensitivity) is the probability an attested cluster is predicted to occur.
-
-\begin{unlabeledexample}
-$\displaystyle \textrm{Recall} = \frac{tp}{tp + fn}$
-\end{unlabeledexample}
-
-\noindent It is possible to maximize recall at the expense of precision, by predicting attestation for a greater number of clusters, or to increase precision at the expense of recall by predicting non-attestation for a greater number of clusters.
-$F_1$, the harmonic mean of these two measures, quantifies this trade-off.
-
-\begin{unlabeledexample}
-$\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$
-\end{unlabeledexample}
-
-A soft-margin support vector machine \citep{Cortes1995} with a linear kernel is used to convert from the per-cluster scores to attestation predictions, by finding a single cutoff which optimally separates attested and unattested clusters according to their model scores.
-This is in no way intended to be a cognitively plausible model for learning phonotactics; it simply represents an upper bound on the performance of these models.
-The results for all four models are summarized in Table \ref{cmresults}.
-
-\begin{table} \centering
-\begin{tabular}{l | r r r r}
+\begin{table} 
+\centering
+\begin{tabular}{l | rrrr}
 \toprule
@@
 \midrule
-Null baseline        & 0.813    & n.a.      & n.a.   & n.a.  \\
-Derived constraints  & 0.857    & 0.401     & 0.708  & 0.512 \\
-Expected frequency   & 0.839    & 0.210     & 0.750  & 0.328 \\
-\citet{Hayes2008a}   & 0.877    & 0.777     & 0.642  & 0.703 \\
+Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
+\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
+Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
+Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
+DC and EF           & 0.861    & 0.868     & 0.988  & 0.924 \\
 \bottomrule
 \end{tabular}
-\caption{The \citeauthor{Hayes2008a} model makes the most accurate predictions about possible and impossible syllable contact clusters, though it is only slightly better than the derived constraint baseline.}
+\caption{FIXME}
 \label{cmresults}
@@
 
-\subsubsection{Null baseline}
+\subsection{Discussion}
 
-18.8\% of the 840 possible clusters are attested.
-Consequently, a null baseline which predicts no clusters to be attested trivially achieves 81.2\% accuracy.
-The other metrics are undefined for this baseline.
-
-\subsubsection{Derived constraints}
-
-The three derived constraints define a structured baseline, in which attestation is the predicted outcome for all clusters which not violate the \emph{SPE}-derived generalizations, and non-attestation otherwise.
-This produces a significant improvement to the null baseline (accuracy = 0.857; sign test $p = 2.16$\e{-05}).
-The high recall (0.708) but low precision (0.401) of this model indicates that there are more false positives (i.e., unattested clusters which do not violate any constraint) than false negatives (i.e., exceptions to the three rules).
-It is not clear that low precision is a flaw, since some of these false positives might be accidental gaps---that is, clusters that are well-formed, but unattested.
-
-\subsubsection{Expected frequency}
-
-\citet{Pierrehumbert1994} proposes a simple model in which the wellformedness of syllable contact clusters is proportional to the joint probability, estimated from word-final coda and word-initial onset frequencies and assuming that coda and onset are statistically independent.
-There are two problems with this procedure, however.
-First, there exist languages, such as Finnish \citep[36]{Fischer-Jorgensen1952}, in which there are medial codas not found in final position; this alone suggests that ``outer'' frequencies are not the best model of what is possible medially.
-More to the point, a pilot study determined that using medial coda and onset frequencies produced a better fit, and this variant is adopted here.
-%\citeauthor{Pierrehumbert1994}'s procedure would assign clusters of this type the lowest score. 
-While \citeauthor{Pierrehumbert1994} reports that this is the best predictor of which complex clusters occur, it imposes no constraints on sequences spanning the syllable boundary, and consequently has lower accuracy (0.839) and $F_1$ (0.328) than the derived constraint baseline, though it is a small but significant improvement over the null baseline (sign test $p = 1.23$\e{-04}).
-
-\subsubsection{\citealt{Hayes2008a}}
-
-\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints.
-Since the model has many experimenter-determined parameters, a direct replication of the experiments reported by \citeauthor{Hayes2008a} is attempted; their software, feature specification, and model settings are all used.
-Since the training of this model is inherently stochastic, producing slightly different outcomes on each run, the results reported here from the lowest accuracy of ten independent runs, a practice also used by \citeauthor{Hayes2008a}.
-The model achieves the highest accuracy (0.877) and $F_1$ (0.703), but lower recall (0.642) than the alternation baseline, a significant improvement over the null baseline (sign test $p = 2.9$\e{-05}), but not significantly different from the performance of the alternation baseline (sign test $p = 0.070$).
-
-The generalizations extracted by the \citeauthor{Hayes2008a} model are in general quite similar to those derived from the \emph{SPE} alternations, but there are some notable differences.
-Unattested (and ill-formed) *[m.kl] is assigned the highest score, though it is ruled out by \textsc{Coda Nasal Place Assimilation}.
-On the other hand, very low scores are assigned to the clusters in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.
-
-\subsubsection{\citealt{McGowan2009}}
-
-\citet{McGowan2009} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2009} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
-However, \citeauthor{McGowan2009} finds that sonority of clusters accounts for a very small of the variance in cluster frequency ($R^2 = 0.077$).
-As a predictor of cluster attestation, sonority distance provided no improvement over the null baseline; consequently, it is not included in Table \ref{cmresults}.
-
-\subsubsection{Local summary}
-
-Compared to phonological alternations, current computational models of phonotactic learning are not significantly better predictors of which syllable contact clusters a
-
-\section{Conclusions}
-
-None of the foregoing results have revealed any evidence for structural constraints on the English syllable contact cluster, whether ``discovered'' by linguist or computer, except those derived from phonological alternations.
-What then is to be said about the countless clusters which do not violate any phonologically derived constraint, but yet are unattested, like [b.z] or [z.n]?
-Since these gaps appear to be quite arbitrary from a phonological perspective, all that can be said at this juncture is that these clusters must be regarded \emph{accidental gaps}: they are well-formed clusters, but simply missing from the English lexicon due to the sparsity thereof, a possibility that has long been recognized.
-
-\begin{quote}
-\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
-\citep[][16]{Fischer-Jorgensen1952}
-
-The material will never be complete. It will always contain accidental gaps\ldots{}partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
-\end{quote}
-
-\noindent Below, attempts are made to quantify the extent of accidental gaps in this domain below.
-
-\subsection{Zipfian distribution}
-
-A number of previous studies \citep[e.g.,][]{Weiss1961,Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phoneme or letter frequencies exhibit sparse distributions, in which a few types account for a large amount of the probability mass, and the remaining is divided up among a large number of rare types.
-The same is true for the type frequencies of the medial codas and onsets which make up syllable contact clusters, as well as cluster frequencies themselves.
-In Figure \ref{clus}, cluster frequencies are plotted against rank in the log-log space, and a near-linear relationship obtains ($R^2 = 0.924$), indicating that the data conform to a generalization of Zipf's (\citeyear{Zipf1949}) Law.
-See Appendix \ref{zr} for estimation details.
-
-\begin{figure} \centering
-\includegraphics{cluster.pdf}
-\caption{Syllable contact clusters exhibit a log-log linear relationship between frequency and rank that is consistent with Zipf's Law.}
-\label{clus}
-\end{figure}
-
-Sparse distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2011b} to word \citep{Baroni2009} and phoneme \citep{Belevitch1956,Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Mandelbrot1954,Miller1957,Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
-The importance of sparsity in this context is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps.
-
-\subsection{Good-Turing estimate}
-
-\citet{Good1953} proposes an estimate for the probability of all accidental gaps, $\hat{p}_0$, as the number of events that occur just once, $n_1$ divided by the number of observations.
+\subsubsection{The probability of accidental gaps}
 
 \begin{unlabeledexample}
-$\displaystyle \hat{p}_0 = \frac{n_1}{\displaystyle\sum N}$
+$\displaystyle \hat{p}_0 = \frac{n_1}{N}$
 \end{unlabeledexample}
 
-\noindent In the CELEX data, $65$ clusters occur only once, and there are 873 cluster tokens, and thus $\hat{p}_0 = 0.074$.
-There is thus a non-trivial probability that many of the ``false positive'' clusters are in fact accidental gaps.
+%67 types occur once
+%67 / 977
+%7\%
 
-\subsection{Sampling simulation}
-
-It is finally possible to ask what the lexicon might look like if it was stochastically generated from independent coda and onset frequencies, \emph{a la} \citeauthor{Pierrehumbert1994}, but also constrained by phonological neutralizations.
-The following method was repeated to generate a simulated lexicon.
+\subsubsection{Simulating medial clusters}
 
@@
 
-The cluster frequencies for a single characteristic simulation are plotted in Figure \ref{sim}, along with observed frequencies; the two distributions are nearly indistinguishable.
-The sparse cluster inventory, previously taken as evidence for static constraints on syllable contact, is approximately what one would expect even if the only constraints on possible clusters are phonologically derived.
-
 \begin{figure}
@@
 \includegraphics{sim.pdf}
-\caption{The observed cluster frequencies are closely matched by a simulated cluster inventory generated by random sampling, assuming independence and applying phonological neutralizations.
-The $Z_r$ transform \citep[][29]{Church1991} has been used to eliminate quantization in lower frequency bands.}
+\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line).
+Both simulated and observed frequencies have been smoothed using the $Z_r$ transform.}
 \label{sim}
 \end{figure}
+
+real ($\alpha = -1.65$, $R^2 = 0.940$)
+fake ($\alpha = -1.58$, $R^2 = 0.945$)
+fitt ($R^2 = 0.712$, $p = 4.5$\e{-05})
+
+\section{Conclusions}
+
+%however, the \emph{communis opinio} still is that that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
+%
+%%FIXME
+%% \citet[255]{Anderson1974}, discussing this fact, writes that \citeauthor{Haugen} ``gives an argument that shows the necessity of including syllables as units in underlying representations''.
+%
+%%FIXME
+%(for dissenting views, see \citealt{Blevins2003} and \citealt{Steriade1999})
+%
+%Linguists have long recognized underlying representations as privileged domains for stating phonotactic generalizations (e.g., \citealt{Bloomfield1930,Jakobson1932,Pike1947b}). 
+%
+%More recently, a similar role for the syllable has been recognized 
+%(e.g., \citet{Haugen1956}, \citet{Hooper1973}, \citealt{Kahn1976})
+%
+%
+%\begin{quote}
+%It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
+\citep[283]{Anderson1974}
+%\end{quote}
+%
+%
+%\begin{quote}
+%It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
+\citep[283]{Anderson1974}
+%\end{quote}
+%
+%English word-medial consonant clusters have been used, by \citet{Pierrehumbert1994} and others, to argue for the necessity of an independent phonotactics module and illustrate proposals for its architecture.
+This chapter argues, however, that the only structural constraints on the concatenation of codas and onsets to form word-medial clusters are derivative of well-documented phonological alternations in English.
+%
+%\section{Background}
+%
+%English admits a variety of word-medial consonant clusters spanning syllable boundaries, known as \emph{syllable contact clusters} or \emph{interludes}.
+These clusters can be as short as two consonants (e.g., \emph{a}[n.t]\emph{ics}; singleton intervocalic consonants are ignored here) or as long as four (e.g., \emph{mi}[n.str]\emph{el}).
+\citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in any given language \citep[cf.][]{Haugen1956}.
+%
+%\begin{quote}
+%\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
+%\end{quote}
+%
+%This chapter is concerned with the nature of these ``provisos'' as reflected by which of the ``possible'' clusters, so defined, are attested.
+While many phonotactic generalizations can be thought of as inventory restrictions imposed by prosodic constituents, additional filters on the set of possible clusters are essentially combinatoric in nature and thus distinct, there is no prosodic constituent consisting of the coda of one consonant-final syllable and the onset of a following consonant-initial syllable \citep[though see][]{Steriade1999}.
+%
+%\section{Evaluation}
+%
+%\citeauthor{Pierrehumbert1994} argues that the inventory of clusters in English shows the effect of a small number of static phonotactic constraints.
+The remainder of this chapter provides an exhaustive quantitative evaluation of this hypothesis.
+The most important constraints on medial clusters are those derived from phonological rules which apply across syllable boundaries; there is no remaining role for static constraints.
+The sparse nature of the lexicon plays a secondary role, giving rise to apparent phonotactic gaps even in the absence of static phonotactic constraints.
+%
+%\subsection{Corpus construction}
+%
+%A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database.
+The resulting list of clusters is reproduced in Appendix \ref{appendixC} and in electronic form at the author's website.
+%
+%\subsubsection{Simplex words}
+%
+%An analysis of lexical constraints requires a corpus of lexical representations.
+So as to remain agnostic about particulars of morphological theory, the goal here is to identify words which bear no overt inflection, and which do not admit any plausible further decomposition, henceforth ``simplex words''.
+For the structuralists, a procedure for parsing words into morphs, just the thing for locating simplex words, held the status of ``philosopher's stone'', a sort of near-mythical objective of the endeavor, but actual attempts to develop such a procedure \citep[e.g.,][]{Harris1955,Nida1948} are now recognized as anything but theory-neutral.
+Yet, there is no alternative to adopting some computational procedure or human-coded database as the ``gold standard'' for a study of this type.
+\citet{Pierrehumbert1994} analyzes a list of entries in the Collins English dictionary by hand, using her intuitions to filter out ``complex'' words as well as those which are not ``reasonably familiar''.
+Unfortunately, the resulting list was neither published nor circulated, and the possibility of replicating \citeauthor{Pierrehumbert1994}'s' sensations of morphological complexity is remote.
+Following prior studies of syllable contact clusters by \citet[ chapter 3]{Hammond1999a} and \citet[ chapter 8]{Duanmu2009}, a list of simplex words was derived from the ``lemmas'' (that is, uninflected forms) in the English portion of CELEX \citep{CELEX}, a database of morphological, phonological, and syntactic annotations based on the COBUILD corpus, a resource that was also used to construct the Collins English dictionary employed by \citeauthor{Pierrehumbert1994}.
+%
+%Two further filters were applied to this list.
+First, any lemma coded as an unassimilated loanword is excluded.
+Secondly, all lemmas that have a ``morphological status'' other than ``monomorpheme'' are excluded.
+This is crucial because CELEX makes no guarantee that lemmas are simplex; indeed, many lemmas are products of derivation (e.g., \emph{abusive}, from \emph{abuse}).
+These stringent criteria have a profound effect on the makeup of the data.
+For instance, nearly all exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) cited by \citet[74]{Hammond1999a} are excluded as unassimilated loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as morphologically complex (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}).
+%
+%``Neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}, have an ambiguous and long-debated morphological status.
+\citeauthor{Pierrehumbert1994} attempts to distinguish between words like \emph{complete}, \emph{extreme}, \emph{inspect}, \emph{obtuse}, which she assumes to be simplex, and \emph{excrete} and \emph{transparent}, which she considers complex.
+However, CELEX codes these words as complex and they are thus excluded.
+This is consistent with converging evidence that speakers decompose these Latinate forms.
+The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.
+In a similar vein, \citet[11--13]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.
+%
+%\begin{example}[Bound stem-specific allomorphy]
+%\begin{tabular}{l l l l l l}
+%a. & {adhere}   & {adhesion}   \\
+%   & {cohere}   & {cohesion}   \\
+%b. & {conceive} & {conception} \\
+%   & {perceive} & {perception} \\
+%\end{tabular}
+%\end{example}
+%
+%There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement.
+\citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.
+\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
+%
+%\begin{example}[Latinate verbs and small clauses] 
+%\label{harley}
+%\begin{tabular}{ll@{}ll@{}l}
+%a. & * & {exhibit him the painting} & ~ & {show him the painting} \\
+%b. & * & {imbibe himself stupid}    & ~ & {drink himself stupid}  \\
+%c. & * & {exhibit it off}           & ~ & {show it off}           \\
+%\end{tabular}
+%\end{example}
+%
+%Decomposition is also supported by lexical decision results.
+Among nonce words, those that can be exhaustively decomposed into prefix and bound stem, like *\emph{de}-\emph{juvenate}, are processed slower that contain a licit morph but cannot be exhaustively decomposed, like *\emph{de-pertoire}, and these in turn are processed slower than those which admit no such decomposition \citep{Taft1975}.\footnote{The same contrasts are present in many other domains where decomposition into root and affix(es) is less controversial, for instance, inflected verbs in Italian \citep{Caramazza1988}.} \citet{Emmorey1989} reports facilitative auditory priming between pairs of words which share a bound stem, like \emph{permit}-\emph{submit} \citep[though see][]{Marslen-Wilson1994}.\footnote{It may be the case that full decomposition is only one of the two methods for the processing of complex words, the other being whole word looking, either running in serial \citep{Caramazza1988} or in parallel \citep{Baayen1997b}.
+Whatever evidence that is amassed in support of a whole word lookup procedure, however, does not undermine the considerable evidence for decomposition.}
+%
+%\subsubsection{Syllabification and phonologization}
+%
+%CELEX includes broad, syllabified Received Pronunciation transcriptions of individual wordforms.
+However, a casual inspection of the data reveals the unsystematic nature of these syllabifications.
+Any two words sharing a nucleus-medial cluster-nucleus sequence should be given the same parse, since syllabification is universally non-contrastive.
+Yet many putative syllabification contrasts can be found in CELEX; for instance, compare the parses given to the clusters in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][II.294]{AOE}.}
+%
+%A systematic syllabification requires an automated procedure for delimiting medial clusters and parsing them into coda and onset.
+Since this procedure is applied only to medial clusters, this study remains agnostic on several contentious issues concerning syllabification of English; for instance, there is no need to address the status of so-called ``ambisyllabic'' consonants, since this does not occur with medial clusters, or morphological effects on syllabification, since no complex words are included.
+The technique used here is a variation on the theme of onset maximization (e.g., \citealp[42f.]{Kahn1976}, \citealp{Kurylowicz1948}, \citealp[75]{Pulgram1970}, \citealp[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset.
+Initial onsets are used to define what is a ``possible'' medial onset \citep[though cf.][36]{Fischer-Jorgensen1952}.
+Medial clusters in words like \emph{neu}[.tɹ]\emph{on} or \emph{bi}[.stɹ]\emph{o} are found in word-initial position (e.g., [tɹ]\emph{ansit}, [stɹ]\emph{ike}), so onset maximization assigns the entire medial cluster to the onset, leaving the medial coda empty.
+In contrast, the cluster in \emph{mi}[n.stɹ]\emph{el} is not found word-initially; the maximal onset here is [stɹ], and the residual [n] is assigned to the coda.
+%
+%%\citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''
+%%\citet[?]{Lombardi1999} says that English lacks devoicing in Level II.
+%
+%\paragraph{Stressed lax vowels} When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
+\citet[][55]{Harris1994} notes that onset maximization produces an incorrect result when, in addition, the medial cluster is a valid onset: in \emph{whisper}, \emph{vestige}, and \emph{musket}, onset maximization incorrectly assigns [sp, st, sk] to the medial onset, leaving the lax vowel unchecked.
+Consequently, the first consonant of a medial cluster is assigned to the coda of a preceding syllable before a stressed lax vowel \citep[cf.][48]{Pulgram1970}, a minimal modification of the onset maximization parse.
+%
+%\paragraph{Affricates} If the ffricates [tʃ, dʒ] are treated as stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990}, the previous principle will incorrectly assign the two segments of the affricate to separate syllables before stressed lax vowels (e.g., *\emph{ra}[t.ʃ]\emph{et} or \emph{a}[d.ʒ]\emph{ile}) unless some ad hoc constraint \citep[e.g.,][]{Wells1990} is in place.
+A more traditional analysis of affricates, in which they are simple segments distinguished from pure stops by stridency \citep[24]{Jakobson1961} or delayed release (\emph{SPE}:321f.) is assumed here, following \citeauthor{Pierrehumbert1994}.
+This assumption can be motivated by the tendency of affricates to pattern with individual segments in phonotactic generalizations.
+For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but bans true onset clusters \citep[9]{Launey2011}.
+%%\citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.
+%
+%\paragraph{The velar nasal} According to some \citep[e.g.,][]{Sapir1925}, [ŋ] in English is a phoneme in its own right, but others (e.g., \emph{SPE}:85, \citealp{Borowsky1986}:65f.) hold that it is simply the allophone of /n/ before underlying /k, ɡ/ (with later deletion of /ɡ/ in some contexts).
+The total absence of onset [ŋ], where it cannot be followed by a dorsal consonant which conditions the velar allophone, is strong evidence that [ŋ] is a pure allophone of /n/, the analysis assumed here.
+%
+%\paragraph{[j] onglides} While some have argued that the front onglide in words such as \emph{val}[j]\emph{ue} is inserted by rule (\emph{SPE}:196, \citealp[][89]{Halle1985a}, \citealp[][217]{McMahon1990}), the very presence or absence of the glide is contrastive (e.g., \emph{coot} \alt{} \emph{cute}, \emph{booty} \alt{} \emph{beauty}) indicating that it is present in underlying representation (\citealp{Anderson1988b}, \citealp[278]{Borowsky1986}).
+The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{ju}[n.j]\emph{or}).
+There is considerable evidence for this assumption.
+When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u], suggesting that the glide is nuclear (\citealp{Davis1995}, \citealp[][61f.]{Harris1994}, \citealp[][232]{Hayes1980}).
+\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters, though they may be followed by [ju] in words like [mj]\emph{use} or [vj]\emph{iew}, also suggesting the nuclear affiliation of the glide.
+There is also external support for this analysis: the [ju] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005}, \emph{shm}-reduplication \citep{Nevins2003}, and speech errors (e.g., [kju]\emph{mor} [h]\emph{omponent}, intended [hju]\emph{mor component}; \citealp[130]{Shattuck-Hufnagel1986}).
+%
+%\paragraph{[w] onglides} The phonotactic properties of the back onglide [w] are the opposite of the front onglide.
+Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
+Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[161]{Davis1995}.
+Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[166]{Davis1995}.
+Taken together, these facts indicate that the back onglide is assigned to the onset.
+%
+%\paragraph{Syllabic \emph{r}} Word-medial syllabic \emph{r} has been lost in Received Pronunciation, the accent used for the CELEX transcriptions.
+Even in \emph{r}-ful dialects, though, there is reason to believe that syllabic \emph{r} is fact nuclear, and thus irrelevant to syllable contact clusters.
+Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}): compare American English \emph{fern}, \emph{fir}, \emph{fur} to \emph{pet}, \emph{pit}, \emph{putt}.
+Syllabic \emph{r} patterns with vowels and not other consonants for several phonological processes: it is the only consonant which does not block variable glottalization of following /t/ in British dialects \citep[258]{Harris1994}.
+%
+%\begin{example}[/t/-\textsc{Glottalization} in British English]
+%\begin{tabular}{l l l l@{} l l l}
+%a. & {des}[ɚt]    & \alt{} &   & {des}[ɚʔ]    \\
+%   & {c}[ɚt]{ain} & \alt{} &   & {c}[ɚʔ]{ain} \\
+%b. & {fi}[st]     & \alt{} & * & {fi}[sʔ]     \\
+%   & {mi}[st]{er} & \alt{} & * & {mi}[sʔ]{er} \\
+%\end{tabular}
+%\end{example}
+%
+%\noindent Syllabic \emph{r} is the only consonant which does not trigger variable deletion of a following /t, d/ in American dialects \citep[8]{Guy1980}.
+%
+%\begin{example}[/t, d/-\textsc{Deletion} in American English]
+%\begin{tabular}{l l l l@{} l l l}
+%a. & {be}[lt] & \alt{} &   & {be}[l] \\
+%   & {me}[nd] & \alt{} &   & {me}[n] \\
+%b. & {sh}[ɚt] & \alt{} & * & {sh}[ɚ] \\
+%   & {c}[ɚd]  & \alt{} & * & {c}[ɚ]  \\
+%\end{tabular}
+%\end{example}
+%
+%\subsubsection{Lexical statistics}
+%
+%CELEX contains 6,876 simplex words of English, comprising 21 unique medial codas and 40 unique medial onsets.
+Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested.
+%
+%\subsection{Evaluating derived constraints}
+%
+%\emph{SPE} describes three phonological alternations which target syllable contact clusters.
+The corresponding constraints on simplex syllable contact clusters are considered below.
+The first such derived constraint, obstruent voice assimilation, is presented as a tutorial introduction.
+The second set of constraints consists of three static phonotactic constraints proposed by \citet{Pierrehumbert1994}.
+%
+%\subsubsection{Obstruent voice assimilation} 
+%\label{ova}
+%
+%The [t \alt{} d] allomorphs of the regular past (e.g.
+\emph{nap}[t] $\sim$ \emph{nab}[d]), and the [s \alt{} z] allomorphs of the regular noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), Saxon genitive, and third person singular active indicative verb agreement, exemplify \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178).\footnote{Underlying /-d, -z/ are argued for by \citet[282]{Hockett1958}, \citet[210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, \citet{Anderson1973a}, \citet[102]{Pinker1988}, and \citet[284f.]{Bakovic2005b}; alternatives are suggested by \citet[210f.]{LANGUAGE}, \citet[426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[135]{Borowsky1986}.}
+%
+%\begin{example}[Inflectional affix voice assimilation]
+%\begin{tabular}{l l l l l l}
+%a. & {nap} & {nap}[t] & & {nab} & {nab}[d] \\
+%b. & {lap} & {lap}[s] & & {lab} & {lab}[z] \\
+%\end{tabular}
+%\end{example}
+%
+%\noindent The devoicing of the two voiced obstruent suffixes can be formalized as a 
+%rule spreading the voicing specification of an obstruent rightward.
+%
+%\begin{example}[\textsc{Obstruent Voice Assimilation}]
+%%\xymatrix@R=24pt@C=24pt{
+%%\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{--}[dr] &                               \\
+%%\txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
+%%\txt{[$-$\textsc{Sonorant}]}                  & \txt{[$-$\textsc{Sonorant}]} \\
+%%}
+%$\begin{bmatrix} -\textsc{Son} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Voi} \end{bmatrix}~/~\gap~\begin{bmatrix} =\textsc{Voi} \\ -\textsc{Son} \end{bmatrix}$
+%\end{example}
+%
+%It is quite apparent that root-internal hetero-voiced obstruent clusters like [s.b] or [z.t] are also rare compared to clusters which are uniformly voiced or uniformly voiceless, as predicted the assimilation rule.
+Yet, \citet{Pierrehumbert1994} does not mention voice assimilation in her study of English syllable contact clusters.
+\citet[][74f.]{Hammond1999a} cites words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain hetero-voiced obstruent clusters, as evidence that the process does not apply root-internally, though few of his examples are in fact simplex according to CELEX.\footnote{The \textsc{Revised Alternation Condition} (RAC) proposed by \citeauthor{Kiparsky1973a} (\citeyear{Kiparsky1973a}:163, \citeyear{Kiparsky1982a}:152) blocks the application of obligatory neutralization processes like \textsc{Obstruent Voice Assimilation} in root-internal (i.e., non-derived) environments.
+Simply because there are exceptions in non-derived environments, \textsc{Obstruent Voice Assimilation} is always consistent with the RAC whether or not it actually applies in non-derived environments.
+This is due to the ``obligatory'' condition of the RAC.
+If the rule applies in non-derived environments, then it has lexical exceptions (e.g., \emph{a}[s.b]\emph{estos}) and is not obligatory and thus free from the RAC.
+On the other hand, if the rule is subject to the RAC, it only applies in derived environments, in which it is obligatory.}
+%
+%The question is: does \textsc{Obstruent Voice Assimilation} contribute a meaningful characterization the English lexicon?
+More specifically, are hetero-voiced clusters underrepresented in a way that is unlikely under the hypothesis of free combination?
+To answer this question, the 720 clusters in which both the final coda consonant and initial onset consonant are obstruent are sorted into four bins, according to whether they are attested or not, and whether or not the conform to the derived constraint---both obstruents are either voiced or voiceless, like in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}, respectively---or violate it, like the aforementioned \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}.
+%
+%This table, shown in (\ref{ovatable}), is then submitted to the Fisher Exact Test. This statistical test is so named because it is isomorphic to the chi-square test 
+%for contingency tables, but computes an exact $p$-value, whereas the chi-square test $p$-value depends on an approximation which is inappropriate for small samples.
+The small $p$-value produced by the Fisher test indicates that the rarity of the disagreeing clusters is unlikely to be due to chance.
+The simplest interpretation is that \textsc{Obstruent Voice Assimilation} applies in non-derived environments, preventing learners from positing underlying representations in which adjacent obstruents do not agree in voice.
+%
+%\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}] \label{ovatable}
+%\begin{tabular}{l r r r r}
+%\toprule
+%           & attested & unattested & \% saturation & $p$-value                   \\
+%\midrule
+%conforming & 80       & 370        & 18\%          & \multirow{2}{*}{1.1\e{-11}} \\
+%violating  &  6       & 264        & 2\%           \\
+%\bottomrule
+%\end{tabular}
+%\end{example}
+%
+%There are two possible analyses of the small number of roots that violate the derived generalization.
+Some hetero-voiced obstruent clusters, like the one in \emph{ja}[k.d]\emph{aw}, for instance, might be analyzed as compounds by native speakers, and this analysis might itself be driven by the rarity of morph-internal adjacent hetero-voiced obstruents (cf.
+\citealp[546]{Rice2009d} on compounds in Slave).
+In fact, \citet{Mattys2001b} find that 9-month-old infants are sensitive to the contrast between [f.t] and exceptional (and unattested) [v.t], treating the latter as an indication of a word boundary and this heuristic is also available to adults \citep{Brown1956,McQueen1998b}.
+Less explanatorily, exceptional roots can be marked [$-$\textsc{Obstruent Voice Assimilation}].
+%
+%\subsubsection{Nasal place assimilation} 
+%\label{npa}
+%
+%In addition to [n $\sim$ ŋ] allophony, \textsc{Nasal Place Assimilation} (\emph{SPE}:85) is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant.
+%
+%\begin{example}[\emph{im-}/\emph{in-} allomorphy] 
+%\label{nparule}
+%\begin{tabular}{l l l l l l l}
+%a. & {polite}   & {i}[m.p]{olite}   & & {balance} & {i}[m.b]{alance} \\
+%b. & {tangible} & {i}[n.t]{angible} & & {decent}  & {i}[n.d]{ecent}  \\
+%\end{tabular}
+%\end{example}
+%
+%% mention Jensen here
+%
+%\begin{example}[\textsc{Nasal Place Assimilation}]
+%%\xymatrix@R=24pt@C=24pt{
+%%                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
+%%\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
+%%\txt{[$+$\textsc{Nasal}]} & \txt{[$-$\textsc{Sonorant}]}              \\
+%%}
+%$\begin{bmatrix} $+$\textsc{Nas} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \end{bmatrix}~/~\gap{}~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ -\textsc{Son} \end{bmatrix}$
+%\end{example}
+%
+%The shape of the prefix before the velar stops /k, g/ is somewhat more complex.
+\citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but that assimilation of \emph{im-}/\emph{in-} is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
+However, the stress condition does not hold for simplex words (e.g., \emph{a}[ŋ.ɡ]\emph{óra}), and is ignored here.
+%
+%There is evidence that even young infants acquiring English are aware of this generalization.
+\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} find that infants at 4.5 months, 9 moths, and 10 months of age, respectively, prefer to listen to nonce words which contain the homorganic nasal-obstruent clusters (e.g., \emph{u}[m.b]\emph{o}) over those which contain heterorganic clusters (e.g., \emph{u}[n.b]\emph{o}).
+Nasal-obstruent clusters arising in speech errors also undergo \textsc{Nasal Place Assimilation} (e.g.
+\emph{ra}[nd] \emph{orker}, intended \emph{ra}[ŋk] \emph{order}; \citealt[228]{Myers1993}).
+%
+%\citet[175]{Pierrehumbert1994} observes that ``nasal-stop sequences agree in labiality'', but makes no reference to \textsc{Nasal Place Assimilation}.
+Exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} do occur, but they are much rarer than homorganic nasal-obstruent clusters such as \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey}.
+%
+%\begin{example}[Lexical effects of \textsc{Nasal Place Assimilation}]
+%\begin{tabular}{l r r r r}
+%\toprule
+%           & attested & unattested & \% saturation & $p$-value                   \\
+%\midrule
+%conforming & 41       & 11         & 79          & \multirow{2}{*}{2.6\e{-05}} \\
+%violating  & 8        & 20         & 29                                    \\
+%\bottomrule
+%\end{tabular}
+%\end{example}
+%
+%\subsubsection{Degemination} 
+%\label{deg}
+%
+%A final alternation targeting English syllable contact clusters is the simplification of geminates derived by certain affixes.
+For instance, the deadjectival adverbial suffix \emph{-ly} rarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{fu}[l]\emph{y}, cf.
+\emph{free}[l]\emph{y}).
+Degemination is also found in /d/-final roots selecting the irregular /-t/ past tense suffixes (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).
+More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (see \emph{SPE}:148 and \citealt[102]{Borowsky1986}).
+Identical adjacent consonants do not occur in the corpus, nor do obstruent clusters which differ only in voice (e.g., *[p.b]), which would be made identical by the operation of \textsc{Obstruent Voice Assimilation}.
+As shown by the results of the Fisher test, this absence is unlikely to be due to chance.
+%
+%\begin{example}[\textsc{Degemination}]
+%$\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ =\textsc{Nas} \\ =\textsc{Cont} \end{bmatrix}~\goesto~\zero~/~\gap~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ =\textsc{Nas} \\ =\textsc{Cont} \end{bmatrix}$
+%\end{example}
+%
+%% consider what coda-final consonants there are
+%
+%\begin{example}[Lexical effects of \textsc{Degemination}]
+%\begin{tabular}{l r r r r}
+%\toprule
+%           & attested & unattested & \% saturation & $p$-value                   \\
+%\midrule
+%conforming & 157      & 596        & 21      & \multirow{2}{*}{7.2\e{-09}} \\
+%violating  & 0        &  87        & 0                                    \\
+%\bottomrule
+%\end{tabular}
+%\end{example}
+%
+%\subsubsection{Local summary}
+%
+%The three \emph{SPE} rules affecting syllable consonant clusters are reliably reflected in the lexicon.
+Possible clusters that are surface-exceptions to these three rules are much more likely to be unattested than those that conform to them.
+%
+%\subsection{Evaluating static constraints}
+%
+%The second set of constraints, proposed by \citet{Pierrehumbert1994}, are ``static'' in that they have no analogue in English morphophonemics.
+Clusters of all length are considered here, not only the triconsonantal clusters considered by \citeauthor{Pierrehumbert1994}, as this restriction appears to be arbitrary.
+%
+%\subsubsection{Dorsal/non-coronal consonants}
+%
+%\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Two-consonant clusters like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} are found, however.
+Despite this, dorsal/non-coronal clusters are no less likely to occur than non-dorsal codas with non-coronal onsets (e.g.
+\emph{oi}[nt.m]\emph{ent}) or dorsal codas with coronal onsets (e.g., \emph{ve}[k.t]\emph{or}).
+%
+%\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}]
+%\begin{tabular}{l r r r r}
+%\toprule
+%           & attested & unattested & \% saturation & $p$-value \\
+%\midrule
+%conforming & 76 & 359 & 18 & \multirow{2}{*}{0.145} \\
+%violating  &  6 &  57 & 10 \\
+%\bottomrule
+%\end{tabular}
+%\end{example}
+%
+%\subsubsection{Coda coronal obstruents}
+%
+%\citet[175]{Pierrehumbert1994} proposes that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many clusters such as \emph{a}[t.l]\emph{as} or \emph{e}[s.t]\emph{er}.
+Once again, coda coronal obstruent clusters are no less likely to occur than non-coronal obstruent clusters like \emph{re}[p.t]\emph{ile} or coda coronal sonorant clusters such as \emph{co}[n.d]\emph{or}.
+%
+%\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}]
+%\begin{tabular}{l r r r r}
+%\toprule
+%           & attested & unattested & \% saturation & $p$-value \\
+%\midrule
+%conforming & 48       & 312        & 13            & \multirow{2}{*}{0.301} \\
+%violating  & 38       & 322        & 11            \\
+%\bottomrule
+%\end{tabular}
+%\end{example}
+%
+%\subsubsection{ABA clusters}
+%
+%Finally, \citet[][176]{Pierrehumbert1994} observes ``a lack of clusters with identical first and third elements''.
+Here, this is operationalized in the same fashion as \textsc{Degemination}, by definining ``identity'' as consonants which share the same place and manner features, but which do not necessarily agree on voicing.
+Though the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters.
+%
+%\begin{example}[Lexical effects of \textsc{*ABA}]
+%\begin{tabular}{l r r r r}
+%\toprule
+%           & attested & unattested & \% saturation & $p$-value \\
+%\midrule
+%conforming & 41       & 478        & 8             & \multirow{2}{*}{0.818} \\
+%violating  &  0       &  11        & 0             \\
+%\bottomrule
+%\end{tabular}
+%\end{example}
+%
+%\subsubsection{Local summary}
+%
+%There is no statistically reliable evidence for the three static constraints proposed by \citeauthor{Pierrehumbert1994}.
+However, it is premature to conclude that no static constraint exists.
+It is possible that an explicit model of ``constraint discovery'', such as those proposed by \citet{Pierrehumbert1994} and \citet{Hayes2008a}, could identify reliable static constraints, and predict which of the possible clusters are attested and unattested beyond the derived constraints.
+%
+%\subsection{Evaluating computational models}
+%
+%The computational models of \citeauthor{Pierrehumbert1994} and \citeauthor{Hayes2008a}, as well as two baselines, are scored for their ability to predict which of the possible medial clusters are attested and which are not.
+For any single cluster, there are four possible outcomes.
+%
+%\begin{example}[Outcomes in cluster classification task]
+%\begin{tabular}{c l | l l}
+%                   & & \multicolumn{2}{c}{\textbf{actual outcome}}            \\
+%                   & & \emph{attested} & \emph{unattested}            \\
+%\midrule
+%\textbf{predicted} & \emph{attested}   & true positive ($tp$)  & false positive ($fp$) \\
+%\textbf{outcome}   & \emph{unattested} & false negative ($fn$) & true negative ($tn$)  \\
+%\end{tabular}
+%\end{example}
+%
+%\noindent Accuracy is the probability of a correct classification, whether a true positive or negative.
+%
+%\begin{unlabeledexample}
+%$\displaystyle \textrm{Accuracy} = \frac{tp + tn}{tp + tn + fp + fn}$
+%\end{unlabeledexample}
+%
+%\noindent Precision is the probability that a cluster predicted to occur is attested.
+%
+%\begin{unlabeledexample}
+%$\displaystyle \textrm{Precision} = \frac{tp}{tp + fp}$
+%\end{unlabeledexample}
+%
+%\noindent Recall (or sensitivity) is the probability an attested cluster is predicted to occur.
+%
+%\begin{unlabeledexample}
+%$\displaystyle \textrm{Recall} = \frac{tp}{tp + fn}$
+%\end{unlabeledexample}
+%
+%\noindent It is possible to maximize recall at the expense of precision, by predicting attestation for a greater number of clusters, or to increase precision at the expense of recall by predicting non-attestation for a greater number of clusters.
+$F_1$, the harmonic mean of these two measures, quantifies this trade-off.
+%
+%\begin{unlabeledexample}
+%$\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$
+%\end{unlabeledexample}
+%
+%A soft-margin support vector machine \citep{Cortes1995} with a linear kernel is used to convert from the per-cluster scores to attestation predictions, by finding a single cutoff which optimally separates attested and unattested clusters according to their model scores.
+This is in no way intended to be a cognitively plausible model for learning phonotactics; it simply represents an upper bound on the performance of these models.
+The results for all four models are summarized in Table \ref{cmresults}.
+%
+%\begin{table} \centering
+%\begin{tabular}{l | r r r r | r}
+%\toprule
+%                     & accuracy & precision & recall & $F_1$ \\
+%\midrule
+%Null baseline        &     .808 &      .808 &  1.000 &  .894 \\
+%\citet{Hayes2008a}   &     .837 &      .948 &   .852 &  .898 \\ %& .002       \\
+%Expected frequency   &     .835 &      .845 &   .983 &  .909 \\ %& 1.1\e{-05} \\
+%Derived constraints  &     .840 &      .922 &   .885 &  .903 \\
+%DC and EF            &     .864 &      .871 &   .989 &  .926 \\ %& 1.9\e{-13} \\
+%\bottomrule
+%\end{tabular}
+%\caption{FIXME}
+%\label{cmresults}
+%\end{table}
+%
+%\subsubsection{Null baseline}
+%
+%19\% of the 880 possible clusters are attested.
+Consequently, a null baseline which predicts that no clusters are possible will be 81\% accurate, trivially.
+Because this baseline FIXME RECALL FIXME PRECISION
+%
+%\subsubsection{Derived constraints}
+%
+%The three derived constraints define a structured baseline, in which attestation is the predicted outcome for all clusters which not violate the \emph{SPE}-derived generalizations, and non-attestation otherwise.
+This results in a significant improvement to the null baseline (sign test $p = 0.006$).
+This model has poor recall, however, because many well-formed clusters are unattested.
+This is not a flaw, but rather the signature of accidental gaps.
+%
+%\subsubsection{Expected frequency}
+%
+%\citet{Pierrehumbert1994} proposes a simple model in which the wellformedness of syllable contact clusters is proportional to the joint probability, estimated from word-final coda and word-initial onset frequencies and assuming that coda and onset are statistically independent.
+There are two problems with this procedure, however.
+First, there exist languages, such as Finnish \citep[36]{Fischer-Jorgensen1952}, in which there are medial codas not found in final position; this alone suggests that ``outer'' frequencies are not the best model of what is possible medially.
+More to the point, a pilot study determined that using medial coda and onset frequencies produced a better fit, and this variant is adopted here.
+%%\citeauthor{Pierrehumbert1994}'s procedure would assign clusters of this type the lowest score. 
+%While \citeauthor{Pierrehumbert1994} reports that this is the best predictor of which complex clusters occur, it imposes no constraints on sequences spanning the syllable boundary, and consequently has lower accuracy (0.839) and $F_1$ (0.328) than the derived constraint baseline, though it is a small but significant improvement over the null baseline (sign test $p = 1.23$\e{-04}).
+%
+%\subsubsection{\citealt{Hayes2008a}}
+%
+%\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints.
+Since the model has many experimenter-determined parameters, a direct replication of the experiments reported by \citeauthor{Hayes2008a} is attempted; their software, feature specification, and model settings are all used.
+Since the training of this model is inherently stochastic, producing slightly different outcomes on each run, the results reported here from the lowest accuracy of ten independent runs, a practice also used by \citeauthor{Hayes2008a}.
+The model achieves the highest accuracy (0.877) and $F_1$ (0.703), but lower recall (0.642) than the alternation baseline, a significant improvement over the null baseline (sign test $p = 2.9$\e{-05}), but not significantly different from the performance of the alternation baseline (sign test $p = 0.070$).
+%
+%The generalizations extracted by the \citeauthor{Hayes2008a} model are in general quite similar to those derived from the \emph{SPE} alternations, but there are some notable differences.
+Unattested (and ill-formed) *[m.kl] is assigned the highest score, though it is ruled out by \textsc{Coda Nasal Place Assimilation}.
+On the other hand, very low scores are assigned to the clusters in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.
+%
+%\subsubsection{\citealt{McGowan2009}}
+%
+%\citet{McGowan2009} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2009} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
+However, \citeauthor{McGowan2009} finds that sonority of clusters accounts for a very small of the variance in cluster frequency ($R^2 = 0.077$).
+As a predictor of cluster attestation, sonority distance provided no improvement over the null baseline; consequently, it is not included in Table \ref{cmresults}.
+%
+%\subsubsection{Local summary}
+%
+%Compared to phonological alternations, current computational models of phonotactic learning are not significantly better predictors of which syllable contact clusters a
+%
+%\section{Conclusions}
+%
+%None of the foregoing results have revealed any evidence for structural constraints on the English syllable contact cluster, whether ``discovered'' by linguist or computer, except those derived from phonological alternations.
+What then is to be said about the countless clusters which do not violate any phonologically derived constraint, but yet are unattested, like [b.z] or [z.n]?
+Since these gaps appear to be quite arbitrary from a phonological perspective, all that can be said at this juncture is that these clusters must be regarded \emph{accidental gaps}: they are well-formed clusters, but simply missing from the English lexicon due to the sparsity thereof, a possibility that has long been recognized.
+%
+%\begin{quote}
+%\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
+\citep[][16]{Fischer-Jorgensen1952}
+%
+%The material will never be complete. It will always contain accidental gaps\ldots{}partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
+%\end{quote}
+%
+%\noindent Below, attempts are made to quantify the extent of accidental gaps in this domain below.
+%
+%\subsection{Zipfian distribution}
+%
+%A number of previous studies \citep[e.g.,][]{Weiss1961,Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phoneme or letter frequencies exhibit sparse distributions, in which a few types account for a large amount of the probability mass, and the remaining is divided up among a large number of rare types.
+The same is true for the type frequencies of the medial codas and onsets which make up syllable contact clusters, as well as cluster frequencies themselves.
+In Figure \ref{clus}, cluster frequencies are plotted against rank in the log-log space, and a near-linear relationship obtains ($R^2 = 0.924$), indicating that the data conform to a generalization of Zipf's (\citeyear{Zipf1949}) Law.
+See Appendix \ref{zr} for estimation details.
+%
+%\begin{figure} \centering
+%\includegraphics{cluster.pdf}
+%\caption{Syllable contact clusters exhibit a log-log linear relationship between frequency and rank that is consistent with Zipf's Law.}
+%\label{clus}
+%\end{figure}
+%
+%Sparse distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2011b} to word \citep{Baroni2009} and phoneme \citep{Belevitch1956,Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Mandelbrot1954,Miller1957,Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
+The importance of sparsity in this context is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps.
+%
+%\subsection{Good-Turing estimate}
+%
+%\citet{Good1953} proposes an estimate for the probability of all accidental gaps, $\hat{p}_0$, as the number of events that occur just once, $n_1$ divided by the number of observations.
+%
+%\begin{unlabeledexample}
+%$\displaystyle \hat{p}_0 = \frac{n_1}{\displaystyle\sum N}$
+%\end{unlabeledexample}
+%
+%\noindent In the CELEX data, $65$ clusters occur only once, and there are 873 cluster tokens, and thus $\hat{p}_0 = 0.074$.
+There is thus a non-trivial probability that many of the ``false positive'' clusters are in fact accidental gaps.
+%
+%\subsection{Sampling simulation}
+%
+%It is finally possible to ask what the lexicon might look like if it was stochastically generated from independent coda and onset frequencies, \emph{a la} \citeauthor{Pierrehumbert1994}, but also constrained by phonological neutralizations.
+The following method was repeated to generate a simulated lexicon.
+%
+%\begin{example}[Simulation procedure]
+%\begin{tabular}{l l}
+%a. & Sample a medial coda according to the observed probabilities  \\
+%b. & Sample a medial onset according to the observed probabilities \\
+%c. & Apply all matching \emph{SPE} rules to cluster formed by their concatenation \\
+%\end{tabular}
+%\end{example}
+%
+%The cluster frequencies for a single characteristic simulation are plotted in Figure \ref{sim}, along with observed frequencies; the two distributions are nearly indistinguishable.
+The sparse cluster inventory, previously taken as evidence for static constraints on syllable contact, is approximately what one would expect even if the only constraints on possible clusters are phonologically derived.
+%
+%\begin{figure}
+%\centering
+%\includegraphics{sim.pdf}
+%\caption{The observed cluster frequencies are closely matched by a simulated cluster inventory generated by random sampling, assuming independence and applying phonological neutralizations.
+The $Z_r$ transform \citep[][29]{Church1991} has been used to eliminate quantization in lower frequency bands.}
+%\label{sim}
+%\end{figure}

### 2012-10-29 abs.tex
Writing on 2-4 and appendices

@@
 
-Chapter \ref{clusters} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English.
+Chapter \ref{gaps} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English.
 Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound, whereas phonological alternations impose robust restrictions on the cluster inventory.

### 2012-10-30 2.tex
Graphs and tables

@@
 \noindent
-The position that the well-formedness of nonce words is consistent the view of syntactic grammaticality taken in \emph{LSLT} and \emph{Aspects} \citep{LSLT,ASPECTS}, where it is claimed that different syntactic vioaltions result in different degrees of ungrammaticality.
+The position that the well-formedness of nonce words is consistent the view of syntactic grammaticality taken in \emph{LSLT} and \emph{Aspects} \citep{LSLT,ASPECTS}, where it is claimed that different syntactic violations result in different degrees of ungrammaticality.
 
@@
 
+\begin{quote}
+\ldots{}A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
+\citep[371]{Shademan2006}
+\end{quote}
+
+\citet[382]{Hayes2008a} ``consider the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models''.
+
+\begin{quote}
+No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
+\end{quote}
+
+\begin{quote}
+Some have responded to these findings very consistently, by asserting that the experimental findings are to be interpreted as before: that, psychologically speaking, odd numbers as well as birds and vegetables are graded concepts\ldots{} We reject this conclusion just because we could not explain how a person could compute with integers who believed that 7 was odder than 23.
+We assert confidently that the facts about subjects being able to compute and about their being able to give the definition of odd number, etc., are the more important, highly entrenched, facts we want to preserve and explain\ldots{} we ourselves are prepared to give up the seeming fact that some odd numbers appear, as shown by their behavior in certain experimental paradigms, to be odder than others\ldots{}we do not give it up by saying that it was no fact; rather, by saying it must have been a fact about something other than the structure of concepts.
+\citep[284]{Armstrong1983}
+\end{quote}
+
+\begin{quote}
+\ldots{}we hold that \emph{fruit} and \emph{odd number} have different structures, and yet we obtain the same experimental outcome for both.
+But if the same result is achieved regardless of the concept struture, then the experimental design is not pertinent to the determination of concept structure.
+\citep[284--5]{Armstrong1983}
+\end{quote}
+
 At first blush, this would seem to argue against a naïve account of wellformedness judgements in which ill-formedness results when prosodic parsing fails (e.g., \citealt{Ito1989a}, \citealt{Noske1992}, \citealt{OT}).
 Consider the possibility that a nonce word is ill-formed if it cannot be syllabified without modification.
-There is a long history for the proposal that impossible words are those whose surface form cannot be parsed into prosodic structures like syllables without further phonological modification (e.g., \citealt[10f.]{Hooper1973}, \citealt[57f.]{Kahn1976}).
-There is no doubt that prosodic context is involved in determining where segments may occur.
-For instance, [ŋ] is a possible coda in English, but not a possible onset.
-Similarly, the [bn] of \emph{bnick} is judged to be admissible (though unattested) in medial position in English (e.g., \emph{sta}[b.n]\emph{ick}; \citealt[97]{Hooper1973}) but not as an onset.
-English is unable to sylalbify initial onset /bn/ without further modification, so one might expect underlying /bnɪk/ to be subject to deletion \citep[19f.]{Wolf2009}, prothesis, or anaptyxis.
-Indeed, \citet{Davidson2006b} finds that English speakers use all three of these repairs when asked to mimic foreign pronunciations of obstruent-nasal onsets impossible in English.\footnote{However, \citet{Davidson2005,Davidson2006a} finds acoustic and articulatory distinctions between the ``exscrescent'' schwa used to resolve non-native onset clusters, and the ``lexical'' schwa found in real words.
-For example, the anaptyctic [z${^\textrm{ə}}$k] produced by English speakers attempting to mimic word-initial [zk] more closely resembles the onset of \emph{scum} than the start of \emph{succumb}.
-This may undermine any link between non-native realizations and epenthesis of schwa in native words (e.g., in the syllabic allomorphs of the regular past tense and noun plural).}
+There is a long history for the proposal that impossible words are those whose surface form cannot be parsed into prosodic structures like syllables without further phonological modification (e.g.,
 
-However, prosodic well-formedness distinctions do not provide a ready explanation for the ``numerous other distinctions'' posited in \emph{SPE}.
-This is thought to rule out simple prosodic account of wordlikeness judgements.
+\begin{figure}
+\centering
+\includegraphics{density.pdf}
+\caption{Average ratings of individual nonce words, linearly transformed to fit the interval [0, 1]}
+\label{dsn}
+\end{figure}
 
-\begin{quote}
-A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements. \citep[371]{Shademan2006}
-\end{quote}
+\begin{table}
+\centering
+\begin{tabular}{l rrr}
+\toprule
+                        & subjects & items & trials \\
+\midrule
+\citealt{Albright2007}  & 68       & 40    & 2,720  \\
+\citealt{Albright2003a} & 24       & 87    & 2,064  \\
+\citealt{Scholes1966}   & 33       & 63    & 2,178  \\
+\midrule
+\textsc{Total}          & 125      & 187   & 6,962  \\
+\bottomrule
+\end{tabular}
+\caption{Subject and item counts}
+\label{counts}
+\end{table}
 
-Implicit in this claim is the assumption that the graded judgements obtained from wordlikeness tasks which permit gradient responses require a gradient model of well-formedness.
-However, it is argued here this is not pertinent to the question of whether wordlikeness is gradient or categorical, because graded judgements are a property of all graded tasks, even when applied to judgements of discrete categories.
+\begin{table}
+\centering
+\begin{tabular}{l rrrr}
+\toprule
+Pearson $r$               & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
+\midrule
+\citealt{Albright2007}    & \textbf{.725}    & {.703}          & {.463}          & {.695}           \\
+\citealt{Albright2003b}   & {.594}           & {.208}          & \textbf{.746}   & {.688}           \\
+\citealt{Scholes1966}     & {.803}           & {.534}          & {.737}          & \textbf{.834}    \\
+\midrule
+Spearman $\rho$           & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
+\midrule
+\citealt{Albright2007}    & \textbf{.819}    & {.661}          & {.338}          & {.612}           \\
+\citealt{Albright2003b}   & {.662}           & {.389}          & {.699}          & \textbf{.739}    \\
+\citealt{Scholes1966}     & {.769}           & {.575}          & {.794}          & \textbf{.817}    \\
+\midrule
+Kendall $\tau_{b}$        & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
+\midrule
+\citealt{Albright2007}    & {.666}           & \textbf{.683}   & {.245}          & {.450}           \\
+\citealt{Albright2003b}   & {.474}           & {.159}          & {.499}          & \textbf{.558}    \\
+\citealt{Scholes1966}     & {.619}           & {.484}          & {.614}          & \textbf{.671}    \\
+\midrule
+Goodman-Kruskal $\gamma$  & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
+\midrule
+\citealt{Albright2007}    & \textbf{.874}    & {.849}          & {.502}          & {.574}           \\
+\citealt{Albright2003b}   & \textbf{.929}    & {.606}          & {.246}          & {.488}           \\
+\citealt{Scholes1966}     & \textbf{.907}    & {.558}          & {.628}          & {.738}           \\
+\bottomrule
+\end{tabular}
+\caption{FIXME}
+\label{scores}
+\end{table}
 
-\citet[382]{Hayes2008a} ``consider the ability to model gradient intuitions to be an important criterion for evaluation phonotactic models''.
-However, they fail to ask consider whether categorical models provide a reasonable fit to gradient judgements.
-Rather than directly evaluating the categorical model of English phonotactics proposed by \citet{Clements1983}, \citeauthor{Hayes2008a} convert these constraints, many of which are exceptionless, into probabilitistic constraints.
-This suggests that \citeauthor{Hayes2008a} equate ``the ability to model gradient intuitions'' with the ability to make continuous predictions about well-formedness.
-But if gradient judgements are nothing more than a task effect, a model need not make gradient predictions to fit gradient judgements.
-
-Despite many claims that categorical models are empirically insufficient, there is not a single quantitative investigation of categorical wordlikeness models, and no study which compares categorical and gradient wordlikeness models on an equal footing.
-This chapter represents a first attempt to fill that lacuna.
-
-OUTLINE HERE
-
-\section{What wordlikeness might not be}
-
-\subsection{Gradience in wordlikeness judgements}
-
-\subsection{Gradience as a task effect}
-
-\section{Evaluation} 
-\label{2evaluation}
-
-\subsection{Method}
-
-\subsubsection{Materials and subjects}
-
-\subsubsection{Computational models}
-
-\subsection{Results}
-
-\subsection{Discussion}
-
-\section{Conclusions}
-
-%\citet{Armstrong1983} cut the Gordian knot by asserting the possibility of observing a false consequent; they find the existence of all-or-nothing concepts like ``odd number'' or (more controversially) ``female'' to be apparent.
-%
-%\begin{quote}
-%Are there definitional concepts?
-Of course.
-For example, consider the concept \emph{odd number}.
-This seems to have a clear definition, a precise description.
-[\ldots{}] No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd.
-No odd number seems odder than any other odd number.
-\citep[274]{Armstrong1983}
-%\end{quote}
-%
-%This proposition has the form of an indicative conditional.
-While the antecedent of this condition, the presence or absence of intermediate ratings in a graded judgement task, is readily observable, it has no bearing on the truth or falsity of the conditional proposition.
-However, whereas a false consequent (e.g., an all-or-nothing concept) would potentially falsify the proposition, the consequent, a mental state, cannot be directly observed.
-If this is the case, then both the proposition and the consequent (the proposition that wordlikeness is itself gradient) fail the test of falsification, placing it beyond the scope of scientific inquiry.
-%
-%\subsection{Gradience as a liability}
-%
-%One might reasonably ask whether 
-%representations 
-%
-%
-%\citeauthor{Armstrong1983} note that if subjects produce intermediate ratings for these categories, then the naïve proposition is false, and this is what \citeauthor{Armstrong1983} find in their experiments.
-They ask subjects to rate, on a seven-point scale, the extent to which, e.g., certain odd counting numbers represent the concept ``odd number'', and find that speakers do make use of intermediate values, as shown in Figure \ref{agg}.
-In summary, the proposition that intermediate ratings indicate mental gradience is either untestable, or it is false, and is thus rejected.
-%
-%\begin{figure} 
-%\centering
-%\includegraphics{agg.pdf}
-%\caption{Subjects asked to rate the degree to which certain even and odd numbers represent the categories ``even'' and ``odd'', respectively, freely use intermediate ratings.}
-%\label{agg} 
-%\end{figure}
-%
-%\citeauthor{Armstrong1983} do not view their results as evidence that subjects have an internal model of odd numbers which is at odds with the formal, all-or-nothing definition, and in fact, several additional experiments show that subjects deploy the extension of the formal definition in other tasks.
-Reviewing this study, \citet[215]{Schutze2011} writes that the experiments show that judgements are ``sensitive to factors other than our underlying competence''.
-In the case of wordlikeness, a case can be made that one such factor is similarity to existing words (e.g., \citealt{Bailey2001}, \citealt[][151, fn.
-27]{LSLT}, \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Schutze2005}, \citealt{Sendlmeier1987}, \citealt{Vitz1973}).
-However, the mere fact that subjects use intermediate ratings does not show that they do so with any systematicity, or that, e.g., the contrast between \emph{447} and \emph{7} with respect could be given any satisfying explanation.
-\citeauthor{Schutze2011} further suggests that gradient responses are the product of the task itself:
-%
-%\begin{quote}
-%Putting it another way, when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different.
-\citep[215]{Schutze2011}
-%\end{quote}
-%
-%This introduces a troubling possibility, that subjects ``oblige'' experimenters by introducing random noise to their categorical judgements when presented with a Likert task.
-Were this the case, the observed intermediate judgements are unlikely to yield information about speakers' knowledge of possible and impossible words.
-The alternative considered here is that there exists some model in which the gradience in ratings is systematic.
-%
-%The evaluation makes exclusive use of previously published English wordlikeness data.
-For inclusion in the evaluation, a study must conform to all three of the following conditions.
-First, the stimuli must consist only of monosyllabic words presented auditorily.
-Secondly, a significant portion of the stimuli must contain gross phonotactic violations (e.g., \emph{bnick}).
-Finally, the ratings, averaged across subjects, must be publicly available.
-The large-scale studies by \citet{Bailey2001} and \citet{Shademan2006,Shademan2007} are ineligible because the former lack stimuli with gross phonotactic violations and the latter data are not available to the public in any form.
-All data are reproduced in Appendix \ref{appendixA}.
-%
-%\subsubsection{\citealt{Scholes1966}}
-%
-%\citet{Scholes1966} conducted a number of English wordlikeness judgement with middle school children.
-The data used here come from his ``experiment 5'', in which 63 monosyllabic items were presented to 33 seventh-grade students.
-For each stimulus, the subjects produced forced choice ``yes''/``no'' answers to the question of whether the item ``is likely to be usable as a word of English''.
-\citet{Hayes2008a} and \citet{Albright2009a} analyze this data as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus \citep[see also][]{Pierrehumbert1994,Coleman1997,Frisch2000}.
-For instance, 22 of the 33 students answered ``yes'' for \emph{shlerk} [ʃlɚk], so it is assigned a score of $0.666$.\footnote{This procedure conflates intraspeaker variation, which in some cases is considerable \citep{Shademan2007} with speaker-internal gradience, and its adoption here should not be construed as an endorsement.
-Rather, it is provided solely for comparison with prior studies using the \citeauthor{Scholes1966} data.} These stimuli are all ostensibly non-words, but include \emph{clung} [klʌŋ], the preterite and past participle of the verb \emph{cling}, and \emph{brung} [brʌŋ], a dialectical past participle for \emph{bring}.
-These two words were excluded and the remaining 61 stimuli were submitted to analysis.
-%
-%\subsubsection{\citealt{Albright2003b}}
-%
-%\citet{Albright2003b} gathered wordlikeness judgements to serve as norms for a \emph{wug}-test.
-87 items were presented to 24 undergraduate subjects, who rated each word on a seven-point Likert scale.
-The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``completely normal, would make a fine English word''.
-%
-%\begin{table}
-%\centering
-%\begin{tabular}{l r r r}
-%\toprule
-%                        & subjects & items & trials \\
-%\midrule
-%\citealt{Albright2003a} & 24       & 87    & 2,088  \\
-%\citealt{Albright2007}  & 68       & 40    & 2,720  \\
-%\citealt{HayesInPress}  & 29       & 80    & 2,320  \\
-%\citealt{Scholes1966}   & 33       & 63    & 2,178  \\
-%\midrule
-%\textsc{Total}          & 154      & 270   & 9,306  \\
-%\bottomrule
-%\end{tabular}
-%\caption{Subject and item counts}
-%\label{counts}
-%\end{table}
-%
-%\subsection{Model comparison}
-%
-%The four models used here consist of a binary baseline and three computationally implemented gradient models.
-These three models are chosen because prior studies have shown they are correlated with wordlikeness judgements.
-Non-parametric rank correlation statistics are used evaluate the correlation between model predictions and wordlikeness ratings.
-Rank correlation are used rather than the parametric Pearson correlation statistic that is used by \citet{Hayes2008a}, for instance, because they make none of the potentially troublesome assumptions of the latter method \citep[see][23, fn.
-12]{Albright2009a}.
-%
-%The Spearman $\rho$ is most widely known rank correlation statistic, but it is difficult to give a natural interpretation to this quantity.
-On the other hand, the Kendall $\tau_b$ and Goodman-Kruskal $\gamma$ can be interpreted as fractions of the number of \emph{concordant} and \emph{discordant} pairs \citep{Noether1981}.
-Consider the case here, in which model scores are compared with wordlikeness ratings.
-If a model rates \emph{dresp} [dɹɛsp] more wordlike than *\emph{srest} [sɹɛst] and if speakers further rate \emph{dresp} more English-like than \emph{srest}, than \emph{dresp}/\emph{srest} are a concordant pair.
-If, however, a model and speakers disagree on the relative ranking of \emph{dresp} and \emph{srest}, the pair is discordant.
-The $\tau_b$ and $\gamma$ differ only in their treatment of ``ties'' (i.e., if e.g., *\emph{dresp} and *\emph{srest} are scored the same, or rated the same); $\tau_b$ includes a correction for ties, whereas $\gamma$ ignores tied pairs.
-Much like the familiar Pearson correlation, $\rho$, $\tau_b$ and $\gamma$ are all in the range [$-1$, $1$].
-Correlations of the four models are given in Table \ref{cor}.
-%
-%\subsubsection{Binary baseline}
-%
-%The binary baseline used here is a crude implementation of the null hypothesis that there are no gradient effects in wordlikeness judgements.
-To create such a baseline, it is necessary to distinguish between nonce words which contain a gross phonotactic violation and those which do not.
-As all stimuli here are monosyllables, this task can be further simplified by separately considering the two major subcomponents of the syllable, the onset and rime.
-Speakers are particularly adept as separating onsets from rimes \citep{Treiman1986,Treiman1995,Fowler1993}, and a large portion of phonotactic generalizations and violations can be localized to either onset or rime \citep[e.g.,][]{Borowsky1989,Fudge1969,Treiman1995,Kessler1997,Treiman2000}.
-%
-%The baseline considers a nonce syllable to be well-formed if it consists of both an attested onset and an attested rime; the free combination of these two components is the only mechanism by which this model can generalize beyond attested words.
-It is surely the case that this is an ultimately insufficient model: \citet{Albright2009a} observes, for instance, that [ɛsp] is a well-formed rime, even though it is found in no word of English.
-This problem is more acute in languages with more permissive syllable structures than those of English.
-For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.\footnote{
-%Chapter \ref{clusters} 
-%%\citet{Gorman2012c}
-%reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
-%
-%The rating densities from the three studies, linearly transformed to fit within the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{dsn}.
-In all three studies, it is possible to discern a relatively sharp separation between valid and invalid clusters, but also the presence of intermediate values.
-%
-%\begin{figure} 
-%\centering
-%\includegraphics{density.pdf}
-%\caption{Average ratings of individual nonce words, linearly transformed to fit the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
-%\label{dsn}
-%\end{figure}
-%
-%\begin{table} 
-%\centering
-%\begin{tabular}{l l l l l}
-%\toprule
-%Spearman $\rho$          & baseline         & maxent  & bigram           & density          \\
-%\midrule
-%\citealt{Greenberg1964}  & {0.845}          & {0.765} & {\textbf{0.863}} & {0.648}          \\
-%\citealt{Scholes1966}    & {0.791}          & {0.762} & {\textbf{0.827}} & {\textbf{0.827}} \\
-%\citealt{Albright2003b}  & {0.725}          & {0.429} & {0.708}          & {\textbf{0.742}} \\
-%\midrule
-%Kendall $\tau_b$         & baseline         & maxent  & bigram           & density          \\
-%\midrule
-%\citealt{Greenberg1964}  & {\textbf{0.716}} & {0.585} & {0.692}          & {0.462}          \\
-%\citealt{Scholes1966}    & {\textbf{0.664}} & {0.597} & {0.652}          & {0.565}          \\
-%\citealt{Albright2003b}  & {\textbf{0.599}} & {0.343} & {0.506}          & {0.556}          \\
-%\midrule
-%Goodman-Kruskal $\gamma$ & baseline         & maxent  & bigram           & density          \\
-%\midrule
-%\citealt{Greenberg1964}  & {\textbf{1.000}} & {0.684} & {0.692}          & {0.462}          \\
-%\citealt{Scholes1966}    & {\textbf{0.995}} & {0.634} & {0.667}          & {0.614}          \\
-%\citealt{Albright2003b}  & {\textbf{0.953}} & {0.656} & {0.509}          & {0.575}          \\
-%\bottomrule
-%\end{tabular}
-%\caption{Rank correlations between wordlikeness ratings and phonotactic models surprisingly reveal that the binary baseline meeds or exceeds the coverage of three state-of-the-art phonotactic models.
-All correlations are significant at $p = 0.05$.}
-%\label{cor}
-%\end{table}
-%
-%
 %\subsubsection{Maximum entropy phonotactics}
@@
 %
-%\begin{table} 
-%\centering
-%\begin{tabular}{l r r r r}
-%\toprule
-%                         & \multicolumn{1}{c}{featural bigrams} & \multicolumn{2}{c}{segmental bigrams}  \\
-%                         & maximum likelihood                   & maximum likelihood & Laplace smoothing \\
-%\midrule
-%%Pearson $\rho
-%Spearman $\rho$          & 0.638                                & 0.660              & \textbf{0.708}    \\
-%Kendall $\tau_b$         & 0.457                                & 0.467              & \textbf{0.506}    \\
-%Goodman-Kruskal $\gamma$ & 0.462                                & 0.473              & \textbf{0.509}    \\
-%\bottomrule
-%\end{tabular}
-%\caption{Segmental bigrams outperform ``featural bigrams'', and Laplace smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms.
-All correlations are significant at $p = 0.05$.}
-%\label{albrightimproved}
-%\end{table}
-%
 %\subsubsection{Neighborhood density} 
@@
 This shows quite starkly that these models fail to reliably predict intermediate ratings.
-%
-%\begin{table} 
-%\centering
-%\begin{tabular}{l r r r}
-%\toprule
-%Kendall $\tau_b$          & maxent         & bigram         & density  \\
-%\midrule
-%\citealt{Greenberg1964}   & 0.670          & \textbf{0.680} & 0.501 \\
-%\citealt{Scholes1966}     & \textbf{0.685} & 0.632          & 0.639 \\
-%\citealt{Albright2003b}   & 0.542          & 0.603          & \textbf{0.623} \\
-%\bottomrule
-%\end{tabular}
-%\caption{The binary baseline is strongly correlated with the three gradient model scores; all correlations are significant at $p = 0.05$.}
-%\label{bcor}
-%\end{table}
-%
-%\begin{table} 
-%\centering
-%\begin{tabular}{l r r r}
-%\toprule
-%$\Delta$ Spearman $\rho$          & maxent            & bigram            & density  \\
-%\midrule
-%\citealt{Greenberg1964}  & $-0.060$          &  \textbf{0.038} & $-0.017$ \\
-%\citealt{Scholes1966}    & $-0.029$          &  \textbf{0.047} & $-0.035$ \\
-%\citealt{Albright2003b}  & $-0.008$          & $-0.015$          & \textbf{0.018} \\
-%\midrule
-%$\Delta$ Kendall $\tau_b$         & maxent            & bigram            & density  \\
-%\midrule
-%\citealt{Greenberg1964}  & $-0.114$          & \textbf{$-$0.007} & $-0.084$ \\
-%\citealt{Scholes1966}    & $-0.067$          & \textbf{0.003}  & $-0.061$ \\
-%\citealt{Albright2003b}  & \textbf{$-$0.038} & $-0.092$          & $-0.049$ \\
-%\midrule
-%$\Delta$ Goodman-Kruskal $\gamma$ & maxent            & bigram            & density  \\
-%\midrule
-%\citealt{Greenberg1964}  & $-0.268$          & \textbf{$-$0.260} & $-0.337$ \\
-%\citealt{Scholes1966}    & $-0.361$          & \textbf{$-$0.313} & $-0.345$ \\
-%\citealt{Albright2003b}  & \textbf{$-$0.137} & $-0.443$          & $-0.386$ \\
-%\bottomrule
-%\end{tabular}
-%\caption{The change in rank correlation generated by augmenting the purely binary model with gradient predictions is small and in most cases it is negative.}
-%\label{controlled}
-%\end{table}
 %

### 2012-10-30 3.tex
Graphs and tables

@@
-\chapter{Static and derived phonotactic preferences in Turkish}
+\chapter{Static and derived Turkish phonotactics}
 \label{turkish}
@@
 First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules.
-Directional application provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}).
-Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
+Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{Anderson1974}: chap.~9, \citealt{Gorman2012e}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{Kenstowicz1977}:189f., \citeyear{Kenstowicz1979}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
 
@@
 Several other ``external'' facts suggest that this is not the case.
-The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.
+The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.\footnote{Thanks to Bert Vaux and Kie Zuraw for bringing these studies to my attention.}
 
-The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} Some speakers are said to be capable of pronouncing these non-native clusters, but in fast speech the cluster is split by anaptyxis.
+The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}.
+Some speakers are said to be capable of pronouncing these non-native clusters, but in fast speech the cluster is split by anaptyxis.
 In most cases, this vowel matches the following root vowel for backness.
@@
 
-Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} 
+Similar evidence comes from a language game discussed by \citet{Harrison2001}.
 The game is native to the related language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; it is not indigenous to Turkish, but can be taught quickly to children or adults.
@@
 However, a few trisyllabic verbs exceptionally show full reduplication.
-Exceptionality appears to be a lexical property (since it is specific to certain lexical entries), but it surfaces only in the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.
+This is a lexical property, but one that is only seen on the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.
 
@@
 \section{Evaluation}
+\label{3evaluation}
 
@@
 For the purposes below, `no preference' results are discarded.
-The second experiment used 32 native adults, none of whom appeared in the preceding study, and used a forced binary choice with some minor changes to the stimulus set.
+The second experiment used 32 native adults, none of whom appeared in the preceding study, and used a forced binary choice.
 

### 2012-10-30 4.tex
Graphs and tables

@@
-\chapter{Structural and accidental gaps} 
+\chapter{Structural and accidental phonotatic gaps}
 \label{gaps}
@@
 Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}),
+
+\begin{quote}
+It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
+\citep[283]{Anderson1974}
+\end{quote}
 
@@
 This is done in response to converging evidence that English speakers regard such words as complex.
+
+The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.
+
+In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.
+
+\begin{example}[Bound stem-specific allomorphy]
+\begin{tabular}{l l l l l l}
+a. & {adhere}   & {adhesion}   \\
+   & {cohere}   & {cohesion}   \\
+b. & {conceive} & {conception} \\
+   & {perceive} & {perception} \\
+\end{tabular}
+\end{example}
+
 This assumption has long been made to simplify phonological \citep{SPE}, allomorphic \citep[11f.]{Aronoff1976}, and syntactic \citep{Harley2009} generalizations.
+
 Results from lexical decision experiments also suggest that neoclassical compounds are complex.
@@
 Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect suggestive of a shared morphological identity.
+
+There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement.
+\citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.
+\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
+
+\begin{example}[Latinate verbs and small clauses] 
+\label{harley}
+\begin{tabular}{l l l l@{} l}
+a. & {show him the painting} & \alt{} & * & {exhibit him the painting} \\
+b. & {drink himself stupid}  & \alt{} & * & {imbibe himself stupid}    \\
+c. & {show it off}           & \alt{} & * & {exhibit it off}           \\
+\end{tabular}
+\end{example}
 
@@
 
-In summary, there is no statistical support for any of the three static constraints proposed by \citeauthor{Pierrehumbert1994}. However, it is premature to conclude that no static constraint exists.
-
-%It is possible that an explicit model of ``constraint discovery'', could identify reliable static constraints,
+There is no statistical support for any of the three static constraints proposed by \citeauthor{Pierrehumbert1994}. 
+%However, it is premature to conclude that no static constraint exists. It is possible that an explicit model of ``constraint discovery'', could identify reliable static constraints,
 
@@
 
-The allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and of the regular noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]) illustrate the well-known process of \textsc{Obstruent Voice Assimilation} (e.g., \emph{SPE}:178) in which the voicing specification of the first of two adjacent obstruents perserveres.
+\paragraph{Obstruent voice assimilation}
+
+Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), which take the voicing specification of a preceding obstruent.\footnote{Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.}
 
 \begin{example}[\textsc{Obstruent Voice Assimilation}]
-$\begin{bmatrix} -\alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\goesto{}~\begin{bmatrix} \alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\big /~\begin{bmatrix} \alpha~\textsc{Voiced} \\ -\textsc{Sonorant} \end{bmatrix}~\gap{}$
+\label{ovarule}
+$\begin{bmatrix} -\textsc{Son} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Voi} \end{bmatrix}~/~\gap~\begin{bmatrix} =\textsc{Voi} \\ -\textsc{Son} \end{bmatrix}$
 \end{example}
 
-Though it is not mentioned by \citet{Pierrehumbert1994} in her study of medial clusters, the vast majority of medial obstruent clusters are either uniformly voiced or voiceless, as in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}.
-``Disagreeing'' clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rare
+\noindent
+\citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice.
+However, the vast majority of medial obstruents clusters in simplex wors are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}.
+Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rare in comparison.
+
+\begin{table}
+\centering
+\begin{tabular}{l rrrr}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 35       & 329        & 10\%       & \multirow{2}{*}{0.002} \\
+violating  & 11       & 305        &  3\%       \\
+\bottomrule
+\end{tabular}
+\caption{FIXME}
+\label{ovatab}
+\end{table}
+
+\paragraph{Nasal place assimilation}
+\label{npa}
+
+\textsc{Nasal Place Assimilation} (\emph{SPE}:85) accounts for [n $\sim$ ŋ] allophony as well as
+
+is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant (though there are additional complexities; see \citealt{Borowsky1986}:?, \citealt{Halle1985a}, \citealt{Jensen2000}).
+
+\begin{example}[\emph{im-}/\emph{in-} allomorphy]
+\label{nparule}
+\begin{tabular}{l l l l l l l}
+a. & {polite}   & {i}[m.p]{olite}   & & {balance} & {i}[m.b]{alance} \\
+b. & {tangible} & {i}[n.t]{angible} & & {decent}  & {i}[n.d]{ecent}  \\
+\end{tabular}
+\end{example}
+
+\begin{example}[\textsc{Nasal Place Assimilation}]
+$\begin{bmatrix} $+$\textsc{Nas} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \end{bmatrix}~/~\gap{}~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ -\textsc{Son} \end{bmatrix}$
+\end{example}
+
+\begin{table}
+\centering
+\begin{tabular}{l rrrr}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.37\e{-07}} \\
+violating  & 11       & 22         & 33\%       \\
+\bottomrule
+\end{tabular}
+\caption{FIXME}
+\label{npatab}
+\end{table}
+
+\paragraph{Degemination}
+
+The final alternation found in English medial clusters is the simplification of geminates characteristic of ``level I'' morphology.
+Verb roots ending in /d/ which select the irregular /-t/ past tense suffix exhibit degemination (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).
+
+A final alternation targeting English syllable contact clusters is the simplification of geminates certain affixes.
+
+The deadjectival suffix \emph{-ly} does not result in a geminate \emph{l} when it attaches to
+
+krarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}).
+
+More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}).
+
+\begin{example}[\textsc{Degemination}]
+$\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ \ldots{} \end{bmatrix}~\goesto~\emptyset~\big /~\gap~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ \ldots{} \end{bmatrix}$
+\end{example}
+
+\begin{table}
+\centering
+\begin{tabular}{l rrrr}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 173      & 643        & 21\%       & \multirow{2}{*}{1.2\e{-10}} \\
+violating  & 0        & 104        & 0\%        \\
+\bottomrule
+\end{tabular}
+\caption{FIXME}
+\label{degemtab}
+\end{table}
 
@@
 This is consistent with converging evidence that speakers decompose these Latinate forms.
-The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.
-In a similar vein, \citet[11--13]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.
-%
-%\begin{example}[Bound stem-specific allomorphy]
-%\begin{tabular}{l l l l l l}
-%a. & {adhere}   & {adhesion}   \\
-%   & {cohere}   & {cohesion}   \\
-%b. & {conceive} & {conception} \\
-%   & {perceive} & {perception} \\
-%\end{tabular}
-%\end{example}
-%
-%There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement.
-\citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.
-\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
-%
-%\begin{example}[Latinate verbs and small clauses] 
-%\label{harley}
-%\begin{tabular}{ll@{}ll@{}l}
-%a. & * & {exhibit him the painting} & ~ & {show him the painting} \\
-%b. & * & {imbibe himself stupid}    & ~ & {drink himself stupid}  \\
-%c. & * & {exhibit it off}           & ~ & {show it off}           \\
-%\end{tabular}
-%\end{example}
-%
+
 %Decomposition is also supported by lexical decision results.
@@
 %
-%The [t \alt{} d] allomorphs of the regular past (e.g.
-\emph{nap}[t] $\sim$ \emph{nab}[d]), and the [s \alt{} z] allomorphs of the regular noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), Saxon genitive, and third person singular active indicative verb agreement, exemplify \textsc{Obstruent Voice Assimilation} (\emph{SPE}:178).\footnote{Underlying /-d, -z/ are argued for by \citet[282]{Hockett1958}, \citet[210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, \citet{Anderson1973a}, \citet[102]{Pinker1988}, and \citet[284f.]{Bakovic2005b}; alternatives are suggested by \citet[210f.]{LANGUAGE}, \citet[426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[135]{Borowsky1986}.}
 %
@@
 %%}
-%$\begin{bmatrix} -\textsc{Son} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Voi} \end{bmatrix}~/~\gap~\begin{bmatrix} =\textsc{Voi} \\ -\textsc{Son} \end{bmatrix}$
-%\end{example}
 %
@@
 %\subsubsection{Nasal place assimilation} 
-%\label{npa}
-%
-%In addition to [n $\sim$ ŋ] allophony, \textsc{Nasal Place Assimilation} (\emph{SPE}:85) is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant.
-%
-%\begin{example}[\emph{im-}/\emph{in-} allomorphy] 
-%\label{nparule}
-%\begin{tabular}{l l l l l l l}
-%a. & {polite}   & {i}[m.p]{olite}   & & {balance} & {i}[m.b]{alance} \\
-%b. & {tangible} & {i}[n.t]{angible} & & {decent}  & {i}[n.d]{ecent}  \\
-%\end{tabular}
-%\end{example}
-%
-%% mention Jensen here
-%
-%\begin{example}[\textsc{Nasal Place Assimilation}]
-%%\xymatrix@R=24pt@C=24pt{
-%%                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
-%%\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
-%%\txt{[$+$\textsc{Nasal}]} & \txt{[$-$\textsc{Sonorant}]}              \\
-%%}
-%$\begin{bmatrix} $+$\textsc{Nas} \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \end{bmatrix}~/~\gap{}~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ -\textsc{Son} \end{bmatrix}$
-%\end{example}
-%
 %The shape of the prefix before the velar stops /k, g/ is somewhat more complex.

### 2012-10-30 4.tex
more new graphs

@@
 \citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice.
-However, the vast majority of medial obstruents clusters in simplex wors are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}.
+As shown in Table \ref{ovatab} however, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}.
 Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rare in comparison.

### 2012-10-30 3.tex
writing on 4

@@
 A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone.
-However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.
+However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} The the underspecification analysis is rejected here on these grounds.
 

### 2012-10-30 4.tex
writing on 4

--- 2012-10-30 4.tex
@@
 The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.
-
-In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which is presented as evidence of prefix/bound stem decomposition.
+In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.
 
@@
 
-This assumption has long been made to simplify phonological \citep{SPE}, allomorphic \citep[11f.]{Aronoff1976}, and syntactic \citep{Harley2009} generalizations.
-
-Results from lexical decision experiments also suggest that neoclassical compounds are complex.
-\citet{Taft1986} and \citet{Taft1975,Taft1976} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure such as *\emph{refant}).
-Bound stems also show frequency effects which are independent of their whole word frequency \citep{Taft1979,Taft2006}.
-Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect suggestive of a shared morphological identity.
+\noindent
+\citeauthor{Aronoff1976} takes this to be evidence that \emph{adhere} and \emph{cohere}, for instance, share a bound stem.
 
@@
 \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.
-\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
+%\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
 
@@
 \end{example}
+
+Results from lexical decision experiments also suggest that neoclassical compounds are complex.
+\citet{Taft1975,Taft1976} and \citet{Taft1986} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure such as *\emph{refant}).
+Bound stems also show frequency effects which are independent of their whole word frequency \citep{Taft1979,Taft2006}.
+Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect thought to indicate of morphological relatedness.
 
@@
 
-There is no statistical support for any of the three static constraints proposed by \citeauthor{Pierrehumbert1994}. 
-%However, it is premature to conclude that no static constraint exists. It is possible that an explicit model of ``constraint discovery'', could identify reliable static constraints,
+So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters.
+These constraints are ``static'' in the sense that they have no analogues in the morphophonemics of English.
+The investigation considers clusters of all sizes, not merely the triconsontal clusters studied by \citeauthor{Pierrehumbert1994}, as she gives no justification for excluding biconsontal or quadraconsonantal clusters.
+
+\paragraph{Dorsal-labial clusters} \citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S?).
+Two-consonant velar-labial clusters are found in words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} occur.
+Table \ref{dltab} shows that velar-labial clusters are somewhat less likely to occur than  are velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance.
+
+\begin{table}
+\centering
+\begin{tabular}{l rrrr}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 25       & 91         & 22\%       & \multirow{2}{*}{0.106} \\
+violating  &  4       & 40         &  9\%       \\
+\bottomrule
+\end{tabular}
+\caption{FIXME}
+\label{dltab}
+\end{table}
+
+\paragraph{Coronal codas} \citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}.
+Table \ref{cctab} shows that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}).
+Restricting attention to triconsonantal clusters does not produce a significant lexical effect, either ($p = 0.129$).
+
+\begin{table}
+\centering
+\begin{tabular}{l rrrr}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 56       & 304        & 15\%       & \multirow{2}{*}{0.430} \\
+violating  & 37       & 243        & 13\%       \\
+\bottomrule
+\end{tabular}
+\caption{FIXME}
+\label{cctab}
+\end{table}
+
+\noindent
+
+\paragraph{ABA clusters} \citet[][176]{Pierrehumbert1994} notes the ``lack of clusters with identical first and third elements'', ignoring presence or absence of voicing.
+Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters (Table \ref{abatab}).
+
+\begin{table}
+\centering
+\begin{tabular}{l rrrr}
+\toprule
+           & attested & unattested & saturation & $p$-value \\
+\midrule
+conforming & 47       & 512        &  8\%       & \multirow{2}{*}{0.250} \\
+violating  &  0       &  25        &  0\%       \\
+\bottomrule
+\end{tabular}
+\caption{FIXME}
+\label{abatab}
+\end{table}
+
+\paragraph{Summary} There is no statistical support for \citeauthor{Pierrehumbert1994}'s static constraints.
 
 \subsubsection{Derived constraints}
+
+FIXME
 
@@
 \citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice.
-However, the vast majority of medial obstruents clusters in simplex wors are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}.
-Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rare in comparison.
+As shown in Table \ref{ovatab} however, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}, and hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rarer than would be expected from chance.
 
@@
 
-\textsc{Nasal Place Assimilation} (\emph{SPE}:85) accounts for [n $\sim$ ŋ] allophony as well as
-
-is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix, determined by the place of the following consonant (though there are additional complexities; see \citealt{Borowsky1986}:?, \citealt{Halle1985a}, \citealt{Jensen2000}).
+\textsc{Nasal Place Assimilation} (\emph{SPE}:85) accounts for [n $\sim$ ŋ] allophony and Latinate prefix allomorphy, such as that of the ``negative'' refix \emph{im-}/\emph{in-}.\footnote{\citealt{Borowsky1986}:FIXME and \citealt{Jensen2000} propose additional restrictions on \textsc{Nasal Place Assimilation}.}
 
@@
 
+%Furthermore, \citet[228]{Myers1993} reports speech errors which create nasal-obstruent clusters and undergo \textsc{Nasal Place Assimilation}; e.g., \emph{ra}[nd] \emph{orker}, intended \emph{ra}[ŋk] \emph{order}.
+
+\noindent
+The rule is formalized below.
+
 \begin{example}[\textsc{Nasal Place Assimilation}]
@@
 \end{example}
+
+As shown in Table \ref{npatab}, virtually all (94\%) clusters of a nasal coda followed by a homorganic obstruent (e.g., \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, \emph{mo}[ŋ.k]\emph{ey}) are attested.
+As \citet[175]{Pierrehumbert1994} observes, heterorganic clusters, like those \emph{pli}[m.s]\emph{oll} or \emph{scri}[m.ʃ]\emph{aw}, do occur, but they are significantly more rare.
 
@@
 
-\paragraph{Degemination}
-
-The final alternation found in English medial clusters is the simplification of geminates characteristic of ``level I'' morphology.
-Verb roots ending in /d/ which select the irregular /-t/ past tense suffix exhibit degemination (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).
-
-A final alternation targeting English syllable contact clusters is the simplification of geminates certain affixes.
-
-The deadjectival suffix \emph{-ly} does not result in a geminate \emph{l} when it attaches to
-
-krarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}).
-
-More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}).
+\paragraph{Degemination} The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.
+\emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}).
+\textsc{Degemination} is formalized here as a rule deleting the first of two identical segments, ignoring voice.
 
@@
 \end{example}
+
+\noindent
+As can be seen from Table \ref{degemtab}, the sample includes no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance.
 
@@
 \end{table}
+
+\paragraph{Summary} All three of \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters.
 

### 2012-11-05 1.tex
Writing on chapter 2 primarily, also added the two "MSes" and new graphs for chapter 2.

@@
-\chapter{A generative theory of phonotactic knowledge and acquisition}
 \label{intro}
+
+I would like to suggest that there are two types of problems that arise in developing a theory of phonotactics free of duplication.
+The first type of problem consists of conflicts between theoretical assumptions and the desire to eliminate duplication.
+If we continue to view duplication as a sort of negative heuristic, then it may be the case that the theoretical assumptions are wrong.
+Such a case arises in Chapter \ref{turkish} in the discussion of archiphonemic underspecification analysis of Turkish vowel harmony proposed by \citet{Clements1982}.
+\citeauthor{Clemenst1982} propose that all but the first vowel of a harmonic root is underspecified for backness (and in high vowels, roundness) and is filled in by rule.
+Since there are disharmonic roots, this rule must be ``structure-filling''.
+Consequently, this rule cannot account for the apparent markedness of disharmonic roots revealed by wordlikeness judgements (among other psycholinguistic tasks).
+If duplication is a pathology, this analysis is wrong.
+
+Another type of problem is the ``conundrum'', an apparent fact which poses a problem for virtually any generative theorists.
+
+pure allophony
+
+There are a few contexts which this interacts:
+
+- prosody
+- exceptionality
+- wordlikeness
 
@@
 
+Nothing will be said about what \citet{Stanley1967} calls \emph{segment structure rules}, or language-specific constraints on the inventory of (archi)phonemes, since there is a great deal of overlap between competing theories.
+Consider possible explanations for the apparent absense of ejectives in English, It might be that English does, in a certain sense, permit underlying ejectives, but such segments are uniformly realized as plain (i.e., pulmoni) voiceless stops: in other words, the inventory is an epiphenomenon of phonological processes.
+Or, perhaps English lacks the features needed to contrast ejectives with plain voiceless stops, in which case the inventory is an epiphenomenon of the system of contrast.
+Or, perhaps the absense of ejectives follows from no particular fact, and has the grammatical status of an accidental gap.
+Further assumptions are needed to distinguish these analyses.
+
 \subsubsection{Underlying representations}
+
+The term \emph{sequence structure constraint} is now somewhat imprecise given the apparent obsolescence of boundary segments: the ``sequence structure onstraint'' that rules out onset /bn/ in English must actually be stated as a constraint militating against /\#bn/.
+\citep[149]{Kenstowicz1977}
 

### 2012-11-05 2.tex
more new graphs
writing on 4
Writing on chapter 2 primarily, also added the two "MSes" and new graphs for chapter 2.

@@
-\chapter{Categorical and gradient aspects of wordlikeness judgements} 
 \label{gradience}
 
-Two nonce words, \emph{blick} [blɪk] and \emph{bnick} [bnɪk], underlie the claim that speakers can rapidly distinguish between possible and impossible words \citet{Halle1962}.
-In \emph{SPE}, \citet{SPE} use a third word, \emph{bnzk}, to argue that nonce words fall on a cline of well-formedness.
-Neither \emph{bnick} nor \emph{bnzk} are possible words of English, they are possible words in other languages: stop-nasal onsets are found in Moroccan Arabic (e.g., \emph{bniqa} `closet') and stop-nasal-fricative-stop words in Imdlawn Tashlhiyt Berber \citep{Dell1985}.
-However, there is some sense in which \emph{bnzk} is even less English-like than \emph{bnick}.
-Of this, \citeauthor{SPE} write:
-
-\begin{quote}
-Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
-\end{quote}
-
-\noindent
-The position that the well-formedness of nonce words is consistent the view of syntactic grammaticality taken in \emph{LSLT} and \emph{Aspects} \citep{LSLT,ASPECTS}, where it is claimed that different syntactic violations result in different degrees of ungrammaticality.
-
-Most recent discussions of the ``problem of admissibility'' focus on this cline of wellformedness as it is evidenced in wordlikeness tasks.
-
-\begin{quote}
-When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing.
+FIXME
+%\begin{quote}
+%Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
+%\end{quote}
+%
+%\noindent
+%The position that the well-formedness of nonce words is consistent the view of syntactic grammaticality taken in \emph{LSLT} and \emph{Aspects} \citep{LSLT,ASPECTS}, where it is claimed that different syntactic violations result in different degrees of ungrammaticality.
+%
+%Most recent discussions of the ``problem of admissibility'' focus on this cline of wellformedness as it is evidenced in wordlikeness tasks.
+%
+%\begin{quote}
+%When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing.
 In the usual case, novel items fall along a gradient cline of acceptability.
 \citep[][9]{Albright2009a}
+%
+%In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
+%\citep[][382]{Hayes2008a}
+%
+%\ldots{}when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
+%\end{quote}
+%
+%\begin{quote}
+%\ldots{}A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
+\citep[371]{Shademan2006}
+%\end{quote}
 
-In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
-\citep[][382]{Hayes2008a}
-
-\ldots{}when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
-\end{quote}
-
-\begin{quote}
-\ldots{}A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
-\citep[371]{Shademan2006}
-\end{quote}
-
-\citet[382]{Hayes2008a} ``consider the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models''.
-
-\begin{quote}
-No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
-\end{quote}
-
-\begin{quote}
-Some have responded to these findings very consistently, by asserting that the experimental findings are to be interpreted as before: that, psychologically speaking, odd numbers as well as birds and vegetables are graded concepts\ldots{} We reject this conclusion just because we could not explain how a person could compute with integers who believed that 7 was odder than 23.
-We assert confidently that the facts about subjects being able to compute and about their being able to give the definition of odd number, etc., are the more important, highly entrenched, facts we want to preserve and explain\ldots{} we ourselves are prepared to give up the seeming fact that some odd numbers appear, as shown by their behavior in certain experimental paradigms, to be odder than others\ldots{}we do not give it up by saying that it was no fact; rather, by saying it must have been a fact about something other than the structure of concepts.
-\citep[284]{Armstrong1983}
-\end{quote}
-
-\begin{quote}
-\ldots{}we hold that \emph{fruit} and \emph{odd number} have different structures, and yet we obtain the same experimental outcome for both.
-But if the same result is achieved regardless of the concept struture, then the experimental design is not pertinent to the determination of concept structure.
-\citep[284--5]{Armstrong1983}
-\end{quote}
-
-At first blush, this would seem to argue against a naïve account of wellformedness judgements in which ill-formedness results when prosodic parsing fails (e.g., \citealt{Ito1989a}, \citealt{Noske1992}, \citealt{OT}).
-Consider the possibility that a nonce word is ill-formed if it cannot be syllabified without modification.
-There is a long history for the proposal that impossible words are those whose surface form cannot be parsed into prosodic structures like syllables without further phonological modification (e.g.,
+\section{What some wellformedness judgements might not be}
 
@@
 \centering
-\includegraphics{density.pdf}
-\caption{Average ratings of individual nonce words, linearly transformed to fit the interval [0, 1]}
-\label{dsn}
+\includegraphics{agg.pdf}
+\caption{Subjects freely use intermediate ratings when asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively \citep{Armstrong1983}}
+\label{agg}
 \end{figure}
+
+%The gradedness of a category cannot be observed.
+Scientists choose the granularity at which they observe.
+For example, a linguist may choose to record speakers' judgements on a 7-point scale, for example, or to use a forced binary choice between two minimally different items.
+The level of granularity at which the observations are made have consequences for how statistical inference over these observations should be performed, or for the theory of the \emph{task model}, the mechanism by which speakers perform the task, but it does not have any consequences for the \emph{theory} of phnotactics.
+
+%\begin{quote}
+%No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
+%\end{quote}
+
+%\begin{quote}
+%Some have responded to these findings very consistently, by asserting that the experimental findings are to be interpreted as before: that, psychologically speaking, odd numbers as well as birds and vegetables are graded concepts\ldots{} We reject this conclusion just because we could not explain how a person could compute with integers who believed that 7 was odder than 23.
+We assert confidently that the facts about subjects being able to compute and about their being able to give the definition of odd number, etc., are the more important, highly entrenched, facts we want to preserve and explain\ldots{} we ourselves are prepared to give up the seeming fact that some odd numbers appear, as shown by their behavior in certain experimental paradigms, to be odder than others\ldots{}we do not give it up by saying that it was no fact; rather, by saying it must have been a fact about something other than the structure of concepts.
+\citep[284]{Armstrong1983}
+%\end{quote}
+
+%\begin{quote}
+%\ldots{}we hold that \emph{fruit} and \emph{odd number} have different structures, and yet we obtain the same experimental outcome for both.
+But if the same result is achieved regardless of the concept struture, then the experimental design is not pertinent to the determination of concept structure.
+\citep[284--5]{Armstrong1983}
+%\end{quote}
+
+\section{Evaluation}
+\label{2evaluation}
+
+There have been no serious attempts to evaluate categorical and gradient models of wordlikeness on an equal footing.
+This study represents a first attempt to fill this gap.
+It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared.
+For instance, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, transform the \citeauthor{Clements1983} constraints, many of which are exceptionless, into probabilities.
+While this is consistent with their claim that ``the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, such a principle precludes any attempt to confirm or falsify the hypothesis that underlies it.
+
+It is unknown whether the intermediate ratings in gradient wordlikeness tasks are reliably predicted by the computational models that have been proposed.
+If some model poorly accounts for these intermediate ratings, it may be the case that the model is improperly specified, or it may indicate that a considerable amount of variance in wordlikeness ratings is not a product of the phonotactic system per se.
+
+Imagine, alternatively, that wordlikeness judgements could be effectively modeled with a gross contrast between possible and impossible words.
+
+%[st] [bl]
+
+\subsection{Materials}
+
+This evaluation uses a large sample of three previously published studies on English wordlikeness comprising 125 subjects and 187 items.
+Two criteria were used to select these three studies.
+First, it was required that the stimuli be presented aurally so as to eliminate any possibility of orthographic effects.
+Secondly, the data must be sufficiently ``phonotactically diverse'': that is, it must include both items like \emph{blick} and \emph{bnick}.
+This excludes studies like that of \citet{Bailey2001}, in which few if any items contain gross phonotactic violations of the type represented by \emph{bnick}.
+The data is summarized in Table \ref{counts}.
 
@@
 \midrule
-\citealt{Albright2007}  & 68       & 40    & 2,720  \\
-\citealt{Albright2003a} & 24       & 87    & 2,064  \\
-\citealt{Scholes1966}   & 33       & 63    & 2,178  \\
+\citeauthor{Albright2007}  & 68       & 40    & 2,720  \\
+\citeauthor{Albright2003b} & 24       & 86    & 2,064  \\
+\citeauthor{Scholes1966}   & 33       & 63    & 2,178  \\
 \midrule
@@
 
+\subsubsection{\citealt{Albright2007}}
+
+\citet{Albright2007} administers a wordlikeness task in which 68 adult speakers rate 40 monosyllabic nonce words, presented aurally, on a 7-point Likert scale with endpoints labeled  ``completely impossible as an English word'' and ``would make a fine English word''.
+\citealt{Albright2007}'s study study is primarily concerned with the effects of different onset types (e.g., well-formed /bl/, marginal /bw/, unattested /bn, bd, bz/), and there is less diversity among the choice of rimes, none of which are obviously ill-formed.
+
+\subsubsection{\citealt{Albright2003b} (norming experiment)}
+
+\citet{Albright2003b} have 24 adult speakers rate 87 aurally presented monosyllabic nonce words on a 7-point Likert scale with endpoints labeled ``completely bizarre, impossible as an English word'' and ``completely normal, would make a fine English word''.
+This task was administered to establish phonotactic norms for a later nonce word inflection task.
+Their item [raɪf] is excluded in this study, since this is an actual word of English, \emph{rife}.
+\citet{Albright2009a} uses this data to evaluate several computational models of wordlikeness.
+
+\subsubsection{\citealt{Scholes1966} (experiment 5)}
+
+\citet{Scholes1966} conducts several wordlikeness tasks with 7th-grade students (approximately 12--13 years of age).
+The data used here is his experiment 5, in which 33 speakers provide a ``yes'' or ``no'' as to whether each of the 63 items, presented aurally, are ``likely to be usable as a word of English''.
+Like the study by \citet{Albright2007}, the focus was on onset well-formedness and there is minimal diversity in rime type.
+Two items, [klʌŋ] \emph{clung} and [bɹʌŋ] \emph{brung} (a dialectical past participle of \emph{bring}), are excluded here as actual words of English.
+\citet{Albright2009a} and \citet{Hayes2008a} also use this data for the purposes of model evaluation; following \citet{Frisch2000}, they use the proportion of ``yes'' responses for each item so as to derive a continuous measure of well-formedness.
+
+\subsection{Method}
+
+Models are evaluated by comparing their scores to the average rating of each word using four correlation statistics.
+\citet{Hayes2008a} evaluate their model using the Pearson $r$, a parametric measure of correlation.
+\citet{Stevens1946} argues that statistics of this type are inappropriate for analysis of Likert scale data, like those used by \citet{Albright2009a} and \citet{Albright2003b}.
+The reason is that this makes a \emph{linearity assumption}.
+That is, it assumes that nonce words rated ``1'' and  ``3'', for instance, are just as different as those ``4'' and a ``6''.
+A weaker assumption, more appropriate for Likert scale data, is the ``monotonicity assumption'': that ``1'' is less English-like than ``3'', which is less English-like than ``4'', and so on.
+However, it also has been claimed that $r$ is particularly robust to violations of the linearity assumption \citep[e.g.,][]{Havlicek1976}.
+Pearson $r$ is reported here, but no stance is taken on its appropriateness for this data.
+%\citet{Jamieson2004}
+\citeauthor{Hayes2008a} also report Spearman $\rho$; this statistic requires only the weaker assumption of monotonicity, but it is difficult to give a simple interpretation to the coefficient.
+Much easier to interpret are two non-parametric statistics, the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ \citep{Noether1981}; a variant of the latter is used by \citet{Albright2009a}.
+These statistics are computed by comparing every model score/wordlikeness rating pair to every other: a comparison is counted as \emph{concordant} if the greater of the two model scores is the one associated with the greater of the two wordlikeness ratings (that is, the model ranks these two nonce words the in accordance with speakers' ratings), and as \emph{discordant} otherwise.
+$\gamma$ and $\tau_b$ differ in the treatment of ``ties'', pairs where either the model score or wordlikeness rating are identical.
+For $\gamma$, ties are ignored, and the coefficient is
+
+\begin{unlabeledexample}
+$\gamma = \displaystyle\frac{c - d}{c + d}$
+\end{unlabeledexample}
+
+\noindent
+where $c$ and $d$ represent the number of concordant and discordant pairs, respectively.
+The $\tau_b$ statistic uses a similar formula, but also incorporates a penalty for ties in model score which are not also paired with ties in wordlikeness ratings, or vis versa.
+
+\subsection{Models}
+
+The nonce word stimuli from these three studies are scored automatically using four computational models.
+The first two models represent baselines for comparison to the latter two state-of-the-art gradient models.
+The scores for each model are provided in Appendix \ref{ratings}.
+
+\subsubsection{Gross phonotactic violation}
+
+A simple baseline is constructed by separating nonce words into those which contain a phonotactic violation and those which do not.
+As all nonce words here are monosyllabic, this task can be localized to two subcomponents of the syllable, the onset and the rime.
+This is not to imply that these are the only domains over which phonotactic violations might be stated, but there are prior claims that onset and rime are particularly important domains for stating phonotactic constraints (e.g., \citealt{Fudge1969}, \citealt{Kessler1997}, \citealt{Treiman2000}).
+Furthermore, these units are implicated by language games \citep{Treiman1983} and speech errors \citep{Fowler1993}, and adult speakers are adept at separating syllables into these units \citep{Treiman1986,Treiman1995}.
+
+Operationalizing ``phonotactic violation'' is somewhat more difficult.
+The simplest possible mechanism is chosen here: an onset or rime is identified as well-formed if it occurs with non-zero frequency in a representative sample, and is identified as ill-formed otherwise.
+The sample is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms, the latter though to be particualrly strongly correlated with behavioral measures \citep{Brysbaert2009}.
+These pronunciations are then syllabified, and individual syllables parsed into onset and rime, according to a process described in detail in Appendix \ref{syllabification}.
+This is not is not to imply that all unattested onsets or rimes should be regarded as ill-formed, or that all onsets or rimes with non-zero frequency in this data are well-formed.
+For instance, \citet{Albright2009a} judges [drEsp] to be an unobjectionable possible word of English, despite the total absense of [Esp] rimes in English, and similar observations have been made regarding English onsets \citep[e.g.,][]{Cairns1972,Moreton2002}.
+
+In Figure \ref{boxplot}, wordlikeness ratings from the three studies are plotted according to this gross contrast.
+While there are a considerable number of outliers, there can be little doubt that the contrast is reflected in wordlikeness judgements in all three studies.
+
+\begin{figure}
+\centering
+\includegraphics{boxplot.pdf}
+\caption{Gross phonotactic status and item-averaged wordlikeness ratings}
+\label{boxplot}
+\end{figure}
+
+\subsubsection{Lexical neighborhood density}
+
+A second baseline is provided by measures of similarity to existing English words items, which has long been applied in the study of wordlikeness (e.g., \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Shademan2006,Shademan2007}, \citealt{Vitevitch1998,Vitevitch1999a}).
+\citeauthor{LSLT} (\citeyear{LSLT}: 151, fn.~27) suggests that grammaticality judgements in general might be influenced by similarity to existing grammatical structures.
+\citet[417f.]{SPE} outline a similarity-based wordlikeness model.
+More recently, it has been observed (e.g., \citealt[51]{Coleman1997}, \citealt{Hay2004a}) that nonce words which flagrantly violate English sonority restrictions but which bear common affixes (e.g., *\emph{mrupation}) are rated highly English-like.
+
+A wide variety of lexical similarity measures were considered, including a variant of including a variant of the Generalized Neighborhood model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a set of measures provided by the Irvine Phonotactic Calculator \citep{Vaden2009}.
+The most reliable measure is also the most veneraable measure of lexical similarity: Coltheart's $N$ \citep{Coltheart1977}, which is defined as the number of words in some representative sample which can be changed into a target nonce word by a single insertion, deletion, or substitution of a phone.
+\citet{Greenberg1964} find a correlation between wordlikeness ratings and a variant of this measure which only counts words differing by a single substitution.
+This measure is plotted against ratings from the three studies in Figure \ref{neighborhood}, where it can be seen that it accounts for much of the variance in ratings.
+
+\begin{figure}
+\centering
+\includegraphics{neighborhood.pdf}
+\caption{Correlation between neighborhood density and item-averaged wordlikeness ratings}
+\label{neighborhood}
+\end{figure}
+
+While there is nothing inherently ``phonotactic'' about Coltheart's $N$, it indirectly incorporates much of the information present in the gross phonotactic baseline.
+Consider [blIk]: since there is nothing marked about any part of this nonce word, a ``neighbor'' might be found by modifying any phone: e.g., \emph{click}, \emph{brick}, \emph{bloke}, \emph{bliss}.
+However, since [bn] onsets are unattested in English, a neighbor of [bnIk] must somehow modify this cluster: this leaves only \emph{brick} and \emph{nick}.
+\citet{Frauenfelder1993} find that neighborhood density is also strongly correlated with measures like bigram probability.
+
+\subsubsection{Segmental bigram probability}
+
+Bigrams probabilities have been applied to model results of a number of psycholinguistic tasks \citep[e.g.,][]{Pitt1998,Vitevitch2005}, and \citet{Albright2009a} applies them as a model of wordlikeness judgements.
+The bigram probability of a sequence $ijk$, for instance, is:
+
+\begin{unlabeledexample}
+$\displaystyle \hat{p}(ijk) = p(i|\textrm{start}) \cdot p(j|i) \cdot p(k|j) \cdot p(\textrm{stop}|k)$
+\end{unlabeledexample}
+
+\noindent
+That is, it is the product of sequence-initial $i$, the probability of $j$ following $i$, the probability of $k$ following $j$, and the probability of the sequence ending after $k$.
+
+\citet{Albright2009a} compares two variants of this model, the first operating over segments, the second over sets of features.
+Unfortunately, the latter model is not described in sufficient detail to allow it to be implemented directly, and there is no publicly available implementation.
+However, \citeauthor{Albright2009a}'s evaluation, which includes the \citet{Scholes1966} and \citet{Albright2003b} data, finds an advantage for segmental bigrams.
+In implementing this model, it was found that a slight improvement could be made by preventing any phone-to-phone transition from have zero probability.
+This is accomplished by adding 1 to the count of every transition, a technique known as Laplace, or ``add one'', smoothing.
+As can be seen in Table \ref{bigramcomparison}, this results in a slight increase in the correlation between the scores from this model and wellformedness ratings.
+This smoothed segmental bigram score is adopted below.
+In Figure \ref{bigram}, it is plotted against wordlikeness ratings from the three studies.
+
+\begin{table}
+\centering
+\begin{tabular}{l rrr}
+\toprule
+                         & features & segments & smoothing \\
+\midrule
+Pearson $r$              & {.711}   & {.742}   & \uline{.746} \\
+Spearman $\rho$          & {.628}   & {.667}   & \uline{.699} \\
+Goodman-Kruskal $\gamma$ & {.453}   & {.479}   & \uline{.502} \\
+Kendall $\tau_{b}$       & {.448}   & {.473}   & \uline{.499} \\
+\bottomrule
+\end{tabular}
+\caption{Correlation between item-averaged wordlikeness ratings for the \citet{Albright2003b} norming study and three variants of bigram probability; ``features'': featural bigram probability reported by \citet{Albright2009a}, ``segments'': segment bigram probability reported by \citet{Albright2009a}, ``smoothing'': segment bigram probabilities estimated using Laplace smoothing}
+\label{bigramcomparison}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{bigram.pdf}
+\caption{Correlation between segmental bigram score and item-averaged wordlikeness ratings}
+\label{bigram}
+\end{figure}
+
+\subsubsection{Maximum entropy phonotactics}
+
+\citet{Hayes2008a} propose a sophisticated model in which well-formedness is related to a probability distribution over phone sequences, estimated according to the principle of maximum entropy, or ``MaxEnt'' \citep[e.g.,][]{Goldwater2003,Jager2007}.
+\citeauthor{Hayes2008a} use a complex method to evaluate their model.
+First, they extract onset sequences from the CMU pronunciation dictionary which they use to train the model.
+The model is then used to score the onsets of the \citet{Scholes1966} nonce words.
+Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting corelation.\footnote{This is contrary to standard practices in natural language process, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter), and thus there is no indication that the model will generalize to new data.
+Consequently, no transformation is used here.
+This only has an effect on the Pearson $r$ coefficient, since they use a transformation that preserves monotonicity.}
+
+An attempt was made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications were all used.
+However, \citet{Albright2009a} finds that the maximum entropy model, training and testing only on onsets, does not generalize to the \citet{Albright2003b} data, presumably because of the considerable diversity of rimes in the latter data.
+Consequently, the model was trained to score whole words, not just onsets, using the subset of the CMU dictionary described above.
+Following the procedure of \citet{HayesInPress}, dictionary entries were syllabified and the resulting syllables were parsed into onset, nucleus, and coda, and a feature [$\pm$\textsc{Coda}] was added to contrast onset and coda consonants.
+Since the maximum entropy model, producing slightly different predictions on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}.
+The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to large variety of nonce words, though many of these words appear to have a low rating.
+This suggests that the model is not robust enough to score whole words.
+
+\begin{figure}
+\centering
+\includegraphics{maxent.pdf}
+\caption{Correlation between MaxEnt score and item-averaged wordlikeness ratings}
+\label{maxent}
+\end{figure}
+
+\subsection{Results}
+
+Table \ref{scores} displays the full set of correlation coefficients, for each of the three data sets, and for each of the four models.
+For each pair of statistic and data set, the highest coefficient is underlined.
+The first observation is that in general, there is a positive correlation between model score and ratings in each pair.
+However, the two baselines, gross phonotactic status and neighborhood density, are by far the strongest models across statistics and and data, with gross phonotactic status performing the strongest under the Goodman-Kruskal $\gamma$ and on the \citet{Albright2007} data, and neighborhood density performing strongly under nearly all other statistics and data sets.
+
 \begin{table}
@@
 \toprule
-Pearson $r$               & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
+                           & gross status & density ($N$) & bigram score & MaxEnt score \\
 \midrule
-\citealt{Albright2007}    & \textbf{.725}    & {.703}          & {.463}          & {.695}           \\
-\citealt{Albright2003b}   & {.594}           & {.208}          & \textbf{.746}   & {.688}           \\
-\citealt{Scholes1966}     & {.803}           & {.534}          & {.737}          & \textbf{.834}    \\
+Pearson $r$               \\
 \midrule
-Spearman $\rho$           & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
+\citeauthor{Albright2007}  & \uline{.725} & {.673}        & {.463}       & {.703} \\
+\citeauthor{Albright2003b} & {.594}       & \uline{.793}  & {.746}       & {.208} \\
+\citeauthor{Scholes1966}   & {.803}       & \uline{.855}  & {.737}       & {.534} \\
 \midrule
-\citealt{Albright2007}    & \textbf{.819}    & {.661}          & {.338}          & {.612}           \\
-\citealt{Albright2003b}   & {.662}           & {.389}          & {.699}          & \textbf{.739}    \\
-\citealt{Scholes1966}     & {.769}           & {.575}          & {.794}          & \textbf{.817}    \\
+Spearman $\rho$           \\
 \midrule
-Kendall $\tau_{b}$        & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
+
+\citeauthor{Albright2007}  & \uline{.819} & {.612}        & {.338}       & {.661} \\
+\citeauthor{Albright2003b} & {.662}       & \uline{.739}  & {.699}       & {.389} \\
+\citeauthor{Scholes1966}   & {.803}       & \uline{.817}  & {.794}       & {.575} \\
 \midrule
-\citealt{Albright2007}    & {.666}           & \textbf{.683}   & {.245}          & {.450}           \\
-\citealt{Albright2003b}   & {.474}           & {.159}          & {.499}          & \textbf{.558}    \\
-\citealt{Scholes1966}     & {.619}           & {.484}          & {.614}          & \textbf{.671}    \\
+Goodman-Kruskal $\gamma$  \\
 \midrule
-Goodman-Kruskal $\gamma$  & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
+\citeauthor{Albright2007}  & \uline{.874} & {.488}        & {.246}       & {.849} \\
+\citeauthor{Albright2003b} & \uline{.929} & {.574}        & {.502}       & {.606} \\
+\citeauthor{Scholes1966}   & \uline{.907} & {.738}        & {.628}       & {.558} \\
 \midrule
-\citealt{Albright2007}    & \textbf{.874}    & {.849}          & {.502}          & {.574}           \\
-\citealt{Albright2003b}   & \textbf{.929}    & {.606}          & {.246}          & {.488}           \\
-\citealt{Scholes1966}     & \textbf{.907}    & {.558}          & {.628}          & {.738}           \\
+Kendall $\tau_{b}$        \\
+\midrule
+\citeauthor{Albright2007}  & {.666}       & {.450}        & {.245}       & \uline{.683} \\
+\citeauthor{Albright2003b} & {.474}       & \uline{.558}  & {.499}       & {.159} \\
+\citeauthor{Scholes1966}   & {.619}       & \uline{.671}  & {.614}       & {.484} \\
 \bottomrule
 \end{tabular}
-\caption{FIXME}
+\caption{Correlations between item-averaged wordlikeness ratings and model scores}
 \label{scores}
@@
 
-%\subsubsection{Maximum entropy phonotactics}
-%
-%\citeauthor{Hayes2008a} (\citeyear{Hayes2008a}; henceforth H\&W) develop a sophisticated model of phonotactic grammaticality which estimates a probability distribution over phoneme sequences by weighing constraints according to the principle of maximum entropy, following \citet{Goldwater2003} and \citet{Jager2007}.
-H\&W report that the predictions of their model are closely correlated with the \citet{Scholes1966} wordlikeness ratings.
-A direct replication of their predictions was attempted by using the software, model parameters, and training data as described in that study.
-Since the training of the maximum entropy model is inherently stochastic, producing slightly different outcomes on each run, the lowest scoring of ten runs is reported (H\&W:396), though in general there is not a great deal of variation between individual runs.
-One limitation of this model is that it is not feasible to score whole words, as the number of constraints which must be inspected grows exponentially as the span of possible constraints increases.
-Following H\&W and of \citet{Albright2009a}, who also applies the maximum entropy model to the \citet{Albright2003b} norms, the model is trained and scored only on stimulus onsets.
-However, as a consequence, the maximum entropy model performs particularly poorly on this data set, as many stimuli contain phonotactic violations in rime positions.
-%
-%\subsubsection{Segment bigram probability} 
-%\label{bigram}
-%
-%The bigram probability of a sequence $ijk$ is the product of the probability of an sequence-initial $i$, the probability that $j$ follows $i$, and the probability that $k$ follows $j$, and the product of sequence-final \emph{k}.
-%
-%\begin{unlabeledexample}
-%$\displaystyle \hat{p}(ijk) = p(i|\textrm{start}) \cdot p(j|i) \cdot p(k|j) \cdot p(\textrm{stop}|k)$
-%\end{unlabeledexample}
-%
-%\noindent \citet{Albright2009a} employs bigram probability to model wordlikeness judgements.
-While the focus of \citeauthor{Albright2009a}'s study is on developing a model which uses bigrams over phonological features rather than segments themselves, \citeauthor{Albright2009a}'s evaluation, which includes both the \citeauthor{Scholes1966} and \citeauthor{Albright2003b} data sets, finds an advantage for segmental bigrams.
-\citeauthor{Albright2009a} does not provide an implementation of the featural bigram model, nor does his study describe it in sufficient detail to allow for a new implementation, but segmental bigram scores for the \citeauthor{Albright2003a} data are reported in the appendix.
-As reported by \citeauthor{Albright2009a}, segmental bigrams outperform featural bigrams (see Table \ref{albrightimproved}).
-%
-%\citeauthor{Albright2009a} estimates bigram probabilities using the method of maximum likelihood over types in the lexicon.
-The variant of segmental bigrams used here computes probabilities with a simple type of smoothing in which the count of all possible bigrams (including those never observed) are incremented by one.
-This technique is known as Laplace, or ``add one'' smoothing.
-This has the desirable effect that no nonce word is ever assigned a zero probability, and produces a small increase in the correlation between the \citeauthor{Albright2003b} wordlikeness norms compared with the maximum likelihood estimate (Table \ref{albrightimproved}).
-For all three data sets, this model also consistenly outperforms positional probability models defined by \citet{Vitevitch2004} and \citet{Vaden2009}; given that these model scores are highly correlatd with bigram probability \citep[][54]{Vitevitch1997}, they are not considered further.
-The bigram model consistently performs well in all the evaluations, and has the highest Spearman correlation with the \citeauthor{Greenberg1964} and \citeauthor{Scholes1966} data, and is frequently second place model to the binary baseline elsewhere.
-%
-%\subsubsection{Neighborhood density} 
-%\label{density}
-%
-%There are now many methods for computing similarity between nonce words and existing words, long thought to be reflected in wordlikeness judgements. 
-%For this study, a number of such methods were evaluated, including the Generalized Neighborhood Model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a number of variations on neighborhood density \citep{Coltheart1977} provided by \citet{Vaden2009}.
-The best performance was obtained with the simplest version of neighborhood density, which is defined as the number of real monomorphemic words which can be changed into the target nonce word by a single insertion, deletion, or substitution of a phone.\footnote{\citet{Greenberg1964} use a variant in which only substitutions are counted.} For instance, the neighbors of \emph{blick} include \emph{blink} (insertion), \emph{lick} (deletion) and \emph{black} (substitution).
-While many studies \citep[e.g.,][]{Bailey2001} report robust lexical similarity effects, it may be that the relatively weak performance of neighborhood density is the result of the presence of gross phonotactic violations.
-%
-%\subsection{Modeling residual gradience}
-%
-%The primary result is that no gradient model reliably exceeds the accuracy of the binary baseline.
+\subsection{Discussion}
+
+FIXME
+
+\section{Conclusions}
+
+FIXME
+
+%%The primary result is that no gradient model reliably exceeds the accuracy of the binary baseline.
 Despite this, there are relatively strong correlations between the binary baseline and these gradient models (see Table \ref{bcor}).
@@
 In most cases, including the gradient models on top of the binary baseline produces a worse correlation than is obtained with the binary baseline alone.
-%
-%For $\tau_b$ and $\gamma$, the interpretation of this result is clear.
+%%
+%%For $\tau_b$ and $\gamma$, the interpretation of this result is clear.
 The gradient models assign rankings to the sets of phonotactically valid and invalid clusters, respectively.
@@
 This shows quite starkly that these models fail to reliably predict intermediate ratings.
-%
-%\subsection{The gradience hypothesis}
-%
-%This chapter has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
+%%
+%%\subsection{The gradience hypothesis}
+%%
+%%This chapter has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
 The surprising result is that virtually all of the apparent coverage of state-of-the-art gradient phonotactic models is simply a reflection of their ability to distinguish between the possible and the totally impossible; beyond this, they are unreliable.
@@
 It remains to be seen is whether any model can be put forth which accurately predicts these intermediate ratings.
-%
-%These result provide support for recent findings that speakers asked to perform gradient syntactic judgements produce responses closely corresponding to a widely recognized categorical grammatical/ungrammatical distinction \citep{Sprouse2007}.
-%
-%\subsection{Extensions to the binary baseline}
-%
-%The strong performance of the binary baseline should not be taken as evidence either that wordlikenesss judgements are binary, or that the binary baseline is a plausible model.
+%%
+%%These result provide support for recent findings that speakers asked to perform gradient syntactic judgements produce responses closely corresponding to a widely recognized categorical grammatical/ungrammatical distinction \citep{Sprouse2007}.
+%%
+%%\subsection{Extensions to the binary baseline}
+%%
+%%The strong performance of the binary baseline should not be taken as evidence either that wordlikenesss judgements are binary, or that the binary baseline is a plausible model.
 The most serious limitation of this evaluation is the primitive nature of the binary baseline.
@@
 On the other hand, a cognitively plausible version of this model might need to entertain phonotactic generalizations that are larger than these units, since syllable-sized phonotactic generalizations have been proposed for English \citep[e.g.,][]{Berkley1994a,Berkley1994b,Coetzee2008b,Fudge1969}.
-%
-%A possible further extension to the binary baseline would be the introduction of additional levels of wellformedness.
+%%
+%%A possible further extension to the binary baseline would be the introduction of additional levels of wellformedness.
 While the evaluation has shown that current gradient models do not reliably identify intermediate wellformedness, it does seem possible to identify at least three levels of grammaticality: for instance, one might encode the intuition that \emph{zhlick} [ʒlɪk] is more English-like than \emph{bnick}, though both have unattested onsets.
@@
 The bigram and maxent models make this prediction, as do many others \citep[e.g.,][]{Legendre1990,Levelt2000,Albright2008,Anttila2008a,Pater2009b} but despite this, there is still little data demonstrating cumulative effects in wordlikeness tasks.
-%
-%\subsection{Language acquisition}
-%
-%The weak empirical status of gradient phonotactic knowledge as reflected in adults has rammifications for language acquisition under the hypothesis that that infants acquiring language deploy the same representations as adults \citep[e.g.,][]{Macnamara1982,Pinker1984,Crain1991,Carey1995,deVilliers2001,Legate2007}.
+%%
+%%\subsection{Language acquisition}
+%%
+%%The weak empirical status of gradient phonotactic knowledge as reflected in adults has rammifications for language acquisition under the hypothesis that that infants acquiring language deploy the same representations as adults \citep[e.g.,][]{Macnamara1982,Pinker1984,Crain1991,Carey1995,deVilliers2001,Legate2007}.
 Gradient wordlikeness judgements in adults would provide support for claims that infants recognize statistical(inherently gradient) dependencies between segments \citep{Jusczyk1994} and use these to segment words \citep{Saffran1996}.
@@
 The vacuous nature of current evidence for gradient phonotactic knowledge in adults further weakens any hypothesis that would link statistical learning in infants to adults' behaviors.
+
+%% Hayes2000
 % 
-%\section{Conclusion}
+% Bard1996
 % 
-%%For instance, infants as young as 4.5 months seem to be aware that English nasal codas agree in place with following obstruents \citep{Mattys2001b,Jusczyk2002,Davidson2004}. 
-%%It might be the case that syllable co-occurrence statistics might be little more than a reflection of infants' learning of categorical ``lexical viability'' constraints \citep{Johnson2003} of the sort also seen in adult speech processing \citep[e.g.,][]{Norris1997}.
+% Koo2009
 % 
-%% other ratings, but por qué?:
-%% 
-%% Hayes2000
-%% 
-%% Albright2003a
-%% Albright2003b
-%% Prasada1993 
-%% 
-%% Bard1996
-%% 
-%% Koo2009
-%% 
-%% Treiman2000
-%% 
-%% Warker2006
-%% 
-%% Massaro1983
-%% 
-%% Rusaw2009
-%do not explicitly state why this data is relevant to the construction of models of wordlikeness.
+%
+
+% 
+% Warker2006
+% 
+% Massaro1983
+% 
+% Rusaw2009
+
+%%do not explicitly state why this data is relevant to the construction of models of wordlikeness.
 Presumably, these authors believe that these patterns of judgements demonstrate that wordlikeness, as an internal state, is gradient simply because subjects make use of intermediate degrees of wordlikeness in judgement tasks.
 This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, an influential view in the cognitive sciences in general (see \citealt{Fodor1981a} for a critique).
-%
-%\citeauthor{Chomsky1965} were not the first to consider the notion of possible and impossible words.
+
+%%\citeauthor{Chomsky1965} were not the first to consider the notion of possible and impossible words.
 Their primary contribution is that their mentalist perspective: they recognize that naïve speakers effortlessly acquire language-specific generalizations about possible and impossible words and can report them without any explicit training.
-%
-%However, not all early literature is concerned with gradience.
+
+%%However, not all early literature is concerned with gradience.
 \citet[31]{Vogt1954}, for instance, recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
@@
 \citeauthor{Vogt1954} concludes that additional grammatical machinery will be needed to account for possible and impossible words.
-%
-%Most relevant to the question at hand, \citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated wtih the positional probailities of the constituent syllables.
+
+%%Most relevant to the question at hand, \citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated wtih the positional probailities of the constituent syllables.
 Unfortunately, none of these researchers make any effort to eliminate the possibility that the low positional probability stimuli are ``impossible'' words of English.
 In fact,
+
 %Chapter \ref{clusters} argues
-%%the author has argued elsewhere \citep{Gorman2012c}
 %that many of the stimuli used by \citeauthor{Frisch2000} and \citeauthor{Vitevitch1997} contain illicit word-medial consonant clusters.
 While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and wordlikeness judgements, but \citeauthor{Vitevitch1997} ultimately conclude this cannot explain all the variation in wordlikeness.
-%
+
 %Using the head-term preference paradigm, \citet{Jusczyk1993b} and \citet{Friederici1993} find that typically-developing children as young as 9 months of age distinguish between nonce words which are and are not phonotactically valid in their target language.
@@
 Faciliatory effects of positional probability (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements \citep{Vitevitch1999a,Luce2001,Lipinski2005,Vitevitch2005}, and lexical decision \citep{Pylkkanen2002a}.
-%
-%The aforementioned studies all conclude that the gradient measure of positional probability correlates with behavioral results.
+
+%%The aforementioned studies all conclude that the gradient measure of positional probability correlates with behavioral results.
 As the flaws of the \citet{Vitevitch1997} study demonstrate, the aforementioned studies do little to tease apart the gradient and categorical aspects of phonotactics.
@@
 This is particularly troublesome given that no justification has ever been given for the positional probability measure in the first place; it appears to have been created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production \citep[e.g.,][]{Luce1998,Luce2000} and perception \citep{Marslen-Wilson1984,Marslen-Wilson1987,McClelland1986,Norris1994,Norris2000}.
-%\subsubsection{\citealt{Greenberg1964}}
-%
-%\citet{Greenberg1964} investigated wordlikeness using the technique of free magnitude estimation, a mechanism which has become increasingly popular among syntacticians \citep[e.g.,][]{Bard1996}.
-At the beginning of the experiment, the subject heard a recording of the word \emph{stick}.
-In subsequent trials, the subjects heard a nonce word and were asked to report ``how far would you say that is from English?'', with \emph{stick} at ``1''; subjects are told that a word that is ``twice as far from English'' as \emph{stick} should be scored ``2''.
-The data used here are from \citeauthor{Greenberg1964}'s Experiment B, in which 17 undergraduates were presented 17 stimuli in all.
-In addition to \emph{stick}, the stimuli include three other English words; these four items were excluded from further analyses, leaving 13 stimuli.
-As is standard practice in psychophysics \citep[e.g.,][]{Butler1987}, magnitudes were log-transformed prior to analysis.
-%

### 2012-11-05 3.tex
Writing on chapter 2 primarily, also added the two "MSes" and new graphs for chapter 2.

@@
-\chapter{Static and derived Turkish phonotactics}
 \label{turkish}
@@
 First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules.
-Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{Anderson1974}: chap.~9, \citealt{Gorman2012e}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{Kenstowicz1977}:189f., \citeyear{Kenstowicz1979}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
+Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{A74}: chap.~9, \citealt{Gorman2012e}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{KK77}:189f., \citeyear{KK79}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
 
@@
 
-\citet[212]{Anderson1974} and \citet{Iverson1978} argue that suffix harmony in disharmonic roots found in (\ref{turkexcept}a) requires the rule governing suffix harmony alternations to be distinguished from a sequence structure constraint governing root harmony.
+\citet[212]{A74} and \citet{Iverson1978} argue that suffix harmony in disharmonic roots found in (\ref{turkexcept}a) requires the rule governing suffix harmony alternations to be distinguished from a sequence structure constraint governing root harmony.
 Since the rule and sequence structure constraint are otherwise identical, this constitutes an undesirable ``duplication'' in the sense of \citealt{Kisseberth1970b}.
@@
 
-\citeauthor{Anderson1974} also notes that roots which fail to undergo suffix harmony, like (\ref{turkexcept}b), may themselves be harmonic. He takes to be evidence for the necessity of duplication:
+\citeauthor{A74} also notes that roots which fail to undergo suffix harmony, like (\ref{turkexcept}b), may themselves be harmonic. He takes to be evidence for the necessity of duplication:
 
@@
 Since the morpheme structure condition and the phonological rule in this case have distinct classes of exceptions, it is clear that they cannot be identified.
-\citep[289]{Anderson1974}
+\citep[289]{A74}
 \end{quote}
@@
 
-One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, either by non-derived environment blocking \citep{Mascaro1976} or structure preservation condition \citep{Kiparsky1985}.
+One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, perhaps by structure preservation condition \citep{Kiparsky1985}.
 However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default.
@@
 
-%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
-It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails.
-If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
-%A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements.
-This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).
-%Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed.
-In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.
-%In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word. 
-%The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements. 
-%The data below, drawn from Turkish, shows that 
-%This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish.
-Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements.
-Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.
-%As a quick validation of this claim, consider the results reported by \citet{Albright2007} and discussed in the previous chapter.
-English speakers asked to rate non-words on a scale from 1-7, where 7 indicates ``most like English'', favor non-words [blʌs, blɑd] (average scores 4.67 and 5.13) over [bnʌs, bnɑd] (average scores 2.06, 2.00).
-%This hypothesis is not ahdered to by early research on sequence structure constraints \citep[e.g.,][]{SPR,Chomsky1965,SPE,Stanley1967}.
-Work in this vein sought to use sequence structure constraints to compress the lexicon as much as possible.
-This work has a second interest, namely in ``compressing'' the lexicon as much as possible,
-%motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps.
-\citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible.
-\emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy.
-In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities
-%Generative phonological theory at the time of \emph{SPE} lacked the representational vocabulary to directly distinguish between word-initial /bn/, which is disallowed, and /bn/ as a licit syllable contact cluster (e.g., \emph{o}[b.n]\emph{oxious}), a point first made by \citep{Hooper1973}.
-It is not the introduction of prosodic primitives into generative phonology that explains the contrast between \emph{blick} and \emph{bnick}, though, but rather the principle that surface forms must be syllabified \citep[e.g.,][63f.]{Kiparsky1982b} which \emph{bnick} fails.
-If, as is standardly assumed, phonology repairs unsyllabifiable outputs \citep[e.g.][]{Ito1989a,Noske1992}, then \emph{bnick} is an impossible output, doomed to be modified (perhaps to \emph{nick}; \citealt[][19f.]{Wolf2009}), and /bnɪk/ is an impossible UR by Stampean occultation.
-%Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion).
-This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.
-%\citet{Berent2007a}
-%\citet{Berent2007b}
-%\citet{Berent2008a}
-% 598: Highly marked inputs are repaired in perception to abide by the grammatical restrictions of the language
-% Exps 1-2: syllable count task (misperception due to linguistic experience; "disyllables were more likely to be perceived as monosyllables if their monosyllabic counterparts had a less marked cluster" in both Russian and English, not what OT predicts)
-% Exps 3-4: AB task
-% Exps 5-6: priming effect (non-epenthetic and epenethetic things represented the same)
-% Berent and Lennertz 2007: (640: If the high rate of monosyllabic errors to unmarked onsets is only due to phonetic failures to encode the input, then it is puzzling why the same onsets also yield the highest rate of accurate response.)
-% Berent et al. 2008a:
-%\citet{Peperkamp2007}
-% Peperkamp 2007: during one stage of processing phonology is undone with help from the lexicon...it only takes licit inputs and is irrelevant (p. 633)
-%\citet{Dupoux1999}
-% Dupoux et al. 1999: Japanese speakers still can't distinguish between ebzo and ebuzo in an ABX task using the SAME a or b.
-%Once additional source of evidence on root (dis)harmony is inconclusive.
-There is a small class of bisyllabic words in which the second vowel, always [$+$\textsc{High}, $-$\textsc{Back}], alternates with zero.
 %\ex High-vowel/zero alternations \citep[][243]{Clements1982}: \\
@@
 %\end{tabular} \xe
-%It is possible that \textsc{Backness Harmony} might produce a fluctuating \emph{ı} after root \emph{a}, but this does not obtain (\lastx b).
-However, this might simply indicate that the fluctuating vowel is epenthetic and that harmony applies before epenthesis (see \citealt{Clements1982} for both sides of this argument), making it less than a counterexample.
-%duplication problem: \citep{Kisseberth1970b} (see also \citealp[][401]{Stanley1967} and \emph{SPE}:382)
-%(e.g., \emph{deve} `camel' vs. \emph{deva} `medicine', \emph{sene} `year' vs. \emph{sena} `praise').
 %This is not to imply uncritical acceptance of \citeauthor{Becker2011}'s account.
@@
 It is not immediately clear that this is evidence that the generalizations in question are external to the synchronic grammar, however.
-
+% 
 %  & {harf}         & {harfler}  & `(alphabetic) letter' \\ 

### 2012-11-05 4.tex
Writing on chapter 2 primarily, also added the two "MSes" and new graphs for chapter 2.

@@
-\chapter{Structural and accidental phonotatic gaps}
 \label{gaps}
 
-Early work in generative phonology notes that neutralizing, surface-true phonological alternations impose constraints both on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kisseberth1970b}, \citealp{Kenstowicz1977}: chap.~3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}).
+%\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984} or not \citep[e.g.,][]{Port1985}, it is not strictly relevant to the discussion at hand.}
 
-Two domains are essential to defining
+Since at least \citet{Pike1947b}, linguists have posited language-specific constraints on the contents of underlying representations.
+These \emph{morpheme structure constraints} were at one time thought to fully account for speakers' knowledge of possible and impossible words (e.g., \citealt{Chomsky1965}, \citeyear[382]{SPE}, \citealt[22f.]{SPR}, \citeyear{Halle1962}, \citealt{Stanley1967}).
+\citet{Shibatani1973}, however, argues that morpheme structure constraints cannot account for all wordlikeness constraints.
+In German, for instance, final obstruents devoice, and as a result, there are pairs such  [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' and [ɡʀaːt]-[ɡʀaːdə] `degree(s)' differing only in the plural shape of the root.
+Since the voicing of final obstruents is contrastive, the restriction on the voicing of obstruents cannot be a morpheme structure constraint. 
+Yet, \citeauthor{Shibatani1973} claims that native speakers reject nonce words ending in final voiced obstruents ``\emph{on the ground that they end in voiced obstruents}'' (95). 
+Thus, not all constrasts in possible wordhood can be stated as morpheme structure constraints.
 
-Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}),
+Whereas \citeauthor{Shibatani1973} maintains that morpheme structure constraints are insufficient to account for speakers' knowledge of possible wordhood, others argue that morpheme structure constraints are also unnecessary.
+\citet[297]{Hale1965}, \citet{Kisseberth1970b}, and \citet[212f.]{Postal1968} observe that the structural descriptions of phonological processes often are often reflected in the lexical redundancies in the same language.
+In Russian, there are alternations implicating a process of obstruent voice assimilation, and tautomorphemic obstruent clusters have uniform voicing \citep[283]{A74}.
+\citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that morpheme structure constraints are otiose, as phonological rules triggering alternations impose restrictions on the contents of URs, an effect known as \emph{Stampean occultation}.
 
-\begin{quote}
-It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
-\citep[283]{Anderson1974}
-\end{quote}
+There are some apparent constraints on URs which are not reflected in the system of alternations; these generalizations seem to require reference to the units of prosody (\citealt{Hooper1973}, \citealt{Kahn1976}; see \citealt{Blevins2003} and \citealt{Steriade1999} for possible alternatives).
+Since word-level prosodic structure is generally thought to be largely redundant and thus absent from underlying representations, these cannot be equated with morpheme structure constraints either.
+\citeauthor{Haugen1956} notes, for instance, that word-medial consonant cluster inventory is not arbitrary in prosodic terms.
 
-\section{English syllable contact clusters}
+\begin{example}[\textsc{Medial Cluster Law} (after \citealt{Haugen1956})]
+\label{mcl}
+A word-medial consonant cluster consists of at most a well-formed medial coda followed by a well-formed medial onset
+\end{example}
 
-English admits a variety of word-medial consonant clusters spanning syllable boundaries; these are known as \emph{syllable contact clusters} or \emph{interludes}.
-\citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in some language, echoing a similar proposal by \citet{Haugen1956}.
+Reviewing the inventory of word-medial consonant clusters in English, 
+\citet{Pierrehumbert1994} notes that the vast majority of ``possible clusters'', those which can be parsed into a coda and onset, are unattested.
+\citeauthor{Pierrehumbert1994} argues that the cluster inventory is further filtered by what are essentially morpheme structure constraints: static co-occurrence restrictions with no support from either prosody or the system of alternations.
+This study will argue, however, that the only structural constraints on word-medial clusters in English are the \textsc{Medial Cluster Law} and Stampean occultation.
+Consequently, this domain provides no evidence for the existence of morpheme structure constraints.
 
-\begin{quote}
-\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
-\end{quote}
+\section{English syllable contact clusters \citep{Pierrehumbert1994}}
 
-Many phonotactic generalization can be thought of as restrictions on the inventory of different prosodic constituents defining what is, e.g., a ``well-formed onset''.
-Since few linguists posit the interlude as a prosodic constituent \citep[though cf.][]{Steriade1999}, further constraints on syllable contact clusters, however, are combinatoric.
+English admits a variety of word-medial consonant clusters spanning syllable boundaries which are also known as \emph{syllable contact clusters} or \emph{interludes}.
+These clusters can be as short as two consonants, as in \emph{a}[n.t]\emph{ics}, or as long as four, as in \emph{mi}[n.str]\emph{el}.
+There are three
+
+\subsection{The role of Stampean occultation}
+
+Finally, \citeauthor{Pierrehumbert1994} does not fully consider the possible effects of Stampean occultation, as some of the static constraints she proposes may be related to English morphophonemics.
+For instance, she writes that ``nasal-stop sequences agree in labiality'' (175), and attributes this the status of a static constraint.
+However, this generalization is just a narrower form of the one derived by \textsc{Nasal Place Assimilation} (see \S\ref{npa}), and which requires no further explanation.
+Other derived constraints go unnoticed: for instance, she does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}).
+The evaluation below considers the role of both types of constraints on the English lexicon on an equal footing.
+
+\subsection{The role of sparsity}
+
+\citet{Pierrehumbert1994} infers static constraints from near-exceptionless gaps in the lexicon, but little effort is made to show that the patterns of lexical underrepresentation from which the static constraints are inferred are not due to chance.
+This admits the possibility that the gaps in the consonant cluster inventory are accidental rather than structural in nature \citep{Fischer-Jorgensen1952,Vogt1954}.
+This is made all the more likely given the tendency of segment and cluster frequency distributions to be highly skewed \citep[e.g.,][]{Pande2010,Sigurd1968,Tambovtsev2007,Weiss1961}.
+In Figure \ref{cao}, the rank and type frequency (i.e., frequency in the lexicon) frequency of medial codas and onsets in the lexical sample described below are graphed in log-log space.
+Both codas and onsets show a linear relationship between log rank and log frequency characteristic of Zipfian distributions (see Appendix \ref{zr}).
+As a result, an enormous lexical sample would be needed to realize clusters consistent with the \textsc{Medial Cluster Law} even if there were no constraints on the combination of medial codas and onsets in English.
+
+\begin{figure}
+\centering
+\includegraphics{co.pdf}
+\caption{Medial coda and onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
+\label{cao}
+\end{figure}
+
+\subsection{The role of morphological segmentation}
+
+Many components of \citeauthor{Pierrehumbert1994}'s study cannot be directly replicated.
+\citeauthor{Pierrehumbert1994} focuses her attention on those words she judges to be ``morphologically simple'' and ``reasonably familiar''.
+Since this list has neither been published nor circulated, any replication will have to be indirect.
+Further, it is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias.
+A sequence which is rare in simplex words is sometimes said to mark a juncture even in the absense of other evidence for segmentation; for instance, \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur morph-internally.
+There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings.
+Applied indiscriminately, though, such a heuristic has the possibility to trivializes both morphological segmentation and phonotactic generalization.
 
@@
 
+\citeauthor{Pierrehumbert1994} focuses only on triconsonantal clusters, and no reason is given for the exclusion of other sizes of clusters.
+
 \subsection{Method}
+
+%The same practice is sometimes applied to preserve phonological generalizations; for instance, otherwise unmotivated morphological boundaries are used in \emph{SPE} to simplify principles of stress assignment.
+Such reasoning is found in \citeauthor{Pierrehumbert1994}'s discussion of the word \emph{a}[nt.l]\emph{er}, however: this contains a cluster which violates her constraint on coda coronal obstruents, and therefore must be a complex word, she argues (see \citealt[164]{Borowsky1989} for additional exceptions to this generalization).
 
@@
 
-The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations.
+The treatment of Latinate forms in \emph{The sound pattern of English} (\citealt{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition.
 In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.
@@
 \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.
-%\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.
 
@@
 
+\subsection{Results}
+
 This filtering results in a list of 6,619 simplex words.
@@
 
-\subsection{Results}
-
 \subsubsection{Static constraints}
@@
 These constraints are ``static'' in the sense that they have no analogues in the morphophonemics of English.
-The investigation considers clusters of all sizes, not merely the triconsontal clusters studied by \citeauthor{Pierrehumbert1994}, as she gives no justification for excluding biconsontal or quadraconsonantal clusters.
 
@@
 \end{tabular}
-\caption{FIXME}
+\caption{Dorsal-labial cluster attestation in the lexical sample}
 \label{dltab}
@@
 
-\paragraph{Coronal codas} \citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}.
-Table \ref{cctab} shows that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}).
+\paragraph{Coronal obstruent codas} \citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}.
+Table \ref{cctab} reveals that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}).
 Restricting attention to triconsonantal clusters does not produce a significant lexical effect, either ($p = 0.129$).
@@
 \end{tabular}
-\caption{FIXME}
+\caption{Coda coronal obstruent cluster attestation in the lexical sample}
 \label{cctab}
 \end{table}
-
-\noindent
 
@@
 \end{tabular}
-\caption{FIXME}
+\caption{ABA cluster attestation in the lexical sample}
 \label{abatab}
@@
 \paragraph{Obstruent voice assimilation}
+\label{ova}
 
 Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), which take the voicing specification of a preceding obstruent.\footnote{Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.}
+%Bizarely, \citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''.
 
@@
 \end{tabular}
-\caption{FIXME}
+\caption{Obstruent voice assimilation cluster attestation in the lexical sample}
 \label{ovatab}
@@
 
-\textsc{Nasal Place Assimilation} (\emph{SPE}:85) accounts for [n $\sim$ ŋ] allophony and Latinate prefix allomorphy, such as that of the ``negative'' refix \emph{im-}/\emph{in-}.\footnote{\citealt{Borowsky1986}:FIXME and \citealt{Jensen2000} propose additional restrictions on \textsc{Nasal Place Assimilation}.}
+\textsc{Nasal Place Assimilation} (e.g., \citealt[149]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}) accounts for [n $\sim$ ŋ] allophony as well as Latinate prefix allomorphy.
 
@@
 \end{tabular}
-\caption{FIXME}
+\caption{Nasal place assimilation cluster attestation in the lexical sample}
 \label{npatab}
@@
 \paragraph{Degemination} The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.
-\emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}).
+\emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}).
 \textsc{Degemination} is formalized here as a rule deleting the first of two identical segments, ignoring voice.
@@
 \end{tabular}
-\caption{FIXME}
+\caption{Degemination in the lexical sample}
 \label{degemtab}
@@
 
-\paragraph{Summary} All three of \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters.
+\paragraph{Summary} All three of \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters; possible clusters which are surface exceptions to these three rules are much less likely to be attested than those which conform to them.
 
@@
 
-A soft-margin support vector machine \citep{Cortes1995} with a linear kerne
-l is used to convert from the per-cluster scores to attestation predictions
+A soft-margin support vector machine \citep{Cortes1995} with a linear kernel is used to convert from the per-cluster scores to attestation predictions
 , by finding a single cutoff which optimally separates attested and unattes
-ted clusters according to their model scores. This is in no way intended to
- be a cognitively plausible model for learning phonotactics; it simply repr
-esents an upper bound on the performance of these models. The results for a
-ll four models are summarized in Table \ref{cmresults}.
+ted clusters according to their model scores.
+
+It is possible to maximize recall at the expense of precision, by predicting all clusters to be unattested, or to increase precision at the expense of recall, by predicting non-attestation for a great number of clusters.
+$F_1$, the harmonic mean of these two measures, is a balanced measure of this tradeoff.
+
+This is not put forth as a cognitively plausible model for learning of phonotactics; it simply represents an upper bound for classifying possible clusters.
+
+The results for are summarized in Table \ref{cmresults}.
+
+\paragraph{Null baseline} Since the majority of possible clusters are unattested, a null baseline which predicts all clusters to be unattested achieves classifies 81\% of clusters correctly, achieving perfect recall at the price of precision.
+
+\paragraph{Expected frequency} \citet{Pierrehumbert1994} proposes that the well-formedness of a syllable contact clusters is proportional to the independent probability of the coda and onset that make it up.
+\citeauthor{Pierrehumbert1994} reports that this is the best predictor of which complex clusters occur, and while it is a significant improvement on the baseline (sign test $p = 4.5$\e{-05}), it is not distinguished in performance from the other models under consideration.
+
+\paragraph{Derived constraints} The three derived generalizations define a simple classifier, predicting attestation for all clusters which do not violate these three generalizations.
+This is a significant improvement on the null baseline (sign test $p = 4.0$\e{-09}), preserving high recall while improving precision.
+The predictions of the derived constraints classifier can also be combined with information on expected frequency.
+As expected, this boosts precision, since FIXME.
+This combination produces a model with the highest accuracy and $F_1$.
+
+\paragraph{Maximum entropy phonotactics} \citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints.
+Since this model has many experimenter-determined parameters, a direct replication of the experiments reported by \citealt{Hayes2008a} is attempted; their software, feature specification, and model settings are all used.
+Training in this model is inherently stochastic, producing slightly different constraints and weights on each run, so the least accurate of ten independent runs is used for evaluation; this is also the practice of \citeauthor{Hayes2008a}.
+This model performs approximately as well as other classifiers, but has the poorest recall; this suggests that the constraints learned are overly narrow and fail to filter out many impossible clusters.
+FIXME
+
+\paragraph{Summary} FIXME
 
@@
 Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
-\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
 Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
 Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
-DC and EF           & 0.861    & 0.868     & 0.988  & 0.924 \\
+DC \& EF            & 0.861    & 0.868     & 0.988  & 0.924 \\
+\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
 \bottomrule
 \end{tabular}
-\caption{FIXME}
+\caption{Results of the cluster classification task; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
 \label{cmresults}
@@
 \subsection{Discussion}
+
+FIXME
 
@@
 \subsubsection{Simulating medial clusters}
+\label{simulate}
 
@@
 \caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line).
-Both simulated and observed frequencies have been smoothed using the $Z_r$ transform.}
+Simulated and observed frequencies have been smoothed using the $Z_r$ transform.}
 \label{sim}
@@
 
+In summary, the sparse cluster inventory, previously taken as evidence for static constraints on syllable contact clusters, is expected even if the only constraints on possible clusters are those which are phonologically derived.
+
 \section{Conclusions}
 
-%however, the \emph{communis opinio} still is that that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
-%
-%%FIXME
-%% \citet[255]{Anderson1974}, discussing this fact, writes that \citeauthor{Haugen} ``gives an argument that shows the necessity of including syllables as units in underlying representations''.
-%
-%%FIXME
-%(for dissenting views, see \citealt{Blevins2003} and \citealt{Steriade1999})
-%
-%Linguists have long recognized underlying representations as privileged domains for stating phonotactic generalizations (e.g., \citealt{Bloomfield1930,Jakobson1932,Pike1947b}). 
-%
-%More recently, a similar role for the syllable has been recognized 
-%(e.g., \citet{Haugen1956}, \citet{Hooper1973}, \citealt{Kahn1976})
-%
-%
-%\begin{quote}
-%It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
-\citep[283]{Anderson1974}
-%\end{quote}
-%
-%
-%\begin{quote}
-%It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained.
-\citep[283]{Anderson1974}
-%\end{quote}
-%
-%English word-medial consonant clusters have been used, by \citet{Pierrehumbert1994} and others, to argue for the necessity of an independent phonotactics module and illustrate proposals for its architecture.
-This chapter argues, however, that the only structural constraints on the concatenation of codas and onsets to form word-medial clusters are derivative of well-documented phonological alternations in English.
-%
-%\section{Background}
-%
-%English admits a variety of word-medial consonant clusters spanning syllable boundaries, known as \emph{syllable contact clusters} or \emph{interludes}.
-These clusters can be as short as two consonants (e.g., \emph{a}[n.t]\emph{ics}; singleton intervocalic consonants are ignored here) or as long as four (e.g., \emph{mi}[n.str]\emph{el}).
-\citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in any given language \citep[cf.][]{Haugen1956}.
-%
-%\begin{quote}
-%\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
-%\end{quote}
-%
-%This chapter is concerned with the nature of these ``provisos'' as reflected by which of the ``possible'' clusters, so defined, are attested.
-While many phonotactic generalizations can be thought of as inventory restrictions imposed by prosodic constituents, additional filters on the set of possible clusters are essentially combinatoric in nature and thus distinct, there is no prosodic constituent consisting of the coda of one consonant-final syllable and the onset of a following consonant-initial syllable \citep[though see][]{Steriade1999}.
-%
-%\section{Evaluation}
-%
-%\citeauthor{Pierrehumbert1994} argues that the inventory of clusters in English shows the effect of a small number of static phonotactic constraints.
-The remainder of this chapter provides an exhaustive quantitative evaluation of this hypothesis.
-The most important constraints on medial clusters are those derived from phonological rules which apply across syllable boundaries; there is no remaining role for static constraints.
-The sparse nature of the lexicon plays a secondary role, giving rise to apparent phonotactic gaps even in the absence of static phonotactic constraints.
-%
-%\subsection{Corpus construction}
-%
-%A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database.
-The resulting list of clusters is reproduced in Appendix \ref{appendixC} and in electronic form at the author's website.
-%
-%\subsubsection{Simplex words}
-%
-%An analysis of lexical constraints requires a corpus of lexical representations.
-So as to remain agnostic about particulars of morphological theory, the goal here is to identify words which bear no overt inflection, and which do not admit any plausible further decomposition, henceforth ``simplex words''.
-For the structuralists, a procedure for parsing words into morphs, just the thing for locating simplex words, held the status of ``philosopher's stone'', a sort of near-mythical objective of the endeavor, but actual attempts to develop such a procedure \citep[e.g.,][]{Harris1955,Nida1948} are now recognized as anything but theory-neutral.
-Yet, there is no alternative to adopting some computational procedure or human-coded database as the ``gold standard'' for a study of this type.
-\citet{Pierrehumbert1994} analyzes a list of entries in the Collins English dictionary by hand, using her intuitions to filter out ``complex'' words as well as those which are not ``reasonably familiar''.
-Unfortunately, the resulting list was neither published nor circulated, and the possibility of replicating \citeauthor{Pierrehumbert1994}'s' sensations of morphological complexity is remote.
-Following prior studies of syllable contact clusters by \citet[ chapter 3]{Hammond1999a} and \citet[ chapter 8]{Duanmu2009}, a list of simplex words was derived from the ``lemmas'' (that is, uninflected forms) in the English portion of CELEX \citep{CELEX}, a database of morphological, phonological, and syntactic annotations based on the COBUILD corpus, a resource that was also used to construct the Collins English dictionary employed by \citeauthor{Pierrehumbert1994}.
-%
-%Two further filters were applied to this list.
-First, any lemma coded as an unassimilated loanword is excluded.
-Secondly, all lemmas that have a ``morphological status'' other than ``monomorpheme'' are excluded.
-This is crucial because CELEX makes no guarantee that lemmas are simplex; indeed, many lemmas are products of derivation (e.g., \emph{abusive}, from \emph{abuse}).
-These stringent criteria have a profound effect on the makeup of the data.
-For instance, nearly all exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) cited by \citet[74]{Hammond1999a} are excluded as unassimilated loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as morphologically complex (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}).
-%
-%``Neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}, have an ambiguous and long-debated morphological status.
-\citeauthor{Pierrehumbert1994} attempts to distinguish between words like \emph{complete}, \emph{extreme}, \emph{inspect}, \emph{obtuse}, which she assumes to be simplex, and \emph{excrete} and \emph{transparent}, which she considers complex.
-However, CELEX codes these words as complex and they are thus excluded.
-This is consistent with converging evidence that speakers decompose these Latinate forms.
+None of the foregoing results have revealed any evidence for structural constraints on the English syllable contact cluster except those derived from phonological alternations.
+What then is to be said of the 366 clusters which do not violate a derived constraint, but which are yet unattested, like [b.z] or [z.n]?
+Since these gaps appear to be quite arbitrary from a phonological perspective---they cannot be reliably discovered either by linguist or computer---it seems reasonable to identify these clusters as accidental gaps, resulting from the sparsity of the sample and of the English lexicon.
 
-%Decomposition is also supported by lexical decision results.
-Among nonce words, those that can be exhaustively decomposed into prefix and bound stem, like *\emph{de}-\emph{juvenate}, are processed slower that contain a licit morph but cannot be exhaustively decomposed, like *\emph{de-pertoire}, and these in turn are processed slower than those which admit no such decomposition \citep{Taft1975}.\footnote{The same contrasts are present in many other domains where decomposition into root and affix(es) is less controversial, for instance, inflected verbs in Italian \citep{Caramazza1988}.} \citet{Emmorey1989} reports facilitative auditory priming between pairs of words which share a bound stem, like \emph{permit}-\emph{submit} \citep[though see][]{Marslen-Wilson1994}.\footnote{It may be the case that full decomposition is only one of the two methods for the processing of complex words, the other being whole word looking, either running in serial \citep{Caramazza1988} or in parallel \citep{Baayen1997b}.
-Whatever evidence that is amassed in support of a whole word lookup procedure, however, does not undermine the considerable evidence for decomposition.}
-%
-%\subsubsection{Syllabification and phonologization}
-%
-%CELEX includes broad, syllabified Received Pronunciation transcriptions of individual wordforms.
-However, a casual inspection of the data reveals the unsystematic nature of these syllabifications.
-Any two words sharing a nucleus-medial cluster-nucleus sequence should be given the same parse, since syllabification is universally non-contrastive.
-Yet many putative syllabification contrasts can be found in CELEX; for instance, compare the parses given to the clusters in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][II.294]{AOE}.}
-%
-%A systematic syllabification requires an automated procedure for delimiting medial clusters and parsing them into coda and onset.
-Since this procedure is applied only to medial clusters, this study remains agnostic on several contentious issues concerning syllabification of English; for instance, there is no need to address the status of so-called ``ambisyllabic'' consonants, since this does not occur with medial clusters, or morphological effects on syllabification, since no complex words are included.
-The technique used here is a variation on the theme of onset maximization (e.g., \citealp[42f.]{Kahn1976}, \citealp{Kurylowicz1948}, \citealp[75]{Pulgram1970}, \citealp[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset.
-Initial onsets are used to define what is a ``possible'' medial onset \citep[though cf.][36]{Fischer-Jorgensen1952}.
-Medial clusters in words like \emph{neu}[.tɹ]\emph{on} or \emph{bi}[.stɹ]\emph{o} are found in word-initial position (e.g., [tɹ]\emph{ansit}, [stɹ]\emph{ike}), so onset maximization assigns the entire medial cluster to the onset, leaving the medial coda empty.
-In contrast, the cluster in \emph{mi}[n.stɹ]\emph{el} is not found word-initially; the maximal onset here is [stɹ], and the residual [n] is assigned to the coda.
-%
-%%\citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''
-%%\citet[?]{Lombardi1999} says that English lacks devoicing in Level II.
-%
-%\paragraph{Stressed lax vowels} When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}.
-\citet[][55]{Harris1994} notes that onset maximization produces an incorrect result when, in addition, the medial cluster is a valid onset: in \emph{whisper}, \emph{vestige}, and \emph{musket}, onset maximization incorrectly assigns [sp, st, sk] to the medial onset, leaving the lax vowel unchecked.
-Consequently, the first consonant of a medial cluster is assigned to the coda of a preceding syllable before a stressed lax vowel \citep[cf.][48]{Pulgram1970}, a minimal modification of the onset maximization parse.
-%
-%\paragraph{Affricates} If the ffricates [tʃ, dʒ] are treated as stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990}, the previous principle will incorrectly assign the two segments of the affricate to separate syllables before stressed lax vowels (e.g., *\emph{ra}[t.ʃ]\emph{et} or \emph{a}[d.ʒ]\emph{ile}) unless some ad hoc constraint \citep[e.g.,][]{Wells1990} is in place.
-A more traditional analysis of affricates, in which they are simple segments distinguished from pure stops by stridency \citep[24]{Jakobson1961} or delayed release (\emph{SPE}:321f.) is assumed here, following \citeauthor{Pierrehumbert1994}.
-This assumption can be motivated by the tendency of affricates to pattern with individual segments in phonotactic generalizations.
-For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but bans true onset clusters \citep[9]{Launey2011}.
-%%\citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.
-%
-%\paragraph{The velar nasal} According to some \citep[e.g.,][]{Sapir1925}, [ŋ] in English is a phoneme in its own right, but others (e.g., \emph{SPE}:85, \citealp{Borowsky1986}:65f.) hold that it is simply the allophone of /n/ before underlying /k, ɡ/ (with later deletion of /ɡ/ in some contexts).
-The total absence of onset [ŋ], where it cannot be followed by a dorsal consonant which conditions the velar allophone, is strong evidence that [ŋ] is a pure allophone of /n/, the analysis assumed here.
-%
-%\paragraph{[j] onglides} While some have argued that the front onglide in words such as \emph{val}[j]\emph{ue} is inserted by rule (\emph{SPE}:196, \citealp[][89]{Halle1985a}, \citealp[][217]{McMahon1990}), the very presence or absence of the glide is contrastive (e.g., \emph{coot} \alt{} \emph{cute}, \emph{booty} \alt{} \emph{beauty}) indicating that it is present in underlying representation (\citealp{Anderson1988b}, \citealp[278]{Borowsky1986}).
-The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{ju}[n.j]\emph{or}).
-There is considerable evidence for this assumption.
-When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u], suggesting that the glide is nuclear (\citealp{Davis1995}, \citealp[][61f.]{Harris1994}, \citealp[][232]{Hayes1980}).
-\citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters, though they may be followed by [ju] in words like [mj]\emph{use} or [vj]\emph{iew}, also suggesting the nuclear affiliation of the glide.
-There is also external support for this analysis: the [ju] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005}, \emph{shm}-reduplication \citep{Nevins2003}, and speech errors (e.g., [kju]\emph{mor} [h]\emph{omponent}, intended [hju]\emph{mor component}; \citealp[130]{Shattuck-Hufnagel1986}).
-%
-%\paragraph{[w] onglides} The phonotactic properties of the back onglide [w] are the opposite of the front onglide.
-Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}).
-Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[161]{Davis1995}.
-Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[166]{Davis1995}.
-Taken together, these facts indicate that the back onglide is assigned to the onset.
-%
-%\paragraph{Syllabic \emph{r}} Word-medial syllabic \emph{r} has been lost in Received Pronunciation, the accent used for the CELEX transcriptions.
-Even in \emph{r}-ful dialects, though, there is reason to believe that syllabic \emph{r} is fact nuclear, and thus irrelevant to syllable contact clusters.
-Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}): compare American English \emph{fern}, \emph{fir}, \emph{fur} to \emph{pet}, \emph{pit}, \emph{putt}.
-Syllabic \emph{r} patterns with vowels and not other consonants for several phonological processes: it is the only consonant which does not block variable glottalization of following /t/ in British dialects \citep[258]{Harris1994}.
-%
-%\begin{example}[/t/-\textsc{Glottalization} in British English]
-%\begin{tabular}{l l l l@{} l l l}
-%a. & {des}[ɚt]    & \alt{} &   & {des}[ɚʔ]    \\
-%   & {c}[ɚt]{ain} & \alt{} &   & {c}[ɚʔ]{ain} \\
-%b. & {fi}[st]     & \alt{} & * & {fi}[sʔ]     \\
-%   & {mi}[st]{er} & \alt{} & * & {mi}[sʔ]{er} \\
-%\end{tabular}
-%\end{example}
-%
-%\noindent Syllabic \emph{r} is the only consonant which does not trigger variable deletion of a following /t, d/ in American dialects \citep[8]{Guy1980}.
-%
-%\begin{example}[/t, d/-\textsc{Deletion} in American English]
-%\begin{tabular}{l l l l@{} l l l}
-%a. & {be}[lt] & \alt{} &   & {be}[l] \\
-%   & {me}[nd] & \alt{} &   & {me}[n] \\
-%b. & {sh}[ɚt] & \alt{} & * & {sh}[ɚ] \\
-%   & {c}[ɚd]  & \alt{} & * & {c}[ɚ]  \\
-%\end{tabular}
-%\end{example}
-%
-%\subsubsection{Lexical statistics}
-%
-%CELEX contains 6,876 simplex words of English, comprising 21 unique medial codas and 40 unique medial onsets.
-Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested.
-%
-%\subsection{Evaluating derived constraints}
-%
-%\emph{SPE} describes three phonological alternations which target syllable contact clusters.
-The corresponding constraints on simplex syllable contact clusters are considered below.
-The first such derived constraint, obstruent voice assimilation, is presented as a tutorial introduction.
-The second set of constraints consists of three static phonotactic constraints proposed by \citet{Pierrehumbert1994}.
-%
-%\subsubsection{Obstruent voice assimilation} 
-%\label{ova}
-%
-%
-%\begin{example}[Inflectional affix voice assimilation]
-%\begin{tabular}{l l l l l l}
-%a. & {nap} & {nap}[t] & & {nab} & {nab}[d] \\
-%b. & {lap} & {lap}[s] & & {lab} & {lab}[z] \\
-%\end{tabular}
-%\end{example}
-%
-%\noindent The devoicing of the two voiced obstruent suffixes can be formalized as a 
-%rule spreading the voicing specification of an obstruent rightward.
-%
-%\begin{example}[\textsc{Obstruent Voice Assimilation}]
-%%\xymatrix@R=24pt@C=24pt{
-%%\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{--}[dr] &                               \\
-%%\txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
-%%\txt{[$-$\textsc{Sonorant}]}                  & \txt{[$-$\textsc{Sonorant}]} \\
-%%}
-%
 %It is quite apparent that root-internal hetero-voiced obstruent clusters like [s.b] or [z.t] are also rare compared to clusters which are uniformly voiced or uniformly voiceless, as predicted the assimilation rule.
@@
 To answer this question, the 720 clusters in which both the final coda consonant and initial onset consonant are obstruent are sorted into four bins, according to whether they are attested or not, and whether or not the conform to the derived constraint---both obstruents are either voiced or voiceless, like in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}, respectively---or violate it, like the aforementioned \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}.
-%
-%This table, shown in (\ref{ovatable}), is then submitted to the Fisher Exact Test. This statistical test is so named because it is isomorphic to the chi-square test 
-%for contingency tables, but computes an exact $p$-value, whereas the chi-square test $p$-value depends on an approximation which is inappropriate for small samples.
-The small $p$-value produced by the Fisher test indicates that the rarity of the disagreeing clusters is unlikely to be due to chance.
-The simplest interpretation is that \textsc{Obstruent Voice Assimilation} applies in non-derived environments, preventing learners from positing underlying representations in which adjacent obstruents do not agree in voice.
-%
-%\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}] \label{ovatable}
-%\begin{tabular}{l r r r r}
-%\toprule
-%           & attested & unattested & \% saturation & $p$-value                   \\
-%\midrule
-%conforming & 80       & 370        & 18\%          & \multirow{2}{*}{1.1\e{-11}} \\
-%violating  &  6       & 264        & 2\%           \\
-%\bottomrule
-%\end{tabular}
-%\end{example}
-%
-%There are two possible analyses of the small number of roots that violate the derived generalization.
-Some hetero-voiced obstruent clusters, like the one in \emph{ja}[k.d]\emph{aw}, for instance, might be analyzed as compounds by native speakers, and this analysis might itself be driven by the rarity of morph-internal adjacent hetero-voiced obstruents (cf.
-\citealp[546]{Rice2009d} on compounds in Slave).
-In fact, \citet{Mattys2001b} find that 9-month-old infants are sensitive to the contrast between [f.t] and exceptional (and unattested) [v.t], treating the latter as an indication of a word boundary and this heuristic is also available to adults \citep{Brown1956,McQueen1998b}.
-Less explanatorily, exceptional roots can be marked [$-$\textsc{Obstruent Voice Assimilation}].
-%
-%\subsubsection{Nasal place assimilation} 
-%The shape of the prefix before the velar stops /k, g/ is somewhat more complex.
-\citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but that assimilation of \emph{im-}/\emph{in-} is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}.
-However, the stress condition does not hold for simplex words (e.g., \emph{a}[ŋ.ɡ]\emph{óra}), and is ignored here.
-%
-%There is evidence that even young infants acquiring English are aware of this generalization.
-\citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} find that infants at 4.5 months, 9 moths, and 10 months of age, respectively, prefer to listen to nonce words which contain the homorganic nasal-obstruent clusters (e.g., \emph{u}[m.b]\emph{o}) over those which contain heterorganic clusters (e.g., \emph{u}[n.b]\emph{o}).
-Nasal-obstruent clusters arising in speech errors also undergo \textsc{Nasal Place Assimilation} (e.g.
-\emph{ra}[nd] \emph{orker}, intended \emph{ra}[ŋk] \emph{order}; \citealt[228]{Myers1993}).
-%
-%\citet[175]{Pierrehumbert1994} observes that ``nasal-stop sequences agree in labiality'', but makes no reference to \textsc{Nasal Place Assimilation}.
-Exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} do occur, but they are much rarer than homorganic nasal-obstruent clusters such as \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey}.
-%
-%\begin{example}[Lexical effects of \textsc{Nasal Place Assimilation}]
-%\begin{tabular}{l r r r r}
-%\toprule
-%           & attested & unattested & \% saturation & $p$-value                   \\
-%\midrule
-%conforming & 41       & 11         & 79          & \multirow{2}{*}{2.6\e{-05}} \\
-%violating  & 8        & 20         & 29                                    \\
-%\bottomrule
-%\end{tabular}
-%\end{example}
-%
-%\subsubsection{Degemination} 
-%\label{deg}
-%
-%A final alternation targeting English syllable contact clusters is the simplification of geminates derived by certain affixes.
-For instance, the deadjectival adverbial suffix \emph{-ly} rarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{fu}[l]\emph{y}, cf.
-\emph{free}[l]\emph{y}).
-Degemination is also found in /d/-final roots selecting the irregular /-t/ past tense suffixes (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]).
-More complicated cases of \textsc{Degemination} can be found in Latinate prefix allomorphy (see \emph{SPE}:148 and \citealt[102]{Borowsky1986}).
-Identical adjacent consonants do not occur in the corpus, nor do obstruent clusters which differ only in voice (e.g., *[p.b]), which would be made identical by the operation of \textsc{Obstruent Voice Assimilation}.
-As shown by the results of the Fisher test, this absence is unlikely to be due to chance.
-%
-%\begin{example}[\textsc{Degemination}]
-%$\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ =\textsc{Nas} \\ =\textsc{Cont} \end{bmatrix}~\goesto~\zero~/~\gap~\begin{bmatrix} =\textsc{Lab} \\ =\textsc{Cor} \\ =\textsc{Dor} \\ =\textsc{Son} \\ =\textsc{Nas} \\ =\textsc{Cont} \end{bmatrix}$
-%\end{example}
-%
-%% consider what coda-final consonants there are
-%
-%\begin{example}[Lexical effects of \textsc{Degemination}]
-%\begin{tabular}{l r r r r}
-%\toprule
-%           & attested & unattested & \% saturation & $p$-value                   \\
-%\midrule
-%conforming & 157      & 596        & 21      & \multirow{2}{*}{7.2\e{-09}} \\
-%violating  & 0        &  87        & 0                                    \\
-%\bottomrule
-%\end{tabular}
-%\end{example}
-%
-%\subsubsection{Local summary}
-%
-%The three \emph{SPE} rules affecting syllable consonant clusters are reliably reflected in the lexicon.
-Possible clusters that are surface-exceptions to these three rules are much more likely to be unattested than those that conform to them.
-%
-%\subsection{Evaluating static constraints}
-%
-%The second set of constraints, proposed by \citet{Pierrehumbert1994}, are ``static'' in that they have no analogue in English morphophonemics.
-Clusters of all length are considered here, not only the triconsonantal clusters considered by \citeauthor{Pierrehumbert1994}, as this restriction appears to be arbitrary.
-%
-%\subsubsection{Dorsal/non-coronal consonants}
-%
-%\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Two-consonant clusters like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} are found, however.
-Despite this, dorsal/non-coronal clusters are no less likely to occur than non-dorsal codas with non-coronal onsets (e.g.
-\emph{oi}[nt.m]\emph{ent}) or dorsal codas with coronal onsets (e.g., \emph{ve}[k.t]\emph{or}).
-%
-%\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}]
-%\begin{tabular}{l r r r r}
-%\toprule
-%           & attested & unattested & \% saturation & $p$-value \\
-%\midrule
-%conforming & 76 & 359 & 18 & \multirow{2}{*}{0.145} \\
-%violating  &  6 &  57 & 10 \\
-%\bottomrule
-%\end{tabular}
-%\end{example}
-%
-%\subsubsection{Coda coronal obstruents}
-%
-%\citet[175]{Pierrehumbert1994} proposes that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many clusters such as \emph{a}[t.l]\emph{as} or \emph{e}[s.t]\emph{er}.
-Once again, coda coronal obstruent clusters are no less likely to occur than non-coronal obstruent clusters like \emph{re}[p.t]\emph{ile} or coda coronal sonorant clusters such as \emph{co}[n.d]\emph{or}.
-%
-%\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}]
-%\begin{tabular}{l r r r r}
-%\toprule
-%           & attested & unattested & \% saturation & $p$-value \\
-%\midrule
-%conforming & 48       & 312        & 13            & \multirow{2}{*}{0.301} \\
-%violating  & 38       & 322        & 11            \\
-%\bottomrule
-%\end{tabular}
-%\end{example}
-%
-%\subsubsection{ABA clusters}
-%
-%Finally, \citet[][176]{Pierrehumbert1994} observes ``a lack of clusters with identical first and third elements''.
-Here, this is operationalized in the same fashion as \textsc{Degemination}, by definining ``identity'' as consonants which share the same place and manner features, but which do not necessarily agree on voicing.
-Though the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters.
-%
-%\begin{example}[Lexical effects of \textsc{*ABA}]
-%\begin{tabular}{l r r r r}
-%\toprule
-%           & attested & unattested & \% saturation & $p$-value \\
-%\midrule
-%conforming & 41       & 478        & 8             & \multirow{2}{*}{0.818} \\
-%violating  &  0       &  11        & 0             \\
-%\bottomrule
-%\end{tabular}
-%\end{example}
-%
-%\subsubsection{Local summary}
-%
-%There is no statistically reliable evidence for the three static constraints proposed by \citeauthor{Pierrehumbert1994}.
-However, it is premature to conclude that no static constraint exists.
-It is possible that an explicit model of ``constraint discovery'', such as those proposed by \citet{Pierrehumbert1994} and \citet{Hayes2008a}, could identify reliable static constraints, and predict which of the possible clusters are attested and unattested beyond the derived constraints.
-%
-%\subsection{Evaluating computational models}
-%
-%The computational models of \citeauthor{Pierrehumbert1994} and \citeauthor{Hayes2008a}, as well as two baselines, are scored for their ability to predict which of the possible medial clusters are attested and which are not.
-For any single cluster, there are four possible outcomes.
-%
-%\begin{example}[Outcomes in cluster classification task]
-%\begin{tabular}{c l | l l}
-%                   & & \multicolumn{2}{c}{\textbf{actual outcome}}            \\
-%                   & & \emph{attested} & \emph{unattested}            \\
-%\midrule
-%\textbf{predicted} & \emph{attested}   & true positive ($tp$)  & false positive ($fp$) \\
-%\textbf{outcome}   & \emph{unattested} & false negative ($fn$) & true negative ($tn$)  \\
-%\end{tabular}
-%\end{example}
-%
-%\noindent Accuracy is the probability of a correct classification, whether a true positive or negative.
-%
-%\begin{unlabeledexample}
-%$\displaystyle \textrm{Accuracy} = \frac{tp + tn}{tp + tn + fp + fn}$
-%\end{unlabeledexample}
-%
-%\noindent Precision is the probability that a cluster predicted to occur is attested.
-%
-%\begin{unlabeledexample}
-%$\displaystyle \textrm{Precision} = \frac{tp}{tp + fp}$
-%\end{unlabeledexample}
-%
-%\noindent Recall (or sensitivity) is the probability an attested cluster is predicted to occur.
-%
-%\begin{unlabeledexample}
-%$\displaystyle \textrm{Recall} = \frac{tp}{tp + fn}$
-%\end{unlabeledexample}
-%
-%\noindent It is possible to maximize recall at the expense of precision, by predicting attestation for a greater number of clusters, or to increase precision at the expense of recall by predicting non-attestation for a greater number of clusters.
-$F_1$, the harmonic mean of these two measures, quantifies this trade-off.
-%
-%\begin{unlabeledexample}
-%$\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$
-%\end{unlabeledexample}
-%
-%A soft-margin support vector machine \citep{Cortes1995} with a linear kernel is used to convert from the per-cluster scores to attestation predictions, by finding a single cutoff which optimally separates attested and unattested clusters according to their model scores.
-This is in no way intended to be a cognitively plausible model for learning phonotactics; it simply represents an upper bound on the performance of these models.
-The results for all four models are summarized in Table \ref{cmresults}.
-%
-%\begin{table} \centering
-%\begin{tabular}{l | r r r r | r}
-%\toprule
-%                     & accuracy & precision & recall & $F_1$ \\
-%\midrule
-%Null baseline        &     .808 &      .808 &  1.000 &  .894 \\
-%\citet{Hayes2008a}   &     .837 &      .948 &   .852 &  .898 \\ %& .002       \\
-%Expected frequency   &     .835 &      .845 &   .983 &  .909 \\ %& 1.1\e{-05} \\
-%Derived constraints  &     .840 &      .922 &   .885 &  .903 \\
-%DC and EF            &     .864 &      .871 &   .989 &  .926 \\ %& 1.9\e{-13} \\
-%\bottomrule
-%\end{tabular}
-%\caption{FIXME}
-%\label{cmresults}
-%\end{table}
-%
-%\subsubsection{Null baseline}
-%
-%19\% of the 880 possible clusters are attested.
-Consequently, a null baseline which predicts that no clusters are possible will be 81\% accurate, trivially.
-Because this baseline FIXME RECALL FIXME PRECISION
-%
-%\subsubsection{Derived constraints}
-%
-%The three derived constraints define a structured baseline, in which attestation is the predicted outcome for all clusters which not violate the \emph{SPE}-derived generalizations, and non-attestation otherwise.
-This results in a significant improvement to the null baseline (sign test $p = 0.006$).
-This model has poor recall, however, because many well-formed clusters are unattested.
-This is not a flaw, but rather the signature of accidental gaps.
-%
-%\subsubsection{Expected frequency}
-%
-%\citet{Pierrehumbert1994} proposes a simple model in which the wellformedness of syllable contact clusters is proportional to the joint probability, estimated from word-final coda and word-initial onset frequencies and assuming that coda and onset are statistically independent.
-There are two problems with this procedure, however.
-First, there exist languages, such as Finnish \citep[36]{Fischer-Jorgensen1952}, in which there are medial codas not found in final position; this alone suggests that ``outer'' frequencies are not the best model of what is possible medially.
-More to the point, a pilot study determined that using medial coda and onset frequencies produced a better fit, and this variant is adopted here.
-%%\citeauthor{Pierrehumbert1994}'s procedure would assign clusters of this type the lowest score. 
-%While \citeauthor{Pierrehumbert1994} reports that this is the best predictor of which complex clusters occur, it imposes no constraints on sequences spanning the syllable boundary, and consequently has lower accuracy (0.839) and $F_1$ (0.328) than the derived constraint baseline, though it is a small but significant improvement over the null baseline (sign test $p = 1.23$\e{-04}).
-%
-%\subsubsection{\citealt{Hayes2008a}}
-%
-%\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints.
-Since the model has many experimenter-determined parameters, a direct replication of the experiments reported by \citeauthor{Hayes2008a} is attempted; their software, feature specification, and model settings are all used.
-Since the training of this model is inherently stochastic, producing slightly different outcomes on each run, the results reported here from the lowest accuracy of ten independent runs, a practice also used by \citeauthor{Hayes2008a}.
-The model achieves the highest accuracy (0.877) and $F_1$ (0.703), but lower recall (0.642) than the alternation baseline, a significant improvement over the null baseline (sign test $p = 2.9$\e{-05}), but not significantly different from the performance of the alternation baseline (sign test $p = 0.070$).
-%
-%The generalizations extracted by the \citeauthor{Hayes2008a} model are in general quite similar to those derived from the \emph{SPE} alternations, but there are some notable differences.
-Unattested (and ill-formed) *[m.kl] is assigned the highest score, though it is ruled out by \textsc{Coda Nasal Place Assimilation}.
-On the other hand, very low scores are assigned to the clusters in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.
-%
-%\subsubsection{\citealt{McGowan2009}}
-%
+
 %\citet{McGowan2009} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2009} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
@@
 As a predictor of cluster attestation, sonority distance provided no improvement over the null baseline; consequently, it is not included in Table \ref{cmresults}.
-%
-%\subsubsection{Local summary}
-%
-%Compared to phonological alternations, current computational models of phonotactic learning are not significantly better predictors of which syllable contact clusters a
-%
-%\section{Conclusions}
-%
-%None of the foregoing results have revealed any evidence for structural constraints on the English syllable contact cluster, whether ``discovered'' by linguist or computer, except those derived from phonological alternations.
-What then is to be said about the countless clusters which do not violate any phonologically derived constraint, but yet are unattested, like [b.z] or [z.n]?
-Since these gaps appear to be quite arbitrary from a phonological perspective, all that can be said at this juncture is that these clusters must be regarded \emph{accidental gaps}: they are well-formed clusters, but simply missing from the English lexicon due to the sparsity thereof, a possibility that has long been recognized.
-%
-%\begin{quote}
-%\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language.
-\citep[][16]{Fischer-Jorgensen1952}
-%
-%The material will never be complete. It will always contain accidental gaps\ldots{}partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
-%\end{quote}
-%
-%\noindent Below, attempts are made to quantify the extent of accidental gaps in this domain below.
-%
-%\subsection{Zipfian distribution}
-%
-%A number of previous studies \citep[e.g.,][]{Weiss1961,Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phoneme or letter frequencies exhibit sparse distributions, in which a few types account for a large amount of the probability mass, and the remaining is divided up among a large number of rare types.
-The same is true for the type frequencies of the medial codas and onsets which make up syllable contact clusters, as well as cluster frequencies themselves.
-In Figure \ref{clus}, cluster frequencies are plotted against rank in the log-log space, and a near-linear relationship obtains ($R^2 = 0.924$), indicating that the data conform to a generalization of Zipf's (\citeyear{Zipf1949}) Law.
-See Appendix \ref{zr} for estimation details.
-%
-%\begin{figure} \centering
-%\includegraphics{cluster.pdf}
-%\caption{Syllable contact clusters exhibit a log-log linear relationship between frequency and rank that is consistent with Zipf's Law.}
-%\label{clus}
-%\end{figure}
-%
-%Sparse distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2011b} to word \citep{Baroni2009} and phoneme \citep{Belevitch1956,Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Mandelbrot1954,Miller1957,Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}.
-The importance of sparsity in this context is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps.
-%
-%\subsection{Good-Turing estimate}
-%
-%\citet{Good1953} proposes an estimate for the probability of all accidental gaps, $\hat{p}_0$, as the number of events that occur just once, $n_1$ divided by the number of observations.
-%
-%\begin{unlabeledexample}
-%$\displaystyle \hat{p}_0 = \frac{n_1}{\displaystyle\sum N}$
-%\end{unlabeledexample}
-%
-%\noindent In the CELEX data, $65$ clusters occur only once, and there are 873 cluster tokens, and thus $\hat{p}_0 = 0.074$.
-There is thus a non-trivial probability that many of the ``false positive'' clusters are in fact accidental gaps.
-%
-%\subsection{Sampling simulation}
-%
-%It is finally possible to ask what the lexicon might look like if it was stochastically generated from independent coda and onset frequencies, \emph{a la} \citeauthor{Pierrehumbert1994}, but also constrained by phonological neutralizations.
-The following method was repeated to generate a simulated lexicon.
-%
-%\begin{example}[Simulation procedure]
-%\begin{tabular}{l l}
-%a. & Sample a medial coda according to the observed probabilities  \\
-%b. & Sample a medial onset according to the observed probabilities \\
-%c. & Apply all matching \emph{SPE} rules to cluster formed by their concatenation \\
-%\end{tabular}
-%\end{example}
-%
-%The cluster frequencies for a single characteristic simulation are plotted in Figure \ref{sim}, along with observed frequencies; the two distributions are nearly indistinguishable.
-The sparse cluster inventory, previously taken as evidence for static constraints on syllable contact, is approximately what one would expect even if the only constraints on possible clusters are phonologically derived.
-%
-%\begin{figure}
-%\centering
-%\includegraphics{sim.pdf}
-%\caption{The observed cluster frequencies are closely matched by a simulated cluster inventory generated by random sampling, assuming independence and applying phonological neutralizations.
-The $Z_r$ transform \citep[][29]{Church1991} has been used to eliminate quantization in lower frequency bands.}
-%\label{sim}
-%\end{figure}

### 2012-11-05 ack.tex
Writing on 2-4 and appendices
Writing on chapter 2 primarily, also added the two "MSes" and new graphs for chapter 2.

@@
 So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy.
-More than anyone else, I owe a debt of gratitude to Charles Yang, who has done (and put up with) so much for me.
-I am grateful to my committee members, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment for quantitative and theoretical approaches to meet.
-And though his name does not appear again in this document, so much here has been inspired by Bill Labov.
 
-The faculty, staff, and students of the Institute For Research in Cognitive Science have provided me with not merely an office, but a home, complete with dated early-90s decor.
+The ideas I first encountered under the tutelage of Jennifer Cole and Richard Sproat at the University of Illinois continue to guide my research.
+
+The faculty, staff, and students of the Institute For Research in Cognitive Science provided me with not merely an office, but something close to a home.
 Thanks to my officemates, Chandan Narayan, John Trueswell, Laurel Sweeney and the rest of the staff there, as well as Amy Forsyth back at Williams Hall.
@@
 Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar.
+John Murdock fearlessly proofread everything, twice.
 
-The ideas I first encountered under the tutelage of Jennifer Cole and Richard Sproat at the University of Illinois continue to guide my research.
-Insofar as I have acknowledged historical precursors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee.
-Morris Halle taught me to speak clearly and to never apologize.
+I am grateful to my committee members, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment for quantitative and theoretical approaches to meet.
+
+Insofar as I have acknowledged my historical precursors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee.
+
+More than anyone else, I owe a debt of gratitude to Charles Yang: I don't know how he puts up with me.
+
+I would like to conclude by thanking three people who were \emph{not} involved in the preparation of this dissertation, but without which it could never exist: Lila Gleitman, Morris Halle, and Bill Labov.

### 2012-11-05 ms2.tex
Writing on chapter 2 primarily, also added the two "MSes" and new graphs for chapter 2.

\documentclass[12pt]{article}
\usepackage[hmargin=1in,top=1in,bottom=1in]{geometry}

% bibliography stuff
\usepackage{url}
\usepackage{natbib}
\bibpunct[:]{(}{)}{,}{a}{}{,}

% normalem
\usepackage[normalem]{ulem}

% graphics
\usepackage{graphicx}

% tables
\usepackage{longtable}
\usepackage{multirow}
\usepackage{booktabs}

% my example environments
\usepackage{simplex}

% nice feature matrices
\usepackage{amsmath}

% fonts
\usepackage{mathspec}
\setmainfont[Mapping=tex-text]{Linux Libertine}
\setmathfont(Digits,Greek,Latin){Linux Libertine}

%% KG's shortcuts
\newcommand{\buf}{\hspace{0pt}}
\newcommand{\gap}{\rule{1em}{0.5pt}}
\newcommand{\alt}{\ensuremath{\sim}}
\newcommand{\asp}{\textsuperscript{h}}
\newcommand{\pal}{\textsuperscript{j}}
\newcommand{\zero}{\ensuremath{\emptyset}}
\newcommand{\goesto}{\ensuremath{\longrightarrow}}
\providecommand{\e}[1]{\textsc{e}{$#1$}}

\title{Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}} 
\author{Kyle Gorman \\ University of Pennsylvania}
\date{November 2012 (comments welcome)}

\begin{document}
\maketitle

\abstract{Gradient ratings are a property of ratings tasks, not of the grammatical system itself.}

\input{2}
\appendix
\renewcommand{\arraystretch}{0.25}

\section{English wordlikeness ratings}
\label{ratings}

\subsection{\citet{Albright2007}}

\begin{longtable}{r@{ }r@{ }c@{ }l@{ }l rrrrrr}
\toprule
   &   &     &   &    & lexical & $-$log $p$ & $-$log $p$ & gross  & rating    \\
   &   &     &   &    & density & (bigram)   & (MaxEnt)   & status & (7-point) \\
\midrule
P & L & IY1 & N &    & 13 & 13.585 & 0.000 &   valid & 5.32 \\
B & L & AA1 & D &    & 13 & 17.609 & 0.000 &   valid & 5.13 \\
P & L & IY1 & K &    & 11 & 14.200 & 0.000 &   valid & 5.06 \\
P & L & EY1 & K &    & 14 & 15.576 & 0.000 &   valid & 4.94 \\
P & R & AH1 & N & JH & 3 & 16.546 & 0.000 &   valid & 4.94 \\
B & L & UW1 & T &    & 14 & 15.692 & 0.000 &   valid & 4.84 \\
P & L & IH1 & M &    & 5 & 15.126 & 0.000 &   valid & 4.71 \\
B & L & EH1 & M &  P & 1 & 19.447 & 0.000 &   valid & 4.69 \\
B & L & AH1 & S &    & 14 & 13.806 & 0.000 &   valid & 4.67 \\
B & L & AE1 & D &    & 15 & 16.259 & 0.000 &   valid & 4.65 \\
B & L & IH1 & G &    & 4 & 16.347 & 0.000 &   valid & 4.58 \\
P & R & EH1 & S &  P & 4 & 17.214 & 0.000 & invalid & 4.50 \\
B & R & EH1 & N & TH & 4 & 21.255 & 0.000 &   valid & 4.11 \\
P & R & AH1 & P &  T & 4 & 18.487 & 0.000 &   valid & 4.07 \\
B & R & EH1 & L & TH & 2 & 23.014 & 0.000 &   valid & 3.14 \\
P & W & IH1 & S &  T & 4 & 21.499 & 0.000 &   valid & 2.94 \\
B & W & AH1 & D &    & 2 & 20.596 & 0.000 &   valid & 2.94 \\
B & W & AA1 & D &    & 3 & 21.329 & 0.000 &   valid & 2.94 \\
P & W & AE1 & D &    & 2 & 24.103 & 0.000 &   valid & 2.89 \\
P & W & AH1 & S &    & 4 & 20.684 & 0.000 &   valid & 2.61 \\
P & W & EH1 & T &    & 6 & 20.998 & 0.000 &   valid & 2.53 \\
P & T & IY1 & N &    & 4 & 15.440 & 6.762 & invalid & 2.44 \\
B & W & AE1 & D &    & 2 & 23.365 & 0.000 &   valid & 2.41 \\
B & N & IY1 & N &    & 2 & 21.180 & 7.296 & invalid & 2.39 \\
P & W & AH1 & D &  Z & 0 & 25.210 & 0.000 &   valid & 2.17 \\
P & N & IY1 & N &    & 2 & 21.181 & 6.019 & invalid & 2.16 \\
B & N & AH1 & S &    & 6 & 19.732 & 7.296 & invalid & 2.06 \\
P & N & EH1 & P &    & 2 & 23.587 & 6.019 & invalid & 2.00 \\
B & N & AA1 & D &    & 2 & 24.066 & 7.296 & invalid & 2.00 \\
B & Z & IY1 & N &    & 1 & 16.896 & 19.097 & invalid & 2.00 \\
P & T & AH1 & S &    & 3 & 14.169 & 6.762 & invalid & 1.94 \\
P & T & EH1 & P &    & 3 & 17.228 & 6.762 & invalid & 1.86 \\
B & Z & AH1 & S &    & 2 & 15.237 & 19.097 & invalid & 1.81 \\
P & N & IY1 & K &    & 2 & 21.796 & 6.019 & invalid & 1.76 \\
B & D & IY1 & K &    & 2 & 18.773 & 13.131 & invalid & 1.72 \\
B & D & UW1 & T &    & 3 & 19.781 & 13.131 & invalid & 1.71 \\
B & D & AH1 & S &    & 4 & 17.041 & 13.131 & invalid & 1.71 \\
P & T & AE1 & D &    & 3 & 17.622 & 6.762 & invalid & 1.67 \\
B & Z & AA1 & D &    & 1 & 20.151 & 19.097 & invalid & 1.63 \\
B & Z & AY1 & K &    & 1 & 19.118 & 19.097 & invalid & 1.28 \\
\bottomrule
\end{longtable}

\subsection{\citet{Albright2003b}, norming study}

\begin{longtable}{r@{ }r@{ }r@{ }c@{ }l@{ }l@{ } rrrrr}
\toprule
   &    &    &    &    &   & lexical & $-$log $p$ & $-$log $p$ & gross  & rating    \\
   &    &    &    &    &   & density & (bigram)   & (MaxEnt)   & status & (7-point) \\
\midrule
  &  S &  L & EY1 & M  &    & 15 & 17.469 & 0.000 &   valid & 5.84 \\
  &    &  W & IH1 & S  &    & 34 & 11.208 & 0.000 &   valid & 5.84 \\
  &    &  P & IH1 & N  & T  & 26 & 13.046 & 0.000 &   valid & 5.67 \\
  &    &  P & AE1 & NG & K  & 18 & 13.723 & 0.000 &   valid & 5.63 \\
  &  S &  T & IH1 & P  &    & 18 & 12.599 & 0.000 &   valid & 5.53 \\
  &    &  M & IH1 & P  &    & 33 & 12.345 & 0.000 &   valid & 5.47 \\
  &  S &  T & AY1 & R  &    & 11 & 15.118 & 0.000 &   valid & 5.47 \\
  &    &  M & ER1 & N  &    & 34 & 12.872 & 0.000 &   valid & 5.42 \\
  &  P &  L & EY1 & K  &    & 14 & 15.576 & 0.000 &   valid & 5.39 \\
  &  S &  N & EH1 & L  &    & 10 & 18.582 & 0.000 &   valid & 5.32 \\
  &  S &  T & IH1 & N  &    & 18 & 10.899 & 0.000 &   valid & 5.28 \\
  &    &  R & AE1 & S  & K  & 11 & 15.544 & 0.000 &   valid & 5.21 \\
  &  T &  R & IH1 & S  & K  &  5 & 17.980 & 0.000 &   valid & 5.21 \\
  &  S &  P & AE1 & K  &    & 17 & 14.205 & 0.000 &   valid & 5.16 \\
  &    &  D & EY1 & P  &    & 22 & 14.193 & 0.000 &   valid & 5.11 \\
  &    &  G & EH1 & R  &    & 25 & 13.044 & 0.000 &   valid & 5.11 \\
  &  G &  L & IH1 & T  &    & 14 & 16.830 & 0.000 &   valid & 5.11 \\
  &  S &  K & EH1 & L  &    & 16 & 16.356 & 0.000 &   valid & 5.11 \\
  &    & SH & ER1 & N  &    & 23 & 15.913 & 0.000 &   valid & 5.11 \\
  &    &  T & AA1 & R  & K  & 18 & 17.702 & 0.000 &   valid & 5.11 \\
  &    & CH & EY1 & K  &    & 28 & 15.023 & 0.000 &   valid & 5.05 \\
  &  G &  L & IY1 & D  &    & 14 & 16.118 & 0.000 &   valid & 5.05 \\
  &  G &  R & AY1 & N  & T  &  4 & 17.626 & 0.000 &   valid & 5.00 \\
  &  P &  R & IY1 & K  &    & 11 & 13.396 & 0.000 &   valid & 5.00 \\
  &    & SH & IH1 & L  & K  &  8 & 21.270 & 0.000 &   valid & 4.89 \\
  &    &  D & AY1 & Z  &    & 39 & 12.730 & 0.000 &   valid & 4.84 \\
  &    &  N & EY1 & S  &    & 23 & 14.952 & 0.000 &   valid & 4.84 \\
  &    &  T & AH1 & NG & K  & 18 & 15.046 & 0.000 &   valid & 4.84 \\
S &  K &  W & IH1 & L  &    &  6 & 18.210 & 0.000 &   valid & 4.83 \\
  &    &  L & AH1 & M  &    & 35 & 11.569 & 0.000 &   valid & 4.79 \\
  &    &  P & AH1 & M  &    & 30 & 11.121 & 0.000 &   valid & 4.79 \\
S &  P &  L & IH1 & NG &    & 14 & 15.573 & 0.000 &   valid & 4.72 \\
  &  G &  R & EH1 & L  &    &  3 & 14.624 & 0.000 &   valid & 4.63 \\
  &    &  T & EH1 & SH &    & 12 & 14.517 & 0.000 &   valid & 4.63 \\
  &    &  T & IY1 & P  &    & 32 & 12.980 & 0.000 &   valid & 4.63 \\
  &    &  B & AY1 & Z  &    & 35 & 12.821 & 0.000 &   valid & 4.58 \\
  &  G &  L & IH1 & P  &    & 11 & 17.377 & 0.000 &   valid & 4.53 \\
  &    & CH & AY1 & N  & D  & 18 & 17.747 & 0.000 &   valid & 4.37 \\
  &  P &  L & IH1 & M  &    &  5 & 15.126 & 0.000 &   valid & 4.37 \\
  &    &  G & UW1 & D  &    & 29 & 15.448 & 0.000 &   valid & 4.32 \\
  &  B &  L & EY1 & F  &    &  6 & 19.485 & 0.000 &   valid & 4.21 \\
  &    &  G & EH1 & Z  &    & 17 & 16.466 & 0.000 &   valid & 4.21 \\
  &  D &  R & IH1 & T  &    &  8 & 15.563 & 0.000 &   valid & 4.16 \\
  &  F &  L & IY1 & P  &    & 10 & 15.292 & 0.000 &   valid & 4.16 \\
  &    &  Z & EY1 &    &    & 23 & 15.208 & 0.000 &   valid & 4.16 \\
S &  K &  R & AY1 & D  &    &  5 & 18.722 & 0.000 &   valid & 4.11 \\
  &    &  K & IH1 & V  &    & 16 & 12.591 & 0.000 &   valid & 4.05 \\
  &  F &  L & EH1 & T  &    & 17 & 16.490 & 0.000 &   valid & 4.00 \\
  &    &  N & OW1 & L  & D  & 19 & 19.101 & 0.000 &   valid & 4.00 \\
  &  S &  K & IH1 & K  &    & 13 & 14.628 & 0.000 &   valid & 4.00 \\
  &  B &  R & EH1 & JH &    &  7 & 17.318 & 0.000 &   valid & 3.95 \\
  &  K &  W & IY1 & D  &    & 10 & 16.039 & 0.000 &   valid & 3.95 \\
  &  S &  K & OY1 & L  &    &  9 & 19.350 & 0.000 &   valid & 3.89 \\
  &  D &  R & AY1 & S  &    & 12 & 17.758 & 0.000 &   valid & 3.84 \\
  &  F &  L & IH1 & JH &    &  8 & 17.312 & 0.000 &   valid & 3.79 \\
  &  B &  L & IH1 & G  &    &  4 & 16.347 & 0.000 &   valid & 3.53 \\
  &    &  Z & EY1 & P  & S  &  7 & 24.825 & 0.000 &   valid & 3.47 \\
  &    & CH & UW1 & L  &    & 17 & 14.492 & 0.000 &   valid & 3.42 \\
  &    & SH & AY1 & N  & T  &  8 & 18.503 & 0.000 &   valid & 3.42 \\
  & SH &  R & UH1 & K  & S  &  5 & 26.733 & 0.000 &   valid & 3.32 \\
  &  G &  W & EH1 & N  & JH &  0 & 22.722 & 0.000 &   valid & 3.32 \\
  &    &  N & AH1 & NG &    & 19 & 15.754 & 0.000 &   valid & 3.28 \\
S &  K &  W & AA1 & L  & K  &  1 & 25.752 & 0.000 & invalid & 3.26 \\
  &  T &  W & UW1 &    &    &  5 & 17.918 & 0.000 &   valid & 3.17 \\
  &  S &  M & AH1 & M  &    &  8 & 14.940 & 0.000 &   valid & 3.05 \\
  &  S &  N & OY1 & K  & S  &  4 & 32.283 & 4.136 & invalid & 3.00 \\
  &  S &  F & UW1 & N  & D  &  1 & 23.241 & 3.507 &   valid & 2.94 \\
  &  P &  W & IH1 & P  &    &  4 & 20.928 & 0.000 &   valid & 2.89 \\
  &    &  R & AY1 & N  & T  &  8 & 14.412 & 0.000 &   valid & 2.89 \\
S &  K &  L & UW1 & N  & D  &  0 & 22.661 & 0.000 & invalid & 2.83 \\
  &  S &  M & IY1 & R  & G  &  0 & 27.601 & 0.000 & invalid & 2.79 \\
  &  F &  R & IH1 & L  & G  &  3 & 24.299 & 0.000 & invalid & 2.68 \\
  & SH &  W & UW1 & JH &    &  0 & 28.270 & 0.000 & invalid & 2.68 \\
  & TH &  R & OY1 & K  & S  &  0 & 32.485 & 4.136 & invalid & 2.68 \\
  &  T &  R & IH1 & L  & B  &  4 & 22.097 & 0.000 & invalid & 2.63 \\
  &  K &  R & IH1 & L  & G  &  1 & 23.719 & 0.000 & invalid & 2.58 \\
  &  S &  M & EH1 & R  & G  &  0 & 22.473 & 0.000 & invalid & 2.58 \\
  & TH &  W & IY1 & K  & S  &  2 & 23.984 & 0.000 & invalid & 2.53 \\
  &  S &  M & EH1 & R  & F  &  0 & 23.136 & 0.000 & invalid & 2.47 \\
  &  S &  M & IY1 & L  & TH &  0 & 26.377 & 0.000 & invalid & 2.47 \\
  &  P &  L & OW1 & M  & F  &  0 & 23.336 & 0.000 & invalid & 2.42 \\
  &  P &  L & OW1 & N  & TH &  0 & 22.805 & 0.000 & invalid & 2.26 \\
  &    & TH & EY1 & P  & T  &  4 & 23.380 & 0.000 &   valid & 2.26 \\
  &  S &  M & IY1 & N  & TH &  0 & 25.043 & 0.000 &   valid & 2.06 \\
S &  P &  R & AA1 & R  & F  &  0 & 24.031 & 0.000 &   valid & 2.05 \\
  &  P &  W & AH1 & JH &    &  0 & 23.205 & 0.000 &   valid & 1.74 \\
\bottomrule
\end{longtable}

\subsection{\citet{Scholes1966}, experiment 5}

\begin{longtable}{r@{ }r@{ }c@{ }l@{ } rrrrr}
\toprule
  &   &    &     & lexical & $-$log $p$ & $-$log $p$ & gross  & rating   \\
  &   &    &     & density & (bigram)   & (MaxEnt)   & status & (binary) \\
\midrule 
 G & R & AH1 & N & 18 & 11.799 & 0.000 &   valid & 33 \\
 K & R & AH1 & N & 21 & 11.597 & 0.000 &   valid & 33 \\
 S & T & IH1 & N & 18 & 10.899 & 0.000 &   valid & 33 \\
 S & M & AE1 & T & 13 & 16.654 & 0.000 &   valid & 32 \\
 P & R & AH1 & N & 11 & 10.845 & 0.000 &   valid & 32 \\
 S & L & ER1 & K & 12 & 17.846 & 0.000 &   valid & 31 \\
 F & L & ER1 & K & 11 & 17.456 & 0.000 &   valid & 31 \\
 B & L & AH1 & NG & 8 & 17.156 & 0.000 &   valid & 31 \\
 D & R & AH1 & NG & 7 & 17.753 & 0.000 &   valid & 31 \\
 T & R & AH1 & N & 12 & 11.975 & 0.000 &   valid & 31 \\
 F & R & AH1 & N & 12 & 12.177 & 0.000 &   valid & 29 \\
 S & P & EY1 & L & 16 & 15.851 & 0.000 &   valid & 29 \\
 S & N & EH1 & T & 7 & 19.384 & 0.000 &   valid & 28 \\
 P & L & AH1 & NG & 11 & 16.960 & 0.000 &   valid & 28 \\
SH & R & AH1 & K & 8 & 19.734 & 0.000 &   valid & 27 \\
 G & L & AH1 & NG & 9 & 18.990 & 0.000 &   valid & 27 \\
 M & R & AH1 & NG & 1 & 22.888 & 3.365 & invalid & 27 \\
SH & L & ER1 & K & 4 & 23.711 &  0.000 & invalid & 22 \\
 S & K & IY1 & P & 15 & 16.845 & 0.000 &   valid & 20 \\
 V & R & AH1 & N & 4 & 17.087 &  0.000 & invalid & 19 \\
 S & R & AH1 & N & 9 & 16.626 &  0.000 & invalid & 14 \\
 V & L & ER1 & K & 2 & 21.777 &  0.000 & invalid & 14 \\
 M & L & AH1 & NG & 4 & 21.300 & 10.164 & invalid & 13 \\
SH & T & IH1 & N & 3 & 17.106 & 0.000 & invalid & 13 \\
 F & P & EY1 & L & 4 & 24.250 & 3.685 & invalid & 13 \\
ZH & R & AH1 & N & 4 & 28.305 & 4.042 & invalid & 11 \\
F & SH & IH1 & P & 2 & 22.640 & 10.198 & invalid & 11 \\
SH & N & EH1 & T & 2 & 24.044 &  0.000 &   valid & 10 \\
 F & T & IH1 & N & 2 & 14.767 &  3.685 & invalid & 10 \\
 Z & R & AH1 & N & 5 & 21.556 &  4.042 & invalid & 8 \\
 N & R & AH1 & N & 5 & 18.588 &  3.365 & invalid & 8 \\
SH & M & AE1 & T & 1 & 20.389 &  0.000 &   valid & 7 \\
 S & F & IY1 & D & 7 & 18.656 &  3.701 &   valid & 7 \\
 Z & L & ER1 & K & 2 & 24.578 &  5.678 & invalid & 6 \\
 Z & T & IH1 & N & 1 & 23.600 &  5.678 & invalid & 6 \\
 F & S & EH1 & T & 4 & 19.079 & 10.198 & invalid & 6 \\
 V & Z & IH1 & P & 1 & 17.401 & 19.601 & invalid & 6 \\
 V & Z & AH1 & T & 1 & 15.806 & 19.601 & invalid & 6 \\
ZH & L & ER1 & K & 2 & 33.442 &  5.678 & invalid & 5 \\
SH & F & IY1 & D & 1 & 23.258 &  3.701 & invalid & 5 \\
 Z & N & AE1 & T & 1 & 25.541 &  5.678 & invalid & 4 \\
 F & N & EH1 & T & 2 & 23.969 &  3.315 & invalid & 3 \\
 F & K & IY1 & P & 1 & 23.905 &  3.685 & invalid & 3 \\
 V & T & IH1 & N & 2 & 22.639 &  3.685 & invalid & 3 \\
 Z & V & IY1 & L & 2 & 26.018 & 15.023 & invalid & 3 \\
 Z & M & AE1 & T & 1 & 21.983 &  5.678 & invalid & 2 \\
ZH & M & AE1 & T & 1 & 26.800 &  5.678 & invalid & 2 \\
 F & M & AE1 & T & 4 & 21.800 &  3.315 & invalid & 2 \\
SH & P & EY1 & L & 2 & 26.172 &  0.000 & invalid & 2 \\
 V & M & AE1 & T & 2 & 20.388 &  3.315 & invalid & 1 \\
 V & N & EH1 & T & 2 & 24.017 &  3.315 & invalid & 1 \\
SH & K & IY1 & P & 2 & 26.976 &  0.000 & invalid & 1 \\
 Z & P & EY1 & L & 1 & 25.421 &  5.678 & invalid & 1 \\
ZH & P & EY1 & L & 1 & 32.906 &  5.678 & invalid & 1 \\
ZH & T & IH1 & N & 1 & 29.763 &  5.678 & invalid & 1 \\
ZH & K & IY1 & P & 1 & 33.710 &  5.678 & invalid & 1 \\
ZH & N & EH1 & T & 1 & 33.775 &  5.678 & invalid & 0 \\
 Z & K & IY1 & P & 1 & 27.547 &  5.678 & invalid & 0 \\
 V & P & EY1 & L & 2 & 25.782 &  3.685 & invalid & 0 \\
 V & K & IY1 & P & 1 & 26.586 &  3.685 & invalid & 0 \\
ZH & V & IY1 & L & 1 & 32.181 & 15.023 & invalid & 0 \\
\bottomrule
\end{longtable}

\section{English syllabification}
\label{syllabification}

For every entry in the CELEX database, there is a corresponding broad syllabified transcription of the word in a Received Pronunciation accent. This appendix describes an automated procedure used to process these transcripts and to separate medial clusters from their flanking nuclei, parsing the resulting sequences into coda and onset, and reversing allophonic processes targeting medial clusters.

While the segmental content of these transcriptions is precise, the CELEX syllabifications are unsystematic. Given the absence of contrastive syllabification in English (if not all languages: see \citealp[221]{Blevins1995}, \citealp{Elfner2006}), any sequence of a medial consonant cluster and its flanking nuclei should receive the same syllabification in all words in which it occurs. This is not always the case with the CELEX transcriptions, however. For instance, the sequence [ɪstɹɪ] receives a different parse in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is lax in Received Pronunciation \citep[][II.294]{AOE}.}

\subsection{Ambiguous segments}

The syllabification procedure begins by separating sequences of vocalic and consonantal segments. In English, \emph{r} and onglides pattern with consonants or with vowels depending on the context in which they occur. The heuristic adopted here is that ambiguous segments which impose restrictions on adjacent vowels are themselves vocalic, and those which impose restrictions on adjacent consonants are consonantal.

Initially, between two vowels, or finally, \emph{r} is consonantal. Before another consonant, however, \emph{r} has been lost in Received Pronunciation. Even in \emph{r}-ful dialects, though, post-vocalic non-onset \emph{r} patterns with vowels, not coda consonants. Before non-onset \emph{r} many vowel contrasts are suspended (e.g., \citealp[269f.]{Fudge1969}, \citealp[][255]{Harris1994}): compare American English \emph{fern}/\emph{fir}/\emph{fur} to \emph{pet}/\emph{pit}/\emph{putt}. In this position, \emph{r} is the only consonant which permits variable glottalization of a following /t/ in \emph{r}-ful British dialects \citep[258]{Harris1994}, and the only consonant which does does not trigger variable deletion of a following word-final /t, d/ in American dialects \citep[8]{Guy1980}. This is shown in (\ref{tglottalization}--\ref{tddeletion}) below.

\begin{example}[/t/-\textsc{Glottalization} in \emph{r}-ful British dialects]
\label{tglottalization}
\begin{tabular}{l l l l@{} l l l}
a. & {des}[ɚt]    & \alt{} &   & {des}[ɚʔ]    \\
   & {c}[ɚt]{ain} & \alt{} &   & {c}[ɚʔ]{ain} \\
b. & {fi}[st]     & \alt{} & * & {fi}[sʔ]     \\
   & {mi}[st]{er} & \alt{} & * & {mi}[sʔ]{er} \\
\end{tabular}
\end{example}

\begin{example}[/t, d/-\textsc{Deletion} in American English]
\label{tddeletion}
\begin{tabular}{l l l l@{} l l l}
a. & {be}[lt] & \alt{} &   & {be}[l] \\
   & {me}[nd] & \alt{} &   & {me}[n] \\
b. & {sk}[ɚt] & \alt{} & * & {sk}[ɚ] \\
   & {th}[ɚd] & \alt{} & * & {th}[ɚ] \\
\end{tabular}
\end{example}

\noindent
Consequently, pre-consonantal \emph{r} is assigned to the preceding nucleus.

The front onglide is assigned to onset position when initial or preceded by a single consonant, as in [j]\emph{arn} or \emph{ju}[n.j]\emph{or}. When the glide is preceded by two or more consonants, it is assigned to the nucleus. There is considerable evidence in support of this assumption. When [j] is assigned to the onset, it may be followed by any vowel \citep[276]{Borowsky1986}, but when it is nuclear, the following vowel is always [u], suggesting a nuclear affiliation (\citealp[61f.]{Harris1994}, \citealp[232]{Hayes1980}). \citet[42]{Clements1983} note that [j] is the only consonant which can follow onset /m/ and /v/: [mj]\emph{use}, [vj]\emph{iew}. Finally, [ju] sequences in words such as \emph{spew} behave as a unit in language games \citep{Davis1995,Nevins2003} and speech errors \citep[130]{Shattuck-Hufnagel1986}.\footnote{The glide is also assumed to be present in underlying representation (e.g., \citealp{Anderson1988b}, \citealp[278]{Borowsky1986}) rather than inserted by rule (e.g., \emph{SPE}:196, \citealp[][89]{Halle1985a}, \citealp[][217]{McMahon1990}) since presence or absense of the glide is contrastive (e.g., \emph{booty}/\emph{beauty}, \emph{coot}/\emph{cute}).}

The phonotactic properties of the back onglide [w] are quite different than those of the front onglide, and it is consequently assigned to the onset portion of medial clusters. Whereas [j] shows only limited selectivity for preceding tautosyllabic consonants \citep{Kaye1996}, [w] only rarely occurs after onset consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}), and never after tautosyllabic labials in the native vocabulary. Whereas [kj] is always followed by [u], [kw] may precede nearly any vowel \citep[161]{Davis1995}.

\subsection{Parsing medial consonant clusters}

Medial consonant clusters are segmented into coda and onset using a heuristic version of the principle of onset maximization (e.g., \citealp[42f.]{Kahn1976}, \citealp{Kurylowicz1948}, \citealp[75]{Pulgram1970}, \citealp[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset. A medial onset is defined to be ``possible'' simply if it occurs word-initially (according to the rules defined above). As an example, the medial clusters in words such as \emph{neu}[.tɹ]\emph{on} or \emph{bi}[.stɹ]\emph{o} also occur in word-initial position (e.g., [tɹ]\emph{ain}, [stɹ]\emph{ike}), so the entire cluster is assigned ot the onset. In contrast, the cluster in \emph{mi}[n.stɹ]\emph{el} is not found word-initially; the maximal onset here is [stɹ] and the remaining [n] is assigned to the preceding coda.

In English, when a medial consonant cluster is preceded by a stressed lax vowel, as \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[3]{Hammond1997}. As \citet[55]{Harris1994} notes, however, when the medial cluster is also a valid onset, as in \emph{whi}[s.p]\emph{er}, \emph{ve}[s.ti]\emph{ge}, and \emph{mu}[s.k]\emph{et}, onset maximization will incorrectly assign the entire cluster to the onset and leave the lax vowel unchecked. For this reason, onset maximization parses are modified to assign the first consonant of a complex medial consonant cluster to the coda before a stressed lax vowel \citep[48]{Pulgram1970}.

\subsection{Phonologization}

The traditional analysis of affricates as single segments (e.g., \emph{SPE}:321f., \citealp[24]{Jakobson1961}) rather than stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990} is assumed here. In many languages, affricates pattern with simple onsets; for instance, Classical Nahua bans true onset clusters but permits the affricate series [ts, tʃ, tɬ] \citep[9]{Launey2011}. In English, however, affricates do not form complex onsets. Yet, the stop and release phase of affricates cannot be separated by a syllable boundary, as predicted from the assumption they are single phonological units.

In English, [ŋ] has been analyzed as a pure allophone of /n/ before underlying /k, ɡ/ (with later deletion of /ɡ/ in some contexts; \citealt[65f.]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}), or as a phoneme in its own right \citep[e.g.,][]{Jusczyk2002,Sapir1925}. Onset [ŋ] is totally absent in onset position, where it cannot be followed by a /k, ɡ/ needed to derive the velar allophone, a fact predicted only by the former account, and English speakers have considerable difficulty producing initial [ŋ] \citep{Rusaw2009}. The allophonic analysis is assumed here. When followed by /k, ɡ/, [ŋ] is mapped to /n/. When not followed by a velar stop (i.e., finally), [ŋ] is analyzed as underlying /nɡ/.

%\section{Exemplification}
% 
%In (\ref{exemplification}), a few segmentations are given so as to demonstrate the principles the principles outlined above.
% 
%\begin{example}[Some examples of medial cluster segmentation]
%\label{exemplification}
%\begin{tabular}{l l l}
%a. & \emph{tranquil} & /\ldots{}n.kw\ldots{}/ & (coda nasal place assimilation, [Cw] onsets) \\
%b. & \emph{
%c. & \emph{
%d. & \emph{
%e. & \emph{
%f. & \emph{
%g. & \emph{
%h. & \emph{
%i. & \emph{
%j. & \emph{
%\end{tabular}
%\end{example}

\bibliography{gorman_diss.bib}
\bibliographystyle{pwpl}
\end{document}

### 2012-11-11 1.tex
Completed writing on 2-4!

@@
 \label{intro}
+
+\section{A program for phonotactic theory}
+
+Theories of phonotactic knowledge should be evaluated by the same stringentcriteria applied to other formal arenas: neither overgeneration nor undergeneration should be permitted.
+It will be shown that the orthodoxy that phonotactic knowledge is in some sense ``probabilistic'' suffers from quite severe over- and undergeneration.
+
+This heuristic has the greatest impact regarding debates about ``abstractness'' of underlying representations. 
+It cannot be said, precisely, that it either excludes or requires any particular type of abstractness.
+
+Proposed restrictions on UR abstractness have been faulted for a number of reasons.
+In some cases, they preclude otherwise-desirable analyses of alternations.
+Another objection that could reasonably made is that any restriction on abstract URs that introduces an assymetry in URs is bad.
+\citep[~chap.~1]{KK77}
 
@@
 The positional probability of a nonce monosyllabic word is derived from the combined probabilities at which segments occur in onset, nucleus, and coda positions in the lexicon of a given language.
+
+\citet{Shibatani1973}, however, argues that morpheme structure constraints cannot account for all wordlikeness constraints.
+In German, for instance, final obstruents devoice, and as a result, there a
+re pairs such  [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' and [ɡʀaːt]-[ɡʀaːdə] `degree(s)'
+ differing only in the plural shape of the root.
+Since the voicing of final obstruents is contrastive, the restriction on th
+e voicing of obstruents cannot be a morpheme structure constraint. 
+Yet, \citeauthor{Shibatani1973} claims that native speakers reject nonce wo
+rds ending in final voiced obstruents ``\emph{on the ground that they end i
+n voiced obstruents}'' (95). 
+Thus, not all constrasts in possible wordhood can be stated as morpheme str
+ucture constraints.
+
+Whereas \citeauthor{Shibatani1973} maintains that morpheme structure constraints are insufficient to account for speakers' knowledge of possible wordhood, others argue that morpheme structure constraints are also unnecessary.
+\citet[297]{Hale1965}, \citet{Kisseberth1970b}, and \citet[212f.]{Postal1968} observe that the structural descriptions of phonological processes often are often reflected in the lexical redundancies in the same language.
+In Russian, there are alternations implicating a process of obstruent voice assimilation, and tautomorphemic obstruent clusters have uniform voicing \citep[283]{A74}.
+\citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that morpheme structure constraints are otiose, as phonological rules triggering alternations impose restrictions on the contents of URs, an effect known as \emph{Stampean occultation}.
+
+Since at least \citet{Pike1947b}, linguists have posited language-specific constraints on the contents of underlying representations.
+These \emph{morpheme structure constraints} were at one time thought to fully account for speakers' knowledge of possible and impossible words (e.g., \citealt{Chomsky1965}, \citeyear[382]{SPE}, \citealt[22f.]{SPR}, \citeyear{
+Halle1962}, \citealt{Stanley1967}).

### 2012-11-11 2.tex
Completed writing on 2-4!

@@
 
-FIXME
-%\begin{quote}
-%Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
-%\end{quote}
-%
-%\noindent
-%The position that the well-formedness of nonce words is consistent the view of syntactic grammaticality taken in \emph{LSLT} and \emph{Aspects} \citep{LSLT,ASPECTS}, where it is claimed that different syntactic violations result in different degrees of ungrammaticality.
-%
-%Most recent discussions of the ``problem of admissibility'' focus on this cline of wellformedness as it is evidenced in wordlikeness tasks.
-%
-%\begin{quote}
-%When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing.
-In the usual case, novel items fall along a gradient cline of acceptability.
-\citep[][9]{Albright2009a}
-%
-%In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
-%\citep[][382]{Hayes2008a}
-%
-%\ldots{}when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
-%\end{quote}
-%
-%\begin{quote}
-%\ldots{}A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
+Recent developments in the theory of phonotactic knowledge are motivated by the claim that speakers' wordlikeness intuitions are inherently \emph{gradient}, i.e., that such intuitions are more granular than implied by a binary contrast between ``possible'' and ``impossible'' words.
+
+\begin{quote}
+When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing. 
+In the usual case, novel items fall along a gradient cline of acceptability. \citep[][9]{Albright2009a}
+
+In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
+\citep[][382]{Hayes2008a}
+
+\ldots{}when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
+\end{quote}
+
+This is not a novel claim.
+Early generative discussions of wordlikeness \citep[e.g.,][]{Chomsky1965,Halle1962} are best remembered for the famous examples [blɪk] and [bnɪk], the former representing a ``possible word'' of English and the latter representing an ``impossible word''.
+A naïve account of this contrast follows from the assumption that segments must be parsed into syllables or subject to further phonological repair (e.g., \citealt[10f.]{Hooper1973}, \citealt[57f.]{Kahn1976}, \citealt{Ito1989a}, \citealt[19f.]{Wolf2009}).
+Unlike some languages (e.g., Morroccan Arabic: \emph{bniqa} `closet'), English does not permit stop-nasal onsets like [bn], so the latter nonce word cannot surface as such. 
+In other words, [bnɪk] is an impossible surface representation in English. 
+However, in \emph{The Sound Pattern of English}, \citet{SPE} introduce a third nonce word, [bznk], and claim that last form is in some sense even less English-like than [bnɪk].\footnote{
+    That wordlikeness judgements depend on language-specific knowledge is apparent given that [bznk] is not impossible in all languages: Imdlawn Tashlhiyt Berber permits whole words consisting of a stop-fricative-nasal-stop sequence (e.g., [tzmt] `it is stifling'; \citealt[112]{Dell1985}).
+    This suggests that wordlikeness judgements depend on language-specific knowledge.
+    It has more recently been suggested some contrasts between unattested clusters can be viewed as cross-linguistic implicational universals \citep[e.g.,][]{Berent2007a,Berent2008a,Berent2009,Berent2007b}, so that, for instance, the acceptance of [bznk] might imply acceptance of [bnɪk].
+    If this is correct, it undermines \citeauthor{SPE}'s position.}
+From this they too conclude that wordlikeness intuitions are gradient.
+
+\begin{quote}
+Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
+\end{quote}
+
+\noindent
+This brings the theory of wordlikeness in line with the view of syntactic grammaticality presented by foundational documents like \emph{The Logical Structure of Linguistic Theory} \citep{LSLT} and \emph{Aspects in the Theory of Syntax} \citep{ASPECTS}, which posits multiple degrees of ungrammaticality arising from different types of syntactic violations.
+\citet{BARRIERS} and \citet{Huang1982} propose further elaborations of this type of theory (see \citealt[43f.]{Schutze1996} for a critique).
+
+It has been recognized, however, that the naïve account of the [blɪk] and [bnɪk] cannot easily be extended to account for these ``numerous other distinctions''.
+
+\begin{quote}
+A defect of current grammatical accounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
 \citep[371]{Shademan2006}
-%\end{quote}
+\end{quote}
 
-\section{What some wellformedness judgements might not be}
+Scientific observations do not necessarily arrive at the appropriate granularity for analysis.
+Since the granularity of measurement and analysis is determined by prior hypotheses about the nature of the system under study, disputes are inevitable as scientific paradigms evolve. 
+This study argues that there are both theoretical and empirical arguments to doubt the implicit hypothesis linking scalar wordlikeness judgements and a gradient model of phonotactics.
+First, intermediate judgements are characteristic of all gradient rating tasks, and therefore these judgements are irrelevant to the question of whether well-formedness is categorical or gradient.
+Secondly, simple baselines better account for gradient well-formedness judgements than current computational models of phonotactic knowledge, suggesting that the gradience observed in these tasks do not derive from grammatical mechanisms.
+
+%are ``\ldots{}under the obligation to say why \emph{what it takes to be} data relevant to the confirmation of its theories \emph{are} data relevant to the confirmation of its theories'' \citep[200]{Fodor1981b}.
+%The naïve account of the [blɪk]/[bnɪk] contrast presented above is a special case of this
+
+\section{Aspects in the theory of gradient grammaticality}
+
+The aforementioned discussions of gradient aspects of wordlikeness judgements take for granted that intermediate judgements require a model of gradient grammaticality. 
+The general position that such data can be taken at face value is known in the cognitive sciences as \emph{naïve realism} \citep{Fodor1981a}, and is not uncontroversial. 
+With respect to wordlikeness, there are numerous reasons to question the wisdom of this position.
+
+\subsection{What some linguistic intuitions might not be}
+
+As first noted by \citet{Chomsky1963b}, speakers experience difficulty processing sentences with multiple center embeddings.
+\citet{Gibson1999} find that speakers rate sentences like (\ref{gibson}a), which is well-formed, less grammatical than (\ref{gibson}b), which is nonsensical.
+
+\begin{example}[A well-formedness illusion]
+\label{gibson}
+\begin{tabular}{l l@{}l}
+a. &    & The patient who the nurse who the clinic had hired admitted met Jack. \\
+b. & *  & The patient who the nurse who the clinic had hired met Jack. \\
+\end{tabular}
+\end{example}
+
+\noindent
+It is informative to consider that this well-known result has had no effect on the theory of syntactic representations, only on the theory of linguistic memory; it is recognized as the product of cognitive restrictions found in non-linguistic domains, as a \emph{task effect}.
+This contrasts with the argument made by \citet{Hayes2000}, that gradient wordlikeness judgements demand an all-encompassing revision to the grammatical architecture, reviewed below.
+
+The results of controlled experiments are often biased by subtle details that seem orthogonal to the task: for instance, certain types of duration judgements are systematically biased by consumption of caffeine \citep{Gruber2005}.
+It should come as no surprise, then, that a highly salient aspect of a judgement task, the scale used for responses, also influences the results obtained.
+\citet*{Armstrong1983} argue that the use of many-valued scales may cause intermediate ratings, and should be interpreted as a task effect.
+
+\citet{Armstrong1983} are concerned with experimental evidence for the nature of cognitive concepts. 
+While they do not attempt to dispute that certain concepts (e.g., \emph{fruit}) have a family-resemblance structure \citep[e.g.,][]{Rosch1975a}, they assert that it is apparent that other concepts are ``definitional'' (i.e., all-or-nothing), a notion which they illustrate with \emph{odd number}.
+
+\begin{quote}
+No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
+\end{quote}
+
+\noindent
+However, when subjects are asked to rate, using a 7-point Likert scale, how representative individual odd counting numbers are of the concept \emph{odd number}, they freely use intermediate ratings; their results with \emph{odd number} and \emph{even number} are shown in Figure \ref{agg}.
 
@@
 
-%The gradedness of a category cannot be observed.
-Scientists choose the granularity at which they observe.
-For example, a linguist may choose to record speakers' judgements on a 7-point scale, for example, or to use a forced binary choice between two minimally different items.
-The level of granularity at which the observations are made have consequences for how statistical inference over these observations should be performed, or for the theory of the \emph{task model}, the mechanism by which speakers perform the task, but it does not have any consequences for the \emph{theory} of phnotactics.
+This suggests that the gradience observed is primarily an artifact of the task itself.
+\citeauthor{Schutze2011} suggests that the nature of this effect might be understood as the result of speakers' attempts to reconcile bizarre experimental tasks with their knowledge.
 
-%\begin{quote}
-%No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
-%\end{quote}
+\begin{quote}
+Putting it another way, when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different.
+\citep[24]{Schutze2011}
+\end{quote}
 
-%\begin{quote}
-%Some have responded to these findings very consistently, by asserting that the experimental findings are to be interpreted as before: that, psychologically speaking, odd numbers as well as birds and vegetables are graded concepts\ldots{} We reject this conclusion just because we could not explain how a person could compute with integers who believed that 7 was odder than 23.
-We assert confidently that the facts about subjects being able to compute and about their being able to give the definition of odd number, etc., are the more important, highly entrenched, facts we want to preserve and explain\ldots{} we ourselves are prepared to give up the seeming fact that some odd numbers appear, as shown by their behavior in certain experimental paradigms, to be odder than others\ldots{}we do not give it up by saying that it was no fact; rather, by saying it must have been a fact about something other than the structure of concepts.
+As \citeauthor{Armstrong1983} observe, these results show that the scalar judgement tasks provide no evidence as to whether the category being rated is categorical or gradient.
+
+\begin{quote}
+\ldots{}we hold that \emph{fruit} and \emph{odd number} have different structures, and yet we obtain the same experimental outcome for both.
+But if the same result is achieved regardless of the concept structure, then the experimental design is not pertinent to the determination of concept structure.
+\citep[284--5]{Armstrong1983}
+\end{quote}
+
+It might be said that these results reveal something about the representation of odd numbers, but no scientist of mathematical cognition has risen to the challenge of constructing a theory that might account for the fact that 447 is rated more odd than 3.
+\citeauthor{Armstrong1983} anticipate this objection.
+
+\begin{quote}
+Some have responded to these findings very consistently, by asserting that the experimental findings are to be interpreted as before: that, psychologically speaking, odd numbers as well as birds and vegetables are graded concepts\ldots{} We reject this conclusion just because we could not explain how a person could compute with integers who believed that 7 was odder than 23.
+We assert confidently that the facts about subjects being able to compute and about their being able to give the definition of odd number, etc., are the more important, highly entrenched, facts we want to preserve and explain\ldots{} \citep[284]{Armstrong1983}
+%we ourselves are prepared to give up the seeming fact that some odd numbers appear, as shown by their behavior in certain experimental paradigms, to be odder than others\ldots{}we do not give it up by saying that it was no fact; rather, by saying it must have been a fact about something other than the structure of concepts.
 \citep[284]{Armstrong1983}
-%\end{quote}
+\end{quote}
 
-%\begin{quote}
-%\ldots{}we hold that \emph{fruit} and \emph{odd number} have different structures, and yet we obtain the same experimental outcome for both.
-But if the same result is achieved regardless of the concept struture, then the experimental design is not pertinent to the determination of concept structure.
-\citep[284--5]{Armstrong1983}
-%\end{quote}
+\subsection{A model of gradient intuitions}
+
+It is possible to draw an analogy: can it be the case that [st] is a significantly ``better'' onset than [bl] (a prediction of the wordlikeness model proposed by \citealt{Albright2009a}, for instance), but ensure that both are treated the same with respect to syllabification?
+Current research on gradient well-formedness is concerned with specifying the component by which a scalar value is assigned to linguistic structures. 
+This is only one part of any model of gradient grammaticality, however.
+Further assumptions are necessary, as follows.
+When presented with a linguistic item (of whatever type) in a well-formedness task, the grammar assigns a parse.
+This grammar must be able to parse an enormous range of linguistic structures, including many which it cannot be permitted to generate; independent perception and production grammars may be necessary.
+A scalar value is then assigned to this parse. 
+Then, speakers consciously access this scalar value, then transform it in accordance with the numerical scale chosen by the experimenter.
+
+Each step of this procedure is dubious, however.
+First, speakers have difficulty perceiving \citep{Brown1956} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
+Secondly, the computation of a scalar value serves no further purpose than to provide for gradient well-formedness judgements; even if this is the result of probabilistic parsing \citep[e.g.,][]{Coleman1997},
+it need not be computed overtly.\footnote{
+    An objection might be made here on evolutionary grounds. 
+    It is still quite mysterious why the human language endowment includes constraints on pronominal binding, for instance, but these constraints are implicated in everyday language use.
+    Far more bizarre is the suggestion that the linguistic endowment includes mechanisms only used in certain experimental tasks.}
+Next, speakers must be able to consciously access and report the magnitude of this value (it must be \emph{cognitively penetrable} in the sense of \citealt{Pylyshyn1984}), an ability which is limited in many other domains.
+Finally, \citet{Sprouse2011} argues that speakers do not (or cannot) scale well-formedness values to satisfy the assumptions magnitude estimation, a type of gradient rating task.
+%\citep{Bard1996}
+
+It is informative to compare this baroque architecture to a model of what is necessary to make a binary well-formedness judgement.
+When presented with a linguistic item in a judgement task, the grammar attempts to assign a parse.
+Speakers then access whether or not parsing was successful.
+There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}.
+As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures may not stored in linguistic memory. 
+%This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
+The fact that requests for repetition and clarification are ubiquituous in spontaneous speech illustrates further that speakers are frequently aware when parsing has failed.
+Consequently, a great deal of evidence is needed to reject this simple model in favor of the gradient grammaticality architecture.
+
+\subsection{Evidencing gradience}
+
+\citet{Hayes2000} argues that it is ``uninsightful'' to attribute gradience to task effects, insofar as these effects call upon grammatical representations.
+
+%\citep[?]{Hayes2000}
+%``much of the patterning of gradient judgments is based on authentic structural aspects of the linguistic material being judged''
+
+\begin{quote}
+\ldots{}patterns of gradient well-formedness often seem to be driven by the very same principles that govern absolute well-formedness\ldots{} I conclude that the proposed attribution of gradient well-formedness judgments to performance mechanisms would be uninsightful.
+Whatever ``performance'' mechanisms we adopted would look startlingly like the grammatical mechanisms that account for non-gradient judgments.
+\citep[99]{Hayes2000}
+\end{quote}
+
+\noindent
+This is indisputable. 
+However, there have been no serious attempts to evaluate categorical and gradient models of wordlikeness on an equal footing.
+In light of the complexities of gradient models, such an evaluation requires strong quantitative evidence for the superiority of gradient grammatical models.
+This study represents a first attempt to fill this gap.
+
+It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared. 
+\citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables.
+Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words.
+Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, many of which are exceptionless, into probabilities.
+While this is consistent with their claim, that ``the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.
+
+%(see \citealt{Sproat1993} and \citealt{Wrench2003} for articulatory investigations).
 
@@
 
-There have been no serious attempts to evaluate categorical and gradient models of wordlikeness on an equal footing.
-This study represents a first attempt to fill this gap.
-It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared.
-For instance, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, transform the \citeauthor{Clements1983} constraints, many of which are exceptionless, into probabilities.
-While this is consistent with their claim that ``the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, such a principle precludes any attempt to confirm or falsify the hypothesis that underlies it.
-
-It is unknown whether the intermediate ratings in gradient wordlikeness tasks are reliably predicted by the computational models that have been proposed.
-If some model poorly accounts for these intermediate ratings, it may be the case that the model is improperly specified, or it may indicate that a considerable amount of variance in wordlikeness ratings is not a product of the phonotactic system per se.
-
-Imagine, alternatively, that wordlikeness judgements could be effectively modeled with a gross contrast between possible and impossible words.
-
-%[st] [bl]
+It is unknown whether the intermediate ratings in gradient wordlikeness tasks are reliably predicted by computational models that have been proposed. 
+If some model poorly accounts for these intermediate ratings, it may be the case that the model is improperly specified, or it may indicate that a considerable amount of variance in wordlikeness ratings is not a product of the grammatical system per se.
+It seems particularly plausible that speakers might differentiate, in a regular fashion, between different types of ``impossible'' words, as proposed by \citet{SPE}. 
+There are also claims that speakers distinguish between different types of ``possible'' words, so that, for instance, [stɪn] \emph{stin} is rated more English-like than [blɪn] \emph{blin} \citep{Albright2009a}.
+Alternatively, wordlikeness judgements could be effectively modeled with a gross contrast between possible and impossible words; even then, a gradient model might show a residual correlation.
+All of these possibilities are considered here.
 
@@
 Two criteria were used to select these three studies.
-First, it was required that the stimuli be presented aurally so as to eliminate any possibility of orthographic effects.
+First, it was required that the stimuli be presented aurally so as to eliminate any possibility of orthographic effects \citep[e.g.,][]{Berent2001b,Berent2008b}.
 Secondly, the data must be sufficiently ``phonotactically diverse'': that is, it must include both items like \emph{blick} and \emph{bnick}.
-This excludes studies like that of \citet{Bailey2001}, in which few if any items contain gross phonotactic violations of the type represented by \emph{bnick}.
+This excludes studies like that of \citet{Bailey2001}, in which few if any items contain gross phonotactic violations of the type represented by \emph{bnick}, since this is likely to minimize the coverage of any grammatical model.
 The data is summarized in Table \ref{counts}.
@@
 \citet{Albright2007} administers a wordlikeness task in which 68 adult speakers rate 40 monosyllabic nonce words, presented aurally, on a 7-point Likert scale with endpoints labeled  ``completely impossible as an English word'' and ``would make a fine English word''.
-\citealt{Albright2007}'s study study is primarily concerned with the effects of different onset types (e.g., well-formed /bl/, marginal /bw/, unattested /bn, bd, bz/), and there is less diversity among the choice of rimes, none of which are obviously ill-formed.
+\citeauthor{Albright2007}'s study is primarily concerned with the effects of different onset types (e.g., well-formed /bl/, marginal /bw/, unattested /bn, bd, bz/), and there is less diversity among the choice of rimes, none of which are obviously ill-formed.
 
@@
 The data used here is his experiment 5, in which 33 speakers provide a ``yes'' or ``no'' as to whether each of the 63 items, presented aurally, are ``likely to be usable as a word of English''.
-Like the study by \citet{Albright2007}, the focus was on onset well-formedness and there is minimal diversity in rime type.
+Like the study by \citet{Albright2007}, the focus is on onset well-formedness and there is minimal diversity in rime type.
 Two items, [klʌŋ] \emph{clung} and [bɹʌŋ] \emph{brung} (a dialectical past participle of \emph{bring}), are excluded here as actual words of English.
@@
 Models are evaluated by comparing their scores to the average rating of each word using four correlation statistics.
-\citet{Hayes2008a} evaluate their model using the Pearson $r$, a parametric measure of correlation.
-\citet{Stevens1946} argues that statistics of this type are inappropriate for analysis of Likert scale data, like those used by \citet{Albright2009a} and \citet{Albright2003b}.
-The reason is that this makes a \emph{linearity assumption}.
+Each of these range between $[-1, 1]$, where $1$ indicates a perfect positive correlation and $-1$ denotes a perfect negative correlation.
+\citet{Hayes2008a} evaluate their model using the Pearson (``product-moment'') $r$, a parametric correlation measure.
+\citet{Stevens1946} argues that statistics of this type are inappropriate for analysis of Likert scale data, like those used by \citet{Albright2007} and \citet{Albright2003b}.
+The reason is that the Pearson $r$ makes a \emph{linearity assumption}.
 That is, it assumes that nonce words rated ``1'' and  ``3'', for instance, are just as different as those ``4'' and a ``6''.
-A weaker assumption, more appropriate for Likert scale data, is the ``monotonicity assumption'': that ``1'' is less English-like than ``3'', which is less English-like than ``4'', and so on.
+A weaker assumption, more appropriate for Likert scale data, is the \emph{monotonicity assumption}: that ``1'' is less English-like than ``3'', which is less English-like than ``4'', and so on.
 However, it also has been claimed that $r$ is particularly robust to violations of the linearity assumption \citep[e.g.,][]{Havlicek1976}.
@@
 \citeauthor{Hayes2008a} also report Spearman $\rho$; this statistic requires only the weaker assumption of monotonicity, but it is difficult to give a simple interpretation to the coefficient.
-Much easier to interpret are two non-parametric statistics, the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ \citep{Noether1981}; a variant of the latter is used by \citet{Albright2009a}.
+
+Much easier to interpret are two other non-parametric statistics, the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ \citep{Noether1981}, as follows.
 These statistics are computed by comparing every model score/wordlikeness rating pair to every other: a comparison is counted as \emph{concordant} if the greater of the two model scores is the one associated with the greater of the two wordlikeness ratings (that is, the model ranks these two nonce words the in accordance with speakers' ratings), and as \emph{discordant} otherwise.
-$\gamma$ and $\tau_b$ differ in the treatment of ``ties'', pairs where either the model score or wordlikeness rating are identical.
+These two statistics differ only in the treatment of ``ties'', pairs where either the model score or wordlikeness rating are identical.
 For $\gamma$, ties are ignored, and the coefficient is
 
-\begin{unlabeledexample}
-$\gamma = \displaystyle\frac{c - d}{c + d}$
-\end{unlabeledexample}
+\begin{equation*}
+\gamma = \displaystyle\frac{c - d}{c + d}
+\end{equation*}
 
@@
 The $\tau_b$ statistic uses a similar formula, but also incorporates a penalty for ties in model score which are not also paired with ties in wordlikeness ratings, or vis versa.
+\citet{Albright2009a} uses a variant of this statistic to evaluate wordlikeness models.
 
@@
 This is not to imply that these are the only domains over which phonotactic violations might be stated, but there are prior claims that onset and rime are particularly important domains for stating phonotactic constraints (e.g., \citealt{Fudge1969}, \citealt{Kessler1997}, \citealt{Treiman2000}).
-Furthermore, these units are implicated by language games \citep{Treiman1983} and speech errors \citep{Fowler1993}, and adult speakers are adept at separating syllables into these units \citep{Treiman1986,Treiman1995}.
+Speakers are adept at separating syllables into these units \citep{Treiman1983,Treiman1986,Treiman1995}, and they are implicated by patterns of speech errors \citep{Fowler1987,Fowler1993}.
 
@@
 The simplest possible mechanism is chosen here: an onset or rime is identified as well-formed if it occurs with non-zero frequency in a representative sample, and is identified as ill-formed otherwise.
-The sample is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms, the latter though to be particualrly strongly correlated with behavioral measures \citep{Brysbaert2009}.
+The sample is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms, the latter though to be particularly strongly correlated with behavioral measures \citep{Brysbaert2009}.
 These pronunciations are then syllabified, and individual syllables parsed into onset and rime, according to a process described in detail in Appendix \ref{syllabification}.
 This is not is not to imply that all unattested onsets or rimes should be regarded as ill-formed, or that all onsets or rimes with non-zero frequency in this data are well-formed.
-For instance, \citet{Albright2009a} judges [drEsp] to be an unobjectionable possible word of English, despite the total absense of [Esp] rimes in English, and similar observations have been made regarding English onsets \citep[e.g.,][]{Cairns1972,Moreton2002}.
+For instance, \citet{Albright2009a} judges [dɹɛsp] \emph{dresp} to be phonotactically well-formed, despite the total lack of [ɛsp] rimes in English.
+Similar observations have been made concerning English onsets \citep[e.g.,][]{Cairns1972,Moreton2002}.
+In the other direction, there are precedents for labeling certain attested words as phonotactically ``peripheral'' (e.g., \citealt{Myers1987}, \citealt{Borowsky1989}), as lexical exceptions to language-specific prosodic constraints, without labeling these words as absolutely ungrammatical.
+Neither of these dissociations between attestation and well-formedness are implemented here, however.
 
@@
 
-A second baseline is provided by measures of similarity to existing English words items, which has long been applied in the study of wordlikeness (e.g., \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Shademan2006,Shademan2007}, \citealt{Vitevitch1998,Vitevitch1999a}).
-\citeauthor{LSLT} (\citeyear{LSLT}: 151, fn.~27) suggests that grammaticality judgements in general might be influenced by similarity to existing grammatical structures.
-\citet[417f.]{SPE} outline a similarity-based wordlikeness model.
+A second baseline is provided by measures of similarity to existing English words items, which has long been applied to model wordlikeness judgements (e.g., \citealt{Bailey2001}, \citealt{Greenberg1964}, \citealt{Kirby2007}, \citealt{Ohala1986b}, \citealt{Shademan2006,Shademan2007}, \citealt{Vitevitch1998,Vitevitch1999a}).
+\citeauthor{LSLT} (\citeyear{LSLT}: 151, fn.~27) suggests that grammaticality judgements in general might be influenced by similarity to existing grammatical structures, and \citet[417f.]{SPE} outline a similarity-based wordlikeness model.
 More recently, it has been observed (e.g., \citealt[51]{Coleman1997}, \citealt{Hay2004a}) that nonce words which flagrantly violate English sonority restrictions but which bear common affixes (e.g., *\emph{mrupation}) are rated highly English-like.
@@
 A wide variety of lexical similarity measures were considered, including a variant of including a variant of the Generalized Neighborhood model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a set of measures provided by the Irvine Phonotactic Calculator \citep{Vaden2009}.
-The most reliable measure is also the most veneraable measure of lexical similarity: Coltheart's $N$ \citep{Coltheart1977}, which is defined as the number of words in some representative sample which can be changed into a target nonce word by a single insertion, deletion, or substitution of a phone.
+The most reliable measure is also the most venerable measure of lexical similarity: Coltheart's $N$ \citep{Coltheart1977}, which is defined as the number of words in some representative sample which can be changed into a target nonce word by a single insertion, deletion, or substitution of a phone.
 \citet{Greenberg1964} find a correlation between wordlikeness ratings and a variant of this measure which only counts words differing by a single substitution.
 This measure is plotted against ratings from the three studies in Figure \ref{neighborhood}, where it can be seen that it accounts for much of the variance in ratings.
+
+%The effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production \citep[e.g.,][]{Luce1998,Luce2000} and perception \citep{Marslen-Wilson1984,Marslen-Wilson1987,McClelland1986,Norris1994,Norris2000}.
 
@@
 While there is nothing inherently ``phonotactic'' about Coltheart's $N$, it indirectly incorporates much of the information present in the gross phonotactic baseline. 
-Consider [blIk]: since there is nothing marked about any part of this nonce word, a ``neighbor'' might be found by modifying any phone: e.g., \emph{click}, \emph{brick}, \emph{bloke}, \emph{bliss}.
-However, since [bn] onsets are unattested in English, a neighbor of [bnIk] must somehow modify this cluster: this leaves only \emph{brick} and \emph{nick}.
-\citet{Frauenfelder1993} find that neighborhood density is also strongly correlated with measures like bigram probability.
+Consider [blɪk]: since there is nothing marked about any part of this nonce word, a ``neighbor'' might be found by modifying any phone: e.g., \emph{click}, \emph{brick}, \emph{bloke}, \emph{bliss}. 
+However, since [bn] onsets are unattested in English, a neighbor of [bnɪk] must somehow modify this cluster: this leaves only \emph{brick} and \emph{nick}. 
+\citet{Bailey2001} and \citet{Frauenfelder1993} note that neighborhood density is also strongly correlated with measures like bigram probability, but it is argued elsewhere that phonotactic measures and neighborhood density have distinct effects \citep[e.g.,][]{Berent2003,Pitt1998b,Vitevitch1998,Vitevitch1999a}.
 
@@
 
-Bigrams probabilities have been applied to model results of a number of psycholinguistic tasks \citep[e.g.,][]{Pitt1998,Vitevitch2005}, and \citet{Albright2009a} applies them as a model of wordlikeness judgements.
+Faciliatory effects of bigram probabilities (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements \citep{Lipinski2005,Luce2001,Vitevitch1999a,Vitevitch2005}, and lexical decision \citep{Pylkkanen2002a}.
+\citet{Albright2009a} applies them as a model of wordlikeness judgements.
 The bigram probability of a sequence $ijk$, for instance, is:
 
-\begin{unlabeledexample}
-$\displaystyle \hat{p}(ijk) = p(i|\textrm{start}) \cdot p(j|i) \cdot p(k|j) \cdot p(\textrm{stop}|k)$
-\end{unlabeledexample}
+\begin{equation*}
+\displaystyle \hat{p}(ijk) = p(i|\textsc{start}) \cdot p(j|i) \cdot p(k|j) \cdot p(\textsc{stop}|k)
+\end{equation*}
 
@@
 However, \citeauthor{Albright2009a}'s evaluation, which includes the \citet{Scholes1966} and \citet{Albright2003b} data, finds an advantage for segmental bigrams.
-In implementing this model, it was found that a slight improvement could be made by preventing any phone-to-phone transition from have zero probability.
-This is accomplished by adding 1 to the count of every transition, a technique known as Laplace, or ``add one'', smoothing.
+In implementing this model, it was found that a slight improvement could be made by preventing any phone-to-phone transition from having zero probability.
+This is accomplished by adding 1 to the count of every transition, a technique used in natural language processing and known as Laplace, or ``add one'', smoothing.
 As can be seen in Table \ref{bigramcomparison}, this results in a slight increase in the correlation between the scores from this model and wellformedness ratings.
@@
 \centering
-\begin{tabular}{l rrr}
+\begin{tabular}{l rrrr}
 \toprule
-                         & features & segments & smoothing \\
+                                 & Pearson $r$ & Spearman $\rho$ & G-K $\gamma$ & Kendall $\tau_{b}$ \\
 \midrule
-Pearson $r$              & {.711}   & {.742}   & \uline{.746} \\
-Spearman $\rho$          & {.628}   & {.667}   & \uline{.699} \\
-Goodman-Kruskal $\gamma$ & {.453}   & {.479}   & \uline{.502} \\
-Kendall $\tau_{b}$       & {.448}   & {.473}   & \uline{.499} \\
+featural  bigrams                &       {.71} &       {.64}     &       {.45}  &       {.45} \\
+segmental bigrams                &       {.74} &       {.67}     &       {.48}  &       {.47} \\
+segmental bigrams with smoothing & \uline{.75} & \uline{.70}     & \uline{.50}  & \uline{.50} \\
 \bottomrule
 \end{tabular}
-\caption{Correlation between item-averaged wordlikeness ratings for the \citet{Albright2003b} norming study and three variants of bigram probability; ``features'': featural bigram probability reported by \citet{Albright2009a}, ``segments'': segment bigram probability reported by \citet{Albright2009a}, ``smoothing'': segment bigram probabilities estimated using Laplace smoothing}
+\caption{Correlation between item-averaged wordlikeness ratings for the \citet{Albright2003b} norming study and three variants of bigram probability}
 \label{bigramcomparison}
@@
 \citeauthor{Hayes2008a} use a complex method to evaluate their model.
-First, they extract onset sequences from the CMU pronunciation dictionary which they use to train the model.
+First, they extract onset sequences from the CMU pronunciation dictionary, and use these to train the model.
 The model is then used to score the onsets of the \citet{Scholes1966} nonce words.
-Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting corelation.\footnote{This is contrary to standard practices in natural language process, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter), and thus there is no indication that the model will generalize to new data.
-Consequently, no transformation is used here.
-This only has an effect on the Pearson $r$ coefficient, since they use a transformation that preserves monotonicity.}
+Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting correlation.\footnote{This is contrary to standard practices in natural language process, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter); when this is the case, there is reason to suspect the parameter values will not generalize to new data.
+No transformation is used here; this only has an effect on the Pearson $r$ coefficient, since they use a transformation that preserves monotonicity.}
 
-An attempt was made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications were all used.
+An attempt is made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications are all used.
 However, \citet{Albright2009a} finds that the maximum entropy model, training and testing only on onsets, does not generalize to the \citet{Albright2003b} data, presumably because of the considerable diversity of rimes in the latter data.
 Consequently, the model was trained to score whole words, not just onsets, using the subset of the CMU dictionary described above.
-Following the procedure of \citet{HayesInPress}, dictionary entries were syllabified and the resulting syllables were parsed into onset, nucleus, and coda, and a feature [$\pm$\textsc{Coda}] was added to contrast onset and coda consonants.
-Since the maximum entropy model, producing slightly different predictions on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}.
+Following the procedure of \citet{HayesInPress}, dictionary entries were syllabified and the resulting syllables were parsed into onset, nucleus, and coda, and a novel feature [$\pm$\textsc{Coda}] is added to contrast onset and coda consonants.
+Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}.
 The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to large variety of nonce words, though many of these words appear to have a low rating.
@@
 Table \ref{scores} displays the full set of correlation coefficients, for each of the three data sets, and for each of the four models.
-For each pair of statistic and data set, the highest coefficient is underlined.
 The first observation is that in general, there is a positive correlation between model score and ratings in each pair.
-However, the two baselines, gross phonotactic status and neighborhood density, are by far the strongest models across statistics and and data, with gross phonotactic status performing the strongest under the Goodman-Kruskal $\gamma$ and on the \citet{Albright2007} data, and neighborhood density performing strongly under nearly all other statistics and data sets.
+The two baselines, gross phonotactic status and neighborhood density, are by far the strongest models across statistics and studies, with gross phonotactic status performing the strongest under the Goodman-Kruskal $\gamma$ and on the \citet{Albright2007} data, and neighborhood density performing strongly under nearly all other statistics and data sets.
 
@@
 \centering
-\begin{tabular}{l rrrr}
+\begin{tabular}{l rrr c rrr c rrr c rrr}
 \toprule
-                           & gross status & density ($N$) & bigram score & MaxEnt score \\
+              &         \multicolumn{3}{c}{Pearson $r$} &&     \multicolumn{3}{c}{Spearman $\rho$} && \multicolumn{3}{c}{G-K $\gamma$} && \multicolumn{3}{c}{Kendall $\tau_b$} \\
+\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12} \cmidrule{14-16}
+              &           A &         AH &            S &&           A &       AH    &           S &&           A &          AH &           S &&           A &          AH & S \\
 \midrule
-Pearson $r$               \\
-\midrule
-\citeauthor{Albright2007}  & \uline{.725} & {.673}        & {.463}       & {.703} \\
-\citeauthor{Albright2003b} & {.594}       & \uline{.793}  & {.746}       & {.208} \\
-\citeauthor{Scholes1966}   & {.803}       & \uline{.855}  & {.737}       & {.534} \\
-\midrule
-Spearman $\rho$           \\
-\midrule
-
-\citeauthor{Albright2007}  & \uline{.819} & {.612}        & {.338}       & {.661} \\
-\citeauthor{Albright2003b} & {.662}       & \uline{.739}  & {.699}       & {.389} \\
-\citeauthor{Scholes1966}   & {.803}       & \uline{.817}  & {.794}       & {.575} \\
-\midrule
-Goodman-Kruskal $\gamma$  \\
-\midrule
-\citeauthor{Albright2007}  & \uline{.874} & {.488}        & {.246}       & {.849} \\
-\citeauthor{Albright2003b} & \uline{.929} & {.574}        & {.502}       & {.606} \\
-\citeauthor{Scholes1966}   & \uline{.907} & {.738}        & {.628}       & {.558} \\
-\midrule
-Kendall $\tau_{b}$        \\
-\midrule
-\citeauthor{Albright2007}  & {.666}       & {.450}        & {.245}       & \uline{.683} \\
-\citeauthor{Albright2003b} & {.474}       & \uline{.558}  & {.499}       & {.159} \\
-\citeauthor{Scholes1966}   & {.619}       & \uline{.671}  & {.614}       & {.484} \\
+Gross status  & \uline{.73} &       {.60} &       {.80} && \uline{.82} &       {.66} &       {.80} && \uline{.87} & \uline{.93} & \uline{.91} &&       {.67} &       {.47} &       {.62} \\
+Density ($N$) &       {.67} & \uline{.79} & \uline{.86} &&       {.61} & \uline{.74} & \uline{.82} &&       {.49} &       {.57} &       {.74} &&       {.45} & \uline{.56} & \uline{.67} \\
+Bigram score  &       {.46} &       {.75} &       {.74} &&       {.34} &       {.70} &       {.79} &&       {.25} &       {.50} &       {.63} &&       {.25} &       {.50} &       {.61} \\
+MaxEnt score  &       {.70} &       {.21} &       {.53} &&       {.66} &       {.39} &       {.58} &&       {.85} &       {.61} &       {.56} && \uline{.68} &       {.16} &       {.48} \\ 
 \bottomrule
 \end{tabular}
-\caption{Correlations between item-averaged wordlikeness ratings and model scores}
+\caption{Correlations between item-averaged wordlikeness ratings and model scores; the largest coefficient for each statistic/data set pair is underlined.}
 \label{scores}
+\end{table}
+
+It is also possible to consider whether there is any residual correlation between bigram and MaxEnt model scores, and wordlikeness ratings within the ``valid'' and ``invalid'' groups defined by the gross phonotactic status measure.
+Kendall $\tau_{b}$ correlations within these subgroups for each data set are shown in Table \ref{resid}.
+The only reliable positive correlation is present among the ``valid'' items as rated by the smoothed segmental bigram model.
+This model is somewhat capable of accounting for contrasts between different ``possible'' nonce words: for instance, it favors [plin] \emph{plean} over [brɛlθ] \emph{brelth} just as subjects in the \citet{Albright2007} study do.
+Within the set of ``invalid'' items, however, neither grammatical model reliably distinguishes among items; both models, for instance, rate [ptʌs] \emph{ptus} more well-formed than [bnʌs] \emph{bnus}, but speakers have the opposite preference.
+
+\begin{table}
+\centering
+\begin{tabular}{l rrr c rrr}
+\toprule                
+             & \multicolumn{3}{c}{``Valid'' items} && \multicolumn{3}{c}{``Invalid'' items} \\
+\cmidrule{2-4} \cmidrule{6-8}
+             %& \multicolumn{3}{c}{Kendall $\tau_{b}$} && \multicolumn{3}{c}{Kendall $\tau_{b}$} \\
+             &         A &       AH &        S &&     A  &       AH &          S \\
+\midrule
+Bigram score &     {.65} &    {.34} &    {.60} &&    {.03} & {$-$.17} &    {.47} \\
+MaxEnt score &     {.00} & {$-$.15} & {$-$.32} && {$-$.42} &    {.29} & {$-$.16} \\
+\bottomrule
+\end{tabular}
+\caption{Kendall $\tau_{b}$ correlation between item-averaged wordlikeness ratings (rated ``valid'' or ``invalid'' by the gross phonotactic status baseline) and model scores}
+\label{resid}
 \end{table}
@@
 
-FIXME
+The bigram and MaxEnt models do not reliably outperform simple baselines.
+From this it can be inferred that the gradient models do not reliably predict intermediate ratings.
+These models do not reliably distinguish within classes of ``valid'' and ``invalid'' words either.
+
+A serious limitation of this evaluation is the primitive nature of the gross phonotactic status baseline.
+It does not allow for any way to state constraints on onset-nucleus sequences, which have been proposed for some languages (e.g., \citealt{Kirby2007}),\footnote{\citet{Kessler1997}, however, argue there are no clear restrictions on English onset-nuclei pairs.} or constraints spanning whole syllables, which have been proposed for English \citep[e.g.,][]{Berkley1994b,Berkley1994a,Coetzee2008b,Fudge1969}.
+Furthermore, the gross phonotactic baseline does not have any mechanism for generalizing the wellformedness of [ɛsp] rimes from \emph{clasp}, \emph{lisp}, and other rimes consisting of a lax vowel followed by [sp] found in English.
+\citet{Borowsky1989}, for instance, proposes a theory of possible rimes in English which makes the correct prediction regarding [ɛsp].
+This is not embedded in an acquisitional model, but models of syllable type acquisition have been proposed \citep[e.g.,][]{Levelt2000}.
+
+The gross phonotactic baseline could also be extended so as to include additional levels of wellformedness.
+While the bigram and MaxEnt models do not appear to be able to reliably distinguish intermediate levels of well-formedness, it might be desirable to encode the intuition that, for example, [ʒlɪk] \emph{zhlick}, is more English-like than [bnɪk], though both have unattested onsets.
+It is also possible to imagine that phonotactic violations would have a cumulative effect on well-formedness; for instance, a nonce word with an unattested onset and an unattested rime, like [tsɪlb], might be less English-like than either [tsɪb] or [sɪlb].
+This prediction is made by the bigram and MaxEnt models, among others \citep[e.g.,][]{Albright2008,Anttila2008a}, but could easily be incorporated into a simple baseline by counting the number of violations.
+However, there is not yet any convincing evidence for cumulativity effects in wordlikeness tasks.
 
@@
 
-FIXME
+State-of-the-art computational models of wellformedness do not reliably predict intermediate ratings in wordlikeness tasks.
+To the degree to which the bigram or MaxEnt models are correlated with speakers' judgements, these judgements are more precisely modeled by similarity to existing words, or by a gross contrast between attested and unattested onsets and rimes.
+While it remains an open question whether future gradient models will account for intermediate judgements, the current evidence suggests that gradient grammaticality is not crucial for modeling gradient wordlikeness judgements.
 
-%%The primary result is that no gradient model reliably exceeds the accuracy of the binary baseline.
-Despite this, there are relatively strong correlations between the binary baseline and these gradient models (see Table \ref{bcor}).
-From the strong performance of the categorical model one can infer that the gradient models do not reliably predict intermediate ratings, or contrasts in ratings between words which are grouped together.
-To quantify this, the following method was used to estimate the residual contribution of the three gradient models once gross phonotactic violations are taken into account.
-Instead of calculating rank correlations directly on the model scores as in Table \ref{cor}, the model scores are mapped to ranks with the additional constraint that all ``valid'' stimuli be ranked above all ``invalid'' stimuli.
-The resulting ranks are used to compute new correlation statistics.
-Finally, the binary baseline correlation is subtracted from this number, so that the resulting value is the amount of improvement derived from augmenting the binary model with gradience.
-These difference numbers are shown in Table \ref{controlled}.
-In most cases, including the gradient models on top of the binary baseline produces a worse correlation than is obtained with the binary baseline alone.
-%%
-%%For $\tau_b$ and $\gamma$, the interpretation of this result is clear.
-The gradient models assign rankings to the sets of phonotactically valid and invalid clusters, respectively.
-For instance, the bigram model favors \emph{troog} [tɹuːɡ] over \emph{swach} [swætʃ], though neither contains any gross phonotactic violation.
-Similarly, the bigram model favors \emph{chwoop} [tʃwuːp] over \emph{zhrick} [ʒɹɪk], even though both contain ill-formed onsets.
-However, the majority of such predicted contrasts are not reflected in speakers' judgements; for instance, \emph{troog} is rated less English-like than \emph{swach} \citep{Greenberg1964}, contrary to the model predictions.
-This shows quite starkly that these models fail to reliably predict intermediate ratings.
-%%
-%%\subsection{The gradience hypothesis}
-%%
-%%This chapter has evaluated the axiom of gradience as a falsifiable alternative hypothesis.
-The surprising result is that virtually all of the apparent coverage of state-of-the-art gradient phonotactic models is simply a reflection of their ability to distinguish between the possible and the totally impossible; beyond this, they are unreliable.
-A trivial baseline, endowed with few abilities to project beyond the observed data, generally outperforms the state of the art.
-The projections made by the state-of-the-art gradient models are not like those made by speakers.
-It remains to be seen is whether any model can be put forth which accurately predicts these intermediate ratings.
-%%
-%%These result provide support for recent findings that speakers asked to perform gradient syntactic judgements produce responses closely corresponding to a widely recognized categorical grammatical/ungrammatical distinction \citep{Sprouse2007}.
-%%
-%%\subsection{Extensions to the binary baseline}
-%%
-%%The strong performance of the binary baseline should not be taken as evidence either that wordlikenesss judgements are binary, or that the binary baseline is a plausible model.
-The most serious limitation of this evaluation is the primitive nature of the binary baseline.
-The inability to generalization within onsets and rimes is a serious flaw, as is the assumption of independence of onset and rime.
-Regarding the rime, \citet{Borowsky1989} proposes a theory of possible rimes in English, which does make the correct prediction regarding the unattested but well-formed [ɛsp].
-On the other hand, a cognitively plausible version of this model might need to entertain phonotactic generalizations that are larger than these units, since syllable-sized phonotactic generalizations have been proposed for English \citep[e.g.,][]{Berkley1994a,Berkley1994b,Coetzee2008b,Fudge1969}.
-%%
-%%A possible further extension to the binary baseline would be the introduction of additional levels of wellformedness.
-While the evaluation has shown that current gradient models do not reliably identify intermediate wellformedness, it does seem possible to identify at least three levels of grammaticality: for instance, one might encode the intuition that \emph{zhlick} [ʒlɪk] is more English-like than \emph{bnick}, though both have unattested onsets.
-There are precedents for labeling certain attested words as phonotactically ``peripheral'' (see, e.g., the appendices in \citealt{Myers1987} and \citealt{Borowsky1989}); such words are regarded as lexical exceptions to language-general principles of syllabification.
-If this extends to nonce words, then an intermediate level of grammaticality could be assigned to ``possible'' but formally marked words.
-Another likely source of additional levels of grammaticality is the cumulative effect of multiple phonotactic violations.
-While, as \citet{Coleman1997} note, classical Optimality Theory predicts that a nonce word is as ill-formed as its worst deviation from syllable structure, it is possible to imagine that multiple phonotactic violations would result in greater degrees of ill-formedness.
-The bigram and maxent models make this prediction, as do many others \citep[e.g.,][]{Legendre1990,Levelt2000,Albright2008,Anttila2008a,Pater2009b} but despite this, there is still little data demonstrating cumulative effects in wordlikeness tasks.
-%%
-%%\subsection{Language acquisition}
-%%
-%%The weak empirical status of gradient phonotactic knowledge as reflected in adults has rammifications for language acquisition under the hypothesis that that infants acquiring language deploy the same representations as adults \citep[e.g.,][]{Macnamara1982,Pinker1984,Crain1991,Carey1995,deVilliers2001,Legate2007}.
-Gradient wordlikeness judgements in adults would provide support for claims that infants recognize statistical(inherently gradient) dependencies between segments \citep{Jusczyk1994} and use these to segment words \citep{Saffran1996}.
-An emerging consensus suggests, however, that infants attend to transitional probabilities primarily in the absence of grammatical cues \citep{Gambell2005,Hohne1994,Johnson2001,Jusczyk1999c,Lignos2012b,Mattys2001a,Shukla2007,Lew-Williams2012}.
+%An emerging consensus suggests, however, that infants attend to transitional probabilities primarily in the absence of grammatical cues \citep{Gambell2005,Hohne1994,Johnson2001,Jusczyk1999c,Lignos2012b,Mattys2001a,Shukla2007,Lew-Williams2012}.
 The vacuous nature of current evidence for gradient phonotactic knowledge in adults further weakens any hypothesis that would link statistical learning in infants to adults' behaviors.
-
-%% Hayes2000
-% 
-% Bard1996
-% 
-% Koo2009
-% 
-%
-
-% 
-% Warker2006
-% 
-% Massaro1983
-% 
-% Rusaw2009
-
-%%do not explicitly state why this data is relevant to the construction of models of wordlikeness.
-Presumably, these authors believe that these patterns of judgements demonstrate that wordlikeness, as an internal state, is gradient simply because subjects make use of intermediate degrees of wordlikeness in judgement tasks.
-This proposition, generalized below, is ``naïve'' not because it lacks sophistication, but because it is rooted in a belief in naïve realism, a philosophy which holds that perception provides a relatively direct picture of the nature of the world, an influential view in the cognitive sciences in general (see \citealt{Fodor1981a} for a critique).
-
-%%\citeauthor{Chomsky1965} were not the first to consider the notion of possible and impossible words.
-Their primary contribution is that their mentalist perspective: they recognize that naïve speakers effortlessly acquire language-specific generalizations about possible and impossible words and can report them without any explicit training.
-
-%%However, not all early literature is concerned with gradience.
-\citet[31]{Vogt1954}, for instance, recognizes that the taxonomic phoneme is insufficient to account for many wordlikeness contrasts.
-\citeauthor{Vogt1954} observes that allophony may account for the absence of certain phone sequences, but it does not provide a suitable explanation for the absence of initial [bn] in English, nor does it make correct predictions about the surface realization of an underlying initial /bn/.
-\citeauthor{Vogt1954} concludes that additional grammatical machinery will be needed to account for possible and impossible words.
-
-%%Most relevant to the question at hand, \citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated wtih the positional probailities of the constituent syllables.
-Unfortunately, none of these researchers make any effort to eliminate the possibility that the low positional probability stimuli are ``impossible'' words of English.
-In fact,
-
-%Chapter \ref{clusters} argues
-%that many of the stimuli used by \citeauthor{Frisch2000} and \citeauthor{Vitevitch1997} contain illicit word-medial consonant clusters.
-While \citeauthor{Vitevitch1997} neither control nor manipulate the well-formedness of medial clusters, in a post-hoc test they consider a probabilistic measure of cluster well-formedness, which reveals that cluster well-formedness is correlated with syllable-internal positional probabilities and wordlikeness judgements, but \citeauthor{Vitevitch1997} ultimately conclude this cannot explain all the variation in wordlikeness.
-
 %Using the head-term preference paradigm, \citet{Jusczyk1993b} and \citet{Friederici1993} find that typically-developing children as young as 9 months of age distinguish between nonce words which are and are not phonotactically valid in their target language.
 \citet{Jusczyk1994} report that 9-month-old children acquiring English also show preferences for nonce words with high positional probability over those with low positional probability.
-Faciliatory effects of positional probability (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements \citep{Vitevitch1999a,Luce2001,Lipinski2005,Vitevitch2005}, and lexical decision \citep{Pylkkanen2002a}.
-
-%%The aforementioned studies all conclude that the gradient measure of positional probability correlates with behavioral results.
-As the flaws of the \citet{Vitevitch1997} study demonstrate, the aforementioned studies do little to tease apart the gradient and categorical aspects of phonotactics.
-More generally, they do little to distinguish between positional probability and closely correlated measures like bigram probability (see \S\ref{bigram} below) or neighborhood density, since these studies carefully select stimuli which either have high or low values for all of positional probability, bigram probablility, and neighborhood density.
-This is particularly troublesome given that no justification has ever been given for the positional probability measure in the first place; it appears to have been created \emph{ex nihilo}; in contrast, the effects of neighborhood density in various psycholinguistic tasks are emergent properties of many models of speech production \citep[e.g.,][]{Luce1998,Luce2000} and perception \citep{Marslen-Wilson1984,Marslen-Wilson1987,McClelland1986,Norris1994,Norris2000}.
+%``Suppose that linguistic well-formedness really is an all-or-nothing matter, but that the judgments we get are filtered through various performance mechanisms.
+It is the performance mechanisms, not the grammar itself, which results in the gradient intuitions.
+If this is so\ldots{}we should be reducing them (somehow) to two categories and developing a model of the judgment process itself to account for the numbers observed.''

### 2012-11-11 3.tex
Completed writing on 2-4!

@@
 
-\section{Conclusions}
+\section{Conclusion}
 
-It has been shown that statistical reliability is not a sufficient condition for a sequence structure generalization to be reflected in psycholinguistic tasks.
-However, in the case of constraints on Turkish vowels, it is a necessary condition that the constraint in question be derived from an alternation.
+Statistical reliability is not a sufficient condition for a sequence structure generalization internalized by speakers or reflected in psycholinguistic tasks.
+This case suggests that it is a necessary condition that the constraint in question be derived from an alternation, however.
 Even alternations that have large numbers of lexical exceptions have robust effects on a constrained wordlikeness task.
-Purely statistical approaches to phonotactic knowledge fail to draw this important distinction.
+Purely statistical approaches to phonotactic knowledge fail to draw the distinction between static and derived phonotactic generalizations.
 

### 2012-11-11 4.tex
Completed writing on 2-4!

@@
 
-%\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984} or not \citep[e.g.,][]{Port1985}, it is not strictly relevant to the discussion at hand.}
-
-Since at least \citet{Pike1947b}, linguists have posited language-specific constraints on the contents of underlying representations.
-These \emph{morpheme structure constraints} were at one time thought to fully account for speakers' knowledge of possible and impossible words (e.g., \citealt{Chomsky1965}, \citeyear[382]{SPE}, \citealt[22f.]{SPR}, \citeyear{Halle1962}, \citealt{Stanley1967}).
-\citet{Shibatani1973}, however, argues that morpheme structure constraints cannot account for all wordlikeness constraints.
-In German, for instance, final obstruents devoice, and as a result, there are pairs such  [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' and [ɡʀaːt]-[ɡʀaːdə] `degree(s)' differing only in the plural shape of the root.
-Since the voicing of final obstruents is contrastive, the restriction on the voicing of obstruents cannot be a morpheme structure constraint. 
-Yet, \citeauthor{Shibatani1973} claims that native speakers reject nonce words ending in final voiced obstruents ``\emph{on the ground that they end in voiced obstruents}'' (95). 
-Thus, not all constrasts in possible wordhood can be stated as morpheme structure constraints.
-
-Whereas \citeauthor{Shibatani1973} maintains that morpheme structure constraints are insufficient to account for speakers' knowledge of possible wordhood, others argue that morpheme structure constraints are also unnecessary.
-\citet[297]{Hale1965}, \citet{Kisseberth1970b}, and \citet[212f.]{Postal1968} observe that the structural descriptions of phonological processes often are often reflected in the lexical redundancies in the same language.
-In Russian, there are alternations implicating a process of obstruent voice assimilation, and tautomorphemic obstruent clusters have uniform voicing \citep[283]{A74}.
-\citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that morpheme structure constraints are otiose, as phonological rules triggering alternations impose restrictions on the contents of URs, an effect known as \emph{Stampean occultation}.
-
-There are some apparent constraints on URs which are not reflected in the system of alternations; these generalizations seem to require reference to the units of prosody (\citealt{Hooper1973}, \citealt{Kahn1976}; see \citealt{Blevins2003} and \citealt{Steriade1999} for possible alternatives).
-Since word-level prosodic structure is generally thought to be largely redundant and thus absent from underlying representations, these cannot be equated with morpheme structure constraints either.
-\citeauthor{Haugen1956} notes, for instance, that word-medial consonant cluster inventory is not arbitrary in prosodic terms.
+It is a curious fact that many constraints on underlying representations are elegantly described using the units of prosody.
+In some cases, these are part of the system of contrast, such as long segments in moraic theory \citep[e.g.,][]{Davis1999}.
+For instance, Latin has constrastive consonant and vowel length (e.g., \emph{anus} `ring' vs.~\emph{annus} `year', \emph{os} `bone' vs.~\emph{ōs} `mouth'), but there is no vowel length contrast before geminate consonants.
+However, constraints on underlying representations may also involve references to non-contrastive prosodic structures, such as syllables (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample is the result of an underlying moraic, not syllabic, contrast.}
+For instance, as \citeauthor{Haugen1956} first notes, the constraints on word-medial consonant clusters are not arbitrary in prosodic terms:
 
@@
 \label{mcl}
-A word-medial consonant cluster consists of at most a well-formed medial coda followed by a well-formed medial onset
+A medial cluster must consist of a well-formed medial coda and a well-formed medial onset
 \end{example}
 
-Reviewing the inventory of word-medial consonant clusters in English, 
-\citet{Pierrehumbert1994} notes that the vast majority of ``possible clusters'', those which can be parsed into a coda and onset, are unattested.
-\citeauthor{Pierrehumbert1994} argues that the cluster inventory is further filtered by what are essentially morpheme structure constraints: static co-occurrence restrictions with no support from either prosody or the system of alternations.
-This study will argue, however, that the only structural constraints on word-medial clusters in English are the \textsc{Medial Cluster Law} and Stampean occultation.
-Consequently, this domain provides no evidence for the existence of morpheme structure constraints.
+\noindent
+It is not necessary for predictable prosodic structure to be present in underlying form to achieve this effect, contrary to what has sometimes been stated \citep[e.g.,][255]{A74}.
+All that is necesary is that no phonological processes intervene between underlying representation and the surface form of medial clusters.
+Under the further assumption that underlying and surface representations are as similar as alternations allow (e.g., \citealt[205f.]{Dell1973}, \citealt[28f.]{Stampe1973}), this is the null hypothesis.
 
-\section{English syllable contact clusters \citep{Pierrehumbert1994}}
+In a study of English word-medial consonant clusters, \citet{Pierrehumbert1994} notes that the vast majority of ``possible clusters'', those which can be parsed into coda and onset in accordance with the \textsc{Medial Cluster Law}, are unattested.
+\citeauthor{Pierrehumbert1994} argues that this can be accounted for by positing static co-occurrence restrictions---unrelated to any phonological alternation in English.
+This study argues that well-known phonological processes, but not the static constraints proposed by \citeauthor{Pierrehumbert1994}, impose restrictions on this inventory.
+%\citeauthor{Pierrehumbert1994}'s static constraints are not statistically valid
 
-English admits a variety of word-medial consonant clusters spanning syllable boundaries which are also known as \emph{syllable contact clusters} or \emph{interludes}.
-These clusters can be as short as two consonants, as in \emph{a}[n.t]\emph{ics}, or as long as four, as in \emph{mi}[n.str]\emph{el}.
-There are three
+\section{English syllable contact clusters}
 
-\subsection{The role of Stampean occultation}
+English syllable contact clusters have been used, by \citeauthor{Pierrehumbert1994} and others, to argue for the necessity of admitting static phonotactic constraints, and to illustrate proposals for the architecture of the phonotactic system.
+However, there are a number of reasons to reconsider the facts of this domain.
 
-Finally, \citeauthor{Pierrehumbert1994} does not fully consider the possible effects of Stampean occultation, as some of the static constraints she proposes may be related to English morphophonemics.
+\subsection{The role of phonological processes}
+
+\citeauthor{Pierrehumbert1994} does not fully consider the effects of phonological processes targeting word-medial clusters in English, and consequently some of the static constraints she proposes may be related to English morphophonemics.
 For instance, she writes that ``nasal-stop sequences agree in labiality'' (175), and attributes this the status of a static constraint. 
@@
 
-\citet{Pierrehumbert1994} infers static constraints from near-exceptionless gaps in the lexicon, but little effort is made to show that the patterns of lexical underrepresentation from which the static constraints are inferred are not due to chance.
-This admits the possibility that the gaps in the consonant cluster inventory are accidental rather than structural in nature \citep{Fischer-Jorgensen1952,Vogt1954}.
+\citet{Pierrehumbert1994} infers static constraints from near-exceptionless gaps in the lexicon, but little effort is made to show that the patterns of lexical underrepresentation are not due to chance.
+Consquently, it is possible to suggest that some of these gaps 
+are accidental rather than structural in nature.
+%\citep{Fischer-Jorgensen1952,Vogt1954}. 
 This is made all the more likely given the tendency of segment and cluster frequency distributions to be highly skewed \citep[e.g.,][]{Pande2010,Sigurd1968,Tambovtsev2007,Weiss1961}. 
-In Figure \ref{cao}, the rank and type frequency (i.e., frequency in the lexicon) frequency of medial codas and onsets in the lexical sample described below are graphed in log-log space.
-Both codas and onsets show a linear relationship between log rank and log frequency characteristic of Zipfian distributions (see Appendix \ref{zr}).
-As a result, an enormous lexical sample would be needed to realize clusters consistent with the \textsc{Medial Cluster Law} even if there were no constraints on the combination of medial codas and onsets in English.
-
-\begin{figure}
-\centering
-\includegraphics{co.pdf}
-\caption{Medial coda and onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
-\label{cao}
-\end{figure}
+Furthermore, \citeauthor{Pierrehumbert1994} considers only triconsonantal clusters, but medial clusters may be as short as two consonants, as in \emph{a}[n.t]\emph{ics}, or as long as four, as in \emph{mi}[n.str]\emph{el}, and \citeauthor{Pierrehumbert1994} offers no justification for a narrower focus.
+If this has any effect on the distribution of clusters, it would be to impose further sparsity.
 
@@
 Many components of \citeauthor{Pierrehumbert1994}'s study cannot be directly replicated.
-\citeauthor{Pierrehumbert1994} focuses her attention on those words she judges to be ``morphologically simple'' and ``reasonably familiar''.
-Since this list has neither been published nor circulated, any replication will have to be indirect.
-Further, it is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias.
-A sequence which is rare in simplex words is sometimes said to mark a juncture even in the absense of other evidence for segmentation; for instance, \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur morph-internally.
-There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings.
-Applied indiscriminately, though, such a heuristic has the possibility to trivializes both morphological segmentation and phonotactic generalization.
+\citeauthor{Pierrehumbert1994} focuses her attention on those words she judges to be ``morphologically simple'' and ``reasonably familiar'', but this list has neither been published nor circulated.
+It is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias; 
+\citet{Labov1975}, for instance
+argues for what he calls the \emph{Experimenter Principle}, 
+that the intuitions of all those familiar with the formal matters should be excluded.
+Otherwise-unmotivated morphological junctures are sometimes posited simply to preserve phonological or phonotactic generalizations; 
+%The same practice is sometimes applied to preserve phonological generalizations; for instance, otherwise unmotivated morphological boundaries are used in \emph{SPE} to simplify principles of stress assignment.
+this is done by \citet{SPE}, for instance, to simplify principles of English stress assignment, and \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur morph-internally.
+%There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings. 
+Applied indiscriminately, however, such a heuristic trivializes both morphological segmentation and phonotactic generalization.
+For these reasons, the wordlist used in this study is derived from a publicly available database, and no experimenter intuitions are used for establishing morphological boundaries.
+
+%Such reasoning is found in \citeauthor{Pierrehumbert1994}'s discussion of the word \emph{a}[nt.l]\emph{er}, however: this contains a cluster which violates her constraint on coda coronal obstruents, and therefore must be a complex word, she argues (see \citealt[164]{Borowsky1989} for additional exceptions to this generalization).
 
@@
 
-\citeauthor{Pierrehumbert1994} focuses only on triconsonantal clusters, and no reason is given for the exclusion of other sizes of clusters.
+After constructing a sample of syllable contact clusters in English simplex words, this data is used to evaluate the coverage of static and phonologically derived constraints.
 
 \subsection{Method}
-
-%The same practice is sometimes applied to preserve phonological generalizations; for instance, otherwise unmotivated morphological boundaries are used in \emph{SPE} to simplify principles of stress assignment.
-Such reasoning is found in \citeauthor{Pierrehumbert1994}'s discussion of the word \emph{a}[nt.l]\emph{er}, however: this contains a cluster which violates her constraint on coda coronal obstruents, and therefore must be a complex word, she argues (see \citealt[164]{Borowsky1989} for additional exceptions to this generalization).
 
@@
 Following \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, 
-a wordlist was produced using the English portion of the CELEX database \citep{CELEX}.
-Any words which were coded as borrowings were excluded, as were all words not coded as ``monomorphemic''.
-The use of this stringent criterion eliminates many of the exceptions noted by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies.
-For instance, many of the exceptions to \textsc{Obstruent Voice Assimilation} (see \S?
-below) noted by \citet[74]{Hammond1999a} are coded either as loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as complex words (e.g., \emph{jurisdiction}, \emph{madcap} \emph{tadpole}, \emph{scapegoat}, \emph{magpie}).
+a wordlist is generated using the English portion of the CELEX database \citep{CELEX}. 
+Only words coded as ``monomorphemic'' are used, and all words coded as non-native are excluded.
+This stringent criterion for inclusion eliminates many of the exceptions noted by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies.
+For instance, nearly all the exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) noted by \citet[74]{Hammond1999a} are
+either complex words (e.g., \emph{jurisdiction}, \emph{madcap} \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).
 
-Unlike prior studies, ``neo-classical compounds'', words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from this wordlist.
-This is done in response to converging evidence that English speakers regard such words as complex.
-
-The treatment of Latinate forms in \emph{The sound pattern of English} (\citealt{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition.
-In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.
+Unlike prior studies, however, words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from the sample, as there is broad formal and experimental support for the claim that these words are complex.
+This assumption allows for numerous simplifications to accounts of the phonological shape of these forms. 
+For instance, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.
 
@@
 \citeauthor{Aronoff1976} takes this to be evidence that \emph{adhere} and \emph{cohere}, for instance, share a bound stem.
-
-There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement.
-\citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.
+There is also an interaction between Latinate prefixes on verbs and the makeup of the verbal complement. 
+Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep{Gropen1989,Harley2009}.
+%\citep[see also][]{Gropen1989,Coppock2008}.
 
@@
 
-Results from lexical decision experiments also suggest that neoclassical compounds are complex.
-\citet{Taft1975,Taft1976} and \citet{Taft1986} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure such as *\emph{refant}).
-Bound stems also show frequency effects which are independent of their whole word frequency \citep{Taft1979,Taft2006}.
-Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect thought to indicate of morphological relatedness.
+Lexical decision experiments also suggest that Latinate prefixed forms are complex.
+\citet{Taft1975,Taft1976} and \citet{Taft1986} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure, such as *\emph{refant}.
+Bound stems also show frequency effects which are independent of whole word frequency \citep{Taft1979,Taft2006}. 
+Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect thought to indicate morphological relatedness.
 
@@
 
-This filtering results in a list of 6,619 simplex words.
+This results in a list of 6,619 simplex words. 
+The full set of clusters and their frequencies are given in Appendix \ref{clusters}.
 The CELEX transcriptions of these words were then syllabified and phonologized using a technique described in Appendix \ref{syllabification}. 
@@
 Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination, 174 (19\%) are attested.
-The full set of clusters and their frequencies are reproduced in Appendix \ref{clusters}.
 
@@
 So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters.
-These constraints are ``static'' in the sense that they have no analogues in the morphophonemics of English.
+These constraints are static, in the sense that they have no analogues in the morphophonemics of English.
 
-\paragraph{Dorsal-labial clusters} \citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S?).
+\paragraph{Dorsal-labial clusters} 
+\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S?).
 Two-consonant velar-labial clusters are found in words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} occur. 
-Table \ref{dltab} shows that velar-labial clusters are somewhat less likely to occur than  are velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance.
+Table \ref{dltab} shows that velar-labial clusters are somewhat less likely to occur than  are velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance ($p = .106$).
 
@@
 \midrule
-conforming & 25       & 91         & 22\%       & \multirow{2}{*}{0.106} \\
+conforming & 25       & 91         & 22\%       & \multirow{2}{*}{.106} \\
 violating  &  4       & 40         &  9\%       \\
@@
 
-\paragraph{Coronal obstruent codas} \citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}.
+\paragraph{Coronal obstruent codas} 
+\citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}.
 Table \ref{cctab} reveals that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}). 
-Restricting attention to triconsonantal clusters does not produce a significant lexical effect, either ($p = 0.129$).
+The same is true if attention is restricted to triconsonantal clusters ($p = .129$).
 
@@
 \midrule
-conforming & 56       & 304        & 15\%       & \multirow{2}{*}{0.430} \\
+conforming & 56       & 304        & 15\%       & \multirow{2}{*}{.430} \\
 violating  & 37       & 243        & 13\%       \\
@@
 
-\paragraph{ABA clusters} \citet[][176]{Pierrehumbert1994} notes the ``lack of clusters with identical first and third elements'', ignoring presence or absence of voicing.
-Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters (Table \ref{abatab}).
+\paragraph{ABA clusters} 
+\citet[176]{Pierrehumbert1994} notes the ``lack of clusters with identical first and third elements'', ignoring presence or absence of voicing. 
+Despite the fact that there are no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters (Table \ref{abatab}).
 
@@
 \midrule
-conforming & 47       & 512        &  8\%       & \multirow{2}{*}{0.250} \\
+conforming & 47       & 512        &  8\%       & \multirow{2}{*}{.250} \\
 violating  &  0       &  25        &  0\%       \\
@@
 
-\paragraph{Summary} There is no statistical support for \citeauthor{Pierrehumbert1994}'s static constraints.
+\paragraph{Summary} 
+There is no statistical support for \citeauthor{Pierrehumbert1994}'s static constraints.
 
@@
 
-FIXME
+In \emph{The Sound Pattern of English} (henceforth, \emph{SPE}), \citet{SPE} describe three phonological processes which target medial consonant clusters.
+As will be shown, the neutralizing effect of these three processes has a major influence on the cluster inventory.
 
@@
 \citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice. 
-As shown in Table \ref{ovatab} however, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}, and hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} are rarer than would be expected from chance.
+As shown in Table \ref{ovatab} however, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}.
+Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, are rarer than would be expected from chance.
 
@@
 \midrule
-conforming & 35       & 329        & 10\%       & \multirow{2}{*}{0.002} \\
+conforming & 35       & 329        & 10\%       & \multirow{2}{*}{.002} \\
 violating  & 11       & 305        &  3\%       \\
@@
 
-\textsc{Nasal Place Assimilation} (e.g., \citealt[149]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}) accounts for [n $\sim$ ŋ] allophony as well as Latinate prefix allomorphy.
+\textsc{Nasal Place Assimilation} (e.g., \citealt[65f.]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}) permits [ŋ] to be described as a pure allophone of /n/, and accounts for certain patterns in Latinate prefix allomorphy.
 
@@
 \begin{tabular}{l l l l l l l}
-a. & {polite}   & {i}[m.p]{olite}   & & {balance} & {i}[m.b]{alance} \\
-b. & {tangible} & {i}[n.t]{angible} & & {decent}  & {i}[n.d]{ecent}  \\
+a. & {polite}   & {i}[m.p]{olite}   \\
+   & {balance}  & {i}[m.b]{alance}  \\
+b. & {tangible} & {i}[n.t]{angible} \\
+   & {decent}   & {i}[n.d]{ecent}   \\
 \end{tabular}
@@
 
-As shown in Table \ref{npatab}, virtually all (94\%) clusters of a nasal coda followed by a homorganic obstruent (e.g., \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, \emph{mo}[ŋ.k]\emph{ey}) are attested.
+\noindent
+As shown in Table \ref{npatab}, virtually all clusters consisting of a nasal coda followed by a homorganic obstruent (e.g., \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, \emph{mo}[ŋ.k]\emph{ey}) are attested.
 As \citet[175]{Pierrehumbert1994} observes, heterorganic clusters, like those \emph{pli}[m.s]\emph{oll} or \emph{scri}[m.ʃ]\emph{aw}, do occur, but they are significantly more rare.
@@
 \midrule
-conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.37\e{-07}} \\
+conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.4\e{-07}} \\
 violating  & 11       & 22         & 33\%       \\
@@
 \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}).
-\textsc{Degemination} is formalized here as a rule deleting the first of two identical segments, ignoring voice.
+\textsc{Degemination} is formalized here as a rule deleting the first of two segments agreeing on all feature values (except voice, possibly).
 
@@
 
-A soft-margin support vector machine \citep{Cortes1995} with a linear kernel is used to convert from the per-cluster scores to attestation predictions
-, by finding a single cutoff which optimally separates attested and unattes
-ted clusters according to their model scores.
+Current computational models of phonotactic knowledge ``rate'' possible clusters by producing a numerical value for each cluster. 
+To apply these models to predicting the cluster inventory, these 
+these numerical values must be transformed into a categorical prediction, of either attestation or non-attestation.
+This is accomplished with a soft-margin support vector machine \citep{Cortes1995} with a linear kernel, which optimally splits the ratings distribution into a set of attested and unattested clusters.
+This classifier is does not correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.
 
-It is possible to maximize recall at the expense of precision, by predicting all clusters to be unattested, or to increase precision at the expense of recall, by predicting non-attestation for a great number of clusters.
-$F_1$, the harmonic mean of these two measures, is a balanced measure of this tradeoff.
-
-This is not put forth as a cognitively plausible model for learning of phonotactics; it simply represents an upper bound for classifying possible clusters.
-
-The results for are summarized in Table \ref{cmresults}.
-
-\paragraph{Null baseline} Since the majority of possible clusters are unattested, a null baseline which predicts all clusters to be unattested achieves classifies 81\% of clusters correctly, achieving perfect recall at the price of precision.
-
-\paragraph{Expected frequency} \citet{Pierrehumbert1994} proposes that the well-formedness of a syllable contact clusters is proportional to the independent probability of the coda and onset that make it up.
-\citeauthor{Pierrehumbert1994} reports that this is the best predictor of which complex clusters occur, and while it is a significant improvement on the baseline (sign test $p = 4.5$\e{-05}), it is not distinguished in performance from the other models under consideration.
-
-\paragraph{Derived constraints} The three derived generalizations define a simple classifier, predicting attestation for all clusters which do not violate these three generalizations.
-This is a significant improvement on the null baseline (sign test $p = 4.0$\e{-09}), preserving high recall while improving precision.
-The predictions of the derived constraints classifier can also be combined with information on expected frequency.
-As expected, this boosts precision, since FIXME.
-This combination produces a model with the highest accuracy and $F_1$.
-
-\paragraph{Maximum entropy phonotactics} \citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints.
-Since this model has many experimenter-determined parameters, a direct replication of the experiments reported by \citealt{Hayes2008a} is attempted; their software, feature specification, and model settings are all used.
-Training in this model is inherently stochastic, producing slightly different constraints and weights on each run, so the least accurate of ten independent runs is used for evaluation; this is also the practice of \citeauthor{Hayes2008a}.
-This model performs approximately as well as other classifiers, but has the poorest recall; this suggests that the constraints learned are overly narrow and fail to filter out many impossible clusters.
-FIXME
-
-\paragraph{Summary} FIXME
+Four metrics are used to score the models. 
+\emph{Accuracy} represents the probability that a cluster is correctly classified as attested or unattested.
+Two additional metrics break down accuracy into its constituent parts;
+\emph{precision} represents the probability that a cluster which is predicted to be unattested is in fact unattested, and \emph{recall} is the probability that an unattested cluster is predicted as such.
+It is possible to increase precision at the expense of recall, by predicting non-attestation for a greater number of clusters, or to maximize recall at the expense of precision by predicting all clusters to be unattested.
+$F_1$, the harmonic mean of these two measures, is a standard metric for quantifying this tradeoff; increases in either precision or recall produce a  non-linear increase in $F_1$.
+The results are summarized in Table \ref{cmresults}.
 
@@
 \end{tabular}
-\caption{Results of the cluster classification task; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
+\caption{Results for the cluster classification task; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
 \label{cmresults}
@@
 
+\paragraph{Null baseline} 
+In a classification task, the simplest baseline is one which uniformly predicts the most common outcome. 
+Since only 19\% of clusters are attested, 
+81\% accuracy is achieved by predicting all clusters to be unattested.
+
+\paragraph{Expected frequency} 
+\citet{Pierrehumbert1994} proposes that the well-formedness of a syllable contact cluster is proportional to the product of the independent probabilities of the coda and of the onset that make it up; this is the cluster's \emph{expected frequency}.
+\citeauthor{Pierrehumbert1994} reports that this is an excellent predictor of which complex clusters occur and which do not. 
+This does not impose any constraints which span the syllable boundary; rather, it is a model of which clusters might be expected to represent accidental gaps in the sample.
+This represents a significant improvement compared to the null baseline (sign test $p = 4.5$\e{-05}).
+
+\paragraph{Derived constraints} 
+By hypothesis, \textsc{Obstruent Voice Assimilation}, \textsc{Nasal Place Assimilation}, and \textsc{Degemination} rule out a large number of possible clusters. 
+In all, they target 316 out of 920 possible clusters (34\%), and only 11 (3\%) of these clusters has non-zero frequency in the sample.
+Together, these three processes define a simple classifier in which a cluster is predicted to be unattested if and only if it would be neutralized by one of these processes. 
+Unsurprisingly, this results in increased precision and a significant improvement in accuracy compared to the null baseline (sign test $p = 4.0$\e{-09}).
+
+\paragraph{Expected frequency and derived constraints}
+It is possible to combine into a single classifier the intuitions of the expected frequency and derived constraint models, the former accounting for accidental gaps and the latter for structural gaps imposed by neutralizing phonological processes.
+By simultaneously accounting for both sources of cluster inventory gaps, this model outperforms all others in accuracy and $F_1$.
+
+\paragraph{Maximum entropy phonotactics} 
+\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints. 
+In one sense, this is isomorphic to the expected frequency/derived constraint model in that the constraint discovery mechanism is sensitive to the expected frequency of clusters: it favors constraints which rule out clusters with high expected frequency (but which are unattested) over those which have low expected frequency.
+Unlike the expected frequency/derived constraint model, however, the constraints discovered are not limited to those which are illustrated by alternations; it may also posit static constraints, much like those proposed by \citet{Pierrehumbert1994}.
+Since this model has many experimenter-determined parameters, 
+an attempt is made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications are all used. 
+Following the procedure of \citet{HayesInPress}, a novel feature [$\pm$\textsc{Coda}] is added to allow the model to distinguish coda and onset consonants. 
+Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}.
+This model has the poorest recall of any model; compared to the derived constraints baseline, the constraints induced by the maximum entropy model are narrower.
+This is particularly clear regarding possible clusters with a nasal coda followed by a non-homorganic obstruent, like *[m.kl]: the vast majority of such clusters, which would be neutralized by \textsc{Nasal Place Assimilation}, are unattested, but many are erroneously assigned the highest possible score by the maximum entropy model.
+
+\paragraph{Summary}
+Expected frequency and derived constraints effectively account for 
+accidental and structural gaps in the cluster inventory.
+The \citet{Hayes2008a} computational model only provides an inferior approximation of the derived constraints.
+
 \subsection{Discussion}
 
-FIXME
+Insofar as the \citet{Hayes2008a} model discovers static constraints, these fare no better than those proposed by \citet{Pierrehumbert1994}.
+It appears that further gaps in the cluster inventory cannot be described in phonological, structural terms; they can only be accidental gaps.
+In what remains, the nature of these accidental gaps is considered.
 
@@
 
-\begin{unlabeledexample}
-$\displaystyle \hat{p}_0 = \frac{n_1}{N}$
-\end{unlabeledexample}
+\citet{Good1953} proposes a method for estimating the extent of accidental gaps in a distribution under simple assumptions.
+More specifically, this takes the form of $p_0$, a probability representing the possibility of an unseen cluster.
+Were it possible to extend the lexical sample, this would be the probability that the next cluster would be one not yet observed.
+The value of $p_0$ is the ratio of clusters that occur exactly once in the sample to the size of the sample.
 
-%67 types occur once
-%67 / 977
-%7\%
+\begin{equation*}
+\displaystyle p_0 = \frac{n_1}{N}
+\end{equation*}
+
+\noindent
+Here, $n_1 = 67$ and there are 997 clusters in all, so there is approximately a 7\% chance that the next cluster would fill in what is now an accidental gap.
 
@@
 \label{simulate}
+
+It can be shown that the large number of possible-but-unattested clusters is a logical necessity, given the sparse nature of the coda and onset frequencies. 
+In Figure \ref{cao}, the rank and type frequency (i.e., frequency in the lexicon) of medial codas and onsets in the lexical sample described below are graphed in log-log space. 
+Both codas and onsets show a linear relationship between log rank and log frequency characteristic of Zipfian distributions (see Appendix \ref{zr}). 
+As a result, an enormous lexical sample would be needed to realize clusters consistent with the \textsc{Medial Cluster Law} even if there were no constraints on the combination of medial codas and onsets in English.
+To illustrate this fact, a simulations is used to create new ``samples'' of the same size as the lexical sample used here.
+The following procedure is repeated so as to generate each observation in the simulated sample.
+
+\begin{figure}
+\centering
+\includegraphics{co.pdf}
+\caption{Medial coda and medial onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
+\label{cao}
+\end{figure}
 
@@
 b. & Sample a medial onset according to the observed probabilities \\
-c. & Apply all matching \emph{SPE} rules to cluster formed by their concatenation \\
+c. & Apply the \emph{SPE} rules to the cluster formed by their concatenation \\
 \end{tabular}
 \end{example}
+
+This procedure corresponds to the assumption that the derived constraints impose the only structural restrictions on the cluster inventory.
+Cluster frequencies in one simulated sample are shown in Figure \ref{sim}; dots represent simulated frequencies, and the line represents actual cluster frequencies.
+As can be seen, the observed and simulated frequencies are quite similar 
+($R^2 = 0.712$, $p = 4.5$\e{-05}).
+The sparse cluster inventory, previously taken as evidence for static constraints on syllable contact clusters, is expected even if the only constraints on possible clusters are those which are phonologically derived.
 
@@
 \caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line).
-Simulated and observed frequencies have been smoothed using the $Z_r$ transform.}
+Simulated and observed frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
 \label{sim}
@@
 
-real ($\alpha = -1.65$, $R^2 = 0.940$)
-fake ($\alpha = -1.58$, $R^2 = 0.945$)
-fitt ($R^2 = 0.712$, $p = 4.5$\e{-05})
-
-In summary, the sparse cluster inventory, previously taken as evidence for static constraints on syllable contact clusters, is expected even if the only constraints on possible clusters are those which are phonologically derived.
-
 \section{Conclusions}
 
-None of the foregoing results have revealed any evidence for structural constraints on the English syllable contact cluster except those derived from phonological alternations.
+The foregoing results demonstrate that the only structural constraints on English syllable contact are derived from the phonological system: there is no evidence for static constraints in this domain.
 What then is to be said of the 366 clusters which do not violate a derived constraint, but which are yet unattested, like [b.z] or [z.n]? 
-Since these gaps appear to be quite arbitrary from a phonological perspective---they cannot be reliably discovered either by linguist or computer---it seems reasonable to identify these clusters as accidental gaps, resulting from the sparsity of the sample and of the English lexicon.
+These gaps appear to be phonologically arbitrary, since they are not reliably identified either by phonologist or computional model.
+They can only be understood as accidental gaps resulting from the sparsity of the sample and the finite nature of the English lexicon.
 

### 2012-11-11 abs.tex
Completed writing on 2-4!

@@
-This dissertation argues for that phonotactic knowledge is primarily derived not from static generalizations but rather from phonological alternations and prosodic parsing.
+This dissertation outlines a program for phonotactic theory---the theory of speakers' language-specific knowledge of possible and impossible (or likely and unlikely) words---minimizing undue \emph{duplication} (in the sense of \citeboth{KK77}:136f.) between phonotactics and phonology, and shows that the alternative view of phonotactics as stochastic, and of phonotactic learning as probabilistic inference, is incapable of accounting for the facts of this domain.
 Chapter \ref{intro} outlines the proposal, historical precursors, and predictions. 

### 2012-11-11 ms2.tex
Completed writing on 2-4!

@@
 
-\abstract{Gradient ratings are a property of ratings tasks, not of the grammatical system itself.}
-
+\abstract{
+Gradient wordlikeness judgements do not necessarily imply that a gradient well-formedness system underlies them; gradient judgements may be an artifact of gradient rating tasks.
+Dubious architectural assumptions are needed for speakers to report gradient well-formedness judgements.
+Simple baselines better account for gradient well-formedness judgements than state-of-the-art computational models of gradient phonotactic knowledge.
+}
+\section{Introduction}
 \input{2}
@@
 
-For every entry in the CELEX database, there is a corresponding broad syllabified transcription of the word in a Received Pronunciation accent.
-This appendix describes an automated procedure used to process these transcripts and to separate medial clusters from their flanking nuclei, parsing the resulting sequences into coda and onset, and reversing allophonic processes targeting medial clusters.
-
-While the segmental content of these transcriptions is precise, the CELEX syllabifications are unsystematic.
-Given the absence of contrastive syllabification in English (if not all languages: see \citealp[221]{Blevins1995}, \citealp{Elfner2006}), any sequence of a medial consonant cluster and its flanking nuclei should receive the same syllabification in all words in which it occurs.
-This is not always the case with the CELEX transcriptions, however.
-For instance, the sequence [ɪstɹɪ] receives a different parse in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is lax in Received Pronunciation \citep[][II.294]{AOE}.}
+This appendix describes an automated procedure used to process the CMU pronunciation dictionary entries, separating medial clusters from their flanking nuclei, parsing the resulting sequences into coda and onset, and reversing allophonic processes targeting medial clusters.
 
@@
 
-In English, when a medial consonant cluster is preceded by a stressed lax vowel, as \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[3]{Hammond1997}.
+In English, when a medial consonant cluster is preceded by a stressed lax vowel, as \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel (\citealp[3]{Hammond1997}, \citealp{Treiman1990}).
 As \citet[55]{Harris1994} notes, however, when the medial cluster is also a valid onset, as in \emph{whi}[s.p]\emph{er}, \emph{ve}[s.ti]\emph{ge}, and \emph{mu}[s.k]\emph{et}, onset maximization will incorrectly assign the entire cluster to the onset and leave the lax vowel unchecked.
@@
 In English, however, affricates do not form complex onsets.
-Yet, the stop and release phase of affricates cannot be separated by a syllable boundary, as predicted from the assumption they are single phonological units.
+Yet the stop and release phase of affricates cannot be separated by a syllable boundary, as predicted from the assumption they are single phonological units.
 

### 2012-11-16 1.tex
Moving stuff around, mostly.

@@
 Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature.
-At the risk of explaining what might be no more than an an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms.
+At the risk of explaining what might be no more than an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms.
 \citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.

### 2012-11-16 2.tex
Moving stuff around, mostly.

@@
 Since the granularity of measurement and analysis is determined by prior hypotheses about the nature of the system under study, disputes are inevitable as scientific paradigms evolve. 
-This study argues that there are both theoretical and empirical arguments to doubt the implicit hypothesis linking scalar wordlikeness judgements and a gradient model of phonotactics.
+This study argues that there are both theoretical and empirical reasons to doubt the implicit hypothesis linking scalar wordlikeness judgements and a gradient model of phonotactics.
 First, intermediate judgements are characteristic of all gradient rating tasks, and therefore these judgements are irrelevant to the question of whether well-formedness is categorical or gradient.
@@
 
-\section{Aspects in the theory of gradient grammaticality}
+\section{Aspects of the theory of gradient grammaticality}
 
@@
 \includegraphics{agg.pdf}
-\caption{Subjects freely use intermediate ratings when asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively \citep{Armstrong1983}}
+\caption{Subjects freely use intermediate ratings when asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively (\citealp{Armstrong1983}; ``1'' indicates ``most representative'')}
 \label{agg}
@@
 Each step of this procedure is dubious, however.
-First, speakers have difficulty perceiving \citep{Brown1956} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
+First, speakers have difficulty perceiving \citep{Berent2007a,Brown1956,Dupoux1999,Kabak2007a} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
 Secondly, the computation of a scalar value serves no further purpose than to provide for gradient well-formedness judgements; even if this is the result of probabilistic parsing \citep[e.g.,][]{Coleman1997},
@@
 Next, speakers must be able to consciously access and report the magnitude of this value (it must be \emph{cognitively penetrable} in the sense of \citealt{Pylyshyn1984}), an ability which is limited in many other domains.
-Finally, \citet{Sprouse2011} argues that speakers do not (or cannot) scale well-formedness values to satisfy the assumptions magnitude estimation, a type of gradient rating task.
+Finally, \citet{Sprouse2011} argues that speakers do not (or cannot) scale well-formedness values to satisfy the assumptions of magnitude estimation, a type of gradient rating task.
 %\citep{Bard1996}
@@
 There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}.
-As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures may not stored in linguistic memory. 
+As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures may not be stored in linguistic memory. 
 %This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
@@
 Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, many of which are exceptionless, into probabilities.
-While this is consistent with their claim, that ``the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.
+While this is consistent with their claim, that ``the ability to model gradient intuitions [are] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.
 
@@
 Like the study by \citet{Albright2007}, the focus is on onset well-formedness and there is minimal diversity in rime type.
-Two items, [klʌŋ] \emph{clung} and [bɹʌŋ] \emph{brung} (a dialectical past participle of \emph{bring}), are excluded here as actual words of English.
+Two items, [klʌŋ] \emph{clung} and [bɹʌŋ] \emph{brung} (a dialectal past participle of \emph{bring}), are excluded here as actual words of English.
 \citet{Albright2009a} and \citet{Hayes2008a} also use this data for the purposes of model evaluation; following \citet{Frisch2000}, they use the proportion of ``yes'' responses for each item so as to derive a continuous measure of well-formedness.
@@
 The reason is that the Pearson $r$ makes a \emph{linearity assumption}.
-That is, it assumes that nonce words rated ``1'' and  ``3'', for instance, are just as different as those ``4'' and a ``6''.
+That is, it assumes that nonce words rated ``1'' and  ``3'', for instance, are just as different as are those rated ``4'' and ``6''.
 A weaker assumption, more appropriate for Likert scale data, is the \emph{monotonicity assumption}: that ``1'' is less English-like than ``3'', which is less English-like than ``4'', and so on.
@@
 Much easier to interpret are two other non-parametric statistics, the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ \citep{Noether1981}, as follows.
-These statistics are computed by comparing every model score/wordlikeness rating pair to every other: a comparison is counted as \emph{concordant} if the greater of the two model scores is the one associated with the greater of the two wordlikeness ratings (that is, the model ranks these two nonce words the in accordance with speakers' ratings), and as \emph{discordant} otherwise.
+These statistics are computed by comparing every model score/wordlikeness rating pair to every other: a comparison is counted as \emph{concordant} if the greater of the two model scores is the one associated with the greater of the two wordlikeness ratings (that is, the model ranks these two nonce words in accordance with speakers' ratings), and as \emph{discordant} otherwise.
 These two statistics differ only in the treatment of ``ties'', pairs where either the model score or wordlikeness rating are identical.
@@
 where $c$ and $d$ represent the number of concordant and discordant pairs, respectively.
-The $\tau_b$ statistic uses a similar formula, but also incorporates a penalty for ties in model score which are not also paired with ties in wordlikeness ratings, or vis versa.
+The $\tau_b$ statistic uses a similar formula, but also incorporates a penalty for ties in model score which are not also paired with ties in wordlikeness ratings, or vice versa.
 \citet{Albright2009a} uses a variant of this statistic to evaluate wordlikeness models.
@@
 The simplest possible mechanism is chosen here: an onset or rime is identified as well-formed if it occurs with non-zero frequency in a representative sample, and is identified as ill-formed otherwise.
-The sample is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms, the latter though to be particularly strongly correlated with behavioral measures \citep{Brysbaert2009}.
+The sample is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms, the latter thought to be particularly strongly correlated with behavioral measures \citep{Brysbaert2009}.
 These pronunciations are then syllabified, and individual syllables parsed into onset and rime, according to a process described in detail in Appendix \ref{syllabification}.
-This is not is not to imply that all unattested onsets or rimes should be regarded as ill-formed, or that all onsets or rimes with non-zero frequency in this data are well-formed.
+This is not to imply that all unattested onsets or rimes should be regarded as ill-formed, or that all onsets or rimes with non-zero frequency in this data are well-formed.
 For instance, \citet{Albright2009a} judges [dɹɛsp] \emph{dresp} to be phonotactically well-formed, despite the total lack of [ɛsp] rimes in English.
@@
 
-A second baseline is provided by measures of similarity to existing English words items, which has long been applied to model wordlikeness judgements (e.g., \citealt{Bailey2001}, \citealt{Greenberg1964}, \citealt{Kirby2007}, \citealt{Ohala1986b}, \citealt{Shademan2006,Shademan2007}, \citealt{Vitevitch1998,Vitevitch1999a}).
+A second baseline is provided by measures of similarity to existing English words, which has long been applied to model wordlikeness judgements (e.g., \citealt{Bailey2001}, \citealt{Greenberg1964}, \citealt{Kirby2007}, \citealt{Ohala1986b}, \citealt{Shademan2006,Shademan2007}, \citealt{Vitevitch1998,Vitevitch1999a}).
 \citeauthor{LSLT} (\citeyear{LSLT}: 151, fn.~27) suggests that grammaticality judgements in general might be influenced by similarity to existing grammatical structures, and \citet[417f.]{SPE} outline a similarity-based wordlikeness model.
@@
 
-A wide variety of lexical similarity measures were considered, including a variant of including a variant of the Generalized Neighborhood model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a set of measures provided by the Irvine Phonotactic Calculator \citep{Vaden2009}.
+A wide variety of lexical similarity measures were considered, including a variant of the Generalized Neighborhood model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a set of measures provided by the Irvine Phonotactic Calculator \citep{Vaden2009}.
 The most reliable measure is also the most venerable measure of lexical similarity: Coltheart's $N$ \citep{Coltheart1977}, which is defined as the number of words in some representative sample which can be changed into a target nonce word by a single insertion, deletion, or substitution of a phone.
@@
 The model is then used to score the onsets of the \citet{Scholes1966} nonce words.
-Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting correlation.\footnote{This is contrary to standard practices in natural language process, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter); when this is the case, there is reason to suspect the parameter values will not generalize to new data.
+Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting correlation.\footnote{This is contrary to standard practices in natural language processing, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter); when this is the case, there is reason to suspect the parameter values will not generalize to new data.
 No transformation is used here; this only has an effect on the Pearson $r$ coefficient, since they use a transformation that preserves monotonicity.}
@@
 Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}.
-The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to large variety of nonce words, though many of these words appear to have a low rating.
+The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to a large variety of nonce words, though many of these words appear to have a low rating.
 This suggests that the model is not robust enough to score whole words.
@@
 Density ($N$) &       {.67} & \uline{.79} & \uline{.86} &&       {.61} & \uline{.74} & \uline{.82} &&       {.49} &       {.57} &       {.74} &&       {.45} & \uline{.56} & \uline{.67} \\
-Bigram score  &       {.46} &       {.75} &       {.74} &&       {.34} &       {.70} &       {.79} &&       {.25} &       {.50} &       {.63} &&       {.25} &       {.50} &       {.61} \\
-MaxEnt score  &       {.70} &       {.21} &       {.53} &&       {.66} &       {.39} &       {.58} &&       {.85} &       {.61} &       {.56} && \uline{.68} &       {.16} &       {.48} \\ 
+Bigram $p$ &       {.46} &       {.75} &       {.74} &&       {.34} &       {.70} &       {.79} &&       {.25} &       {.50} &       {.63} &&       {.25} &       {.50} &       {.61} \\
+MaxEnt $p$ &       {.70} &       {.21} &       {.53} &&       {.66} &       {.39} &       {.58} &&       {.85} &       {.61} &       {.56} && \uline{.68} &       {.16} &       {.48} \\ 
 \bottomrule
@@
 \citet{Borowsky1989}, for instance, proposes a theory of possible rimes in English which makes the correct prediction regarding [ɛsp].
-This is not embedded in an acquisitional model, but models of syllable type acquisition have been proposed \citep[e.g.,][]{Levelt2000}.
+This is not embedded in an acquisitional model, but models of syllable type acquisition have been proposed \citep[e.g.,][]{Fikkert1994,Levelt2000,Pan2003,Pan2004}.
 
@@
 While the bigram and MaxEnt models do not appear to be able to reliably distinguish intermediate levels of well-formedness, it might be desirable to encode the intuition that, for example, [ʒlɪk] \emph{zhlick}, is more English-like than [bnɪk], though both have unattested onsets.
-It is also possible to imagine that phonotactic violations would have a cumulative effect on well-formedness; for instance, a nonce word with an unattested onset and an unattested rime, like [tsɪlb], might be less English-like than either [tsɪb] or [sɪlb].
-This prediction is made by the bigram and MaxEnt models, among others \citep[e.g.,][]{Albright2008,Anttila2008a}, but could easily be incorporated into a simple baseline by counting the number of violations.
+It is also possible to imagine that phonotactic violations would have a cumulative effect on well-formedness; for instance, a nonce word with an unattested onset and an unattested rime, like [tsɪlm], might be less English-like than either [tsɪl] or [sɪlm].
+This prediction is made by the bigram and MaxEnt models, among others \citep[e.g.,][]{Albright2008,Anttila1997}, but could easily be incorporated into a simple baseline by counting the number of violations.
 However, there is not yet any convincing evidence for cumulativity effects in wordlikeness tasks.

### 2012-11-16 3.tex
Moving stuff around, mostly.

@@
 The assumption appears to be that static co-occurrence restrictions which have a statistically significant trace in the lexicon will also be reflected in psycholinguistic tasks such as wordlikeness judgements.
-While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}.
+While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,GraffInPress,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}.
 In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:
@@
 As a result, Classical Latin intervocalic \emph{s} is rare.
-However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.}
+However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{GormanInPressc} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.}
 
@@
 \midrule
-\gap{}ʃ\# & 78                & 9                 & 8\%      & \multirow{2}{*}{0.026} \\
+\gap{}ʃ\# & 78                & 9                 & 8\%      & \multirow{2}{*}{.026} \\
 \gap{}s\# & 410               & 107               & 16\%     & \\
@@
 First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules.
-Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{A74}: chap.~9, \citealt{Gorman2012e}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{KK77}:189f., \citeyear{KK79}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
+Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{A74}: chap.~9, \citealt{GormanInPressc}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{KK77}:189f., \citeyear{KK79}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
 
@@
 \noindent
-\citet[196]{Inkelas2001} propose instead that the trigger must contain a labial consonant but does not require a preceding \emph{a} vowel.
+\citet{Inkelas2001} propose instead that the trigger must contain a labial consonant but does not require a preceding \emph{a} vowel.
 
@@
 Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
-\citep[][196]{Inkelas2001}
+\citep[196]{Inkelas2001}
 \end{quote}
@@
 
-\begin{unlabeledexample}
-$\displaystyle \gamma = \frac{c - d}{c + d}$
-\end{unlabeledexample}
+\begin{equation*}
+\displaystyle \gamma = \frac{c - d}{c + d}
+\end{equation*}
 
@@
 \multicolumn{4}{c|}{Experiment 1} & \multicolumn{4}{c}{Experiment 2} \\
-\multicolumn{2}{c}{aPu} & \multicolumn{2}{c|}{\textsc{disharmonic}} & \multicolumn{2}{c}{\textsc{harmonic}} & \multicolumn{2}{c}{\textsc{disharmonic}} \\
+\multicolumn{2}{c}{aPu} & \multicolumn{2}{c|}{aPı} & \multicolumn{2}{c}{aPu} & \multicolumn{2}{c}{aPı} \\
 \midrule
@@
 
-\section{Conclusion}
+\section{Conclusions}
 

### 2012-11-16 4.tex
Moving stuff around, mostly.

@@
 English syllable contact clusters have been used, by \citeauthor{Pierrehumbert1994} and others, to argue for the necessity of admitting static phonotactic constraints, and to illustrate proposals for the architecture of the phonotactic system.
-However, there are a number of reasons to reconsider the facts of this domain.
+However, there are a number of reasons to reconsider this domain.
 
@@
 For instance, she writes that ``nasal-stop sequences agree in labiality'' (175), and attributes this the status of a static constraint. 
-However, this generalization is just a narrower form of the one derived by \textsc{Nasal Place Assimilation} (see \S\ref{npa}), and which requires no further explanation. 
+However, this generalization is just a narrower form of the one derived by \textsc{Nasal Place Assimilation} (see \S\ref{npa}).
 Other derived constraints go unnoticed: for instance, she does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}). 
@@
 It is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias; 
-\citet{Labov1975}, for instance
+\citet{Labov1975} for instance
 argues for what he calls the \emph{Experimenter Principle}, 
@@
 For instance, nearly all the exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) noted by \citet[74]{Hammond1999a} are
-either complex words (e.g., \emph{jurisdiction}, \emph{madcap} \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).
+either complex words (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).
 
-Unlike prior studies, however, words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from the sample, as there is broad formal and experimental support for the claim that these words are complex.
+Unlike prior studies, words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from the sample, as there is broad formal and experimental support for the claim that these words are complex.
 This assumption allows for numerous simplifications to accounts of the phonological shape of these forms. 
@@
 There is also an interaction between Latinate prefixes on verbs and the makeup of the verbal complement. 
-Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep{Gropen1989,Harley2009}.
+Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all of which are acceptable with similar Anglo-Saxon verbs \citep{Gropen1989,Harley2009}.
 %\citep[see also][]{Gropen1989,Coppock2008}.
@@
 b. & {drink himself stupid}  & \alt{} & * & {imbibe himself stupid}    \\
-c. & {show it off}           & \alt{} & * & {exhibit it off}           \\
+%c. & {show it off}           & \alt{} & * & {exhibit it off}           \\
+c. & {break it off}           & \alt{} & * & {terminate it off}        \\
 \end{tabular}
@@
 
-This results in a list of 6,619 simplex words. 
+The filters discussed above result in a list of 6,619 simplex words.
 The full set of clusters and their frequencies are given in Appendix \ref{clusters}.
@@
 So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters.
-These constraints are static, in the sense that they have no analogues in the morphophonemics of English.
+These constraints are static, in that they have no analogues in the morphophonemics of English.
 
 \paragraph{Dorsal-labial clusters} 
-\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S?).
+\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S\ref{degemx}).
 Two-consonant velar-labial clusters are found in words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} occur. 
@@
 
-\paragraph{Degemination} The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.
+\paragraph{Degemination} 
+The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.
 \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}).
 \textsc{Degemination} is formalized here as a rule deleting the first of two segments agreeing on all feature values (except voice, possibly).
+\label{degemx}
 
@@
 
-\paragraph{Summary} All three of \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters; possible clusters which are surface exceptions to these three rules are much less likely to be attested than those which conform to them.
+\paragraph{Summary} All three of the \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters; possible clusters which are surface exceptions to these three rules are much less likely to be attested than those which conform to them.
 
@@
 Current computational models of phonotactic knowledge ``rate'' possible clusters by producing a numerical value for each cluster. 
-To apply these models to predicting the cluster inventory, these 
-these numerical values must be transformed into a categorical prediction, of either attestation or non-attestation.
+To apply these models to predicting the cluster inventory, these numerical values must be transformed into a categorical prediction of either attestation or non-attestation.
 This is accomplished with a soft-margin support vector machine \citep{Cortes1995} with a linear kernel, which optimally splits the ratings distribution into a set of attested and unattested clusters.
-This classifier is does not correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.
+This classifier does not correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.
 
@@
 As a result, an enormous lexical sample would be needed to realize clusters consistent with the \textsc{Medial Cluster Law} even if there were no constraints on the combination of medial codas and onsets in English.
-To illustrate this fact, a simulations is used to create new ``samples'' of the same size as the lexical sample used here.
+To illustrate this fact, a simulation is used to create new ``samples'' of the same size as the lexical sample used here.
 The following procedure is repeated so as to generate each observation in the simulated sample.

### 2012-11-16 E.tex
Moving stuff around, mostly.

\label{zr}

\citet{Zipf1949} famously observes the linear relationship between log rank $r$ and log word frequency $f(r)$ in several linguistic samples. This is generalized below:

\begin{unlabeledexample} 
$\displaystyle f(C, \alpha) = \frac{C}{r^\alpha}$ 
\end{unlabeledexample}

\noindent 
$C$ is a constant, sensitive to sample size. \citeauthor{Zipf1949} assumes a 1-to-1 relationship, implying $\alpha = -1$, but it is possible to compute an optimal estimate for this parameter by taking the logarithm of both sides of the equation and solving for values of $C$ and $\alpha$ that minimize the error term $\epsilon$; this can be done efficiently with linear regression:

\begin{unlabeledexample} 
$\displaystyle \textrm{log}~f(r) \sim \textrm{log}~C + \alpha~\textrm{log}~r + \epsilon$  
\end{unlabeledexample}

\citet{Good1953} notes that sparse distributions exhibit quantization at low frequencies, resulting in an artificially long flat right tail imposing an upward bias on estimates of $\alpha$. \citet[][29]{Church1991} propose a transform which eliminates this quantization. The vectors $r, n$ are defined so such that $n_i$ is the number of types which occur at frequency $r_i$ (that is, $n$ is a vector of frequencies of individual type frequencies). $Z$ contains the elements of $n$ by normalized by the points to the left and right.

\begin{unlabeledexample}
$\displaystyle Z_i = \frac{2 n_i}{r_{i + 1} - r_{i - 1}}$
\end{unlabeledexample}

\noindent \citeauthor{Church1991} do not define this transform for the lowest and highest points (i.e., when $i = 1$ or $N$), but a natural extension of their definition is to scale the endpoints according to the next intermost point, as defined below.

\begin{unlabeledexample}
$\displaystyle Z_1 = \frac{n_1}{r_2 - r_1}$
\end{unlabeledexample}

\begin{unlabeledexample}
$\displaystyle Z_N = \frac{n_N}{r_N - r_{N - 1}}$
\end{unlabeledexample}

\noindent The effect of applying this transform to sparse frequency data is shown in Figure \ref{subtlex}.

\begin{figure}
\centering
\includegraphics{zr.pdf}
\caption{The left panel shows the empirical word frequencies from the SUBTLEX-US frequency norms \citep{Brysbaert2009}. The right panel shows the same frequencies smoothed with the $Z_r$ transform.}
\label{subtlex}
\end{figure}

### 2012-11-16 abs.tex
Moving stuff around, mostly.

--- 2012-10-29 abs.tex
@@
-This dissertation argues for that phonotactic knowledge is primarily derived not from static generalizations but rather from phonological alternations and prosodic parsing.
-Chapter \ref{intro} outlines the proposal, historical precursors, and predictions.
+This dissertation outlines a program for the theory of phonotactics---the theory of speakers' knowledge of possible and impossible (or likely and unlikely) words---minimizing undue \emph{duplication} (in the sense of \citealt{KK77}:136f.) of phonotactics and phonology, and argues that the alternative view of phonotactics as stochastic, and of phonotactic learning as probabilistic inference, is not capable of accounting for the facts of this domain.
+Chapter \ref{intro} outlines the proposal, precursors, and predictions. 
 %especially with regards to order of acquisition.
@@
 It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
-A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements at least as accurately as state-of-the-art gradient wellformedness models.
+A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as accurately as state-of-the-art gradient wellformedness models.
 Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.

### 2012-11-16 ms2.tex
Moving stuff around, mostly.

@@
 \abstract{
-Gradient wordlikeness judgements do not necessarily imply that a gradient well-formedness system underlies them; gradient judgements may be an artifact of gradient rating tasks.
+Gradient wordlikeness judgements do not necessarily imply that there is a gradient well-formedness system underlying them; gradient judgements may be an artifact of gradient rating tasks.
 Dubious architectural assumptions are needed for speakers to report gradient well-formedness judgements.
@@
 Before non-onset \emph{r} many vowel contrasts are suspended (e.g., \citealp[269f.]{Fudge1969}, \citealp[][255]{Harris1994}): compare American English \emph{fern}/\emph{fir}/\emph{fur} to \emph{pet}/\emph{pit}/\emph{putt}.
-In this position, \emph{r} is the only consonant which permits variable glottalization of a following /t/ in \emph{r}-ful British dialects \citep[258]{Harris1994}, and the only consonant which does does not trigger variable deletion of a following word-final /t, d/ in American dialects \citep[8]{Guy1980}.
+In this position, \emph{r} is the only consonant which permits variable glottalization of a following /t/ in \emph{r}-ful British dialects \citep[258]{Harris1994}, and the only consonant which does not trigger variable deletion of a following word-final /t, d/ in American dialects \citep[8]{Guy1980}.
 This is shown in (\ref{tglottalization}--\ref{tddeletion}) below.

### 2012-11-18 2.tex
Small amount of writing

@@
 \noindent
-This brings the theory of wordlikeness in line with the view of syntactic grammaticality presented by foundational documents like \emph{The Logical Structure of Linguistic Theory} \citep{LSLT} and \emph{Aspects in the Theory of Syntax} \citep{ASPECTS}, which posits multiple degrees of ungrammaticality arising from different types of syntactic violations.
+This brings the theory of wordlikeness in line with the view of syntactic grammaticality presented by foundational documents like \emph{The Logical Structure of Linguistic Theory} \citep{LSLT} and \emph{Aspects of the Theory of Syntax} \citep{ASPECTS}, which posits multiple degrees of ungrammaticality arising from different types of syntactic violations.
 \citet{BARRIERS} and \citet{Huang1982} propose further elaborations of this type of theory (see \citealt[43f.]{Schutze1996} for a critique).

### 2012-11-18 4.tex
Small amount of writing

@@
 \paragraph{Degemination} 
-The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.
+The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]), in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.
 \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}).

### 2012-11-29 1.tex
Brought 4 up to speed with CLS 47 paper changes.

--- 2012-11-05 1.tex
@@
 
-I would like to suggest that there are two types of problems that arise in developing a theory of phonotactics free of duplication.
+This dissertation has two distinct but complementary aims. 
+The first is to outline the empirical scope of \emph{phonotactic theory}, the theory of speakers' knowledge of possible and impossible sounds, sound sequences, and words.
+The second is to show that the facts of this domain are consonant with a traditional view of phonotactic knowledge as categorical, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of statistical inference over lexical entries.
+
+If MSCs are epiphenomenal, there are no substantive constraints on underlying representations which are not derived from phonological processes.
+This principle---\emph{no static phonotactic constraints}---has interesting ramifications for evaluating certain competing phonological analyses.
+Consider Sanskrit aspiration alternations such as \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
+%According to one analysis, which has precedents going back to Pāṇini, the root /budh/ undergoes a process shifting aspiration leftward in certain contexts (e.g., \citeboth{Borowsky1983}, \citeboth{Hoenigswald1965}, \citeboth{Kaye1985}, \citeboth{Sag1974}, \citeyear{Sag1976}, \citeboth{Schindler1976}, \citeboth{Stemberger1980}, \citeboth{Whitney1889}:\S141f.).
+%An alternative analysis posits an underlying /bhudh/ and a process of aspirate dissimilation, a synchronic analogue of Grassman's Law (e.g., \citeboth{Anderson1970}, \citeboth{Hoard1975}, \citeboth{Kiparsky1965}:\S3.2, \citeboth{Phelps1973}, \citeboth{Phelps1975b}, \citeboth{Zwicky1965}:109f.).
+%Under the latter analysis, multiple surface aspirates (e.g., hypothetical *\emph{bhodhati}) are phonotactically marked; the former account makes no such prediction.
+%If the principle of no static phonotactic constraints can be maintained, psycholinguistic tasks could in theory be used to decide between these two accounts.
+
+%While lexical entries are thought to consist of segments, and segments to consist of feature specifications, the phonological form of a word like \emph{spectre} is not ``generated'', in any relevant sense, by the concatenation of segments /s/, /p/, and so on, or by the concatenation of syllables \{spɛk\}$_{\sigma}$ and \{tɚ\}$_{\sigma}$.
+%It would be possible to posit ``lexical concatenation rules'', on analogy with the phrase structure rules of syntax, but it is unclear how the strings produced would be linked with meanings, or how such a system would distinguish between, e.g., \emph{brick} and meaningless \emph{blick}.
+%The more appealing alternative is that the phonological form of \emph{spectre}, whatever it is, is an atom of unit of lexical memory. 
+%It remains to account for any restrictions on the contents of underlying forms. 
+%The theory of \emph{phonotactics} (which is related to \emph{syntax} through the Greek root \emph{táxis} `order') is concerned with these restrictions and how they are related to the mappings between underlying and surface forms for which the phonological component is responsible.
+%\footnote{It is tempting, however, to describe binding principles as ``syntactics''.}
+
+\section{The empirical scope of phonotactic theory}
+
+\section{The architectural basis of phonotactic knowledge}
+
+With the primary evidence for phonotactic theory now established, it is possible to consider the grammatical architecture that underlies this knowledge.
+
+\subsection{The insufficiency of morpheme structure constraints}
+
+Early generative phonologists posited that phonotactic ill-formedness derives solely from \emph{morpheme structure constraints}, restrictions on underlying representations \citep[e.g.,][]{Chomsky1966,SPE,SPR,Halle1962}.
+These come in two types \citep{Stanley1967}.
+\emph{Segment structure constraints} impose restrictions on the underlying segment inventory: e.g., in Russian, voicing is contrastive for all obstruents except /ts, tʃ, x/ \citep[22]{SPR}: [dz, dʒ, ɣ] appear in surface, but not underlying, representations.
+\emph{Sequence structure constraints} apply to underlying sequences: an example is given below.
+
+\begin{example}[An English MSC (adapted from \citealt{Chomsky1965}:100)]
+$\begin{bmatrix} +\textsc{Cons} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Liquid} \end{bmatrix}~/~\#~\begin{bmatrix} +\textsc{Cons} \end{bmatrix}~\gap$
+\end{example}
+
+\noindent
+This sequence structure constraint specifies the second of a sequence of word-initial consonants as a liquid, so as to preclude underlying /bnɪk/, for example.
+
+However, \citet{Shibatani1973} argues convincingly that not all wordlikeness contrasts can be expressed as constraints on URs.
+In German, for instance, obstruent voicing is contrastive, but neutralizes finally: e.g., [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' vs.~[ɡʀaːt]-[ɡʀaːdə] `degree(s)'
+.
+By hypothesis, the non-alternating `ridge(s)' indicates a root-final /t/, and the [t]-[d] alternation in `degree(s)' a final /d/. 
+Since the voicing of root-final obstruents is contrastive, the surface restriction on obstruent voicing is not also a sequence structure constraint.
+Yet, \citeauthor{Shibatani1973} claims, German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
+    Voicing of final obstruents is also lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \cite[506]{Ussishkin2003}.}
+
+\subsection{The duplication problem}
+\label{ss:dp}
+
+Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' knowledge of possible words, other authors suggest that the distinction btween morpheme structure constraints and phonological processes may be artifactual.
+\citet[297]{Hale1965}, \citet{Kisseberth1970b}, and \citet[212f.]{Postal1968} all observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \citeboth{Hale1965}:297, \citeboth{Kisseberth1970b}, \citeboth{Postal1968}:212f.).
+In Russian, for instance, a process of anticipatory assimilation ensures that derived clusters of obstruents agree in voice.\footnote{For sake of discussion, the somewhat more complex behavior of of [v] is ignored here.}
+
+\begin{example}[Russian voice assimilation alternations (adapted from \citealt{SPR}:22f.)]
+\label{rovs}
+\begin{tabular}{l ll ll}
+a. & [ˈʒed.ʒbɨ] & `were one to burn'      & [ˈʒet.ʃl\pal{}i] & `should one burn?'      \\
+b. & [ˈmoɡ.bɨ]  & `were (he) getting wet' & [ˈmok.l\pal{}i]  & `was (he) getting wet?' \\
+\end{tabular}
+\end{example}
+
+\noindent
+Similarly, underlying voicing is ``nondistinctive in all but the last member of a cluster of obstruents'' (\citeboth{A74}:283): e.g., [proz\pal{}bə] `request'.
+Despite their tantalizing similarity, these two facts are treated as distinct under the traditional view.
+This is what 
+\citet[?]{KK77} call the \emph{duplication} problem,
+
+In some prior sense, the alternation facts are more privileged.
+It is certainly possible to deny that the restrictions on underlying representations are psychologically real since they are ``computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for'' \citep[18]{PE}.
+%Somewhat more to the point, the derivations will come out the same whether or not the grammar includes any statement of the redundancy of obstruent voicing. 
+On the other hand, voice assimilation is essential to a concise statement of the surface forms of Russian; the only alternative would result in a huge number of allomorph; the only alternative would be to posit an enormous amount of phonologically sensitive suppletive allomorphy, so that, for instance, /moɡ/ occurs before voiced obstruents and /mok/ elsewhere.
+\namecite[205f.]{Dell1973} and \namecite[28f.]{Stampe1973} suggest that the problem is that the distinction between constraints on URs and alternations is unmotivated, and that these different levels are related by a principle now known as \emph{Stampean occultation} \cite[54]{OT}.
+In a language like Russian, in which surface obstruent clusters exceptionlessly agree in voicing, there is simply no reason for the language acquisition device to posit underlying hetero-voiced obstruent clusters: obstruent voice assimilation ``occults'' underlying */kb/, for instance.
+Were such an underlying form posited, it would surface as [gb] in all contexts.
+Unless there is some alternation (e.g., epenthesis) which would reveal /kb/, there is simply no reason for learners to posit this abstract underlying representation.\footnote{
+    \emph{Lexicon Optimization} \cite[209]{OT} implements a form of Stampean occultation notable in that it projects all non-alternating surface segments directly into URs.
+    For instance, in English, Lexicon Optimization demands underlying /ŋ/ in words like \emph{bank}, even though [ŋ] can be analyzed as an allophone of /n/ before velar consonants (e.g., \citeboth[65f.]{Borowsky1986}:65f.), simplifying the phoneme inventory.
+    However, this is not a necessary component of an implementation of \citeauthor{Dell1973} and \citeauthor{Stampe1973}'s insight about the relationship between surface and underlying sequence structure restrictions.
+    For instance, the hypothetical /bæŋk/ posited by Lexicon Optimization could be revised to /bænk/, and take a \emph{free ride} (in the sense of \citealt{Zwicky1970}) on the process of nasal place assimilation found elsewhere in English.
+    Indeed, this seems desirable, since Lexicon Optimization forces a duplication between underlying and surface constraints: for instance, [ŋ] does not appear word-initially and English speakers have considerable difficulty producing it in this position \cite{Rusaw2009}.
+    The allophonic analysis of [ŋ] predicts this fact, since there is no way to derive the [ŋ] allophone in onset position.
+    Assuming Lexicon Optimization, the only alternative is to posit an ad-hoc constraint against onset [ŋ] \citep[39]{Jusczyk2002}.
+    Future work will consider how purely allophonic relationships could be acquired in a framework like Lexicon Optimization.}
+In an architecture like Lexical Phonology, it is even possible to apply a process to individual underlying representations (i.e., at the ``lexical level'') to enforce constraints which are not surface-true.
+Consequently, MSCs are otiose, their empirical converage entirely subsumed by phonological processes.
+
+This is similar in principle to the context in which Russian obstruent voicing is discussed by \citet{SPR}.
+The principle of biuniquness in vogue at that time distinguishes between, and separates, neutralizing (morphophonemic) and non-neutralizing (phonemic) phonological processes.
+In Russian, obstruents participate in voice assimilation whether this neutralizes a phonemic distinction (\ref{rovs}a) or not (\ref{rovs}b): recall that there is no underlying /dʒ/ in Russian.
+\citeauthor{SPR} famously argues that biuniqueness (and the distinction between the morphophonemic and phonemic levels that follows from it) entails ``a significant increase in the complexity of the representation'' (24).
+While \namecite{Anderson2000} sketches an analysis which preserves biuniqueness without morphophonemic/phonemic duplication, this requires further contested assumptions---contrastive underspecification (against which, see \citeboth{Steriade1995}) and a Duke-of-York derivation, and so is consistent with \citeauthor{SPR}'s claim that biuniquness imposes unnecessary complexities.
+
+Even constraints on underlying representations may be elegantly described in terms of non-contrastive prosodic structures like the syllable \citep[e.g.,][]{Hooper1973,Kahn1976}.
+For instance, as \citet{Haugen1956} notes, many restrictions on medial consonant clusters follow from the requirement that word-medial consonant clusters be decomposable into a (possibly null) well-formed coda and a well-formed onset.
+Assuming Stampean occultation, syllable structure need not be underlyingly present to derive this constraint, contrary to what has sometimes been claimed (e.g., \citeboth{A74}:255).
+
+If some underlying cluster does not surface faithfully, it
+
+: an underlying representation containing a cluster which could not be decomposed in this fashion would be occulted.
+
+Chapter \ref{gaps} FIXME
+Similarly, initial /bn/ in English is occulted by FIXME
+
+\section{Problems for probabilistic phonotactics}
+
+\section{Outline of the dissertation}
+
+The remainder of the dissertation consists of three case studies which provide support for the novel and contentious predictions of the minimal phonotactic model sketched above.
+
+Chapter \ref{gradience} considers the status of intermediate well-formedness ratings in wordlikeness tasks. 
+It is argued that intermediate wellformedness ratings are obtained whether or not the categories being rated are graded or not.
+A primitive categorical model of wordlikeness using simple prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as well as state-of-the-art gradient wellformedness models; a simple model of similarity to existing words is also a reliable predictor of wellformedness ratings.
+Once categorical effects are controlled for, current gradient phonotactic models are largely uncorrelated with wellformedness ratings.
+
+Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns.
+It is shown that even exception-filled phonological generalizations may have a robust effect on wellformedness judgements.
+However, phonotactic generaliztaions go unlearned when they are not derived from phonological alternations; this fact is independent of ``naturalness''.
+
+Chapter \ref{clusters} investigates medial consonant clusters in English.
+Static phonotactic constraints previously proposed by \citet{Pierrehumbert1994} to describe gaps in the inventory of medial clusters are shown to be statistically unsound.
+Neutralizing phonological alternations impose robust restrictions on the cluster inventory and the remaining gaps can be attributed to the sparse nature of the lexicon.
+
+% JUNK
+
+%Whereas \citeauthor{Shibatani1973} maintains that morpheme structure constraints are insufficient to account for speakers' knowledge of possible wordhood, others argue that morpheme structure constraints are also unnecessary.
+\citet[297]{Hale1965}, \citet{Kisseberth1970b}, and \citet[212f.]{Postal1968} observe that the structural descriptions of phonological processes often are often reflected in the lexical redundancies in the same language.
+In Russian, there are alternations implicating a process of obstruent voice assimilation, and tautomorphemic obstruent clusters have uniform voicing \citep[283]{A74}.
+\citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that morpheme structure constraints are otiose, as phonological rules triggering alternations impose restrictions on the contents of URs, an effect known as \emph{Stampean occultation}.
+%\citet{Silverman2000}
+%\citet{Kiparsky1995}
+%\citet[352f.]{Hale2003a}
+%Theories of phonotactic knowledge should be evaluated by the same stringentcriteria applied to other formal arenas: neither overgeneration nor undergeneration should be permitted.
+It will be shown that the orthodoxy that phonotactic knowledge is in some sense ``probabilistic'' suffers from quite severe over- and undergeneration.
+%This heuristic has the greatest impact regarding debates about ``abstractness'' of underlying representations. 
+%It cannot be said, precisely, that it either excludes or requires any particular type of abstractness. 
+%Proposed restrictions on UR abstractness have been faulted for a number of reasons.
+In some cases, they preclude otherwise-desirable analyses of alternations.
+Another objection that could reasonably made is that any restriction on abstract URs that introduces an assymetry in URs is bad.
+%\citep[~chap.~1]{KK77}
+%I would like to suggest that there are two types of problems that arise in developing a theory of phonotactics free of duplication.
 The first type of problem consists of conflicts between theoretical assumptions and the desire to eliminate duplication.
@@
 If duplication is a pathology, this analysis is wrong.
-
-Another type of problem is the ``conundrum'', an apparent fact which poses a problem for virtually any generative theorists.
-
-pure allophony
-
-There are a few contexts which this interacts:
-
-- prosody
-- exceptionality
-- wordlikeness
-
-Syntax can hardly be described as a theory of word ``arrangement''; it would be inconceivable for syntacticians to engage in a lengthy debate about whether some word sequence is permitted in some language.
-Yet it is possible for two reasonable phonologists to dispute whether [k.p] is a permissible medial consonant cluster in English.
-While lexical entries are thought to consist of segments, and segments to consist of feature specifications, the phonological form of a word like \emph{spectre} is not ``generated'', in any relevant sense, by the concatenation of segments /s/, /p/, and so on, or by the concatenation of syllables \{spɛk\}$_{\sigma}$ and \{tɚ\}$_{\sigma}$.
-It would be possible to posit ``lexical concatenation rules'', on analogy with the phrase structure rules of syntax, but it is unclear how the strings produced would be linked with meanings, or how such a system would distinguish between, e.g., \emph{brick} and meaningless \emph{blick}.
-
-The more appealing alternative is that the phonological form of \emph{spectre}, whatever it is, is an atom of unit of lexical memory.
-It remains to account for any restrictions on the contents of underlying forms.
-The theory of \emph{phonotactics} (which is related to \emph{syntax} through the Greek root \emph{táxis} `order') is concerned with these restrictions and how they are related to the mappings between underlying and surface forms for which the phonological component is responsible.
-
-%\footnote{It is tempting, however, to describe binding principles as ``syntactics''.}
-
-I believe that the boundaries of this knowledge
-
-The boundaries of this knowledge
-and what might be acquired.
-These are not unrelated issues:
-
-One question is: why bother?
-
-It is certainly possible to imagine otherwise.
-\citet[][320]{Zimmer1969} writes that phonotactic generalizations ``have no observable consequences in the course of the normal use of the language'', and \citeauthor{PE} echo this sentiment more recently:
-
-\begin{quote}
-Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for.
-\citep[][18]{PE}
-\end{quote}
-
-This represents a principled null hypothesis, but can be quickly dismissed in light of speaker's ready judgements of possible and impossible words observed by \citeauthor{Chomsky1965}.
-Metalinguistic judgements of these sorts, reviewed in Chapter \ref{wordlikeness}, are not the only task which exemplify phonotactic knowledge; infants and adults are also thought to make use \emph{possible word constraint} in more quotidien tasks like recognizing words in running speech \citep[e.g.,][]{Brown1956,Mattys1999,McQueen1998b,Norris1997}.
-
-\subsection{Units of phonotactic description}
-
-\subsubsection{Segments}
-
-Nothing will be said about what \citet{Stanley1967} calls \emph{segment structure rules}, or language-specific constraints on the inventory of (archi)phonemes, since there is a great deal of overlap between competing theories.
-Consider possible explanations for the apparent absense of ejectives in English, It might be that English does, in a certain sense, permit underlying ejectives, but such segments are uniformly realized as plain (i.e., pulmoni) voiceless stops: in other words, the inventory is an epiphenomenon of phonological processes.
-Or, perhaps English lacks the features needed to contrast ejectives with plain voiceless stops, in which case the inventory is an epiphenomenon of the system of contrast.
-Or, perhaps the absense of ejectives follows from no particular fact, and has the grammatical status of an accidental gap.
-Further assumptions are needed to distinguish these analyses.
-
-\subsubsection{Underlying representations}
-
-The term \emph{sequence structure constraint} is now somewhat imprecise given the apparent obsolescence of boundary segments: the ``sequence structure onstraint'' that rules out onset /bn/ in English must actually be stated as a constraint militating against /\#bn/.
-\citep[149]{Kenstowicz1977}
-
-\subsubsection{Surface sequences}
-
-% discussion of Ernestus & Baayen 2003, Becker et al. 2011
-
-The notion of blocking in non-derived environments (\citealp[163]{Kiparsky1973a}, \citeyear[152]{Kiparsky1982a}, \citealp{Mascaro1976}) is well-known, but \citet{Hall2006} also amasses evidence for phonological processes restricted to non-derived environments.
-One of the most famous examples is the distribution of \emph{ich-laut} [ç] and \emph{ach-laut} [x] in German \citep{Bloomfield1930}.
-The dorsal fricative is [x] before back vowels and[ç] in other contexts.
-
-\begin{example}[German \emph{ich}- and \emph{ach-laut}]
-\begin{tabular}{l l l}
-a. & [buːx]   & `book'           \\
-   & [tɔxtər] & `daughter'       \\
-   & [naxt]   & `night'          \\
-b. & [siçt]   & `view'           \\
-   & [ʃpeçt]  & `woodpecker'     \\
-   & [ɡərʏçt] & `rumor'          \\
-   & [knøçəl] & `ankle, knuckle' \\
-   & [flɛçə]  & `surface'        \\
-\end{tabular}
-\end{example}
-
-\emph{Umlaut}, the fronting (and raising) of back vowels in certain morphological contexts, produces the front variant of the dorsal fricative; e.g., [lɔx]-[løçər] `hole-holes'.
-%\emph{B}[uːx]-\emph{B}[yːç]\emph{er} `book-books',
-
-\emph{K}[uːx]\emph{en}
-\emph{K}[uːç]\emph{en}
-
-%\emph{Mas}[oːx]-\emph{Mas}
-\footnote{Examples like \emph{Mas}[oːç]\emph{ist} `masochist', \emph{Eun}[uːç]\emph{ismus} `eunichism', first noted by \citet{Merchant1994}, suggest that this should also be restricted to assimilation within the same foot \citep[226f.]{Jensen2000}.}
-%\emph{Mas}[oːx]-\emph{Mas}[oːç]\emph{ist}
-
-\section{The model}
-
-\subsection{Syllabification}
-
-\subsection{Rule application and occultation}
-
-\subsection{Detecting violations}
-
-This is precisely the case for
-
-\subsection{Exceptionality}
-% autosegments
-
-\section{Predictions of the model}
-
-\subsection{Gradience}
-\subsubsection{Probabilistic well-formedness}
-\subsubsection{External factors}
-
-\subsection{Static phonotactics}
-\subsubsection{Insensitivity to statistics}
-\subsubsection{Saussurean arbitrariness}
-
-\subsection{Order of acquisition}
-
-A point of departure is a review of phonotactic acquisition by \citet{Hayes2004b}.
-
-\emph{pure phonotactic learner}
-%...
-However, \citeauthor{Hayes2004b} ignores the considerable evidence that infants younger than nine months of age---the \emph{terminus ante quem} for the onset of phonotactic knowledge---already have relatively rich and detailed lexical knowledge
-
-A word of caution is in order.
-An experiment which finds typically-developing infants of a certain age insensitive to an adult-like contrast using a certain task is by no means sufficient evidence to conclude that infants of this age lack this ability.
-A negative result may simply indicate that the task lacks ecological validity or requires domain-general cognitive resources beyond those of the infants.
-For instance, \citet{Werker2002} report that 14-month-old infants are insensitive to small phonological differences between novel words, but \citet{Fennell2006} shows that they are sensitive to the same contrasts when the novel words are situated in a more naturalistic naming scenario and cognitive demands are minimized.
-Other negative findings might be attributed to failure to control for properties of the stimulus, or to an experimental design which lacks sufficient statistical power.
-
-\citet{Jusczyk1993b} native vs. non-native
-\citet{Jusczyk1994} no preference for high prob
-\citet{Friederici1993}
-
-Finally, \citet{Mattys1999} find that 9-month-old infants treat hetero-organic nasal-stop clusters in nonce words as indications of word juncture, though this effect is small compared to the effect of other cues like primary stress: they do not test younger infants.
-
-\subsubsection{Prosodic parsing}
-
-Newborn infants are already able to distinguish between monosyllabic and bisyllabic words, but these infants are thought to recognize syllables ``holistically'' rather than as sequences at least until 4 months \citep[e.g.,]{Bertoncini1981,Eimas1999,Jusczyk1987}.
-The earliest evidence for segmental representations comes not from phonotactic preferences, but from infants' ability to dicriminate between familiar words like \emph{cup} and mispronunciations like \emph{*tup} at 7.5 months of age \citep{Jusczyk1995}.
-
-\subsubsection{Phonological processes}
-
-Very few studies have investigated young infants' knowledge of phonological alternations.
-One exception is a fascinating study by \citet{White2008}.
-Simplifying somewhat, the experimenters expose 8.5-month-old and 10-month-old infants to an artificial language in which the voicing of fricatives is contrastive, but voiced and voiceless variants of plosives are in a complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively.
-After familiarization, infants at both ages prefer to listen to stimuli which preserve the complementary distribution over those which disrupt it (e.g., \emph{na-poli}, \emph{rot-boli}).
-As \citeauthor{White2008} note, this suggests that infants have both extracted the plosive voicing alternation and grouped \emph{pevi} and \emph{bevi} together.
-
-\subsubsection{Lexical acquisition}
-
-As first observed by \citet{Darwin1877}, the first words infants recognize are names, their own and those of caretakers.
-Recent studies suggest that infants learn these names as early as 4 months of age \citep{Bortfeld2005,Mandel1995,Tincoff1999}.
-Infants as young as 6 months of age have learned the meaning of familiar words \citep{Bergelson2012}.
-By 8 months of age, infants are able to locate familiar and novel words in utterances \citep{Jusczyk1997,Seidl2006}.
-
-\section{Outline of the dissertation}
-
-The remainder of the dissertation consists of three case studies which provide support for the novel and contentious predictions of the minimal phonotactic model sketched above.
-
-Chapter \ref{gradience} considers evidence from wordlikeness rating tasks.
-It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
-A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements at least as acccurately as state-of-the-art gradient wellformedness models.
-Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.
-
-Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns.
-It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.
-
-Chapter \ref{clusters} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English.
-Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound, whereas phonological alternations impose robust restrictions on the cluster inventory.
-The remaining gaps in the cluster inventory are attributed to the sparse nature of the lexicon, not static phonotactic restrictions.
-
-%% NOTES HERE
-%
-%a. & \emph{B}[uːx]          & \emph{B}[yːç]\emph{er} & `book(s)' \\
-%   & \emph{L}[ɔx]           & \emph{L}[øç]\emph{er}  & `hole(s)' \\
-%   &\emph{T}[ɔx]\emph{ter} & \emph{T}[øç]\emph{ter}
-%   & \emph{N}[ax]\emph{t}   & \emph{N}[ɛç]\emph{te} & `night' \\
-%
-% While the bulk of evidence for detailed phonological entries is found in infants over a year of age \citep[e.g.,]{Werker2002,Fennell2003,Fennell2006,Stager1997,Swingley2000,Zamuner2006},
-%9-month-old infants already have learned the predominant stress pattern of their language \citep{Jusczyk1993a}.
-%\emph{τάξις}
-%Whether phonological computations or representations themselves are graded \citep[e.g.,][]{Lakoff1973} is besides the point, as metalinguistic judgements are behaviors, not mental states; they can no more be compared than can ``fear'' and ``flight response''.
-
-\citet{Silverman2000}
-\citet{Kiparsky1995}
-\citet[352f.]{Hale2003a}
-%$\begin{bmatrix} +\textsc{Dorsal} \\ +\textsc{Continuant \end{bmatrix}~\goesto~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~/~\begin{bmatrix} =\textsc{Back} \end{bmatrix}~\gap$
-\citeauthor{SPE} propose to derive gradience from the complexity (using a simple feature-counting metric) of the redundancy rules needed to map a nonce word to an attested word, and they show that this derives the increasing cline of wordlikeness running from ``consonant soup'' \emph{bnzk} to ill-formed \emph{bnick}, possible \emph{blick}, and finally lexical \emph{brick}.
-
-Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature.
-At the risk of explaining what might be no more than an an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms.
-\citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
-An example from German is shown in (\ref{fd}) below.
-
-\begin{example}[German final devoicing] 
-\label{fd}
-\begin{tabular}{l l l l}
-   & nom.sg. & nom.pl.    \\
-a. & [piːp]    & [piːpə]  & `cheep(s)'      \\
-   & [diːp]    & [diːbə]  & `thief/thieves' \\
-b. & [ɡʀaːt]   & [ɡʀaːtə] & `ridge(s)'      \\
-   & [ɡʀaːt]   & [ɡʀaːdə] & `degree(s)'     \\
-\end{tabular}
-\end{example}
-
-\noindent The plural appears to preserve a contrast in final obstruent voicing which is absent in the singular:\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984} or not \citep[e.g.,][]{Port1985}, it is irrelevant to the discussion at hand.} word-final voiced stops are never found in German, and \citet[95]{Shibatani1973} reports that ``it is easy to show that a native speaker of German rejects those forms \emph{on the ground that they end in voiced obstruents}'' (emphasis in original).
-However, given that root-final voicing is not predictable (e.g., /ɡʀaːt \alt{} ɡʀaːd/), the process of \textsc{Final Devoicing} cannot be any kind of lexical redundancy and is mysterious under the \emph{SPE} account.
-The alternative proposed by \citeauthor{Shibatani1973} and by \citet{Clayton1976}, is that to say that nonce words like [ɡʀaːd] are ill-formed not because of any underlying property, but since they fail to undergo an otherwise-exceptionless phonological process, \textsc{Final Devoicing}, or equivalently, the surface-true generalization it derives.
-
-\citet{Sommerstein1974} further notes that the computation of hypothetical underlying forms from nonce surface forms which is implied by the \emph{SPE} theory is non-trivial in the presence of phonological opacity \citep[see][528f.]{Anderson1988a}, and the rejection of the biuniqueness principle means that there is not always a unique solution, as is illustrated by the two underlying forms corresponding to [ɡʀaːt] in (\ref{fd}) above.
-
-\subsubsection{Autosegmental phonology and beyond}
-
-The arguments of \citeauthor{Shibatani1973} and others led theorists to focus their attention on properties of surface representations as determinants of wordlikeness.
+%The arguments of \citeauthor{Shibatani1973} and others led theorists to focus their attention on properties of surface representations as determinants of wordlikeness.
 Though syllabification plays no role in \emph{SPE}, it is crucial to many earlier studies \citep[for a review, see][]{Goldsmith2011b}, and it received particular attention in the 1970s.
@@
 Further enrichments to the theory are provided by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as first envisioned by][]{Pike1947a}, and theories like prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
+%Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature.
+At the risk of explaining what might be no more than an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms.
+\citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
+An example from German is shown in (\ref{fd}) below.
 
-While none of these authors discuss gradience, the syllable plays a role in the definition of a gradient measure, ``positional probability'', though to correlate closely with human judgements of wordlikeness.
-The positional probability of a nonce monosyllabic word is derived from the combined probabilities at which segments occur in onset, nucleus, and coda positions in the lexicon of a given language.
+%\emph{K}[uːx]\emph{en}
+%\emph{K}[uːç]\emph{en}
+%\footnote{Examples like \emph{Mas}[oːç]\emph{ist} `masochist', \emph{Eun}[uːç]\emph{ismus} `eunichism', first noted by \citet{Merchant1994}, suggest that this should also be restricted to assimilation within the same foot \citep[226f.]{Jensen2000}.}
+
+%Whether phonological computations or representations themselves are graded \citep[e.g.,][]{Lakoff1973} is besides the point, as metalinguistic judgements are behaviors, not mental states; they can no more be compared than can ``fear'' and ``flight response''.
+
+%\begin{example}[German \emph{ich}- and \emph{ach-laut}]
+%\begin{tabular}{l l l}
+%a. & [buːx]   & `book'           \\
+%   & [tɔxtər] & `daughter'       \\
+%   & [naxt]   & `night'          \\
+%b. & [siçt]   & `view'           \\
+%   & [ʃpeçt]  & `woodpecker'     \\
+%   & [ɡərʏçt] & `rumor'          \\
+%   & [knøçəl] & `ankle, knuckle' \\
+%   & [flɛçə]  & `surface'        \\
+%\end{tabular}
+%\end{example}
+%
+%\emph{Umlaut}, the fronting (and raising) of back vowels in certain morphological contexts, produces the front variant of the dorsal fricative; e.g., [lɔx]-[løçər] `hole-holes', \emph{B}[uːx]-\emph{B}[yːç]\emph{er} `book-books',

### 2012-11-29 2.tex
Brought 4 up to speed with CLS 47 paper changes.

@@
 
+One of the major  for the development of stochastic models o
+motivations for the 
 Recent developments in the theory of phonotactic knowledge are motivated by the claim that speakers' wordlikeness intuitions are inherently \emph{gradient}, i.e., that such intuitions are more granular than implied by a binary contrast between ``possible'' and ``impossible'' words.
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \centering
@@
 
-\begin{table}
+\begin{table}[ht]
 \centering
@@
 
-\citet{Hayes2008a} propose a sophisticated model in which well-formedness is related to a probability distribution over phone sequences, estimated according to the principle of maximum entropy, or ``MaxEnt'' \citep[e.g.,][]{Goldwater2003,Jager2007}.
+\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints (e.g., \citealt{Goldwater2003}, \citealt{Jager2007}). 
 \citeauthor{Hayes2008a} use a complex method to evaluate their model.
@@
 No transformation is used here; this only has an effect on the Pearson $r$ coefficient, since they use a transformation that preserves monotonicity.}
-
-An attempt is made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications are all used.
 However, \citet{Albright2009a} finds that the maximum entropy model, training and testing only on onsets, does not generalize to the \citet{Albright2003b} data, presumably because of the considerable diversity of rimes in the latter data.
 Consequently, the model was trained to score whole words, not just onsets, using the subset of the CMU dictionary described above.
-Following the procedure of \citet{HayesInPress}, dictionary entries were syllabified and the resulting syllables were parsed into onset, nucleus, and coda, and a novel feature [$\pm$\textsc{Coda}] is added to contrast onset and coda consonants.
-Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}.
+
+Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original paper is attempted: both their implementation and phonological feature specifications are used here.
+Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in \S\ref{syllabification} and a novel feature [$\pm$\textsc{Coda}] is added to allow the model to distinguish coda and onset consonants.
+Also, following \citeauthor{Hayes2008a}, constraints are limited to those spanning as many as three segments and an ``accuracy schedule'' of [$.001$, $.01$, $.1$, $.2$, $.3$] is used.
+Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citeauthor{Hayes2008a}.
 The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to a large variety of nonce words, though many of these words appear to have a low rating.

### 2012-11-29 3.tex
Brought 4 up to speed with CLS 47 paper changes.

@@
 
-There is a long tradition of distinguishing between structural phonotactic gaps and those which are simply accidental \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.
-In purely statistical terms, an accidental gap is one which is likely to arise without any antecedent cause, i.e., by chance.
-The contribution of early generative discussions of this contrast \citep[e.g.,][]{Halle1962,Chomsky1965,SPE} is to ground the definitions of \emph{structural} and \emph{accidental} gaps not in statistical expectations, but rather in mentalism.
-For \citeauthor{SPE}, an accidental gap is one which is not learned, one which lacks a mental antecedent.
+There are many precedents for the idea that statistical criteria could be used to differentiate \emph{accidental phonotactic gaps}, those which could arise without any antecedent cause, from \emph{structural phonotactic gaps} \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.
+
+``directly determine the mental representation of the phonotactic constraints'' \citep[180]{Frisch2004}.
+
+In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:
+
+\begin{quote}
+\ldots{} the patterns outlined above are statistically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
+\end{quote}
 
@@
 While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,GraffInPress,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}.
-In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:
-
-\begin{quote}
-\ldots{} the patterns outlined above are statistically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
-\end{quote}
 
@@
 
-\begin{table}
+\begin{table}[ht]
 \center
@@
 %\begin{example}[Wordlikeness comparisons, labial attraction]
-%\begin{table}
+%\begin{table}[ht]
 %\centering

### 2012-11-29 4.tex
Brought 4 up to speed with CLS 47 paper changes.

@@
 All that is necesary is that no phonological processes intervene between underlying representation and the surface form of medial clusters.
-Under the further assumption that underlying and surface representations are as similar as alternations allow (e.g., \citealt[205f.]{Dell1973}, \citealt[28f.]{Stampe1973}), this is the null hypothesis.
+This is little more than the definition of a ``faithful'' cluster.
+Under the further assumption that underlying and surface representations are as similar as phonological processes allow (e.g., \citealt[205f.]{Dell1973}, \citealt[28f.]{Stampe1973}), faithful realization is the null hypothesis.
 
@@
 As can be seen from Table \ref{degemtab}, the sample includes no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance.
+Stampean occultation by \textsc{Degemination} provides a natural explanation for this gap.
+It is informative to compare this to the constraint against ``ABA'' clusters proposed by \citeauthor{Pierrehumbert1994}: both are exceptionless, but only the former is unlikely to arise by chance.
 
-\begin{table}
+\begin{table}[ht]
 \centering
@@
 Unlike the expected frequency/derived constraint model, however, the constraints discovered are not limited to those which are illustrated by alternations; it may also posit static constraints, much like those proposed by \citet{Pierrehumbert1994}.
-Since this model has many experimenter-determined parameters, 
-an attempt is made to replicate the details of \citeauthor{Hayes2008a}'s evaluation as closely as is feasible: their software, model parameters, and feature specifications are all used. 
-Following the procedure of \citet{HayesInPress}, a novel feature [$\pm$\textsc{Coda}] is added to allow the model to distinguish coda and onset consonants. 
-Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citet{Hayes2008a}.
+
+Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original paper is attempted: both their implementation and phonological feature specifications are used here.
+Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in \S\ref{syllabification} and a novel feature [$\pm$
+\textsc{Coda}] is added to allow the model to distinguish coda and onset co
+nsonants.
+Also, following \citeauthor{Hayes2008a}, constraints are limited to those s
+panning as many as three segments and an ``accuracy schedule'' of [$.001$, 
+$.01$, $.1$, $.2$, $.3$] is used.
+Since the maximum entropy model produces slightly different scores on each 
+run, the worst-performing of 10 runs is reported here, following \citeauthor{Hayes2008a}.
 This model has the poorest recall of any model; compared to the derived constraints baseline, the constraints induced by the maximum entropy model are narrower.

### 2012-12-02 1.tex
First complete draft!

@@
 The first is to outline the empirical scope of \emph{phonotactic theory}, the theory of speakers' knowledge of possible and impossible sounds, sound sequences, and words.
-The second is to show that the facts of this domain are consonant with a traditional view of phonotactic knowledge as categorical, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of statistical inference over lexical entries.
+The second is to show that the core facts of this domain are compatible with a traditional view of phonotactic knowledge as independent of the lexicon, categorical, and closely related to phonological processes, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of probabilistic inference over the lexicon, and therefore gradient and independent of phonological processes.\footnote{
+    Throughout, the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory.
+    See \citealt{LignosInPressa} for a review of experimental evidence on this question.}
 
-If MSCs are epiphenomenal, there are no substantive constraints on underlying representations which are not derived from phonological processes.
-This principle---\emph{no static phonotactic constraints}---has interesting ramifications for evaluating certain competing phonological analyses.
-Consider Sanskrit aspiration alternations such as \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
-%According to one analysis, which has precedents going back to Pāṇini, the root /budh/ undergoes a process shifting aspiration leftward in certain contexts (e.g., \citeboth{Borowsky1983}, \citeboth{Hoenigswald1965}, \citeboth{Kaye1985}, \citeboth{Sag1974}, \citeyear{Sag1976}, \citeboth{Schindler1976}, \citeboth{Stemberger1980}, \citeboth{Whitney1889}:\S141f.).
-%An alternative analysis posits an underlying /bhudh/ and a process of aspirate dissimilation, a synchronic analogue of Grassman's Law (e.g., \citeboth{Anderson1970}, \citeboth{Hoard1975}, \citeboth{Kiparsky1965}:\S3.2, \citeboth{Phelps1973}, \citeboth{Phelps1975b}, \citeboth{Zwicky1965}:109f.).
-%Under the latter analysis, multiple surface aspirates (e.g., hypothetical *\emph{bhodhati}) are phonotactically marked; the former account makes no such prediction.
-%If the principle of no static phonotactic constraints can be maintained, psycholinguistic tasks could in theory be used to decide between these two accounts.
-
-%While lexical entries are thought to consist of segments, and segments to consist of feature specifications, the phonological form of a word like \emph{spectre} is not ``generated'', in any relevant sense, by the concatenation of segments /s/, /p/, and so on, or by the concatenation of syllables \{spɛk\}$_{\sigma}$ and \{tɚ\}$_{\sigma}$.
-%It would be possible to posit ``lexical concatenation rules'', on analogy with the phrase structure rules of syntax, but it is unclear how the strings produced would be linked with meanings, or how such a system would distinguish between, e.g., \emph{brick} and meaningless \emph{blick}.
-%The more appealing alternative is that the phonological form of \emph{spectre}, whatever it is, is an atom of unit of lexical memory. 
-%It remains to account for any restrictions on the contents of underlying forms. 
-%The theory of \emph{phonotactics} (which is related to \emph{syntax} through the Greek root \emph{táxis} `order') is concerned with these restrictions and how they are related to the mappings between underlying and surface forms for which the phonological component is responsible.
-%\footnote{It is tempting, however, to describe binding principles as ``syntactics''.}
+Despite a recent surge of interest in phonotactic theory, the empirical scope of the theory remains poorly defined.
+%and consequently, it is not uncommon to find suggestions that speakers do not have any type of phonotactic knowledge at all \citep[e.g.,][17f.]{PE}. 
+The first task in developing a theory of phonotactic knowledge, then, is to outline the types of facts that the theory should account for.
 
 \section{The empirical scope of phonotactic theory}
+\label{s:espt}
 
-\section{The architectural basis of phonotactic knowledge}
+Phonotactic knowledge is evidenced in quite different ways than phonological knowledge.
+In Russian, for instance, a process of anticipatory assimilation ensures that derived clusters of obstruents agree in voice.\footnote{
+    For sake of discussion, the complex behavior of [v] is ignored here.}
+
+\begin{example}[Russian voice assimilation alternations (adapted from \citealt{SPR}:22f.)]
+\label{rovs}
+\begin{tabular}{l ll ll}
+a. & [ˈʒedʒbɨ] & `were one to burn'      & [ˈʒetʃl\pal{}i] & `should one burn?'      \\
+b. & [ˈmoɡbɨ]  & `were (he) getting wet' & [ˈmokl\pal{}i]  & `was (he) getting wet?' \\
+\end{tabular}
+\end{example}
+
+There is only one concise explanation for the [dʒ]-[tʃ] and [k]-[ɡ] alternations, namely that Russian speakers have internalized a process of voice assimilation.
+Now, it is also the case that voicing in underlying representations in Russian is ``nondistinctive in all but the last member of a cluster of obstruents'' \citep[283]{A74}.
+It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters is part of the grammatical knowledge of Russian speakers. 
+For instance, \citeauthor{PE} write:
+
+\begin{quote}
+Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their langauge, since they are computationally inert and thus irrelevant to the input-output mappings that the grammar is responsible for.
+\citep[17f.]{PE}
+\end{quote}
+
+\noindent
+\citeauthor{PE} go on to label constraints on underlying representations as ``extensional'', and thus irrelevant to generative grammar, in contrast to ``intensional'' statements like phonological rules.
+Regarding the facts about Russian obstruents discussed so far, it seems to be clear that the alternation facts are in some sense more privileged.
+Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
+In contrast, were there no contraint on underlying obstruent voicing, however, it is not immediately obvious that \emph{anything} would be different. 
+Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
+There are also many cases where restrictions on underlying representations hold of underlying representations but not of derived forms.
+For instance, hiatus is exceptionally rare in native Turkish roots; of the handful of examples in the Turkish Electronic Living Lexicon \citep{TELL}, many appear to be compounds (e.g., \emph{ısıalan} `endothermic', cf.~\emph{ısı} `heat', \emph{alan} `taker').
+Yet, numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
+\citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of constraints on underlying representations.
+However, there are many facts which strongly suggest that speakers internalize phonotactic constraints.
+
+\subsection{Wordlikeness judgements}
+
+As \citet{Halle1962} and \citet{Chomsky1965} note, speakers can distinguish between a well-formed and an ill-formed word, neither of which is an actual word.
+Neither [blɪk] \emph{blick} nor [bnɪk] \emph{bnick} is a word of English, yet English speakers immediately report that only the former is ``possible'', an \emph{accidental} gap in the lexicon, whereas the latter is judged to be impossible (i.e., structurally excluded).
+There can be no question that it is part of speakers' knowledge.
+Elicited in a controlled fashion, these \emph{wordlikeness judgements} are perhaps the most important (and least controversial) source of phonotactic evidence, and they play a major role in chapters \ref{gradience}--\ref{turkish} of this dissertation.
+
+\subsection{Word production and recognition}
+\label{ss:wpr}
+
+Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
+For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; this suggests that
+multiple ejective words are not merely absent, but actually difficult for these speakers to perceive, represent, or produce.
+
+In English, sequences of adjacent obstruents which do not also agree in voice (e.g., \emph{a}[b.s]\emph{inth}) are quite rare within a word, and therefore a hetero-voiced obstruent cluster is a clue to the presence of a word boundary in running speech.
+Infants \citep[e.g.,][]{Mattys2001b} and \citep{McQueen1998b,Norris1997} are both thought to use this heuristic for word recognition in experimental settings.\footnote{
+    There is reason to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
+    Whatever the locus of [bnɪk]'s ill-formedness, for example, no segmentation into multiple words or morphs renders it a well-formed sequence.
+    See \S\ref{ss:bh} for further discussion.}
+
+\citet{Berent2001b} and \citet{Coetzee2008b} claim that non-word recognition latencies in lexical decision tasks reflect speakers' phonotactic knowledge, the hypothesis being that a phonotactically illicit nonce word will be rejected more quickly than a well-formed nonce word.
+However, phonotactic constraints are often confounded with independent predictors of lexical decision latencies.
+For instance, \citet{Coetzee2008b} finds that English speakers recognize [sp\ldots{}p] and [sk\ldots{}k] nonce monosyllables faster than [st\ldots{}t] nonce monosyllables in an auditory lexical decision task.
+\citeauthor{Coetzee2008b} attributes this to ad hoc phonotactic constraints against the former sequences, but another explanation is available.
+Even at an early stage of recognition, [stVt] is distinguished from [spVp, skVk] by its higher \emph{cohort density}: there are far more English words starting with initial [st] than with [sp] or [sk].
+High cohort density is known to inhibit auditory processing of non-words \citep[e.g.,][]{Marslen-Wilson1978} and this alone could account for the processing difference.
+
+\subsection{Loanword adaptation}
+
+Loanword adaptation may provide further evidence for the psychological reality of phonotactic knowledge.
+In Desano \citep{Kaye1974}, for instance, all underlying representations (URs) are either totally oral (e.g., [yaha] `to hear') or totally nasal (e.g., [ñãhã] `to enter'), and loanwords are made to conform to this generalization: Portuguese \emph{martelo} `hammer' is adapted as [barateru] and Spanish \emph{naranja} `orange' as [nãnãnã].
+Presumably, some component of the synchronic grammar is responsible for the fact that the restriction over native vocabulary is extended to loanwords.
+
+There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
+Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.
+
+\subsection{Alternate phonologies}
+
+Language games or speech disguises may also provide evidence for phonotactic knowledge \citep[e.g.,][]{Vaux2011}; an example appears in \S\ref{ss:bh} below.
+
+\subsection{Lexical statistics}
+
+Finally, phonotactic gaps or tendencies in the lexicon are often taken as evidence for phonotactic knowledge, under the hypothesis that grammatical constraints are the cause of these lexical generalizations.
+Chapters \ref{turkish}--\ref{gaps} consider in detail the evidentiary status of these lexical tendencies.
+
+\section{The grammatical basis of phonotactic knowledge}
 
@@
 
-Early generative phonologists posited that phonotactic ill-formedness derives solely from \emph{morpheme structure constraints}, restrictions on underlying representations \citep[e.g.,][]{Chomsky1966,SPE,SPR,Halle1962}.
-These come in two types \citep{Stanley1967}.
-\emph{Segment structure constraints} impose restrictions on the underlying segment inventory: e.g., in Russian, voicing is contrastive for all obstruents except /ts, tʃ, x/ \citep[22]{SPR}: [dz, dʒ, ɣ] appear in surface, but not underlying, representations.
-\emph{Sequence structure constraints} apply to underlying sequences: an example is given below.
+Early generative phonologists posited that phonotactic ill-formedness derives solely from \emph{morpheme structure constraints}, restrictions on underlying representations \citep{Chomsky1965,SPE,SPR,Halle1962}.
+\citet{Stanley1967} distinguishes between two types.
+\emph{Segment structure constraints} impose restrictions on the underlying segment inventory.
+For example, voicing is non-contrastive for /ts, tʃ, x/ in Russian \cite[22]{SPR}: [dz, dʒ, ɣ] appear in surface, but not underlying, representations.
+\emph{Sequence structure constraints} apply to underlying sequences; an example is given below.
 
@@
 \noindent
-This sequence structure constraint specifies the second of a sequence of word-initial consonants as a liquid, so as to preclude underlying /bnɪk/, for example.
+This sequence structure constraint specifies the second of a sequence of word-initial consonants as a liquid, so as to preclude underlying */bnɪk/, for example.
 
-However, \citet{Shibatani1973} argues convincingly that not all wordlikeness contrasts can be expressed as constraints on URs.
-In German, for instance, obstruent voicing is contrastive, but neutralizes finally: e.g., [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' vs.~[ɡʀaːt]-[ɡʀaːdə] `degree(s)'
-.
-By hypothesis, the non-alternating `ridge(s)' indicates a root-final /t/, and the [t]-[d] alternation in `degree(s)' a final /d/. 
-Since the voicing of root-final obstruents is contrastive, the surface restriction on obstruent voicing is not also a sequence structure constraint.
+However, \citet{Shibatani1973} shows quite convincingly that not all wordlikeness contrasts can be expressed as constraints on URs.
+In German, for instance, obstruent voicing is contrastive, but neutralizes finally: e.g., [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' vs.~[ɡʀaːt]-[ɡʀaːdə] `degree(s)'.
+By hypothesis, the latter root ends in /d/, so the constraint against final voiced obstruents is specific to surface representations.
 Yet, \citeauthor{Shibatani1973} claims, German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
-    Voicing of final obstruents is also lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \cite[506]{Ussishkin2003}.}
+    Voicing of final obstruents is usually lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \citep[506]{Ussishkin2003}.}
 
@@
 
-Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' knowledge of possible words, other authors suggest that the distinction btween morpheme structure constraints and phonological processes may be artifactual.
-\citet[297]{Hale1965}, \citet{Kisseberth1970b}, and \citet[212f.]{Postal1968} all observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \citeboth{Hale1965}:297, \citeboth{Kisseberth1970b}, \citeboth{Postal1968}:212f.).
-In Russian, for instance, a process of anticipatory assimilation ensures that derived clusters of obstruents agree in voice.\footnote{For sake of discussion, the somewhat more complex behavior of of [v] is ignored here.}
+Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' phonotactic knowledge, other authors observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \citealt{Hale1965}:297, \citealt{Kisseberth1970b}, \citealt{Postal1968}:212f.).
+Russian obstruent voice assimilation, discussed above, provides an example of this type: there are no hetero-voiced obstruent clusters in either underlying or surface representations.
+These two facts are tantalizingly similar, but are treated as separate if a distinction between morpheme structure constraints and phonological processes is drawn.
+This is a special case of what \citet{Kisseberth1970b} calls \emph{conspiracies}, which \citet[136]{KK77} term the \emph{duplication problem}.
+\citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that the distinction between constraints on URs and alternations is artificial, and that these different levels of description are related by a principle now known as \emph{Stampean occultation} \citep[54]{OT}.
+In a language like Russian, in which surface obstruent clusters exceptionlessly agree in voicing, there is simply no reason for the language acquisition device to posit underlying hetero-voiced obstruent clusters: obstruent voice assimilation ``occults'' underlying */kb/, for instance.
+Were such an underlying form posited, it would surface as [gb] in all contexts.\footnote{
+    \emph{Lexicon Optimization} \citep[209]{OT} implements a form of Stampean occultation notable in that it projects all non-alternating surface segments directly into URs.
+    For instance, in English, Lexicon Optimization demands underlying /ŋ/ in words like \emph{bank}, though [ŋ] could be otherwise be analyzed as an allophone of /n/ before velar consonants (e.g., \citealt[65f.]{Borowsky1986}, \citealt[85]{SPE}, \citealt[62]{Halle1985a}), eliminating /ŋ/ from the phoneme inventory.
+    However, this is not core to \citeauthor{Dell1973} and \citeauthor{Stampe1973}'s insight about the relationship between surface and underlying sequence structure restrictions.
+    For instance, the hypothetical \{bæŋk\} posited by Lexicon Optimization could be revised to /bænk/, and take a \emph{free ride} (in the sense of \citealt{Zwicky1970}) on the process of nasal place assimilation found elsewhere in English (see \S\ref{npa}).
+    Indeed, this seems desirable, since Lexicon Optimization forces a duplication between underlying and surface constraints.
+%this point is made by \citet{Chomsky1957} and \citet[69f.]{Harris1960} with respect to the principle of biuniqueness, which enforces a similar restriction.
+    For instance, [ŋ] does not appear word-initially and English speakers have considerable difficulty producing it in this position \citep{Rusaw2009}.
+    The allophonic analysis of [ŋ] predicts this fact, since there is no way to derive the [ŋ] allophone in onset position.
+    Future work will consider the how purely allophonic relationships might be acquired.}
+    %Assuming Lexicon Optimization, the only alternative is to stipulate an ad hoc constraint against onset [ŋ] \citep[39]{Jusczyk2002}.}
+In chapter \ref{gaps}, it is argued that constraints described in terms of non-contrastive prosodic structures can also be derived by Stampean occultation.
+In an architecture like Lexical Phonology, it is even possible to apply a process to individual underlying representations (i.e., at the ``lexical level'') to enforce constraints which are not surface-true.
+Consequently, morpheme structure constraints are otiose, their empirical converage entirely subsumed by phonological processes.
 
-\begin{example}[Russian voice assimilation alternations (adapted from \citealt{SPR}:22f.)]
-\label{rovs}
-\begin{tabular}{l ll ll}
-a. & [ˈʒed.ʒbɨ] & `were one to burn'      & [ˈʒet.ʃl\pal{}i] & `should one burn?'      \\
-b. & [ˈmoɡ.bɨ]  & `were (he) getting wet' & [ˈmok.l\pal{}i]  & `was (he) getting wet?' \\
-\end{tabular}
-\end{example}
+This argument against the distinction between morpheme structure constraints and phonological processes is quite similar to the famous argument (\citeyear{SPR}) against the morphophonemic/phonemic distinction.
+The principle of biuniquness in vogue at that time separates neutralizing (morphophonemic) and non-neutralizing (phonemic) processes.
+In Russian, obstruents participate in voice assimilation whether this neutralizes a phonemic distinction (\ref{rovs}a) or not (\ref{rovs}b): recall that there is no underlying /dʒ/ in Russian.
+Under biuniqueness, there must be separate neutralizing and non-neutralizing variants of this process, however.
+From this, \citeauthor{SPR} argues that biuniqueness (and the distinction between the morphophonemic and phonemic levels that follows from it) entails ``a significant increase in the complexity of the representation'' (24).
+While \citet{Anderson2000} sketches an analysis which preserves biuniqueness without morphophonemic/phonemic duplication (see also \citealt{Kiparsky1985}), this requires further contested assumptions---contrastive underspecification (against which, see \citealt{Steriade1995}) and a Duke-of-York derivation.
+This does not seem inconsistent with \citeauthor{SPR}'s claim that biuniquness imposes unnecessary additional complexities: under \citeauthor{Anderson2000}'s analysis the complexity is not a duplication, but rather a dependency on contentious theoretical assumptions.
 
-\noindent
-Similarly, underlying voicing is ``nondistinctive in all but the last member of a cluster of obstruents'' (\citeboth{A74}:283): e.g., [proz\pal{}bə] `request'.
-Despite their tantalizing similarity, these two facts are treated as distinct under the traditional view.
-This is what 
-\citet[?]{KK77} call the \emph{duplication} problem,
+\subsection{Static constraints}
 
-In some prior sense, the alternation facts are more privileged.
-It is certainly possible to deny that the restrictions on underlying representations are psychologically real since they are ``computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for'' \citep[18]{PE}.
-%Somewhat more to the point, the derivations will come out the same whether or not the grammar includes any statement of the redundancy of obstruent voicing. 
-On the other hand, voice assimilation is essential to a concise statement of the surface forms of Russian; the only alternative would result in a huge number of allomorph; the only alternative would be to posit an enormous amount of phonologically sensitive suppletive allomorphy, so that, for instance, /moɡ/ occurs before voiced obstruents and /mok/ elsewhere.
-\namecite[205f.]{Dell1973} and \namecite[28f.]{Stampe1973} suggest that the problem is that the distinction between constraints on URs and alternations is unmotivated, and that these different levels are related by a principle now known as \emph{Stampean occultation} \cite[54]{OT}.
-In a language like Russian, in which surface obstruent clusters exceptionlessly agree in voicing, there is simply no reason for the language acquisition device to posit underlying hetero-voiced obstruent clusters: obstruent voice assimilation ``occults'' underlying */kb/, for instance.
-Were such an underlying form posited, it would surface as [gb] in all contexts.
-Unless there is some alternation (e.g., epenthesis) which would reveal /kb/, there is simply no reason for learners to posit this abstract underlying representation.\footnote{
-    \emph{Lexicon Optimization} \cite[209]{OT} implements a form of Stampean occultation notable in that it projects all non-alternating surface segments directly into URs.
-    For instance, in English, Lexicon Optimization demands underlying /ŋ/ in words like \emph{bank}, even though [ŋ] can be analyzed as an allophone of /n/ before velar consonants (e.g., \citeboth[65f.]{Borowsky1986}:65f.), simplifying the phoneme inventory.
-    However, this is not a necessary component of an implementation of \citeauthor{Dell1973} and \citeauthor{Stampe1973}'s insight about the relationship between surface and underlying sequence structure restrictions.
-    For instance, the hypothetical /bæŋk/ posited by Lexicon Optimization could be revised to /bænk/, and take a \emph{free ride} (in the sense of \citealt{Zwicky1970}) on the process of nasal place assimilation found elsewhere in English.
-    Indeed, this seems desirable, since Lexicon Optimization forces a duplication between underlying and surface constraints: for instance, [ŋ] does not appear word-initially and English speakers have considerable difficulty producing it in this position \cite{Rusaw2009}.
-    The allophonic analysis of [ŋ] predicts this fact, since there is no way to derive the [ŋ] allophone in onset position.
-    Assuming Lexicon Optimization, the only alternative is to posit an ad-hoc constraint against onset [ŋ] \citep[39]{Jusczyk2002}.
-    Future work will consider how purely allophonic relationships could be acquired in a framework like Lexicon Optimization.}
-In an architecture like Lexical Phonology, it is even possible to apply a process to individual underlying representations (i.e., at the ``lexical level'') to enforce constraints which are not surface-true.
-Consequently, MSCs are otiose, their empirical converage entirely subsumed by phonological processes.
+Under the assumptions of Stampean occultation, there is no way to express substantive constraints on underlying representations except via phonological process.
+In lexical phonology, such constraints are the consequences of phonological processes operating at the ``lexical'' level, and in an Optimality Theory grammar, these are high ranked markedness constraints.\footnote{In Optimality Theory, it is often assumed that all markedness constraints are ranked above all faithfulness constraints in the ``initial state'' of the learner \citep[e.g.,][]{Smolensky1996a}.
+    This admits the possibility that markedness constraints not implicated by alternations will remain undominated \citep[e.g.,][]{Coetzee2008b}.
+    It is difficult to evaluate this proposal in the absence of any substantive proposal for the contents of \textsc{Con}, the universal constraint sets, but this has the potential to blur the traditional distinction between accidental and structural phonotactic gaps: any exceptionless gap which corresponds to a markedness constraint \emph{in any language} will be accorded structural, inviolable status, a quite powerful prediction.
+    If \textsc{Con} is sufficiently elaborate to incorporate constraints like the *[spVp] proposed by \citet{Coetzee2008b}, it might also incorporate a constraint like *[s.w], perhaps as a subcomponent of the so-called \emph{syllable contact law} \citep[e.g.,][]{Gouskova2004,Murray1983}, which disfavors syllable contact clusters with increasing sonority
+    The *[s.w] constraint is without exception in English, yet nonce words like \emph{teeswa} [tis.wa] seem quite unobjectionable.
+    Numerous other examples of this type could be adduced.
+    Consequently, this specific model is rejected here.}
+A corrolary is that if some phonotactic constraint is not (or cannot be) identified with a phonological process, it does not exist in the minds of speakers at all.
 
-This is similar in principle to the context in which Russian obstruent voicing is discussed by \citet{SPR}.
-The principle of biuniquness in vogue at that time distinguishes between, and separates, neutralizing (morphophonemic) and non-neutralizing (phonemic) phonological processes.
-In Russian, obstruents participate in voice assimilation whether this neutralizes a phonemic distinction (\ref{rovs}a) or not (\ref{rovs}b): recall that there is no underlying /dʒ/ in Russian.
-\citeauthor{SPR} famously argues that biuniqueness (and the distinction between the morphophonemic and phonemic levels that follows from it) entails ``a significant increase in the complexity of the representation'' (24).
-While \namecite{Anderson2000} sketches an analysis which preserves biuniqueness without morphophonemic/phonemic duplication, this requires further contested assumptions---contrastive underspecification (against which, see \citeboth{Steriade1995}) and a Duke-of-York derivation, and so is consistent with \citeauthor{SPR}'s claim that biuniquness imposes unnecessary complexities.
+This principle---\emph{no static phonotactic constraints}---has interesting ramifications for evaluating certain competing phonological analyses.
+Consider Sanskrit aspiration alternations such as \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
+According to one analysis, which has precedents as far back as the grammar of Pāṇini, the root /budh/ undergoes a process shifting aspiration leftward in certain contexts (e.g., \citealt{Borowsky1983}, \citealt{Hoenigswald1965}, \citealt{Kaye1985}, \citealt{Sag1974}, \citeyear{Sag1976}, \citealt{Schindler1976}, \citealt{Stemberger1980}, \citealt{Whitney1889}:\S141f.).
+An alternative analysis posits an underlying /bhudh/ and a process of aspirate dissimilation, a synchronic analogue of Grassman's Law (e.g., \citealt{Anderson1970}, \citealt{Hoard1975}, \citealt{Kiparsky1965}:\S3.2, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt{Zwicky1965}:109f.).
+Under the latter analysis, multiple surface aspirates (e.g., hypothetical *\emph{bhodhati}) are phonotactically marked; the former account makes no such prediction.
+The principle of no static phonotactic constraints, if it can be maintained, could in theory adjudicate between these competing analyses: if the postulated constraint on multiple surface aspirates is active in wordlikeness judgements or loanword adaptation, for instance, this rules out the former account.
 
-Even constraints on underlying representations may be elegantly described in terms of non-contrastive prosodic structures like the syllable \citep[e.g.,][]{Hooper1973,Kahn1976}.
-For instance, as \citet{Haugen1956} notes, many restrictions on medial consonant clusters follow from the requirement that word-medial consonant clusters be decomposable into a (possibly null) well-formed coda and a well-formed onset.
-Assuming Stampean occultation, syllable structure need not be underlyingly present to derive this constraint, contrary to what has sometimes been claimed (e.g., \citeboth{A74}:255).
+\subsection{Acquisition and the (in)dependence of phonotactic knowledge}
+\label{ss:aerbpp}
 
-If some underlying cluster does not surface faithfully, it
+\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and is therefore must be independent of other types of grammatical knowledge.
+This is not fully consistent with the available experimental evidence available, however.
 
-: an underlying representation containing a cluster which could not be decomposed in this fashion would be occulted.
+Typically-developing infants know their names and the names of their caretakers as early as 4 months of age \citep{Bortfeld2005,Mandel1995,Tincoff1999}.
+While infants are born able to distinguish between monosyllabic and bisyllabic words, the youngest infants are thought to recognize syllables ``holistically'' rather than as segment sequences \citep{Bertoncini1981,Eimas1999,Jusczyk1987}.
+As early as 6 months of age, infants know the visual referents of familiar words presented auditorily \citep{Bergelson2012}.
+At 7.5 months, phonological representations are sufficiently detailed to allow infants to discriminate between words like \emph{cup} and mispronunciations like \emph{*tup} \citep{Jusczyk1995}.
+By 8 months, infants are able to locate both familiar and novel words in continuous speech \citep{Jusczyk1997,Seidl2006}.
 
-Chapter \ref{gaps} FIXME
-Similarly, initial /bn/ in English is occulted by FIXME
+While very few studies have investigated young infants' knowledge of phonological alternations, this is the subject of a fascinating study by \citet{White2008}.
+Simplifying somewhat, the experimenters expose 8.5-month-old infants to an artificial language in which fricative voicing is contrastive, but voiced and voiceless variants of plosives are in complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively.
+After familiarization, infants prefer to listen to nonce words preserving this complementary distribution of stops over nonce words which disrupt this distribution (e.g., \emph{na-poli}, \emph{rot-boli}), suggesting that the infants have extracted an allophonic generalization concerning plosive voicing.
 
-\section{Problems for probabilistic phonotactics}
+Crucially, all of this occurs \emph{before} there is any evidence for phonotactic knowledge, for which the earliest evidence begins at approximately 9 months of age \citep{Friederici1993,Jusczyk1993b,Jusczyk1994}.
+While there are many gaps in current understanding (and keeping in mind the usual caveats about a strong interpretation of negative results), there is no reason to think that phonotactic knowledge is acquired before non-trivial amounts of lexical and phonological acquisition.
+If anything, infants's phonotactic knowledge appears to come online at approximately the same point as highly specific phonological representations including subsyllabic units and the ability to extract allophonic generalizations, as predicted by the firm link between phonological and phonotactic knowledge advocated for here.
 
@@
 
-Chapter \ref{gradience} considers the status of intermediate well-formedness ratings in wordlikeness tasks. 
-It is argued that intermediate wellformedness ratings are obtained whether or not the categories being rated are graded or not.
-A primitive categorical model of wordlikeness using simple prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as well as state-of-the-art gradient wellformedness models; a simple model of similarity to existing words is also a reliable predictor of wellformedness ratings.
-Once categorical effects are controlled for, current gradient phonotactic models are largely uncorrelated with wellformedness ratings.
+Chapter \ref{gradience} considers evidence from wordlikeness rating tasks.
+It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
+A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as accurately as state-of-the-art gradient wellformedness models.
+Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.
 
 Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns.
-It is shown that even exception-filled phonological generalizations may have a robust effect on wellformedness judgements.
-However, phonotactic generaliztaions go unlearned when they are not derived from phonological alternations; this fact is independent of ``naturalness''.
+It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.
 
-Chapter \ref{clusters} investigates medial consonant clusters in English.
-Static phonotactic constraints previously proposed by \citet{Pierrehumbert1994} to describe gaps in the inventory of medial clusters are shown to be statistically unsound.
-Neutralizing phonological alternations impose robust restrictions on the cluster inventory and the remaining gaps can be attributed to the sparse nature of the lexicon.
-
-% JUNK
+Chapter \ref{gaps} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English.
+Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound, whereas phonological alternations impose robust restrictions on the cluster inventory.
+The remaining gaps in the cluster inventory are attributed to the sparse nature of the lexicon, not static phonotactic restrictions.
 
@@
 %\emph{Umlaut}, the fronting (and raising) of back vowels in certain morphological contexts, produces the front variant of the dorsal fricative; e.g., [lɔx]-[løçər] `hole-holes', \emph{B}[uːx]-\emph{B}[yːç]\emph{er} `book-books',
+% 
+%Throughout the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory (thought see \citealt{LignosInPressa} for some recent experimental evidence bearing on this question).
+%I reject the possibility that phonotactics do exist altogether.

### 2012-12-02 2.tex
First complete draft!

@@
 motivations for the 
-Recent developments in the theory of phonotactic knowledge are motivated by the claim that speakers' wordlikeness intuitions are inherently \emph{gradient}, i.e., that such intuitions are more granular than implied by a binary contrast between ``possible'' and ``impossible'' words.
+Recent developments in the theory of phonotactic knowledge are motivated by the claim that speakers' wordlikeness intuitions are inherently \emph{gradient}, i.e., that such intuitions are more granular than a mere binary contrast between ``possible'' and ``impossible'' words.
 
@@
 Since the granularity of measurement and analysis is determined by prior hypotheses about the nature of the system under study, disputes are inevitable as scientific paradigms evolve. 
-This study argues that there are both theoretical and empirical reasons to doubt the implicit hypothesis linking scalar wordlikeness judgements and a gradient model of phonotactics. 
+This chapter argues that there are both theoretical and empirical reasons to doubt the implicit hypothesis linking scalar wordlikeness judgements and a gradient model of phonotactics. 
 First, intermediate judgements are characteristic of all gradient rating tasks, and therefore these judgements are irrelevant to the question of whether well-formedness is categorical or gradient.
@@
 As first noted by \citet{Chomsky1963b}, speakers experience difficulty processing sentences with multiple center embeddings. 
-\citet{Gibson1999} find that speakers rate sentences like (\ref{gibson}a), which is well-formed, less grammatical than (\ref{gibson}b), which is nonsensical.
+\citet{Gibson1999} perform a controlled experiment which reveals that speakers rate sentences like (\ref{gibson}a), which is well-formed, less grammatical than (\ref{gibson}b), despite the fact that a moment of reflection reveals that the latter sentence is nonsensical.
 
@@
 It is informative to consider that this well-known result has had no effect on the theory of syntactic representations, only on the theory of linguistic memory; it is recognized as the product of cognitive restrictions found in non-linguistic domains, as a \emph{task effect}.
-This contrasts with the argument made by \citet{Hayes2000}, that gradient wordlikeness judgements demand an all-encompassing revision to the grammatical architecture, reviewed below.
+This contrasts with the argument made by \citet{Hayes2000}, that gradient wordlikeness judgements demand a considerable and unmotivated revision to the grammatical architecture, as is discussed below.
 
@@
 \noindent
-However, when subjects are asked to rate, using a 7-point Likert scale, how representative individual odd counting numbers are of the concept \emph{odd number}, they freely use intermediate ratings; their results with \emph{odd number} and \emph{even number} are shown in Figure \ref{agg}.
+However, when subjects are asked to rate, using a 7-point Likert scale, how representative individual odd counting numbers are of the concept \emph{odd number}, they freely use intermediate ratings; the ratings they obtain with instances of \emph{odd number} and \emph{even number} are shown in Figure \ref{agg}.
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
@@
 In light of the complexities of gradient models, such an evaluation requires strong quantitative evidence for the superiority of gradient grammatical models.
-This study represents a first attempt to fill this gap.
+The evaluation below represents a first attempt to fill this gap.
 
@@
 
-\begin{table}[ht]
+\begin{table}[t]
 \centering
@@
 \end{tabular}
-\caption{Subject and item counts}
+\caption{Subject and item counts for the wordlikeness study}
 \label{counts}
@@
 Their item [raɪf] is excluded in this study, since this is an actual word of English, \emph{rife}.
-\citet{Albright2009a} uses this data to evaluate several computational models of wordlikeness.
+\citet{Albright2009a} uses this data to compare computational models of wordlikeness.
 
@@
 
-\citet{Scholes1966} conducts several wordlikeness tasks with 7th-grade students (approximately 12--13 years of age).
+\citet{Scholes1966} conducts several wordlikeness tasks with 
+students in 7th grade (approximately 12--13 years of age).
 The data used here is his experiment 5, in which 33 speakers provide a ``yes'' or ``no'' as to whether each of the 63 items, presented aurally, are ``likely to be usable as a word of English''. 
@@
 
-Much easier to interpret are two other non-parametric statistics, the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ \citep{Noether1981}, as follows.
+Two other non-parametric statistics the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ are much easier to interpret, as follows \citep{Noether1981}. 
 These statistics are computed by comparing every model score/wordlikeness rating pair to every other: a comparison is counted as \emph{concordant} if the greater of the two model scores is the one associated with the greater of the two wordlikeness ratings (that is, the model ranks these two nonce words in accordance with speakers' ratings), and as \emph{discordant} otherwise.
@@
 The first two models represent baselines for comparison to the latter two state-of-the-art gradient models.
-The scores for each model are provided in Appendix \ref{ratings}.
+The scores are reproduced in Appendix \ref{ratings}.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
@@
 \citet{Greenberg1964} find a correlation between wordlikeness ratings and a variant of this measure which only counts words differing by a single substitution.
-This measure is plotted against ratings from the three studies in Figure \ref{neighborhood}, where it can be seen that it accounts for much of the variance in ratings.
+This is plotted against ratings from the three studies in Figure \ref{neighborhood}, where it can be seen that it accounts for much of the variance in ratings.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
 \includegraphics{neighborhood.pdf}
-\caption{Correlation between neighborhood density and item-averaged wordlikeness ratings}
+\caption{Correlation between Coltheart's $N$ and item-averaged wordlikeness ratings}
 \label{neighborhood}
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
 \includegraphics{bigram.pdf}
-\caption{Correlation between segmental bigram score and item-averaged wordlikeness ratings}
+\caption{Correlation between smoothed segmental bigram score and item-averaged wordlikeness ratings}
 \label{bigram}
@@
 The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to a large variety of nonce words, though many of these words appear to have a low rating.
-This suggests that the model is not robust enough to score whole words.
+Apparently, this model is still not robust enough for extracting phonotactic generalizations from whole words.
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
@@
 \end{tabular}
-\caption{Correlations between item-averaged wordlikeness ratings and model scores; the largest coefficient for each statistic/data set pair is underlined.}
+\caption{Correlation between item-averaged wordlikeness ratings and model scores}
+%; the largest coefficient for each statistic/data set pair is underlined.}
 \label{scores}
@@
 \end{tabular}
-\caption{Kendall $\tau_{b}$ correlation between item-averaged wordlikeness ratings (rated ``valid'' or ``invalid'' by the gross phonotactic status baseline) and model scores}
+\caption{Kendall $\tau_{b}$ correlation between item-averaged wordlikeness ratings (rated the gross phonotactic status baseline) and model scores}
 \label{resid}
@@
 From this it can be inferred that the gradient models do not reliably predict intermediate ratings.
-These models do not reliably distinguish within classes of ``valid'' and ``invalid'' words either.
+Nor do these models reliably distinguish within classes of ``valid'' and ``invalid'' words in a way that conforms with wordlikeness ratings.
 

### 2012-12-02 3.tex
First complete draft!

@@
 
-There are many precedents for the idea that statistical criteria could be used to differentiate \emph{accidental phonotactic gaps}, those which could arise without any antecedent cause, from \emph{structural phonotactic gaps} \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.
+It has long been suggested that statistical criteria could distinguish between \emph{accidental phonotactic gaps}, those which could arise without any antecedent cause, and those gaps which are \emph{structural} in nature \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.
+Following seminal work by \citet{Mester1988} and \citet{McCarthy1988}, it is now commonly assumed that the phonotactic patterns in lexical entries ``directly determine the mental representation of the phonotactic constraints'' \citep[180]{Frisch2004} acquired by speakers, and therefore phonotactic constraints, whether categorical or gradient, can be directly inferred from statistical analysis of the lexicon.
+There is by now an enormous amount of research that adopts this assumption.
+Semitic languages, especially Arabic \citep{Coetzee2008a,Frisch2004,McCarthy1988,Pierrehumbert1993} but also Berber \citep{Elmedlaoui1995} and Tigrinya \citep{Buckley1997}, and Austronesian languages, including Javanese \citep{GraffInPress,Mester1988}, Muna \citep{Anttila2008a,Coetzee2008a}, and Samoan \citep{AldereteInPress}, have been of particular interest to linguists adopting this hypothesis; Aymara \citep{GraffInPress}, Cantonese \citep{Yip1989}, English \citep{Berkley1994a,Berkley1994b,Berkley2000,Coetzee2008b,Martin2007,Martin2011}, Gitksan \citep{Brown2010}, Japanese \citep{Kawahara2006}, Ju|'hoansi, \citep{Kinney2005}, Navajo \citep{Martin2007,Martin2011}, Ofo \citep[38f.]{MacEachern1999}, Russian \citep{Padgett1991,Padgett1992}, Shona \citep{Hayes2008a}, and Wargamay \citep{Hayes2008a} have also received a statistical treatment.
+\citet{Pozdniakov2007} analyze lexical tendencies in a diverse sample of 30 languages.
 
-``directly determine the mental representation of the phonotactic constraints'' \citep[180]{Frisch2004}.
-
-In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:
+\citet{Brown2010} presents a strong form of this hypothesis, implying that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:
 
@@
 
-Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, takes less interest in the question of what generalizations speakers internalize.
-In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique to validate a generalization about Arabic root consonant co-occurrence facts originally due to \citet{Greenberg1950}.
-The assumption appears to be that static co-occurrence restrictions which have a statistically significant trace in the lexicon will also be reflected in psycholinguistic tasks such as wordlikeness judgements.
-While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,GraffInPress,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}.
+It would be a result of great interest were it shown that statistical significance is both necessary and sufficient to identify linguistic generalizations which are internalized by speakers, but there is no reason to grant this assumption with respect to phonotactic knowledge.
+Nothing demands that the antecedent cause of a statistically reliable lexical tendency be grammatical. 
+There are reasons to suspect that many static phonotactic constraints identified in this manner have no synchronic reality.
 
-There are several arguments against this inference.
-First, negative results in formal learnability theory (see \citealt{Yang2012} for a recent review) show that there are whole classes of rules and constraints which cannot be learned under standard assumptions about acquisition (e.g., that the data is finite, can appear in any order, and that speakers do not have to request grammatical judgements from adults to achieve adult-like competence).
+By hypothesis, the phonological component may impose constraints on sound segments and sequence on underlying representations.
+Beyond this, 
+a principled null hypothesis (implied by the principle of \textsc{No Static Phonotactics}) is that there is no synchronic, grammatical explanation for underlying representations a language chooses to instantiate.
+It is a practical necessity that numerous well-formed underlying representations will be uninstantiated: the lexicon is finite but there are an infinitude of possible URs.
+The synchronic grammar cannot reasonably be expected to account for all non-existent underlying representations: for instance, the phonology of English has little to say about absence of */blɪk/ (cf.~\emph{flick}, \emph{brick}, \emph{block}, \emph{blink}).
 
-Secondly, there are a number of results which statistically reliable patterns are not learned by speakers \citep[e.g.,][]{Becker2011,Hayes2006,Hayes2009,HayesInPress}.
-This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations.
-The danger is that grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may overstate the phonotactic generalizations internalized by speakers.
+This much seems uncontroversial.
+However, what has not been appreciated is that even when a phonotactic generalization over underlying representations can be given a phonological characterization, there is a plausible alternative to the assumption that it is part of the synchronic grammar: namely, the generalization may be the result of now-complete diachronic change.
+Since sound changes often begin as phonological processes, it is no surprise that phonotactic gaps or tendencies can be explained with reference to phonological representations.
+But this suggests that the structural nature of a gap is not pertinent to determining whether the constraint is synchronically real.
 
-Another type of non-learning occurs when generalizations are statistically reliable but demonstrably incorrect.
-\citet{Legate2012}, for instance, note that the overwhelming majority of English words have initial primary stress.
-While this suggests a quantity-insensitive stress system, English stress is in fact quantity-sensitive (see \citealt{Halle1998c} and references therein).
+Saussure (\citeyear{CLG}) presents an example of phonotactic underrepresentation caused by sound change.
+With only sporadic exceptions, Old Latin intervocalic \emph{s} undergoes a conditioned phonemic merger with \emph{r}.
+This has two consequences.
+Second, it introduces many \emph{s}-\emph{r} alternations: e.g., \emph{honōs}-\emph{honōris} `honor'.
+The traditional analysis (e.g., \citealt[62]{Foley1965}, \citealt[142]{Gruber2006}, \citealt[134]{Heslin1987}, \citealt[377]{Kenstowicz1996}, \citealt[314]{Klausenburger1976}, \citealt[19]{Matthews1972a}, \citealt[88]{Roberts2012}, \citealt[526]{Watkins1970}) treats \emph{r} as the intervocalic allophone of /s/, and derives \emph{honōris} from underlying /honoːs-is/.
 
-Finally, there are many cases of statistical generalizations with a known diachronic cause but lacking any synchronic reality.
-One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}) in a discussion of rhotacism, a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}.
-As a result, Classical Latin intervocalic \emph{s} is rare.
-However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{GormanInPressc} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.}
+\begin{example}[\textsc{Rhotacism}]
+$\textrm{s}~\goesto~\textrm{r}~/~\begin{bmatrix} +\textsc{Voc} \end{bmatrix}~\gap~\begin{bmatrix} +\textsc{Voc} \end{bmatrix}$
+\end{example}
 
-Similarly, [ʃ] in Modern English derives from Old English [sk] (e.g., \emph{fisc} `fish'), and since long vowels before complex codas are not found in Old English, [ʃ] is rarely preceded by long vowels in word-final syllables to this day according to \citet{Iverson2005}.
-The type frequencies of short and long vowels before final [ʃ] and a similar segment, [s] are shown in Table \ref{ssh}.
-As can be seen, long vowels before final [s] are twice as common before final [ʃ]; according to the appropriate statistical test (the Fisher exact test, described below), this is significant.
-Yet \citet{Iverson2005} label this constraint ``phonologically accidental'' since a millennium of speakers have produced coinages (e.g., \emph{posh}, \emph{swoosh}) and  borrowings (e.g., \emph{douche}, \emph{capiche}) that disregard this generalization.\footnote{In the loanword adaptation literature, there are many reports that statistically reliable generalizations over the native lexicon are ignored in nativization processes \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}.
-On the other hand, \citet{Frisch2001} suggest that novel borrowings in Egyptian Arabic rarely violate co-occurrence restrictions in that language; however, the generalization in question was exceptionless from Proto-Semitic \citep{Ehret1989} through Classical Arabic \citep{Greenberg1950}, so any violations of this generalization are unexpected.}
+\noindent
+However, subsequent sound changes \citep[e.g.,][]{Baldi1994,Safarewicz1932}, particularly the degemination of Old Latin \emph{ss} before diphthongs and long monophthongs (e.g., \emph{caussa} $>$ \emph{causa} `cause'), introduce numerous exceptions to \textsc{Rhotacism}.
+In Classical Latin, intervocalic \emph{s} is found root-internally (\emph{asellus} `donkey', \emph{casa} `hut'), in environments derived by inflectional suffixes (\emph{uās}-\emph{uāsis} `vase', \emph{uisēre} `to view'), prefixation (\emph{dēsecāre} `to cut off'; cf.~\emph{dē} `from', \emph{secāre} `to cut'), compounding (\emph{olusātrum} `parsnip'; cf.~\emph{olus} `vegetable', \emph{ātrum} `black'), and denominal adjective formation (\emph{uentōsus} `windy'; cf.~\emph{uentus} `wind'), and is tolerated in nativized loanwords from Celtic (\emph{omāsum} `tripe'), Germanic (\emph{glaesum} `amber'), and Greek (\emph{basis} `pedestal').
+These facts lead Saussure to conclude that \textsc{Rhotacism} is no longer ``inhérente à la nature de la langue'' (202).
+While some of the apparent exceptions may be the result of opaque phonological application \citep{Heslin1987}---an explanation not yet available in Saussure's time---any formulation of rhotacization will admit nearly as many lexical exceptions as there are roots exhibiting \emph{s}-\emph{r} alternations \citep{GormanInPressc}.
+Any synchronic account of the underrepresentation of intervocalic \emph{s} must confront and the unproductive nature (as indicated by the accumulation of exceptions) of the proximate explanation for this tendency.
 
-\begin{table}[ht]
+Other cases show that constraints identified by statistical analysis are inert: that is, there is no evidence (of any of type described in \S\ref{s:espt} above) that they are internalized by speakers.
+For instance, most instances of Modern English [ʃ] derive from Old English [sk] (e.g., \emph{fisc} `fish') via sound change.
+Since Old English does not permit long vowels before complex codas, compared to similar segments like [s], [ʃ] is still rarely preceded by long vowels in word-final syllables.\footnote{
+    This sample is drawn from the CMU pronunciation dictionary, filtered by excluding words with a token frequency of less than 1 per million words in the SUBTLEX-US frequency norms.
+    Less restrictive samples give similar results.
+    The 9 words ending with a long vowel-[ʃ] sequence are \emph{douche}, \emph{leash}, \emph{gosh}, \emph{josh}, \emph{posh}, \emph{squash}, \emph{unleash}, \emph{wash}, and \emph{woosh}.}
+As can be seen, long vowels are twice as common before [s] as before [ʃ], a significant generalization according to the Fisher exact test.
+Similarly, \citet{HayesInPress} report that this constraint is discovered by the \citet{Hayes2008a} phonotactic learning model, which uses a related statistical criterion to identify constraints.
+Despite this, \citet{Iverson2005} label the constraint on long vowels before [ʃ] as ``phonologically accidental'', as a millennium of coinages (e.g., \emph{posh}) and loanwords (e.g., \emph{douche}) disregard this generalization.
+\citet{HayesInPress} find that a variant of this restriction has little to no effect on wordlikeness judgements.
+
+\begin{table}[t]
 \centering
@@
 \end{tabular}
-\caption{Type frequencies of short and long vowels before word-final [ʃ] and [s] in a subset of the CMU pronunciation dictionary containing all words with token frequency greater than 1 per million words in the SUBTLEX-US frequency norms; the 9 words containing ending in a long vowel followed by ʃ are \emph{douche}, \emph{leash}, \emph{gosh}, \emph{josh}, \emph{posh}, \emph{squash}, \emph{unleash}, \emph{wash}, and \emph{woosh}.}
+\caption{Type frequencies of short and long vowels before word-final [ʃ] and [s]}
 \label{ssh}
@@
 
-This chapter is a comprehensive investigation of three phonotactic generalizations in Turkish, comparing lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}.
-Both the lexical statistics and \citeauthor{Zimmer1969}'s experimental results merit reconsideration, because prior discussions do not relate the lexical statistics to competing formalizations of the generalizations involved, and there is widespread disagreement concerning the experimental findings.
+This latter example makes it clear that a purely statistical criterion overgenerates in the sense that it predicts phonotactic constraints which speakers do not seem to internalize.
+To account for the cases like the one above, \citet{HayesInPress} propose that speakers are biased in favor of ``natural generalizations'' in probabilistic phonotactic learning.\footnote{
+    \citeauthor{HayesInPress} do not provide an operational definition of ``naturalness'', so it is difficult to evaluate their specific results or to determine whether they apply the classification of ``natural'' and ''unnatural'' is applied in an ad hoc (or post hoc) fashion.}
+However, this chapter argues that this overgeneration gives the lie to the broader assumption that phonotactics are extracted from patterns in the lexicon.
+Rather, the only restrictions which speakers internalize are those which derive from phonological processes in the language.
+This is independent of ``naturalness'', since both ``natural'' and ``unnatural'' variants of statistically reliable static constraints are shown to be inert.
 
-One of the three generalizations, \textsc{Labial Attraction}, is shown to   state a statistically reliable generalization about the Turkish lexicon, but has no effect on speakers' wordlikeness judgements.
-This further undermines the widespread use of statistical criteria to determine which of the vast number of statistically reliable phonotactic generalizations are internalized by speakers and which are not.
-The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is static, corresponding to no phonological process in the language.
+This chapter focuses on three phonotactic generalizations in Turkish, comparing lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}. 
+Both the lexical statistics and \citeauthor{Zimmer1969}'s experimental results merit reconsideration, because prior discussions do not relate the lexical statistics to experimental data or competing formalizations of the generalizations involved.
 
@@
 A structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].
-This is more restrictive than Greek-letter notation, in that it prevents the value of one feature being applied to another.
-\citet{Odden2012} argues that the two counterexamples against this restriction adduced in \emph{SPE} (352-353) are not probative.
+This is more restrictive than Greek-letter notation, in that it prevents the value of one feature being applied to different feature; \citet{Odden2012} argues that the two counterexamples against this restriction presented in \emph{SPE} (352-353) are not probative.
 
 \subsection{Backness harmony}
+\label{ss:bh}
 
-\citeauthor{Lees1966b} (\citeyear[35]{Lees1966b}, \citeyear[284]{Lees1966a}) models the Turkish vowel harmony system with three rules.
-The most general of these rules spreads the specification [\textsc{Back}] rightward.
+\citeauthor{Lees1966b} (\citeyear[35]{Lees1966b}, \citeyear[284]{Lees1966a}) models the Turkish vowel harmony system with three rules; the most general of these rules spreads the specification [\textsc{Back}] rightward.
 
@@
 There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler}, although their final root vowel is [$+$\textsc{Back}].
-Interestingly, they may themselves be harmonic.
+Interestingly, the roots themselves may be harmonic.
 
@@
 \citet[212]{A74} and \citet{Iverson1978} argue that suffix harmony in disharmonic roots found in (\ref{turkexcept}a) requires the rule governing suffix harmony alternations to be distinguished from a sequence structure constraint governing root harmony.
-Since the rule and sequence structure constraint are otherwise identical, this constitutes an undesirable ``duplication'' in the sense of \citealt{Kisseberth1970b}.
-However, the theory of exceptionality proposed in \emph{SPE} can account for these facts without duplication \citep[197f.]{Zonneveld1978}.\footnote{\citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process.
-\citet[171f.]{Howard1972} correctly observes that this duplication is unnecessary.}
+Since the rule and sequence structure constraint are otherwise identical, this constitutes a ``duplication''. 
+However, the theory of exceptionality proposed in \emph{SPE} can account for these facts without duplication \citep[197f.]{Zonneveld1978}.\footnote{
+    \citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process.
+    \citet[171f.]{Howard1972} correctly observes that this duplication is also unnecessary.}
 
@@
 A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}]. 
-In other words, no form is never truly an ``exception''; rather, some underlying representations have non-default features which do not match the extended structural description of $R$ which requires that the target be [$+R$].
+In other words, no representation is never truly an exception; rather, some underlying representations have non-default features which do not match the extended structural description of $R$, which requires that the target be [$+R$].
 If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.
@@
 Under the assumptions so far, there is no reason to think that \emph{kabahat} is [$+$\textsc{Backness Harmony}], however: even roots with all back vowels need not undergo this rule.
-Whatever the ultimate analysis of those few words which fail to show suffix harmony, nothing depends on whether or not they are harmonic.
-
-%\begin{example}[Turkish progressives]
-%\begin{tabular}{l l l l l}
-%   & \emph{1sg. present}  & \emph{1sg. progressive} & \\
-%a. & gelirim              & geliyorum               & `come'  \\
-%b. & görürüm              & görüyorum               & `see'   \\
-%c. & atarım               & atıyorum                & `throw' \\
-%d. & bulurum              & buluyorum               & `find'  \\
-%\end{tabular}
-%\end{example}
+Whatever the ultimate analysis of those few words which fail to show suffix harmony, nothing depends on whether or not the roots in question are harmonic.
 
 It is necessary to dispense with an alternative analysis proposed by \citet{Clements1982} and \citet{Inkelas1997}. 
-This analysis has desirable properties, but a subtlety of it reintroduces the duplication of rule and sequence structure constraint.
+This analysis has desirable properties, but it subtlely reintroduces the duplication of phonological process and sequence structure constraint. 
 Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
 \emph{kul} `servant', \emph{kepek} `bran' vs.
-\emph{kapak} `lid'), whereas backness of vowels in non-initial syllables is predictable in harmonic roots; there are no prefixes in Turkish.
+\emph{kapak} `lid'), whereas backness of vowels in non-initial syllables is predictable in harmonic roots (there are no prefixes in Turkish).
 \citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots and of certain exceptional suffixes are fully specified.
@@
 One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, perhaps by structure preservation condition \citep{Kiparsky1985}.
-However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default.
+However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{
+    On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default. 
     A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone.
-However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} The the underspecification analysis is rejected here on these grounds.
+    However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} 
+    The underspecification analysis is rejected here on these grounds.
 
 While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show that Turkish speakers internalize the tendency for roots to conform to backness harmony.
-If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it.
-Several other ``external'' facts suggest that this is not the case.
-The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.\footnote{Thanks to Bert Vaux and Kie Zuraw for bringing these studies to my attention.}
+    If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it. Several other ``external'' facts suggest that this is not the case. 
+    The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games, or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.\footnote{
+    Thanks to Bert Vaux and Kie Zuraw for bringing these studies to my attention.}
 
 The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}. 
-Some speakers are said to be capable of pronouncing these non-native clusters, but in fast speech the cluster is split by anaptyxis.
-In most cases, this vowel matches the following root vowel for backness.
+    Some speakers pronounce these non-native clusters, but in fast speech the cluster is split by anaptyxis. 
+    In the majority of cases, this vowel matches the following root vowel for backness.
 
@@
     In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the inserted [$+$\textsc{Back}] vowel. 
-The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}
+    The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{
+    A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}
 
@@
 In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant. 
-On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{There is reason to believe that reduplicants bear lexical diacritics.
-Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, provides such a case.
-Verb reduplication requires a bisyllabic reduplicant.
-This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots actually have three copies of the root, and since reduplicated forms of many trisyllable verbs are ineffable.
+    On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{
+    There is reason to suspect that reduplicants bear the lexical diacritics of the roots they are derived from.
+    In Kinande, verb reduplication, 
+    Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, requires a bisyllabic reduplicant. 
+    This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots contain three copies of the root so as to satisfy this requirement, and reduplicated forms of many trisyllabic verbs are ineffable.
     However, a few trisyllabic verbs exceptionally show full reduplication.
-This is a lexical property, but one that is only seen on the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.
+    This is a lexical property, but presence or absense of this exceptionalty feature is only apparent in the form of the reduplicant.}
+    Under this definition, root harmony is not static, and so this is not an exception to the claim that static phonotactic generalizations are not preserved by operations like reduplication and truncation \citep{Silverman2000}.
 
@@
     \citet{Kabak2010} find that Turkish \textsc{Backness Harmony} has a similar effect: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}).
-\citet{Kabak2010} find that this effect does not obtain for speakers of French, a language which lacks vowel harmony.
+    \citeauthor{Kabak2010} report that this effect does not obtain for speakers of French, a language which lacks vowel harmony. 
     This implies that Turkish speakers have internalized the tendency of harmonic sequences to be root-internal and of disharmonic transitions to cross word boundaries.
@@
 The Turkish word-spotting experiment is adapted for infants by \citet{Kampen2008}.
-In this study, 9-month-old infants are familiarized with an auditory stream containing harmonic disyllabic nonce words.
+In this study, 9-month-old infants are familiarized with recordings of harmonic, disyllabic nonce words presented in isolation.
 At test time, the infants listen to the disyllabic nonce words in isolation using the head turn preference paradigm. 
@@
 Similarly, \citeauthor{Kampen2008} report that Turkish 6-month-old infants prefer to listen to harmonic nonce words such as \emph{paroz} over disharmonic nonce words like \emph{nelok}, but German 6-month-old infants show no such preference.
-
-It is not clear, however, that the segmentation effects are evidence for grammatical generalizations; for instance, while the effect is defined in terms of harmony, it operates over a domain much larger than the prosodic word over which harmony is computed.
+However, there are some caveats for identifying these effects with grammatical computations: the domain for segmentation effects is considerably larger than the domain for harmony, so the two cannot be easily identified.
+All that can be said is that there is an obvious similarity between computing the contexts for vowel harmony and for the word segmentation heuristic.
 
@@
 
-This rule is quite similar to \textsc{Backness Harmony}, but imposes an additional restriction, that targets be [$+$\textsc{High}].
+\textsc{Roundness Harmony} is quite similar to \textsc{Backness Harmony}, but imposes an additional restriction, that targets be [$+$\textsc{High}].
 
@@
 
-If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only [$+$\textsc{Rnd}] or [$-$\textsc{Rnd}] if the non-initial vowels are [$+$\textsc{High}].
-In concert with \textsc{Backness Harmony}, \textsc{Roundness Harmony} also triggers alternations which account for the shape of the dative singular (dat.sg.) and genitive singular (gen.sg.) among other suffixes.
+If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only [$+$\textsc{Rnd}] or only [$-$\textsc{Rnd}] [$+$\textsc{High}] vowels.
+In concert with \textsc{Backness Harmony}, \textsc{Roundness Harmony} also triggers alternations which account for the shape of the dative singular (dat.sg.) and genitive singular (gen.sg.) suffixes, among others.suffixes, among others.
 As was the case for \textsc{Backness Harmony}, disharmonic roots are found, and the final vowel of disharmonic roots triggers suffix harmony.
@@
 Few studies have directly investigated whether speakers are aware of the tendency for roots to conform to \textsc{Roundness Harmony}. 
-However, two pieces of the external evidence on \textsc{Backness Harmony} bear on this question.
-First, the cluster-splitting vowel found in non-native onset clusters
-, discussed above, tends to agree in roundness with following high vowels (e.g., \emph{prusya}-\emph{purusya} `Prussia').
+However, two of the external sources of evidence for \textsc{Backness Harmony} also bear on this question.
+First, the cluster-splitting vowel found in non-native onset clusters, discussed above, tends to agree in roundness with following high vowels (e.g., \emph{prusya}-\emph{purusya} `Prussia'). 
 Secondly, \textsc{Roundness Harmony} participates in reharmonization in the language game described by \citeauthor{Harrison2001}: the second \emph{ü} in \emph{bütün} `whole' reharmonizes to the [$-$\textsc{Rnd}] vowel \emph{ı} in reduplicated \emph{bütün-batın}.
@@
 
-\citet{Lees1966b} describes \textsc{Labial Attraction} as a phonological process by which ``a high, short harmonic vowel is rounded in the second syllable of a disyllabic word whose first vowel is /a/, and whose medial consonant cluster contains a labial /p, b, m, v/, and then it is de-harmonified'' (36).
+\citet[35]{Lees1966b} notes the tendency of Turkish high back vowels to be round after \emph{a}-labial consonant sequences, and formalizes this as a phonological process.
 
@@
 of \emph{sap} `stalk is \emph{sapın} rather than *\emph{sapun} that would be predicted if \textsc{Labial Attraction} triggered alternations.
-\citet[286]{Lees1966a} and \citet[311]{Zimmer1969} note the existence of root-internal exceptions to this rule (e.g., \emph{tavır} `mode') but they agree that they are surprisingly rare.
+\citeauthor{Lees1966b} and \citet{Zimmer1969} cite roots which do not conform to \textsc{Labial Attraction} (e.g., \emph{tavır} `mode') but agree that they are surprisingly rare.
+Others \citep[e.g.,][]{Clements1982} dispute this.
 
@@
 
-Others dispute that \textsc{Labial Attraction} is a reliable generalization about Turkish roots.
-\citet{Clements1982} point to additional exceptions to the rule, and \citet{Inkelas2001} provide lexical counts from the Turkish Electronic Living Lexicon \citep[TELL;][]{TELL}, a database of words known to two native Turkish speakers.
-However, no prior study reports any inferential statistics.
-
 \citet[311]{Zimmer1969} administers two paired wordlikeness tasks designed to evaluate native speakers' knowledge of \textsc{Backness Harmony}, \textsc{Roundness Harmony}, and \textsc{Labial Attraction} in roots.
-Speakers are presented with two nonce words and simply have to indicate the nonce word that is more Turkish-like.\footnote{Compared to the unpaired ratings tasks  commonly used in wordlikeness research, paired rating tasks have considerably more statistical power \citep{Gigerenzer2004}.
-When items under comparison differ only in whether or not they conform to the phonotactic generalizations, it is less likely that any contrast between the members of the pairs are caused by some omitted variable.
-Also, there is no unnecessary deception regarding the purpose of the experiment, which is known to reduce spurious response variability.
-Consequently, paired tasks are particularly sensitive to small phonotactic contrasts and are ideal for collecting wordlikeness judgements.} \citeauthor{Zimmer1969} concludes that the former two rules are reflected in wordlikeness judgements, whereas \textsc{Labial Attraction} is not.
-
+Speakers are presented with a pair of nonce words, differing only in whether they obey or violate one of these three constraints, and then indicate the nonce word that is more Turkish-like.\footnote{
+    Compared to the unpaired ratings tasks  commonly used in wordlikeness research, paired rating tasks have considerably more statistical power \citep[e.g.][]{Gigerenzer2004}, since there is little chance that any contrast between the phonotactically licit and illicit members of an otherwise-identical nonce words pair is caused by an omitted variable.
+    Consequently, paired rating tasks are ideal for collecting wordlikeness judgements.
+    The use of paired stimuli also makes more overt the purpose of the experiment from speakers, as opposed to the standard practice of concealing the purpose, the latter thought to increase response variability (see \citealt[398f.]{Hertwig2001} for discussion).}
+\citeauthor{Zimmer1969} concludes that the former two rules are reflected in wordlikeness judgements, whereas \textsc{Labial Attraction} is not. 
 Below, both lexical statistics and \citeauthor{Zimmer1969}'s wordlikeness results are analyzed statistically; \textsc{Labial Attraction} is shown to be a statistically robust generalization over the Turkish lexicon, but no variant of \textsc{Labial Attraction} is reflected in \citeauthor{Zimmer1969}'s wordlikeness study.
 In contrast, the two harmony processes have robust effects both on the lexicon and on wordlikeness.  
-This dissociation between lexical generalizations and wordlikeness results provides further evidence against the common practice of inferring phonotactic knowledge directly from lexical statistics.
+This dissociation between statistical tendencies  generalizations and wordlikeness results provides further evidence against the assumption that phonotactic knowledge can be inferred directly from lexical statistics.
 
@@
 
-Counts were computed by regular expression matching on a 9,601-root subset of the TELL database which consists of words which show no surface variation between the two informants in any inflected form.
+Counts were computed by regular expression matching on a 9,601-root subset of the TELL database which consists of roots which show no surface variation in any inflected form.
 
@@
 The counts in this table are not expected to add up to 9,601, since many roots neither could exemplify nor violate the process in question; for instance, monosyllabic words are irrelevant to root harmony.
-The Fisher exact test is used to compute a $p$-value representing the probability of the observed data arising under the null hypothesis that there is no association between the process and type frequency.
+The Fisher exact test is used to compute a $p$-value representing the probability of the observed patterns arising under the null hypothesis that there is no association between the constraint and type frequency.
 
@@
 
-\textsc{Backness Harmony} is exemplified in the lexicon insofar as there is an association between the backness of vowels in all adjacent syllables.
-Any disagreement on the backness specification of vowels in adjacent syllables are counted solely as exceptions, even if other vowel transitions in the root are harmonic.
+\textsc{Backness Harmony} is exemplified in the lexicon insofar as there is an positive association between the backness of vowels in all adjacent syllables.
+Any disagreement on the backness specification of vowels in adjacent syllables are counted solely as exceptions, even if other vowel transitions in the root are harmonic; further assumptions are necessary to determine whether a root which has one disharmonic transition may in any sense ``obey'' harmony elsewhere.
 %This follows from \emph{SPE} theory of lexical exceptionality (see also \citealt{Gouskova2012}), in which phonological exceptionality is a property of underlying representations, not individual segments.
@@
 
-Reviewing the results of wordlikeness experiments, \citet{Zimmer1969} proposes a third variant of \textsc{Labial Attraction} which is insensitive to intervening consonant place.
-This is formalized below.
-
-\begin{example}[\textsc{Labial Attraction} (revised by \citealt{Zimmer1969})]
-$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\textrm{ɑ}~\textrm{C}_0~\gap{}$
-\end{example}
-
-\noindent
-\citet{Clements1982} also  to \textsc{Labial Attraction}, arguing that it is not ``systematic''.
+\citet{Clements1982} also object to \textsc{Labial Attraction} and argue that it is not ``systematic''.
 
@@
 Whereas the sequence \emph{a}P\emph{u} is more than twice as likely as \emph{a}P\emph{ı}, the sequence \emph{a}T\emph{u} is 5 times less likely than \emph{a}T\emph{ı}. 
-This interaction is significant, as predicted by \textsc{Labial Attraction}, but contradicting a statistical interpretation of the \citeauthor{Clements1982}'s claim.
-In fact, \emph{a\ldots{}u} sequences are less, not more, common than \emph{a\ldots{}ı}.
+This interaction is significant, as predicted by \textsc{Labial Attraction}, but contrary to \citeauthor{Clements1982}'s claim.
+In fact, \emph{a\ldots{}u} sequences are less, not more, common than \emph{a\ldots{}ı} sequences, presumably a consequence of \textsc{Roundness Harmony}.
 
-\begin{table}[ht]
+\begin{table}[t]
 \centering
@@
 
-\noindent
-\citet{Inkelas2001} propose instead that the trigger must contain a labial consonant but does not require a preceding \emph{a} vowel.
-
-\begin{quote}
-Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
-\citep[196]{Inkelas2001}
-\end{quote}
-
-\noindent
-This is formalized below.
-
-\begin{example}[\textsc{Labial Attraction} (revised by \citealt{Inkelas2001})]
-$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} +\textsc{Cons} \\ +\textsc{Labial} \end{bmatrix}~\gap{}$
-\end{example}
-
-Counts from TELL for this formulation of the process are shown in Table \ref{lasi}.
-This triples the number of roots which exemplify \textsc{Labial Attraction} (top left cell) but also slightly increases the number of exceptions.
-The interaction remains significant.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{lrrr}
-\toprule
-       & \ldots{}u  & \ldots{}ı & $p$-value                      \\
-\midrule
-P\ldots{}  & 371    & 71        & \multirow{2}{*}{$6.98$\e{-49}} \\
-T\ldots{}  & 811    & 922       &                                \\
-\bottomrule
-\end{tabular}
-\caption{TELL roots sorted according to the reformulation of \textsc{Labial Attraction} proposed by \citet{Inkelas2001}}
-\label{lasi}
-\end{table}
-
 \subsection{Wordlikeness ratings}
@@
 The first used 23 native adult speakers who were permitted to select either nonce word as more like Turkish, or to indicate `no preference'.
-For the purposes below, `no preference' results are discarded.
+For the purposes below, `no preference' results are of.
 The second experiment used 32 native adults, none of whom appeared in the preceding study, and used a forced binary choice.
@@
 \noindent
-The $\gamma$ statistic ranges between -1, in the case that all paired choices are discordant, and 1, if all paired choices are concordant.\footnote{It also is possible to perform statistical tests aggregating over items, but for this small number of items, such tests have very low statistical power.}
+The $\gamma$ statistic ranges between -1, in the case that all paired choices are discordant, and 1, if all paired choices are concordant.\footnote{It also is possible to perform statistical tests aggregating over items, but for small number of items, such tests have very poor power.}
 
@@
 Both experiments include 5 pairs which differ in the presence or absence of \textsc{Roundness Harmony}. 
-As shown in Table \ref{rhw}, there is an approximately 5-to-1 preference for harmonic nonce words, and as was the case above, no disharmonic member of any pair is overall preferred.
+As shown in Table \ref{rhw}, there is an approximately 5-to-1 preference for harmonic nonce words, and as was the case above, no disharmonic member of any pair is preferred overall, across all speakers.
 Turkish speakers have a reliable preference for nonce words to conform to \textsc{Roundness Harmony} ($\gamma = 0.680$, $p = 1.1$\e{-47}).
 
-\begin{table}[ht]
+\begin{table}[t]
 \center
@@
 Both experiments include 5 pairs which either conform to \textsc{Labial Attraction} and violate \textsc{Roundness Harmony}, or vice versa; the preferences are shown in Table \ref{law}.
-There is a small preference against \textsc{Labial Attraction}, though this is non-significant ($\gamma = -0.043$, $p = 0.305$).
+There is a small preference against \textsc{Labial Attraction} (and therefore in favor of \textsc{Roundness Harmony}, though this is non-significant ($\gamma = -0.043$, $p = 0.305$).
+
 Speakers do not have the preferences predicted by \textsc{Labial Attraction}. 
-This result also extends for the variants of \textsc{Labial Attraction} proposed by \citet{Zimmer1969} and \citet{Inkelas2001}, since these variants have structural descriptions targeting a superset of the original formulation by \citet{Lees1966a}.
-It is interesting to note, however, that speakers do not have the clear opposite preference, as is predicted by \textsc{Roundness Harmony}.
+%This result also extends for the variants of \textsc{Labial Attraction} proposed by \citet{Zimmer1969} and \citet{Inkelas2001}, since these variants have structural descriptions targeting a superset of the original formulation by \citet{Lees1966a}.
+It is of interest that speakers have no apparent preference at all: one might have expected that they would prefer nonce items conforming to \textsc{Roundness Harmony}.
 
-\begin{table}[ht]
+\begin{table}[t]
 \centering
@@
 In contrast, harmony processes have a similar statistical profile, but have large effects on wordlikeness.
-The most plausible explanation for this is that  \textsc{Labial Attraction} does not trigger alternations; indeed, it is counter-exemplified by the effects of \textsc{Roundness Harmony} in suffixes.
+The most plausible explanation for this is that  \textsc{Labial Attraction} does not trigger alternations; indeed, it is counter-exemplified by the effects of \textsc{Roundness Harmony} in suffix allomorphy.
 As is noted by \citet[412f.]{Inkelas1997}, the lexicon of Turkish will, under the assumptions here, remain as it is whether or not \textsc{Labial Attraction} has a synchronic reality.
-A principle of simplicity suggests that it does not exist at all.
+The law of parsimony suggests that it does not exist at all.
 
-\citet{NiChiosain1993} and \citet{Ito1995b} argue that \textsc{Labial Attraction} is true only of native vocabulary.
-This would be a potential confound for the experimental results, since it is not implausible that the subject's in \citeauthor{Zimmer1969}'s study treated nonce words as if they were loanwords.
-%\footnote{Indeed, many wordlikeness studies include instructions to the participants to treat the stimuli much as if they were loanwords \citep[e.g.,][]{Hay2004a}.
-This may influence the choice of task model used by participants.}
-This was not the opinion of \citet[266]{Lees1966a}, though, and it is conclusively refuted by  \citet{Inkelas2001}.
-They observe that in the subset of the TELL database with etymological coding, a full 75\% of foreign vocabulary conform to \textsc{Labial Attraction}.
-Among native lexical items, only 52\% do, a significant trend but in an unexpected direction (Fisher exact test, $p = 0.042$).
-One possible explanation is that many of the languages in contact with Turkish lack the \emph{ı} phoneme, and therefore cannot contribute exceptions to \textsc{Labial Attraction}.
+\citet{Ito1995a,Ito1995b} and \citet{NiChiosain1993} claim that \textsc{Labial Attraction} holds only over the native vocabulary.
+This is a potential confound for \citeauthor{Zimmer1969}'s experiment, since it is not implausible that speakers in \citeauthor{Zimmer1969}'s study would treat nonce words much like loanwords; indeed, many wordlikeness studies include instructions to the participants to treat the stimuli much as if they were loanwords \citep[e.g.,][]{Hay2004a}.
+However, \citet{Inkelas2001} find that foreign words are more---not less---likely to conform to \textsc{Labial Attraction} than native words.
+One possible explanation is that many of the languages in contact with Turkish (including English,  Farsi, and French) lack the /ɯ/ (\emph{ı}) phoneme needed to contribute exceptions to a hypothetical \textsc{Labial Attraction}.
 
-\citet{Becker2011} attribute the inactivity of \textsc{Labial Attraction} in wordlikeness to a claim that it is phonologically unnatural, and therefore goes unlearned.
-This claim is difficult to evaluate insofar as phonotactic naturalness has continually eluded formalization and \citeauthor{Becker2011} neither propose nor refer to any theory thereof.
-\citeauthor{Becker2011} use the lack of phonetic precursors as a diagnostic of unnaturalness; however, there is extensive formal and experimental evidence for the learning of ``unnatural'' generalizations \citep[e.g.,][]{Anderson1981,Bach1972,Blevins2003,Buckley2000a,Hayes2009,Pierrehumbert2006c,Seidl2005}.
-Furthermore, other functionally-oriented theorists have considered \textsc{Labial Attraction} sufficiently natural \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995b}.
+It is not obvious that it is desirable to exclude \textsc{Labial Attraction} as a possible rule, as \citet[394, fn.
+2]{Inkelas1997} suggest that \textsc{Labial Attraction} may have even induced alternations at one point in the history of Turkish.
+However, \citeauthor{Becker2011} cite the inertness of \textsc{Labial Attraction} as evidence that naturalness constrains phonotactic learning.
 
-It is not even obvious that it is desirable to exclude \textsc{Labial Attraction} as a possible rule.
-\citet[394, fn.
-2]{Inkelas1997} suggest that \textsc{Labial Attraction} may have even induced alternations at one point in the history of Turkish.
-Beyond Turkish, the rounding of a high back vowel after a labial consonant, as in the alternative formulation of \textsc{Labial Attraction} proposed by \citet{Inkelas2001}, is widely attested \citep[e.g.,][]{Vaux1993}.
-It is also phonetically natural, as both labial consonants and high round vowels are distinguished by a very low first and second formant.
-As shown in Table \ref{lasi}, it is even a statistically reliable generalization about the Turkish lexicon.
-However, it is not reflected in alternations or in wordlikeness judgements.
+\begin{quote}
+This is clearly a complex and somewhat unnatural phonotactic, both in terms of the nonlocality of environment and the conjunction of features from two distinct triggers, and it is therefore a welcome result that not all speakers readily encoded it into a generalizable constraint.
+\citep[118]{Becker2011}
+\end{quote}
+
+If this is correct, it should be possible to show that a more ``natural'' variant of \textsc{Labial Attraction} is in fact reflected in wordlikeness judgements, assuming it too is statistically valid.
+\citeauthor{Inkelas2001} make a similar observation.
+
+\begin{quote}
+Vowel labialization following labials is not a synchronic alternation in Turkish, yet it (unlike \textsc{Labial Attraction} per se) \emph{is} a statistically supported tendency worthy of further research.
+\citep[196]{Inkelas2001}
+\end{quote}
+
+\noindent
+This proposal is formalized below as \textsc{Labial Attraction}$'$.\footnote{
+    \citet{Zimmer1969} proposes another variant of \textsc{Labial Attraction} which ignores intermediate consonantal place but requires an \emph{a} trigger in the preceding syllable. 
+    However, this is neither supported by lexical statistics or the results of his study.}
+
+\begin{shortexample}[\textsc{Labial Attraction}$'$]
+$\begin{bmatrix} -\textsc{Cons} \\ +\textsc{Back} \\ +\textsc{High} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Rnd} \end{bmatrix}~/~\begin{bmatrix} +\textsc{Lab} \\ +\textsc{Cons} \end{bmatrix}~\gap{}$
+\end{shortexample}
+
+\noindent
+The environment is now strictly local.
+Rounding of high vowels after labial consonants is also acoustically natural, as both are distinguished by low first and second formants.
+Finally, the rounding of a high back vowel after a labial consonant is widely attested \citep[e.g.,][]{Vaux1993}.
+Furthermore, as shown in Table \ref{lasi}, \textsc{Labial Attraction}$'$ is even more statistically reliable than \citeauthor{Lees1966a}'s original formulation.
+Since this reformulation targets a superset of \citeauthor{Lees1966b}'s original rule, it should be reflected in the paired rating task, but this does not obtain.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{lrrr}
+\toprule
+       & \ldots{}u  & \ldots{}ı & $p$-value                      \\
+\midrule
+P\ldots{}  & 371    & 71        & \multirow{2}{*}{$6.98$\e{-49}} \\
+T\ldots{}  & 811    & 922       &                                \\
+\bottomrule
+\end{tabular}
+\caption{TELL roots sorted according to \textsc{Labial Attraction}$'$}
+\label{lasi}
+\end{table}
 
@@
 
-Statistical reliability is not a sufficient condition for a sequence structure generalization internalized by speakers or reflected in psycholinguistic tasks.
-This case suggests that it is a necessary condition that the constraint in question be derived from an alternation, however.
-Even alternations that have large numbers of lexical exceptions have robust effects on a constrained wordlikeness task.
-Purely statistical approaches to phonotactic knowledge fail to draw the distinction between static and derived phonotactic generalizations.
+Statistical reliability is neither necessary nor sufficient to predict what speakers know about possible and impossible words in their language.
+Phonotactic constraints may go unlearned whether or not they are ``natural''.
+In the case of Turkish vowel restrictions, at least, it is precisely those constraints which are derive from phonological processes---albeit processes with a considerable number of exceptions--- which are reflected in psycholinguistic tasks.
+This should not be taken to imply that all phonotactic constraints inferred from the lexicon are illusory: for instance, \citet{Frisch2001} present psycholinguistic support for co-occurrence restrictions in Arabic posited by \citeauthor{Frisch2004} (\citeyear{Frisch2004} [1995]) on the basis of lexical data.
+As a general principle, though, it should be apparent that lexical statistics do not contribute reliable evidence for the theory of phonotactics.
+This is even more serious when a strong relationship between phonotactic and phonological representations is assumed (as it is here): contrary to common practices \citep[e.g.,][]{Mester1988,Padgett1991,Padgett1992}, lexical statistics do not provide principled evidence for the architecture of phonological grammar.
 
@@
 %  & {pul}          & {pulu}         & {pulun}        & `stamp'   \\
+%\begin{example}[Turkish progressives]
+%\begin{tabular}{l l l l l}
+%   & \emph{1sg. present}  & \emph{1sg. progressive} & \\
+%a. & gelirim              & geliyorum               & `come'  \\
+%b. & görürüm              & görüyorum               & `see'   \\
+%c. & atarım               & atıyorum                & `throw' \\
+%d. & bulurum              & buluyorum               & `find'  \\
+%\end{tabular}
+%\end{example}

### 2012-12-02 4.tex
First complete draft!

@@
 For instance, Latin has constrastive consonant and vowel length (e.g., \emph{anus} `ring' vs.~\emph{annus} `year', \emph{os} `bone' vs.~\emph{ōs} `mouth'), but there is no vowel length contrast before geminate consonants.
-However, constraints on underlying representations may also involve references to non-contrastive prosodic structures, such as syllables (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample is the result of an underlying moraic, not syllabic, contrast.}
+However, constraints on underlying representations may also involve references to non-contrastive prosodic structures, such as syllables (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{
+    See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample is the result of an underlying moraic, not syllabic, contrast.}
 For instance, as \citeauthor{Haugen1956} first notes, the constraints on word-medial consonant clusters are not arbitrary in prosodic terms:
@@
 \label{mcl}
-A medial cluster must consist of a well-formed medial coda and a well-formed medial onset
+A medial cluster consists maximally of a well-formed medial coda and a well-formed medial onset
 \end{example}
@@
 
-\begin{table}[ht]
+\begin{table}
 \centering
@@
 \end{tabular}
-\caption{Results for the cluster classification task; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
+\caption{Results for the cluster classification task}
 \label{cmresults}
 \end{table}
+%; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
 \includegraphics{co.pdf}
-\caption{Medial coda and medial onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
+\caption{Medial coda and medial onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr})}
 \label{cao}
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
 \includegraphics{sim.pdf}
-\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line).
-Simulated and observed frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
+\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line); simulated and observed frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr})}
 \label{sim}

### 2012-12-07 1.tex
Added mention of Werker's group.
Added discussion of 1) early doubts about biuniqueness and 2) early phonological category (un)learning.

@@
 
-This argument against the distinction between morpheme structure constraints and phonological processes is quite similar to the famous argument (\citeyear{SPR}) against the morphophonemic/phonemic distinction.
+This argument against the distinction between morpheme structure constraints and phonological processes is quite similar to the famous argument (\citeyear{SPR}) against the morphophonemic/phonemic distinction.\footnote{
+    This is not to imply that \citeauthor{SPR} was the first to make this argument, or to otherwise disregard this distinction: similar ideas can be found in work by \citet{Bloch1941}, \citeauthor{Bloomfield1926} (\citeyear{Bloomfield1926}:160, \citeyear{Bloomfield1962}:5f.), \citet{Chao1934}, \citet[244f.]{Hamp1953}, and \citet[47f.]{Sapir1930a}, among others (see \citealt{Anderson1985} \emph{passim}).}
 The principle of biuniquness in vogue at that time separates neutralizing (morphophonemic) and non-neutralizing (phonemic) processes.
@@
 Under biuniqueness, there must be separate neutralizing and non-neutralizing variants of this process, however.
-From this, \citeauthor{SPR} argues that biuniqueness (and the distinction between the morphophonemic and phonemic levels that follows from it) entails ``a significant increase in the complexity of the representation'' (24).
-While \citet{Anderson2000} sketches an analysis which preserves biuniqueness without morphophonemic/phonemic duplication (see also \citealt{Kiparsky1985}), this requires further contested assumptions---contrastive underspecification (against which, see \citealt{Steriade1995}) and a Duke-of-York derivation.
+From this, \citeauthor{SPR} argues that biuniqueness (and the distinction between the morphophonemic and phonemic levels that follows from it) entails ``a significant increase in the complexity of the representation\ldots{}an unwarranted complication which has no place in a scientific description of language'' (24).
+While \citeauthor{Anderson1985} (\citeyear{Anderson1985}:110f., \citeyear{Anderson2000}) sketches an analysis which preserves biuniqueness without morphophonemic/phonemic duplication (see also \citealt{Kiparsky1985}), this requires further contested assumptions---contrastive underspecification (against which, see \citealt{Steriade1995}) and a Duke-of-York derivation.
 This does not seem inconsistent with \citeauthor{SPR}'s claim that biuniquness imposes unnecessary additional complexities: under \citeauthor{Anderson2000}'s analysis the complexity is not a duplication, but rather a dependency on contentious theoretical assumptions.
@@
 By 8 months, infants are able to locate both familiar and novel words in continuous speech \citep{Jusczyk1997,Seidl2006}.
+Perception of non-native phonetic contrasts for vowels and consonants begins to decline at this age \citep{Best1994,Polka1994,Werker1981,Werker1984,Werker1988}.
 

### 2012-12-09 0ther/4old.tex
Small edits to chapter 1

\chapter{Structural and accidental gaps} \label{clusters}

Early work in generative phonology recognized that neutralizing, surface-true phonological alternations constrain possible surface forms and possibly, lexical entries (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1979}, \citealp[410f.]{Stanley1967}). Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the curent \emph{communis opinio} holds that speakers' knowledge of phonotactics---that is, language-specific sound sequence co-occurrence restrictions---is partially independent of the phonology proper.

\begin{quote}
It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained. \citep[283]{Anderson1974}
\end{quote}

This chapter argues that the only structural constraints on possible syllable contact clusters in English are derived from phonological alternations. This result serves to undermine the hypothesis of an independent phonotactics module, insofar as this particular domain has been used, by \citet{Pierrehumbert1994} and others, to illustrate the necessity of such a module and to outline proposals for its architecture.

\section{Background}

\subsection{Phonotactics and the morph}

Linguists have long recognized the morph as a privileged domain for phonotactic generalizations.

A major result of American structuralism is that generalizations which hold of the sound sequences found in underlying forms may not hold across morph boundaries. In his history of phonology, \citet[][267]{Anderson1985} indirectly credits this insight to \citet{Bloomfield1930}, who proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries. \citeauthor{Bloomfield1930} notes that [ç] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [kuːxən] `cake' $\sim$ \emph{Kuhchen} [kuːçən] `cow-let'. \citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf. \emph{Kuh} `cow'). Under the standard analysis that [x] is the allophone of /ç/ before back vowels, all that needs to be added is a restriction that the conditioning vowel belongs to the same word (or perhaps morph).

A more sophisticated approach to problems of this type, making no reference to \citeauthor{Bloomfield1930}, was published only two years later. \citet{Jakobson1932}, then a member of the Prague circle, attributes a number of phonological properties of the Russian imperative to morph juncture. Nearly all Russian infinitives are marked with a final theme vowel and /-t\pal{}/; in reflexives, this is followed by a suffix realized as [s\pal{}ə], which triggers a general process of regressive voicing assimilation. In the reflexive infinitive, the palatalization of infinitive /-t\pal/ is lost, but in the reflexive imperative, palatalization of a stem-final consonant is preserved, even when the root-final consonant is /t\pal/, as in (\ref{rusreflex}c).\footnote{Thanks to Lev Blumenfeld for assistance with this data.}

\begin{example}[Russian reflexives] \label{rusreflex}
\begin{tabular}{l l l l}
   &  infinitive       & imperative \\
a. & [slav\pal{}itsə]  & [slaf\pal{}s\pal{}ə]  & `be glorious'   \\
   & [upram\pal{}itsə] & [upram\pal{}s\pal{}ə] & `be stubborn'   \\
b. & [kras\pal{}itsə]  & [kras\pal{}s\pal{}ə]  & `put on makeup' \\
   & [ʒar\pal{}itsə]   & [ʒar\pal{}{}s\pal{}ə] & `roast'         \\
c. & [zəbytsə]         & [zəbut\pal{}s\pal{}ə] & `forget'        \\
\end{tabular}
\end{example}

\noindent \citeauthor{Jakobson1932} proposes that the preservation of palatalization in /\ldots{}t\pal{}-s\ldots{}/ is a special property of the imperative. Both (\ref{rusreflex}c) forms contain a /t\pal-s/ juncture, though palatalization is only preserved in the imperative. \citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization. A few years later, \citet{Trnka1936} makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.

This is apparent from a discussion of constraints on vowel sequences in Mixteco (a dialect cluster also known as Mixtec) given by \citet{Pike1947b}. It is apparent that \citeauthor{Pike1947b}'s constraints on vowel sequences are not intended to hold across morph boundaries, according to the phonological analysis of a Mixteco text published a few years earlier \citep{Pike1944}. A few examples are given below; the transcriptions are \citeauthor{Pike1944,Pike1947b}'s.

\begin{example}[Mixteco MSCs and complex words \citep{Pike1944,Pike1947b}]
\begin{tabular}{l l@{} l l l}
a. & * & {C}a{C}e & [ká\textsuperscript{n}dee]    & `kept inside'   \\
b. & * & {C}ə{C}e & [nìk\`əbəde]                  & `who entered'   \\
c. & * & {C}e{C}i & [teníke\textsuperscript{n}da] & `was walking    \\
d. & * & {C}i{C}e & [tenìkeet\`ə]                 & `and went away' \\
\end{tabular}
\end{example}

\noindent \citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a discovery procedure for morphs (see also \citet{Hafer1974}).

\subsubsection{Early generativism}

\citet{SPR,Halle1962} and \citet{Chomsky1965,SPE} posit a system of a system of \emph{morpheme structure constraints}---more precisely, \emph{sequence structure constraints} \citep{Stanley1967}---used to fill out redundant specifications in underlying forms, and to account for speakers' knowledge of possible words.

\subsubsection{Gradient patterns}

\section{Evaluation}

This study

% The subject of this study are medial clusters

\subsection{Data sources}

A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database. The resulting list of clusters is reproduced in Appendix \ref{appendixB} and in electronic form at the author's website.

\subsubsection{Identifying simplex words}

An analysis of lexical constraints requires a corpus of lexical representations. So as to remain agnostic about particulars of morphological theory, the goal here is to identify words which bear no overt inflection, and which do not admit any plausible further decomposition, henceforth ``simplex words''. For the structuralists, a procedure for parsing words into morphs, just the thing for locating simplex words, held the status of ``philosopher's stone'', a sort of near-mythical objective of the endeavour, but actual attempts to develop such a procedure \citep[e.g.,][]{Harris1955,Nida1948} are now recognized as anything but theory-neutral. Yet, there is no alternative to adopting some computational procedure or human-coded database as the ``gold standard'' for a study of this type. \citet{Pierrehumbert1994} analyzes a list of entries in the Collins English dictionary by hand, using her intuitions to filter out ``complex'' words as well as those which are not ``reasonably familiar''. Unfortunately, the resulting list was neither published nor circulated, and the possibility of replicating \citeauthor{Pierrehumbert1994}'s' sensations of morphological complexity is remote. Following prior studies of syllable contact clusters by \citet[ chapter 3]{Hammond1999a} and \citet[ chapter 8]{Duanmu2009}, a list of simplex words was derived from the ``lemmas'' (that is, uninflected forms) in the English portion of CELEX \citep{CELEX}, a database of morphological, phonological, and syntactic annotations based on the COBUILD corpus, a resource that was also used to construct the Collins English dictionary employed by \citeauthor{Pierrehumbert1994}.

Two further filters were applied to this list. First, any lemma coded as an unassimilated loanword is excluded. Secondly, all lemmas that have a ``morphological status'' other than ``monomorpheme'' are excluded. This is crucial because CELEX makes no guarantee that lemmas are simplex; indeed, many lemmas are products of derivation (e.g., \emph{abusive}, from \emph{abuse}). These stringent criteria have a profound effect on the makeup of the data. For instance, nearly all exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) cited by \citet[74]{Hammond1999a} are excluded as unassimilated loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as morphologically complex (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}).

One problematic case are the ``neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}.

%However, it is troublesome to use these principles to argue for an otherwise-unmotivated morph juncture if the argument is only being made to preserve the very phonotactic generalization used to motivate the juncture in the first place. Unfortunately, this is the structure of \citeauthor{Pierrehumbert1994}'s discussion of \emph{antler}: her theory rules out coda coronal obstruents (like /t/) in trisyllabic clusters, and complex medial codas like /nt/ \citep[though cf.][]{Borowsky1989}, and therefore she assumes that \emph{antler} is polymorphemic and thus /nt.l/ is unattested in simplex words. Repeated indefinitely, this sort of argument would produce morphological analyses that were little more than syllabifications.

In this vein, \citet[11--13]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation, which he takes as evidence for prefix/bound stem decomposition. Similar behaviors are found in Russian verbs which share roots but which take different prefixes or the reflexive \citep{Pesetsky1977}.

\begin{example}[Bound stem-specific allomorphy]
\begin{tabular}{l l l}
a. & \emph{adhere}   & \emph{adhesion}   \\
   & \emph{cohere}   & \emph{cohesion}   \\
b. & \emph{conceive} & \emph{conception} \\
   & \emph{perceive} & \emph{perception} \\
\end{tabular}
\end{example}

There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement. \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}. \citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.

\begin{example}[Latinate verbs and small clauses] \label{harley}
\begin{tabular}{ll@{}ll@{}l}
a. & * & \emph{exhibit him the painting} & ~ & \emph{show him the painting} \\
b. & * & \emph{imbibe himself stupid}    & ~ & \emph{drink himself stupid}  \\
c. & * & \emph{exhibit it off}           & ~ & \emph{show it off}           \\
\end{tabular}
\end{example}

Furthermore the decomposition of neo-classical compounds is observed in a number of lexical decision experiments. In general, it is known that simplex words are recognized faster than complex words \citep{Caramazza1988}. Among nonce words, those that can be exhaustively decomposed into prefix and bound stem (like *\emph{de-juvenate}) are recognized slower that contain a licit morph but cannot be exhaustively decomposed (like *\emph{de-pertoire}), and these in turn are recognized slower than those which admit no such decomposition \citep{Taft1975}.\footnote{The same contrasts are present in many other domains where decomposition into root and affix(es) is far less controversial, for instance, inflected verbs in Italian \citep{Caramazza1988,Chialant1995}.} \citet{Emmorey1989} reports facilitative auditory priming between \emph{permit} and \emph{submit} \citep[though cf.][]{Marslen-Wilson1994}, and \citet{Taft2004a} find facilitative masked priming between orthographically presented pairs like \emph{virus}--\emph{viral}, but not \emph{future}--\emph{futile}, which do not share a possible bound stem.\footnote{Many have argued that full decomposition is only one of the two methods for the processing of complex words, the other being whole word looking, either running in serial \citep{Caramazza1988} or in parallel \citep{Baayen1997b}. Whatever evidence that is amassed in support of a whole word lookup procedure, however, does not undermine the considerable evidence for decomposition.}

\subsubsection{Syllabification and phonologization}

CELEX provides broad transcriptions of individual wordforms which also include syllable boundaries. However, the syllabification procedure used is not described in the documentation, and an inspection of the data reveals various unsystematic elements. In a list of English simplex words, for instance, a medial cluster and adjacent nuclei should always be syllabified in the same fashion, since there is no convincing evidence for truly ``contrastive'' syllabification in any language \citep[see]{Elfner2006}. But many syllabification contrasts can be found in the CELEX data, for instance between \emph{chemistry} [ˈkɛ.mɪ.strɪ] and \emph{ministry} [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][II.294]{AOE}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast does not conform to any native speaker intuition.

For this reason, an automated syllabification procedure was developed to separate medial clusters in the CELEX  monomorphs into coda and onset. Since this procedure is applied only to medial clusters, it is possible to remain agnostic on several contentious issues concerning syllabification of English; for instance, there is no need to address the status of so-called ``ambisyllabic'' consonants, since this does not occur with medial clusters, or morphological effects on syllabification, since only monomorphs are used.

The technique used is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948}, \citealt[75]{Pulgram1970}, \citealt[42f.]{Kahn1976}, \citealt[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset. Just what is a ``possible'' medial onset is defined by the onsets that occur in initial position \citep[though cf.][36]{Fischer-Jorgensen1952}. Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), so onset maximization assigns the entire medial cluster to the onset, leaving the medial coda empty. In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} is not found word-initially; the maximal onset here is [str], and the residual [n] is assigned to the coda.

There is one case where unchecked maximization of medial onsets produces incorrect syllabifications. When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}. \citet[][55]{Harris1994} notes that onset maximization produces an incorrect result when a stressed lax vowel is followed by a medial cluster which is also a valid onset: in \emph{whisper}, \emph{vestige}, and \emph{musket}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty. The approach adopted here is is to assign the first consonant of a medial cluster to the coda of a preceding syllable if it follows a stressed lax vowel \citep[cf.][48]{Pulgram1970}.

The previous principle has the potential to introduce a related issue for parsing medial affricates. Since the \emph{Grundzüge der Phonologie} \citep{Trubetzkoy1958}, first published in 1939, affricates are understood as single plosive segments distinguished from pure stops by the presence of stridency \citep[24]{Jakobson1961} or a delayed release (\emph{SPE}:321f). If however, the English affricates [tʃ, dʒ] are treated as stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990}, this procedure will incorrectly assign the two components of the medial affricate to separate syllables before stressed lax vowels (e.g., \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}) unless there is an ad-hoc constraint that treats these sequences as indivisible \citep[e.g.,][]{Wells1990}. Here, the traditional treatment of affricates as single segments is assumed. This assumption can be motivated by the tendency of affricates to pattern with individual segments in phonotactic generalizations. For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but bans all other onset clusters \citep[9]{Launey2011}.%, and \citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Geor

This potentially introduces a problem with medial affricates. At least as far back as the 1939 first edition of the \emph{Grundzüge der Phonologie} \citep{Trubetzkoy1958}, affricates have been analyzed as single plosive segments distinguished from stops by the presence of stridency \citep[24]{Jakobson1961} or a delayed release (\emph{SPE}:321f). If however, the English affricates [tʃ, dʒ] are treated as stop-fricative sequences (e.g., \citealt{Hualde1988}, \citealt{Lombardi1990}), this procedure will incorrectly assign the two components of the medial affricate to separate syllables before stressed lax vowels (e.g., \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}) unless there is an ad-hoc constraint that treats these sequences as indivisible \citep[e.g.,][]{Wells1990}. Here, the traditional treatment of affricates as single segments is assumed. This assumption can be motivated by the tendency of affriates to pattern with individual segments in phonotactic generalizations. For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but otherwise does not permit other onset clusters \citep[9]{Launey2011}, and \citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.

\subsubsection{Subsyllabic parsing}

It is also necessary to separate the interlude from the nuclei on either side. Distributional heuristics are used to develop a subsyllabic parsing procedure for English below.

\paragraph{The velar nasal} There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right \citep[e.g.,][]{Sapir1925}, an analysis which is a consequence of principles like the Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[53]{OT}, or simply the allophone of /n/ when followed by underlying /k, ɡ/ (e.g., \emph{SPE}:85, \citealt{Borowsky1986}:65f.), sometimes followed by the deletion of /ɡ/. The strongest piece of evidence for the latter position, in which the velar nasal is a pure allophone, is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. Following \citet{Pierrehumbert1994}, the pure allophonic analysis is adopted here, and English nasal place allophony is formalized in Section \ref{npasection} below.

\paragraph{[j] onglides} While some have assumed that the front onglide in words such as \emph{val}[j]\emph{ue} is inserted by rule (\emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), the very presence or absence of the glide is in fact contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}) and thus assumed to be present in underlying representation (\citealt{Anderson1988b}, \citealt[278]{Borowsky1986}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}). There is considerable evidence for this assumption. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː]. This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of an underlyingly complex vowel segment (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters, though they may be followed by [ju] in words like [mj]\emph{use} or [vj]\emph{iew}, also suggesting the nuclear affiliation of the glide. There is also strong external support for this principle. The [ju] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005}, \emph{shm}-reduplication \citep{Nevins2003}, and speech errors (e.g., [kjuː]\emph{mor} [h]\emph{omponent} for intended [hju]\emph{mor component}; \citealt[130]{Shattuck-Hufnagel1986}).

\paragraph{[w] onglides} The phonotactic properties of the back onglide [w] are the opposite of the front onglide. Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[161]{Davis1995}. Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[166]{Davis1995}. Taken together, these facts indicate that the back onglide is assigned to the onset.

\paragraph{Post-vocalic \emph{r}} Both CELEX and the dictionary used by \citet{Pierrehumbert1994} transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost. Even in \emph{r}-full dialects, though, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters. Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}); for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}. Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which it patterns differently than internal coda consonants. \citet[258]{Harris1994} reports that post-vocalic \emph{r} is the only consonant which does not block a variable process of /t/-\textsc{Glottalization} found in many British dialects.

\begin{example}[/t/-\textsc{Glottalization} in British English]
\begin{tabular}{l l l l@{} l}
a. & \emph{des}[ɚt]         & \alt{} &   & \emph{des}[ɚʔ]         \\
   & \emph{c}[ɚt]\emph{ain} & \alt{} &   & \emph{c}[ɚʔ]\emph{ain} \\
b. & \emph{fi}[st]          & \alt{} & * & \emph{fi}[sʔ]          \\
   & \emph{mi}[st]\emph{er} & \alt{} & * & \emph{mi}[sʔ]\emph{er} \\
\end{tabular}
\end{example}

\noindent Whereas post-vocalic \emph{r} is the only consonant which fails to block /t/-\textsc{Glottalization} in British dialects, it is the only consonant which fails to feed /t, d/-\textsc{Deletion} in American English \citep[8]{Guy1980}.

\begin{example}[/t, d/-\textsc{Deletion} in American English] \label{td}
\begin{tabular}{l l l l@{} l}
a. & \emph{be}[lt] & \alt{} &   & \emph{be}[l] \\
   & \emph{me}[nd] & \alt{} &   & \emph{me}[n] \\
b. & \emph{sh}[ɚt] & \alt{} & * & \emph{sh}[ɚ] \\
   & \emph{c}[ɚd]  & \alt{} & * & \emph{c}[ɚ]  \\
\end{tabular}
\end{example}

\subsubsection{Lexical statistics}

The English portion of CELEX contains 6,876 simplex words with 21 unique medial codas and 40 unique medial onsets. Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, an 18.8\% saturation rate. The data set is reproduced in Appendix \ref{appendixB}.

\subsection{Static and derived constraints}

\emph{SPE} describes three phonological alternations which target syllable contact clusters. The corresponding constraints on simplex syllable contact clusters are considered below. The first such derived constraint, obstruent voice assimilation, is presented as a tutorial introduction. The second set of constraints consists of three static phonotactic constraints proposed by \citet{Pierrehumbert1994}.

\subsubsection{Obstruent voice assimilation} \label{ova}

Voice assimilation in English is exemplified by the allomorphs of /-d/ suffix, forming regular pasts and denominal adjectives, and the /-z/ suffix, forming regular noun plurals, genitives, and the 3sg. present active indicative.\footnote{The URs /-d, -z/ are argued for by \citet[282]{Hockett1958}, \citet[210]{SPE}, \citet{Basboll1972}, \citet{Shibatani1972}, \citet{Anderson1973a}, \citet[102]{Pinker1988}, and \citet[284f.]{Bakovic2005b}; alternatives are suggested by \citet[210f.]{LANGUAGE}, \citet[426]{Nida1948}, \citet{Luelsdorff1969}, \citet{Lightner1970}, \citet{Hoard1971}, \citet{Miner1975}, \citet{Zwicky1975}, \citet{Kiparsky1985}, and \citet[135]{Borowsky1986}.}

\begin{example}[Inflectional affix voice assimilation]
\begin{tabular}{llllll}
a. & \emph{nap} & \emph{nap}[t] \\
   & \emph{nab} & \emph{nab}[d] \\
b. & \emph{lap} & \emph{lap}[s] \\
   & \emph{lab} & \emph{lab}[z] \\
\end{tabular}
\end{example}

\noindent The devoicing of the two voiced obstruent suffixes can be formalized as a rule spreading the voicing specification of an obstruent rightward.

\begin{example}[\textsc{Obstruent Voice Assimilation} (after \emph{SPE}:178)]
\xymatrix@R=24pt@C=24pt{
\txt{[α \textsc{Voice}]}\ar@{-}[d]\ar@{--}[dr] &                               \\
\txt{C}\ar@{-}[d]                              & \txt{C}\ar@{-}[d]             \\
\txt{[$+$\textsc{Obstruent}]}                  & \txt{[$+$\textsc{Obstruent}]} \\
}
\end{example}

It is quite apparent that root-internal hetero-voiced obstruent clusters like [s.b] or [z.t] are rare compared to clusters which are uniformly voiced or uniformly voiceless, as predicted the assimilation rule. Yet, \citet{Pierrehumbert1994} does not mention voice assimilation in her study of English syllable contact clusters. \citet[]
[74f.]{Hammond1999a} cites words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain hetero-voiced obstruent clusters, as evidence that the proce
ss does not apply to simplex words. Indeed, the \textsc{Revised Alternation Condition} (RAC) proposed by \citeauthor{Kiparsky1973a} (\citeyear{Kiparsky1973a}:163, \citeyear{Kiparsky1982a}:152) blocks the application of obligatory neutralization processes like \textsc{Obstruent Voice Assimilation} in root-internal (i.e., non-derived) environments.\footnote{The ``obligatory'' condition on the RAC has the unforunate result that \textsc{Obstruent Voice Assimilation} is always consistent with the RAC whether or not it actually applies in non-derived environments. If the rule applies in non-derived environments, then it has lexical exceptions (e.g., \emph{a}[s.b]\emph{estos}) and is not obligatory and thus free from the RAC. On the other hand, if the rule is subject to the RAC, it only applies in derived environments, in which it is obligatory.}

The question can be stated in the form of a statistical test: does \textsc{Obstruent Voice Assimilation} (posited to apply in non-derived environments) account for a meaningful generalization about the lexicon (henceforth, the alternative hypothesis) or is it indistinguishable from noise (the null hpypothesis)?

This question is answered by considering the subset of the 840 ``possible'' clusters in which the final coda consonant and the first onset consonant are both obstruents: there are 720 in all. These are then sorted into 4 bins, according to whether or not they are attested or not, and whether or not they conform to the generalization---that is, whether both are voiced or voiceless, like in \emph{hu}[z.b]\emph{and} or \emph{ma}[s.t]\emph{iff}, respectively---or violate it, like the aforementioned \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}. The resulting table is shown in (\ref{ova}) below. 18\% of the possible clusters containing adjacent obstruents agreeing in voicing are found in the corpus, whereas only 2\% of those which disagree in voicing are attested. The table is then submitted to the Fisher Exact Test. The resulting very small $p$-value indicates that the rarity of the disagreeing clusters is unlikely to be due to chance. The simplest interpretation of this fact is that \textsc{Obstruent Voice Assimilation} applies in non-derived environments, preventing learners from positing URs in which adjacent obstruents do not agree in voice.

\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}] \label{ova}
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value                   \\
\midrule
conforming & 80       & 370        & 0.178      & \multirow{2}{*}{1.1\e{-11}} \\
violating  &  6       & 264        & 0.022                                    \\
\bottomrule
\end{tabular}
\end{example}

There are two possibilities for the small number of roots that violate the derived generalization. Some hetero-voiced obstruent clusters, like the one in \emph{ja}[k\#d]\emph{aw}, for instance, might be analyzed as compounds by native speakers (cf. \S\ref{slave} above). Less informatively, exceptionality can be derived from the \emph{SPE} rule exception convention, by marking exceptional roots with the feature [$-$\textsc{Obstruent Voice Assimilation}].

\subsubsection{Nasal place assimilation} \label{npasection}

In addition to the [ŋ] allophone, \textsc{Nasal Place Assimilation} is responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix. In (\ref{nparule}a), the prefix-final consonant takes on the same major place of articulation as the following obstruent. The vowel-initial bases in (\ref{nparule}b) demonstrate that the prefix is underlyingly /ɪn-/.

\begin{example}[\emph{im-}/\emph{in-} allomorphy] \label{nparule}
\begin{tabular}{l l l l}
a. & \emph{balance}   & \emph{i}[m.b]\emph{alance}  \\
   & \emph{decent}    & \emph{i}[n.d]\emph{ecent}   \\
b. & \emph{elegant}   & \emph{i}[n]\emph{elegant}   \\
   & \emph{ability}   & \emph{i}[n]\emph{ability}   \\
c. & \emph{mature}    & \emph{i}[m]\emph{ature}     \\
   & \emph{numerable} & \emph{i}[n]\emph{umerable}  \\
\end{tabular}
\end{example}

\citet[][175]{Pierrehumbert1994} observes that ``nasal-stop sequences agree in labiality'' in simplex English words; \citeauthor{Pierrehumbert1994} presumably makes no statement about other place features as she also assumes the velar nasal is derived from underlying /n/. While \citeauthor{Pierrehumbert1994} makes no connection between this sequence structure constraint and the any phonological alternation, the nasal place assimilation that derives [ŋ] is also responsible for the allomorphs of the \emph{im-}/\emph{in-} prefix.

\begin{example}[\textsc{Nasal Place Assimilation} (after \emph{SPE}:85)]
\xymatrix@R=24pt@C=24pt{
                          & \txt{\textsc{Place}}\ar@{-}[d]\ar@{--}[dl] \\
\txt{C}\ar@{-}[d]         & \txt{C}\ar@{-}[d]                          \\
\txt{[$+$\textsc{Nasal}]} & \txt{[$+$\textsc{Obstruent}]}              \\
}
\end{example}

The shape of the prefix before the velar stops /k, g/ is somewhat more complex. \citet[][62]{Halle1985a} and \citet[][90]{Borowsky1986} report that coda nasals assimilate to [ŋ] before dorsal consonants, but that assimilation of \emph{im-}/\emph{in-} is blocked by stress on the following syllable, citing contrasts like the one between \emph{í}[ŋ.k]\emph{ubate} and \emph{i}[n.k]\emph{lúde}. However, the stress condition does not hold in simplex words (e.g., \emph{a}[ŋ.ɡ]\emph{óra}), and thus can be safely ignored here.

There is some evidence that even young infants acquiring English are aware of this generalization. \citet{Davidson2004}, \citet{Mattys1999}, and \citet{Jusczyk2002} find that 4.5-, 9-, and 10-month old infants (respectively) prefer to listen to nonce words which contain the homorganic nasal-obstruent clusters (e.g., \emph{u}[m.b]\emph{o}) over those which contain heterorganic clusters (e.g., \emph{u}[n.b]\emph{o}). Furthermore, \citet[228]{Myers1993} reports on speech errors which create nasal-obstruent clusters which then undergo \textsc{Nasal Place Assimilation}.

\begin{example}[Speech errors feeding \textsc{Nasal Place Assimilation} (\citealp{Myers1993}:228)]
\begin{tabular}{l l l}
a. & \emph{pri}[mp]\emph{s}      & (\emph{prints})      \\
b. & \emph{ra}[nd] \emph{orker}  & (\emph{rank order})  \\
c. & \emph{po}[ŋk]\emph{utation} & (\emph{computation}) \\
\end{tabular}
\end{example}

\citet[175]{Pierrehumbert1994} observes that ``nasal-stop sequences agree in labiality'', but draws no connection between this and \textsc{Nasal Place Assimilation}. Exceptions like \emph{pli}[m.s]\emph{oll}, \emph{da}[m.z]\emph{el}, \emph{scri}[m.ʃ]\emph{aw}, and \emph{ra}[m.k]\emph{in} do occur, they are much rarer than homorganic nasal-obstruent clusters such as \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, and \emph{mo}[ŋ.k]\emph{ey}, a preference which is unlikely to be due to chance.\footnote{There are also a small class of simplex words which contain clusters like \emph{oi}[nt.m]\emph{ent}, in which a nasal-obstruent cluster is found within a complex coda, but place assimilation is exceptionless in this context, and thus these clusters are ignored here.}

\begin{example}[Lexical effects of \textsc{Nasal Place Assimilation}]
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value                   \\
\midrule
conforming & 41       & 11         & 0.788      & \multirow{2}{*}{2.6\e{-05}} \\
violating  & 8        & 20         & 0.286                                    \\
\bottomrule
\end{tabular}
\end{example}

\subsubsection{Degemination}

A final alternation targeting English syllable contact clusters consists of the simplification of geminates derived by certain affixes. For instance, the deadjectival adverbial suffix \emph{-ly} rarely produces a geminate when it attaches to /l/-final roots (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}). Degemination is also found in /d/-final roots selecting the irregular /-t/ past tense suffix.

\begin{example}[Degemination in /-t/ pasts (\citealp{Myers1987}:492)]
\begin{tabular}{l l l l l}
a. & \emph{burn}  & \emph{burn}[t] \\
   & \emph{spill} & \emph{spil}[t] \\
b. & \emph{bend}  & \emph{ben}[t]  \\
   & \emph{build} & \emph{buil}[t] \\
\end{tabular}
\end{example}

More complicated cases of \textsc{Degemination} are can be found in prefix allomorphy (\citealt[102]{Borowsky1986}, \emph{SPE}:148); one example is (\ref{nparule}c) above, in which \textsc{Nasal Place Assimilation} feeds \textsc{Degemination}. To formalize this process of degemination, it is assumed that geminates are adjacent similar segments. The alternative, that geminates are single feature matrices linked to multiple timing slots, places this constraint in the inventory and makes it difficult to link degemination and the absence of geminates in underlying representations. Geminates are identified using quantification over features \citep{Reiss2003b}.

\begin{example}[\textsc{Degemination} (\emph{SPE}:253)]
\xymatrix@R=24pt@C=24pt{
\txt{C}_1       & \rightarrow & \emptyset & / & \gap\gap & \txt{C}_2 \\
}

\noindent Condition: $\forall F_i \in \{\textsc{Place}, \textsc{Sonorant}, \textsc{Nasal}, \textsc{Continuant}, \textsc{Lateral}\}$ . $(F_i)_1 = (F_i)_2$
\end{example}

Identical adjacent consonants do not occur in the CELEX data, nor do those which differ only in voice, such as *[p.b], which would be made identical by the operation of \textsc{Obstruent Voice Assimilation} and this absence is unlikely to be due to chance.\footnote{It is not the case that the English lexicon envisioned in \emph{SPE} is free of underlying geminates; \citeauthor{SPE} use underlying (but never surfacing) geminates to produce exceptions to the stress rules (see also \citealt{Burzio1994}:52f., \citealt{Halle1998c}:551, fn. 9). In addition to the usual objections to absolute neutralization (i.e., that it is a learning conundrum mascarading as a phonological solution), the phonological particulars of this account are critiqued by \citet{Ross1972a}.}

\begin{example}[Lexical effects of \textsc{Degemination}] 
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value                   \\
\midrule
conforming & 157      & 596        & 0.208      & \multirow{2}{*}{7.2\e{-09}} \\
violating  & 0        &  87        & 0.000                                    \\
\bottomrule
\end{tabular}
\end{example}

\subsubsection{Local summary}

All three \emph{SPE} rules affecting syllable consonant clusters are reliable predictors of attested and unattested medial clusters.

\subsubsection{Static constraints}

The second set of constraints consists of three phonotactic constraints proposed by \citet{Pierrehumbert1994}. These constraints are ``static'' in that they have no analogue in English morphophonemics. \citeauthor{Pierrehumbert1994} focuses only on triconsonantal clusters. However, this is not a privileged subset, since syllable contact clusters can be as short a single coda and onset consonant or as long as four consonants (e.g., \emph{mi}[n.str]\emph{el}), and therefore clusters of all lengths are considered here.

\subsubsection{Dorsal/non-coronal clusters}

\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars.'' Two-consonant clusters like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} are found, however. Despite this, dorsal/non-coronal clusters are no less likely to occur than non-dorsal codas with non-coronal onsets (e.g. \emph{oi}[nt.m]\emph{ent}) or dorsal codas with coronal onsets (e.g., \emph{ve}[k.t]\emph{or}).

\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}, triconsonantal clusters only]
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value              \\
\midrule
conforming & 20       & 148        & 0.119      & \multirow{2}{*}{0.136} \\
violating  &  0       &  22        & 0.000                               \\
\bottomrule
\end{tabular}
\end{example}

\begin{example}[Lexical effects of \textsc{*[$+$Dorsal][$-$Coronal]}, all clusters]
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 76 & 359 & 0.175 & \multirow{2}{*}{0.145} \\
violating  &  6 &  57 & 0.095 \\
\bottomrule
\end{tabular}
\end{example}

\subsubsection{Coda coronal obstruents}

\citet[][175]{Pierrehumbert1994} proposes that ``clusters with a coronal obstruent in the coda do not occur'', though \citet{Pierrehumbert1994} notes an exception in \emph{a}[nt.l]\emph{er}, and one can add the triconsonantal clusters in \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}, as well as many biconsonantal clusters like \emph{a}[t.l\emph{as} or \emph{e}[s.t]\emph{er}. Triconsonantal coda coronal obstruent clusters are not significantly less likely to occur than ``conforming'' clusters with non-coronal or non-obstruents. The same is true if the arbitrary restriction to triconsonantal clusters is lifted.

\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}, triconsonantal clusters only]
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value              \\
\midrule
conforming & 20       & 148        & 0.119      & \multirow{2}{*}{0.136} \\
violating  & 0        & 22         & 0.000                               \\
\bottomrule
\end{tabular}
\end{example}

\begin{example}[Lexical effects of \textsc{*Coda Coronal Obstruent}, all clusters]
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value              \\
\midrule
conforming & 48       & 312        & 0.133      & \multirow{2}{*}{0.301} \\
violating  & 38       & 322        & 0.106                               \\
\bottomrule
\end{tabular}
\end{example}

\subsubsection{ABA clusters}

Finally, \citet[][176]{Pierrehumbert1994} observes ``a lack of clusters with identical first and third elements''. Here, this is operationalized in the same fashion as \textsc{Degemination}, by definining ``identity'' as consonants which share the same place and manner features, but which do not necessarily agree on voicing. \citeauthor{Pierrehumbert1994} notes one potential exception to this generalization, the cluster in words such as \emph{e}[ks.kl]\emph{ude}, which contains two /k/, does not arise here since such words are marked as morphologically complex in CELEX. Despite the fact that the corpus contains no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters.

\begin{example}[Lexical effects of \textsc{*ABA}]
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 41       & 478        & 0.079      & \multirow{2}{*}{0.818} \\
violating  &  0       &  11        & 0.000                               \\
\bottomrule
\end{tabular}
\end{example}

\subsubsection{Local summary: static constraints}

The inventory of attested clusters shows no statistically reliable traces of the three static constraints proposed by \citet{Pierrehumbert1994}.

\subsection{Computational models}

It is possible, however, that an explicit model of ``constraint discovery'' could identify reliable constraints on underlying representations, whether static or derived by alternation. The goal of \citet{Hayes2008a} is to develop a computational model which discovers and weighs constraints using observed data. A very simple model of this type is also proposed by \citeauthor{Pierrehumbert1994}, in which the well-formednesss of a medial cluster is proportional to its expected probability, computed from the joint probability of coda and onset under the assumption that these are independent (see \S?). These two models as well as two baselines are scored for their ability to predict which of the possible clusters are attested and which are not. There are four possible outcomes for each of the 840 possible clusters.

\subsubsection{Performance metrics}

\begin{example}[Outcomes in cluster classification task]
%\vspace{0.5\baselineskip}
%\centerline{\noindent\ignorespaces
\begin{tabular}{c l | l l}
                   & & \multicolumn{2}{c}{\textbf{actual outcome}}            \\
                   & & \emph{attested}   & \emph{unattested}            \\
\midrule
\textbf{predicted} & \emph{attested}   & true positive ($tp$)  & false positive ($fp$) \\
\textbf{outcome}   & \emph{unattested} & false negative ($fn$) & true negative ($tn$)  \\
\end{tabular}
%}%\vspace{0.5\baselineskip}
\end{example}

\noindent Accuracy is the probability of a correct classification, whether a true positive or negative.

\begin{unlabeledexample}
$\displaystyle \textrm{Accuracy} = \frac{tp + tn}{tp + tn + fp + fn}$
\end{unlabeledexample}

\noindent Precision is the probability that a cluster predicted to occur is attested.

\begin{unlabeledexample}
$\displaystyle \textrm{Precision} = \frac{tp}{tp + fp}$
\end{unlabeledexample}

\noindent Recall (or sensitivity) is the probability a observed cluster is predicted to occur.

\begin{unlabeledexample}
$\displaystyle \textrm{Recall} = \frac{tp}{tp + fn}$
\end{unlabeledexample}

\noindent It is possible to maximize recall at the expense of precision, by predicting attestation for a greater number of clusters, or to increase precision at the expense of recall by predicting non-attestation for a greater number of clusters. $F_1$, the harmonic mean of these two measures, quantifies this trade-off.

\begin{unlabeledexample}
$\displaystyle F_1 = 2 \left( \frac{\textrm{Precision} \times \textrm{Recall}}{\textrm{Precision} + \textrm{Recall}}\right)$
\end{unlabeledexample}

A soft-margin support vector machine \citep{Cortes1995} with a linear kernel is used to convert the model scores of individual clusters into predictions. This technique finds a single split which separates clusters into those which it predicts to be attested, and unattested, in an optimal fashion. This is presented as part of a cognitively plausible model for learning phonotactics; it simply represents an upper bound on the performance of these models. The results for all four models are summarized in Table \ref{cmresults}.

\begin{table} \centering
\begin{tabular}{l | r r r r}
\toprule
                          & accuracy & precision & recall & $F_1$ \\ 
\midrule
Null baseline             & 0.813    & n.a.      & n.a.   & n.a.  \\
Derived constraints       & 0.857    & 0.401     & 0.708  & 0.512 \\
Expected frequency        & 0.839    & 0.210     & 0.750  & 0.328 \\
\citet{Hayes2008a}        & 0.877    & 0.777     & 0.642  & 0.703 \\
\bottomrule
\end{tabular}
\caption{The \citeauthor{Hayes2008a} model makes the most accurate predictions about possible and impossible syllable contact clusters, though it is only slightly better than the derived constraint baseline.}
\label{cmresults}
\end{table}

\subsubsection{Null baseline}

18.8\% of the 840 possible clusters are attested. As a consequence of this sparsity, a null baseline which classifies all clusters as unattested will achieve 81.2\% accuracy. Precision, recall, and $F_1$ are undefined for this baseline.

\subsubsection{Derived constraints}

The three constraints derived by \emph{SPE} rules have already been seen to be reliable predictors of cluster attestation. A slightly more complex baseline can be produced by predicting any cluster which does not violate these generalizations to be attested. This results in a significant improvement to the null baseline (accuracy = 0.857). Given the sparse nature of the syllable contact inventory, it is no surprise t
hat there are more false positives (i.e., unattested clusters which do not run afoul
 of any of these constraints) than false negatives (i.e., exceptions to the three ru
les), and this is reflected in the fact that recall (0.708) is somewhat higher than
precision (0.401).

\subsubsection{Expected frequency}

\citet{Pierrehumbert1994} proposes a simple model in which the wellformedness of syllable contact clusters is proportional to the joint probability, estimated from word-final coda and word-initial onset frequencies and assuming that coda and onset are statistically independent. \citeauthor{Coleman1997} summarize the results as follows.

\begin{quote}
\citet{Pierrehumbert1994} showed that the independent probabilities of the coda and the following onset was the single biggest factor in predicting which complex clusters exist. \citep[50]{Coleman1997}
\end{quote}

\noindent A pilot study revealed, however, that using only medial coda and onset frequencies provided a much better fit to the data, and the results reported above use medial frequencies. While \citeauthor{Pierrehumbert1994} reports that this is the best predictor of which complex clusters occur, it provides only a small improvement over the null baseline (accuracy = 0.839). It imposes no constraints on sequences spanning the syllable boundary, and in fact has both lower accuracy and $F_1$ (0.328) than the derived constraint baseline.

\subsubsection{\citet{Hayes2008a}}

\citet{Hayes2008a} provide a software implementation of their phonotactic learning system, which uses the principle of maximum entropy to weigh competing phonotactic constraints. Since this software exposes many parameters of the model to the experimenter, a direct replication of the experiments reported by \citeauthor{Hayes2008a} is attempted; their software, feature specification, and model settings were all used. Since the training of this model is inherently stochastic, producing slightly different outcomes on each run, the results reported here are from the lowest accuracy run of ten, a practice used by \citeauthor{Hayes2008a}. The model achieves greater accuracy (0.877) and $F_1$ (0.703), but lower recall (0.642), indicating the presence of additional false negatives. The generalizations extracted by the \citeauthor{Hayes2008a} model are in general quite similar to those derived from the \emph{SPE} alternations (as indicated by similar performance), but the lower recall indicates that the model formulates somewhat narrower generalizations. A few interesting divergences fro the alternation baseline are apparent. Unattested *[m.kl] is assigned the highest probability, though it is ruled out by \textsc{Coda Nasal Place Assimilation}. On the other hand, very low scores are assigned to the clusters in \emph{hu}[z.b]\emph{and} and \emph{pla}[t.f]\emph{orm}, though they are both attested and consistent with all three alternation-derived constraints.\footnote{\citet{Berent2012} discuss a potential defect of the H\&W model: it does not use of variables over features and thus fails to unify constraints against geminate clusters. This does not appear to be relevant to this evaluation, as the H\&W model correctly predicts the total absence of clusters with geminates.}

\subsubsection{\citet{McGowan2009}}

\citet{McGowan2009} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2009} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority. However, \citeauthor{McGowan2009} finds that sonority of clusters accounts for a very small of the variance in cluster frequency ($R^2 = 0.077$). As a predictor of cluster attestedness, sonority distance provided no improvement over the null baseline, and therefore is not reported in Table \ref{cmresults}.

\section{Conclusions}

Though the H\&W model has the overall best performance, it is not a large improvement on than the derived constraint baseline. This suggests that most, if not all, of the 71.2\% of possible clusters which are unattested are not ill-formed, but are simply missing due to the sparsity of the data: that is, they are accidental gaps. Below, the role of sparsity in the syllable contact inventory is quantified in a number of ways.

\subsection{The Zipfian distribution}

A number of previous studies \citep[e.g.,][]{Weiss1961,Sigurd1968,Good1969,Borodovsky1989,Witten1990,Martindale1996,Tambovtsev2007} observe that phoneme or letter frequencies exhibit sparse distributions, in which a few types account for a large amount of the probability mass, and the remaining is divided up among a large number of rare types. The same is true for the type frequencies of the medial codas ond onsets which make up syllable contact clusters, as well as cluster frequencies themselves. In Figure \ref{clus}, cluster frequencies are plotted against rank in the log-log space, and a near-linear relationship obtains ($R^2 = 0.924$), indicating that the data conform to a generalization of Zipf's (\citeyear{Zipf1949}) Law. See Appendix \ref{zr} for estimation details.

\begin{figure} \centering
\includegraphics{cluster.pdf}
\caption{Syllable contact clusters exhibit a log-log linear relationship between frequency and rank that is consistent with Zipf's Law.}
\label{clus}
\end{figure}

Sparse distributions characterize the frequencies of many linguistic objects, from syntactic rules \citep{Yang2009} to word \citep{Teahan1998,Baroni2009} and phoneme \citep{Belevitch1956,Daland2011a} $n$-grams, but are also found in non-linguistic symbol systems \citep{Mandelbrot1954,Miller1957,Chomsky1958,Sproat2010} and randomly generated texts \citep{Miller1957,Li1992}. The importance of sparsity in this context is that it entails a long right tail, making it difficult to determine on statistical grounds alone which unobserved events are impossible and which are accidental gaps.

\subsection{The Good-Turing estimate}

One way to quantify the possibility of accidental gaps is proposed by \citet{Good1953}, who proposes an estimate of $\hat{p}_0$, the probability of all unseen events. This probability is simply $n_1$, the number of events which occur just once, divided by the number of observations.

\begin{unlabeledexample}
$\displaystyle \hat{p}_0 = \frac{n_1}{\displaystyle\sum N}$
\end{unlabeledexample}

\noindent In CELEX, $65$ clusters occur only once, and there are 873 cluster tokens, so $\hat{p}_0 = 0.074$. This indicates that were it possible to generate a new corpus of syllable contact clusters with the same statistical properties as the English data, approximately 7\% of the tokens would be clusters not found in the previous sample.

\subsection{Sampling simulation}

One possible way to generate a new corpus is by simulation, in which new clusters are generated using the following procedure, which combines the derived constraint baseline with the model proposed by \citet{Pierrehumbert1994}.

\begin{example}[Simulation procedure]
\begin{tabular}{l l}
a. & Sample a medial coda according to the observed probabilities  \\
b. & Sample a medial onset according to the observed probabilities \\
c. & Apply all matching \emph{SPE} rules to the resulting clusters \\
\end{tabular}
\end{example}

While this procedure can be repeated indefinitely, it is worthwhile to consider the results of a single characteristic ``lexicon'' produced in this fashion. The slope of the log-rank/log-frequency relationship for the simulated data ($\alpha = -0.524$) is also close to the observed slope for the CELEX data ($\alpha = -0.588$). And, as shown in Figure \ref{sim}, the two distributions are nearly indistinguishable.

\begin{figure}
\centering
\includegraphics{sim.pdf}
\caption{The observed cluster frequencies are closely matched by a simulated cluster inventory generated by random sampling.}
\label{sim}
\end{figure}

This result should not be construed as an endorsement of the \citet{Pierrehumbert1994} model. Rather, it shows that even if the only constraints on syllable contact clusters are that they must consist of valid medial codas and onsets which further conform to the three derived constraints, there still will be many gaps in the cluster inventory, gaps which are quite arbitrary from the perspective of the phonology and phonological representations in general.

%as in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos} CELEX also contains \emph{jo}[d.p]\emph{urs}, \emph{pi}[ntʃ.b]\emph{eck}, \emph{sa}[k.b]\emph{ut}, and \emph{ja}[k.d]\emph{aw}. 
%There is no duplication between constraints on sequence structure and surface forms if both are derived from the application of the phonological rule (\citealt[][401f.]{Stanley1967}, \citealt[][382]{SPE}). 
%\citet{Wright1975} taught several nonce words to a group of adolescents, and asked them to repeat the words to each other in a game of ``Telephone''. In these nonse words, nasals do not agree in \textsc{Place} with the following obstruent in these nonsense words, and \citeauthor{Wright1975} observes that after a few rounds of the game, the nonce words had been adapted to conform to \textsc{Coda Nasal Place Assimilation}.
%\begin{example}[Artificial language adaptations (\citealt{Wright1975}:394, his transcriptions)]
%\begin{tabular}{l l l l}
%a. & [ɡownp] & \textgreater{} & [ɡump]  \\
%b. & [ǰumɡ]  & \textgreater{} & [ǰúŋɡuə] \\
%c. & [ðʌŋd]  & \textgreater{} & [tɔŋɡ]  \\
%\end{tabular}
%\end{example}
%\noindent If this game is akin to the process of loanword adaptation, it is possible that various extragrammatical pressures are at play \citep[e.g.,][]{Halle1998b,Dupoux1999,Ussishkin2003,Peperkamp2008}. Yet, the independent evidence for a phonological process of nasal place assimilation provides the most parsiminious account of the adaptations seen in \citeauthor{Wright1975}'s study. 
%\citet{Hay2004a} demonstrates that [n.p, mθ] clusters, which are not found in English, are in fact rated better than many attested clusters.
%The same is true of the prefix \emph{com-}/\emph{con-}, as shown by the presence of assimilation in \emph{có}[ŋ.ɡ]\emph{ress} but not in \emph{co}[n.g]\emph{réssional}. 
%(e.g., \citealt{Bakovic2005b}, \citealt{Fruehwald2011}),

%\ex \textsc{Coda Nasal Place Assimilation} feeds \textsc{Degemination} \citep[][116]{Borowsky1986}: \\
%\begin{tabular}{l l l l}
%a. & mature    & i[m]ature    & (cf. \emph{i}[m.b]\emph{alance})   \\
%b. & numerable & i[n]umerable & (cf. \emph{i}[n.d]\emph{ependent}) \\
%\end{tabular}
%\xe
%(\citealp{Halle1985a}:105, 
%; presumably, this generalization is limited to this one place feature because  \citeauthor{Pierrehumbert1994} also adopts the assumption that the velar nasal is derived from /n/. 
%Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets. 
%
%\begin{example}[VBR's complex onsets (\citealp{Buchwald2005}:79--80, his transcriptions)] \label{VBR}
%\begin{tabular}{l l l}
%a. & kəræb  & `crab'  \\
%   & bəlid  & `bleed' \\
%b. & kəwin  & `queen' \\
%   & kəwoʊt & `quote' \\
%c. & kut    & `cute'  \\
%   & musɪk  & `music  \\ 
%\end{tabular}
%\end{example}
%
%\noindent VBR breaks up complex onsets with epenthesis, including those that consist of a consonant and a back onglide cluster (\ref{VBR}b). However, no epenthesis occurs between a consonant and a front onglide; rather, the glide is absent (\ref{VBR}c). The failure of the front onglide to pattern with other consonant clusters suggests once again that the glide is part of the nucleus. 
%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [juː] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant.This is the only sign that [juː] shows an affiliation for the onset. 
%In fact, [t.ʃ, d.ʒ] clusters are not found in simplex English words, despite the fact that that affricates occur as medial onsets in clusters (e.g., \emph{tru}[n.tʃ]\emph{eon}, \emph{so}[l.dʒ]\emph{er}).
%Finally, \citet[][251]{Fromkin1973} also presents evidence that post-vocalic \emph{r} may behave as if nuclear in speech errors. 
%\citep[e.g.,][142]{Mester1988}.
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or a part of a contrast in any known language \citep{Rubach2002}.}
% Note that [tʃ] has not been included since it so rarely occurs in clusters.
%\subsubsection{Summary}
%
%The above results are summarized in Table \ref{constraints}.
%
%\begin{table}
%\centering
%\begin{tabular}{l r r r r r}
%\toprule
%                                       & \multicolumn{2}{c}{attested} & \multicolumn{2}{c}{unattested} & \multirow{2}{*}{$p$-value} \\
%                                       & conforming & violating & conforming & violating \\
%\midrule
%\textsc{O.V.A.} (\ref{ovar})        &  80 &   6 & 370 & 264 & 1.1\e{-11} \\
%\textsc{C.N.P.A} (\ref{CNPA})       &  41 &  25 &  12 &  42 & 1.7\e{-05} \\ 
%\textsc{Degemination} (\ref{degem}) & 161 & 632 &   0 &  87 & 1.4\e{-08} \\
%\midrule
%\textsc{*Dorsal-Labial} ()          &  68 &   4 & 246 &  40 & 0.069 \\
%\textsc{*C.C.O.} ()                 &  48 &  38 & 312 & 322 & 0.301 \\
%\textsc{*A,BA} ()                   & 161 &   0 & 708 &  11 & 0.231 \\
%\bottomrule
%\end{tabular}
%\caption{Counts and Fisher exact test $p$-values for the constraints discussed above.}
%\label{constraints}
%\end{table}
%Whereas 18.8.\% of the ``possible'' clusters are attested, 28.8\% of those which conform to these three constraints are found. 
%Binary classification contingency table: \\
%\begin{tabular}{c | l l} \toprule
%                             & \multicolumn{2}{c}{prediction} \\
%\midrule
%\multirow{2}{*}{observation} & true positive ($tp$)  & false positive ($fp$) \\
%                             & false negative ($fn$) & true negative ($tn$) \\
%\end{tabular}
%\begin{tabular}{|l l|}
%\toprule
%true positive ($tp$)  & false positive ($fp$) \\
%false negative ($fn$) & true negative ($tn$) \\
%\bottomrule
%\end{tabular}
%\begin{figure}
%\centering
%\includegraphics{coda.pdf} 
%\includegraphics{onset.pdf}
%\caption{Medial coda and medial onset lexical frequency exhibit the Zipfian log-log-linear relationship between rank and frequency.}
%\label{codaonset}
%\end{figure}
%Sparsity alone is not the only reason for cluster underrepresentation which is external to the phonology. For instance, a cluster could be underrepresented due to language change. \citet{Iverson2005} discuss a case of this type: In English, long vowels before word-final [ʃ] in English are generally rare (though note ``affective'' \emph{swoosh}, \emph{sheesh}, etc.). This is ultimately due to the fact that native [ʃ] is descended from Old English [sk], and long vowels are not fonud not occur before complex codas. \citeauthor{Iverson2005} conclude that this gap is accidental because it does not inhibit borrowing and coinage. 
%On the other hand, patterns created by sound change are not guaranteed to persist over time. One example of non-persistence is discussed by \citet{Iverson2005}. Around 1100 CE, Old English \emph{sk} became [ʃ]. This sound change introduced no alternations. Since long vowels were not found before tautosyllabic syllable clusters at this time, there were no \emph{Vːsk\#} words when the change was actuated, and \emph{Vːsh\#} continues to be rare in Modern English. What \citeauthor{Iverson2005} observe, however, is that there is nothing apparently peripheral about words like \emph{leash} or \emph{whoosh}, and loanwords and coinages have readily filled the gap.
%\citet[][140]{Frisch2004} suggest that the strong tendency for the first and second consonants of the Arabic root to be non-identical is the ``a diachronic result of a processing constraint that disfavors repetition.'' Unfortunately, there is no evidence that this pattern is diachronic other than in the sense that it appears to be inherited from the proto-language: there is simply no Proto-Semitic verb roots with identical first and second consonants \citep[][178]{Greenberg1950}. In other Semitic languages, the inherited patern has experienced considerable erosion. 
%\begin{example}[Tigrinya roots with identical first and second consonants \citep{Buckley1990a}]
%\begin{tabular}{l l l l}
%a. & lʌlʌw     & `scorch'                   & (\textless{} Ge'ez \emph{lʌwlʌw} `inflame')     \\
%   & mʌmʌy     & `winnow'                   & (\textless{} Ge'ez \emph{mʌymʌy} `distinguish') \\
%   & mʌmʌt     & `pick out loot'            \\
%b. & s’ʌs’ʌw   & `finish off a drink'       & (cf. \emph{s’ʌws’ʌw} `gulp down')           \\
%   & t’ʌt’ʌf   & `prune tree'               & (cf. \emph{t’ʌft’ʌf} `smear wall with mud') \\
%c. & kʷakʷkʷʌr & `waste away, be emaciated' & (cf. \emph{kʷarkʷʌr} `interrogate')         \\
%   & kakʷkʷɨʕ  & `clean wax from ears'      & (cf. \emph{kaʕkʷɨʕ} `start to form pods')   \\
%\end{tabular}
%\end{example}
%Similar exceptions are found in 
%Amharic (\citealp[][?]{Broselow1984}, \citealp[][?]{McCarthy1985}) and 
%Hebrew \citep[][29]{Bat-El2005}.
%And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.
%Before the dawn of generative phonology, many linguists concerned themselves with documenting co-occurrence in various languages. \citet[][28]{Vogt1954} declares that a ``phonemic description of a language \ldots should also comprise a description of the phoneme combinations that occur or may be expected to occur in the language''. Studies in this tradition include the discussions of consononantal co-occurrence in Arabic and Javanese by \citet{Greenberg1950} and \citet{Uhlenbeck1950}, respectively. \citeauthor{Uhlenbeck1950}, for instance, puts forth co-occurrence dispreferences, rather than exceptionless generalizations of the sort given by \citet{Greenberg1950}.
%\citet{McCarthy1988}
%Years later, \citet{Mester1988} revisits \citeauthor{Uhlenbeck1950}'s study and proposes a statistical technique, based on the chi-square test, to distinguish between accidental and structural lexical dispreferences in the Javanese lexicon. 
%\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000} on English). The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented. Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.
%\begin{example}[null hypothesis testing]
%$\displaystyle \textrm{Reject } H_0 \textrm{ iff } \frac{(O E) ^ 2}{E} > χ^2$
%\end{example}
%\footnote{This is the one-tailed version of the test.}
%In this study, I use the \citet{Fisher1934} exact test, which is isomorphic to this chi-square test. Compared to the chi-square test, the Fisher exact test $p$-value is somewhat more difficult to compute, but both are automated by modern statistical packages and can be computed rapidly. The major advantage of the Fisher test over the chi-square test is that it is appropriate for small samples or rare events, whereas the chi-square test depends on an approximation which is only exact in the limit (i.e., with a hypothetically infinite sample), and therefore is not appropriate for small samples.
%On the other end of the spectrum, the Monte Carlo statistical procedure developed by \citet{Kessler2001} and applied to co-occurrence statistics by \citet{Martin2007,Martin2011} and \citet{Brown2010} does not scale to large samples. The Monte Carlo procedure is isomorphic to the one-tailed Fisher Exact test, and involves comparing observed co-occurrence counts to those produced by randomly generated samples.  This requires the researcher to generate random permutations, something which is particularly difficult for large samples, as I will explain; disinterested readers may wish to skip the following paragraph.
%Any permutation of a sequence $L$ can be represented as a sequence $S$ of the same length, where the value of $S_i$ corresponds to the permuted position of $L_i$. Generating random permutations requires an unbiased method to generate all $N!$ of these lists with equal probability. Unfortunately, this is near impossible for $N$s of moderate size with current computational resources. Any pseudo-random number generator can be characterized by its ``period'', the number of pseduo-random numbers it can emit before repeating itself; this is also the upper limit for the number of random permutations that can be generated. The pseudo-random number generator used by both \citeauthor{Martin2011} and \citeauthor{Brown2010} has a period of approximately $2^{48}$, a number which turns out to be smaller than $17!$. For the large samples, sometimes in the thousands, used by these researchers, virtually all the possible permutations can never be generated by this method. As a consequence, the shuffling procedure is far from random and introducing the possibility that the statistical test is biased in unpredictable ways.
%As an example of the Fisher exact test as it is applied here, consider some of the Javanese co-occurrence facts considered by \citet{Mester1988}. \citeauthor{Mester1988} is at pains to show that there is a gradient dispreference for first and second root consonants which share the same major place. However, perfectly identical first and second consonants are exempt. Since 
% APPLY THIS TO JAVANESE
%\begin{example}[Javanese root co-occurrence] %(\citealp[][264]{Uhlenbeck1950}, \citealp[][139]{Mester1988})]
%\begin{tabular}{l r r r r}
%\toprule
%           & attested & unattested & saturation & $p$-value \\
%\midrule
%conforming & 134      & 12         & 0.918      & \multirow{2}{*}{6.095\e{-06}} \\
%violating  & 21       & 15         & 0.583 \\
%\bottomrule
%\end{tabular}
%\end{example}
%\noindent It is evident that roots with identical labial stops are far more common than those in which the roots disagree in voicing. 
%This data is input to a chi-square test, which reports that this highly skewed distribution is unlikely to be due to chance 
%That this is highly unlikely to be due to chance is apparent both from the chi-square ($\chi^2 = 25.572$, two-tailed $p = 4.3$\e{-07}) and Fisher exaact tests ($p =  6.095$\e{-06}). 
%\citep[e.g.,]{Brown2010} have argued for phonotactic constraints giving rise to overrepresentation, so I admit this possibility here and use two-tailed tests throughout. 
%\footnote{For example, \citet{Martin2007} applies this technique to a list of 4,758 noun-noun compounds. The number of permutations of a list of this length can be estimated using Stirling's approximation.
%\ex $ \lim_{n \rightarrow \infty} n! = e ^ {n \ln{n} - n} = 2 ^ {\log_2{e} (n \ln{n} - n)} $ \xe
%$4,758!$ is approximately $2 ^ {51260}$. The author is unaware of any system that provides 51,260 bits of entropy; and some popular programming languages, such as ANSI C, provide as few as 15 bits.}
%With this database of English syllable contact clusters it is now possible to evaluate constraints on sequences of underlying phonemes. 
%I now consider constraints on sequences of URs as predictors of attested and unattested contact clusters. 
%Two-tailed statistical tests are used throughout.
%\citet[56]{Coppock2008} observes that there is every reason to believe that the ditransitive is productive in the sense that it can be applied to new words; consider recent conversion \emph{text} (e.g., \emph{text me your number}). 
%\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}
%Despite this, six of the eight words that \citet[172]{Pierrehumbert1994} labels ``reasonably familiar'' (and thus included) are included in this study (\emph{velcro} and \emph{eclampsia} are not found in CELEX), and none of the five examples of words which are not ``reasonably familiar''.
%between irregular verbs and their base forms \citep{Allen2002,Stockall2006}
%\begin{quote}
%These regularities, although not the products of alternations, are described and defined by elements of phonological theories: segments\ldots{}and distinctive features. \citep[13]{Berkley2000}
%\end{quote}
%One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}. This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a}) but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.
%\begin{quote}
%\ldots{}the fact that some [clusters--KG] are not found must be due to accidental gaps in the inventory of signs, and cannot be explained by structural laws of the language. \citep[][16]{Fischer-Jorgensen1952}
% 
%Although my material is drawn from a fairly extensive corpus---all accessible dictionaries and vocabularies, printed texts of tens of thousands of pages as well as ordinary speech---there is every reason to believe, as experience has shown, that additional material would yield new clusters. The material will never be complete. It will always contain accidental gaps \ldots partly because some clusters by pure chance do not occur in the vocabulary. \citep[][30]{Vogt1954}
%\end{quote}
%\begin{quote}
%Of course, to obtain the desired results, we must guarantee that each sequence structure rule reflects a general systematic fact about the language, and not a fact which is due merely to the existence of accidental gaps in the lexicon. \citep[][401, fn.~8]{Stanley1967}
%
%In describing and explaining the phonotactic patterns in any language, phonologists recognize a distinction between accidental and systematic gaps in the lexicon\ldots{}A gap is taken to be systematic if it belongs to a natural class of examples according to the current theory of the investigator. Otherwise it is viewed as accidental.\citep[][180]{Frisch2004}
%\end{quote}
%\footnote{At this juncture it is necessary to clarify a persistent confusion in the literature: commitment to a negative heuristic like Stampean occultation or Richness of the Base \citep{OT} does not entail an adoption of a particular algorithm, lke Lexicon Optimization, which produces a grammar consistent with these heuristics. For instance, Lexicon Optimization is incapable of converging on a grammar of English in which [k] derives from an underlying dorsal click \citep[73]{PE}. Yet, such a grammar is entirely consistent with Richness of the Base.}

### 2012-12-09 0ther/C.tex
Small edits to chapter 1

\chapter{Problems with the random shuffle procedure} 
\label{zr}

\citet{Kessler2001,Kessler2008} proposes a novel

which is applied to co-occurrence restrictions by \citet{Brown2010} and \citet{Martin2007,2011}.

it is in theory simply a Monte Carlo approximation of a one-tailed chi-square test.

This method requires a way to randomly generate all $N!$ permutations of a list of $N$ words with equal probability. Unfortunately, for reasonably large samples, this is all but impossible with current computational resources. Random shuffling is accomplished by a \emph{pseudo-random number generator} (PRNG). PRNGs are characterized by their \emph{period}, which is the number of times they can be sampled independently; this is also the upper bound on the number of unique random shuffles they can generate. Since factorials grow so quickly, even small samples require a period much larger than is currently available. The PRNG in Perl, which is used by both \citeauthor{Martin2011} and \citeauthor{Brown2010}, has a period of approximately $2^{48}$, a number which is between $16!$ and $17!$. For samples of 17 or more words, this PRNG will generate a huge proportion of the possible permutations with zero probability, introducing bias. Large samples easily exceed the resources of even the best PRNGs. For instance, \citet{Martin2007} applies the Monte Carlo technique to a list of 4,758 noun-noun compounds. By Stirling's approximation, $4,758!$ is approximately $2 ^ {51260}$. This is many times larger than the period of the best PRNGS, including the $2^{19937} - 1$ period of the Mersenne Twister \citep{Matsumoto1998}.

### 2012-12-09 1.tex
Small edits to chapter 1

@@
 There is only one concise explanation for the [dʒ]-[tʃ] and [k]-[ɡ] alternations, namely that Russian speakers have internalized a process of voice assimilation.
-Now, it is also the case that voicing in underlying representations in Russian is ``nondistinctive in all but the last member of a cluster of obstruents'' \citep[283]{A74}.
+It is also the case that voicing in underlying representations in Russian is ``nondistinctive in all but the last member of a cluster of obstruents'' \citep[283]{A74}.
 It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters is part of the grammatical knowledge of Russian speakers. 
@@
 Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
-There are also many cases where restrictions on underlying representations hold of underlying representations but not of derived forms.
+There are also many cases where restrictions on underlying representations are specific to underlying representations and do not apply to surface representations.
 For instance, hiatus is exceptionally rare in native Turkish roots; of the handful of examples in the Turkish Electronic Living Lexicon \citep{TELL}, many appear to be compounds (e.g., \emph{ısıalan} `endothermic', cf.~\emph{ısı} `heat', \emph{alan} `taker').
-Yet, numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
+Yet numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
 \citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of constraints on underlying representations.
@@
 In English, sequences of adjacent obstruents which do not also agree in voice (e.g., \emph{a}[b.s]\emph{inth}) are quite rare within a word, and therefore a hetero-voiced obstruent cluster is a clue to the presence of a word boundary in running speech.
-Infants \citep[e.g.,][]{Mattys2001b} and \citep{McQueen1998b,Norris1997} are both thought to use this heuristic for word recognition in experimental settings.\footnote{
+Infants \citep[e.g.,][]{Mattys2001b} and adults \citep{McQueen1998b,Norris1997} are both thought to use this heuristic for word recognition in experimental settings.\footnote{
     There is reason to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
@@
 By hypothesis, the latter root ends in /d/, so the constraint against final voiced obstruents is specific to surface representations.
-Yet, \citeauthor{Shibatani1973} claims, German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
+\citeauthor{Shibatani1973} claims, however, that German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
     Voicing of final obstruents is usually lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \citep[506]{Ussishkin2003}.}
@@
 
-\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and is therefore must be independent of other types of grammatical knowledge.
-This is not fully consistent with the available experimental evidence available, however.
+\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and it is therefore must be independent of other types of grammatical knowledge.
+This is not fully consistent with the available experimental evidence, however.
 
@@
 While there are many gaps in current understanding (and keeping in mind the usual caveats about a strong interpretation of negative results), there is no reason to think that phonotactic knowledge is acquired before non-trivial amounts of lexical and phonological acquisition.
-If anything, infants's phonotactic knowledge appears to come online at approximately the same point as highly specific phonological representations including subsyllabic units and the ability to extract allophonic generalizations, as predicted by the firm link between phonological and phonotactic knowledge advocated for here.
+If anything, infants's phonotactic knowledge appears to come online at approximately the same point as highly specific phonological representations including subsyllabic units and the ability to extract allophonic generalizations, as predicted by the firm link between phonological and phonotactic knowledge argued for here.
 

### 2013-01-10 1.tex
immediate pre-defense writing

@@
 It is also the case that voicing in underlying representations in Russian is ``nondistinctive in all but the last member of a cluster of obstruents'' \citep[283]{A74}.
-It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters is part of the grammatical knowledge of Russian speakers. 
-For instance, \citeauthor{PE} write:
+It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters is part of the grammatical knowledge of Russian speakers, as some linguists have done.
 
 \begin{quote}
-Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their langauge, since they are computationally inert and thus irrelevant to the input-output mappings that the grammar is responsible for.
+MSCs are merely artifacts of the grammar, and thus play no part in the phonological component of a language. \citep[302]{Clayton1976}
+
+Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mappings that the grammar is responsible for.
 \citep[17f.]{PE}
+
+[MSCs] have no observable consequences in the course of the normal use of the language\ldots{} \citep[320]{Zimmer1969}
 \end{quote}
@@
 
-There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
+%FIXME
+
+\begin{example}[Adaptation of onset clusters in Finnish loanwords]
+\begin{tabular}{l llll}
+a. & toːri    & `store'   & ($<$ \emph{store})              & \citep[English:][89]{Hellstrom1976}   \\
+   & rosseri  & `grocery' & ($<$ \emph{grocery})            &                             \\
+b. & ranska   & `French'  & ($<$ \emph{Franska})            & \citep[Swedish:][67]{Campbell2004}    \\
+   & ruːvi    & `screw'   & ($<$ \emph{skruv})              &                             \\
+c. & risti    & `cross'   & ($<$ \emph{kristĭ})             & \citep[Old Russian:][60]{Bjornflaten2006} \\
+   & raːmattu & `bible'   & ($<$ \emph{gramota} `document') &                             \\
+\end{tabular}
+\end{example}
+
+There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt{Fries1949}, \citealt{Holden1976}, \citealt{Ito1995a,Ito1995b}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
+For instance, native words in San Mateo Huave all end in a consonant, but final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
+%FIXME example here
 Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.

### 2013-01-10 3.tex
immediate pre-defense writing

@@
 There is by now an enormous amount of research that adopts this assumption.
-Semitic languages, especially Arabic \citep{Coetzee2008a,Frisch2004,McCarthy1988,Pierrehumbert1993} but also Berber \citep{Elmedlaoui1995} and Tigrinya \citep{Buckley1997}, and Austronesian languages, including Javanese \citep{GraffInPress,Mester1988}, Muna \citep{Anttila2008a,Coetzee2008a}, and Samoan \citep{AldereteInPress}, have been of particular interest to linguists adopting this hypothesis; Aymara \citep{GraffInPress}, Cantonese \citep{Yip1989}, English \citep{Berkley1994a,Berkley1994b,Berkley2000,Coetzee2008b,Martin2007,Martin2011}, Gitksan \citep{Brown2010}, Japanese \citep{Kawahara2006}, Ju|'hoansi, \citep{Kinney2005}, Navajo \citep{Martin2007,Martin2011}, Ofo \citep[38f.]{MacEachern1999}, Russian \citep{Padgett1991,Padgett1992}, Shona \citep{Hayes2008a}, and Wargamay \citep{Hayes2008a} have also received a statistical treatment.
+Semitic languages, especially Arabic \citep{Coetzee2008a,Frisch2004,McCarthy1988,Pierrehumbert1993} but also Berber \citep{Elmedlaoui1995} and Tigrinya \citep{Buckley1997}, and Austronesian languages, including Javanese \citep{GraffInPress,Mester1988}, Muna \citep{Anttila2008a,Coetzee2008a}, and Samoan \citep{AldereteInPress}, have been of particular interest to linguists adopting this hypothesis; Aymara \citep{GraffInPress}, Cantonese \citep{Yip1989}, English \citep{Berkley1994a,Berkley1994b,Berkley2000,Coetzee2008b,Davis1989,Martin2007,Martin2011}, Gitksan \citep{Brown2010}, Hungarian \citep{Grimes2010}, Japanese \citep{Kawahara2006}, Ju|'hoansi \citep{Kinney2005}, Navajo \citep{Martin2007,Martin2011}, Ofo \citep[38f.]{MacEachern1999}, Russian \citep{Padgett1991,Padgett1992}, Shona \citep{Hayes2008a}, and Wargamay \citep{Hayes2008a} have also received a statistical treatment.
 \citet{Pozdniakov2007} analyze lexical tendencies in a diverse sample of 30 languages.

### 2013-01-17 1.tex
Pre-"revisions", with lots of material in the TODO

@@
     Throughout, the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory.
-    See \citealt{LignosInPressa} for a review of experimental evidence on this question.}
+    For a review of recent experimental evidence on this question, see \citealt{LignosInPressa}.}
 
@@
 Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
-In contrast, were there no contraint on underlying obstruent voicing, however, it is not immediately obvious that \emph{anything} would be different. 
+In contrast, were there no constraint on underlying obstruent voicing, however, it is not immediately obvious that \emph{anything} would be different. 
 Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
@@
 Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
-For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; this suggests that
-multiple ejective words are not merely absent, but actually difficult for these speakers to perceive, represent, or produce.
+For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language;
+in other words, multiple ejective sequences are not merely absent, but also difficult for these speakers to repeat.
 
 In English, sequences of adjacent obstruents which do not also agree in voice (e.g., \emph{a}[b.s]\emph{inth}) are quite rare within a word, and therefore a hetero-voiced obstruent cluster is a clue to the presence of a word boundary in running speech.
-Infants \citep[e.g.,][]{Mattys2001b} and adults \citep{McQueen1998b,Norris1997} are both thought to use this heuristic for word recognition in experimental settings.\footnote{
+Infants \citep[e.g.,][]{Mattys2001b} and adults \citep{McQueen1998b,Norris1997} are thought to use this heuristic for word recognition in experimental settings.\footnote{
     There is reason to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
@@
 
-Loanword adaptation may provide further evidence for the psychological reality of phonotactic knowledge.
+Loanword adaptation may provide further evidence for the grammatical relevance of phonotactic knowledge.
 In Desano \citep{Kaye1974}, for instance, all underlying representations (URs) are either totally oral (e.g., [yaha] `to hear') or totally nasal (e.g., [ñãhã] `to enter'), and loanwords are made to conform to this generalization: Portuguese \emph{martelo} `hammer' is adapted as [barateru] and Spanish \emph{naranja} `orange' as [nãnãnã].
-Presumably, some component of the synchronic grammar is responsible for the fact that the restriction over native vocabulary is extended to loanwords.
-
-%FIXME
+It is not a stretch to imagine that some component of the synchronic grammar is responsible for the fact that the restriction over native vocabulary is extended to loanwords.
+% FIXME discuss evidence that Desano has nasal spreading
+% FIXME mention \citet{Jacobs2000}
 
@@
 Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.
+% FIXME mention hyman
+% peperkamp & dupoux 2003
+% peperkamp 2005
 
@@
 
-Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' phonotactic knowledge, other authors observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \citealt{Hale1965}:297, \citealt{Kisseberth1970b}, \citealt{Postal1968}:212f.).
+Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' phonotactic knowledge, other authors observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \emph{SPE}:382, \citealt{Hale1965}:297, \citealt{Kisseberth1970b},
+%\citealt{PasterInPress},
+\citealt{Postal1968}:212f., \citet{Stanley1967}:401).
 Russian obstruent voice assimilation, discussed above, provides an example of this type: there are no hetero-voiced obstruent clusters in either underlying or surface representations.
@@
     This admits the possibility that markedness constraints not implicated by alternations will remain undominated \citep[e.g.,][]{Coetzee2008b}.
-    It is difficult to evaluate this proposal in the absence of any substantive proposal for the contents of \textsc{Con}, the universal constraint sets, but this has the potential to blur the traditional distinction between accidental and structural phonotactic gaps: any exceptionless gap which corresponds to a markedness constraint \emph{in any language} will be accorded structural, inviolable status, a quite powerful prediction.
+    It is difficult to evaluate this proposal in the absence of any complete proposal for the contents of \textsc{Con}, the universal constraint sets, but this has the potential to blur the traditional distinction between accidental and structural phonotactic gaps: any exceptionless gap which corresponds to a markedness constraint \emph{in any language} will be accorded structural, inviolable status, a quite powerful prediction.
     If \textsc{Con} is sufficiently elaborate to incorporate constraints like the *[spVp] proposed by \citet{Coetzee2008b}, it might also incorporate a constraint like *[s.w], perhaps as a subcomponent of the so-called \emph{syllable contact law} \citep[e.g.,][]{Gouskova2004,Murray1983}, which disfavors syllable contact clusters with increasing sonority
-    The *[s.w] constraint is without exception in English, yet nonce words like \emph{teeswa} [tis.wa] seem quite unobjectionable.
+    The *[s.w] constraint is without exception in English, yet nonce words like \emph{teeswa} [tes.wa] seem quite unobjectionable.
+    % FIXME tilwa?
     Numerous other examples of this type could be adduced.
@@
 
-\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and it is therefore must be independent of other types of grammatical knowledge.
+\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and it therefore must be independent of other types of grammatical knowledge.
 This is not fully consistent with the available experimental evidence, however.
@@
 Crucially, all of this occurs \emph{before} there is any evidence for phonotactic knowledge, for which the earliest evidence begins at approximately 9 months of age \citep{Friederici1993,Jusczyk1993b,Jusczyk1994}.
+% Sketch what this is.
 While there are many gaps in current understanding (and keeping in mind the usual caveats about a strong interpretation of negative results), there is no reason to think that phonotactic knowledge is acquired before non-trivial amounts of lexical and phonological acquisition.
@@
 The remaining gaps in the cluster inventory are attributed to the sparse nature of the lexicon, not static phonotactic restrictions.
+
+Chapter \ref{conclusions} summarizes the findings and describes future directions.
 

### 2013-01-17 2.tex
Pre-"revisions", with lots of material in the TODO

@@
 One of the major  for the development of stochastic models o
+%FIXME
 motivations for the 
@@
 %\citep{Bard1996}
+% FIXME short discussion of the utility of gradient judgements in syntax
 
@@
 There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}.
-As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures may not be stored in linguistic memory. 
+As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures cannot be stored in linguistic memory. 
 %This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
@@
 Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, many of which are exceptionless, into probabilities.
-While this is consistent with their claim, that ``the ability to model gradient intuitions [are] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.
+While this is consistent with their claim, that ``the ability to model gradient intuitions [is] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.
+
+% FIXME make this work: As \citet{Newmeyer2007} writes, ``the idea that categoricity is not represented in the data itself is a truism.
+Whether distinctions of grammaticality (as opposed to acceptability) are binary is a difficult question.'' (398)
 
@@
 
+% FIXME note that it might still be useful to use continuous judgements to gather things, since it might have more statistical power (sprouse)
+% featherston 2007
+
 %An emerging consensus suggests, however, that infants attend to transitional probabilities primarily in the absence of grammatical cues \citep{Gambell2005,Hohne1994,Johnson2001,Jusczyk1999c,Lignos2012b,Mattys2001a,Shukla2007,Lew-Williams2012}.

### 2013-01-17 3.tex
Pre-"revisions", with lots of material in the TODO

@@
 \noindent
-However, subsequent sound changes \citep[e.g.,][]{Baldi1994,Safarewicz1932}, particularly the degemination of Old Latin \emph{ss} before diphthongs and long monophthongs (e.g., \emph{caussa} $>$ \emph{causa} `cause'), introduce numerous exceptions to \textsc{Rhotacism}.
+However, subsequent sound changes \citep[e.g.,][]{Baldi1994,Safarewicz1932}, particularly the degemination of Old Latin \emph{ss} after diphthongs and long monophthongs (e.g., \emph{caussa} $>$ \emph{causa} `cause'), introduce numerous exceptions to \textsc{Rhotacism}.
 In Classical Latin, intervocalic \emph{s} is found root-internally (\emph{asellus} `donkey', \emph{casa} `hut'), in environments derived by inflectional suffixes (\emph{uās}-\emph{uāsis} `vase', \emph{uisēre} `to view'), prefixation (\emph{dēsecāre} `to cut off'; cf.~\emph{dē} `from', \emph{secāre} `to cut'), compounding (\emph{olusātrum} `parsnip'; cf.~\emph{olus} `vegetable', \emph{ātrum} `black'), and denominal adjective formation (\emph{uentōsus} `windy'; cf.~\emph{uentus} `wind'), and is tolerated in nativized loanwords from Celtic (\emph{omāsum} `tripe'), Germanic (\emph{glaesum} `amber'), and Greek (\emph{basis} `pedestal').
@@
 While some of the apparent exceptions may be the result of opaque phonological application \citep{Heslin1987}---an explanation not yet available in Saussure's time---any formulation of rhotacization will admit nearly as many lexical exceptions as there are roots exhibiting \emph{s}-\emph{r} alternations \citep{GormanInPressc}.
-Any synchronic account of the underrepresentation of intervocalic \emph{s} must confront and the unproductive nature (as indicated by the accumulation of exceptions) of the proximate explanation for this tendency.
+Any synchronic account of the underrepresentation of intervocalic \emph{s} must confront the unproductive nature (as indicated by the accumulation of exceptions) of the proximate explanation for this tendency.
+
+% FIXME mention Fudge's analysis of short-a before voiceless fricatives
+% \citet{Trager1930,Trager1934,Trager1940},
+% \citet{Cohen1970}
+% \citet{Ferguson1975}
+% \citet{PLC1} 429f.
+% \citet{Fudge1987}
+% \citet{AOE} I.203f., 515f. for an attempt to write a rule
+% \citet{Payne1976,Payne1980} who finds that whereas speakers from out of state acquire 5 vocalic sound changes, only one out of 34 interviewees acquired the core pattern
+% Fudge(!)
 
@@
 A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}]. 
-In other words, no representation is never truly an exception; rather, some underlying representations have non-default features which do not match the extended structural description of $R$, which requires that the target be [$+R$].
+In other words, no representation is ever truly an exception; rather, some underlying representations have non-default features which do not match the extended structural description of $R$, which requires that the target be [$+R$].
 If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.
@@
 
-\citeauthor{A74} also notes that roots which fail to undergo suffix harmony, like (\ref{turkexcept}b), may themselves be harmonic. He takes to be evidence for the necessity of duplication:
+\citeauthor{A74} also notes that roots which fail to undergo suffix harmony, like (\ref{turkexcept}b), may themselves be harmonic. He takes this to be evidence for the necessity of duplication:
 
 \begin{quote}
-\ldots{}there are words which are exceptions to harmony across boundaries (e.g., \emph{kabahat} `fault', \emph{kabahattı} `his fault') but which are perfectly regular internally.
+\ldots{}there are words which are exceptions to harmony across boundaries (e.g., \emph{kabahat} `fault', \emph{kabahatti} `his fault') but which are perfectly regular internally.
 Since the morpheme structure condition and the phonological rule in this case have distinct classes of exceptions, it is clear that they cannot be identified.
@@
 It is unclear whether the cluster-splitting vowel is deleted in the non-native variant or epenthesized in the fast speech variant. 
-    Under either analysis, there is no ready explanation for the tendency of the cluster-splitting vowel to have the backness features of following consonants; if anything, one might expect it to determine the backness features of following consonants.
-    All that can be said with certainty is that the adaptation of non-native onset clusters appears to proceed in such a fashion so that the lexical items in question are [$-$\textsc{Backness Harmony}].
+    Under either analysis, there is no ready explanation for the tendency of the cluster-splitting vowel to have the backness features of following vowels; if anything, one might expect it to determine the backness features of following vowels.
+    All that can be said with certainty is that the adaptation of non-native onset clusters appears to proceed in such a fashion so that the lexical items in question are [$+$\textsc{Backness Harmony}].
 
@@
     There is reason to suspect that reduplicants bear the lexical diacritics of the roots they are derived from.
-    In Kinande, verb reduplication, 
-    Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, requires a bisyllabic reduplicant. 
+    In Kinande, verb reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, requires a bisyllabic reduplicant. 
     This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots contain three copies of the root so as to satisfy this requirement, and reduplicated forms of many trisyllabic verbs are ineffable.
@@
 
-A number of studies have investigated the role of harmony in word-spotting tasks, though to mimic auditory word recognition and segmentation in natural settings. 
+A number of studies have investigated the role of harmony in word-spotting tasks, thought to mimic auditory word recognition and segmentation in natural settings. 
     Many of these studies have been carried out in Finnish, which has a vowel harmony system similar to that of Turkish. 
@@
 To test for associations between the process (more specifically, the constraint that it imposes on roots) and type frequency in this database, each root was sorted into a $2 \times 2$ contingency table; the contents of each cell are specific to the process in question.
-The counts in this table are not expected to add up to 9,601, since many roots neither could exemplify nor violate the process in question; for instance, monosyllabic words are irrelevant to root harmony.
+The counts in this table are not expected to add up to 9,601, since many roots neither exemplify nor violate the process in question; for instance, monosyllabic words are irrelevant to root harmony.
 The Fisher exact test is used to compute a $p$-value representing the probability of the observed patterns arising under the null hypothesis that there is no association between the constraint and type frequency.
@@
 
-\textsc{Backness Harmony} is exemplified in the lexicon insofar as there is an positive association between the backness of vowels in all adjacent syllables.
-Any disagreement on the backness specification of vowels in adjacent syllables are counted solely as exceptions, even if other vowel transitions in the root are harmonic; further assumptions are necessary to determine whether a root which has one disharmonic transition may in any sense ``obey'' harmony elsewhere.
+\textsc{Backness Harmony} is exemplified in the lexicon insofar as there is a positive association between the backness of vowels in all adjacent syllables.
+Any disagreements on the backness specification of vowels in adjacent syllables are counted solely as exceptions, even if other vowel transitions in the root are harmonic; further assumptions are necessary to determine whether a root which has one disharmonic transition may in any sense ``obey'' harmony elsewhere.
 %This follows from \emph{SPE} theory of lexical exceptionality (see also \citealt{Gouskova2012}), in which phonological exceptionality is a property of underlying representations, not individual segments.
@@
 The first used 23 native adult speakers who were permitted to select either nonce word as more like Turkish, or to indicate `no preference'. 
-For the purposes below, `no preference' results are of.
+For the purposes below, `no preference' results are ignored. 
 The second experiment used 32 native adults, none of whom appeared in the preceding study, and used a forced binary choice.

### 2013-01-17 4.tex
Pre-"revisions", with lots of material in the TODO

@@
 a wordlist is generated using the English portion of the CELEX database \citep{CELEX}. 
-Only words coded as ``monomorphemic'' are used, and all words coded as non-native are excluded.
+Only words marked in CELEX as ``monomorphemic'' are used, and all words 
+labeled non-native are excluded.
 This stringent criterion for inclusion eliminates many of the exceptions noted by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies.
@@
 
-\begin{example}[Latinate verbs and small clauses] 
+\begin{example}[Syntactic restrictions on Latinate verbs]
 \label{harley}
@@
 
-\textsc{Nasal Place Assimilation} (e.g., \citealt[65f.]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}) permits [ŋ] to be described as a pure allophone of /n/, and accounts for certain patterns in Latinate prefix allomorphy.
+\textsc{Nasal Place Assimilation} (e.g., \citealt[65f.]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}) 
+%\citealt[269]{Fudge1969}) 
+permits [ŋ] to be described as a pure allophone of /n/, and accounts for certain patterns in Latinate prefix allomorphy.
 
@@
 By hypothesis, \textsc{Obstruent Voice Assimilation}, \textsc{Nasal Place Assimilation}, and \textsc{Degemination} rule out a large number of possible clusters. 
-In all, they target 316 out of 920 possible clusters (34\%), and only 11 (3\%) of these clusters has non-zero frequency in the sample.
+In all, they target 316 out of 920 possible clusters (34\%), and only 11 (3\%) of these clusters have non-zero frequency in the sample.
 Together, these three processes define a simple classifier in which a cluster is predicted to be unattested if and only if it would be neutralized by one of these processes. 
@@
 Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original paper is attempted: both their implementation and phonological feature specifications are used here.
-Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in \S\ref{syllabification} and a novel feature [$\pm$
+Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in Appendix \ref{syllabification} and a novel feature [$\pm$
 \textsc{Coda}] is added to allow the model to distinguish coda and onset co

### 2013-01-28 1.tex
Revisions! Lots of 'em.

@@
 The first is to outline the empirical scope of \emph{phonotactic theory}, the theory of speakers' knowledge of possible and impossible sounds, sound sequences, and words.
-The second is to show that the core facts of this domain are compatible with a traditional view of phonotactic knowledge as independent of the lexicon, categorical, and closely related to phonological processes, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of probabilistic inference over the lexicon, and therefore gradient and independent of phonological processes.\footnote{
+The second is to suggest that the core core facts of this domain are compatible with a traditional view of phonotactic knowledge as independent of the lexicon, categorical, and closely related to phonological processes, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of probabilistic inference over the lexicon, and therefore gradient and independent of phonological processes.\footnote{
     Throughout, the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory.
@@
 It is also the case that voicing in underlying representations in Russian is ``nondistinctive in all but the last member of a cluster of obstruents'' \citep[283]{A74}.
-It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters is part of the grammatical knowledge of Russian speakers, as some linguists have done.
+It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters (and many other constraints on underlying representations that have been proposed) is part of the grammatical knowledge of Russian speakers, as some linguists have done.
 
@@
 \citep[17f.]{PE}
-
-[MSCs] have no observable consequences in the course of the normal use of the language\ldots{} \citep[320]{Zimmer1969}
+%[MSCs] have no observable consequences in the course of the normal use of the language\ldots{} \citep[320]{Zimmer1969}
 \end{quote}
@@
 Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
-In contrast, were there no constraint on underlying obstruent voicing, however, it is not immediately obvious that \emph{anything} would be different. 
+In contrast, were there no constraint on underlying obstruent voicing, however, it is not immediately obvious that anything would be different.
 Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
@@
 Yet numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
-\citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of constraints on underlying representations.
+\citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of restrictions that hold of underlying representations.
 However, there are many facts which strongly suggest that speakers internalize phonotactic constraints.
@@
 As \citet{Halle1962} and \citet{Chomsky1965} note, speakers can distinguish between a well-formed and an ill-formed word, neither of which is an actual word.
-Neither [blɪk] \emph{blick} nor [bnɪk] \emph{bnick} is a word of English, yet English speakers immediately report that only the former is ``possible'', an \emph{accidental} gap in the lexicon, whereas the latter is judged to be impossible (i.e., structurally excluded).
+Neither [blɪk] \emph{blick}\footnote{
+    In fact, philospher R.M.~\citet{Hare1955} uses the term \emph{blik} to describe \emph{a priori}, unfalsifiable frames of reference through which experiences are filtered; for instance, materialism can be thought of as the \emph{blik} interpreting all experience in terms of the properties of matter.
+    Since this term has some currency in the philosophy of science and religion, this is likely a ``real word'' rather than a lexical gap for a significant English-speaking minority. 
+    Thanks to Steve Anderson for brining this to my attention.}
+nor [bnɪk] \emph{bnick} is a word of English, yet speakers immediately report that only the former is ``possible'', an \emph{accidental} gap in the lexicon, whereas the latter is judged to be impossible (i.e., structurally excluded).
 There can be no question that it is part of speakers' knowledge.
-Elicited in a controlled fashion, these \emph{wordlikeness judgements} are perhaps the most important (and least controversial) source of phonotactic evidence, and they play a major role in chapters \ref{gradience}--\ref{turkish} of this dissertation.
+Elicited in a controlled fashion, these \emph{wordlikeness judgements} are perhaps the most important (and least controversial) source of phonotactic evidence, and they play a major role in Chapters \ref{gradience}--\ref{turkish} of this dissertation.
 
@@
 Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
-For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language;
-in other words, multiple ejective sequences are not merely absent, but also difficult for these speakers to repeat.
+For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; in other words, multiple ejective sequences are not merely absent, but also difficult for these speakers to repeat.
 
@@
 
-Loanword adaptation may provide further evidence for the grammatical relevance of phonotactic knowledge.
-In Desano \citep{Kaye1974}, for instance, all underlying representations (URs) are either totally oral (e.g., [yaha] `to hear') or totally nasal (e.g., [ñãhã] `to enter'), and loanwords are made to conform to this generalization: Portuguese \emph{martelo} `hammer' is adapted as [barateru] and Spanish \emph{naranja} `orange' as [nãnãnã].
-It is not a stretch to imagine that some component of the synchronic grammar is responsible for the fact that the restriction over native vocabulary is extended to loanwords.
-% FIXME discuss evidence that Desano has nasal spreading
-% FIXME mention \citet{Jacobs2000}
+Loanword adaptation may provide further evidence for the grammatical relevance of phonotactic knowledge \citep[e.g.,][]{Fischer-Jorgensen1952}.
+In Desano \citep{Kaye1971,Kaye1974}, for instance, all underlying representations (URs) are either totally oral (e.g., [jaha] `to hear') or totally nasal (e.g., [ñãhã] `to enter'), and loanwords are made to conform to this generalization: Portuguese \emph{martelo} `hammer' is adapted as [barateru] and Spanish \emph{naranja} `orange' as [nãnãnã].\footnote{
+    \citeauthor{Kaye1971}'s transcriptions have been converted to IPA notation.}
+It is not difficult to imagine that some component of the synchronic grammar is responsible for the fact that the restriction over native vocabulary is extended to loanwords.
+Furthermore, many Desano affixes have have distinct oral and nasal allomorphs. 
+For instance, a noun classifier suffix for round objects takes the form [ru] after oral roots (e.g., [goru] `ball') and [nũ] after nasal roots (e.g., [sẽnãnũ] `pineapple').
+%\citep[38]{Kaye1971}. 
+As \citet[38]{Kaye1971} notes, this distribution implies the existence of a process of nasal harmony, and it is not implausible that the same process is responsible for the above-mentioned adaptations.
 
-\begin{example}[Adaptation of onset clusters in Finnish loanwords]
-\begin{tabular}{l llll}
-a. & toːri    & `store'   & ($<$ \emph{store})              & \citep[English:][89]{Hellstrom1976}   \\
-   & rosseri  & `grocery' & ($<$ \emph{grocery})            &                             \\
-b. & ranska   & `French'  & ($<$ \emph{Franska})            & \citep[Swedish:][67]{Campbell2004}    \\
-   & ruːvi    & `screw'   & ($<$ \emph{skruv})              &                             \\
-c. & risti    & `cross'   & ($<$ \emph{kristĭ})             & \citep[Old Russian:][60]{Bjornflaten2006} \\
-   & raːmattu & `bible'   & ($<$ \emph{gramota} `document') &                             \\
+However, it is not possible to equate all patterns of adaptation with phonological alternations targeting native vocabulary.
+\citet{Peperkamp2005} highlights several cases where native alternations are distinct from loanword adaptations.
+For instance, in Korean, [s] does not surface in codas.
+As \citet{Kenstowicz2001} report, native /s/ is realized as [t] in codas (e.g., [nat]-[nasɨl] `sickle \textsc{nom.}-\textsc{acc.}').
+In loanwords, however, final [s] becomes an onset by epenthesis (e.g., \emph{boss} $>$ [posɨ]).
+Evidence of this sort has lead many \citep[e.g.,][]{Dupoux1999,Peperkamp2003,Peperkamp2005} to suggest that loanword adaptations are the result of speech perception, not phonological computations.
+
+There are other cases in which loanword adaptations are not easily identified with any synchronic process.
+Consider the case of loanwords which begin in onset clusters and which are borrowed into languages which do not permit complex onsets.
+In the Wikchamni dialect of Yokuts, for instance, Spanish loanwords with complex onsets are adapted either via anaptyxis or deletion of the first consonant.
+
+\begin{example}[Wikchamni Yokuts adaptations \citep{Gamble1989}]
+\begin{tabular}{l lll l}
+a. & \emph{cruz}     & $>$ & k\asp{}uluʃ & `cross' \\
+   & \emph{frijoles} & $>$ & pilhoːliʃ   & `beans' \\
+b. & \emph{plato}    & $>$ & laːto       & `plate' \\
+   & \emph{clavo}    & $>$ & laːwu       & `nail'  \\
 \end{tabular}
@@
 
-There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt{Fries1949}, \citealt{Holden1976}, \citealt{Ito1995a,Ito1995b}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
-For instance, native words in San Mateo Huave all end in a consonant, but final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
-%FIXME example here
-Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.
-% FIXME mention hyman
-% peperkamp & dupoux 2003
-% peperkamp 2005
+\noindent
+However, there is no reason to regard these adaptations as a product of the synchronic phonology: no Yokuts root begins with a consonant, and there is no way to derive an initial CC cluster.
+Furthermore, as \citet{Gamble1989} notes, there is no reason to believe that the adaptations actually occurred within Yokuts at all, since there was at best quite limited contact between Spanish and Yokuts speakers; the adaptations may have occurred in another language altogether.
+
+%\begin{example}[Adaptation of onset clusters in Finnish loanwords]
+%\begin{tabular}{l llll}
+%a. & toːri    & `store'   & ($<$ \emph{store})              & \citep[English:][89]{Hellstrom1976}   \\
+%   & rosseri  & `grocery' & ($<$ \emph{grocery})            &                             \\
+%b. & ranska   & `French'  & ($<$ \emph{Franska})            & \citep[Swedish:][67]{Campbell2004}    \\
+%   & ruːvi    & `screw'   & ($<$ \emph{skruv})              &                             \\
+%c. & risti    & `cross'   & ($<$ \emph{kristĭ})             & \citep[Old Russian:][60]{Bjornflaten2006} \\
+%   & raːmattu & `bible'   & ($<$ \emph{gramota} `document') &                             \\
+%\end{tabular}
+%\end{example}
+
+Finally, there are many cases where putative phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt{Fries1949}, \citealt{Holden1976}, \citealt{Ito1995a,Ito1995b}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
+For instance, native words in San Mateo Huave all end in a consonant, but 
+\citet{Davidson1997} note that final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
+Given the considerable disagreement about the nature of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.
 
@@
 
-Language games or speech disguises may also provide evidence for phonotactic knowledge \citep[e.g.,][]{Vaux2011}; an example appears in \S\ref{ss:bh} below.
+Language games or speech disguises may also provide evidence for phonotactic knowledge \citep[e.g.,][]{Vaux2011}, assuming that these ``alternative phonologies'' are not qualitatively different from naturally-occurring language processes \citep[e.g.,][697]{Bagemihl1995}.
+An example of language game evidence bearing on phonotactic questions can be found in \S\ref{ss:bh}, where a language game is used to argue for the hypothesis that root-internal vowel sequences in Turkish are subject to vowel harmony.
 
@@
 For example, voicing is non-contrastive for /ts, tʃ, x/ in Russian \cite[22]{SPR}: [dz, dʒ, ɣ] appear in surface, but not underlying, representations.
-\emph{Sequence structure constraints} apply to underlying sequences; an example is given below.
+This dissertation will have little to say about the nature of segment structure constraints.
+Of more interest here are \emph{Sequence structure constraints} apply to underlying sequences; an example is given below.
 
 \begin{example}[An English MSC (adapted from \citealt{Chomsky1965}:100)]
-$\begin{bmatrix} +\textsc{Cons} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Liquid} \end{bmatrix}~/~\#~\begin{bmatrix} +\textsc{Cons} \end{bmatrix}~\gap$
+$\begin{bmatrix} +\textsc{Cons} \end{bmatrix}~\goesto~\begin{bmatrix} +\textsc{Liquid} \end{bmatrix}~/~\#~\begin{bmatrix} -\textsc{Cont} \end{bmatrix}~\gap$
 \end{example}
@@
 \noindent
-This sequence structure constraint specifies the second of a sequence of word-initial consonants as a liquid, so as to preclude underlying */bnɪk/, for example.
+This sequence structure constraint specifies that a consonant immediately after a word-initial stop is a liquid; this would preclude underlying */bnɪk/, for example.
 
@@
 \citeauthor{Shibatani1973} claims, however, that German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
-    Voicing of final obstruents is usually lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \citep[506]{Ussishkin2003}.}
+    Furthermore, voicing of final obstruents is usually lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \citep[506]{Ussishkin2003}.}
 
@@
 
-Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' phonotactic knowledge, other authors observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \emph{SPE}:382, \citealt{Hale1965}:297, \citealt{Kisseberth1970b},
+Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' phonotactic knowledge, other authors observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \citealt{SPE}:382, \citealt{Hale1965}:297, \citealt{Kisseberth1970b},
 %\citealt{PasterInPress},
-\citealt{Postal1968}:212f., \citet{Stanley1967}:401).
+\citealt{Postal1968}:212f., \citealt{Stanley1967}:401).
 Russian obstruent voice assimilation, discussed above, provides an example of this type: there are no hetero-voiced obstruent clusters in either underlying or surface representations.
 These two facts are tantalizingly similar, but are treated as separate if a distinction between morpheme structure constraints and phonological processes is drawn.
-This is a special case of what \citet{Kisseberth1970b} calls \emph{conspiracies}, which \citet[136]{KK77} term the \emph{duplication problem}.
+This is a special case of what \citet{Kisseberth1970b} calls \emph{conspiracies} and what \citet[136]{KK77} term the \emph{duplication problem}.
 \citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that the distinction between constraints on URs and alternations is artificial, and that these different levels of description are related by a principle now known as \emph{Stampean occultation} \citep[54]{OT}.
@@
     %Assuming Lexicon Optimization, the only alternative is to stipulate an ad hoc constraint against onset [ŋ] \citep[39]{Jusczyk2002}.}
-In chapter \ref{gaps}, it is argued that constraints described in terms of non-contrastive prosodic structures can also be derived by Stampean occultation.
+In Chapter \ref{gaps}, it is argued that constraints described in terms of non-contrastive prosodic structures can also be derived by Stampean occultation.
 In an architecture like Lexical Phonology, it is even possible to apply a process to individual underlying representations (i.e., at the ``lexical level'') to enforce constraints which are not surface-true.
-Consequently, morpheme structure constraints are otiose, their empirical converage entirely subsumed by phonological processes.
+Consequently, it is impossible to construct an argument which would distinguish morpheme structure constraints from other types of phonological computations.
 
-This argument against the distinction between morpheme structure constraints and phonological processes is quite similar to the famous argument (\citeyear{SPR}) against the morphophonemic/phonemic distinction.\footnote{
+This is in ways similar to \citeauthor{SPR}'s famous argument (\citeyear{SPR}) against the morphophonemic/phonemic distinction.\footnote{
     This is not to imply that \citeauthor{SPR} was the first to make this argument, or to otherwise disregard this distinction: similar ideas can be found in work by \citet{Bloch1941}, \citeauthor{Bloomfield1926} (\citeyear{Bloomfield1926}:160, \citeyear{Bloomfield1962}:5f.), \citet{Chao1934}, \citet[244f.]{Hamp1953}, and \citet[47f.]{Sapir1930a}, among others (see \citealt{Anderson1985} \emph{passim}).}
@@
 In Russian, obstruents participate in voice assimilation whether this neutralizes a phonemic distinction (\ref{rovs}a) or not (\ref{rovs}b): recall that there is no underlying /dʒ/ in Russian.
-Under biuniqueness, there must be separate neutralizing and non-neutralizing variants of this process, however.
+Under biuniqueness, there must be separate, non-overlapping variants of this process, one applying in neutralizing contexts and another in non-neutralizing contexts.
 From this, \citeauthor{SPR} argues that biuniqueness (and the distinction between the morphophonemic and phonemic levels that follows from it) entails ``a significant increase in the complexity of the representation\ldots{}an unwarranted complication which has no place in a scientific description of language'' (24).
-While \citeauthor{Anderson1985} (\citeyear{Anderson1985}:110f., \citeyear{Anderson2000}) sketches an analysis which preserves biuniqueness without morphophonemic/phonemic duplication (see also \citealt{Kiparsky1985}), this requires further contested assumptions---contrastive underspecification (against which, see \citealt{Steriade1995}) and a Duke-of-York derivation.
-This does not seem inconsistent with \citeauthor{SPR}'s claim that biuniquness imposes unnecessary additional complexities: under \citeauthor{Anderson2000}'s analysis the complexity is not a duplication, but rather a dependency on contentious theoretical assumptions.
+While \citeauthor{Anderson1985} (\citeyear{Anderson1985}:110f., \citeyear{Anderson2000}) correctly observes that it is in principle possible to sketch an analysis of Russian obstruent voicing which preserves biuniqueness without the morphophonemic/phonemic duplication \citeauthor{SPR} objects to (see also \citealt{Kiparsky1985}), this requires further contested assumptions---contrastive underspecification (against which, see \citealt{Steriade1995}) and a Duke-of-York derivation.
+This does not seem inconsistent with \citeauthor{SPR}'s claim that biuniquness imposes unnecessary additional complexities: under \citeauthor{Anderson2000}'s analysis the complexity is not two variants of a single process, but rather a dependency on contentious theoretical assumptions.
 
@@
 
-Under the assumptions of Stampean occultation, there is no way to express substantive constraints on underlying representations except via phonological process.
-In lexical phonology, such constraints are the consequences of phonological processes operating at the ``lexical'' level, and in an Optimality Theory grammar, these are high ranked markedness constraints.\footnote{In Optimality Theory, it is often assumed that all markedness constraints are ranked above all faithfulness constraints in the ``initial state'' of the learner \citep[e.g.,][]{Smolensky1996a}.
-    This admits the possibility that markedness constraints not implicated by alternations will remain undominated \citep[e.g.,][]{Coetzee2008b}.
-    It is difficult to evaluate this proposal in the absence of any complete proposal for the contents of \textsc{Con}, the universal constraint sets, but this has the potential to blur the traditional distinction between accidental and structural phonotactic gaps: any exceptionless gap which corresponds to a markedness constraint \emph{in any language} will be accorded structural, inviolable status, a quite powerful prediction.
-    If \textsc{Con} is sufficiently elaborate to incorporate constraints like the *[spVp] proposed by \citet{Coetzee2008b}, it might also incorporate a constraint like *[s.w], perhaps as a subcomponent of the so-called \emph{syllable contact law} \citep[e.g.,][]{Gouskova2004,Murray1983}, which disfavors syllable contact clusters with increasing sonority
-    The *[s.w] constraint is without exception in English, yet nonce words like \emph{teeswa} [tes.wa] seem quite unobjectionable.
-    % FIXME tilwa?
-    Numerous other examples of this type could be adduced.
-    Consequently, this specific model is rejected here.}
-A corrolary is that if some phonotactic constraint is not (or cannot be) identified with a phonological process, it does not exist in the minds of speakers at all.
+Stampean occultation provides a mechanism for many sequence structure constraints constraints on underlying representations to be expressed as phonological processes. 
+However, not all restrictions on underlying representation have an obvious reflex in the system of alternations.
+Of the residue that remains once these \emph{derived constraints} are identified, many can be attributed to the language-specific prosodic inventory.
+For instance, a language which does not permit onset clusters may do so without there being evidence for processes that simplify initial consonant clusters; Yokuts, discussed above, is a clear example.
+There is a sense in which prosodic parsing operations like syllabification or footing can be thought of as phonological computations, and the boundary between prosodic restrictions and alternations is not always clear.
+For instance, Latin does not permit a geminate consonant after a long vowel; since geminates always span syllable boundaries, this appears at first glance to be a restriction on the Latin syllable template.
+However, Latin seems to enforce this restriction via a process of degemination \citep{GormanInPressc}; assibilation of \emph{t, d} in perfect passive participles produces a geminate \emph{ss} (e.g., 
+\emph{fossus} `dug', cf.~\emph{fodere} `to dig') except after diphthongs and long monophthongs, where a singleton \emph{s} appears (e.g., \emph{lūsus} `played', cf.~\emph{lūdere} `to play').
+This restriction is simultaneously a component of the prosodic inventory and of the alternation system.
 
-This principle---\emph{no static phonotactic constraints}---has interesting ramifications for evaluating certain competing phonological analyses.
+A question which is central to phonotactic theory is whether there are additional types of sequence structure constraints beyond those which are derived from alternations or restrictions on the prosodic inventory.
+Recent answers to the affirmative have generally been presented with proposals for how such constraints are learned.
+
+One possibility will be carefully considered in this dissertation.
+According to this view, much of phonotactic learning is accomplished by statistical inference performed over the lexicon, and phonotactic knowledge is the sum of these lexical statistical patterns.
+Two predictions immediately follow from this position.
+First, since statistical generalizations may be more or less true, phonotactic knowledge may be gradient, reflecting the strength of the many competing patterns involved.
+Secondly, lexical statistical patterns need not be reflected in the system of alternations or in the prosodic system at all: they may be ``static''.
+These predictions will be taken up in some detail in Chapters \ref{gradience} and \ref{turkish}, respectively.
+
+Another possibility is posed by traditional thinking in Optimality Theory.
+According to a standard proposal, at the ``initial state'' all markedness constraints are ranked above all faithfulness constraints \citep[e.g.,][]{Smolensky1996a}.
+If learning proceeds via constraint demotion, markedness constraints which are not implicated by alternations will remain undominated \citep[e.g.,][]{Coetzee2008b}.
+It is difficult to make concrete predictions from this proposal in the absence of a complete proposal for the contents of the universal constraint set \textsc{Con}, but it has the potential to blur the distinction between accidental and structural phonotactic gaps, a distinction which is at the heart of phonotactic theory and is the subject of Chapter \ref{gaps}.
+If the constraint set is in fact universal, any exceptionless gap which corresponds to a markedness constraint \emph{in any langauge} will be accorded the status of an inviolable phonotactic restriction.
+If \textsc{Con} is sufficiently rich to incorporate constraints like the *[spVp] proposed by \citet{Coetzee2008b}, it seems quite likely that it will also contain markedness constraints rulling out English syllable contact clusters which are regarded as accidental gaps in Chapter \ref{gaps}.
+For instance, a constraint *[s.w] seems plausible as a subcomponent of the so-called \emph{syllable contact law} \citep[e.g.,][]{Gouskova2004,Murray1983} which disfavors syllable contact clusters with increasing sonority.
+This *[s.w] constraint is without exception in English, yet nonce words like \emph{teeswa} [tes.wa] seem quite unobjectionable to native speakers.
+Numerous other examples of this type could be adduced, and pose a serious problem for the markedness-over-faithfulness model of phonotactic learning.
+
+The merits of these two models are the subject of the rest of this dissertation.
+Whatever these merits, there is some value also in the null hypothesis, that there are no static phonotactic constraints at all, if it can be maintained.
+Specifically, the principle of \emph{no static phonotactics} has interesting ramifications for evaluating certain competing phonological analyses.
 Consider Sanskrit aspiration alternations such as \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
@@
 
-\subsection{Acquisition and the (in)dependence of phonotactic knowledge}
-\label{ss:aerbpp}
-
-\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and it therefore must be independent of other types of grammatical knowledge.
-This is not fully consistent with the available experimental evidence, however.
-
-Typically-developing infants know their names and the names of their caretakers as early as 4 months of age \citep{Bortfeld2005,Mandel1995,Tincoff1999}.
-While infants are born able to distinguish between monosyllabic and bisyllabic words, the youngest infants are thought to recognize syllables ``holistically'' rather than as segment sequences \citep{Bertoncini1981,Eimas1999,Jusczyk1987}.
-As early as 6 months of age, infants know the visual referents of familiar words presented auditorily \citep{Bergelson2012}.
-At 7.5 months, phonological representations are sufficiently detailed to allow infants to discriminate between words like \emph{cup} and mispronunciations like \emph{*tup} \citep{Jusczyk1995}.
-By 8 months, infants are able to locate both familiar and novel words in continuous speech \citep{Jusczyk1997,Seidl2006}.
-Perception of non-native phonetic contrasts for vowels and consonants begins to decline at this age \citep{Best1994,Polka1994,Werker1981,Werker1984,Werker1988}.
-
-While very few studies have investigated young infants' knowledge of phonological alternations, this is the subject of a fascinating study by \citet{White2008}.
-Simplifying somewhat, the experimenters expose 8.5-month-old infants to an artificial language in which fricative voicing is contrastive, but voiced and voiceless variants of plosives are in complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively.
-After familiarization, infants prefer to listen to nonce words preserving this complementary distribution of stops over nonce words which disrupt this distribution (e.g., \emph{na-poli}, \emph{rot-boli}), suggesting that the infants have extracted an allophonic generalization concerning plosive voicing.
-
-Crucially, all of this occurs \emph{before} there is any evidence for phonotactic knowledge, for which the earliest evidence begins at approximately 9 months of age \citep{Friederici1993,Jusczyk1993b,Jusczyk1994}.
-% Sketch what this is.
-While there are many gaps in current understanding (and keeping in mind the usual caveats about a strong interpretation of negative results), there is no reason to think that phonotactic knowledge is acquired before non-trivial amounts of lexical and phonological acquisition.
-If anything, infants's phonotactic knowledge appears to come online at approximately the same point as highly specific phonological representations including subsyllabic units and the ability to extract allophonic generalizations, as predicted by the firm link between phonological and phonotactic knowledge argued for here.
-
 \section{Outline of the dissertation}
@@
 
-Chapter \ref{gradience} considers evidence from wordlikeness rating tasks.
-It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
+Chapter \ref{gradience} considers evidence from wordlikeness rating tasks. It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded. 
 A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as accurately as state-of-the-art gradient wellformedness models.
-Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.
+Once categorical effects are controlled for, gradient models are largely uncorrelated with well-formedness ratings.
 
 Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns. 
-It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.
+It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations may go unlearned when they are not derived from phonological alternations.
 
@@
 
-Chapter \ref{conclusions} summarizes the findings and describes future directions.
+Chapter \ref{conclusions} summarizes the findings, considers their relation to order of acquisition, and proposes directions for future research.
 

### 2013-01-28 2.tex
Revisions! Lots of 'em.

--- 2012-12-02 2.tex
@@
 
-One of the major  for the development of stochastic models o
-motivations for the 
-Recent developments in the theory of phonotactic knowledge are motivated by the claim that speakers' wordlikeness intuitions are inherently \emph{gradient}, i.e., that such intuitions are more granular than a mere binary contrast between ``possible'' and ``impossible'' words.
+Much of the recent work in phonotactic theory has attempted to incorporate the intuition that phonotactic wellformedness is not an ``all-or-nothing'' matter.
+Rather, it is alleged, wellformedness judgements have more fidelity than is implied by a simple contrast between ``possible'' and ``impossible'', and therefore must be measured and studied at a greater degree of granularity.
+
+This is hardly a novel claim, though it has taken on greater import with the emergence of computational models of wordlikeness.
+Early generative discussions of wordlikeness \citep[e.g.,][]{Chomsky1965,Halle1962} are best remembered for the famous examples [blɪk] and [bnɪk], the former representing a ``possible word'' of English and the latter representing an ``impossible word''.
+A naïve account of this contrast would be to derive it from the assumption that segments must be parsed into syllables or subject to further phonological repair (e.g., \citealt[10f.]{Hooper1973}, \citealt[57f.]{Kahn1976}, \citealt{Ito1989a}, \citealt[19f.]{Wolf2009}).
+Unlike some languages (e.g., Morroccan Arabic: \emph{bniqa} `closet'), English does not permit stop-nasal onsets like [bn], so the latter nonce word cannot surface as such.
+In other words, [bnɪk] is an impossible surface representation in English.
+However, in \emph{The Sound Pattern of English} (henceforth, \emph{SPE}), \citet{SPE} introduce a third nonce word, [bznk], constructed so as to be even less English-like than [bnɪk].\footnote{
+    That wordlikeness judgements depend on language-specific knowledge is apparent given that [bznk] is not impossible in all languages: Imdlawn Tashlhiyt Berber permits whole words consisting of a stop-fricative-nasal-stop sequence (e.g., [tzmt] `it is stifling'; \citealt[112]{Dell1985}).
+    This is clear evidence for the assumption that wordlikeness depends on language-specific knowledge.}
+This leads \citeauthor{SPE} to conclude that wordlikeness intuitions are gradient.
+
+\begin{quote}
+Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
+\end{quote}
+
+This brings the theory of wordlikeness in line with the view of syntactic grammaticality presented by foundational documents like \emph{The Logical Structure of Linguistic Theory} \citep{LSLT} and \emph{Aspects of the Theory of Syntax} \citep{ASPECTS}, and reflexes can be found in later work, such as the proposals of \citet{BARRIERS} and \citet{Huang1982}; see \citealt[43f.]{Schutze1996} for a critique.
+
+\citeauthor{SPE}'s claim about the gradient nature of phonotactic wellformedness does not seem to have had much of an impact on practices of the time---as can be seen from discussion in the previous chapter, contemporary critiques focused on other elements of the \emph{SPE} phonotactic theory---but reflexes can once again be found in later work: for instance, \citet{Borowsky1989}, \citet[50f.]{Clements1983}, and \citet{Myers1987} all assume a contrast between ``peripheral'' and absolutely ungrammatical sound sequences.
+
+Recent discussions of gradient grammaticality in wordlikeness attempt to present experimental support for \citeauthor{SPE}'s intuitions:
 
@@
 
-\ldots{}when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
-\end{quote}
-
-This is not a novel claim.
-Early generative discussions of wordlikeness \citep[e.g.,][]{Chomsky1965,Halle1962} are best remembered for the famous examples [blɪk] and [bnɪk], the former representing a ``possible word'' of English and the latter representing an ``impossible word''.
-A naïve account of this contrast follows from the assumption that segments must be parsed into syllables or subject to further phonological repair (e.g., \citealt[10f.]{Hooper1973}, \citealt[57f.]{Kahn1976}, \citealt{Ito1989a}, \citealt[19f.]{Wolf2009}).
-Unlike some languages (e.g., Morroccan Arabic: \emph{bniqa} `closet'), English does not permit stop-nasal onsets like [bn], so the latter nonce word cannot surface as such. 
-In other words, [bnɪk] is an impossible surface representation in English. 
-However, in \emph{The Sound Pattern of English}, \citet{SPE} introduce a third nonce word, [bznk], and claim that last form is in some sense even less English-like than [bnɪk].\footnote{
-    That wordlikeness judgements depend on language-specific knowledge is apparent given that [bznk] is not impossible in all languages: Imdlawn Tashlhiyt Berber permits whole words consisting of a stop-fricative-nasal-stop sequence (e.g., [tzmt] `it is stifling'; \citealt[112]{Dell1985}).
-    This suggests that wordlikeness judgements depend on language-specific knowledge.
-    It has more recently been suggested some contrasts between unattested clusters can be viewed as cross-linguistic implicational universals \citep[e.g.,][]{Berent2007a,Berent2008a,Berent2009,Berent2007b}, so that, for instance, the acceptance of [bznk] might imply acceptance of [bnɪk].
-    If this is correct, it undermines \citeauthor{SPE}'s position.}
-From this they too conclude that wordlikeness intuitions are gradient.
-
-\begin{quote}
-Hence, a real solution to the problem of ``admissibility'' will not simply define a tripartite categorization of occurring, accidental gap, and inadmissible, but will define the `degree of admissibility' of each potential lexical matrix in such a way as to\ldots{}make numerous other distinctions of this sort (\emph{SPE}:416--417)
-\end{quote}
-
-\noindent
-This brings the theory of wordlikeness in line with the view of syntactic grammaticality presented by foundational documents like \emph{The Logical Structure of Linguistic Theory} \citep{LSLT} and \emph{Aspects of the Theory of Syntax} \citep{ASPECTS}, which posits multiple degrees of ungrammaticality arising from different types of syntactic violations.
-\citet{BARRIERS} and \citet{Huang1982} propose further elaborations of this type of theory (see \citealt[43f.]{Schutze1996} for a critique).
-
-It has been recognized, however, that the naïve account of the [blɪk] and [bnɪk] cannot easily be extended to account for these ``numerous other distinctions''.
-
-\begin{quote}
 A defect of current grammatical accounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements.
+But when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values.
 \citep[371]{Shademan2006}
@@
 
-Scientific observations do not necessarily arrive at the appropriate granularity for analysis.
-Since the granularity of measurement and analysis is determined by prior hypotheses about the nature of the system under study, disputes are inevitable as scientific paradigms evolve. 
-This chapter argues that there are both theoretical and empirical reasons to doubt the implicit hypothesis linking scalar wordlikeness judgements and a gradient model of phonotactics. 
-First, intermediate judgements are characteristic of all gradient rating tasks, and therefore these judgements are irrelevant to the question of whether well-formedness is categorical or gradient.
-Secondly, simple baselines better account for gradient well-formedness judgements than current computational models of phonotactic knowledge, suggesting that the gradience observed in these tasks do not derive from grammatical mechanisms.
+If the presence of intermediate values in wordlikeness tasks is evidence for the gradient nature of phonotactic wellformedness, then it follows that wordlikeness intuitions should be measured and modeled with a high degree of granularity.
+For instance, this would be strong evidence against the naïve account of the [blɪk]-[bnɪk] contrast just alluded to, since it cannot easily be extended to account for ``numerous other distinctions''.
+However, this chapter argues that there are both theoretical and empirical reasons to doubt the implicit hypothesis linking scalar wordlikeness ratings and gradient wellformedness.
+First, intermediate ratings are characteristic of all gradient rating tasks, and therefore are irrelevant to the question of whether the internal phonotactic system is categorical or gradient.
+Secondly, simple baselines better account for gradient well-formedness judgements than current computational models of phonotactic knowledge, suggesting that the gradience observed in these tasks do not derive from known grammatical mechanisms.
 
 %are ``\ldots{}under the obligation to say why \emph{what it takes to be} data relevant to the confirmation of its theories \emph{are} data relevant to the confirmation of its theories'' \citep[200]{Fodor1981b}.
-%The naïve account of the [blɪk]/[bnɪk] contrast presented above is a special case of this
 
@@
 
-The aforementioned discussions of gradient aspects of wordlikeness judgements take for granted that intermediate judgements require a model of gradient grammaticality. 
-The general position that such data can be taken at face value is known in the cognitive sciences as \emph{naïve realism} \citep{Fodor1981a}, and is not uncontroversial. 
-With respect to wordlikeness, there are numerous reasons to question the wisdom of this position.
+The aforementioned discussions of gradient aspects of wordlikeness judgements takes for granted that intermediate ratings are the product of an internal system of gradient grammaticality.
+This view is itself an instance of what is known as \emph{common-sense} or \emph{naïve realism}; in the cognitive sciences, this often takes the form of the assumption that experimental data can be taken at face value, without mediation from other sources of information.
+However, there are several arguments for \emph{a priori} skepticism about the (a) linguistic abilities required for reporting gradient grammaticality judgements, (b) intermediate acceptability ratings as evidence for gradient grammaticality, and (c) the total lack of previous attempts to consider categorical models of wellformedness.
+
+\subsection{A model of gradient intuitions}
+
+Current research into gradient wellformedness is concerned with specifying a function from sound sequences to scalar judgements, and thus describes the wellformedness system at a level of some abstraction, corresponding roughly to what \citet{Marr1982} calls the ``computational'' level of description.
+This is only one part of any model of gradient grammaticality, however; further assumptions are needed to articulate the internal representations and algorithms by which this function computes.
+
+First, consider the architecture which is implied by any system of gradient grammaticality.
+It is essential that a system of gradient grammaticality have access to a relatively faithful representation of stimuli in a wellformedness task, and therefore it must be able to parse an enormous range of linguistic structures, including many which are not generated by the grammar itself; independent perception and production grammars may be necessary.
+A scalar value, representing wellformedness, must then be assigned to this parse.
+To report wellformedness, the speaker must transform this scalar value in accordance with the numerical scale chosen by the experimenter.
+
+Each step of this procedure merits scrutiny, however.
+First, speakers have difficulty perceiving \citep{Berent2007a,Brown1956,Dupoux1999,Kabak2007a} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
+Secondly, the computation of a scalar value serves no further purpose than to provide for gradient well-formedness judgements, so an objection might be made here on evolutionary grounds.
+It is quite mysterious why the human language endowment includes constraints on pronominal binding, for instance, but no one can doubt that these constraints are implicated in everyday language use; far more bizarre is the suggestion that the linguistic endowment includes mechanisms only implicated in certain experimental tasks.
+Next, speakers must be able to consciously access and report the magnitude of this value (it must be \emph{cognitively penetrable} in the sense of \citealt{Pylyshyn1984}), an ability which is limited in many other domains.
+Finally, there is some evidence that speakers do not (or perhaps, cannot) respect the numerical scales chosen by experimenters \citep{Sprouse2011}.
+
+It is informative to compare this baroque model to the architecture implied by a binary well-formedness judgement.
+When presented with a linguistic item in a judgement task, the grammar attempts to assign a parse. 
+Speakers then access whether or not parsing was successful.
+There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}.
+As priming of linguistic structures is thought to implicate shared representations in memory, this suggests different memory mechanisms for grammatical and ungrammatical linguistic objects.
+%This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
+Finally, the fact that requests for repetition and clarification are ubiquituous in spontaneous speech illustrates further that speakers are frequently aware when parsing has failed.
+Consequently, a great deal of evidence is needed to reject this simple model in favor of the gradient grammaticality architecture.
 
@@
 It should come as no surprise, then, that a highly salient aspect of a judgement task, the scale used for responses, also influences the results obtained.
-\citet*{Armstrong1983} argue that the use of many-valued scales may cause intermediate ratings, and should be interpreted as a task effect.
+\citet*{Armstrong1983} argue that rating tasks using many-valued scales may induce intermediate ratings as a task effect.
 
@@
 
-It might be said that these results reveal something about the representation of odd numbers, but no scientist of mathematical cognition has risen to the challenge of constructing a theory that might account for the fact that 447 is rated more odd than 3.
+It might be said that these results reveal something about the representation of odd numbers.
 \citeauthor{Armstrong1983} anticipate this objection.
@@
 
-\subsection{A model of gradient intuitions}
-
-It is possible to draw an analogy: can it be the case that [st] is a significantly ``better'' onset than [bl] (a prediction of the wordlikeness model proposed by \citealt{Albright2009a}, for instance), but ensure that both are treated the same with respect to syllabification?
-Current research on gradient well-formedness is concerned with specifying the component by which a scalar value is assigned to linguistic structures. 
-This is only one part of any model of gradient grammaticality, however.
-Further assumptions are necessary, as follows.
-When presented with a linguistic item (of whatever type) in a well-formedness task, the grammar assigns a parse.
-This grammar must be able to parse an enormous range of linguistic structures, including many which it cannot be permitted to generate; independent perception and production grammars may be necessary.
-A scalar value is then assigned to this parse. 
-Then, speakers consciously access this scalar value, then transform it in accordance with the numerical scale chosen by the experimenter.
-
-Each step of this procedure is dubious, however.
-First, speakers have difficulty perceiving \citep{Berent2007a,Brown1956,Dupoux1999,Kabak2007a} and producing \citep{Davidson2005,Davidson2006a,Davidson2006b,Davidson2010,GallagherInPress,Rose2007,Vitevitch1998,Vitevitch2005} phonotactically illicit non-words, suggesting that speakers' ability to faithfully parse illicit representations is at best quite limited.
-Secondly, the computation of a scalar value serves no further purpose than to provide for gradient well-formedness judgements; even if this is the result of probabilistic parsing \citep[e.g.,][]{Coleman1997},
-it need not be computed overtly.\footnote{
-    An objection might be made here on evolutionary grounds. 
-    It is still quite mysterious why the human language endowment includes constraints on pronominal binding, for instance, but these constraints are implicated in everyday language use.
-    Far more bizarre is the suggestion that the linguistic endowment includes mechanisms only used in certain experimental tasks.}
-Next, speakers must be able to consciously access and report the magnitude of this value (it must be \emph{cognitively penetrable} in the sense of \citealt{Pylyshyn1984}), an ability which is limited in many other domains.
-Finally, \citet{Sprouse2011} argues that speakers do not (or cannot) scale well-formedness values to satisfy the assumptions of magnitude estimation, a type of gradient rating task.
-%\citep{Bard1996}
-
-It is informative to compare this baroque architecture to a model of what is necessary to make a binary well-formedness judgement.
-When presented with a linguistic item in a judgement task, the grammar attempts to assign a parse.
-Speakers then access whether or not parsing was successful.
-There are reasons to think that parsing of ungrammatical structures does in fact result in a ``crash'': whereas syntactic priming increases the acceptability of grammatical structures \citep{Luka2005}, ungrammatical structures show no priming effects \citep{Sprouse2007b}.
-As priming of linguistic structures is thought to implicate shared representations in memory, ungrammatical structures may not be stored in linguistic memory. 
-%This cannot be reduced to the fact that ungrammatical structures fail to denote, since well-formed but non-existent structures do produce facilitory priming \citep[e.g.,][]{Longtin2005}. 
-The fact that requests for repetition and clarification are ubiquituous in spontaneous speech illustrates further that speakers are frequently aware when parsing has failed.
-Consequently, a great deal of evidence is needed to reject this simple model in favor of the gradient grammaticality architecture.
+No scientist has risen to the challenge of constructing a theory that might account for the fact that 447 is rated more odd than 3, and as \citeauthor{Armstrong1983} suggest, it is unclear whether such a theory could preserve more central facts about oddness.
+Here it is possible to draw an analogy to phonotactic theory.
+According to current orthodoxy, the wellformedness of a sequence is closely related to its type frequency (i.e., frequency in the lexicon).
+Is it then the case that [bl] is a significantly ``better'' onset than [kl], simply because the former is approximately twice as frequent, and if so, is it possible to also ensure that these sequences are treated the same with respect to syllabification?
 
@@
 
-\citet{Hayes2000} argues that it is ``uninsightful'' to attribute gradience to task effects, insofar as these effects call upon grammatical representations.
+\citet{Hayes2000} argues that it is ``uninsightful'' to attribute gradience to task effects, insofar as these effects implicate grammatical representations.
 
@@
 \noindent
-This is indisputable. 
-However, there have been no serious attempts to evaluate categorical and gradient models of wordlikeness on an equal footing.
+The logic of this implication is indisputable.
+However, there is little empirical support for the claims that absolute and gradient well-formedness derive from similar principles; indeed, there have been no prior attempts to evaluate categorical and gradient models of wordlikeness on an equal footing.
 In light of the complexities of gradient models, such an evaluation requires strong quantitative evidence for the superiority of gradient grammatical models.
 The evaluation below represents a first attempt to fill this gap.
+%FIXME clements and keyser
 
@@
 \citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables.
-Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words.
-Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, many of which are exceptionless, into probabilities.
-While this is consistent with their claim, that ``the ability to model gradient intuitions [are] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, this principle would preclude any attempt to test the hypotheses that underlies it.
+Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words (see \S\ref{ova}).
+Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of English phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, several of which are without exception, into probabilities.
+While this is consistent with their claim, that ``the ability to model gradient intuitions [is] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, little insight can be gained by annotating an exceptionless rule ``$p = 1.0$''.
+\citeauthor{Hayes2008a}'s principle precludes any attempt to test the hypotheses that underlies it, and therefore must be rejected.
 
+% FIXME make this work: 
 %(see \citealt{Sproat1993} and \citealt{Wrench2003} for articulatory investigations).
@@
 
-It is unknown whether the intermediate ratings in gradient wordlikeness tasks are reliably predicted by computational models that have been proposed. 
-If some model poorly accounts for these intermediate ratings, it may be the case that the model is improperly specified, or it may indicate that a considerable amount of variance in wordlikeness ratings is not a product of the grammatical system per se.
-It seems particularly plausible that speakers might differentiate, in a regular fashion, between different types of ``impossible'' words, as proposed by \citet{SPE}. 
-There are also claims that speakers distinguish between different types of ``possible'' words, so that, for instance, [stɪn] \emph{stin} is rated more English-like than [blɪn] \emph{blin} \citep{Albright2009a}.
-Alternatively, wordlikeness judgements could be effectively modeled with a gross contrast between possible and impossible words; even then, a gradient model might show a residual correlation.
-All of these possibilities are considered here.
+As \citet{Newmeyer2007} writes ``the idea that categoricity is not represented in the data itself is a truism.
+Whether distinctions of grammaticality (as opposed to acceptability) are binary is a difficult question.'' (398)
+The mere presence of gradience in judgements cannot falsify the claim of gradient grammaticality; another method is needed to evaluate this claim.
+As a first step towards a falsifiable theory of gradient wordlikeness, the remainder of this chapter considers intermediate ratings in gradient wordlikeness tasks are reliably predicted by the computational models that have been proposed.
+If a model is incapable of accounting for intermediate ratings, there are two possibilities: either the model itself is improperly specified and therefore at fault, or the inputs and outputs of the model are unrelated to the actual causes of the intermediate ratings, \emph{contra} \citet{Hayes2000}.
+
+It is plausible that speakers might differentiate, in a regular fashion, between different types of ``impossible'' words, and a gradient model should reliably predict the distinctions that speakers make.
+There are also claims that speakers distinguish between different types of ``possible'' words, so that, for instance, [stɪn] \emph{stin} is rated more English-like than [blɪn] \emph{blin} \citep[e.g.,][]{Albright2009a}, because the former onset is more frequent in the English lexicon.
+Even if wordlikeness judgements can be effectively modeled with a gross contrast between possible and impossible words, a gradient model might show a correlation with the residual ratings.
+All of these possibilities are considered below.
 
@@
 Two criteria were used to select these three studies.
-First, it was required that the stimuli be presented aurally so as to eliminate any possibility of orthographic effects \citep[e.g.,][]{Berent2001b,Berent2008b}.
+First, the stimuli must be presented aurally so as to eliminate any possibility of orthographic effects \citep[e.g.,][]{Berent2001b,Berent2008b}. 
 Secondly, the data must be sufficiently ``phonotactically diverse'': that is, it must include both items like \emph{blick} and \emph{bnick}. 
-This excludes studies like that of \citet{Bailey2001}, in which few if any items contain gross phonotactic violations of the type represented by \emph{bnick}, since this is likely to minimize the coverage of any grammatical model.
-The data is summarized in Table \ref{counts}.
+This excludes studies like that of \citet{Bailey2001}, in which few if any items contain gross phonotactic violations of the type represented by \emph{bnick}.
+In the absence of phonotactic violences, little variance in wordlikeness ratings can be attributed to phonotactic wellformedness, making it difficult to determine the coverage of gradient wellformedness models.
+The data used here is summarized in Table \ref{counts}.
 
@@
 
-\citet{Scholes1966} conducts several wordlikeness tasks with 
-students in 7th grade (approximately 12--13 years of age).
+\citet{Scholes1966} conducts several wordlikeness tasks with students in 7th grade (approximately 12--13 years of age).
 The data used here is his experiment 5, in which 33 speakers provide a ``yes'' or ``no'' as to whether each of the 63 items, presented aurally, are ``likely to be usable as a word of English''. 
@@
 \citet{Hayes2008a} evaluate their model using the Pearson (``product-moment'') $r$, a parametric correlation measure.
-\citet{Stevens1946} argues that statistics of this type are inappropriate for analysis of Likert scale data, like those used by \citet{Albright2007} and \citet{Albright2003b}.
-The reason is that the Pearson $r$ makes a \emph{linearity assumption}.
+It has long been argued \citep[e.g.,][]{Stevens1946} that parametric statistics are inappropriate for analysis of Likert scale data, like those used by \citet{Albright2007} and \citet{Albright2003b}, since the Pearson $r$ makes a \emph{linearity assumption}.
 That is, it assumes that nonce words rated ``1'' and  ``3'', for instance, are just as different as are those rated ``4'' and ``6''. 
@@
 However, it also has been claimed that $r$ is particularly robust to violations of the linearity assumption \citep[e.g.,][]{Havlicek1976}.
-Pearson $r$ is reported here, but no stance is taken on its appropriateness for this data.
-%\citet{Jamieson2004}
+Pearson $r$ is reported here, but this should not be taken to imply an endorsement of its use for Likert scale data.
+
 \citeauthor{Hayes2008a} also report Spearman $\rho$; this statistic requires only the weaker assumption of monotonicity, but it is difficult to give a simple interpretation to the coefficient.
-
-Two other non-parametric statistics the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ are much easier to interpret, as follows \citep{Noether1981}. 
+Two other non-parametric statistics, the Goodman-Kruskal $\gamma$ and the Kendall $\tau_b$ are much easier to interpret, as follows \citep{Noether1981}. 
 These statistics are computed by comparing every model score/wordlikeness rating pair to every other: a comparison is counted as \emph{concordant} if the greater of the two model scores is the one associated with the greater of the two wordlikeness ratings (that is, the model ranks these two nonce words in accordance with speakers' ratings), and as \emph{discordant} otherwise.
@@
 The simplest possible mechanism is chosen here: an onset or rime is identified as well-formed if it occurs with non-zero frequency in a representative sample, and is identified as ill-formed otherwise.
-The sample is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms, the latter thought to be particularly strongly correlated with behavioral measures \citep{Brysbaert2009}.
+This is not to imply that all unattested onsets or rimes should be regarded as ill-formed, or that all onsets or rimes with non-zero frequency in this data are well-formed.
+For instance, \citet{Albright2009a} judges [dɹɛsp] \emph{dresp} to be phonotactically well-formed, despite the total lack of [ɛsp] rimes in English; similar observations have been made concerning English onsets \citep[e.g.,][]{Cairns1972,Moreton2002}.
+
+The representative sample used to define the phonotactic baseline is derived from those entries of the CMU pronunciation dictionary which occur at least once per million words in the SUBTLEX-US frequency norms; these norms are known to be particularly strongly correlated with other behavioral measures \citep{Brysbaert2009}.
 These pronunciations are then syllabified, and individual syllables parsed into onset and rime, according to a process described in detail in Appendix \ref{syllabification}.
-This is not to imply that all unattested onsets or rimes should be regarded as ill-formed, or that all onsets or rimes with non-zero frequency in this data are well-formed.
-For instance, \citet{Albright2009a} judges [dɹɛsp] \emph{dresp} to be phonotactically well-formed, despite the total lack of [ɛsp] rimes in English.
-Similar observations have been made concerning English onsets \citep[e.g.,][]{Cairns1972,Moreton2002}.
-In the other direction, there are precedents for labeling certain attested words as phonotactically ``peripheral'' (e.g., \citealt{Myers1987}, \citealt{Borowsky1989}), as lexical exceptions to language-specific prosodic constraints, without labeling these words as absolutely ungrammatical.
-Neither of these dissociations between attestation and well-formedness are implemented here, however.
 
-In Figure \ref{boxplot}, wordlikeness ratings from the three studies are plotted according to this gross contrast.
-While there are a considerable number of outliers, there can be little doubt that the contrast is reflected in wordlikeness judgements in all three studies.
+Wordlikeness ratings from the three studies are plotted against this gross contrast in Figure \ref{boxplot}.
+While there are several outliers, there can be little doubt that gross phonotactic status accounts for a considerable amount of variance in wordlikeness judgements.
 
@@
 A wide variety of lexical similarity measures were considered, including a variant of the Generalized Neighborhood model \citep{Bailey2001}, PLD20 \citep{Suarez2011}, and a set of measures provided by the Irvine Phonotactic Calculator \citep{Vaden2009}.
-The most reliable measure is also the most venerable measure of lexical similarity: Coltheart's $N$ \citep{Coltheart1977}, which is defined as the number of words in some representative sample which can be changed into a target nonce word by a single insertion, deletion, or substitution of a phone.
+The measure best correlated with wellformedness judgements is also the most venerable measure of lexical similarity: Coltheart's $N$ \citep{Coltheart1977}, which is defined as the number of words in some representative sample which can be changed into a target nonce word by a single insertion, deletion, or substitution of a phoneme.
 \citet{Greenberg1964} find a correlation between wordlikeness ratings and a variant of this measure which only counts words differing by a single substitution.
-This is plotted against ratings from the three studies in Figure \ref{neighborhood}, where it can be seen that it accounts for much of the variance in ratings.
+This is plotted against ratings from the three studies in Figure \ref{neighborhood} with a superimposed local regression (LOESS; \citealt{Cleveland1988}) curve; neighborhood density accounts for much of the variance in ratings.
 
@@
 \includegraphics{neighborhood.pdf}
-\caption{Correlation between Coltheart's $N$ and item-averaged wordlikeness ratings}
+\caption{Correlation between Coltheart's $N$ and item-averaged wordlikeness ratings, with LOESS curve}
 \label{neighborhood}
@@
 However, since [bn] onsets are unattested in English, a neighbor of [bnɪk] must somehow modify this cluster: this leaves only \emph{brick} and \emph{nick}. 
-\citet{Bailey2001} and \citet{Frauenfelder1993} note that neighborhood density is also strongly correlated with measures like bigram probability, but it is argued elsewhere that phonotactic measures and neighborhood density have distinct effects \citep[e.g.,][]{Berent2003,Pitt1998b,Vitevitch1998,Vitevitch1999a}.
+\citet{Bailey2001} and \citet{Frauenfelder1993} note that neighborhood density is also strongly correlated with measures like bigram probability, but it has been argued elsewhere that phonotactic measures and neighborhood density have distinct effects \citep[e.g.,][]{Berent2003,Pitt1998b,Vitevitch1998,Vitevitch1999a}.
 
@@
 Faciliatory effects of bigram probabilities (i.e., shorter latencies) are reported for other nonce word tasks conducted with adults, including single-word shadowing \citep{Vitevitch1997,Vitevitch1998}, same/different judgements \citep{Lipinski2005,Luce2001,Vitevitch1999a,Vitevitch2005}, and lexical decision \citep{Pylkkanen2002a}.
-\citet{Albright2009a} applies them as a model of wordlikeness judgements.
-The bigram probability of a sequence $ijk$, for instance, is:
+\citet{Albright2009a} applies bigram probability as a model of wordlikeness judgements. 
+The bigram probability of a sequence $ijk$, for instance, is defined as
 
@@
 Unfortunately, the latter model is not described in sufficient detail to allow it to be implemented directly, and there is no publicly available implementation.
-However, \citeauthor{Albright2009a}'s evaluation, which includes the \citet{Scholes1966} and \citet{Albright2003b} data, finds an advantage for segmental bigrams.
-In implementing this model, it was found that a slight improvement could be made by preventing any phone-to-phone transition from having zero probability.
+However, \citeauthor{Albright2009a}'s evaluation, which includes the \citet{Scholes1966} and \citet{Albright2003b} data, finds an advantage for the segment-based model.
+In implementing this model, it was found that a slight improvement could be made by preventing any phoneme-to-phoneme transition from having zero probability. 
 This is accomplished by adding 1 to the count of every transition, a technique used in natural language processing and known as Laplace, or ``add one'', smoothing.
-As can be seen in Table \ref{bigramcomparison}, this results in a slight increase in the correlation between the scores from this model and wellformedness ratings.
+As can be seen in Table \ref{bigramcomparison}
+this results in a slight increase in the correlation between the scores from this model and wellformedness ratings 
 This smoothed segmental bigram score is adopted below.
-In Figure \ref{bigram}, it is plotted against wordlikeness ratings from the three studies.
+In Figure \ref{bigram}, it is plotted against wordlikeness ratings from three studies.
 
@@
 \includegraphics{bigram.pdf}
-\caption{Correlation between smoothed segmental bigram score and item-averaged wordlikeness ratings}
+\caption{Correlation between smoothed segmental bigram score and item-averaged wordlikeness ratings, with LOESS curve}
 \label{bigram}
@@
 \citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints (e.g., \citealt{Goldwater2003}, \citealt{Jager2007}). 
-\citeauthor{Hayes2008a} use a complex method to evaluate their model.
-First, they extract onset sequences from the CMU pronunciation dictionary, and use these to train the model.
+\citeauthor{Hayes2008a} use a complex method to evaluate their model. First, they extract onset sequences from the CMU pronunciation dictionary, and use these to train the model. 
 The model is then used to score the onsets of the \citet{Scholes1966} nonce words. 
-Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting correlation.\footnote{This is contrary to standard practices in natural language processing, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter); when this is the case, there is reason to suspect the parameter values will not generalize to new data.
+Then they compute a parameter for transforming their model scores so as to maximize the correlation between these transformed scores and wordlikeness ratings, then report the resulting correlation.\footnote{
+    This is contrary to standard practices in natural language processing, in that the data used for evaluation is also used to fit the model (namely, the transformation's parameter); when this is the case, there is reason to suspect the parameter values will not generalize to new data.
     No transformation is used here; this only has an effect on the Pearson $r$ coefficient, since they use a transformation that preserves monotonicity.}
-However, \citet{Albright2009a} finds that the maximum entropy model, training and testing only on onsets, does not generalize to the \citet{Albright2003b} data, presumably because of the considerable diversity of rimes in the latter data.
+\citet{Albright2009a} reports that the maximum entropy model, training and testing only on onsets, performs well on the \citet{Scholes1966} data, but does not generalize well to the \citet{Albright2003b} sample.
 Consequently, the model was trained to score whole words, not just onsets, using the subset of the CMU dictionary described above.
@@
 Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original paper is attempted: both their implementation and phonological feature specifications are used here.
-Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in \S\ref{syllabification} and a novel feature [$\pm$\textsc{Coda}] is added to allow the model to distinguish coda and onset consonants.
+Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in Appendix \ref{syllabification}, and a novel feature [$\pm$\textsc{Coda}] is added to allow the model to distinguish coda and onset consonants.
 Also, following \citeauthor{Hayes2008a}, constraints are limited to those spanning as many as three segments and an ``accuracy schedule'' of [$.001$, $.01$, $.1$, $.2$, $.3$] is used.
 Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citeauthor{Hayes2008a}.
-The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to a large variety of nonce words, though many of these words appear to have a low rating.
-Apparently, this model is still not robust enough for extracting phonotactic generalizations from whole words.
+The resulting scores are plotted against wordlikeness ratings in Figure \ref{maxent}; it can be seen that the model assigns the highest possible score to a large variety of nonce words, though many words with a low rating receive the highest MaxEnt probability score.
+It appears that this model is still not robust enough to reliably extracting phonotactic generalizations from monosyllabic words.
 
@@
 The only reliable positive correlation is present among the ``valid'' items as rated by the smoothed segmental bigram model. 
-This model is somewhat capable of accounting for contrasts between different ``possible'' nonce words: for instance, it favors [plin] \emph{plean} over [brɛlθ] \emph{brelth} just as subjects in the \citet{Albright2007} study do.
+This model is somewhat capable of accounting for contrasts between different ``possible'' nonce words: for instance, it favors [plin] \emph{plean} over [brɛlθ] \emph{brelth} just as subjects in the \citet{Albright2007} study do; this can also be seen in the top three panels of Figure \ref{bigramresid}.
 Within the set of ``invalid'' items, however, neither grammatical model reliably distinguishes among items; both models, for instance, rate [ptʌs] \emph{ptus} more well-formed than [bnʌs] \emph{bnus}, but speakers have the opposite preference.
@@
 \end{tabular}
-\caption{Kendall $\tau_{b}$ correlation between item-averaged wordlikeness ratings (rated the gross phonotactic status baseline) and model scores}
+\caption{Kendall $\tau_{b}$ correlation between model scores and item-averaged wordlikeness ratings, sorted according to gross phonotactic status}
 \label{resid}
 \end{table}
+
+\begin{figure}[t]
+\centering
+\includegraphics{bigramresid.pdf}
+\caption{Correlation between smoothed segmental bigram score and item-averaged wordlikeness ratings, sorted according to the gross phonotactic status, with LOESS curve}
+\label{bigramresid}
+\end{figure}
 
@@
 A serious limitation of this evaluation is the primitive nature of the gross phonotactic status baseline. 
-It does not allow for any way to state constraints on onset-nucleus sequences, which have been proposed for some languages (e.g., \citealt{Kirby2007}),\footnote{\citet{Kessler1997}, however, argue there are no clear restrictions on English onset-nuclei pairs.} or constraints spanning whole syllables, which have been proposed for English \citep[e.g.,][]{Berkley1994b,Berkley1994a,Coetzee2008b,Fudge1969}.
-Furthermore, the gross phonotactic baseline does not have any mechanism for generalizing the wellformedness of [ɛsp] rimes from \emph{clasp}, \emph{lisp}, and other rimes consisting of a lax vowel followed by [sp] found in English.
-\citet{Borowsky1989}, for instance, proposes a theory of possible rimes in English which makes the correct prediction regarding [ɛsp].
-This is not embedded in an acquisitional model, but models of syllable type acquisition have been proposed \citep[e.g.,][]{Fikkert1994,Levelt2000,Pan2003,Pan2004}.
+It does not allow for any way to state constraints on onset-nucleus sequences, which have been proposed for some languages (e.g., \citealt{Kirby2007} on Cantonese), or constraints spanning whole syllables \citep[e.g.,][]{Berkley1994b,Berkley1994a,Clements1983,Coetzee2008b,Fudge1969}.\footnote{
+    It is disputed whether English in particular exhibits onset-nucleus restrictions.
+    \citet{Clements1983} claim that ``cooccurrence restrictions holding between the nucleus and preceding elements of the syllable appear to be just as common as cooccurrence restrictions holding between the nucleus and following elements'' (20), but admit that at least some of these generalizations may represent accidental gaps.
+    However, \citet{Kessler1997}, argue there are no clear restrictions on English onset-nuclei pairs.}
+Furthermore, the gross phonotactic baseline does not have any mechanism for generalizing the wellformedness of [ɛsp] rimes from \emph{clasp}, \emph{lisp}, and other rimes consisting of a lax vowel followed by [sp] found in English, but \citet{Borowsky1989}, for instance, proposes a theory of possible rimes in English which makes the correct prediction regarding [ɛsp].
+This is not embedded in an acquisitional model, but many models of syllable type acquisition have been proposed \citep[e.g.,][]{Fikkert1994,Levelt2000,Pan2003,Pan2004}.
+As observed by \citet{Smith1973} in a careful study of a single child acquiring English, children's productions are at first highly restricted but progress systematically to stages with fewer and fewer restrictions.
+Assuming productive competence is an appropriate measure of syllable acqusition, this suggests that syllable types are acquired like many other linguistic phenomena in that the child progresses from subset to superset.
+The difficulty here is that the typology of syllables must be delineated so that, for instance, the robust presence of [æsp] and [ɪsp] implies acceptance of [ɛsp].
 
-The gross phonotactic baseline could also be extended so as to include additional levels of wellformedness.
-While the bigram and MaxEnt models do not appear to be able to reliably distinguish intermediate levels of well-formedness, it might be desirable to encode the intuition that, for example, [ʒlɪk] \emph{zhlick}, is more English-like than [bnɪk], though both have unattested onsets.
-It is also possible to imagine that phonotactic violations would have a cumulative effect on well-formedness; for instance, a nonce word with an unattested onset and an unattested rime, like [tsɪlm], might be less English-like than either [tsɪl] or [sɪlm].
-This prediction is made by the bigram and MaxEnt models, among others \citep[e.g.,][]{Albright2008,Anttila1997}, but could easily be incorporated into a simple baseline by counting the number of violations.
-However, there is not yet any convincing evidence for cumulativity effects in wordlikeness tasks.
+The gross phonotactic baseline could also be extended so as to recognize more than two levels of wellformedness, without introducing the infinite amount of contrast implied by fully gradient models.
+While the bigram and MaxEnt models do not appear to be able to reliably distinguish intermediate levels of well-formedness, it might be desirable to encode the intuition that, for example, [ʒlɪk] \emph{zhlick}, is more English-like than [bnɪk], though both have unattested onsets \citep[e.g.,][50f.]{Clements1983}.
+It is also possible to imagine that phonotactic violations would have a cumulative effect on well-formedness.
+For instance, a nonce word with an unattested onset and an unattested rime, like [tsɪlm], might be less English-like than either [tsɪl] or [sɪlm], an ability that could easily be extended to the gross phonotactic baseline.
+Cumulativity effects are predicted by the bigram and MaxEnt models, among others \citep[e.g.,][]{Albright2008,Anttila1997}, but could easily be incorporated into a simple baseline by counting the number of violations.
+However, as of yet there is no convincing evidence for cumulativity effects in wordlikeness tasks, and the stimuli used here are not suited to test this hypothesis.
 
@@
 While it remains an open question whether future gradient models will account for intermediate judgements, the current evidence suggests that gradient grammaticality is not crucial for modeling gradient wordlikeness judgements.
+This does not imply that wordlikeness judgements collected using Likert scales or magnitude estimation are tainted; \citet{SprouseSubmitted} argue that gradient wellformedness measures are better able to detect syntactic violations thought to be categorical than are binary judgements, and it seems likely this result would also hold for wordlikeness tasks.
+However, intermediate ratings can no longer be taken at face value.
+
+%\citep{Bard1996}
+%FIXME short discussion of the utility of gradient judgements in syntax
+%\citep[e.g.,][]{Hale2003c,Hale2006,Levy2008,Roark2009}
+%\citet{Pereira2000}
 

### 2013-01-28 3.tex
Revisions! Lots of 'em.

@@
 
-It has long been suggested that statistical criteria could distinguish between \emph{accidental phonotactic gaps}, those which could arise without any antecedent cause, and those gaps which are \emph{structural} in nature \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.
+It has long been speculated that statistical criteria could distinguish between \emph{accidental phonotactic gaps}, those which could arise without any antecedent cause, and those gaps which are \emph{structural} in nature \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.\footnote{
+    Some of this early work is, however, critical of any attempt to establish a firm distinction between structural and accidental gap.
+    For instance, \citet[3]{Fischer-Jorgensen1952} writes that ``it is theoretically impossible to fix a non-arbitrary borderline between law and accident.''}
 Following seminal work by \citet{Mester1988} and \citet{McCarthy1988}, it is now commonly assumed that the phonotactic patterns in lexical entries ``directly determine the mental representation of the phonotactic constraints'' \citep[180]{Frisch2004} acquired by speakers, and therefore phonotactic constraints, whether categorical or gradient, can be directly inferred from statistical analysis of the lexicon.
 There is by now an enormous amount of research that adopts this assumption.
-Semitic languages, especially Arabic \citep{Coetzee2008a,Frisch2004,McCarthy1988,Pierrehumbert1993} but also Berber \citep{Elmedlaoui1995} and Tigrinya \citep{Buckley1997}, and Austronesian languages, including Javanese \citep{GraffInPress,Mester1988}, Muna \citep{Anttila2008a,Coetzee2008a}, and Samoan \citep{AldereteInPress}, have been of particular interest to linguists adopting this hypothesis; Aymara \citep{GraffInPress}, Cantonese \citep{Yip1989}, English \citep{Berkley1994a,Berkley1994b,Berkley2000,Coetzee2008b,Davis1989,Martin2007,Martin2011}, Gitksan \citep{Brown2010}, Hungarian \citep{Grimes2010}, Japanese \citep{Kawahara2006}, Ju|'hoansi \citep{Kinney2005}, Navajo \citep{Martin2007,Martin2011}, Ofo \citep[38f.]{MacEachern1999}, Russian \citep{Padgett1991,Padgett1992}, Shona \citep{Hayes2008a}, and Wargamay \citep{Hayes2008a} have also received a statistical treatment.
+Afroasiatic languages, especially Arabic \citep{Coetzee2008a,Frisch2004,McCarthy1988,Pierrehumbert1993} but also Berber \citep{Elmedlaoui1995} and Tigrinya \citep{Buckley1997}, and Austronesian languages, including Javanese \citep{GraffInPress,Mester1988}, Muna \citep{Anttila2008a,Coetzee2008a}, and Samoan \citep{AldereteInPress}, have been of particular interest to linguists adopting this hypothesis; Aymara \citep{GraffInPress}, Cantonese \citep{Yip1989}, English \citep{Berkley1994a,Berkley1994b,Berkley2000,Coetzee2008b,Davis1989,Martin2007,Martin2011}, Gitksan \citep{Brown2010}, Hungarian \citep{Grimes2010}, Japanese \citep{Kawahara2006}, Ju|'hoansi \citep{Kinney2005}, Navajo \citep{Martin2007,Martin2011}, Ofo \citep[38f.]{MacEachern1999}, Russian \citep{Padgett1991,Padgett1992}, Shona \citep{Hayes2008a}, and Wargamay \citep{Hayes2008a} have also received a statistical treatment.
 \citet{Pozdniakov2007} analyze lexical tendencies in a diverse sample of 30 languages.
@@
 Nothing demands that the antecedent cause of a statistically reliable lexical tendency be grammatical. 
-There are reasons to suspect that many static phonotactic constraints identified in this manner have no synchronic reality.
+On the contrary: there are reasons to suspect that many static phonotactic constraints identified in this manner have no synchronic reality.
 
 By hypothesis, the phonological component may impose constraints on sound segments and sequence on underlying representations.
-Beyond this, 
-a principled null hypothesis (implied by the principle of \textsc{No Static Phonotactics}) is that there is no synchronic, grammatical explanation for underlying representations a language chooses to instantiate.
+Beyond this, a principled null hypothesis (implied by the principle of \textsc{No Static Phonotactics}) is that there is no synchronic, grammatical explanation for underlying representations a language chooses to instantiate.
 It is a practical necessity that numerous well-formed underlying representations will be uninstantiated: the lexicon is finite but there are an infinitude of possible URs.
@@
 However, subsequent sound changes \citep[e.g.,][]{Baldi1994,Safarewicz1932}, particularly the degemination of Old Latin \emph{ss} after diphthongs and long monophthongs (e.g., \emph{caussa} $>$ \emph{causa} `cause'), introduce numerous exceptions to \textsc{Rhotacism}.
-In Classical Latin, intervocalic \emph{s} is found root-internally (\emph{asellus} `donkey', \emph{casa} `hut'), in environments derived by inflectional suffixes (\emph{uās}-\emph{uāsis} `vase', \emph{uisēre} `to view'), prefixation (\emph{dēsecāre} `to cut off'; cf.~\emph{dē} `from', \emph{secāre} `to cut'), compounding (\emph{olusātrum} `parsnip'; cf.~\emph{olus} `vegetable', \emph{ātrum} `black'), and denominal adjective formation (\emph{uentōsus} `windy'; cf.~\emph{uentus} `wind'), and is tolerated in nativized loanwords from Celtic (\emph{omāsum} `tripe'), Germanic (\emph{glaesum} `amber'), and Greek (\emph{basis} `pedestal').
+In Classical Latin, intervocalic \emph{s} is found root-internally (\emph{asellus} `donkey', \emph{casa} `hut'), in environments derived by inflectional suffixes (\emph{uāsis} `vase' \textsc{gen.sg.}, \emph{uisēre} `to view'), prefixation (\emph{dēsecāre} `to cut off'; cf.~\emph{dē} `from', \emph{secāre} `to cut'), compounding (\emph{olusātrum} `parsnip'; cf.~\emph{olus} `vegetable', \emph{ātrum} `black'), and denominal adjective formation (\emph{uentōsus} `windy'; cf.~\emph{uentus} `wind'), and is tolerated in nativized loanwords from Celtic (\emph{omāsum} `tripe'), Germanic (\emph{glaesum} `amber'), and Greek (\emph{basis} `pedestal').
 These facts lead Saussure to conclude that \textsc{Rhotacism} is no longer ``inhérente à la nature de la langue'' (202).
@@
 
-% FIXME mention Fudge's analysis of short-a before voiceless fricatives
-% \citet{Trager1930,Trager1934,Trager1940},
-% \citet{Cohen1970}
-% \citet{Ferguson1975}
-% \citet{PLC1} 429f.
-% \citet{Fudge1987}
-% \citet{AOE} I.203f., 515f. for an attempt to write a rule
-% \citet{Payne1976,Payne1980} who finds that whereas speakers from out of state acquire 5 vocalic sound changes, only one out of 34 interviewees acquired the core pattern
-% Fudge(!)
+In other cases, it is possible to observe once robust phonotactic tendencies induced by sound change become increasingly moribund in real time.
+Perhaps the most famous example concerns the distribution of short \emph{a} in English.
+Early Modern English short \emph{a} underwent irregular lengthening and tensing before voiceless fricatives /f, θ, s/ and nasals /m, n/ \citep[I.203f.]{AOE}, introducing the first traces of the complex pattern now found in many American dialects (as well as in various British dialects, e.g., \citealp[74f.]{Jones1964}).
+A great deal of variation in Mid-Atlantic speech can be summarized as follows: the low front vowel is lax except before tautosyllabic /f, θ, s, m, n/ (and perhaps other consonants), where the low front vowel is tense (and perhaps raised), with various exceptions in both directions \citep[e.g.,][]{Cohen1970,Ferguson1975,Labov1981,Trager1930,Trager1934,Trager1940}.
+\citet{Cohen1970}, \citet{Labov1981}, and \citet{Trager1940} all argue that this tendency is no longer to be attributed to synchronic allophony, but to a largely-complete lexical/phonemic split.
+\citeauthor{Labov1981} presents three arguments for this analysis.
+First, there are minimal pairs (tense modal \emph{can} vs.~lax noun/verb \emph{can}) and extensive individual variation.
+Secondly, \citet{Payne1980} finds that children born out-of-state are unable to acquire the short \emph{a} pattern of the Philadelphia metropolitian area in later childhood, though they participate fully in other sound changes specific to the area.
+Finally, Mid-Atlantic speakers have a greater ability to discriminate tense and lax variants of short \emph{a} than speakers of dialects with allophonic short \emph{a} tensing/laxing.
+In summary, the historical facts that account for this distribution no longer play a clear role in the synchronic phonology.
 
-Other cases show that constraints identified by statistical analysis are inert: that is, there is no evidence (of any of type described in \S\ref{s:espt} above) that they are internalized by speakers.
+Other cases show that many constraints identified by statistical analysis are inert: that is, there is no evidence that they are internalized by speakers.
 For instance, most instances of Modern English [ʃ] derive from Old English [sk] (e.g., \emph{fisc} `fish') via sound change.
-Since Old English does not permit long vowels before complex codas, compared to similar segments like [s], [ʃ] is still rarely preceded by long vowels in word-final syllables.\footnote{
-    This sample is drawn from the CMU pronunciation dictionary, filtered by excluding words with a token frequency of less than 1 per million words in the SUBTLEX-US frequency norms.
-    Less restrictive samples give similar results.
-    The 9 words ending with a long vowel-[ʃ] sequence are \emph{douche}, \emph{leash}, \emph{gosh}, \emph{josh}, \emph{posh}, \emph{squash}, \emph{unleash}, \emph{wash}, and \emph{woosh}.}
+Since Old English does not permit long vowels before complex codas, compared to similar segments like [s], [ʃ] is still rarely preceded by long vowels in word-final syllables (Table \ref{ssh}).\footnote{
+    This sample is drawn from the CMU pronunciation dictionary, filtered by excluding words with a token frequency of less than 1 per million words in the SUBTLEX-US frequency norms; similar results can be obtained with less restrictive samples.
+    The four words ending in a long vowel-[ʃ] sequence are \emph{douche}, \emph{leash}, \emph{unleash}, and \emph{woosh}.}
 As can be seen, long vowels are twice as common before [s] as before [ʃ], a significant generalization according to the Fisher exact test.
 Similarly, \citet{HayesInPress} report that this constraint is discovered by the \citet{Hayes2008a} phonotactic learning model, which uses a related statistical criterion to identify constraints.
-Despite this, \citet{Iverson2005} label the constraint on long vowels before [ʃ] as ``phonologically accidental'', as a millennium of coinages (e.g., \emph{posh}) and loanwords (e.g., \emph{douche}) disregard this generalization.
-\citet{HayesInPress} find that a variant of this restriction has little to no effect on wordlikeness judgements.
+Despite this, \citet{Iverson2005} label the constraint on long vowels before [ʃ] as ``phonologically accidental'', as a millennium of coinages (e.g., 
+`affective'' \emph{woosh}) and loanwords (e.g., \emph{douche}) disregard this generalization.
+\citet{HayesInPress} find that a variant of this restriction has little or no effect on wordlikeness judgements.
 
@@
 \toprule
-          & \{ɪ, ɛ, æ, ʌ, ʊ\}\gap{}\# & \{i, e, ɑ, ɔ, u\}\gap{}\# & \% long & $p$-value \\
+          & \{ɪ, ʊ\}\gap{}\# & \{i, u\}\gap{}\# & \% long & $p$-value \\
 \midrule
-\gap{}ʃ\# & 78                & 9                 & 8\%      & \multirow{2}{*}{.026} \\
-\gap{}s\# & 410               & 107               & 16\%     & \\
+\gap{}ʃ\# & 46               & 4                & 8\%     & \multirow{2}{*}{1.67\e{-06}} \\
+\gap{}s\# & 40               & 36               & 53\%    & \\
 \bottomrule
 \end{tabular}
-\caption{Type frequencies of short and long vowels before word-final [ʃ] and [s]}
+\caption{Type frequencies of high vowels before word-final [ʃ] and [s]}
 \label{ssh}
@@
 To account for the cases like the one above, \citet{HayesInPress} propose that speakers are biased in favor of ``natural generalizations'' in probabilistic phonotactic learning.\footnote{
-    \citeauthor{HayesInPress} do not provide an operational definition of ``naturalness'', so it is difficult to evaluate their specific results or to determine whether they apply the classification of ``natural'' and ''unnatural'' is applied in an ad hoc (or post hoc) fashion.}
+    \citeauthor{HayesInPress} do not provide an operational definition of ``naturalness'', so it is difficult to evaluate their specific results despite their relevance to the argument at hand.}
 However, this chapter argues that this overgeneration gives the lie to the broader assumption that phonotactics are extracted from patterns in the lexicon.
@@
 In this section, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence bearing on speakers' knowlege of the restrictions.
-The following feature specification for the eight vowels of Turkish is assumed throughout.
+The following feature specification for the eight vowels of Turkish is assumed throughout \citep[e.g.,][298]{Dresher2009}.\footnote{\citeauthor{Dresher2009} in fact deploys a feature labeled \textsc{Labial} rather than \textsc{Rnd}.}
 
@@
 
-Two less familiar notations are used in this chapter.
+Two less common notations are used in this chapter.
 First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules.
-Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{A74}: chap.~9, \citealt{GormanInPressc}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{KK77}:189f., \citeyear{KK79}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
+Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{A74}: chap.~9, \citealt{GormanInPressc}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{KK77}:189f., \citeyear{KK79}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.)
+Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.
 
@@
 A structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].
-This is more restrictive than Greek-letter notation, in that it prevents the value of one feature being applied to different feature; \citet{Odden2012} argues that the two counterexamples against this restriction presented in \emph{SPE} (352-353) are not probative.
+This is more restrictive than Greek-letter notation, in that it prevents the value of one feature being applied to different feature.\footnote{
+    \citet{Odden2012} argues that the two counterexamples against this restriction presented in \emph{SPE} (352-353) are not probative.}
 
@@
 \citet[212]{A74} and \citet{Iverson1978} argue that suffix harmony in disharmonic roots found in (\ref{turkexcept}a) requires the rule governing suffix harmony alternations to be distinguished from a sequence structure constraint governing root harmony.
-Since the rule and sequence structure constraint are otherwise identical, this constitutes a ``duplication''. 
-However, the theory of exceptionality proposed in \emph{SPE} can account for these facts without duplication \citep[197f.]{Zonneveld1978}.\footnote{
+Since the rule and sequence structure constraint are otherwise identical, this constitutes a ``duplication'' (in the sense of \citealt{KK77}:136f.): harmony is stated at two points in the grammar.
+
+Another possiblity, however, is to understand root disharmony as a type of lexical exceptionality.
+Under such an analysis, suffix harmony in disharmonic roots is entirely consistent with the theory of exceptionality proposed in \emph{SPE} \citep[197f.]{Zonneveld1978}.\footnote{
     \citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process.
-    \citet[171f.]{Howard1972} correctly observes that this duplication is also unnecessary.}
-
+    \citet[171f.]{Howard1972} suggests an analysis of Finnish that is quite similar to the one proposed for Turkish below.}
 \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention.
 A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}]. 
-In other words, no representation is ever truly an exception; rather, some underlying representations have non-default features which do not match the extended structural description of $R$, which requires that the target be [$+R$].
+In other words, no representation is ever truly an exception to a rule; rather, some underlying representations have non-default features which do not match the extended structural description of $R$, which requires that the target be [$+R$].
 If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.
+Under this account, root and suffix harmony are both derived from \textsc{Backness Harmony}, but they do not have the same ontology: suffix harmony is direct result of phonological rule application, whereas the tendency for harmonic roots arises from a dispreference for lexical exceptionality.
+Under this analysis, there is no need to state a sequence structure constraint producing root harmony.
 
-Under this account, root and suffix harmony are derived from \textsc{Backness Harmony}, but they do not have the same ontology: suffix harmony is direct result of phonological rule application, whereas any ``effects'' of the  generalization concerning harmonic roots arises from a dispreference for lexical exceptionality.
-
-\citeauthor{A74} also notes that roots which fail to undergo suffix harmony, like (\ref{turkexcept}b), may themselves be harmonic. He takes this to be evidence for the necessity of duplication:
+\citeauthor{A74} also notes that a small number of roots which fail to undergo suffix harmony, like (\ref{turkexcept}b), may themselves be harmonic.
+He takes this to be evidence for the necessity of duplication:
 
@@
 \noindent
-Under the assumptions so far, there is no reason to think that \emph{kabahat} is [$+$\textsc{Backness Harmony}], however: even roots with all back vowels need not undergo this rule.
-Whatever the ultimate analysis of those few words which fail to show suffix harmony, nothing depends on whether or not the roots in question are harmonic.
+Under the assumptions so far, however, there is no reason to assume that \emph{kabahat} (or any other root which fails to trigger suffix harmony) is [$+$\textsc{Backness Harmony}].
+The tendency for root harmony holds of the Turkish lexicon as a whole, but there is no evidence that it holds of those few roots which select exceptional suffixes; indeed, there are disharmonic roots which select disharmonic suffixes (e.g., \emph{etol} `fur stole', nom.pl.~\emph{etoller}).
 
-It is necessary to dispense with an alternative analysis proposed by \citet{Clements1982} and \citet{Inkelas1997}. 
-This analysis has desirable properties, but it subtlely reintroduces the duplication of phonological process and sequence structure constraint. 
-Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.
-\emph{kul} `servant', \emph{kepek} `bran' vs.
-\emph{kapak} `lid'), whereas backness of vowels in non-initial syllables is predictable in harmonic roots (there are no prefixes in Turkish).
-\citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots and of certain exceptional suffixes are fully specified.
+This analysis compares favorably to an alternative proposed by \citet{Clements1982} and \citet{Inkelas1997} which has other desirable properties but
+which makes no predictions about root harmony.
+Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.~\emph{kul} `servant', \emph{kepek} `bran' vs.~\emph{kapak} `lid'), whereas backness of vowels in non-initial syllables is predictable in harmonic roots (note that there are no prefixes in Turkish).
+\citeauthor{Clements1982} and \citeauthor{Inkelas1997} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots and of certain exceptional suffixes are fully specified.
 This is schematized below.
@@
 One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, perhaps by structure preservation condition \citep{Kiparsky1985}.
-However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{
+However, such a condition makes it impossible to relate \textsc{Backness Harmony} to the tendency for harmonic roots, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{
     On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default.
@@
     However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.}
-    The underspecification analysis is rejected here on these grounds.
 
@@
     If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it. Several other ``external'' facts suggest that this is not the case. 
-    The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games, or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.\footnote{
+    The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games, or particular psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.\footnote{
     Thanks to Bert Vaux and Kie Zuraw for bringing these studies to my attention.}
@@
     The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{
-    A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}
+    A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated language game in Finnish \citep{Campbell1986}.}
 
@@
 \noindent
-In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant. 
-    On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{
-    There is reason to suspect that reduplicants bear the lexical diacritics of the roots they are derived from.
-    In Kinande, verb reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, requires a bisyllabic reduplicant. 
-    This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots contain three copies of the root so as to satisfy this requirement, and reduplicated forms of many trisyllabic verbs are ineffable.
+\citeauthor{Harrison2001} propose that those roots which fail to reharmonize are prespecified for backness throughout; this also requires that reharmonization is subject to a non-derived environment condition.
+In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant; the lack of reharmonization in the reduplicants of disharmonic roots requires that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{
+    There are other cases which suggest that lexical diacritics are copied in reduplication.
+    In Kinande, verbal reduplication requires a bisyllabic reduplicant \citep{Downing2000,Mutaka1990}.
+    This requirement has synchronic force in the grammar, since ``reduplicated'' monosyllabic roots in fact contain three copies of the root (so that the reduplicant is bisyllabic); furthermore, reduplicated forms of many trisyllabic verbs are ineffable.
     However, a few trisyllabic verbs exceptionally show full reduplication.
-    This is a lexical property, but presence or absense of this exceptionalty feature is only apparent in the form of the reduplicant.}
-    Under this definition, root harmony is not static, and so this is not an exception to the claim that static phonotactic generalizations are not preserved by operations like reduplication and truncation \citep{Silverman2000}.
+    This is a lexical property, but presence or absence of the exceptionalty feature that permits non-bisyllabic reduplicants only surfaces on the reduplicant itself.}
 
@@
 The Turkish word-spotting experiment is adapted for infants by \citet{Kampen2008}.
-In this study, 9-month-old infants are familiarized with recordings of harmonic, disyllabic nonce words presented in isolation.
-At test time, the infants listen to the disyllabic nonce words in isolation using the head turn preference paradigm. 
+In this study, 9-month-old infants are familiarized with recordings of harmonic, disyllabic nonce words presented in isolation and tested with the head turn preference paradigm.
 Infants acquiring Turkish listen longer to nonce words preceded by a disharmonic juncture during familiarization (e.g., \emph{lo-NETIS}), whereas infants acquiring German, a language which lacks vowel harmony, do not exhibit this preference.
@@
 
-If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only [$+$\textsc{Rnd}] or only [$-$\textsc{Rnd}] [$+$\textsc{High}] vowels.
+If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only round or unround high vowels.
 In concert with \textsc{Backness Harmony}, \textsc{Roundness Harmony} also triggers alternations which account for the shape of the dative singular (dat.sg.) and genitive singular (gen.sg.) suffixes, among others.suffixes, among others.
-As was the case for \textsc{Backness Harmony}, disharmonic roots are found, and the final vowel of disharmonic roots triggers suffix harmony.
+As was the case for \textsc{Backness Harmony}, disharmonic roots occur, but generally trigger suffix harmony.
 
@@
 of \emph{sap} `stalk is \emph{sapın} rather than *\emph{sapun} that would be predicted if \textsc{Labial Attraction} triggered alternations.
-\citeauthor{Lees1966b} and \citet{Zimmer1969} cite roots which do not conform to \textsc{Labial Attraction} (e.g., \emph{tavır} `mode') but agree that they are surprisingly rare.
-Others \citep[e.g.,][]{Clements1982} dispute this.
+\citeauthor{Lees1966b} and \citet{Zimmer1969} cite roots which do not conform to \textsc{Labial Attraction} (e.g., \emph{tavır} `mode') but agree that they are surprisingly rare; so as to dispute the claim that such exceptions are rare, \citet{Clements1982} presents many additional examples of exceptions.
 
@@
 
-Counts were computed by regular expression matching on a 9,601-root subset of the TELL database which consists of roots which show no surface variation in any inflected form.
+Counts are computed by regular expression matching on a 9,601-root subset of the TELL database, consisting of roots which show no surface variation in any inflected form.
 
@@
 \textsc{Backness Harmony} is exemplified in the lexicon insofar as there is a positive association between the backness of vowels in all adjacent syllables.
-Any disagreements on the backness specification of vowels in adjacent syllables are counted solely as exceptions, even if other vowel transitions in the root are harmonic; further assumptions are necessary to determine whether a root which has one disharmonic transition may in any sense ``obey'' harmony elsewhere.
+Any root which contains vowels in adjacent syllables which disagree in backness is counted as disharmonic, even if other vowel transitions in the root are harmonic; further assumptions are necessary to determine whether a root which has one disharmonic transition may in any sense ``obey'' harmony elsewhere.
 %This follows from \emph{SPE} theory of lexical exceptionality (see also \citealt{Gouskova2012}), in which phonological exceptionality is a property of underlying representations, not individual segments.
-To construct the contingency table, roots are binned according to the backness specification of the first nucleus, and that of following nuclei.
+To construct the contingency table, roots are binned first according to the backness specification of the first nucleus, and then according to the backness of all following nuclei.
 For example, the first two syllables of \emph{adalet} `justice' are harmonic, but it is coded as disharmonic because there is a \emph{a\ldots{}e} transition later in the word.
 The resulting counts are shown in Table \ref{bhs}. 
-61\% of roots conform to \textsc{Backness Harmony}, and the interaction between the backness of the first and of the subsequent vowels
-%predicted by root-internal \textsc{Backness Harmony} 
-is significant.
+61\% of roots conform to \textsc{Backness Harmony}, and the interaction between the backness of the first and of the subsequent vowels is significant.
 
@@
 Any root for which a vowel does not agree in roundness with a high vowel in the following syllable (e.g., \emph{ümit}) is considered to be an exception. 
-Roots are binned according to the roundness of non-initial high vowels, and according to the roundness of the preceding vowel.
-%Roots lacking non-initial high vowels do not bear on the status of root \textsc{Roundness Harmony}. 
+Roots are binned first according to the roundness of a vowel followed by a high vowel in the next syllable, and then according to the roundness of all following high vowels.
 The resulting counts are shown in Table \ref{rhs}. 
@@
 Each response is coded as \emph{concordant} if the nonce word conforming to the process is preferred, and \emph{discordant} if the disharmonic word is selected. 
-To test for an association between the constraints, a non-parametric statistic, the Goodman-Kruskal (\citeyear{Goodman1954}) $\gamma$ is computed from the count of concordant ($c$) and discordant ($d$) pairs:
-
-\begin{equation*}
-\displaystyle \gamma = \frac{c - d}{c + d}
-\end{equation*}
-
-\noindent
-The $\gamma$ statistic ranges between -1, in the case that all paired choices are discordant, and 1, if all paired choices are concordant.\footnote{It also is possible to perform statistical tests aggregating over items, but for small number of items, such tests have very poor power.}
+To test for an association between the constraints, a non-parametric statistic, the Goodman-Kruskal (\citeyear{Goodman1954}) $\gamma$ is used. 
+The $\gamma$ statistic ranges between -1 (which would indicate that non-conforming nonce words are always preferred to conforming nonce words) and 1 (which indicates that conforming nonce words are always preferred).\footnote{
+    It also is possible to perform statistical tests aggregating over items, but for small number of items, such tests have very poor power.}
 
@@
 There is a small preference against \textsc{Labial Attraction} (and therefore in favor of \textsc{Roundness Harmony}, though this is non-significant ($\gamma = -0.043$, $p = 0.305$).
-
-Speakers do not have the preferences predicted by \textsc{Labial Attraction}. 
 %This result also extends for the variants of \textsc{Labial Attraction} proposed by \citet{Zimmer1969} and \citet{Inkelas2001}, since these variants have structural descriptions targeting a superset of the original formulation by \citet{Lees1966a}.
-It is of interest that speakers have no apparent preference at all: one might have expected that they would prefer nonce items conforming to \textsc{Roundness Harmony}.
 
@@
 
+Speakers do not have the preferences predicted by \textsc{Labial Attraction}.
+At the same time, they do not have a preference for \textsc{Roundness Harmony} either, as one might expect.
+At most, it could be said that \textsc{Labial Attraction} is sufficiently robust so as to suspend the effect of \textsc{Roundness Harmony}, with which it is in competition.
+
 \subsection{Discussion}
 
-It has been shown that while \textsc{Labial Attraction} is a highly reliable generalization about Turkish roots, it is not reflected in wordlikeness judgements.
-In contrast, harmony processes have a similar statistical profile, but have large effects on wordlikeness.
+It has been shown that while \textsc{Labial Attraction} is a highly reliable generalization about Turkish roots, it has at best a minimal effect on wordlikeness judgements.
+In contrast, harmony processes have a similar statistical profile, but have quite robust effects on wordlikeness.
 The most plausible explanation for this is that \textsc{Labial Attraction} does not trigger alternations; indeed, it is counter-exemplified by the effects of \textsc{Roundness Harmony} in suffix allomorphy.
-As is noted by \citet[412f.]{Inkelas1997}, the lexicon of Turkish will, under the assumptions here, remain as it is whether or not \textsc{Labial Attraction} has a synchronic reality.
-The law of parsimony suggests that it does not exist at all.
+As is noted by \citet[412f.]{Inkelas1997}, the lexicon of Turkish will, under the assumptions here, remain as it is whether or not \textsc{Labial Attraction} has a synchronic reality.\footnote{
+    Another possibility is that such a constraint could result in a lexical ``trend'' over time.
+    Even if such a trend is observed, however, it is quite difficult to establish that it is the result of a synchronic constraint at every intermediate stage of the language.}
+It might even be possible to suggest that it does not exist at all, though under such analysis the absence of a preference in the direction of \textsc{Roundness Harmony} is without ready explanation.
 
-\citet{Ito1995a,Ito1995b} and \citet{NiChiosain1993} claim that \textsc{Labial Attraction} holds only over the native vocabulary.
-This is a potential confound for \citeauthor{Zimmer1969}'s experiment, since it is not implausible that speakers in \citeauthor{Zimmer1969}'s study would treat nonce words much like loanwords; indeed, many wordlikeness studies include instructions to the participants to treat the stimuli much as if they were loanwords \citep[e.g.,][]{Hay2004a}.
+\citet{Ito1995a,Ito1995b} and \citet{NiChiosain1993} claim that \textsc{Labial Attraction} holds only over the native vocabulary, and that the Turkish lexicon is ``stratified''.
+This would be a potential confound for \citeauthor{Zimmer1969}'s experiment, since it is not implausible that speakers in \citeauthor{Zimmer1969}'s study would treat nonce words much like loanwords; indeed, many wordlikeness studies include instructions to the participants to treat the stimuli much as if they were loanwords \citep[e.g.,][]{Hay2004a}.
 However, \citet{Inkelas2001} find that foreign words are more---not less---likely to conform to \textsc{Labial Attraction} than native words.
@@
 
-It is not obvious that it is desirable to exclude \textsc{Labial Attraction} as a possible rule, as \citet[394, fn.
-2]{Inkelas1997} suggest that \textsc{Labial Attraction} may have even induced alternations at one point in the history of Turkish.
-However, \citeauthor{Becker2011} cite the inertness of \textsc{Labial Attraction} as evidence that naturalness constrains phonotactic learning.
+It is not obvious that it is desirable to exclude \textsc{Labial Attraction} as a possible rule.
+\citet[394, fn. 2]{Inkelas1997} suggest that \textsc{Labial Attraction} may have even induced alternations at one point in the history of Turkish.
+However, \citeauthor{Becker2011} cite the apparent inertness of \textsc{Labial Attraction} as evidence that naturalness constrains phonotactic learning.
 
@@
 
-If this is correct, it should be possible to show that a more ``natural'' variant of \textsc{Labial Attraction} is in fact reflected in wordlikeness judgements, assuming it too is statistically valid.
+If this is correct, it should be possible to show that a more ``natural'' variant of \textsc{Labial Attraction} is in fact better reflected in wordlikeness judgements, assuming it too is statistically valid.
 \citeauthor{Inkelas2001} make a similar observation.
@@
 Furthermore, as shown in Table \ref{lasi}, \textsc{Labial Attraction}$'$ is even more statistically reliable than \citeauthor{Lees1966a}'s original formulation.
-Since this reformulation targets a superset of \citeauthor{Lees1966b}'s original rule, it should be reflected in the paired rating task, but this does not obtain.
+This reformulation targets a superset of \citeauthor{Lees1966b}'s original rule, and therefore distinguishes the stimulus pairs in Table \ref{law}; i.e., any \emph{aPı} sequence violates \textsc{Labial Attraction}$'$ just as much as it violates the original formulation.
+Yet, it is not clear that either form of the generalization is strongly reflected in wordlikeness judgements.
 
@@
 Statistical reliability is neither necessary nor sufficient to predict what speakers know about possible and impossible words in their language.
-Phonotactic constraints may go unlearned whether or not they are ``natural''.
-In the case of Turkish vowel restrictions, at least, it is precisely those constraints which are derive from phonological processes---albeit processes with a considerable number of exceptions--- which are reflected in psycholinguistic tasks.
+Further, phonotactic constraints may go unlearned whether or not they are ``natural''.
+In the case of Turkish vowel restrictions, at least, it is precisely those constraints which are derive from phonological processes---albeit processes with a considerable number of exceptions---which are most clearly reflected in psycholinguistic tasks.
 This should not be taken to imply that all phonotactic constraints inferred from the lexicon are illusory: for instance, \citet{Frisch2001} present psycholinguistic support for co-occurrence restrictions in Arabic posited by \citeauthor{Frisch2004} (\citeyear{Frisch2004} [1995]) on the basis of lexical data.
-As a general principle, though, it should be apparent that lexical statistics do not contribute reliable evidence for the theory of phonotactics.
-This is even more serious when a strong relationship between phonotactic and phonological representations is assumed (as it is here): contrary to common practices \citep[e.g.,][]{Mester1988,Padgett1991,Padgett1992}, lexical statistics do not provide principled evidence for the architecture of phonological grammar.
+As a general principle, though, it should be apparent that lexical statistics do not contribute reliable evidence for the theory of phonotactics; the linguist who has identified a statistical tendency in the lexicon has much more work to do before it can be identified with the synchronic grammar.
+This is even more serious when a strong relationship between phonotactic and phonological representations is assumed (as it is here): contrary to common practices \citep[e.g.,][]{Mester1988,Padgett1991,Padgett1992}, lexical statistics can not be taken as provide principled evidence for the nature of phonological features or of other components of phonological grammar.
 

### 2013-01-28 4.tex
Revisions! Lots of 'em.

@@
 
-It is a curious fact that many constraints on underlying representations are elegantly described using the units of prosody.
-In some cases, these are part of the system of contrast, such as long segments in moraic theory \citep[e.g.,][]{Davis1999}.
-For instance, Latin has constrastive consonant and vowel length (e.g., \emph{anus} `ring' vs.~\emph{annus} `year', \emph{os} `bone' vs.~\emph{ōs} `mouth'), but there is no vowel length contrast before geminate consonants.
-However, constraints on underlying representations may also involve references to non-contrastive prosodic structures, such as syllables (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{
-    See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample is the result of an underlying moraic, not syllabic, contrast.}
-For instance, as \citeauthor{Haugen1956} first notes, the constraints on word-medial consonant clusters are not arbitrary in prosodic terms:
+It is indisputable that many phonotactic restrictions are easily described with reference to prosodic primitives.
+In some cases, prosodic factors constrain the system of contrast.
+For instance, Latin has contrastive vowel and consonant length (e.g., \emph{os} `bone' vs.~\emph{ōs} `mouth', \emph{anus} `ring' vs.~\emph{annus} `year'), but the latter contrast is suspended in codas preceded by a diphthong or long monophthong; a syllable may contain a long vowel, or be checked by the first half a geminate consonant, but not both.
+However, constraints on underlying representations may also involve references to non-contrastive prosodic structures such as the syllable (e.g., \citealt{Hooper1973}, \citealt{Kahn1976}).\footnote{
+    See \citealt{Blevins1995} for the claim that syllable structure is universally non-contrastive, and \citealt{Elfner2006} for arguments that a putative counterexample derives from an underlying vowel length contrast.}
+For instance, as noted by \citet{Haugen1956}, 
+numerous restrictions on word-medial consonant clusters have a unified statement in prosodic terms:
 
-\begin{example}[\textsc{Medial Cluster Law} (after \citealt{Haugen1956})]
+\begin{example}[\textsc{Medial Cluster Law}]
 \label{mcl}
-A medial cluster consists maximally of a well-formed medial coda and a well-formed medial onset
+A medial cluster can consist maximally of a well-formed medial coda and a well-formed medial onset
 \end{example}
 
+As an illustration of this tautology, consider languages like Yokuts, which forbid complex codas and complex onsets, and which enforce these restrictions in complex words via processes such as vowel epenthesis.
+\citet[26f.]{Newman1944} notes that this imposes an an upper bound on the size of medial clusters: no cluster of more than two consonants can be parsed into a simple coda and simple onset.
+In the case of Yokuts, it is certainly possible to state this restriction without reference to syllable structure, as *CCC, a constraint on trisyllabic clusters (e.g., \citealt[92f.]{Ettlinger2008}, \citealt[820f.]{Zuraw2003a}).
+However, the aforementioned constraints on complex onsets and codas find independent motivation from the total absence of initial and final clusters in Yokuts;\footnote{
+    The tendency of consonants to pattern with word boundaries and morph junctures has been long been noted \citep[e.g.,][]{Hill1954,Lass1971,Moulton1947}; \citet[24f.]{Kahn1976} takes this as evidence for the syllable.}
+with these two restrictions in place, a further constraint on medial triconsonantal clusters is otiose.\footnote{
+    \citet[31f.]{Cote2000} alludes to another critique of constraints like *CCC, namely that this constraint requires ``counting'': see \citealt[64f.]{Isac2008} for further discussion of the comparative merits of ``counting'' and ``grouping'' analyses in phonology.}
+While the medial cluster law is certainly consistent with the hypothesis that syllable structure may be present in underlying representations (e.g., \citealt{A74}:255, \citealt{Vaux2003}), this need not be the case under Stampean occultation.
+If an underlying medial consonant sequence appears on the surface, it satisfies the medial cluster law by definition.
+If, however, an underlying cluster is modified by consonant deletion, coalescence, or vowel epenthesis, then it need not consist of a licit medial onset and medial coda.
+
+\citet{Pierrehumbert1994} begins a study of English word-medial consonant clusters with a restatement of the medial cluster law:
+
+\begin{quote}
+That is, in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[168]{Pierrehumbert1994}
+\end{quote}
+
 \noindent
-It is not necessary for predictable prosodic structure to be present in underlying form to achieve this effect, contrary to what has sometimes been stated \citep[e.g.,][255]{A74}.
-All that is necesary is that no phonological processes intervene between underlying representation and the surface form of medial clusters.
-This is little more than the definition of a ``faithful'' cluster.
-Under the further assumption that underlying and surface representations are as similar as phonological processes allow (e.g., \citealt[205f.]{Dell1973}, \citealt[28f.]{Stampe1973}), faithful realization is the null hypothesis.
-
-In a study of English word-medial consonant clusters, \citet{Pierrehumbert1994} notes that the vast majority of ``possible clusters'', those which can be parsed into coda and onset in accordance with the \textsc{Medial Cluster Law}, are unattested.
-\citeauthor{Pierrehumbert1994} argues that this can be accounted for by positing static co-occurrence restrictions---unrelated to any phonological alternation in English.
-This study argues that well-known phonological processes, but not the static constraints proposed by \citeauthor{Pierrehumbert1994}, impose restrictions on this inventory.
-%\citeauthor{Pierrehumbert1994}'s static constraints are not statistically valid
+However, \citeauthor{Pierrehumbert1994} reports that the vast majority of the ``possible'' clusters (i.e., those which conform to the medial cluster law) are in fact unattested. 
+So as to acount for this, \citeauthor{Pierrehumbert1994} presents ``provisos'' in the forms of static co-occurrence restrictions, unrelated to any phonological alternation in English.\footnote{
+    Given the arguments presented in the previous chapter, that not all lexical tendencies have a synchronic basis, one may question the intuition that any gaps in the English cluster inventory (or those that go beyond the medial cluster law) must be accounted for by the synchronic grammar.
+    In addition to some informal statistical evidence (of the sort problematized in the previous chapter), \citeauthor{Pierrehumbert1994} administers a wordlikeness task to validate the static constraints she proposes.
+    However, this experiment is of a quite informal nature and the results are given only a superficial analysis, so it is less than probative.} 
+This chapter will argue, however, the static constraints proposed by \citeauthor{Pierrehumbert1994} are unnecessary, and that the only restrictions on the inventory of medial clusters in English (beyond the medial cluster law) are those which derive from well-known phonological processes.
 
@@
 
-English syllable contact clusters have been used, by \citeauthor{Pierrehumbert1994} and others, to argue for the necessity of admitting static phonotactic constraints, and to illustrate proposals for the architecture of the phonotactic system.
-However, there are a number of reasons to reconsider this domain.
+The aforementioned study by \citeauthor{Pierrehumbert1994}, as well as further investigations of this domain by \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, argue for the necessity of admitting static phonotactic constraints, and illustrate proposals for the architecture of the phonotactic system.
+However, there are a number of reasons to reconsider the findings of these authors in light of the proposals made in previous chapters.
 
@@
 
-\citeauthor{Pierrehumbert1994} does not fully consider the effects of phonological processes targeting word-medial clusters in English, and consequently some of the static constraints she proposes may be related to English morphophonemics.
-For instance, she writes that ``nasal-stop sequences agree in labiality'' (175), and attributes this the status of a static constraint. 
-However, this generalization is just a narrower form of the one derived by \textsc{Nasal Place Assimilation} (see \S\ref{npa}).
-Other derived constraints go unnoticed: for instance, she does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}). 
-The evaluation below considers the role of both types of constraints on the English lexicon on an equal footing.
+\citeauthor{Duanmu2009}, \citeauthor{Hammond1999a}, and \citeauthor{Pierrehumbert1994} do not generally take into account the effects of phonological processes which target medial clusters in English.
+As a consequence, some of the static constraints these authors identify may be in fact the product of English morphophonemics and Stampean occultation. 
+For instance, \citeauthor{Pierrehumbert1994} writes that ``nasal-stop sequences agree in labiality'' (175) and posits a static constraint to account for this fact.
+However, this generalization is merely a narrower form of a restriction deriving from a process of \textsc{Nasal Place Assimilation} (see \S\ref{npa}).
+Other derived constraints are simply not mentioned; for instance, \citeauthor{Pierrehumbert1994} does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}); while \citet{Hammond1999a} does allude to this restriction, it is dismissed in light of a few apparent counterexamples (though see \S\ref{sss:map} below).
+In contrast, this chapter attempts to evaluate derived and static constraints on an equal footing.
 
@@
 \citet{Pierrehumbert1994} infers static constraints from near-exceptionless gaps in the lexicon, but little effort is made to show that the patterns of lexical underrepresentation are not due to chance.
-Consquently, it is possible to suggest that some of these gaps 
-are accidental rather than structural in nature.
-%\citep{Fischer-Jorgensen1952,Vogt1954}. 
-This is made all the more likely given the tendency of segment and cluster frequency distributions to be highly skewed \citep[e.g.,][]{Pande2010,Sigurd1968,Tambovtsev2007,Weiss1961}. 
-Furthermore, \citeauthor{Pierrehumbert1994} considers only triconsonantal clusters, but medial clusters may be as short as two consonants, as in \emph{a}[n.t]\emph{ics}, or as long as four, as in \emph{mi}[n.str]\emph{el}, and \citeauthor{Pierrehumbert1994} offers no justification for a narrower focus.
-If this has any effect on the distribution of clusters, it would be to impose further sparsity.
+Consequently, it is possible to suggest that some of these gaps are accidental rather than structural in nature.
+This is made all the more likely given the tendency of segment and cluster frequency distributions to be highly skewed \citep[e.g.,][]{Pande2010,Sigurd1968,Tambovtsev2007,Weiss1961} so that it is difficult to distinguish between structural and accidental gaps.
+Furthermore, \citeauthor{Pierrehumbert1994} considers only triconsonantal clusters, but medial clusters may be as short as two consonants, as in \emph{a}[n.t]\emph{ics}, or as long as four, as in \emph{mi}[n.str]\emph{el}, and no justification is given for ignoring clusters of other lengths.
+If there is any effect of this focus, it is presumably to produce further sparsity in the distribution observed.
 
@@
 
-Many components of \citeauthor{Pierrehumbert1994}'s study cannot be directly replicated.
-\citeauthor{Pierrehumbert1994} focuses her attention on those words she judges to be ``morphologically simple'' and ``reasonably familiar'', but this list has neither been published nor circulated.
-It is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias; 
-\citet{Labov1975} for instance
-argues for what he calls the \emph{Experimenter Principle}, 
-that the intuitions of all those familiar with the formal matters should be excluded.
-Otherwise-unmotivated morphological junctures are sometimes posited simply to preserve phonological or phonotactic generalizations; 
-%The same practice is sometimes applied to preserve phonological generalizations; for instance, otherwise unmotivated morphological boundaries are used in \emph{SPE} to simplify principles of stress assignment.
-this is done by \citet{SPE}, for instance, to simplify principles of English stress assignment, and \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur morph-internally.
+Many components of \citeauthor{Pierrehumbert1994}'s study cannot be replicated.
+\citeauthor{Pierrehumbert1994} limits her study to words she judges to be ``morphologically simple'' and ``reasonably familiar''; the author's sensations thereof are not replicable, nor are they available to other researchers in any form.
+It has been suggested \citep[e.g.,][]{Labov1975,Schutze1996} that the sensations (as well as cognitive limitations) of concerned parties should not be granted evidential status in the first place, given the potential for implicit bias; \citeauthor{Labov1975} calls the \emph{Experimenter Principle}.
+
+It is not uncommon for analysts to propose otherwise-unmotivated morphological junctures simply to preserve phonological or phonotactic generalizations.
+This is done by \citet{SPE}, for instance, to simplify principles of English stress assignment.
+Similarly, \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur in morph-internal contexts.
 %There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings.
-Applied indiscriminately, however, such a heuristic trivializes both morphological segmentation and phonotactic generalization.
-For these reasons, the wordlist used in this study is derived from a publicly available database, and no experimenter intuitions are used for establishing morphological boundaries.
-
-%Such reasoning is found in \citeauthor{Pierrehumbert1994}'s discussion of the word \emph{a}[nt.l]\emph{er}, however: this contains a cluster which violates her constraint on coda coronal obstruents, and therefore must be a complex word, she argues (see \citealt[164]{Borowsky1989} for additional exceptions to this generalization).
+Applied indiscriminately, however, this heuristic trivializes both morphological segmentation and phonotactic generalization.
+For these reasons, the wordlist used in this study is derived from a publicly available database, and no experimenter intuitions are used.
 
@@
 
-After constructing a sample of syllable contact clusters in English simplex words, this data is used to evaluate the coverage of static and phonologically derived constraints.
+After constructing a sample of syllable contact clusters in English simplex words, this sample is used to evaluate the coverage of static and derived constraints.
 
@@
 \subsubsection{Materials and procedure}
+\label{sss:map}
 
-Following \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, 
-a wordlist is generated using the English portion of the CELEX database \citep{CELEX}. 
-Only words marked in CELEX as ``monomorphemic'' are used, and all words 
-labeled non-native are excluded.
-This stringent criterion for inclusion eliminates many of the exceptions noted by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies.
-For instance, nearly all the exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) noted by \citet[74]{Hammond1999a} are
-either complex words (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).
+Following \citet[ chap.~8]{Duanmu2009} and \citet[ chap.~3]{Hammond1999a}, who also consider restrictions on English medial clusters, a wordlist is generated using the English portion of the CELEX database \citep{CELEX}.
+Only words marked in CELEX as ``monomorphemic'' are used, and all words labeled in CELEX as non-native are excluded.\footnote{
+    The first criterion results in the exclusion of proper names, which have long been noted to push the bounds of native language phonotactics \citep[e.g.,][254]{Trubetzkoy1939}.}
+These more-stringent criteria exclude many words labeled exceptions in the studies by \citeauthor{Duanmu2009} or \citeauthor{Hammond1999a} in their studies.
+For instance, nearly all the exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) noted by \citet[74]{Hammond1999a} are excluded either as complex words (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}) or non-native (e.g., \emph{vodka}, \emph{smorgasbord}).
 
-Unlike prior studies, words which appear to consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}) are also excluded from the sample, as there is broad formal and experimental support for the claim that these words are complex.
-This assumption allows for numerous simplifications to accounts of the phonological shape of these forms. 
+In contrast to prior studies, these criteria also exclude words which consist of a Latinate prefix and a bound stem (e.g., \emph{inspect}, \emph{excrete}). 
+While \citeauthor{Pierrehumbert1994} rejects this analysis as as unmotcontroversial, it in fact has extensive formal and experimental support.
+First, Latinate prefixes simplify the statement of many morphophonemic details in English.
 For instance, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.
@@
 \citeauthor{Aronoff1976} takes this to be evidence that \emph{adhere} and \emph{cohere}, for instance, share a bound stem.
-There is also an interaction between Latinate prefixes on verbs and the makeup of the verbal complement. 
-Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all of which are acceptable with similar Anglo-Saxon verbs \citep{Gropen1989,Harley2009}.
-%\citep[see also][]{Gropen1989,Coppock2008}.
+There is also an interaction between Latinate prefixes on verbs and the complements they select.
+Latinate verbs do not generally allow ditransitive, verb participle, or adjectival resultative constructions, all of which are acceptable with similar Anglo-Saxon verbs \citep[e.g.,][]{Gropen1989,Harley2009}.
 
-\begin{example}[Syntactic restrictions on Latinate verbs]
+\begin{example}[Restrictions on Latinate verbal complements]
 \label{harley}
@@
 b. & {drink himself stupid}  & \alt{} & * & {imbibe himself stupid}    \\
-%c. & {show it off}           & \alt{} & * & {exhibit it off}           \\
 c. & {break it off}          & \alt{} & * & {terminate it off}         \\
@@
 
-Lexical decision experiments also suggest that Latinate prefixed forms are complex.
+Lexical decision also provide evidence for the segmentation of Latinate prefixed forms.
 \citet{Taft1975,Taft1976} and \citet{Taft1986} find that nonce words like \emph{*re-sert}, which appear to be composed of a prefix and a bound stem, take longer to reject that non-words which lack apparent morphological structure, such as *\emph{refant}.
-Bound stems also show frequency effects which are independent of whole word frequency \citep{Taft1979,Taft2006}. 
-Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming between pairs like \emph{permit}-\emph{submit}, an effect thought to indicate morphological relatedness.
+Bound stems also show frequency effects independent of whole word frequency \citep{Taft1979,Taft2006}.
+Finally, \citet{Emmorey1989} and \citet{Forster2000} report facilitative priming, thought to implicating morphological relatedness, between pairs like \emph{permit}-\emph{submit}, which appear to share a bound stem.
 
@@
 
-The filters discussed above result in a list of 6,619 simplex words.
-The full set of clusters and their frequencies are given in Appendix \ref{clusters}.
-The CELEX transcriptions of these words were then syllabified and phonologized using a technique described in Appendix \ref{syllabification}. 
-This produces 23 unique medial codas and 40 unique medial onsets. 
-Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination, 174 (19\%) are attested.
+Filtering the CELEX data according to the above criteria results in a list of 6,619 simplex words.
+The full set of clusters and their frequencies are listed in Appendix \ref{clusters}.
+The CELEX transcriptions of these words are then syllabified and phonologized using the same procedure as used in Chapter \ref{gradience} and described in Appendix \ref{syllabification}. 
+In all, the sample contains 23 different medial coda and 40 different medial onsets.
+Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination of medial coda and medial onset, 174 (19\%) are attested.
 
@@
 
-So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters.
-These constraints are static, in that they have no analogues in the morphophonemics of English.
+To account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three static constraints on English medial clusters.
 
 \paragraph{Dorsal-labial clusters}
-\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', noting that velar-velar clusters are independently excluded by constraints on geminates (see \S\ref{degemx}).
-Two-consonant velar-labial clusters are found in words like \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent} occur. 
-Table \ref{dltab} shows that velar-labial clusters are somewhat less likely to occur than  are velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance ($p = .106$).
+\citet[173]{Pierrehumbert1994} writes that ``velar obstruents occurred only before coronals in the clusters studied, never before labials or other velars'', while noting that absence of velar-velar clusters is expected due to a separate constraint on geminate clusters (see \S\ref{degemx}).
+However, biliteral velar-labial clusters are found in words such as \emph{a}[k.m]\emph{e}, \emph{ru}[ɡ.b]\emph{y}, or \emph{pi}[ɡ.m]\emph{ent}.
+Velar-labial clusters are somewhat less common than velar-coronal clusters (e.g., \emph{ve}[k.t]\emph{or}), but such underrepresentation is not unlikely to occur by chance according to the Fisher exact test (Table \ref{dltab}).
 
@@
 \paragraph{Coronal obstruent codas}
-\citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}.
-Table \ref{cctab} reveals that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}). 
-The same is true if attention is restricted to triconsonantal clusters ($p = .129$).
+\citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', but at the same time observes exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}.
+In the CELEX sample (Table \ref{cctab}), coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}).
+While not shown in tabular form, the same is true if attention is restricted to triconsonantal clusters ($p = .129$).
 
@@
 \paragraph{ABA clusters} 
-\citet[176]{Pierrehumbert1994} notes the ``lack of clusters with identical first and third elements'', ignoring presence or absence of voicing. 
-Despite the fact that there are no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than other triconsonantal and quadraconsonantal clusters (Table \ref{abatab}).
+\citet[176]{Pierrehumbert1994} observes a ``lack of clusters with identical first and third elements'', ignoring presence or absence of voicing. 
+Despite the fact that there are no exceptions to this generalization, these \textsc{ABA} clusters are not significantly less common than any other triconsonantal and quadraconsonantal clusters (Table \ref{abatab}).
 
@@
 \paragraph{Summary} 
-There is no statistical support for \citeauthor{Pierrehumbert1994}'s static constraints.
+There is no statistical support for any of \citeauthor{Pierrehumbert1994}'s static constraints.
 
@@
 
-In \emph{The Sound Pattern of English} (henceforth, \emph{SPE}), \citet{SPE} describe three phonological processes which target medial consonant clusters.
-As will be shown, the neutralizing effect of these three processes has a major influence on the cluster inventory.
+In \emph{SPE}, \citet{SPE} describe three phonological processes which target medial consonant clusters. 
+As will be shown, these three processes have a profound influence on the English cluster inventory.
 
@@
 
-Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), which take the voicing specification of a preceding obstruent.\footnote{Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.}
+Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t]-\emph{nab}[d]) and noun plural (e.g., \emph{lap}[s]-\emph{lab}[z]), which take the voicing specification of a preceding obstruent;\footnote{
+    Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.}
+voice assimilation is also claimed to operate across prefix and compound junctures \citep{Davidsen-Nielsen1974}.
 %Bizarely, \citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''.
@@
 \citet{Pierrehumbert1994} does not discuss a constraint against adjacent obstruents disagreeing in voice. 
-As shown in Table \ref{ovatab} however, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy}.
-Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, are rarer than would be expected from chance.
+However, the vast majority of medial obstruents clusters in simplex words are either uniformly voiced, as in \emph{hu}[z.b]\emph{and}, or uniformly voiceless, as in or \emph{rha}[p.s]\emph{osdy} (Table \ref{ovatab}).
+Hetero-voiced clusters, like those in \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, are far rarer than would be expected from chance.
 
@@
 %\citealt[269]{Fudge1969})
-permits [ŋ] to be described as a pure allophone of /n/, and accounts for certain patterns in Latinate prefix allomorphy.
+permits [ŋ] to be described as an allophone of /n/ (see \S\ref{s:phonologization}), and furthermore accounts for allomorphy in certain Latinate prefixes.
 
@@
 \noindent
-As shown in Table \ref{npatab}, virtually all clusters consisting of a nasal coda followed by a homorganic obstruent (e.g., \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, \emph{mo}[ŋ.k]\emph{ey}) are attested.
-As \citet[175]{Pierrehumbert1994} observes, heterorganic clusters, like those \emph{pli}[m.s]\emph{oll} or \emph{scri}[m.ʃ]\emph{aw}, do occur, but they are significantly more rare.
+Virtually all clusters consisting of a nasal coda followed by a homorganic obstruent (e.g., \emph{pi}[m.p]\emph{le}, \emph{sta}[n.z]\emph{a}, \emph{mo}[ŋ.k]\emph{ey}) are attested (Table \ref{npatab}).
+As \citet[175]{Pierrehumbert1994} observes, heterorganic clusters, like those \emph{pli}[m.s]\emph{oll} or \emph{scri}[m.ʃ]\emph{aw}, do occur, but in this sample they are significantly more rare.
 
@@
 \paragraph{Degemination} 
-The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]), in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.
-\emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}).
-\textsc{Degemination} is formalized here as a rule deleting the first of two segments agreeing on all feature values (except voice, possibly).
+The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]), in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf.~\emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\citealt[102]{Borowsky1986}, \emph{SPE}:148).
+\textsc{Degemination} is formalized here as a rule deleting the first of two segments agreeing on all feature values (except for voice, possibly).
 \label{degemx}
@@
 \noindent
-As can be seen from Table \ref{degemtab}, the sample includes no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance.
-Stampean occultation by \textsc{Degemination} provides a natural explanation for this gap.
-It is informative to compare this to the constraint against ``ABA'' clusters proposed by \citeauthor{Pierrehumbert1994}: both are exceptionless, but only the former is unlikely to arise by chance.
+The sample contains no sequences of identical segments, or of identical segments differing only in voice, something highly unlikely to arise by chance (Table \ref{degemtab}); \textsc{Degemination} and Stampean occultation provide a natural explanation for this gap.
+It is interesting to compare the absense of geminates to the constraint against ``ABA'' clusters proposed by \citeauthor{Pierrehumbert1994} in this regardl: both are exceptionless, but only the former imposes a lexical tendency unlikely to arise by chance.
 
@@
 
-Current computational models of phonotactic knowledge ``rate'' possible clusters by producing a numerical value for each cluster. 
-To apply these models to predicting the cluster inventory, these numerical values must be transformed into a categorical prediction of either attestation or non-attestation.
-This is accomplished with a soft-margin support vector machine \citep{Cortes1995} with a linear kernel, which optimally splits the ratings distribution into a set of attested and unattested clusters.
-This classifier does not correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.
+Current computational models of phonotactic knowledge can ``rate'' possible clusters, assigning a numerical wellformedness score to any input.
+These models can be applied to a task of predicting which clusters are and are not attested by transforming these numerical values into a categorical prediction of either attestation or non-attestation.
+This is accomplished here with a soft-margin support vector machine \citep{Cortes1995} with a linear kernel, which attempts to find a single optimal numerical value about which to split attested and unattested clusters.
+This classifier is not intended to correspond to any component of a cognitively plausible model of phonotactic learning: it simply represents an upper bound for predicting the cluster inventory from positive data.
 
-Four metrics are used to score the models. 
+The models are scored using a ``leave-one-out'' scheme, in which each observation (a cluster) is scored using a model trained on all other observations.
+Four metrics are used to evaluate model fit.
 \emph{Accuracy} represents the probability that a cluster is correctly classified as attested or unattested.
-Two additional metrics break down accuracy into its constituent parts;
+Two additional metrics break down accuracy into constituent parts;
 \emph{precision} represents the probability that a cluster which is predicted to be unattested is in fact unattested, and \emph{recall} is the probability that an unattested cluster is predicted as such.
 It is possible to increase precision at the expense of recall, by predicting non-attestation for a greater number of clusters, or to maximize recall at the expense of precision by predicting all clusters to be unattested.
-$F_1$, the harmonic mean of these two measures, is a standard metric for quantifying this tradeoff; increases in either precision or recall produce a  non-linear increase in $F_1$.
+$F_1$, the harmonic mean of these two measures, is a standard metric for quantifying this tradeoff; any increase in either precision or recall will result in an increase in $F_1$.
 The results are summarized in Table \ref{cmresults}.
@@
 Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
-Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
-Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
-DC \& EF            & 0.861    & 0.868     & 0.988  & 0.924 \\
-\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
+Expected frequency  & 0.835    & 0.837     & 0.960  & 0.894 \\
+Derived constraints & 0.838    & 0.835     & 0.967  & 0.897 \\
+DC \& EF            & 0.861    & 0.866     & 0.969  & 0.914 \\
+\citet{Hayes2008a}  & 0.835    & 0.964     & 0.833  & 0.894 \\
 \bottomrule
@@
 In a classification task, the simplest baseline is one which uniformly predicts the most common outcome.
-Since only 19\% of clusters are attested, 
-81\% accuracy is achieved by predicting all clusters to be unattested.
+Since only 19\% of clusters are attested, 81\% accuracy can be achieved simply by predicting all clusters to be unattested.
 
@@
 \citeauthor{Pierrehumbert1994} reports that this is an excellent predictor of which complex clusters occur and which do not.
-This does not impose any constraints which span the syllable boundary; rather, it is a model of which clusters might be expected to represent accidental gaps in the sample.
-This represents a significant improvement compared to the null baseline (sign test $p = 4.5$\e{-05}).
+This model does not impose any constraints which span the syllable boundary; rather, it is a model of which clusters might be expected to represent accidental gaps in the sample.
+This produces a small but significant improvement in accuracy over the null baseline (sign test, $p = 4.5$\e{-05}).
 
@@
 By hypothesis, \textsc{Obstruent Voice Assimilation}, \textsc{Nasal Place Assimilation}, and \textsc{Degemination} rule out a large number of possible clusters. 
-In all, they target 316 out of 920 possible clusters (34\%), and only 11 (3\%) of these clusters have non-zero frequency in the sample.
-Together, these three processes define a simple classifier in which a cluster is predicted to be unattested if and only if it would be neutralized by one of these processes. 
-Unsurprisingly, this results in increased precision and a significant improvement in accuracy compared to the null baseline (sign test $p = 4.0$\e{-09}).
+In all, they target 316 out of 920 possible clusters (34\%) for neutralization;
+of these clusters, only 11 (3\%) occur in the sample.
+Together, these three processes define a simple classifier in which a cluster is predicted to be attested only if it would not be neutralized by one of these processes.
+This results in increased precision and a small but significant improvement in accuracy compared to the null baseline (sign test, $p = 4.0$\e{-09}).
 
@@
 It is possible to combine into a single classifier the intuitions of the expected frequency and derived constraint models, the former accounting for accidental gaps and the latter for structural gaps imposed by neutralizing phonological processes.
-By simultaneously accounting for both sources of cluster inventory gaps, this model outperforms all others in accuracy and $F_1$.
+Simultaneously accounting for both sources of cluster inventory gaps, this model outperforms all others in accuracy and $F_1$.
 
 \paragraph{Maximum entropy phonotactics}
-\citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints. 
-In one sense, this is isomorphic to the expected frequency/derived constraint model in that the constraint discovery mechanism is sensitive to the expected frequency of clusters: it favors constraints which rule out clusters with high expected frequency (but which are unattested) over those which have low expected frequency.
-Unlike the expected frequency/derived constraint model, however, the constraints discovered are not limited to those which are illustrated by alternations; it may also posit static constraints, much like those proposed by \citet{Pierrehumbert1994}.
+\citet{Hayes2008a} present a model using the principle of maximum entropy to weigh a large number of competing phonotactic constraints.
+In one sense, this is isomorphic to the expected frequency model in that the constraint discovery mechanism is sensitive to the expected frequency of clusters: it favors constraints which rule out clusters with high expected frequency (but which are unattested) over those which have low expected frequency.
+Also like the expected frequency model, alternations play no role and static constraints like those proposed by \citet{Pierrehumbert1994} may be posited.
 
-Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original paper is attempted: both their implementation and phonological feature specifications are used here.
-Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in Appendix \ref{syllabification} and a novel feature [$\pm$
-\textsc{Coda}] is added to allow the model to distinguish coda and onset co
-nsonants.
+Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original study is attempted: both their implementation and phonological feature specifications are used here, just as in Chapter \ref{gradience}.
+Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in Appendix \ref{syllabification}, and a novel feature [$\pm$ \textsc{Coda}] is added to allow the model to distinguish coda and onset consonants.
 Also, following \citeauthor{Hayes2008a}, constraints are limited to those s
-panning as many as three segments and an ``accuracy schedule'' of [$.001$, 
-$.01$, $.1$, $.2$, $.3$] is used.
-Since the maximum entropy model produces slightly different scores on each 
-run, the worst-performing of 10 runs is reported here, following \citeauthor{Hayes2008a}.
+panning as many as three segments and the suggested ``accuracy schedule'' is used.
+Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citeauthor{Hayes2008a}.
 This model has the poorest recall of any model; compared to the derived constraints baseline, the constraints induced by the maximum entropy model are narrower.
@@
 \paragraph{Summary}
-Expected frequency and derived constraints effectively account for 
-accidental and structural gaps in the cluster inventory.
-The \citet{Hayes2008a} computational model only provides an inferior approximation of the derived constraints.
+Expected frequency and derived constraints effectively account for accidental and structural gaps in the cluster inventory.
+The \citet{Hayes2008a} computational model only provides an inferior approximation of the derived constraints.\footnote{
+    \citet{McGowan2009} claims that the frequency of individual syllable contact clusters in English is proportional to the change of sonority over the syllable boundary.
+    However, \citeauthor{McGowan2009} reports that the change in sonority in fact accounts for only a small portion of the variance in cluster frequency.
+    A pilot study showed that sonority change was not useful as a predictor of cluster attestation.}
 
@@
 
-Insofar as the \citet{Hayes2008a} model discovers static constraints, these fare no better than those proposed by \citet{Pierrehumbert1994}.
-It appears that further gaps in the cluster inventory cannot be described in phonological, structural terms; they can only be accidental gaps.
-In what remains, the nature of these accidental gaps is considered.
+What then is to be said of the 366 clusters which do not violate a derived constraint, but which are yet unattested, like [b.z] or [z.n]?
+Insofar as the ``discovery procedure'' used by linguists (e.g., \citeauthor{Pierrehumbert1994}) and computers (e.g., the \citeauthor{Hayes2008a} model) are based on principled phonological primitives, yet fail to find meaningful gaps, the absence of these clusters appears to be phonologically arbitrary.
+It appears that further gaps in the cluster inventory cannot be described in phonological, structural terms.
+The remainder of this chapter is concerned with the nature of these gaps.
 
@@
 
-\citet{Good1953} proposes a method for estimating the extent of accidental gaps in a distribution under simple assumptions.
-More specifically, this takes the form of $p_0$, a probability representing the possibility of an unseen cluster.
-Were it possible to extend the lexical sample, this would be the probability that the next cluster would be one not yet observed.
-The value of $p_0$ is the ratio of clusters that occur exactly once in the sample to the size of the sample.
+\citet{Good1953} proposes a method for estimating the probability of accidental gaps in a sample distribution. 
+This takes the form of an estimate of $p_0$, the probability of outcomes with have zero frequency in the sample.
+%Were it possible to extend the lexical sample, $p_0$ respresents an estimate of the probability that the next cluster would be one not yet observed.
 
@@
 \noindent
-Here, $n_1 = 67$ and there are 997 clusters in all, so there is approximately a 7\% chance that the next cluster would fill in what is now an accidental gap.
+In prose, the value of $p_0$ is the ratio of clusters that occur exactly once in the sample ($n_1$) to the size of the sample ($N$). 
+In the data here, $n_1 = 67$ and there are 997 clusters in all, so were it possible to extend the lexical sample, there is an approximately 7\% chance that the next cluster would ``fill in'' what is now a gap.
+This alone indicates the non-trivial amount of ``missingness'' and the high likelihood of accidental gaps in such a sample.
 
@@
 
-It can be shown that the large number of possible-but-unattested clusters is a logical necessity, given the sparse nature of the coda and onset frequencies. 
-In Figure \ref{cao}, the rank and type frequency (i.e., frequency in the lexicon) of medial codas and onsets in the lexical sample described below are graphed in log-log space. 
+It can be shown that the large number of possible-but-unattested clusters is a logical necessity given the sparse distribution of codas and onsets.
+The rank and type frequency (i.e., frequency in the lexicon) of medial codas and onsets in the lexical sample described below are displayed in log-log space in Figure \ref{cao}.
 Both codas and onsets show a linear relationship between log rank and log frequency characteristic of Zipfian distributions (see Appendix \ref{zr}).
-As a result, an enormous lexical sample would be needed to realize clusters consistent with the \textsc{Medial Cluster Law} even if there were no constraints on the combination of medial codas and onsets in English.
+As a result, an enormous lexical sample would be needed to realize all clusters predicted by the medial cluster law, even if there were no constraints on the combination of medial codas and onsets in English.
 To illustrate this fact, a simulation is used to create new ``samples'' of the same size as the lexical sample used here.
-The following procedure is repeated so as to generate each observation in the simulated sample.
+The following procedure is repeated so as to generate new ``observations'' for the simulated sample.
 
@@
 
-This procedure corresponds to the assumption that the derived constraints impose the only structural restrictions on the cluster inventory.
-Cluster frequencies in one simulated sample are shown in Figure \ref{sim}; dots represent simulated frequencies, and the line represents actual cluster frequencies.
-As can be seen, the observed and simulated frequencies are quite similar 
-($R^2 = 0.712$, $p = 4.5$\e{-05}).
-The sparse cluster inventory, previously taken as evidence for static constraints on syllable contact clusters, is expected even if the only constraints on possible clusters are those which are phonologically derived.
+This procedure corresponds to the assumption that the medial cluster law the derived constraints impose the only structural restrictions on the cluster inventory.
+Cluster frequencies in one simulated sample are shown in Figure \ref{sim}; points represent simulated frequencies and the line actual cluster frequencies.
+As can be seen, the observed and simulated frequencies are quite similar (i.e., $R^2 = 0.712$; $p = 4.5$\e{-05}).
+To summarize, the sparse cluster inventory, which \citeauthor{Pierrehumbert1994} takes as evidence for static constraints on syllable contact clusters, would result even if said static constraints do not exist.
 
@@
 \includegraphics{sim.pdf}
-\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line); simulated and observed frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr})}
+\caption{A simulated cluster inventory closely matches the observed cluster frequencies (represented by the smoothing line); frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr})}
 \label{sim}
@@
 
-The foregoing results demonstrate that the only structural constraints on English syllable contact are derived from the phonological system: there is no evidence for static constraints in this domain.
-What then is to be said of the 366 clusters which do not violate a derived constraint, but which are yet unattested, like [b.z] or [z.n]? 
-These gaps appear to be phonologically arbitrary, since they are not reliably identified either by phonologist or computional model.
-They can only be understood as accidental gaps resulting from the sparsity of the sample and the finite nature of the English lexicon.
+The foregoing results suggest that the only structural constraints on English syllable contact are derived from the phonological system: this study finds no evidence for static constraints.
+The many other unattested clusters can only be understood as accidental gaps which are a consequence of the finite nature of the English lexicon.
 
@@
 To answer this question, the 720 clusters in which both the final coda consonant and initial onset consonant are obstruent are sorted into four bins, according to whether they are attested or not, and whether or not the conform to the derived constraint---both obstruents are either voiced or voiceless, like in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}, respectively---or violate it, like the aforementioned \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}.
-
-%\citet{McGowan2009} claims that the type frequency of individual syllable contact clusters in English is predicted by the change of sonority in the clusters.\footnote{Thanks to Maria Gouskova for bringing this study to my attention.} Using a sonority scale proposed by \citet{Jespersen1904}, \citeauthor{McGowan2009} reports that clusters like [m.p], with sharply falling sonority, are slightly more frequent than those like [t.j], with rising sonority.
-However, \citeauthor{McGowan2009} finds that sonority of clusters accounts for a very small of the variance in cluster frequency ($R^2 = 0.077$).
-As a predictor of cluster attestation, sonority distance provided no improvement over the null baseline; consequently, it is not included in Table \ref{cmresults}.

### 2013-01-28 5.tex
Revisions! Lots of 'em.

@@
-\chapter{Inflectional gaps in generative grammar}
+\label{conclusions}
 
-\section{A brief history of inflectional gaps} \input{5.1}
-\section{The null hypothesis}                  \input{5.2}
-\section{Outline of Part II}                   \input{5.3}
+\section{Summary of dissertation and future directions}
+
+\subsection{Chapter 1}
+
+Chapter \ref{intro} began with the claim that phonotactic theorists have not been sufficiently explicit about what they take as evidence for phonotactic knowledge.
+A core challenge in phonotactic theory is to determine which of the countless competing generalizations are internalized by speakers.
+Wordlikeness judgements, production and recognition experiments, loanword adaptation, alternative phonologies, and lexical statistics all are potential sources of phonotactic evidence, though there are important caveats associated with each source of evidence.
+
+Some questions remain open regarding the relationship between loanword adaptation and phonotactic generalizations.
+A crucial question is what determines which phonotactic generalizations will be implicated in loanword adaptation. 
+As indicated in the chapter, a naïve hypothesis that those generalizations which are phonologically ``derived'' cannot be maintained in light of evidence put forth by \citet{Peperkamp2005}.
+On the other hand, the model of loanword adapation advocated by \citeauthor{Peperkamp2005} is has little to say regarding how phonotactic knowledge and how it might influence non-native speech perception.
+
+\citeauthor{Shibatani1973}'s classic argument for the insufficiency of morpheme structure constraints was then presented.
+It was suggested that much of the distinction between \emph{sequence structure constraints} and phonological processes is moot in light of \emph{Stampean occultation}.
+Other sequence structure constraints may be identified with language-specific restrictions on the prosodic inventory.
+It therefore remains to be seen whether there are ``static'' constraints: those which are neither derived by phonological process or phonotactic parsing. 
+According to one hypothesis, static constraints are inferred from statistical trends in the lexicon; this account makes predictions about gradient grammaticality and the independence of phonology and phonotactics which are taken up in in Chapters \ref{gradience} and \ref{turkish}.
+Another account identifies phonotactic knowledge with inviolate markedness constraints and is considered in Chapter \ref{gaps}.
+
+Finally, it was argued that whatever the merits of these accounts of static constraints, there is also some potential value to the null hypothesis that static constraints do not exist at all: such a principle would allow for a resolution of certain intransigent debates in phonology, such as the proper analysis of the Sanskrit ``diaspirates''.
+Other cases of this type, where one analysis derives a phonotactic constraint and another makes no such prediction, surely could be adduced and investigated experimentally.
+
+\subsection{Chapter 2}
+
+Chapter \ref{gradience} is concerned with the claim that phonotactic wellformedness is not an ``all-or-nothing'' affair.
+This claim is first placed in a historical context.
+Then, three arguments are presented for \emph{a priori} skepticism regarding the claims of gradience in wordlikeness judgements.
+First, a model of gradient grammatical judgements seems to require extraordinary abilities for which there is little evidence, whereas reporting categorical judgements calls only on well-established capacities.
+Secondly, intermediate ratings in wordlikeness tasks, long taken as evidence for gradient models, are in fact present in any rating task which permits intermediate ratings, even when this is nonsensical.
+Finally, there have been no serious attempts to compare current gradient models to simple categorical baselines.
+
+The second half of this chapter is concerned with providing an evaluation which implements this comparison.
+A simple categorical baseline, as well as another baseline model measuring similarity to existing items, reliably outperform state-of-the-art models of gradient wellformedness.
+Once the categorical baseline effect is controlled for, gradient models are largely uncorrelated with the ratings.
+It is concluded that there is no evidence that intermediate ratings in wordlikeness tasks are the result of gradient grammaticality.
+
+This chapter suggests many directions for further research.
+First, it throws down a gauntlet to computational modelers who are proponents of gradient grammaticality, and the evaluation therein must be kept up to date when the challenge is answered.
+Furthermore, the categorical baseline presented in this model is intentionally quite primitive, and some of the proposed improvements discussed in the chapter are worthy of implementation and evaluation.
+The evaluation in this chapter is based on a small amount of published wordlikeness data. 
+It should be clear that the quality and quantity of this data is quite limited, but that it can be cheaply gathered.
+On analogy with undertakings like the English Lexicon Project \citep{ELP}, a publicly available database of lexical decision latencies, a carefully designed ``English Wordlikeness Project'' would be of great value to other phonotactic theorists, and would provide useful norming data for many other psycholinguistic tasks.
+
+\subsection{Chapter 3}
+
+The lexical/statistical model of phonotactic knowledge is the subject of Chapter 3. 
+While it has long been speculated that statistical criteria could be used to determine which phonotactic generalizations are internalized and which are ignored, it is argued that lexical statistics are neither necessary nor sufficient to determine this fact.
+On the contray, there are many static phonotactic constraints which are statistically reliable but synchronically inert: case studies include Latin rhotacism and restrictions on the English vowel system.
+A longer case study considers constraints on Turkish vowels.
+A host of evidence, phonological, statistical, and ``external'', suggests that vowel harmony is active throughout the language, despite a significant amount of lexical exceptions.
+Considerably less clear is the status of the static constraint known as \textsc{Labial Attraction}; at best, it counteracts the effect of harmony.
+
+Future work should further consider instances of ``statistically reliable/synchronically inert'' restrictions.
+Two obvious targets would be Mid-Atlantic short \emph{a} and Turkish \textsc{Labial Attraction}, which would benefit from further careful psycholinguistic investigation.
+
+\subsection{Chapter 4}
+
+Chapter 4 begins with the observation that many phonotactic constraints can be described with reference to prosodic primitives, and that this is in fact consistent with Stampean occultation.
+\citeauthor{Pierrehumbert1994}, however, has argued that not all constraints on English syllable contact clusters can be attributed to prosodic restrictions, and that various static constraints are necessary.
+However, \citeauthor{Pierrehumbert1994}'s study suffers from several methodological flaws.
+Principle among them is the failure to distinguish between derived and static constraints.
+A restudy shows that the former impose robust restrictions on the English syllable contact cluster inventory, whereas the static constraints proposed by \citeauthor{Pierrehumbert1994} lack statistical validity.
+
+Many of the clusters which appear to be prosodically well-formed are not attested, but state-of-the-art computational models are not able to detect any systematicity in the patterns of missingness. 
+Consequently, it is argued that many unattested clusters must be viewed as accidentap gaps, and that such is to be expected given the statistical properties of distributions of codas and onsets that make up medial clusters.
+
+Further psycholinguistic research could be used to further evaluate the claim made in this chapter, that many unattested syllable contact clusters are accidental gaps.
+The analysis could profitably be extended to medial clusters in other languages.
+
+\section{A note on architectural matters}
+
+A major finding of this dissertation is that the ``lexical/statistical'' theory of phonotactics is not sufficiently in evidence.
+Consequently, it is premature to assume the existence of a grammatical module which performs the task of computing phonotactic wellformedness.
+
+The alternative argued for here is that phonotactic knowledge is derived from familiar components of the (``narrow'') phonological grammar: phonological processes and prosodic structures like syllables, feet, etc.
+These too must be compiled into a module capable of recognizing ``possible'' and ``impossible' words, and work will need to be done to clarify how this is accomplished. 
+But, unlike the ``lexical/statistical'' theory, this knowledge comes online only in response to phonological acquisition. 
+As will be shown in what remains, this is in fact the view from research into phonological development.
+
+\section{Acquisition of phonotactic knowledge}
+\label{s:aopk}
+
+As a sort of final summary of the theories under evaluation in this dissertation, consider the evidence provided by the order in which phonotactic, phonological, and lexical knowledge is acquired.
+\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and it therefore must be independent of other types of grammatical knowledge.
+
+At 9 months of age, several studies imply that typically-develping infants have internalized langauge-specific preferences for native language phonotactics \citep{Friederici1993,Jusczyk1994}.
+A particularly telling study is performed by \citet{Jusczyk1993b}, who find that Dutch and English infants prefer to listen to lists of (likely unfamiliar) words in their native languages over those in English or Dutch, respectively.
+Crucially, these studies do not find the same preferences at earlier ages.
+Keeping in mind the usual caveats about a strong interpretation of negative results, this actually suggests that phonotactic acquisition occurs relatively late.
+
+Some non-trivial amount of lexical learning occurs at an earlier age, for instance.
+Typically-developing infants know their names and the names of their caretakers as early as 4 months of age \citep{Bortfeld2005,Mandel1995,Tincoff1999}.
+As early as 6 months of age, infants know the visual referents of familiar words presented auditorily \citep{Bergelson2012}.
+At 7.5 months, phonological representations are sufficiently detailed to allow infants to discriminate between words like \emph{cup} and mispronunciations like \emph{*tup} \citep{Jusczyk1995}.
+By 8 months, infants are able to locate both familiar and novel words in continuous speech \citep{Jusczyk1997,Seidl2006}.
+It is approximately at this time that the ability to discriminate non-native phonetic contrasts for vowels and consonants begins to decline \citep{Best1994,Polka1994,Werker1981,Werker1984,Werker1988}.
+
+While very few studies have investigated young infants' knowledge of phonological alternations, this is the subject of a fascinating study by \citet{White2008}.
+Simplifying somewhat, the experimenters expose 8.5-month-old infants to an artificial language in which fricative voicing is contrastive, but voiced and voiceless variants of plosives are in complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively.
+After familiarization, infants prefer to listen to nonce words preserving this complementary distribution of stops over nonce words which disrupt this distribution (e.g., \emph{na-poli}, \emph{rot-boli}), suggesting that the infants have extracted an allophonic generalization concerning plosive voicing.
+This too occurs before the earliest evidence for a more traditional type of phonotactic learning.
+
+It is not until much later that phonotactic knowledge can be investigated using productive language skills; the seminal study by \citet{Smith1973}, for instance, begins when the target child, Amahl, is more than 2 years of age.
+Given the familiar assymetry between comprehension and production (the former leading), it seems unlikely that much can be learned about the chronology of acquisition from this ``lagging indicator''.
+
+While there are many gaps in current understanding which merit future research, there is no reason to think that phonotactic knowledge is acquired before a considerable amount of lexical acquisition has occurred, or before children are capable of extracting phonological alternations and applying them to novel words.
+This observation that infants' phonotactic knowledge comes online only after highly specific phonological entries, subsyllabic representations, and alternation learning are available, is precisely what is predicted by the null hypothesis that phonotactic knowledge is derived from phonological processes and prosodic restrictions.
+
+%I would like to suggest that there are two types of problems that arise in developing a theory of phonotactics free of duplication.
+The first type of problem consists of conflicts between theoretical assumptions and the desire to eliminate duplication.
+If we continue to view duplication as a sort of negative heuristic, then it may be the case that the theoretical assumptions are wrong.
+Such a case arises in Chapter \ref{turkish} in the discussion of archiphonemic underspecification analysis of Turkish vowel harmony proposed by \citet{Clements1982}.
+\citeauthor{Clemenst1982} propose that all but the first vowel of a harmonic root is underspecified for backness (and in high vowels, roundness) and is filled in by rule.
+Since there are disharmonic roots, this rule must be ``structure-filling''.
+Consequently, this rule cannot account for the apparent markedness of disharmonic roots revealed by wordlikeness judgements (among other psycholinguistic tasks).
+If duplication is a pathology, this analysis is wrong.
+%The arguments of \citeauthor{Shibatani1973} and others led theorists to focus their attention on properties of surface representations as determinants of wordlikeness.
+Though syllabification plays no role in \emph{SPE}, it is crucial to many earlier studies \citep[for a review, see][]{Goldsmith2011b}, and it received particular attention in the 1970s.
+\citet{Hooper1973} and \citet{Kahn1976} argue that the syllable is useful for defining wordlikeness generalizations.\footnote{\citet{Steriade1999} and \citet{Blevins2003}, however, argue that a number of phonotactic generalizations previously stated in syllabic terms can be reanalyzed without making reference to syllables.} \citeauthor{Hooper1973} argues, for instance, that [bn], impossible as an English onset, is unobjectionable as a syllable contact cluster in nonce words like \emph{stabnik} (or in names like \emph{Abner}), and that this demonstrates the superiority of syllable-based wordlikeness generalizations.
+This already signals further trouble for alternative accounts which focus on underlying forms.
+Syllabification may span morphs, is generally predictable, and is universally non-contrastive, and as a consequence, few posit in to be present in underlying representations \citep[though see, e.g.,][]{Vaux2003}.
+\citeauthor{Hooper1973} also points to loanword adaptations which produce native syllable structure \citep[e.g.,][]{Carlisle1991} as evidence that syllabification is part of the phonological computation.
+Further enrichments to the theory are provided by the autosegmental theory of the syllable \citep{McCarthy1979b}, which envisions the syllable as an articulated tree structure \citep[as first envisioned by][]{Pike1947a}, and theories like prosodic licensing \citep{Ito1989a}, in which syllabification triggers phonological repairs.
+%Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature.
+At the risk of explaining what might be no more than an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms.
+\citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms.
+An example from German is shown in (\ref{fd}) below.
+
+%\emph{K}[uːx]\emph{en}
+%\emph{K}[uːç]\emph{en}
+%\footnote{Examples like \emph{Mas}[oːç]\emph{ist} `masochist', \emph{Eun}[uːç]\emph{ismus} `eunichism', first noted by \citet{Merchant1994}, suggest that this should also be restricted to assimilation within the same foot \citep[226f.]{Jensen2000}.}
+
+%Whether phonological computations or representations themselves are graded \citep[e.g.,][]{Lakoff1973} is besides the point, as metalinguistic judgements are behaviors, not mental states; they can no more be compared than can ``fear'' and ``flight response''.
+
+%\begin{example}[German \emph{ich}- and \emph{ach-laut}]
+%\begin{tabular}{l l l}
+%a. & [buːx]   & `book'           \\
+%   & [tɔxtər] & `daughter'       \\
+%   & [naxt]   & `night'          \\
+%b. & [siçt]   & `view'           \\
+%   & [ʃpeçt]  & `woodpecker'     \\
+%   & [ɡərʏçt] & `rumor'          \\
+%   & [knøçəl] & `ankle, knuckle' \\
+%   & [flɛçə]  & `surface'        \\
+%\end{tabular}
+%\end{example}
+%
+%\emph{Umlaut}, the fronting (and raising) of back vowels in certain morphological contexts, produces the front variant of the dorsal fricative; e.g., [lɔx]-[løçər] `hole-holes', \emph{B}[uːx]-\emph{B}[yːç]\emph{er} `book-books',
+% 
+%Throughout the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory (thought see \citealt{LignosInPressa} for some recent experimental evidence bearing on this question).
+%I reject the possibility that phonotactics do exist altogether.

### 2013-01-28 abs.tex
Pre-"revisions", with lots of material in the TODO
Revisions! Lots of 'em.

@@
-This dissertation outlines a program for the theory of phonotactics---the theory of speakers' knowledge of possible and impossible (or likely and unlikely) words---minimizing undue \emph{duplication} (in the sense of \citealt{KK77}:136f.) of phonotactics and phonology, and argues that the alternative view of phonotactics as stochastic, and of phonotactic learning as probabilistic inference, is not capable of accounting for the facts of this domain.
+% FIXME 350 word limit!
+
+This dissertation outlines a program for the theory of phonotactics---the theory of speakers' knowledge of possible and impossible (or likely and unlikely) words---and argues that the alternative view of phonotactics as stochastic, and of phonotactic learning as probabilistic inference, is not capable of accounting for the facts of this domain.
 Chapter \ref{intro} outlines the proposal, precursors, and predictions.
-%especially with regards to order of acquisition.
 
-Chapter \ref{gradience} considers evidence from wordlikeness rating tasks.
-It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded.
+Chapter \ref{gradience} considers evidence from wordlikeness rating tasks. It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded. 
 A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as accurately as state-of-the-art gradient wellformedness models.
@@
 Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns. 
-It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.
+It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations may go unlearned when they are not derived from phonological alternations.
 
@@
 The remaining gaps in the cluster inventory are attributed to the sparse nature of the lexicon, not static phonotactic restrictions.
+
+Chapter \ref{conclusions} summarizes the findings, considers their relation to order of acquisition, and proposes directions for future research.

### 2013-01-28 ack.tex
Revisions! Lots of 'em.

@@
-So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy.
-
-The ideas I first encountered under the tutelage of Jennifer Cole and Richard Sproat at the University of Illinois continue to guide my research.
+While many of my fellow students at Penn have had a hand in this document, none so much as my cohort: Toni Cook and Laurel MaKenzie.
+Without Joe Fruehwald, Marielle Lerner, Constantine Lignos, Brittany McLaughlin, and Hilary Prichard, it wouldn't have been nearly as much fun.
+Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar.
+John Murdock and Hilary Prichard proofread the entire document numerous times.
 
 The faculty, staff, and students of the Institute For Research in Cognitive Science provided me with not merely an office, but something close to a home.
-Thanks to my officemates, Chandan Narayan, John Trueswell, Laurel Sweeney and the rest of the staff there, as well as Amy Forsyth back at Williams Hall.
+Thanks to my officemates, Jessica Marcus, Chandan Narayan, John Trueswell, and especially Laurel Sweeney.
 
-So many of my fellow students at Penn have had a hand in this document, none so much as Toni Cook and Laurel MaKenzie, who have supported me so much over the past few years.
-Without Joe Fruehwald, Marielle Lerner, Constantine Lignos, Brittany McLaughlin, and Hilary Prichard, it wouldn't have been nearly as much fun.
-Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar.
-John Murdock fearlessly proofread everything, twice.
+The ideas I first encountered as an undergraduate at the University of Illinois under the tutelage of Jennifer Cole and Richard Sproat, continue to guide my research.
+I am grateful to the members of the committee, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment where quantitative and theoretical approaches can meet.
+Insofar as I have acknowledged and understood the arguments of my predecessors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee and to show me the wonders of New Haven.
+I do not expect to ever again have such a fearless advocate in the incomporable Charles Yang; the scientist he is is the one I strive to be.
 
-I am grateful to my committee members, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment for quantitative and theoretical approaches to meet.
-
-Insofar as I have acknowledged my historical precursors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee.
-
-More than anyone else, I owe a debt of gratitude to Charles Yang: I don't know how he puts up with me.
-
-I would like to conclude by thanking three people who were \emph{not} involved in the preparation of this dissertation, but without which it could never exist: Lila Gleitman, Morris Halle, and Bill Labov.
+While they were not involved in the preparation of this dissertation, it is hard to imagine that it would exist without Lila Gleitman, Morris Halle, and Bill Labov.

### 2013-01-28 pre.tex
First complete draft!
Revisions! Lots of 'em.

--- 2012-10-23 pre.tex
@@
 
-By
-some
-
-I can only hope that this humble theory has even a fraction of the influence on phonotactic thinking that
-
 There are two ways to read this thesis.
@@
 
-This negative heuristic also allows for the possibility of using psycholinguistic evidence to decide between competing phonological analyses.
-For instance, there is a long-standing debate concerning the representation of Sanskrit aspiration alternations like those represented by \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
-One analysis, going back to the ancient grammarian Pāṇini \citep{Sag1974} and to \citet[\S141f.]{Whitney1889}, posits an underlying representation /bud\asp{}-/ with aspiration shifting leftward in certain contexts: modern proponents of this analysis include \citet{Borowsky1983}, \citet[][59f.]{Hoenigswald1965}, \citet{Kaye1985}, \citet{Sag1974,Sag1976} \citet{Schindler1976}, \citet{Stemberger1980}, and \citet[][\S141f.]{Whitney1889}.
-An alternative analysis posits an underlying representation /b\asp{}ud\asp{}-/ and an apsirate dissimilation process (\citealt{Anderson1970}, \citealt{Hoard1975}, \citealt[\S3.2]{Kiparsky1965}, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt[109f.]{Zwicky1965}).
-It is not my intention to adjudicate between these competing analyses, or to evaluate the claim that both accounts are deeply flawed \citep{Janda1989}, but simply to note that they make different predictions under the theory proposed here.
-While the former, aspirate shift analysis denies the existence of diaspirate roots, this fact lacks synchronic motivation, so the co-occurrence restriction go unlearned.
-On the other hand, under the aspirate dissimilation account, underlying diaspirate roots exist but speakers internalize synchronic processes responsible for their absence on the surface.
-
 %\citet[38f.]{MacEachern1999} discusses a similar restriction on co-occurign aspirate roots in Ofo, but also notes, following \citet{DeReuse1981}, that this generalization holds both within roots and also within complex words.

### 2013-01-31 1.tex
Looks like this is wrapped up...!

@@
 In Russian, for instance, a process of anticipatory assimilation ensures that derived clusters of obstruents agree in voice.\footnote{
-    For sake of discussion, the complex behavior of [v] is ignored here.}
+    For sake of discussion, the complex behavior of [v] has been ignored here.}
 
@@
 \citeauthor{PE} go on to label constraints on underlying representations as ``extensional'', and thus irrelevant to generative grammar, in contrast to ``intensional'' statements like phonological rules.
-Regarding the facts about Russian obstruents discussed so far, it seems to be clear that the alternation facts are in some sense more privileged.
+Regarding the facts about Russian obstruents discussed so far, it is clear  that the alternation facts are in some sense more privileged.
 Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
@@
 Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
-For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; in other words, multiple ejective sequences are not merely absent, but also difficult for these speakers to repeat.
+For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; in other words, multiple ejective sequences are not merely absent, but also difficult for these speakers to perceive or produce.
 
@@
 Infants \citep[e.g.,][]{Mattys2001b} and adults \citep{McQueen1998b,Norris1997} are thought to use this heuristic for word recognition in experimental settings.\footnote{
-    There is reason to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
+    It is important to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
     Whatever the locus of [bnɪk]'s ill-formedness, for example, no segmentation into multiple words or morphs renders it a well-formed sequence.
@@
 However, there is no reason to regard these adaptations as a product of the synchronic phonology: no Yokuts root begins with a consonant, and there is no way to derive an initial CC cluster.
-Furthermore, as \citet{Gamble1989} notes, there is no reason to believe that the adaptations actually occurred within Yokuts at all, since there was at best quite limited contact between Spanish and Yokuts speakers; the adaptations may have occurred in another language altogether.
+Furthermore, as \citet{Gamble1989} notes, there is no reason to believe that the adaptations actually occurred within Yokuts at all, since there was at best quite limited contact between Spanish and Yokuts speakers; the adaptations may have occurred to conform to the phonotactic restrictions of another language, also in contact with Yokuts.
 
@@
 \citet{Davidson1997} note that final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
-Given the considerable disagreement about the nature of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.
+Given the considerable disagreement about the nature of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be ultimately be a useful diagnostic.
 
@@
 Language games or speech disguises may also provide evidence for phonotactic knowledge \citep[e.g.,][]{Vaux2011}, assuming that these ``alternative phonologies'' are not qualitatively different from naturally-occurring language processes \citep[e.g.,][697]{Bagemihl1995}.
-An example of language game evidence bearing on phonotactic questions can be found in \S\ref{ss:bh}, where a language game is used to argue for the hypothesis that root-internal vowel sequences in Turkish are subject to vowel harmony.
+An example of language game evidence bearing on phonotactic questions can be found in \S\ref{ss:bh}, where a language game is used to argue that root-internal vowel sequences in Turkish are subject to vowel harmony.
 
@@
 This dissertation will have little to say about the nature of segment structure constraints.
-Of more interest here are \emph{Sequence structure constraints} apply to underlying sequences; an example is given below.
+Of more interest here are \emph{sequence structure constraints}, restrictions on underlying phoneme sequences. 
+An example is given below.
 
@@
 \noindent
-This sequence structure constraint specifies that a consonant immediately after a word-initial stop is a liquid; this would preclude underlying */bnɪk/, for example.
+This sequence structure constraint specifies that a consonant immediately after a word-initial stop is a liquid; this would preclude underlying */bnɪk/, for example (though it would erroneously rule out attested *[sn], *[sm] onsets).
 
-However, \citet{Shibatani1973} shows quite convincingly that not all wordlikeness contrasts can be expressed as constraints on URs.
+However, \citet{Shibatani1973} argues that not all wordlikeness contrasts can be expressed as constraints on URs.
 In German, for instance, obstruent voicing is contrastive, but neutralizes finally: e.g., [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' vs.~[ɡʀaːt]-[ɡʀaːdə] `degree(s)'.

### 2013-01-31 2.tex
Looks like this is wrapped up...!

@@
 \includegraphics{agg.pdf}
-\caption{Subjects freely use intermediate ratings when asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively (\citealp{Armstrong1983}; ``1'' indicates ``most representative'')}
+\caption{When asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively, subjects use intermediate ratings (from \citealp{Armstrong1983}; ``1'' indicates ``most representative'')}
 \label{agg}
@@
 The evaluation below represents a first attempt to fill this gap.
-%FIXME clements and keyser
 
@@
 \citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables.
-Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words (see \S\ref{ova}).
+Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words
+%\citep{GormanInPressb}.
+(see \S\ref{ova}). %FIXME
 Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of English phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, several of which are without exception, into probabilities.
@@
 \citeauthor{Hayes2008a}'s principle precludes any attempt to test the hypotheses that underlies it, and therefore must be rejected.
-
-% FIXME make this work: 
-%(see \citealt{Sproat1993} and \citealt{Wrench2003} for articulatory investigations).
 

### 2013-01-31 4.tex
Looks like this is wrapped up...!

@@
 The full set of clusters and their frequencies are listed in Appendix \ref{clusters}.
-The CELEX transcriptions of these words are then syllabified and phonologized using the same procedure as used in Chapter \ref{gradience} and described in Appendix \ref{syllabification}. 
+The CELEX transcriptions of these words are then syllabified and phonologized using a procedure
+%the same procedure as used in Chapter \ref{gradience} and
+described in Appendix \ref{syllabification}. 
 In all, the sample contains 23 different medial coda and 40 different medial onsets.
@@
 
-Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original study is attempted: both their implementation and phonological feature specifications are used here, just as in Chapter \ref{gradience}.
+Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original study is attempted: both their implementation and phonological feature specifications are used here.
+%just as in Chapter \ref{gradience}. FIXME
 Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in Appendix \ref{syllabification}, and a novel feature [$\pm$ \textsc{Coda}] is added to allow the model to distinguish coda and onset consonants.

### 2013-01-31 ack.tex
Looks like this is wrapped up...!

@@
 Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar. 
-John Murdock and Hilary Prichard proofread the entire document numerous times.
+John Murdock and Hilary Prichard graciously proofread the entire document numerous times.
 
@@
 
-The ideas I first encountered as an undergraduate at the University of Illinois under the tutelage of Jennifer Cole and Richard Sproat, continue to guide my research.
+The ideas I first encountered as an undergraduate at the University of Illinois under the tutelage of Jennifer Cole and Richard Sproat continue to guide my research. 
 I am grateful to the members of the committee, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment where quantitative and theoretical approaches can meet. 
-Insofar as I have acknowledged and understood the arguments of my predecessors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee and to show me the wonders of New Haven.
+Insofar as I have acknowledged and understood the arguments of my predecessors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee and showed me the wonders of New Haven. 
 I do not expect to ever again have such a fearless advocate in the incomporable Charles Yang; the scientist he is is the one I strive to be.
-
 While they were not involved in the preparation of this dissertation, it is hard to imagine that it would exist without Lila Gleitman, Morris Halle, and Bill Labov.

### 2013-01-31 ms2.tex
Brought 4 up to speed with CLS 47 paper changes.
Looks like this is wrapped up...!

@@
 \setmathfont(Digits,Greek,Latin){Linux Libertine}
+%\setmainfont[Mapping=tex-text]{Times New Roman}
+%\setmathfont(Digits,Greek,Latin){Times New Roman}
 
@@
 
-\title{Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}} 
+\title{Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Constantine Lignos, Hilary Prichard, Charles Yang, and audiences at NELS 43 for helpful comments.}}
 \author{Kyle Gorman \\ University of Pennsylvania}
-\date{November 2012 (comments welcome)}
+\date{Revised January 2013 (comments welcome)}
 
@@
 \input{2}
+
 \appendix
 \renewcommand{\arraystretch}{0.25}
-
 \section{English wordlikeness ratings}
@@
 
-This appendix describes an automated procedure used to process the CMU pronunciation dictionary entries, separating medial clusters from their flanking nuclei, parsing the resulting sequences into coda and onset, and reversing allophonic processes targeting medial clusters.
+For every entry in the CELEX database, there is a corresponding broad syllabified transcription of the word in a Received Pronunciation accent.
+This appendix describes an automated procedure used to process these transcripts and to separate medial clusters from their flanking nuclei, parsing the resulting sequences into coda and onset, and reversing allophonic processes targeting medial clusters.
+
+While the segmental content of these transcriptions is precise, the CELEX syllabifications are unsystematic.
+Given the absence of contrastive syllabification in English (if not all languages: see \citealp[221]{Blevins1995}, \citealp{Elfner2006}), any sequence of a medial consonant cluster and its flanking nuclei should receive the same syllabification in all words in which it occurs.
+This is not always the case with the CELEX transcriptions, however.
+For instance, the sequence [ɪstɹɪ] receives a different parse in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is usually lax in Received Pronunciation \citep[][II.294]{AOE}.} Consequently, these syllabifications are not used here.
 
@@
 \noindent
-Consequently, pre-consonantal \emph{r} is assigned to the preceding nucleus.
+Following \citet{Pierrehumbert1994}, pre-consonantal \emph{r} is assigned to the preceding nucleus.
 
@@
 
-In English, when a medial consonant cluster is preceded by a stressed lax vowel, as \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel (\citealp[3]{Hammond1997}, \citealp{Treiman1990}).
+In English, when a medial consonant cluster is preceded by a stressed lax vowel, as \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel (\citealt[3]{Hammond1997}, \citealt{Treiman1990}).
 As \citet[55]{Harris1994} notes, however, when the medial cluster is also a valid onset, as in \emph{whi}[s.p]\emph{er}, \emph{ve}[s.ti]\emph{ge}, and \emph{mu}[s.k]\emph{et}, onset maximization will incorrectly assign the entire cluster to the onset and leave the lax vowel unchecked.
@@
 \subsection{Phonologization}
+\label{s:phonologization}
 
-The traditional analysis of affricates as single segments (e.g., \emph{SPE}:321f., \citealp[24]{Jakobson1961}) rather than stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990} is assumed here.
+Following \citet{Pierrehumbert1994}, the traditional analysis of affricates as single segment (e.g., \emph{SPE}:321f., \citealp[24]{Jakobson1961}) rather than sequences of a stop and fricative \citep[e.g.,][]{Hualde1988,Lombardi1990} is adopted here.
 In many languages, affricates pattern with simple onsets; for instance, Classical Nahua bans true onset clusters but permits the affricate series [ts, tʃ, tɬ] \citep[9]{Launey2011}.
-In English, however, affricates do not form complex onsets.
-Yet the stop and release phase of affricates cannot be separated by a syllable boundary, as predicted from the assumption they are single phonological units.
+Other languages, such as Polish, distinguish affricates and stop-fricative sequences \citep{Brooks1965}, providing further evidence that ``true'' affricates are represented as single segments (or single timing units), and in contrast with stop-fricative clusters \citep[34f.]{Clements1983}.
 
@@
 Onset [ŋ] is totally absent in onset position, where it cannot be followed by a /k, ɡ/ needed to derive the velar allophone, a fact predicted only by the former account, and English speakers have considerable difficulty producing initial [ŋ] \citep{Rusaw2009}.
-The allophonic analysis is assumed here.
-When followed by /k, ɡ/, [ŋ] is mapped to /n/.
+Following \citet{Pierrehumbert1994}, the allophonic analysis is assumed here. When followed by /k, ɡ/, [ŋ] is mapped to /n/. 
 When not followed by a velar stop (i.e., finally), [ŋ] is analyzed as underlying /nɡ/.
 
-%\section{Exemplification}
-% 
-%In (\ref{exemplification}), a few segmentations are given so as to demonstrate the principles the principles outlined above.
-% 
-%\begin{example}[Some examples of medial cluster segmentation]
-%\label{exemplification}
-%\begin{tabular}{l l l}
-%a. & \emph{tranquil} & /\ldots{}n.kw\ldots{}/ & (coda nasal place assimilation, [Cw] onsets) \\
-%b. & \emph{
-%c. & \emph{
-%d. & \emph{
-%e. & \emph{
-%f. & \emph{
-%g. & \emph{
-%h. & \emph{
-%i. & \emph{
-%j. & \emph{
-%\end{tabular}
-%\end{example}
+% what about Fromkin 1973 on sprig time for hintler
 
